### 2017-01-20 REIJ_article.tex
first commit

\documentclass[11.5pt,two column]{llncs}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{amssymb, amsmath}
\setcounter{tocdepth}{3}
\usepackage{multicol}
\usepackage{multirow}
\usepackage{url}
%\usepackage[margin=2pt]{subfig}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{authblk}
\usetikzlibrary{arrows}
\usepackage{enumitem}
\usepackage{sectsty}
\usepackage{array,url}
\usepackage{tikz}
\usetikzlibrary{arrows}
% change the appearance of the tikz arrow for the argumentation networks
% and some other settings to make all the graphs look similar
\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}

\addtolength{\oddsidemargin}{-0.85in}
\addtolength{\evensidemargin}{-0.85in}
\addtolength{\textwidth}{1.7in}
\addtolength{\topmargin}{-.7in}
\addtolength{\textheight}{1.3in}
\setlength{\columnsep}{0.3in}

% TODO command
\newcommand\todo[4][]{%
	\ifthenelse{\equal{#1}{resolved}}{%
		% nothing
	}{%
		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
	}%
}
\newcommand{\Mark}[1]{\textsuperscript{#1}}

% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
% \renewcommand\todo[4][]{}

% use like this:
% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex

\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}

\sectionfont{\fontsize{12}{15}\selectfont}
\subsectionfont{\fontsize{10}{15}\selectfont}

\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
\author{Marc van Zee\\
\vspace{-0.35cm}
Computer Science and Communication (CSC), University of Luxembourg\\
2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
\and
\vspace{-0.4cm}
Floris Bex\\
\vspace{-0.35cm}
Department of Information and Computing Science, Utrecht University\\
Princetonplein 5, 3584 CC Utrecht, Netherlands\\
florisbex@gmail.com
\and
\vspace{-0.4cm}
Sepideh Ghanavati\\
\vspace{-0.35cm}
Department of Computer Science, Texas Tech University\\
2500 Broadway, Lubbock, TX 79409, United States\\
sepideh.ghanavati@ttu.edu
}
\date{}

\begin{document}
\twocolumn[{%
 \centering
 \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
 \large Marc van Zee\Mark{1},
        Floris Bex\Mark{2},
        and Sepideh Ghanavati\Mark{3}\\[1em]
 \normalsize
 \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
 \end{tabular}\\[3em] % some more space after the title part
}]

\begin{abstract}
Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks. The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders. While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model. For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured. In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling. We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL). We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
\end{abstract}

\newpage

\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering

\section{Introduction}
\label{sect:introduction}
Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment. An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.

One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}. This is generally referred to as goal modeling. Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE. Several goal modeling languages have been developed in the last two decades. The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.

A goal model is often the result of a discussion process between a group of stakeholders. For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background. Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model. %SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops. In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.

\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}

\begin{enumerate}
\item
It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
\item
Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage. If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
\item
Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}. 
\item
It is not possible to reason about changing arguments and their effect on the goal model, and vice versa. A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model. Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments. This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
\end{enumerate}

Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above. We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.

The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture. However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models. The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed. %SGNew2: I am not sure if I understand the last sentence.

In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question. \todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done. If we write it this way, it means we are unsure of the result.}

\subsection{Contributions}

The contributions of this paper are aligned with our five success criteria. In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}. This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}. By combining arguments and critical questions, one can create a structured representation of a discussion. We, first, propose a set of arguments and questions we expect to find in those discussions. We, then, validate this initial set with transcripts containing discussions about the architecture of an information system. After this, we obtain our final set of arguments and critical questions.

In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework. We formalize the critical questions as algorithms modifying the argumentation framework. We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected. We propose an algorithm to generate a GRL model based on accepted arguments. This creates traceability links from GRL elements to underlying arguments (requirement 3).

We implement our framework into an online tool called RationalGRL. The tool is implemented using Javascript. It contains  two parts, goal modeling and argumentation. The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4). The argumentation part is new, and we develop a modeling language for the arguments and critical questions. %SGNew3: ADDED 
The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.

Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions. The methodology consists of two parts, namely argumentation and goal modeling. In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions. When switching to the goal modeling part, the accepted arguments are used to create a goal model. In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments. This might mean that the underlying arguments are no longer consistent with the goal models.

\subsection{Organization}

The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.

The rest of the article is in line with the main contributions. We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme. In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions. In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}. In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.

\section{Background: Goal-oriented Requirements Language and Argument Schemes}
\label{sect:background}

In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article. Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.

\subsection{Goal-oriented Requirements Language (GRL)}
\label{sect:background:grl}
The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs). %such as performance, cost,  security, and usability. 
URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness. URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements. In this article, we focus on GRL models and their links to argumentation.

GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders. It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions. A GRL model is a connected graph of intentional elements that optionally are part of actors. Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g. In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}). Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals (e.g. \textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g. \textsf{Agreement})

\begin{figure*}[ht]
\centering
\includegraphics[scale=0.35]{img/SchedulerModel}
\caption{GRL Model for Online Meeting Scheduler}
\label{fig:SchedulerModel}
\end{figure*}

Different links connect the elements in a GRL model. AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors. For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.

GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.

\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)? It used to be a separate section but I think it fits best here, have a look at what we need. I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}

GRL is an international standard with well defined syntax and semantics. This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction. Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation. Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.

According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details). Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models. This traceability enables direct impact analysis of goal changes on a design and vice versa. This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments. \todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain. This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture. }

\subsubsection{Online Meeting Scheduler Example}
\label{sect:background:casestudy}

In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times. The meeting scheduler has been modeled with Tropos~\cite{} and $i*$. We remodel this example with GRL in jUCMNav for the purpose of our framework.

The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}. The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e. \textbf{Meeting Participant}) can attend the meeting. The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available. Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend. Participants should then accept or reject the event.

The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model. However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model. Some of the questions that might be interesting to know about are the following:

\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
\begin{itemize}
	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}? Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
	\item What does \textbf{Quality of Proposed Date} mean?
	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
	\item What does \textbf{Richer Medium} mean? 
\end{itemize}

\todo[resolved]{Sepideh}{all}{We need to further improve this part}
\todo{Marc}{all}{Improved this part, what do you think?}
These are the type of the questions that we cannot answer just by looking at the GRL models. The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships. In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS). GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.

\subsection{Argument Scheme for Practical Reasoning (PRAS)}
\label{sect:background:pras}

\input{section22pras.tex}

\section{Preview of the Framework}
\label{sect:preview}

\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}

\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
\label{sect:gmas}

In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section. In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.

\begin{table*}[h]
\centering
\begin{tabular}{|l|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
\hline
AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
\hline
AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
\hline
AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
\hline
AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
\hline
AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
\hline
\hline
AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
& & CQ5b & Are there alternative ways of realizing the same goal?\\
\hline
AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
&& CQ6d & Does the task contribute to some other softgoal?\\
\hline
AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
&& CQ7b & Does the goal contribute to some other softgoal?\\
\hline
AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
\hline
AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
\hline
\end{tabular}
\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
\label{table:argument-schemes}
\end{table*}

\subsection{Argument schemes and critical questions}

Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model. Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions. We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).

The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}. Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question. CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS. AS5-AS9 are argument schemes for various GRL relationships between elements. As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship. Here we merely use the critical questions developed by Atkinson et al.

Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered. For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available. In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well. We will return to this issue in more detail in the next section. In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.

\section{GMAS: Second iteration}
\label{sect:gmas:2}

The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL. However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice. In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.

\subsection{Empirical evaluation with transcripts}

The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.

\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course. Two teams consist of three students, and one team consists of two students.

\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator. Designers are provided a problem description, requirements, and a description of the desired outcomes. The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}. Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging. Participants were asked to use a think-aloud method during the design session. For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}. The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}. All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software. The students had two hours for the tasks, and the transcripts document the entire discussion. The details of the transcripts are shown in table~\ref{table:transcripts:info}.

\begin{table}[ht]
\centering
\begin{tabular}{|l|l|l|l|}
\hline
& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
\hline
participants & 2 & 3 & 3\\
\hline
duration & 1h34m52s & 1h13m39s & 1h17m20s\\
\hline
\end{tabular}
\caption{Number of participants and duration of the transcripts.}
\label{table:transcripts:info}
\end{table}

\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}. If we found arguments or critical questions that did not appear in the list, we added them and counted them as well. Most of the occurrences were not literally found back, but had to be inferred from the context. For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.

It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts. Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself. Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts. In this way, we aim to make our annotation process as transparent as possible.

\subsubsection{Results}

Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}. We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts. The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}). Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).

Of our critical questions, we annotated 9 instantiations. Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).

\begin{table}[ht]
\centering
\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
\hline
\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
\hline 
AS0: Actor & 2 & 2 & 5 & 9\\
\hline
AS1: Resource & 2 & 4 & 5 & 11\\
\hline
AS2: Task/action & 20 & 21 & 17 & 58\\
\hline
AS3: Goal & 0 & 2 & 2 & 4\\
\hline
AS4: Softgoal & 3 & 4 & 2 & 9\\
\hline
AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
\hline
AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
\hline
AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
\hline
AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
\hline
AS9: Actor depends on actor &0& 1 & 3 & 4\\
\hline
\hline
\textbf{TOTAL} & 37& 41 & 42 & 120\\
\hline
\end{tabular}
\caption{Number of occurrences of AS0-AS9 in the transcripts.}
\label{table:transcripts:results:argumentschemes}

\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
\hline
\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
\hline 			
CQ2: Task is possible? & 2 & 2 & 1 & 5\\
\hline		
CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
\hline
CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
\hline
CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
\hline
\hline
\textbf{TOTAL} & 5 & 2 & 1 & 9\\
\hline
\end{tabular}
\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
\label{table:transcripts:results:criticalquestions}
\end{table}

Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences). The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.

\subsubsection{Analysis}

The analysis of the results of our empirical evaluation consists of three parts. First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.

\paragraph{Analysis of the argument schemes}
The difference between the initial list of argument schemes and those found back in the transcripts is quite small. We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system. This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent. However, we also found the following additional argument schemes:
\begin{itemize}
\item
\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition. We found that are large part of the discussions were focused around the tasks that the information system should provide. In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
\item 
\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
\end{itemize}
More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS. Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL. GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5. Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).

\begin{table*}[ht]
\centering
\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
\hline
\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
\hline
AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
\hline
- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
\hline
- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
\hline
AS11: Task contributes negatively to softgoal&2&1	&0&3\\
\hline
\hline
TI: Topic introduction & 5 & 3 & 7 &15\\
\hline
REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
\hline
CLAR: Clarifying an element &3 &10 & 3 & 16\\
\hline
GEN: Generic counterargument	& 0& 2 & 2 & 4\\
\hline
\textbf{TOTAL}&24&34&25&87\\
\hline
\end{tabular}
\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
\label{table:transcripts:results:new}
\end{table*}

\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions. On the one hand, we found back few of the critical questions we initially proposed. However, this does not mean that they weren't implicitly used in the minds of the participants. If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''. However, many of these critical questions are not mentioned explicitly. If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework. In the context of tool support, we feel that having these critical questions available may stimulate discussions.

On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
\begin{itemize}
\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
\item \emph{CQ11b: Is the decomposition correct? (AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur. However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
\end{itemize}
Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
\begin{itemize}
\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often. We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way. Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}. See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.

\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
\end{itemize}

\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments. For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system. In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''. In general, we found the following four operations that may be associated with the arguments:

\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
\begin{itemize}
\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything. Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition). An example is \emph{Clarifying an element}.
\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument. It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it. We explain this in more detail in the next section where we introduce our algorithms.
\end{itemize}

\subsection{Final argument schemes and critical questions}

Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}. Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.

In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question. In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations. In the next section, we formalize these operations as algorithms.

\begin{table}[ht]
\begin{tabular}{|l|l|}
\hline
\textbf{Argument} & \textbf{Operation}\\
\hline
AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
\hline
CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
\hline
CLAR & REPLACE\\
\hline
GEN & GENERIC\\
\hline
\end{tabular}
\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
\label{table:operation-mappings}
\end{table}

\section{The Formal Framework}
\label{sect:formalframework}

We briefly summarize our results so far and our tasks for the current section. In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations. In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}). We then validated this set using transcripts in section~\ref{sect:gmas:2}. This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}). We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC). In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.

In this section we develop a formal framework for the arguments and operations on them. In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements. In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model. In the third subsection we formalize the critical questions.

\subsection{Formal Language for a GRL Model}

In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.

In our language, each element of a goal model is denoted by a non-negative integer. We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.

\begin{definition}[Elements]
Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
\end{definition}

We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.

\begin{definition}[Name of IE or Actor]
The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
\end{definition}

All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.

\begin{definition}[Intentional Element]
Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
\end{definition}

In the same way, we distinguish between the different types of links between IEs using the following definition.

\begin{definition}[Links]
Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp. negative contribution. We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition. We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
\end{definition}

In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.

\begin{definition}[Elements of an actor]
Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
\end{definition}

\todo{Marc}{Marc}{Provide example here}

\subsection{Formalizing the argument schemes}

In order to formalize the argument schemes we first note that an

\section{The RationalGRL Tool}
\label{sect:implementation}

\section{The RationalGRL Methodology}
\label{sect:methodology}

The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}. There are two main activities, depicted in grey, namely "goal modeling" and "argumentation". These are two separate activities that are being done in parallel.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.4]{img/methodology}
\caption{The RationalGRL Methodology}
\label{fig:rationalgrl-methodology}
\end{figure}

\section{Related Work}
\label{sect:relatedwork}

There are several contributions that relate argumentation-based techniques with goal modeling. The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}. This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models. One of the elements of GAM is the translation of formal argument models to goal models (similar to ours). In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM. One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.  \textsf{RationalGRL} also provides tool support for argumentation, i.e. Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e. jUCMNav~\cite{jUCMNav}. Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation. Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.

\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact. DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs. One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.

Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements. However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks. Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented. Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.

\section{Conclusions and Future Work}
\label{sect:conclusion}

In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders. We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.

There are many directions of future work. There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not. It would be very interesting to explore the effect of these semantics on goal models. Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments. Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized. We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it. In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}). That is, the translation from goal models to argument diagrams is still missing. The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram. Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool. By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.

\section*{Acknowledgments}
Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.

\bibliographystyle{abbrv}
\bibliography{references}

\newpage

\onecolumn
\appendix

\section{Transcripts excerpts}
\label{sect:transcripts:excerpts}

\begin{table}[!htbp]
\centering
\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
\hline
0:11:00.8 (P2) & Development team, I don't know. Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
\hline
\end{tabular}
\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
\label{table:transcript:as0-cq0}

\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
\hline
0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
\hline
\end{tabular}
\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
\label{table:transcript:as1-as8}

\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:15:11.2 (P1) & And then, we have a set of actions. Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
\cline{1-2}
0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
\hline
0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
\hline
0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
\hline
\end{tabular}
\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
\label{table:transcript:as2-cq_star_1-cq2}

\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
\hline
0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled. (...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough? \newline
\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
\hline
\end{tabular}
\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
\label{table:transcript:as2-cq_star_2}

\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
\hline
\end{tabular}
\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
\label{table:transcript:as4-as5-as7}

\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context. That we have to determine who the users of the system are gonna be, stakeholders. & \textbf{[1 topic]} What are the actors?\\
\hline
\end{tabular}
\caption{AS*: Topic introduction (transcript $t_1$)}
\label{table:transcript:as-star}
\end{table}

\begin{table}[!htbp]
\centering
\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.  (...) we would have traffic light behavior gives you, I guess two options then. & \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
\cline{1-2}
0:30:23.6 (P1) & Sequences and timing schemes &\\
\cline{1-2}
0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
\cline{1-2}
0:30:30.9 (P1) & Or timing schemes&\\
\hline	
\end{tabular}
\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
\label{table:transcript:as2-as10}

\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
\cline{1-2}
0:30:14.3 (P3) &I think it's an OR&\\
\cline{1-2}
0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
\hline	
\end{tabular}
\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
\label{table:transcript:cq11b}

\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:06:29.3 (P2) & So, is that a trade-off. I think so. &\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\	
\cline{1-2}
0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality. Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional. Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic]. Is there around Utrecht always the same amount of cars? &\\
\hline	
\end{tabular}
\caption{AS11: Negative decomposition (transcript $t_1$)}
\label{table:transcript:as11}

\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
\hline
Respondent & Text & Annotation\\
\hline
0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
\textbf{[69 answer to 68]} Yes, type of cars.}\\
\cline{1-2}
0:49:20.1 (P1) & Maybe type of cars&\\	
\cline{1-2}
0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
\hline	
\end{tabular}
\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
\label{table:transcript:cq:task_decomp}

\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
\hline
Respondent & Text & Annotation\\
\hline

0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
\hline
0:11:00.8 (P2)&Development team, I don't know. Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
\hline
..&..&..\\
\hline
0:18:13.4 (P2) & I think we can still do developers here. To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
\cline{1-2}
0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
\hline	
\end{tabular}
\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
\label{table:transcript:as0-cq0-cq_counterarg}

\end{table}

\end{document}

### 2017-01-20 section22pras.tex
first commit

Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g. \cite{Raz1978-RAZPR,walton1990}) and Artificial Intelligence \cite{Bratman1987,atkinson2007}. One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}. The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action. This argument can then be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.

A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS). PRAS follows the following basic argument structure. 
\begin{equation}
\label{eq:eq1}
  \begin{aligned}
 \qquad&\text{I have goal } G&\\
&\text{Doing action }A \text{ will realize goal }G&\\
&\text{Which will promote value }V&\\
&\text{\emph{Therefore} I should do action }A.&
  \end{aligned}
\end{equation}

So, for example, we can say that 
\begin{itemize}
\item[] \textbf{Find agreeable meeting date} is a goal,
\item[] \textbf{Find agreeable date by talking to initiator} will realize the goal \textbf{(Find) agreeable meeting date},
\item[] Which will promote the value \textbf{User friendliness}
\item[] \textit{Therefore} 
\item[] We should perform action \textbf{Find agreeable date by talking to initiator}
\end{itemize}

Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information. Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements. Some examples of critical questions are as follows.

\begin{enumerate}
\item Will the action bring about the desired goal?
%\item Does the goal promote the value stated?
\item Are there alternative ways of realizing the same goal?
\item Are there alternative ways of promoting the same value?
%\item Does doing the action have a side effect which demotes the value?
\item Does doing the action have a side effect which demotes some other value?
\item Does doing the action promote some other value?
%\item Does doing the action preclude some other action which would promote some other value?
\item Is the action possible?
\item Can the desired goal be realized?
\item Is the value indeed a legitimate value?
\end{enumerate}

These critical questions can point to new arguments that might counter the original argument. Take, for example, critical question 8. If, for example, there is evidence that schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textsf{user friendliness} is not a legitimate value'. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3). For example, we can argue that \textbf{Find agreeable date using scheduler} also realizes the goal \textbf{Find agreeable meeting date}, which gives us a counterargument to the original argument -- to find a date by talking to the initiator -- that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}. In this section, we will briefly discuss the intuitions behind these semantics.}. Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for \textbf{Find agreeable date by talking to initiator} (A1), the argument for \textbf{Find agreeable date using scheduler} (A3), and the argument that \textbf{User friendliness} is not a legitimate value (A2). The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as \textbf{Find agreeable date by talking to initiator}  and \textbf{Find agreeable date using scheduler} are considered to be mutually exclusive. Argument A2 attacks A1, as it questions the legitimacy of the value \textbf{User friendliness}.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.8]{img/Fig1}
\caption{Argumentation framework}
\label{fig:pras:example}
\end{figure}

Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics. The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated. Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}. Argument A2 is undefeated because it has no attackers. This makes A1 defeated, because one of its attackers, A2, is undefeated. A3 is then also undefeated, since its only attacker, A1, is defeated by A2. Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.

In some cases, it is more difficult to determine whether or not an argument is defeated. Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two. It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}. Take, for example, A1 and A3. If we say that we prefer the action \textbf{Find agreeable date by talking to initiator} (A1) over the action \textbf{Find agreeable date using scheduler} (A3), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left). This makes A1 the only undefeated argument, whereas A3 is now defeated. It is also possible to give explicit arguments for preferences \cite{modgil2009}. These arguments are then essentially attacks on attacks. For example, say we prefer A1 over A3 because `the \textbf{Meeting participant} does not like working with a scheduler' (A4). This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.8]{img/Fig2}
\caption{Preferences between arguments}
\label{fig:pras:example2}
\end{figure}

\subsubsection{Practical Argumentation and Goal Modeling}
\label{sect:background:pras:motivation}

Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{amgoud2002reasoning,modgil2009}. The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}. The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).

PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) have some obvious similarities, and in previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016,vanZee-etal:comma2016} we presented arguments based on PRAS of the form ``G is a goal, Action A realizes G \emph{Therefore} perform action A'' (see Figure~\ref{fig:pras:example3}, left, where the arrow denotes an inference step). Such arguments, which can be constructed using the online OVA tool\footnote{\url{http://ova.arg-tech.org/}}, can be combined with further arguments providing, for example, expert opinions about the practical reasoning elements (Figure~\ref{fig:pras:example3}, left).

\begin{figure}[ht]
\centering
\includegraphics[width=\columnwidth]{img/Fig3}
\caption{Structured argument based on PRAS (left) and the corresponding GRL diagram (right)}
\label{fig:pras:example3}
\end{figure}

Arguments built in OVA can automatically be translated to GRL diagrams using further online tooling\footnote{\url{https://github.com/RationalArchitecture/RationalGRL}} \cite{vanZee-etal:comma2016}. This translation takes the elements of the arguments and directly translates this to GRL elements. Tasks/actions, goals and softgoals/values can be directly translated from the arguments to the GRL diagrams, as can realization statements. Elements of arguments that have nothing to do with goal modelling (e.g. the expert opinion premise in Figure~\ref{fig:pras:example3}) can be included in the GRL diagram as \emph{beliefs}. Finally, attacks between arguments are captured as negative contribution links. Take, for example, \todo{example}. A very similar approach to argumentation and goal models had already been taken by Jureta et al.~\cite{Jureta:RE2008} who proposed a formal argumentation model that can be used to justify goal-modelling decisions. This argumentation model, while not explicitly based on a practical reasoning scheme, is essentially the same as our previous argumentation model based on PRAS: a formal model of structured arguments with goals as premises and tasks as conclusions, and a translation function of these arguments to goal diagrams.

The core problem of the use of argumentation in the above-mentioned work \cite{Jureta:RE2008,vanzee-etal:renext2015,vanZee-etal:er2016}, is that it focuses mainly on structured arguments, which are then directly translated to goal models as in Figure~\ref{fig:pras:example3}. The question is what the added value of this exercise is. Firstly, the arguments and goal models contain many of the same elements. The idea is that arguments show the rationales behind goal models and the alternative or possibly conflicting views that were lost in the goal-modeling process. But GRL already contains a \emph{belief} element that allow one to provide beliefs underlying a goal model, an \emph{XOR decomposition} link that allows for modeling of alternative ways to realize goals, and \emph{negative contribution} links for capturing conflicting goals.

The point is that argumentation models are not needed to precisely render goals, tasks, beliefs and the relations between them, as GRL is already perfectly suited for this. Rather, argumentation should guide the goal modelling \emph{process}. Instead of static, formal argument trees, we need a dynamic argumentation process in which goals and tasks are critically analysed and which captures the dialectical nature of justification. Jureta et al.~\cite{Jureta:RE2008} take a step towards such a process by defining a high-level decision process that organizes argumentation and clarification techniques in relation to goal modelling. However, the argumentation techniques in this process are focused on building precisely the type of arguments that we argue are redundant next to GRL diagrams, namely structured arguments like in Figure~\ref{fig:pras:example3} which can be translated more or less directly to goal models. The decision process itself is never formally defined, and how to go from a discussion on goals and tasks to a goal model remains largely implicit.

What is needed is a formal framework that captures the discussions about goal models, and allows for (semi-)automated construction of goal models given this discussion. In section XX, we argue that a combination of Atkinson et al.'s~\cite{atkinson2007} critical questions and the dialectical semantics as formally captured by Dung's work \cite{Dung1995} are ideally suited to provide such a formal, dynamic framework.

### 2017-01-20 versions floris/REIJ_article_FB.tex
first commit

\documentclass[11.5pt,two column]{llncs}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{amssymb, amsmath}
\setcounter{tocdepth}{3}
\usepackage{multicol}
\usepackage{url}
%\usepackage[margin=2pt]{subfig}
\usepackage{xcolor}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{authblk}
\usetikzlibrary{arrows}
\usepackage{enumitem}
\usepackage{sectsty}
\usepackage{array,url}
\usepackage{tikz}
\usetikzlibrary{arrows}
% change the appearance of the tikz arrow for the argumentation networks
% and some other settings to make all the graphs look similar
\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}

\addtolength{\oddsidemargin}{-0.85in}
\addtolength{\evensidemargin}{-0.85in}
\addtolength{\textwidth}{1.7in}
\addtolength{\topmargin}{-.7in}
\addtolength{\textheight}{1.3in}
\setlength{\columnsep}{0.3in}

\newcommand{\todo}[1]{{\color{red}\textbf{[TODO: } #1 \textbf{]}}}
\newcommand{\Mark}[1]{\textsuperscript{#1}}
\newcommand{\floris}[1]{{\color{red}\textbf{Floris: }#1}}

\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}

\sectionfont{\fontsize{12}{15}\selectfont}
\subsectionfont{\fontsize{10}{15}\selectfont}

\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
\author{Marc van Zee\\
\vspace{-0.35cm}
Computer Science and Communication (CSC), University of Luxembourg\\
2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
\and
\vspace{-0.4cm}
Diana Marosin\\
\vspace{-0.35cm}
Luxembourg Institute of Science and Technology (LIST)\\
5 Avenue des Hauts-Fourneaux, 4362 Esch-sur-Alzette, Luxembourg\\ diana.marosin@list.lu
\and 
\vspace{-0.4cm}
Floris Bex\\
\vspace{-0.35cm}
Department of Information and Computing Science, Utrecht University\\
Princetonplein 5, 3584 CC Utrecht, Netherlands\\
florisbex@gmail.com
\and
\vspace{-0.4cm}
Sepideh Ghanavati\\
\vspace{-0.35cm}
Department of Computer Science, Texas Tech University\\
2500 Broadway, Lubbock, TX 79409, United States\\
sepideh.ghanavati@ttu.edu
}
\date{}

\begin{document}
\twocolumn[{%
 \centering
 \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
 \large Marc van Zee\Mark{1},
        Diana Marosin\Mark{2},
        Floris Bex\Mark{3},
        and Sepideh Ghanavati\Mark{4}\\[1em]
 \normalsize
 \begin{tabular}{*{4}{>{\centering}p{.22\textwidth}}}
  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Luxembourg Institute of Science and Technology & \Mark{3}Information and Computing Science & \Mark{4}Department of Computer Science \tabularnewline
  University of Luxembourg & Luxembourg & Utrecht University & Texas Tech University\tabularnewline
  \url{marcvanzee@gmail.com} & \url{diana.marosin@list.lu} & \url{florisbex@gmail.com} & \url{sepideh.ghanavati@ttu.edu}
 \end{tabular}\\[3em] % some more space after the title part
}]

%SG1: Abstract is good for now but when we are done with the paper, we should add our findings wrt the case study in the abstract too
\begin{abstract}
Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks. The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders. While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model. For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured. In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling. We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL). We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL. We evaluate the framework via a study among 20 subjects, showing that it is superior to traditional goal modeling with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects.  
\end{abstract}

\newpage

\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering

\section{Introduction}
\label{sect:introduction}
Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment. An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.

One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}. This is generally referred to as goal modeling. Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE. Several goal modeling languages have been developed in the last two decades. The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{URN}, which is part of an ITU-T standard, the User Requirements Notation (URN).

A goal model is usually the result of a discussion process between a group of stakeholders and a requirements engineer. For goal models are constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model. %SG1: This sentence doesn't make sense.. Do you mean for the small systems? The next sentence seemed to follow that logic..
However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops. In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.

\floris{These four reasons are very interesting. They could be improved by mentioning references in each: 2 and 3 are now without references, which makes it less strong.}

\begin{enumerate}
\item
It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}. Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model. If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions. As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
\item
Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage. If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
\item
Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost. For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''. Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost. This leads to a poor understanding of the problem and solution domain. In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}. 
\item
It is not possible to reason about changing arguments and their effect on the goal model, and vice versa. A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model. Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments. This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
\end{enumerate}

%SG2: This is good but make them a bit more clear and concrete so they can be used as our base for the evaluation. Agreed with floris too. I am a bit unsure about the first criteria. Why this is needed? The main goal is to help better decision making and have more informed decision based on the arguments. Try to consider that in your criteria (comparing goal models with or without argumentation)
Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above. We identified several important requirements for such an argumentation theory: (1) it must have prior successful use in the field of software development, (2) it should allow to formally reason about the consistency of discussions and the goal model, (3) it must be adaptable for use in goal models, (4) the adapted theory must produce repeatable results that can be formalized and implemented, (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools, and (6) the adapted theory must identify arguments and rationales that might not have been found, or might have been lost, otherwise.

\floris{interesting! ad (1) why must the argumentation theory have prior successful use in software development? What is `successful use'? An academic paper? A system prototype? A large-scale evaluation? Implementation in a system that is used daily in companies? ad(4) what do you mean by `repeatable results?'}

In this context, the main research question is: \emph{Can AI theories of practical reasoning and argumentation be adapted, formalized, and combined with an existing goal modeling methodology to rationalize goal model with underlying arguments, and to reason about their consistency?} The satisfaction of the above six requirements will result in a positive answer to the research question. %SG3: The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done. If we write it this way, it means we are unsure of the result.

\subsection{Contributions} %SG4:this part is a bit hard to read. I think a diagram helps and also make the contributions more aligned with the goals you have above.

The contribution of this paper is to show that the Argument Scheme for Practical Reasoning (PRAS) fits all of the six requirements above. We briefly discuss how PRAS meets the first two requirements with an example from the medical domain, particularly in reasoning about medical treatment for a patient. \floris{medical domain is not software development, so how does this address req. 1?} We, then, focus on issue three and four by modifying and implementing PRAS: We develop the Argument Schemes for Goal Modeling (GMAS), an extension to PRAS, which can be used to construct arguments for parts of a goal model. These arguments can be attacked with so-called ``critical questions'', which are natural language statements that question a certain aspect of the argument. We develop nine argument schemes and 18 critical questions. \floris{why does the preceding development of GMAS and the cq's address 3 in that it produces \emph{repeatable} results?} In order to reason about the consistency of the goal model with underlying arguments, we develop three \emph{consistency conditions} \floris{does this not address req. 2?}. We integrate GMAS with the Goal-oriented Requirements Language (GRL) and we implement it in jUCMNav, the Eclipse-based open source tool for GRL modeling. To address issue five, we develop the RationalGRL methodology that combines goal modeling with argumentation. Finally, we answer issue six positively through an empirical study \floris{is the explanation of the study copied from Singh et al>?} in which 20 subjects apply RationalGRL or GRL to develop a goal modeling in a information system concerning national security. The study shows RationalGRL is superior to GRL with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects. As a side effect, RationalGRL requires more time than GRL, but is justified by improvements in quality.

\subsection{Organization}

The rest of this paper is organized as follows: In Section~\ref{sect:background}, we introduce a goal modeling language, i.e. Goal-oriented Requirements Language (GRL)~\cite{} and the practical reasoning framework PRAS. We also provide a motivation and examples for both formalisms. In Section~\ref{sect:gmas}, we first analyze the differences between PRAS and GRL in terms of concepts and relationships, after which we introduce the Argument Schemes for Goal Modeling (GMAS) and associated critical questions. We, then, illustrate GMAS through several examples. In Section~\ref{sect:consistency}, we specify a consistency condition on the goal model and underlying arguments, while in Section~\ref{sect:implementation}, we provide a metamodel for GMAS and we link it to the GRL metamodel. We also discuss implementation in jUCMNav~\cite{} and the visualization details of our language.%SG5: I don't think we can call it our language. It can be an extenstion. For creating a new modeling language you need ontologies etc and we are not doing that. 
In Section~\ref{sect:methodology}, we explain how GMAS can be used by practitioners by introducing the RationalGRL methodology. It specifies how goal modeling and argumentation can be used to construct a GRL model. In Section~\ref{sect:evaluation}, we describe the design of our empirical study and the findings from it. Section~\ref{sect:relatedwork} presents related work and Section~\ref{sect:conclusion} discusses the conclusions and the future work.

\section{Background: Goal-oriented Requirements Language and Practical Argumentation}
\label{sect:background}

In this section, we introduce the Goal-oriented Requirements Language (GRL) and the Argument Scheme for Practical Reasoning (PRAS).

\subsection{Goal-oriented Requirements Language (GRL)}
\label{sect:background:grl}

The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs). %such as performance, cost,  security, and usability. 
URN allows software and requirements engineers to identify  requirements for a proposed or an evolving system and to review such requirements for correctness and completeness. URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements. In this article, we restrict ourselves to GRL.

GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders. It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions. A GRL model is a connected graph of intentional elements that optionally are part of actors. An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself. Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied. Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way. Softgoals are often more related to NFRs, whereas goals are more related to functional requirements. Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals. In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available.

A variety of links connect the elements in a GRL model. AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements. Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element. A contribution link has a qualitative contribution type or a quantitative contribution. Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.

GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$. Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.

\subsubsection{Motivation for using GRL}
\label{sect:background:grl:motivation}

GRL has the existing tools that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented. (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools. GRL is an international standard with well defined syntax and semantics. This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying issue 4. Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation. Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.  According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details). Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models. This traceability enables direct impact analysis of goal changes on a design and vice versa. This traceability feature seems to be a good fit with the current research: We aim to add traceability links between intentional elements and their underlying arguments. %SG8: what do you mean by current research?

\subsection{Argument Scheme for Practical Reasoning (PRAS)} %SG10: This section is good but make the examples clearer. It has too much concepts but it makes me a bit lost.
\label{sect:background:pras}

Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g. \cite{raz1978,walton1990}) and Artificial Intelligence \cite{bratman1987,atkinson2007}. One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}. The idea is that an instantiation of such a scheme gives a presumptive argument in favour of, for example, taking an action. This argument can then be tested by posing critical questions about, for example, if the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.

A formal approach to argumentative about actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS), as follows. 
\begin{itemize}[leftmargin=22pt]
\item[PRAS]
In the circumstances $R$,\\
agent $Ag$ should perform action $A$,\\
to achieve new circumstances $S$,\\
which will realize goal $G$,\\
which will promote some value $V$.
\end{itemize}

This scheme distinguishes between the state of affairs brought about by an action, the goal (i.e., the desired features in that state of affairs), and the value (i.e., the motivation why those features are desirable). Take, for instance, the example from \cite{vanZee-etal:er2016}: the company (actor) does not have a standard corporate data model (current circumstances), and therefore decides to define such a model (action), leading to a new data model (new circumstances), which realizes the goal to adhere to a corporate data model, which promotes lower costs (value).

In practical reasoning, conclusions which are at one point acceptable can later be rejected because of new information. For example, if we find out that `adhere to corporate data model' is not a realistic goal, we can argue that we should not define such a new data model. Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized. An unfavourable\footnote{Unfavourably can be either 'no', or 'yes', depending on what the question is.} answer to a critical question then identifies a possible counterargument. One critical question is `Does the goal realize the value stated?'. For our example, we can answer this question favourably with a `yes', as one way to lower costs is to adhere to a company's corporate data model \cite{vanZee-etal:er2016}. Another critical question is `Can the desired goal be realized?'. If the answer to this is `no' - for example, there is evidence that `adhere to corporate data model' is not a realistic goal, then we would say the question has been answered unfavourably, and we have a counterargument: `evidence shows that there is not much attention for adherence to a corporate data model, so it can probably not be realized'. Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal or an alternative goal that promotes the same value. For example, in \cite{vanZee-etal:er2016} it is argued that the action `use data model of package applications' realizes the goal `use canonical data model', which also promotes lower costs. This gives us a counterargument that also follows PRAS.

In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa). In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an emph{argumentation framework} $(Args, Attacks)$, as introduced by Dung~\cite{Dung1995}, where $Args$ is the set of arguments and $Attack$ is an attack relation between arguments. Figure \ref{fig:pras:example1} shows that an argumentation framework that includes the arguments from the above example can be rendered as a graph. The two alternative PRAS instantiations are A1 and A3. These arguments mutually attack each other, as `adhere to corporate data model' (A1) is an alternative to `adhere to canonical data model' (A3) and vice versa. Argument A2 attacks A1, as it questions the validity of the goal `adhere to corporate data model'.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{img/Fig1}
\caption{Argumentation framework}
\label{fig:pras:example}
\end{figure}

Given an argumentation framework, the acceptability of arguments can be determined according to the semantics of Dung~\cite{Dung1995}. The idea is that we are justified in accepting an argument if it can be successfully defended against attacking counterarguments. In other words, an argument $A \in Args$ is \emph{defeated} by an attacker $B \in Args$ -- where $(A, B) \in Attacks$ -- unless $B$ itself can be defeated. Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}. Argument A2 is undefeated or justified, because it has no attackers. This makes A1 defeated, because one of its attackers, A2, is undefeated. A3 is then also justified, because its only attacker, A1, is defeated by A2. Thus, the set of justified (undefeated) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.

attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two. However, A2 also attacks A1

 where $Args$ is the set of arguments and $Attack$ is an attack relation between arguments. Dung proposes that an argument $A_1\in Args$ is \emph{defeated} by an attacker $A_2\in Args$, unless $A_2$ itself can be defeated. Although this is an appropriate definition of reason about beliefs, it is not for reasoning about action since we are, to some extent, free to choose what we attempt to bring about. Therefore, we may choose to reject an attacker, even if it cannot be defeated, if we believe that the purpose motivating the attacked argument is more important. For instance, suppose an argument $A_2$ stating that a particular resource is expensive attacks an argument $A_1$ for using the resource. However, we may still choose to use it (i.e., accept argument $A_1$), if that would promote a value that we judge more importantly than expense. In order to accommodate this, \emph{valued-based argumentation frameworks} (VAF)~\cite{bench2003persuasion} are used. The idea is that we are given an \emph{argumentation framework}, a set of values $V$, a function $val$ mapping each argument to a value $v\in V$, and a set of \emph{audiences} $a$ to which the arguments are addressed. Each audience is then represented by a strict partial ordering $>_a$ on the values. Then, for a given audience $>_a$, $A_2$ defeats $A_1$ iff $A_2$ attacks $A_1$ and $val(A_2)\not>_a val(A_1)$. \floris{you do not give formal definitions here}

It is then possible to determine which arguments in $Args$ are acceptable for an audience by determining the preferred extension. This is the maximal (under set inclusion) subset $S$ of $Args$ such that no two arguments in $S$ defeat each other, and all arguments in $S$ are acceptable with respect to $S$, i.e., for any argument $A_1$ in $S$, if $A_1$ is defeated by an argument $A_2$ that is not in $S$, then there exists an argument in $S$ that defeats $A_2$. The preferred extension represents the maximal consistent set of acceptable arguments with respect to the argumentation framework and a given audience or value ordering.

\subsubsection{PRAS Medical Treatment Example} %SG11: Isn't it better to use an example more relevant to what we do?
\label{sect:background:pras:example}

A concrete example of using PRAS comes from the medical domain, where Atkinson and colleagues use practical reasoning to formalize a system for reasoning about the treatment of a patient~\cite{atkinson2006argumentation}. The system in this example consists of a group of experts in different areas, such as \emph{treatment knowledge} and \emph{patient knowledge}.\footnote{The original application uses a single expert using different knowledge bases, but we changed this to a group of experts, each with a different expertise, in order to make the comparison with goal modeling more apparent.} There are five critical questions relevant to the application:
\begin{enumerate}
\item[CQ1] Are there alternative ways of realizing the same effects?
\item[CQ2] Are there alternative ways of realizing the same goal?
\item[CQ3] Are the assumptions on which the argument is based true?
\item[CQ4] Does performing the action have side effects which demotes some other value?
\item[CQ5] Will the action have the effects described?
\end{enumerate}

The experts can construct an argumentation framework for a medical treatment by instantiating PAS and posing critical questions CQ1-CQ5. Consider for instance a patient whose health is threatened by blood clotting. The experts start with the null option \--- do nothing (EA0). Then, suppose that the treatment knowledge experts suggests blood clotting can be prevented by reducing platelet adhesion, which can, assuming aspirin is not contraindicated, be achieved by prescribing aspirin. This information is used to instantiate PAS:\\

\noindent
\textbf{EA1.} \emph{Assuming no contradictions, we should prescribe aspirin, which will reduce platelet adhesion, preventing blood clotting, and so is an efficacious course of action.}\\

\noindent
This argument is then subjected to the critical questions. CQ1 asks for alternative solutions to reduce platelet adhesion. The expert may reply by stating chlopidogrel also reduces platelet adhesion. Asking CQ2 forces the expert to seek further solutions for preventing blood clotting, which results in an alternative action, namely to administer streptokinase, which has the same goal of preventing blood clotting. These are formed into two arguments, EA2 and EA3:\\

\noindent
\textbf{EA2.} \emph{Assuming no contradictions, we should prescribe chlopidogrel, which will reduce platelet adhesion, preventing blood clotting, and so is an efficacious course of action.}\\\\
\textbf{EA3}. \emph{Assuming no contradictions, we should prescribe streptokinase, which will increase blood clot dispersal, preventing blood clotting, and so is an efficacious course of action.}\\

\noindent
These three arguments all mutually attack one another, resulting in the argumentation framework shown in figure~\ref{fig:pras:example1}.

Each of the arguments are in mutual conflict, so it is not clear which action to pursue. This depends on the preferences of the experts. For this, suppose the experts vote and agree the effect `reduce platelet adhesion' is a preferred means by which the goal can be realized. Moreover, they prefer aspirin to chlopidogrel because it is cheaper.

These preferences are represented in the argumentation framework by removing attacks of unfavorable actions, so EA1 is no longer attacked, and EA3 no longer attacks EA2. In this case, EA1 forms the preferred extension, and its action is currently the best candidate. However, there are still remaining critical questions that can be asked. For instance, EA1 assumes that aspirin was not contraindicated, and CQ3 instructs us to test this assumption. Suppose that the patient knowledge expert notes the patient has a history of gastritis, and infers that aspirin is contraindicated because its acidity may result in gastric ulceration. The expert forms this into an argument motivated by the value of safety:\\

\noindent
\textbf{EA4.} \emph{Where there is a history of gastritis and no acid reducing therapy, we should not prescribe aspirin, which would cause excess acidity, which would risk ulceration, and so is unsafe.}\\

\begin{figure}[ht]
\centering
\includegraphics[scale=0.4]{img/pras-example2}
\caption{Final argumentation framework showing all critiques from~\cite{atkinson2006argumentation}}
\label{fig:pras:example2}
\end{figure}

\floris{please explain EA5, 6 and 7 as well}
In a similar manner the experts construct three more arguments that attack other arguments in various ways, resulting in the argumentation framework shown in figure~\ref{fig:pras:example2}. In order to determine which arguments are accepted, the preferred extension is computed by first adding the arguments with no attackers, which are EA5, EA6, and EA7. EA7 defeats EA4, which means that EA1 can be included, since its only attacker is defeated. EA1 therefore defeats EA2 and EA4, and also excludes EA0. Thus, in the end, the winning argument is EA1, which prescribes aspirin.

\subsubsection{Motivation for using PRAS}
\label{sect:background:pras:motivation}

In the introduction, we discussed six requirements for a practical reasoning and argumentation framework to be used in our research. The first requirement is: (1) it must have prior successful use in the field of software development. The example of the previous subsection is just one of many where PRAS has been successfully used by computer scientists to formalize arguments in a computer system. \floris{has it been implemented? and does that mean that it has prior successful use in softw dev?} The usefulness of the argument scheme approach for the analysis of practical reasoning situations has been shown in various other areas including e-democracy~\cite{Atkinson2006}, law~\cite{bench2005persuasion} and reasoning about the morally correct course of action~\cite{bench2006automating} \floris{Safety critical actions, explaining character motives in stories, planning}. The second requirement from the introduction is: (2) it should allow to formally reason about the consistency of arguments. The formal semantics in terms of value-based argumentation frameworks allows for the computation of accepted arguments. This semantics is extensively studied in the literature, and various different semantics exists, which lead to different flavors of acceptability of arguments.~\cite{} The third requirement from the introduction is: (3) it must be adaptable for use in goal modeling. In the next section we adapt PRAS in order to do so, but already at first glance the concepts that are used in PRAS and in goal modeling are very similar. In GRL, ``tasks'' can be performed by actors, while in argument scheme PAS these are called ``actions''. Both PAS and GRL have the concept of a ``goal''. Finally, the GRL notion of a ``softgoal'' seems semantically identical to PAS's ``value'', namely that it is a underlying motivation for which goals are desirable. We therefore decided that PRAS does indeed meet our first three requirements of a formal argumentation theory to be used in our research.

%SG12: Read up to the section below.
\section{Argument Schemes for Goal Modeling (GMAS)}
\label{sect:gmas}

PRAS is not directly applicable to goal modeling in general, and GRL in particular. There are differences in the concepts used, and the relationships between these concepts. Some elements appear in PRAS that do not appear in GRL, some elements appear in GRL but not in PRAS, and some concepts appear in both formalisms but with different names. Moreover, GRL allows more relations between elements than the ones appearing in PAS. In this section, we first analyze these differences in more detail. After this, we introduce and motivate Argument Schemes for Goal Modeling (GMAS) and associated critical questions, which are an extension to PRAS. While PRAS contains a single argument scheme PAS, GMAS instead consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model. With each argument scheme we associate critical questions that can be used to attack arguments or provide alternatives. This enables focused discussions about elements and relationships of a goal model. In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.

\subsection{Differences between GRL and PRAS}

\begin{table*}[h]
\centering
\begin{tabular}{|l|l|l|l|l|l|}
\hline
\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Attacks?}\\
\hline
AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & Yes\\
\hline
AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & Yes\\
\hline
AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & Yes\\
\hline
AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal? & Yes\\
\hline
\hline
AS5 & Task $T$ realizes goal $G$ & CQ5a & Will the task realize the desired goal? & Yes\\
& & CQ5b & Are there alternative ways of realizing the same goal? \floris{How do you handle OR and XOR decompositions?} & No\\
\hline
AS6 & Task $T$ contributes to softgoal $S$ & CQ6a & Are there alternative ways of contributing to the same softgoal? & No\\
&& CQ6b & Does the task have a side effect which contribute negatively to the softgoal?&Yes\\
&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal? & No\\
&& CQ6d & Does the task contribute to some other softgoal?&No\\
&& CQ6e & Does the task preclude some other task which would contribute to some other softgoal?&Yes\\
\hline
AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal? & Yes\\
&& CQ7b & Does the goal contribute to some other softgoal? & No\\
\hline
AS8 & Resource $R$ contributes to task $T$ & CQ8a & Is the resource required in order to perform the task? & Yes\\
&&CQ8b & Can other resources be used to perform the same task? & No\\
&&CQ8c & Is the resource required in order to perform the task? & Yes\\
&&CQ8d & Does using the resource make other resources unavailable? & Yes\\
\hline
AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors? & No\\
\hline
\end{tabular}
\caption{Argument schemes and critical questions for GRL elements (CQ1-CQ4) and relationships (CQ5-CQ16), as well as whether answering the critical question unfavorably results in an attack on the original argument (right column).}
\label{table:argument-schemes}
\end{table*}

The concept that appears in PRAS and not in GRL is \emph{circumstance}. In PRAS, an action $a$ is performed in current circumstances $R$, and by performing this action new circumstances $S$ are obtained. From this new circumstance $S$, another action may be performed again, and so on. In the AI literature, this type of reasoning usually referred to as ``AI planning''~\cite{weld1999recent}: from some initial state, find a sequence of actions such that a goal state is reached if these actions are executed. This planning approach is taken by a number of contributions using PRAS to create joint plans for a group of actors, often using a transition system as the underlying semantics~\cite{medellin2014}. However, in the early RE phase of an IS one generally does not consider temporal planning. Decisions about the temporal order in which tasks should be executed are generally postponed until a later stage. In other words, goal modeling is not concerned with making plans, scenarios, or business processes for the information system. In URN (see section~\ref{sect:background:grl}) the distinction between goal modeling and business process modeling is clearly visible in the distinction between two separate modeling languages for the two activities. First, goal models are created in GRL, while after this use case maps (similar to business processes) are created in UCM. Therefore, when extending PRAS for GRL we do not consider the notion of \emph{circumstances}.

The concept that appears in GRL but not in PRAS is \emph{resource}. A resource may be used by a part of the system, or an actor, in order to perform a task in order to reach a goal. Therefore, PRAS has to be extended with the notion of a resource in order for it to be applicable to GRL.

Finally, there are two concepts appearing in both formalisms under different names. Where PRAS uses the notions ``action'' and ``value'', GRL uses the concepts ``task'' respectively ``softgoal''. However, these two concepts are semantically identical.

With respect to relationships between concepts, the argument scheme PAS from Section~\ref{sect:background:pras} contains the following relationships:\\

\noindent
R1. Actor $AG$ \emph{performs} action $A$\\
R2. Action $A$ \emph{realizes} goal $G$\\
R3. Goal $G$ \emph{promotes} value $S$\\
R4. Actor $AG$ \emph{has} goal $G$\\
R5. Actor $AG$ \emph{has} value $S$\\

\noindent
Relationships R4 and R5 are implicit in PAS, but follow directly from its formulation (see~\cite{atkinson2007} for more details). The relationship ``promotes'' does not exist in GRL, but the relationship ``contributes to'' does, which is conceptually very similar. Therefore, we choose to use this relationship to formalize the ``promotes'' relation.

As mentioned in Section~\ref{sect:background:grl}, GRL is a very non-restrictive modeling language, allowing intentional elements and links to be combined freely. Thus, GRL allows any links between any type of intentional element. PRAS, in contrast, only considers relationship R1-R5 above. As a result, the critical questions of PRAS are only applicable to a subset of all possible relationships in GRL. This, however, does not mean it is not possible to construct arguments and formulate critical questions for other relationship.  In section~\ref{sect:implementation:implementation}, where we discuss the implementation of our tool, we explain in more detail that a user can straightforwardly add additional critical questions for other relationships as well.

\subsection{The Argument Schemes and Critical Questions}

Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model. Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions. We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).

We developed a total of 9 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}. Argument schemes AS1-AS4 can be instantiated for individual GRL elements, and each have a single critical question. CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS. AS5-AS9 are argument schemes for various GRL relationships between elements. As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship. Here we merely use the critical questions developed by Atkinson et al.

Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered. For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available. In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well. In general, answering a critical questions can have two effects:
\begin{enumerate}
\item
Answering the critical question unfavorably results in a new argument attacking the original argument (\emph{attack}).
\item
Answering the critical question unfavorably results in a new argument not attacking the original argument, but representing an alternative (\emph{no attack}).
\end{enumerate}
The last column of table~\ref{table:argument-schemes} shows for each critical question whether it results in an attack on the original argument or not.

\floris{you do not provide a formal definition of GMAS}

\subsection{Example: Airport Assistant}

We now provide a simple scenario of arguments created by instantiating AS1-AS9, and demonstrate how new arguments can be generated through critical questions, possibly attacking the original arguments.

Consider a system for assisting users at the airport. The assistant should be able to guide users through the airport, inform them about delays, and provide other information such as the location of the check-in register for each airline company. When constructing a goal model for this system, one stakeholders argues that the user interface of the assistant should by easy to use, and that this can be realized by using a Graphical User Interface (GUI). We can formalize this as an instantiation of AS6:\\

\noindent
\textbf{$A_1$.} \emph{Task ``Use GUI'' contributes to softgoal ``easy to use'' (AS6)}.\\

\noindent
This argument can be subjected to critical questions. In this case, there are five critical questions that can be used on AS6, namely CQ6a-CQ6e. Critical question CQ6a asks for alternative solutions to contribute to softgoal ``easy to use''. One of the stakeholders may reply by stating a natural language interface is easy to use as well.\\

\noindent
\textbf{$A_2$.} \emph{Task ``Use NL interface'' contributes to softgoal ``easy to use'' (CQ6a/AS6)}.\\

\noindent
Note that $A_2$ does not create an attack on $A_1$, since both are valid tasks to contribute to the softgoal. Instead, $A_2$ actually generates new elements that will appear in the goal model.  Now both $A_1$ and $A_2$ can be questioned. Suppose CQ6b is used by a stakeholder to argue that a side effect of using a natural language interface is that not all users may be able to understand it. When using a GUI, it may be easier to let the user change its language.\\

\noindent
\textbf{$A_3$.} \emph{Task ``Use NL interface'' contributes negatively to softgoal ``easy to use'' because choosing a language is difficult (CQ6b).}\\

Suppose another stakeholder uses critical question CQ6c to question $A_1$: she notes that implementing a GUI is probably quite expensive, because they don't have a graphical designer, so they will have to outsource this.\\

\noindent
\textbf{$A_4$.} \emph{Task ``Use GUI'' contributes negatively to softgoal ``low costs''}.\\

\noindent
The resulting argumentation framework of these four arguments is shown in figure~\ref{fig:example:airport:arguments}. As we can see, argument $A_1$ and $A_4$ are both not attacked by any argument. However, $A_2$ and $A_3$ are in direct conflict. Recall from sectin~\ref{sect:background:pras} that the stakeholders now have to decide about their preferences. Suppose the stakeholders decide that argument $A_3$ is stronger, or more preferred, than $A_2$. As a result, the accepted arguments are $A_1, A_3$, and $A_4$.

\begin{figure}
\centering
\begin{tikzpicture}[scale=2]
\node (a1) at (0,1) {$A_1$};
\node (a2) at (1,1) {$A_2$};
\node (a3) at (1,0) {$A_3$};
\node (a4) at (0,0) {$A_4$};
\draw [->] (a2) to [out=-60, in=60] (a3);
\draw [->] (a3) to [out=120, in=-120] (a2);
\end{tikzpicture}
\caption{Argumentation framework for Airport Assistant example.}
\label{fig:example:airport:arguments}
\end{figure}

It is now possible to translate this argumentation framework into a GRL model, which is shown in figure~\ref{fig:example:airport:grl}. Argument $A_1$ results in a contribution link from task ``use GUI'' to softgoal ``lower costs'', similarly for $A_4$. Since the conflict between $A_2$ and $A_3$ has been resolved in favor of $A_3$, there is a negative contribution link from task ``use NL interface'' to softgoal ``Easy to use''.

We can make a number of observations about the obtained GRL model and the underlying arguments:
\begin{itemize}
\item The underlying arguments contain information that is not represented in the GRL model. For instance, the decision to reject $A_2$ (i.e., that a natural language interface is easy to use) is not visible in the GRL model, but it is in the underlying arguments.
\item There is a direct relation between the accepted arguments and the elements that appear in the goal model. If an element or relation has an underlying accepted argument, it appears in the goal model, while rejected ones do not appear. Therefore, if the underlying arguments change \--- for instance if the stakeholders change their preference from $A_2$ to $A_3$ \-- then the resulting goal model changes as well.
\item In the current example each GRL element and relation is supported by an argument. However, this is not required. One could imagine that the stakeholders choose to add additional elements to the goal model without rationalizing them by arguments. We explain this in more detail in the RationalGRL methodology in section~\ref{sect:methodology}. 
\item Although it may be intuitively understandable how to translate argument $A_1-A_4$ and the argumentation framework of figure~\ref{fig:example:airport:arguments} into the GRL model of figure~\ref{fig:example:airport:grl}, we have not yet formally specified how to make the translation from GMAS to GRL, and visa versa.
\end{itemize}

In this section we have fulfilled requirement three of the introduction: (3) it (the practical reasoning and argumentation theory) must be adaptable for use in goal modeling. However, we have not yet considered the fourth requirement: (4) the adapted theory must produce repeatable results that can be formalized and implemented. This is what we do in the next section by providing formal mappings from GMAS to GRL, and discussing implementation details.

\begin{figure}
\centering
\includegraphics[scale=0.35]{img/grl_example_airport}
\caption{GRL model for Airport Assistant example.}
\label{fig:example:airport:grl}
\end{figure}

\section{Mapping GMAS to GRL: Metamodel and Implementation}
\label{sect:implementation}

\begin{figure*}[h]
\includegraphics[scale=0.45]{img/ClassDiagram}
\caption{Metamodel specification of GMAS and GRL}
\label{fig:metamodel}
\end{figure*}
\floris{we have no formal definition of GMAS, so proposing a formal mapping will be difficult}
In the previous section we adapted the practical reasoning framework PRAS for use in goal modeling, obtaining nine Argument Schemes for Goal Modeling (GMAS). These argument schemes can be instantiated to create arguments for elements and relationships in a GRL model. We introduced critical questions that can be used to construct counter arguments or alternatives, and we provided an example in which we obtained a GRL model from a GMAS specification. However, we did not yet create a formal language for GMAS, nor did we map it to GRL. This is what we do in the current section. We provide a formal specification of GMAS in terms of a metamodel, and we link it with the GRL metamodel. Using this mapping, we detail how to translate an argumentation framework into a GRL model and visa versa. The metamodel is the starting point for the tool we developed as an extension of jUCMNav. We present this tool in the last subsection, together with the visual language we developed for our extension.

\subsection{GMAS Metamodel Specification}
\label{sect:implementation:metamodel}

The UML metamodel of GMAS is shown on the left hand side of figure~\ref{fig:metamodel}.

\subsection{Mappings between GMAS and GRL}
\label{sect:implementation:mappings}

\subsubsection{Information Mappings}

\subsubsection{Formal Mappings}

\subsection{Implementation into jUCMNAv}
\label{sect:implementation:implementation}

\section{The RationalGRL Methodology}
\label{sect:methodology}

The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}. There are two main activities, depicted in grey, namely "goal modeling" and "argumentation". These are two separate activities that are being done in parallel.

\begin{figure}[ht]
\centering
\includegraphics[scale=0.4]{img/methodology}
\caption{The RationalGRL Methodology}
\label{fig:rationalgrl-methodology}
\end{figure}

\section{Empirical Evaluation}
\label{sect:evaluation}

\section{Related Work}
\label{sect:relatedwork}

There are several contributions that relate argumentation-based techniques with goal modeling. The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}. This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models. One of the elements of GAM is the translation of formal argument models to goal models (similar to ours). In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM. One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.  \textsf{RationalGRL} also provides tool support for argumentation, i.e. Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e. jUCMNav~\cite{jUCMNav}. Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation. Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.

\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact. DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs. One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.

Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements. However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks. Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented. Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.

\section{Conclusions and Future Work}
\label{sect:conclusion}

In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders. We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.

There are many directions of future work. There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not. It would be very interesting to explore the effect of these semantics on goal models. Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments. Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized. We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it. In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}). That is, the translation from goal models to argument diagrams is still missing. The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram. Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool. By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.

\section*{Acknowledgments}
Marc van Zee and Diana Marosin are funded by the National Research Fund (FNR), Luxembourg, by the  RationalArchitecture project.

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

### 2017-01-20 versions floris/REIJ_article_FB2.tex
first commit

--- 2017-01-20 versions floris/REIJ_article_FB.tex
@@
 Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{URN}, which is part of an ITU-T standard, the User Requirements Notation (URN).
+The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
 
 A goal model is usually the result of a discussion process between a group of stakeholders and a requirements engineer.
-For goal models are constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SG1: This sentence doesn't make sense..
-Do you mean for the small systems?
-The next sentence seemed to follow that logic..
+For small-sized systems that their goal models are usually constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
+%SG1: I rephrased this since the sentence didn't make any sense logically and grammatically.
 However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
@@
 Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
+If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
 \item
@@
 
-\section{Background: Goal-oriented Requirements Language and Practical Argumentation}
+\section{Background: Argument Schemes and Goal-oriented Requirements Language}
 \label{sect:background}
@@
 URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we restrict ourselves to GRL.
+In this article, we focus on GRL for capturing and documenting argumentation.
 
@@
 
-A variety of links connect the elements in a GRL model.
+A variety of links connect the intentional elements in a GRL model.
 AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
+Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one intentional element on another one.
+A contribution link has a qualitative contribution type or a quantitative contribution values.
+Dependency links (\includegraphics[scale=1]{img/dependency}) model the relationships between actors or intentional elements.
 
@@
 Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
+%SGNew1: I don't think we need this paragraph at all.
+
+\floris{include an example here about the corporate data model etc, which is also used in section \ref{sect:background:pras}. This will make it easier for people to understand the PRAS reasoning}
 
 \subsubsection{Motivation for using GRL}
-\label{sect:background:grl:motivation}
+\label{sect:background:grl:motivation} %SGNew3: I will rewrite this later.
 
-GRL has the existing tools that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
+GRL has the existing tools %SGNew2:? what do you mean existing tools? has some means? potential? 
+that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
 (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools.
 GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying issue 4.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
+This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying criterion 4.
+Moreover, GRL has an open source tool called jUCMNav~\cite{jUCMNav}, which simplifies an implementation.
 Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
+According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
 Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
@@
 
-\subsection{Argument Scheme for Practical Reasoning (PRAS)} %SG10: This section is good but make the examples clearer. It has too much concepts but it makes me a bit lost.
+\subsection{Argument Scheme for Practical Reasoning (PRAS)} 
 \label{sect:background:pras}
@@
 Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g.
-\cite{raz1978,walton1990}) and Artificial Intelligence \cite{bratman1987,atkinson2007}.
+\cite{raz1978, walton1990}) and Artificial Intelligence \cite{bratman1987, atkinson2007}.
 One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}.
-The idea is that an instantiation of such a scheme gives a presumptive argument in favour of, for example, taking an action.
-This argument can then be tested by posing critical questions about, for example, if the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
+The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action.
+This argument can then be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
 
-A formal approach to argumentative about actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS), as follows. 
-\begin{itemize}[leftmargin=22pt]
-\item[PRAS]
-In the circumstances $R$,\\
-agent $Ag$ should perform action $A$,\\
-to achieve new circumstances $S$,\\
-which will realize goal $G$,\\
-which will promote some value $V$.
-\end{itemize}
+A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS).
+PRAS follows the following basic argument structure.
+\begin{equation}
+\label{eq:eq1}
+  \begin{aligned}
+ \qquad&\text{I have goal } G&\\
+&\text{Doing action }A \text{ will realize goal }G&\\
+&\text{Which will promote value }V&\\
+&\text{\emph{Therefore} I should do action }A.&
+  \end{aligned}
+\end{equation}
 
-This scheme distinguishes between the state of affairs brought about by an action, the goal (i.e., the desired features in that state of affairs), and the value (i.e., the motivation why those features are desirable).
-Take, for instance, the example from \cite{vanZee-etal:er2016}: the company (actor) does not have a standard corporate data model (current circumstances), and therefore decides to define such a model (action), leading to a new data model (new circumstances), which realizes the goal to adhere to a corporate data model, which promotes lower costs (value).
+So, for example, we can say that XXXX. The argument scheme can be slightly amended to capture subgoals (\emph{i.e.}, realizing goals $G_1, \ldots, G_n$ will allow me to realize goal $G_i$).
 
-In practical reasoning, conclusions which are at one point acceptable can later be rejected because of new information.
-For example, if we find out that `adhere to corporate data model' is not a realistic goal, we can argue that we should not define such a new data model.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized.
-An unfavourable\footnote{Unfavourably can be either 'no', or 'yes', depending on what the question is.} answer to a critical question then identifies a possible counterargument.
-One critical question is `Does the goal realize the value stated?'.
-For our example, we can answer this question favourably with a `yes', as one way to lower costs is to adhere to a company's corporate data model \cite{vanZee-etal:er2016}.
-Another critical question is `Can the desired goal be realized?'.
-If the answer to this is `no' - for example, there is evidence that `adhere to corporate data model' is not a realistic goal, then we would say the question has been answered unfavourably, and we have a counterargument: `evidence shows that there is not much attention for adherence to a corporate data model, so it can probably not be realized'.
-Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal or an alternative goal that promotes the same value.
+Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
+Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
+Some examples of critical questions are as follows.
+
+\begin{enumerate}
+\item Will the action bring about the desired goal?
+\item Does the goal promote the value stated?
+\item Are there alternative ways of realizing the same goal?
+\item Are there alternative ways of promoting the same value?
+\item Does doing the action have a side effect which demotes the value?
+\item Does doing the action have a side effect which demotes some other value?
+\item Does doing the action promote some other value?
+\item Does doing the action preclude some other action which would promote some other value?
+\item Is the action possible?
+\item Can the desired goal be realized?
+\item Is the value indeed a legitimate value?
+\end{enumerate}
+
+An unfavorable\footnote{Unfavorable can be either 'no', or 'yes', depending on what the question is.} answer to a critical question then identifies a possible counterargument.
+Take, for example, question 2.
+In the case of our example, we can answer this question favourably with a `yes', as a common way to adhere to a company's corporate data model is obviously to define such a data model.
+However, take critical question 11: if this is answered negatively -- for example, there is evidence that `adhere to corporate data model' was hardly used in project evaluations over the last four years \cite{vanZee-etal:er2016} -- then we have a counterargument: `evidence shows that there is not much attention for adherence to a corporate data model, so it can probably not be realized'.
+Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question X) or an alternative goal that promotes the same value (question Z).
 For example, in \cite{vanZee-etal:er2016} it is argued that the action `use data model of package applications' realizes the goal `use canonical data model', which also promotes lower costs.
-This gives us a counterargument that also follows PRAS.
+This essentially gives us a counterargument to the original argument -- to define a corporate data model -- that also follows PRAS.
 
 In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa).
-In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an emph{argumentation framework} $(Args, Attacks)$, as introduced by Dung~\cite{Dung1995}, where $Args$ is the set of arguments and $Attack$ is an attack relation between arguments.
+In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}.
+In this section, we will briefly discuss the intuitions behind these semantics.}.
 Figure \ref{fig:pras:example1} shows that an argumentation framework that includes the arguments from the above example can be rendered as a graph.
@@
 
-Given an argumentation framework, the acceptability of arguments can be determined according to the semantics of Dung~\cite{Dung1995}.
-The idea is that we are justified in accepting an argument if it can be successfully defended against attacking counterarguments.
-In other words, an argument $A \in Args$ is \emph{defeated} by an attacker $B \in Args$ -- where $(A, B) \in Attacks$ -- unless $B$ itself can be defeated.
+Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics.
+The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
 Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}.
-Argument A2 is undefeated or justified, because it has no attackers.
+Argument A2 is undefeated because it has no attackers.
 This makes A1 defeated, because one of its attackers, A2, is undefeated.
-A3 is then also justified, because its only attacker, A1, is defeated by A2.
-Thus, the set of justified (undefeated) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.
+A3 is then also undefeated, since its only attacker, A1, is defeated by A2.
+Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.
 
-attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two. However, A2 also attacks A1
-
- where $Args$ is the set of arguments and $Attack$ is an attack relation between arguments.
-Dung proposes that an argument $A_1\in Args$ is \emph{defeated} by an attacker $A_2\in Args$, unless $A_2$ itself can be defeated.
-Although this is an appropriate definition of reason about beliefs, it is not for reasoning about action since we are, to some extent, free to choose what we attempt to bring about.
-Therefore, we may choose to reject an attacker, even if it cannot be defeated, if we believe that the purpose motivating the attacked argument is more important.
-For instance, suppose an argument $A_2$ stating that a particular resource is expensive attacks an argument $A_1$ for using the resource.
-However, we may still choose to use it (i.e., accept argument $A_1$), if that would promote a value that we judge more importantly than expense.
-In order to accommodate this, \emph{valued-based argumentation frameworks} (VAF)~\cite{bench2003persuasion} are used.
-The idea is that we are given an \emph{argumentation framework}, a set of values $V$, a function $val$ mapping each argument to a value $v\in V$, and a set of \emph{audiences} $a$ to which the arguments are addressed.
-Each audience is then represented by a strict partial ordering $>_a$ on the values.
-Then, for a given audience $>_a$, $A_2$ defeats $A_1$ iff $A_2$ attacks $A_1$ and $val(A_2)\not>_a val(A_1)$.
-\floris{you do not give formal definitions here}
-
-It is then possible to determine which arguments in $Args$ are acceptable for an audience by determining the preferred extension.
-This is the maximal (under set inclusion) subset $S$ of $Args$ such that no two arguments in $S$ defeat each other, and all arguments in $S$ are acceptable with respect to $S$, i.e., for any argument $A_1$ in $S$, if $A_1$ is defeated by an argument $A_2$ that is not in $S$, then there exists an argument in $S$ that defeats $A_2$.
-The preferred extension represents the maximal consistent set of acceptable arguments with respect to the argumentation framework and a given audience or value ordering.
-
-\subsubsection{PRAS Medical Treatment Example} %SG11: Isn't it better to use an example more relevant to what we do?
-\label{sect:background:pras:example}
-
-A concrete example of using PRAS comes from the medical domain, where Atkinson and colleagues use practical reasoning to formalize a system for reasoning about the treatment of a patient~\cite{atkinson2006argumentation}.
-The system in this example consists of a group of experts in different areas, such as \emph{treatment knowledge} and \emph{patient knowledge}.\footnote{The original application uses a single expert using different knowledge bases, but we changed this to a group of experts, each with a different expertise, in order to make the comparison with goal modeling more apparent.} There are five critical questions relevant to the application:
-\begin{enumerate}
-\item[CQ1] Are there alternative ways of realizing the same effects?
-\item[CQ2] Are there alternative ways of realizing the same goal?
-\item[CQ3] Are the assumptions on which the argument is based true?
-\item[CQ4] Does performing the action have side effects which demotes some other value?
-\item[CQ5] Will the action have the effects described?
-\end{enumerate}
-
-The experts can construct an argumentation framework for a medical treatment by instantiating PAS and posing critical questions CQ1-CQ5.
-Consider for instance a patient whose health is threatened by blood clotting.
-The experts start with the null option \--- do nothing (EA0).
-Then, suppose that the treatment knowledge experts suggests blood clotting can be prevented by reducing platelet adhesion, which can, assuming aspirin is not contraindicated, be achieved by prescribing aspirin.
-This information is used to instantiate PAS:\\
-
-\noindent
-\textbf{EA1.} \emph{Assuming no contradictions, we should prescribe aspirin, which will reduce platelet adhesion, preventing blood clotting, and so is an efficacious course of action.}\\
-
-\noindent
-This argument is then subjected to the critical questions.
-CQ1 asks for alternative solutions to reduce platelet adhesion.
-The expert may reply by stating chlopidogrel also reduces platelet adhesion.
-Asking CQ2 forces the expert to seek further solutions for preventing blood clotting, which results in an alternative action, namely to administer streptokinase, which has the same goal of preventing blood clotting.
-These are formed into two arguments, EA2 and EA3:\\
-
-\noindent
-\textbf{EA2.} \emph{Assuming no contradictions, we should prescribe chlopidogrel, which will reduce platelet adhesion, preventing blood clotting, and so is an efficacious course of action.}\\\\
-\textbf{EA3}. \emph{Assuming no contradictions, we should prescribe streptokinase, which will increase blood clot dispersal, preventing blood clotting, and so is an efficacious course of action.}\\
-
-\noindent
-These three arguments all mutually attack one another, resulting in the argumentation framework shown in figure~\ref{fig:pras:example1}.
-
-Each of the arguments are in mutual conflict, so it is not clear which action to pursue.
-This depends on the preferences of the experts.
-For this, suppose the experts vote and agree the effect `reduce platelet adhesion' is a preferred means by which the goal can be realized.
-Moreover, they prefer aspirin to chlopidogrel because it is cheaper.
-
-These preferences are represented in the argumentation framework by removing attacks of unfavorable actions, so EA1 is no longer attacked, and EA3 no longer attacks EA2.
-In this case, EA1 forms the preferred extension, and its action is currently the best candidate.
-However, there are still remaining critical questions that can be asked.
-For instance, EA1 assumes that aspirin was not contraindicated, and CQ3 instructs us to test this assumption.
-Suppose that the patient knowledge expert notes the patient has a history of gastritis, and infers that aspirin is contraindicated because its acidity may result in gastric ulceration.
-The expert forms this into an argument motivated by the value of safety:\\
-
-\noindent
-\textbf{EA4.} \emph{Where there is a history of gastritis and no acid reducing therapy, we should not prescribe aspirin, which would cause excess acidity, which would risk ulceration, and so is unsafe.}\\
+In some cases, it is more difficult to determine whether or not an argument is defeated.
+Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two.
+It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{Amgoud}.
+Take, for example, A1 and A3.
+If we say that we prefer the goal ``XXX'' (A1) over the goal ``YYY'' (A3), we remove the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, left).
+This makes A1 the only undefeated argument, whereas A3 is now defeated.
+It is also possible to give explicit arguments for preferences \cite{Modgil2009}.
+These arguments are then essentially attacks on attacks.
+For example, say we prefer A1 over A3 because ``XXX''.
+This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.
 
@@
 \centering
-\includegraphics[scale=0.4]{img/pras-example2}
-\caption{Final argumentation framework showing all critiques from~\cite{atkinson2006argumentation}}
+\includegraphics[scale=0.6]{img/Fig2}
+\caption{Preferences between arguments}
 \label{fig:pras:example2}
@@
 
-\floris{please explain EA5, 6 and 7 as well}
-In a similar manner the experts construct three more arguments that attack other arguments in various ways, resulting in the argumentation framework shown in figure~\ref{fig:pras:example2}.
-In order to determine which arguments are accepted, the preferred extension is computed by first adding the arguments with no attackers, which are EA5, EA6, and EA7.
-EA7 defeats EA4, which means that EA1 can be included, since its only attacker is defeated.
-EA1 therefore defeats EA2 and EA4, and also excludes EA0.
-Thus, in the end, the winning argument is EA1, which prescribes aspirin.
-
-\subsubsection{Motivation for using PRAS}
+\subsubsection{Practical Reasoning and Goal Modeling}
 \label{sect:background:pras:motivation}
 
-In the introduction, we discussed six requirements for a practical reasoning and argumentation framework to be used in our research.
-The first requirement is: (1) it must have prior successful use in the field of software development.
-The example of the previous subsection is just one of many where PRAS has been successfully used by computer scientists to formalize arguments in a computer system.
-\floris{has it been implemented?
-and does that mean that it has prior successful use in softw dev?} The usefulness of the argument scheme approach for the analysis of practical reasoning situations has been shown in various other areas including e-democracy~\cite{Atkinson2006}, law~\cite{bench2005persuasion} and reasoning about the morally correct course of action~\cite{bench2006automating} \floris{Safety critical actions, explaining character motives in stories, planning}.
-The second requirement from the introduction is: (2) it should allow to formally reason about the consistency of arguments.
-The formal semantics in terms of value-based argumentation frameworks allows for the computation of accepted arguments.
-This semantics is extensively studied in the literature, and various different semantics exists, which lead to different flavors of acceptability of arguments.~\cite{} The third requirement from the introduction is: (3) it must be adaptable for use in goal modeling.
-In the next section we adapt PRAS in order to do so, but already at first glance the concepts that are used in PRAS and in goal modeling are very similar.
-In GRL, ``tasks'' can be performed by actors, while in argument scheme PAS these are called ``actions''.
-Both PAS and GRL have the concept of a ``goal''.
-Finally, the GRL notion of a ``softgoal'' seems semantically identical to PAS's ``value'', namely that it is a underlying motivation for which goals are desirable.
-We therefore decided that PRAS does indeed meet our first three requirements of a formal argumentation theory to be used in our research.
+Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{Amgoud, Modgil2009}.
+The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{Atkinson2006}, law~\cite{???}, planning \cite{MedellinXXX} and choosing between safety critical actions \cite{}.
+The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).
 
-%SG12: Read up to the section below.
+PRAS  and GRL have some obvious similarities in that PRAS defines elements as actions, goals, values and GRL includes tasks, goals, softgoals.
+In previous work, we have already discussed how to translate practical reasoning arguments to GRL diagrams~\cite{}.
+For example, ...
+A similar approach, albeit without the explicit focus on practical reasoning, was taken by Jureta et al.~\cite{}, in which it translates similar structured arguments to goal diagrams.
+One of the problems of these approaches is that ...
+
+The concept that appears in PRAS and not in GRL is \emph{circumstance}.
+In PRAS, an action $a$ is performed in current circumstances $R$, and by performing this action new circumstances $S$ are obtained.
+From this new circumstance $S$, another action may be performed again, and so on.
+In the AI literature, this type of reasoning usually referred to as ``AI planning''~\cite{weld1999recent}: from some initial state, find a sequence of actions such that a goal state is reached if these actions are executed.
+This planning approach is taken by a number of contributions using PRAS to create joint plans for a group of actors, often using a transition system as the underlying semantics~\cite{medellin2014}.
+However, in the early RE phase of an IS, one generally does not consider temporal planning.
+Decisions about the temporal order in which tasks should be executed are generally postponed until a later stage.
+In other words, goal modeling is not concerned with making plans, scenarios, or business processes for the information system.
+In URN (see section~\ref{sect:background:grl}) the distinction between goal modeling and business process modeling is clearly visible in the distinction between two separate modeling languages for the two activities.
+First, goal models are created in GRL and then, the use case maps (similar to business processes) are created in UCM.
+Therefore, when extending PRAS for GRL, we do not consider the notion of \emph{circumstances}.
+
+The concept that appears in GRL but not in PRAS is \emph{resource}.
+A resource may be used by part of the system, or an actor, in order to perform a task or to reach a goal.
+Therefore, PRAS has to be extended with the notion of a resource in order for it to be applicable to GRL.
+%SGNew: to be applicable?
+don't understand the sentence.
+
+Finally, there are two concepts which appear in both models under different names.
+Where PRAS uses the notions ``action'' and ``value'', GRL uses the concepts ``task'' and respectively ``softgoal''.
+However, these two concepts are semantically identical.
+
+With respect to relationships between concepts, the argument scheme PRAS from Section~\ref{sect:background:pras} contains the following relationships:\\
+
+\noindent
+R1. Actor $AG$ \emph{performs} action $A$\\
+R2. Action $A$ \emph{realizes} goal $G$\\
+R3. Goal $G$ \emph{promotes} value $S$\\
+R4. Actor $AG$ \emph{has} goal $G$\\
+R5. Actor $AG$ \emph{has} value $S$\\
+
+\noindent
+
+Relationships R4 and R5 are implicit in PRAS, but follow directly from its formulation (see~\cite{atkinson2007} for more details).
+The relationship ``promotes'' does not exist in GRL, but the relationship ``contributes to'' does, which is conceptually very similar.
+Therefore, we choose to use this relationship to formalize ``promotes'' relation.
+
+As mentioned in Section~\ref{sect:background:grl}, GRL is a very non-restrictive modeling language, allowing intentional elements and links to be combined freely.
+Thus, GRL allows any links between any type of intentional element.
+PRAS, in contrast, only considers relationship R1-R5 above.
+As a result, the critical questions of PRAS are only applicable to a subset of all possible relationships in GRL.
+This, however, does not mean it is not possible to construct arguments and formulate critical questions for other relationship.
+In section~\ref{sect:implementation:implementation}, where we discuss the implementation of our tool, we explain in more detail that a user can straightforwardly add additional critical questions for other relationships as well.
+
 \section{Argument Schemes for Goal Modeling (GMAS)}
@@
 
-PRAS is not directly applicable to goal modeling in general, and GRL in particular.
+%SGNew: This pragprah seems redundant.  
+%PRAS is not directly applicable to goal modeling in general, and GRL in particular.
 There are differences in the concepts used, and the relationships between these concepts.
 Some elements appear in PRAS that do not appear in GRL, some elements appear in GRL but not in PRAS, and some concepts appear in both formalisms but with different names.
-Moreover, GRL allows more relations between elements than the ones appearing in PAS.
-In this section, we first analyze these differences in more detail.
-After this, we introduce and motivate Argument Schemes for Goal Modeling (GMAS) and associated critical questions, which are an extension to PRAS.
+Moreover, GRL allows more relations between elements than the ones appearing in PRAS.
+
+%SGNew: What is the difference between PRAS and PAS?
+I think you should start this section with saying that based on what you defined in the last section, we developed a new argument schemes for goal modeling which is called GMAS.
+Then you can go ahead and talk that this is  the extension of PRAS, etc.
+I tried to reword.
+%In this section, we first analyze these differences in more detail. A
+
+In this section, we define our new argument schemes for GRL which is called Argument Schemes for Goal Modeling (GMAS) and its associated critical questions.
+GMAS is an extension of PRAS.
+PRAS  contains a single argument scheme PAS while GMAS  consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
+Each argument scheme is associated with critical questions that can be used to attack arguments or provide alternatives.
+This enables focused discussions about intentional elements and their relationships in a goal model.
+In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
+
+%After this, we introduce and motivate Argument Schemes for Goal Modeling (GMAS) and associated critical questions, which are an extension to PRAS.
 While PRAS contains a single argument scheme PAS, GMAS instead consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
@@
 In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
-
-\subsection{Differences between GRL and PRAS}
 
@@
 \hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & Yes\\
+AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & Yes\\ %SGNew: tasks may also require resources to be available.. how do you show that?
 \hline
@@
 AS5 & Task $T$ realizes goal $G$ & CQ5a & Will the task realize the desired goal? & Yes\\
-& & CQ5b & Are there alternative ways of realizing the same goal? \floris{How do you handle OR and XOR decompositions?} & No\\
+& & CQ5b & Are there alternative ways of realizing the same goal? \floris{How do you handle OR and XOR decompositions?} & No\\ %SGNew:What about decomposition of goals into goals?
 \hline
@@
 AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal? & Yes\\
-&& CQ7b & Does the goal contribute to some other softgoal? & No\\
+&& CQ7b & Does the goal contribute to some other softgoal? & No\\ %SGNew: What about contribution of goals to goals or softgoals to softgoals?
 \hline
@@
 &&CQ8c & Is the resource required in order to perform the task? & Yes\\
-&&CQ8d & Does using the resource make other resources unavailable? & Yes\\
+&&CQ8d & Does using the resource make other resources unavailable?
+& Yes\\ %SGNew: Tasks need resources..
+that is why tasks depends on resources..
+I can't understand the contribution link between them..
+resource doesn't have impact on the task..it should be dependency link.
 \hline
@@
 
-The concept that appears in PRAS and not in GRL is \emph{circumstance}.
-In PRAS, an action $a$ is performed in current circumstances $R$, and by performing this action new circumstances $S$ are obtained.
-From this new circumstance $S$, another action may be performed again, and so on.
-In the AI literature, this type of reasoning usually referred to as ``AI planning''~\cite{weld1999recent}: from some initial state, find a sequence of actions such that a goal state is reached if these actions are executed.
-This planning approach is taken by a number of contributions using PRAS to create joint plans for a group of actors, often using a transition system as the underlying semantics~\cite{medellin2014}.
-However, in the early RE phase of an IS one generally does not consider temporal planning.
-Decisions about the temporal order in which tasks should be executed are generally postponed until a later stage.
-In other words, goal modeling is not concerned with making plans, scenarios, or business processes for the information system.
-In URN (see section~\ref{sect:background:grl}) the distinction between goal modeling and business process modeling is clearly visible in the distinction between two separate modeling languages for the two activities.
-First, goal models are created in GRL, while after this use case maps (similar to business processes) are created in UCM.
-Therefore, when extending PRAS for GRL we do not consider the notion of \emph{circumstances}.
-
-The concept that appears in GRL but not in PRAS is \emph{resource}.
-A resource may be used by a part of the system, or an actor, in order to perform a task in order to reach a goal.
-Therefore, PRAS has to be extended with the notion of a resource in order for it to be applicable to GRL.
-
-Finally, there are two concepts appearing in both formalisms under different names.
-Where PRAS uses the notions ``action'' and ``value'', GRL uses the concepts ``task'' respectively ``softgoal''.
-However, these two concepts are semantically identical.
-
-With respect to relationships between concepts, the argument scheme PAS from Section~\ref{sect:background:pras} contains the following relationships:\\
-
-\noindent
-R1. Actor $AG$ \emph{performs} action $A$\\
-R2. Action $A$ \emph{realizes} goal $G$\\
-R3. Goal $G$ \emph{promotes} value $S$\\
-R4. Actor $AG$ \emph{has} goal $G$\\
-R5. Actor $AG$ \emph{has} value $S$\\
-
-\noindent
-Relationships R4 and R5 are implicit in PAS, but follow directly from its formulation (see~\cite{atkinson2007} for more details).
-The relationship ``promotes'' does not exist in GRL, but the relationship ``contributes to'' does, which is conceptually very similar.
-Therefore, we choose to use this relationship to formalize the ``promotes'' relation.
-
-As mentioned in Section~\ref{sect:background:grl}, GRL is a very non-restrictive modeling language, allowing intentional elements and links to be combined freely.
-Thus, GRL allows any links between any type of intentional element.
-PRAS, in contrast, only considers relationship R1-R5 above.
-As a result, the critical questions of PRAS are only applicable to a subset of all possible relationships in GRL.
-This, however, does not mean it is not possible to construct arguments and formulate critical questions for other relationship.
-In section~\ref{sect:implementation:implementation}, where we discuss the implementation of our tool, we explain in more detail that a user can straightforwardly add additional critical questions for other relationships as well.
-
 \subsection{The Argument Schemes and Critical Questions}
 
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
+PRAS consists of the single argument scheme PAS.
+However, our approach is to split this single scheme into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
 Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
+We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing the ones that question the elements which are not part of GRL (\emph{circumstances}), by adding questions for additional elements in GRL (\emph{resources}), and finally by renaming concepts and relationships (e.g., the \emph{promotes} relationship).
 
@@
 AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
+As we mentioned before%SGNew: where?,
+this list is not meant to be exhaustive, since GRL does not put restrictions on any of the relationships and  in theory any type of elements can be combined with any other type of elements, using any relationship.
+Here, we merely use the critical questions developed by Atkinson et al.
+%SGNew: merely = only?
+I thought you modified them?
 
@@
 In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-In general, answering a critical questions can have two effects:
+In general, answering a critical questions can have two effects: %SGNew: why unfavorably?
 \begin{enumerate}
@@
 
-\floris{you do not provide a formal definition of GMAS}
+\floris{you do not provide a formal definition of GMAS} %SGNew: Agreed.
+What is missing here is a big picture or as we can say a methodology.
+I think it is important to introduce the methodology earlier and show how everything fit together.
+Also, to show what are the elements of GMAS and how it can be connected to GRL.
+You said, you extended PRAS and that is GMAS but can you show the extension?
 
@@
 Therefore, if the underlying arguments change \--- for instance if the stakeholders change their preference from $A_2$ to $A_3$ \-- then the resulting goal model changes as well.
-\item In the current example each GRL element and relation is supported by an argument.
+\item In the current example, each GRL element and relation is supported by an argument.
 However, this is not required.
@@
 
-In this section we have fulfilled requirement three of the introduction: (3) it (the practical reasoning and argumentation theory) must be adaptable for use in goal modeling.
+In this section, we have fulfilled requirement three of the introduction: (3) it (the practical reasoning and argumentation theory) must be adaptable for use in goal modeling.
 However, we have not yet considered the fourth requirement: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
@@
 \floris{we have no formal definition of GMAS, so proposing a formal mapping will be difficult}
-In the previous section we adapted the practical reasoning framework PRAS for use in goal modeling, obtaining nine Argument Schemes for Goal Modeling (GMAS).
+In the previous section, we adapted and extended the practical reasoning framework PRAS for use in goal modeling and developed  nine Argument Schemes for Goal Modeling (GMAS).
 These argument schemes can be instantiated to create arguments for elements and relationships in a GRL model.
@@
 We provide a formal specification of GMAS in terms of a metamodel, and we link it with the GRL metamodel.
-Using this mapping, we detail how to translate an argumentation framework into a GRL model and visa versa.
+Using this mapping, we detail how to translate an argumentation framework into a GRL model and vice versa.
 The metamodel is the starting point for the tool we developed as an extension of jUCMNav.

### 2017-01-20 versions floris/section22pras.tex
first commit

--- 2017-01-20 versions floris/REIJ_article_FB2.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-\newcommand{\todo}[1]{{\color{red}\textbf{[TODO: } #1 \textbf{]}}}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-\newcommand{\floris}[1]{{\color{red}\textbf{Floris: }#1}}
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Diana Marosin\\
-\vspace{-0.35cm}
-Luxembourg Institute of Science and Technology (LIST)\\
-5 Avenue des Hauts-Fourneaux, 4362 Esch-sur-Alzette, Luxembourg\\ diana.marosin@list.lu
-\and 
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Diana Marosin\Mark{2},
-        Floris Bex\Mark{3},
-        and Sepideh Ghanavati\Mark{4}\\[1em]
- \normalsize
- \begin{tabular}{*{4}{>{\centering}p{.22\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Luxembourg Institute of Science and Technology & \Mark{3}Information and Computing Science & \Mark{4}Department of Computer Science \tabularnewline
-  University of Luxembourg & Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{diana.marosin@list.lu} & \url{florisbex@gmail.com} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-%SG1: Abstract is good for now but when we are done with the paper, we should add our findings wrt the case study in the abstract too
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-We evaluate the framework via a study among 20 subjects, showing that it is superior to traditional goal modeling with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is usually the result of a discussion process between a group of stakeholders and a requirements engineer.
-For small-sized systems that their goal models are usually constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SG1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\floris{These four reasons are very interesting. They could be improved by mentioning references in each: 2 and 3 are now without references, which makes it less strong.}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-%SG2: This is good but make them a bit more clear and concrete so they can be used as our base for the evaluation.
-Agreed with floris too.
-I am a bit unsure about the first criteria.
-Why this is needed?
-The main goal is to help better decision making and have more informed decision based on the arguments.
-Try to consider that in your criteria (comparing goal models with or without argumentation)
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for such an argumentation theory: (1) it must have prior successful use in the field of software development, (2) it should allow to formally reason about the consistency of discussions and the goal model, (3) it must be adaptable for use in goal models, (4) the adapted theory must produce repeatable results that can be formalized and implemented, (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools, and (6) the adapted theory must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-\floris{interesting!
-ad (1) why must the argumentation theory have prior successful use in software development?
-What is `successful use'?
-An academic paper?
-A system prototype?
-A large-scale evaluation?
-Implementation in a system that is used daily in companies?
-ad(4) what do you mean by `repeatable results?'}
-
-In this context, the main research question is: \emph{Can AI theories of practical reasoning and argumentation be adapted, formalized, and combined with an existing goal modeling methodology to rationalize goal model with underlying arguments, and to reason about their consistency?} The satisfaction of the above six requirements will result in a positive answer to the research question.
-%SG3: The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.
-
-\subsection{Contributions} %SG4:this part is a bit hard to read. I think a diagram helps and also make the contributions more aligned with the goals you have above.
-
-The contribution of this paper is to show that the Argument Scheme for Practical Reasoning (PRAS) fits all of the six requirements above.
-We briefly discuss how PRAS meets the first two requirements with an example from the medical domain, particularly in reasoning about medical treatment for a patient.
-\floris{medical domain is not software development, so how does this address req.
-1?} We, then, focus on issue three and four by modifying and implementing PRAS: We develop the Argument Schemes for Goal Modeling (GMAS), an extension to PRAS, which can be used to construct arguments for parts of a goal model.
-These arguments can be attacked with so-called ``critical questions'', which are natural language statements that question a certain aspect of the argument.
-We develop nine argument schemes and 18 critical questions.
-\floris{why does the preceding development of GMAS and the cq's address 3 in that it produces \emph{repeatable} results?} In order to reason about the consistency of the goal model with underlying arguments, we develop three \emph{consistency conditions} \floris{does this not address req.
-2?}.
-We integrate GMAS with the Goal-oriented Requirements Language (GRL) and we implement it in jUCMNav, the Eclipse-based open source tool for GRL modeling.
-To address issue five, we develop the RationalGRL methodology that combines goal modeling with argumentation.
-Finally, we answer issue six positively through an empirical study \floris{is the explanation of the study copied from Singh et al>?} in which 20 subjects apply RationalGRL or GRL to develop a goal modeling in a information system concerning national security.
-The study shows RationalGRL is superior to GRL with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects.
-As a side effect, RationalGRL requires more time than GRL, but is justified by improvements in quality.
-
-\subsection{Organization}
-
-The rest of this paper is organized as follows: In Section~\ref{sect:background}, we introduce a goal modeling language, i.e.
-Goal-oriented Requirements Language (GRL)~\cite{} and the practical reasoning framework PRAS.
-We also provide a motivation and examples for both formalisms.
-In Section~\ref{sect:gmas}, we first analyze the differences between PRAS and GRL in terms of concepts and relationships, after which we introduce the Argument Schemes for Goal Modeling (GMAS) and associated critical questions.
-We, then, illustrate GMAS through several examples.
-In Section~\ref{sect:consistency}, we specify a consistency condition on the goal model and underlying arguments, while in Section~\ref{sect:implementation}, we provide a metamodel for GMAS and we link it to the GRL metamodel.
-We also discuss implementation in jUCMNav~\cite{} and the visualization details of our language.%SG5: I don't think we can call it our language.
-It can be an extenstion.
-For creating a new modeling language you need ontologies etc and we are not doing that.
-In Section~\ref{sect:methodology}, we explain how GMAS can be used by practitioners by introducing the RationalGRL methodology.
-It specifies how goal modeling and argumentation can be used to construct a GRL model.
-In Section~\ref{sect:evaluation}, we describe the design of our empirical study and the findings from it.
-Section~\ref{sect:relatedwork} presents related work and Section~\ref{sect:conclusion} discusses the conclusions and the future work.
-
-\section{Background: Argument Schemes and Goal-oriented Requirements Language}
-\label{sect:background}
-
-In this section, we introduce the Goal-oriented Requirements Language (GRL) and the Argument Scheme for Practical Reasoning (PRAS).
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify  requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL for capturing and documenting argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself.
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals are often more related to NFRs, whereas goals are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available.
-
-A variety of links connect the intentional elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one intentional element on another one.
-A contribution link has a qualitative contribution type or a quantitative contribution values.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model the relationships between actors or intentional elements.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-%SGNew1: I don't think we need this paragraph at all.
-
-\floris{include an example here about the corporate data model etc, which is also used in section \ref{sect:background:pras}. This will make it easier for people to understand the PRAS reasoning}
-
-\subsubsection{Motivation for using GRL}
-\label{sect:background:grl:motivation} %SGNew3: I will rewrite this later.
-
-GRL has the existing tools %SGNew2:? what do you mean existing tools? has some means? potential? 
-that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
-(5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools.
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying criterion 4.
-Moreover, GRL has an open source tool called jUCMNav~\cite{jUCMNav}, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with the current research: We aim to add traceability links between intentional elements and their underlying arguments.
-%SG8: what do you mean by current research?
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)} 
-\label{sect:background:pras}
+\todo{Floris: I will fill in the XXX in the examples once we have set a running example in the GRL section}
 
@@
 \centering
-\includegraphics[scale=0.6]{img/Fig1}
+\includegraphics[width=\columnwidth]{img/Fig1}
 \caption{Argumentation framework}
@@
 Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two.
-It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{Amgoud}.
+It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}.
 Take, for example, A1 and A3.
@@
 This makes A1 the only undefeated argument, whereas A3 is now defeated.
-It is also possible to give explicit arguments for preferences \cite{Modgil2009}.
+It is also possible to give explicit arguments for preferences \cite{modgil2009}.
 These arguments are then essentially attacks on attacks.
@@
 \centering
-\includegraphics[scale=0.6]{img/Fig2}
+\includegraphics[width=\columnwidth]{img/Fig2}
 \caption{Preferences between arguments}
@@
 
-Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{Amgoud, Modgil2009}.
-The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{Atkinson2006}, law~\cite{???}, planning \cite{MedellinXXX} and choosing between safety critical actions \cite{}.
+Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{amgoud2002reasoning, modgil2009}.
+The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
 The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).
 
-PRAS  and GRL have some obvious similarities in that PRAS defines elements as actions, goals, values and GRL includes tasks, goals, softgoals.
-In previous work, we have already discussed how to translate practical reasoning arguments to GRL diagrams~\cite{}.
+PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) have some obvious similarities, and in previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016} we already presented arguments based on PRAS of the form ``G is a goal, Action A realizes G \emph{Therefore} perform action A'' (see Figure~\ref{fig:pras:example3}, left, where the arrow denotes an inference step).
+These arguments, which can be constructed using the online OVA tool\footnote{\url{http://ova.arg-tech.org/}}, can be combined with further arguments providing, for example, expert opinions about the practical reasoning elements (Figure~\ref{fig:pras:example3}, right).
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{img/Fig3}
+\caption{Structured argument based on PRAS}
+\label{fig:pras:example3}
+\end{figure}
+
+The arguments built in OVA can automatically be translated to GRL diagrams using further online tooling\footnote{\url{https://github.com/RationalArchitecture/RationalGRL}}.
+This translation takes the elements of the arguments and directly translates this to GRL elements - any elements of arguments that have nothing to do with goal modelling (e.g.
+the expert opinion premise in Figure~\ref{fig:pras:example3}, right) can be included in the GRL diagram as \emph{beliefs}.
+
 For example, ...
@@
 In section~\ref{sect:implementation:implementation}, where we discuss the implementation of our tool, we explain in more detail that a user can straightforwardly add additional critical questions for other relationships as well.
-
-\section{Argument Schemes for Goal Modeling (GMAS)}
-\label{sect:gmas}
-
-%SGNew: This pragprah seems redundant.  
-%PRAS is not directly applicable to goal modeling in general, and GRL in particular.
-There are differences in the concepts used, and the relationships between these concepts.
-Some elements appear in PRAS that do not appear in GRL, some elements appear in GRL but not in PRAS, and some concepts appear in both formalisms but with different names.
-Moreover, GRL allows more relations between elements than the ones appearing in PRAS.
-
-%SGNew: What is the difference between PRAS and PAS?
-I think you should start this section with saying that based on what you defined in the last section, we developed a new argument schemes for goal modeling which is called GMAS.
-Then you can go ahead and talk that this is  the extension of PRAS, etc.
-I tried to reword.
-%In this section, we first analyze these differences in more detail. A
-
-In this section, we define our new argument schemes for GRL which is called Argument Schemes for Goal Modeling (GMAS) and its associated critical questions.
-GMAS is an extension of PRAS.
-PRAS  contains a single argument scheme PAS while GMAS  consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
-Each argument scheme is associated with critical questions that can be used to attack arguments or provide alternatives.
-This enables focused discussions about intentional elements and their relationships in a goal model.
-In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
-
-%After this, we introduce and motivate Argument Schemes for Goal Modeling (GMAS) and associated critical questions, which are an extension to PRAS.
-While PRAS contains a single argument scheme PAS, GMAS instead consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
-With each argument scheme we associate critical questions that can be used to attack arguments or provide alternatives.
-This enables focused discussions about elements and relationships of a goal model.
-In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Attacks?}\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & Yes\\ %SGNew: tasks may also require resources to be available.. how do you show that?
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & Yes\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & Yes\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal? & Yes\\
-\hline
-\hline
-AS5 & Task $T$ realizes goal $G$ & CQ5a & Will the task realize the desired goal? & Yes\\
-& & CQ5b & Are there alternative ways of realizing the same goal? \floris{How do you handle OR and XOR decompositions?} & No\\ %SGNew:What about decomposition of goals into goals?
-\hline
-AS6 & Task $T$ contributes to softgoal $S$ & CQ6a & Are there alternative ways of contributing to the same softgoal? & No\\
-&& CQ6b & Does the task have a side effect which contribute negatively to the softgoal?&Yes\\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal? & No\\
-&& CQ6d & Does the task contribute to some other softgoal?&No\\
-&& CQ6e & Does the task preclude some other task which would contribute to some other softgoal?&Yes\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal? & Yes\\
-&& CQ7b & Does the goal contribute to some other softgoal? & No\\ %SGNew: What about contribution of goals to goals or softgoals to softgoals?
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8a & Is the resource required in order to perform the task? & Yes\\
-&&CQ8b & Can other resources be used to perform the same task? & No\\
-&&CQ8c & Is the resource required in order to perform the task? & Yes\\
-&&CQ8d & Does using the resource make other resources unavailable?
-& Yes\\ %SGNew: Tasks need resources..
-that is why tasks depends on resources..
-I can't understand the contribution link between them..
-resource doesn't have impact on the task..it should be dependency link.
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors? & No\\
-\hline
-\end{tabular}
-\caption{Argument schemes and critical questions for GRL elements (CQ1-CQ4) and relationships (CQ5-CQ16), as well as whether answering the critical question unfavorably results in an attack on the original argument (right column).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{The Argument Schemes and Critical Questions}
-
-PRAS consists of the single argument scheme PAS.
-However, our approach is to split this single scheme into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing the ones that question the elements which are not part of GRL (\emph{circumstances}), by adding questions for additional elements in GRL (\emph{resources}), and finally by renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-We developed a total of 9 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS1-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before%SGNew: where?,
-this list is not meant to be exhaustive, since GRL does not put restrictions on any of the relationships and  in theory any type of elements can be combined with any other type of elements, using any relationship.
-Here, we merely use the critical questions developed by Atkinson et al.
-%SGNew: merely = only?
-I thought you modified them?
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-In general, answering a critical questions can have two effects: %SGNew: why unfavorably?
-\begin{enumerate}
-\item
-Answering the critical question unfavorably results in a new argument attacking the original argument (\emph{attack}).
-\item
-Answering the critical question unfavorably results in a new argument not attacking the original argument, but representing an alternative (\emph{no attack}).
-\end{enumerate}
-The last column of table~\ref{table:argument-schemes} shows for each critical question whether it results in an attack on the original argument or not.
-
-\floris{you do not provide a formal definition of GMAS} %SGNew: Agreed.
-What is missing here is a big picture or as we can say a methodology.
-I think it is important to introduce the methodology earlier and show how everything fit together.
-Also, to show what are the elements of GMAS and how it can be connected to GRL.
-You said, you extended PRAS and that is GMAS but can you show the extension?
-
-\subsection{Example: Airport Assistant}
-
-We now provide a simple scenario of arguments created by instantiating AS1-AS9, and demonstrate how new arguments can be generated through critical questions, possibly attacking the original arguments.
-
-Consider a system for assisting users at the airport.
-The assistant should be able to guide users through the airport, inform them about delays, and provide other information such as the location of the check-in register for each airline company.
-When constructing a goal model for this system, one stakeholders argues that the user interface of the assistant should by easy to use, and that this can be realized by using a Graphical User Interface (GUI).
-We can formalize this as an instantiation of AS6:\\
-
-\noindent
-\textbf{$A_1$.} \emph{Task ``Use GUI'' contributes to softgoal ``easy to use'' (AS6)}.\\
-
-\noindent
-This argument can be subjected to critical questions.
-In this case, there are five critical questions that can be used on AS6, namely CQ6a-CQ6e.
-Critical question CQ6a asks for alternative solutions to contribute to softgoal ``easy to use''.
-One of the stakeholders may reply by stating a natural language interface is easy to use as well.\\
-
-\noindent
-\textbf{$A_2$.} \emph{Task ``Use NL interface'' contributes to softgoal ``easy to use'' (CQ6a/AS6)}.\\
-
-\noindent
-Note that $A_2$ does not create an attack on $A_1$, since both are valid tasks to contribute to the softgoal.
-Instead, $A_2$ actually generates new elements that will appear in the goal model.
-Now both $A_1$ and $A_2$ can be questioned.
-Suppose CQ6b is used by a stakeholder to argue that a side effect of using a natural language interface is that not all users may be able to understand it.
-When using a GUI, it may be easier to let the user change its language.\\
-
-\noindent
-\textbf{$A_3$.} \emph{Task ``Use NL interface'' contributes negatively to softgoal ``easy to use'' because choosing a language is difficult (CQ6b).}\\
-
-Suppose another stakeholder uses critical question CQ6c to question $A_1$: she notes that implementing a GUI is probably quite expensive, because they don't have a graphical designer, so they will have to outsource this.\\
-
-\noindent
-\textbf{$A_4$.} \emph{Task ``Use GUI'' contributes negatively to softgoal ``low costs''}.\\
-
-\noindent
-The resulting argumentation framework of these four arguments is shown in figure~\ref{fig:example:airport:arguments}.
-As we can see, argument $A_1$ and $A_4$ are both not attacked by any argument.
-However, $A_2$ and $A_3$ are in direct conflict.
-Recall from sectin~\ref{sect:background:pras} that the stakeholders now have to decide about their preferences.
-Suppose the stakeholders decide that argument $A_3$ is stronger, or more preferred, than $A_2$.
-As a result, the accepted arguments are $A_1, A_3$, and $A_4$.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}[scale=2]
-\node (a1) at (0,1) {$A_1$};
-\node (a2) at (1,1) {$A_2$};
-\node (a3) at (1,0) {$A_3$};
-\node (a4) at (0,0) {$A_4$};
-\draw [->] (a2) to [out=-60, in=60] (a3);
-\draw [->] (a3) to [out=120, in=-120] (a2);
-\end{tikzpicture}
-\caption{Argumentation framework for Airport Assistant example.}
-\label{fig:example:airport:arguments}
-\end{figure}
-
-It is now possible to translate this argumentation framework into a GRL model, which is shown in figure~\ref{fig:example:airport:grl}.
-Argument $A_1$ results in a contribution link from task ``use GUI'' to softgoal ``lower costs'', similarly for $A_4$.
-Since the conflict between $A_2$ and $A_3$ has been resolved in favor of $A_3$, there is a negative contribution link from task ``use NL interface'' to softgoal ``Easy to use''.
-
-We can make a number of observations about the obtained GRL model and the underlying arguments:
-\begin{itemize}
-\item The underlying arguments contain information that is not represented in the GRL model.
-For instance, the decision to reject $A_2$ (i.e., that a natural language interface is easy to use) is not visible in the GRL model, but it is in the underlying arguments.
-\item There is a direct relation between the accepted arguments and the elements that appear in the goal model.
-If an element or relation has an underlying accepted argument, it appears in the goal model, while rejected ones do not appear.
-Therefore, if the underlying arguments change \--- for instance if the stakeholders change their preference from $A_2$ to $A_3$ \-- then the resulting goal model changes as well.
-\item In the current example, each GRL element and relation is supported by an argument.
-However, this is not required.
-One could imagine that the stakeholders choose to add additional elements to the goal model without rationalizing them by arguments.
-We explain this in more detail in the RationalGRL methodology in section~\ref{sect:methodology}.
-\item Although it may be intuitively understandable how to translate argument $A_1-A_4$ and the argumentation framework of figure~\ref{fig:example:airport:arguments} into the GRL model of figure~\ref{fig:example:airport:grl}, we have not yet formally specified how to make the translation from GMAS to GRL, and visa versa.
-\end{itemize}
-
-In this section, we have fulfilled requirement three of the introduction: (3) it (the practical reasoning and argumentation theory) must be adaptable for use in goal modeling.
-However, we have not yet considered the fourth requirement: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
-This is what we do in the next section by providing formal mappings from GMAS to GRL, and discussing implementation details.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.35]{img/grl_example_airport}
-\caption{GRL model for Airport Assistant example.}
-\label{fig:example:airport:grl}
-\end{figure}
-
-\section{Mapping GMAS to GRL: Metamodel and Implementation}
-\label{sect:implementation}
-
-\begin{figure*}[h]
-\includegraphics[scale=0.45]{img/ClassDiagram}
-\caption{Metamodel specification of GMAS and GRL}
-\label{fig:metamodel}
-\end{figure*}
-\floris{we have no formal definition of GMAS, so proposing a formal mapping will be difficult}
-In the previous section, we adapted and extended the practical reasoning framework PRAS for use in goal modeling and developed  nine Argument Schemes for Goal Modeling (GMAS).
-These argument schemes can be instantiated to create arguments for elements and relationships in a GRL model.
-We introduced critical questions that can be used to construct counter arguments or alternatives, and we provided an example in which we obtained a GRL model from a GMAS specification.
-However, we did not yet create a formal language for GMAS, nor did we map it to GRL.
-This is what we do in the current section.
-We provide a formal specification of GMAS in terms of a metamodel, and we link it with the GRL metamodel.
-Using this mapping, we detail how to translate an argumentation framework into a GRL model and vice versa.
-The metamodel is the starting point for the tool we developed as an extension of jUCMNav.
-We present this tool in the last subsection, together with the visual language we developed for our extension.
-
-\subsection{GMAS Metamodel Specification}
-\label{sect:implementation:metamodel}
-
-The UML metamodel of GMAS is shown on the left hand side of figure~\ref{fig:metamodel}.
-
-\subsection{Mappings between GMAS and GRL}
-\label{sect:implementation:mappings}
-
-\subsubsection{Information Mappings}
-
-\subsubsection{Formal Mappings}
-
-\subsection{Implementation into jUCMNAv}
-\label{sect:implementation:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Empirical Evaluation}
-\label{sect:evaluation}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee and Diana Marosin are funded by the National Research Fund (FNR), Luxembourg, by the  RationalArchitecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}

### 2017-01-20 main.tex
x

--- 2017-01-20 REIJ_article.tex
@@
 
-\input{section22pras.tex}
+Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g.
+\cite{Raz1978-RAZPR,walton1990}) and Artificial Intelligence \cite{Bratman1987,atkinson2007}.
+One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}.
+The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action.
+This argument can then be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
+
+A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS).
+PRAS follows the following basic argument structure.
+\begin{equation}
+\label{eq:eq1}
+  \begin{aligned}
+ \qquad&\text{I have goal } G&\\
+&\text{Doing action }A \text{ will realize goal }G&\\
+&\text{Which will promote value }V&\\
+&\text{\emph{Therefore} I should do action }A.&
+  \end{aligned}
+\end{equation}
+
+So, for example, we can say that 
+\begin{itemize}
+\item[] \textbf{Find agreeable meeting date} is a goal,
+\item[] \textbf{Find agreeable date by talking to initiator} will realize the goal \textbf{(Find) agreeable meeting date},
+\item[] Which will promote the value \textbf{User friendliness}
+\item[] \textit{Therefore} 
+\item[] We should perform action \textbf{Find agreeable date by talking to initiator}
+\end{itemize}
+
+Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
+Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
+Some examples of critical questions are as follows.
+
+\begin{enumerate}
+\item Will the action bring about the desired goal?
+%\item Does the goal promote the value stated?
+\item Are there alternative ways of realizing the same goal?
+\item Are there alternative ways of promoting the same value?
+%\item Does doing the action have a side effect which demotes the value?
+\item Does doing the action have a side effect which demotes some other value?
+\item Does doing the action promote some other value?
+%\item Does doing the action preclude some other action which would promote some other value?
+\item Is the action possible?
+\item Can the desired goal be realized?
+\item Is the value indeed a legitimate value?
+\end{enumerate}
+
+These critical questions can point to new arguments that might counter the original argument.
+Take, for example, critical question 8.
+If, for example, there is evidence that schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textsf{user friendliness} is not a legitimate value'.
+Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).
+For example, we can argue that \textbf{Find agreeable date using scheduler} also realizes the goal \textbf{Find agreeable meeting date}, which gives us a counterargument to the original argument -- to find a date by talking to the initiator -- that also follows PRAS.
+
+In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa).
+In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}.
+In this section, we will briefly discuss the intuitions behind these semantics.}.
+Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for \textbf{Find agreeable date by talking to initiator} (A1), the argument for \textbf{Find agreeable date using scheduler} (A3), and the argument that \textbf{User friendliness} is not a legitimate value (A2).
+The two alternative PRAS instantiations are A1 and A3.
+These arguments mutually attack each other, as \textbf{Find agreeable date by talking to initiator}  and \textbf{Find agreeable date using scheduler} are considered to be mutually exclusive.
+Argument A2 attacks A1, as it questions the legitimacy of the value \textbf{User friendliness}.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.8]{img/Fig1}
+\caption{Argumentation framework}
+\label{fig:pras:example}
+\end{figure}
+
+Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics.
+The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
+Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}.
+Argument A2 is undefeated because it has no attackers.
+This makes A1 defeated, because one of its attackers, A2, is undefeated.
+A3 is then also undefeated, since its only attacker, A1, is defeated by A2.
+Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.
+
+In some cases, it is more difficult to determine whether or not an argument is defeated.
+Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two.
+It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}.
+Take, for example, A1 and A3.
+If we say that we prefer the action \textbf{Find agreeable date by talking to initiator} (A1) over the action \textbf{Find agreeable date using scheduler} (A3), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left).
+This makes A1 the only undefeated argument, whereas A3 is now defeated.
+It is also possible to give explicit arguments for preferences \cite{modgil2009}.
+These arguments are then essentially attacks on attacks.
+For example, say we prefer A1 over A3 because `the \textbf{Meeting participant} does not like working with a scheduler' (A4).
+This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.8]{img/Fig2}
+\caption{Preferences between arguments}
+\label{fig:pras:example2}
+\end{figure}
+
+\subsubsection{Practical Argumentation and Goal Modeling}
+\label{sect:background:pras:motivation}
+
+Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{amgoud2002reasoning,modgil2009}.
+The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
+The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).
+
+PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) have some obvious similarities, and in previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016,vanZee-etal:comma2016} we presented arguments based on PRAS of the form ``G is a goal, Action A realizes G \emph{Therefore} perform action A'' (see Figure~\ref{fig:pras:example3}, left, where the arrow denotes an inference step).
+Such arguments, which can be constructed using the online OVA tool\footnote{\url{http://ova.arg-tech.org/}}, can be combined with further arguments providing, for example, expert opinions about the practical reasoning elements (Figure~\ref{fig:pras:example3}, left).
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{img/Fig3}
+\caption{Structured argument based on PRAS (left) and the corresponding GRL diagram (right)}
+\label{fig:pras:example3}
+\end{figure}
+
+Arguments built in OVA can automatically be translated to GRL diagrams using further online tooling\footnote{\url{https://github.com/RationalArchitecture/RationalGRL}} \cite{vanZee-etal:comma2016}.
+This translation takes the elements of the arguments and directly translates this to GRL elements.
+Tasks/actions, goals and softgoals/values can be directly translated from the arguments to the GRL diagrams, as can realization statements.
+Elements of arguments that have nothing to do with goal modelling (e.g.
+the expert opinion premise in Figure~\ref{fig:pras:example3}) can be included in the GRL diagram as \emph{beliefs}.
+Finally, attacks between arguments are captured as negative contribution links.
+Take, for example, \todo{example}.
+A very similar approach to argumentation and goal models had already been taken by Jureta et al.~\cite{Jureta:RE2008} who proposed a formal argumentation model that can be used to justify goal-modelling decisions.
+This argumentation model, while not explicitly based on a practical reasoning scheme, is essentially the same as our previous argumentation model based on PRAS: a formal model of structured arguments with goals as premises and tasks as conclusions, and a translation function of these arguments to goal diagrams.
+
+The core problem of the use of argumentation in the above-mentioned work \cite{Jureta:RE2008,vanzee-etal:renext2015,vanZee-etal:er2016}, is that it focuses mainly on structured arguments, which are then directly translated to goal models as in Figure~\ref{fig:pras:example3}.
+The question is what the added value of this exercise is.
+Firstly, the arguments and goal models contain many of the same elements.
+The idea is that arguments show the rationales behind goal models and the alternative or possibly conflicting views that were lost in the goal-modeling process.
+But GRL already contains a \emph{belief} element that allow one to provide beliefs underlying a goal model, an \emph{XOR decomposition} link that allows for modeling of alternative ways to realize goals, and \emph{negative contribution} links for capturing conflicting goals.
+
+The point is that argumentation models are not needed to precisely render goals, tasks, beliefs and the relations between them, as GRL is already perfectly suited for this.
+Rather, argumentation should guide the goal modelling \emph{process}.
+Instead of static, formal argument trees, we need a dynamic argumentation process in which goals and tasks are critically analysed and which captures the dialectical nature of justification.
+Jureta et al.~\cite{Jureta:RE2008} take a step towards such a process by defining a high-level decision process that organizes argumentation and clarification techniques in relation to goal modelling.
+However, the argumentation techniques in this process are focused on building precisely the type of arguments that we argue are redundant next to GRL diagrams, namely structured arguments like in Figure~\ref{fig:pras:example3} which can be translated more or less directly to goal models.
+The decision process itself is never formally defined, and how to go from a discussion on goals and tasks to a goal model remains largely implicit.
+
+What is needed is a formal framework that captures the discussions about goal models, and allows for (semi-)automated construction of goal models given this discussion.
+In section XX, we argue that a combination of Atkinson et al.'s~\cite{atkinson2007} critical questions and the dialectical semantics as formally captured by Dung's work \cite{Dung1995} are ideally suited to provide such a formal, dynamic framework.
 

### 2017-01-21 main.tex
m

@@
 
-Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{amgoud2002reasoning,modgil2009}.
+Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung1995} and its various extensions \cite{amgoud2002reasoning,modgil2009}.
 The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.

### 2017-01-21 abstract.tex
separated files

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
 \begin{abstract}
@@
 \end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-01-21 background.tex
separated files

--- 2017-01-21 main.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
 \section{Background: Goal-oriented Requirements Language and Argument Schemes}
@@
 In section XX, we argue that a combination of Atkinson et al.'s~\cite{atkinson2007} critical questions and the dialectical semantics as formally captured by Dung's work \cite{Dung1995} are ideally suited to provide such a formal, dynamic framework.
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-01-21 conclusion.tex
separated files

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
 \section{Conclusions and Future Work}
@@
 By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-01-21 formalframework.tex
separated files

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
 \section{The Formal Framework}
@@
 In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-01-21 gmas.tex
separated files

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
 \section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
@@
 \end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-01-21 introduction.tex
separated files

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
 \section{Introduction}
 \label{sect:introduction}
+
 Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
@@
 In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-01-21 main.tex
m2
separated files

--- 2017-01-20 REIJ_article.tex
@@
 \documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
+
+\setcounter{tocdepth}{3}
+\usepackage{algorithm}
+\usepackage{algpseudocode}
+\usepackage{amssymb, amsmath}
+\usepackage{array,url}
+\usepackage{authblk}
 \usepackage{courier}
@@
 \usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
+\usepackage{helvet}
 \usepackage{multicol}
 \usepackage{multirow}
+\usepackage{sectsty}
+\usepackage{tikz}\usetikzlibrary{arrows}
+\usepackage{times}
 \usepackage{url}
-%\usepackage[margin=2pt]{subfig}
 \usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
+
 % change the appearance of the tikz arrow for the argumentation networks
@@
 \setlength{\columnsep}{0.3in}
+\newcommand{\Mark}[1]{\textsuperscript{#1}}
 
@@
 }
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
 
@@
 \subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
 
@@
 
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
+\input{introduction}
+\input{background}
+\input{preview}
+\input{gmas}
+\input{formalframework}
+\input{tool}
+\input{methodology}
+\input{relatedwork}
+\input{conclusion}
 
@@
 
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
+\input{transcripts}
 

### 2017-01-21 methodology.tex
separated files

--- 2017-01-20 versions floris/REIJ_article_FB2.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-\newcommand{\todo}[1]{{\color{red}\textbf{[TODO: } #1 \textbf{]}}}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-\newcommand{\floris}[1]{{\color{red}\textbf{Floris: }#1}}
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Diana Marosin\\
-\vspace{-0.35cm}
-Luxembourg Institute of Science and Technology (LIST)\\
-5 Avenue des Hauts-Fourneaux, 4362 Esch-sur-Alzette, Luxembourg\\ diana.marosin@list.lu
-\and 
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Diana Marosin\Mark{2},
-        Floris Bex\Mark{3},
-        and Sepideh Ghanavati\Mark{4}\\[1em]
- \normalsize
- \begin{tabular}{*{4}{>{\centering}p{.22\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Luxembourg Institute of Science and Technology & \Mark{3}Information and Computing Science & \Mark{4}Department of Computer Science \tabularnewline
-  University of Luxembourg & Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{diana.marosin@list.lu} & \url{florisbex@gmail.com} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-%SG1: Abstract is good for now but when we are done with the paper, we should add our findings wrt the case study in the abstract too
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-We evaluate the framework via a study among 20 subjects, showing that it is superior to traditional goal modeling with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is usually the result of a discussion process between a group of stakeholders and a requirements engineer.
-For small-sized systems that their goal models are usually constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SG1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\floris{These four reasons are very interesting. They could be improved by mentioning references in each: 2 and 3 are now without references, which makes it less strong.}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-%SG2: This is good but make them a bit more clear and concrete so they can be used as our base for the evaluation.
-Agreed with floris too.
-I am a bit unsure about the first criteria.
-Why this is needed?
-The main goal is to help better decision making and have more informed decision based on the arguments.
-Try to consider that in your criteria (comparing goal models with or without argumentation)
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for such an argumentation theory: (1) it must have prior successful use in the field of software development, (2) it should allow to formally reason about the consistency of discussions and the goal model, (3) it must be adaptable for use in goal models, (4) the adapted theory must produce repeatable results that can be formalized and implemented, (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools, and (6) the adapted theory must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-\floris{interesting!
-ad (1) why must the argumentation theory have prior successful use in software development?
-What is `successful use'?
-An academic paper?
-A system prototype?
-A large-scale evaluation?
-Implementation in a system that is used daily in companies?
-ad(4) what do you mean by `repeatable results?'}
-
-In this context, the main research question is: \emph{Can AI theories of practical reasoning and argumentation be adapted, formalized, and combined with an existing goal modeling methodology to rationalize goal model with underlying arguments, and to reason about their consistency?} The satisfaction of the above six requirements will result in a positive answer to the research question.
-%SG3: The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.
-
-\subsection{Contributions} %SG4:this part is a bit hard to read. I think a diagram helps and also make the contributions more aligned with the goals you have above.
-
-The contribution of this paper is to show that the Argument Scheme for Practical Reasoning (PRAS) fits all of the six requirements above.
-We briefly discuss how PRAS meets the first two requirements with an example from the medical domain, particularly in reasoning about medical treatment for a patient.
-\floris{medical domain is not software development, so how does this address req.
-1?} We, then, focus on issue three and four by modifying and implementing PRAS: We develop the Argument Schemes for Goal Modeling (GMAS), an extension to PRAS, which can be used to construct arguments for parts of a goal model.
-These arguments can be attacked with so-called ``critical questions'', which are natural language statements that question a certain aspect of the argument.
-We develop nine argument schemes and 18 critical questions.
-\floris{why does the preceding development of GMAS and the cq's address 3 in that it produces \emph{repeatable} results?} In order to reason about the consistency of the goal model with underlying arguments, we develop three \emph{consistency conditions} \floris{does this not address req.
-2?}.
-We integrate GMAS with the Goal-oriented Requirements Language (GRL) and we implement it in jUCMNav, the Eclipse-based open source tool for GRL modeling.
-To address issue five, we develop the RationalGRL methodology that combines goal modeling with argumentation.
-Finally, we answer issue six positively through an empirical study \floris{is the explanation of the study copied from Singh et al>?} in which 20 subjects apply RationalGRL or GRL to develop a goal modeling in a information system concerning national security.
-The study shows RationalGRL is superior to GRL with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects.
-As a side effect, RationalGRL requires more time than GRL, but is justified by improvements in quality.
-
-\subsection{Organization}
-
-The rest of this paper is organized as follows: In Section~\ref{sect:background}, we introduce a goal modeling language, i.e.
-Goal-oriented Requirements Language (GRL)~\cite{} and the practical reasoning framework PRAS.
-We also provide a motivation and examples for both formalisms.
-In Section~\ref{sect:gmas}, we first analyze the differences between PRAS and GRL in terms of concepts and relationships, after which we introduce the Argument Schemes for Goal Modeling (GMAS) and associated critical questions.
-We, then, illustrate GMAS through several examples.
-In Section~\ref{sect:consistency}, we specify a consistency condition on the goal model and underlying arguments, while in Section~\ref{sect:implementation}, we provide a metamodel for GMAS and we link it to the GRL metamodel.
-We also discuss implementation in jUCMNav~\cite{} and the visualization details of our language.%SG5: I don't think we can call it our language.
-It can be an extenstion.
-For creating a new modeling language you need ontologies etc and we are not doing that.
-In Section~\ref{sect:methodology}, we explain how GMAS can be used by practitioners by introducing the RationalGRL methodology.
-It specifies how goal modeling and argumentation can be used to construct a GRL model.
-In Section~\ref{sect:evaluation}, we describe the design of our empirical study and the findings from it.
-Section~\ref{sect:relatedwork} presents related work and Section~\ref{sect:conclusion} discusses the conclusions and the future work.
-
-\section{Background: Argument Schemes and Goal-oriented Requirements Language}
-\label{sect:background}
-
-In this section, we introduce the Goal-oriented Requirements Language (GRL) and the Argument Scheme for Practical Reasoning (PRAS).
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify  requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL for capturing and documenting argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself.
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals are often more related to NFRs, whereas goals are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available.
-
-A variety of links connect the intentional elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one intentional element on another one.
-A contribution link has a qualitative contribution type or a quantitative contribution values.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model the relationships between actors or intentional elements.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-%SGNew1: I don't think we need this paragraph at all.
-
-\floris{include an example here about the corporate data model etc, which is also used in section \ref{sect:background:pras}. This will make it easier for people to understand the PRAS reasoning}
-
-\subsubsection{Motivation for using GRL}
-\label{sect:background:grl:motivation} %SGNew3: I will rewrite this later.
-
-GRL has the existing tools %SGNew2:? what do you mean existing tools? has some means? potential? 
-that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
-(5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools.
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying criterion 4.
-Moreover, GRL has an open source tool called jUCMNav~\cite{jUCMNav}, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with the current research: We aim to add traceability links between intentional elements and their underlying arguments.
-%SG8: what do you mean by current research?
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)} 
-\label{sect:background:pras}
-
-Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g.
-\cite{raz1978, walton1990}) and Artificial Intelligence \cite{bratman1987, atkinson2007}.
-One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}.
-The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action.
-This argument can then be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
-
-A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS).
-PRAS follows the following basic argument structure.
-\begin{equation}
-\label{eq:eq1}
-  \begin{aligned}
- \qquad&\text{I have goal } G&\\
-&\text{Doing action }A \text{ will realize goal }G&\\
-&\text{Which will promote value }V&\\
-&\text{\emph{Therefore} I should do action }A.&
-  \end{aligned}
-\end{equation}
-
-So, for example, we can say that XXXX. The argument scheme can be slightly amended to capture subgoals (\emph{i.e.}, realizing goals $G_1, \ldots, G_n$ will allow me to realize goal $G_i$).
-
-Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
-Some examples of critical questions are as follows.
-
-\begin{enumerate}
-\item Will the action bring about the desired goal?
-\item Does the goal promote the value stated?
-\item Are there alternative ways of realizing the same goal?
-\item Are there alternative ways of promoting the same value?
-\item Does doing the action have a side effect which demotes the value?
-\item Does doing the action have a side effect which demotes some other value?
-\item Does doing the action promote some other value?
-\item Does doing the action preclude some other action which would promote some other value?
-\item Is the action possible?
-\item Can the desired goal be realized?
-\item Is the value indeed a legitimate value?
-\end{enumerate}
-
-An unfavorable\footnote{Unfavorable can be either 'no', or 'yes', depending on what the question is.} answer to a critical question then identifies a possible counterargument.
-Take, for example, question 2.
-In the case of our example, we can answer this question favourably with a `yes', as a common way to adhere to a company's corporate data model is obviously to define such a data model.
-However, take critical question 11: if this is answered negatively -- for example, there is evidence that `adhere to corporate data model' was hardly used in project evaluations over the last four years \cite{vanZee-etal:er2016} -- then we have a counterargument: `evidence shows that there is not much attention for adherence to a corporate data model, so it can probably not be realized'.
-Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question X) or an alternative goal that promotes the same value (question Z).
-For example, in \cite{vanZee-etal:er2016} it is argued that the action `use data model of package applications' realizes the goal `use canonical data model', which also promotes lower costs.
-This essentially gives us a counterargument to the original argument -- to define a corporate data model -- that also follows PRAS.
-
-In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa).
-In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}.
-In this section, we will briefly discuss the intuitions behind these semantics.}.
-Figure \ref{fig:pras:example1} shows that an argumentation framework that includes the arguments from the above example can be rendered as a graph.
-The two alternative PRAS instantiations are A1 and A3.
-These arguments mutually attack each other, as `adhere to corporate data model' (A1) is an alternative to `adhere to canonical data model' (A3) and vice versa.
-Argument A2 attacks A1, as it questions the validity of the goal `adhere to corporate data model'.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.6]{img/Fig1}
-\caption{Argumentation framework}
-\label{fig:pras:example}
-\end{figure}
-
-Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics.
-The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
-Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}.
-Argument A2 is undefeated because it has no attackers.
-This makes A1 defeated, because one of its attackers, A2, is undefeated.
-A3 is then also undefeated, since its only attacker, A1, is defeated by A2.
-Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.
-
-In some cases, it is more difficult to determine whether or not an argument is defeated.
-Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two.
-It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{Amgoud}.
-Take, for example, A1 and A3.
-If we say that we prefer the goal ``XXX'' (A1) over the goal ``YYY'' (A3), we remove the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, left).
-This makes A1 the only undefeated argument, whereas A3 is now defeated.
-It is also possible to give explicit arguments for preferences \cite{Modgil2009}.
-These arguments are then essentially attacks on attacks.
-For example, say we prefer A1 over A3 because ``XXX''.
-This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.6]{img/Fig2}
-\caption{Preferences between arguments}
-\label{fig:pras:example2}
-\end{figure}
-
-\subsubsection{Practical Reasoning and Goal Modeling}
-\label{sect:background:pras:motivation}
-
-Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{Amgoud, Modgil2009}.
-The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{Atkinson2006}, law~\cite{???}, planning \cite{MedellinXXX} and choosing between safety critical actions \cite{}.
-The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).
-
-PRAS  and GRL have some obvious similarities in that PRAS defines elements as actions, goals, values and GRL includes tasks, goals, softgoals.
-In previous work, we have already discussed how to translate practical reasoning arguments to GRL diagrams~\cite{}.
-For example, ...
-A similar approach, albeit without the explicit focus on practical reasoning, was taken by Jureta et al.~\cite{}, in which it translates similar structured arguments to goal diagrams.
-One of the problems of these approaches is that ...
-
-The concept that appears in PRAS and not in GRL is \emph{circumstance}.
-In PRAS, an action $a$ is performed in current circumstances $R$, and by performing this action new circumstances $S$ are obtained.
-From this new circumstance $S$, another action may be performed again, and so on.
-In the AI literature, this type of reasoning usually referred to as ``AI planning''~\cite{weld1999recent}: from some initial state, find a sequence of actions such that a goal state is reached if these actions are executed.
-This planning approach is taken by a number of contributions using PRAS to create joint plans for a group of actors, often using a transition system as the underlying semantics~\cite{medellin2014}.
-However, in the early RE phase of an IS, one generally does not consider temporal planning.
-Decisions about the temporal order in which tasks should be executed are generally postponed until a later stage.
-In other words, goal modeling is not concerned with making plans, scenarios, or business processes for the information system.
-In URN (see section~\ref{sect:background:grl}) the distinction between goal modeling and business process modeling is clearly visible in the distinction between two separate modeling languages for the two activities.
-First, goal models are created in GRL and then, the use case maps (similar to business processes) are created in UCM.
-Therefore, when extending PRAS for GRL, we do not consider the notion of \emph{circumstances}.
-
-The concept that appears in GRL but not in PRAS is \emph{resource}.
-A resource may be used by part of the system, or an actor, in order to perform a task or to reach a goal.
-Therefore, PRAS has to be extended with the notion of a resource in order for it to be applicable to GRL.
-%SGNew: to be applicable?
-don't understand the sentence.
-
-Finally, there are two concepts which appear in both models under different names.
-Where PRAS uses the notions ``action'' and ``value'', GRL uses the concepts ``task'' and respectively ``softgoal''.
-However, these two concepts are semantically identical.
-
-With respect to relationships between concepts, the argument scheme PRAS from Section~\ref{sect:background:pras} contains the following relationships:\\
-
-\noindent
-R1. Actor $AG$ \emph{performs} action $A$\\
-R2. Action $A$ \emph{realizes} goal $G$\\
-R3. Goal $G$ \emph{promotes} value $S$\\
-R4. Actor $AG$ \emph{has} goal $G$\\
-R5. Actor $AG$ \emph{has} value $S$\\
-
-\noindent
-
-Relationships R4 and R5 are implicit in PRAS, but follow directly from its formulation (see~\cite{atkinson2007} for more details).
-The relationship ``promotes'' does not exist in GRL, but the relationship ``contributes to'' does, which is conceptually very similar.
-Therefore, we choose to use this relationship to formalize ``promotes'' relation.
-
-As mentioned in Section~\ref{sect:background:grl}, GRL is a very non-restrictive modeling language, allowing intentional elements and links to be combined freely.
-Thus, GRL allows any links between any type of intentional element.
-PRAS, in contrast, only considers relationship R1-R5 above.
-As a result, the critical questions of PRAS are only applicable to a subset of all possible relationships in GRL.
-This, however, does not mean it is not possible to construct arguments and formulate critical questions for other relationship.
-In section~\ref{sect:implementation:implementation}, where we discuss the implementation of our tool, we explain in more detail that a user can straightforwardly add additional critical questions for other relationships as well.
-
-\section{Argument Schemes for Goal Modeling (GMAS)}
-\label{sect:gmas}
-
-%SGNew: This pragprah seems redundant.  
-%PRAS is not directly applicable to goal modeling in general, and GRL in particular.
-There are differences in the concepts used, and the relationships between these concepts.
-Some elements appear in PRAS that do not appear in GRL, some elements appear in GRL but not in PRAS, and some concepts appear in both formalisms but with different names.
-Moreover, GRL allows more relations between elements than the ones appearing in PRAS.
-
-%SGNew: What is the difference between PRAS and PAS?
-I think you should start this section with saying that based on what you defined in the last section, we developed a new argument schemes for goal modeling which is called GMAS.
-Then you can go ahead and talk that this is  the extension of PRAS, etc.
-I tried to reword.
-%In this section, we first analyze these differences in more detail. A
-
-In this section, we define our new argument schemes for GRL which is called Argument Schemes for Goal Modeling (GMAS) and its associated critical questions.
-GMAS is an extension of PRAS.
-PRAS  contains a single argument scheme PAS while GMAS  consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
-Each argument scheme is associated with critical questions that can be used to attack arguments or provide alternatives.
-This enables focused discussions about intentional elements and their relationships in a goal model.
-In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
-
-%After this, we introduce and motivate Argument Schemes for Goal Modeling (GMAS) and associated critical questions, which are an extension to PRAS.
-While PRAS contains a single argument scheme PAS, GMAS instead consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
-With each argument scheme we associate critical questions that can be used to attack arguments or provide alternatives.
-This enables focused discussions about elements and relationships of a goal model.
-In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Attacks?}\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & Yes\\ %SGNew: tasks may also require resources to be available.. how do you show that?
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & Yes\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & Yes\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal? & Yes\\
-\hline
-\hline
-AS5 & Task $T$ realizes goal $G$ & CQ5a & Will the task realize the desired goal? & Yes\\
-& & CQ5b & Are there alternative ways of realizing the same goal? \floris{How do you handle OR and XOR decompositions?} & No\\ %SGNew:What about decomposition of goals into goals?
-\hline
-AS6 & Task $T$ contributes to softgoal $S$ & CQ6a & Are there alternative ways of contributing to the same softgoal? & No\\
-&& CQ6b & Does the task have a side effect which contribute negatively to the softgoal?&Yes\\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal? & No\\
-&& CQ6d & Does the task contribute to some other softgoal?&No\\
-&& CQ6e & Does the task preclude some other task which would contribute to some other softgoal?&Yes\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal? & Yes\\
-&& CQ7b & Does the goal contribute to some other softgoal? & No\\ %SGNew: What about contribution of goals to goals or softgoals to softgoals?
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8a & Is the resource required in order to perform the task? & Yes\\
-&&CQ8b & Can other resources be used to perform the same task? & No\\
-&&CQ8c & Is the resource required in order to perform the task? & Yes\\
-&&CQ8d & Does using the resource make other resources unavailable?
-& Yes\\ %SGNew: Tasks need resources..
-that is why tasks depends on resources..
-I can't understand the contribution link between them..
-resource doesn't have impact on the task..it should be dependency link.
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors? & No\\
-\hline
-\end{tabular}
-\caption{Argument schemes and critical questions for GRL elements (CQ1-CQ4) and relationships (CQ5-CQ16), as well as whether answering the critical question unfavorably results in an attack on the original argument (right column).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{The Argument Schemes and Critical Questions}
-
-PRAS consists of the single argument scheme PAS.
-However, our approach is to split this single scheme into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing the ones that question the elements which are not part of GRL (\emph{circumstances}), by adding questions for additional elements in GRL (\emph{resources}), and finally by renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-We developed a total of 9 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS1-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before%SGNew: where?,
-this list is not meant to be exhaustive, since GRL does not put restrictions on any of the relationships and  in theory any type of elements can be combined with any other type of elements, using any relationship.
-Here, we merely use the critical questions developed by Atkinson et al.
-%SGNew: merely = only?
-I thought you modified them?
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-In general, answering a critical questions can have two effects: %SGNew: why unfavorably?
-\begin{enumerate}
-\item
-Answering the critical question unfavorably results in a new argument attacking the original argument (\emph{attack}).
-\item
-Answering the critical question unfavorably results in a new argument not attacking the original argument, but representing an alternative (\emph{no attack}).
-\end{enumerate}
-The last column of table~\ref{table:argument-schemes} shows for each critical question whether it results in an attack on the original argument or not.
-
-\floris{you do not provide a formal definition of GMAS} %SGNew: Agreed.
-What is missing here is a big picture or as we can say a methodology.
-I think it is important to introduce the methodology earlier and show how everything fit together.
-Also, to show what are the elements of GMAS and how it can be connected to GRL.
-You said, you extended PRAS and that is GMAS but can you show the extension?
-
-\subsection{Example: Airport Assistant}
-
-We now provide a simple scenario of arguments created by instantiating AS1-AS9, and demonstrate how new arguments can be generated through critical questions, possibly attacking the original arguments.
-
-Consider a system for assisting users at the airport.
-The assistant should be able to guide users through the airport, inform them about delays, and provide other information such as the location of the check-in register for each airline company.
-When constructing a goal model for this system, one stakeholders argues that the user interface of the assistant should by easy to use, and that this can be realized by using a Graphical User Interface (GUI).
-We can formalize this as an instantiation of AS6:\\
-
-\noindent
-\textbf{$A_1$.} \emph{Task ``Use GUI'' contributes to softgoal ``easy to use'' (AS6)}.\\
-
-\noindent
-This argument can be subjected to critical questions.
-In this case, there are five critical questions that can be used on AS6, namely CQ6a-CQ6e.
-Critical question CQ6a asks for alternative solutions to contribute to softgoal ``easy to use''.
-One of the stakeholders may reply by stating a natural language interface is easy to use as well.\\
-
-\noindent
-\textbf{$A_2$.} \emph{Task ``Use NL interface'' contributes to softgoal ``easy to use'' (CQ6a/AS6)}.\\
-
-\noindent
-Note that $A_2$ does not create an attack on $A_1$, since both are valid tasks to contribute to the softgoal.
-Instead, $A_2$ actually generates new elements that will appear in the goal model.
-Now both $A_1$ and $A_2$ can be questioned.
-Suppose CQ6b is used by a stakeholder to argue that a side effect of using a natural language interface is that not all users may be able to understand it.
-When using a GUI, it may be easier to let the user change its language.\\
-
-\noindent
-\textbf{$A_3$.} \emph{Task ``Use NL interface'' contributes negatively to softgoal ``easy to use'' because choosing a language is difficult (CQ6b).}\\
-
-Suppose another stakeholder uses critical question CQ6c to question $A_1$: she notes that implementing a GUI is probably quite expensive, because they don't have a graphical designer, so they will have to outsource this.\\
-
-\noindent
-\textbf{$A_4$.} \emph{Task ``Use GUI'' contributes negatively to softgoal ``low costs''}.\\
-
-\noindent
-The resulting argumentation framework of these four arguments is shown in figure~\ref{fig:example:airport:arguments}.
-As we can see, argument $A_1$ and $A_4$ are both not attacked by any argument.
-However, $A_2$ and $A_3$ are in direct conflict.
-Recall from sectin~\ref{sect:background:pras} that the stakeholders now have to decide about their preferences.
-Suppose the stakeholders decide that argument $A_3$ is stronger, or more preferred, than $A_2$.
-As a result, the accepted arguments are $A_1, A_3$, and $A_4$.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}[scale=2]
-\node (a1) at (0,1) {$A_1$};
-\node (a2) at (1,1) {$A_2$};
-\node (a3) at (1,0) {$A_3$};
-\node (a4) at (0,0) {$A_4$};
-\draw [->] (a2) to [out=-60, in=60] (a3);
-\draw [->] (a3) to [out=120, in=-120] (a2);
-\end{tikzpicture}
-\caption{Argumentation framework for Airport Assistant example.}
-\label{fig:example:airport:arguments}
-\end{figure}
-
-It is now possible to translate this argumentation framework into a GRL model, which is shown in figure~\ref{fig:example:airport:grl}.
-Argument $A_1$ results in a contribution link from task ``use GUI'' to softgoal ``lower costs'', similarly for $A_4$.
-Since the conflict between $A_2$ and $A_3$ has been resolved in favor of $A_3$, there is a negative contribution link from task ``use NL interface'' to softgoal ``Easy to use''.
-
-We can make a number of observations about the obtained GRL model and the underlying arguments:
-\begin{itemize}
-\item The underlying arguments contain information that is not represented in the GRL model.
-For instance, the decision to reject $A_2$ (i.e., that a natural language interface is easy to use) is not visible in the GRL model, but it is in the underlying arguments.
-\item There is a direct relation between the accepted arguments and the elements that appear in the goal model.
-If an element or relation has an underlying accepted argument, it appears in the goal model, while rejected ones do not appear.
-Therefore, if the underlying arguments change \--- for instance if the stakeholders change their preference from $A_2$ to $A_3$ \-- then the resulting goal model changes as well.
-\item In the current example, each GRL element and relation is supported by an argument.
-However, this is not required.
-One could imagine that the stakeholders choose to add additional elements to the goal model without rationalizing them by arguments.
-We explain this in more detail in the RationalGRL methodology in section~\ref{sect:methodology}.
-\item Although it may be intuitively understandable how to translate argument $A_1-A_4$ and the argumentation framework of figure~\ref{fig:example:airport:arguments} into the GRL model of figure~\ref{fig:example:airport:grl}, we have not yet formally specified how to make the translation from GMAS to GRL, and visa versa.
-\end{itemize}
-
-In this section, we have fulfilled requirement three of the introduction: (3) it (the practical reasoning and argumentation theory) must be adaptable for use in goal modeling.
-However, we have not yet considered the fourth requirement: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
-This is what we do in the next section by providing formal mappings from GMAS to GRL, and discussing implementation details.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.35]{img/grl_example_airport}
-\caption{GRL model for Airport Assistant example.}
-\label{fig:example:airport:grl}
-\end{figure}
-
-\section{Mapping GMAS to GRL: Metamodel and Implementation}
-\label{sect:implementation}
-
-\begin{figure*}[h]
-\includegraphics[scale=0.45]{img/ClassDiagram}
-\caption{Metamodel specification of GMAS and GRL}
-\label{fig:metamodel}
-\end{figure*}
-\floris{we have no formal definition of GMAS, so proposing a formal mapping will be difficult}
-In the previous section, we adapted and extended the practical reasoning framework PRAS for use in goal modeling and developed  nine Argument Schemes for Goal Modeling (GMAS).
-These argument schemes can be instantiated to create arguments for elements and relationships in a GRL model.
-We introduced critical questions that can be used to construct counter arguments or alternatives, and we provided an example in which we obtained a GRL model from a GMAS specification.
-However, we did not yet create a formal language for GMAS, nor did we map it to GRL.
-This is what we do in the current section.
-We provide a formal specification of GMAS in terms of a metamodel, and we link it with the GRL metamodel.
-Using this mapping, we detail how to translate an argumentation framework into a GRL model and vice versa.
-The metamodel is the starting point for the tool we developed as an extension of jUCMNav.
-We present this tool in the last subsection, together with the visual language we developed for our extension.
-
-\subsection{GMAS Metamodel Specification}
-\label{sect:implementation:metamodel}
-
-The UML metamodel of GMAS is shown on the left hand side of figure~\ref{fig:metamodel}.
-
-\subsection{Mappings between GMAS and GRL}
-\label{sect:implementation:mappings}
-
-\subsubsection{Information Mappings}
-
-\subsubsection{Formal Mappings}
-
-\subsection{Implementation into jUCMNAv}
-\label{sect:implementation:implementation}
-
 \section{The RationalGRL Methodology}
@@
 \end{figure}
-
-\section{Empirical Evaluation}
-\label{sect:evaluation}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee and Diana Marosin are funded by the National Research Fund (FNR), Luxembourg, by the  RationalArchitecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}

### 2017-01-21 relatedwork.tex
separated files

--- 2017-01-20 versions floris/REIJ_article_FB2.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-\newcommand{\todo}[1]{{\color{red}\textbf{[TODO: } #1 \textbf{]}}}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-\newcommand{\floris}[1]{{\color{red}\textbf{Floris: }#1}}
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Diana Marosin\\
-\vspace{-0.35cm}
-Luxembourg Institute of Science and Technology (LIST)\\
-5 Avenue des Hauts-Fourneaux, 4362 Esch-sur-Alzette, Luxembourg\\ diana.marosin@list.lu
-\and 
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Diana Marosin\Mark{2},
-        Floris Bex\Mark{3},
-        and Sepideh Ghanavati\Mark{4}\\[1em]
- \normalsize
- \begin{tabular}{*{4}{>{\centering}p{.22\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Luxembourg Institute of Science and Technology & \Mark{3}Information and Computing Science & \Mark{4}Department of Computer Science \tabularnewline
-  University of Luxembourg & Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{diana.marosin@list.lu} & \url{florisbex@gmail.com} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-%SG1: Abstract is good for now but when we are done with the paper, we should add our findings wrt the case study in the abstract too
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-We evaluate the framework via a study among 20 subjects, showing that it is superior to traditional goal modeling with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is usually the result of a discussion process between a group of stakeholders and a requirements engineer.
-For small-sized systems that their goal models are usually constructed in a short amount of time, involving stakeholders with a similar background, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SG1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\floris{These four reasons are very interesting. They could be improved by mentioning references in each: 2 and 3 are now without references, which makes it less strong.}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-%SG2: This is good but make them a bit more clear and concrete so they can be used as our base for the evaluation.
-Agreed with floris too.
-I am a bit unsure about the first criteria.
-Why this is needed?
-The main goal is to help better decision making and have more informed decision based on the arguments.
-Try to consider that in your criteria (comparing goal models with or without argumentation)
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for such an argumentation theory: (1) it must have prior successful use in the field of software development, (2) it should allow to formally reason about the consistency of discussions and the goal model, (3) it must be adaptable for use in goal models, (4) the adapted theory must produce repeatable results that can be formalized and implemented, (5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools, and (6) the adapted theory must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-\floris{interesting!
-ad (1) why must the argumentation theory have prior successful use in software development?
-What is `successful use'?
-An academic paper?
-A system prototype?
-A large-scale evaluation?
-Implementation in a system that is used daily in companies?
-ad(4) what do you mean by `repeatable results?'}
-
-In this context, the main research question is: \emph{Can AI theories of practical reasoning and argumentation be adapted, formalized, and combined with an existing goal modeling methodology to rationalize goal model with underlying arguments, and to reason about their consistency?} The satisfaction of the above six requirements will result in a positive answer to the research question.
-%SG3: The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.
-
-\subsection{Contributions} %SG4:this part is a bit hard to read. I think a diagram helps and also make the contributions more aligned with the goals you have above.
-
-The contribution of this paper is to show that the Argument Scheme for Practical Reasoning (PRAS) fits all of the six requirements above.
-We briefly discuss how PRAS meets the first two requirements with an example from the medical domain, particularly in reasoning about medical treatment for a patient.
-\floris{medical domain is not software development, so how does this address req.
-1?} We, then, focus on issue three and four by modifying and implementing PRAS: We develop the Argument Schemes for Goal Modeling (GMAS), an extension to PRAS, which can be used to construct arguments for parts of a goal model.
-These arguments can be attacked with so-called ``critical questions'', which are natural language statements that question a certain aspect of the argument.
-We develop nine argument schemes and 18 critical questions.
-\floris{why does the preceding development of GMAS and the cq's address 3 in that it produces \emph{repeatable} results?} In order to reason about the consistency of the goal model with underlying arguments, we develop three \emph{consistency conditions} \floris{does this not address req.
-2?}.
-We integrate GMAS with the Goal-oriented Requirements Language (GRL) and we implement it in jUCMNav, the Eclipse-based open source tool for GRL modeling.
-To address issue five, we develop the RationalGRL methodology that combines goal modeling with argumentation.
-Finally, we answer issue six positively through an empirical study \floris{is the explanation of the study copied from Singh et al>?} in which 20 subjects apply RationalGRL or GRL to develop a goal modeling in a information system concerning national security.
-The study shows RationalGRL is superior to GRL with respect to completeness and coverage of goal search; size of the goal model; ease of use; explicitness in the assumptions made; and repeatability of conclusions across subjects.
-As a side effect, RationalGRL requires more time than GRL, but is justified by improvements in quality.
-
-\subsection{Organization}
-
-The rest of this paper is organized as follows: In Section~\ref{sect:background}, we introduce a goal modeling language, i.e.
-Goal-oriented Requirements Language (GRL)~\cite{} and the practical reasoning framework PRAS.
-We also provide a motivation and examples for both formalisms.
-In Section~\ref{sect:gmas}, we first analyze the differences between PRAS and GRL in terms of concepts and relationships, after which we introduce the Argument Schemes for Goal Modeling (GMAS) and associated critical questions.
-We, then, illustrate GMAS through several examples.
-In Section~\ref{sect:consistency}, we specify a consistency condition on the goal model and underlying arguments, while in Section~\ref{sect:implementation}, we provide a metamodel for GMAS and we link it to the GRL metamodel.
-We also discuss implementation in jUCMNav~\cite{} and the visualization details of our language.%SG5: I don't think we can call it our language.
-It can be an extenstion.
-For creating a new modeling language you need ontologies etc and we are not doing that.
-In Section~\ref{sect:methodology}, we explain how GMAS can be used by practitioners by introducing the RationalGRL methodology.
-It specifies how goal modeling and argumentation can be used to construct a GRL model.
-In Section~\ref{sect:evaluation}, we describe the design of our empirical study and the findings from it.
-Section~\ref{sect:relatedwork} presents related work and Section~\ref{sect:conclusion} discusses the conclusions and the future work.
-
-\section{Background: Argument Schemes and Goal-oriented Requirements Language}
-\label{sect:background}
-
-In this section, we introduce the Goal-oriented Requirements Language (GRL) and the Argument Scheme for Practical Reasoning (PRAS).
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify  requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL for capturing and documenting argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself.
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals are often more related to NFRs, whereas goals are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available.
-
-A variety of links connect the intentional elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one intentional element on another one.
-A contribution link has a qualitative contribution type or a quantitative contribution values.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model the relationships between actors or intentional elements.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-%SGNew1: I don't think we need this paragraph at all.
-
-\floris{include an example here about the corporate data model etc, which is also used in section \ref{sect:background:pras}. This will make it easier for people to understand the PRAS reasoning}
-
-\subsubsection{Motivation for using GRL}
-\label{sect:background:grl:motivation} %SGNew3: I will rewrite this later.
-
-GRL has the existing tools %SGNew2:? what do you mean existing tools? has some means? potential? 
-that allow us to achieve requirements four and five of a formal argumentation theory as described in the introduction: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
-(5) a methodology must be identified that allows the adapted theory results to be used with existing goal modeling frameworks and tools.
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying criterion 4.
-Moreover, GRL has an open source tool called jUCMNav~\cite{jUCMNav}, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provide support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with the current research: We aim to add traceability links between intentional elements and their underlying arguments.
-%SG8: what do you mean by current research?
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)} 
-\label{sect:background:pras}
-
-Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g.
-\cite{raz1978, walton1990}) and Artificial Intelligence \cite{bratman1987, atkinson2007}.
-One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}.
-The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action.
-This argument can then be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
-
-A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS).
-PRAS follows the following basic argument structure.
-\begin{equation}
-\label{eq:eq1}
-  \begin{aligned}
- \qquad&\text{I have goal } G&\\
-&\text{Doing action }A \text{ will realize goal }G&\\
-&\text{Which will promote value }V&\\
-&\text{\emph{Therefore} I should do action }A.&
-  \end{aligned}
-\end{equation}
-
-So, for example, we can say that XXXX. The argument scheme can be slightly amended to capture subgoals (\emph{i.e.}, realizing goals $G_1, \ldots, G_n$ will allow me to realize goal $G_i$).
-
-Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
-Some examples of critical questions are as follows.
-
-\begin{enumerate}
-\item Will the action bring about the desired goal?
-\item Does the goal promote the value stated?
-\item Are there alternative ways of realizing the same goal?
-\item Are there alternative ways of promoting the same value?
-\item Does doing the action have a side effect which demotes the value?
-\item Does doing the action have a side effect which demotes some other value?
-\item Does doing the action promote some other value?
-\item Does doing the action preclude some other action which would promote some other value?
-\item Is the action possible?
-\item Can the desired goal be realized?
-\item Is the value indeed a legitimate value?
-\end{enumerate}
-
-An unfavorable\footnote{Unfavorable can be either 'no', or 'yes', depending on what the question is.} answer to a critical question then identifies a possible counterargument.
-Take, for example, question 2.
-In the case of our example, we can answer this question favourably with a `yes', as a common way to adhere to a company's corporate data model is obviously to define such a data model.
-However, take critical question 11: if this is answered negatively -- for example, there is evidence that `adhere to corporate data model' was hardly used in project evaluations over the last four years \cite{vanZee-etal:er2016} -- then we have a counterargument: `evidence shows that there is not much attention for adherence to a corporate data model, so it can probably not be realized'.
-Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question X) or an alternative goal that promotes the same value (question Z).
-For example, in \cite{vanZee-etal:er2016} it is argued that the action `use data model of package applications' realizes the goal `use canonical data model', which also promotes lower costs.
-This essentially gives us a counterargument to the original argument -- to define a corporate data model -- that also follows PRAS.
-
-In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa).
-In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}.
-In this section, we will briefly discuss the intuitions behind these semantics.}.
-Figure \ref{fig:pras:example1} shows that an argumentation framework that includes the arguments from the above example can be rendered as a graph.
-The two alternative PRAS instantiations are A1 and A3.
-These arguments mutually attack each other, as `adhere to corporate data model' (A1) is an alternative to `adhere to canonical data model' (A3) and vice versa.
-Argument A2 attacks A1, as it questions the validity of the goal `adhere to corporate data model'.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.6]{img/Fig1}
-\caption{Argumentation framework}
-\label{fig:pras:example}
-\end{figure}
-
-Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics.
-The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
-Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}.
-Argument A2 is undefeated because it has no attackers.
-This makes A1 defeated, because one of its attackers, A2, is undefeated.
-A3 is then also undefeated, since its only attacker, A1, is defeated by A2.
-Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.
-
-In some cases, it is more difficult to determine whether or not an argument is defeated.
-Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two.
-It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{Amgoud}.
-Take, for example, A1 and A3.
-If we say that we prefer the goal ``XXX'' (A1) over the goal ``YYY'' (A3), we remove the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, left).
-This makes A1 the only undefeated argument, whereas A3 is now defeated.
-It is also possible to give explicit arguments for preferences \cite{Modgil2009}.
-These arguments are then essentially attacks on attacks.
-For example, say we prefer A1 over A3 because ``XXX''.
-This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.6]{img/Fig2}
-\caption{Preferences between arguments}
-\label{fig:pras:example2}
-\end{figure}
-
-\subsubsection{Practical Reasoning and Goal Modeling}
-\label{sect:background:pras:motivation}
-
-Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung95} and its various extensions \cite{Amgoud, Modgil2009}.
-The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{Atkinson2006}, law~\cite{???}, planning \cite{MedellinXXX} and choosing between safety critical actions \cite{}.
-The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).
-
-PRAS  and GRL have some obvious similarities in that PRAS defines elements as actions, goals, values and GRL includes tasks, goals, softgoals.
-In previous work, we have already discussed how to translate practical reasoning arguments to GRL diagrams~\cite{}.
-For example, ...
-A similar approach, albeit without the explicit focus on practical reasoning, was taken by Jureta et al.~\cite{}, in which it translates similar structured arguments to goal diagrams.
-One of the problems of these approaches is that ...
-
-The concept that appears in PRAS and not in GRL is \emph{circumstance}.
-In PRAS, an action $a$ is performed in current circumstances $R$, and by performing this action new circumstances $S$ are obtained.
-From this new circumstance $S$, another action may be performed again, and so on.
-In the AI literature, this type of reasoning usually referred to as ``AI planning''~\cite{weld1999recent}: from some initial state, find a sequence of actions such that a goal state is reached if these actions are executed.
-This planning approach is taken by a number of contributions using PRAS to create joint plans for a group of actors, often using a transition system as the underlying semantics~\cite{medellin2014}.
-However, in the early RE phase of an IS, one generally does not consider temporal planning.
-Decisions about the temporal order in which tasks should be executed are generally postponed until a later stage.
-In other words, goal modeling is not concerned with making plans, scenarios, or business processes for the information system.
-In URN (see section~\ref{sect:background:grl}) the distinction between goal modeling and business process modeling is clearly visible in the distinction between two separate modeling languages for the two activities.
-First, goal models are created in GRL and then, the use case maps (similar to business processes) are created in UCM.
-Therefore, when extending PRAS for GRL, we do not consider the notion of \emph{circumstances}.
-
-The concept that appears in GRL but not in PRAS is \emph{resource}.
-A resource may be used by part of the system, or an actor, in order to perform a task or to reach a goal.
-Therefore, PRAS has to be extended with the notion of a resource in order for it to be applicable to GRL.
-%SGNew: to be applicable?
-don't understand the sentence.
-
-Finally, there are two concepts which appear in both models under different names.
-Where PRAS uses the notions ``action'' and ``value'', GRL uses the concepts ``task'' and respectively ``softgoal''.
-However, these two concepts are semantically identical.
-
-With respect to relationships between concepts, the argument scheme PRAS from Section~\ref{sect:background:pras} contains the following relationships:\\
-
-\noindent
-R1. Actor $AG$ \emph{performs} action $A$\\
-R2. Action $A$ \emph{realizes} goal $G$\\
-R3. Goal $G$ \emph{promotes} value $S$\\
-R4. Actor $AG$ \emph{has} goal $G$\\
-R5. Actor $AG$ \emph{has} value $S$\\
-
-\noindent
-
-Relationships R4 and R5 are implicit in PRAS, but follow directly from its formulation (see~\cite{atkinson2007} for more details).
-The relationship ``promotes'' does not exist in GRL, but the relationship ``contributes to'' does, which is conceptually very similar.
-Therefore, we choose to use this relationship to formalize ``promotes'' relation.
-
-As mentioned in Section~\ref{sect:background:grl}, GRL is a very non-restrictive modeling language, allowing intentional elements and links to be combined freely.
-Thus, GRL allows any links between any type of intentional element.
-PRAS, in contrast, only considers relationship R1-R5 above.
-As a result, the critical questions of PRAS are only applicable to a subset of all possible relationships in GRL.
-This, however, does not mean it is not possible to construct arguments and formulate critical questions for other relationship.
-In section~\ref{sect:implementation:implementation}, where we discuss the implementation of our tool, we explain in more detail that a user can straightforwardly add additional critical questions for other relationships as well.
-
-\section{Argument Schemes for Goal Modeling (GMAS)}
-\label{sect:gmas}
-
-%SGNew: This pragprah seems redundant.  
-%PRAS is not directly applicable to goal modeling in general, and GRL in particular.
-There are differences in the concepts used, and the relationships between these concepts.
-Some elements appear in PRAS that do not appear in GRL, some elements appear in GRL but not in PRAS, and some concepts appear in both formalisms but with different names.
-Moreover, GRL allows more relations between elements than the ones appearing in PRAS.
-
-%SGNew: What is the difference between PRAS and PAS?
-I think you should start this section with saying that based on what you defined in the last section, we developed a new argument schemes for goal modeling which is called GMAS.
-Then you can go ahead and talk that this is  the extension of PRAS, etc.
-I tried to reword.
-%In this section, we first analyze these differences in more detail. A
-
-In this section, we define our new argument schemes for GRL which is called Argument Schemes for Goal Modeling (GMAS) and its associated critical questions.
-GMAS is an extension of PRAS.
-PRAS  contains a single argument scheme PAS while GMAS  consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
-Each argument scheme is associated with critical questions that can be used to attack arguments or provide alternatives.
-This enables focused discussions about intentional elements and their relationships in a goal model.
-In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
-
-%After this, we introduce and motivate Argument Schemes for Goal Modeling (GMAS) and associated critical questions, which are an extension to PRAS.
-While PRAS contains a single argument scheme PAS, GMAS instead consists of nine argument schemes, which each can be used to instantiate an argument for particular part of a goal model.
-With each argument scheme we associate critical questions that can be used to attack arguments or provide alternatives.
-This enables focused discussions about elements and relationships of a goal model.
-In the last subsection we illustrate applications of the argument schemes and critical questions with a simple example.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Attacks?}\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & Yes\\ %SGNew: tasks may also require resources to be available.. how do you show that?
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & Yes\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & Yes\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal? & Yes\\
-\hline
-\hline
-AS5 & Task $T$ realizes goal $G$ & CQ5a & Will the task realize the desired goal? & Yes\\
-& & CQ5b & Are there alternative ways of realizing the same goal? \floris{How do you handle OR and XOR decompositions?} & No\\ %SGNew:What about decomposition of goals into goals?
-\hline
-AS6 & Task $T$ contributes to softgoal $S$ & CQ6a & Are there alternative ways of contributing to the same softgoal? & No\\
-&& CQ6b & Does the task have a side effect which contribute negatively to the softgoal?&Yes\\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal? & No\\
-&& CQ6d & Does the task contribute to some other softgoal?&No\\
-&& CQ6e & Does the task preclude some other task which would contribute to some other softgoal?&Yes\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal? & Yes\\
-&& CQ7b & Does the goal contribute to some other softgoal? & No\\ %SGNew: What about contribution of goals to goals or softgoals to softgoals?
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8a & Is the resource required in order to perform the task? & Yes\\
-&&CQ8b & Can other resources be used to perform the same task? & No\\
-&&CQ8c & Is the resource required in order to perform the task? & Yes\\
-&&CQ8d & Does using the resource make other resources unavailable?
-& Yes\\ %SGNew: Tasks need resources..
-that is why tasks depends on resources..
-I can't understand the contribution link between them..
-resource doesn't have impact on the task..it should be dependency link.
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors? & No\\
-\hline
-\end{tabular}
-\caption{Argument schemes and critical questions for GRL elements (CQ1-CQ4) and relationships (CQ5-CQ16), as well as whether answering the critical question unfavorably results in an attack on the original argument (right column).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{The Argument Schemes and Critical Questions}
-
-PRAS consists of the single argument scheme PAS.
-However, our approach is to split this single scheme into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing the ones that question the elements which are not part of GRL (\emph{circumstances}), by adding questions for additional elements in GRL (\emph{resources}), and finally by renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-We developed a total of 9 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS1-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before%SGNew: where?,
-this list is not meant to be exhaustive, since GRL does not put restrictions on any of the relationships and  in theory any type of elements can be combined with any other type of elements, using any relationship.
-Here, we merely use the critical questions developed by Atkinson et al.
-%SGNew: merely = only?
-I thought you modified them?
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-In general, answering a critical questions can have two effects: %SGNew: why unfavorably?
-\begin{enumerate}
-\item
-Answering the critical question unfavorably results in a new argument attacking the original argument (\emph{attack}).
-\item
-Answering the critical question unfavorably results in a new argument not attacking the original argument, but representing an alternative (\emph{no attack}).
-\end{enumerate}
-The last column of table~\ref{table:argument-schemes} shows for each critical question whether it results in an attack on the original argument or not.
-
-\floris{you do not provide a formal definition of GMAS} %SGNew: Agreed.
-What is missing here is a big picture or as we can say a methodology.
-I think it is important to introduce the methodology earlier and show how everything fit together.
-Also, to show what are the elements of GMAS and how it can be connected to GRL.
-You said, you extended PRAS and that is GMAS but can you show the extension?
-
-\subsection{Example: Airport Assistant}
-
-We now provide a simple scenario of arguments created by instantiating AS1-AS9, and demonstrate how new arguments can be generated through critical questions, possibly attacking the original arguments.
-
-Consider a system for assisting users at the airport.
-The assistant should be able to guide users through the airport, inform them about delays, and provide other information such as the location of the check-in register for each airline company.
-When constructing a goal model for this system, one stakeholders argues that the user interface of the assistant should by easy to use, and that this can be realized by using a Graphical User Interface (GUI).
-We can formalize this as an instantiation of AS6:\\
-
-\noindent
-\textbf{$A_1$.} \emph{Task ``Use GUI'' contributes to softgoal ``easy to use'' (AS6)}.\\
-
-\noindent
-This argument can be subjected to critical questions.
-In this case, there are five critical questions that can be used on AS6, namely CQ6a-CQ6e.
-Critical question CQ6a asks for alternative solutions to contribute to softgoal ``easy to use''.
-One of the stakeholders may reply by stating a natural language interface is easy to use as well.\\
-
-\noindent
-\textbf{$A_2$.} \emph{Task ``Use NL interface'' contributes to softgoal ``easy to use'' (CQ6a/AS6)}.\\
-
-\noindent
-Note that $A_2$ does not create an attack on $A_1$, since both are valid tasks to contribute to the softgoal.
-Instead, $A_2$ actually generates new elements that will appear in the goal model.
-Now both $A_1$ and $A_2$ can be questioned.
-Suppose CQ6b is used by a stakeholder to argue that a side effect of using a natural language interface is that not all users may be able to understand it.
-When using a GUI, it may be easier to let the user change its language.\\
-
-\noindent
-\textbf{$A_3$.} \emph{Task ``Use NL interface'' contributes negatively to softgoal ``easy to use'' because choosing a language is difficult (CQ6b).}\\
-
-Suppose another stakeholder uses critical question CQ6c to question $A_1$: she notes that implementing a GUI is probably quite expensive, because they don't have a graphical designer, so they will have to outsource this.\\
-
-\noindent
-\textbf{$A_4$.} \emph{Task ``Use GUI'' contributes negatively to softgoal ``low costs''}.\\
-
-\noindent
-The resulting argumentation framework of these four arguments is shown in figure~\ref{fig:example:airport:arguments}.
-As we can see, argument $A_1$ and $A_4$ are both not attacked by any argument.
-However, $A_2$ and $A_3$ are in direct conflict.
-Recall from sectin~\ref{sect:background:pras} that the stakeholders now have to decide about their preferences.
-Suppose the stakeholders decide that argument $A_3$ is stronger, or more preferred, than $A_2$.
-As a result, the accepted arguments are $A_1, A_3$, and $A_4$.
-
-\begin{figure}
-\centering
-\begin{tikzpicture}[scale=2]
-\node (a1) at (0,1) {$A_1$};
-\node (a2) at (1,1) {$A_2$};
-\node (a3) at (1,0) {$A_3$};
-\node (a4) at (0,0) {$A_4$};
-\draw [->] (a2) to [out=-60, in=60] (a3);
-\draw [->] (a3) to [out=120, in=-120] (a2);
-\end{tikzpicture}
-\caption{Argumentation framework for Airport Assistant example.}
-\label{fig:example:airport:arguments}
-\end{figure}
-
-It is now possible to translate this argumentation framework into a GRL model, which is shown in figure~\ref{fig:example:airport:grl}.
-Argument $A_1$ results in a contribution link from task ``use GUI'' to softgoal ``lower costs'', similarly for $A_4$.
-Since the conflict between $A_2$ and $A_3$ has been resolved in favor of $A_3$, there is a negative contribution link from task ``use NL interface'' to softgoal ``Easy to use''.
-
-We can make a number of observations about the obtained GRL model and the underlying arguments:
-\begin{itemize}
-\item The underlying arguments contain information that is not represented in the GRL model.
-For instance, the decision to reject $A_2$ (i.e., that a natural language interface is easy to use) is not visible in the GRL model, but it is in the underlying arguments.
-\item There is a direct relation between the accepted arguments and the elements that appear in the goal model.
-If an element or relation has an underlying accepted argument, it appears in the goal model, while rejected ones do not appear.
-Therefore, if the underlying arguments change \--- for instance if the stakeholders change their preference from $A_2$ to $A_3$ \-- then the resulting goal model changes as well.
-\item In the current example, each GRL element and relation is supported by an argument.
-However, this is not required.
-One could imagine that the stakeholders choose to add additional elements to the goal model without rationalizing them by arguments.
-We explain this in more detail in the RationalGRL methodology in section~\ref{sect:methodology}.
-\item Although it may be intuitively understandable how to translate argument $A_1-A_4$ and the argumentation framework of figure~\ref{fig:example:airport:arguments} into the GRL model of figure~\ref{fig:example:airport:grl}, we have not yet formally specified how to make the translation from GMAS to GRL, and visa versa.
-\end{itemize}
-
-In this section, we have fulfilled requirement three of the introduction: (3) it (the practical reasoning and argumentation theory) must be adaptable for use in goal modeling.
-However, we have not yet considered the fourth requirement: (4) the adapted theory must produce repeatable results that can be formalized and implemented.
-This is what we do in the next section by providing formal mappings from GMAS to GRL, and discussing implementation details.
-
-\begin{figure}
-\centering
-\includegraphics[scale=0.35]{img/grl_example_airport}
-\caption{GRL model for Airport Assistant example.}
-\label{fig:example:airport:grl}
-\end{figure}
-
-\section{Mapping GMAS to GRL: Metamodel and Implementation}
-\label{sect:implementation}
-
-\begin{figure*}[h]
-\includegraphics[scale=0.45]{img/ClassDiagram}
-\caption{Metamodel specification of GMAS and GRL}
-\label{fig:metamodel}
-\end{figure*}
-\floris{we have no formal definition of GMAS, so proposing a formal mapping will be difficult}
-In the previous section, we adapted and extended the practical reasoning framework PRAS for use in goal modeling and developed  nine Argument Schemes for Goal Modeling (GMAS).
-These argument schemes can be instantiated to create arguments for elements and relationships in a GRL model.
-We introduced critical questions that can be used to construct counter arguments or alternatives, and we provided an example in which we obtained a GRL model from a GMAS specification.
-However, we did not yet create a formal language for GMAS, nor did we map it to GRL.
-This is what we do in the current section.
-We provide a formal specification of GMAS in terms of a metamodel, and we link it with the GRL metamodel.
-Using this mapping, we detail how to translate an argumentation framework into a GRL model and vice versa.
-The metamodel is the starting point for the tool we developed as an extension of jUCMNav.
-We present this tool in the last subsection, together with the visual language we developed for our extension.
-
-\subsection{GMAS Metamodel Specification}
-\label{sect:implementation:metamodel}
-
-The UML metamodel of GMAS is shown on the left hand side of figure~\ref{fig:metamodel}.
-
-\subsection{Mappings between GMAS and GRL}
-\label{sect:implementation:mappings}
-
-\subsubsection{Information Mappings}
-
-\subsubsection{Formal Mappings}
-
-\subsection{Implementation into jUCMNAv}
-\label{sect:implementation:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Empirical Evaluation}
-\label{sect:evaluation}
-
 \section{Related Work}
@@
 Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee and Diana Marosin are funded by the National Research Fund (FNR), Luxembourg, by the  RationalArchitecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\end{document}

### 2017-01-21 tool.tex
separated files

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
 \section{The RationalGRL Tool}
 \label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-01-21 transcripts.tex
separated files

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
-\label{sect:preview}
-
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
-
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
-
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
-
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
-
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
 \section{Transcripts excerpts}
@@
 \end{table}
-
-\end{document}

### 2017-01-21 background.tex
update

@@
 These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
+The model in figure~\ref{fig:SchedulerModel} does not contain information about discussions that let up to the resulting elements of the model, such as by various clarification steps for the naming, or alternatives that have been considered for the relationships.
+In order to address this, we use Argument Schemes for Goal Modeling (GMAS).
 GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
@@
 Take, for example, critical question 8.
-If, for example, there is evidence that schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textsf{user friendliness} is not a legitimate value'.
+If new evidence comes up indicating that meeting schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textsf{user friendliness} is not a legitimate value'.
 Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).
@@
 The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
-Take, for example, the argumentation framework in Figure~\ref{fig:pras:example}.
-Argument A2 is undefeated because it has no attackers.
+In the argumentation framework in Figure~\ref{fig:pras:example}, argument A2 is undefeated because it has no attackers.
 This makes A1 defeated, because one of its attackers, A2, is undefeated.
@@
 The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
-The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{}).
+The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \todo{Marc}{Floris}{Which article do you want to cite here?}\cite{}).
+
+\todo{Marc}{Floris}{From here on you seem to be discussing related work.
+This kind of distracts us from the main point of this article, and the section becomes quite long.
+Perhaps it is better to make the main arguments about why our approach is useful here, and move the discussion of our previous work and Jureta to related work?}
 

### 2017-01-21 gmas.tex
update

@@
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
+\section{Argument Schemes for Goal Modeling (GMAS)}
 \label{sect:gmas}
 
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
+In this section we develop a set of argument schemes and critical questions for goal modeling.
+The list of argument schemes and critical questions is shown in table~\ref{table:argument-schemes}.
+The first four argument schemes (AS0-AS4) are argument for an element of a goal model, the next seven (AS5-AS11) are about relationships, and the remain two (AS12-AS13) are about intentional elements in general.
+For each critical question, the right column shows the effect of answering the critical questions affirmatively, which can be DISABLE (disable the element of the original argument), INTRO (introduce a new element and argument), or REPLACE (replace the element of the original argument).
+For instance, if the critical question CQ0 ``Is the actor relevant?'' is answered with ``No'', then the actor of the original argument AS0 will be disabled in the goal model.
+We come back to these effects in more detail in subsection~\ref{sect:?} of this section.
+
+We obtained the list in table~\ref{table:argument-schemes} by analyzing transcripts of discussions between students about the development of an information system.
+In this section, we provide details of this analysis process with concrete examples.
 
@@
 \centering
-\begin{tabular}{|l|l|l|l|l|}
+\begin{tabularx}{\textwidth}{|l|l|l|X|l|l|}
 \hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
+\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Effect}\\
 \hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
+AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant? & DISABLE\\
 \hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
+AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & DISABLE\\
 \hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
+AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & DISABLE\\
 \hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
+AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & DISABLE\\
 \hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
+AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?& DISABLE\\
 \hline
 \hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
+AS5 & goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
+& & CQ5b & Does the goal decompose into any other tasks?& INTRO\\
 \hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
+AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
+&& CQ6b & Are there alternative ways of contributing to the same softgoal?& INTRO \\
+&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?& INTRO\\
+&& CQ6d & Does the task contribute to some other softgoal?& INTRO\\
 \hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
+AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?& DISABLE\\
+&& CQ7b & Does the goal contribute to some other softgoal?& INTRO\\
 \hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
+AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?& DISABLE\\
 \hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
+AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?& INTRO\\
 \hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
+AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& INTRO\\
+ &  & CQ10b & Is the decomposition type correct? (AND/OR/XOR)& REPLACE\\
+\hline
+AS11 & Task $T$ contributes negatively to softgoal $S$& CQ11 & Does the task contribute negatively to the softgoal?& DISABLE\\
+\hline
+\hline
+AS12 & Element $IE$ is relevant & CQ12a & Is the element relevant? & DISABLE\\
+&  & CQ12b & Is the element redundant? & DISABLE\\
+\hline
+AS13 & Element $IE$ has name $n$ & CQ12a & Is the name clear? & REPLACE\\
+&  & CQ12b & Is the name ambiguous? & REPLACE\\
+\hline
+\end{tabularx}
+\caption{List of argument schemes (AS0-AS13, left column), critical questions (CQ0-CQ12, middle column), and the effect of answering them (right column).}
 \label{table:argument-schemes}
@@
 
-\subsection{Argument schemes and critical questions}
-
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
-
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
+\subsection{Details experiment}
 
@@
 The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
+All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software.
 The students had two hours for the tasks, and the transcripts document the entire discussion.
@@
 
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
+\paragraph{Annotation method}
+
+We started with an initial list of 10 argument schemes and 18 critical questions that we derived from PRAS.
+We annotated transcripts with the arguments and critical questions from this list.
+If we found arguments or critical questions that did not appear in the original list, we added them and counted them as well.
+Argument schemes that did not appear were removed from the list.
 Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
+This can be seen in the various examples we will discuss.
 
@@
 
-\subsubsection{Results}
+\paragraph{Results}
 
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
 We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
 The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
 
@@
 \centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
+\begin{tabularx}{0.5\textwidth}{|l|X|l|l|l|>{\bfseries}l|}
 \hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
+\multicolumn{2}{|c|}{\textbf{Scheme/Question}} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
 \hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
+AS0 & Actor & 2 & 2 & 5 & 9\\
 \hline
-AS1: Resource & 2 & 4 & 5 & 11\\
+AS1 & Resource & 2 & 4 & 5 & 11\\
 \hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
+AS2 & Task/action & 20 & 21 & 17 & 58\\
 \hline
-AS3: Goal & 0 & 2 & 2 & 4\\
+AS3 & Goal & 0 & 2 & 2 & 4\\
 \hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
+AS4 & Softgoal & 3 & 4 & 2 & 9\\
 \hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
+AS5 & Goal decomposes into tasks & 4 &0& 4 & 8\\
 \hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
+AS6 & Task contributes to softgoal & 6 & 2 &0& 8\\
 \hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
+AS7 & Goal contributes to softgoal &0& 1 & 1 & 2\\
 \hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
+AS8 & Resource contributes to task & 0 & 4 & 3 & 7\\
 \hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
+AS9 & Actor depends on actor &0& 1 & 3 & 4\\
+\hline
+AS10 & Task decomposes into tasks & 11 &14 &11 &36\\ 
+\hline
+AS11 & Task contributes negatively to softgoal & 2 & 1 & 0 & 3\\
 \hline
 \hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
+CQ2 & Task is possible? & 2 & 2 & 1 & 5\\
 \hline		
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
+CQ5a & Does the goal decompose into the tasks? & 0 & 1 & 0 & 1\\
 \hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
+CQ5b & Alternative ways to realize goal? & 1 & 0 & 0 & 1\\
 \hline
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
+CQ6b & Task has negative side effects? & 2 & 0 & 0 & 2\\
 \hline
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
+CQ10a & Task decompose into other tasks? & 1 &2 &0&3\\
 \hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
+CQ10b & Decomposition type correct? &1 &0& 1 &2\\
 \hline
 \hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
+CQ12 & Useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
 \hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
+CQ13 & Clarifying an element &3 &10 & 3 & 16\\
+\hline
+\hline
+\multicolumn{2}{|c|}{\textbf{TOTAL}}&69&80&69&218\\
+\hline
+\end{tabularx}
+\caption{Number of occurrences of AS0-AS9, CQ0-CQ12 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
+\label{table:transcripts:results:argumentschemes}
 \end{table}
 
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
+We found that answering a critical questions can have three effects on the original argument:
+\begin{itemize}
+\item \emph{INTRO}: Introduce a new element with a corresponding argument. This operation does not attack the original argument, but rather creates a new argument.
+\item \emph{DISABLE:} Disable the element of the argument scheme to which the critical questions pertains.
+This operation does not create a new argument, but only disables (i.e., attacks) the original one.
+\item \emph{REPLACE:} Replace the element of the argument scheme with a new element. This operation both introduces a new argument and attacks the original one.
+\end{itemize}
 
-\subsubsection{Analysis}
+\subsection{Analysis}
 
 The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
+First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of answering a critical question affirmatively.
 
 \paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
+Our initial list of argument schemes consists of AS1-AS4, AS5-AS9 (figure~\ref{table:argument-schemes}).
+Therefore, the difference between the initial list of argument schemes and those found back in the transcripts is quite small.
 We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
 This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
+
 More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
 Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
+GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 11.
+Our analysis however shows that many of these schemes are not often used, and thus gives us some confidence in the resulting list.
 
@@
 If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
+In the context of tool support, we believe that having these critical questions available may stimulate discussions.
 
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
+\subsection{Examples}
 
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
+For each transcript, we manually create a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
+An example of such a model is shown in figure~\ref{fig:transcripts:grl}.
+We added green and red dots to various elements and relationships in the figure.
+A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
+Note that if the underlying argument is rejected, the corresponding GRL element has been disabled.
+Some elements do not have a corresponding green or red dot.
+In that case, we have inferred the elements from the discussion, but we could not explicitly find back arguments for it.
 
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
+We now discuss various instantiations of argument schemes and the result of answering critical questions in more detail.
+The examples below come from all three transcripts, and therefore not always correspond to the goal model in figure~\ref{fig:transcripts:grl}.
+For each example we provide transcript excerpts, a visualization of the arguments, and the corresponding goal model elements.
 
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
+\subsubsection{Irrelevant actor}
 
-\subsection{Final argument schemes and critical questions}
+The transcript excerpt of the first example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix.
+It consists of participant P1 putting forth the suggestion to include actor \emph{Development team} into the model.
+This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.
 
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
+We formalized the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}.
+This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
+This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
+This is shown in figure~\ref{fig:examples:relevant-actor:arguments}.
 
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
+Argument scheme AS0 is an argument to include actor \emph{Development team} into the GRL model.
+Therefore, by attacking it the argument is no longer accepted, and as a result the actor is disabled in the resulting GRL model (figure~\ref{fig:examples:relevant-actor:grl}).
+Note we added a red traceability link from the rejected argument to the disabled GRL element.
 
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
+\begin{figure}[ht!]
+\centering
+\begin{subfigure}[b]{0.22\textwidth}
+        \centering
+        \begin{tikzpicture}[->]
+        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (0,0) {
+            \begin{tabularx}{\textwidth}{X}
+                \textbf{(AS0)} Development team is relevant 
+            \end{tabularx}
+        } ;
+        \node (a1) [shape=rectangle,draw] at (0,-2){
+            \begin{tabularx}{\textwidth}{X}
+                \textbf{(CQ0)} Development team is not relevant (\emph{The professor makes the software})
+            \end{tabularx}
+        };
+        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace) at (1.2,-0.3) {};
+
+         \path[line width=1mm]
+    (a1) edge[draw=black] (a0);
+\end{tikzpicture}
+        \caption{Arguments}
+        \label{fig:examples:relevant-actor:arguments}
+    \end{subfigure}\hfill
+    \begin{subfigure}[b]{0.22\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{img/actor_disabled}
+        \caption{GRL elements}
+        \label{fig:examples:relevant-actor:grl}
+    \end{subfigure}
+\caption{Argument schemes and critical questions (left) and GRL model (right) of a discussion about the relevance of actor Development team.}
+\label{fig:examples:relevant-actor}
+\end{figure}
+
+We see from this example that answering critical question CQ0 affirmatively has the effect of disabling the corresponding actor \emph{Development team}.
+Therefore, critical question CQ0 has the associated effect DISABLE in table~\ref{table:argument-schemes}.
+
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{img/transcript_grl}
+\caption{The GRL model manually constructed from transcript 1.
+Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
+Elements and relationships with no dot have been inferred by us.}
+\label{fig:transcripts:grl}
+\end{figure*}

### 2017-01-21 main.tex
update

@@
 \usepackage{authblk}
+\usepackage{caption}
 \usepackage{courier}
@@
 \usepackage{sectsty}
+\usepackage{subcaption}\captionsetup{compatibility=false}
+\usepackage{tabularx}
 \usepackage{tikz}\usetikzlibrary{arrows}
@@
 
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
+\input{abstract}
 

### 2017-01-21 gmas.tex
update

@@
 
-\subsubsection{Irrelevant actor}
+\subsubsection{Example 1: Introducing a new element (INTRO)}
 
-The transcript excerpt of the first example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix.
-It consists of participant P1 putting forth the suggestion to include actor \emph{Development team} into the model.
-This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.
+\subsubsection{Example 2: Disabling an element (DISABLE)}
 
-We formalized the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}.
-This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
-This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
-This is shown in figure~\ref{fig:examples:relevant-actor:arguments}.
-
-Argument scheme AS0 is an argument to include actor \emph{Development team} into the GRL model.
-Therefore, by attacking it the argument is no longer accepted, and as a result the actor is disabled in the resulting GRL model (figure~\ref{fig:examples:relevant-actor:grl}).
-Note we added a red traceability link from the rejected argument to the disabled GRL element.
-
-\begin{figure}[ht!]
-\centering
-\begin{subfigure}[b]{0.22\textwidth}
-        \centering
-        \begin{tikzpicture}[->]
-        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (0,0) {
-            \begin{tabularx}{\textwidth}{X}
-                \textbf{(AS0)} Development team is relevant 
-            \end{tabularx}
-        } ;
-        \node (a1) [shape=rectangle,draw] at (0,-2){
-            \begin{tabularx}{\textwidth}{X}
-                \textbf{(CQ0)} Development team is not relevant (\emph{The professor makes the software})
-            \end{tabularx}
-        };
-        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace) at (1.2,-0.3) {};
-
-         \path[line width=1mm]
-    (a1) edge[draw=black] (a0);
-\end{tikzpicture}
-        \caption{Arguments}
-        \label{fig:examples:relevant-actor:arguments}
-    \end{subfigure}\hfill
-    \begin{subfigure}[b]{0.22\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{img/actor_disabled}
-        \caption{GRL elements}
-        \label{fig:examples:relevant-actor:grl}
-    \end{subfigure}
-\caption{Argument schemes and critical questions (left) and GRL model (right) of a discussion about the relevance of actor Development team.}
-\label{fig:examples:relevant-actor}
-\end{figure}
-
-We see from this example that answering critical question CQ0 affirmatively has the effect of disabling the corresponding actor \emph{Development team}.
-Therefore, critical question CQ0 has the associated effect DISABLE in table~\ref{table:argument-schemes}.
+\subsubsection{Example 3: Change decomposition type (REPLACE)}
 
@@
 \end{figure*}
+
+\subsubsection{Example 4: Clarifying an element (REPLACE)}
+
+The transcript excerpt of the second example is shown in table~\ref{table:transcript:task-clarification} in the appendix and comes from transcript $t_3$.
+It consists of a number of clarification steps, resulting in the task \emph{Choose a road pattern}.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}[->]
+        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (-2,0) {
+            \begin{tabularx}{0.2\textwidth}{X}
+                \textbf{(AS2)} Actor \emph{student} has task \emph{create road}
+            \end{tabularx}
+        };
+        \node (a1) [shape=rectangle, draw=black!60, fill=white, text=black!40] at (-2,-2) {
+            \begin{tabularx}{0.2\textwidth}{X}
+                \textbf{(AS2)} Actor \emph{student} has task \emph{choose a pattern}
+            \end{tabularx}
+        };
+        \node (a2) [shape=rectangle, draw=black!60, fill=white, text=black!40] at (-2,-4.2) {
+            \begin{tabularx}{0.2\textwidth}{X}
+                \textbf{(AS2)} Actor \emph{student} has task \emph{choose a pattern preference}
+            \end{tabularx}
+        };
+        \node (a3) [shape=rectangle] at (-2,-6.7) {
+            \begin{tabularx}{0.2\textwidth}{X}
+                \textbf{(AS2)} Actor \emph{student} has task \emph{choose a road pattern}
+            \end{tabularx}
+        };
+        \node[draw=none] (actor) at (2.3,-4) { \includegraphics[scale=0.5]{img/task_choose_road_pattern} };
+        \node[circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt] (trace1) at (-.5,-6.5) {};
+        \node[circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt] (trace2) at (2.2,-4.4) {};
+
+\begin{pgfonlayer}{background}
+         \path[line width=1mm]
+    (a1) edge[draw=black] (a0)
+    (a2) edge[draw=black] (a1)
+    (a2) edge[draw=black, bend right=20] (a0)
+    (a3) edge[draw=black] (a2)
+    (a3) edge[draw=black, bend right=20] (a1)
+    (a3) edge[draw=black, bend right=40] (a0);
+\end{pgfonlayer}
+     \path[line width=0.5mm, dotted]
+    (a0) edge [bend right=50] node [left,draw=none] {CQ12a} (a1)
+    (a1) edge [bend right=50] node [left,draw=none] {CQ12a} (a2)
+    (a2) edge [bend right=50] node [left,draw=none] {CQ12a} (a3);
+
+    \path[draw, densely dotted,-,line width=0.5mm] (trace1) edge[draw=green] (trace2);
+\end{tikzpicture}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the clarification of a task.}
+\label{fig:examples:clarification}
+\end{figure}
+
+The formalized argument schemes and critical questions are shown in figure~\ref{fig:examples:clarification}.
+The discussion starts with the first instantiation of argument scheme AS2: \emph{Actor student has task create road}.
+This argument is then challenged with critical question CQ12a: \emph{Is the task create road clear?}.
+Answering this question results in a new instantiation of argument scheme AS2: \emph{Actor student has task choose a pattern}.
+This process is repeated three times, resulting in the final argument scheme \emph{Actor student has task choose a road pattern}.
+This final argument is unattacked and has a corresponding intentional element (right image).
+
+What is clearly shown in this example is that a clarification step does not only attacks the previous argument, but it also attacks all arguments previously used to describe the element.
+For instance, the final argument on the bottom of figure~\ref{fig:examples:clarification} attacks all previous arguments.
+If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.
+
+\subsubsection{Example 5: Generic counterargument (ATTACK)}
+
+\todo{Marc}{Marc}{Add the additional counterargument and add to critical questions}
+
+The transcript excerpt of the first example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
+It consists of participant P1 putting forth the suggestion to include actor \emph{Development team} into the model.
+This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}[->]
+        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (-2,0) {
+            \begin{tabularx}{0.2\textwidth}{X}
+                \textbf{(AS0)} Development team is relevant 
+            \end{tabularx}
+        } ;
+        \node (a1) [shape=rectangle,draw] at (-2,-2.5){
+            \begin{tabularx}{0.2\textwidth}{X}
+                Development team is not relevant (\emph{The professor makes the software})
+            \end{tabularx}
+        };
+        \node[draw=none] (actor) at (2.3,-1) { \includegraphics[scale=0.8]{img/actor_disabled} };
+        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace1) at (-.5,-.3) {};
+        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace2) at (2.2,-0.1) {};
+
+         \path[line width=1mm]
+    (a1) edge[draw=black] (a0);
+    \path[line width=0.5mm, dotted]
+    (a0) edge [bend right=50] node [left,draw=none] {CQ0} (a1);
+    \path[draw, densely dotted,-,line width=0.5mm] (trace1) edge[draw=red] (trace2);
+\end{tikzpicture}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
+\label{fig:examples:relevant-actor}
+\end{figure}
+
+We formalized the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}.
+This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
+This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
+This is shown in figure~\ref{fig:examples:relevant-actor}, left image.
+
+Argument scheme AS0 is an argument to include actor \emph{Development team} into the GRL model.
+Therefore, by attacking it the argument is no longer accepted, and as a result the actor is disabled in the resulting GRL model (figure~\ref{fig:examples:relevant-actor}, right image).
+Note we added a red traceability link from the rejected argument to the disabled GRL element.
+
+We see from this example that answering critical question CQ0 affirmatively has the effect of disabling the corresponding actor \emph{Development team}.
+Therefore, critical question CQ0 has the associated effect DISABLE in table~\ref{table:argument-schemes}.

### 2017-01-21 transcripts.tex
update

@@
 \end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
+\caption{Discussion about the relevance of an actor (transcript $t_3$)}
+\label{table:transcript:irrelevant-actor}
 
+\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
+\hline
+Respondent & Text & Annotation\\
+\hline
+0:17:39.5 (P1) & And in that process there are activities like create a visual map, create a road& \textbf{[14 task (AS2)]} Student has task ``Create road''\\
+\hline
+0:24:36.0 (P3)	& And, well interaction.
+Visualization sorry.
+Or interaction, I don't know.
+So create a visual map would have laying out roads and a pattern of their choosing.
+So this would be first, would be choose a pattern.
+& \multirow{3}{60mm}{\textbf{[31 critical question CQ??
+for 14]} Is Task ``Create road'' clear?\\
+\textbf{[32 answer to 31]} no, according to the specification the student should choose a pattern.\\
+\textbf{[32a REPLACE]} ``Create road'' becomes ``Choose a pattern''}\\
+\cline{1-2}
+0:24:55.4 (P1) &	How do you mean, choose a pattern	&\\
+\cline{1-2}
+0:24:57.5 (P3)	& Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing	&\\
+\hline
+0:25:07.5 (P1)	& Yeah I'm not sure if they mean that.
+I don't know what they mean by pattern in this case.
+I thought you could just pick roads, varying sizes and like, broads of roads.
+& \textbf{[33 critical question CQ??
+for 32a]} Is ``Choose a pattern'' clear?\newline
+\textbf{[34 answer to 33]} No, not sure what they mean by a pattern.\\
+\hline
+0:25:26.0 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \multirow{3}{60mm}{\textbf{[34a REPLACE]} ``Choose a pattern'' becomes ``Choose a pattern preference''}\\
+\cline{1-2}
+0:25:36.3 (P1) & Yeah, selecting a- I don't know &\\
+\cline{1-2}
+0:25:38.0 (P3)	& Pattern preference maybe? As in, maybe we can explain this in the documentation &\\
+\hline
+0:25:43.9 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ?? for 34a]} Is ``Choose a pattern preference'' clear?\newline
+\textbf{[36 answer to 35]} no, what kind of pattern?\\
+\hline
+0:25:47.4 (P3) & Maybe, I don't know it's- & \multirow{2}{60mm}{\textbf{[36a rename]} ``Choose a pattern preference'' becomes ``Choose a road pattern''}\\
+\cline{1-2}
+0:25:48.5 (P1)	& [inaudible] a road pattern&\\
+\hline
+\end{tabular}
+\caption{Clarifying the name of a task (transcript $t_3$)}
+\label{table:transcript:task-clarification}
+\end{table}
+
+\begin{table}[!htbp]
 \begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}

### 2017-01-22 gmas.tex
GMAS finished

--- 2017-01-21 gmas.tex
@@
 In this section we develop a set of argument schemes and critical questions for goal modeling.
-The list of argument schemes and critical questions is shown in table~\ref{table:argument-schemes}.
-The first four argument schemes (AS0-AS4) are argument for an element of a goal model, the next seven (AS5-AS11) are about relationships, and the remain two (AS12-AS13) are about intentional elements in general.
-For each critical question, the right column shows the effect of answering the critical questions affirmatively, which can be DISABLE (disable the element of the original argument), INTRO (introduce a new element and argument), or REPLACE (replace the element of the original argument).
+The final list of argument schemes and critical questions is shown in table~\ref{table:argument-schemes}.
+The first four argument schemes (AS0-AS4) are argument for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
+For each critical question, the right column shows the effect of answering the critical questions affirmatively, which can be DISABLE (disable the element of the original argument), INTRO (introduce a new element and argument), REPLACE (replace the element of the original argument), ATTACK (attack an argument directly).
 For instance, if the critical question CQ0 ``Is the actor relevant?'' is answered with ``No'', then the actor of the original argument AS0 will be disabled in the goal model.
-We come back to these effects in more detail in subsection~\ref{sect:?} of this section.
 
 We obtained the list in table~\ref{table:argument-schemes} by analyzing transcripts of discussions between students about the development of an information system.
-In this section, we provide details of this analysis process with concrete examples.
+In the first subsection we provide details of this analysis process with concrete examples, after which we analyze our results in the next subsection.
+In the last subsection we provide several examples of the effect of answering a critical question.
+In the next section, we develop a formal language for our argument schemes, critical questions, and the effect of answering them.
 
@@
 \hline
-AS5 & goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
+AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
 & & CQ5b & Does the goal decompose into any other tasks?& INTRO\\
@@
 \hline
-AS12 & Element $IE$ is relevant & CQ12a & Is the element relevant? & DISABLE\\
-&  & CQ12b & Is the element redundant? & DISABLE\\
+AS12 & Element $IE$ is relevant & CQ12 & Is the element relevant/useful? & DISABLE\\
 \hline
-AS13 & Element $IE$ has name $n$ & CQ12a & Is the name clear? & REPLACE\\
-&  & CQ12b & Is the name ambiguous? & REPLACE\\
+AS13 & Element $IE$ has name $n$ & CQ13 & Is the name clear/unambiguous? & REPLACE\\
+\hline
+\hline
+- & - & Att & Generic counterargument & ATTACK\\
 \hline
@@
 Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
+The assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
 The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
@@
 If we found arguments or critical questions that did not appear in the original list, we added them and counted them as well.
-Argument schemes that did not appear were removed from the list.
+Argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in section~\ref{sect:gmas:transcripts:analysis}).
 Most of the occurrences were not literally found back, but had to be inferred from the context.
@@
 
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
+\todo{Marc}{Floris}{I'd like to say something in the next paragraph about the difficulty of extracting arguments from text. Do you know any references?}
+It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.
+Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
 Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
@@
 
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
+\begin{figure*}[h]
+\includegraphics[width=\textwidth]{img/transcript_grl}
+\caption{The GRL model manually constructed from transcript 1.
+Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
+Elements and relationships with no dot have been inferred by us.}
+\label{fig:transcripts:grl}
+\end{figure*}
+
+We found a total of 159 instantiations of the argument schemes AS0-AS11 in the transcripts.
 The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
+A large portion (about 60\%) of the argument schemes we found involved discussions around tasks of the information system (AS2, AS10).
 
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
+We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing about the relevance of it (CQ12, CQ13).
 
@@
 \hline
-CQ5b & Alternative ways to realize goal? & 1 & 0 & 0 & 1\\
+CQ5b & Goes decomposes into other tasks? & 1 & 0 & 0 & 1\\
 \hline
@@
 \hline
-CQ12 & Useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
+CQ12 & Is the element relevant/useful? & 2 & 3 & 2 &7\\
 \hline
-CQ13 & Clarifying an element &3 &10 & 3 & 16\\
+CQ13 & Is the name clear/unambiguous? &3 &10 & 3 & 16\\
 \hline
 \hline
-\multicolumn{2}{|c|}{\textbf{TOTAL}}&69&80&69&218\\
+- & Generic counterargument & 0& 2 & 2 & 4\\
+\hline
+\hline
+\multicolumn{2}{|c|}{\textbf{TOTAL}}&69&80&69&222\\
 \hline
@@
 
-We found that answering a critical questions can have three effects on the original argument:
+For each transcript, we manually create a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
+An example of such a model is shown in figure~\ref{fig:transcripts:grl}.
+We added green and red dots to various elements and relationships in the figure.
+A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
+Note that if the underlying argument is rejected, the corresponding GRL element has been disabled.
+Some elements do not have a corresponding green or red dot.
+In that case, we have inferred the elements from the discussion, but we could not explicitly find back arguments for it.
+
+We found that answering a critical questions can have four different effects on the original argument:
 \begin{itemize}
-\item \emph{INTRO}: Introduce a new element with a corresponding argument. This operation does not attack the original argument, but rather creates a new argument.
-\item \emph{DISABLE:} Disable the element of the argument scheme to which the critical questions pertains.
+\item \emph{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
+This operation does not attack the original argument of the critical question, but rather creates a new argument.
+For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler} and \emph{Find agreeable date by talking to initiator}'' (figure~\ref{fig:SchedulerModel}).
+Suppose now the critical question CQ5b: ``Does the goal \emph{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \emph{Find agreeable date by contacting secretary}''.
+This results in a new instantiation of AS5, namely: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler}, \emph{Find agreeable date by talking to initiator}, and \emph{Find agreeable date by contacting secretary}.
+As a result, the goal model will contain the corresponding task \emph{Find agreeable date by contacting secretary}, as well as an OR-decomposition relation from the goal \emph{Agreeable meeting dates} to that task.
+\item \emph{DISABLE:} Disable the element or relationship of the argument scheme to which the critical questions pertains.
 This operation does not create a new argument, but only disables (i.e., attacks) the original one.
-\item \emph{REPLACE:} Replace the element of the argument scheme with a new element. This operation both introduces a new argument and attacks the original one.
+For instance, suppose argument scheme AS0 is instantiated with: ``Actor \emph{meeting initiator} is relevant'' (figure~\ref{fig:SchedulerModel}).
+This argument can be attack with critical question CQ0: ``Actor \emph{meeting initiator} is not relevant''.
+As a result, the argument for the actor is attack, and actor \emph{meeting initiator} is disabled in the goal model.
+\item \emph{REPLACE:} Replace the element of the argument scheme with a new element.
+This operation both introduces a new argument and attacks the original one.
+For instance, suppose argument scheme AS2 is instantiated wit: ``Actor \emph{meeting initiator} can perform task \emph{schedule meeting}.
+This argument can be attacked with critical question CQ13: ``The task \emph{schedule meeting} is unclear, it should be \emph{schedule board meeting}''.
+This results in replacing the original argument with the new argument ``Actor \emph{meeting initiator} can perform task \emph{schedule board meeting}.
+In the goal model, the description of the task should change accordingly.
+\item \emph{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
+For instance, suppose argument scheme AS4 is instantiated as follows: ``Actor \emph{meeting initiator} has goal \emph{quick}.
+For instance, suppose argument scheme AS0 is instantiated with ``Actor \emph{meeting scheduler} is relevant''.
+Suppose this argument is attacked with critical question CQ0: ``Actor \emph{meeting initiator} is not relevant'', because the participants decide to fully automate the meeting scheduling process.
+However, this critical question may in turn be attacked again, if new evidence comes up.
+For instance, it may turn out that it is sometimes more desirable to schedule meeting by hand.
+In this case, the generic counter argument ``Actor \emph{meeting initiator} is relevant, because it is sometimes easier to schedule meetings manually'' attack the argument ``Actor \emph{meeting scheduler} is not relevant''.
+As a result, the original argument ``Actor \emph{meeting scheduler} is relevant'' is accepted again, and as a result is shown in the goal model.
 \end{itemize}
 
+In section~\ref{sect:gmas:examples} we provide examples we found in the transcripts for all of these four effects.
+
 \subsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of answering a critical question affirmatively.
+\label{sect:gmas:transcripts:analysis}
 
@@
 \subsection{Examples}
-
-For each transcript, we manually create a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
-An example of such a model is shown in figure~\ref{fig:transcripts:grl}.
-We added green and red dots to various elements and relationships in the figure.
-A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
-Note that if the underlying argument is rejected, the corresponding GRL element has been disabled.
-Some elements do not have a corresponding green or red dot.
-In that case, we have inferred the elements from the discussion, but we could not explicitly find back arguments for it.
+\label{sect:gmas:examples}
 
 We now discuss various instantiations of argument schemes and the result of answering critical questions in more detail.
-The examples below come from all three transcripts, and therefore not always correspond to the goal model in figure~\ref{fig:transcripts:grl}.
 For each example we provide transcript excerpts, a visualization of the arguments, and the corresponding goal model elements.
-
-\subsubsection{Irrelevant actor}
-
-The transcript excerpt of the first example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix.
-It consists of participant P1 putting forth the suggestion to include actor \emph{Development team} into the model.
-This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.
-
-We formalized the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}.
-This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
-This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
-This is shown in figure~\ref{fig:examples:relevant-actor:arguments}.
-
-Argument scheme AS0 is an argument to include actor \emph{Development team} into the GRL model.
-Therefore, by attacking it the argument is no longer accepted, and as a result the actor is disabled in the resulting GRL model (figure~\ref{fig:examples:relevant-actor:grl}).
-Note we added a red traceability link from the rejected argument to the disabled GRL element.
+We provide a legend for our visualization notation in figure~\ref{fig:legend}.
 
@@
 \centering
-\begin{subfigure}[b]{0.22\textwidth}
+\begin{tikzpicture}
+        \node (att1) [argNodeIN] at (-5,0) {$A_1$};
+        \node (att2) [argNodeOUT] at (-2,0) {$A_2$};
+        \node (attDescription) [draw=none, align=center] at (0.5,0) {Argument $A_1$ (accepted)\\ attacks argument $A_2$\\ (rejected)};
+        \node[traceNodeGREEN] (trace1a) at (-5,-1.5) {};
+        \node[traceNodeGREEN] (trace1b) at (-2,-1.5) {};
+        \node (traceDescription) [draw=none,align=center] at (0.5,-1.5) {Traceability link from\\ accepted argument to\\ enabled GRL element};
+        \node[traceNodeRED] (trace2a) at (-5,-3) {};
+        \node[traceNodeRED] (trace2b) at (-2,-3) {};
+        \node (traceDescription) [draw=none,align=center] at (0.5,-3) {Traceability link from\\ rejected argument to\\ disabled GRL element};
+        \node (cq1) [argNodeIN] at (-5,-4.5) {$A_1$};
+        \node (cq2) [argNodeIN] at (-2,-4.5) {$A_2$};
+        \node (cqDescription) [draw=none,align=center] at (0.5,-4.5) {Argument $A_2$ is the result of\\ answering critical question\\ $CQ$ of argument $A_1$};
+         \path
+    (att1) edge [attackLink] (att2)
+    (cq1) edge [CQLink] node [above,draw=none] {CQ} (cq2)
+    (trace1a) edge[traceLinkGREEN] (trace1b)
+    (trace2a) edge[traceLinkRED] (trace2b);
+\end{tikzpicture}
+\caption{Legend of the various elements and relationships we use for the examples in this article.}
+\label{fig:legend}
+\end{figure}
+
+\subsubsection{Example 1: Disable task ``Traffic light''}
+
+The transcript excerpt of this example is shown in table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$.
+In this example, participants are summing up functionality of the traffic simulator, which are tasks that the student can perform in the simulator.
+All these tasks can be formalized are instantiations of argument scheme AS2: ``Actor \emph{Student} has task $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
+
+Once all these tasks are summed up, participant P1 notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \emph{Add traffic light} is not useful.
+We formalized this using the critical question CQ12: ``Is task \emph{Add traffic light} useful/relevant?''.
+
+We visualized the some of the argument schemes, critical questions, and traceability links with the GRL model in figure~\ref{fig:examples:traffic-light}.
+On the left side of the image, we see three of the instantiated argument schemes AS2.
+The bottom one, ``Actor \emph{Student} has task \emph{Add traffic light}'', is attack by another argument generated from applying critical question CQ12: ``\emph{Add traffic light} is useless (\emph{All intersections have traffic lights}).
+As a result, the corresponding GRL task is disabled.
+The other two tasks are enabled and have green traceability links.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeIN] at (-2,0) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Save map}}
+        };
+        \node (a1) [argNodeIN] at (-2,-2) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Add road}}
+        };
+        \node (a2) [argNodeOUT] at (-2,-4) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Add traffic light}}
+        };
+        \node (a3) [argNodeIN] at (-2,-6){
+        	\argtext{}{\emph{Add traffic light} is useless (\emph{All intersections have traffic lights})}
+        };
+        \node[grl] (task1) at (2.3,0) { \includegraphics[scale=0.5]{img/task_save_map} };
+        \node[grl] (task2) at (2.3,-2) { \includegraphics[scale=0.5]{img/task_add_road} };
+        \node[grl, disabled] (task3) at (2.3,-4) { \includegraphics[scale=0.5]{img/task_add_traffic_light} };
+        \node[traceNodeGREEN] (trace1a) at (-.4,-.3) {};
+        \node[traceNodeGREEN] (trace1b) at (2.2,-.3) {};
+        \node[traceNodeGREEN] (trace2a) at (-.4,-2.3) {};
+        \node[traceNodeGREEN] (trace2b) at (2.2,-2.3) {};
+        \node[traceNodeRED] (trace3a) at (-.4,-4.3) {};
+        \node[traceNodeRED] (trace3b) at (2.2,-4.3) {};
+         \path
+    (a3) edge [attackLink] (a2)
+    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ12} (a3)
+    (trace1a) edge[traceLinkGREEN] (trace1b)
+    (trace2a) edge[traceLinkGREEN] (trace2b)
+    (trace3a) edge[traceLinkRED] (trace3b);
+\end{tikzpicture}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted lines) for the traffic light example.}
+\label{fig:examples:traffic-light}
+\end{figure}
+
+\subsubsection{Example 2: Clarify task ``Road pattern''}
+
+The transcript excerpt of the second example is shown in table~\ref{table:transcript:task-clarification} in the appendix and comes from transcript $t_3$.
+It consists of a number of clarification steps, resulting in the task \emph{Choose a road pattern}.
+
+\begin{figure}[ht!]
 \centering
         \begin{tikzpicture}[->]
-        \node (a0) [shape=rectangle, draw=black!60, text=black!40] at (0,0) {
-            \begin{tabularx}{\textwidth}{X}
-                \textbf{(AS0)} Development team is relevant 
-            \end{tabularx}
+        \node (a0) [argNodeOUT] at (-2,0) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Create road}}
+        };
+        \node (a1) [argNodeOUT] at (-2,-2) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a pattern}}
+        };
+        \node (a2) [argNodeOUT] at (-2,-4.2) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a pattern preference}}
+        };
+        \node (a3) [argNodeIN] at (-2,-6.7) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a road pattern}}
+        };
+        \node[grl] (actor) at (2.3,-4) { 
+        	\includegraphics[scale=0.5]{img/task_choose_road_pattern} 
+        };
+        \node[traceNodeGREEN] (trace1) at (-.5,-7.1) {};
+        \node[traceNodeGREEN] (trace2) at (2.2,-4.4) {};
+
+\begin{pgfonlayer}{background}
+         \path
+    (a1) edge[attackLink] (a0)
+    (a2) edge[attackLink] (a1)
+    (a2) edge[attackLink, bend right=20] (a0)
+    (a3) edge[attackLink] (a2)
+    (a3) edge[attackLink, bend right=20] (a1)
+    (a3) edge[attackLink, bend right=40] (a0);
+\end{pgfonlayer}
+
+	\path
+	(a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a1)
+    (a1) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a2)
+    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a3)
+    (trace1) edge[traceLinkGREEN] (trace2);
+\end{tikzpicture}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the road pattern example.}
+\label{fig:examples:clarification}
+\end{figure}
+
+The formalized argument schemes and critical questions are shown in figure~\ref{fig:examples:clarification}.
+The discussion starts with the first instantiation of argument scheme AS2: ``Actor \emph{Student} has task \emph{Create road}''.
+This argument is then challenged with critical question CQ12: ``Is the task \emph{Create road} clear?''.
+Answering this question results in a new instantiation of argument scheme AS2: ``Actor \emph{Student} has task \emph{Choose a pattern}''.
+This process is repeated two more times, resulting in the final argument ``Actor \emph{Student} has task \emph{Choose a road pattern}''.
+This final argument is unattacked and has a corresponding intentional element (right image).
+
+What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element.
+For instance, the final argument on the bottom of figure~\ref{fig:examples:clarification} attacks all previous arguments.
+If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.
+
+\subsubsection{Example 3: Decompose goal ``Simulate''}
+
+The transcript excerpt of this example is shown in table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
+It consists of a discussion about the type of decomposition relationship for the goal \emph{Simulate}.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}[->]
+        \node (a0) [argNodeIN] at (-2,0) {
+        	\argtext{AS3}{Actor \emph{System} has goal \emph{Simulate}}
+        };
+        \node (a1) [argNodeIN] at (-2,-1.5) {
+        	\argtext{AS2}{Actor \emph{System} has task \emph{Dynamic simulation}}
+        };
+        \node (a2) [argNodeIN] at (-2,-3) {
+        	\argtext{AS2}{Actor \emph{System} has task \emph{Static simulation}}
+        };
+        \node (a3) [argNodeOUT] at (-2,-5) {
+        	\argtext{AS5}{Goal \emph{Simulate} AND-decomposes into \emph{Static simulation} and \emph{Dynamic simulation}}
+        };
+        \node (a4) [argNodeIN] at (-2,-7.5) {
+        	\argtext{AS5}{Goal \emph{Simulate} OR-decomposes into \emph{Static simulation} and \emph{Dynamic simulation}}
+        };
+        \node[grl] (actor) at (2.3,-4) { 
+        	\includegraphics[scale=0.5]{img/simulate_decomposition} 
+        };
+        \node[traceNodeGREEN] (trace1a) at (-.4,-.3) {};
+        \node[traceNodeGREEN] (trace1b) at (2.5,-3.1) {};
+        \node[traceNodeGREEN] (trace2a) at (-.4,-1.8) {};
+        \node[traceNodeGREEN] (trace2b) at (3.5,-5.6) {};
+        \node[traceNodeGREEN] (trace3a) at (-.4,-3.3) {};
+        \node[traceNodeGREEN] (trace3b) at (1.3,-5.6) {};
+        \node[traceNodeGREEN] (trace4a) at (-.4,-8.2) {};
+        \node[traceNodeGREEN] (trace4b) at (2.5,-3.9) {};
+
+\begin{pgfonlayer}{background}
+         \path
+    (a4) edge[attackLink] (a3);
+\end{pgfonlayer}
+
+	\path
+	(a3) edge [CQLink, bend right=50] node [left,draw=none] {CQ10b} (a4)
+    (trace1a) edge[traceLinkGREEN] (trace1b)
+    (trace2a) edge[traceLinkGREEN] (trace2b)
+    (trace3a) edge[traceLinkGREEN] (trace3b)
+    (trace4a) edge[traceLinkGREEN, bend right] (trace4b);
+\end{tikzpicture}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the goal decomposition example.}
+\label{fig:examples:decomposition}
+\end{figure}
+
+The visualization of this discussion is shown in figure~\ref{fig:examples:decomposition}.
+Each GRL element on the right has a corresponding argument on the left.
+Moreover, the original argument for an AND-decomposition is attacked by the argument for the OR decomposition, and the new argument is linked to the decomposition relation in the GRL model.
+
+\subsubsection{Example 4: Reinstate actor ``Development team''}
+
+The transcript excerpt of this example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
+It consists of two parts: first participant P1 puts forth the suggestion to include actor \emph{Development team} into the model.
+This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.
+However, in the second part, participant P2 argue that the development team should be considered, since the professor doesn't develop the software.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeOUT] at (-2,0) {
+        	\argtext{AS0}{Development team is relevant}
         } ;
-        \node (a1) [shape=rectangle,draw] at (0,-2){
-            \begin{tabularx}{\textwidth}{X}
-                \textbf{(CQ0)} Development team is not relevant (\emph{The professor makes the software})
-            \end{tabularx}
+        \node (a1) [argNodeIN] at (-2,-2.5){
+        	\argtext{}{Development team is not relevant (\emph{The professor makes the software})}
         };
-        \node[circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt] (trace) at (1.2,-0.3) {};
+        \node[grl, disabled] (actor) at (2.3,-1) { \includegraphics[scale=0.5]{img/actor_development_team} };
+        \node[traceNodeRED] (trace1) at (-.5,-.3) {};
+        \node[traceNodeRED] (trace2) at (2.2,-0.1) {};
 
-         \path[line width=1mm]
-    (a1) edge[draw=black] (a0);
+         \path
+    (a1) edge[attackLink] (a0)
+    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
+    (trace1) edge[traceLinkRED] (trace2);
 \end{tikzpicture}
-        \caption{Arguments}
-        \label{fig:examples:relevant-actor:arguments}
-    \end{subfigure}\hfill
-    \begin{subfigure}[b]{0.22\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{img/actor_disabled}
-        \caption{GRL elements}
-        \label{fig:examples:relevant-actor:grl}
-    \end{subfigure}
-\caption{Argument schemes and critical questions (left) and GRL model (right) of a discussion about the relevance of actor Development team.}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
 \label{fig:examples:relevant-actor}
@@
 
-We see from this example that answering critical question CQ0 affirmatively has the effect of disabling the corresponding actor \emph{Development team}.
-Therefore, critical question CQ0 has the associated effect DISABLE in table~\ref{table:argument-schemes}.
+We formalize this using a \emph{generic counterargument}, attacking the critical question.
+The first part of the discussion is shown in figure~\ref{fig:examples:relevant-actor}.
+We formalize the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}.
+This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
+This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
+This is shown in figure~\ref{fig:examples:relevant-actor}, left image.
 
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{img/transcript_grl}
-\caption{The GRL model manually constructed from transcript 1.
-Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
-Elements and relationships with no dot have been inferred by us.}
-\label{fig:transcripts:grl}
-\end{figure*}
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeIN] at (-2,0) {
+        	\argtext{AS0}{Development team is relevant}
+        };
+        \node (a1) [argNodeOUT] at (-2,-2.5){
+        	\argtext{}{Development team is not relevant (\emph{The professor makes the software})}
+        };
+        \node (a2) [argNodeIN] at (-2,-5) {
+        	\argtext{}{The professor doesn't develop the software}
+        } ;
+        \node[grl] (actor) at (2.3,-1) { \includegraphics[scale=0.5]{img/actor_development_team} };
+        \node[traceNodeGREEN] (trace1) at (-.5,-.3) {};
+        \node[traceNodeGREEN] (trace2) at (2.2,-0.1) {};
+
+         \path
+    (a1) edge[attackLink] (a0)
+    (a2) edge[attackLink] (a1)
+    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
+    (a1) edge [CQLink, bend right=50] node [left,draw=none] {Att} (a2)
+    (trace1) edge[traceLinkGREEN] (trace2);
+\end{tikzpicture}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
+\label{fig:examples:relevant-actor2}
+\end{figure}
+
+Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counter argument has been put forward.
+The argument ``The professor doesn't develop the software'' now attacks the argument ``\emph{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\emph{Development team} is relevant''.
+As a result, the first and the last argument are both accepted, which causes the actor in the GRL model to be enabled again.

### 2017-01-22 main.tex
update
GMAS finished

@@
 \usepackage{tabularx}
-\usepackage{tikz}\usetikzlibrary{arrows}
+\usepackage{tikz}\usetikzlibrary{arrows, backgrounds}
 \usepackage{times}
@@
 \tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
+
+% Tikz stuff for arguments with GRL
+\tikzset{
+  argNodeIN/.style={shape=rectangle,draw,fill=white,align=center},
+  argNodeOUT/.style={shape=rectangle, draw=black!60, text=black!40,fill=white},
+  traceNodeRED/.style={circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt},
+  traceNodeGREEN/.style={circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt},
+  attackLink/.style={draw=black,->,line width=1mm},
+  traceLinkRED/.style={draw, densely dotted,-,line width=0.5mm, draw=red},
+  traceLinkGREEN/.style={draw, densely dotted,-,line width=0.5mm, draw=green},
+  CQLink/.style={line width=0.5mm, dotted,->},
+  ldots/.style={draw=none,rotate=90,font=\fontsize{20}{22.4}\selectfont},
+  grl/.style={draw=none},
+  disabled/.style={opacity=0.5}
+}
+\newcommand{\argtext}[2]{
+\begin{tabularx}{0.2\textwidth}{X}
+\ifx&#1&%
+\else
+\textbf{(#1)}
+\fi
+#2
+\end{tabularx}
+}
 

### 2017-01-22 transcripts.tex
GMAS finished

@@
 \hline
-Respondent & Text & Annotation\\
+\textbf{Respondent} & \textbf{Text} & \textbf{Annotation}\\
 \hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
+0:15:11.2 (P1) & And then, we have a set of actions.
+Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map", ``add intersection", ``remove intersection", ``add road", ``add traffic light"}\\
+\cline{1-2}
+0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights &\\
 \hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
+0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{60mm}{\textbf{[21 critical question CQ12 for 20]} Is the task ``Add traffic light" useful/relevant?\newline
+\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.}\\
+\cline{1-2}
+0:15:44.9 (P2) & Yeah &\\
+\cline{1-2}
+0:15:45.2 (P1) & It's, you don't have to specifically add a traffic light because if you have &\\
+\cline{1-2}
+0:15:51.4 (P2)	& They need-&\\
 \hline
 \end{tabular}
-\caption{Discussion about the relevance of an actor (transcript $t_3$)}
-\label{table:transcript:irrelevant-actor}
+\caption{Adding tasks, disabling useless task ``Add traffic light'' (transcript $t_1$)}
+\label{table:transcripts:traffic-light}
 
@@
 \label{table:transcript:task-clarification}
+
+\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
+\hline
+Respondent & Text & Annotation\\
+\hline
+0:18:55.7 (P1) &Yeah. And then two processes, static, dynamic and they belong to the goal simulate. & \textbf{[17 goal (AS3)]} Actor ``System'' has goal ``Simulate"\newline
+\textbf{[18 task (AS2)]} Actor ``System" has task ``Static simulation"\newline
+\textbf{[19 task (AS2)]} Actor ``System has task ``Dynamic simulation"\newline
+\textbf{[20 decomposition (AS?)]} Goal ``Simulation" AND-decomposes into "Static simulation" and ``Dynamic simulation"\\
+\hline
+0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{5}{80mm}{\textbf{[26 critical question CQ10b for 20]} Is the decomposition type of ``simulate'' correct?\newline
+\textbf{[27 answer to 26]} No, it should be an OR decomposition.}\\
+\cline{1-2}
+0:30:12.6 (P2) & That's and OR &\\
+\cline{1-2}
+0:30:14.3 (P3) & I think it's an OR &\\
+\cline{1-2}
+0:30:15.4 (P1) & It's for the data, it's an OR &\\
+\cline{1-2}
+0:30:18.1 (P3) & Yep &\\
+\hline	
+\end{tabular}
+\caption{Incorrect decomposition type for goal \emph{Simulate} (transcript $t_3$)}
+\label{table:transcript:decomposition}
 \end{table}
 
+\begin{table}[!htbp]
+\centering
+\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
+\hline
+Respondent & Text & Annotation\\
+\hline
+0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
+\hline
+0:11:00.8 (P2) & Development team, I don't know.
+Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
+\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
+\hline
+0:18:13.4 (P2) & I think we can still do developers here.
+To the system & \multirow{4}{60mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
+\cline{1-2}
+0:18:18.2 (P1)  & Yeah?&\\
+\cline{1-2}
+0:18:19.8 (P2) & Yeah, it isn't mentioned but, the professor does-&\\
+\cline{1-2}
+0:18:22.9 (P1) & Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
+\hline
+\end{tabular}
+\caption{Discussion about the relevance of an actor (transcript $t_3$)}
+\label{table:transcript:irrelevant-actor}
+\end{table}
+
+\iffalse
 \begin{table}[!htbp]
@@
 
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
 \begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
@@
 \label{table:transcript:as0-cq0-cq_counterarg}
-
 \end{table}
+\fi

### 2017-01-23 background.tex
Small modifications

@@
 
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
+In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language used to connect to the argumentation framework.
 Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
@@
 %such as performance, cost,  security, and usability.
+
 URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.

### 2017-01-23 formalframework.tex
formal framework

@@
-\section{The Formal Framework}
+\section{The Logical Framework}
 \label{sect:formalframework}
 
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
+In the previous section we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of an information system.
+The resulting list is shown in table~\ref{table:argument-schemes}.
+We also discussed various examples of applications of the different critical questions and all four different effects (right column of table~\ref{table:argument-schemes}): INTRO, DISABLE, REPLACE, and ATTACK.
 
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
+The examples and corresponding visualizations of the previous section provided some insight into how to formalize argument schemes, critical questions, and their relationship with goal models.
+However, if we are to implement our framework into a tool, we require a more precise formalisation of these concepts.
+We choose to use formal logic to specify this, because argumentation has been studied in formal logic extensively for the past decades.
 
-\subsection{Formal Language for a GRL Model}
+In the first subsection we develop a formal language to specify a GRL model.
+This language consists of propositions only, and a GRL model specified in this language can be specified as a logic program directly.
+In the second subsection we then provide formal definitions of all argument schemes.
+An instantiation of an argument scheme consists of a set of statement about a GRL model.
+In the third subsection we formalize the critical questions as algorithms.
 
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
+\subsection{Formal Language for RationalGRL}
 
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
+Basic concepts set theory are adequate to define the entities in a GRL model and the relations between them. We begin with the representation for the elements.
 
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
+\begin{definition}[RationalGRL elements]
+Each element of a goal model is identified with a non-negative integer. Formally: let $Actors\cup IEs \cup Links \subseteq \mathbb{N}$ denote the set of elements of a goal model, where 
+\begin{itemize}
+\item $Actors$ is the set of actors, 
+\item $IEs$ is the set of intentional elements (softgoals, goals, tasks, and resources),
+\item $Links$ is the set of links (decomposition, contribution, and dependency). 
+\end{itemize}
+
+The three sets are pairwise disjoint, i.e. $Actors \cap IEs = Actors \cap Links = IEs \cap Links = \emptyset$. 
 \end{definition}
 
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
+\rationale{} We identify GRL element with an identifier and not with a name, because it is possible that the name of the element changes, for instance by applying critical question CQ12a (clarification), see the example ``Clarify task \emph{Road pattern}'' above.
+The GRL model of figure~\ref{fig:examples:decomposition} can be formalized as follows:
+\begin{align*}
+&Actors = \{1\},\\
+&IEs = \{2, 3, 4\},\\ 
+&Links = \{5\}.
+\end{align*}
+
+\begin{definition}[GRL intentional element type]
+For an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
+\end{definition}
+
+\rationale{} All intentional elements are collected in the set $IEs$.
+In order to distinguish softgoals, goals, tasks, and resources, we use the corresponding propositions.
+In our example this looks as follows:
+\begin{align*}
+&goal(2),\\
+&task(3),\\
+&task(4).
+\end{align*}
+
+Note we do not require such propositions for actors, since there is only one type of actor, and all actors are contained in the set $Actors$.
+
+\begin{definition}[Links]
+Given a link $i\in Links$, we denote the links types as follows:
+\begin{itemize}
+\item $contrib(i,src,dest,type)$ is a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{+,-\}$ means a positive resp. negative contribution. 
+\item $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ is a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
+\item $dep(i,src,dest)$ is a dependency from $dest\in IEs$ to $dest\in IEs$.
+\end{itemize}
+\end{definition}
+
+\rationale{} Similarly to the intentional elements, we distinguish between the different types of links between IEs using the corresponding propositions.
+The formalisation of the link in figure~\ref{fig:examples:decomposition} looks as follows:
+\begin{align*}
+&decomp(5, 2, \{3,4\}, OR)
+\end{align*}
 
@@
 
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
+\rationale{} We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Continuing our example of formalizing figure~\ref{fig:examples:decomposition}, we have 
+\begin{align*}
+&name(1,\text{system}),&\\
+&name(2,\text{simulate}),&\\
+&name(3,\text{static\ simulation}),&\\
+&name(4, \text{dynamic simulation}).&
+\end{align*}
+Note the decomposition link with identifier 5 does not have a name, since decompositions are already distinguished by their type (OR, XOR, AND).
 
@@
 
-\todo{Marc}{Marc}{Provide example here}
+\rationale{} In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements. This looks as follows in our example:
+\begin{align*}
+&has(1, 2),\\
+&has(1, 3),\\
+&has(1, 4),\\
+&has(1, 5).
+\end{align*}
+
+\begin{definition}[Disabled GRL element]
+A disabled element $i\in Actors\cup IEs$ is denoted by $disabled(i)$.
+\end{definition}
+
+\rationale{} If a GRL element corresponds to an argument that is rejected, then this argument should be disabled.
+For instance, the GRL model in figure~\ref{fig:examples:relevant-actor} can be formalized as follows:
+\begin{align*}
+&Actors = \{1\},\\
+&name(1, \text{development team}),\\
+&disabled(1)
+\end{align*}
+
+This concludes our definitions to formalize a GRL model.
+An advantage of our approach is that we can formalize a GRL model using a set of facts, which can be directly be formalized as a logic program.
+We will come back to this in the last part of this section.
 
@@
 
-In order to formalize the argument schemes we first note that an
+We formalize an argument as a set of statements about a GRL model in the language we introduced in the previous section.
+For instance, $A= \{actor(0), name(0,student)\}$ is an argument stating that identifier 0 represent an actor, and that the name of this actor is \emph{Student}.
+An argument scheme then corresponds to adding one or more arguments to the set of arguments.
+
+\begin{table*}[h]
+\centering
+\begin{tabularx}{\textwidth}{|l|l|l|}
+\hline
+\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \textbf{Formal arguments}\\
+\hline
+$AS0(a_{id},a)$ & Actor $a$ is relevant & $\{actor(id), name(id, a)\}$\\
+\hline
+$AS1(a_{id},r_{id},R$ & Actor $a_{id}$ has resource $R$ & $\{resource(r_{id}), name(r_{id},R), has(a_{id},r_{id})\}$\\
+\hline
+$AS2(a_{id},t_{id},T)$ & Actor $a_{id}$ can perform task &$T$ $\{task(t_{id}), name(t_{id},T), has(a_{id},t_{id})\}$\\
+\hline
+$AS3(a_{id},g_{id},G)$ & Actor $a_{id}$ has goal $G$ &$\{goal(g_{id}), name(g_{id},G), has(a_{id},g_{id})\}$\\
+\hline
+$AS4(a_{id},s_{id},S)$ & Actor $a$ has softgoal $S$ & $\{softgoal(s_{id}), name(s_{id},S), has(a_{id}, s_{id}\}$\\
+\hline
+\hline
+$AS5(g_{id},(t_{id_1},\ldots,t_{id_n}),(T_1,\ldots,T_n),d_{id})$ & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & $\{decomp(d_{id},g_{id},\{t_{id_1},\ldots,t_{id_n}\})\}$\\
+&&$\{task(t_{id_1}),name(t_{id_1},T_1)\}$\\
+&&...\\
+&&$\{task(t_{id_n}),name(t_{id_n},T_n)\}$\\
+\hline
+$AS6(t_{id},s_{id},$ & Task $T$ contributes to softgoal $S$& \\
+\hline
+AS7 & Goal $G$ contributes to softgoal $S$ & \\
+\hline
+AS8 & Resource $R$ contributes to task $T$ & \\
+\hline
+AS9 & Actor $a$ depends on actor $b$ & \\
+\hline
+AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & \\
+\hline
+AS11 & Task $T$ contributes negatively to softgoal $S$&\\
+\hline
+\end{tabularx}
+\caption{List of argument schemes (AS0-AS13, left column), critical questions (CQ0-CQ12, middle column), and the effect of answering them (right column).}
+\label{table:argument-schemes}
+\end{table*}
+
+We start by formalizing an argumentation framework.
+
+\begin{definition}[Argumentation framework]
+An argumentation framework $(Args,Att)$ consists of a set of arguments $Args$ and an attack relationship $Att:Args\times Args$, where $(A_1,A_2)\in Att$ means that argument $A_1\in Args$ attacks arguments $A_2\in Args$.
+\end{definition}
+
+\begin{definition}[AS0] Argument scheme $AS0$ (Actor $a$ is relevant) is: $$AS_0(i,n) = \{actor(i), name(i,n)\},$$ where $i$ is an element identifier and $n$ is a name.
+\end{definition}
+
+\rationale{} The argument scheme takes two arguments, the id $i$ and the name $n$ of an actor.
+In figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS(0,\text{Development team})$, results in the argument $\{actor(0), name(0, \text{Development team})\}$.
+
+\begin{definition}[AS1] Argument scheme $AS1$ (Actor $a$ has resource $R$) is: $$AS_1(a,r,R) = \{resource(r), name(r,R), has(a,r)\}.$$ where $a$ is an actor identifier, $r$ is an element identifier, and $R$ is a resource name.
+\end{definition}
+
+\rationale{} This argument scheme takes three arguments, the actor id $a$, the resource id $r$, and the resource name $R$.
+
+The formalisations of argument scheme AS2 to AS4 are exactly the same as AS1 but then for respectively tasks, goals, and softgoals.
+Therefore, we have omitted them here.
+We continue with argument scheme AS5.
+
+\begin{definition}[AS5] Argument scheme $AS5$ (Goal $G$ decomposes into tasks $T_1,\ldots,T_n$) is formalized as follows: $$AS_5(d, g_id,T_{id},T) = \{decomp(d, g_{id}, T_{id}, type)\},$$
+where $d$ is the decomposition link identifier, $g$ is a goal identifier, $T_{id}=(t_1,\ldots,t_n)$ is a list of task identifiers, $T=(T_1,\ldots,T_n)$ is a list of task names, and $type\in\{and,or,xor\}$ denotes the decomposition type.
+\end{definition}
+
+\rationale{} Note $AS_5$ requires the list of task names $T$ as an argument as well, while it does not show up in the argument.
+This is because when applying the argument scheme, it may be the case that the tasks does not exist yet.
+We return to this in the next subsection when we provide algorithms for applying the argument schemes and critical questions.
+Consider as an example figure~\ref{fig:examples:decomposition}.
+Suppose the goal is formalized with $AS_3(0, 1, simulate)=\{goal(1), name(1, simulate), has(0,1)\}$.
+The OR-decomposition can be formalized as an instantiation of $AS_5$ as follows: $AS_5(2, 1, (3,4), (\text{static simulation, dynamic simulation}))$.
+
+The remaining argument schemes AS6-AS11 are all very similar to AS5, so we have omitted them here.
+
+\subsection{Algorithms for applying argument schemes}
+
+In the previous subsection we formalized a static representation of argument schemes.
+In this subsection we provide algorithms for applying them.
+We keep track of an identifier $ID$, representing the currently highest identifier number.
+This number is increased with each application of an argument scheme.
+
+\begin{definition}[Application of AS0-AS4] Applying argument schemes $AS_0$ to $AS_4$ adds the new argument to the set of arguments: 
+\begin{align*}
+&ID\text{++}\\
+&Args := Args \cup AS_i(ID,\ldots)
+\end{align*}
+for $i\in[0,4]$.
+\end{definition}

### 2017-01-24 formalframework.tex
formal framework

@@
 
-\subsection{Formalizing the argument schemes}
+\subsection{Formal argumentation}
+
+In the next section, we formalize an argument as a set of statements about a GRL model in the language we introduced in the previous section.
+First, we formalize an argumentation framework and its semantics, which is used to determine sets of acceptable arguments.
+
+\begin{definition}[Argumentation framework]
+An argumentation framework $AF=(Args,Att)$ consists of a set of arguments $Args$ and an attack relationship $Att:Args\times Args$, where $(A_1,A_2)\in Att$ means that argument $A_1\in Args$ attacks arguments $A_2\in Args$.
+\end{definition}
+
+\rationale{} This is the standard definition used by Dung~\cite{Dung1995}. For instance, the argumentation framework from figure~\ref{fig:pras:example} can be formalized as:
+\begin{align*}
+&Args=\{A_1,A_2,A_3\}\\
+&Att = \{(A_2,A_1),(A_1,A_3),(A_3,A_1)\}
+\end{align*}
+
+\todo{Marc}{Marc}{explain semantics}
+
+\subsection{Algorithms for instantiating the argument schemes}
 
@@
 
-\begin{table*}[h]
-\centering
-\begin{tabularx}{\textwidth}{|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \textbf{Formal arguments}\\
-\hline
-$AS0(a_{id},a)$ & Actor $a$ is relevant & $\{actor(id), name(id, a)\}$\\
-\hline
-$AS1(a_{id},r_{id},R$ & Actor $a_{id}$ has resource $R$ & $\{resource(r_{id}), name(r_{id},R), has(a_{id},r_{id})\}$\\
-\hline
-$AS2(a_{id},t_{id},T)$ & Actor $a_{id}$ can perform task &$T$ $\{task(t_{id}), name(t_{id},T), has(a_{id},t_{id})\}$\\
-\hline
-$AS3(a_{id},g_{id},G)$ & Actor $a_{id}$ has goal $G$ &$\{goal(g_{id}), name(g_{id},G), has(a_{id},g_{id})\}$\\
-\hline
-$AS4(a_{id},s_{id},S)$ & Actor $a$ has softgoal $S$ & $\{softgoal(s_{id}), name(s_{id},S), has(a_{id}, s_{id}\}$\\
-\hline
-\hline
-$AS5(g_{id},(t_{id_1},\ldots,t_{id_n}),(T_1,\ldots,T_n),d_{id})$ & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & $\{decomp(d_{id},g_{id},\{t_{id_1},\ldots,t_{id_n}\})\}$\\
-&&$\{task(t_{id_1}),name(t_{id_1},T_1)\}$\\
-&&...\\
-&&$\{task(t_{id_n}),name(t_{id_n},T_n)\}$\\
-\hline
-$AS6(t_{id},s_{id},$ & Task $T$ contributes to softgoal $S$& \\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & \\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & \\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & \\
-\hline
-AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & \\
-\hline
-AS11 & Task $T$ contributes negatively to softgoal $S$&\\
-\hline
-\end{tabularx}
-\caption{List of argument schemes (AS0-AS13, left column), critical questions (CQ0-CQ12, middle column), and the effect of answering them (right column).}
-\label{table:argument-schemes}
-\end{table*}
+In the following algorithms, we assume the following global variables:
+\begin{itemize}
+\item $id$ the current highest identifier of the elements. This variable is increased for each new element that is added.
+\item $Args$ contains the set of all arguments.
+\item $Att$ contains the set of all attack relations.
+\end{itemize}
 
-We start by formalizing an argumentation framework.
+\begin{algorithm}[h]
+  \caption{Applying AS0: Actor $a$ is relevant}\label{alg:as0}
+  \begin{algorithmic}[1]
+    \Procedure{$AS_0$}{$a$}
+    \State $id\gets id+1$
+    \State $A \gets \{actor(id), name(id,a)\}$
+    \State $Args \gets Args \cup A$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
 
-\begin{definition}[Argumentation framework]
-An argumentation framework $(Args,Att)$ consists of a set of arguments $Args$ and an attack relationship $Att:Args\times Args$, where $(A_1,A_2)\in Att$ means that argument $A_1\in Args$ attacks arguments $A_2\in Args$.
-\end{definition}
+\rationale{} Algorithm~\ref{alg:as0} takes one argument, namely the name of the actor $a$.
+In figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_(\text{Development team})$, results in one argument: $Args = \{\{actor(0), name(0, \text{Development team})\}\}$.
 
-\begin{definition}[AS0] Argument scheme $AS0$ (Actor $a$ is relevant) is: $$AS_0(i,n) = \{actor(i), name(i,n)\},$$ where $i$ is an element identifier and $n$ is a name.
-\end{definition}
+\begin{algorithm}[h]
+  \caption{Applying AS1: Actor $a_{id}$ has resource $n$}\label{alg:as1}
+  \begin{algorithmic}[1]
+    \Procedure{$AS_1$}{$a_{id}, n$}
+    \State $id\gets id+1$
+    \State $A\gets \{resource(id), name(id,n), has(a_{id},id)\}$
+    \State $Args \gets Args\cup A$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
 
-\rationale{} The argument scheme takes two arguments, the id $i$ and the name $n$ of an actor.
-In figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS(0,\text{Development team})$, results in the argument $\{actor(0), name(0, \text{Development team})\}$.
-
-\begin{definition}[AS1] Argument scheme $AS1$ (Actor $a$ has resource $R$) is: $$AS_1(a,r,R) = \{resource(r), name(r,R), has(a,r)\}.$$ where $a$ is an actor identifier, $r$ is an element identifier, and $R$ is a resource name.
-\end{definition}
-
-\rationale{} This argument scheme takes three arguments, the actor id $a$, the resource id $r$, and the resource name $R$.
+\rationale{}  This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
 
@@
 
-\begin{definition}[AS5] Argument scheme $AS5$ (Goal $G$ decomposes into tasks $T_1,\ldots,T_n$) is formalized as follows: $$AS_5(d, g_id,T_{id},T) = \{decomp(d, g_{id}, T_{id}, type)\},$$
-where $d$ is the decomposition link identifier, $g$ is a goal identifier, $T_{id}=(t_1,\ldots,t_n)$ is a list of task identifiers, $T=(T_1,\ldots,T_n)$ is a list of task names, and $type\in\{and,or,xor\}$ denotes the decomposition type.
-\end{definition}
+\begin{algorithm}[h]
+  \caption{Applying AS5: Goal $g_{id}$ decomposes into tasks $T_1,\ldots,T_n$}\label{alg:as5}
+  \begin{algorithmic}[1]
+    \Procedure{$AS_5$}{$g_{id}, \{T_1,\ldots,T_n\}, type$}
+    \State $T_{id} = \emptyset$
+    \For{$T_i$ in $\{T_1,\ldots,T_n\}$}
+      \If{$\exists_{A\in Args} \{task(t_{id}),name(t_{id},T_i)\}\subseteq A$}
+        \State $T_{id} \gets T_{id} \cup \{t_{id}\}$
+      \Else
+        \State $id\gets id+1$
+        \State $A \gets \{task(id),name(id,T_i)\}$
+        \State $Args \gets Args\cup A$
+        \State $T_{id} \gets T_{id} \cup \{id\}$
+      \EndIf
+    \EndFor
+    \State $id\gets id+1$
+    \State $A\gets \{decomp(id, g_{id}, T_{id}, type)\}$
+    \State $Args \gets Args\cup A$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
 
-\rationale{} Note $AS_5$ requires the list of task names $T$ as an argument as well, while it does not show up in the argument.
-This is because when applying the argument scheme, it may be the case that the tasks does not exist yet.
-We return to this in the next subsection when we provide algorithms for applying the argument schemes and critical questions.
-Consider as an example figure~\ref{fig:examples:decomposition}.
-Suppose the goal is formalized with $AS_3(0, 1, simulate)=\{goal(1), name(1, simulate), has(0,1)\}$.
-The OR-decomposition can be formalized as an instantiation of $AS_5$ as follows: $AS_5(2, 1, (3,4), (\text{static simulation, dynamic simulation}))$.
+\rationale{} The procedure in algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
+The for loop in the algorithm iterates over all tasks in list $T$, and if a task does not exist yet, an argument is added for it.
+This ensures that all tasks have corresponding arguments.
+Note that the algorithm assumes the goal already exists.
 
-The remaining argument schemes AS6-AS11 are all very similar to AS5, so we have omitted them here.
+\begin{algorithm}[h]
+  \caption{Applying AS6: Task $t_{id}$ contributes to softgoal $s$}\label{alg:as6}
+  \begin{algorithmic}[1]
+    \Procedure{$AS_6$}{$t_{id}, s$}
+    \If{$\exists_{A\in Args} \{softgoal(i),name(i,s)\}\subseteq A$}
+        \State $s_{id} \gets i$
+    \Else
+      \State $id\gets id+1$
+      \State $A \gets \{softgoal(id),name(id,t)\}$
+      \State $Args \gets Args\cup A$
+      \State $s_{id} \gets id$
+    \EndIf
+    \State $id\gets id+1$
+    \State $A\gets \{contr(id, t_{id}, s_{id}, pos)\}$
+    \State $Args \gets Args\cup A$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
 
-\subsection{Algorithms for applying argument schemes}
+\rationale{} The procedure in algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
+The if statements check if the softgoal exists already, and if not, an argument is added for it.
+This ensures that all softgoals have corresponding arguments.
+Note that the algorithm assumes the task already exists.
 
-In the previous subsection we formalized a static representation of argument schemes.
-In this subsection we provide algorithms for applying them.
-We keep track of an identifier $ID$, representing the currently highest identifier number.
-This number is increased with each application of an argument scheme.
+The algorithms for $AS7$ to $AS11$ all have a very similar structure as algorithm~\ref{alg:as6} and have therefore been omitted from this article.
 
-\begin{definition}[Application of AS0-AS4] Applying argument schemes $AS_0$ to $AS_4$ adds the new argument to the set of arguments: 
-\begin{align*}
-&ID\text{++}\\
-&Args := Args \cup AS_i(ID,\ldots)
-\end{align*}
-for $i\in[0,4]$.
-\end{definition}
+\subsection{Algorithms for answering the critical questions}
+
+We now develop algorithms for our critical questions.
+We start with the critical questions that have as effect DISABLE.
+All critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12 fall into this category.
+The algorithm below applies when the critical question is answered affirmatively.
+
+\begin{algorithm}[h]
+  \caption{Applying DISABLE: Element $i$ is disabled}\label{alg:disable}
+  \begin{algorithmic}[1]
+    \Procedure{DISABLE}{$i$}
+    \State $id\gets id+1$
+    \State $A\gets \{disabled(i))\}$
+    \State $Args \gets Args\cup A$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\rationale{} We the see disable operation is quite straightforward. It simply consists of adding an argument stating argument $i$ is disabled.
+
+Next we consider the REPLACE operation. This operation is less straightforward and cannot be captured by a single algorithm, since the operation depends on the type of argument scheme.
+
+\begin{algorithm}[h]
+  \caption{Answering CQ5b: ``Does goal $G$ decompose into any other tasks?'' With: ``Yes, namey into tasks $t_{n+1},\ldots,t_k$''}\label{alg:cq5b}
+  \begin{algorithmic}[1]
+    \Procedure{CQ5b}{$g_{id},\{i_1,\ldots,i_n\}, type, \{t_1,\ldots,t_k\}$}
+    \State $T_{id} = \{i_1,\ldots,i_n\}$
+    \For{$t_i$ in $\{t_1,\ldots,t_k\}$}\label{alg:cq5b:for1}
+      \If{$\exists_{A\in Args} \{task(t_{id}),name(t_{id},t_i)\}\subseteq A$}
+        \State $T_{id} \gets T_{id} \cup \{t_{id}\}$
+      \Else
+        \State $id\gets id+1$
+        \State $A \gets \{task(id),name(id,t_i)\}$
+        \State $Args \gets Args\cup A$
+        \State $T_{id} \gets T_{id} \cup \{id\}$
+      \EndIf
+    \EndFor
+    \State $id\gets id+1$
+    \State $A_{new} = \{decomp(id, g_{id}, T_{id}, type)\}$\label{alg:cq5b:newarg}
+    \For{$A$ in $\{decomp(d,g_{id},x,y)\}\subseteq A\mid A\in Args\}$}\label{alg:cq5b:for2}
+      \State $Att\gets Att \cup \{(A_{new},A)\}$
+    \EndFor
+    \State $Args \gets Args \cup \{A_{new}\}$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\rationale{} Algorithm~\ref{alg:cq5b} is executed when critical question CQ5b is answered, which is a critical question for argument scheme AS5.
+Therefore, it assumes an argument for a goal decomposition of the form $decomp(d, g_{id}, \{i_1,\ldots,i_n\},type)$ as already been generated (see algorithm~\ref{alg:as5}).
+The goal of the algorithm is to generate a new argument of the form $decomp(d, g_{id}, \{i_1,\ldots,i_5\}\cup \{j_1,\ldots,j_k\}, type)$, where $\{j_1,\ldots,j_k\}$ are the identifiers of the additional decomposing tasks $\{t_1,\ldots,t_k\}$.
+The first for loop in line~\ref{alg:cq5b:for1} check whether any of these tasks already exist, and if not, adds corresponding arguments for them.
+The list of all decomposing tasks is stored in $T_{id}$.
+The new argument for the decomposition link is then created in line~\ref{alg:cq5b:newarg}.
+Finally, the last for loop on line~\ref{alg:cq5b:for2} ensures that all previous arguments for decomposition links are attacked by the new argument.
+This ensures that only the new decomposition link will be accepted.
+
+The REPLACE operations of CQ10a and CQ10b have a very similar structure as algorithm~\ref{alg:cq5b} and have therefore been omitted. We continue with the last REPLACE operation left, namely CQ13.
+
+\begin{algorithm}[h]
+  \caption{Answering CQ13: ``Is the name of element $i$ clear?'' With: ``No, it should be $n$''}\label{alg:cq13}
+  \begin{algorithmic}[1]
+    \Procedure{CQ13}{$i, n$}
+    \State $id\gets id+1$
+    \State $ArgsN \gets\{A\backslash\{name(i,x)\}\mid$\par\label{alg:cq13:line3}
+    \hskip\algorithmicindent$A\in Args, name(i,x)\in A\}$
+    \State $A \gets B \cup \{name(i,n)\}$ with $B\in ArgsN$
+    \State $Args \gets Args \cup \{A\}$
+    \For{$B$ in $ArgsN$}
+      \State $Att\gets Att \cup \{(A,B)\}$
+    \EndFor
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\rationale{} Algorithm~\ref{alg:cq13} first collects all arguments contain a statement of the form $name(i,x)$ into the set $ArgsN$, where $x$ can by any name.
+These are all the arguments that the new argument will attack, to ensure this is the only argument for a name of this element type.
+Since all such arguments are the same, except for the name proposition, we can take any such argument to construct a new one.
+The argument is added to the set of arguments, and the for loop ensures that all previous arguments are attacked.
+
+We next turn to the INTRO operations, which are very similar to the algorithms for argument schemes.
+The main difference is that when answering a critical question, we can already assume an argument has been created, so we do not have to create it again.
+
+\begin{algorithm}[h]
+  \caption{Answering CQ6b: ``Are there alternative ways of contributing to the same softgoal $s_{id}$?'' With: ``Yes, namely by doing task $t$''}\label{alg:cq6b}
+  \begin{algorithmic}[1]
+    \Procedure{CQ6b}{$s_{id}, t$}
+    \If{$\exists_{A\in Args} \{task(i),name(i,t)\}\subseteq A$}
+        \State $t_{id} \gets i$
+    \Else
+      \State $id\gets id+1$
+      \State $A \gets \{task(id),name(id,t)\}$
+      \State $Args \gets Args\cup A$
+      \State $t_{id} \gets id$
+    \EndIf
+    \State $id\gets id+1$
+    \State $A\gets \{contr(id, t_{id}, s_{id}, pos)\}$
+    \State $Args \gets Args\cup A$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\rationale{} Comparing algorithm~\ref{alg:cq6b} with algorithm~\ref{alg:as6} shows that the two algorithms are indeed very similar.
+
+The other critical questions with effect INTRO, namely CQ6c, CQ6d, CQ7b, and CQ9 all have a very similar structure to algorithm~\ref{alg:cq6b} and have therefore been omitted.
+
+Our final critical question is the generic counterargument, which is very simple.
+
+\begin{algorithm}[h]
+  \caption{Generic counterargument to argument $A$}\label{alg:cq6b}
+  \begin{algorithmic}[1]
+    \Procedure{Attack}{$A$}
+    \State $A_{new} = \{\}$
+    \State $Args \gets Args \cup \{A_{new}\}$
+    \State $Att \gets Att \cup \{(A_{new},A)\}$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\subsection{Some examples}
+
+\todo{Marc}{Marc}{Revisit the examples}

### 2017-01-24 gmas.tex

@@
 AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
-& & CQ5b & Does the goal decompose into any other tasks?& INTRO\\
+& & CQ5b & Does the goal decompose into any other tasks?& REPLACE\\
 \hline
@@
 \hline
-AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& INTRO\\
+AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& REPLACE\\
  &  & CQ10b & Is the decomposition type correct? (AND/OR/XOR)& REPLACE\\
@@
 \includegraphics[width=\textwidth]{img/transcript_grl}
-\caption{The GRL model manually constructed from transcript 1.
+\caption{The GRL model manually constructed from transcript $t_1$.
 Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
@@
 
-\subsubsection{Example 1: Disable task ``Traffic light''}
+\subsubsection{Example 1: Disable task \emph{Traffic light}}
 
@@
 
-\subsubsection{Example 2: Clarify task ``Road pattern''}
+\subsubsection{Example 2: Clarify task \emph{Road pattern}}
 
@@
 
-\subsubsection{Example 3: Decompose goal ``Simulate''}
+\subsubsection{Example 3: Decompose goal \emph{Simulate}}
 
@@
 
-\subsubsection{Example 4: Reinstate actor ``Development team''}
+\subsubsection{Example 4: Reinstate actor \emph{Development team}}
 

### 2017-01-25 introduction.tex
small things intro

@@
 
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
+Requirements Engineering (RE) is an approach to assess the role of a future information system within a human or automated environment.
 An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
@@
 Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
+The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}.
 
@@
 For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
+Therefore, it is often not necessary to record all of the details of the discussion process that led to the final goal model.
+However, most real-world information systems -- e.g., air-traffic management, industrial production processes, or government and healthcare services -- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons.
 
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
 \todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
@@
 As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
+\todo{Floris}{all}{Depending on whether we use preferences in argumentation models, we might want to rephrase this}
 \item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
+Other stakeholders, who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
 If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
@@
 \item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
+It is not possible to reason about changing beliefs and opinions, and their effect on the goal model.
 A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
+Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying beliefs and arguments.
+This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs put forward by another group of stakeholders.
 \end{enumerate}
 
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
+The aim of this research is to resolve the above issues by designing a framework, which combines a methodology for the goal modeling process \todo{Floris}{Sepideh}{Do you know of any generic methodologies or heuristics for goal modeling (akin to Juretas GAM)?} with ideas from Artificial Intelligence on practical reasoning and argumentation \cite{atkinson2007} and simple software tool support.
+We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc.
+allows us to find rationales that would otherwise not have been found -- for this we would need an experiment.
+Tailor it towards our empirical work: ``the framework should be close to on the actual discussions stakeholders or designers have in the early requirements engineering phase''}.
 
@@
 The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
+In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from argumentation and discourse modelling called \emph{argument schemes} \cite{walton-etal2004}.
 This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
 By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
+We first propose a set of argument schemes and critical questions we expect to find in the discussions that take place during the goal modelling process.
+We then validate this initial set with transcripts containing discussions about the architecture of an information system.
+This then gives us the final set of arguments and critical questions.
 
@@
 
-We implement our framework into an online tool called RationalGRL.
+We implement our framework into an online tool called RationalGRL (requirement 4).
 The tool is implemented using Javascript.
 It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
+The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators .
 The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
+The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the Eclipse-based tool for GRL modeling, for further evaluation and  analysis.
 

### 2017-01-26 background.tex
GRL section and example

@@
 \label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
+GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
 It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
 A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
+Figure~\ref{fig:trafficsim} illustrates a GRL diagram.
+An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\textbf{User}, Figure~\ref{fig:trafficsim}), or the system itself (\textbf{Simulation}, Figure~\ref{fig:trafficsim}).
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
@@
 Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
+\textbf{Realistic simulation)} are often more related to NFRs, whereas goals (such as \textbf{Generate cars}) are more related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
+In Figure~\ref{fig:trafficsim}, some of the tasks are \textbf{Create new cars} and \textbf{Keep same cars}.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
+\textbf{******})
 
@@
 \centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
+\includegraphics[scale=0.35]{img/Example}
+\caption{GRL Model for the traffic simulator}
+\label{fig:trafficsim}
 \end{figure*}
@@
 AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
+In Figure~\ref{fig:trafficsim}, the goal \textbf{Generate cars} is XOR-decomposed to the tasks \textbf{Create new cars} and \textbf{Keep same cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
 A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
+Task \textbf{Create new cars} has a \textbf{help} qualitative contribution to the softgoal \textbf{Realistic simulation}.
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
+For example, actor \textbf{Simulation} depends on the actor \textbf{User} to perform the task \textbf{Adjust car spawning rate} to fulfil its task \textbf{Create new cars}.
 
@@
 Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
+GRL has a well-defined syntax and semantics, which are necessary if we want to incorporate it into a formal framework (requirements 1 and 2 as described in the introduction).
+Furthermore,GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
+For example, traceability links are provided between between concepts and instances of goal modeling and behavioral design models \cite{}.
+Multiple views and traceability are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
 \todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
@@
 
-\subsubsection{Online Meeting Scheduler Example}
+\subsubsection{Traffic simulator example}
 \label{sect:background:casestudy}
@@
 
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
+The GRL model in Figure~\ref{fig:trafficsim} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
 However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
@@
 These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain information about discussions that let up to the resulting elements of the model, such as by various clarification steps for the naming, or alternatives that have been considered for the relationships.
+The model in figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as by various clarification steps for the naming, or alternatives that have been considered for the relationships.
 In order to address this, we use Argument Schemes for Goal Modeling (GMAS).
@@
 Take, for example, critical question 8.
-If new evidence comes up indicating that meeting schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textsf{user friendliness} is not a legitimate value'.
+If new evidence comes up indicating that meeting schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textbf{user friendliness} is not a legitimate value'.
 Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).

### 2017-01-26 introduction.tex
GRL section and example

@@
 The aim of this research is to resolve the above issues by designing a framework, which combines a methodology for the goal modeling process \todo{Floris}{Sepideh}{Do you know of any generic methodologies or heuristics for goal modeling (akin to Juretas GAM)?} with ideas from Artificial Intelligence on practical reasoning and argumentation \cite{atkinson2007} and simple software tool support.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc.
+We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system; (2) it must be able to generate formal goal models based on the discussions; (3) it should have formal traceability links between goal elements and arguments in the discussions; (4) it must have tool support; (5) a methodology must be identified that allows the framework to be used by practitioners; and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc.
 allows us to find rationales that would otherwise not have been found -- for this we would need an experiment.

### 2017-01-26 background.tex
example

@@
 Softgoals (e.g.
-\textbf{Realistic simulation)} are often more related to NFRs, whereas goals (such as \textbf{Generate cars}) are more related to functional requirements.
+\textbf{Realistic simulation)} are often more related to NFRs, whereas goals (such as \textbf{Generate traffic}) are more related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
@@
 AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:trafficsim}, the goal \textbf{Generate cars} is XOR-decomposed to the tasks \textbf{Create new cars} and \textbf{Keep same cars}.
+In Figure~\ref{fig:trafficsim}, the goal \textbf{Generate traffic} is XOR-decomposed to the tasks \textbf{Create new cars} and \textbf{Keep same cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
@@
 
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
+In this paper, we use for our examples materials from our case study on the architecture for a traffic simulator \cite{}.
+\todo{Floris}{all}{briefly explain the traffic simulation case and the GRL model}
 
@@
 
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
+\todo{Floris}{all}{These have to be amended}
 \begin{itemize}
@@
 
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
 These are the type of the questions that we cannot answer just by looking at the GRL models.
@@
 \begin{itemize}
-\item[] \textbf{Find agreeable meeting date} is a goal,
-\item[] \textbf{Find agreeable date by talking to initiator} will realize the goal \textbf{(Find) agreeable meeting date},
-\item[] Which will promote the value \textbf{User friendliness}
+\item[] \textbf{Generate traffic} is a goal,
+\item[] \textbf{Keep same cars} will realize the goal \textbf{Generate traffic},
+\item[] Which will promote the value \textbf{Simple design}
 \item[] \textit{Therefore} 
-\item[] We should perform action \textbf{Find agreeable date by talking to initiator}
+\item[] We should perform action \textbf{Keep same cars}
 \end{itemize}
@@
 These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 8.
-If new evidence comes up indicating that meeting schedulers do not care about user friendliness we have a counterargument to the above argument, `evidence shows that \textbf{user friendliness} is not a legitimate value'.
+Take, for example, critical question 4: if we find that \textbf{Keep same cars} actually negatively influences the value \textbf{Realistic simulation}, we have a counterargument to the above argument.
 Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).
-For example, we can argue that \textbf{Find agreeable date using scheduler} also realizes the goal \textbf{Find agreeable meeting date}, which gives us a counterargument to the original argument -- to find a date by talking to the initiator -- that also follows PRAS.
+For example, we can argue that \textbf{Create new cars} also realizes the goal \textbf{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
 
@@
 In this section, we will briefly discuss the intuitions behind these semantics.}.
-Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for \textbf{Find agreeable date by talking to initiator} (A1), the argument for \textbf{Find agreeable date using scheduler} (A3), and the argument that \textbf{User friendliness} is not a legitimate value (A2).
+Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for \textbf{Keep same cars} (A1), the argument for \textbf{Create new cars} (A3), and the argument that \textbf{Keep same cars} demotes the value \textbf{Realistic simulation} (A2).
 The two alternative PRAS instantiations are A1 and A3.
-These arguments mutually attack each other, as \textbf{Find agreeable date by talking to initiator}  and \textbf{Find agreeable date using scheduler} are considered to be mutually exclusive.
-Argument A2 attacks A1, as it questions the legitimacy of the value \textbf{User friendliness}.
+These arguments mutually attack each other, as \textbf{Keep same cars}  and \textbf{Create new cars} are considered to be mutually exclusive.
+Argument A2 attacks A1, as it points to a negative side-effect of \textbf{Keep same cars}.
 
@@
 Take, for example, A1 and A3.
-If we say that we prefer the action \textbf{Find agreeable date by talking to initiator} (A1) over the action \textbf{Find agreeable date using scheduler} (A3), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left).
-This makes A1 the only undefeated argument, whereas A3 is now defeated.
+If we say that we prefer the action \textbf{Create new cars} (A3) over the action \textbf{Keep same cars} (A1), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left).
+This makes A3 the only undefeated argument, whereas A1 is now defeated.
 It is also possible to give explicit arguments for preferences \cite{modgil2009}.
 These arguments are then essentially attacks on attacks.
-For example, say we prefer A1 over A3 because `the \textbf{Meeting participant} does not like working with a scheduler' (A4).
-This can be rendered as a separate argument that attacks the attack from A3 to A1 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A1, A4$\}$ the undefeated justified set of arguments.
+For example, say we prefer A3 over A1 because `it is important to have realistic traffic flows' (A4).
+This can be rendered as a separate argument that attacks the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A3, A4$\}$ the undefeated set of arguments.
 

### 2017-01-27 background.tex
PRAS example & section changed

@@
 \centering
-\includegraphics[scale=0.35]{img/Example}
+\includegraphics[]{img/Example}
 \caption{GRL Model for the traffic simulator}
@@
 PRAS follows the following basic argument structure.
-\begin{equation}
-\label{eq:eq1}
-  \begin{aligned}
- \qquad&\text{I have goal } G&\\
-&\text{Doing action }A \text{ will realize goal }G&\\
-&\text{Which will promote value }V&\\
-&\text{\emph{Therefore} I should do action }A.&
-  \end{aligned}
-\end{equation}
+
+\begin{itemize}
+\item[] We have goal $G$,
+\item[] Doing action $A$ will realize goal $G$,
+\item[] Which will promote the value $V$
+\item[] \textit{Therefore} 
+\item[] We should perform action $A$
+\end{itemize}
 
@@
 \begin{itemize}
-\item[] \textbf{Generate traffic} is a goal,
-\item[] \textbf{Keep same cars} will realize the goal \textbf{Generate traffic},
+\item[] We have goal \textbf{Generate traffic},
+\item[] \textbf{Keep same cars} will realize goal \textbf{Generate traffic},
 \item[] Which will promote the value \textbf{Simple design}
@@
 The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
-The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \todo{Marc}{Floris}{Which article do you want to cite here?}\cite{}).
+The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{Amyot:2010:EGM:1841349.1841356}).
 
-\todo{Marc}{Floris}{From here on you seem to be discussing related work.
-This kind of distracts us from the main point of this article, and the section becomes quite long.
-Perhaps it is better to make the main arguments about why our approach is useful here, and move the discussion of our previous work and Jureta to related work?}
-
-PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) have some obvious similarities, and in previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016,vanZee-etal:comma2016} we presented arguments based on PRAS of the form ``G is a goal, Action A realizes G \emph{Therefore} perform action A'' (see Figure~\ref{fig:pras:example3}, left, where the arrow denotes an inference step).
-Such arguments, which can be constructed using the online OVA tool\footnote{\url{http://ova.arg-tech.org/}}, can be combined with further arguments providing, for example, expert opinions about the practical reasoning elements (Figure~\ref{fig:pras:example3}, left).
+As can be seen in the above examples, PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) have some obvious similarities.
+One way to connect goal modelling and argumentation is to build practical reasoning arguments based on PRAS and then translate these to goal models, or vice versa (Figure~\ref{fig:pras:example3}).
+This is essentially what happens in much of the existing work on argumentation and goal modelling \cite{Jureta:RE2008,vanzee-etal:renext2015,vanZee-etal:er2016}
 
@@
 \includegraphics[width=\columnwidth]{img/Fig3}
-\caption{Structured argument based on PRAS (left) and the corresponding GRL diagram (right)}
+\caption{Translating between practical reasoning argument (left) and GRL diagram (right)}
 \label{fig:pras:example3}
@@
 
-Arguments built in OVA can automatically be translated to GRL diagrams using further online tooling\footnote{\url{https://github.com/RationalArchitecture/RationalGRL}} \cite{vanZee-etal:comma2016}.
-This translation takes the elements of the arguments and directly translates this to GRL elements.
-Tasks/actions, goals and softgoals/values can be directly translated from the arguments to the GRL diagrams, as can realization statements.
-Elements of arguments that have nothing to do with goal modelling (e.g.
-the expert opinion premise in Figure~\ref{fig:pras:example3}) can be included in the GRL diagram as \emph{beliefs}.
-Finally, attacks between arguments are captured as negative contribution links.
-Take, for example, \todo{example}.
-A very similar approach to argumentation and goal models had already been taken by Jureta et al.~\cite{Jureta:RE2008} who proposed a formal argumentation model that can be used to justify goal-modelling decisions.
-This argumentation model, while not explicitly based on a practical reasoning scheme, is essentially the same as our previous argumentation model based on PRAS: a formal model of structured arguments with goals as premises and tasks as conclusions, and a translation function of these arguments to goal diagrams.
+One question is what the added value is of the direct translations as shown in Figure~\ref{fig:pras:example3}.
+The idea is that arguments show the rationales behind goal models and the alternative or possibly conflicting views that were lost in the goal-modeling process.
+However, GRL already contains a \emph{belief} element that allow one to provide beliefs underlying a goal model, an \emph{XOR decomposition} link that allows for modeling of alternative ways to realize goals, and \emph{negative contribution} links for capturing conflicting goals.
+So argumentation models are not needed to precisely render goals, tasks, beliefs and the relations between them, as GRL is already perfectly suited for this.
+Rather, argumentation should guide the goal modelling \emph{process}.
 
-The core problem of the use of argumentation in the above-mentioned work \cite{Jureta:RE2008,vanzee-etal:renext2015,vanZee-etal:er2016}, is that it focuses mainly on structured arguments, which are then directly translated to goal models as in Figure~\ref{fig:pras:example3}.
-The question is what the added value of this exercise is.
-Firstly, the arguments and goal models contain many of the same elements.
-The idea is that arguments show the rationales behind goal models and the alternative or possibly conflicting views that were lost in the goal-modeling process.
-But GRL already contains a \emph{belief} element that allow one to provide beliefs underlying a goal model, an \emph{XOR decomposition} link that allows for modeling of alternative ways to realize goals, and \emph{negative contribution} links for capturing conflicting goals.
-
-The point is that argumentation models are not needed to precisely render goals, tasks, beliefs and the relations between them, as GRL is already perfectly suited for this.
-Rather, argumentation should guide the goal modelling \emph{process}.
-Instead of static, formal argument trees, we need a dynamic argumentation process in which goals and tasks are critically analysed and which captures the dialectical nature of justification.
-Jureta et al.~\cite{Jureta:RE2008} take a step towards such a process by defining a high-level decision process that organizes argumentation and clarification techniques in relation to goal modelling.
-However, the argumentation techniques in this process are focused on building precisely the type of arguments that we argue are redundant next to GRL diagrams, namely structured arguments like in Figure~\ref{fig:pras:example3} which can be translated more or less directly to goal models.
-The decision process itself is never formally defined, and how to go from a discussion on goals and tasks to a goal model remains largely implicit.
-
-What is needed is a formal framework that captures the discussions about goal models, and allows for (semi-)automated construction of goal models given this discussion.
-In section XX, we argue that a combination of Atkinson et al.'s~\cite{atkinson2007} critical questions and the dialectical semantics as formally captured by Dung's work \cite{Dung1995} are ideally suited to provide such a formal, dynamic framework.
+In order for an argumentation-based framework to captures discussions about goal models, it should capture the dynamic, question-answer interplay between argument and counterargument.
+Argumentation schemes and their associated critical questions are ideally suited for this: as Murukannaiah et al.~\cite{murukannaiah2015} have shown, argumentation schemes and critical questions can guide users in systematically deriving conclusions and making their assumptions explicit.
+Furthermore, if we want to allow for the (semi-)automated construction of goal models given arguments and counterarguments, the argumentation framework needs to have a proper dialectical semantics \cite{Dung1995}.
+This means that the basic ideas of Atkinson et al.'s~\cite{atkinson2007} practical reasoning -- goal-based arguments and critical questions combined with the acceptability semantics of Dung -- can be used in our RationalGRL framework, as will be discussed in section \ref{sect:gmas}.

### 2017-01-27 relatedwork.tex
PRAS example & section changed

@@
 Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
+
+Such arguments, which can be constructed using the online OVA tool\footnote{\url{http://ova.arg-tech.org/}}, can be combined with further arguments providing, for example, expert opinions about the practical reasoning elements (Figure~\ref{fig:pras:example3}, left).
+Arguments built in OVA can automatically be translated to GRL diagrams using further online tooling\footnote{\url{https://github.com/RationalArchitecture/RationalGRL}} \cite{vanZee-etal:comma2016}.
+This translation takes the elements of the arguments and directly translates this to GRL elements.
+Tasks/actions, goals and softgoals/values can be directly translated from the arguments to the GRL diagrams, as can realization statements.
+Elements of arguments that have nothing to do with goal modelling (e.g.
+the expert opinion premise in Figure~\ref{fig:pras:example3}) can be included in the GRL diagram as \emph{beliefs}.
+Finally, attacks between arguments are captured as negative contribution links.
+Take, for example, \todo{example}.
+A very similar approach to argumentation and goal models had already been taken by Jureta et al.~\cite{Jureta:RE2008} who proposed a formal argumentation model that can be used to justify goal-modelling decisions.
+This argumentation model, while not explicitly based on a practical reasoning scheme, is essentially the same as our previous argumentation model based on PRAS: a formal model of structured arguments with goals as premises and tasks as conclusions, and a translation function of these arguments to goal diagrams.

### 2017-01-27 relatedwork.tex
minor related work

@@
 This argumentation model, while not explicitly based on a practical reasoning scheme, is essentially the same as our previous argumentation model based on PRAS: a formal model of structured arguments with goals as premises and tasks as conclusions, and a translation function of these arguments to goal diagrams.
+
+Instead of static, formal argument trees, we need a dynamic argumentation process in which goals and tasks are critically analysed and which captures the dialectical nature of justification.
+Jureta et al.~\cite{Jureta:RE2008} take a step towards such a process by defining a high-level decision process that organizes argumentation and clarification techniques in relation to goal modelling.
+However, the argumentation techniques in this process are focused on building precisely the type of arguments that we argue are redundant next to GRL diagrams, namely structured arguments like in Figure~\ref{fig:pras:example3} which can be translated more or less directly to goal models.
+The decision process itself is never formally defined, and how to go from a discussion on goals and tasks to a goal model remains largely implicit.

### 2017-02-04 introduction.tex
Fixed/rewrote introduction partially
Addressed some of the comments and rewrote part of the intro

--- 2017-01-25 introduction.tex
@@
 However, most real-world information systems -- e.g., air-traffic management, industrial production processes, or government and healthcare services -- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons, stated below:
 
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
+%TODO {Marc}{Sepideh}{Please improve 2 and 4 by adding references}
+% Sepi_Reply: Item #2 is a very obvious point that is even the point that we use RE and we do documentation.
+I don't think it needs any reference.
+I just added examples and reworded a bit to make it obvious.
+I also reword the forth point.
 
@@
 \item
-Other stakeholders, who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
+Other stakeholders, such as new developers on the team, who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase.
 If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
+%~\cite{castro2002towards}.
+%TODO Marc: if you like you can add this reference but I think it is not necessary.
 \item
@@
 \item
-It is not possible to reason about changing beliefs and opinions, and their effect on the goal model.
+In goal models in general, ``rationale'' behind any decision is static and does not immediately impact the goal models when they change.
+That is, it is not possible to reason about changing beliefs and opinions, and their effect on the goal model.
 A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
@@
 
-The aim of this research is to resolve the above issues by designing a framework, which combines a methodology for the goal modeling process \todo{Floris}{Sepideh}{Do you know of any generic methodologies or heuristics for goal modeling (akin to Juretas GAM)?} with ideas from Artificial Intelligence on practical reasoning and argumentation \cite{atkinson2007} and simple software tool support.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc.
+The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with practical reasoning and argumentation approaches from Artificial Intelligence (AI)~\cite{atkinson2007}.
+We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system; (2) it must be able to generate formal goal models based on the discussions; (3) it should have formal traceability links between goal elements and arguments in the discussions; (4) it must have tool support; (5) a methodology must be identified that allows the framework to be used by practitioners; and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise %TODO Sepi_Reply: we need to come back to this at the end of the paper.
+ \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc.
 allows us to find rationales that would otherwise not have been found -- for this we would need an experiment.
@@
 
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
+In this context, the main research question is: \emph{What are the constructs, mechanisms and rules for developing a framework that formally captures the discussions between stakeholders such that it can generate goal models?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
+%TODO Sepi: Need to refer to research question other parts of the paper when we discuss the constructs, etc.
+
+%TODO \todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done. If we write it this way, it means we are unsure of the result.}
 
@@
 The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from argumentation and discourse modelling called \emph{argument schemes} \cite{walton-etal2004}.
+In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
 This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
 By combining arguments and critical questions, one can create a structured representation of a discussion.
-We first propose a set of argument schemes and critical questions we expect to find in the discussions that take place during the goal modelling process.
-We then validate this initial set with transcripts containing discussions about the architecture of an information system.
-This then gives us the final set of arguments and critical questions.
+We, first, propose a set of argument schemes and critical questions we expect to find in the discussions that take place during the goal modeling process.
+We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
+This, then, gives us the final set of arguments and critical questions.
 
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
+In order to generate goal models based on formalized discussions (requirement 2), we, first, formalize the list of arguments from requirement 1 in an argumentation framework.
 We formalize the critical questions as algorithms modifying the argumentation framework.
@@
 We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
+This creates traceability links from GRL elements to the underlying arguments (requirement 3).
 
-We implement our framework into an online tool called RationalGRL (requirement 4).
+We implement our framework in an online tool called RationalGRL (requirement 4).
 The tool is implemented using Javascript.
 It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators .
+The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators.
 The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
@@
 
-The rest of the article is in line with the main contributions.
+The rest of the article is in line with the main contributions and the research question.
 We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.

### 2017-02-04 background.tex
Rewrote GRL section and fixed other things
Lots of comments left to be answered

@@
 
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language used to connect to the argumentation framework.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
+In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.
+Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS)~\cite{}, which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
 
@@
 GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
+GRL is part of the User Requirements Notation (URN)~\cite{}, an ITU-T standard, that combines goals and non-functional requirements with functional and operational requirements (i.e.
+Use Case Maps (UCM)) in one.
+GRL can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
 A GRL model is a connected graph of intentional elements that optionally are part of actors.
 Figure~\ref{fig:trafficsim} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\textbf{User}, Figure~\ref{fig:trafficsim}), or the system itself (\textbf{Simulation}, Figure~\ref{fig:trafficsim}).
+An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system ( \texttt{User}, Figure~\ref{fig:trafficsim}), or the system itself ( \texttt{Simulation}, Figure~\ref{fig:trafficsim}).
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
@@
 Softgoals (e.g.
-\textbf{Realistic simulation)} are often more related to NFRs, whereas goals (such as \textbf{Generate traffic}) are more related to functional requirements.
+\texttt{Realistic simulation)} are often more related to NFRs, whereas goals (such as  \texttt{Generate traffic}) are more related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:trafficsim}, some of the tasks are \textbf{Create new cars} and \textbf{Keep same cars}.
+In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \textbf{Keep same cars}.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textbf{******})
+\texttt{******}) %TODO Sepideh: The figure doesn't match the text anymore!
 
@@
 AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:trafficsim}, the goal \textbf{Generate traffic} is XOR-decomposed to the tasks \textbf{Create new cars} and \textbf{Keep same cars}.
+In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate traffic} is XOR-decomposed to the tasks \textbf{Create new cars} and  \texttt{Keep same cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
 A contribution link has a qualitative contribution type or a quantitative contribution.
-Task \textbf{Create new cars} has a \textbf{help} qualitative contribution to the softgoal \textbf{Realistic simulation}.
+Task  \texttt{Create new cars} has a  emph{help} qualitative contribution to the softgoal  \texttt{Realistic simulation}.
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textbf{Simulation} depends on the actor \textbf{User} to perform the task \textbf{Adjust car spawning rate} to fulfil its task \textbf{Create new cars}.
+For example, actor  \texttt{Simulation} depends on the actor  \texttt{User} to perform the task  \texttt{Adjust car spawning rate} to fulfill its task  \texttt{Create new cars}.
+\todo{Sepideh}{all} These elements are not shown anywhere.
+Improve the graph.
 
@@
 GRL has a well-defined syntax and semantics, which are necessary if we want to incorporate it into a formal framework (requirements 1 and 2 as described in the introduction).
-Furthermore,GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-For example, traceability links are provided between between concepts and instances of goal modeling and behavioral design models \cite{}.
+Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
+GRL is also linked to Use Case Maps (UCM) via URNLink ((\includegraphics[scale=1]{img/urnlink}) which provides traceability between concepts and instances of the goal model and behavioral design models.
 Multiple views and traceability are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
+GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way.
+The satisfaction values from intentional elements in GRL can also be propagated to UCM elements.
+jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{}.
+GRL also has the capability to be extended through metadata, URL links, and external OCL constraints.
+This allows GRL to be used in many domains without the need to change the whole modeling language.
 This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
 
-\subsubsection{Traffic simulator example}
+\subsubsection{Traffic Simulator Example} %TODO This section needs to be written!
 \label{sect:background:casestudy}
@@
 \begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
+	\item Why are only the two softgoals \texttt{Quick} and \texttt{Low effort} selected for the task  \texttt{Organize Meeting}?
+Why is, for example, a goal \texttt{Selecting the most convenient time} not included in the analysis for the actor \texttt{Meeting Initiator}?
+	\item What does  \texttt{Quality of Proposed Date} mean?
+	\item How can one decide if the quality is low or high? Who is in charge? %\todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
 	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
@@
 	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
+	\item What does \texttt{Richer Medium} mean? 
 \end{itemize}
@@
 The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action.
-This argument can then be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
+This argument can, then, be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
 
@@
 \begin{itemize}
-\item[] We have goal \textbf{Generate traffic},
-\item[] \textbf{Keep same cars} will realize goal \textbf{Generate traffic},
-\item[] Which will promote the value \textbf{Simple design}
+\item[] We have goal  \texttt{Generate traffic},
+\item[]  \texttt{Keep same cars} will realize goal  \texttt{Generate traffic},
+\item[] Which will promote the value  \texttt{Simple design}
 \item[] \textit{Therefore} 
-\item[] We should perform action \textbf{Keep same cars}
+\item[] We should perform action  \texttt{Keep same cars}
 \end{itemize}
@@
 These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 4: if we find that \textbf{Keep same cars} actually negatively influences the value \textbf{Realistic simulation}, we have a counterargument to the above argument.
+Take, for example, critical question 4: if we find that  \texttt{Keep same cars} actually negatively influences the value  \texttt{Realistic simulation}, we have a counterargument to the above argument.
 Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).
-For example, we can argue that \textbf{Create new cars} also realizes the goal \textbf{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
+For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
+\todo{Sepideh}{all} {we don't have any link from create new cars and simple design or realistic simulation.}
 
@@
 In this section, we will briefly discuss the intuitions behind these semantics.}.
-Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for \textbf{Keep same cars} (A1), the argument for \textbf{Create new cars} (A3), and the argument that \textbf{Keep same cars} demotes the value \textbf{Realistic simulation} (A2).
+Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for  \texttt{Keep same cars} (A1), the argument for \texttt{Create new cars} (A3), and the argument that \texttt{Keep same cars} demotes the value  \texttt{Realistic simulation} (A2).
 The two alternative PRAS instantiations are A1 and A3.
-These arguments mutually attack each other, as \textbf{Keep same cars}  and \textbf{Create new cars} are considered to be mutually exclusive.
-Argument A2 attacks A1, as it points to a negative side-effect of \textbf{Keep same cars}.
+These arguments mutually attack each other, as \texttt{Keep same cars}  and \texttt{Create new cars} are considered to be mutually exclusive.
+Argument A2 attacks A1, as it points to a negative side-effect of \texttt{Keep same cars}.
 
@@
 Take, for example, A1 and A3.
-If we say that we prefer the action \textbf{Create new cars} (A3) over the action \textbf{Keep same cars} (A1), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left).
+If we say that we prefer the action  \texttt{Create new cars} (A3) over the action  \texttt{Keep same cars} (A1), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left).
 This makes A3 the only undefeated argument, whereas A1 is now defeated.
-It is also possible to give explicit arguments for preferences \cite{modgil2009}.
+It is also possible to give explicit arguments for preferences~\cite{modgil2009}.
 These arguments are then essentially attacks on attacks.
@@
 The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
-The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{Amyot:2010:EGM:1841349.1841356}).
+As mentioned in Section~\ref{sect:introduction}, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model.
+
+%The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{Amyot:2010:EGM:1841349.1841356}).
 
 As can be seen in the above examples, PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) have some obvious similarities.
-One way to connect goal modelling and argumentation is to build practical reasoning arguments based on PRAS and then translate these to goal models, or vice versa (Figure~\ref{fig:pras:example3}).
-This is essentially what happens in much of the existing work on argumentation and goal modelling \cite{Jureta:RE2008,vanzee-etal:renext2015,vanZee-etal:er2016}
+One way to connect goal models and argumentation is to build practical reasoning arguments based on PRAS and then translate these to goal models, or vice versa (Figure~\ref{fig:pras:example3}).
+This is essentially what happens in much of the existing work on argumentation and goal modeling approaches \cite{Jureta:RE2008,vanzee-etal:renext2015,vanZee-etal:er2016}
 
@@
 
-One question is what the added value is of the direct translations as shown in Figure~\ref{fig:pras:example3}.
-The idea is that arguments show the rationales behind goal models and the alternative or possibly conflicting views that were lost in the goal-modeling process.
-However, GRL already contains a \emph{belief} element that allow one to provide beliefs underlying a goal model, an \emph{XOR decomposition} link that allows for modeling of alternative ways to realize goals, and \emph{negative contribution} links for capturing conflicting goals.
+%Sepideh and Marc: We removed this paragraph as we already reasoned about the needs of argumentation in intro and this cause confusion and also downgrade our work.
+
+%One question is: ``what is the added value of using argumentation models for GRL (as shown in Figure~\ref{fig:pras:example3})?'' The idea is that arguments show the rationales behind goal models and the alternative or possibly conflicting views that were lost in the goal modeling process.
+However, GRL already contains a \emph{belief} element that allow one to provide beliefs underlying a goal model, an \emph{OR} and {XOR} decomposition links that allow for modeling of alternative ways to realize goals, and \emph{negative contribution} links for capturing conflicting goals.
 So argumentation models are not needed to precisely render goals, tasks, beliefs and the relations between them, as GRL is already perfectly suited for this.
-Rather, argumentation should guide the goal modelling \emph{process}.
+Rather, argumentation should guide the goal modeling \emph{process}.
 
@@
 This means that the basic ideas of Atkinson et al.'s~\cite{atkinson2007} practical reasoning -- goal-based arguments and critical questions combined with the acceptability semantics of Dung -- can be used in our RationalGRL framework, as will be discussed in section \ref{sect:gmas}.
+\todo{sepideh}{all}{Did we use  RationalGRL framework before?
+I don't see anywhere that we gave the name to our framework, we have to introduce it before..
+we just had the tool name}

### 2017-02-04 gmas.tex
Fixed several things and also have tons of comments

@@
 
-In this section we develop a set of argument schemes and critical questions for goal modeling.
-The final list of argument schemes and critical questions is shown in table~\ref{table:argument-schemes}.
+In this section, we develop a set of argument schemes and critical questions for goal modeling.
+The final list of argument schemes and critical questions is shown in Table~\ref{table:argument-schemes}.
 The first four argument schemes (AS0-AS4) are argument for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
-For each critical question, the right column shows the effect of answering the critical questions affirmatively, which can be DISABLE (disable the element of the original argument), INTRO (introduce a new element and argument), REPLACE (replace the element of the original argument), ATTACK (attack an argument directly).
+For each critical question, the right column shows the effect of answering the critical questions affirmatively, which can be \textsf{DISABLE} (disable the element of the original argument), \textsf{INTRO} (introduce a new element and argument), \textsf{REPLACE} (replace the element of the original argument), \textsf{ATTACK} (attack an argument directly).
 For instance, if the critical question CQ0 ``Is the actor relevant?'' is answered with ``No'', then the actor of the original argument AS0 will be disabled in the goal model.
 
-We obtained the list in table~\ref{table:argument-schemes} by analyzing transcripts of discussions between students about the development of an information system.
-In the first subsection we provide details of this analysis process with concrete examples, after which we analyze our results in the next subsection.
-In the last subsection we provide several examples of the effect of answering a critical question.
+%SG: If we only obtained them from these transcripts, how can we be sure that they are complete or even sufficient?
+Also, what are the transcripts?
+where did you take them from?
+why are they useful and why are they relevant?
+We obtained the list in Table~\ref{table:argument-schemes} by analyzing transcripts of discussions between students about the development of an information system.
+In the first subsection, we provide details of this analysis process with concrete examples, after which we analyze our results in the next subsection.
+In the last subsection, we provide several examples of the effect of answering a critical question.
 In the next section, we develop a formal language for our argument schemes, critical questions, and the effect of answering them.
+
+%SG: some comments on the table: softgoals can't decompose into goals or tasks?
+tasks don't contribute to goals?
+tasks mainly depend on resources..
+why do we have A6 and A11?
+Note that GRL doesn't have those restrictions some languages such as i* have..
+for example a softgoal can technically contribute to a task or a task can be decompsosed into goals..
+how do you address those?
+We also mentioned in background that this is the benefit of GRL over i* so why do we now don't address it?
 
@@
 
-\paragraph{Annotation method}
+\paragraph{Annotation Method}
 
+%SG: here we say we started from pras and then added more questions.. in the first paragraph we say the questions are coming from the transcripts!This is better... and we should mention it above..
 We started with an initial list of 10 argument schemes and 18 critical questions that we derived from PRAS.
@@
 If we found arguments or critical questions that did not appear in the original list, we added them and counted them as well.
-Argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in section~\ref{sect:gmas:transcripts:analysis}).
+Argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}).
 Most of the occurrences were not literally found back, but had to be inferred from the context.
@@
 We found a total of 159 instantiations of the argument schemes AS0-AS11 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
+The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (Table~\ref{table:transcripts:results:argumentschemes}).
 A large portion (about 60\%) of the argument schemes we found involved discussions around tasks of the information system (AS2, AS10).
@@
 For each transcript, we manually create a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
-An example of such a model is shown in figure~\ref{fig:transcripts:grl}.
-We added green and red dots to various elements and relationships in the figure.
+An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
+We added green and red dots to various elements and relationships in the Figure.
 A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
@@
 
-We found that answering a critical questions can have four different effects on the original argument:
+We found that answering a critical questions can have four different effects on the original argument: \todo{Sepideh}{all}{Examples here are from the meeting scheduler!
+and our figure is about traffic simulator}
 \begin{itemize}
@@
 This operation does not attack the original argument of the critical question, but rather creates a new argument.
-For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler} and \emph{Find agreeable date by talking to initiator}'' (figure~\ref{fig:SchedulerModel}).
+For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler} and \emph{Find agreeable date by talking to initiator}'' (Figure~\ref{fig:SchedulerModel}).
 Suppose now the critical question CQ5b: ``Does the goal \emph{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \emph{Find agreeable date by contacting secretary}''.
@@
 This operation does not create a new argument, but only disables (i.e., attacks) the original one.
-For instance, suppose argument scheme AS0 is instantiated with: ``Actor \emph{meeting initiator} is relevant'' (figure~\ref{fig:SchedulerModel}).
+For instance, suppose argument scheme AS0 is instantiated with: ``Actor \emph{meeting initiator} is relevant'' (Figure~\ref{fig:SchedulerModel}).
 This argument can be attack with critical question CQ0: ``Actor \emph{meeting initiator} is not relevant''.
@@
 
-In section~\ref{sect:gmas:examples} we provide examples we found in the transcripts for all of these four effects.
+In Section~\ref{sect:gmas:examples} we provide examples we found in the transcripts for all of these four effects.
 
@@
 
-\paragraph{Analysis of the argument schemes}
-Our initial list of argument schemes consists of AS1-AS4, AS5-AS9 (figure~\ref{table:argument-schemes}).
+\paragraph{Analysis of the Argument Schemes}
+Our initial list of argument schemes consists of AS1-AS4, AS5-AS9 (Table~\ref{table:argument-schemes}).
 Therefore, the difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
+We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system.
 This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
@@
 GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 11.
-Our analysis however shows that many of these schemes are not often used, and thus gives us some confidence in the resulting list.
+Our analysis however shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
+\todo{Sepideh}{all}{how does the analysis show that?
+also, what if someone says that you mentioned above that the transcripts were not used for goal models and that is why you didn't have those schemes..
+what if someone is doing arguments for goal models, how do you make sure that the list is good enough, especially that many of the relationships are not considered?)}
 
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
+\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
+%SG: I don't get this sentence.
+Also, the ones below.
+On the one hand, we found a few of the critical questions we initially proposed.
+However, this does not mean that they were not implicitly used in the minds of the participants.
 If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
 However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
+If we assume this explanation is at least partially correct, then, this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
 In the context of tool support, we believe that having these critical questions available may stimulate discussions.
@@
 For each example we provide transcript excerpts, a visualization of the arguments, and the corresponding goal model elements.
-We provide a legend for our visualization notation in figure~\ref{fig:legend}.
+We provide a legend for our visualization notation in Figure~\ref{fig:legend}.
 
@@
 
-The transcript excerpt of this example is shown in table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$.
+The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the Appendix and comes from transcript $t_1$.
 In this example, participants are summing up functionality of the traffic simulator, which are tasks that the student can perform in the simulator.
-All these tasks can be formalized are instantiations of argument scheme AS2: ``Actor \emph{Student} has task $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
+All these task can be formalized and are instantiations of argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
 
-Once all these tasks are summed up, participant P1 notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \emph{Add traffic light} is not useful.
-We formalized this using the critical question CQ12: ``Is task \emph{Add traffic light} useful/relevant?''.
+Once all these tasks are summed up, participant \texttt{P1} notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \texttt{Add traffic light} is not useful.
+We formalized this using the critical question CQ12: ``Is task \texttt{Add traffic light} useful/relevant?''.
 
-We visualized the some of the argument schemes, critical questions, and traceability links with the GRL model in figure~\ref{fig:examples:traffic-light}.
+We visualize some of the argument schemes, critical questions, and traceability links with the GRL model in Figure~\ref{fig:examples:traffic-light}.
 On the left side of the image, we see three of the instantiated argument schemes AS2.
-The bottom one, ``Actor \emph{Student} has task \emph{Add traffic light}'', is attack by another argument generated from applying critical question CQ12: ``\emph{Add traffic light} is useless (\emph{All intersections have traffic lights}).
+The bottom one, ``Actor \texttt{Student} has task \texttt{Add traffic light}'', is attacked by another argument generated from applying critical question CQ12: ``\texttt{Add traffic light} is useless (\emph{All intersections have traffic lights}).
 As a result, the corresponding GRL task is disabled.
@@
 
-The transcript excerpt of the second example is shown in table~\ref{table:transcript:task-clarification} in the appendix and comes from transcript $t_3$.
-It consists of a number of clarification steps, resulting in the task \emph{Choose a road pattern}.
+The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in the Appendix and comes from transcript $t_3$.
+It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.
 
@@
 
-The formalized argument schemes and critical questions are shown in figure~\ref{fig:examples:clarification}.
-The discussion starts with the first instantiation of argument scheme AS2: ``Actor \emph{Student} has task \emph{Create road}''.
-This argument is then challenged with critical question CQ12: ``Is the task \emph{Create road} clear?''.
-Answering this question results in a new instantiation of argument scheme AS2: ``Actor \emph{Student} has task \emph{Choose a pattern}''.
-This process is repeated two more times, resulting in the final argument ``Actor \emph{Student} has task \emph{Choose a road pattern}''.
+The formalized argument schemes and critical questions are shown in Figure~\ref{fig:examples:clarification}.
+The discussion starts with the first instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Create road}''.
+This argument is then challenged with critical question CQ12: ``Is the task \texttt{Create road} clear?''.
+Answering this question results in a new instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Choose a pattern}''.
+This process is repeated two more times, resulting in the final argument ``Actor \texttt{Student} has task \texttt{Choose a road pattern}''.
 This final argument is unattacked and has a corresponding intentional element (right image).
@@
 What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element.
-For instance, the final argument on the bottom of figure~\ref{fig:examples:clarification} attacks all previous arguments.
+For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all previous arguments.
 If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.
@@
 
-The transcript excerpt of this example is shown in table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
-It consists of a discussion about the type of decomposition relationship for the goal \emph{Simulate}.
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the Appendix and comes from transcript $t_3$.
+It consists of a discussion about the type of decomposition relationship for the goal \texttt{Simulate}.
 
@@
 
-The visualization of this discussion is shown in figure~\ref{fig:examples:decomposition}.
+The visualization of this discussion is shown in Figure~\ref{fig:examples:decomposition}.
 Each GRL element on the right has a corresponding argument on the left.
-Moreover, the original argument for an AND-decomposition is attacked by the argument for the OR decomposition, and the new argument is linked to the decomposition relation in the GRL model.
+Moreover, the original argument for an AND-decomposition is attacked by the argument for the OR-decomposition, and the new argument is linked to the decomposition relation in the GRL model.
 
@@
 
-The transcript excerpt of this example is shown in table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
-It consists of two parts: first participant P1 puts forth the suggestion to include actor \emph{Development team} into the model.
-This is then questioned by person P2, who argues that the professor will develop the software, so there won't be any development team.
-However, in the second part, participant P2 argue that the development team should be considered, since the professor doesn't develop the software.
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the Appendix and comes from transcript $t_3$.
+It consists of two parts: first participant \texttt{P1} puts forth the suggestion to include actor \texttt{Development team} in the model.
+This is, then, questioned by participant \texttt{P2}, who argues that the professor will develop the software, so there won't be any development team.
+However, in the second part, participant \texttt{P2} argue that the development team should be considered, since the professor does not develop the software.
 
@@
 We formalize this using a \emph{generic counterargument}, attacking the critical question.
-The first part of the discussion is shown in figure~\ref{fig:examples:relevant-actor}.
-We formalize the first statement as an instantiation of argument scheme AS0: \emph{Actor development team is relevant}.
-This argument is then attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
+The first part of the discussion is shown in Figure~\ref{fig:examples:relevant-actor}.
+We formalize the first statement as an instantiation of argument scheme AS0: \texttt{Actor development team is relevant}.
+This argument is, then, attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
 This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
-This is shown in figure~\ref{fig:examples:relevant-actor}, left image.
+This is shown in Figure~\ref{fig:examples:relevant-actor}, left image.
 

### 2017-02-05 formalframework.tex
formal framework

@@
 \begin{algorithm}[h]
-  \caption{Answering CQ5b: ``Does goal $G$ decompose into any other tasks?'' With: ``Yes, namey into tasks $t_{n+1},\ldots,t_k$''}\label{alg:cq5b}
+  \caption{Answering CQ5b: ``Does goal $G$ decompose into any other tasks?'' With: ``Yes, namey into tasks $t_1,\ldots,t_k$''}\label{alg:cq5b}
   \begin{algorithmic}[1]
@@
     \State $A_{new} = \{decomp(id, g_{id}, T_{id}, type)\}$\label{alg:cq5b:newarg}
-    \For{$A$ in $\{decomp(d,g_{id},x,y)\}\subseteq A\mid A\in Args\}$}\label{alg:cq5b:for2}
+    \For{$A$ in $\{decomp(\_,g_{id},\_,\_)\}\subseteq A\mid A\in Args\}$}\label{alg:cq5b:for2}
       \State $Att\gets Att \cup \{(A_{new},A)\}$
@@
 \todo{Marc}{Marc}{Revisit the examples}
+
+\todo{Marc}{Marc}{Do links need identifiers??}

### 2017-02-05 formalframework.tex
modified few things here and there
Also have a few comments

@@
-\section{The Logical Framework}
+\section{Formalization of RationalGRL Framework}
 \label{sect:formalframework}
 
-In the previous section we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of an information system.
+%SG: make sure the first paragraph is aligned with the previous section
+In Section~\ref{sect:gmas}, we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of an information system.
 The resulting list is shown in table~\ref{table:argument-schemes}.
-We also discussed various examples of applications of the different critical questions and all four different effects (right column of table~\ref{table:argument-schemes}): INTRO, DISABLE, REPLACE, and ATTACK.
+We also discussed various examples of applications of the different critical questions and all four different effects (right column of Table~\ref{table:argument-schemes}): INTRO, DISABLE, REPLACE, and ATTACK.
 
-The examples and corresponding visualizations of the previous section provided some insight into how to formalize argument schemes, critical questions, and their relationship with goal models.
-However, if we are to implement our framework into a tool, we require a more precise formalisation of these concepts.
-We choose to use formal logic to specify this, because argumentation has been studied in formal logic extensively for the past decades.
+The examples and corresponding visualizations of the previous section provided some insight on how to formalize argument schemes, critical questions, and their relationship with goal models.
+However, if we are to implement our framework in a tool, we require a more precise formalization of these concepts.
+We choose to use formal logic to specify this, since argumentation has been studied in formal logic extensively for the past decades.
+%SG: any citation?
 
-In the first subsection we develop a formal language to specify a GRL model.
-This language consists of propositions only, and a GRL model specified in this language can be specified as a logic program directly.
-In the second subsection we then provide formal definitions of all argument schemes.
-An instantiation of an argument scheme consists of a set of statement about a GRL model.
-In the third subsection we formalize the critical questions as algorithms.
+The rest of this section is as follows: In the first subsection, we develop a formal language to specify a GRL model.
+This language consists of propositions only, and a GRL model can be specified as a logic program directly.
+In the second subsection, we provide formal definitions of all argument schemes.
+An instantiation of an argument scheme consists of a set of statements about a GRL model.
+Finally, in the third subsection, we formalize the critical questions as several algorithms.
 
@@
 
-Basic concepts set theory are adequate to define the entities in a GRL model and the relations between them. We begin with the representation for the elements.
+Basic concepts of propositional logic are adequate to define the entities in a GRL model and the relations between them. We begin with the representation for the elements.
 
@@
 \item $IEs$ is the set of intentional elements (softgoals, goals, tasks, and resources),
-\item $Links$ is the set of links (decomposition, contribution, and dependency). 
+\item $Links$ is the set of links (decomposition, contribution, and dependency). %what about correlation link?
 \end{itemize}
@@
 
-\rationale{} We identify GRL element with an identifier and not with a name, because it is possible that the name of the element changes, for instance by applying critical question CQ12a (clarification), see the example ``Clarify task \emph{Road pattern}'' above.
-The GRL model of figure~\ref{fig:examples:decomposition} can be formalized as follows:
+\rationale{} We identify a GRL element with an identifier and not with a name, since it is possible that the name of the element changes, for instance by applying critical question CQ12a (clarification), see the example ``Clarify task \texttt{Road pattern}'' above.
+The GRL model of Figure~\ref{fig:examples:decomposition} can be formalized as follows:
 \begin{align*}
@@
 
-Note we do not require such propositions for actors, since there is only one type of actor, and all actors are contained in the set $Actors$.
+Note that we do not require such propositions for actors, since there is only one type of actor, and all actors are contained in the set $Actors$.
 
@@
 \rationale{} Similarly to the intentional elements, we distinguish between the different types of links between IEs using the corresponding propositions.
-The formalisation of the link in figure~\ref{fig:examples:decomposition} looks as follows:
+The formalization of the link in Figure~\ref{fig:examples:decomposition} looks as follows:
 \begin{align*}
@@
 \begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
+The name $p$ of an element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
 \end{definition}
 
-\rationale{} We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Continuing our example of formalizing figure~\ref{fig:examples:decomposition}, we have 
+\rationale{} We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Continuing our example of formalizing Figure~\ref{fig:examples:decomposition}, we have 
 \begin{align*}
@@
 \end{align*}
-Note the decomposition link with identifier 5 does not have a name, since decompositions are already distinguished by their type (OR, XOR, AND).
+Note the decomposition link with identifier 5 does not have a name, since decomposition links are already distinguished by their type (OR, XOR, AND).
 
 \begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
+Given an actor $i\in Actors$ and an element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
 \end{definition}
@@
 \rationale{} If a GRL element corresponds to an argument that is rejected, then this argument should be disabled.
-For instance, the GRL model in figure~\ref{fig:examples:relevant-actor} can be formalized as follows:
+For instance, the GRL model in Figure~\ref{fig:examples:relevant-actor} can be formalized as follows:
 \begin{align*}
@@
 
-In the next section, we formalize an argument as a set of statements about a GRL model in the language we introduced in the previous section.
+In this subsection, we formalize an argument as a set of statements about a GRL model in the language we introduced in the previous section.
 First, we formalize an argumentation framework and its semantics, which is used to determine sets of acceptable arguments.
@@
 
+%SG: a set of statements mean argument schemes?
+I think it is better to also mention the argument scheme in the text and then show the algo...
+so the reader doesn't need to always move between the table and the algo
 We formalize an argument as a set of statements about a GRL model in the language we introduced in the previous section.
@@
 \rationale{} Algorithm~\ref{alg:as0} takes one argument, namely the name of the actor $a$.
-In figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_(\text{Development team})$, results in one argument: $Args = \{\{actor(0), name(0, \text{Development team})\}\}$.
+In Figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_(\text{Development team})$, results in one argument: $Args = \{\{actor(0), name(0, \text{Development team})\}\}$.
 
@@
 
-The formalisations of argument scheme AS2 to AS4 are exactly the same as AS1 but then for respectively tasks, goals, and softgoals.
+The formalization of argument scheme AS2 to AS4 are exactly the same as AS1 but it needs to replace the actor with tasks, goals, and softgoals, respectively.
 Therefore, we have omitted them here.
@@
 
-\rationale{} The procedure in algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
+\rationale{} The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
 The for loop in the algorithm iterates over all tasks in list $T$, and if a task does not exist yet, an argument is added for it.
@@
 
-\rationale{} The procedure in algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
+\rationale{} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
 The if statements check if the softgoal exists already, and if not, an argument is added for it.
@@
 
-The algorithms for $AS7$ to $AS11$ all have a very similar structure as algorithm~\ref{alg:as6} and have therefore been omitted from this article.
+The algorithms for $AS7$ to $AS11$ all have a very similar structure as Algorithm~\ref{alg:as6} and have therefore been omitted from this article.
 
+%SG: up to here.
 \subsection{Algorithms for answering the critical questions}
@@
 We now develop algorithms for our critical questions.
-We start with the critical questions that have as effect DISABLE.
+We start with the critical questions that have an effect DISABLE.
 All critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12 fall into this category.

### 2017-02-12 background.tex
background

@@
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \textbf{Keep same cars}.
+In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
@@
 \includegraphics[]{img/Example}
-\caption{GRL Model for the traffic simulator}
+\caption{GRL Model for the traffic simulator \todo{Marc}{all}{This model is maybe a bit too simplified.
+Why not use Fig 5 completely without the dots?
+Then it is easier to see why argumentation may help to understand the model better.}}
 \label{fig:trafficsim}
@@
 AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate traffic} is XOR-decomposed to the tasks \textbf{Create new cars} and  \texttt{Keep same cars}.
+In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate traffic} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
@@
 For example, actor  \texttt{Simulation} depends on the actor  \texttt{User} to perform the task  \texttt{Adjust car spawning rate} to fulfill its task  \texttt{Create new cars}.
-\todo{Sepideh}{all} These elements are not shown anywhere.
-Improve the graph.
+\todo{Sepideh}{all}{These elements are not shown anywhere.
+Improve the graph.}
 
@@
 This can be rendered as a separate argument that attacks the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A3, A4$\}$ the undefeated set of arguments.
+\todo{Marc}{Floris}{I think we don't do preferences, right?
+Then we can simplify this discussion.
+Also, figure~\ref{fig:pras:example} and \ref{fig:pras:example2} are the same.}
 

### 2017-03-11 gmas.tex
merged thesis chapter with article

@@
-\section{Argument Schemes for Goal Modeling (GMAS)}
+\section{Argument Schemes for Goal Modeling}
 \label{sect:gmas}
-
-In this section, we develop a set of argument schemes and critical questions for goal modeling.
-The final list of argument schemes and critical questions is shown in Table~\ref{table:argument-schemes}.
-The first four argument schemes (AS0-AS4) are argument for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
-For each critical question, the right column shows the effect of answering the critical questions affirmatively, which can be \textsf{DISABLE} (disable the element of the original argument), \textsf{INTRO} (introduce a new element and argument), \textsf{REPLACE} (replace the element of the original argument), \textsf{ATTACK} (attack an argument directly).
-For instance, if the critical question CQ0 ``Is the actor relevant?'' is answered with ``No'', then the actor of the original argument AS0 will be disabled in the goal model.
-
-%SG: If we only obtained them from these transcripts, how can we be sure that they are complete or even sufficient?
-Also, what are the transcripts?
-where did you take them from?
-why are they useful and why are they relevant?
-We obtained the list in Table~\ref{table:argument-schemes} by analyzing transcripts of discussions between students about the development of an information system.
-In the first subsection, we provide details of this analysis process with concrete examples, after which we analyze our results in the next subsection.
-In the last subsection, we provide several examples of the effect of answering a critical question.
-In the next section, we develop a formal language for our argument schemes, critical questions, and the effect of answering them.
-
-%SG: some comments on the table: softgoals can't decompose into goals or tasks?
-tasks don't contribute to goals?
-tasks mainly depend on resources..
-why do we have A6 and A11?
-Note that GRL doesn't have those restrictions some languages such as i* have..
-for example a softgoal can technically contribute to a task or a task can be decompsosed into goals..
-how do you address those?
-We also mentioned in background that this is the benefit of GRL over i* so why do we now don't address it?
 
@@
 
+In this section, we develop a set of \emph{argument schemes for goal modeling} and associated critical questions.
+We start from an initial list that we derive from PRAS, containing argument schemes and critical questions that specific to elements and relationships of GRL.
+We then refine this list incrementally by annotating transcripts, meaning we identify and count instantiations of arguments schemes.
+If we find new argument schemes or critical questions, we add them to the list.
+If some of them do not occur at all, we remove them from the list.
+We then manually construct goal models from the arguments and counter arguments we find in the transcripts, and add traceability links between the goal model and underlying arguments.
+An example is shown in Figure~\ref{fig:transcripts:grl}, which is the same as the example in Figure~\ref{fig:trafficsim}, but now including tracability links to underlying arguments.
+A red dot indicates the underlying argument is rejected, and a green dot means the underlying argument is accepted.
+
+The final list of argument schemes and critical questions that we end up with after our analysis is shown in Table~\ref{table:argument-schemes}.
+The initial list of argument schemes, i.e., those we start with before annotating the transcripts, consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}), and their corresponding critical questions.
+The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
+
+We found that answering critical questions can have varying effects on the model, and for each critical question, the right column in Table~\ref{table:argument-schemes} shows the effect of answering the critical questions affirmatively.
+Answering a critical questions can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
+
+In the first subsection of this section, we provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts), after which we analyze our results in the second subsection.
+
 \subsection{Details experiment}
@@
 \paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
 Participants were asked to use a think-aloud method during the design session.
 The assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
+The full problem descriptions can be found in Appendix~\ref{ch:designprompt}.
 All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software.
 The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
+The details of the transcripts are shown in Table~\ref{table:transcripts:info}.
 
@@
 
-%SG: here we say we started from pras and then added more questions.. in the first paragraph we say the questions are coming from the transcripts!This is better... and we should mention it above..
-We started with an initial list of 10 argument schemes and 18 critical questions that we derived from PRAS.
+We started with an initial list of 8 argument schemes and 18 critical questions that we derived from PRAS (AS1-AS4, AS6-AS9 of Table~\ref{table:argument-schemes}).
 We annotated transcripts with the arguments and critical questions from this list.
@@
 
-\todo{Marc}{Floris}{I'd like to say something in the next paragraph about the difficulty of extracting arguments from text. Do you know any references?}
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.
+It is generally known in the argumentation literature that it can be very difficult to identify arguments in natural language texts~\cite{walton-etal2004}.
 Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
+However, there is hardly any research on on argument extraction in the requirement engineering domains, so despite this potential weakness in our approach, we believe it nevertheless is at least useful as something that others can build further on (see Section~\ref{sect:goalmodeling:openissues}).
 
@@
 
+All original transcripts, annotations, and models are available on the following website:
+\begin{quote} \url{http://www.github.com/marcvanzee/RationalArchitecture}
+\end{quote}
+in the folder ``Ch4 transcript analysis''.
+We also provide excerpts of the annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples we use in this article come from the transcripts.
+
 We found a total of 159 instantiations of the argument schemes AS0-AS11 in the transcripts.
@@
 
-We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing about the relevance of it (CQ12, CQ13).
+We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing the relevance of it (CQ12, CQ13).
 
@@
 
-For each transcript, we manually create a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
+For each transcript, we manually created a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
 An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
-We added green and red dots to various elements and relationships in the Figure.
+We added green and red dots to various elements and relationships in the figure.
 A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
@@
 
-We found that answering a critical questions can have four different effects on the original argument: \todo{Sepideh}{all}{Examples here are from the meeting scheduler!
-and our figure is about traffic simulator}
+We found that answering a critical questions can have four different effects on the original argument and the corresponding GRL element:
 \begin{itemize}
-\item \emph{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
+\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
 This operation does not attack the original argument of the critical question, but rather creates a new argument.
-For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler} and \emph{Find agreeable date by talking to initiator}'' (Figure~\ref{fig:SchedulerModel}).
-Suppose now the critical question CQ5b: ``Does the goal \emph{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \emph{Find agreeable date by contacting secretary}''.
-This results in a new instantiation of AS5, namely: ``Goal \emph{Agreeable meeting dates} OR-decomposes into tasks \emph{Find agreeable date using scheduler}, \emph{Find agreeable date by talking to initiator}, and \emph{Find agreeable date by contacting secretary}.
-As a result, the goal model will contain the corresponding task \emph{Find agreeable date by contacting secretary}, as well as an OR-decomposition relation from the goal \emph{Agreeable meeting dates} to that task.
-\item \emph{DISABLE:} Disable the element or relationship of the argument scheme to which the critical questions pertains.
+For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
+Suppose now the critical question CQ5b: ``Does the goal \texttt{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \texttt{Choose randomly}''.
+This results in a new instantiation of AS5, namely: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.
+As a result, the goal model will contain the corresponding task \texttt{Choose randomly}, as well as an OR-decomposition relation from the goal \texttt{Generate cars} to that task.
+\item \textsf{DISABLE:} Disable the element or relationship of the argument scheme to which the critical questions pertains.
 This operation does not create a new argument, but only disables (i.e., attacks) the original one.
-For instance, suppose argument scheme AS0 is instantiated with: ``Actor \emph{meeting initiator} is relevant'' (Figure~\ref{fig:SchedulerModel}).
-This argument can be attack with critical question CQ0: ``Actor \emph{meeting initiator} is not relevant''.
-As a result, the argument for the actor is attack, and actor \emph{meeting initiator} is disabled in the goal model.
-\item \emph{REPLACE:} Replace the element of the argument scheme with a new element.
+For instance, suppose argument scheme AS0 is instantiated with: ``Actor \texttt{Teacher} is relevant''.
+This argument can be attack with critical question CQ0: ``Actor \texttt{Teacher} is not relevant''.
+As a result, the argument for the actor is attack, and actor \texttt{Teacher} is disabled in the goal model.
+\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
 This operation both introduces a new argument and attacks the original one.
-For instance, suppose argument scheme AS2 is instantiated wit: ``Actor \emph{meeting initiator} can perform task \emph{schedule meeting}.
-This argument can be attacked with critical question CQ13: ``The task \emph{schedule meeting} is unclear, it should be \emph{schedule board meeting}''.
-This results in replacing the original argument with the new argument ``Actor \emph{meeting initiator} can perform task \emph{schedule board meeting}.
+For instance, suppose argument scheme AS2 is instantiated with: ``Actor \texttt{Student} can perform task \texttt{Choose a pattern preference}.
+This argument can be attacked with critical question CQ13: ``The task \texttt{Choose a pattern preference} is unclear, it should be \texttt{Choose a road pattern}''.
+This results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}.
 In the goal model, the description of the task should change accordingly.
-\item \emph{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-For instance, suppose argument scheme AS4 is instantiated as follows: ``Actor \emph{meeting initiator} has goal \emph{quick}.
-For instance, suppose argument scheme AS0 is instantiated with ``Actor \emph{meeting scheduler} is relevant''.
-Suppose this argument is attacked with critical question CQ0: ``Actor \emph{meeting initiator} is not relevant'', because the participants decide to fully automate the meeting scheduling process.
+\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
+For instance, suppose argument scheme AS0 is instantiated with ``Actor \texttt{Teacher} is relevant''.
+Suppose this argument is attacked with critical question CQ0: ``Actor \texttt{Teacher} is not relevant'', because she does not create the system.
 However, this critical question may in turn be attacked again, if new evidence comes up.
-For instance, it may turn out that it is sometimes more desirable to schedule meeting by hand.
-In this case, the generic counter argument ``Actor \emph{meeting initiator} is relevant, because it is sometimes easier to schedule meetings manually'' attack the argument ``Actor \emph{meeting scheduler} is not relevant''.
-As a result, the original argument ``Actor \emph{meeting scheduler} is relevant'' is accepted again, and as a result is shown in the goal model.
+For instance, it may turn out that the teacher will work on developing the application herself.
+In this case, the generic counter argument ``Actor \texttt{Teacher} is relevant, because does develop for the application'' attack the argument ``Actor \texttt{Teacher} is not relevant''.
+As a result, the original argument ``Actor \texttt{Teacher} is relevant'' is accepted again, and as a result is shown in the goal model.
 \end{itemize}
@@
 \paragraph{Analysis of the Argument Schemes}
-Our initial list of argument schemes consists of AS1-AS4, AS5-AS9 (Table~\ref{table:argument-schemes}).
+Recall that our initial list of argument schemes consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}).
 Therefore, the difference between the initial list of argument schemes and those found back in the transcripts is quite small.
 We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
+This gives us an indication that it is possible to capture (parts of the) arguments used in those type of discussions using argument schemes.
 
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
+We observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
 Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
@@
 Our analysis however shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
-\todo{Sepideh}{all}{how does the analysis show that?
-also, what if someone says that you mentioned above that the transcripts were not used for goal models and that is why you didn't have those schemes..
-what if someone is doing arguments for goal models, how do you make sure that the list is good enough, especially that many of the relationships are not considered?)}
+However, if needed, additional argument schemes and critical questions can be added, and our list is not meant to be exhaustive.
 
@@
 Also, the ones below.
-On the one hand, we found a few of the critical questions we initially proposed.
+We found few of the critical questions we initially proposed.
 However, this does not mean that they were not implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
+If a participants for instance forms an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
 However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then, this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
+If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
 In the context of tool support, we believe that having these critical questions available may stimulate discussions.
 
-\subsection{Examples}
+\section{Examples}
 \label{sect:gmas:examples}
@@
 
-\subsubsection{Example 1: Disable task \emph{Traffic light}}
+\subsubsection{Example 1: Disable task \texttt{Traffic light}}
 
-The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the Appendix and comes from transcript $t_1$.
+The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$.
 In this example, participants are summing up functionality of the traffic simulator, which are tasks that the student can perform in the simulator.
@@
 
-\subsubsection{Example 2: Clarify task \emph{Road pattern}}
+\subsubsection{Example 2: Clarify task \texttt{Road pattern}}
 
-The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in the Appendix and comes from transcript $t_3$.
+The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_3$.
 It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.
@@
 What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element.
-For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all previous arguments.
+For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all three other arguments for a name of the element.
 If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.
+Suppose for instance the bottom argument ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' did not attack the second argument: ``Action \texttt{Student} has task \texttt{Choose a pattern}''.
+In that case, this argument would be reinstated, because its only attacker ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' is itself defeated by the bottom argument.
 
-\subsubsection{Example 3: Decompose goal \emph{Simulate}}
+\subsubsection{Example 3: Decompose goal \texttt{Simulate}}
 
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the Appendix and comes from transcript $t_3$.
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
 It consists of a discussion about the type of decomposition relationship for the goal \texttt{Simulate}.
@@
 
-\subsubsection{Example 4: Reinstate actor \emph{Development team}}
+\subsubsection{Example 4: Reinstate actor \texttt{Development team}}
 
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the Appendix and comes from transcript $t_3$.
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
 It consists of two parts: first participant \texttt{P1} puts forth the suggestion to include actor \texttt{Development team} in the model.
@@
 The first part of the discussion is shown in Figure~\ref{fig:examples:relevant-actor}.
-We formalize the first statement as an instantiation of argument scheme AS0: \texttt{Actor development team is relevant}.
-This argument is, then, attacked by answering critical question CQ0: \emph{Is actor development team relevant?} with \emph{No}.
+We formalize the first statement as an instantiation of argument scheme AS0: ``Actor \texttt{development team} is relevant''.
+This argument is, then, attacked by answering critical question CQ0: ``Is actor \texttt{development team} relevant?
+with \emph{No}.
 This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
@@
 Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counter argument has been put forward.
-The argument ``The professor doesn't develop the software'' now attacks the argument ``\emph{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\emph{Development team} is relevant''.
-As a result, the first and the last argument are both accepted, which causes the actor in the GRL model to be enabled again.
+The argument ``Actor \texttt{Professor} doesn't develop the software'' now attacks the argument ``\texttt{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\texttt{Development team} is relevant''.
+As a result, the first argument is reinstated, which causes the actor in the GRL model to be enabled again.

### 2017-03-11 main.tex
minor typo - test
formal framework
merged thesis chapter with article

--- 2017-01-21 main.tex
@@
 \usepackage{algpseudocode}
-\usepackage{amssymb, amsmath}
+\usepackage[fleqn]{amsmath} % flush equations left
+\usepackage{amssymb}
 \usepackage{array,url}
@@
 \usepackage{xcolor}
+\usepackage{fancyvrb}
+\usepackage{algorithm}% http://ctan.org/pkg/algorithms
+\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
+\newcommand{\var}[1]{{\ttfamily#1}}% variable
 
@@
 \tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
+
+% Tikz stuff for arguments with GRL
+\tikzset{
+  argNodeIN/.style={shape=rectangle,draw,fill=white,align=center},
+  argNodeOUT/.style={shape=rectangle, draw=black!60, text=black!40,fill=white},
+  traceNodeRED/.style={circle,draw=black, fill=red, inner sep=0pt,minimum size=8pt},
+  traceNodeGREEN/.style={circle,draw=black, fill=green, inner sep=0pt,minimum size=8pt},
+  attackLink/.style={draw=black,->,line width=0.5mm},
+  traceLinkRED/.style={draw, densely dotted,-,line width=0.5mm, draw=red},
+  traceLinkGREEN/.style={draw, densely dotted,-,line width=0.5mm, draw=green},
+  CQLink/.style={line width=0.5mm, dotted,->},
+  ldots/.style={draw=none,rotate=90,font=\fontsize{20}{22.4}\selectfont},
+  grl/.style={draw=none},
+  disabled/.style={opacity=0.5}
+}
+\newcommand{\argtext}[2]{
+\begin{tabularx}{0.2\textwidth}{X}
+\ifx&#1&%
+\else
+\textbf{(#1)}
+\fi
+#2
+\end{tabularx}
+}
+\newcommand{\smallargtext}[2]{
+\begin{tabularx}{0.25\textwidth}{X}
+\ifx&#1&%
+\else
+\textbf{(#1)}
+\fi
+#2
+\end{tabularx}
+}
+\newcommand{\longargtext}[2]{
+\begin{tabularx}{0.5\textwidth}{X}
+\ifx&#1&%
+\else
+\textbf{(#1)}
+\fi
+#2
+\end{tabularx}
+}
+
+% rationale
+\newcommand{\rationale}{\emph{Rationale and example:}}
 
@@
 % \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
+% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This will show only in the tex
 
@@
 \input{background}
-\input{preview}
+\input{overview}
 \input{gmas}
@@
 \input{tool}
-\input{methodology}
-\input{relatedwork}
-\input{conclusion}
+\input{discussion}
 
@@
 \input{transcripts}
+\input{grl_specification}
 

### 2017-03-11 overview.tex
merged thesis chapter with article

--- 2017-01-20 REIJ_article.tex
@@
-\documentclass[11.5pt,two column]{llncs}
-\usepackage{times}
-\usepackage{helvet}
-\usepackage{courier}
-\usepackage{enumitem}
-\usepackage{graphicx}
-\usepackage{amssymb, amsmath}
-\setcounter{tocdepth}{3}
-\usepackage{multicol}
-\usepackage{multirow}
-\usepackage{url}
-%\usepackage[margin=2pt]{subfig}
-\usepackage{xcolor}
-\usepackage{tikz}
-\usepackage{algorithm}
-\usepackage{algpseudocode}
-\usepackage{authblk}
-\usetikzlibrary{arrows}
-\usepackage{enumitem}
-\usepackage{sectsty}
-\usepackage{array,url}
-\usepackage{tikz}
-\usetikzlibrary{arrows}
-% change the appearance of the tikz arrow for the argumentation networks
-% and some other settings to make all the graphs look similar
-\tikzset{>=latex,every node/.style={circle, minimum size=0.5cm, draw=black}}
-
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
-
-% TODO command
-\newcommand\todo[4][]{%
-	\ifthenelse{\equal{#1}{resolved}}{%
-		% nothing
-	}{%
-		{\bf\color{red}TODO for #3\textcolor{gray}{(by #2)}: #4}%
-	}%
-}
-\newcommand{\Mark}[1]{\textsuperscript{#1}}
-
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
-
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This wiull show only in the tex
-
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
-
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
-
-\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee\\
-\vspace{-0.35cm}
-Computer Science and Communication (CSC), University of Luxembourg\\
-2, Avenue de l'Universite, L-4365 Esch-sur-Alzette, Luxembourg\\ marcvanzee@gmail.com
-\and
-\vspace{-0.4cm}
-Floris Bex\\
-\vspace{-0.35cm}
-Department of Information and Computing Science, Utrecht University\\
-Princetonplein 5, 3584 CC Utrecht, Netherlands\\
-florisbex@gmail.com
-\and
-\vspace{-0.4cm}
-Sepideh Ghanavati\\
-\vspace{-0.35cm}
-Department of Computer Science, Texas Tech University\\
-2500 Broadway, Lubbock, TX 79409, United States\\
-sepideh.ghanavati@ttu.edu
-}
-\date{}
-
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
-
-\begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
-The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
-\end{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
-
-\section{Introduction}
-\label{sect:introduction}
-Requirements Engineering (RE) is an approach to assess the role of a future information system (IS) within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is part of an ITU-T standard, the User Requirements Notation (URN)~\cite{URN}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is not often necessary to record all of the details of the discussion process that led to the final goal model.
-%SGNew1: I rephrased this since the sentence didn't make any sense logically and grammatically.
-However, most real-world information systems \-- e.g., air-traffic management, industrial production processes, or government and healthcare services\-- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons as followed.
-
-\todo[resolved]{Floris}{Sepideh/Marc}{Who of you was going to work on these?}
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
-As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\item
-Other users who were not the original authors of the goal model may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
-This leads to a poor understanding of the problem and solution domain.
-In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
-\item
-It is not possible to reason about changing arguments and their effect on the goal model, and vice versa.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying arguments put forward by another group of stakeholders.
-\end{enumerate}
-
-Our research goal is to apply practical reasoning and argumentation theory from Artificial Intelligence (AI) to the goal modeling process, with the expectation that doing so will help resolve issues 1-4 above.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system, (2) it must be able to generate goal models based on the formalize discussions, (3) it should have formal traceability links between goal elements and arguments in the discussions, (4) it must have tool support, (5) a methodology must be identified that allows the framework to be used by practitioners, and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise.
-
-The first requirement might seem to be a little imprecise, since it does not specify exactly  which parts of the discussion process, the framework should capture.
-However, if we combine this requirement with the second requirement, it get clear that we should capture the parts of the discussions that are required to form goal models.
-The sixth requirement is the validation against other goal modeling methodologies, and has not yet been completed.
-%SGNew2: I am not sure if I understand the last sentence.
-
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
-
-\subsection{Contributions}
-
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from AI called \emph{argument schemes}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We, first, propose a set of arguments and questions we expect to find in those discussions.
-We, then, validate this initial set with transcripts containing discussions about the architecture of an information system.
-After this, we obtain our final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL.
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators (requirement 4).
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-%SGNew3: ADDED
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the eclipsed-based tool support of GRL, for further evaluation analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
-
-\subsection{Organization}
-
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
-
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
-\label{sect:background}
-
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL), which is the goal modeling language we focus on in this article.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS), which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-
-\subsection{Goal-oriented Requirements Language (GRL)}
-\label{sect:background:grl}
-The User Requirements Notation (URN)~\cite{URN}, an ITU-T Standard, is one of the first modeling languages in the area of RE which aims at the standardization of visual notations to analyze functional (behavior) and non-functional requirements (NFRs).
-%such as performance, cost,  security, and usability.
-URN allows software and requirements engineers to identify requirements for a proposed or an evolving system and to review such requirements for correctness and completeness.
-URN combines two complementary notations: the \emph{Goal-oriented Requirement Language} (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} for goals and NFRs, and \emph{Use Case Maps} (UCMs)~\cite{Weiss05designing} for scenarios, business processes and functional requirements.
-In this article, we focus on GRL models and their links to argumentation.
-
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders.
-It can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:SchedulerModel} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system, or the system itself (e.g.
-In Figure~\ref{fig:SchedulerModel}, \textsf{Meeting Participant}).
-Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
-Softgoals (e.g.
-\textsf{User Friendly)} are often more related to NFRs, whereas goals (such as \textsf{Agreeable Meeting Dates}) are more related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:SchedulerModel}, some of the tasks are \textsf{Attend Meetings} and \textsf{Arrange Meetings}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\textsf{Agreement})
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.35]{img/SchedulerModel}
-\caption{GRL Model for Online Meeting Scheduler}
-\label{fig:SchedulerModel}
-\end{figure*}
-
-Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:SchedulerModel}, the task \textsf{Participate in Meetings} is AND-decomposed to the tasks \textsf{Attend Meetings} and \textsf{Arrange Meetings} as well as the softgoal \textsf{Convenient Meeting Date}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Softgoal \textsf{User Friendly} has \textsf{some positive} qualitative contribution to the softgoal \textsf{Low Efforts}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor \textsf{Meeting Scheduler} depends on the actor \textsf{Meeting Participant} to \textsf{Agree to Date} via the resource \textsf{Proposal Dates} to fulfill its task \textsf{Schedule a Meeting}.
-
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-
-\todo{Floris}{Sepideh}{can you haver a look at the motivation for using GRL (below)?
-It used to be a separate section but I think it fits best here, have a look at what we need.
-I agree that we can mention jUCMNav but not use it as motivation since we will use online tool.}
-
-GRL is an international standard with well defined syntax and semantics.
-This allows us to combine the theory of formal argumentation with GRL precisely, hereby satisfying requirement 3 as described in the introduction.
-Moreover, GRL has an open source tool called jUCMNav, which simplifies an implementation.
-Since implementation is a key concern for us, the extensive tool support is a strong argument in favor of GRL.
-
-According to Amyot et al.~\cite{amyot2009lightweight}, GRL has several benefits over other existing goal modeling languages, such as integration with a scenario notation (UCM), support for both qualitative and quantitative attributes (for contributions levels and satisfaction levels), there is a clear separation of GRL model elements from their graphical representation, and it provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for an more details).
-Finally, URN and jUCMNav provide traceability links between concepts and instances of goal modeling and behavioral design models.
-This traceability enables direct impact analysis of goal changes on a design and vice versa.
-This traceability feature seems to be a good fit with our current research: We aim to add traceability links between intentional elements and their underlying arguments.
-\todo{Sepideh}{Sepideh}{I need to improve: GRL has has the capability to be extended via profile, metadata and rules, and due to this capability, it has been extended to many domain.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-}
-
-\subsubsection{Online Meeting Scheduler Example}
-\label{sect:background:casestudy}
-
-In this paper, we use the classic example of an online meeting scheduler that has been used in the literature~\cite{}, ~\cite{}, ~\cite{} several times.
-The meeting scheduler has been modeled with Tropos~\cite{} and $i*$.
-We remodel this example with GRL in jUCMNav for the purpose of our framework.
-
-The meeting scheduler system has three actors, \textbf{Meeting Initiator}, \textbf{Meeting Scheduler} and \textbf{Meeting Participant}.
-The \textbf{Meeting Scheduler} has a goal that for every meeting request, it has to try to find the best time and location in a way that most of the intended participants (i.e.
-\textbf{Meeting Participant}) can attend the meeting.
-The \textbf{Meeting Initiator} organizes the meetings by asking the participants to give their availability for a range of dates and time as well as the dates that they are not available.
-Based on these information, the \textbf{Meeting Scheduler} needs to pick a date/time and a location from the available dates/times given by the participants in a way that the most number of participants can attend.
-Participants should then accept or reject the event.
-
-The GRL model in Figure~\ref{fig:SchedulerModel} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
-Some of the questions that might be interesting to know about are the following:
-
-\todo{Sepideh}{Marc}{I think you can add some more or remove/modify some of these based on your needs in the future sections}
-\begin{itemize}
-	\item Why are only the two softgoals \textbf{Quick} and \textbf{Low effort} selected for the task \textbf{Organize Meeting}?
-Why is, for example, a goal \textbf{Selecting the most convenient time} not included in the analysis for the actor \textbf{Meeting Initiator}?
-	\item What does \textbf{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? \todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \textbf{Richer Medium} mean? 
-\end{itemize}
-
-\todo[resolved]{Sepideh}{all}{We need to further improve this part}
-\todo{Marc}{all}{Improved this part, what do you think?}
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:SchedulerModel} does not contain such information, such as elements or actions that have been discussed, but did not end up in the final model, discussions that led up to the formation of an element (for instance, by various clarification steps), or alternatives that have been considered for the relationships.
-In order to address this shortcoming, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
-
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
-\label{sect:background:pras}
-
-\input{section22pras.tex}
-
-\section{Preview of the Framework}
+\section{RationalGRL overview and metamodel}
 \label{sect:preview}
 
-\todo{Marc}{Marc}{Give an example showing how our framework works without the technical details.}
+\todo{Marc}{all}{I suggest the following subsections, so the methodology becomes part of this section. What do you think of this?}
 
-\section{Argument Schemes for Goal Modeling (GMAS): First Iteration}
-\label{sect:gmas}
+\subsection{RationalGRL overview}
 
-In this section we develop an initial set of argument schemes and critical questions for goal modeling, based by the argument scheme for practical reasoning as described in the previous section.
-In the next section, we validate and improve this list by annotating arguments found in transcripts of discussions about an information system.
+\todo{Marc}{all}{Explain general idea of RationalGRL, perhaps by explaining the difference between Figure 2 and 5.}
 
-\begin{table*}[h]
-\centering
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant?\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available?\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible?\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized?\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?\\
-\hline
-\hline
-AS5 & Task $T$ contributes to goal $G$ & CQ5a & Does the task contribute to the desired goal?\\
-& & CQ5b & Are there alternative ways of realizing the same goal?\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal? \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?\\
-&& CQ6d & Does the task contribute to some other softgoal?\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?\\
-&& CQ7b & Does the goal contribute to some other softgoal?\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?\\
-\hline
-\end{tabular}
-\caption{Initial list of argument schemes and critical questions for GRL elements (AS0-AS4) and relationships (AS5-AS16).}
-\label{table:argument-schemes}
-\end{table*}
+\subsection{Metamodel}
 
-\subsection{Argument schemes and critical questions}
+Extend the GRL metamodel with our concepts.
 
-Where PRAS consists of the single argument scheme PAS, our approach is to split this into a family of argument schemes, such that each separate argument scheme can be applied to a specific part of a goal model.
-Similar to PRAS, someone who does not accept the presumptive argument may challenge it by applying critical questions.
-We have modified Atkinson et al.'s original 16 questions associated with the scheme \cite{atkinson2006argumentation} by removing those questioning elements not part of GRL (\emph{circumstances}), adding questions for additional elements GRL (\emph{resources}), and renaming concepts and relationships (e.g., the \emph{promotes} relationship).
+\subsection{RationalGRL methodology}
 
-The initial list We developed a total of 10 argument schemes and 18 critical questions, which are shown in table~\ref{table:argument-schemes}.
-Argument schemes AS0-AS4 can be instantiated for individual GRL elements, and each have a single critical question.
-CQ2-CQ4 are critical questions by Atkinson rephrased using the GRL terminology, while CQ1 is added in order to question a resource element, which does not appear in PRAS.
-AS5-AS9 are argument schemes for various GRL relationships between elements.
-As we mentioned before, this list is not meant to be exhaustive, since GRL does not put restrictions on any of relationships, so in theory any type of element can be combined with any type of elements, using any relationship.
-Here we merely use the critical questions developed by Atkinson et al.
-
-Answering a critical question results in the creation of a new argument, which may or may not attack the original arguments, depending on which question is answered.
-For instance, answering CQ1 with ``no'' results in a new argument attacking the argument that actor $a$ has resource $R$ available.
-In contrary, answering CQ5b with ``yes'' does not result in an attack on the original argument, but creates a new argument stating that some other task realizes the goal as well.
-We will return to this issue in more detail in the next section.
-In the current section, we restrict our analysis to the development of appropriate argument schemes and critical questions.
-
-\section{GMAS: Second iteration}
-\label{sect:gmas:2}
-
-The argument schemes and critical questions of Table~\ref{table:argument-schemes} are based on the theoretical work by Atkinson, and our own observations about the differences with GRL.
-However, it is unclear whether this list is exhaustive, or even whether the schemes and questions are actually used in practice.
-In order to evaluate this, we annotate transcripts of discussions between students with the arguments occurring in them, and analyze the results in order to refine our list.
-
-\subsection{Empirical evaluation with transcripts}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Designers are provided a problem description, requirements, and a description of the desired outcomes.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-Participants were asked to use a think-aloud method during the design session.
-For the student groups the assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The final problem descriptions can be found in Appendix A of Schriek's master thesis~\cite{masterthesis1}.
-All groups were instructed to apply the functional architecture method (FAM), focusing on developing the Context, the Functional, and the Informational viewpoint of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation method} We annotated transcripts with the arguments and critical questions of table~\ref{table:argument-schemes}.
-If we found arguments or critical questions that did not appear in the list, we added them and counted them as well.
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-For instance, if a participant questions an argument with a statement such as ``I don't know about that'', then we interpret this as a critical question.
-
-It is generally known in the argumentation literature that it can be very difficult to annotate arguments correctly.\todo{Marc}{Floris}{add citation} Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-Still, since research on argument extraction in the requirement engineering domains is in its infancy, we believe that our evaluation is useful by itself.
-Furthermore, our annotation is openly available\footnote{\todo{Marc}{Marc}{provide url}}, we provide parts of our annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples from this article come from the transcripts.
-In this way, we aim to make our annotation process as transparent as possible.
-
-\subsubsection{Results}
-
-Some examples of annotation can be found in appendix~\ref{sect:transcripts:excerpts}.
-We found a total of 120 instantiations of the existing argument schemes AS0-AS9 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (table~\ref{table:transcripts:results:argumentschemes}).
-Examples of argument schemes are AS1, an argument for a resource (table~\ref{table:transcript:as1-as8}); AS2, an argument for a task (tables~\ref{table:transcript:as2-cq_star_1-cq2},~\ref{table:transcript:as2-cq_star_2}, and~\ref{table:transcript:as2-as10}); AS4, an argument for a softgoal, AS5, an argument for a goal, AS7, an argument for a contribution from goal to softgoal (table~\ref{table:transcript:as4-as5-as7}); and AS8, an argument for a contribution from resource to task (table~\ref{table:transcript:as1-as8}).
-
-Of our critical questions, we annotated 9 instantiations.
-Example of critical questions are CQ0, questioning the relevance of an actor (table~\ref{table:transcript:as0-cq0}) and CQ2, questioning the possibility of a task (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Argument Schemes} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0: Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1: Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2: Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3: Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4: Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5: Goal decomposes into Task & 4 &0& 4 & 8\\
-\hline
-AS6: Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7: Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8: Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9: Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-\hline
-\textbf{TOTAL} & 37& 41 & 42 & 120\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of AS0-AS9 in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{Critical questions} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 			
-CQ2: Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a: Does the task contribute to the the goal? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b: Alternative ways to realize the same goal? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b: Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-\hline
-\textbf{TOTAL} & 5 & 2 & 1 & 9\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of critical questions CQ0-CQ9 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:criticalquestions}
-\end{table}
-
-Additionally, we identified 85 statements that did not fit our existing argument schemes and critical questions, leading to 2 new argument schemes (39 occurrences) and 8 new critical questions (29 occurrences).
-The new argument schemes and critical questions are shown in table~\ref{table:transcripts:results:new}.
-
-\subsubsection{Analysis}
-
-The analysis of the results of our empirical evaluation consists of three parts.
-First, we analyze the argument schemes, then the critical questions, and finally we analyze the effect of posing a new argument.
-
-\paragraph{Analysis of the argument schemes}
-The difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion wasn't goal models, but more generally the architecture of an information system.
-This gives us an indication that these argument schemes are able to capture arguments used in those type of discussions to some extent.
-However, we also found the following additional argument schemes:
-\begin{itemize}
-\item
-\emph{AS10: Task decomposes into task.} While our initial list of argument schemes contains a scheme to decompose goals into tasks (AS5), it does not contain one for task decomposition.
-We found that are large part of the discussions were focused around the tasks that the information system should provide.
-In the first phase of the discussion, participants often listed a number of tasks the information system should provide (table~\ref{table:transcript:as2-cq_star_1-cq2}), which were then later refined through task decomposition (table~\ref{table:transcript:as2-as10}).
-\item 
-\emph{AS11: Task contributes negatively to softgoal.}  We found that participants occasionally discuss negative contributions as well (table~\ref{table:transcript:as11}).
-\end{itemize}
-More generally, we observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 5.
-Our analysis shows that many of these schemes are not often used, but we should at least add the possibility for task decomposition (AS10) and negative contribution (AS11).
-
-\begin{table*}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|>{\bfseries}l|}
-\hline
-\textbf{New annotation} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline
-AS10: Task decomposes into task & 11 & 14 & 11 & 36\\
-\hline
-- CQ10a: Does the task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-- CQ10b: Is the decomposition correct (AND/OR/XOR)? &1 &0& 1 &2\\
-\hline
-AS11: Task contributes negatively to softgoal&2&1	&0&3\\
-\hline
-\hline
-TI: Topic introduction & 5 & 3 & 7 &15\\
-\hline
-REM: Removing useless/irrelevant/redundant element & 2 & 3 & 2 &7\\
-\hline
-CLAR: Clarifying an element &3 &10 & 3 & 16\\
-\hline
-GEN: Generic counterargument	& 0& 2 & 2 & 4\\
-\hline
-\textbf{TOTAL}&24&34&25&87\\
-\hline
-\end{tabular}
-\caption{Number of occurrences of new argument schemes and critical questions in the transcripts.}
-\label{table:transcripts:results:new}
-\end{table*}
-
-\paragraph{Analysis of the critical questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-On the one hand, we found back few of the critical questions we initially proposed.
-However, this does not mean that they weren't implicitly used in the minds of the participants.
-If a participants makes an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we feel that having these critical questions available may stimulate discussions.
-
-On the other hand, we found back relatively many new critical questions that were not on our initial list. We found back two critical questions for the new argument schemes AS10:
-\begin{itemize}
-\item \emph{CQ11a: Does the task decompose into other tasks?} (table~\ref{table:transcript:cq:task_decomp}).
-\item \emph{CQ11b: Is the decomposition correct?
-(AND/OR/XOR)} The initial list of critical questions does not distinguish between the type of decomposition that may occur.
-However, we found that this is sometimes discussed by participants (table~\ref{table:transcript:cq11b}).
-\end{itemize}
-Moreover, we found four types arguments that we could not classify as an instantiation of an argument scheme or a critical question:
-\begin{itemize}
-\item \emph{Topic introduction.} Before posing arguments, participants often proposed to discuss a certain topic, for instance by stating ``we have to determine who the users of the system are gonna be'' (table~\ref{table:transcript:as-star})..
-\item \emph{Removing useless/irrelevant/redundant element} We found this critical question especially in relation with tasks (table~\ref{table:transcript:as2-cq_star_1-cq2}).
-\item \emph{Clarifying an element} Given the large body of literature on clarification \todo{Marc}{Marc}{add citations}, it is not surprising that we found this type of argument relatively often.
-We do not aim to provide a detailed analysis of the different types of clarification techniques in this article, but we merely want to be able to capture an argument clarifying a previous argument in a course grained way.
-Such an analysis in the context of argumentation and requirement engineering can be found elsewhere~\cite{Jureta:RE2008}.
-See table~\ref{table:transcript:as2-cq_star_2} for an example of a clarification argument.
-
-\item \emph{CQ: Generic counterargument.} Participants occasionally posed counterarguments against arguments or critical questions as well, which we could not classify further (table~\ref{table:transcript:as0-cq0-cq_counterarg}).
-\end{itemize}
-
-\paragraph{Analysis of the effect of a new argument} When a new argument is put forward, this can have varying effects on the previous arguments.
-For instance, in the excerpt of table~\ref{table:transcript:as0-cq0}, the answer to the critical questions results in an argument attacking the original argument for actor ``Development team'', which ensures that ``Development team'' is no longer considered to be a relevant actor of the system.
-In table~\ref{table:transcript:as2-cq_star_2}, answering the critical question also results in an attack on the original argument for task ``car influx'', but at the same time also generates a new argument for a more specific task ``control car influx per road''.
-In general, we found the following four operations that may be associated with the arguments:
-
-\todo{Marc}{Marc/Floris}{Say that an argument is attacked in different ways and we have to distinguish this}
-\begin{itemize}
-\item \emph{INTRO: Introduce new element/link.} This operation does not attack anything, but generates new elements. Examples are CQ5b, CQ6c, and \emph{Topic introduction}.
-\item \emph{DISABLE: Disable element/link.} This operation generates an attack on an element or link but does not replace it with anything.
-Examples are CQ0-CQ4, CQ5a, and \emph{Clarifying an element}.
-\item \emph{REPLACE: Replace element/link.} This operation replaces the description of the intentional element, or it replaces the type of link (e.g., from positive contribution to negative contribution, or from AND-decomposition or OR-decomposition).
-An example is \emph{Clarifying an element}.
-\item \emph{GENERIC: Generic counterargument.} This operation simply directly attacks an argument.
-It is different from DISABLE in the sense that it only attacks one argument directly, while DISABLE attacks an argument for an element/link, and all possible previous replacements for it.
-We explain this in more detail in the next section where we introduce our algorithms.
-\end{itemize}
-
-\subsection{Final argument schemes and critical questions}
-
-Our final set of argument schemes are AS0-AS9 (table~\ref{table:argument-schemes} and AS10-AS11 (table~\ref{table:transcripts:results:new}, and the final set of critical questions are CQ0-C9 (table~\ref{table:argument-schemes} and CQ10 (table~\ref{table:transcripts:results:new}.
-Additionally, we add the four arguments TI, REM, CLAR, and GEN (table~\ref{table:transcripts:results:new}.
-
-In the end of the previous section we distinguished four operations (REPLACE, DISABLE, INTRO, and GENERIC) that can be applied when instantiating an argument scheme or answering a critical question.
-In table~\ref{table:operation-mappings} we provide a mapping form our argument schemes, critical questions, and additional arguments to these four operations.
-In the next section, we formalize these operations as algorithms.
-
-\begin{table}[ht]
-\begin{tabular}{|l|l|}
-\hline
-\textbf{Argument} & \textbf{Operation}\\
-\hline
-AS0-AS11, CQ5b,CQ6b-d,CQ7b,TI& INTRO\\
-\hline
-CQ0-CQ4,CQ5a,CQ6a,CQ7a,CQ8,CQ9,REM & DISABLE\\
-\hline
-CLAR & REPLACE\\
-\hline
-GEN & GENERIC\\
-\hline
-\end{tabular}
-\caption{Mapping from the final argument schemes AS0-AS11, final critical questions CQ0-CQ10, and additional arguments TI, REM, CLAR, GEN to the operations.}
-\label{table:operation-mappings}
-\end{table}
-
-\section{The Formal Framework}
-\label{sect:formalframework}
-
-We briefly summarize our results so far and our tasks for the current section.
-In the previous two sections we developed argument schemes and critical questions for goal modeling in two iterations.
-In section~\ref{sect:gmas} we proposed an initial list, which we derived from PRAS and the elements and relationships used in GRL (table~\ref{table:argument-schemes}).
-We then validated this set using transcripts in section~\ref{sect:gmas:2}.
-This resulted in new argument schemes and critical questions, and also in four new types of arguments that we distinguished from argument schemes and critical questions, namely Topic Introduction (TI), Removing redundant elements (REM), Clarifying an element (CLAR), and Generic counterargument (GEN) (table~\ref{table:transcripts:results:new}).
-We distinguished four types of operations that can be applied when putting forward an argument, namely Introducing a new element (INTRO), Disabling an element (DISABLE), Replacing an element (REPLACE), and adding a Generic Counterargument (GENERIC).
-In table~\ref{table:operation-mappings} we provided a mapping from our final list of arguments to these operations.
-
-In this section we develop a formal framework for the arguments and operations on them.
-In the first subsection we develop a formal language for a GRL model, consisting of actors, intentional elements (softgoals, goals, tasks, and resources), and links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements.
-In the second subsection we then provide formal definitions of all argument schemes, which are simply statements about a goal model.
-In the third subsection we formalize the critical questions.
-
-\subsection{Formal Language for a GRL Model}
-
-In order to formalize the arguments used when discussing a goal model, we first provide a formal language for a goal model. The arguments are simply statements in using this language.
-
-In our language, each element of a goal model is denoted by a non-negative integer.
-We distinguish the actors, intentional elements (softgoals, goals, tasks, and resources), the links (decomposition, positive contribution, negative contribution, and dependency) between intentional elements in three separate sets.
-
-\begin{definition}[Elements]
-Let $\{i_1,\ldots,i_n\}\subset \mathbb{N}$ denote the set of elements of a goal model.
-The actors are denoted by $Actors=\{i_1,\ldots,i_k\}$, the intentional elements are denoted by $IEs=\{i_{k+1},\ldots,i_m\}$, and the relationships are denoted by $Links=\{i_{m+1},\ldots,i_n\}$.
-\end{definition}
-
-We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Note that links do not have a description.
-
-\begin{definition}[Name of IE or Actor]
-The name $p$ of a element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
-\end{definition}
-
-All intentional elements are collected in the set $IEs$. In order to distinguish softgoals, goals, tasks, and resources, we use the following definition.
-
-\begin{definition}[Intentional Element]
-Given an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-In the same way, we distinguish between the different types of links between IEs using the following definition.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote with $contrib(i,src,dest,type)$ a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{pos,neg\}$ means a positive resp.
-negative contribution.
-We denote with $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-We denote with $dep(i,d_1,d_2,d_3)$ a dependency from $d_1\in IEs$ via $d_2\in IEs$ to $d_3\in IEs$.
-\end{definition}
-
-In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements.
-
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and a element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\todo{Marc}{Marc}{Provide example here}
-
-\subsection{Formalizing the argument schemes}
-
-In order to formalize the argument schemes we first note that an
-
-\section{The RationalGRL Tool}
-\label{sect:implementation}
-
-\section{The RationalGRL Methodology}
-\label{sect:methodology}
+\todo{Marc}{Sepideh}{Finish this}
 
@@
 \end{figure}
-
-\section{Related Work}
-\label{sect:relatedwork}
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-This work proposes ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our \textsf{RationalGRL} framework can be seen as an instantiation and implementation of  part of the GAM.
-One of the main contribution of \textsf{RationalGRL} is that it also takes the acceptability of arguments as determined by the argumentation semantics \cite{Dung1995} into account when translating from arguments to goal models.
-\textsf{RationalGRL} also provides tool support for argumentation, i.e.
-Argument Web toolset, to which OVA belongs \cite{bex2013implementing}, and for goal modeling, i.e.
-jUCMNav~\cite{jUCMNav}.
-Finally, \textsf{RationalGRL} is based on the practical reasoning approach of \cite{Atkinson2014}, which itself is also a specialization of Dung's~\cite{Dung1995} abstract approach to argumentation.
-Thus, the specific critical questions and counterarguments based on these critical question proposed by~\cite{Atkinson2014} can easily be incorporated into \textsf{RationalGRL}.
-
-\textsf{RationalGRL} framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of \textsf{RationalGRL} over DR approaches is that \textsf{RationalGRL} incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
-
-\section{Conclusions and Future Work}
-\label{sect:conclusion}
-
-In this paper, we developed and implemented a framework to trace back elements of GRL models to arguments and evidence that derived from the discussions between stakeholders.
-We created a mapping algorithm from a formal argumentation theory to a goal model, which allows us to compute the evaluation values of the GRL IEs based on the formal semantics of the argumentation theory.
-
-There are many directions of future work.
-There are a large number of different semantics for formal argumentation, that lead to different arguments being acceptable or not.
-It would be very interesting to explore the effect of these semantics on goal models.
-Jureta \emph{et al.} develop a methodology for clarification to address issues such as ambiguity, overgenerality, synonymy, and vagueness in arguments.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a formal set of critical questions that point to typical ways in which a practical argument can be criticized.
-We believe that critical questions are the right way to implement Jureta's methodology, and our framework would benefit from it.
-In addition, currently, we have not considered the \emph{Update} step of our framework (Figure~\ref{fig:overview}).
-That is, the translation from goal models to argument diagrams is still missing.
-The \emph{Update} step helps analysts change parts of the goal model and analyze its impact  on the underlying argument diagram.
-Finally, the implementation is currently a browser-based mapping from an existing argument diagramming tool to an existing goal modeling tool.
-By adding an argumentation component to jUCMNav, the development of goal models can be improved significantly.
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
-
-\bibliographystyle{abbrv}
-\bibliography{references}
-
-\newpage
-
-\onecolumn
-\appendix
-
-\section{Transcripts excerpts}
-\label{sect:transcripts:excerpts}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:10:55.2 (P1) & Maybe developers &\textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2) & Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software & \textbf{[5 critical question CQ0 for 4]} Is actor "development team" relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-\end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:19:08.6 (P3) & Should have a link with an outsource program for the statistical distribution [inaudible] & \textbf{[21 resource (AS1)]} Actor System has resource ``Statistics library''\\
-\hline
-0:35:27.4 (P3) & Maybe before traffic simulation view you can- the outsource package that makes the map	& \textbf{[38 contribution (AS8)]} Resource ``Statistics library'' contributes to task ``Display traffic simulation''\\
-\hline
-\end{tabular}
-\caption{AS1: Resource, AS8: Resource contributes to task (transcript $t_3$)}
-\label{table:transcript:as1-as8}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:15:11.2 (P1) & And then, we have a set of actions.
-Save map, open map, add and remove intersection, roads & \multirow{2}{60mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map'', ``add intersection'', ``add road'', ``add traffic light'', ``remove intersection''}\\
-\cline{1-2}
-0:15:34.7 (P2) & Yeah, road. Intersection, add traffic lights	&\\
-\hline
-0:15:42.3 (P1) & Well, all intersection should have traffic lights so it's & \textbf{[21 critical question CQ*1 for 20]} Is the task ``Add traffic light'' useful/redundant? \newline
-\textbf{[22 answer to 22]} Not useful, because according to the specification all intersections have traffic lights.\\
-\hline
-0:15:52.3 (P1) & (..) And on the technical side it's gonna be a real pain to remove one intersection you're gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then-& \textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?\newline
-\textbf{[24 answer to 22]} It is going to be very difficult to implement.\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*1: Redundant element, CQ2: impossible task (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_1-cq2}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:23:20.4 (P1) & So, sets, yeah, car influx & \textbf{[32 task (AS2)]} Student has task ``car influx''\\
-\hline
-0:23:41.2 (P2) & (..) If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently traveled.
-(...) So setting it per road, I think is something we want & \textbf{[33 critical question CQ*2 on 36]} Is the task description specific/clear enough?
-\newline
-\textbf{[34 answer to 37]} No, it is not clear where the influx is changing. Change to ``control car influx per road''\\
-\hline
-\end{tabular}
-\caption{AS2: Task, CQ*2: Specify/clarify element (transcript $t_1$)}
-\label{table:transcript:as2-cq_star_2}
-
-\begin{tabular}{|p{20mm}|p{50mm}|p{80mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:14:31.2 (P1) & Let's see- she uses the system in her course to explain her lectures about traffic problem thing & \textbf{[11 softgoal (AS4)]} ``Explain lectures traffic theory'' (Professor)\newline
-\textbf{[12 goal (AS5)]} ``Use traffic light system'' (Professor)\newline
-\textbf{[13 contribution (AS7)]} ``use traffic light system'' contributes to ``explain lectures traffic theory''\\
-\hline
-\end{tabular}
-\caption{AS4: softgoal, AS5: goal, AS7: contribution (transcript $t_3$)}
-\label{table:transcript:as4-as5-as7}
-
-\begin{tabular}{|p{20mm}|p{90mm}|p{40mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:00:10.2\newline PERSON 1 & 	So, yeah [pause] I would start with something about the context.
-That we have to determine who the users of the system are gonna be, stakeholders.
-& \textbf{[1 topic]} What are the actors?\\
-\hline
-\end{tabular}
-\caption{AS*: Topic introduction (transcript $t_1$)}
-\label{table:transcript:as-star}
-\end{table}
-
-\begin{table}[!htbp]
-\centering
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:29:59.5 (P3) & (...) a variety of sequences and timing schemes should be allowed.
-(...) we would have traffic light behavior gives you, I guess two options then.
-& \multirow{4}{60mm}{\textbf{[42 task (AS2)]} Student has task ``set sequence scheme''\newline
-\textbf{[43 task (AS2)]} Student has task ``set timing scheme'' \newline
-\textbf{[44 decomposition (AS10)] }Task ``set traffic light behavior'' XOR-decomposes into ``set sequence scheme'' and ''set timing scheme''}\\
-\cline{1-2}
-0:30:23.6 (P1) & Sequences and timing schemes &\\
-\cline{1-2}
-0:30:25.0 (P3) & Sequences and timing schemes. So you can either go for, yeah, sequences-&\\
-\cline{1-2}
-0:30:30.9 (P1) & Or timing schemes&\\
-\hline	
-\end{tabular}
-\caption{AS2: Task, AS10: Task decomposition (transcript $t_2$)}
-\label{table:transcript:as2-as10}
-
-\begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{3}{60mm}{\textbf{[26 critical question CQ*0 for 20]} Is the OR-decomposition of ``simulate'' correct?\newline
-\textbf{[27 answer to 26]} No, it should be an AND because the system can do both.}\\
-\cline{1-2}
-0:30:14.3 (P3) &I think it's an OR&\\
-\cline{1-2}
-0:30:15.4 (P1) & for the data, it's an OR (...) And for the system it's an AND. (...) Static manner or dynamic. But the system can do both&\\
-\hline	
-\end{tabular}
-\caption{CQ11b: Incorrect decomposition (transcript $t_3$)}
-\label{table:transcript:cq11b}
-
-\begin{tabular}{|p{20mm}|p{100mm}|p{30mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:06:29.3 (P2) & So, is that a trade-off.
-I think so.
-&\multirow{2}{30mm}{\textbf{[10 negative contribution (AS11)]}  task ``generate cars randomly'' contributes negatively to softgoal ``dynamic simulation''}\\
-\cline{1-2}
-0:06:36.0 (P1) & Yeah, performance versus, I don't know, functionality.
-Like, what you say, cars come out at the end of the map side [are generated randomly] is performance wise and, I don't know, easier to make but it is less functional.
-Because you can't see traffic flows that easy because, well there's fixed amount of cars so there's not really gonna be jams [the simulation is less dynamic].
-Is there around Utrecht always the same amount of cars?
-&\\
-\hline	
-\end{tabular}
-\caption{AS11: Negative decomposition (transcript $t_1$)}
-\label{table:transcript:as11}
-
-\begin{tabular}{|p{20mm}|p{80mm}|p{50mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-0:49:05.3 (P3)&So, density, speed and, is there anything else.&\multirow{3}{50mm}{\textbf{[68 critical question for 63c]} Does ``set road characteristics'' decompose into any other tasks?\newline
-\textbf{[69 answer to 68]} Yes, type of cars.}\\
-\cline{1-2}
-0:49:20.1 (P1) & Maybe type of cars&\\	
-\cline{1-2}
-0:49:22.0 (P3) & Type of cars, because you could have trucks, you could have personal cars. That would be good because-&\\
-\hline	
-\end{tabular}
-\caption{CQ: Does a task decompose into other tasks? (transcript $t_2$)}
-\label{table:transcript:cq:task_decomp}
-
-\begin{tabular}{|p{20mm}|p{60mm}|p{70mm}|}
-\hline
-Respondent & Text & Annotation\\
-\hline
-
-0:10:55.2 (P1) & Maybe developers or & \textbf{[4 actor (AS0)]} Development team\\
-\hline
-0:11:00.8 (P2)&Development team, I don't know.
-Because that's- in this context it looks like she's gonna make the software&\textbf{[5 critical question CQ0 for 4]} Is actor ``development team'' relevant?\newline
-\textbf{[6 answer to 5]} No, it looks like the professor will develop the software.\\
-\hline
-..&..&..\\
-\hline
-0:18:13.4 (P2) & I think we can still do developers here.
-To the system & \multirow{2}{70mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
-\cline{1-2}
-0:18:22.9 (P1)&Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
-\hline	
-\end{tabular}
-\caption{AS0: Task, CQ0: Relevant task? CQ: Generic counterargument (transcript $t_2$)}
-\label{table:transcript:as0-cq0-cq_counterarg}
-
-\end{table}
-
-\end{document}

### 2017-03-14 background.tex

--- 2017-02-04 background.tex
@@
-\section{Background: Goal-oriented Requirements Language and Argument Schemes}
+\section{Background: Goal-oriented Requirements Language and argument schemes}
 \label{sect:background}
 
-In this section, we first introduce and motivate the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.
-Next, we introduce and motivate \emph{the Argument Scheme for Practical Reasoning} (PRAS)~\cite{}, which is a particular argument scheme that is used to form arguments and counter-arguments about situations involving goals.
+In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.
+Lastly, we introduce argument schemes, and in particular, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, which is an argument scheme that is used to form arguments and counter-arguments about situations involving goals.
+This will be our starting point in the next section.
+
+\subsection{Running example: Traffic Simulator}
+\label{sect:goals:runningexample}
+
+Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise.
+In this exercise, designers are provided a problem description, requirements, and a description of the desired outcomes.
+The problem description is given in full in Appendix~\ref{ch:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI.
+It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory.
+To this end, a piece of software has to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
+The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
+Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
 
@@
 \label{sect:background:grl}
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements (NFRs) of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
-GRL is part of the User Requirements Notation (URN)~\cite{}, an ITU-T standard, that combines goals and non-functional requirements with functional and operational requirements (i.e.
-Use Case Maps (UCM)) in one.
+GRL is a visual modeling language for specifying intentions, business goals, and \emph{non-functional requirements} of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
+GRL is part of the User Requirements Notation, an ITU-T standard, that combines goals and non-functional requirements with functional and operational requirements (i.e.
+use case maps) in one.
 GRL can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
 A GRL model is a connected graph of intentional elements that optionally are part of actors.
-Figure~\ref{fig:trafficsim} illustrates a GRL diagram.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system ( \texttt{User}, Figure~\ref{fig:trafficsim}), or the system itself ( \texttt{Simulation}, Figure~\ref{fig:trafficsim}).
+All the elements and relationships used in GRL are shown in Figure~\ref{fig:grl_legend}.
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[scale=0.6]{img/grl_legend}
+\caption{Basic elements and relationships of GRL}
+\label{fig:grl_legend}
+\end{figure*}
+
+Figure~\ref{fig:trafficsim} illustrates a GRL diagram from the traffic simulator design exercise.
+An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\texttt{Student}, Figure~\ref{fig:trafficsim}), or the system itself ( \texttt{Traffic Tycoon}, Figure~\ref{fig:trafficsim}).
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
@@
 Softgoals (e.g.
-\texttt{Realistic simulation)} are often more related to NFRs, whereas goals (such as  \texttt{Generate traffic}) are more related to functional requirements.
+\texttt{Realistic simulation)} are often more related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are more related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \textbf{Keep same cars}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g.
-\texttt{******}) %TODO Sepideh: The figure doesn't match the text anymore!
+In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}.
+In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}, Figure~\ref{fig:trafficsim}).
 
@@
 \centering
-\includegraphics[]{img/Example}
-\caption{GRL Model for the traffic simulator}
+\includegraphics[width=\textwidth]{img/transcript_grl_0}
+\caption{GRL Model for the traffic simulator.}
 \label{fig:trafficsim}
@@
 AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate traffic} is XOR-decomposed to the tasks \textbf{Create new cars} and  \texttt{Keep same cars}.
+In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
 A contribution link has a qualitative contribution type or a quantitative contribution.
-Task  \texttt{Create new cars} has a  emph{help} qualitative contribution to the softgoal  \texttt{Realistic simulation}.
+Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the softgoal  \texttt{Dynamic simulation}.
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor  \texttt{Simulation} depends on the actor  \texttt{User} to perform the task  \texttt{Adjust car spawning rate} to fulfill its task  \texttt{Create new cars}.
-\todo{Sepideh}{all} These elements are not shown anywhere.
-Improve the graph.
+For example, actor  \texttt{Traffic Tycoon} depends on the actor  \texttt{Student} to perform the task  \texttt{Adjust car spawning rate} to fulfill its task  \texttt{Generate cars}.
 
@@
 Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-GRL is also linked to Use Case Maps (UCM) via URNLink ((\includegraphics[scale=1]{img/urnlink}) which provides traceability between concepts and instances of the goal model and behavioral design models.
+GRL is also linked to Use Case Maps via URNLink ((\includegraphics[scale=1]{img/urnlink}) which provides traceability between concepts and instances of the goal model and behavioral design models.
 Multiple views and traceability are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
+
 GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way.
-The satisfaction values from intentional elements in GRL can also be propagated to UCM elements.
-jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{}.
-GRL also has the capability to be extended through metadata, URL links, and external OCL constraints.
+The satisfaction values from intentional elements in GRL can also be propagated to use case maps elements.
+jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
+GRL also has the capability to be extended through metadata, links, and external OCL constraints.
 This allows GRL to be used in many domains without the need to change the whole modeling language.
-This feature also helps us to apply our argumentation to other domain such as compliance or enterprise architecture.
-
-\subsubsection{Traffic Simulator Example} %TODO This section needs to be written!
-\label{sect:background:casestudy}
-
-In this paper, we use for our examples materials from our case study on the architecture for a traffic simulator \cite{}.
-\todo{Floris}{all}{briefly explain the traffic simulation case and the GRL model}
+This feature also helps us to apply our argumentation to other domain such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
 
 The GRL model in Figure~\ref{fig:trafficsim} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
-However, the rationales and arguments behind certain intentional elements have not been discussed or illustrated in the GRL model.
+However, the rationales and arguments behind certain intentional elements are not shown in the GRL model.
 Some of the questions that might be interesting to know about are the following:
 
-\todo{Floris}{all}{These have to be amended}
 \begin{itemize}
-	\item Why are only the two softgoals \texttt{Quick} and \texttt{Low effort} selected for the task  \texttt{Organize Meeting}?
-Why is, for example, a goal \texttt{Selecting the most convenient time} not included in the analysis for the actor \texttt{Meeting Initiator}?
-	\item What does  \texttt{Quality of Proposed Date} mean?
-	\item How can one decide if the quality is low or high? Who is in charge? %\todo{Floris}{Sepideh}{do you mean the \textbf{Quality of Proposed Date}? Or quality of the GRL diagram?}
-	\item Does the meeting initiator use an email to inform the meeting participants to enter their availabilities? 
-	\item Does the meeting scheduler uses an online calendar to obtain the available dates or an online website?
-	\item How much time do the participants have to enter their availabilities? Is that important? Do they get a reminder from either the initiator or the scheduler? Is the reminder sent via an email? 
-	\item What does \texttt{Richer Medium} mean? 
+	\item Why does actor \texttt{Teacher} have only a single softgoal \texttt{Students learn from practice}? Why is this, for instance, not connected to any of the elements of \texttt{Student}?
+	\item What does  \texttt{Adjust timing schemes of sensorless interactions} mean?
+	\item Why does task \texttt{Keep same cars} contribut positively to \texttt{Realistic simulation} and negatively to \texttt{Dynamic simulation}?
+	\item How does the \texttt{Student} control the \texttt{Traffic Tycoon}?
+	\item Why does \texttt{Map design} have so many decompositions into other tasks?
 \end{itemize}
@@
 These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as by various clarification steps for the naming, or alternatives that have been considered for the relationships.
-In order to address this, we use Argument Schemes for Goal Modeling (GMAS).
-GMAS can help us decide what intentional elements should remain in GRL and what new ones to be added or deleted, provide rationale behind the design decisions and the relationships between the links.
+The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+In this article we aim to address this shortcoming.
 
@@
 Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g.
-\cite{Raz1978-RAZPR,walton1990}) and Artificial Intelligence \cite{Bratman1987,atkinson2007}.
+\cite{Raz1978-RAZPR,walton1990}) and artificial intelligence \cite{bratman1987,atkinson2007}.
 One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}.
@@
 
-A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the Practical Reasoning Argument Scheme (PRAS).
+A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the \emph{practical reasoning argument scheme} (PRAS).
 PRAS follows the following basic argument structure.
@@
 For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
-\todo{Sepideh}{all} {we don't have any link from create new cars and simple design or realistic simulation.}
 
@@
 In this section, we will briefly discuss the intuitions behind these semantics.}.
-Figure \ref{fig:pras:example1} shows an argumentation framework with three arguments from the above example: the argument for  \texttt{Keep same cars} (A1), the argument for \texttt{Create new cars} (A3), and the argument that \texttt{Keep same cars} demotes the value  \texttt{Realistic simulation} (A2).
+Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the above example: the argument for  \texttt{Keep same cars} (A1), the argument for \texttt{Create new cars} (A3), and the argument that \texttt{Keep same cars} demotes the value  \texttt{Realistic simulation} (A2).
 The two alternative PRAS instantiations are A1 and A3.
@@
 
-\begin{figure}[ht]
+\begin{figure}[ht!]
 \centering
-\includegraphics[scale=0.8]{img/Fig1}
-\caption{Argumentation framework}
+\begin{tikzpicture}
+        \node[minimum size=1cm] (att3) [argNodeIN] at (0,-3) {$A_3$};
+        \node[minimum size=1cm] (att1) [argNodeOUT] at (0,0) {$A_1$};
+        \node[minimum size=1cm] (att2) [argNodeIN] at (3,0) {$A_2$};
+         \path
+    (att2) edge [attackLink] (att1)
+    (att1) edge [attackLink,<->] (att3);
+\end{tikzpicture}
+\caption{Example argumentation framework.}
 \label{fig:pras:example}
@@
 A3 is then also undefeated, since its only attacker, A1, is defeated by A2.
-Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$.
+Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$, corresponding to arguments for \texttt{Realistic simulation} and \texttt{Create new cars}.
 
-In some cases, it is more difficult to determine whether or not an argument is defeated.
-Take, for example, the argumentation framework with just A1 and A3: they attack each other, they are alternatives and without any explicit preference, it is impossible to choose between the two.
-It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}.
-Take, for example, A1 and A3.
-If we say that we prefer the action  \texttt{Create new cars} (A3) over the action  \texttt{Keep same cars} (A1), we remove the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, left).
-This makes A3 the only undefeated argument, whereas A1 is now defeated.
-It is also possible to give explicit arguments for preferences~\cite{modgil2009}.
-These arguments are then essentially attacks on attacks.
-For example, say we prefer A3 over A1 because `it is important to have realistic traffic flows' (A4).
-This can be rendered as a separate argument that attacks the attack from A1 to A3 (Figure~\ref{fig:pras:example2}, right), removing this attack and making $\{$A3, A4$\}$ the undefeated set of arguments.
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.8]{img/Fig2}
-\caption{Preferences between arguments}
-\label{fig:pras:example2}
-\end{figure}
-
-\subsubsection{Practical Argumentation and Goal Modeling}
+\subsubsection*{Practical Argumentation and Goal Modeling}
 \label{sect:background:pras:motivation}
 
-Practical reasoning in the PRAS framework as described above provides a formally sound framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung1995} and its various extensions \cite{amgoud2002reasoning,modgil2009}.
+Practical reasoning in the PRAS framework as described above provides a formal framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung1995} and its various extensions \cite{amgoud2002reasoning,modgil2009}.
 The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
-As mentioned in Section~\ref{sect:introduction}, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model.
+In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model.
+This give a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it allows one to understand why certain other elements have been rejected.
 
-%The question in this paper is how PRAS can be adopted for use in goal modeling, so that we can capture the discussions between stakeholders that build a goal model as formal argumentation, thus adding a new evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (as opposed to the \emph{satisfiability} \cite{Amyot:2010:EGM:1841349.1841356}).
+Argumentation schemes and their associated critical questions are very well suited for modeling discussions about a goal model: as Murukannaiah et al.~\cite{murukannaiah2015} have shown, they can guide users in systematically deriving conclusions and making assumptions explicit.
+This can also be seen from the obvious similaries between PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) in the example above.
 
-As can be seen in the above examples, PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) have some obvious similarities.
-One way to connect goal models and argumentation is to build practical reasoning arguments based on PRAS and then translate these to goal models, or vice versa (Figure~\ref{fig:pras:example3}).
-This is essentially what happens in much of the existing work on argumentation and goal modeling approaches \cite{Jureta:RE2008,vanzee-etal:renext2015,vanZee-etal:er2016}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\columnwidth]{img/Fig3}
-\caption{Translating between practical reasoning argument (left) and GRL diagram (right)}
-\label{fig:pras:example3}
-\end{figure}
-
-%Sepideh and Marc: We removed this paragraph as we already reasoned about the needs of argumentation in intro and this cause confusion and also downgrade our work.
-
-%One question is: ``what is the added value of using argumentation models for GRL (as shown in Figure~\ref{fig:pras:example3})?'' The idea is that arguments show the rationales behind goal models and the alternative or possibly conflicting views that were lost in the goal modeling process.
-However, GRL already contains a \emph{belief} element that allow one to provide beliefs underlying a goal model, an \emph{OR} and {XOR} decomposition links that allow for modeling of alternative ways to realize goals, and \emph{negative contribution} links for capturing conflicting goals.
-So argumentation models are not needed to precisely render goals, tasks, beliefs and the relations between them, as GRL is already perfectly suited for this.
-Rather, argumentation should guide the goal modeling \emph{process}.
-
-In order for an argumentation-based framework to captures discussions about goal models, it should capture the dynamic, question-answer interplay between argument and counterargument.
-Argumentation schemes and their associated critical questions are ideally suited for this: as Murukannaiah et al.~\cite{murukannaiah2015} have shown, argumentation schemes and critical questions can guide users in systematically deriving conclusions and making their assumptions explicit.
-Furthermore, if we want to allow for the (semi-)automated construction of goal models given arguments and counterarguments, the argumentation framework needs to have a proper dialectical semantics \cite{Dung1995}.
-This means that the basic ideas of Atkinson et al.'s~\cite{atkinson2007} practical reasoning -- goal-based arguments and critical questions combined with the acceptability semantics of Dung -- can be used in our RationalGRL framework, as will be discussed in section \ref{sect:gmas}.
-\todo{sepideh}{all}{Did we use  RationalGRL framework before?
-I don't see anywhere that we gave the name to our framework, we have to introduce it before..
-we just had the tool name}
+However, there are also some differences between PRAS and GRL.
+Not all elements and relationships of GRL fit into PRAS.
+For instance, PRAS does not have a notion of ``resource'', and many of the relationships of GRL do not occur in PRAS.
+Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson actually apply to GRL.
+Therefore, we develop our own set of argument schemes and critical questions in the next section by analyzing transcripts of discussions about the traffic simulator.

### 2017-03-14 discussion.tex

\section{Discussion}
\label{sect:goalmodeling:discussion}

\subsection{Related work}
\label{sect:goalmodeling:relatedwork}

The need for justifications of modeling choices plays an important role in different requirements engineering methods using goal models. High-level goals are often understood as reasons for representing lower-level goals (in other words, the need for low-level goals is justified by having high-level goals) and other elements in a goal model such as tasks and resources. Various refinements and decomposition techniques,  often used in requirements engineer (See~\cite{EP-2001-Lamsweerde-GORE} for an overview), can be seen as incorporating argumentation and justification, in that sub-goals could be understood as arguments supporting parent goals. In that case, a refinement alternative is justified if there are no conflicts between sub-goals (i.e., it is consistent), as few obstacles as possible sub-goals harm sub-goal achievement, there are no superfluous sub-goals (the refinement is minimal), and the achievement of sub-goals can be verified to lead to achieving the parent goal (if refinement is formal~\cite{Darimont:1996:FRP:239098.239131}). This interpretation is one of the founding ideas of goal modeling. However, while this interpretation may seem satisfactory, argumentation and justification processes differ from and are complementary to refinement in several respects, such as limited possibilities for rationalization and lack of semantics (see Jureta~\cite{Jureta:RE2008} for more details).

There are several contributions that relate argumentation-based techniques with goal modeling. The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}. Jureta \emph{et al.} propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models. One of the elements of GAM is the translation of formal argument models to goal models (similar to ours). In this sense, our RationalGRL framework can be seen as an instantiation and implementation of  part of the GAM. The main difference between our approach and GAM is that we integrate arguments and goal models using argument schemes, and that we develop these argument schemes by analyzing transcripts. GAM instead uses structured argumentation.

The RationalGRL framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact. DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs. One of the main improvements of RationalGRL over DR approaches is that RationalGRL incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.

Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements. However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks. Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented. Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.

\subsection{Open issues}
\label{sect:goalmodeling:openissues}

We see a large number of open issues that we hope will be explored in future research. We discuss five promising directions here.

\subsubsection*{Architecture principles} 
One aspect of enterprise architecture that we did not touch upon in this article are \emph{(enterprise) architecture principles}. Architecture principles are general rules and guidelines, intended to be enduring and seldom amended, that inform and support the way in which an organization sets about fulfilling its mission~\cite{Lankhorst2005,OptLand2007a,OG2009}. They reflect a level of consensus among the various elements of the enterprise, and form the basis for making future IT decisions. Two characteristics of architecture principles are:
\begin{itemize}
\item There are usually a small number of principles (around 10) for an entire organization. These principles are developed by enterprise architecture, through discussions with stakeholders or the executive board. Such a small list is intended to be understood \emph{throughout the entire organization}. All employees should keep these principles in the back of their hard when making a decision.
\item Principles are meant to guide decision making, and if someone decides to deviate from them, he or she should have a good reason for this and explain why this is the case. As such, they play a normative role in the organization.
\end{itemize}

Looking at these two characteristics, we see that argumentation, or justification, plays an important role in both forming the principles and adhering to them:
\begin{itemize}
\item Architecture principles are \emph{formed} based on underlying arguments, which can be the goals and values of the organization, preferences of stakeholders, environmental constraints, etc.
\item If architecture principles are \emph{violated}, this violation has to be explained by underlying arguments, which can be project-specific details or lead to a change in the principle.
\end{itemize}

In a previous paper, we~\cite{marosin-etal:caise2016} propose an extension to GRL based on enterprise architecture principles. We present a set of requirements for improving the clarity of definitions and develop a framework to formalize architecture principles in GRL. We introduce an extension of the language with the required constructs and establish modeling rules and constraints. This allows one to automatically reason about the soundness, completeness and consistency of a set of architecture principles. Moreover, principles can be traced back to high-level goals.

It would be very interesting future work to combine the architecture principles extension with the argumentation extension. This would lead to a framework in which principles cannot only be traced back to goals, but also to underlying arguments by the stakeholders.

\subsubsection*{Extensions for argumentation}

The amount of argumentation theory we used in this article has been rather small. Our intention was to create a bridge between the formal theories in argumentation and the rather practical tools in requirements engineering. Now that the initial framework has been developed, is it worth exploring what tools and variations formal argumentation has to offer in more detail.

For instance, until now we have assumed that every argument put forward by a critical questions always defeats the argument it questions, but this is a rather strong assumption.  In some cases, it is more difficult to determine whether or not an argument is defeated. Take, for example, the argumentation framework in Figure~\ref{fig:goalmodeling:futureargs} with just A1 and A2. These two arguments attack each other, they are alternatives and without any explicit preference, and it is impossible to choose between the two. It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}. If we say that we prefer the action  \texttt{Create new cars} (A2) over the action  \texttt{Keep same cars} (A1), we remove the attack from A1 to A2. This makes A2 the only undefeated argument, whereas A1 is now defeated. It is also possible to give explicit arguments for preferences~\cite{modgil2009}. These arguments are then essentially attacks on attacks. For example, say we prefer A3 over A1 because `it is important to have realistic traffic flows' (A4). This can be rendered as a separate argument that attacks the attack from A1 to A3, removing this attack and making $\{$A3, A4$\}$ the undefeated set of arguments.

Allowing undefeated attacks also make the question of which semantics to choose more interested. In our current (a-cyclic) setting, all semantics coincide, and we always have the same set of accepted arguments. However, once we allow for cycles, we may choose accepted arguments based on semantics which, for instance, try to accept/reject as many arguments as possible (preferred semantics), or just do not make any choice once there are multiple choices (grounded). Another interesting element of having cycles is that one can have multiple extensions. This corresponds to various \emph{positions} are possible, representing various sets of possibly accepted arguments. Such sets can then be shown to the user, who can then argue about which one they deem most appropriate.

\begin{figure}[ht]
\centering
\begin{tikzpicture}
        \node[minimum size=1cm] (a1) [argNodeIN] at (0,0) {$A_1$};
        \node[minimum size=1cm] (a2) [argNodeIN] at (3,0) {$A_2$};

         \path
    (a2) edge [attackLink] (a1)
    (a1) edge [attackLink] (a2);
\end{tikzpicture}
\caption{Preferences between arguments}
\label{fig:goalmodeling:futureargs}
\end{figure}

Finally, in this article we have only explored one single argument scheme, but there are many other around. In his famous book ``Argumentation schemes'', Walton describes a total of 96 schemes. Murukannaiah \emph{et al.}~\cite{murukannaiah2015} already explain how some of these schemes may be use for resolving goal conflicts, and it is worth studying what this would look like in our framework as well.

\subsubsection*{Empirical study}

Although we develop our argument schemes and critical questions with some empirical data, we did not yet validate the outcome. This is an important part, because it will allow us to understand whether adding arguments to goal modeling is actually useful. We have developed an experimental setup for our experiment, which we intend to do during courses at various universities. However, we cannot carry out this experiment until the tool is finished.

\subsubsection*{Formal framework}

The formal framework we present in this article is very simple, and does not provide a lot of detail. We believe it would be interesting to develop a more robust characterization of a GRL model using logical formulas. Right now, we have no way to verify whether the goal models we obtain through out algorithms are actually valid GRL models. This is because we allow any set of atoms to be a GRL model, which is clearly very permissive and incorrect. Once we develop a number of such constraints, we can ensure (and even proof) our algorithms do not generate invalid GRL models.

For instance, suppose we assert that an \emph{intentional element} is a goal, softgoal, task, or resource: \begin{align*}
(softgoal(i)\vee goal(i)\vee task(i)\\
\vee resource(i)\rightarrow IE(i).
\end{align*}
We can then formalize an intuition such as: ``Only intentional elements can be used in contribution relations'' as follows
\begin{align*}
contrib(k,i,j,ctype)\rightarrow \\
(IE(i)\wedge IE(j)\wedge IE(j).
\end{align*}

Interestingly, such constraints are very comparable to \emph{logic programming} rules. We therefore see it as interesting future research to explore this further, specifically in the following two ways:
\begin{itemize}
\item Develop a set of constraints on sets of atoms of our language, which correctly describe a GRL model. Show formally that using our algorithms, each extension of the resulting argumentation framework corresponds to a valid GRL model, i.e., a GRL model that does not violate any of the constraints.
\item Implement the constraints as a logic program, and use a logic programming language to compute the resulting GRL model.
\end{itemize}

\subsection{Conclusion}

In this article, we develop the RationalGRL framework to trace back elements of GRL models to arguments using in discussions between stakeholders. The contributions of this article have become increasingly more formal. First, we analyze transcripts of meetings about the development of an information system. We count occurrences of argument schemes and critical questions, and categorize them by the effect of instantiating them. Then, we create a visual notation to link arguments and questions to elements of a goal model. If an argument for an element of a goal model is rejected, then the corresponding element is disabled. Similarly, if the argument is accepted, then the corresponding element is enabled. Finally, we formalize the argument schemes and critical questions in a logical framework. We propose a formal language to describe a GRL model, and we develop various algorithms for applying our argument schemes and critical questions.

Our framework is one of the first attempts that try to combine the formal theory of argumentation with the practical frameworks of goal modeling. History has shown that it is remarkably difficult to combine formal results with practice. By taking argument schemes as a starting point, we believe we have chosen the right level of formality, but as the open issues section shows, much work is left to be done.

### 2017-03-14 formalframework.tex

@@
-\section{Formalization of RationalGRL Framework}
+\section{RationalGRL: the logical framework}
 \label{sect:formalframework}
 
-%SG: make sure the first paragraph is aligned with the previous section
-In Section~\ref{sect:gmas}, we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of an information system.
-The resulting list is shown in table~\ref{table:argument-schemes}.
-We also discussed various examples of applications of the different critical questions and all four different effects (right column of Table~\ref{table:argument-schemes}): INTRO, DISABLE, REPLACE, and ATTACK.
+In the previous section we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of a traffic simulator.
+The resulting list is shown in Table~\ref{table:argument-schemes}.
+We also discussed various examples of applications of the different critical questions and all four different effects (right column of Table~\ref{table:argument-schemes}): \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE}, and \textsf{ATTACK}.
 
-The examples and corresponding visualizations of the previous section provided some insight on how to formalize argument schemes, critical questions, and their relationship with goal models.
+The examples and corresponding visualizations of the previous section provide some insight on how to formalize argument schemes, critical questions, and their relationship with goal models.
 However, if we are to implement our framework in a tool, we require a more precise formalization of these concepts.
-We choose to use formal logic to specify this, since argumentation has been studied in formal logic extensively for the past decades.
-%SG: any citation?
+We choose to use formal logic to specify this.
+This is also a good fit with formal argumentation, which has flourished in the past decades, leading to a large number of different semantics that we can choose from.
 
 The rest of this section is as follows: In the first subsection, we develop a formal language to specify a GRL model.
-This language consists of propositions only, and a GRL model can be specified as a logic program directly.
-In the second subsection, we provide formal definitions of all argument schemes.
-An instantiation of an argument scheme consists of a set of statements about a GRL model.
-Finally, in the third subsection, we formalize the critical questions as several algorithms.
+This language consists of atomic sentences.
+In the second subsection, we develop our notion of an argument: an argument is simply a set of atoms from our language.
+In the third subsection, we develop algorithms for all the argument schemes and critical questions.
 
-\subsection{Formal Language for RationalGRL}
+\subsection{Logical Language for RationalGRL}
+\label{sect:goalmodeling:logicallanguage}
 
-Basic concepts of propositional logic are adequate to define the entities in a GRL model and the relations between them. We begin with the representation for the elements.
+We define our language as a set of atoms, that is, grounded formulas (without variables) with no logical connectives or negation.
 
-\begin{definition}[RationalGRL elements]
-Each element of a goal model is identified with a non-negative integer. Formally: let $Actors\cup IEs \cup Links \subseteq \mathbb{N}$ denote the set of elements of a goal model, where 
+\begin{definition}[GRL Atoms]
+Let $P$ be a set of sentences representing names for GRL elements. The \emph{set of GRL atoms} $At$ is defined as follows: 
+\begin{align*}
+At = \{&actor(i), softgoal(i), goal(i), task(i),\\
+&resource(i), decomp(k,i,J,dtype),\\
+&dep(k,i,j),contr(k,i,j,ctype),\\
+&has(i,j),disabled(i),name(i,p)\},
+\end{align*}
+where
 \begin{itemize}
-\item $Actors$ is the set of actors, 
-\item $IEs$ is the set of intentional elements (softgoals, goals, tasks, and resources),
-\item $Links$ is the set of links (decomposition, contribution, and dependency). %what about correlation link?
-\end{itemize}
-
-The three sets are pairwise disjoint, i.e. $Actors \cap IEs = Actors \cap Links = IEs \cap Links = \emptyset$. 
-\end{definition}
-
-\rationale{} We identify a GRL element with an identifier and not with a name, since it is possible that the name of the element changes, for instance by applying critical question CQ12a (clarification), see the example ``Clarify task \texttt{Road pattern}'' above.
-The GRL model of Figure~\ref{fig:examples:decomposition} can be formalized as follows:
-\begin{align*}
-&Actors = \{1\},\\
-&IEs = \{2, 3, 4\},\\ 
-&Links = \{5\}.
-\end{align*}
-
-\begin{definition}[GRL intentional element type]
-For an intentional element $i\in IEs$, if $i$ is a softgoal, goal, task, or resource, then this is respectively denoted by $softgoal(i), goal(i), task(i)$, and $resource(i)$.
-\end{definition}
-
-\rationale{} All intentional elements are collected in the set $IEs$.
-In order to distinguish softgoals, goals, tasks, and resources, we use the corresponding propositions.
-In our example this looks as follows:
-\begin{align*}
-&goal(2),\\
-&task(3),\\
-&task(4).
-\end{align*}
-
-Note that we do not require such propositions for actors, since there is only one type of actor, and all actors are contained in the set $Actors$.
-
-\begin{definition}[Links]
-Given a link $i\in Links$, we denote the links types as follows:
-\begin{itemize}
-\item $contrib(i,src,dest,type)$ is a contribution from $src\in IEs$ to $dest\in IEs$, where $type\in\{+,-\}$ means a positive resp. negative contribution. 
-\item $decomp(i,src,\{dest_1,\ldots,dest_n\},type)$ is a decomposition of $src\in IEs$ into $\{dest_1,\dots,dest_n\}\subseteq IEs$, where $type\in\{AND,OR,XOR\}$ means respective an AND, OR, and XOR decomposition.
-\item $dep(i,src,dest)$ is a dependency from $dest\in IEs$ to $dest\in IEs$.
+\item $\{i,j,k\}\cup J \subseteq\mathbb{N}$
+\item $p\in P$
+\item $dtype\in\{and,or,xor\}$
+\item $ctype\in\{+,-\}$
 \end{itemize}
@@
 
-\rationale{} Similarly to the intentional elements, we distinguish between the different types of links between IEs using the corresponding propositions.
-The formalization of the link in Figure~\ref{fig:examples:decomposition} looks as follows:
-\begin{align*}
-&decomp(5, 2, \{3,4\}, OR)
-\end{align*}
+The non-negative integers associates with each element and relation are identifiers.
+We thus choose to identify each element and relationship in a GRL model with an identifier.
+This allows us to change the name of an element while still being able to make different statements about this element, for instance by applying critical question CQ12a (clarification, see the example ``Clarify task \texttt{Road pattern}'' above).
 
-\begin{definition}[Name of IE or Actor]
-The name $p$ of an element $i\in IEs\cup Actors$ is denoted by $name(i,p)$.
+We now briefly explain each atom in more detail.
+\begin{itemize}
+\item $actor(i)$: Identifier $i$ is an actor.
+\item $softgoal(i)$: Identifier $i$ is a softgoal.
+\item $goal(i)$: Identifier $i$ is a goal.
+\item $task(i)$: Identifier $i$ is a task.
+\item $resource(i)$: Identifier $i$ is a resource.
+\item $decomp(k,i,J,dtype)$: Identifier $k$ is a $dtype$-decomposition (and/or/xor) from the element corresponding to identifier $i$ to the set of elements corresponding to identifiers $J$.
+\item $dep(k,i,j)$: Identifier $k$ is a dependency link from the element corresponding to identifier $i$ to the element corresponding to identifier $j$.
+\item $contr(k,i,j,ctype)$: Identifier $k$ is a $ctype$-contribution (+/-) from the element corresponding to identifier $i$ to the element corresponding to identifier $j$.
+\item $has(i,j)$: Identifier $i$ (which is an actor) has the element corresponding to identifier $j$.
+\item $disabled(i)$: The element or relationships corresponding to identifier $i$ is disabled.
+\item $name(i,p)$: The name of the element corresponding to identifier $i$ is $p$.
+\end{itemize}
+
+\begin{remark}
+From these descriptions one can observe that many of the atoms come with a number of assumptions and dependencies.
+For instance, if $has(i,j)$ is true, then $i$ is an actor and $j$ is an element.
+We could formalize this as follows: $$has(i,j)\rightarrow (actor(i) \wedge (softgoal(j)\vee goal(j)\vee task(i) \vee resource(j))).$$ It would be possible to enumerate all such properties in order to correctly specify a GRL model.
+One could then formally verify whether a set of atoms violates these constraints.
+If not, it is a ``valid'' representation of a GRL model.
+Since the focus of this article is not on a logical analysis, but rather on developing a framework for empirical data, we leave such a formal analysis for future work (see Section~\ref{sect:goalmodeling:openissues}).
+\end{remark}
+
+Using this formalization, it is rather straightforward to specify a GRL model.
+An example of the specification of the GRL model in Figure~\ref{fig:examples:decomposition} is shown in Table~\ref{table:grl_atom_spec}.
+The specification has been written in logic programming style.
+In this article we do not make use of any logic programming techniques but this would be interesting future work (see Section~\ref{sect:goalmodeling:openissues}).
+A more elaborate example of a specification is shown in Appendix~\ref{ch:grlspec}, showing a complete specification of the traffic simulator GRL model in Figure~\ref{fig:trafficsim}.
+
+\begin{table}[h!]
+\centering
+\begin{tabularx}{0.4\textwidth}{|X|}
+\hline
+\texttt{goal(0).}\\
+\texttt{task(1).}\\
+\texttt{task(2).}\\
+\texttt{name(0,simulate).}\\
+\texttt{name(1,static\_simulation).}\\
+\texttt{name(2,dynamic\_simulation).}\\
+\texttt{decomp(3,0,[1,2],or).}\\
+\hline
+\end{tabularx}
+\caption{Specification of the GRL model in Figure~\ref{fig:examples:decomposition}}
+\label{table:grl_atom_spec}
+\end{table}
+
+\subsection{Formal argumentation semantics}
+\label{sect:goalmodeling:argumentationsemantics}
+
+In the previous subsection we introduced an atomic language to specify a GRL model.
+In this subsection we give a formal definition of an argument, which is simply a set of atoms from our language.
+We introduce Dung's acceptability semantics as well, which allows us to determine sets of acceptable arguments.
+
+\begin{definition}[Argument]
+An \emph{argument} $A\subseteq Att$ is a set of atoms from $Att$.
 \end{definition}
 
-\rationale{} We formalize the description of an actor or an intentional element with a proposition $name(i,p)$. Continuing our example of formalizing Figure~\ref{fig:examples:decomposition}, we have 
-\begin{align*}
-&name(1,\text{system}),&\\
-&name(2,\text{simulate}),&\\
-&name(3,\text{static\ simulation}),&\\
-&name(4, \text{dynamic simulation}).&
-\end{align*}
-Note the decomposition link with identifier 5 does not have a name, since decomposition links are already distinguished by their type (OR, XOR, AND).
+This simple definition allows us to form arguments about (parts of) a GRL model.
+For instance, $$\{goal(0),name(0,development\_team)\},$$ is an argument.
+We next introduce an argumentation framework, which is a set of arguments an attack relations between arguments.
 
-\begin{definition}[Elements of an actor]
-Given an actor $i\in Actors$ and an element $j\in IEs\cup Links$, we use $has(i,j)$ to denote that element $j$ belongs to actor $j$.
-\end{definition}
-
-\rationale{} In order to denote that an intentional element or a link belongs to an actor, we use $has$ statements. This looks as follows in our example:
-\begin{align*}
-&has(1, 2),\\
-&has(1, 3),\\
-&has(1, 4),\\
-&has(1, 5).
-\end{align*}
-
-\begin{definition}[Disabled GRL element]
-A disabled element $i\in Actors\cup IEs$ is denoted by $disabled(i)$.
-\end{definition}
-
-\rationale{} If a GRL element corresponds to an argument that is rejected, then this argument should be disabled.
-For instance, the GRL model in Figure~\ref{fig:examples:relevant-actor} can be formalized as follows:
-\begin{align*}
-&Actors = \{1\},\\
-&name(1, \text{development team}),\\
-&disabled(1)
-\end{align*}
-
-This concludes our definitions to formalize a GRL model.
-An advantage of our approach is that we can formalize a GRL model using a set of facts, which can be directly be formalized as a logic program.
-We will come back to this in the last part of this section.
-
-\subsection{Formal argumentation}
-
-In this subsection, we formalize an argument as a set of statements about a GRL model in the language we introduced in the previous section.
-First, we formalize an argumentation framework and its semantics, which is used to determine sets of acceptable arguments.
-
-\begin{definition}[Argumentation framework]
+\begin{definition}[Argumentation framework~\cite{Dung1995}]
 An argumentation framework $AF=(Args,Att)$ consists of a set of arguments $Args$ and an attack relationship $Att:Args\times Args$, where $(A_1,A_2)\in Att$ means that argument $A_1\in Args$ attacks arguments $A_2\in Args$.
@@
 
-\rationale{} This is the standard definition used by Dung~\cite{Dung1995}. For instance, the argumentation framework from figure~\ref{fig:pras:example} can be formalized as:
-\begin{align*}
-&Args=\{A_1,A_2,A_3\}\\
-&Att = \{(A_2,A_1),(A_1,A_3),(A_3,A_1)\}
-\end{align*}
+Now that we have defined arguments and their attacks, we are going to define a semantics to determine which arguments are acceptable. The following notions are preliminary to this.
 
-\todo{Marc}{Marc}{explain semantics}
+\begin{definition}[Attack, conflict-freeness, defense, and admissibility~\cite{Dung1995}] Suppose an argumentation framework $AF=(Arg,Att)$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
+We say that
+\begin{itemize}
+\item $S$ \emph{attacks} $A$ if some argument in $S$ attacks $A$,
+\item $S$ \emph{attacks} $S'$ if some argument in $S$ attacks some argument in $S'$,
+\item $S$ is \emph{conflict-free} if it does not attack itself,
+\item $S$ \emph{defends} $A$ if $S$ attacks each attack against $A$.
+\item $S$ is \emph{admissible} if $S$ is conflict-free and defends each argument in it.
+\end{itemize}
+\end{definition}
 
-\subsection{Algorithms for instantiating the argument schemes}
+\begin{figure}[ht!]
+\centering
+\begin{tikzpicture}
+        \node[minimum size=1cm] (a) [argNodeIN] at (-2,-2) {$A_1$};
+        \node[minimum size=1cm] (b) [argNodeIN] at (-2,0) {$A_2$};
+        \node[minimum size=1cm] (c) [argNodeIN] at (0,0) {$A_3$};
+        \node[minimum size=1cm] (d) [argNodeIN] at (2,0) {$A_4$};
+        \node[minimum size=1cm] (e) [argNodeIN] at (2,-2) {$A_5$};
 
-%SG: a set of statements mean argument schemes?
-I think it is better to also mention the argument scheme in the text and then show the algo...
-so the reader doesn't need to always move between the table and the algo
-We formalize an argument as a set of statements about a GRL model in the language we introduced in the previous section.
-For instance, $A= \{actor(0), name(0,student)\}$ is an argument stating that identifier 0 represent an actor, and that the name of this actor is \emph{Student}.
-An argument scheme then corresponds to adding one or more arguments to the set of arguments.
+         \path
+    (a) edge [attackLink] (b)
+    (c) edge [attackLink] (b)
+    (c) edge [attackLink] (d)
+    (d) edge [attackLink] (c)
+    (d) edge [attackLink] (e)
+    (e) edge [attackLink, loop right] (e);
+\end{tikzpicture}
+\caption{Example argumentation framework.}
+\label{fig:goalmodeling:arg2}
+\end{figure}
+
+Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}.
+We say that the set $\{A_1,A_4,A_5\}$ \emph{attacks} argument $A_2$, because $A_1$ attacks $A_2$.
+However, $\{A_1,A_4,A_5\}$ is not \emph{conflict-free}, because $A_4$ attacks $A_5$, so this set of arguments attacks itself.
+If we remove $A_5$ from this set, then it is conflict-free.
+In total, all the following sets are conflict-free in Figure~\ref{fig:goalmodeling:arg2}: $$\{A_1,A_3\},\{A_1,A_4\},\{A_2,A_4\},\{A_1\},\{A_2\},\{A_3\},\{A_4\},\emptyset.$$
+
+However, not all of these sets are \emph{admissible}.
+For instance, the set $\{A_2,A_4\}$ is not admissible, because $A_1$ attacks this set, but the set does not defend itself against this attack.
+The set $\{A_1,A_3\}$ is admissible, because its only attacker, $A_4$ is attacked by $A_3$.
+
+There are a large number of different semantics using these notions to determine which arguments are acceptable, such as the \emph{grounded}, the \emph{preferred}, the \emph{stable}, and the \emph{complete} semantics.
+However, in this article our argumentation frameworks are very simple, in the sense that they do not introduce cycles.
+In future work, we aim to extend this by using preferences (see open issues in Section~\ref{sect:goalmodeling:openissues}).
+An advantage of our current approach is that all semantics coincide when there are no cycles, which simplifies our case.
+We use the preferred semantics here, but we could equivalently have chosen any other version.
+
+\begin{definition}[Preferred semantics~\cite{Dung1995}] 
+Suppose an argumentation framework $AF=(Arg,Att)$. A set of arguments $S\subseteq Arg$ is a \emph{preferred extension} if and only if 
+\begin{itemize}
+\item $S$ is an admissible set of argument,
+\item For each admissible set of arguments $T\subseteq A$: $S\not\subset T$.
+\end{itemize}
+\end{definition}
+
+In our example in Figure~\ref{fig:goalmodeling:arg2}, the preferred extensions are $\{A_1,A_3\}$ and $\{A_1,A_4\}$.
+
+\subsection{Algorithms for argument schemes and critical questions}
+
+Let us briefly review what we have done so far. Up until now we have achieve the following:
+\begin{itemize}
+\item We (informally) developed a set of argument schemes and critical questions in Section~\ref{sect:gmas}, by annotating transcripts.
+\item We developed a visual notation to illustrate some argument schemes and critical questions in Section~\ref{sect:gmas:examples}.
+\item We proposed a logical language consisting of atomic sentence to describe a GRL model in Section~\ref{sect:goalmodeling:logicallanguage}.
+\item We formalized an ``argument'' as a set of atoms from our language, and we introduced \emph{argumentation semantics} to compute sets of accepted arguments in Section~\ref{sect:goalmodeling:argumentationsemantics}.
+\end{itemize}
+
+We see that we are becoming increasingly more formal in each step of the progress.
+What we haven't formalized properly yet are the argument schemes and critical questions, which is what we turn to now.
+
+Both the application of an argument schemes and the answering of critical questions are \emph{procedural} or \emph{dialectic} in nature.
+Therefore, a natural choice of formalizing them seems to develop procedures, or algorithms for them, which is what we do here.
 
 In the following algorithms, we assume the following global variables:
+
+\begin{definition}[Global variables] The following variables are intended to have a \emph{global} scope and have a \emph{static} value, meaning the lifetime of the variable extends across the entire run of the program.
 \begin{itemize}
-\item $id$ the current highest identifier of the elements. This variable is increased for each new element that is added.
-\item $Args$ contains the set of all arguments.
-\item $Att$ contains the set of all attack relations.
+\item $id$: the current highest identifier of the elements. This variable is increased for each new element that is added.
+\item $Args$: contains the set of all arguments.
+\item $Att$: contains the set of all attack relations.
 \end{itemize}
+\end{definition}
+
+\subsubsection*{Algorithms for argument schemes}
+The algorithms for the argument schemes are very simple, because they simply consist of forming a new argument and adding it to the set of arguments. No attack relations are introduced.
 
@@
     \Procedure{$AS_0$}{$a$}
-    \State $id\gets id+1$
-    \State $A \gets \{actor(id), name(id,a)\}$
-    \State $Args \gets Args \cup A$
+    \State $id\gets id+1$\label{alg:as0:1}
+    \State $A \gets \{actor(id), name(id,a)\}$\label{alg:as0:2}
+    \State $Args \gets Args \cup A$\label{alg:as0:3}
     \EndProcedure
@@
 
-\rationale{} Algorithm~\ref{alg:as0} takes one argument, namely the name of the actor $a$.
-In Figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_(\text{Development team})$, results in one argument: $Args = \{\{actor(0), name(0, \text{Development team})\}\}$.
+\emph{Algortihm~\ref{alg:as0} for argument scheme AS0}: The algorithm takes one argument, namely the name of the actor $a$.
+On line~\ref{alg:as0:1} of the algorithm, the global variable $id$ is increased by one.
+This ensures that each new argument has a unique identifier.
+On line~\ref{alg:as0:2}, the argument for the actor is formed, consisting of two statements, stating respectively that $id$ is an actor, and that the name of this actor is $a$.
+Finally, on line~\ref{alg:as0:3} this argument is added to the global set of arguments $Args$.
+In Figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_0(\text{Development team})$, results in one argument: $$Args = \{\{actor(0), name(0, \text{Development team})\}\}.$$
 
@@
 
-\rationale{}  This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
+\emph{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
+The algorithm is similar to the previous one, with the difference that the newly added argument contains the statement $has(a_{id},id)$ as well, meaning that the actor with id $a_{id}$ has element $id$ (which is a resource).
+As an example, let us formalize the argument corresponding to resource \emph{External library} of actor \emph{Traffic tycoon} in Figure~\ref{fig:transcripts:grl}.
+First, we assume some id is associated with the actor: $$\{actor(0),name(0,traffic\_tycoon)\}.$$ Then we can formalize the argument for the resource as follows: $$\{resource(1),name(1,external\_library),has(0,1)\}.$$
 
-The formalization of argument scheme AS2 to AS4 are exactly the same as AS1 but it needs to replace the actor with tasks, goals, and softgoals, respectively.
-Therefore, we have omitted them here.
-We continue with argument scheme AS5.
+Argument scheme AS1 to AS4 are all very similar, in the sense that they all assert that some element belongs to an actor.
+Therefore, we only provide the algorithm for AS1 and we assume the reader can easily construct the remaining algorithms AS2-AS4.
 
@@
     \Procedure{$AS_5$}{$g_{id}, \{T_1,\ldots,T_n\}, type$}
-    \State $T_{id} = \emptyset$
+    \State $T_{id} = \emptyset$\label{alg:as5:1}
     \For{$T_i$ in $\{T_1,\ldots,T_n\}$}
-      \If{$\exists_{A\in Args} \{task(t_{id}),name(t_{id},T_i)\}\subseteq A$}
+      \If{$\exists_{A\in Args}\label{alg:as5:2} \{task(t_{id}),name(t_{id},T_i)\}\subseteq A$}
         \State $T_{id} \gets T_{id} \cup \{t_{id}\}$
-      \Else
+      \Else\label{alg:as5:3}
         \State $id\gets id+1$
@@
     \EndFor
-    \State $id\gets id+1$
+    \State $id\gets id+1$\label{alg:as5:4}
     \State $A\gets \{decomp(id, g_{id}, T_{id}, type)\}$
@@
 
-\rationale{} The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
-The for loop in the algorithm iterates over all tasks in list $T$, and if a task does not exist yet, an argument is added for it.
-This ensures that all tasks have corresponding arguments.
-Note that the algorithm assumes the goal already exists.
+\emph{Algorithm~\ref{alg:as5} for argument scheme AS5:} The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
+The difficulty of this algorithm is that each of the tasks are stated in natural language, and it is not directly clear whether these tasks are already in the GRL model or not.
+Therefore, we have to check for each tasks whether it already exists, and if not, we have to create a new task.
+On line~\ref{alg:as5:1}, the set $T_{id}$ is initialized, which will contain the ids of the tasks $T_1,\ldots,T_n$ to decompose into.
+In the for loop, the if statement on line~\ref{alg:as5:2} checks whether some argument already exists for the current task $T_i$, and if so, it simply adds the identifier of the task ($t_{id}$) to the set of task identifiers $T_{id}$.
+Otherwise (line~\ref{alg:as5:3}), a new task is created and the new identifier $id$ is added to the set of task identifiers.
+After the for loop on line~\ref{alg:as5:4}, an argument for the decomposition link itself is constructed, and it is added to the set of arguments $Args$.
+
+Let us explain this algorithm with the XOR-decomposition of goal \texttt{Generate cars} of Figure~\ref{fig:transcripts:grl}. Suppose the following arguments are constructed already:
+\begin{itemize}
+\item $\{goal(0),name(0,generate\_cars)\},$
+\item $\{taks(1),name(1,keep\_same\_cars\}.$
+\end{itemize}
+Suppose furthermore that someone wants to put forward the argument that goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create news cars}.
+Formally: $AS_5(0,\{generate\_cars,keep\_same\_cars\},xor)$.
+The algorithm will first set $T_{id}=\emptyset$, and then iterate over the two task names.
+For the first task $generate\_cars$, there does not exist an argument $\{task(t_{id}),name(t_{id},generate\_cars)\}$ yet, so a  new argument is created.
+Suppose the following argument is created: $\{task(2),name(2,generate\_cars)\}.$ After this, 2 is added to $T_{id}$.
+For the second task an argument exists already, namely $\{task(1),name(1,keep\_same\_cars)\}$, so 1 is simply added to $T_{id}$.
+After the for loop, we have $T_{id}=\{1,2\}$.
+Next, the decomposition argument is created, which is $\{decomp(3,0,\{1,2\},xor)\}$.
+This argument is added to $Args$ and the algorithm terminates.
 
@@
 
-\rationale{} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
-The if statements check if the softgoal exists already, and if not, an argument is added for it.
+\emph{Algorithm~\ref{alg:as6} for argument scheme AS6:} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
+The idea behind this algorithm is very similar to the previous one.
+First, the if statements check whether the softgoal exists already, and if not, an argument is added for it.
 This ensures that all softgoals have corresponding arguments.
-Note that the algorithm assumes the task already exists.
+After the if statement, the argument for the contribution link is created and it is added to the set of arguments $Args$.
 
-The algorithms for $AS7$ to $AS11$ all have a very similar structure as Algorithm~\ref{alg:as6} and have therefore been omitted from this article.
+Let us again illustrate this with a simple example from Figure~\ref{fig:transcripts:grl}.
+Suppose the following argument exists already: $\{task(0),name(0,keep\_same\_cars\}$, and suppose someone would like to add an argument that the task \texttt{Keep same cars} contributes positively to softgoal \texttt{Dynamic simulation}, i.e.
+$AS_6(0,dynamic\_simulation)$.
+The algorithm first checks whether an argument already exists for the softgoal, and when it finds out it does not exist, creates the argument $\{softgoal(1),name(1,dynamic\_simulation)\}$ and adds it to $Args$.
+Then, the argument for the contribution is added to $Args$ as well, which is $\{contr(2,0,1,pos)\}$.
 
-%SG: up to here.
-\subsection{Algorithms for answering the critical questions}
+\emph{Algorithms for argument schemes AS7-AS11:} The algorithms for $AS7$ to $AS11$ all have a very similar structure as Algorithm~\ref{alg:as6} and have therefore been omitted.
+Again, we assume the reader can reconstruct them straightforwardly.
 
-We now develop algorithms for our critical questions.
-We start with the critical questions that have an effect DISABLE.
-All critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12 fall into this category.
-The algorithm below applies when the critical question is answered affirmatively.
+\subsubsection*{Algorithms for critical questions}
+
+We now develop algorithms for our critical questions. Recall that answering a critical question can have four effects, and we discuss each of these effects separately.
 
@@
 
-\rationale{} We the see disable operation is quite straightforward. It simply consists of adding an argument stating argument $i$ is disabled.
+\emph{Algorithm~\ref{alg:disable} (\textsf{DISABLE}) for critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12:} The disable operation is very straightforward: It simply consists of adding an argument stating the GRL element with identifier $i$ is disabled.
+Let us reconsider the example of Figure~\ref{fig:examples:relevant-actor}.
+This example consists of an instantiation of argument scheme AS0, which is attacked by an argument that resulted from answering critical question.
+The formalization of this scenario is shown in Figure~\ref{fig:examples:relevant-actor:formal}.
+Interestingly, we see that the \textsf{DISABLE} operation no longer leads to an attack between the two arguments.
+Instead, the only thing that this operation does is adding another argument stating the element is disabled.
 
-Next we consider the REPLACE operation. This operation is less straightforward and cannot be captured by a single algorithm, since the operation depends on the type of argument scheme.
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeIN] at (-3.7,-0.2) {
+        	\smallargtext{AS0}{$\{actor(0),name(0,dev\_team\}$}
+        } ;
+        \node (a1) [argNodeIN] at (-3.7,-2.5){
+        	\argtext{DISABLE}{$\{disabled(0)\}$}
+        };
+         \path
+    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1);
+\end{tikzpicture}
+\caption{Formalization of the arguments in Figure~\ref{fig:examples:relevant-actor}.}
+\label{fig:examples:relevant-actor:formal}
+\end{figure}
 
@@
 
-\rationale{} Algorithm~\ref{alg:cq5b} is executed when critical question CQ5b is answered, which is a critical question for argument scheme AS5.
-Therefore, it assumes an argument for a goal decomposition of the form $decomp(d, g_{id}, \{i_1,\ldots,i_n\},type)$ as already been generated (see algorithm~\ref{alg:as5}).
-The goal of the algorithm is to generate a new argument of the form $decomp(d, g_{id}, \{i_1,\ldots,i_5\}\cup \{j_1,\ldots,j_k\}, type)$, where $\{j_1,\ldots,j_k\}$ are the identifiers of the additional decomposing tasks $\{t_1,\ldots,t_k\}$.
-The first for loop in line~\ref{alg:cq5b:for1} check whether any of these tasks already exist, and if not, adds corresponding arguments for them.
-The list of all decomposing tasks is stored in $T_{id}$.
-The new argument for the decomposition link is then created in line~\ref{alg:cq5b:newarg}.
-Finally, the last for loop on line~\ref{alg:cq5b:for2} ensures that all previous arguments for decomposition links are attacked by the new argument.
-This ensures that only the new decomposition link will be accepted.
+\emph{Algorithm~\ref{alg:cq5b} (\textsf{REPLACE}) for critical questions CQ5b:} This algorithm is executed when critical question CQ5b is answered, which is a critical question for argument scheme AS5.
+Therefore, it assumes an argument for a goal decomposition already exists of the following form (see Algorithm~\ref{alg:as5}): $$\{decomp(d, g_{id}, \{i_1,\ldots,i_n\},type).$$  The goal of the algorithm is to generate a new argument of the form $decomp(d, g_{id}, \{i_1,\ldots,i_5\}\cup \{j_1,\ldots,j_k\}, type)$, where $\{j_1,\ldots,j_k\}$ are the identifiers of the additional decomposing tasks $\{t_1,\ldots,t_k\}$.
 
-The REPLACE operations of CQ10a and CQ10b have a very similar structure as algorithm~\ref{alg:cq5b} and have therefore been omitted. We continue with the last REPLACE operation left, namely CQ13.
+The algorithm task as input the goal identifier $g_{id}$, the set of existing decomposing task identifiers $i_1,\ldots,i_n$, the decomposition type, and the names of the new tasks $t_1,\ldots,t_k$ that should be added to the decomposition.
+The first part  of the algorithm is already familiar: For each task name we check whether it already exists as an argument (line 4), and if it doesn't (line 6) we add a new argument for it.
+After the for loop (line 13), a new argument is created for the new decomposition relation (14).
+After this, the for loop on line 15 ensures that the new argument attacks all previous arguments for this decomposition (note that the variable ``$\_$'' means ``do not care'').
+Only at the very end the new argument is added (line 18), to ensure it does not attack itself after the for loop of line 15-17.
+
+An example of this algorithm is shown in Figure~\ref{fig:examples:cq5b}.\footnote{Note that part of the arguments (the statements about actors) have been omitted from the figure for readability.} Before the critical question is applied, the following arguments have been put forward:
+\begin{itemize}
+\item $\{goal(0),name(0,show\_simulation)\}$
+\item $\{task(1),name(1,generate\_traffic)\}$
+\item $\{task(2),name(2,compute\_lights)\}$
+\item $\{decomp(3,0,\{1,2\},and)\}$.
+\end{itemize}
+
+Next, Algorithm~\ref{alg:cq5b} is called as follows: $CQ5b(0,\{1,2\},and,\{show\_controls\})$.
+That is, the existing decomposition is challenged by stating that goal $show\_simulation$ not only decomposes into $generate\_traffic$ and $compute\_lights$, but it also decomposes into $show\_controls$.
+Since this task does not exist yet, it is created by the algorithm, which also ensures the new argument for the decomposition link attacks the previous argument for the decomposition link.
+
+\begin{figure*}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeIN,minimum width=1.5cm] at (-5,0) {
+        	\smallargtext{AS3}{$\{goal(0),$\\$name(0,show\_sim\}$}
+        };
+        \node (a1) [argNodeIN] at (0,0) {
+        \smallargtext{AS2}{$\{task(1),$\\$name(1,gen\_traffic\}$}
+        };
+        \node (a2) [argNodeIN] at (5,0) {
+        \smallargtext{AS2}{$\{task(3),$\\$name(2,show\_ctrls\}$}
+        };
+        \node (a2a) [argNodeIN] at (-2.5,-2) {
+        \smallargtext{AS2}{$\{task(2),$\\$name(2,comp\_lights\}$}
+        };
+        \node (a3) [argNodeOUT] at (-3,-4) {$\{decomp(3,0,\{1,2\},and\}$
+        };
+        \node (a4) [argNodeIN] at (3,-4) {$\{decomp(3,0,\{1,2,3\},and\}$
+        };
+
+         \path
+    (a4) edge [attackLink] (a3)
+    (a3) edge [CQLink, bend right=50] node [below,draw=none] {CQ5b} (a4)
+    (a4) edge [CQLink] node [left,draw=none] {CQ5b} (a2);
+\end{tikzpicture}
+\caption{Example of applying critical question CQ5b (Algorithm~\ref{alg:cq5b})}
+\label{fig:examples:cq5b}
+\end{figure*}
+
+\emph{Algorithms for critical questions CQ10a and CQ10b (REPLACE)}: These algorithms have a very similar structure as Algorithm~\ref{alg:cq5b} and have therefore been omitted.
 
@@
     \Procedure{CQ13}{$i, n$}
-    \State $id\gets id+1$
-    \State $ArgsN \gets\{A\backslash\{name(i,x)\}\mid$\par\label{alg:cq13:line3}
-    \hskip\algorithmicindent$A\in Args, name(i,x)\in A\}$
-    \State $A \gets B \cup \{name(i,n)\}$ with $B\in ArgsN$
+    \State $ArgsN \gets\{ A\in Args \mid name(i,x)\in A\}$
+    \State $B\gets B'\backslash \{name(i,\_)\}$ with $B'\in ArgsN$
+    \State $A \gets B \cup \{name(i,n)\}$
     \State $Args \gets Args \cup \{A\}$
@@
 
-\rationale{} Algorithm~\ref{alg:cq13} first collects all arguments contain a statement of the form $name(i,x)$ into the set $ArgsN$, where $x$ can by any name.
-These are all the arguments that the new argument will attack, to ensure this is the only argument for a name of this element type.
-Since all such arguments are the same, except for the name proposition, we can take any such argument to construct a new one.
-The argument is added to the set of arguments, and the for loop ensures that all previous arguments are attacked.
+\emph{Algorithms for critical question CQ13 (REPLACE):} This algorithm is used to clarify/change the name of an element.
+It takes two parameters: the element identifier $i$ and the new name $n$.
+The idea behind the algorithm is that we construct a new argument for $n$ from the previous arguments, and we only replace the $name$ atom.
+We also have to ensure that we attack all previous arguments for a name.
+On line 2, all arguments that have been put forward for this element and contain $name(i,x)$ are collected into the set $ArgsN$.
+On line 3, some arguments $B'\in ArgsN$ minus the $name$ statement is assigned to $B$, and on line 4 $B$ is joined with the new $name$ statement and stored in $A$, which is then added to the set of arguments $Args$.
+The for loop on lines 6-8 ensures all previous arguments for names of the element are attacked by the new argument.
 
-We next turn to the INTRO operations, which are very similar to the algorithms for argument schemes.
-The main difference is that when answering a critical question, we can already assume an argument has been created, so we do not have to create it again.
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}[->]
+        \node (a0a) [argNodeIN] at (-2,2) {
+        	\longargtext{AS0}{$\{actor(0),name(0,student)\}$}
+        };
+        \node (a0) [argNodeOUT] at (-2,0) {
+        	\longargtext{AS2}{$\{task(1),name(1,create\_road),has(0,1)\}$}
+        };
+        \node (a1) [argNodeOUT] at (-2,-2) {
+        	\longargtext{AS2}{$\{task(1),name(1,choose\_pattern),has(0,1)\}$}
+        };
+        \node (a2) [argNodeOUT] at (-2,-4) {
+        	\longargtext{AS2}{$\{task(1),name(1,pattern\_pref),has(0,1)\}$}
+        };
+        \node (a3) [argNodeIN] at (-2,-6) {
+        	\longargtext{AS2}{$\{task(1),name(1,road\_pattern),has(0,1)\}$}
+        };
+\begin{pgfonlayer}{background}
+         \path
+    (a1) edge[attackLink] (a0)
+    (a2) edge[attackLink] (a1)
+    (a2) edge[attackLink, bend right=20] (a0)
+    (a3) edge[attackLink] (a2)
+    (a3) edge[attackLink, bend right=20] (a1)
+    (a3) edge[attackLink, bend right=40] (a0);
+\end{pgfonlayer}
 
-\begin{algorithm}[h]
-  \caption{Answering CQ6b: ``Are there alternative ways of contributing to the same softgoal $s_{id}$?'' With: ``Yes, namely by doing task $t$''}\label{alg:cq6b}
-  \begin{algorithmic}[1]
-    \Procedure{CQ6b}{$s_{id}, t$}
-    \If{$\exists_{A\in Args} \{task(i),name(i,t)\}\subseteq A$}
-        \State $t_{id} \gets i$
-    \Else
-      \State $id\gets id+1$
-      \State $A \gets \{task(id),name(id,t)\}$
-      \State $Args \gets Args\cup A$
-      \State $t_{id} \gets id$
-    \EndIf
-    \State $id\gets id+1$
-    \State $A\gets \{contr(id, t_{id}, s_{id}, pos)\}$
-    \State $Args \gets Args\cup A$
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
+	\path
+	(a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ13} (a1)
+    (a1) edge [CQLink, bend right=50] node [left,draw=none] {CQ13} (a2)
+    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ13} (a3);
+\end{tikzpicture}
+\caption{Applying critical question CQ13 (Algorithm~\ref{alg:cq13}) to the example in Figure~\ref{fig:examples:clarification}.}
+\label{fig:examples:clarification:formal}
+\end{figure}
 
-\rationale{} Comparing algorithm~\ref{alg:cq6b} with algorithm~\ref{alg:as6} shows that the two algorithms are indeed very similar.
+An example of the working of Algorithm~\ref{alg:cq13} is shown in Figure~\ref{fig:examples:clarification:formal}.
+Let us consider the last application of CQ13 (bottom argument).
+Before this application, the following arguments have been put forward:
+\begin{itemize}
+\item $A_1$: $\{actor(0),name(0,student)\}$
+\item $A_2$:$\{task(1),name(1,create\_road),has(0,1)\}$
+\item $A_3$ $\{task(1),name(1,choose\_pattern),has(0,1)\}$
+\item $A_4$:$\{task(1),name(1,pattern\_pref),has(0,1)\}$
+\end{itemize}
+The algorithm is now called as follows: $CQ13(1,road\_pattern)$, i.e., the new name of the element should be $road\_pattern$.
+Let us briefly run through the algorithm.
+After executing line 2 we obtain $ArgsN=\{A_2,A_3,A_4\}$, since only those arguments contain $name(1,\_)$.
+Next, on line 3, $B=\{task(1),has(0,1)\}$, i.e., $B$ is the general argument for the task without the $name$ statement.
+After line 4 we have $$A=\{task(1),has(0,1),name(1,road\_pattern),$$ which is added to $Args$ and attacks arguments $A_2,A_3$, and $A_4$.
 
-The other critical questions with effect INTRO, namely CQ6c, CQ6d, CQ7b, and CQ9 all have a very similar structure to algorithm~\ref{alg:cq6b} and have therefore been omitted.
-
-Our final critical question is the generic counterargument, which is very simple.
+\emph{Algorithms for critical questions CQ6b, CQ6c, CQ6d, CQ7b, and CQ9 (INTRO):} The introduction algorithms for the critical questions are all very similar to the INTRO algorithms for argument schemes (Algorithm~\ref{alg:as1}).
+They have therefore been omitted.
 
@@
 
-\subsection{Some examples}
+\emph{Algorithm for $Att$ (Generic counter argument:} Applying a generic counter argument is very simple, and simply results on an attack on the original argument.
+We illustrate this by continuing our example from Figure~\ref{fig:examples:relevant-actor2:formal} (Algorithm~\ref{alg:as0}).
+The example is shown in Figure~\ref{fig:examples:relevant-actor2:formal}, where we see that a generic counter argument simply attacks the argument to disable the actor.
 
-\todo{Marc}{Marc}{Revisit the examples}
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeIN] at (-3.7,0) {
+        	\smallargtext{AS0}{$\{actor(0),name(0,dev\_team\}$}
+        } ;
+        \node (a1) [argNodeOUT] at (-3.7,-2){
+        	\argtext{DISABLE}{$\{disabled(0)\}$}
+        };
+        \node (a2) [argNodeIN] at (-3.7,-4){
+        	\argtext{Att}{$\{\}$}
+        };
+         \path
+    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
+    (a1) edge [CQLink, bend right=50] node [left,draw=none] {Att} (a2)
+    (a2) edge [attackLink] (a1);
+\end{tikzpicture}
+\caption{Formalization of the arguments in Figure~\ref{fig:examples:relevant-actor2}.}
+\label{fig:examples:relevant-actor2:formal}
+\end{figure}
 
-\todo{Marc}{Marc}{Do links need identifiers??}
+\subsection{Constructing GRL models}
+
+Constructing GRL models from the arguments is extremely simple: We simply compute the extensions of the argumentation frameworks, and collect all atomic sentences in the accepted arguments.
+This forms out GRL model.
+Let us briefly do so for the examples of the previous subsection:
+\begin{itemize}
+\item
+Figure~\ref{fig:examples:relevant-actor:formal}: Since there are no attacks between the arguments, all atomic sentences are accepted. This results in the following specification: 
+\begin{quote}
+\begin{verbatim}
+actor(0).
+name(0,dev_team).
+disabled(0).
+\end{verbatim}
+\end{quote}
+This again corresponds to the GRL model on the right-hand side of Figure~\ref{fig:examples:relevant-actor}.
+\item
+Figure~\ref{fig:examples:cq5b}: There is one rejected argument and five accepted ones. The resulting specification is:
+\begin{quote}
+\begin{verbatim}
+goal(0). name(0,show_simulation).
+task(1). name(1,generate_traffic).
+task(2). name(2,compute_lights).
+task(3). name(3,show_controls).
+decomp(3,0,{1,2,3},and).
+\end{verbatim}
+\end{quote}
+\item Figure~\ref{fig:examples:clarification:formal}: There are only two accepted arguments. The resulting specification is:
+\begin{quote}
+\begin{verbatim}
+actor(0). name(0, student).
+task(1).  name(1,road_pattern).
+has(0,1).
+\end{verbatim}
+\end{quote}
+This corresponds to the right-hand GRL model of Figure~\ref{fig:examples:clarification}.
+\item Figure~\ref{fig:examples:relevant-actor2:formal}.
+There are two accepted arguments, but the \emph{generic counterargument} does not contain any formulas.
+Therefore the resulting specification is:
+\begin{quote}
+\begin{verbatim}
+actor(0).
+name(0,dev_team).
+\end{verbatim}
+\end{quote}
+This corresponds to the right-hand GRL model of Figure~\ref{fig:examples:relevant-actor2}.
+\end{itemize}

### 2017-03-14 tool.tex

@@
 \label{sect:implementation}
+
+\todo{Marc}{Marc}{Currently this section comes from my thesis where I present the tool as future work. Here we should explain it}
+
+GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}.
+This tool is an extension to Eclipse.
+Although it is a rich tool with many features, we also believe it is not very easy to set it up.
+This seriously harms the exposure of the language, as well as the ability for practitioners to use it.
+We have started to implement a simple version of GRL as an online tool in Javascript.
+This makes it usable from the browser, without requiring the installation of any tool.
+The tool can be used from the following address:
+
+\begin{quote}
+\url{http://marcvanzee.nl/RationalGRL/editor}
+\end{quote}
+
+A screenshot of the tool is shown in Figure~\ref{fig:goalmodeling:tool}.
+As shown, there are two tabs in the tool, one for ``Argument'' and one for ``GRL''.
+The argument part has not been implemented yet, and the GRL part only partly, but the idea behind the tool should be clear.
+Users are able to work on argumentation and on goal modeling in parallel, where the argumentation consists of forming arguments and counterarguments by instantiating argument schemes and answering critical questions.
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{img/tool_screen}
+\caption{Screenshot of the prototype tool}
+\label{fig:goalmodeling:tool}
+\end{figure*}
+
+An important aspect of the tool is that users can switch freely between these two ways of modeling the problem.
+One can model the entire problem in GRL, or one can do everything using argumentation.
+However, we believe the most powerful way to do so is to switch back and forth.
+For instance, one can create a simple goal model in GRL, and then turn to the argumentation part, which the users can look at the various critical questions for the elements, which may trigger discussions.
+These discussions results in new arguments for and against the elements in the goal model.
+Once this process is completed, one may switch to the goal model again, and so on.
+We believe that in this way, there is a close and natural coupling between modeling the goals of an organization as well as rationalizing them with arguments.

### 2017-05-23 abstract.tex
Small modifications
Modified and commented on abstract and intro
Need to address the comments

@@
 The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in the goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
+While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found in the resulting model.
 For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to goal modeling.
+In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to a goal modelingn approach.
 We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
@@
 \end{abstract}
+%SG: what are the findings? or what case study we are using? 
+%SG_New: I still have these comments!

### 2017-05-23 introduction.tex
background
checked the intro
Modified and commented on abstract and intro
Need to address the comments

@@
 As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\todo{Floris}{all}{Depending on whether we use preferences in argumentation models, we might want to rephrase this}
+\todo[resolved]{Floris}{all}{Depending on whether we use preferences in argumentation models, we might want to rephrase this}
+% [Marc] We don't use preferences.
 \item
@@
 The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with practical reasoning and argumentation approaches from Artificial Intelligence (AI)~\cite{atkinson2007}.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system; (2) it must be able to generate formal goal models based on the discussions; (3) it should have formal traceability links between goal elements and arguments in the discussions; (4) it must have tool support; (5) a methodology must be identified that allows the framework to be used by practitioners; and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise %TODO Sepi_Reply: we need to come back to this at the end of the paper.
+We have identified several important requirements for our framework: (1) it must have the means to formally model parts of the discussion process in the early-requirements phase of an information system; (2) it must have the means to generate formal goal models based on the discussions; (3) it should have formal traceability links between goal elements and arguments in the discussions; (4) it must have a tool support; (5) there should be a methodology that allows the framework to be used by practitioners; and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise %TODO Sepi_Reply: we need to come back to this at the end of the paper.
  \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc.
@@
 Tailor it towards our empirical work: ``the framework should be close to on the actual discussions stakeholders or designers have in the early requirements engineering phase''}.
+%TODO SG: What about 6?
+we still don't have it!
 
@@
 
-The contributions of this paper are aligned with our five success criteria.
+The contributions of this paper are aligned with our five %TODO SG: why 5?Not 6?
+ success criteria.
 In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
@@
 We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to the underlying arguments (requirement 3).
+We propose an algorithm to generate a GRL model based on the accepted arguments.
+This helps providing traceability links from GRL elements to the underlying arguments (requirement 3).
 
@@
 The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the Eclipse-based tool for GRL modeling, for further evaluation and analysis.
+The created GRL models in RationalGRL can be exported to jUCMNav~\cite{} %TODO SG: add citation
+ the Eclipse-based tool for GRL modeling, for further evaluation and analysis.
 
@@
 
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
+The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{}  %TODO SG: add citation
+and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
 
@@
 We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
+In Section~\ref{sect:gmas:2} %TODO SG: where is this?
+, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
+In Section~\ref{sect} %TODO SG:which section?
+, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
+In Section~\ref{sect:methodology} %TODO SG: where is this?
+, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}. %TODO SG: where is this?
+
+%TODO SG: In section 3, we talk about the methodology and then you discuss it again in section 7? why?

### 2017-05-23 background.tex
Many many inconsistencies!

@@
 In this exercise, designers are provided a problem description, requirements, and a description of the desired outcomes.
-The problem description is given in full in Appendix~\ref{ch:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI.
+The problem description is given in full in Appendix~\ref{ch:designprompt}, %TODO SG: shown as ??
+ and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI.
+%TODO SG: Make it anonymized.
+For example use a university in California (if this is UC Irvine)
  It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory.
@@
 Softgoals (e.g.
-\texttt{Realistic simulation)} are often more related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are more related to functional requirements.
+\texttt{Realistic simulation)} are often related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
@@
 This allows GRL to be used in many domains without the need to change the whole modeling language.
-This feature also helps us to apply our argumentation to other domain such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
+This feature also helps us to apply our argumentation to other domains such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
 
@@
 These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that leads to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
 In this article we aim to address this shortcoming.
@@
 \item[] We have goal  \texttt{Generate traffic},
-\item[]  \texttt{Keep same cars} will realize goal  \texttt{Generate traffic},
-\item[] Which will promote the value  \texttt{Simple design}
+\item[]  \texttt{Keep same cars} will realize goal  \texttt{Generate traffic}, %TODO SG: where is this in the model?
+\item[] Which will promote the value  \texttt{Simple design} %TODO SG: where is this in the model?
 \item[] \textit{Therefore} 
@@
 %\item Does doing the action have a side effect which demotes the value?
-\item Does doing the action have a side effect which demotes some other value?
-\item Does doing the action promote some other value?
+\item Does performing the action have a side effect which demotes some other value? %TODO SG: do you mean correlation link?
+\item Does performing the action promote some other value?
 %\item Does doing the action preclude some other action which would promote some other value?
@@
 Take, for example, critical question 4: if we find that  \texttt{Keep same cars} actually negatively influences the value  \texttt{Realistic simulation}, we have a counterargument to the above argument.
+%TODO SG: be careful with the concepts and semantics of links and elements in GRL.
+If you talk about side effect this indicates a correlation link.
+If you talk about impact or influence this indicates contribution link.
+In CQ4 you say side effect, here you say influence and in the model you show contribution link.
+Finally which one????
+%TODO SG: Keep the same car in the model positively impacts the realistic simulation. But you say different here... VERY confusing.
 Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).
-For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
+For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, %TODO SG: where is this in the model?
+which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
+%TODO SG: what do you mean?
 
@@
 Argument A2 attacks A1, as it points to a negative side-effect of \texttt{Keep same cars}.
+%TODO SG: again negative side-effect!
 
@@
 \end{figure}
+
+%TODO SG: VERY IMPORTANT!
+I am confused with this section.
+Are you explaining the GRL model and discussing how you are going to argue about it in order to change it?
+Or are you discussing how the links and elements in the GRL are created!
+Either of both, you should explain it clearly.
+You wrote this section and section 4 as if everyone has lots of background and they are in your head.
+Make them clear...
+If for example you are arguing that the + contribution link for keept same cars is not correct and you are changing it to negative side-effect, you should explain it.
+SO SO SO confusing...
+especially that nothing is matched between the model and the text.
 

### 2017-06-26 examples.tex
k

--- 2017-03-11 gmas.tex
@@
-\section{Argument Schemes for Goal Modeling}
-\label{sect:gmas}
-
-\begin{table*}[h]
-\centering
-\begin{tabularx}{\textwidth}{|l|l|l|X|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Effect}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant? & DISABLE\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & DISABLE\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & DISABLE\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & DISABLE\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?& DISABLE\\
-\hline
-\hline
-AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
-& & CQ5b & Does the goal decompose into any other tasks?& REPLACE\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal?& INTRO \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?& INTRO\\
-&& CQ6d & Does the task contribute to some other softgoal?& INTRO\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?& DISABLE\\
-&& CQ7b & Does the goal contribute to some other softgoal?& INTRO\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?& DISABLE\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?& INTRO\\
-\hline
-AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& REPLACE\\
- &  & CQ10b & Is the decomposition type correct? (AND/OR/XOR)& REPLACE\\
-\hline
-AS11 & Task $T$ contributes negatively to softgoal $S$& CQ11 & Does the task contribute negatively to the softgoal?& DISABLE\\
-\hline
-\hline
-AS12 & Element $IE$ is relevant & CQ12 & Is the element relevant/useful? & DISABLE\\
-\hline
-AS13 & Element $IE$ has name $n$ & CQ13 & Is the name clear/unambiguous? & REPLACE\\
-\hline
-\hline
-- & - & Att & Generic counterargument & ATTACK\\
-\hline
-\end{tabularx}
-\caption{List of argument schemes (AS0-AS13, left column), critical questions (CQ0-CQ12, middle column), and the effect of answering them (right column).}
-\label{table:argument-schemes}
-\end{table*}
-
-In this section, we develop a set of \emph{argument schemes for goal modeling} and associated critical questions.
-We start from an initial list that we derive from PRAS, containing argument schemes and critical questions that specific to elements and relationships of GRL.
-We then refine this list incrementally by annotating transcripts, meaning we identify and count instantiations of arguments schemes.
-If we find new argument schemes or critical questions, we add them to the list.
-If some of them do not occur at all, we remove them from the list.
-We then manually construct goal models from the arguments and counter arguments we find in the transcripts, and add traceability links between the goal model and underlying arguments.
-An example is shown in Figure~\ref{fig:transcripts:grl}, which is the same as the example in Figure~\ref{fig:trafficsim}, but now including tracability links to underlying arguments.
-A red dot indicates the underlying argument is rejected, and a green dot means the underlying argument is accepted.
-
-The final list of argument schemes and critical questions that we end up with after our analysis is shown in Table~\ref{table:argument-schemes}.
-The initial list of argument schemes, i.e., those we start with before annotating the transcripts, consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}), and their corresponding critical questions.
-The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
-
-We found that answering critical questions can have varying effects on the model, and for each critical question, the right column in Table~\ref{table:argument-schemes} shows the effect of answering the critical questions affirmatively.
-Answering a critical questions can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
-
-In the first subsection of this section, we provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts), after which we analyze our results in the second subsection.
-
-\subsection{Details experiment}
-
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
-
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
-Participants were asked to use a think-aloud method during the design session.
-The assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The full problem descriptions can be found in Appendix~\ref{ch:designprompt}.
-All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
-The details of the transcripts are shown in Table~\ref{table:transcripts:info}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Annotation Method}
-
-We started with an initial list of 8 argument schemes and 18 critical questions that we derived from PRAS (AS1-AS4, AS6-AS9 of Table~\ref{table:argument-schemes}).
-We annotated transcripts with the arguments and critical questions from this list.
-If we found arguments or critical questions that did not appear in the original list, we added them and counted them as well.
-Argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}).
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-This can be seen in the various examples we will discuss.
-
-It is generally known in the argumentation literature that it can be very difficult to identify arguments in natural language texts~\cite{walton-etal2004}.
-Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-However, there is hardly any research on on argument extraction in the requirement engineering domains, so despite this potential weakness in our approach, we believe it nevertheless is at least useful as something that others can build further on (see Section~\ref{sect:goalmodeling:openissues}).
-
-\paragraph{Results}
-
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{img/transcript_grl}
-\caption{The GRL model manually constructed from transcript $t_1$.
-Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
-Elements and relationships with no dot have been inferred by us.}
-\label{fig:transcripts:grl}
-\end{figure*}
-
-All original transcripts, annotations, and models are available on the following website:
-\begin{quote} \url{http://www.github.com/marcvanzee/RationalArchitecture}
-\end{quote}
-in the folder ``Ch4 transcript analysis''.
-We also provide excerpts of the annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples we use in this article come from the transcripts.
-
-We found a total of 159 instantiations of the argument schemes AS0-AS11 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (Table~\ref{table:transcripts:results:argumentschemes}).
-A large portion (about 60\%) of the argument schemes we found involved discussions around tasks of the information system (AS2, AS10).
-
-We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing the relevance of it (CQ12, CQ13).
-
-\begin{table}[ht]
-\centering
-\begin{tabularx}{0.5\textwidth}{|l|X|l|l|l|>{\bfseries}l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Scheme/Question}} & $t_1$ & $t_2$ & $t_3$ & \textbf{total}\\
-\hline 
-AS0 & Actor & 2 & 2 & 5 & 9\\
-\hline
-AS1 & Resource & 2 & 4 & 5 & 11\\
-\hline
-AS2 & Task/action & 20 & 21 & 17 & 58\\
-\hline
-AS3 & Goal & 0 & 2 & 2 & 4\\
-\hline
-AS4 & Softgoal & 3 & 4 & 2 & 9\\
-\hline
-AS5 & Goal decomposes into tasks & 4 &0& 4 & 8\\
-\hline
-AS6 & Task contributes to softgoal & 6 & 2 &0& 8\\
-\hline
-AS7 & Goal contributes to softgoal &0& 1 & 1 & 2\\
-\hline
-AS8 & Resource contributes to task & 0 & 4 & 3 & 7\\
-\hline
-AS9 & Actor depends on actor &0& 1 & 3 & 4\\
-\hline
-AS10 & Task decomposes into tasks & 11 &14 &11 &36\\ 
-\hline
-AS11 & Task contributes negatively to softgoal & 2 & 1 & 0 & 3\\
-\hline
-\hline
-CQ2 & Task is possible? & 2 & 2 & 1 & 5\\
-\hline		
-CQ5a & Does the goal decompose into the tasks? & 0 & 1 & 0 & 1\\
-\hline
-CQ5b & Goes decomposes into other tasks? & 1 & 0 & 0 & 1\\
-\hline
-CQ6b & Task has negative side effects? & 2 & 0 & 0 & 2\\
-\hline
-CQ10a & Task decompose into other tasks? & 1 &2 &0&3\\
-\hline
-CQ10b & Decomposition type correct? &1 &0& 1 &2\\
-\hline
-\hline
-CQ12 & Is the element relevant/useful? & 2 & 3 & 2 &7\\
-\hline
-CQ13 & Is the name clear/unambiguous? &3 &10 & 3 & 16\\
-\hline
-\hline
-- & Generic counterargument & 0& 2 & 2 & 4\\
-\hline
-\hline
-\multicolumn{2}{|c|}{\textbf{TOTAL}}&69&80&69&222\\
-\hline
-\end{tabularx}
-\caption{Number of occurrences of AS0-AS9, CQ0-CQ12 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
-\label{table:transcripts:results:argumentschemes}
-\end{table}
-
-For each transcript, we manually created a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
-An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
-We added green and red dots to various elements and relationships in the figure.
-A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
-Note that if the underlying argument is rejected, the corresponding GRL element has been disabled.
-Some elements do not have a corresponding green or red dot.
-In that case, we have inferred the elements from the discussion, but we could not explicitly find back arguments for it.
-
-We found that answering a critical questions can have four different effects on the original argument and the corresponding GRL element:
-\begin{itemize}
-\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
-This operation does not attack the original argument of the critical question, but rather creates a new argument.
-For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
-Suppose now the critical question CQ5b: ``Does the goal \texttt{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \texttt{Choose randomly}''.
-This results in a new instantiation of AS5, namely: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.
-As a result, the goal model will contain the corresponding task \texttt{Choose randomly}, as well as an OR-decomposition relation from the goal \texttt{Generate cars} to that task.
-\item \textsf{DISABLE:} Disable the element or relationship of the argument scheme to which the critical questions pertains.
-This operation does not create a new argument, but only disables (i.e., attacks) the original one.
-For instance, suppose argument scheme AS0 is instantiated with: ``Actor \texttt{Teacher} is relevant''.
-This argument can be attack with critical question CQ0: ``Actor \texttt{Teacher} is not relevant''.
-As a result, the argument for the actor is attack, and actor \texttt{Teacher} is disabled in the goal model.
-\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
-This operation both introduces a new argument and attacks the original one.
-For instance, suppose argument scheme AS2 is instantiated with: ``Actor \texttt{Student} can perform task \texttt{Choose a pattern preference}.
-This argument can be attacked with critical question CQ13: ``The task \texttt{Choose a pattern preference} is unclear, it should be \texttt{Choose a road pattern}''.
-This results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}.
-In the goal model, the description of the task should change accordingly.
-\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-For instance, suppose argument scheme AS0 is instantiated with ``Actor \texttt{Teacher} is relevant''.
-Suppose this argument is attacked with critical question CQ0: ``Actor \texttt{Teacher} is not relevant'', because she does not create the system.
-However, this critical question may in turn be attacked again, if new evidence comes up.
-For instance, it may turn out that the teacher will work on developing the application herself.
-In this case, the generic counter argument ``Actor \texttt{Teacher} is relevant, because does develop for the application'' attack the argument ``Actor \texttt{Teacher} is not relevant''.
-As a result, the original argument ``Actor \texttt{Teacher} is relevant'' is accepted again, and as a result is shown in the goal model.
-\end{itemize}
-
-In Section~\ref{sect:gmas:examples} we provide examples we found in the transcripts for all of these four effects.
-
-\subsection{Analysis}
-\label{sect:gmas:transcripts:analysis}
-
-\paragraph{Analysis of the Argument Schemes}
-Recall that our initial list of argument schemes consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}).
-Therefore, the difference between the initial list of argument schemes and those found back in the transcripts is quite small.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system.
-This gives us an indication that it is possible to capture (parts of the) arguments used in those type of discussions using argument schemes.
-
-We observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 11.
-Our analysis however shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
-However, if needed, additional argument schemes and critical questions can be added, and our list is not meant to be exhaustive.
-
-\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-%SG: I don't get this sentence.
-Also, the ones below.
-We found few of the critical questions we initially proposed.
-However, this does not mean that they were not implicitly used in the minds of the participants.
-If a participants for instance forms an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we believe that having these critical questions available may stimulate discussions.
-
 \section{Examples}

### 2017-06-26 gmas.tex
k

@@
 In the context of tool support, we believe that having these critical questions available may stimulate discussions.
-
-\section{Examples}
-\label{sect:gmas:examples}
-
-We now discuss various instantiations of argument schemes and the result of answering critical questions in more detail.
-For each example we provide transcript excerpts, a visualization of the arguments, and the corresponding goal model elements.
-We provide a legend for our visualization notation in Figure~\ref{fig:legend}.
-
-\begin{figure}[ht!]
-\centering
-\begin{tikzpicture}
-        \node (att1) [argNodeIN] at (-5,0) {$A_1$};
-        \node (att2) [argNodeOUT] at (-2,0) {$A_2$};
-        \node (attDescription) [draw=none, align=center] at (0.5,0) {Argument $A_1$ (accepted)\\ attacks argument $A_2$\\ (rejected)};
-        \node[traceNodeGREEN] (trace1a) at (-5,-1.5) {};
-        \node[traceNodeGREEN] (trace1b) at (-2,-1.5) {};
-        \node (traceDescription) [draw=none,align=center] at (0.5,-1.5) {Traceability link from\\ accepted argument to\\ enabled GRL element};
-        \node[traceNodeRED] (trace2a) at (-5,-3) {};
-        \node[traceNodeRED] (trace2b) at (-2,-3) {};
-        \node (traceDescription) [draw=none,align=center] at (0.5,-3) {Traceability link from\\ rejected argument to\\ disabled GRL element};
-        \node (cq1) [argNodeIN] at (-5,-4.5) {$A_1$};
-        \node (cq2) [argNodeIN] at (-2,-4.5) {$A_2$};
-        \node (cqDescription) [draw=none,align=center] at (0.5,-4.5) {Argument $A_2$ is the result of\\ answering critical question\\ $CQ$ of argument $A_1$};
-         \path
-    (att1) edge [attackLink] (att2)
-    (cq1) edge [CQLink] node [above,draw=none] {CQ} (cq2)
-    (trace1a) edge[traceLinkGREEN] (trace1b)
-    (trace2a) edge[traceLinkRED] (trace2b);
-\end{tikzpicture}
-\caption{Legend of the various elements and relationships we use for the examples in this article.}
-\label{fig:legend}
-\end{figure}
-
-\subsubsection{Example 1: Disable task \texttt{Traffic light}}
-
-The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$.
-In this example, participants are summing up functionality of the traffic simulator, which are tasks that the student can perform in the simulator.
-All these task can be formalized and are instantiations of argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
-
-Once all these tasks are summed up, participant \texttt{P1} notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \texttt{Add traffic light} is not useful.
-We formalized this using the critical question CQ12: ``Is task \texttt{Add traffic light} useful/relevant?''.
-
-We visualize some of the argument schemes, critical questions, and traceability links with the GRL model in Figure~\ref{fig:examples:traffic-light}.
-On the left side of the image, we see three of the instantiated argument schemes AS2.
-The bottom one, ``Actor \texttt{Student} has task \texttt{Add traffic light}'', is attacked by another argument generated from applying critical question CQ12: ``\texttt{Add traffic light} is useless (\emph{All intersections have traffic lights}).
-As a result, the corresponding GRL task is disabled.
-The other two tasks are enabled and have green traceability links.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}
-        \node (a0) [argNodeIN] at (-2,0) {
-        	\argtext{AS2}{Actor \emph{Student} has task \emph{Save map}}
-        };
-        \node (a1) [argNodeIN] at (-2,-2) {
-        	\argtext{AS2}{Actor \emph{Student} has task \emph{Add road}}
-        };
-        \node (a2) [argNodeOUT] at (-2,-4) {
-        	\argtext{AS2}{Actor \emph{Student} has task \emph{Add traffic light}}
-        };
-        \node (a3) [argNodeIN] at (-2,-6){
-        	\argtext{}{\emph{Add traffic light} is useless (\emph{All intersections have traffic lights})}
-        };
-        \node[grl] (task1) at (2.3,0) { \includegraphics[scale=0.5]{img/task_save_map} };
-        \node[grl] (task2) at (2.3,-2) { \includegraphics[scale=0.5]{img/task_add_road} };
-        \node[grl, disabled] (task3) at (2.3,-4) { \includegraphics[scale=0.5]{img/task_add_traffic_light} };
-        \node[traceNodeGREEN] (trace1a) at (-.4,-.3) {};
-        \node[traceNodeGREEN] (trace1b) at (2.2,-.3) {};
-        \node[traceNodeGREEN] (trace2a) at (-.4,-2.3) {};
-        \node[traceNodeGREEN] (trace2b) at (2.2,-2.3) {};
-        \node[traceNodeRED] (trace3a) at (-.4,-4.3) {};
-        \node[traceNodeRED] (trace3b) at (2.2,-4.3) {};
-         \path
-    (a3) edge [attackLink] (a2)
-    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ12} (a3)
-    (trace1a) edge[traceLinkGREEN] (trace1b)
-    (trace2a) edge[traceLinkGREEN] (trace2b)
-    (trace3a) edge[traceLinkRED] (trace3b);
-\end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted lines) for the traffic light example.}
-\label{fig:examples:traffic-light}
-\end{figure}
-
-\subsubsection{Example 2: Clarify task \texttt{Road pattern}}
-
-The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_3$.
-It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}[->]
-        \node (a0) [argNodeOUT] at (-2,0) {
-        	\argtext{AS2}{Actor \emph{Student} has task \emph{Create road}}
-        };
-        \node (a1) [argNodeOUT] at (-2,-2) {
-        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a pattern}}
-        };
-        \node (a2) [argNodeOUT] at (-2,-4.2) {
-        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a pattern preference}}
-        };
-        \node (a3) [argNodeIN] at (-2,-6.7) {
-        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a road pattern}}
-        };
-        \node[grl] (actor) at (2.3,-4) { 
-        	\includegraphics[scale=0.5]{img/task_choose_road_pattern} 
-        };
-        \node[traceNodeGREEN] (trace1) at (-.5,-7.1) {};
-        \node[traceNodeGREEN] (trace2) at (2.2,-4.4) {};
-
-\begin{pgfonlayer}{background}
-         \path
-    (a1) edge[attackLink] (a0)
-    (a2) edge[attackLink] (a1)
-    (a2) edge[attackLink, bend right=20] (a0)
-    (a3) edge[attackLink] (a2)
-    (a3) edge[attackLink, bend right=20] (a1)
-    (a3) edge[attackLink, bend right=40] (a0);
-\end{pgfonlayer}
-
-	\path
-	(a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a1)
-    (a1) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a2)
-    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a3)
-    (trace1) edge[traceLinkGREEN] (trace2);
-\end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the road pattern example.}
-\label{fig:examples:clarification}
-\end{figure}
-
-The formalized argument schemes and critical questions are shown in Figure~\ref{fig:examples:clarification}.
-The discussion starts with the first instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Create road}''.
-This argument is then challenged with critical question CQ12: ``Is the task \texttt{Create road} clear?''.
-Answering this question results in a new instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Choose a pattern}''.
-This process is repeated two more times, resulting in the final argument ``Actor \texttt{Student} has task \texttt{Choose a road pattern}''.
-This final argument is unattacked and has a corresponding intentional element (right image).
-
-What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element.
-For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all three other arguments for a name of the element.
-If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.
-Suppose for instance the bottom argument ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' did not attack the second argument: ``Action \texttt{Student} has task \texttt{Choose a pattern}''.
-In that case, this argument would be reinstated, because its only attacker ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' is itself defeated by the bottom argument.
-
-\subsubsection{Example 3: Decompose goal \texttt{Simulate}}
-
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
-It consists of a discussion about the type of decomposition relationship for the goal \texttt{Simulate}.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}[->]
-        \node (a0) [argNodeIN] at (-2,0) {
-        	\argtext{AS3}{Actor \emph{System} has goal \emph{Simulate}}
-        };
-        \node (a1) [argNodeIN] at (-2,-1.5) {
-        	\argtext{AS2}{Actor \emph{System} has task \emph{Dynamic simulation}}
-        };
-        \node (a2) [argNodeIN] at (-2,-3) {
-        	\argtext{AS2}{Actor \emph{System} has task \emph{Static simulation}}
-        };
-        \node (a3) [argNodeOUT] at (-2,-5) {
-        	\argtext{AS5}{Goal \emph{Simulate} AND-decomposes into \emph{Static simulation} and \emph{Dynamic simulation}}
-        };
-        \node (a4) [argNodeIN] at (-2,-7.5) {
-        	\argtext{AS5}{Goal \emph{Simulate} OR-decomposes into \emph{Static simulation} and \emph{Dynamic simulation}}
-        };
-        \node[grl] (actor) at (2.3,-4) { 
-        	\includegraphics[scale=0.5]{img/simulate_decomposition} 
-        };
-        \node[traceNodeGREEN] (trace1a) at (-.4,-.3) {};
-        \node[traceNodeGREEN] (trace1b) at (2.5,-3.1) {};
-        \node[traceNodeGREEN] (trace2a) at (-.4,-1.8) {};
-        \node[traceNodeGREEN] (trace2b) at (3.5,-5.6) {};
-        \node[traceNodeGREEN] (trace3a) at (-.4,-3.3) {};
-        \node[traceNodeGREEN] (trace3b) at (1.3,-5.6) {};
-        \node[traceNodeGREEN] (trace4a) at (-.4,-8.2) {};
-        \node[traceNodeGREEN] (trace4b) at (2.5,-3.9) {};
-
-\begin{pgfonlayer}{background}
-         \path
-    (a4) edge[attackLink] (a3);
-\end{pgfonlayer}
-
-	\path
-	(a3) edge [CQLink, bend right=50] node [left,draw=none] {CQ10b} (a4)
-    (trace1a) edge[traceLinkGREEN] (trace1b)
-    (trace2a) edge[traceLinkGREEN] (trace2b)
-    (trace3a) edge[traceLinkGREEN] (trace3b)
-    (trace4a) edge[traceLinkGREEN, bend right] (trace4b);
-\end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the goal decomposition example.}
-\label{fig:examples:decomposition}
-\end{figure}
-
-The visualization of this discussion is shown in Figure~\ref{fig:examples:decomposition}.
-Each GRL element on the right has a corresponding argument on the left.
-Moreover, the original argument for an AND-decomposition is attacked by the argument for the OR-decomposition, and the new argument is linked to the decomposition relation in the GRL model.
-
-\subsubsection{Example 4: Reinstate actor \texttt{Development team}}
-
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
-It consists of two parts: first participant \texttt{P1} puts forth the suggestion to include actor \texttt{Development team} in the model.
-This is, then, questioned by participant \texttt{P2}, who argues that the professor will develop the software, so there won't be any development team.
-However, in the second part, participant \texttt{P2} argue that the development team should be considered, since the professor does not develop the software.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}
-        \node (a0) [argNodeOUT] at (-2,0) {
-        	\argtext{AS0}{Development team is relevant}
-        } ;
-        \node (a1) [argNodeIN] at (-2,-2.5){
-        	\argtext{}{Development team is not relevant (\emph{The professor makes the software})}
-        };
-        \node[grl, disabled] (actor) at (2.3,-1) { \includegraphics[scale=0.5]{img/actor_development_team} };
-        \node[traceNodeRED] (trace1) at (-.5,-.3) {};
-        \node[traceNodeRED] (trace2) at (2.2,-0.1) {};
-
-         \path
-    (a1) edge[attackLink] (a0)
-    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
-    (trace1) edge[traceLinkRED] (trace2);
-\end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
-\label{fig:examples:relevant-actor}
-\end{figure}
-
-We formalize this using a \emph{generic counterargument}, attacking the critical question.
-The first part of the discussion is shown in Figure~\ref{fig:examples:relevant-actor}.
-We formalize the first statement as an instantiation of argument scheme AS0: ``Actor \texttt{development team} is relevant''.
-This argument is, then, attacked by answering critical question CQ0: ``Is actor \texttt{development team} relevant?
-with \emph{No}.
-This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
-This is shown in Figure~\ref{fig:examples:relevant-actor}, left image.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}
-        \node (a0) [argNodeIN] at (-2,0) {
-        	\argtext{AS0}{Development team is relevant}
-        };
-        \node (a1) [argNodeOUT] at (-2,-2.5){
-        	\argtext{}{Development team is not relevant (\emph{The professor makes the software})}
-        };
-        \node (a2) [argNodeIN] at (-2,-5) {
-        	\argtext{}{The professor doesn't develop the software}
-        } ;
-        \node[grl] (actor) at (2.3,-1) { \includegraphics[scale=0.5]{img/actor_development_team} };
-        \node[traceNodeGREEN] (trace1) at (-.5,-.3) {};
-        \node[traceNodeGREEN] (trace2) at (2.2,-0.1) {};
-
-         \path
-    (a1) edge[attackLink] (a0)
-    (a2) edge[attackLink] (a1)
-    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
-    (a1) edge [CQLink, bend right=50] node [left,draw=none] {Att} (a2)
-    (trace1) edge[traceLinkGREEN] (trace2);
-\end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
-\label{fig:examples:relevant-actor2}
-\end{figure}
-
-Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counter argument has been put forward.
-The argument ``Actor \texttt{Professor} doesn't develop the software'' now attacks the argument ``\texttt{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\texttt{Development team} is relevant''.
-As a result, the first argument is reinstated, which causes the actor in the GRL model to be enabled again.

### 2017-06-27 background-full.tex
many improvements (see email)

--- 2017-03-14 background.tex
@@
 In this exercise, designers are provided a problem description, requirements, and a description of the desired outcomes.
-The problem description is given in full in Appendix~\ref{ch:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI.
+The problem description is given in full in Appendix~\ref{sect:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI.
 It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory.
@@
 
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{img/example_small}
+\caption{Partial GRL Model of Traffic Simulator Example (\todo{Marc}{Sepideh}{Turn into vector image, turn ``Dynamic simulation'' and ``Static Simulation'' into goals, turn ``Simulate'' into softgoal, add some resource}}
+\label{fig:example-small}
+\end{figure}
+
 \begin{figure*}[ht]
@@
 
-Figure~\ref{fig:trafficsim} illustrates a GRL diagram from the traffic simulator design exercise.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system (\texttt{Student}, Figure~\ref{fig:trafficsim}), or the system itself ( \texttt{Traffic Tycoon}, Figure~\ref{fig:trafficsim}).
+Figure~\ref{fig:example-small} illustrates a partial GRL diagram from the traffic simulator design exercise (the full figure is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}).
+An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Tycoon}, Figure~\ref{fig:example-small}).
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
@@
 Softgoals (e.g.
-\texttt{Realistic simulation)} are often more related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are more related to functional requirements.
+\texttt{Simulate}) are often related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}, Figure~\ref{fig:trafficsim}).
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=\textwidth]{img/transcript_grl_0}
-\caption{GRL Model for the traffic simulator.}
-\label{fig:trafficsim}
-\end{figure*}
+In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}.
+In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \todo{Marc}{Sepideh}{Add resource here}).
 
 Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}.
+AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
+In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
 A contribution link has a qualitative contribution type or a quantitative contribution.
-Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the softgoal  \texttt{Dynamic simulation}.
+Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic}.
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor  \texttt{Traffic Tycoon} depends on the actor  \texttt{Student} to perform the task  \texttt{Adjust car spawning rate} to fulfill its task  \texttt{Generate cars}.
+\todo{Marc}{Sepideh}{Shall we add a dependency relation as well?
+They don't play a big role in our work..}
 
@@
 Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-GRL has a well-defined syntax and semantics, which are necessary if we want to incorporate it into a formal framework (requirements 1 and 2 as described in the introduction).
+GRL has a well-defined syntax and semantics.
 Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
@@
 This allows GRL to be used in many domains without the need to change the whole modeling language.
-This feature also helps us to apply our argumentation to other domain such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
+This feature also helps us to apply our argumentation to other domains such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
 
-The GRL model in Figure~\ref{fig:trafficsim} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
+The GRL model in Figure~\ref{fig:example-small} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
 However, the rationales and arguments behind certain intentional elements are not shown in the GRL model.
@@
 \begin{itemize}
-	\item Why does actor \texttt{Teacher} have only a single softgoal \texttt{Students learn from practice}? Why is this, for instance, not connected to any of the elements of \texttt{Student}?
-	\item What does  \texttt{Adjust timing schemes of sensorless interactions} mean?
-	\item Why does task \texttt{Keep same cars} contribut positively to \texttt{Realistic simulation} and negatively to \texttt{Dynamic simulation}?
-	\item How does the \texttt{Student} control the \texttt{Traffic Tycoon}?
-	\item Why does \texttt{Map design} have so many decompositions into other tasks?
+	\item Why does actor \texttt{Traffic Tycoon} have only a single softgoal \texttt{Simulate}? Why is this, for instance, not connected to goal \texttt{Show Simulation}?
+	\item What does \texttt{Keep same cars} mean?
+	\item Why does task \texttt{Keep same cars} contribute positively to \texttt{Static simulation} and negatively to \texttt{Dynamic simulation}?
+	\item Why does \texttt{Generate Cars} XOR-decompose into two tasks?
 \end{itemize}
@@
 These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+The model in Figure~\ref{fig:example-small} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
 In this article we aim to address this shortcoming.
@@
 \cite{Raz1978-RAZPR,walton1990}) and artificial intelligence \cite{bratman1987,atkinson2007}.
-One approach is to capture practical reasoning in terms of arguments schemes and critical questions~\cite{walton1990}.
-The idea is that an instantiation of such a scheme gives a presumptive argument in favor of, for example, taking an action.
-This argument can, then, be tested by posing critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original presumptive argument for the action.
+One approach is to capture practical reasoning using arguments schemes and critical questions~\cite{walton1990}.
+Applying an argument schemes results in an argument in favor of, for example, taking an action.
+This argument can, then, be tested with critical questions about, for instance, whether the action is possible given the situation.
+A negative answer to such a question leads to a counterargument to the original argument for the action.
 
-A formal approach to persuasive and deliberative reasoning about goals and actions has been presented by Atkinson et al.~\cite{atkinson2007}, who define the \emph{practical reasoning argument scheme} (PRAS).
-PRAS follows the following basic argument structure.
+Atkinson et al.~\cite{atkinson2007} develop an argument scheme for practical reasoning, which they call the \emph{practical reasoning argument scheme} (PRAS).
+This argument scheme is as follows\footnote{The original argument schemes uses slightly different terminology: we have replaced \emph{value} with \emph{softgoal}.
+Our results do not depend on these adjustments, but they make the relation between PRAS and GRL more clear.}:
 
@@
 \item[] We have goal $G$,
-\item[] Doing action $A$ will realize goal $G$,
-\item[] Which will promote the value $V$
+\item[] Doing action $A$ realizes to goal $G$,
+\item[] Which will contribute positively to the softgoal $S$
 \item[] \textit{Therefore} 
@@
 
-So, for example, we can say that 
+The capital letters $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals.
+Once instantiated, we obtain an argument.
+We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows:
 \begin{itemize}
-\item[] We have goal  \texttt{Generate traffic},
-\item[]  \texttt{Keep same cars} will realize goal  \texttt{Generate traffic},
-\item[] Which will promote the value  \texttt{Simple design}
+\item[] We have goal \texttt{Static simulation},
+\item[] Doing action \texttt{Keep same cars} realizes goal \texttt{Static simulation}, 
+\item[] Which contribute positively to the softgoal \texttt{Simulate} 
 \item[] \textit{Therefore} 
-\item[] We should perform action  \texttt{Keep same cars}
+\item[] We should perform action \texttt{Keep same cars}.
 \end{itemize}
 
-Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
 Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
@@
 \begin{enumerate}
-\item Will the action bring about the desired goal?
-%\item Does the goal promote the value stated?
+\item Will the action realize the desired goal?
 \item Are there alternative ways of realizing the same goal?
-\item Are there alternative ways of promoting the same value?
-%\item Does doing the action have a side effect which demotes the value?
-\item Does doing the action have a side effect which demotes some other value?
-\item Does doing the action promote some other value?
-%\item Does doing the action preclude some other action which would promote some other value?
+\item Are there alternative ways of contributing to the same softgoal?
+\item Does performing the action contribute negatively to some other softgoal?
 \item Is the action possible?
 \item Can the desired goal be realized?
-\item Is the value indeed a legitimate value?
+\item Is the softgoal indeed a legitimate softgoal?
 \end{enumerate}
@@
 These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 4: if we find that  \texttt{Keep same cars} actually negatively influences the value  \texttt{Realistic simulation}, we have a counterargument to the above argument.
-Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).
-For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
+Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument.
+
+Another way to elaborate an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that contributes to the same softgoal (question 3).
+For example,  can argue that another goal \texttt{Semi-dynamic Simulation} also realizes the softgoal \texttt{Simulate}.
+\todo{Marc}{Sepideh}{Do you think our example is good?
+I think it is confusing.
+Perhaps we should just create an artificial example here, loosely based on the traffic simulator but where the model makes sense to the reader.
+Now the tasks are all a bit vague and I think nobody knows what they actually mean.}
 

### 2017-06-27 designprompt.tex
many improvements (see email)

\section{UCI Design Workshop Prompt}
\label{sect:designprompt}

\subsection*{Design Prompt: Traffic Signal Simulator}

\subsubsection*{Problem Description}

For the next two hours, you will be tasked with designing a traffic flow simulation program. Your client for this project is Professor E, who teaches civil engineering at UCI. One of the courses she teaches has a section on traffic signal timing, and according to her, this is a particularly challenging subject for her students. In short, traffic signal timing involves determining the amount of time that each of an inter- section's traffic lights spend being green, yellow, and red, in order to allow cars in to flow through the intersection from each direction in a fluid manner. In the ideal case, the amount of time that people spend waiting is minimized by the chosen settings for a given intersection's traffic lights. This can be a very subtle matter: changing the timing at a single intersection by a couple of seconds can have far- reaching effects on the traffic in the surrounding areas. There is a great deal of theory on this subject, but Professor E. has found that her students find the topic quite abstract. She wants to provide them with some software that they can use to ``play'' with different traffic signal timing schemes, in different scenarios. She anticipates that this will allow her students to learn from practice, by seeing first-hand some of the patterns that govern the subject.

\subsubsection*{Requirements}
The following broad requirements should be followed when designing this system:
\begin{enumerate}
\item
Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing. The resulting map need not be complex, but should allow for roads of varying length to be placed, and different arrangements of intersections to be created. Your approach should readily accommodate at least six intersections, if not more.
\item
Students must be able to describe the behavior of the traffic lights at each of the intersections. It is up to you to determine what the exact interaction will be, but a variety of sequences and timing schemes should be allowed. Your approach should also be able to accommodate left-hand turns protected by left-hand green arrow lights. In addition:
\begin{enumerate}
\item
Combinations of individual signals that would result in crashes should not be allowed.
\item
Every intersection on the map must have traffic lights (there are not any stop signs, over- passes, or other variations). All intersections will be 4-way: there are no ``T'' intersections, nor one-way roads.
\item
Students must be able to design each intersection with or without the option to have sensors that detect whether any cars are present in a given lane. The intersection's lights' behavior should be able to change based on the input from these sensors, though the exact behavior of this feature is up to you.
\end{enumerate}
\item
Based on the map created, and the intersection timing schemes, the students must be able to simulate traffic flows on the map. The traffic levels should be conveyed visually to the user in a real-time manner, as they emerge in the simulation. The current state of the intersections' traffic lights should also be depicted visually, and updated when they change. It is up to you how to present this information to the students using your program. For example, you may choose to depict individual cars, or to use a more abstract representation.
\item
Students should be able to change the traffic density that enters the map on a given road. For ex- ample, it should be possible to create a busy road, or a seldom used one, and any variation in between. How exactly this is declared by the user and depicted by the system is up to you. Broadly, the tool should be easy to use, and should encourage students to explore multiple alternative approaches. Stu- dents should be able to observe any problems with their map's timing scheme, alter it, and see the results of their changes on the traffic patterns. This program is not meant to be an exact, scientific simulation, but aims to simply illustrate the basic effect that traffic signal timing has on traffic. If you wish, you may assume that you will be able to reuse an existing software package that provides relevant mathematical functionality such as statistical distributions, random number generators, and queuing theory.
\end{enumerate}

You may add additional features and details to the simulation, if you think that they would support these goals.

Your design will primarily be evaluated based on its elegance and clarity – both in its overall solution and envisioned implementation structure.

\subsubsection*{Desired Outcomes}

Your work on this design should focus on two main issues:

\begin{enumerate}
\item
You must design the interaction that the students will have with the system. You should design the basic appearance of the program, as well as the means by which the user creates a map, sets traffic timing schemes, and views traffic simulations.
\item
You must design the basic structure of the code that will be used to implement this system. You should focus on the important design decisions that form the foundation of the implementation, and work those out to the depth you believe is needed.
\end{enumerate}

The result of this session should be: \emph{the ability to present your design to a team of software developers who will be tasked with actually implementing it.} The level of competency you can expect is that of students who just completed a basic computer science or software engineering undergraduate degree. You do not need to create a complete, final diagram to be handed off to an implementation team. But you should have an understanding that is sufficient to explain how to implement the system to competent developers, without requiring them to make many high-level design decisions on their own.

To simulate this hand-off, you will be asked to briefly explain the above two aspects of your design after the design session is over.

\subsubsection*{Timeline}
\begin{itemize}
\item 1 hour and 50 minutes: Design session
\item 10 minutes: Break / collect thoughts
\item 10 minutes: Explanation of your design 
\item 10 minutes: Exit questionnaire
\end{itemize}

### 2017-06-27 discussion.tex
many improvements (see email)

@@
 \section{Discussion}
-\label{sect:goalmodeling:discussion}
+\label{sect:discussion}
 
@@
 
-In this article, we develop the RationalGRL framework to trace back elements of GRL models to arguments using in discussions between stakeholders.
-The contributions of this article have become increasingly more formal.
-First, we analyze transcripts of meetings about the development of an information system.
-We count occurrences of argument schemes and critical questions, and categorize them by the effect of instantiating them.
-Then, we create a visual notation to link arguments and questions to elements of a goal model.
-If an argument for an element of a goal model is rejected, then the corresponding element is disabled.
-Similarly, if the argument is accepted, then the corresponding element is enabled.
-Finally, we formalize the argument schemes and critical questions in a logical framework.
-We propose a formal language to describe a GRL model, and we develop various algorithms for applying our argument schemes and critical questions.
+The introduction of this article contains five requirements we identified for our framework. We use the conclusion to discuss how RationalGRL meets our initial requirements.
 
-Our framework is one of the first attempts that try to combine the formal theory of argumentation with the practical frameworks of goal modeling.
-History has shown that it is remarkably difficult to combine formal results with practice.
-By taking argument schemes as a starting point, we believe we have chosen the right level of formality, but as the open issues section shows, much work is left to be done.
+\paragraph{1. The argumentation techniques should be close to actual discussions stakeholders or designers have.}
+
+We analyze a set of transcripts containing discussions about the architecture of an information system.
+
+\paragraph{2. The framework must have the means to formally model parts of the discussion process.}
+In order to generate goal models based on formalized discussions (requirement 2), we, first, formalize the list of arguments from requirement 1 in an argumentation framework.
+We formalize the critical questions as algorithms modifying the argumentation framework.
+We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
+We propose an algorithm to generate a GRL model based on the accepted arguments.
+This helps providing traceability links from GRL elements to the underlying arguments (requirement 3).
+
+We implement our framework in an online tool called RationalGRL (requirement 4).
+The tool is implemented using Javascript.
+It contains  two parts, goal modeling and argumentation.
+The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators.
+The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
+The created GRL models in RationalGRL can be exported to jUCMNav~\cite{} %TODO SG: add citation
+ the Eclipse-based tool for GRL modeling, for further evaluation and analysis.
+
+Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
+The methodology consists of two parts, namely argumentation and goal modeling.
+In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
+When switching to the goal modeling part, the accepted arguments are used to create a goal model.
+In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
+This might mean that the underlying arguments are no longer consistent with the goal models.

### 2017-06-27 examples.tex
many improvements (see email)

@@
 \section{Examples}
-\label{sect:gmas:examples}
+\label{sect:examples}
 
@@
 
-\begin{figure}[ht!]
-\centering
+\begin{figure}[ht]
+\begin{adjustwidth}{-3em}{}
         \begin{tikzpicture}[->]
-        \node (a0) [argNodeIN] at (-2,0) {
-        	\argtext{AS3}{Actor \emph{System} has goal \emph{Simulate}}
+        \node (a0) [argNodeIN] at (-2,0.5) {
+        	\argtext{CQ3}{Task \emph{Dynamic simulation} is not relevant}
         };
@@
         };
-        \node (a3) [argNodeOUT] at (-2,-5) {
+        \node (a3) [argNodeOUT] at (-2,-4.5) {
         	\argtext{AS5}{Goal \emph{Simulate} AND-decomposes into \emph{Static simulation} and \emph{Dynamic simulation}}
@@
         };
-        \node[grl] (actor) at (2.3,-4) { 
+        \node[grl] (actor) at (2.2,-4) { 
         	\includegraphics[scale=0.5]{img/simulate_decomposition} 
         };
-        \node[traceNodeGREEN] (trace1a) at (-.4,-.3) {};
-        \node[traceNodeGREEN] (trace1b) at (2.5,-3.1) {};
-        \node[traceNodeGREEN] (trace2a) at (-.4,-1.8) {};
-        \node[traceNodeGREEN] (trace2b) at (3.5,-5.6) {};
+        \node[traceNodeRED] (trace2a) at (-.4,-1.8) {};
+        \node[traceNodeRED] (trace2b) at (3.2,-5.6) {};
         \node[traceNodeGREEN] (trace3a) at (-.4,-3.3) {};
-        \node[traceNodeGREEN] (trace3b) at (1.3,-5.6) {};
+        \node[traceNodeGREEN] (trace3b) at (1,-5.6) {};
         \node[traceNodeGREEN] (trace4a) at (-.4,-8.2) {};
-        \node[traceNodeGREEN] (trace4b) at (2.5,-3.9) {};
+        \node[traceNodeGREEN] (trace4b) at (2.2,-3.9) {};
 
@@
          \path
+    (a1) edge[attackLink] (a0)
     (a4) edge[attackLink] (a3);
@@
 	(a3) edge [CQLink, bend right=50] node [left,draw=none] {CQ10b} (a4)
-    (trace1a) edge[traceLinkGREEN] (trace1b)
-    (trace2a) edge[traceLinkGREEN] (trace2b)
+	(a1) edge [CQLink, bend right=50] node [right,draw=none] {CQ3} (a0)
+    (trace2a) edge[traceLinkRED] (trace2b)
     (trace3a) edge[traceLinkGREEN] (trace3b)
@@
 \end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the goal decomposition example.}
+\end{adjustwidth}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the example.}
 \label{fig:examples:decomposition}

### 2017-06-27 formalframework.tex
many improvements (see email)

@@
-\section{RationalGRL: the logical framework}
+\section{RationalGRL: Logical Framework}
 \label{sect:formalframework}
@@
 
-Let us briefly review what we have done so far. Up until now we have achieve the following:
+Let us briefly take stock:
 \begin{itemize}
-\item We (informally) developed a set of argument schemes and critical questions in Section~\ref{sect:gmas}, by annotating transcripts.
-\item We developed a visual notation to illustrate some argument schemes and critical questions in Section~\ref{sect:gmas:examples}.
+\item We developed an initial set of argument schemes and critical questions in Section~\ref{sect:gmas}, by annotating transcripts.
+\item We developed a visual notation to illustrate some argument schemes and critical questions in Section~\ref{sect:examples}.
 \item We proposed a logical language consisting of atomic sentence to describe a GRL model in Section~\ref{sect:goalmodeling:logicallanguage}.
@@
 
-We see that we are becoming increasingly more formal in each step of the progress.
-What we haven't formalized properly yet are the argument schemes and critical questions, which is what we turn to now.
-
-Both the application of an argument schemes and the answering of critical questions are \emph{procedural} or \emph{dialectic} in nature.
-Therefore, a natural choice of formalizing them seems to develop procedures, or algorithms for them, which is what we do here.
-
-In the following algorithms, we assume the following global variables:
+In this subsection we develop algorithms for applying argument schemes and critical questions. In the following algorithms, we assume the following global variables:
 

### 2017-06-27 gmas.tex
many improvements (see email)

@@
 
-In this section, we develop a set of \emph{argument schemes for goal modeling} and associated critical questions.
-We start from an initial list that we derive from PRAS, containing argument schemes and critical questions that specific to elements and relationships of GRL.
-We then refine this list incrementally by annotating transcripts, meaning we identify and count instantiations of arguments schemes.
-If we find new argument schemes or critical questions, we add them to the list.
-If some of them do not occur at all, we remove them from the list.
-We then manually construct goal models from the arguments and counter arguments we find in the transcripts, and add traceability links between the goal model and underlying arguments.
-An example is shown in Figure~\ref{fig:transcripts:grl}, which is the same as the example in Figure~\ref{fig:trafficsim}, but now including tracability links to underlying arguments.
-A red dot indicates the underlying argument is rejected, and a green dot means the underlying argument is accepted.
+In the previous section we provided an overview of RationalGRL.
+In the example of the methodology, we discussed various argument schemes and critical questions.
+In this section, we discuss our argument schemes and critical questions in more detail.
+It is important to note that the list of argument schemes we obtain in this section is not exhaustive.
+It is an initial list that we obtained by annotating transcripts, but our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
 
-The final list of argument schemes and critical questions that we end up with after our analysis is shown in Table~\ref{table:argument-schemes}.
-The initial list of argument schemes, i.e., those we start with before annotating the transcripts, consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}), and their corresponding critical questions.
+The list of argument schemes and critical questions that we obtained from our analysis is shown in Table~\ref{table:argument-schemes}.
 The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
 
-We found that answering critical questions can have varying effects on the model, and for each critical question, the right column in Table~\ref{table:argument-schemes} shows the effect of answering the critical questions affirmatively.
+As we already discussed in the methodology, we found that answering critical questions can have varying effects on the model, and for each critical question, the right column in Table~\ref{table:argument-schemes} shows the effect of answering the critical questions affirmatively.
 Answering a critical questions can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
@@
 The assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The full problem descriptions can be found in Appendix~\ref{ch:designprompt}.
+The full problem descriptions can be found in Appendix~\ref{sect:designprompt}.
 All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software.
@@
 
-All original transcripts, annotations, and models are available on the following website:
+All original transcripts, annotations, and models are available on the following website \todo{Marc}{Marc}{Change URL and put everything together with the tool}:
 \begin{quote} \url{http://www.github.com/marcvanzee/RationalArchitecture}
 \end{quote}
-in the folder ``Ch4 transcript analysis''.
 We also provide excerpts of the annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples we use in this article come from the transcripts.

### 2017-06-27 introduction.tex
many improvements (see email)

--- 2017-01-26 introduction.tex
@@
 
+\subsection{Requirements Engineering and Goal Modeling}
+
 Requirements Engineering (RE) is an approach to assess the role of a future information system within a human or automated environment.
-An important goal in RE is to produce a consistent and comprehensive set of system requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
+An important goal in RE is to produce a consistent and comprehensive set of requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
 
 One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-This is generally referred to as goal modeling.
-Given the number of currently established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}), there is a general consensus that goal models are useful in RE.
-Several goal modeling languages have been developed in the last two decades.
+These activities fall under the umbrella of goal modeling.
+The are a large number of established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}).
+Several goal modeling languages have been developed in the last two decades as well.
 The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}.
+
+\subsection{Problem Statement}
 
@@
 However, most real-world information systems -- e.g., air-traffic management, industrial production processes, or government and healthcare services -- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for several reasons.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons.
 
-\todo{Marc}{Sepideh}{Please improve 2 and 4 by adding references}
-
-\begin{enumerate}
-\item
-It is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
+First, it is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
 Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
@@
 As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
-\todo{Floris}{all}{Depending on whether we use preferences in argumentation models, we might want to rephrase this}
-\item
-Other stakeholders, who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage.
+
+Secondly, other stakeholders, such as new developers on the team who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase.
 If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
-\item
-Alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
+
+Thirdly, alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
 For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
@@
 In fact, empirical data suggest that this is an important reason of RE project failure~\cite{curtis1988field}.
+
+Fourth and lastly, in goal models in general, ``rationale'' behind any decision is static and does not immediately impact the goal models when they change.
+That is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
+A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
+Similarly, a part of the goal model may change, but it is not possible to formally reason about the consistency of this new goal model with the underlying beliefs and arguments in the goal modeling languages.
+This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs put forward by another group of stakeholders.
+
+\subsection{Research Questions}
+
+The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with argumentation approaches techniques from Artificial Intelligence (AI) research\cite{atkinson2007}.
+We have identified five important requirements for our framework:
+\begin{enumerate}
+\item The argumentation techniques should be close to actual discussions stakeholders or designers have in the early requirements engineering phase.
 \item 
-It is not possible to reason about changing beliefs and opinions, and their effect on the goal model.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to reason about the consistency of this new goal model with the underlying beliefs and arguments.
-This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs put forward by another group of stakeholders.
+The framework must have formal traceability links between elements of the goal model and underlying arguments.
+\item 
+Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and visa versa.
+\item 
+The framework must have tool support.
+\item 
+A methodology should be identified that allows the framework to be used by practitioners.
 \end{enumerate}
 
-The aim of this research is to resolve the above issues by designing a framework, which combines a methodology for the goal modeling process \todo{Floris}{Sepideh}{Do you know of any generic methodologies or heuristics for goal modeling (akin to Juretas GAM)?} with ideas from Artificial Intelligence on practical reasoning and argumentation \cite{atkinson2007} and simple software tool support.
-We identified several important requirements for our framework: (1) it must be able to formally model parts of the discussion process in the early-requirements phase of an information system; (2) it must be able to generate formal goal models based on the discussions; (3) it should have formal traceability links between goal elements and arguments in the discussions; (4) it must have tool support; (5) a methodology must be identified that allows the framework to be used by practitioners; and (6) the framework must identify arguments and rationales that might not have been found, or might have been lost, otherwise \todo{Floris}{all}{I think (6) should be replaced with something we actually do in the paper: we don't show that arguments etc.
-allows us to find rationales that would otherwise not have been found -- for this we would need an experiment.
-Tailor it towards our empirical work: ``the framework should be close to on the actual discussions stakeholders or designers have in the early requirements engineering phase''}.
+In this context, the main research question is:
 
-In this context, the main research question is: \emph{What is a suitable framework to formally capture the discussions between stakeholders such that it can generate goal models, and how to implement this framework?} The first five requirements are the success criteria of our approach, that is, the satisfaction of the five requirements will result in a positive answer to the research question.
-\todo{Sepideh}{all}{The last sentence has to be changed in the end to reflect the result when the paper/evaluation is done.
-If we write it this way, it means we are unsure of the result.}
+\begin{quote}
+\textbf{RQ.} What are the constructs, mechanisms and rules for developing a framework that formally captures the discussions between stakeholders such that it can generate goal models?
+\end{quote}
 
-\subsection{Contributions}
+The five requirements above are the success criteria of our approach, that is, the satisfaction of the six requirements will result in a positive answer to the research question.
 
-The contributions of this paper are aligned with our five success criteria.
-In order to formalize discussions of the early requirements phase of an information system (requirement 1), we use a technique from argumentation and discourse modelling called \emph{argument schemes} \cite{walton-etal2004}.
-This allows us to formulate arguments and counter-arguments using so-called \emph{critical questions}.
-By combining arguments and critical questions, one can create a structured representation of a discussion.
-We first propose a set of argument schemes and critical questions we expect to find in the discussions that take place during the goal modelling process.
-We then validate this initial set with transcripts containing discussions about the architecture of an information system.
-This then gives us the final set of arguments and critical questions.
-
-In order to generate goal models based on formalized discussions (requirement 2), we first formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on accepted arguments.
-This creates traceability links from GRL elements to underlying arguments (requirement 3).
-
-We implement our framework into an online tool called RationalGRL (requirement 4).
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators .
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{}, the Eclipse-based tool for GRL modeling, for further evaluation and  analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
+The main contribution of this article is the RationalGRL framework,  which combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
+We develop a metamodel specifying how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}.
+We develop an initial list of argument schemes and critical questions by analyzing a set of transcripts containing discussion about an information system.
+Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains.
+We develop a methodology for using RationalGRL, which consists developing goal models and posing arguments in an integrated way.
+This methodology is intended to be used by practitioners in the field.
+We formalize RationalGRL using propositional logic, and develop algorithms for adding goal elements and arguments.
 
@@
 
-The first two sections are introductory: Section~\ref{sect:background} contains background and introduces the Goal-oriented Requirements Language (GRL)~\cite{} and argument schemes, while Section~\ref{sect:preview} provides a brief and high-level overview of our framework and methodology using an example.
-
-The rest of the article is in line with the main contributions.
-We develop the list of argument schemes in two iterations, corresponding to two sections: In Section~\ref{sect:gmas}, we introduce an initial list of Argument Schemes for Goal Modeling (GMAS) and associated critical questions based on an existing argument scheme.
-In Section~\ref{sect:gmas:2}, we evaluate the initial list of argument schemes by annotating transcripts from discussions about an information system, which results in our final list of argument schemes and critical questions.
-In Section~\ref{sect}, we develop a formal model for argument schemes and critical questions and we describe RationalGRL tool in Section~\ref{sect:implementation}.
-In Section~\ref{sect:methodology}, we propose our methodology while we discuss the related work in Section~\ref{sect:relatedwork}.
+This article is organized as follows.
+Section~\ref{sect:background} contains background and introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{}, and argument schemes.
+Section~\ref{sect:overview} provides a brief and high-level overview of our framework, together with a metamodel and the methodology.
+Section~\ref{sect:gmas} contains an in depth explanation of how we obtained an initial set of argument schemes by annotating transcripts from discussions about an information system, and in Section~\ref{sect:examples} we provide several examples of these arguments schemes.
+In Section~\ref{sect:formalframework} we formalize RationalGRL and our arguments schemes using propositional logic and we use well-known argumentation semantics to compute which arguments are accepted and which are rejected.
+We also develop various algorithms for the argument schemes in this section.
+In Section~\ref{sect:tool} we provide a brief overview of the prototype tool we developed for RationalGRL.
+Finally, Section~\ref{sect:discussion} contains a discussion, covering related work, future work, and a conclusion.

### 2017-06-27 overview.tex
many improvements (see email)

@@
-\section{RationalGRL overview and metamodel}
-\label{sect:preview}
+\section{RationalGRL Methodology and Metamodel}
+\label{sect:overview}
 
-\todo{Marc}{all}{I suggest the following subsections, so the methodology becomes part of this section. What do you think of this?}
+In this section we present a high-level overview of the RationalGRL framework.
+In the first subsection we present a methodology, specifying how practitioners can use RationalGRL to create goal models with traceability links to underlying arguments.
+In the second subsection we link the new argumentation elements and relations to the existing elements and relations of GRL in a metamodel.
+This metamodel serves as the specification for an implementation.
+In Section~\ref{sect:tool} we briefly discuss our prototype implementation based on this metamodel.
 
-\subsection{RationalGRL overview}
+\subsection{RationalGRL Methodology}
 
-\todo{Marc}{all}{Explain general idea of RationalGRL, perhaps by explaining the difference between Figure 2 and 5.}
+As mentioned earlier, the RationalGRL framework aims at integrating GRL models with the underlying discussions and arguments of the stakeholders during the analysis phase by combining PRAS with GRL.
+Our argumentation framework is developed based on the analysis of the CQs, practical reasoning and discussions of the stakeholders of the system.
+Initial GRL models are first created based on the analysis of non-functional requirements in the requirements specification documents, discussions with stakeholders and by refining the high-level goals into operationalized tasks.
 
-\subsection{Metamodel}
-
-Extend the GRL metamodel with our concepts.
-
-\subsection{RationalGRL methodology}
-
-\todo{Marc}{Sepideh}{Finish this}
-
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
+In RationalGRL, both GRL and argumentation models are done iteratively.
+Argumentation models impact GRL models and result in refinement and changes in GRL.
+For example, an answer to the CQ \emph{Is the task possible?} can lead to rejection or acceptance of the task in GRL.
+The analysis of GRL models may result in changes in the underlying argumentation framework as well.
+For example, adding a new IE to the GRL model can result in posing new CQs with respect to the IE and its relationships.
+Figure~\ref{fig:rationalgrl-framework} shows RationalGRL and the relationships between the different parts.
+On the right-hand side of the framework, we have GRL models and on the left-hand side we have argumentation framework.
+The links between the two sides illustrates the impacts each side has on the other side and the changes that occur due to the impacts and the refinements.
 
 \begin{figure}[ht]
+\centering
+\includegraphics[scale=0.4]{img/framework}
+\caption{The RationalGRL Framework}
+\label{fig:rationalgrl-framework}
+\end{figure}
+
+To develop a RationalGRL model, we follow the methodology shown  in Figure~\ref{fig:rationalgrl-methodology}. We already have an initial GRL model. The steps of the methodology are as follows:
+
+\begin{figure*}[ht]
 \centering
@@
 \label{fig:rationalgrl-methodology}
+\end{figure*}
+
+\textbf{Instantiate Argument Schemes (AS)} -- In this step, we instantiate one of the AS from the argumentation framework.
+In the section we develop 13 arguments schemes which we use in our analysis.
+An example of an argument scheme is "Goal \emph{G} contributes to softgoal \emph{S}".
+An instantiation of an argument scheme thus corresponds to an argument for or against part of a goal model.
+
+\textbf{Answer Critical Questions (CQs)} -- Instantiated arguments can be attacked by counter-arguments.
+CQs are ways in which AS can be attacked.
+Each argument scheme includes one or more CQs.
+For example, for the argument scheme mentioned above, we have two CQs as: \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}.
+Note that, the answer to CQ can also result into ``conflict'' situation which we do not consider here.
+Answering a CQ may result in an instantiation of a new AS.
+Thus, it is possible to go back and forth between this step and the previous one.
+
+\textbf{Decide on IE and the Relationships} -- The answers to the CQs can result in one of the four cases which impact the arguments and corresponding GRL IE: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.
+\textsf{INTRO} means that the argument scheme of the CQ does not get attacked and instead it creates a new argument.
+\textsf{DISABLE} means that the IE or the relationship related to the AS needs to be disabled or removed from the models.
+\textsf{REPLACE} introduces a new argument and attack the original argument at the same time.
+It basically replace the original element of the AS with a new one.
+\textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when a new evidence occur.
+
+\textbf{Modify GRL Models} -- Based on the result of step three, the GRL models can be modified.
+In this case, either a new IE or a new relationships is introduced, an existing IE or relationship gets disabled (removed) from the model or finally an existing IE or relationships is replaced by a new one.
+This results in a new modified GRL.
+The new GRL model can then impact the argument schemes and instantiate another argument scheme.
+
+In the next section, we provide a short example to illustrate the relationship between argumentation framework and goal models.
+
+\subsubsection*{Example}
+\label{sect:methodology-example}
+
+Our examples comes from a transcript containing discussions about the development of an information system, and they are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+We provide several more examples in Section~\ref{sect:examples}.
+
+A group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \emph{simulate} of the system using the RationalGRL methodology.
+This proceeds in the following way:
+\begin{itemize}
+\item
+First they start at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), and add the IE \texttt{Simulate} to GRL model (Figure~\ref{fig:examples:decomposition}, GRL model, top IE).
+\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
+They answer CQ \emph{Does Simulate AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic simulation'' and ``Static simulation}.
+\item As a result, two argument schemes are created, namely:
+\begin{itemize}
+\item Actor \emph{System} has task \emph{Dynamic simulation}
+\item Actor \emph{System} has task \emph{Static simulation}
+\end{itemize}
+\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate} to these two tasks.
+\item Next, the stakeholders test the validity of their goal model by answering more CQ. They answer two CQs:
+\begin{itemize}
+\item
+CQ \emph{Is task ``Dynamic simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
+Thus, the corresponding GRL IE is disabled.
+\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
+\end{itemize}
+\end{itemize}
+
+The resulting RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}, which can be found in Section~\ref{sect:examples}, where we will also explain the visualization of the examples in more detail.
+
+\subsection{Metamodel}
+
+\todo{Marc}{all}{This section is copy/pasted from RENext, we should revise the metamodel and the text.}
+
+Figure~\ref{fig:metamodel} depicts the metamodel linking the main elements of our argumentation extension to the main GRL elements.
+The part below the dashed horizontal line depicts GRL elements.
+A \textsf{GRL Diagram} (bottom) contains zero or more \textsf{IntentionalElements}, either a \textsf{Goal}, a \textsf{Softgoal}, a \textsf{Task}, or a \textsf{Belief}.
+An \textsf{ElementLink} is either a \textsf{Contribution}, a \textsf{Decomposition}, or a \textsf{Dependency} and contains an \textsf{IntentionalElements} as source and as target.
+
+The part above the horizontal dashed line depicts the concepts we introduced in the previous section.
+The top-left element \textsf{Argument} (see Section~\ref{sect:arguments}) can attack other arguments (Section~\ref{sect:attacks}) and generalizes both a \textsf{Formula} and an \textsf{Inference}.
+That is, both these elements can be arguments.
+A formula has an \textsf{AcceptStatus} (Section~\ref{sect:extensions}).
+An \textsf{Inference} is from a set of \textsf{Arguments} as premise and a \textsf{Formula} as conclusion.
+The \textsf{InferenceType} is either strict or defeasible (Section~\ref{sect:aspic}).
+A \textsf{Formula} is either a \textsf{Modality} (modal formula), a \textsf{Proposition}, a \textsf{BinaryOperation}, or a \textsf{Negation} (negated proposition).
+The \textsf{Modality} can be either B (belief), G (goal), E (evidence), or A (action).
+A \textsf{BinaryOperation} is either a \textsf{Disjunction}, a \textsf{Conjunction}, or a \emph{contributes\_to} \textsf{Connective}~(\ref{sect:aspic}).
+
+Finally, the red arrows depict how the two metamodels are integrated.
+The left arrow connects a GRL \textsf{IntentionalElement} with an argumentation \textsf{Modality}, where the mapping is denoted with red text.
+Thus, an intention element traces to an argument, which is always a modal formula.
+The right arrow connects the GRL \textsf{ElementLink} with the argumentation \textsf{Connective} (i.e.
+a \emph{contributes\_to} connective).
+Thus, an arrow between elements of a GRL model corresponds to \emph{contributes\_to} connectives in the argumentation framework.
+
+\begin{figure}[h!]
+\includegraphics[scale=0.61]{img/metamodelimg}
+\caption{The Metamodel \todo{Marc}{all}{This has to be revised.}}
+\label{fig:metamodel}
 \end{figure}

### 2017-07-03 background.tex
Section 1, 2.1, 2.2

--- 2017-03-14 background.tex
@@
 In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.
-Lastly, we introduce argument schemes, and in particular, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, which is an argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-This will be our starting point in the next section.
+Next, we introduce argument schemes, and in particular, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, which is an argument scheme that is used to form arguments and counter-arguments about situations involving goals.
+Finally, we give an overview of the integration between PRAS and GRL.
+%This will be our starting point in the next section.
 
@@
 
-Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise.
-In this exercise, designers are provided a problem description, requirements, and a description of the desired outcomes.
-The problem description is given in full in Appendix~\ref{ch:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering at UCI.
-It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory.
-To this end, a piece of software has to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning since it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}, and transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
+We use the traffic simulator design case to explain the concepts and the framework in this paper.
+Traffic simulator design is the topic of discussion in the transcripts used for argumentation framework.
+%Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise. 
+In this exercise, designers are provided with a problem description, requirements, and a description of the desired outcomes.
+The problem description is given in full in Appendix~\ref{sect:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering courses at an American university.
+In order for the professor to teach students how various theories (such as queuing theory) around traffic lights work, a software analyst is hired to specify the goal and requirements of the system.
+%and It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory.
+To this end, a piece of software will be developed in which students can create visual maps of an area, regulate traffic, and so forth.
+The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning and it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}.
+Transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
 Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
@@
 GRL can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
-A GRL model is a connected graph of intentional elements that optionally are part of actors.
+A GRL model is a connected graph of intentional elements that optionally are part of the actors.
 All the elements and relationships used in GRL are shown in Figure~\ref{fig:grl_legend}.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=0.5\textwidth]{img/example_small}
+\caption{Partial GRL Model of Traffic Simulator Example %(\todo{Marc}{Sepideh}{Turn into vector image, turn ``Dynamic simulation'' and ``Static Simulation'' into goals, turn ``Simulate'' into softgoal, add some resource}}
+\label{fig:example-small}
+\end{figure} %SG: I changed the figure a bit so that we won't need to change Figure 10. Let me know if you still want to change them according to the above.
 
@@
 Softgoals (e.g.
-\texttt{Realistic simulation)} are often more related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are more related to functional requirements.
+\texttt{Show Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:trafficsim}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}, Figure~\ref{fig:trafficsim}).
-
-\begin{figure*}[ht]
-\centering
-\includegraphics[width=\textwidth]{img/transcript_grl_0}
-\caption{GRL Model for the traffic simulator.}
-\label{fig:trafficsim}
-\end{figure*}
+In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}.
+In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}).
 
 Different links connect the elements in a GRL model.
-AND, IOR, and XOR decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:trafficsim}, the goal  \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}.
+AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
+In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
 A contribution link has a qualitative contribution type or a quantitative contribution.
-Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the softgoal  \texttt{Dynamic simulation}.
+Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}.
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-For example, actor  \texttt{Traffic Tycoon} depends on the actor  \texttt{Student} to perform the task  \texttt{Adjust car spawning rate} to fulfill its task  \texttt{Generate cars}.
+The goal \texttt{Generate Cars} depends on the goal \texttt{Value Has Changed}.
+%SG: I added it to the model but we can also remove it altogether.
+%\todo{Marc}{Sepideh}{Shall we add a dependency relation as well? They don't play a big role in our work..}
 
@@
 Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
-GRL has a well-defined syntax and semantics, which are necessary if we want to incorporate it into a formal framework (requirements 1 and 2 as described in the introduction).
+GRL has a well-defined syntax and semantics.
 Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
 GRL is also linked to Use Case Maps via URNLink ((\includegraphics[scale=1]{img/urnlink}) which provides traceability between concepts and instances of the goal model and behavioral design models.
-Multiple views and traceability are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
+Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
 
 GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way.
-The satisfaction values from intentional elements in GRL can also be propagated to use case maps elements.
+The satisfaction values from intentional elements in GRL can also be propagated to the elements of Use Case Maps.
 jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
@@
 This allows GRL to be used in many domains without the need to change the whole modeling language.
-This feature also helps us to apply our argumentation to other domain such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
+This feature also helps us apply our argumentation to other domains such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
 
-The GRL model in Figure~\ref{fig:trafficsim} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
+The GRL model in Figure~\ref{fig:example-small} shows the softgoals, goals, tasks and the relationship between the different intentional elements in the model.
 However, the rationales and arguments behind certain intentional elements are not shown in the GRL model.
@@
 \begin{itemize}
-	\item Why does actor \texttt{Teacher} have only a single softgoal \texttt{Students learn from practice}? Why is this, for instance, not connected to any of the elements of \texttt{Student}?
-	\item What does  \texttt{Adjust timing schemes of sensorless interactions} mean?
-	\item Why does task \texttt{Keep same cars} contribut positively to \texttt{Realistic simulation} and negatively to \texttt{Dynamic simulation}?
-	\item How does the \texttt{Student} control the \texttt{Traffic Tycoon}?
-	\item Why does \texttt{Map design} have so many decompositions into other tasks?
+	\item Why does actor \texttt{Traffic Tycoon} have only a single softgoal \texttt{Show Simulation}?
+Why is this, for instance, not connected to goal \texttt{Simulate}?
+%SG: I changed this question but let me know if you want to change them back and then change the graph.
+	\item What does \texttt{Keep same cars} mean?
+	\item Why does task \texttt{Keep same cars} contribute positively to \texttt{Static simulation} and negatively to \texttt{Dynamic simulation}?
+	\item Why does \texttt{Generate Cars} XOR-decompose into two tasks?
 \end{itemize}
 
-These are the type of the questions that we cannot answer just by looking at the GRL models.
-The model in Figure~\ref{fig:trafficsim} does not contain information about discussions that let up to the resulting elements of the model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+These are the type of the questions that we cannot answer by just looking at GRL models.
+The model in Figure~\ref{fig:example-small} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
 In this article we aim to address this shortcoming.
@@
 These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 4: if we find that  \texttt{Keep same cars} actually negatively influences the value  \texttt{Realistic simulation}, we have a counterargument to the above argument.
-Another way to counter an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that promotes the same value (question 3).
-For example, we can argue that  \texttt{Create new cars} also realizes the goal  \texttt{Generate traffic}, which gives us a counterargument to the original argument -- to generate traffic by simply keeping the cars that disappear off the screen and have them wrap around to the other side of the screen -- that also follows PRAS.
+Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument.
+
+Another way to elaborate an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that contributes to the same softgoal (question 3).
+For example,  can argue that another goal \texttt{Semi-dynamic Simulation} also realizes the softgoal \texttt{Simulate}.
+\todo{Marc}{Sepideh}{Do you think our example is good?
+I think it is confusing.
+Perhaps we should just create an artificial example here, loosely based on the traffic simulator but where the model makes sense to the reader.
+Now the tasks are all a bit vague and I think nobody knows what they actually mean.}
 
 In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa).
-In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}\footnote{Full definitions of Dung's~\cite{Dung1995} frameworks and semantics will be given in section \ref{sect:gmas}.
-In this section, we will briefly discuss the intuitions behind these semantics.}.
-Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the above example: the argument for  \texttt{Keep same cars} (A1), the argument for \texttt{Create new cars} (A3), and the argument that \texttt{Keep same cars} demotes the value  \texttt{Realistic simulation} (A2).
-The two alternative PRAS instantiations are A1 and A3.
-These arguments mutually attack each other, as \texttt{Keep same cars}  and \texttt{Create new cars} are considered to be mutually exclusive.
-Argument A2 attacks A1, as it points to a negative side-effect of \texttt{Keep same cars}.
+In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}.
+Given an argumentation framework, once can compute very precisely which arguments are accepted and which are rejected.
 
-\begin{figure}[ht!]
-\centering
-\begin{tikzpicture}
-        \node[minimum size=1cm] (att3) [argNodeIN] at (0,-3) {$A_3$};
-        \node[minimum size=1cm] (att1) [argNodeOUT] at (0,0) {$A_1$};
-        \node[minimum size=1cm] (att2) [argNodeIN] at (3,0) {$A_2$};
-         \path
-    (att2) edge [attackLink] (att1)
-    (att1) edge [attackLink,<->] (att3);
-\end{tikzpicture}
-\caption{Example argumentation framework.}
-\label{fig:pras:example}
-\end{figure}
-
-Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics.
-The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
-In the argumentation framework in Figure~\ref{fig:pras:example}, argument A2 is undefeated because it has no attackers.
-This makes A1 defeated, because one of its attackers, A2, is undefeated.
-A3 is then also undefeated, since its only attacker, A1, is defeated by A2.
-Thus, the set of undefeated (justified) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A2, A3$\}$, corresponding to arguments for \texttt{Realistic simulation} and \texttt{Create new cars}.
-
-\subsubsection*{Practical Argumentation and Goal Modeling}
+\subsection{Combining PRAS and GRL}
 \label{sect:background:pras:motivation}
 
-Practical reasoning in the PRAS framework as described above provides a formal framework for defeasible reasoning about goals and actions that adheres to the acceptability semantics of Dung~\cite{Dung1995} and its various extensions \cite{amgoud2002reasoning,modgil2009}.
-The usefulness of PRAS for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
+The usefulness of argument schemes, and PRAS in particular, for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
 In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model.
@@
 Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson actually apply to GRL.
-Therefore, we develop our own set of argument schemes and critical questions in the next section by analyzing transcripts of discussions about the traffic simulator.
+Therefore, we develop our own set of argument schemes and critical questions in Section~\ref{sect:gmas} by analyzing transcripts of discussions about the traffic simulator.
+Before doing so, we give an overview of our framework in the next section.

### 2017-07-03 introduction.tex
Section 1, 2.1, 2.2

@@
 \subsection{Problem Statement}
+\label{sect:problem_statement}
 
@@
 However, most real-world information systems -- e.g., air-traffic management, industrial production processes, or government and healthcare services -- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons stated below.
 
@@
 Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions, thus risking unnecessary discussions and revisions.
+If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions which results in having unnecessary discussions and revisions.
 As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
 
-Secondly, other stakeholders, such as new developers on the team who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase.
-If these user have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
+Second, other stakeholders, such as new developers on the team who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase.
+If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
 
-Thirdly, alternative different ideas and opposing views that could potentially have led to different goal diagrams are lost.
+Third, alternative and different ideas and opposing views that could potentially have led to different goal diagrams could be lost.
 For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model will merely contain the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
+Thus, the resulting goal model merely contains the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
 This leads to a poor understanding of the problem and solution domain.
@@
 
-The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with argumentation approaches techniques from Artificial Intelligence (AI) research\cite{atkinson2007}.
+The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research\cite{atkinson2007}.
 We have identified five important requirements for our framework:
 \begin{enumerate}
-\item The argumentation techniques should be close to actual discussions stakeholders or designers have in the early requirements engineering phase.
+\item The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
 \item 
@@
 \item 
-Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and visa versa.
+Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.
 \item 
-The framework must have tool support.
+The framework must have a tool support.
 \item 
-A methodology should be identified that allows the framework to be used by practitioners.
+There should be a methodology for the framework to guide the practitioners in its application in real cases.
 \end{enumerate}
+
+The five requirements above serve as the success criteria of our approach. That is, if our framework satisfies these requirements, we reach our goal.
 
@@
 
-The five requirements above are the success criteria of our approach, that is, the satisfaction of the six requirements will result in a positive answer to the research question.
+%The five requirements above are the success criteria of our approach, that is, the satisfaction of the six requirements will result in a positive answer to the research question.
 
-The main contribution of this article is the RationalGRL framework,  which combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
+%SG: We need to write a bit about the framework in past and refer it here. 
+To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL~\cite{}.
+The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
+%SG: we need to add 1-2 sentences about our previous work and then add what we do here...
 We develop a metamodel specifying how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}.
+%SG: Placeholder for the link.
 We develop an initial list of argument schemes and critical questions by analyzing a set of transcripts containing discussion about an information system.
@@
 We formalize RationalGRL using propositional logic, and develop algorithms for adding goal elements and arguments.
+We illustrates the steps of our framework  and how the framework can be applied in real cases with a traffic simulator example.
 
-\subsection{Organization}
+\subsection{Organization} %SG: Will get back to this at the end of the paper.
 
 This article is organized as follows.
-Section~\ref{sect:background} contains background and introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{}, and argument schemes.
+Section~\ref{sect:background} contains background and introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, and argument schemes.
 Section~\ref{sect:overview} provides a brief and high-level overview of our framework, together with a metamodel and the methodology.

### 2017-07-03 background.tex
added a paragraph to intro

@@
 
-Atkinson et al.~\cite{atkinson2007} develop an argument scheme for practical reasoning, which they call the \emph{practical reasoning argument scheme} (PRAS).
+Atkinson et al.~\cite{atkinson2007} develop an argument scheme for practical reasoning, which they call the \emph{Practical Reasoning Argument Scheme} (PRAS).
 This argument scheme is as follows\footnote{The original argument schemes uses slightly different terminology: we have replaced \emph{value} with \emph{softgoal}.
@@
 Once instantiated, we obtain an argument.
-We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows:
+We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows: %SG: Need to update the example.
 \begin{itemize}
@@
 Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument.
+%SG:placeholder for the examples.
 
@@
 In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model.
-This give a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it allows one to understand why certain other elements have been rejected.
+This approach provides a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it helps understanding why certain other elements have been rejected.
 
 Argumentation schemes and their associated critical questions are very well suited for modeling discussions about a goal model: as Murukannaiah et al.~\cite{murukannaiah2015} have shown, they can guide users in systematically deriving conclusions and making assumptions explicit.
-This can also be seen from the obvious similaries between PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) in the example above.
+This can also be seen from the obvious similarities between PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) in the example above.
 
@@
 For instance, PRAS does not have a notion of ``resource'', and many of the relationships of GRL do not occur in PRAS.
-Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson actually apply to GRL.
+Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson can actually apply to GRL.
 Therefore, we develop our own set of argument schemes and critical questions in Section~\ref{sect:gmas} by analyzing transcripts of discussions about the traffic simulator.

### 2017-07-03 introduction.tex
added a paragraph to intro

@@
 
-%SG: We need to write a bit about the framework in past and refer it here. 
-To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL~\cite{}.
+\todo{Sepideh}{Marc}{Read this paragraph and see if it makes sense.}
+To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL.
 The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
-%SG: we need to add 1-2 sentences about our previous work and then add what we do here...
-We develop a metamodel specifying how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}.
+
+In past, we introduced our preliminary steps towards RationalGRL framework~\cite{bnaic2014,vanzee-etal:renext2015,vanZee-etal:er2016} with its tool-support~\cite{vanZee-etal:comma2016}.
+In our previous work, argument diagrams are integrated with GRL models to allow better representation of stakeholder's arguments and discussions.
+In our current paper, we improve RationalGRL framework by adding argument schemes and critical questions to argumentation framework and provide traceability links between the argumentation framework and GRL.
+For this, we modify and extend our metamodel to specify how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}.
 %SG: Placeholder for the link.

### 2017-07-04 background.tex
added a new small example
fixed the error

@@
 \centering
-\includegraphics[width=0.5\textwidth]{img/example_small}
-\caption{Partial GRL Model of Traffic Simulator Example %(\todo{Marc}{Sepideh}{Turn into vector image, turn ``Dynamic simulation'' and ``Static Simulation'' into goals, turn ``Simulate'' into softgoal, add some resource}}
+\includegraphics[width=0.5\textwidth]{img/Example1}
+\caption{Partial GRL Model of Traffic Simulator Example} %(\todo{Marc}{Sepideh}{Turn into vector image, turn ``Dynamic simulation'' and ``Static Simulation'' into goals, turn ``Simulate'' into softgoal, add some resource}}
 \label{fig:example-small}
@@
 These are the type of the questions that we cannot answer by just looking at GRL models.
-The model in Figure~\ref{fig:example-small} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+The model in Figure~\ref{fig:Example1} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
 In this article we aim to address this shortcoming.

### 2017-07-04 background.tex
Section 1, 2 and the example

@@
 Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
+
+In this Section and Section~\ref{}, we focus on the smaller example of traffic simulator case. In Section~\ref{} we provide the detailed example of the traffic simulator and its model.
 
@@
 Figure~\ref{fig:example-small} illustrates a partial GRL diagram from the traffic simulator design exercise (the full figure is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}).
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Tycoon}, Figure~\ref{fig:example-small}).
+An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Simulator}, Figure~\ref{fig:example-small}).
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
@@
 Softgoals (e.g.
-\texttt{Show Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Generate cars}) are related to functional requirements.
+\texttt{Improve Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Simulate Traffic}) are related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{External Library}).
+In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars} which are alternative way of achieving the goal, \texttt{Generate Cars}.
+In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).
 
@@
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
-The goal \texttt{Generate Cars} depends on the goal \texttt{Value Has Changed}.
+Here, the goal \texttt{Generate Cars} depends on the resource \texttt{Car Objects}.
 %SG: I added it to the model but we can also remove it altogether.
@@
 \begin{itemize}
-	\item Why does actor \texttt{Traffic Tycoon} have only a single softgoal \texttt{Show Simulation}?
-Why is this, for instance, not connected to goal \texttt{Simulate}?
+	\item Why does actor \texttt{Traffic Simulator} have softgoal \texttt{High Performance}which is not linked to any of the goals \texttt{Generate Cars} and \texttt{Simulate Traffic}?
 %SG: I changed this question but let me know if you want to change them back and then change the graph.
 	\item What does \texttt{Keep same cars} mean?
-	\item Why does task \texttt{Keep same cars} contribute positively to \texttt{Static simulation} and negatively to \texttt{Dynamic simulation}?
-	\item Why does \texttt{Generate Cars} XOR-decompose into two tasks?
+	\item Why does task \texttt{Create New cars} contribute negatively to \texttt{Static simulation} and positively to \texttt{Dynamic simulation}?
+	\item Why does \texttt{Simulate Traffic} AND-decompose into two tasks?
 \end{itemize}
@@
 These are the type of the questions that we cannot answer by just looking at GRL models.
-The model in Figure~\ref{fig:Example1} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+The model in Figure~\ref{fig:example-small} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
 In this article we aim to address this shortcoming.
@@
 
+\todo{Sepideh}{Marc}{Read this part and see if it makes sense.}
 \begin{itemize}
@@
 Once instantiated, we obtain an argument.
-We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows: %SG: Need to update the example.
+We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows:
 \begin{itemize}
-\item[] We have goal \texttt{Static simulation},
-\item[] Doing action \texttt{Keep same cars} realizes goal \texttt{Static simulation}, 
-\item[] Which contribute positively to the softgoal \texttt{Simulate} 
+\item[] We have goal \texttt{Simulate Traffic},
+\item[] Doing action \texttt{Dynamic Simulation} realizes goal \texttt{Simulate Traffic}, 
+\item[] Which contribute positively to the softgoal \texttt{Improve Simulation} 
 \item[] \textit{Therefore} 
-\item[] We should perform action \texttt{Keep same cars}.
+\item[] We should perform action \texttt{Dynamic Simulation}.
 \end{itemize}
@@
 These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 1: if we find that  \texttt{Keep same cars} actually does not realize goal \texttt{Static Simulation}, we can form a counter-argument.
-%SG:placeholder for the examples.
+Take, for example, critical question 1: if we find that  \texttt{Dynamic Simulation} actually does not realize goal \texttt{Simulate Traffic}, we can form a counter-argument.
 
 Another way to elaborate an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that contributes to the same softgoal (question 3).
-For example,  can argue that another goal \texttt{Semi-dynamic Simulation} also realizes the softgoal \texttt{Simulate}.
-\todo{Marc}{Sepideh}{Do you think our example is good?
+For example,  can argue that another task \texttt{Semi-dynamic Simulation} also realizes the goal \texttt{Simulate Traffic}.
+%\todo{Marc}{Sepideh}{Do you think our example is good?
 I think it is confusing.

### 2017-07-04 gmas.tex
changes

@@
 
-In Section~\ref{sect:gmas:examples} we provide examples we found in the transcripts for all of these four effects.
+In Section~\ref{sect:examples} we provide examples we found in the transcripts for all of these four effects.
 
@@
 Our analysis however shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
-However, if needed, additional argument schemes and critical questions can be added, and our list is not meant to be exhaustive.
+However, additional argument schemes and critical questions can be added freely to our framework, and our list is not meant to be exhaustive.
 

### 2017-07-04 discussion.tex

@@
 
+\todo{Marc}{all}{Review the following notes:
+
+\begin{itemize}
+\item RE17 paper: merging argumentation theory with law
+  "Using Argumentation to Explain Ambiguity in Requirements Elicitation Interviews" Yehia Elrakaiby, Alessio Ferrari, Paola Spoletini, Stefania Gnesi and Bashar Nuseibeh
+\item more details on Murukannaiah [RE15]
+\item related work on argumentation theory
+\item First paper we wrote they also said something about Bashar Nuseibeh
+\item Also mention argumentation in AI and how that relates to our work
+\item Also paper by Daniel Amyot that he once sent to us:
+  "Synergy between Activity Theory and goal/scenario modeling for requirements elicitation, analysis, and evolution"
+\item Maybe this one:
+  "Complete Traceability for Requirements in Satisfaction Arguments"
+  Anitha Murugesan, Michael Whalen, Elaheh Ghassabani and Mats Heimdahl
+  (University of Minnesota, United States)
+
+\item  At least we can say our work is different from this if it is unrelated..
+\item We should make the comparison more in depth.
+\item Look at related work of Jureta and see how it is related with us.
+\item Maybe the papers above (e.g., Bashar) have related work that relates to us as well.
+\end{itemize}
+}
 \subsection{Open issues}
@@
 
+\todo{Marc}{all}{Finish this}
+
 The introduction of this article contains five requirements we identified for our framework. We use the conclusion to discuss how RationalGRL meets our initial requirements.

### 2017-07-04 examples.tex

@@
 
+\todo{Marc}{all}{Check whether the text matches the figure here}
 The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.

### 2017-07-04 overview.tex
Part of section 3.1

@@
 
-\subsection{RationalGRL Methodology}
+\subsection{RationalGRL Methodology} \todo{Sepideh}{Marc}{Review this section and make sure it is correct}
 
-As mentioned earlier, the RationalGRL framework aims at integrating GRL models with the underlying discussions and arguments of the stakeholders during the analysis phase by combining PRAS with GRL.
-Our argumentation framework is developed based on the analysis of the CQs, practical reasoning and discussions of the stakeholders of the system.
-Initial GRL models are first created based on the analysis of non-functional requirements in the requirements specification documents, discussions with stakeholders and by refining the high-level goals into operationalized tasks.
+As mentioned in Section~\ref{sect:introduction}, the RationalGRL framework has been extended with concepts from practical reasoning argument scheme (PRAS) to help integrating goal models with the detailed discussions and arguments the stakeholders pose during the analysis phase.
+That is, the RationalGRL framework includes two main parts: Argumentation model and GRL model.
 
-In RationalGRL, both GRL and argumentation models are done iteratively.
-Argumentation models impact GRL models and result in refinement and changes in GRL.
-For example, an answer to the CQ \emph{Is the task possible?} can lead to rejection or acceptance of the task in GRL.
-The analysis of GRL models may result in changes in the underlying argumentation framework as well.
-For example, adding a new IE to the GRL model can result in posing new CQs with respect to the IE and its relationships.
-Figure~\ref{fig:rationalgrl-framework} shows RationalGRL and the relationships between the different parts.
-On the right-hand side of the framework, we have GRL models and on the left-hand side we have argumentation framework.
-The links between the two sides illustrates the impacts each side has on the other side and the changes that occur due to the impacts and the refinements.
+To develop the argumentation model, we need to analyze critical questions, argument schemes and the stakeholder's discussions.
+For the GRL part of the framework, we first need to create the "initial" GRL models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
+These two parts, GRL and argumentation models are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
+For example, answering a critical question \emph{Is the task \texttt{A} possible?}, instantiated from the argumentation model, can result in removing or adding a task in the GRL model.
+Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships with other intentional elements.
+Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework with its components and their relationships.
+The GRL models  are shown  on the right-hand side of the framework while the argumentation models is on the left-hand side.
+The links between the two sides illustrates the impacts and relationships between two sides.
 
@@
 
-To develop a RationalGRL model, we follow the methodology shown  in Figure~\ref{fig:rationalgrl-methodology}. We already have an initial GRL model. The steps of the methodology are as follows:
+We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework.
+Note that, here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+The rest of the steps are as follows:
 
@@
 
-\textbf{Instantiate Argument Schemes (AS)} -- In this step, we instantiate one of the AS from the argumentation framework.
+%TODO I need to reword up to the example.
+
+\textbf{Instantiate Argument Schemes (AS)} -- After creating the initial GRL models, we In the first step, we instantiate one of the AS from the argumentation framework.
 In the section we develop 13 arguments schemes which we use in our analysis.
@@
 
-\subsubsection*{Example}
+\subsubsection*{Example} \todo{Sepideh}{Marc}{Do we need an example here? I think no...this is from istar also, if we want it, we need to modify it}
 \label{sect:methodology-example}

### 2017-07-10 overview.tex
Section 3.1 and 3.2 has been rewritten

@@
 In this section we present a high-level overview of the RationalGRL framework.
-In the first subsection we present a methodology, specifying how practitioners can use RationalGRL to create goal models with traceability links to underlying arguments.
-In the second subsection we link the new argumentation elements and relations to the existing elements and relations of GRL in a metamodel.
+In the first subsection, we present a methodology, specifying how practitioners can use RationalGRL to create goal models with traceability links to underlying arguments.
+In the second subsection, we link the new argumentation elements and relations to the existing elements and relations of GRL in a metamodel.
 This metamodel serves as the specification for an implementation.
@@
 To develop the argumentation model, we need to analyze critical questions, argument schemes and the stakeholder's discussions.
-For the GRL part of the framework, we first need to create the "initial" GRL models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
+For the GRL part, we first need to create the "initial" GRL models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
 These two parts, GRL and argumentation models are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
 For example, answering a critical question \emph{Is the task \texttt{A} possible?}, instantiated from the argumentation model, can result in removing or adding a task in the GRL model.
-Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships with other intentional elements.
+Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
 Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework with its components and their relationships.
@@
 The links between the two sides illustrates the impacts and relationships between two sides.
+Note that, any of the critical question, their answer and the argument schemes that are instantiated during the analysis phase of the GRL model are documented with the GRL model and can be referred to in the future.
 
@@
 We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework.
-Note that, here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
 The rest of the steps are as follows:
@@
 
-%TODO I need to reword up to the example.
+\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes of the argumentation framework.
+We select an intentional element from the initial GRL model that we want to analyze and we instantiate a relevant argument scheme from the already existing list of argument schemes or by adding a new one.
+For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}".
+When an argument scheme is instantiated, it corresponds to  an argument for or against part of a goal model.
 
-\textbf{Instantiate Argument Schemes (AS)} -- After creating the initial GRL models, we In the first step, we instantiate one of the AS from the argumentation framework.
-In the section we develop 13 arguments schemes which we use in our analysis.
-An example of an argument scheme is "Goal \emph{G} contributes to softgoal \emph{S}".
-An instantiation of an argument scheme thus corresponds to an argument for or against part of a goal model.
+\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions o attack the argument by counter-arguments.
+Each argument scheme includes one or more critical question.
+For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}.
+It is worth mentioning that, answering a critical question may result in a "conflict" situation.
+This is out of the scope of our work here.
+When the analyst answers  a critical question, a new argument scheme may be instantiated.
+Thus, it is possible to go back and forth between this step and the step (1).
 
-\textbf{Answer Critical Questions (CQs)} -- Instantiated arguments can be attacked by counter-arguments.
-CQs are ways in which AS can be attacked.
-Each argument scheme includes one or more CQs.
-For example, for the argument scheme mentioned above, we have two CQs as: \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}.
-Note that, the answer to CQ can also result into ``conflict'' situation which we do not consider here.
-Answering a CQ may result in an instantiation of a new AS.
-Thus, it is possible to go back and forth between this step and the previous one.
-
-\textbf{Decide on IE and the Relationships} -- The answers to the CQs can result in one of the four cases which impact the arguments and corresponding GRL IE: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.
-\textsf{INTRO} means that the argument scheme of the CQ does not get attacked and instead it creates a new argument.
-\textsf{DISABLE} means that the IE or the relationship related to the AS needs to be disabled or removed from the models.
-\textsf{REPLACE} introduces a new argument and attack the original argument at the same time.
-It basically replace the original element of the AS with a new one.
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the four following cases can occur: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.
+Any of these cases can  impact the arguments and corresponding GRL intentional elements.
+\textsf{INTRO} means that
+a new argument scheme is created.
+That means, the current argument scheme related to the critical question does not get attacked.
+In the case of \textsf{DISABLE}, the intentional element or its related links are disabled or removed from the models.
+\textsf{REPLACE} introduces a new argument and attacks the original argument at the same time.
+This means that the original element of the argument scheme is replaced with a new one.
 \textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when a new evidence occur.
 
-\textbf{Modify GRL Models} -- Based on the result of step three, the GRL models can be modified.
-In this case, either a new IE or a new relationships is introduced, an existing IE or relationship gets disabled (removed) from the model or finally an existing IE or relationships is replaced by a new one.
+\textbf{(4) Modify GRL Models} -- In this step, we modify the GRL models based on the situation of step (3).
+That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
 This results in a new modified GRL.
-The new GRL model can then impact the argument schemes and instantiate another argument scheme.
+The new GRL model can then impact the argument schemes and instantiate another argument scheme (Step (1)).
+
+We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
 
@@
 
-\subsubsection*{Example} \todo{Sepideh}{Marc}{Do we need an example here? I think no...this is from istar also, if we want it, we need to modify it}
+\subsubsection*{Example} \todo{Sepideh}{Marc}{Is this part ok?
+I am not sure if we should have this section at all here.
+It may make things more complicated.
+Better to have it all in the example section}
 \label{sect:methodology-example}
 
-Our examples comes from a transcript containing discussions about the development of an information system, and they are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+In this section, we present a small example for the RationalGRL methodology.
+This example, which is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
+We base the argument schemes and the critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
+These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
 We provide several more examples in Section~\ref{sect:examples}.
 
-A group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \emph{simulate} of the system using the RationalGRL methodology.
+As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology.
 This proceeds in the following way:
@@
 \item
-First they start at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), and add the IE \texttt{Simulate} to GRL model (Figure~\ref{fig:examples:decomposition}, GRL model, top IE).
+First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
+with the initial GRL model (Figure~\ref{fig:example-small})  and they invoke the intentional element \texttt{Simulate Traffic}.
+%SG: We say above that we have an initial GRL model, so how do we match it with this sentence?
+I tried to reword it but it doesn't make so much sense...
 \item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
-They answer CQ \emph{Does Simulate AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic simulation'' and ``Static simulation}.
+They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation}.
 \item As a result, two argument schemes are created, namely:
 \begin{itemize}
-\item Actor \emph{System} has task \emph{Dynamic simulation}
-\item Actor \emph{System} has task \emph{Static simulation}
+\item Actor \texttt{Traffic Simulator} has task \texttt{Dynamic Simulation}
+\item Actor \texttt{Traffic Simulator} has task \texttt{Static Simulation}
 \end{itemize}
-\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate} to these two tasks.
-\item Next, the stakeholders test the validity of their goal model by answering more CQ. They answer two CQs:
+\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
+\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
 \begin{itemize}
 \item
-CQ \emph{Is task ``Dynamic simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
-Thus, the corresponding GRL IE is disabled.
+CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
+Thus, the corresponding GRL intentional element is disabled.
 \item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.

### 2017-07-23 background.tex
metamodel

@@
 We use the traffic simulator design case to explain the concepts and the framework in this paper.
-Traffic simulator design is the topic of discussion in the transcripts used for argumentation framework.
+Traffic simulator design is the topic of discussion in the transcripts used for the argumentation framework.
 %Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise. 
@@
 
-In this Section and Section~\ref{}, we focus on the smaller example of traffic simulator case. In Section~\ref{} we provide the detailed example of the traffic simulator and its model.
+In this Section and in Section~\ref{}, we focus on the simplified example of traffic simulator case for simplicity.
+In Section~\ref{} we provide the detailed example of the traffic simulator and its model.
 
@@
 \includegraphics[width=0.5\textwidth]{img/Example1}
-\caption{Partial GRL Model of Traffic Simulator Example} %(\todo{Marc}{Sepideh}{Turn into vector image, turn ``Dynamic simulation'' and ``Static Simulation'' into goals, turn ``Simulate'' into softgoal, add some resource}}
+\caption{Partial GRL Model of Traffic Simulator Example}
 \label{fig:example-small}
 \end{figure} %SG: I changed the figure a bit so that we won't need to change Figure 10. Let me know if you still want to change them according to the above.
+%MvZ: The figure looks good to me!
 
@@
 
-Figure~\ref{fig:example-small} illustrates a partial GRL diagram from the traffic simulator design exercise (the full figure is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}).
+Figure~\ref{fig:example-small} illustrates a simplified GRL diagram from the traffic simulator design exercise (a full example is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}).
+%MvZ: Since the example no longer matches the full figure I reworded it slightly in the sentence above.
 An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Simulator}, Figure~\ref{fig:example-small}).
@@
 %\todo{Marc}{Sepideh}{Shall we add a dependency relation as well? They don't play a big role in our work..}
+%MvZ: It is a bit strange because we say dependencies are between actors but our examples does show this. Perhaps better to leave it out?
 
@@
 Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-GRL is also linked to Use Case Maps via URNLink ((\includegraphics[scale=1]{img/urnlink}) which provides traceability between concepts and instances of the goal model and behavioral design models.
+GRL is also linked to Use Case Maps via URNLinks (\includegraphics[scale=1]{img/urnlink}) which provide traceability between concepts and instances of the goal model and behavioral design models.
 Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
@@
 \begin{itemize}
-	\item Why does actor \texttt{Traffic Simulator} have softgoal \texttt{High Performance}which is not linked to any of the goals \texttt{Generate Cars} and \texttt{Simulate Traffic}?
+	\item Why does actor \texttt{Traffic Simulator} have softgoal \texttt{High Performance}, which is not linked to any of the goals \texttt{Generate Cars} and \texttt{Simulate Traffic}?
 %SG: I changed this question but let me know if you want to change them back and then change the graph.
@@
 
-\todo{Sepideh}{Marc}{Read this part and see if it makes sense.}
+%\todo{Sepideh}{Marc}{Read this part and see if it makes sense.}
+%MvZ: Yes it makes sense, I made some small adjustments but overall it looks good.
 \begin{itemize}
 \item[] We have goal $G$,
-\item[] Doing action $A$ realizes to goal $G$,
+\item[] Doing action $A$ realizes goal $G$,
 \item[] Which will contribute positively to the softgoal $S$
@@
 For example,  can argue that another task \texttt{Semi-dynamic Simulation} also realizes the goal \texttt{Simulate Traffic}.
-%\todo{Marc}{Sepideh}{Do you think our example is good?
-I think it is confusing.
-Perhaps we should just create an artificial example here, loosely based on the traffic simulator but where the model makes sense to the reader.
-Now the tasks are all a bit vague and I think nobody knows what they actually mean.}
 
 In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa).
-In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations as introduced by Dung~\cite{Dung1995}.
-Given an argumentation framework, once can compute very precisely which arguments are accepted and which are rejected.
+In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations.
+Given an argumentation framework, once can compute very precisely which arguments are accepted and which are rejected using argumentation semantics~\cite{Dung1995}.
+These semantics have been studied extensively in AI, and are currently a popular topic of research (see~\cite{argumentationbook} for an overview).
 
@@
 In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model.
-This approach provides a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it helps understanding why certain other elements have been rejected.
+This approach provides a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it helps understanding why certain other elements have been accepted and others have been rejected.
 

### 2017-07-23 examples.tex
metamodel

@@
 
-\subsubsection{Example 1: Disable task \texttt{Traffic light}}
+\subsubsection{Example 1: Methodology example}
+
+In this section, we present a small example for the RationalGRL methodology.
+This example, which is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
+We base the argument schemes and the critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
+These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+We provide several more examples in Section~\ref{sect:examples}.
+
+As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology.
+This proceeds in the following way:
+\begin{itemize}
+\item
+First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
+with the initial GRL model (Figure~\ref{fig:example-small})  and they invoke the intentional element \texttt{Simulate Traffic}.
+%SG: We say above that we have an initial GRL model, so how do we match it with this sentence?
+I tried to reword it but it doesn't make so much sense...
+\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
+They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation}.
+\item As a result, two argument schemes are created, namely:
+\begin{itemize}
+\item Actor \texttt{Traffic Simulator} has task \texttt{Dynamic Simulation}
+\item Actor \texttt{Traffic Simulator} has task \texttt{Static Simulation}
+\end{itemize}
+\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
+\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
+\begin{itemize}
+\item
+CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
+Thus, the corresponding GRL intentional element is disabled.
+\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
+\end{itemize}
+\end{itemize}
+
+The resulting RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
+
+\subsubsection{Example 2: Disable task \texttt{Traffic light}}
 
@@
 
-\subsubsection{Example 2: Clarify task \texttt{Road pattern}}
+\subsubsection{Example 3: Clarify task \texttt{Road pattern}}
 
@@
 
-\subsubsection{Example 3: Decompose goal \texttt{Simulate}}
+\subsubsection{Example 4: Decompose goal \texttt{Simulate}}
 
@@
 
-\subsubsection{Example 4: Reinstate actor \texttt{Development team}}
+\subsubsection{Example 5: Reinstate actor \texttt{Development team}}
 

### 2017-07-23 introduction.tex
metamodel

@@
 Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not possible to remind stakeholders of their previous opinions which results in having unnecessary discussions and revisions.
+If previous preferences and opinions are not stored explicitly, it is not straightforward to remind stakeholders of their previous opinions which results in having unnecessary discussions and revisions.
 As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
@@
 
-The aim of this research is to resolve the above issues by developing a framework with a tool-support, which combines goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research\cite{atkinson2007}.
+The aim of this research is to resolve the above issues by developing a framework with tool-support, which combines goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{atkinson2007}.
 We have identified five important requirements for our framework:
@@
 
-%The five requirements above are the success criteria of our approach, that is, the satisfaction of the six requirements will result in a positive answer to the research question.
-
-\todo{Sepideh}{Marc}{Read this paragraph and see if it makes sense.}
 To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL.
@@
 Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains.
-We develop a methodology for using RationalGRL, which consists developing goal models and posing arguments in an integrated way.
+We develop a methodology for using RationalGRL, which consists of developing goal models and posing arguments in an integrated way.
 This methodology is intended to be used by practitioners in the field.

### 2017-07-23 overview.tex
metamodel

@@
 
-\subsection{RationalGRL Methodology} \todo{Sepideh}{Marc}{Review this section and make sure it is correct}
+\subsection{RationalGRL Methodology} 
+%\todo{Sepideh}{Marc}{Review this section and make sure it is correct}
 
-As mentioned in Section~\ref{sect:introduction}, the RationalGRL framework has been extended with concepts from practical reasoning argument scheme (PRAS) to help integrating goal models with the detailed discussions and arguments the stakeholders pose during the analysis phase.
-That is, the RationalGRL framework includes two main parts: Argumentation model and GRL model.
+As mentioned in Section~\ref{sect:introduction}, the RationalGRL framework uses concepts from practical reasoning argument scheme (PRAS) to help integrating goal models with the detailed discussions and arguments the stakeholders pose during the analysis phase.
+That is, the RationalGRL framework includes two main parts: Argumentation modeling and GRL modeling.
 
-To develop the argumentation model, we need to analyze critical questions, argument schemes and the stakeholder's discussions.
-For the GRL part, we first need to create the "initial" GRL models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
-These two parts, GRL and argumentation models are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
+For the GRL part, we first need to create the ``initial'' GRL model by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
+For the argumentation part, arguments and counterarguments are put forward about various parts of the goal model.
+These two parts, GRL and argumentation, are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
 For example, answering a critical question \emph{Is the task \texttt{A} possible?}, instantiated from the argumentation model, can result in removing or adding a task in the GRL model.
@@
 Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework with its components and their relationships.
-The GRL models  are shown  on the right-hand side of the framework while the argumentation models is on the left-hand side.
-The links between the two sides illustrates the impacts and relationships between two sides.
-Note that, any of the critical question, their answer and the argument schemes that are instantiated during the analysis phase of the GRL model are documented with the GRL model and can be referred to in the future.
+The GRL model is shown  on the right-hand side of the framework while the argumentation model is on the left-hand side.
+The links between the two sides illustrate the impacts and relationships between two sides.
+Note that answer to critical questions and argument schemes that are instantiated during the analysis phase of the GRL model are documented with the GRL model and can be referred to in the future.
 
@@
 
-\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions o attack the argument by counter-arguments.
-Each argument scheme includes one or more critical question.
+\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions to attack the argument by counter-arguments.
+Each argument scheme includes one or more critical questions.
 For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}.
-It is worth mentioning that, answering a critical question may result in a "conflict" situation.
-This is out of the scope of our work here.
-When the analyst answers  a critical question, a new argument scheme may be instantiated.
-Thus, it is possible to go back and forth between this step and the step (1).
+%It is worth mentioning that, answering a critical question may result in a "conflict" situation. This is out of the scope of our work here.  
+%MvZ: removed this sentence because I don't know what you mean with "conflict". I think We should explain it better or leave it out.
+When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and the step (1).
 
@@
 This means that the original element of the argument scheme is replaced with a new one.
-\textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when a new evidence occur.
+\textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when new evidence occurs.
 
@@
 We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
-
-In the next section, we provide a short example to illustrate the relationship between argumentation framework and goal models.
-
-\subsubsection*{Example} \todo{Sepideh}{Marc}{Is this part ok?
-I am not sure if we should have this section at all here.
-It may make things more complicated.
-Better to have it all in the example section}
-\label{sect:methodology-example}
-
-In this section, we present a small example for the RationalGRL methodology.
-This example, which is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
-We base the argument schemes and the critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
-These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-We provide several more examples in Section~\ref{sect:examples}.
-
-As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology.
-This proceeds in the following way:
-\begin{itemize}
-\item
-First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
-with the initial GRL model (Figure~\ref{fig:example-small})  and they invoke the intentional element \texttt{Simulate Traffic}.
-%SG: We say above that we have an initial GRL model, so how do we match it with this sentence?
-I tried to reword it but it doesn't make so much sense...
-\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
-They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation}.
-\item As a result, two argument schemes are created, namely:
-\begin{itemize}
-\item Actor \texttt{Traffic Simulator} has task \texttt{Dynamic Simulation}
-\item Actor \texttt{Traffic Simulator} has task \texttt{Static Simulation}
-\end{itemize}
-\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
-\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
-\begin{itemize}
-\item
-CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
-Thus, the corresponding GRL intentional element is disabled.
-\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
-\end{itemize}
-\end{itemize}
-
-The resulting RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}, which can be found in Section~\ref{sect:examples}, where we will also explain the visualization of the examples in more detail.
 
@@
 
-\begin{figure}[h!]
-\includegraphics[scale=0.61]{img/metamodelimg}
-\caption{The Metamodel \todo{Marc}{all}{This has to be revised.}}
+\begin{figure*}[h!]
+\includegraphics[scale=0.61]{metamodel/metamodel}
+\caption{The RationalGRL metamodel}
 \label{fig:metamodel}
-\end{figure}
+\end{figure*}

### 2017-07-25 overview.tex
metamodel

@@
 
-\todo{Marc}{all}{This section is copy/pasted from RENext, we should revise the metamodel and the text.}
+Figure~\ref{fig:metamodel} depicts the RationalGRL metamodel linking the main elements of our argumentation extension to the main GRL elements.
+We describe the metamodel bottom-up, starting with the GRL package.
 
-Figure~\ref{fig:metamodel} depicts the metamodel linking the main elements of our argumentation extension to the main GRL elements.
-The part below the dashed horizontal line depicts GRL elements.
-A \textsf{GRL Diagram} (bottom) contains zero or more \textsf{IntentionalElements}, either a \textsf{Goal}, a \textsf{Softgoal}, a \textsf{Task}, or a \textsf{Belief}.
-An \textsf{ElementLink} is either a \textsf{Contribution}, a \textsf{Decomposition}, or a \textsf{Dependency} and contains an \textsf{IntentionalElements} as source and as target.
+The GRL package of our metamodel consists of the core GRL concepts, which constitute a part of
+the URN metamodel from Recommendation Z.151~\cite{URN}.
+These concepts represent the abstract grammar of the language, independently of the notation.
+This metamodel also formalizes the GRL concepts and constructs introduced earlier\footnote{Note that for readability, some GRL concepts ave been omitted from the figure.
+For instance, a GRL contribution can have a qualitative strength, ranging from ``Make'' to ``Break''.
+Since these concepts are not relevant to our framework, they have been omitted, but none of the results of this article depend on this.}.
 
-The part above the horizontal dashed line depicts the concepts we introduced in the previous section.
-The top-left element \textsf{Argument} (see Section~\ref{sect:arguments}) can attack other arguments (Section~\ref{sect:attacks}) and generalizes both a \textsf{Formula} and an \textsf{Inference}.
-That is, both these elements can be arguments.
-A formula has an \textsf{AcceptStatus} (Section~\ref{sect:extensions}).
-An \textsf{Inference} is from a set of \textsf{Arguments} as premise and a \textsf{Formula} as conclusion.
-The \textsf{InferenceType} is either strict or defeasible (Section~\ref{sect:aspic}).
-A \textsf{Formula} is either a \textsf{Modality} (modal formula), a \textsf{Proposition}, a \textsf{BinaryOperation}, or a \textsf{Negation} (negated proposition).
-The \textsf{Modality} can be either B (belief), G (goal), E (evidence), or A (action).
-A \textsf{BinaryOperation} is either a \textsf{Disjunction}, a \textsf{Conjunction}, or a \emph{contributes\_to} \textsf{Connective}~(\ref{sect:aspic}).
+The GRL specification consists of \textsf{GRLModelElements}, which can be either \textsf{GRLLinkableElements} or \textsf{ElementLinks}.
+A \textsf{GRLLinkableElement} can again be specialized into an \textsf{Actor} or an \textsf{IntentionalElement} (which is either a \textsf{Softgoal}, \textsf{Goal}, \textsf{Task}, \textsf{Resource}, or a \textsf{Belief}).
+Intentional elements can be part of an actor, and \textsf{GRLLinkableElements} are connected through \textsf{ElementLinks} of different types (i.e., \textsf{Contribution, Decomposition}, or \textsf{Dependency}).
+Note that actors can be connected through links as well, which is done wth \textsf{Dependency} links.
 
-Finally, the red arrows depict how the two metamodels are integrated.
-The left arrow connects a GRL \textsf{IntentionalElement} with an argumentation \textsf{Modality}, where the mapping is denoted with red text.
-Thus, an intention element traces to an argument, which is always a modal formula.
-The right arrow connects the GRL \textsf{ElementLink} with the argumentation \textsf{Connective} (i.e.
-a \emph{contributes\_to} connective).
-Thus, an arrow between elements of a GRL model corresponds to \emph{contributes\_to} connectives in the argumentation framework.
+The Argumentation package depicts the concepts we introduced in the previous sections.
+An \textsf{ArgumentScheme} represents an (uninstantiated) scheme containing variables that can be replaced with intentional elements.
+\textsf{CriticalQuestions} are possible ways to attack or elaborate an argument scheme.
+As such, each critical question applies to exactly one scheme, but each scheme can be elaborated or attack through multiple critical questions.
+When an argument scheme is instantiated, we obtain an \textsf{Argument}.
+Therefore, each argument is associated with exactly one scheme, but a scheme can be instantiated in multiple ways.
+When a critical question is answered, we obtain an \textsf{AttackLink}.
+Each \textsf{AttackLink} is associated with at most one critical question, but a critical question can be used to attack multiple arguments.
+Note that an \textsf{AttackLink} can also be associated with no critical questions.
+This allows the user to create attacks between arguments, which do not necessarily correspond to one of the critical questions.
+A \textsf{RationalGRLDiagram} is composed out of arguments and attack relations.
+
+There is only one link between the \textsf{GRL} package and the \textsf{Argumentation} package, but it is a very important one.
+The link specifies that each \textsf{GRLModelElement} is in fact an argument.
+This means that each model element inherits the \textsf{AcceptStatus} as well, allowing GRL elements to be accepted, rejected, or undecided.
+This furthermore means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme.
+Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements, but can come from an external source, for instance a piece of evidence or an expert opinion.
+We will see various examples of such arguments in the next sections.
 
 \begin{figure*}[h!]
-\includegraphics[scale=0.61]{metamodel/metamodel}
+\includegraphics[width=\textwidth]{metamodel/metamodel}
 \caption{The RationalGRL metamodel}

### 2017-08-03 overview.tex
few typos

@@
 Intentional elements can be part of an actor, and \textsf{GRLLinkableElements} are connected through \textsf{ElementLinks} of different types (i.e., \textsf{Contribution, Decomposition}, or \textsf{Dependency}).
-Note that actors can be connected through links as well, which is done wth \textsf{Dependency} links.
+Note that actors can be connected through links as well, which is done with \textsf{Dependency} links.
 
@@
 \textsf{CriticalQuestions} are possible ways to attack or elaborate an argument scheme.
-As such, each critical question applies to exactly one scheme, but each scheme can be elaborated or attack through multiple critical questions.
+As such, each critical question applies to exactly one scheme, but each scheme can be elaborated or attacked through multiple critical questions.
 When an argument scheme is instantiated, we obtain an \textsf{Argument}.
@@
 This means that each model element inherits the \textsf{AcceptStatus} as well, allowing GRL elements to be accepted, rejected, or undecided.
-This furthermore means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme.
+This, furthermore, means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme.
 Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements, but can come from an external source, for instance a piece of evidence or an expert opinion.

### 2017-08-11 gmas.tex
changed many things

--- 2017-06-27 gmas.tex
@@
 
-In the previous section we provided an overview of RationalGRL.
-In the example of the methodology, we discussed various argument schemes and critical questions.
+In previous section, we provided an overview of RationalGRL methodology.
 In this section, we discuss our argument schemes and critical questions in more detail.
-It is important to note that the list of argument schemes we obtain in this section is not exhaustive.
-It is an initial list that we obtained by annotating transcripts, but our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
+It is important to note that the list of argument schemes we present in this section is not exhaustive.
+It is an initial list that we have obtained by annotating transcripts, however, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
 
-The list of argument schemes and critical questions that we obtained from our analysis is shown in Table~\ref{table:argument-schemes}.
-The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are about relationships, the next two (AS12-AS13) are about intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
+The list of argument schemes and critical questions that we have obtained from our analysis is shown in Table~\ref{table:argument-schemes}.
+The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are related to relationships, the next two (AS12-AS13) are for intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
 
-As we already discussed in the methodology, we found that answering critical questions can have varying effects on the model, and for each critical question, the right column in Table~\ref{table:argument-schemes} shows the effect of answering the critical questions affirmatively.
-Answering a critical questions can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
+As we have already discussed in Section~\ref{sect:overview}, answering critical questions can have different impacts on the model.
+The right column in Table~\ref{table:argument-schemes} shows the impact of answering each critical question affirmatively.
+%SG: what does affirmatively mean?
+As mentioned earlier, answering a critical question can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
 
-In the first subsection of this section, we provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts), after which we analyze our results in the second subsection.
+In the following subsections, we first provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts) and then, we analyze our results.
 
@@
 
-\paragraph{Experimental Setup} The assignment used for the experiments is to design a traffic simulator.
+\paragraph{Experimental Setup} The goal of the experiment is to design a traffic simulator.
 Participants were asked to use a think-aloud method during the design session.
-The assignment was slightly adjusted to include several viewpoints as end products in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
+The assignment was slightly adjusted to include several viewpoints as an end product in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
 The full problem descriptions can be found in Appendix~\ref{sect:designprompt}.
 All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software.
-The students had two hours for the tasks, and the transcripts document the entire discussion.
+The students had two hours for the tasks.
+The entire discussion was documented in transcripts.
 The details of the transcripts are shown in Table~\ref{table:transcripts:info}.
@@
 
-\paragraph{Annotation Method}
+\paragraph{Process of Extracting Argument Schemes and Critical Questions} %TODO I changed the title from annotation method.
 
-We started with an initial list of 8 argument schemes and 18 critical questions that we derived from PRAS (AS1-AS4, AS6-AS9 of Table~\ref{table:argument-schemes}).
-We annotated transcripts with the arguments and critical questions from this list.
-If we found arguments or critical questions that did not appear in the original list, we added them and counted them as well.
-Argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}).
-Most of the occurrences were not literally found back, but had to be inferred from the context.
-This can be seen in the various examples we will discuss.
+To develop the list of argument schemes and critical questions, we started with an initial list of 8 argument schemes and 18 critical questions that we derived from PRAS (AS1-AS4, AS6-AS9 of Table~\ref{table:argument-schemes}).
+We, then, annotated transcripts with arguments and critical questions from this initial list.
+If there were arguments or critical questions that did not appear in the initial list, we added them to the list.
+Existing argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}).
+%TODO I don't understand it!
+why?
+
+Most of the occurrences were not literally found back, but had to be inferred from the context. This can be seen in the various examples we will discuss.
 
 It is generally known in the argumentation literature that it can be very difficult to identify arguments in natural language texts~\cite{walton-etal2004}.
-Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in the transcripts.
-However, there is hardly any research on on argument extraction in the requirement engineering domains, so despite this potential weakness in our approach, we believe it nevertheless is at least useful as something that others can build further on (see Section~\ref{sect:goalmodeling:openissues}).
+Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in transcripts.
+However, there is hardly any research on argument extraction in the requirement engineering domain.
+Thus, despite this potential weakness in our approach, we believe it nevertheless is at least useful as something that others can build further on (see Section~\ref{sect:goalmodeling:openissues}).
+%TODO I don't understand the last few sentences...
+what does argument extraction mean?
+why is it needed?
+don't use "something that others" it is very informal for a journal paper.
 
@@
 \label{fig:transcripts:grl}
-\end{figure*}
+\end{figure*} %TODO Figure doesn't match the text!
 
@@
 
-We found a total of 159 instantiations of the argument schemes AS0-AS11 in the transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', but each argument scheme has been found back in the transcripts at least twice (Table~\ref{table:transcripts:results:argumentschemes}).
-A large portion (about 60\%) of the argument schemes we found involved discussions around tasks of the information system (AS2, AS10).
+We found a total of 159 instantiation of argument schemes AS0-AS11 in transcripts.
+The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice (Table~\ref{table:transcripts:results:argumentschemes}).
+A large portion (about 60\%) of argument schemes involved discussions about tasks of the information system (AS2, AS10).
 
-We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing the relevance of it (CQ12, CQ13).
+We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, CQ13).
 
@@
 
-For each transcript, we manually created a GRL model from the argument schemes and critical questions we found in them, in order to verify whether the arguments put forward by the participants were sufficiently informative.
+We manually created GRL models from argument schemes and critical questions found in each transcript, to verify whether the arguments put forward by the participants were sufficiently informative.
 An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
-We added green and red dots to various elements and relationships in the figure.
-A green dot indicates there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
-Note that if the underlying argument is rejected, the corresponding GRL element has been disabled.
+%TODO Again the text and the figure don't match.
+Depending on whether the underlying argument for an element is accepted or rejected, we added a green or a red dot to the elemenet. 
+%We added green and red dots to various elements and relationships in the figure. 
+A green dot indicates that there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
+Note that if the underlying argument is rejected, the corresponding GRL element is disabled.
 Some elements do not have a corresponding green or red dot.
-In that case, we have inferred the elements from the discussion, but we could not explicitly find back arguments for it.
+In that case, we have inferred the elements from the discussion, but we could not explicitly find an argument for them.
 
-We found that answering a critical questions can have four different effects on the original argument and the corresponding GRL element:
+Answering a critical question can have four different effects on the original argument and the corresponding GRL element:
 \begin{itemize} 
@@
 For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
+%TODO In figure it is XOR
 Suppose now the critical question CQ5b: ``Does the goal \texttt{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \texttt{Choose randomly}''.
+%TODO Where can we find this goal and its task?
 This results in a new instantiation of AS5, namely: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.
 As a result, the goal model will contain the corresponding task \texttt{Choose randomly}, as well as an OR-decomposition relation from the goal \texttt{Generate cars} to that task.
-\item \textsf{DISABLE:} Disable the element or relationship of the argument scheme to which the critical questions pertains.
-This operation does not create a new argument, but only disables (i.e., attacks) the original one.
+%TODO ?????
+\item \textsf{DISABLE:} Disable the element or relationship of argument scheme to which critical questions pertains.
+This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
 For instance, suppose argument scheme AS0 is instantiated with: ``Actor \texttt{Teacher} is relevant''.
-This argument can be attack with critical question CQ0: ``Actor \texttt{Teacher} is not relevant''.
-As a result, the argument for the actor is attack, and actor \texttt{Teacher} is disabled in the goal model.
+This argument can be attacked with critical question CQ0: ``Actor \texttt{Teacher} is not relevant''.
+As a result, the argument related to the actor is attacked as well, and actor \texttt{Teacher} is disabled in the goal model.
 \item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
 This operation both introduces a new argument and attacks the original one.
-For instance, suppose argument scheme AS2 is instantiated with: ``Actor \texttt{Student} can perform task \texttt{Choose a pattern preference}.
+For instance, suppose argument scheme AS2 is instantiated with: ``Actor \texttt{Student}" can perform task \texttt{Choose a pattern preference}.
 This argument can be attacked with critical question CQ13: ``The task \texttt{Choose a pattern preference} is unclear, it should be \texttt{Choose a road pattern}''.
-This results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}.
+This analysis results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}.
 In the goal model, the description of the task should change accordingly.
 \item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-For instance, suppose argument scheme AS0 is instantiated with ``Actor \texttt{Teacher} is relevant''.
-Suppose this argument is attacked with critical question CQ0: ``Actor \texttt{Teacher} is not relevant'', because she does not create the system.
-However, this critical question may in turn be attacked again, if new evidence comes up.
+Suppose argument scheme AS0 is instantiated with ``Actor \texttt{Teacher} is relevant'' and it is attacked by critical question CQ0: ``Actor \texttt{Teacher} is not relevant'', since the teacher does not create the system.
+However,  if a new evidence comes up, this critical question may in return be attacked again.
 For instance, it may turn out that the teacher will work on developing the application herself.
-In this case, the generic counter argument ``Actor \texttt{Teacher} is relevant, because does develop for the application'' attack the argument ``Actor \texttt{Teacher} is not relevant''.
-As a result, the original argument ``Actor \texttt{Teacher} is relevant'' is accepted again, and as a result is shown in the goal model.
+In this case, the generic counter argument ``Actor \texttt{Teacher} is relevant, because he/she does the development of the application'' attack the argument ``Actor \texttt{Teacher} is not relevant''.Thus, the original argument ``Actor \texttt{Teacher} is relevant'' is accepted again, and it is shown in the goal model.
 \end{itemize}
 
-In Section~\ref{sect:gmas:examples} we provide examples we found in the transcripts for all of these four effects.
+In Section~\ref{sect:examples} we provide examples for all of these four effects.
 
@@
 
-\paragraph{Analysis of the Argument Schemes}
+\paragraph{Analysis of Argument Schemes}
 Recall that our initial list of argument schemes consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}).
-Therefore, the difference between the initial list of argument schemes and those found back in the transcripts is quite small.
+Therefore, the difference between the initial list of argument schemes and those found in transcripts is quite small.
+%TODO The next few sentences are weird... they make our work very trivial.. rewrite.
 We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system.
@@
 GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 11.
-Our analysis however shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
-However, if needed, additional argument schemes and critical questions can be added, and our list is not meant to be exhaustive.
+%TODO which analysis? why aren't they used? How did we empirically validate it? How are we confident in the result?
+Our analysis, however, shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
+However, additional argument schemes and critical questions can be easily added to our framework.
+Beside, our list is not meant to be exhaustive.
 
-\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found back in the transcripts is much larger than for the critical questions.
-%SG: I don't get this sentence.
+\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found in transcripts is much larger than for the critical questions.
+%TODO  I don't get this sentence.
 Also, the ones below.
-We found few of the critical questions we initially proposed.
+We found few of the critical questions we initially proposed. %TODO ???
 However, this does not mean that they were not implicitly used in the minds of the participants.
-If a participants for instance forms an argument for a contribution from a task to a softgoal, it may very well be that she was asking herself the question ``Does the task contribute to some other softgoal?''.
+If a participant, for instance, forms an argument for a contribution from a task to a softgoal, it may very well be that she/he was asking her/him-self the question ``Does the task contributes to some other softgoals?''.
 However, many of these critical questions are not mentioned explicitly.
 If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of tool support, we believe that having these critical questions available may stimulate discussions.
+In the context of a tool support, we believe that having these critical questions available may stimulate discussions.
+%TODO what does it mean in the context of a tool support?

### 2017-08-12 examples.tex
fixed many things

@@
 We now discuss various instantiations of argument schemes and the result of answering critical questions in more detail.
-For each example we provide transcript excerpts, a visualization of the arguments, and the corresponding goal model elements.
+For each example, we provide transcript excerpts, a visualization of arguments, and the corresponding goal model elements.
 We provide a legend for our visualization notation in Figure~\ref{fig:legend}.
@@
 \end{tikzpicture}
-\caption{Legend of the various elements and relationships we use for the examples in this article.}
+\caption{Legends of Various Elements and Relationships}
 \label{fig:legend}
@@
 In this section, we present a small example for the RationalGRL methodology.
-This example, which is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
-We base the argument schemes and the critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
-These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-We provide several more examples in Section~\ref{sect:examples}.
+This example is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
+We base argument schemes and critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
+%These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+%TODO again?
+%We provide several more examples in Section~\ref{sect:examples}.
 
-As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders is developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology.
+As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders are developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology.
 This proceeds in the following way:
@@
 with the initial GRL model (Figure~\ref{fig:example-small}) and they invoke the intentional element \texttt{Simulate Traffic}. 
-%SG: We say above that we have an initial GRL model, so how do we match it with this sentence?
-I tried to reword it but it doesn't make so much sense...
 \item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
-They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation}.
+They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation"}.
 \item As a result, two argument schemes are created, namely:
@@
 The resulting RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
+%TODO You need to update goal "simulate" in this figure to "Simulate Traffic".
+Also, I don't think we should really call it the rational GRL model.
+it is an excerpts of the model.
 
@@
 The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$.
-In this example, participants are summing up functionality of the traffic simulator, which are tasks that the student can perform in the simulator.
-All these task can be formalized and are instantiations of argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
+In this example, participants sum up functionality of the traffic simulator, which are tasks that the students can perform in the simulator.
+All these tasks can be formalized and are instantiated from  argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
 
@@
 We formalized this using the critical question CQ12: ``Is task \texttt{Add traffic light} useful/relevant?''.
+%TODO I think it is better to say "necessary" and "not necessary" --> change figure 8: useless --> not necessary
 
-We visualize some of the argument schemes, critical questions, and traceability links with the GRL model in Figure~\ref{fig:examples:traffic-light}.
+We visualize some of the argument schemes, critical questions, and traceability links with GRL model in Figure~\ref{fig:examples:traffic-light}.
 On the left side of the image, we see three of the instantiated argument schemes AS2.
-The bottom one, ``Actor \texttt{Student} has task \texttt{Add traffic light}'', is attacked by another argument generated from applying critical question CQ12: ``\texttt{Add traffic light} is useless (\emph{All intersections have traffic lights}).
-As a result, the corresponding GRL task is disabled.
-The other two tasks are enabled and have green traceability links.
+The bottom one, ``Actor \texttt{Student} has task \texttt{Add traffic light}'', is attacked by another argument generated from applying critical question CQ12: ``\texttt{Add traffic light} is %useless
+not necessary (\emph{All intersections have traffic lights}). As a result, the corresponding GRL task is disabled. The other two tasks are enabled and have green traceability links.
 
@@
 \end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted lines) for the traffic light example.}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted lines) for the Traffic Light Example.}
 \label{fig:examples:traffic-light}
@@
 It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.
+%TODO where is this task in the model?
 
@@
 \end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the road pattern example.}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted line) for the Road Pattern Example.}
 \label{fig:examples:clarification}
@@
 The discussion starts with the first instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Create road}''.
+%TODO I don't see it in Figure 6.
 This argument is then challenged with critical question CQ12: ``Is the task \texttt{Create road} clear?''.
 Answering this question results in a new instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Choose a pattern}''.
+%TODO same
 This process is repeated two more times, resulting in the final argument ``Actor \texttt{Student} has task \texttt{Choose a road pattern}''.
-This final argument is unattacked and has a corresponding intentional element (right image).
+This final argument is "un-attacked" and has a corresponding intentional element (right image).
 
@@
 For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all three other arguments for a name of the element.
-If this was not the case, then it may occur that a previous argument is \emph{reinstatiated}, meaning that it becomes accepted again because the argument attacking it is itself attacked.
-Suppose for instance the bottom argument ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' did not attack the second argument: ``Action \texttt{Student} has task \texttt{Choose a pattern}''.
-In that case, this argument would be reinstated, because its only attacker ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' is itself defeated by the bottom argument.
+If this is not the case, then it may occur that a previous argument is \emph{re-instantiated}, meaning that it becomes accepted again because the argument attacking it, is itself attacked.
+Suppose for instance that the bottom argument ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' did not attack the second argument: ``Action \texttt{Student} has task \texttt{Choose a pattern}''.
+In that case, this argument would be re-instantiated, because its only attacker ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' is itself defeated by the bottom argument.
+%TODO check these tasks with the GRL model of Figure 6.
+Also, I recommend to have the changed full  GRL model after we finish the examples to show how the model improves.
 
-\subsubsection{Example 4: Decompose goal \texttt{Simulate}}
+\subsubsection{Example 4: Decompose goal \texttt{Simulate}} %TODO In fig 1, the goal is called simulate traffic which makes more sense.
+goal "simulate" is a bit meaningless.
+I suggest to update fig 6 and all the other ones accordingly.
+Also, we are refering to this example in example 1.
+Why can't we merge them?
 
@@
 \end{adjustwidth}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of the example.}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted line) of the example.} %TODO which example? 
 \label{fig:examples:decomposition}
@@
 Moreover, the original argument for an AND-decomposition is attacked by the argument for the OR-decomposition, and the new argument is linked to the decomposition relation in the GRL model.
+%TODO I think we can merge this easily with example 1.
 
@@
 It consists of two parts: first participant \texttt{P1} puts forth the suggestion to include actor \texttt{Development team} in the model.
-This is, then, questioned by participant \texttt{P2}, who argues that the professor will develop the software, so there won't be any development team.
+This is, then, questioned by participant \texttt{P2}, who argues that the professor will develop the software, so there will not be any development team.
 However, in the second part, participant \texttt{P2} argue that the development team should be considered, since the professor does not develop the software.
@@
 \end{tikzpicture}
-\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted line) of a Discussion about the Relevance of Actor Development Team.}
 \label{fig:examples:relevant-actor2}
@@
 Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counter argument has been put forward.
-The argument ``Actor \texttt{Professor} doesn't develop the software'' now attacks the argument ``\texttt{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\texttt{Development team} is relevant''.
-As a result, the first argument is reinstated, which causes the actor in the GRL model to be enabled again.
+The argument ``Actor \texttt{Professor} does not develop the software'' now attacks the argument ``\texttt{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\texttt{Development team} is relevant''.
+As a result, the first argument is re-instantiated, which causes the actor in the GRL model to be enabled again.

### 2017-08-15 examples.tex
Updated section 4+5+6

@@
 For each example, we provide transcript excerpts, a visualization of arguments, and the corresponding goal model elements.
+Figure~\ref{fig:transcripts:grl} in the previous section shows an example of how a RationalGRL may look like when it is developed in our prototype tool.
+However, the argument schemes that are instantiated are left implicit in this figure (in the tool, they can be seen in the details pane, see Section~\ref{sect:tool}).
+Since we would like to make these instantiated argument schemes explicit in the examples below, we use a different visual language here.
 We provide a legend for our visualization notation in Figure~\ref{fig:legend}.
@@
 
-The resulting RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
+An excerpt of the RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
 %TODO You need to update goal "simulate" in this figure to "Simulate Traffic".
@@
 
-Once all these tasks are summed up, participant \texttt{P1} notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \texttt{Add traffic light} is not useful.
+Once all these tasks are summed up, participant \texttt{P1} notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \texttt{Add traffic light} is not necessary.
 We formalized this using the critical question CQ12: ``Is task \texttt{Add traffic light} useful/relevant?''.
@@
 The bottom one, ``Actor \texttt{Student} has task \texttt{Add traffic light}'', is attacked by another argument generated from applying critical question CQ12: ``\texttt{Add traffic light} is %useless
-not necessary (\emph{All intersections have traffic lights}). As a result, the corresponding GRL task is disabled. The other two tasks are enabled and have green traceability links.
+not necessary (\emph{All intersections have traffic lights})''. As a result, the corresponding GRL task is disabled. The other two tasks are enabled and have green traceability links.
 
@@
         \node (a3) [argNodeIN] at (-2,-6){
-        	\argtext{}{\emph{Add traffic light} is useless (\emph{All intersections have traffic lights})}
+        	\argtext{}{\emph{Add traffic light} is not necessary (\emph{All intersections have traffic lights})}
         };
@@
 %TODO where is this task in the model?
+%MvZ: It is in Figure 9, but not in Figure 6.
+Figure 6 is just a single model from our transcripts, but Figure 9 comes from another transcript.
+Do you think it is better to take all of the examples from the same transcript?
+I feel that may look a bit strange since we are completely disregarding the other transcripts then.
 
@@
 %TODO I don't see it in Figure 6.
+%MvZ: see previous comment, Figure 9 is not part of Figure 6
 This argument is then challenged with critical question CQ12: ``Is the task \texttt{Create road} clear?''.
 Answering this question results in a new instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Choose a pattern}''.
-%TODO same
+%TODO same %MvZ same
 This process is repeated two more times, resulting in the final argument ``Actor \texttt{Student} has task \texttt{Choose a road pattern}''.
@@
 with \emph{No}.
-This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.
-This is shown in Figure~\ref{fig:examples:relevant-actor}, left image.
+This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.'' This is shown in Figure~\ref{fig:examples:relevant-actor}, left image.
 

### 2017-08-15 formalframework.tex
Updated section 4+5+6

@@
 
-In the previous section we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of a traffic simulator.
+In Section~\ref{sect:gmas} we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of a traffic simulator.
 The resulting list is shown in Table~\ref{table:argument-schemes}.
@@
 
-The examples and corresponding visualizations of the previous section provide some insight on how to formalize argument schemes, critical questions, and their relationship with goal models.
-However, if we are to implement our framework in a tool, we require a more precise formalization of these concepts.
-We choose to use formal logic to specify this.
-This is also a good fit with formal argumentation, which has flourished in the past decades, leading to a large number of different semantics that we can choose from.
+The examples and corresponding visualizations of the previous section provide some insight on how our RationalGRL framework can be used.
+However, since we would like to implement our framework in a tool, we require a more precise formalization.
+In this section, we present a formalization of our framework using formal logic.
+This is the main approach using in formal argumentation, so it allows us to use many of the techniques developed in that area directly (see Section~\ref{sect:goalmodeling:relatedwork} for more details).
 
-The rest of this section is as follows: In the first subsection, we develop a formal language to specify a GRL model.
-This language consists of atomic sentences.
-In the second subsection, we develop our notion of an argument: an argument is simply a set of atoms from our language.
-In the third subsection, we develop algorithms for all the argument schemes and critical questions.
+In the first two subsections we formalize a static representation of our framework: We develop a formal language to specify a GRL model (independently of arguments) in the fist subsection, and we formalize the notion of an argument in the second subsection.
+In the third subsection we formalize the dynamics of our framwork: we develop algorithms for instantiating argument schemes and for answering critical questions.
+The formal framework we develop in this section serves as the underlying model for the prototype implementation we present in the next section.
 
@@
 
-We define our language as a set of atoms, that is, grounded formulas (without variables) with no logical connectives or negation.
+We assume familiarity with basic notions from propositonal logic and set theory.
+A \emph{propositional atom} is a grounded formula (a formula without variables) with no logical connectives ($\vee, \and, \rightarrow, \leftrightarrow$) or negation ($\not$).
+We use integers $i,j\in\mathbb{N}$ as identifiers for actors and GRL elements.
 
 \begin{definition}[GRL Atoms]
-Let $P$ be a set of sentences representing names for GRL elements. The \emph{set of GRL atoms} $At$ is defined as follows: 
-\begin{align*}
-At = \{&actor(i), softgoal(i), goal(i), task(i),\\
-&resource(i), decomp(k,i,J,dtype),\\
-&dep(k,i,j),contr(k,i,j,ctype),\\
-&has(i,j),disabled(i),name(i,p)\},
-\end{align*}
-where
-\begin{itemize}
-\item $\{i,j,k\}\cup J \subseteq\mathbb{N}$
-\item $p\in P$
-\item $dtype\in\{and,or,xor\}$
-\item $ctype\in\{+,-\}$
-\end{itemize}
-\end{definition}
-
-The non-negative integers associates with each element and relation are identifiers.
-We thus choose to identify each element and relationship in a GRL model with an identifier.
-This allows us to change the name of an element while still being able to make different statements about this element, for instance by applying critical question CQ12a (clarification, see the example ``Clarify task \texttt{Road pattern}'' above).
-
-We now briefly explain each atom in more detail.
+The \emph{set of GRL atoms} $At$ contains the following atoms:
 \begin{itemize}
@@
 
-\begin{remark}
-From these descriptions one can observe that many of the atoms come with a number of assumptions and dependencies.
-For instance, if $has(i,j)$ is true, then $i$ is an actor and $j$ is an element.
-We could formalize this as follows: $$has(i,j)\rightarrow (actor(i) \wedge (softgoal(j)\vee goal(j)\vee task(i) \vee resource(j))).$$ It would be possible to enumerate all such properties in order to correctly specify a GRL model.
-One could then formally verify whether a set of atoms violates these constraints.
-If not, it is a ``valid'' representation of a GRL model.
-Since the focus of this article is not on a logical analysis, but rather on developing a framework for empirical data, we leave such a formal analysis for future work (see Section~\ref{sect:goalmodeling:openissues}).
-\end{remark}
+For all of the above definitions we assume that all identifiers used in each atom are different, i.e.,
+\begin{itemize}
+\item $k\not=i$,$k\not=i$,$k\not=j$
+\item forall $j\in J: k\not=j, i\not=j$
+\end{itemize}
+\end{definition}
 
-Using this formalization, it is rather straightforward to specify a GRL model.
 An example of the specification of the GRL model in Figure~\ref{fig:examples:decomposition} is shown in Table~\ref{table:grl_atom_spec}.
@@
 
-In the previous subsection we introduced an atomic language to specify a GRL model.
-In this subsection we give a formal definition of an argument, which is simply a set of atoms from our language.
-We introduce Dung's acceptability semantics as well, which allows us to determine sets of acceptable arguments.
+In the previous subsection we introduced an atomic language to specify a GRL model. In this subsection we give a formal definition of an argument, which is simply a set of atoms from our language.
 
-\begin{definition}[Argument]
+\begin{definition}[Argument] \label{def:argument}
 An \emph{argument} $A\subseteq Att$ is a set of atoms from $Att$.
@@
 This simple definition allows us to form arguments about (parts of) a GRL model.
-For instance, $$\{goal(0),name(0,development\_team)\},$$ is an argument.
-We next introduce an argumentation framework, which is a set of arguments an attack relations between arguments.
+For instance, $$\{goal(0),name(0,development\_team)\},$$ is an argument stating that identifier 0 is a goal element with name ``development team''.
 
-\begin{definition}[Argumentation framework~\cite{Dung1995}]
+Note that this definition does not put any restrictions on the content of an argument.
+This means it is possible to form arguments that do not correspond to correct or sensible GRL models.
+Consider for instance the following argument: $$\{goal(0), task(0), name(0,static\_simulation)\}.$$ This argument states that identifier 0 is both a goal and a task with name ``statis simulation'', which is clearly undesirable since it does not correspond to a valid GRL model.
+While we could have chosen to put logical constraints on the content of an argument, we choose not to do so here.
+Instead, we put these restrictions on the algorithms in section~\ref{sect:algorithms}.
+The algorithms are constructed in such a way that they only allow the construction of argumetns that correspond to valid GRL models.\footnote{Note that we do not provide formal proofs for these claims in this paper.
+This is an interesting direction of future work, but it is not the focus of this article.} The advantage of this approach is that it keeps the presentation of the formal part relatively simple.
+
+In line with the main approaches in formal argumentation, we next introduce an argumentation framework as a set of arguments and an attack relation between the arguments.
+
+\begin{definition}[Argumentation framework~\cite{Dung1995}] \label{def:argumentationframework}
 An argumentation framework $AF=(Args,Att)$ consists of a set of arguments $Args$ and an attack relationship $Att:Args\times Args$, where $(A_1,A_2)\in Att$ means that argument $A_1\in Args$ attacks arguments $A_2\in Args$.
@@
 
-Now that we have defined arguments and their attacks, we are going to define a semantics to determine which arguments are acceptable. The following notions are preliminary to this.
+In order to determine which arguments are accepted or not, we define \emph{argumentation semantics}.
+We used the standard approach here, which is known as \emph{Dung's semantics}.
+The following notions are preliminary.
 
@@
 In future work, we aim to extend this by using preferences (see open issues in Section~\ref{sect:goalmodeling:openissues}).
-An advantage of our current approach is that all semantics coincide when there are no cycles, which simplifies our case.
+An advantage of our current approach is that all the standard semantics coincide when there are no cycles, which simplifies our case.
 We use the preferred semantics here, but we could equivalently have chosen any other version.
 
-\begin{definition}[Preferred semantics~\cite{Dung1995}] 
+\begin{definition}[Preferred semantics~\cite{Dung1995}] \label{def:semantics}
 Suppose an argumentation framework $AF=(Arg,Att)$. A set of arguments $S\subseteq Arg$ is a \emph{preferred extension} if and only if 
@@
 \subsection{Algorithms for argument schemes and critical questions}
+\label[sect:algorithms]
 
@@
 
-In this subsection we develop algorithms for applying argument schemes and critical questions. In the following algorithms, we assume the following global variables:
+In this subsection we develop algorithms for applying argument schemes and critical questions.
+The process of applying these algorithms and generating goal models from them is shown in context of the RationalGRL methodology in Figure~\ref{fig:methodologyalgorithm} \todo{Marc}{Marc}{Create an extended version of the methodology, containing where which parts of the formal framework are used.} These algorithms produce arguments (Definition~\ref{def:argument}) and attack relations, forming an argumentation framework (Definition~\ref{def:argumentationframework}).
+One can then use argumentation semantics (Definition~\ref{def:semantics}) to compute sets of accepted arguments.
+The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.
+
+In the following algorithms, we assume the following global variables:
 
@@
 \subsubsection*{Algorithms for argument schemes}
-The algorithms for the argument schemes are very simple, because they simply consist of forming a new argument and adding it to the set of arguments. No attack relations are introduced.
+The algorithms for the argument schemes consist of forming a new argument and adding it to the set of arguments. No attack relations are introduced.
 
@@
 On line~\ref{alg:as5:1}, the set $T_{id}$ is initialized, which will contain the ids of the tasks $T_1,\ldots,T_n$ to decompose into.
-In the for loop, the if statement on line~\ref{alg:as5:2} checks whether some argument already exists for the current task $T_i$, and if so, it simply adds the identifier of the task ($t_{id}$) to the set of task identifiers $T_{id}$.
-Otherwise (line~\ref{alg:as5:3}), a new task is created and the new identifier $id$ is added to the set of task identifiers.
+In the for-loop, the if-statement on line~\ref{alg:as5:2} checks whether some argument already exists for the current task $T_i$, and if so, it simply adds the identifier of the task ($t_{id}$) to the set of task identifiers $T_{id}$.\footnote{Note that the existing task may be disabled, but this does not matter, since the decomposition relation will be suppressed for disabled elements.} Otherwise (line~\ref{alg:as5:3}), a new task is created and the new identifier $id$ is added to the set of task identifiers.
 After the for loop on line~\ref{alg:as5:4}, an argument for the decomposition link itself is constructed, and it is added to the set of arguments $Args$.
@@
     \Procedure{$AS_6$}{$t_{id}, s$}
-    \If{$\exists_{A\in Args} \{softgoal(i),name(i,s)\}\subseteq A$}
+    \If{$\exists_{A\in Args} \{softgoal(i),name(i,s)\}\subseteq A$} \label{alg:as6:if}
         \State $s_{id} \gets i$
@@
 \emph{Algorithm~\ref{alg:as6} for argument scheme AS6:} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
-The idea behind this algorithm is very similar to the previous one.
-First, the if statements check whether the softgoal exists already, and if not, an argument is added for it.
+The idea behind this algorithm is very similar to the previous one, with the difference that in the current algorithm we create a single relation, while we created a set of relations in the previous algorithm.
+First, the if-statement on line~\ref{alg:as6:if} checks whether the softgoal exists already, and if not, an argument is added for it.
 This ensures that all softgoals have corresponding arguments.
-After the if statement, the argument for the contribution link is created and it is added to the set of arguments $Args$.
+After the if-statement, the argument for the contribution link is created and it is added to the set of arguments $Args$.
 
@@
 
-\emph{Algorithm~\ref{alg:disable} (\textsf{DISABLE}) for critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12:} The disable operation is very straightforward: It simply consists of adding an argument stating the GRL element with identifier $i$ is disabled.
+\emph{Algorithm~\ref{alg:disable} (\textsf{DISABLE}) for critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12:} The disable operation consists of adding an argument stating the GRL element with identifier $i$ is disabled.
 Let us reconsider the example of Figure~\ref{fig:examples:relevant-actor}.
 This example consists of an instantiation of argument scheme AS0, which is attacked by an argument that resulted from answering critical question.
-The formalization of this scenario is shown in Figure~\ref{fig:examples:relevant-actor:formal}.
-Interestingly, we see that the \textsf{DISABLE} operation no longer leads to an attack between the two arguments.
-Instead, the only thing that this operation does is adding another argument stating the element is disabled.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}
-        \node (a0) [argNodeIN] at (-3.7,-0.2) {
-        	\smallargtext{AS0}{$\{actor(0),name(0,dev\_team\}$}
-        } ;
-        \node (a1) [argNodeIN] at (-3.7,-2.5){
-        	\argtext{DISABLE}{$\{disabled(0)\}$}
-        };
-         \path
-    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1);
-\end{tikzpicture}
-\caption{Formalization of the arguments in Figure~\ref{fig:examples:relevant-actor}.}
-\label{fig:examples:relevant-actor:formal}
-\end{figure}
+The instantiation of AS0 leads to the argument $A=\{actor(0),name(0,dev\_team\}$.
+After applying $DISABLE(0)$ we obtain the arguments: $$Args=\{\{actor(0),name(0,dev\_team\}, \{disabled(0)\}\}.$$ Note that our implementation of this attack does not lead to an actual attack in the argumentation framework.
 
@@
 
-The algorithm task as input the goal identifier $g_{id}$, the set of existing decomposing task identifiers $i_1,\ldots,i_n$, the decomposition type, and the names of the new tasks $t_1,\ldots,t_k$ that should be added to the decomposition.
-The first part  of the algorithm is already familiar: For each task name we check whether it already exists as an argument (line 4), and if it doesn't (line 6) we add a new argument for it.
-After the for loop (line 13), a new argument is created for the new decomposition relation (14).
-After this, the for loop on line 15 ensures that the new argument attacks all previous arguments for this decomposition (note that the variable ``$\_$'' means ``do not care'').
+The algorithm takes as input the goal identifier $g_{id}$, the set of existing decomposing task identifiers $i_1,\ldots,i_n$, the decomposition type, and the names of the new tasks $t_1,\ldots,t_k$ that should be added to the decomposition.
+The first part of the algorithm is familiar from Algorithm~\ref{alg:as5}: For each task name we check whether it already exists as an argument (line 4), and if it doesn't (line 6) we add a new argument for it.
+After the for-loop (line 13), a new argument is created for the new decomposition relation (14).
+After this, the for-loop on line 15 ensures that the new argument attacks all previous arguments for this decomposition (note that the variable ``$\_$'' means ``do not care'').
 Only at the very end the new argument is added (line 18), to ensure it does not attack itself after the for loop of line 15-17.
@@
     \State $Args \gets Args \cup \{A\}$
-    \For{$B$ in $ArgsN$}
-      \State $Att\gets Att \cup \{(A,B)\}$
+    \For{$C$ in $ArgsN$}
+      \State $Att\gets Att \cup \{(A,C)\}$
     \EndFor
@@
 It takes two parameters: the element identifier $i$ and the new name $n$.
-The idea behind the algorithm is that we construct a new argument for $n$ from the previous arguments, and we only replace the $name$ atom.
-We also have to ensure that we attack all previous arguments for a name.
+The idea behind the algorithm is that we construct a new argument for $n$, and to ensure that this argument attacks all previous arguments for giving a name to this element.
+Here we can use the following fact: Suppose $Args$ is a set of arguments, each containing an atom about the name for an element $i$, i.e.
+for all $A\in Args: name(i,_)\in Args$, then all arguments in $Args$ are the same except for the $name$ atoms, i.e.
+for all $A,B\in Args: A\backslash\{name(i,_)\} = B\backslash\{name(i,_)\}$.
+This means that if we would like to attack every argument of $Args$ with a new argument that replaces the $name$ atom, we can simply take any argument in $Args$, remove the previous $name$ atom, add the new one and attack all arguments in $Args$.
+This is exactly what the algorithm does.
+
 On line 2, all arguments that have been put forward for this element and contain $name(i,x)$ are collected into the set $ArgsN$.
-On line 3, some arguments $B'\in ArgsN$ minus the $name$ statement is assigned to $B$, and on line 4 $B$ is joined with the new $name$ statement and stored in $A$, which is then added to the set of arguments $Args$.
-The for loop on lines 6-8 ensures all previous arguments for names of the element are attacked by the new argument.
+On line 3, some arguments $B'\in ArgsN$ minus the $name$ statement is assigned to $B$ (note that it does not matter which one we pick), and on line 4 $B$ is joined with the new $name$ statement and stored in $A$, which is then added to the set of arguments $Args$.
+The for-loop on lines 6-8 ensures all previous arguments for names of the element are attacked by the new argument.
 
@@
 
-An example of the working of Algorithm~\ref{alg:cq13} is shown in Figure~\ref{fig:examples:clarification:formal}.
+An example of of Algorithm~\ref{alg:cq13} is shown in Figure~\ref{fig:examples:clarification:formal}.
 Let us consider the last application of CQ13 (bottom argument).

### 2017-08-15 gmas.tex
modified a small typo
Updated section 4+5+6

@@
 
-In previous section, we provided an overview of RationalGRL methodology.
+In the previous section, we provided an overview of RationalGRL methodology.
 In this section, we discuss our argument schemes and critical questions in more detail.
 It is important to note that the list of argument schemes we present in this section is not exhaustive.
-It is an initial list that we have obtained by annotating transcripts, however, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
+It is an initial list that we have obtained by annotating transcripts, but our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
 
@@
 
-In the following subsections, we first provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts) and then, we analyze our results.
+In the following subsections, we first provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts) and then we analyze our results.
 
@@
 
-\paragraph{Process of Extracting Argument Schemes and Critical Questions} %TODO I changed the title from annotation method.
+\paragraph{Process of Extracting Argument Schemes and Critical Questions} %TODO I changed the title from annotation method. %MvZ: ok, looks good.
 
@@
 If there were arguments or critical questions that did not appear in the initial list, we added them to the list.
-Existing argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}).
-%TODO I don't understand it!
-why?
+% Existing argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}). 
+%TODO I don't understand it! why?
+%MvZ: I agree it is confusing.
+The idea is that critical questions are often implicit, so we kept them in since we believe it may still stimulate discussions.
+I now simply left it out since it doesn't add much and will only be confusing.
 
@@
 
-It is generally known in the argumentation literature that it can be very difficult to identify arguments in natural language texts~\cite{walton-etal2004}.
-Arguments are often imprecise, lack conclusion, and may be supported by non verbal communication that is not captured in transcripts.
+%It is generally known in the argumentation literature that it can be very difficult to identify arguments in natural language texts~\cite{walton-etal2004}.
+Arguments are often imprecise, lack conclusion, and may be supported by nonverbal communication that is not captured in transcripts.
 However, there is hardly any research on argument extraction in the requirement engineering domain.
 Thus, despite this potential weakness in our approach, we believe it nevertheless is at least useful as something that others can build further on (see Section~\ref{sect:goalmodeling:openissues}).
-%TODO I don't understand the last few sentences...
-what does argument extraction mean?
-why is it needed?
-don't use "something that others" it is very informal for a journal paper.
+%TODO I don't understand the last few sentences... what does argument extraction mean? why is it needed? don't use "something that others" it is very informal for a journal paper. 
+%MvZ: I was also sceptical about this paragraph when I wrote it. I think it belongs in the related work section where we can discuss it in more detail. I now removed it here.
 
-\paragraph{Results}
+\subection{Experiment Results}
 
@@
 %TODO Again the text and the figure don't match.
-Depending on whether the underlying argument for an element is accepted or rejected, we added a green or a red dot to the elemenet. 
-%We added green and red dots to various elements and relationships in the figure. 
+%MvZ: What doesn't match? We only say here we give an example in Figure 6, so there is not problem, right?
+Depending on whether the underlying argument for an element is accepted or rejected, we added a green or a red dot to the element.
 A green dot indicates that there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
@@
 
+%MvZ: The examples below were not meant to correspond to anything, I just wrote them as ficitious examples. Looking at your remarks I think this is confusing.
 Answering a critical question can have four different effects on the original argument and the corresponding GRL element:
@@
 This operation does not attack the original argument of the critical question, but rather creates a new argument.
-For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
-%TODO In figure it is XOR
-Suppose now the critical question CQ5b: ``Does the goal \texttt{Agreeable meeting dates} decompose into other tasks?'' is answered with ``yes, namely \texttt{Choose randomly}''.
-%TODO Where can we find this goal and its task?
-This results in a new instantiation of AS5, namely: ``Goal \texttt{Generate cars} OR-decomposes into tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.
-As a result, the goal model will contain the corresponding task \texttt{Choose randomly}, as well as an OR-decomposition relation from the goal \texttt{Generate cars} to that task.
+For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}'' (see Figure~\ref{fig:transcripts:grl}).
+%TODO In figure it is XOR %MvZ: Fixed
+Suppose now the critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'' is answered with ``Yes, namely \texttt{Choose randomly}''.
+This results in a new instantiation of AS5, namely: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' As a result, the goal model will contain the corresponding task \texttt{Choose randomly}, as well as an OR-decomposition relation from the goal \texttt{Generate cars} to that task.
 %TODO ?????
@@
 This argument can be attacked with critical question CQ13: ``The task \texttt{Choose a pattern preference} is unclear, it should be \texttt{Choose a road pattern}''.
-This analysis results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}.
-In the goal model, the description of the task should change accordingly.
+This analysis results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}.'' In the goal model, the description of the task should change accordingly.
 \item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
@@
 
-\subsection{Analysis}
-\label{sect:gmas:transcripts:analysis}
+%MvZ: I completely removed the subsection below. I think it doesn't add much and it is confusing. What do you think?
 
-\paragraph{Analysis of Argument Schemes}
-Recall that our initial list of argument schemes consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}).
+%\subsection{Analysis}
+%\label{sect:gmas:transcripts:analysis}
+
+%\paragraph{Analysis of Argument Schemes}
+%Recall that our initial list of argument schemes consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}).
 Therefore, the difference between the initial list of argument schemes and those found in transcripts is quite small.
 %TODO The next few sentences are weird... they make our work very trivial.. rewrite.
-We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system.
+%We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system.
 This gives us an indication that it is possible to capture (parts of the) arguments used in those type of discussions using argument schemes.
 
-We observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
+%We observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
 Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
@@
 %TODO which analysis? why aren't they used? How did we empirically validate it? How are we confident in the result?
-Our analysis, however, shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
+%Our analysis, however, shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
 However, additional argument schemes and critical questions can be easily added to our framework.
@@
 
-\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found in transcripts is much larger than for the critical questions.
+%\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found in transcripts is much larger than for the critical questions.
 %TODO  I don't get this sentence.
 Also, the ones below.
-We found few of the critical questions we initially proposed. %TODO ???
-However, this does not mean that they were not implicitly used in the minds of the participants.
+%We found few of the critical questions we initially proposed. %TODO ???
+%However, this does not mean that they were not implicitly used in the minds of the participants.
 If a participant, for instance, forms an argument for a contribution from a task to a softgoal, it may very well be that she/he was asking her/him-self the question ``Does the task contributes to some other softgoals?''.

### 2017-08-15 formalframework.tex
Some minor changes
Previous version didn’t compile

@@
 We assume familiarity with basic notions from propositonal logic and set theory.
-A \emph{propositional atom} is a grounded formula (a formula without variables) with no logical connectives ($\vee, \and, \rightarrow, \leftrightarrow$) or negation ($\not$).
-We use integers $i,j\in\mathbb{N}$ as identifiers for actors and GRL elements.
+A \emph{propositional atom} is a grounded formula (a formula without variables) with no logical connectives ($\vee, \wedge, \rightarrow, \leftrightarrow$) or negation ($\neg$).
+We use integers $i,j,k\in\mathbb{N}$ as identifiers for actors and GRL elements.
 

### 2017-08-16 gmas.tex
Some minor changes
Previous version didn’t compile
New figure and rewrote Section 4

@@
 
-In the following subsections, we first provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts) and then we analyze our results.
+The RationalGRL language is an extension of GRL.
+The prototype tool we developed therefore also contains various new elements and a new relationship.
+The legend of the these additions is shown in Figure~\ref{fig:rationalgrllegend}.
+
+In the following subsections, we first explain the new visual language in more detail, after which we provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts).
+
+\subsection{RationalGRL Language}
+
+RationalGRL is an extension of GRL, and adds the following new elements (Figure~\ref{fig:rationalgrllegend}):
+\begin{itemize}
+\item \emph{Argument}: This represents an argument created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
+\item \emph{Disabled GRL element}: If a GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be disabled, meaning that it does not play a role in the analysis of the GRL model.
+\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
+It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
+In these cases, the corresponding GRL element is shown with a striped background.
+If the user clicks on this element, a details pane shows up containing the history of refinements (see Section~\ref{sect:tool}).
+\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument attacks the target GRL element or argument.
+\end{itemize}
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.35\textwidth]{img/legend}
+\caption{The new elements and link of RationalGRL}
+\label{fig:rationalgrllegend}
+\end{figure}
 
@@
 
-\subection{Experiment Results}
+\subsection{Experiment Results}
 
-\begin{figure*}[h]
-\includegraphics[width=\textwidth]{img/transcript_grl}
+\begin{figure*}[h!]
+\includegraphics[width=\textwidth]{img/Fig6}
 \caption{The GRL model manually constructed from transcript $t_1$.
@@
 
-We manually created GRL models from argument schemes and critical questions found in each transcript, to verify whether the arguments put forward by the participants were sufficiently informative.
+We manually created RationalGRL models from argument schemes and critical questions found in each transcript, to verify whether the arguments put forward by the participants were sufficiently informative.
 An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
-%TODO Again the text and the figure don't match.
-%MvZ: What doesn't match? We only say here we give an example in Figure 6, so there is not problem, right?
-Depending on whether the underlying argument for an element is accepted or rejected, we added a green or a red dot to the element.
-A green dot indicates that there is an underlying argument for the element that is accepted, while a red dot indicates a rejected underlying argument.
-Note that if the underlying argument is rejected, the corresponding GRL element is disabled.
-Some elements do not have a corresponding green or red dot.
-In that case, we have inferred the elements from the discussion, but we could not explicitly find an argument for them.
+This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
+\todo{Marc}{all}{Perhaps we should explain a bit about the model here?}
 
-%MvZ: The examples below were not meant to correspond to anything, I just wrote them as ficitious examples. Looking at your remarks I think this is confusing.
 Answering a critical question can have four different effects on the original argument and the corresponding GRL element:
@@
 This operation does not attack the original argument of the critical question, but rather creates a new argument.
-For instance, suppose argument scheme AS5 is instantiated as follows:: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}'' (see Figure~\ref{fig:transcripts:grl}).
-%TODO In figure it is XOR %MvZ: Fixed
-Suppose now the critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'' is answered with ``Yes, namely \texttt{Choose randomly}''.
-This results in a new instantiation of AS5, namely: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' As a result, the goal model will contain the corresponding task \texttt{Choose randomly}, as well as an OR-decomposition relation from the goal \texttt{Generate cars} to that task.
-%TODO ?????
+In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
+For instance, the XOR-decomposition from ``Generate Cars'' is an instantiation of AS5 as follows: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
+Suppose now that the modelers that created Figure~\ref{fig:transcripts:grl} would continue their analysis by discussing critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'', and that they would answer this with ``Yes, namely \texttt{Choose randomly}''.
+This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.''
 \item \textsf{DISABLE:} Disable the element or relationship of argument scheme to which critical questions pertains.
 This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
-For instance, suppose argument scheme AS0 is instantiated with: ``Actor \texttt{Teacher} is relevant''.
-This argument can be attacked with critical question CQ0: ``Actor \texttt{Teacher} is not relevant''.
-As a result, the argument related to the actor is attacked as well, and actor \texttt{Teacher} is disabled in the goal model.
+In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
+The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
+What we also see from Figure~\ref{fig:transcripts:grl} is that disabling actor ``Teacher'' also disables all elements belonging to this actor.
+Furthermore, disabling task ``Dynamic simulation'' also disables all incoming and outgoing links with this task.
+
 \item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
-This operation both introduces a new argument and attacks the original one.
-For instance, suppose argument scheme AS2 is instantiated with: ``Actor \texttt{Student}" can perform task \texttt{Choose a pattern preference}.
-This argument can be attacked with critical question CQ13: ``The task \texttt{Choose a pattern preference} is unclear, it should be \texttt{Choose a road pattern}''.
-This analysis results in replacing the original argument with the new argument ``Actor \texttt{Student} can perform task \texttt{Choose a road pattern}.'' In the goal model, the description of the task should change accordingly.
+In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
+In this case, the participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
+While the previous names are not shown in the figure, they show up in the details pane of the corresponding element.
 \item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-Suppose argument scheme AS0 is instantiated with ``Actor \texttt{Teacher} is relevant'' and it is attacked by critical question CQ0: ``Actor \texttt{Teacher} is not relevant'', since the teacher does not create the system.
-However,  if a new evidence comes up, this critical question may in return be attacked again.
-For instance, it may turn out that the teacher will work on developing the application herself.
-In this case, the generic counter argument ``Actor \texttt{Teacher} is relevant, because he/she does the development of the application'' attack the argument ``Actor \texttt{Teacher} is not relevant''.Thus, the original argument ``Actor \texttt{Teacher} is relevant'' is accepted again, and it is shown in the goal model.
+In Figure~\ref{fig:transcripts:grl} we see one example of such a counter-attack.
+First, task ``Control car influx per road'' is attack by answering CQ2 (irrelevant task).
+However, after discussing this, the participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
+Therefore, the arguments that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
+We provide a precise semantics for this in Figure~\ref{sect:formalframework}.
 \end{itemize}
-
-In Section~\ref{sect:examples} we provide examples for all of these four effects.
 

### 2017-08-17 gmas.tex
Updated section 4 and gave comments

@@
 It is important to note that the list of argument schemes we present in this section is not exhaustive.
-It is an initial list that we have obtained by annotating transcripts, but our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
+It is an initial list that we have obtained by annotating transcripts.
+However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
 
@@
 The RationalGRL language is an extension of GRL.
-The prototype tool we developed therefore also contains various new elements and a new relationship.
-The legend of the these additions is shown in Figure~\ref{fig:rationalgrllegend}.
+The prototype tool we developed, thus, contains various new elements and a new relationship.
+The legends of the these additions are shown in Figure~\ref{fig:rationalgrllegend}.
 
@@
 \item \emph{Argument}: This represents an argument created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
+%SG: I thought arguments can also do replace, or intro GRL elements.
+So why in this description we are very limiting?
 \item \emph{Disabled GRL element}: If a GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be disabled, meaning that it does not play a role in the analysis of the GRL model.
@@
 If the user clicks on this element, a details pane shows up containing the history of refinements (see Section~\ref{sect:tool}).
+%SG: Can we have a separate element for the intro?
+or at least explain it in this text more explicitly?
 \item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument attacks the target GRL element or argument.
-\end{itemize}
+\end{itemize} %SG: what about the links? The links can also get attacked right? or have some arguments? what do we do with those?
 
@@
 
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from the University of Utrecht, following a Software Architecture course.
+\paragraph{Subjects} The subjects for the case study are three teams of Master students from Utrecht University, following the Software Architecture course.
 Two teams consist of three students, and one team consists of two students.
@@
 
-\paragraph{Process of Extracting Argument Schemes and Critical Questions} %TODO I changed the title from annotation method. %MvZ: ok, looks good.
+\paragraph{Process of Extracting Argument Schemes and Critical Questions}
 
@@
 % Existing argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}). 
-%TODO I don't understand it! why?
+% I don't understand it! why?
 %MvZ: I agree it is confusing.
@@
 
-Most of the occurrences were not literally found back, but had to be inferred from the context. This can be seen in the various examples we will discuss.
+Most of the occurrences were not literally found back, but had to be inferred from the context. This can be seen in the various examples we will discuss in Section~\ref{sect:examples}.
 
@@
 Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
-Elements and relationships with no dot have been inferred by us.}
+Elements and relationships with no dot have been inferred by us.} %TODO We need to update the caption
 \label{fig:transcripts:grl}
-\end{figure*} %TODO Figure doesn't match the text!
+\end{figure*}
 
@@
 \end{tabularx}
-\caption{Number of occurrences of AS0-AS9, CQ0-CQ12 in the transcripts. Critical questions not appearing in this table were not found back in the transcripts.}
+\caption{Number of Occurrences of AS0-AS9, CQ0-CQ12 in Transcripts. Critical Questions not Appearing in This Table were not Found in Transcripts.}
 \label{table:transcripts:results:argumentschemes}
@@
 Suppose now that the modelers that created Figure~\ref{fig:transcripts:grl} would continue their analysis by discussing critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'', and that they would answer this with ``Yes, namely \texttt{Choose randomly}''.
-This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.''
-\item \textsf{DISABLE:} Disable the element or relationship of argument scheme to which critical questions pertains.
+This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' %SG: Very good example.
+But should we also show it in the model as a new element?
+\item \textsf{DISABLE:} Disable the element or relationship of an argument scheme to which critical questions pertains.
 This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
@@
 The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
-What we also see from Figure~\ref{fig:transcripts:grl} is that disabling actor ``Teacher'' also disables all elements belonging to this actor.
-Furthermore, disabling task ``Dynamic simulation'' also disables all incoming and outgoing links with this task.
+What we also see from Figure~\ref{fig:transcripts:grl} is that actor ``Teacher'' is disabled and thus all the elements that are bound to this actor are disabled as well.
+Furthermore, disabled task ``Dynamic Simulation'' also disables all incoming and outgoing links with this task.
 
@@
 In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
-In this case, the participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
+In this case, participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
 While the previous names are not shown in the figure, they show up in the details pane of the corresponding element.
+%SG: why generate car has a star but not this one?
 \item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-In Figure~\ref{fig:transcripts:grl} we see one example of such a counter-attack.
+In Figure~\ref{fig:transcripts:grl}, we see one example of such a counter-attack.
 First, task ``Control car influx per road'' is attack by answering CQ2 (irrelevant task).
-However, after discussing this, the participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
-Therefore, the arguments that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
+However, after discussing this, participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
+Therefore, the argument that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
 We provide a precise semantics for this in Figure~\ref{sect:formalframework}.
@@
 %MvZ: I completely removed the subsection below. I think it doesn't add much and it is confusing. What do you think?
+
+%SG: I agree... It was more confusing and downgrade our approach
 

### 2017-08-17 main.tex
k
many improvements (see email)
Updated section 4 and gave comments

@@
 \usepackage{fancyvrb}
+\usepackage{changepage}
 \usepackage{algorithm}% http://ctan.org/pkg/algorithms
@@
  \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
-        and Sepideh Ghanavati\Mark{3}\\[1em]
+        Sepideh Ghanavati\Mark{2},
+        and Floris Bex\Mark{3}\\[1em]
  \normalsize
  \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
+  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Computer Science  & \Mark{3}Department of Information and Computing Sciences\tabularnewline
+  University of Luxembourg & Texas Tech University, USA & Utrecht University\tabularnewline
+  \url{marcvanzee@gmail.com} & \url{sepideh.ghanavati@ttu.edu} & \url{f.j.bex@uu.nl}
  \end{tabular}\\[3em] % some more space after the title part
@@
 \input{gmas}
+\input{examples}
 \input{formalframework}
@@
 
+\input{designprompt}
 \input{transcripts}

### 2017-08-21 abstract.tex
Small edits and few comments abstract and intro

@@
 While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found in the resulting model.
-For instance, reasons for accepting or rejecting an element or a relation between two elements are not captured.
-In this paper, we investigate to what extent argumentation techniques from artificial intelligence can be applied to a goal modelingn approach.
-We apply the argument scheme for practical reasoning (PRAS), which is used in AI to reason about goals to the Goal-oriented Requirements Language (GRL).
-We develop a formal metamodel for the new language, link it to the GRL metamodel, and we implement our extension into jUCMNav, the Eclipse-based open source tool for GRL.
+For instance, the discussion on whether to accept or reject a certain goal and the ultimate rationale for acceptance or rejection cannot be captured in current goal modeling languages.
+Based on a case study in which stakeholders discuss requirements for a Traffic Simulator, we apply argumentation techniques from artificial intelligence to a goal modeling approach.
+Thus, we combine a traditional goal modelling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).
+RationalGRL provides a methodology, formal semantics and tool support to capture the discussions and outcomes of the argumentation process that leads to a goal model.
 \end{abstract}
-%SG: what are the findings? or what case study we are using? 
-%SG_New: I still have these comments!

### 2017-08-21 introduction.tex
Small edits and few comments abstract and intro

@@
 \subsection{Requirements Engineering and Goal Modeling}
-
+\todo{Floris}{all}{Starting section 1 directly with subsection 1.1 is generally frowned upon. Maybe we can get rid of the subsections in the intro?} 
 Requirements Engineering (RE) is an approach to assess the role of a future information system within a human or automated environment.
@@
 
-One of the first activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
+Among the initial activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what the implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
 These activities fall under the umbrella of goal modeling.
-The are a large number of established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}).
+There are a large number of established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}).
 Several goal modeling languages have been developed in the last two decades as well.
@@
 Therefore, it is often not necessary to record all of the details of the discussion process that led to the final goal model.
-However, most real-world information systems -- e.g., air-traffic management, industrial production processes, or government and healthcare services -- are complex and are not constructed in a short amount of time, but rather over the course of several workshops.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons stated below.
+However, goal models for many complex, real-world information systems -- e.g., air-traffic management systems, systems that support industrial production processes, or government and healthcare services -- are not constructed in a short amount of time, but rather over the course of several workshops with stakeholders and requirements engineers.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons.
 
@@
 Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
-If previous preferences and opinions are not stored explicitly, it is not straightforward to remind stakeholders of their previous opinions which results in having unnecessary discussions and revisions.
+If previous preferences and opinions are not stored explicitly, it is not straightforward to remind stakeholders of their previous opinions which can result in unnecessary discussions and revisions.
 As the number of participants increases, revising the goal model based on changing preferences can take up a significant amount of time.
@@
 
-Third, alternative and different ideas and opposing views that could potentially have led to different goal diagrams could be lost.
-For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce softgoals ``easy to use'' and ``fast'' to one softgoal ``easy to use''.
-Thus, the resulting goal model merely contains the softgoal ``easy to use'', but the discussion as well as the decision to reject ``fast'' are lost.
+Third, alternative and different ideas and opposing views that could potentially have led to different goal models could be lost.
+For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce the two goals ``easy to use'' and ``fast'' to one goal ``easy to use''.
+Thus, the resulting goal model merely contains the goal ``easy to use'', but the discussion as well as the decision to reject the goal ``fast'' are lost.
 This leads to a poor understanding of the problem and solution domain.
@@
 
-Fourth and lastly, in goal models in general, ``rationale'' behind any decision is static and does not immediately impact the goal models when they change.
+Fourth and finally, in goal models in general the rationale behind any modeling decision is static and changing these rationales does not immediately impact the goal models.
 That is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
 A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, a part of the goal model may change, but it is not possible to formally reason about the consistency of this new goal model with the underlying beliefs and arguments in the goal modeling languages.
+Similarly, with existing goal modelling languages one can change a part of the goal model, but it is not possible to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
 This becomes more problematic if the participants constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs put forward by another group of stakeholders.
+
+\todo{Floris}{Floris}{I think the above 4 reasons overlap in some cases. If it's OK, I'll rewrite this.}
 
@@
 \item 
-The framework must have a tool support.
+The framework must have software tool support.
 \item 
@@
 \end{quote}
+\todo{Floris}{Marc,Sepideh}{While I like an explicit research question, you introduce new terminology here ``construct, mechanisms and rules''.
+Do you explicitly refer to this later in the paper (e.g.
+we found mechanisms A, B, rules R1, R2, etc.)?}
+To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL.
+The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation theory and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
+\todo{Floris}{Floris}{mention Singh et al.
+argument schemes in RE here.}
 
-To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL.
-The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
-
-In past, we introduced our preliminary steps towards RationalGRL framework~\cite{bnaic2014,vanzee-etal:renext2015,vanZee-etal:er2016} with its tool-support~\cite{vanZee-etal:comma2016}.
-In our previous work, argument diagrams are integrated with GRL models to allow better representation of stakeholder's arguments and discussions.
-In our current paper, we improve RationalGRL framework by adding argument schemes and critical questions to argumentation framework and provide traceability links between the argumentation framework and GRL.
-For this, we modify and extend our metamodel to specify how our framework extends GRL, and we use this as the basis for a prototype web-based implementation~\footnote{The tool is available at <GITHUB LINK>}.
+In previous work, we have taken preliminary steps towards the RationalGRL framework~\cite{bnaic2014,vanzee-etal:renext2015,vanZee-etal:er2016} with its tool-support~\cite{vanZee-etal:comma2016}.
+In our previous work, argument diagrams are integrated with GRL models to allow better representation of stakeholders' arguments and discussions.
+In our current paper, we improve the RationalGRL framework by adding argument schemes and critical questions to argumentation framework and provide traceability links between the argumentation framework and GRL.
+For this, we modify and extend our metamodel to specify how our framework extends GRL, and we use this as the basis for a prototype web-based implementation\footnote{The tool is available at <GITHUB LINK>}.
 %SG: Placeholder for the link.
-We develop an initial list of argument schemes and critical questions by analyzing a set of transcripts containing discussion about an information system.
+Inspired by work on practical reasoning in Artificial Intelligence~\cite{atkinson2007}, we develop an initial list of argument schemes and critical questions for reasoning about goals in RE.
+We do this by analyzing a set of transcripts containing discussions about a traffic simulator information system.
 Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains.
@@
 This methodology is intended to be used by practitioners in the field.
-We formalize RationalGRL using propositional logic, and develop algorithms for adding goal elements and arguments.
-We illustrates the steps of our framework  and how the framework can be applied in real cases with a traffic simulator example.
+We provide formal semantics for RationalGRL by translating the elements of the GRL language into simple propositional logical formulas and subsequently capturing arguments about these elements of a GRL model in a formal argumentation framework \cite{dung1995}.
+We then develop a set of algorithms for applying argument schemes and critical questions about goal models.
+These algorithms further formally ground the RationalGRL methodology and form the basis of the RationalGRL tool.
+We illustrate the steps of our framework and how RationalGRL can be applied in a real case with the traffic simulator case study.
 
@@
 This article is organized as follows.
-Section~\ref{sect:background} contains background and introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, and argument schemes.
+Section~\ref{sect:background} introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, the Practical Reasoning Argument Scheme (PRAS)~\cite{atkinson2007} and some general background about .
 Section~\ref{sect:overview} provides a brief and high-level overview of our framework, together with a metamodel and the methodology.
-Section~\ref{sect:gmas} contains an in depth explanation of how we obtained an initial set of argument schemes by annotating transcripts from discussions about an information system, and in Section~\ref{sect:examples} we provide several examples of these arguments schemes.
-In Section~\ref{sect:formalframework} we formalize RationalGRL and our arguments schemes using propositional logic and we use well-known argumentation semantics to compute which arguments are accepted and which are rejected.
+Section~\ref{sect:gmas} contains an in depth explanation of how we obtained an initial set of argument schemes and critical questions by annotating transcripts from discussions about an information system, and in Section~\ref{sect:examples} we provide several examples of these schemes and questions.
+In Section~\ref{sect:formalframework} we provide formal semantics for GRL and show how argumentation semantics \cite{dung1995} can be used to compute which arguments are accepted and which are rejected.
 We also develop various algorithms for the argument schemes in this section.

### 2017-08-21 background.tex
metamodel
Small changes to 2.1, 2.2 and extended example in 2.3

@@
 
-We use the traffic simulator design case to explain the concepts and the framework in this paper.
-Traffic simulator design is the topic of discussion in the transcripts used for the argumentation framework.
-%Most of the examples in this article, as well as the topic of discussion in the transcripts we analyze, come from the traffic simulator design exercise. 
-In this exercise, designers are provided with a problem description, requirements, and a description of the desired outcomes.
-The problem description is given in full in Appendix~\ref{sect:designprompt}, and is summarized as follows: The client of the project is Professor E, who teaches civil engineering courses at an American university.
+We use a traffic simulator design case to explain the concepts and framework in this paper.
+Our examples and case study are based on a recent series of experiments by Schriek et al.
+\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning problem in software engineering.
+In the original exercise (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes: The client of the project is Professor E, who teaches civil engineering courses at an American university.
 In order for the professor to teach students how various theories (such as queuing theory) around traffic lights work, a software analyst is hired to specify the goal and requirements of the system.
-%and It is the task of the designer to specify a system in which the professor can teach students how various theories around traffic lights works, such as queuing theory.
-To this end, a piece of software will be developed in which students can create visual maps of an area, regulate traffic, and so forth.
-The original version of the problem descrption~\cite{UCIworkshop} is well known in the field of design reasoning and it has been used in a workshop\footnote{\url{http://www.ics.uci.edu/design-workshop/}}.
-Transcripts of this workshop have been analyzed in detail~\cite{Petre:2013:SDA:2535028}.
+To this end, a piece of software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
 Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
 
-In this Section and in Section~\ref{}, we focus on the simplified example of traffic simulator case for simplicity.
-In Section~\ref{} we provide the detailed example of the traffic simulator and its model.
+For their experiments, Schriek et al.
+\cite{SchriekEtal2016} gave the traffic simulator assignment to 12 groups of students in a Software Architecture course at MSc level.
+They asked the groups to record their two-hour design session, and the recordings were subsequently transcribed (see Appendix~\ref{sect:transcripts:excerpts}).
+We used these transcripts to develop our argument schemes and critical questions (Section \ref{sect:gmas}), and we also use the traffic simulator case for simple (Figures \ref{fig:example-small}, \ref{fig:args}) and more extensive examples (Section \ref{sect:gmas}).
 
@@
 \label{sect:background:grl}
-GRL is a visual modeling language for specifying intentions, business goals, and \emph{non-functional requirements} of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
+GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
 GRL is part of the User Requirements Notation, an ITU-T standard, that combines goals and non-functional requirements with functional and operational requirements (i.e.
-use case maps) in one.
+use case maps).
 GRL can be used to specify alternatives that have to be considered, decisions that have been made, and rationales for making decisions.
@@
 \includegraphics[width=0.5\textwidth]{img/Example1}
-\caption{Partial GRL Model of Traffic Simulator Example}
+\caption{Partial GRL Model of the traffic simulator example}
 \label{fig:example-small}
@@
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars} which are alternative way of achieving the goal, \texttt{Generate Cars}.
+In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars} which are alternative way of achieving the goal \texttt{Generate Cars}.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).
+\todo{Floris}{Sepideh}{Belief elements are not explained and not in the example.
+I think it would be good to include them, as beliefs allow one to give some sort of rationale for e.g.
+goals (in fact, one of the reviewers of one of our previous papers argued that beliefs can also be used to capture reasons).}
 
@@
 A contribution link has a qualitative contribution type or a quantitative contribution.
-Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}.
+Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}, and a \emph{hurt} qualitative contribution to the task \texttt{Static Simulation}.
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
@@
 %SG: I changed this question but let me know if you want to change them back and then change the graph.
-	\item What does \texttt{Keep same cars} mean?
-	\item Why does task \texttt{Create New cars} contribute negatively to \texttt{Static simulation} and positively to \texttt{Dynamic simulation}?
+	\item What does \texttt{Keep Same Cars} mean?
+	\item Why does task \texttt{Create New Cars} contribute negatively to \texttt{Static Simulation} and positively to \texttt{Dynamic Simulation}?
 	\item Why does \texttt{Simulate Traffic} AND-decompose into two tasks?
@@
 These are the type of the questions that we cannot answer by just looking at GRL models.
-The model in Figure~\ref{fig:example-small} does not contain information about discussions that lead to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
-In this article we aim to address this shortcoming.
+The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+With our RationalGRL framework we aim to address this shortcoming.
 
@@
 \cite{Raz1978-RAZPR,walton1990}) and artificial intelligence \cite{bratman1987,atkinson2007}.
-One approach is to capture practical reasoning using arguments schemes and critical questions~\cite{walton1990}.
-Applying an argument schemes results in an argument in favor of, for example, taking an action.
-This argument can, then, be tested with critical questions about, for instance, whether the action is possible given the situation.
-A negative answer to such a question leads to a counterargument to the original argument for the action.
+One approach is to capture practical reasoning with argumentation or argument schemes~\cite{walton1990}.
+Applying an argument scheme results in an argument in favor of, for example, taking an action.
+This argument can then be tested with critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original argument for the action.
 
-Atkinson et al.~\cite{atkinson2007} develop an argument scheme for practical reasoning, which they call the \emph{Practical Reasoning Argument Scheme} (PRAS).
-This argument scheme is as follows\footnote{The original argument schemes uses slightly different terminology: we have replaced \emph{value} with \emph{softgoal}.
-Our results do not depend on these adjustments, but they make the relation between PRAS and GRL more clear.}:
+Atkinson and Bench-Capon~\cite{atkinson2007} develop and formalize the \emph{Practical Reasoning Argument Scheme} (PRAS).
+A slightly adapted version\footnote{The original argument scheme uses different terminology: Atkinson and Bench-Capon say that realizing a goal contributes to some \emph{value}, an interest which an agent may or may not wish to uphold or subscribe to, such as \texttt{High Performance}.
+In this paper, we use the term \emph{softgoal} to indicate such interests.} of this argument scheme is as follows:
 
-%\todo{Sepideh}{Marc}{Read this part and see if it makes sense.}
-%MvZ: Yes it makes sense, I made some small adjustments but overall it looks good.
 \begin{itemize}
 \item[] We have goal $G$,
-\item[] Doing action $A$ realizes goal $G$,
+\item[] Performing action $A$ realizes goal $G$,
 \item[] Which will contribute positively to the softgoal $S$
@@
 
-The capital letters $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals.
-Once instantiated, we obtain an argument.
-We can for instance instantiate the argument scheme above with the elements of Figure~\ref{fig:example-small} as follows:
+Here, $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals to provide a specific practical argument.
+For example, a concrete argument about the traffic simulator (cf.
+Figure \ref{fig:example-small}) is as follows:
 \begin{itemize}
 \item[] We have goal \texttt{Simulate Traffic},
-\item[] Doing action \texttt{Dynamic Simulation} realizes goal \texttt{Simulate Traffic}, 
-\item[] Which contribute positively to the softgoal \texttt{Improve Simulation} 
+\item[] Performing action \texttt{Dynamic Simulation} realizes goal \texttt{Simulate Traffic}, 
+\item[] Which contributes positively to the softgoal \texttt{Improve Simulation} 
 \item[] \textit{Therefore} 
@@
 
+Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
 Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
@@
 \item Will the action realize the desired goal?
-\item Are there alternative ways of realizing the same goal?
-\item Are there alternative ways of contributing to the same softgoal?
-\item Does performing the action contribute negatively to some other softgoal?
-\item Is the action possible?
-\item Can the desired goal be realized?
-\item Is the softgoal indeed a legitimate softgoal?
+\item Are there alternative ways of realizing the same goal or contributing to the same softgoal?
+\item Does performing the action have a negative side effect?
 \end{enumerate}
@@
 These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 1: if we find that  \texttt{Dynamic Simulation} actually does not realize goal \texttt{Simulate Traffic}, we can form a counter-argument.
-
-Another way to elaborate an argument for an action is to suggest an alternative action that realizes the same goal (question 2) or an alternative goal that contributes to the same softgoal (question 3).
-For example,  can argue that another task \texttt{Semi-dynamic Simulation} also realizes the goal \texttt{Simulate Traffic}.
+Take, for example, critical question 1: if we find that the task \texttt{Dynamic Simulation} actually does not realize goal \texttt{Simulate Traffic}, we can form a counter-argument.
+Another way to criticize an argument for an action is to suggest an alternative action that realizes the same goal (question 2).
+For example, we can argue that task \texttt{Static Simulation} also realizes the goal \texttt{Simulate Traffic}.
+It is also possible that performing an action has a negative side effect (critical question 3).
+For example, while the action \texttt{Create New Cars} realizes the goal \texttt{Generate Cars}, it has a negative side effect, namely hurting \texttt{Static Simulation} (cf.
+Figure \ref{fig:example-small}).
 
@@
 In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations.
-Given an argumentation framework, once can compute very precisely which arguments are accepted and which are rejected using argumentation semantics~\cite{Dung1995}.
-These semantics have been studied extensively in AI, and are currently a popular topic of research (see~\cite{argumentationbook} for an overview).
+Given an argumentation framework, once can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Full definitions of argumentation frameworks and semantics will be given in section \ref{sect:gmas}.
+In this section, we will briefly discuss the intuitions behind these semantics.}.
+Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the traffic simulation example, where arguments are rendered as boxes and attack relations as arrows.
+There is an argument (A1) for \texttt{Keep Same Cars}, an argument (A2) for \texttt{Create New Cars}, and an argument A3 that \texttt{Create New Cars} has a negative effect on \texttt{Static Simulation} (A3).
+The two alternative PRAS instantiations are A1 and A2.
+These arguments mutually attack each other, denoted by the double-headed arrow between A1 and A2.
+Argument A3 attacks A2, as it points to a negative side-effect of \texttt{Create new cars}.
+
+\begin{figure}[ht!]
+\centering
+\begin{tikzpicture}
+        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.5) {\argtext{A1}{Action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}\\ \emph{Therefore} we should perform \texttt{Keep Same Cars}}};
+        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \texttt{Create New Cars} realizes goal \texttt{Generate Cars}\\ \emph{Therefore} we should perform \texttt{Create New Cars}}};
+        \node[minimum size=1cm] (att2) [argNodeIN] at (3.5,0) {\argtext{A3}{\texttt{Create Same Cars} has a negative effect on \texttt{Static Simulation}}};
+         \path
+    (att2) edge [attackLink] (att1)
+    (att1) edge [attackLink,<->] (att3);
+\end{tikzpicture}
+\caption{Arguments and attacks in the traffic simulation example.}
+\label{fig:pras:example}
+\end{figure}
+
+Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics.
+The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
+In the argumentation framework in Figure~\ref{fig:pras:example}, argument A3 is undefeated because it has no attackers.
+This makes A2 defeated, because one of its attackers, A3, is undefeated.
+A1 is then also undefeated, since its only attacker, A2, is defeated by A3.
+Thus, the set of undefeated (acceptable) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A1, A3$\}$.
 

### 2017-08-21 introduction.tex
Small changes to 2.1, 2.2 and extended example in 2.3

@@
 
-The aim of this research is to resolve the above issues by developing a framework with tool-support, which combines goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{atkinson2007}.
+The aim of this research is to resolve the above issues by developing a framework with tool-support, which combines goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{BenchCaponDunne2007}.
 We have identified five important requirements for our framework:

### 2017-08-22 background.tex
Major update section 1, some changes section 2.4, references added

@@
 \cite{SchriekEtal2016} gave the traffic simulator assignment to 12 groups of students in a Software Architecture course at MSc level.
-They asked the groups to record their two-hour design session, and the recordings were subsequently transcribed (see Appendix~\ref{sect:transcripts:excerpts}).
-We used these transcripts to develop our argument schemes and critical questions (Section \ref{sect:gmas}), and we also use the traffic simulator case for simple (Figures \ref{fig:example-small}, \ref{fig:args}) and more extensive examples (Section \ref{sect:gmas}).
+They asked the groups to record their design session, and the recordings were subsequently transcribed (see Appendix~\ref{sect:transcripts:excerpts}).
+We used three of these transcripts as an extensive case study along which we developed our argument schemes and critical questions (Sections \ref{sect:gmas} and \ref{sect:examples}).
+Furthermore, we also use the traffic simulator case for a simple  running example in this section (Figures \ref{fig:example-small}, \ref{fig:pras:example}).
 
@@
 
-Figure~\ref{fig:example-small} illustrates a simplified GRL diagram from the traffic simulator design exercise (a full example is shown in Figure~\ref{fig:transcripts:grl} and will be discussed in Section~\ref{sect:gmas}).
-%MvZ: Since the example no longer matches the full figure I reworded it slightly in the sentence above.
+Figure~\ref{fig:example-small} illustrates a simplified GRL diagram from the traffic simulator design exercise.
 An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Simulator}, Figure~\ref{fig:example-small}).
@@
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, some of the tasks are  \texttt{Create new cars} and \texttt{Keep same cars} which are alternative way of achieving the goal \texttt{Generate Cars}.
+In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).
@@
 AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}.
+In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}, as they are alternative ways of achieving the goal \texttt{Generate Cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
@@
 Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}, and a \emph{hurt} qualitative contribution to the task \texttt{Static Simulation}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors.
+Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors or resources.
 Here, the goal \texttt{Generate Cars} depends on the resource \texttt{Car Objects}.
@@
 \cite{Raz1978-RAZPR,walton1990}) and artificial intelligence \cite{bratman1987,atkinson2007}.
-One approach is to capture practical reasoning with argumentation or argument schemes~\cite{walton1990}.
+One approach is to capture practical reasoning with argument schemes~\cite{walton1990}.
 Applying an argument scheme results in an argument in favor of, for example, taking an action.
@@
 Here, $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals to provide a specific practical argument.
-For example, a concrete argument about the traffic simulator (cf.
-Figure \ref{fig:example-small}) is as follows:
+For example, a concrete argument about the traffic simulator is as follows:
 \begin{itemize}
 \item[] We have goal \texttt{Simulate Traffic},
-\item[] Performing action \texttt{Dynamic Simulation} realizes goal \texttt{Simulate Traffic}, 
+\item[] Performing action \texttt{Static Simulation} realizes goal \texttt{Simulate Traffic}, 
 \item[] Which contributes positively to the softgoal \texttt{Improve Simulation} 
 \item[] \textit{Therefore} 
-\item[] We should perform action \texttt{Dynamic Simulation}.
+\item[] We should perform action \texttt{Static Simulation}.
 \end{itemize}
@@
 Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
-Atkinson \emph{et al.}~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
+Atkinson et al.~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
 Some examples of critical questions are as follows.
@@
 These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 1: if we find that the task \texttt{Dynamic Simulation} actually does not realize goal \texttt{Simulate Traffic}, we can form a counter-argument.
+Take, for example, critical question 1: if we find that performing a \texttt{Static Simulation} actually does not realize goal \texttt{Simulate Traffic}, we can form a counter-argument.
 Another way to criticize an argument for an action is to suggest an alternative action that realizes the same goal (question 2).
-For example, we can argue that task \texttt{Static Simulation} also realizes the goal \texttt{Simulate Traffic}.
-It is also possible that performing an action has a negative side effect (critical question 3).
-For example, while the action \texttt{Create New Cars} realizes the goal \texttt{Generate Cars}, it has a negative side effect, namely hurting \texttt{Static Simulation} (cf.
-Figure \ref{fig:example-small}).
+For example, we can argue that performing a \texttt{Dynamic Simulation} also realizes the goal \texttt{Simulate Traffic} on its own.
+Finally, it is possible that performing an action has a negative side effect (critical question 3).
+For example, while the action \texttt{Create New Cars} realizes the goal \texttt{Generate Cars}, it has a negative side effect, namely hurting \texttt{Static Simulation}: having the simulation constantly create new cars is a functionality that does not allow for a static simulation.
 
@@
 In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations.
-Given an argumentation framework, once can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Full definitions of argumentation frameworks and semantics will be given in section \ref{sect:gmas}.
-In this section, we will briefly discuss the intuitions behind these semantics.}.
+Given an argumentation framework, we can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Formal definitions of argumentation frameworks and semantics will be given in section \ref{sect:gmas}.
+In this section, we will briefly discuss the intuitions behind these concepts.}.
 Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the traffic simulation example, where arguments are rendered as boxes and attack relations as arrows.
-There is an argument (A1) for \texttt{Keep Same Cars}, an argument (A2) for \texttt{Create New Cars}, and an argument A3 that \texttt{Create New Cars} has a negative effect on \texttt{Static Simulation} (A3).
-The two alternative PRAS instantiations are A1 and A2.
-These arguments mutually attack each other, denoted by the double-headed arrow between A1 and A2.
-Argument A3 attacks A2, as it points to a negative side-effect of \texttt{Create new cars}.
+There are two slightly simplified practical reasoning arguments based on PRAS: argument A1 for \texttt{Keep Same Cars} and argument A2 for \texttt{Create New Cars}.
+Argument A2 proposes an alternative way of realizing the same goal \texttt{Generate Cars} with respect to argument A1 and vice versa (cf.
+critical question 2), so A1 and A2 mutually attack each other, denoted by the double-headed arrow between A1 and A2.
+Argument A3 says that \texttt{Create New Cars} has a negative effect on \texttt{Static Simulation}, so A3 attacks A2, as it points to a negative side-effect of \texttt{Create new cars} (critical question 3).
 
@@
 \begin{tikzpicture}
-        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.5) {\argtext{A1}{Action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}\\ \emph{Therefore} we should perform \texttt{Keep Same Cars}}};
-        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \texttt{Create New Cars} realizes goal \texttt{Generate Cars}\\ \emph{Therefore} we should perform \texttt{Create New Cars}}};
-        \node[minimum size=1cm] (att2) [argNodeIN] at (3.5,0) {\argtext{A3}{\texttt{Create Same Cars} has a negative effect on \texttt{Static Simulation}}};
+        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.4) {\argtext{A1}{Action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} we should perform \texttt{Keep Same Cars}}};
+        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \texttt{Create New Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} we should perform \texttt{Create New Cars}}};
+        \node[minimum size=1cm] (att2) [argNodeIN] at (3.5,0) {\argtext{A3}{\texttt{Create New Cars} has a negative effect on \texttt{Static Simulation}}};
          \path
@@
 
-Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics.
+Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics~\cite{Dung1995}.
 The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
 In the argumentation framework in Figure~\ref{fig:pras:example}, argument A3 is undefeated because it has no attackers.
-This makes A2 defeated, because one of its attackers, A3, is undefeated.
+This makes A2 defeated (indicated by the lighter grey color of A2), because its attacker A3 is undefeated.
 A1 is then also undefeated, since its only attacker, A2, is defeated by A3.
@@
 
-The usefulness of argument schemes, and PRAS in particular, for the analysis of practical reasoning situations has been shown in different areas such as e-democracy~\cite{cartwright2009IS}, law~\cite{atkinson2005legal}, planning \cite{medellin2013planning} and choosing between safety critical actions \cite{tolchinsky2012deliberation}.
-In this article, we aim at capturing the stakeholder's discussions as formal argumentation based on PRAS to decide whether intentional elements and their relationships are shown in the resulting goal model.
-This approach provides a rationalization to the elements of the goal model in terms of underlying arguments, and furthermore, it helps understanding why certain other elements have been accepted and others have been rejected.
+In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we have explored one way of combining PRAS and GRL.
+In this work, which takes a similar approach to \cite{Jureta:RE2008}, argument diagrams are translated to GRL models (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
+The argument diagrams are complex practical reasoning arguments, where the premises are the goals and the conclusion is some action we need to perform to realize those goals.
+Thus, we have essentially two complex diagrams, a practical reasoning argument diagram and a GRL goal diagram, and a mapping between them.
 
-Argumentation schemes and their associated critical questions are very well suited for modeling discussions about a goal model: as Murukannaiah et al.~\cite{murukannaiah2015} have shown, they can guide users in systematically deriving conclusions and making assumptions explicit.
-This can also be seen from the obvious similarities between PRAS (actions, goals, values) and GRL (tasks, goals, softgoals) in the example above.
+Connecting argumentation and goal modelling by providing a mapping between two types of diagram was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
+One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion thus far has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
+Furthermore, in our previous work we did not provide the specific critical questions for goal models (see Section \ref{sect:gmas}).
+This meant that any counterargument had to be constructed from scratch, as no guidance was given as to possible ways to criticize a GRL model.
+So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged.
+Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling.
 
-However, there are also some differences between PRAS and GRL.
-Not all elements and relationships of GRL fit into PRAS.
-For instance, PRAS does not have a notion of ``resource'', and many of the relationships of GRL do not occur in PRAS.
-Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson can actually apply to GRL.
+In the current version of GRL, we do not literally take the Practical Reasoning Argument Scheme and its critical questions - this would be impossible given that there are elements of the GRL language, such as resources, which cannot be found in PRAS.
+Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson and Bench-Capon~\cite{atkinson2007} actually apply to GRL models.
 Therefore, we develop our own set of argument schemes and critical questions in Section~\ref{sect:gmas} by analyzing transcripts of discussions about the traffic simulator.

### 2017-08-22 introduction.tex
Major update section 1, some changes section 2.4, references added

--- 2017-08-21 introduction.tex
@@
 
-\subsection{Requirements Engineering and Goal Modeling}
-\todo{Floris}{all}{Starting section 1 directly with subsection 1.1 is generally frowned upon. Maybe we can get rid of the subsections in the intro?} 
 Requirements Engineering (RE) is an approach to assess the role of a future information system within a human or automated environment.
@@
 
-\subsection{Problem Statement}
-\label{sect:problem_statement}
-
 A goal model is often the result of a discussion process between a group of stakeholders.
@@
 However, goal models for many complex, real-world information systems -- e.g., air-traffic management systems, systems that support industrial production processes, or government and healthcare services -- are not constructed in a short amount of time, but rather over the course of several workshops with stakeholders and requirements engineers.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development for various reasons.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development.
 
-First, it is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
+The first difficulty is that the goal modeling phase, particularly in large projects, is dynamic and that goal models continuously change and evolve.
+Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}, and stakeholders may change their opinion about a modeling decision in between two goal modeling sessions, which in turn may require revisions of the goal model.
+If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially have led to different goal models are lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
+Furthermore, other stakeholders, such as developers who were not the original authors of the goal model, may have to make sense of a goal model in order to, for example, use it as input in a later RE stage or in the development phase.
+If preferences, opinions and rationales behind the goal models are not stored explicitly, it may not only be more difficult to understand the model, but the stakeholders may also end up having the same unnecessary discussions throughout the goal modeling phase.
+
+A further problem is that in goal models in general the rationale behind any modeling decision is static, that is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
+A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
+Similarly, with existing goal modelling languages one can change a part of the goal model, but it is not possible to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
+This becomes even more problematic if the stakeholders constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs of another group of stakeholders.
+The disconnect between the goal models and their underlying beliefs and opinions may further lead to a poor understanding of the problem and solution domain, which is an important reason of RE project failure~\cite{curtis1988field}.
+
+%First, it is well-known that stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
 Therefore, it is possible that a stakeholder changes his or her opinion about a modeling decision in between two goal modeling sessions, which may require revisions of the goal model.
@@
 
-Second, other stakeholders, such as new developers on the team who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase.
+%Second, other stakeholders, such as new developers on the team who were not the original authors of the goal model, may have to make sense of the goal model, for instance, to use it as an input in a later RE stage or at the design and development phase.
 If these users have no knowledge of the underlying rationale of the goal model, it may not only be more difficult to understand the model, but they may also end up having the same discussions as the previous group of stakeholders.
 
-Third, alternative and different ideas and opposing views that could potentially have led to different goal models could be lost.
+%Third, alternative and different ideas and opposing views that could potentially have led to different goal models could be lost.
 For instance, a group of stakeholders specifying a goal model for a user interface may decide to reduce the two goals ``easy to use'' and ``fast'' to one goal ``easy to use''.
@@
 
-Fourth and finally, in goal models in general the rationale behind any modeling decision is static and changing these rationales does not immediately impact the goal models.
+%Fourth and finally, in goal models in general the rationale behind any modeling decision is static and changing these rationales does not immediately impact the goal models.
 That is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
@@
 
-\todo{Floris}{Floris}{I think the above 4 reasons overlap in some cases. If it's OK, I'll rewrite this.}
+To summarize, what is needed is a way of recording the rationales (beliefs, opinions, discussions, ideas) underlying a goal model.
+It should be possible to see how these rationales changed during the goal modeling process, and the rationales hould be clearly linked to the various elements of the resulting goal model.
+In order to be able to do this, we propose a framework with tool-support that combines traditional goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{BenchCaponDunne2007}.
+We have identified \textbf{five important requirements} for our framework:
 
-\subsection{Research Questions}
-
-The aim of this research is to resolve the above issues by developing a framework with tool-support, which combines goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{atkinson2007}.
-We have identified five important requirements for our framework:
 \begin{enumerate}
-\item The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
+\item 
+The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
 \item 
@@
 \item 
+There should be a methodology for the framework to guide the practitioners in its application in real cases.
+\item 
 The framework must have software tool support.
-\item 
-There should be a methodology for the framework to guide the practitioners in its application in real cases.
 \end{enumerate}
 
-The five requirements above serve as the success criteria of our approach. That is, if our framework satisfies these requirements, we reach our goal.
+Our framework, which we call \emph{RationalGRL}, combines the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory and discourse modeling called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
+Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason, and they are very well suited for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}.
+One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for capturing multi-agent planning \cite{medellin2013planning}, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.
 
-In this context, the main research question is:
+Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyse and guide stakeholders' discussions about goal models.
+Our approach thus provides a rationalization to the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted and others have been rejected.
+Our list of argument schemes was constructed by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
+This ensures that the argumentation schemes we propose are close to actual real-world discussions stakeholders have (\textbf{requirement 1}).
 
-\begin{quote}
-\textbf{RQ.} What are the constructs, mechanisms and rules for developing a framework that formally captures the discussions between stakeholders such that it can generate goal models?
-\end{quote}
-\todo{Floris}{Marc,Sepideh}{While I like an explicit research question, you introduce new terminology here ``construct, mechanisms and rules''.
-Do you explicitly refer to this later in the paper (e.g.
-we found mechanisms A, B, rules R1, R2, etc.)?}
-To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL.
-The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation theory and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
-\todo{Floris}{Floris}{mention Singh et al.
-argument schemes in RE here.}
+The meta-model of the RationalGRL framework clearly specifies the traceability links between the arguments based on the schemes and the GRL models (\textbf{requirement 2}).
+In addition to this meta-model, we provide formal semantics for RationalGRL by formalising the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
+We then formally capture the link between argumentation and goal modelling as a set of algorithms for applying argument schemes and critical questions about goal models.
+These formal traceability links allow us to compute the effect of the arguments and counterarguments proposed in a discussion on a GRL model (\textbf{requirement 3}).
+In other words, we can determine whether the elements of a GRL model are acceptable given potentially contradictory opinions of stakeholders.
+Thus, we add a new formal evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (in addition to their \emph{satisfiability}~\cite{Amyot:2010:EGM:1841349.1841356}).
 
-In previous work, we have taken preliminary steps towards the RationalGRL framework~\cite{bnaic2014,vanzee-etal:renext2015,vanZee-etal:er2016} with its tool-support~\cite{vanZee-etal:comma2016}.
-In our previous work, argument diagrams are integrated with GRL models to allow better representation of stakeholders' arguments and discussions.
-In our current paper, we improve the RationalGRL framework by adding argument schemes and critical questions to argumentation framework and provide traceability links between the argumentation framework and GRL.
-For this, we modify and extend our metamodel to specify how our framework extends GRL, and we use this as the basis for a prototype web-based implementation\footnote{The tool is available at <GITHUB LINK>}.
-%SG: Placeholder for the link.
-Inspired by work on practical reasoning in Artificial Intelligence~\cite{atkinson2007}, we develop an initial list of argument schemes and critical questions for reasoning about goals in RE.
-We do this by analyzing a set of transcripts containing discussions about a traffic simulator information system.
-Our framework is fully extensible, meaning that argument schemes and critical questions can be added by users for specific problem domains.
-We develop a methodology for using RationalGRL, which consists of developing goal models and posing arguments in an integrated way.
-This methodology is intended to be used by practitioners in the field.
-We provide formal semantics for RationalGRL by translating the elements of the GRL language into simple propositional logical formulas and subsequently capturing arguments about these elements of a GRL model in a formal argumentation framework \cite{dung1995}.
-We then develop a set of algorithms for applying argument schemes and critical questions about goal models.
-These algorithms further formally ground the RationalGRL methodology and form the basis of the RationalGRL tool.
-We illustrate the steps of our framework and how RationalGRL can be applied in a real case with the traffic simulator case study.
+Because we want RE practitioners in the field to be able to use our RationalGRL framework (\textbf{requirement 4}), we also propose a methodology for using RationalGRL, which consists of developing goal models and posing arguments based on schemes in an integrated way.
+To show that the RationalGRL methodology can be used in a real case, we illustrate the steps of our methodology with the traffic simulator case study.
+Finally, we have developed a web-based prototype\footnote{\url{insertURL}} for building goal models and arguing about them, which acts as a supporting tool to the RationalGRL methodology (\textbf{requirement 5}).
 
-\subsection{Organization} %SG: Will get back to this at the end of the paper.
-
-This article is organized as follows.
-Section~\ref{sect:background} introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, the Practical Reasoning Argument Scheme (PRAS)~\cite{atkinson2007} and some general background about .
+The rest of this article is organized as follows.
+Section~\ref{sect:background} introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, the Practical Reasoning Argument Scheme (PRAS)~\cite{atkinson2007} and discusses some of our previous work on combining GRL and PRAS.
 Section~\ref{sect:overview} provides a brief and high-level overview of our framework, together with a metamodel and the methodology.
 Section~\ref{sect:gmas} contains an in depth explanation of how we obtained an initial set of argument schemes and critical questions by annotating transcripts from discussions about an information system, and in Section~\ref{sect:examples} we provide several examples of these schemes and questions.
-In Section~\ref{sect:formalframework} we provide formal semantics for GRL and show how argumentation semantics \cite{dung1995} can be used to compute which arguments are accepted and which are rejected.
+In Section~\ref{sect:formalframework} we provide formal semantics for GRL and show how argumentation semantics \cite{Dung1995} can be used to compute which arguments are accepted and which are rejected.
 We also develop various algorithms for the argument schemes in this section.

### 2017-08-24 background.tex
Comments
Added small comments in TeX (%FB) and todo's

@@
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).
-\todo{Floris}{Sepideh}{Belief elements are not explained and not in the example.
-I think it would be good to include them, as beliefs allow one to give some sort of rationale for e.g.
+In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}.\todo{F}{S}{Later in the paper this goal is called  \texttt{Simulate} , not \texttt{Simulate Traffic}.
+\texttt{Simulate Traffic} seems more clear to me.} In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).
+\todo{F}{S,M}{Belief elements are not explained and not in the example.
+I think it would be good to include them (or at least discuss them), as beliefs allow one to give some sort of rationale for e.g.
 goals (in fact, one of the reviewers of one of our previous papers argued that beliefs can also be used to capture reasons).}
@@
 
+%FB: I deleted a few of the critical questions
 \begin{enumerate}
@@
 For example, while the action \texttt{Create New Cars} realizes the goal \texttt{Generate Cars}, it has a negative side effect, namely hurting \texttt{Static Simulation}: having the simulation constantly create new cars is a functionality that does not allow for a static simulation.
+
+%FB: I extended the example and put the figure back in. Note that this is just to explain argumentation, not to reason about the model in 2.2. Putting explicit text in the example makes it more clear.
 
@@
 
+%FB: I discuss our previous work on RationalGRL here now (used to be in intro), and I discuss why the approach we take in this paper is different from the approach in our previous work.
+
 In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we have explored one way of combining PRAS and GRL.
@@
 So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged.
-Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling.
+Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling (requirement 4).
 

### 2017-08-24 formalframework.tex
Comments
Added small comments in TeX (%FB) and todo's

@@
 
+\todo{F}{F,M}{Strictly speaking you don't define a set of atoms here, but a set of predicates...
+A neat solution would be to do it as follows: Def 1 is the \emph{GRL language} L, and we add a new Def for \emph{GRL models} M, which is then a set of atoms based on L.
+The definition of Argument is then the same as Model, but this is correct (cf.
+the metamodel: a model is an argument).}
+
 An example of the specification of the GRL model in Figure~\ref{fig:examples:decomposition} is shown in Table~\ref{table:grl_atom_spec}.
@@
 \begin{definition}[Argument] \label{def:argument}
-An \emph{argument} $A\subseteq Att$ is a set of atoms from $Att$.
+An \emph{argument} $A\subseteq Args$ is a set of atoms from $At$.
 \end{definition}
@@
 
+\todo{F}{F,M}{It would be good to have a formal definition of how to construct models given an argumentation framework.
+If we define a model M at the start of this section, this should be fairly easy.}
+
 Constructing GRL models from the arguments is extremely simple: We simply compute the extensions of the argumentation frameworks, and collect all atomic sentences in the accepted arguments.

### 2017-08-24 gmas.tex
Comments
Added small comments in TeX (%FB) and todo's

@@
-\section{Argument Schemes for Goal Modeling}
+\section{Argument Schemes for Goal Modeling: a Case Study}
+%FB: I added "a case study" to emphasize that we perform a case study on the transcripts (see also the introduction)
 \label{sect:gmas}
@@
 
-\subsection{Details experiment}
-
+\subsection{Details Case Study}
+\todo{F}{F}{I think it would be better to call what Marc did (i.e.
+annotating the transcripts with arguments and schemes) a case study instead of an experiment.
+This also avoids the confusion between our case study (annotating transcripts) and the original experiment.
+The way it is written now makes it seem like we performed the experiments with the students.
+I can amend the below text early next week.}
 The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.

### 2017-08-24 introduction.tex
Comments
Added small comments in TeX (%FB) and todo's

@@
 \label{sect:introduction}
+
+%FB: I deleted the subsection headings 1.1, 1.2 because the paper started with heading 1.1 immediately following heading 1.
+(which is frowned upon).
+The intro is not overly long now so I don't think we need subsections
 
@@
 In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development.
-
-The first difficulty is that the goal modeling phase, particularly in large projects, is dynamic and that goal models continuously change and evolve.
-Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}, and stakeholders may change their opinion about a modeling decision in between two goal modeling sessions, which in turn may require revisions of the goal model.
-If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially have led to different goal models are lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
-Furthermore, other stakeholders, such as developers who were not the original authors of the goal model, may have to make sense of a goal model in order to, for example, use it as input in a later RE stage or in the development phase.
-If preferences, opinions and rationales behind the goal models are not stored explicitly, it may not only be more difficult to understand the model, but the stakeholders may also end up having the same unnecessary discussions throughout the goal modeling phase.
-
-A further problem is that in goal models in general the rationale behind any modeling decision is static, that is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, with existing goal modelling languages one can change a part of the goal model, but it is not possible to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
-This becomes even more problematic if the stakeholders constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs of another group of stakeholders.
-The disconnect between the goal models and their underlying beliefs and opinions may further lead to a poor understanding of the problem and solution domain, which is an important reason of RE project failure~\cite{curtis1988field}.
 
@@
 
+%FB: I changed the text above to that below. The above 4 reasons overlap in some cases, I tried to make this more concise.
+
+The first difficulty is that the goal modeling phase, particularly in large projects, is dynamic and that goal models continuously change and evolve.
+Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}, and stakeholders may change their opinion about a modeling decision in between two goal modeling sessions, which in turn may require revisions of the goal model.
+If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially have led to different goal models are lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
+Furthermore, other stakeholders, such as developers who were not the original authors of the goal model, may have to make sense of a goal model in order to, for example, use it as input in a later RE stage or in the development phase.
+If preferences, opinions and rationales behind the goal models are not stored explicitly, it may not only be more difficult to understand the model, but the stakeholders may also end up having the same unnecessary discussions throughout the goal modeling phase.
+
+A further problem is that in goal models in general the rationale behind any modeling decision is static, that is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
+A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
+Similarly, with existing goal modelling languages one can change a part of the goal model, but it is not possible to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
+This becomes even more problematic if the stakeholders constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs of another group of stakeholders.
+The disconnect between the goal models and their underlying beliefs and opinions may further lead to a poor understanding of the problem and solution domain, which is an important reason of RE project failure~\cite{curtis1988field}.
+
 To summarize, what is needed is a way of recording the rationales (beliefs, opinions, discussions, ideas) underlying a goal model.
@@
 \item 
-Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.
+Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.\todo{F}{M,S}{Given the current logical framework, is it possible to compute effect of changes in GM on ARG?
+It seems to me that we can only do the other way around with the algorithms.}
 \item 
@@
 
+%In this context, the main research question is:
+
+%\begin{quote}
+%\textbf{RQ.} What are the constructs, mechanisms and rules for developing a framework that formally captures the discussions between stakeholders such that it can generate goal models?
+%\end{quote}
+
+%FB: I deleted the research question.
+I think that the requirements already nicely capture exactly what our goal is in the paper.
+We can rephrase this goal as an RQ, but what does this buy us?
+Also, the RQ introduces new terminology (construct, mechanisms, rules), to which we have to explicitly refer later in the paper (e.g.
+we found mechanisms A, B, rules R1, R2, etc.).
+
+To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL.
+The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation theory and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
+\todo{Floris}{Floris}{mention Singh et al.
+argument schemes in RE here.}
+
+%FB: Some of the text below explaining our use of argument schemes was taken from 2.4. The discussion of previous work on RationalGRL was moved to 2.4.
 Our framework, which we call \emph{RationalGRL}, combines the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory and discourse modeling called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
@@
 One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for capturing multi-agent planning \cite{medellin2013planning}, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.
+
+%FB: the below text is largely new/reworked stuff. I link what we do in the paper to each of the 5 requirements
 
@@
 
+\todo{F}{all}{check text below when we finish the paper}
+
 The rest of this article is organized as follows.

### 2017-08-24 overview.tex
fixed many things
Comments
Added small comments in TeX (%FB) and todo's

@@
 \subsection{RationalGRL Methodology}
-%\todo{Sepideh}{Marc}{Review this section and make sure it is correct}
 
@@
 \includegraphics[scale=0.4]{img/framework}
-\caption{The RationalGRL Framework}
+\caption{The RationalGRL Framework\todo{F}{M,S}{Can you make this a vector image?}}
 \label{fig:rationalgrl-framework}
@@
 \includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology}
+\caption{The RationalGRL Methodology\todo{F}{M,S}{Can you make this a vector image?}}
 \label{fig:rationalgrl-methodology}
@@
 Figure~\ref{fig:metamodel} depicts the RationalGRL metamodel linking the main elements of our argumentation extension to the main GRL elements.
-We describe the metamodel bottom-up, starting with the GRL package.
+We describe the metamodel bottom-up, starting with the GRL package.\todo{F}{all}{Fig.
+6 is now far removed from where it is mentioned in the text, we should fix this when paper is nearly finished}
 

### 2017-08-29 formalframework.tex
added some comments and minor things

@@
 the metamodel: a model is an argument).}
+
+\todo{M}{F,M}{\url{https://en.wikipedia.org/wiki/Atomic_formula}
+I think atom just means there are no connectives or negations.
+I used this terminology since it is used in logic programming which is very concrete and clearly defined in terms of syntax.
+Predicates have such a wide meaning that it may confuse the reader: \url{https://en.wikipedia.org/wiki/Predicate_(mathematical_logic)}.
+Anyway I am also not very happy with these definitions yet.
+I'll give it another try taking your suggestions into account.}
 

### 2017-08-29 introduction.tex
added some comments and minor things

@@
 
-A further problem is that in goal models in general the rationale behind any modeling decision is static, that is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
+A further problem is that the rationale behind goal modeling decisions is usually static, that is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
 A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
@@
 To summarize, what is needed is a way of recording the rationales (beliefs, opinions, discussions, ideas) underlying a goal model.
-It should be possible to see how these rationales changed during the goal modeling process, and the rationales hould be clearly linked to the various elements of the resulting goal model.
+It should be possible to see how these rationales changed during the goal modeling process, and the rationales should be clearly linked to the various elements of the resulting goal model.
 In order to be able to do this, we propose a framework with tool-support that combines traditional goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{BenchCaponDunne2007}.
@@
 \item 
-Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.\todo{F}{M,S}{Given the current logical framework, is it possible to compute effect of changes in GM on ARG?
-It seems to me that we can only do the other way around with the algorithms.}
+Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.
+\todo{F}{M,S}{Given the current logical framework, is it possible to compute effect of changes in GM on ARG?
+It seems to me that we can only do the other way around with the algorithms.
+(M): In a sense this is possible.
+In our formalization, each GRL element and link can be translated into an instantiation of an argument scheme, so adding GRL element simply means adding more instantiate argument schemes, who can then be attack by critical questions.
+But I think it is a good idea to make this more explicit.}
 \item 
@@
 \todo{Floris}{Floris}{mention Singh et al.
-argument schemes in RE here.}
+argument schemes in RE here.
+(Marc): I wouldn't go into much detail here.
+Perhaps one sentence and then refer to the related work section.}
 

### 2017-09-01 background.tex
Small changes, typos etc.

@@
 Our examples and case study are based on a recent series of experiments by Schriek et al.
-\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning problem in software engineering.
-In the original exercise (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes: The client of the project is Professor E, who teaches civil engineering courses at an American university.
-In order for the professor to teach students how various theories (such as queuing theory) around traffic lights work, a software analyst is hired to specify the goal and requirements of the system.
-To this end, a piece of software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
-Although the concepts of traffic lights, lanes, and intersections are common and appear to be simple, building a traffic simulator to represent these relationships and events in real time turns out to be challenging.
-
-For their experiments, Schriek et al.
-\cite{SchriekEtal2016} gave the traffic simulator assignment to 12 groups of students in a Software Architecture course at MSc level.
-They asked the groups to record their design session, and the recordings were subsequently transcribed (see Appendix~\ref{sect:transcripts:excerpts}).
-We used three of these transcripts as an extensive case study along which we developed our argument schemes and critical questions (Sections \ref{sect:gmas} and \ref{sect:examples}).
+\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning assignment in software engineering.
+In this assignment (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes: The client of the project is Professor E, who teaches civil engineering courses at an American university.
+In order for the professor to teach students the various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
+Schriek and colleagues asked designers (groups of students) to discuss the goal and requirements of this traffic simulator.
+These discussions were recorded and transcribed, and we used these transcripts for an extensive case study in which we developed our argument schemes and critical questions (Sections \ref{sect:gmas} and \ref{sect:examples}).
 Furthermore, we also use the traffic simulator case for a simple  running example in this section (Figures \ref{fig:example-small}, \ref{fig:pras:example}).
@@
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}.\todo{F}{S}{Later in the paper this goal is called  \texttt{Simulate} , not \texttt{Simulate Traffic}.
+In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}.\todo{F}{S}{Later in the paper this goal is called  \texttt{Simulate}, not \texttt{Simulate Traffic}.
 \texttt{Simulate Traffic} seems more clear to me.} In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).
@@
 
-%FB: I discuss our previous work on RationalGRL here now (used to be in intro), and I discuss why the approach we take in this paper is different from the approach in our previous work.
-
 In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we have explored one way of combining PRAS and GRL.

### 2017-09-01 discussion.tex
Small changes, typos etc.

@@
 \label{sect:goalmodeling:relatedwork}
+
+In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we have explored one way of combining PRAS and GRL.
+In this work, which takes a similar approach to \cite{Jureta:RE2008}, argument diagrams are translated to GRL models (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
+The argument diagrams are complex practical reasoning arguments, where the premises are the goals and the conclusion is some action we need to perform to realize those goals.
+Thus, we have essentially two complex diagrams, a practical reasoning argument diagram and a GRL goal diagram, and a mapping between them.
+
+Connecting argumentation and goal modelling by providing a mapping between two types of diagram was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
+One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion thus far has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
+Furthermore, in our previous work we did not provide the specific critical questions for goal models (see Section \ref{sect:gmas}).
+This meant that any counterargument had to be constructed from scratch, as no guidance was given as to possible ways to criticize a GRL model.
+So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged.
+Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling (requirement 4).
 

### 2017-09-01 gmas.tex
Small changes, typos etc.

--- 2017-08-17 gmas.tex
@@
 The list of argument schemes and critical questions that we have obtained from our analysis is shown in Table~\ref{table:argument-schemes}.
-The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are related to relationships, the next two (AS12-AS13) are for intentional elements in general, and the last is (Att) is a generic counterargument for any type of argument that has been put forward.
+The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are related to relationships, the next two (AS12-AS13) are for intentional elements in general, and the last (Att) is a generic counterargument against any type of argument that has been put forward.
 
@@
 
-\subsection{Details experiment}
+\subsection{Case Study Details}
 
-The transcripts we used are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+In order to come up with the list of argument schemes and critical questions in Table~\ref{table:argument-schemes}, we performed an extensive case study based on a recent series of experiments by Schriek et al.
+\cite{SchriekEtal2016}.
+In these experiments, Schriek and colleagues gave the traffic simulator assignment (Appendix~\ref{sect:designprompt}) to 12 groups of two or three students in a Software Architecture course at MSc level.
+These groups had a maximum of two hours to design a traffic simulator, which included a discussion of the requirements of this traffic simulator.
+The students were asked to record their design sessions, and the recordings were subsequently transcribed.
+We used three of these transcripts totalling 153 pages of discussion for our case study.
+In the case study, we annotated the transcripts with argument schemes and critical questions, and we created RationalGRL models based on the discussions.
+In this section we present examples of these annotations and models.
+All original transcripts, annotations, and models are available online\footnote{\url{insertURL}}.
 
-\paragraph{Subjects} The subjects for the case study are three teams of Master students from Utrecht University, following the Software Architecture course.
-Two teams consist of three students, and one team consists of two students.
+To develop the list of argument schemes and critical questions, we started with an initial list of 8 argument schemes and 18 critical questions that we derived from the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}.
+We then annotated the three transcripts with arguments and critical questions from this initial list.
+If there were arguments or critical questions that did not appear in the initial list, we added them to the list.
+Argument schemes and critical questions that did not appear were removed from the list.
+Appendix~\ref{sect:transcripts:excerpts} contains some examples of annotations.
 
-\paragraph{Experimental Setup} The goal of the experiment is to design a traffic simulator.
-Participants were asked to use a think-aloud method during the design session.
-The assignment was slightly adjusted to include several viewpoints as an end product in order to conform to the course material~\cite{Bass:2012:SAP:2392670}.
-The full problem descriptions can be found in Appendix~\ref{sect:designprompt}.
-All groups were instructed to apply the \emph{functional architecture method}, focusing on developing the \emph{context}, the \emph{functional}, and the \emph{informational} viewpoints of the traffic simulator software.
-The students had two hours for the tasks.
-The entire discussion was documented in transcripts.
-The details of the transcripts are shown in Table~\ref{table:transcripts:info}.
+\subsection{Case Study Results}
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{|l|l|l|l|}
-\hline
-& transcript $t_1$ & transcript $t_2$ & transcript $t_3$\\
-\hline
-participants & 2 & 3 & 3\\
-\hline
-duration & 1h34m52s & 1h13m39s & 1h17m20s\\
-\hline
-\end{tabular}
-\caption{Number of participants and duration of the transcripts.}
-\label{table:transcripts:info}
-\end{table}
-
-\paragraph{Process of Extracting Argument Schemes and Critical Questions}
-
-To develop the list of argument schemes and critical questions, we started with an initial list of 8 argument schemes and 18 critical questions that we derived from PRAS (AS1-AS4, AS6-AS9 of Table~\ref{table:argument-schemes}).
-We, then, annotated transcripts with arguments and critical questions from this initial list.
-If there were arguments or critical questions that did not appear in the initial list, we added them to the list.
-% Existing argument schemes that did not appear were removed from the list, but critical questions were not removed (see discussion in Section~\ref{sect:gmas:transcripts:analysis}). 
-% I don't understand it! why?
-%MvZ: I agree it is confusing.
-The idea is that critical questions are often implicit, so we kept them in since we believe it may still stimulate discussions.
-I now simply left it out since it doesn't add much and will only be confusing.
-
-Most of the occurrences were not literally found back, but had to be inferred from the context. This can be seen in the various examples we will discuss in Section~\ref{sect:examples}.
-
-%It is generally known in the argumentation literature that it can be very difficult to identify arguments in natural language texts~\cite{walton-etal2004}.
-Arguments are often imprecise, lack conclusion, and may be supported by nonverbal communication that is not captured in transcripts.
-However, there is hardly any research on argument extraction in the requirement engineering domain.
-Thus, despite this potential weakness in our approach, we believe it nevertheless is at least useful as something that others can build further on (see Section~\ref{sect:goalmodeling:openissues}).
-%TODO I don't understand the last few sentences... what does argument extraction mean? why is it needed? don't use "something that others" it is very informal for a journal paper. 
-%MvZ: I was also sceptical about this paragraph when I wrote it. I think it belongs in the related work section where we can discuss it in more detail. I now removed it here.
-
-\subsection{Experiment Results}
-
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{img/Fig6}
-\caption{The GRL model manually constructed from transcript $t_1$.
-Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
-Elements and relationships with no dot have been inferred by us.} %TODO We need to update the caption
-\label{fig:transcripts:grl}
-\end{figure*}
-
-All original transcripts, annotations, and models are available on the following website \todo{Marc}{Marc}{Change URL and put everything together with the tool}:
-\begin{quote} \url{http://www.github.com/marcvanzee/RationalArchitecture}
-\end{quote}
-We also provide excerpts of the annotation in Appendix~\ref{sect:transcripts:excerpts}, and most of the examples we use in this article come from the transcripts.
-
-We found a total of 159 instantiation of argument schemes AS0-AS11 in transcripts.
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice (Table~\ref{table:transcripts:results:argumentschemes}).
+We found a total of 159 instantiation of argument schemes AS0-AS11 in transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
+The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice .
 A large portion (about 60\%) of argument schemes involved discussions about tasks of the information system (AS2, AS10).
-
-We annotated 41 applications of critical questions. Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, CQ13).
+We annotated 41 applications of critical questions.
+Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, CQ13).
 
@@
 This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
-\todo{Marc}{all}{Perhaps we should explain a bit about the model here?}
+
+\begin{figure*}[h!]
+\includegraphics[width=\textwidth]{img/Fig6}
+\caption{The GRL model manually constructed from transcript $t_1$.
+Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
+Elements and relationships with no dot have been inferred by us.} %TODO We need to update the caption
+\label{fig:transcripts:grl}
+\end{figure*}
 

### 2017-09-01 introduction.tex
Small changes, typos etc.

@@
 \label{sect:introduction}
-
-%FB: I deleted the subsection headings 1.1, 1.2 because the paper started with heading 1.1 immediately following heading 1.
-(which is frowned upon).
-The intro is not overly long now so I don't think we need subsections
 
@@
 
-To answer the research question above and solve some of the concerns mentioned in Subsection~\ref{sect:problem_statement}, we propose a framework called RationalGRL.
-The RationalGRL framework combines the Goal-oriented Requirements Language (GRL) with a technique from argumentation theory and discourse modeling called \emph{argument schemes}~\cite{walton-etal2004}.
-\todo{Floris}{Floris}{mention Singh et al.
-argument schemes in RE here.
-(Marc): I wouldn't go into much detail here.
-Perhaps one sentence and then refer to the related work section.}
+Following on from our previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop a framework called \emph{RationalGRL}, which combines the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory and discourse modeling called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
+Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.
+Associated with argumentation schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way.
+Argument schemes they are very well suited for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}.
 
-%FB: Some of the text below explaining our use of argument schemes was taken from 2.4. The discussion of previous work on RationalGRL was moved to 2.4.
-Our framework, which we call \emph{RationalGRL}, combines the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory and discourse modeling called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
-Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason, and they are very well suited for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}.
 One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for capturing multi-agent planning \cite{medellin2013planning}, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.
-
-%FB: the below text is largely new/reworked stuff. I link what we do in the paper to each of the 5 requirements
-
 Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyse and guide stakeholders' discussions about goal models.

### 2017-09-01 overview.tex
Small changes, typos etc.

@@
 
-\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions to attack the argument by counter-arguments.
+\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions to attack the argument with counter-arguments.
 Each argument scheme includes one or more critical questions.
-For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}.
+For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contribute to the softgoal?} and \emph{Does the goal contribute to some other softgoals?}.
 %It is worth mentioning that, answering a critical question may result in a "conflict" situation. This is out of the scope of our work here.  
 %MvZ: removed this sentence because I don't know what you mean with "conflict". I think We should explain it better or leave it out.
-When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and the step (1).
+When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and step (1).
 
@@
 For instance, a GRL contribution can have a qualitative strength, ranging from ``Make'' to ``Break''.
-Since these concepts are not relevant to our framework, they have been omitted, but none of the results of this article depend on this.}.
+Since these concepts are not relevant to our framework, they have been omitted.}.
 

### 2017-09-02 formalframework.tex
Small updates sect. 6
Added definition, simplified formal argument example.

--- 2017-08-15 formalframework.tex
@@
 
-The examples and corresponding visualizations of the previous section provide some insight on how our RationalGRL framework can be used.
-However, since we would like to implement our framework in a tool, we require a more precise formalization.
-In this section, we present a formalization of our framework using formal logic.
-This is the main approach using in formal argumentation, so it allows us to use many of the techniques developed in that area directly (see Section~\ref{sect:goalmodeling:relatedwork} for more details).
-
-In the first two subsections we formalize a static representation of our framework: We develop a formal language to specify a GRL model (independently of arguments) in the fist subsection, and we formalize the notion of an argument in the second subsection.
-In the third subsection we formalize the dynamics of our framwork: we develop algorithms for instantiating argument schemes and for answering critical questions.
+The examples and corresponding visualizations of the previous sections provide some insight on how our RationalGRL framework can be used.
+In this section, we present a more precise formalization of our framework.
+In the first two subsections we formalize a static representation of our framework: We develop a formal language to specify a GRL model (Section \ref{sect:goalmodeling:logicallanguage}) and we formalize the notion of an argument (Section \ref{sect:goalmodeling:argumentationsemantics}).
+In the third subsection we formalize the dynamics of our framework: we develop algorithms for instantiating argument schemes and for answering critical questions.
 The formal framework we develop in this section serves as the underlying model for the prototype implementation we present in the next section.
@@
 
-We assume familiarity with basic notions from propositonal logic and set theory.
-A \emph{propositional atom} is a grounded formula (a formula without variables) with no logical connectives ($\vee, \wedge, \rightarrow, \leftrightarrow$) or negation ($\neg$).
-We use integers $i,j,k\in\mathbb{N}$ as identifiers for actors and GRL elements.
+We assume familiarity with basic notions from propositional logic and set theory.
+An \emph{atom} is a formula with no logical connectives ($\vee, \wedge, \rightarrow, \leftrightarrow$) or negation ($\neg$).
+We first define the logical language for RationalGRL.
 
-\begin{definition}[GRL Atoms]
-The \emph{set of GRL atoms} $At$ contains the following atoms:
+\begin{definition}[RationalGRL Language]
+The \emph{RationalGRL language} $\mathcal{L}$ contains the following atoms:
 \begin{itemize}
@@
 
-For all of the above definitions we assume that all identifiers used in each atom are different, i.e.,
+We use integers $i,j,k\in\mathbb{N}$ as identifiers for actors and GRL elements. For all of the above definitions we assume that all identifiers used in each atom are different, i.e.,
 \begin{itemize}
@@
 
-An example of the specification of the GRL model in Figure~\ref{fig:examples:decomposition} is shown in Table~\ref{table:grl_atom_spec}.
-The specification has been written in logic programming style.
+Using the language from Definition 1, we can define GRL models. A \emph{ground atom} is an atom without variables.
+
+\begin{definition}[GRL model] \label{def:argument}
+A \emph{GRL model} $\mathcal{M}$ is a set of ground atoms based on $\mathcal{L}$.
+\end{definition}
+
+An example of the specification of the GRL model in Figure~\ref{fig:examples:decomposition} is shown in Table~\ref{table:grl_atom_spec}, written in logic programming style.
 In this article we do not make use of any logic programming techniques but this would be interesting future work (see Section~\ref{sect:goalmodeling:openissues}).
@@
 
-\subsection{Formal argumentation semantics}
+\subsection{Formal argumentation}
 \label{sect:goalmodeling:argumentationsemantics}
@@
 \begin{definition}[Argument] \label{def:argument}
-An \emph{argument} $A\subseteq Att$ is a set of atoms from $Att$.
+An \emph{argument} $A\subseteq Args$ is a set of ground atoms based on $\mathcal{L}$.
 \end{definition}
@@
 For instance, $$\{goal(0),name(0,development\_team)\},$$ is an argument stating that identifier 0 is a goal element with name ``development team''.
+Notice that an argument and a GRL model are essentially the same: as was discussed in Section \ref{sect:metamodel}, each element of a GRL model is an argument.
 
-Note that this definition does not put any restrictions on the content of an argument.
+Definition \ref{def:argument} does not put any restrictions on the content of an argument.
 This means it is possible to form arguments that do not correspond to correct or sensible GRL models.
-Consider for instance the following argument: $$\{goal(0), task(0), name(0,static\_simulation)\}.$$ This argument states that identifier 0 is both a goal and a task with name ``statis simulation'', which is clearly undesirable since it does not correspond to a valid GRL model.
+Consider for instance the following argument: $$\{goal(0), task(0), name(0,static\_simulation)\}.$$ This argument states that identifier 0 is both a goal and a task with name ``static simulation'', which is clearly undesirable since it does not correspond to a valid GRL model.
 While we could have chosen to put logical constraints on the content of an argument, we choose not to do so here.
 Instead, we put these restrictions on the algorithms in section~\ref{sect:algorithms}.
-The algorithms are constructed in such a way that they only allow the construction of argumetns that correspond to valid GRL models.\footnote{Note that we do not provide formal proofs for these claims in this paper.
-This is an interesting direction of future work, but it is not the focus of this article.} The advantage of this approach is that it keeps the presentation of the formal part relatively simple.
+The algorithms are constructed in such a way that they only allow the construction of arguments that correspond to valid GRL models.
+The advantage of this approach is that it keeps the presentation of the formal part relatively simple.
 
@@
 \item $S$ is \emph{conflict-free} if it does not attack itself,
-\item $S$ \emph{defends} $A$ if $S$ attacks each attack against $A$.
+\item $S$ \emph{defends} $A$ if for each $B$ such that $B$ attacks $A$, $S$ attacks $B$,
 \item $S$ is \emph{admissible} if $S$ is conflict-free and defends each argument in it.
@@
 \begin{tikzpicture}
-        \node[minimum size=1cm] (a) [argNodeIN] at (-2,-2) {$A_1$};
-        \node[minimum size=1cm] (b) [argNodeIN] at (-2,0) {$A_2$};
-        \node[minimum size=1cm] (c) [argNodeIN] at (0,0) {$A_3$};
-        \node[minimum size=1cm] (d) [argNodeIN] at (2,0) {$A_4$};
-        \node[minimum size=1cm] (e) [argNodeIN] at (2,-2) {$A_5$};
+        \node[minimum size=1cm] (b) [argNodeIN] at (-2,0) {$A_1$};
+        \node[minimum size=1cm] (c) [argNodeIN] at (0,0) {$A_2$};
+        \node[minimum size=1cm] (d) [argNodeIN] at (2,0) {$A_3$};
 
          \path
-    (a) edge [attackLink] (b)
     (c) edge [attackLink] (b)
-    (c) edge [attackLink] (d)
     (d) edge [attackLink] (c)
-    (d) edge [attackLink] (e)
-    (e) edge [attackLink, loop right] (e);
+
 \end{tikzpicture}
@@
 
-Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}.
-We say that the set $\{A_1,A_4,A_5\}$ \emph{attacks} argument $A_2$, because $A_1$ attacks $A_2$.
-However, $\{A_1,A_4,A_5\}$ is not \emph{conflict-free}, because $A_4$ attacks $A_5$, so this set of arguments attacks itself.
-If we remove $A_5$ from this set, then it is conflict-free.
-In total, all the following sets are conflict-free in Figure~\ref{fig:goalmodeling:arg2}: $$\{A_1,A_3\},\{A_1,A_4\},\{A_2,A_4\},\{A_1\},\{A_2\},\{A_3\},\{A_4\},\emptyset.$$
+Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, where $A_2$ attacks $A_1$ and $A_3$ attacks $A_2$.
+There are two admissible sets: $\{A_3\}$ and $\{A_1, A_3\}$, in which $A_3$ defends $A_1$ against its attacker $A_2$.
+Sets containing both $A_2$ and either $A_1$ or $A_3$ are not conflict free, and the sets $\{A_1\}$ and $\{A_2\}$ do not defend themselves against $A_2$ and $A_3$, respectively.
 
-However, not all of these sets are \emph{admissible}.
-For instance, the set $\{A_2,A_4\}$ is not admissible, because $A_1$ attacks this set, but the set does not defend itself against this attack.
-The set $\{A_1,A_3\}$ is admissible, because its only attacker, $A_4$ is attacked by $A_3$.
-
-There are a large number of different semantics using these notions to determine which arguments are acceptable, such as the \emph{grounded}, the \emph{preferred}, the \emph{stable}, and the \emph{complete} semantics.
-However, in this article our argumentation frameworks are very simple, in the sense that they do not introduce cycles.
-In future work, we aim to extend this by using preferences (see open issues in Section~\ref{sect:goalmodeling:openissues}).
-An advantage of our current approach is that all the standard semantics coincide when there are no cycles, which simplifies our case.
-We use the preferred semantics here, but we could equivalently have chosen any other version.
+Given the notion of admissible sets, we can then define our argumentation semantics.
+There are a large number of different semantics to determine which arguments are acceptable; in this article, we choose preferred semantics.
 
 \begin{definition}[Preferred semantics~\cite{Dung1995}] \label{def:semantics}
-Suppose an argumentation framework $AF=(Arg,Att)$. A set of arguments $S\subseteq Arg$ is a \emph{preferred extension} if and only if 
-\begin{itemize}
-\item $S$ is an admissible set of argument,
-\item For each admissible set of arguments $T\subseteq A$: $S\not\subset T$.
-\end{itemize}
+A preferred extension of an argumentation framework $AF$ is a maximal (w.r.t. set inclusion) admissible set of $AF$.
 \end{definition}
 
-In our example in Figure~\ref{fig:goalmodeling:arg2}, the preferred extensions are $\{A_1,A_3\}$ and $\{A_1,A_4\}$.
+In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_1,A_3\}$.
 
 \subsection{Algorithms for argument schemes and critical questions}
-\label[sect:algorithms]
+\label[sect:algorithms}
 

### 2017-09-11 background.tex
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!

@@
-\section{Background: Goal-oriented Requirements Language and argument schemes}
+\section{Background: Goal-oriented Requirements Language and Argumentation}
 \label{sect:background}
@@
 In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.
-Next, we introduce argument schemes, and in particular, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, which is an argument scheme that is used to form arguments and counter-arguments about situations involving goals.
-Finally, we give an overview of the integration between PRAS and GRL.
+Next, we introduce argumentation: we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
+Finally, we briefly discuss the possibilities integrating PRAS and GRL.
 %This will be our starting point in the next section.
@@
 In order for the professor to teach students the various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
-Schriek and colleagues asked designers (groups of students) to discuss the goal and requirements of this traffic simulator.
-These discussions were recorded and transcribed, and we used these transcripts for an extensive case study in which we developed our argument schemes and critical questions (Sections \ref{sect:gmas} and \ref{sect:examples}).
+Schriek and colleagues asked designers (groups of students) to discuss the requirements of this traffic simulator.
+These discussions were recorded and transcribed, and we used these transcripts for an extensive case study (Section \ref{sect:} on the basis of which we developed our RationalGRL framework (Section \ref{sect:gmas}).
 Furthermore, we also use the traffic simulator case for a simple running example in this section (Figures \ref{fig:example-small}, \ref{fig:pras:example}).
@@
 
-\subsection{Argument Scheme for Practical Reasoning (PRAS)}
+\subsection{Practical Reasoning Argument Scheme (PRAS)}
 \label{sect:background:pras}
 
-Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy (e.g.
-\cite{Raz1978-RAZPR,walton1990}) and artificial intelligence \cite{bratman1987,atkinson2007}.
+Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy and artificial intelligence.
 One approach is to capture practical reasoning with argument schemes~\cite{walton1990}.
@@
 
-Atkinson and Bench-Capon~\cite{atkinson2007} develop and formalize the \emph{Practical Reasoning Argument Scheme} (PRAS).
-A slightly adapted version\footnote{The original argument scheme uses different terminology: Atkinson and Bench-Capon say that realizing a goal contributes to some \emph{value}, an interest which an agent may or may not wish to uphold or subscribe to, such as \texttt{High Performance}.
-In this paper, we use the term \emph{softgoal} to indicate such interests.} of this argument scheme is as follows:
+Atkinson and Bench-Capon~\cite{atkinson2007} develop and formalize the \emph{Practical Reasoning Argument Scheme} (PRAS). A simplified version of this argument scheme is as follows:
 
 \begin{itemize}
-\item[] We have goal $G$,
+\item[] $G$ is a goal,
 \item[] Performing action $A$ realizes goal $G$,
-\item[] Which will contribute positively to the softgoal $S$
+%\item[] Which will contribute positively to the softgoal $S$
 \item[] \textit{Therefore} 
-\item[] We should perform action $A$
+\item[] Action $A$ should be performed
 \end{itemize}
 
-Here, $G$, $A$, and $S$ are variables, which can be instantiated with concrete goals, actions, and softgoals to provide a specific practical argument.
+Here, $G$ and $A$ are variables, which can be instantiated with concrete goals and actions to provide a specific practical argument.
 For example, a concrete argument about the traffic simulator is as follows:
 \begin{itemize}
-\item[] We have goal \texttt{Simulate Traffic},
-\item[] Performing action \texttt{Static Simulation} realizes goal \texttt{Simulate Traffic}, 
-\item[] Which contributes positively to the softgoal \texttt{Improve Simulation} 
+\item[] \texttt{Generate Cars} is a goal,
+\item[] Performing action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}, 
+%\item[] Which contributes positively to the softgoal \texttt{Improve Simulation} 
 \item[] \textit{Therefore} 
-\item[] We should perform action \texttt{Static Simulation}.
+\item[] Action \texttt{Keep same Cars} should be performed
 \end{itemize}
 
-Practical reasoning is defeasible, in that conclusions which are at one point acceptable can later be rejected because of new information.
-Atkinson et al.~\cite{atkinson2007} define a set of critical questions that point to typical ways in which a practical argument can be criticized by, for example, questioning the validity of the elements in the scheme or the connections between the elements.
+Note that PRAS is an argument scheme that captures a full inference step: `$G$, $A$ realizes $G$ \emph{Therefore} $A$'.
+There are, however, also schemes that capture simpler reasoning patterns, such as claims of the form `$A$ does not realize $G$'.
+We will discuss these schemes below.
+
+In argumentation, conclusions which are at one point acceptable can later be rejected because of new information.
+For example, we may argue that, in fact, performing action \texttt{Keep Same Cars} does not realize goal \texttt{Generate cars}, thus giving a counterargument to the above instantiation of PRAS.
+Atkinson et al.~\cite{atkinson2007} define a set of so-called \emph{critical questions} that point to typical ways in which an argument based on PRAS can be criticized by.
 Some examples of critical questions are as follows.
 
-%FB: I deleted a few of the critical questions
 \begin{enumerate}
 \item Will the action realize the desired goal?
-\item Are there alternative ways of realizing the same goal or contributing to the same softgoal?
+\item Are there alternative ways of realizing the same goal?
 \item Does performing the action have a negative side effect?
@@
 
-These critical questions can point to new arguments that might counter the original argument.
-Take, for example, critical question 1: if we find that performing a \texttt{Static Simulation} actually does not realize goal \texttt{Simulate Traffic}, we can form a counter-argument.
+The idea is that answers to critical questions are counterarguments to the the original PRAS argument.
+These counterarguments also follow a scheme; for example, a negative answer to critical question 1 follows the scheme `Action $A$ will not realize goal $G$', which can be instantiated (e.g.
+`\texttt{Keep Same Cars} does not realize \texttt{Generate Cars'}) to forma counterargument to the original argument.
+
 Another way to criticize an argument for an action is to suggest an alternative action that realizes the same goal (question 2).
-For example, we can argue that performing a \texttt{Dynamic Simulation} also realizes the goal \texttt{Simulate Traffic} on its own.
-Finally, it is possible that performing an action has a negative side effect (critical question 3).
+For example, we can argue that performing a \texttt{Create New Cars} also realizes the goal \texttt{Generate Cars} on its own.
+Also, it is possible that performing an action has a negative side effect (critical question 3).
 For example, while the action \texttt{Create New Cars} realizes the goal \texttt{Generate Cars}, it has a negative side effect, namely hurting \texttt{Static Simulation}: having the simulation constantly create new cars is a functionality that does not allow for a static simulation.
-
-%FB: I extended the example and put the figure back in. Note that this is just to explain argumentation, not to reason about the model in 2.2. Putting explicit text in the example makes it more clear.
 
@@
 \begin{tikzpicture}
-        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.4) {\argtext{A1}{Action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} we should perform \texttt{Keep Same Cars}}};
-        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \texttt{Create New Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} we should perform \texttt{Create New Cars}}};
+        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.4) {\argtext{A1}{Action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} Action \texttt{Keep Same Cars} should be performed}};
+        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \texttt{Create New Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} Action  \texttt{Create New Cars} should be performed}};
         \node[minimum size=1cm] (att2) [argNodeIN] at (3.5,0) {\argtext{A3}{\texttt{Create New Cars} has a negative effect on \texttt{Static Simulation}}};
@@
 
-\subsection{Combining PRAS and GRL}
-\label{sect:background:pras:motivation}
-
-In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we have explored one way of combining PRAS and GRL.
-In this work, which takes a similar approach to \cite{Jureta:RE2008}, argument diagrams are translated to GRL models (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
-The argument diagrams are complex practical reasoning arguments, where the premises are the goals and the conclusion is some action we need to perform to realize those goals.
-Thus, we have essentially two complex diagrams, a practical reasoning argument diagram and a GRL goal diagram, and a mapping between them.
-
-Connecting argumentation and goal modelling by providing a mapping between two types of diagram was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
-One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion thus far has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
-Furthermore, in our previous work we did not provide the specific critical questions for goal models (see Section \ref{sect:gmas}).
-This meant that any counterargument had to be constructed from scratch, as no guidance was given as to possible ways to criticize a GRL model.
-So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged.
-Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling (requirement 4).
-
-In the current version of GRL, we do not literally take the Practical Reasoning Argument Scheme and its critical questions - this would be impossible given that there are elements of the GRL language, such as resources, which cannot be found in PRAS.
+Looking at PRAS and its critical questions, one can see how it could be used to argue about goals and actions or, more specifically, about goal models.
+However, we cannot literally use PRAS and its critical questions, as there are elements of the GRL language, such as actors and resources, which cannot be found in PRAS.
 Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson and Bench-Capon~\cite{atkinson2007} actually apply to GRL models.
-Therefore, we develop our own set of argument schemes and critical questions in Section~\ref{sect:gmas} by analyzing transcripts of discussions about the traffic simulator.
-Before doing so, we give an overview of our framework in the next section.
+In fact, our case study (Section \ref{sect:gmas}) shows that when discussing requirements, people very often do not structure their reasoning nicely in the way that PRAS presents it.
+That is, you do not see the discussants setting up an argument `We have goal $G$, $A$ realizes $G$ \emph{Therefore} we should perform $A$'.
+A typical discussion is much more unstructured, as is clear from the transcript excerpts in Appendix~\ref{sect:transcripts:excerpts}.
+So if we would use the version of PRAS presented in this section for our argumentation, we would violate requirement 1: The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
+Our solution was to develop our own set of argument schemes and critical questions by analyzing transcripts of discussions about the traffic simulator.
+This set of schemes and questions and our case study are described in the next section.

### 2017-09-11 discussion.tex
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!

@@
 High-level goals are often understood as reasons for representing lower-level goals (in other words, the need for low-level goals is justified by having high-level goals) and other elements in a goal model such as tasks and resources.
-Various refinements and decomposition techniques,  often used in requirements engineer (See~\cite{EP-2001-Lamsweerde-GORE} for an overview), can be seen as incorporating argumentation and justification, in that sub-goals could be understood as arguments supporting parent goals.
+Various refinements and decomposition techniques,  often used in requirements engineer (See~\cite{van2001goal} for an overview), can be seen as incorporating argumentation and justification, in that sub-goals could be understood as arguments supporting parent goals.
 In that case, a refinement alternative is justified if there are no conflicts between sub-goals (i.e., it is consistent), as few obstacles as possible sub-goals harm sub-goal achievement, there are no superfluous sub-goals (the refinement is minimal), and the achievement of sub-goals can be verified to lead to achieving the parent goal (if refinement is formal~\cite{Darimont:1996:FRP:239098.239131}).

### 2017-09-11 gmas.tex
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!

@@
 \section{Argument Schemes for Goal Modeling: a Case Study}
-%FB: I added "a case study" to emphasize that we perform a case study on the transcripts (see also the introduction)
 \label{sect:gmas}
 
-\begin{table*}[h]
-\centering
-\begin{tabularx}{\textwidth}{|l|l|l|X|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Effect}\\
-\hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant? & DISABLE\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & DISABLE\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & DISABLE\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & DISABLE\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?& DISABLE\\
-\hline
-\hline
-AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
-& & CQ5b & Does the goal decompose into any other tasks?& REPLACE\\
-\hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal?& INTRO \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?& INTRO\\
-&& CQ6d & Does the task contribute to some other softgoal?& INTRO\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?& DISABLE\\
-&& CQ7b & Does the goal contribute to some other softgoal?& INTRO\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?& DISABLE\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?& INTRO\\
-\hline
-AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& REPLACE\\
- &  & CQ10b & Is the decomposition type correct? (AND/OR/XOR)& REPLACE\\
-\hline
-AS11 & Task $T$ contributes negatively to softgoal $S$& CQ11 & Does the task contribute negatively to the softgoal?& DISABLE\\
-\hline
-\hline
-AS12 & Element $IE$ is relevant & CQ12 & Is the element relevant/useful? & DISABLE\\
-\hline
-AS13 & Element $IE$ has name $n$ & CQ13 & Is the name clear/unambiguous? & REPLACE\\
-\hline
-\hline
-- & - & Att & Generic counterargument & ATTACK\\
-\hline
-\end{tabularx}
-\caption{List of argument schemes (AS0-AS13, left column), critical questions (CQ0-CQ12, middle column), and the effect of answering them (right column).}
-\label{table:argument-schemes}
-\end{table*}
+The objective of our case study was to see which types of discourse are used during a discussion of system requirements, and how these discourse types can be captured as argument schemes and critical questions regarding goals and tasks.
+In order to study this, we manually coded transcripts of such discussions.
+For our coding, we used a list of argument schemes and critical questions based on GRL and PRAS.
+In this section we present our case study.
+All original transcripts, codings, and models are available in our online repository\footnote{\label{foot:url}\url{insertURL}}.
 
-In the previous section, we provided an overview of RationalGRL methodology.
-In this section, we discuss our argument schemes and critical questions in more detail.
-It is important to note that the list of argument schemes we present in this section is not exhaustive.
-It is an initial list that we have obtained by annotating transcripts.
-However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
-
-The list of argument schemes and critical questions that we have obtained from our analysis is shown in Table~\ref{table:argument-schemes}.
-The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are related to relationships, the next two (AS12-AS13) are for intentional elements in general, and the last (Att) is a generic counterargument against any type of argument that has been put forward.
-
-As we have already discussed in Section~\ref{sect:overview}, answering critical questions can have different impacts on the model.
-The right column in Table~\ref{table:argument-schemes} shows the impact of answering each critical question affirmatively.
-%SG: what does affirmatively mean?
-As mentioned earlier, answering a critical question can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
-
-The RationalGRL language is an extension of GRL.
-The prototype tool we developed, thus, contains various new elements and a new relationship.
-The legends of the these additions are shown in Figure~\ref{fig:rationalgrllegend}.
-
-In the following subsections, we first explain the new visual language in more detail, after which we provide details of the transcript annotation process with concrete examples (see Appendix~\ref{sect:transcripts:excerpts} for transcript excerpts).
-
-\subsection{RationalGRL Language}
-
-RationalGRL is an extension of GRL, and adds the following new elements (Figure~\ref{fig:rationalgrllegend}):
-\begin{itemize}
-\item \emph{Argument}: This represents an argument created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
-%SG: I thought arguments can also do replace, or intro GRL elements.
-So why in this description we are very limiting?
-\item \emph{Disabled GRL element}: If a GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be disabled, meaning that it does not play a role in the analysis of the GRL model.
-\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
-It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
-In these cases, the corresponding GRL element is shown with a striped background.
-If the user clicks on this element, a details pane shows up containing the history of refinements (see Section~\ref{sect:tool}).
-%SG: Can we have a separate element for the intro?
-or at least explain it in this text more explicitly?
-\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument attacks the target GRL element or argument.
-\end{itemize} %SG: what about the links? The links can also get attacked right? or have some arguments? what do we do with those?
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.35\textwidth]{img/legend}
-\caption{The new elements and link of RationalGRL}
-\label{fig:rationalgrllegend}
-\end{figure}
-
-\subsection{Case Study Details}
-
-In order to come up with the list of argument schemes and critical questions in Table~\ref{table:argument-schemes}, we performed an extensive case study based on a recent series of experiments by Schriek et al.
+In order to obtain actual requirements discussions, we turned to a recent series of experiments by Schriek et al.
 \cite{SchriekEtal2016}.
@@
 These groups had a maximum of two hours to design a traffic simulator, which included a discussion of the requirements of this traffic simulator.
+The students did not use any goal modeling technique in the course or during the discussions.
 The students were asked to record their design sessions, and the recordings were subsequently transcribed.
-We used three of these transcripts totalling 153 pages of discussion for our case study.
-In the case study, we annotated the transcripts with argument schemes and critical questions, and we created RationalGRL models based on the discussions.
-In this section we present examples of these annotations and models.
-All original transcripts, annotations, and models are available online\footnote{\url{insertURL}}.
+We used three of these transcripts, totaling 153 pages, for our case study.
 
-To develop the list of argument schemes and critical questions, we started with an initial list of 8 argument schemes and 18 critical questions that we derived from the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}.
-We then annotated the three transcripts with arguments and critical questions from this initial list.
-If there were arguments or critical questions that did not appear in the initial list, we added them to the list.
-Argument schemes and critical questions that did not appear were removed from the list.
-Appendix~\ref{sect:transcripts:excerpts} contains some examples of annotations.
+Before we started coding the transcripts, we drew up an initial list of 10 argument schemes (AS0-AS9 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
+AS0 to AS4 are schemes that concern a single element of the goal model.
+So, for example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'.
+AS5 to AS9 are claims about the links between GRL elements.
+Our initial list also contained 18 critical questions, inspired by the questions associated with the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}.
+CQ2 to CQ6b (Table~\ref{table:transcripts:results:argumentschemes}) are examples of these critical questions, other examples are `Is the softgoal legitimate?' and `Are there alternative ways to contribute to the same softgoal?'.
 
-\subsection{Case Study Results}
-
-We found a total of 159 instantiation of argument schemes AS0-AS11 in transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
-The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice .
-A large portion (about 60\%) of argument schemes involved discussions about tasks of the information system (AS2, AS10).
-We annotated 41 applications of critical questions.
-Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, CQ13).
-
-\begin{table}[ht]
+\begin{table}[t]
 \centering
@@
 \hline
-CQ5b & Goes decomposes into other tasks? & 1 & 0 & 0 & 1\\
+CQ5b & Goal decomposes into other tasks? & 1 & 0 & 0 & 1\\
 \hline
@@
 \end{tabularx}
-\caption{Number of Occurrences of AS0-AS9, CQ0-CQ12 in Transcripts. Critical Questions not Appearing in This Table were not Found in Transcripts.}
+\caption{Occurrences of argument schemes and critical questions in the transcripts.}
 \label{table:transcripts:results:argumentschemes}
@@
 
-We manually created RationalGRL models from argument schemes and critical questions found in each transcript, to verify whether the arguments put forward by the participants were sufficiently informative.
-An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
-This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
+Using the initial list of arguments and critical questions, we coded three transcripts of requirements discussions.
+The codings were performed by one author and subsequently checked by another author.
+As the transcripts contain spoken language, the codings involved some interpretation.
+For example, the students almost never literally say `actor $a$ has task $T$'.
+Rather, they say things such as `...we have a set of actions.
+Save map, open map, ...' (Table~\ref{table:transcripts:traffic-light}, Appendix~\ref{sect:transcripts:excerpts}) and `...
+in that process there are activities like create a visual map, create a road' (Table~\ref{table:transcript:task-clarification}, Appendix~\ref{sect:transcripts:excerpts}).
+Furthermore, in some cases the critical questions are explicit.
+For example, CQ10b is found in the transcripts as `...is this an OR or an AND?' (Table \ref{table:transcript:decomposition}, Appendix~\ref{sect:transcripts:excerpts}).
+In other cases, however, the question remains implicit but we added it in the coding.
+For example, CQ12 is not found in the transcripts, but the related counterargument is:  `...you don't have to specifically add a traffic light' (Table~\ref{table:transcripts:traffic-light}, Appendix~\ref{sect:transcripts:excerpts}).
 
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{img/Fig6}
-\caption{The GRL model manually constructed from transcript $t_1$.
-Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
-Elements and relationships with no dot have been inferred by us.} %TODO We need to update the caption
-\label{fig:transcripts:grl}
-\end{figure*}
+In order to further verify the coherence of our codings, we constructed complete RationalGRL models based on the argument schemes and critical questions found in each transcript.
+An example model is shown in Figure~\ref{fig:transcripts:grl}, and the full models can be found in the repository.
+More examples of RationalGRL models based on our codings will be given in section \ref{sect:overview}.
 
-Answering a critical question can have four different effects on the original argument and the corresponding GRL element:
-\begin{itemize} 
-\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
-This operation does not attack the original argument of the critical question, but rather creates a new argument.
-In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
-For instance, the XOR-decomposition from ``Generate Cars'' is an instantiation of AS5 as follows: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
-Suppose now that the modelers that created Figure~\ref{fig:transcripts:grl} would continue their analysis by discussing critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'', and that they would answer this with ``Yes, namely \texttt{Choose randomly}''.
-This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' %SG: Very good example.
-But should we also show it in the model as a new element?
-\item \textsf{DISABLE:} Disable the element or relationship of an argument scheme to which critical questions pertains.
-This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
-In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
-The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
-What we also see from Figure~\ref{fig:transcripts:grl} is that actor ``Teacher'' is disabled and thus all the elements that are bound to this actor are disabled as well.
-Furthermore, disabled task ``Dynamic Simulation'' also disables all incoming and outgoing links with this task.
-
-\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
-In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
-In this case, participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
-While the previous names are not shown in the figure, they show up in the details pane of the corresponding element.
-%SG: why generate car has a star but not this one?
-\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-In Figure~\ref{fig:transcripts:grl}, we see one example of such a counter-attack.
-First, task ``Control car influx per road'' is attack by answering CQ2 (irrelevant task).
-However, after discussing this, participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
-Therefore, the argument that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
-We provide a precise semantics for this in Figure~\ref{sect:formalframework}.
-\end{itemize}
+During the coding, the list of argument schemes and critical questions was updated: some schemes were removed or amended, and new ones were added.
+For example, we found that the discussants often talk about tasks decomposing into sub-tasks, so we added AS10 and CQ10a.
+Furthermore, because we noticed that there were many discussions on both the relevance and the clarity of the names of elements (goals, tasks, etc.), two generic critical questions CQ12 and CQ13 were added.
+The final results regarding the argument schemes and the critical questions found in the transcripts can be found in Table~\ref{table:transcripts:results:argumentschemes}; for results per transcript, please consult the online repository.
+We found a total of 159 instantiation of argument schemes AS0-AS11 in transcripts.
+The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice.
+A large portion (about 60\%) of argument schemes involved discussions about tasks of the information system (AS2, AS10).
+We coded 41 applications of critical questions.
+Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, CQ13).
 

### 2017-09-11 introduction.tex
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!

@@
 \item 
-Using these traceability links, it must be possible to compute the effect of changes in the goal model on the underlying arguments, and vice versa.
-\todo{F}{M,S}{Given the current logical framework, is it possible to compute effect of changes in GM on ARG?
-It seems to me that we can only do the other way around with the algorithms.
-(M): In a sense this is possible.
-In our formalization, each GRL element and link can be translated into an instantiation of an argument scheme, so adding GRL element simply means adding more instantiate argument schemes, who can then be attack by critical questions.
-But I think it is a good idea to make this more explicit.}
+Using these traceability links, it must be possible to compute the effect of changes in the underlying argumentation on the goal model, and vice versa.
 \item 
@@
 
-One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for capturing multi-agent planning \cite{medellin2013planning}, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.
+One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for, among other things, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.
 Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyse and guide stakeholders' discussions about goal models.

### 2017-09-11 overview.tex
Small updates sect. 6
Added definition, simplified formal argument example.
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!

@@
 
+Answering a critical question can have four different effects on the original argument and the corresponding GRL element:
+\begin{itemize} 
+\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
+This operation does not attack the original argument of the critical question, but rather creates a new argument.
+In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
+For instance, the XOR-decomposition from ``Generate Cars'' is an instantiation of AS5 as follows: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
+Suppose now that the modelers that created Figure~\ref{fig:transcripts:grl} would continue their analysis by discussing critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'', and that they would answer this with ``Yes, namely \texttt{Choose randomly}''.
+This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' %SG: Very good example.
+But should we also show it in the model as a new element?
+\item \textsf{DISABLE:} Disable the element or relationship of an argument scheme to which critical questions pertains.
+This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
+In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
+The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
+What we also see from Figure~\ref{fig:transcripts:grl} is that actor ``Teacher'' is disabled and thus all the elements that are bound to this actor are disabled as well.
+Furthermore, disabled task ``Dynamic Simulation'' also disables all incoming and outgoing links with this task.
+
+\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
+In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
+In this case, participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
+While the previous names are not shown in the figure, they show up in the details pane of the corresponding element.
+%SG: why generate car has a star but not this one?
+\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
+In Figure~\ref{fig:transcripts:grl}, we see one example of such a counter-attack.
+First, task ``Control car influx per road'' is attack by answering CQ2 (irrelevant task).
+However, after discussing this, participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
+Therefore, the argument that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
+We provide a precise semantics for this in Figure~\ref{sect:formalframework}.
+\end{itemize}
+
+The list of argument schemes and critical questions that we have obtained from our analysis is shown in Table~\ref{table:argument-schemes}.
+The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are related to relationships, the next two (AS12-AS13) are for intentional elements in general, and the last (Att) is a generic counterargument against any type of argument that has been put forward.
+
+It is important to note that the list of argument schemes we present in this section is not exhaustive.
+It is an initial list that we have obtained by coding transcripts.
+However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
+
+As we have already discussed in Section~\ref{sect:overview}, answering critical questions can have different impacts on the model.
+The right column in Table~\ref{table:argument-schemes} shows the impact of answering each critical question affirmatively.
+%SG: what does affirmatively mean?
+As mentioned earlier, answering a critical question can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
+
+\begin{table*}[h]
+\centering
+\begin{tabularx}{\textwidth}{|l|l|l|X|l|l|}
+\hline
+\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Effect}\\
+\hline
+AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant? & DISABLE\\
+\hline
+AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & DISABLE\\
+\hline
+AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & DISABLE\\
+\hline
+AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & DISABLE\\
+\hline
+AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?& DISABLE\\
+\hline
+\hline
+AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
+& & CQ5b & Does the goal decompose into any other tasks?& REPLACE\\
+\hline
+AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
+&& CQ6b & Are there alternative ways of contributing to the same softgoal?& INTRO \\
+&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?& INTRO\\
+&& CQ6d & Does the task contribute to some other softgoal?& INTRO\\
+\hline
+AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?& DISABLE\\
+&& CQ7b & Does the goal contribute to some other softgoal?& INTRO\\
+\hline
+AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?& DISABLE\\
+\hline
+AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?& INTRO\\
+\hline
+AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& REPLACE\\
+ &  & CQ10b & Is the decomposition type correct? (AND/OR/XOR)& REPLACE\\
+\hline
+AS11 & Task $T$ contributes negatively to softgoal $S$& CQ11 & Does the task contribute negatively to the softgoal?& DISABLE\\
+\hline
+\hline
+AS12 & Element $IE$ is relevant & CQ12 & Is the element relevant/useful? & DISABLE\\
+\hline
+AS13 & Element $IE$ has name $n$ & CQ13 & Is the name clear/unambiguous? & REPLACE\\
+\hline
+\hline
+- & - & Att & Generic counterargument & ATTACK\\
+\hline
+\end{tabularx}
+\caption{List of argument schemes (AS0-AS13, left column), critical questions (CQ0-CQ12, middle column), and the effect of answering them (right column).}
+\label{table:argument-schemes}
+\end{table*}
+
 \subsection{Metamodel}
+\label{sect:metamodel}
 
@@
 \end{figure*}
+
+\subsection{RationalGRL Language}
+
+The RationalGRL language is an extension of GRL.
+The prototype tool we developed, thus, contains various new elements and a new relationship.
+The legends of the these additions are shown in Figure~\ref{fig:rationalgrllegend}.
+
+In the following subsections, we first explain the new visual language in more detail
+
+RationalGRL is an extension of GRL, and adds the following new elements (Figure~\ref{fig:rationalgrllegend}):
+\begin{itemize}
+\item \emph{Argument}: This represents an argument created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
+%SG: I thought arguments can also do replace, or intro GRL elements.
+So why in this description we are very limiting?
+\item \emph{Disabled GRL element}: If a GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be disabled, meaning that it does not play a role in the analysis of the GRL model.
+\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
+It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
+In these cases, the corresponding GRL element is shown with a striped background.
+If the user clicks on this element, a details pane shows up containing the history of refinements (see Section~\ref{sect:tool}).
+%SG: Can we have a separate element for the intro?
+or at least explain it in this text more explicitly?
+\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument attacks the target GRL element or argument.
+\end{itemize} %SG: what about the links? The links can also get attacked right? or have some arguments? what do we do with those?
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=0.35\textwidth]{img/legend}
+\caption{The new elements and link of RationalGRL}
+\label{fig:rationalgrllegend}
+\end{figure}
+
+We manually created RationalGRL models from argument schemes and critical questions found in each transcript, to verify whether the arguments put forward by the participants were sufficiently informative.
+An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
+This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
+
+\begin{figure*}[h!]
+\includegraphics[width=\textwidth]{img/Fig6}
+\caption{The GRL model manually constructed from transcript $t_1$.
+Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
+Elements and relationships with no dot have been inferred by us.} %TODO We need to update the caption
+\label{fig:transcripts:grl}
+\end{figure*}

### 2017-09-11 transcripts.tex
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!

--- 2017-01-21 transcripts.tex
@@
 \label{sect:transcripts:excerpts}
+
+\begin{table}[!htbp]
+\centering
+\begin{tabular}{|p{17mm}|p{63mm}|p{70mm}|}
+\hline
+\textit{Speaker} & \textit{Text} & \textit{Annotation}\\
+\hline
+(P1) & And then, we have a set of actions.
+Save map, open map, add and remove intersection, roads & \multirow{2}{70mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map", ``add intersection", ``remove intersection", ``add road", ``add traffic light" \textsf{[INTRO]}}\\
+\cline{1-2}
+(P2) & Yeah, road. Intersection, add traffic lights &\\
+\hline
+(P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{70mm}{\textbf{[21 critical question CQ12 for 20]} Is the task ``Add traffic light" useful/relevant?\newline
+\textbf{[22 answer to 21]} Not useful, because according to the specification all intersections have traffic lights. \textsf{[ATTACK]}}\\
+\cline{1-2}
+(P2) & Yeah &\\
+\cline{1-2}
+(P1) & It's, you don't have to specifically add a traffic light because if you have &\\
+\cline{1-2}
+(P2)	& They need-&\\
+\hline
+\end{tabular}
+\caption{Adding tasks, disabling useless task ``Add traffic light'' (transcript $t_1$)}
+\label{table:transcripts:traffic-light}
+
+\begin{tabular}{|p{17mm}|p{63mm}|p{70mm}|}
+\hline
+\textit{Speaker} & \textit{Text} & \textit{Annotation}\\
+\hline
+0:17:39 (P1) & And in that process there are activities like create a visual map, create a road& \textbf{[14 task (AS2)]} Student has task ``Create road'' \textsf{[INTRO]}\\
+\hline
+0:24:36.0 (P3)	& And, well interaction.
+Visualization sorry.
+Or interaction, I don't know.
+So create a visual map would have laying out roads and a pattern of their choosing.
+So this would be first, would be choose a pattern.
+& \textbf{[31 critical question CQ13 for 14]} Is Task ``Create road'' clear?\newline
+\textbf{[32a answer to 31]} no, according to the specification the student should choose a pattern. \newline
+\textbf{[32b answer to 31]} ``Create road'' becomes ``Choose a pattern'' \textsf{[REPLACE]}\\
+\hline
+0:24:55.4 (P1) &	How do you mean, choose a pattern	& \textbf{[33 critical question CQ13 for 32b]} Is ``Choose a pattern'' clear? \\
+\cline{1-2}
+0:24:57.5 (P3)	& Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing	&\\
+\cline{1-2}
+0:25:07.5 (P1)	& Yeah I'm not sure if they mean that. I don't know what they mean by pattern in this case. I thought you could just pick roads, varying sizes and like, broads of roads. & 
+\textbf{[34a answer to 33]} No, not sure what they mean by a pattern.\\
+\cline{1-2}
+0:25:26.0 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \multirow{3}{70mm}{\textbf{[34b answer to 33]} ``Choose a pattern'' becomes ``Choose a pattern preference'' \textsf{[REPLACE]}}\\
+\cline{1-2}
+0:25:36.3 (P1) & Yeah, selecting a- I don't know &\\
+\cline{1-2}
+0:25:38.0 (P3)	& Pattern preference maybe? As in, maybe we can explain this in the documentation &\\
+\hline
+0:25:43.9 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ13 for 34b]} Is ``Choose a pattern preference'' clear?\newline
+\textbf{[36a answer to 35]} no, what kind of pattern?\\
+\cline{1-2}
+0:25:47.4 (P3) & Maybe, I don't know it's- & \\
+\cline{1-2}
+0:25:48.5 (P1)	& [inaudible] a road pattern& \multirow{2}{70mm}{\textbf{[36b answer to 35]} ``Choose a pattern preference'' becomes ``Choose a road pattern'' \textsf{[REPLACE]}}\\
+\hline
+\end{tabular}
+\caption{Clarifying the name of a task (transcript $t_3$)}
+\label{table:transcript:task-clarification}
+
+\begin{tabular}{|p{17mm}|p{43mm}|p{90mm}|}
+\hline
+\textit{Speaker} & \textit{Text} & \textit{Annotation}\\
+\hline
+0:18:55.7 (P1) &Yeah. And then two processes, static, dynamic and they belong to the goal simulate. & \textbf{[17 goal (AS3)]} System has goal ``Simulate'' \textsf{[INTRO]}\newline
+\textbf{[18 task (AS2)]} System has tasks ``Static simulation'', ``Dynamic simulation'' \textsf{[INTRO]}\newline  
+\textbf{[20 decomposition (AS5)]} Goal ``Simulation" AND-decomposes into "Static simulation" and ``Dynamic simulation" \textsf{[INTRO]}\\
+\hline
+0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{5}{80mm}{\textbf{[26 critical question CQ10b for 20]} Is the decomposition type of ``simulate'' correct?\newline
+\textbf{[27 answer to 26]} No, it should be an OR decomposition. \textsf{[REPLACE]}}\\
+\cline{1-2}
+0:30:12.6 (P2) & That's and OR &\\
+\cline{1-2}
+0:30:14.3 (P3) & I think it's an OR &\\
+\cline{1-2}
+0:30:15.4 (P1) & It's for the data, it's an OR &\\
+\cline{1-2}
+0:30:18.1 (P3) & Yep &\\
+\hline	
+\end{tabular}
+\caption{Incorrect decomposition type for goal \emph{Simulate} (transcript $t_3$)}
+\label{table:transcript:decomposition}
+\end{table}
 
@@
 \hline
+0:18:13.4 (P2) & I think we can still do developers here.
+To the system & \multirow{4}{60mm}{\textbf{[16 counter argument for 6]} According to the specification the professor doesn't actually develop the software.}\\
+\cline{1-2}
+0:18:18.2 (P1)  & Yeah?&\\
+\cline{1-2}
+0:18:19.8 (P2) & Yeah, it isn't mentioned but, the professor does-&\\
+\cline{1-2}
+0:18:22.9 (P1) & Yeah, when the system gets stuck they also have to be [inaudible] ok. So development team&\\	
+\hline
 \end{tabular}
-\caption{AS0: Actor, CQ0: Relevant actor? (transcript $t_3$)}
-\label{table:transcript:as0-cq0}
+\caption{Discussion about the relevance of an actor (transcript $t_3$)}
+\label{table:transcript:irrelevant-actor}
+\end{table}
 
+\iffalse
+\begin{table}[!htbp]
 \begin{tabular}{|p{20mm}|p{70mm}|p{60mm}|}

### 2017-09-13 gmas.tex
Section 4 (RationalGRL Framework)

@@
 \hline
-AS6 & Task contributes to softgoal & 6 & 2 &0& 8\\
+AS6 & Task contributes (negatively) to softgoal & 8 & 3 &0& 11\\
 \hline
-AS7 & Goal contributes to softgoal &0& 1 & 1 & 2\\
+AS7 & Goal contributes (negatively) to softgoal &0& 1 & 1 & 2\\
 \hline
@@
 AS10 & Task decomposes into tasks & 11 &14 &11 &36\\ 
-\hline
-AS11 & Task contributes negatively to softgoal & 2 & 1 & 0 & 3\\
 \hline
@@
 \hline
-CQ13 & Is the name clear/unambiguous? &3 &10 & 3 & 16\\
+CQ13 & Is the element clear/unambiguous? &3 &10 & 3 & 16\\
 \hline
@@
 \end{tabularx}
-\caption{Occurrences of argument schemes and critical questions in the transcripts.}
+\caption{Occurrences of argument schemes and critical questions in the transcripts.\todo{F}{F}{check numbers against original annotation docs}}
 \label{table:transcripts:results:argumentschemes}
@@
 
-During the coding, the list of argument schemes and critical questions was updated: some schemes were removed or amended, and new ones were added.
+During the coding, new argument schemes and critical questions were added to the list.
 For example, we found that the discussants often talk about tasks decomposing into sub-tasks, so we added AS10 and CQ10a.
-Furthermore, because we noticed that there were many discussions on both the relevance and the clarity of the names of elements (goals, tasks, etc.), two generic critical questions CQ12 and CQ13 were added.
-The final results regarding the argument schemes and the critical questions found in the transcripts can be found in Table~\ref{table:transcripts:results:argumentschemes}; for results per transcript, please consult the online repository.
+Furthermore, because there were many discussions on the relevance and the clarity of the names of elements (goals, tasks, etc.), two generic critical questions CQ12 and CQ13 were added.
+The final results of the coding can be found in Table~\ref{table:transcripts:results:argumentschemes}; for results per transcript, please consult the online repository.
 We found a total of 159 instantiation of argument schemes AS0-AS11 in transcripts.

### 2017-09-13 overview.tex
Section 4 (RationalGRL Framework)

@@
-\section{RationalGRL Methodology and Metamodel}
+\section{The RationalGRL Framework}
 \label{sect:overview}
 
-In this section we present a high-level overview of the RationalGRL framework.
-In the first subsection, we present a methodology, specifying how practitioners can use RationalGRL to create goal models with traceability links to underlying arguments.
-In the second subsection, we link the new argumentation elements and relations to the existing elements and relations of GRL in a metamodel.
-This metamodel serves as the specification for an implementation.
-In Section~\ref{sect:tool} we briefly discuss our prototype implementation based on this metamodel.
+In this section we present our RationalGRL framework, which allows stakeholders to discuss, analyse and argue about goal models.
+The RationalGRL framework includes two main parts: Argumentation and GRL goal modeling.
 
-\subsection{RationalGRL Methodology}
+The GRL part of RationalGRL allows for the creation of GRL goal models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
+For the argumentation part, arguments and counterarguments can be put forward about various parts of this goal model.
+These two parts, GRL and argumentation, are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
+For example, answering a critical question \emph{Is the task \texttt{A} possible?} can result in removing or adding a task in the GRL model.
+Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
 
-As mentioned in Section~\ref{sect:introduction}, the RationalGRL framework uses concepts from practical reasoning argument scheme (PRAS) to help integrating goal models with the detailed discussions and arguments the stakeholders pose during the analysis phase.
-That is, the RationalGRL framework includes two main parts: Argumentation modeling and GRL modeling.
-
-For the GRL part, we first need to create the ``initial'' GRL model by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
-For the argumentation part, arguments and counterarguments are put forward about various parts of the goal model.
-These two parts, GRL and argumentation, are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
-For example, answering a critical question \emph{Is the task \texttt{A} possible?}, instantiated from the argumentation model, can result in removing or adding a task in the GRL model.
-Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
-Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework with its components and their relationships.
-The GRL model is shown  on the right-hand side of the framework while the argumentation model is on the left-hand side.
-The links between the two sides illustrate the impacts and relationships between two sides.
-Note that answer to critical questions and argument schemes that are instantiated during the analysis phase of the GRL model are documented with the GRL model and can be referred to in the future.
+Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework.
+At the bottom there are two activities: \emph{practical reasoning \& argumentation} and \emph{goal model construction}.
+As already explained, these two activities influence each other: adding elements to a goal model gives rise to new critical questions about these elements, and answering critical questions with arguments may add or delete elements from the GRL model.
+The activities thus provide two models: a RationalGRL argumentation model (left-hand side), which includes arguments for and against goal model elements, and a GRL goal model (right-hand side).
+The two types of models are connected by traceability links.
+In this way, it becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
 
@@
 
-We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework.
-Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
-The rest of the steps are as follows:
+In the rest of this section, we discuss the individual parts of the GRL framework.
+First we discuss the argument schemes and critical questions that form the basis of our framework (Section~\ref{sect:overview:as}.
+After this, we explain the RationalGRL language and metamodel (Section~\ref{sect:metamodel}).
+\todo{F}{all}{finish once paper done}
 
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology\todo{F}{M,S}{Can you make this a vector image?}}
-\label{fig:rationalgrl-methodology}
-\end{figure*}
+\subsection{RationalGRL Argument Schemes and Critical Questions}
+\label{sect:overview:as}
 
-\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes of the argumentation framework.
-We select an intentional element from the initial GRL model that we want to analyze and we instantiate a relevant argument scheme from the already existing list of argument schemes or by adding a new one.
-For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}".
-When an argument scheme is instantiated, it corresponds to  an argument for or against part of a goal model.
-
-\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions to attack the argument with counter-arguments.
-Each argument scheme includes one or more critical questions.
-For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contribute to the softgoal?} and \emph{Does the goal contribute to some other softgoals?}.
-%It is worth mentioning that, answering a critical question may result in a "conflict" situation. This is out of the scope of our work here.  
-%MvZ: removed this sentence because I don't know what you mean with "conflict". I think We should explain it better or leave it out.
-When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and step (1).
-
-\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the four following cases can occur: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.
-Any of these cases can  impact the arguments and corresponding GRL intentional elements.
-\textsf{INTRO} means that
-a new argument scheme is created.
-That means, the current argument scheme related to the critical question does not get attacked.
-In the case of \textsf{DISABLE}, the intentional element or its related links are disabled or removed from the models.
-\textsf{REPLACE} introduces a new argument and attacks the original argument at the same time.
-This means that the original element of the argument scheme is replaced with a new one.
-\textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when new evidence occurs.
-
-\textbf{(4) Modify GRL Models} -- In this step, we modify the GRL models based on the situation of step (3).
-That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
-This results in a new modified GRL.
-The new GRL model can then impact the argument schemes and instantiate another argument scheme (Step (1)).
-
-We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
-
-Answering a critical question can have four different effects on the original argument and the corresponding GRL element:
-\begin{itemize} 
-\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
-This operation does not attack the original argument of the critical question, but rather creates a new argument.
-In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
-For instance, the XOR-decomposition from ``Generate Cars'' is an instantiation of AS5 as follows: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
-Suppose now that the modelers that created Figure~\ref{fig:transcripts:grl} would continue their analysis by discussing critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'', and that they would answer this with ``Yes, namely \texttt{Choose randomly}''.
-This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' %SG: Very good example.
-But should we also show it in the model as a new element?
-\item \textsf{DISABLE:} Disable the element or relationship of an argument scheme to which critical questions pertains.
-This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
-In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
-The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
-What we also see from Figure~\ref{fig:transcripts:grl} is that actor ``Teacher'' is disabled and thus all the elements that are bound to this actor are disabled as well.
-Furthermore, disabled task ``Dynamic Simulation'' also disables all incoming and outgoing links with this task.
-
-\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
-In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
-In this case, participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
-While the previous names are not shown in the figure, they show up in the details pane of the corresponding element.
-%SG: why generate car has a star but not this one?
-\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-In Figure~\ref{fig:transcripts:grl}, we see one example of such a counter-attack.
-First, task ``Control car influx per road'' is attack by answering CQ2 (irrelevant task).
-However, after discussing this, participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
-Therefore, the argument that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
-We provide a precise semantics for this in Figure~\ref{sect:formalframework}.
-\end{itemize}
-
-The list of argument schemes and critical questions that we have obtained from our analysis is shown in Table~\ref{table:argument-schemes}.
-The first four argument schemes (AS0-AS4) are arguments for an element of a goal model, the next seven (AS5-AS11) are related to relationships, the next two (AS12-AS13) are for intentional elements in general, and the last (Att) is a generic counterargument against any type of argument that has been put forward.
-
-It is important to note that the list of argument schemes we present in this section is not exhaustive.
-It is an initial list that we have obtained by coding transcripts.
-However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
-
-As we have already discussed in Section~\ref{sect:overview}, answering critical questions can have different impacts on the model.
-The right column in Table~\ref{table:argument-schemes} shows the impact of answering each critical question affirmatively.
-%SG: what does affirmatively mean?
-As mentioned earlier, answering a critical question can create an argument disabling the corresponding GRL element of the attacked argument scheme (\textsf{DISABLE}); it can create an argument introducing a new GRL element (\textsf{INTRO}); it can replace the GRL element corresponding to the original argument (\textsf{REPLACE}), or it can simply attack an argument directly (\textsf{ATTACK}).
-
-\begin{table*}[h]
+\begin{table*}[t]
 \centering
@@
 \end{tabularx}
-\caption{List of argument schemes (AS0-AS13, left column), critical questions (CQ0-CQ12, middle column), and the effect of answering them (right column).}
+\caption{List of argument schemes (AS0-AS13), critical questions (CQ0-CQ12), and the effect of answering them (right column).}
 \label{table:argument-schemes}
@@
 
-\subsection{Metamodel}
+A core aspect of the RationalGRL framework are the argument schemes that allow us to talk about the goal models.
+Recall from section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
+Using this list as a basis, we further refined our set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
+
+The argument schemes AS0-AS4 and AS12-AS13 are arguments for an element of a goal model, and AS5-AS11 are related to links in a goal model.
+The last scheme (Att) is a scheme for a generic counterargument against any type of argument that has been put forward.
+
+The critical questions presented in Table~\ref{table:argument-schemes} are related to their respective argument schemes.
+Answering a critical question in a certain way can have four different effects on the original argument and the corresponding GRL element:
+
+\begin{itemize} 
+\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
+This operation does not attack the original argument to which the question pertains, but rather creates a new argument.
+In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
+For instance, the XOR-decomposition from ``Generate Cars'' is an instantiation of AS5 as follows: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
+Suppose now that the modelers that created Figure~\ref{fig:transcripts:grl} would continue their analysis by discussing critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'', and that they would answer this with ``Yes, namely \texttt{Choose randomly}''.
+This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' %SG: Very good example.
+But should we also show it in the model as a new element?
+\item \textsf{DISABLE:} Disable the element or relationship of an argument scheme to which critical questions pertains.
+This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
+In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
+The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
+What we also see from Figure~\ref{fig:transcripts:grl} is that actor ``Teacher'' is disabled and thus all the elements that are bound to this actor are disabled as well.
+Furthermore, disabled task ``Dynamic Simulation'' also disables all incoming and outgoing links with this task.
+
+\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
+In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
+In this case, participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
+While the previous names are not shown in the figure, they show up in the details pane of the corresponding element.
+%SG: why generate car has a star but not this one?
+\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
+In Figure~\ref{fig:transcripts:grl}, we see one example of such a counter-attack.
+First, task ``Control car influx per road'' is attack by answering CQ2 (irrelevant task).
+However, after discussing this, participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
+Therefore, the argument that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
+We provide a precise semantics for this in Section~\ref{sect:formalframework}.
+\end{itemize}
+
+It is important to note that the list of argument schemes and critical questions we present in this section is not exhaustive.
+It is an initial list that we have obtained by coding transcripts.
+However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
+
+\subsection{RationalGRL Metamodel}
 \label{sect:metamodel}
@@
 
-The GRL package of our metamodel consists of the core GRL concepts, which constitute a part of
-the URN metamodel from Recommendation Z.151~\cite{URN}.
+The GRL package of our metamodel consists of the core GRL concepts, which constitute a part of the URN metamodel from Recommendation Z.151~\cite{URN}.
 These concepts represent the abstract grammar of the language, independently of the notation.
-This metamodel also formalizes the GRL concepts and constructs introduced earlier\footnote{Note that for readability, some GRL concepts ave been omitted from the figure.
-For instance, a GRL contribution can have a qualitative strength, ranging from ``Make'' to ``Break''.
-Since these concepts are not relevant to our framework, they have been omitted.}.
+This metamodel also further defines the GRL concepts and constructs introduced earlier\footnote{Note that for readability, some GRL concepts, such as contribution strength, have been omitted from the figure.}.
 
@@
 This allows the user to create attacks between arguments, which do not necessarily correspond to one of the critical questions.
-A \textsf{RationalGRLDiagram} is composed out of arguments and attack relations.
+A \textsf{RationalGRLmodel} is composed out of arguments and attack relations.
 
@@
 The link specifies that each \textsf{GRLModelElement} is in fact an argument.
-This means that each model element inherits the \textsf{AcceptStatus} as well, allowing GRL elements to be accepted, rejected, or undecided.
+This means that each model element inherits the \textsf{AcceptStatus} as well, allowing GRL elements to be accepted or rejected.
 This, furthermore, means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme.
@@
 \includegraphics[width=\textwidth]{metamodel/metamodel}
-\caption{The RationalGRL metamodel}
+\caption{The RationalGRL metamodel \todo{F}{M}{Change RationalGRLDiagram to RationalGRLmodel}}
 \label{fig:metamodel}
@@
 
+As an example of a RationalGRL model, consider Figure~\ref{fig:transcripts:grl}, which was created on the basis of transcript $t_1$.
+This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
+
+\begin{figure*}[h!]
+\includegraphics[width=\textwidth]{img/Fig6}
+\caption{The GRL model constructed from transcript $t_1$.} 
+\label{fig:transcripts:grl}
+\end{figure*}
+
+%I wasn't really sure what to do with the below section. 
+\iffalse
 \subsection{RationalGRL Language}
+The RationalGRL language is an extension of GRL, which means that it includes all the GRL elements shown in Figure~\ref{fig:grl_legend}.
+However, there are also new elements (Figure~\ref{fig:rationalgrllegend}):
 
-The RationalGRL language is an extension of GRL.
-The prototype tool we developed, thus, contains various new elements and a new relationship.
-The legends of the these additions are shown in Figure~\ref{fig:rationalgrllegend}.
-
-In the following subsections, we first explain the new visual language in more detail
-
-RationalGRL is an extension of GRL, and adds the following new elements (Figure~\ref{fig:rationalgrllegend}):
 \begin{itemize}
-\item \emph{Argument}: This represents an argument created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
-%SG: I thought arguments can also do replace, or intro GRL elements.
-So why in this description we are very limiting?
-\item \emph{Disabled GRL element}: If a GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be disabled, meaning that it does not play a role in the analysis of the GRL model.
+\item \emph{Argument}: This represents an argument that does not correspond to a GRL element.
+created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
+\item \emph{Rejected GRL element}: If an argument or GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be rejected, meaning that it does not play a role in the analysis of the GRL model.
+The status of arguments and
 \item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
@@
 In these cases, the corresponding GRL element is shown with a striped background.
-If the user clicks on this element, a details pane shows up containing the history of refinements (see Section~\ref{sect:tool}).
-%SG: Can we have a separate element for the intro?
-or at least explain it in this text more explicitly?
-\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument attacks the target GRL element or argument.
-\end{itemize} %SG: what about the links? The links can also get attacked right? or have some arguments? what do we do with those?
+\todo{F}{S,M}{Disabled and Refined GRL elements are not in the metamodel}
+\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument or element attacks the target argument or element.
+\end{itemize}
 
@@
 \end{figure}
+\fi
 
-We manually created RationalGRL models from argument schemes and critical questions found in each transcript, to verify whether the arguments put forward by the participants were sufficiently informative.
-An example of such a model is shown in Figure~\ref{fig:transcripts:grl}.
-This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
+\subsection{RationalGRL Methodology}
 
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{img/Fig6}
-\caption{The GRL model manually constructed from transcript $t_1$.
-Green dots indicate accepted underlying arguments, red dots indicate rejected underlying arguments.
-Elements and relationships with no dot have been inferred by us.} %TODO We need to update the caption
-\label{fig:transcripts:grl}
+We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework.
+Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+The rest of the steps are as follows:
+
+\begin{figure*}[ht]
+\centering
+\includegraphics[scale=0.4]{img/methodology}
+\caption{The RationalGRL Methodology\todo{F}{M,S}{Can you make this a vector image?}}
+\label{fig:rationalgrl-methodology}
 \end{figure*}
+
+\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes of the argumentation framework.
+We select an intentional element from the initial GRL model that we want to analyze and we instantiate a relevant argument scheme from the already existing list of argument schemes or by adding a new one.
+For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}".
+When an argument scheme is instantiated, it corresponds to  an argument for or against part of a goal model.
+
+\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions to attack the argument with counter-arguments.
+Each argument scheme includes one or more critical questions.
+For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contribute to the softgoal?} and \emph{Does the goal contribute to some other softgoals?}.
+%It is worth mentioning that, answering a critical question may result in a "conflict" situation. This is out of the scope of our work here.  
+%MvZ: removed this sentence because I don't know what you mean with "conflict". I think We should explain it better or leave it out.
+When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and step (1).
+
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the four following cases can occur: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.
+Any of these cases can  impact the arguments and corresponding GRL intentional elements.
+\textsf{INTRO} means that
+a new argument scheme is created.
+That means, the current argument scheme related to the critical question does not get attacked.
+In the case of \textsf{DISABLE}, the intentional element or its related links are disabled or removed from the models.
+\textsf{REPLACE} introduces a new argument and attacks the original argument at the same time.
+This means that the original element of the argument scheme is replaced with a new one.
+\textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when new evidence occurs.
+
+\textbf{(4) Modify GRL Models} -- In this step, we modify the GRL models based on the situation of step (3).
+That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
+This results in a new modified GRL.
+The new GRL model can then impact the argument schemes and instantiate another argument scheme (Step (1)).
+
+We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.

### 2017-09-13 transcripts.tex
Section 4 (RationalGRL Framework)

@@
 \hline
-\textit{Speaker} & \textit{Text} & \textit{Annotation}\\
+\textit{Speaker} & \textit{Text} & \textit{Coding}\\
 \hline
-(P1) & And then, we have a set of actions.
+0:15:11 (P1) & And then, we have a set of actions.
 Save map, open map, add and remove intersection, roads & \multirow{2}{70mm}{\textbf{[20 task (AS2)]} Student has tasks ``save map'', ``open map", ``add intersection", ``remove intersection", ``add road", ``add traffic light" \textsf{[INTRO]}}\\
 \cline{1-2}
-(P2) & Yeah, road. Intersection, add traffic lights &\\
+0:15:34 (P2) & Yeah, road. Intersection, add traffic lights &\\
 \hline
-(P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{70mm}{\textbf{[21 critical question CQ12 for 20]} Is the task ``Add traffic light" useful/relevant?\newline
+0:15:42 (P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{70mm}{\textbf{[21 critical question CQ12 for 20]} Is the task ``Add traffic light" useful/relevant?\newline
 \textbf{[22 answer to 21]} Not useful, because according to the specification all intersections have traffic lights. \textsf{[ATTACK]}}\\
 \cline{1-2}
-(P2) & Yeah &\\
+0:15:44 (P2) & Yeah &\\
 \cline{1-2}
-(P1) & It's, you don't have to specifically add a traffic light because if you have &\\
+0:15:45 (P1) & It's, you don't have to specifically add a traffic light because if you have &\\
 \cline{1-2}
@@
 \hline
-\textit{Speaker} & \textit{Text} & \textit{Annotation}\\
+\textit{Speaker} & \textit{Text} & \textit{Coding}\\
 \hline
@@
 \hline
-0:24:36.0 (P3)	& And, well interaction.
+0:24:36 (P3)	& And, well interaction.
 Visualization sorry.
@@
 \hline
-0:24:55.4 (P1) &	How do you mean, choose a pattern	& \textbf{[33 critical question CQ13 for 32b]} Is ``Choose a pattern'' clear? \\
+0:24:55 (P1) &	How do you mean, choose a pattern	& \textbf{[33 critical question CQ13 for 32b]} Is ``Choose a pattern'' clear? \\
 \cline{1-2}
-0:24:57.5 (P3)	& Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing	&\\
+0:24:57 (P3)	& Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing	&\\
 \cline{1-2}
-0:25:07.5 (P1)	& Yeah I'm not sure if they mean that. I don't know what they mean by pattern in this case. I thought you could just pick roads, varying sizes and like, broads of roads. & 
+0:25:07 (P1)	& Yeah I'm not sure if they mean that. I don't know what they mean by pattern in this case. I thought you could just pick roads, varying sizes and like, broads of roads. & 
 \textbf{[34a answer to 33]} No, not sure what they mean by a pattern.\\
 \cline{1-2}
-0:25:26.0 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \multirow{3}{70mm}{\textbf{[34b answer to 33]} ``Choose a pattern'' becomes ``Choose a pattern preference'' \textsf{[REPLACE]}}\\
+0:25:26 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \multirow{3}{70mm}{\textbf{[34b answer to 33]} ``Choose a pattern'' becomes ``Choose a pattern preference'' \textsf{[REPLACE]}}\\
 \cline{1-2}
-0:25:36.3 (P1) & Yeah, selecting a- I don't know &\\
+0:25:36 (P1) & Yeah, selecting a- I don't know &\\
 \cline{1-2}
-0:25:38.0 (P3)	& Pattern preference maybe? As in, maybe we can explain this in the documentation &\\
+0:25:38 (P3)	& Pattern preference maybe? As in, maybe we can explain this in the documentation &\\
 \hline
-0:25:43.9 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ13 for 34b]} Is ``Choose a pattern preference'' clear?\newline
+0:25:43 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ13 for 34b]} Is ``Choose a pattern preference'' clear?\newline
 \textbf{[36a answer to 35]} no, what kind of pattern?\\
 \cline{1-2}
-0:25:47.4 (P3) & Maybe, I don't know it's- & \\
+0:25:47 (P3) & Maybe, I don't know it's- & \\
 \cline{1-2}
-0:25:48.5 (P1)	& [inaudible] a road pattern& \multirow{2}{70mm}{\textbf{[36b answer to 35]} ``Choose a pattern preference'' becomes ``Choose a road pattern'' \textsf{[REPLACE]}}\\
+0:25:48 (P1)	& [inaudible] a road pattern& \multirow{2}{70mm}{\textbf{[36b answer to 35]} ``Choose a pattern preference'' becomes ``Choose a road pattern'' \textsf{[REPLACE]}}\\
 \hline
@@
 \hline
-\textit{Speaker} & \textit{Text} & \textit{Annotation}\\
+\textit{Speaker} & \textit{Text} & \textit{Coding}\\
 \hline
-0:18:55.7 (P1) &Yeah. And then two processes, static, dynamic and they belong to the goal simulate. & \textbf{[17 goal (AS3)]} System has goal ``Simulate'' \textsf{[INTRO]}\newline
+0:18:55 (P1) &Yeah. And then two processes, static, dynamic and they belong to the goal simulate. & \textbf{[17 goal (AS3)]} System has goal ``Simulate'' \textsf{[INTRO]}\newline
 \textbf{[18 task (AS2)]} System has tasks ``Static simulation'', ``Dynamic simulation'' \textsf{[INTRO]}\newline  
@@
 \hline
-0:30:10.3 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{5}{80mm}{\textbf{[26 critical question CQ10b for 20]} Is the decomposition type of ``simulate'' correct?\newline
+0:30:10 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{5}{80mm}{\textbf{[26 critical question CQ10b for 20]} Is the decomposition type of ``simulate'' correct?\newline
 \textbf{[27 answer to 26]} No, it should be an OR decomposition. \textsf{[REPLACE]}}\\
 \cline{1-2}
-0:30:12.6 (P2) & That's and OR &\\
+0:30:12 (P2) & That's and OR &\\
 \cline{1-2}
-0:30:14.3 (P3) & I think it's an OR &\\
+0:30:14 (P3) & I think it's an OR &\\
 \cline{1-2}
-0:30:15.4 (P1) & It's for the data, it's an OR &\\
+0:30:15 (P1) & It's for the data, it's an OR &\\
 \cline{1-2}
-0:30:18.1 (P3) & Yep &\\
+0:30:18 (P3) & Yep &\\
 \hline	

### 2017-09-14 gmas.tex
Finished section 4 (for now)

@@
 In this section we present our case study.
-All original transcripts, codings, and models are available in our online repository\footnote{\label{foot:url}\url{insertURL}}.
+All original transcripts, codings, and models are available in our online repository\footnote{\label{foot:url}\todo{F}{M}{\url{insertURL}}}.
 
@@
 \hline
-CQ12 & Is the element relevant/useful? & 2 & 3 & 2 &7\\
+CQ11 & Is the element relevant/useful? & 2 & 3 & 2 &7\\
 \hline
-CQ13 & Is the element clear/unambiguous? &3 &10 & 3 & 16\\
+CQ12 & Is the element clear/unambiguous? &3 &10 & 3 & 16\\
 \hline
@@
 
+Our coding further led us to identify three different operations, that is, different effects an argument or critical question can have on a goal model: an argument can introduce a new goal model element (\textsf{INTRO}); it can disable (i.e.
+attack) a goal model element (\textsf{DISABLE}); or it can replace a goal model element (\textsf{REPLACE}).
+Consider, for example, Table~\ref{table:transcripts:traffic-light}.
+First, an argument is posed that introduces a number of tasks.
+Counterarguments are then given against two of these tasks (\emph{add traffic light} and \emph{remove intersection}), which are subsequently disabled.
+An example of \textsf{REPLACE} is given in Table~\ref{table:transcript:decomposition}: what used to be an AND-decomposition is changed into an OR-decomposition.
+We will further discuss the possible operations in Section~\ref{sect:overview}.
+
 %MvZ: I completely removed the subsection below. I think it doesn't add much and it is confusing. What do you think?

### 2017-09-14 overview.tex
Finished section 4 (for now)

@@
 In this section we present our RationalGRL framework, which allows stakeholders to discuss, analyse and argue about goal models.
+
 The RationalGRL framework includes two main parts: Argumentation and GRL goal modeling.
-
 The GRL part of RationalGRL allows for the creation of GRL goal models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
@@
 
-\begin{figure}[ht]
+\begin{figure}[t]
 \centering
-\includegraphics[scale=0.4]{img/framework}
-\caption{The RationalGRL Framework\todo{F}{M,S}{Can you make this a vector image?}}
+\includegraphics[width=\columnwidth]{img/framework.pdf}
+\caption{The RationalGRL Framework}
 \label{fig:rationalgrl-framework}
@@
 
-\subsection{RationalGRL Argument Schemes and Critical Questions}
+\subsection{RationalGRL: Argument Schemes and Critical Questions}
 \label{sect:overview:as}
@@
 \hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant? & DISABLE\\
+AS0 & $a$ is an actor & CQ0 & Is the actor relevant? & DISABLE\\
 \hline
@@
 \hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & DISABLE\\
+AS2 & Actor $a$ can perform task $T$ & CQ2a &Is the task possible? & DISABLE\\
+&& CQ2b & Does the task contribute negatively to some softgoal? & DISABLE\\
 \hline
@@
 \hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
+AS6 & Task $T$ contributes (negatively) to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
 && CQ6b & Are there alternative ways of contributing to the same softgoal?& INTRO \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?& INTRO\\
-&& CQ6d & Does the task contribute to some other softgoal?& INTRO\\
+&& CQ6c & Does the task contribute (negatively) to some other softgoal?& INTRO\\
 \hline
@@
 \hline
-AS11 & Task $T$ contributes negatively to softgoal $S$& CQ11 & Does the task contribute negatively to the softgoal?& DISABLE\\
+AS11 & Element $IE$ is relevant & CQ11 & Is the element relevant/useful? & DISABLE\\
 \hline
-\hline
-AS12 & Element $IE$ is relevant & CQ12 & Is the element relevant/useful? & DISABLE\\
-\hline
-AS13 & Element $IE$ has name $n$ & CQ13 & Is the name clear/unambiguous? & REPLACE\\
+AS12 & Element $IE$ has name $n$ & CQ12 & Is the name clear/unambiguous? & REPLACE\\
 \hline
@@
 Using this list as a basis, we further refined our set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
+Note that this list of argument schemes and critical questions is not exhaustive.
+It is an initial list that we have obtained by coding transcripts.
+However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
 
-The argument schemes AS0-AS4 and AS12-AS13 are arguments for an element of a goal model, and AS5-AS11 are related to links in a goal model.
+Schemes AS0-AS4 and AS12-AS13 are arguments for an element of a goal model, and AS5-AS11 are related to links in a goal model.
 The last scheme (Att) is a scheme for a generic counterargument against any type of argument that has been put forward.
-
 The critical questions presented in Table~\ref{table:argument-schemes} are related to their respective argument schemes.
-Answering a critical question in a certain way can have four different effects on the original argument and the corresponding GRL element:
+As was already discussed at the end of Section~\ref{sect:gmas}, answering a critical question in a certain way can have three different effects on the original argument and the corresponding GRL element:
 
@@
 This operation does not attack the original argument to which the question pertains, but rather creates a new argument.
-In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
+%In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
 For instance, the XOR-decomposition from ``Generate Cars'' is an instantiation of AS5 as follows: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
@@
 \item \textsf{DISABLE:} Disable the element or relationship of an argument scheme to which critical questions pertains.
-This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
-In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
+This operation creates a new argument that disables (i.e., attacks) the original one.
+%In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
 The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
@@
 Furthermore, disabled task ``Dynamic Simulation'' also disables all incoming and outgoing links with this task.
-
 \item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
-In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
+%In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
 In this case, participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
@@
 %SG: why generate car has a star but not this one?
-\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
+%\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
 In Figure~\ref{fig:transcripts:grl}, we see one example of such a counter-attack.
@@
 
-It is important to note that the list of argument schemes and critical questions we present in this section is not exhaustive.
-It is an initial list that we have obtained by coding transcripts.
-However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
+%As an example of a RationalGRL model, consider Figure~\ref{fig:transcripts:grl}, which was created on the basis of transcript $t_1$.
+This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
+
+%\begin{figure*}[h!]
+%\includegraphics[width=\textwidth]{img/Fig6}
+%\caption{The GRL model constructed from transcript $t_1$.} 
+%\label{fig:transcripts:grl}
+%\end{figure*}
+
+\subsection{RationalGRL: Examples from the transcripts}
+Having provided the basics of the RationalGRL framework in previous sections -- Section~\ref{sect:background:grl} discusses GRL and Sections \ref{sect:background:pras} and \ref{sect:overview:as} discuss argumentation -- we now turn to some informal examples of the interactions between the argumentative discussions (i.e.
+the left side of the framework in Figure~\ref{fig:rationalgrl-framework}) on the one hand and GRL models (i.e.
+the right side of the framework in Figure~\ref{fig:rationalgrl-framework}).
+The formal grounding for these examples can be found in the RationalGRL Metamodel (Section~\ref{sect:metamodel}) and the logical formalization of RationalGRL (Section~\ref{sect:formalframework}).
+
+\subsubsection{Example 1 - Introducing GRL elements through argument schemes}
+In RationalGRL, each GRL element can be seen as an argument, that is, the instantiation of an argument scheme.
+Take the example in Figure~\ref{fig:example_AS}, which is based on the transcript excerpt shown in Table~\ref{table:transcript:decomposition}.
+On the left side, the arguments found in the transcript are shown, together with the argumentation scheme they are based on (cf.
+Table~\ref{table:argument-schemes}).
+The participants in the discussion argue that \emph{System} has a goal and two tasks, and that the goal AND-decomposes into the two tasks.
+By arguing in this way, new GRL elements are introduced (right side of Figure~\ref{fig:example_AS}).
+The dashed lines indicate the traceability links between the RationalGRL argumentation model on the left and the GRL model on the right.
+
+\begin{figure}[h]
+\centering
+\includegraphics[width=\columnwidth]{img/fig_example_AS.pdf}
+\caption{Arguments (left) and GRL elements (right). Dashed lines indicate traceability links.}
+\label{fig:example_AS}
+\end{figure}
+
+\subsubsection{Example 2: Disable task \texttt{Traffic light}}
+
+The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$.
+In this example, participants sum up functionality of the traffic simulator, which are tasks that the students can perform in the simulator.
+All these tasks can be formalized and are instantiated from  argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
+
+Once all these tasks are summed up, participant \texttt{P1} notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \texttt{Add traffic light} is not necessary.
+We formalized this using the critical question CQ12: ``Is task \texttt{Add traffic light} useful/relevant?''.
+%TODO I think it is better to say "necessary" and "not necessary" --> change figure 8: useless --> not necessary
+
+We visualize some of the argument schemes, critical questions, and traceability links with GRL model in Figure~\ref{fig:examples:traffic-light}.
+On the left side of the image, we see three of the instantiated argument schemes AS2.
+The bottom one, ``Actor \texttt{Student} has task \texttt{Add traffic light}'', is attacked by another argument generated from applying critical question CQ12: ``\texttt{Add traffic light} is %useless
+not necessary (\emph{All intersections have traffic lights})''. As a result, the corresponding GRL task is disabled. The other two tasks are enabled and have green traceability links.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeIN] at (-2,0) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Save map}}
+        };
+        \node (a1) [argNodeIN] at (-2,-2) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Add road}}
+        };
+        \node (a2) [argNodeOUT] at (-2,-4) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Add traffic light}}
+        };
+        \node (a3) [argNodeIN] at (-2,-6){
+        	\argtext{}{\emph{Add traffic light} is not necessary (\emph{All intersections have traffic lights})}
+        };
+        \node[grl] (task1) at (2.3,0) { \includegraphics[scale=0.5]{img/task_save_map} };
+        \node[grl] (task2) at (2.3,-2) { \includegraphics[scale=0.5]{img/task_add_road} };
+        \node[grl, disabled] (task3) at (2.3,-4) { \includegraphics[scale=0.5]{img/task_add_traffic_light} };
+        \node[traceNodeGREEN] (trace1a) at (-.4,-.3) {};
+        \node[traceNodeGREEN] (trace1b) at (2.2,-.3) {};
+        \node[traceNodeGREEN] (trace2a) at (-.4,-2.3) {};
+        \node[traceNodeGREEN] (trace2b) at (2.2,-2.3) {};
+        \node[traceNodeRED] (trace3a) at (-.4,-4.3) {};
+        \node[traceNodeRED] (trace3b) at (2.2,-4.3) {};
+         \path
+    (a3) edge [attackLink] (a2)
+    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ12} (a3)
+    (trace1a) edge[traceLinkGREEN] (trace1b)
+    (trace2a) edge[traceLinkGREEN] (trace2b)
+    (trace3a) edge[traceLinkRED] (trace3b);
+\end{tikzpicture}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted lines) for the Traffic Light Example.}
+\label{fig:examples:traffic-light}
+\end{figure}
+
+\subsubsection{Example 3: Clarify task \texttt{Road pattern}}
+
+The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_3$.
+It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.
+%TODO where is this task in the model?
+%MvZ: It is in Figure 9, but not in Figure 6.
+Figure 6 is just a single model from our transcripts, but Figure 9 comes from another transcript.
+Do you think it is better to take all of the examples from the same transcript?
+I feel that may look a bit strange since we are completely disregarding the other transcripts then.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}[->]
+        \node (a0) [argNodeOUT] at (-2,0) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Create road}}
+        };
+        \node (a1) [argNodeOUT] at (-2,-2) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a pattern}}
+        };
+        \node (a2) [argNodeOUT] at (-2,-4.2) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a pattern preference}}
+        };
+        \node (a3) [argNodeIN] at (-2,-6.7) {
+        	\argtext{AS2}{Actor \emph{Student} has task \emph{Choose a road pattern}}
+        };
+        \node[grl] (actor) at (2.3,-4) { 
+        	\includegraphics[scale=0.5]{img/task_choose_road_pattern} 
+        };
+        \node[traceNodeGREEN] (trace1) at (-.5,-7.1) {};
+        \node[traceNodeGREEN] (trace2) at (2.2,-4.4) {};
+
+\begin{pgfonlayer}{background}
+         \path
+    (a1) edge[attackLink] (a0)
+    (a2) edge[attackLink] (a1)
+    (a2) edge[attackLink, bend right=20] (a0)
+    (a3) edge[attackLink] (a2)
+    (a3) edge[attackLink, bend right=20] (a1)
+    (a3) edge[attackLink, bend right=40] (a0);
+\end{pgfonlayer}
+
+	\path
+	(a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a1)
+    (a1) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a2)
+    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ12a} (a3)
+    (trace1) edge[traceLinkGREEN] (trace2);
+\end{tikzpicture}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted line) for the Road Pattern Example.}
+\label{fig:examples:clarification}
+\end{figure}
+
+The formalized argument schemes and critical questions are shown in Figure~\ref{fig:examples:clarification}.
+The discussion starts with the first instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Create road}''.
+%TODO I don't see it in Figure 6.
+%MvZ: see previous comment, Figure 9 is not part of Figure 6
+This argument is then challenged with critical question CQ12: ``Is the task \texttt{Create road} clear?''.
+Answering this question results in a new instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Choose a pattern}''.
+%TODO same %MvZ same
+This process is repeated two more times, resulting in the final argument ``Actor \texttt{Student} has task \texttt{Choose a road pattern}''.
+This final argument is "un-attacked" and has a corresponding intentional element (right image).
+
+What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element.
+For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all three other arguments for a name of the element.
+If this is not the case, then it may occur that a previous argument is \emph{re-instantiated}, meaning that it becomes accepted again because the argument attacking it, is itself attacked.
+Suppose for instance that the bottom argument ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' did not attack the second argument: ``Action \texttt{Student} has task \texttt{Choose a pattern}''.
+In that case, this argument would be re-instantiated, because its only attacker ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' is itself defeated by the bottom argument.
+%TODO check these tasks with the GRL model of Figure 6.
+Also, I recommend to have the changed full  GRL model after we finish the examples to show how the model improves.
+
+\subsubsection{Example 4: Decompose goal \texttt{Simulate}} %TODO In fig 1, the goal is called simulate traffic which makes more sense.
+goal "simulate" is a bit meaningless.
+I suggest to update fig 6 and all the other ones accordingly.
+Also, we are refering to this example in example 1.
+Why can't we merge them?
+
+\todo{Marc}{all}{Check whether the text matches the figure here}
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
+It consists of a discussion about the type of decomposition relationship for the goal \texttt{Simulate}.
+
+\begin{figure}[ht]
+\begin{adjustwidth}{-3em}{}
+        \begin{tikzpicture}[->]
+        \node (a0) [argNodeIN] at (-2,0.5) {
+        	\argtext{CQ3}{Task \emph{Dynamic simulation} is not relevant}
+        };
+        \node (a1) [argNodeIN] at (-2,-1.5) {
+        	\argtext{AS2}{Actor \emph{System} has task \emph{Dynamic simulation}}
+        };
+        \node (a2) [argNodeIN] at (-2,-3) {
+        	\argtext{AS2}{Actor \emph{System} has task \emph{Static simulation}}
+        };
+        \node (a3) [argNodeOUT] at (-2,-4.5) {
+        	\argtext{AS5}{Goal \emph{Simulate} AND-decomposes into \emph{Static simulation} and \emph{Dynamic simulation}}
+        };
+        \node (a4) [argNodeIN] at (-2,-7.5) {
+        	\argtext{AS5}{Goal \emph{Simulate} OR-decomposes into \emph{Static simulation} and \emph{Dynamic simulation}}
+        };
+        \node[grl] (actor) at (2.2,-4) { 
+        	\includegraphics[scale=0.5]{img/simulate_decomposition} 
+        };
+        \node[traceNodeRED] (trace2a) at (-.4,-1.8) {};
+        \node[traceNodeRED] (trace2b) at (3.2,-5.6) {};
+        \node[traceNodeGREEN] (trace3a) at (-.4,-3.3) {};
+        \node[traceNodeGREEN] (trace3b) at (1,-5.6) {};
+        \node[traceNodeGREEN] (trace4a) at (-.4,-8.2) {};
+        \node[traceNodeGREEN] (trace4b) at (2.2,-3.9) {};
+
+\begin{pgfonlayer}{background}
+         \path
+    (a1) edge[attackLink] (a0)
+    (a4) edge[attackLink] (a3);
+\end{pgfonlayer}
+
+	\path
+	(a3) edge [CQLink, bend right=50] node [left,draw=none] {CQ10b} (a4)
+	(a1) edge [CQLink, bend right=50] node [right,draw=none] {CQ3} (a0)
+    (trace2a) edge[traceLinkRED] (trace2b)
+    (trace3a) edge[traceLinkGREEN] (trace3b)
+    (trace4a) edge[traceLinkGREEN, bend right] (trace4b);
+\end{tikzpicture}
+\end{adjustwidth}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted line) of the example.} %TODO which example? 
+\label{fig:examples:decomposition}
+\end{figure}
+
+The visualization of this discussion is shown in Figure~\ref{fig:examples:decomposition}.
+Each GRL element on the right has a corresponding argument on the left.
+Moreover, the original argument for an AND-decomposition is attacked by the argument for the OR-decomposition, and the new argument is linked to the decomposition relation in the GRL model.
+%TODO I think we can merge this easily with example 1.
+
+\subsubsection{Example 5: Reinstate actor \texttt{Development team}}
+
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
+It consists of two parts: first participant \texttt{P1} puts forth the suggestion to include actor \texttt{Development team} in the model.
+This is, then, questioned by participant \texttt{P2}, who argues that the professor will develop the software, so there will not be any development team.
+However, in the second part, participant \texttt{P2} argue that the development team should be considered, since the professor does not develop the software.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeOUT] at (-2,0) {
+        	\argtext{AS0}{Development team is relevant}
+        } ;
+        \node (a1) [argNodeIN] at (-2,-2.5){
+        	\argtext{}{Development team is not relevant (\emph{The professor makes the software})}
+        };
+        \node[grl, disabled] (actor) at (2.3,-1) { \includegraphics[scale=0.5]{img/actor_development_team} };
+        \node[traceNodeRED] (trace1) at (-.5,-.3) {};
+        \node[traceNodeRED] (trace2) at (2.2,-0.1) {};
+
+         \path
+    (a1) edge[attackLink] (a0)
+    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
+    (trace1) edge[traceLinkRED] (trace2);
+\end{tikzpicture}
+\caption{Argument schemes and critical questions (left), GRL model (right), and traceability link (dotted line) of a discussion about the relevance of actor Development team.}
+\label{fig:examples:relevant-actor}
+\end{figure}
+
+We formalize this using a \emph{generic counterargument}, attacking the critical question.
+The first part of the discussion is shown in Figure~\ref{fig:examples:relevant-actor}.
+We formalize the first statement as an instantiation of argument scheme AS0: ``Actor \texttt{development team} is relevant''.
+This argument is, then, attacked by answering critical question CQ0: ``Is actor \texttt{development team} relevant?
+with \emph{No}.
+This results in two arguments, AS0 and CQ0, where CQ0 attacks AS0.'' This is shown in Figure~\ref{fig:examples:relevant-actor}, left image.
+
+\begin{figure}[ht!]
+\centering
+        \begin{tikzpicture}
+        \node (a0) [argNodeIN] at (-2,0) {
+        	\argtext{AS0}{Development team is relevant}
+        };
+        \node (a1) [argNodeOUT] at (-2,-2.5){
+        	\argtext{}{Development team is not relevant (\emph{The professor makes the software})}
+        };
+        \node (a2) [argNodeIN] at (-2,-5) {
+        	\argtext{}{The professor doesn't develop the software}
+        } ;
+        \node[grl] (actor) at (2.3,-1) { \includegraphics[scale=0.5]{img/actor_development_team} };
+        \node[traceNodeGREEN] (trace1) at (-.5,-.3) {};
+        \node[traceNodeGREEN] (trace2) at (2.2,-0.1) {};
+
+         \path
+    (a1) edge[attackLink] (a0)
+    (a2) edge[attackLink] (a1)
+    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
+    (a1) edge [CQLink, bend right=50] node [left,draw=none] {Att} (a2)
+    (trace1) edge[traceLinkGREEN] (trace2);
+\end{tikzpicture}
+\caption{Argument Schemes and Critical Questions (left), GRL Model (right), and Traceability Link (dotted line) of a Discussion about the Relevance of Actor Development Team.}
+\label{fig:examples:relevant-actor2}
+\end{figure}
+
+Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counter argument has been put forward.
+The argument ``Actor \texttt{Professor} does not develop the software'' now attacks the argument ``\texttt{Development team} is not relevant (\emph{The professor makes the software})'', which in turn attacks the original argument ``\texttt{Development team} is relevant''.
+As a result, the first argument is re-instantiated, which causes the actor in the GRL model to be enabled again.
 
@@
 
+\begin{figure*}[t]
+\includegraphics[width=\textwidth]{metamodel/metamodel}
+\caption{The RationalGRL metamodel \todo{F}{M}{Change RationalGRLDiagram to RationalGRLmodel}}
+\label{fig:metamodel}
+\end{figure*}
+
 There is only one link between the \textsf{GRL} package and the \textsf{Argumentation} package, but it is a very important one.
@@
 
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{metamodel/metamodel}
-\caption{The RationalGRL metamodel \todo{F}{M}{Change RationalGRLDiagram to RationalGRLmodel}}
-\label{fig:metamodel}
-\end{figure*}
-
-As an example of a RationalGRL model, consider Figure~\ref{fig:transcripts:grl}, which was created on the basis of transcript $t_1$.
-This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
-
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{img/Fig6}
-\caption{The GRL model constructed from transcript $t_1$.} 
-\label{fig:transcripts:grl}
-\end{figure*}
-
-%I wasn't really sure what to do with the below section. 
-\iffalse
+%I wasn't really sure what to do with the below section, so I commented it out. 
+\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsection{RationalGRL Language}
@@
 \end{figure}
-\fi
+\fi%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 \centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology\todo{F}{M,S}{Can you make this a vector image?}}
+\includegraphics{img/methodology.pdf}
+\caption{The RationalGRL Methodology}
 \label{fig:rationalgrl-methodology}
@@
 We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
+
+\subsubsection{Example 6: Methodology example}
+
+In this section, we present a small example for the RationalGRL methodology.
+This example is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
+We base argument schemes and critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
+%These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+%TODO again?
+%We provide several more examples in Section~\ref{sect:examples}.
+
+As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders are developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology.
+This proceeds in the following way:
+\begin{itemize}
+\item
+First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
+with the initial GRL model (Figure~\ref{fig:example-small}) and they invoke the intentional element \texttt{Simulate Traffic}. 
+\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
+They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation"}.
+\item As a result, two argument schemes are created, namely:
+\begin{itemize}
+\item Actor \texttt{Traffic Simulator} has task \texttt{Dynamic Simulation}
+\item Actor \texttt{Traffic Simulator} has task \texttt{Static Simulation}
+\end{itemize}
+\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
+\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
+\begin{itemize}
+\item
+CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
+Thus, the corresponding GRL intentional element is disabled.
+\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
+\end{itemize}
+\end{itemize}
+
+An excerpt of the RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
+%TODO You need to update goal "simulate" in this figure to "Simulate Traffic".
+Also, I don't think we should really call it the rational GRL model.
+it is an excerpts of the model.

### 2017-09-14 transcripts.tex
Finished section 4 (for now)

@@
 \hline
-0:15:42 (P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{70mm}{\textbf{[21 critical question CQ12 for 20]} Is the task ``Add traffic light" useful/relevant?\newline
-\textbf{[22 answer to 21]} Not useful, because according to the specification all intersections have traffic lights. \textsf{[ATTACK]}}\\
+0:15:42 (P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{70mm}{\textbf{[21 critical question CQ12 for 20]} Is the task ``Add traffic light" useful/relevant?\newline\newline
+\textbf{[22 answer to 21]} Not useful, because according to the specification all intersections have traffic lights. \textsf{[DISABLE]}}\\
 \cline{1-2}
@@
 \cline{1-2}
-(P2)	& They need-&\\
+0:15:51 (P2)	& They need-&\\
+\hline
+0:15:52 (P1) & [...] Alright.
+And on the technical side it’s gonna be a real pain to remove one intersection you’re gonna have to remove a lot more because there are only four-ways allowed and if you remove one intersection then- & \multirow{2}{70mm}{\textbf{[23 critical question CQ2 for 20]} Is the task ``Remove intersection'' possible?
+\newline \textbf{[24 answer to 22]} No, it is going to be very difficult to implement.
+\textsf{[DISABLE]} }\\
+\cline{1-2}
+0:16:16 (P2) & Then this road is going nowhere. & \\
 \hline
@@
 \cline{1-2}
-0:25:26 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \multirow{3}{70mm}{\textbf{[34b answer to 33]} ``Choose a pattern'' becomes ``Choose a pattern preference'' \textsf{[REPLACE]}}\\
+0:25:26 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \\
 \cline{1-2}
@@
 \cline{1-2}
-0:25:38 (P3)	& Pattern preference maybe? As in, maybe we can explain this in the documentation &\\
+0:25:38 (P3)	& Pattern preference maybe?
+As in, maybe we can explain this in the documentation & \textbf{[34b answer to 33]} ``Choose a pattern'' becomes ``Choose a pattern preference'' \textsf{[REPLACE]}\\
 \hline
-0:25:43 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ13 for 34b]} Is ``Choose a pattern preference'' clear?\newline
-\textbf{[36a answer to 35]} no, what kind of pattern?\\
+0:25:43 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ13 for 34b]} Is ``Choose a pattern preference'' clear?\\
 \cline{1-2}
-0:25:47 (P3) & Maybe, I don't know it's- & \\
+0:25:47 (P3) & Maybe, I don't know it's- & \textbf{[36a answer to 35]} No \\
 \cline{1-2}
-0:25:48 (P1)	& [inaudible] a road pattern& \multirow{2}{70mm}{\textbf{[36b answer to 35]} ``Choose a pattern preference'' becomes ``Choose a road pattern'' \textsf{[REPLACE]}}\\
+0:25:48 (P1)	& [inaudible] a road pattern& \textbf{[36b answer to 35]} ``Choose a pattern preference'' becomes ``Choose a road pattern'' \textsf{[REPLACE]}\\
 \hline
@@
 \hline
-0:30:10 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{5}{80mm}{\textbf{[26 critical question CQ10b for 20]} Is the decomposition type of ``simulate'' correct?\newline
+0:30:10 (P1) & 	Yeah. But this is- is this an OR or an AND? & \multirow{5}{90mm}{\textbf{[26 critical question CQ10b for 20]} Is the decomposition type of ``simulate'' correct?\newline
 \textbf{[27 answer to 26]} No, it should be an OR decomposition. \textsf{[REPLACE]}}\\

### 2017-09-16 formalframework.tex
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!
small update

@@
 
-The examples and corresponding visualizations of the previous sections provide some insight on how our RationalGRL framework can be used.
-In this section, we present a more precise formalization of our framework.
+The examples and corresponding visualizations of the previous section provide some insight on how our RationalGRL framework can be used.
+However, since we would like to implement our framework in a tool, we require a more precise formalization.
+In this section, we present a formalization of our framework using formal logic.
+This is the main approach using in formal argumentation, so it allows us to use many of the techniques developed in that area directly (see Section~\ref{sect:goalmodeling:relatedwork} for more details).
+
 In the first two subsections we formalize a static representation of our framework: We develop a formal language to specify a GRL model (Section \ref{sect:goalmodeling:logicallanguage}) and we formalize the notion of an argument (Section \ref{sect:goalmodeling:argumentationsemantics}).
@@
 The formal framework we develop in this section serves as the underlying model for the prototype implementation we present in the next section.
+
+\subsection{Logic Programming}
+\label{sect:logicprogramming}
+
+Logic programming is a programming technique largely based on formal logic.
+Any program written in a logic programming language is a set of sentences in logical form, expressing facts and rules about some problem domain.
+Well-known logic programming language families include Prolog, Answer set programming (ASP) and Datalog.
+In all of these languages, rules are written in the form of clauses:
+\begin{align*}
+\texttt{H :- B1, ..., Bn.}.
+\end{align*}
+These clauses are understood declaratively as logical implications:
+\begin{align*}
+\texttt{H if B1 and ... and Bn.}.
+\end{align*}
+\texttt{H} is called the head of the rule and \texttt{B1, ..., Bn} is called the body. Facts are rules that have no body, and are written in the simplified form:
+\begin{align}
+\texttt{H.}.
+\end{align}
+In this article we only consider the simplest case of logic programs, where \texttt{H}, \texttt{B1}, ..., \texttt{Bn} are all atomic formulae, meaning that they are formulae without logical connectives ($\vee, \wedge, \rightarrow, \leftrightarrow$) or negation ($\neg$).
+Such clauses are called definite clauses or Horn clauses.
+
+As we saw in the example above, conjunction (``and'') is denoted with a comma.
+Disjunction is denoted with a semi-colon.
+In logic programs, \emph{variables} are usually denoted as words starting with a capital letter, while atoms start with a lower-case letter.
+
+A simple example of a RationalGRL model as a logic program is as follows:
+
+\begin{verbatim}
+1. actor(0).
+2. name(0, student).
+3. softgoal(1).
+4. name(1, learn).
+5. has(0, 1).
+6. disabled(0).
+7. disabled(Y) :- actor(X), disabled(X), 
+               has(X,Y).
+\end{verbatim}
+
+The first six lines of this logic program are all facts, and the last line is a rule, which means it can be used to infer new facts.
+The first two lines assert that identifier 0 is an actor, and that the name of this actor is ``student''.
+The third and fourth line assert that identifier 1 is a softgoal with name ``learn''.
+The fifth line assert a dependency between these two identifiers, namely that actor ``student'' has softgoal ``learn''.
+The sixth line states that the actor is disabled (perhaps because of some underlying argument).
+Finally the last line states that if \texttt{X} is an actor that is disabled, and this actor has element \texttt{Y}, then this element will also be disabled.
+
+If we run this program and we ask it whether element 1 is disabled, it will answer with \texttt{Yes}, since element 0 is disabled and element 0 has element 1.
+>>>>>>> Stashed changes
 
@@
     (c) edge [attackLink] (b)
-    (d) edge [attackLink] (c)
+    (d) edge [attackLink] (c);
 
@@
 \subsection{Algorithms for argument schemes and critical questions}
-\label[sect:algorithms}
+\label{sect:algorithms}
 

### 2017-09-26 background.tex
changes up to formal framework, publisher files

@@
 \centering
-\includegraphics[width=0.5\textwidth]{img/Example1}
+\includegraphics[width=\columnwidth]{img/Example1.pdf}
 \caption{Partial GRL Model of the traffic simulator example}
 \label{fig:example-small}
-\end{figure} %SG: I changed the figure a bit so that we won't need to change Figure 10. Let me know if you still want to change them according to the above.
-%MvZ: The figure looks good to me!
+\end{figure}
 
@@
 Figure~\ref{fig:example-small} illustrates a simplified GRL diagram from the traffic simulator design exercise.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\texttt{Traffic Simulator}, Figure~\ref{fig:example-small}).
+An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\emph{Traffic Simulator}, Figure~\ref{fig:example-small}).
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
@@
 Softgoals (e.g.
-\texttt{Improve Simulation}) are often related to non-functional requirements, whereas goals (such as  \texttt{Simulate Traffic}) are related to functional requirements.
+\emph{Realistic simulation}) are often related to non-functional requirements, whereas goals (such as \emph{Generate Cars}) are related to functional requirements.
 Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, we have the two tasks \texttt{Static Simulation} and \texttt{Dynamic Simulation}: if the system can perform both a static and a dynamic simulation, it can achieve goal \texttt{Simulate Traffic}.\todo{F}{S}{Later in the paper this goal is called  \texttt{Simulate}, not \texttt{Simulate Traffic}.
-\texttt{Simulate Traffic} seems more clear to me.} In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \texttt{Car Objects}).
-\todo{F}{S,M}{Belief elements are not explained and not in the example.
-I think it would be good to include them (or at least discuss them), as beliefs allow one to give some sort of rationale for e.g.
-goals (in fact, one of the reviewers of one of our previous papers argued that beliefs can also be used to capture reasons).}
+In Figure~\ref{fig:example-small}, we have the two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve goal \emph{Generate cars}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off screen.
+In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \emph{Car Objects}).
+Finally, design rationale can be captured using beliefs (\includegraphics[scale=0.3]{img/belief}).
+For example, the idea that the simulation should be \emph{Not scientifically correct} is captured as a belief in the model.
 
@@
 AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:example-small}, the goal \texttt{Generate cars} is XOR-decomposed to the tasks \texttt{Create new cars} and \texttt{Keep same cars}, as they are alternative ways of achieving the goal \texttt{Generate Cars}.
+In Figure~\ref{fig:example-small}, the goal \emph{Generate cars} is XOR-decomposed to the tasks \emph{Create new cars} and \emph{Keep same cars}, as they are alternative ways of achieving the goal \emph{Generate cars}.
 Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
 A contribution link has a qualitative contribution type or a quantitative contribution.
-Task  \texttt{Create new cars} has a \emph{help} qualitative contribution to the task \texttt{Dynamic Simulation}, and a \emph{hurt} qualitative contribution to the task \texttt{Static Simulation}.
+Task \emph{Create new cars} has a \emph{help} qualitative contribution to the task \emph{Realistic simulation}, and a \emph{hurt} qualitative contribution to the task \emph{Simple design}.
 Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors or resources.
-Here, the goal \texttt{Generate Cars} depends on the resource \texttt{Car Objects}.
-%SG: I added it to the model but we can also remove it altogether.
-%\todo{Marc}{Sepideh}{Shall we add a dependency relation as well? They don't play a big role in our work..}
-%MvZ: It is a bit strange because we say dependencies are between actors but our examples does show this. Perhaps better to leave it out?
+Here, the goal \emph{Generate cars} depends on the resource \emph{Car objects}.
 
@@
 Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-GRL is also linked to Use Case Maps via URNLinks (\includegraphics[scale=1]{img/urnlink}) which provide traceability between concepts and instances of the goal model and behavioral design models.
+GRL is also linked to Use Case Maps via URNLinks (\includegraphics[scale=1]{img/urnlink}), which provide traceability between concepts and instances of the goal model and behavioral design models.
 Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
 
-GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements across multiple diagrams quantitatively, qualitatively or in a hybrid way.
-The satisfaction values from intentional elements in GRL can also be propagated to the elements of Use Case Maps.
+GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements quantitatively, qualitatively or in a hybrid way.
 jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
@@
 This allows GRL to be used in many domains without the need to change the whole modeling language.
-This feature also helps us apply our argumentation to other domains such as compliance, which we explain in more detail in Section~\ref{sect:goalmodeling:openissues}.
 
@@
 \begin{itemize}
-	\item Why does actor \texttt{Traffic Simulator} have softgoal \texttt{High Performance}, which is not linked to any of the goals \texttt{Generate Cars} and \texttt{Simulate Traffic}?
-%SG: I changed this question but let me know if you want to change them back and then change the graph.
-	\item What does \texttt{Keep Same Cars} mean?
-	\item Why does task \texttt{Create New Cars} contribute negatively to \texttt{Static Simulation} and positively to \texttt{Dynamic Simulation}?
-	\item Why does \texttt{Simulate Traffic} AND-decompose into two tasks?
+	\item Why is belief \emph{Not scientifically correct} not linked to any of the goals or tasks? 
+	\item What does \emph{Keep same cars} mean?
+	\item Why does task \emph{Create new cars} contribute negatively to \emph{Simple design} and positively to \emph{Realistic simulation}?
+	\item Why does \emph{Generate cars} XOR-decompose into two tasks?
 \end{itemize}
@@
 The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
-With our RationalGRL framework we aim to address this shortcoming.
+The idea behind the original GRL specification is that beliefs can be used to capture such design rationales that make later justification and review of a model easier.
+However, beliefs cannot be connected to links - this makes answering the third and fourth question above impossible.
+Furthermore, beliefs are after-the-fact design rationales and do not capture the types of critical questions given above.
 
@@
 \begin{itemize}
-\item[] \texttt{Generate Cars} is a goal,
-\item[] Performing action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}, 
-%\item[] Which contributes positively to the softgoal \texttt{Improve Simulation} 
+\item[] \emph{Generate Cars} is a goal,
+\item[] Performing action \emph{Keep same cars} realizes goal \emph{Generate cars}, 
 \item[] \textit{Therefore} 
-\item[] Action \texttt{Keep same Cars} should be performed
+\item[] Action \emph{Keep same cars} should be performed
 \end{itemize}
 
-Note that PRAS is an argument scheme that captures a full inference step: `$G$, $A$ realizes $G$ \emph{Therefore} $A$'.
-There are, however, also schemes that capture simpler reasoning patterns, such as claims of the form `$A$ does not realize $G$'.
+Note that PRAS is an argument scheme that captures a full inference step: ``$G$, $A$ realizes $G$ \emph{Therefore} $A$''.
+There are, however, also schemes that capture simpler reasoning patterns, such as claims of the form ``$A$ does not realize $G$''.
 We will discuss these schemes below.
@@
 In argumentation, conclusions which are at one point acceptable can later be rejected because of new information.
-For example, we may argue that, in fact, performing action \texttt{Keep Same Cars} does not realize goal \texttt{Generate cars}, thus giving a counterargument to the above instantiation of PRAS.
-Atkinson et al.~\cite{atkinson2007} define a set of so-called \emph{critical questions} that point to typical ways in which an argument based on PRAS can be criticized by.
+For example, we may argue that, in fact, performing action \emph{Keep same cars} does not realize goal \emph{Generate cars}, thus giving a counterargument to the above instantiation of PRAS.
+Atkinson et al.~\cite{atkinson2007} define a set of so-called critical questions that point to typical ways in which an argument based on PRAS can be criticized by.
 Some examples of critical questions are as follows.
@@
 \begin{enumerate}
-\item Will the action realize the desired goal?
-\item Are there alternative ways of realizing the same goal?
-\item Does performing the action have a negative side effect?
+\item[CQ1] Will the action realize the desired goal?
+\item[CQ2] Are there alternative ways of realizing the same goal?
+\item[CQ3] Does performing the action have a negative side effect?
 \end{enumerate}
 
-The idea is that answers to critical questions are counterarguments to the the original PRAS argument.
-These counterarguments also follow a scheme; for example, a negative answer to critical question 1 follows the scheme `Action $A$ will not realize goal $G$', which can be instantiated (e.g.
-`\texttt{Keep Same Cars} does not realize \texttt{Generate Cars'}) to forma counterargument to the original argument.
+The idea is that answers to critical questions are counterarguments to the original PRAS argument.
+These counterarguments also follow a scheme; for example, a negative answer to CQ1 follows the scheme ``Action $A$ will not realize goal $G$'', which can be instantiated (e.g.
+``\emph{Keep same cars} does not realize \emph{Generate cars''}) to forma counterargument to the original argument.
 
-Another way to criticize an argument for an action is to suggest an alternative action that realizes the same goal (question 2).
-For example, we can argue that performing a \texttt{Create New Cars} also realizes the goal \texttt{Generate Cars} on its own.
-Also, it is possible that performing an action has a negative side effect (critical question 3).
-For example, while the action \texttt{Create New Cars} realizes the goal \texttt{Generate Cars}, it has a negative side effect, namely hurting \texttt{Static Simulation}: having the simulation constantly create new cars is a functionality that does not allow for a static simulation.
+Another way to criticize an argument for an action is to suggest an alternative action that realizes the same goal (CQ2).
+For example, we can argue that performing \emph{Create new cars} also realizes the goal \emph{Generate cars}.
+Also, it is possible that performing an action has a negative side effect (CQ3).
+For example, while the action \emph{Create new cars} realizes the goal \emph{Generate cars}, it has a negative side effect, namely hurting \emph{Simple design}: having the simulation constantly create new cars is fairly complex design choice.
 
-In argumentation, counterarguments are said to \emph{attack} the original arguments (and sometimes vice versa).
-In the work of Atkinson et al.~\cite{atkinson2007}, arguments and their attacks are captured as an \emph{argumentation framework} of arguments and attack relations.
-Given an argumentation framework, we can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Formal definitions of argumentation frameworks and semantics will be given in section \ref{sect:gmas}.
+In argumentation, counterarguments are said to \emph{attack} the original arguments.
+Given a set of arguments and attacks between these arguments, we can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Formal definitions of argumentation frameworks and semantics will be given in section \ref{sect:gmas}.
 In this section, we will briefly discuss the intuitions behind these concepts.}.
-Figure \ref{fig:pras:example} shows an argumentation framework with three arguments from the traffic simulation example, where arguments are rendered as boxes and attack relations as arrows.
-There are two slightly simplified practical reasoning arguments based on PRAS: argument A1 for \texttt{Keep Same Cars} and argument A2 for \texttt{Create New Cars}.
-Argument A2 proposes an alternative way of realizing the same goal \texttt{Generate Cars} with respect to argument A1 and vice versa (cf.
-critical question 2), so A1 and A2 mutually attack each other, denoted by the double-headed arrow between A1 and A2.
-Argument A3 says that \texttt{Create New Cars} has a negative effect on \texttt{Static Simulation}, so A3 attacks A2, as it points to a negative side-effect of \texttt{Create new cars} (critical question 3).
+Figure \ref{fig:pras:example} shows three arguments from the traffic simulation example, where arguments are rendered as boxes and attack relations as arrows.
+There are two arguments based on PRAS: argument A1 for \emph{Keep Same Cars} and argument A2 for \emph{Create new cars}.
+Argument A2 proposes an alternative way of realizing the same goal \emph{Generate cars} with respect to argument A1 and vice versa (cf.
+CQ2), so A1 and A2 mutually attack each other, denoted by the arrows between A1 and A2.
+Argument A3 says that \emph{Create new cars} has a negative effect on \emph{Static Simulation}, so A3 attacks A2, as it points to a negative side-effect of \emph{Create new cars} (CQ3).
+The intuition here is that an argument is acceptable if any argument that attacks it is itself rejected.
+In Figure~\ref{fig:pras:example}, argument A3 is accepted because it has no attackers.
+This makes A2 rejected (indicated by the lighter grey color), because its attacker A3 is accepted.
+A1 is then also accepted, since its only attacker, A2, is rejected.
 
+\begin{figure}[h]
+\centering
+\includegraphics[]{img/fig_AF1.pdf}
+\caption{PRAS arguments and attacks in the traffic simulation example.}
+\label{fig:pras:example}
+\end{figure}
+
+\iffalse
 \begin{figure}[ht!]
@@
 \begin{tikzpicture}
-        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.4) {\argtext{A1}{Action \texttt{Keep Same Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} Action \texttt{Keep Same Cars} should be performed}};
-        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \texttt{Create New Cars} realizes goal \texttt{Generate Cars}\\ \emph{\underline{Therefore}} Action  \texttt{Create New Cars} should be performed}};
-        \node[minimum size=1cm] (att2) [argNodeIN] at (3.5,0) {\argtext{A3}{\texttt{Create New Cars} has a negative effect on \texttt{Static Simulation}}};
+        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.4) {\argtext{A1}{Action \emph{Keep Same Cars} realizes goal \emph{Generate Cars}\\ \emph{\underline{Therefore}} Action \emph{Keep Same Cars} should be performed}};
+        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \emph{Create New Cars} realizes goal \emph{Generate Cars}\\ \emph{\underline{Therefore}} Action  \emph{Create New Cars} should be performed}};
+        \node[minimum size=1cm] (att2) [argNodeIN] at (3.5,0) {\argtext{A3}{\emph{Create New Cars} has a negative effect on \emph{Static Simulation}}};
          \path
@@
 \end{figure}
-
-Given an argumentation framework, the acceptability of arguments can be determined according to the appropriate argumentation semantics~\cite{Dung1995}.
-The intuition is that an argument is acceptable if it is \emph{undefeated}, that is, any argument that attacks it, is itself defeated.
-In the argumentation framework in Figure~\ref{fig:pras:example}, argument A3 is undefeated because it has no attackers.
-This makes A2 defeated (indicated by the lighter grey color of A2), because its attacker A3 is undefeated.
-A1 is then also undefeated, since its only attacker, A2, is defeated by A3.
-Thus, the set of undefeated (acceptable) arguments given the argumentation framework in Figure~\ref{fig:pras:example} is $\{$A1, A3$\}$.
+\fi
 
@@
 In fact, our case study (Section \ref{sect:gmas}) shows that when discussing requirements, people very often do not structure their reasoning nicely in the way that PRAS presents it.
-That is, you do not see the discussants setting up an argument `We have goal $G$, $A$ realizes $G$ \emph{Therefore} we should perform $A$'.
+That is, you do not see the discussants setting up an argument ``We have goal $G$, $A$ realizes $G$ \emph{Therefore} we should perform $A$''.
 A typical discussion is much more unstructured, as is clear from the transcript excerpts in Appendix~\ref{sect:transcripts:excerpts}.

### 2017-09-26 gmas.tex
changes up to formal framework, publisher files

@@
 
+Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
+In order to get a sense of such discussions, we performed an extensive case study.
 The objective of our case study was to see which types of discourse are used during a discussion of system requirements, and how these discourse types can be captured as argument schemes and critical questions regarding goals and tasks.
@@
 \end{tabularx}
-\caption{Occurrences of argument schemes and critical questions in the transcripts.\todo{F}{F}{check numbers against original annotation docs}}
+\caption{Occurrences of argument schemes and critical questions in the transcripts.}
 \label{table:transcripts:results:argumentschemes}
 \end{table}
-
+%%TODO check numbers against original annotation docs
 Using the initial list of arguments and critical questions, we coded three transcripts of requirements discussions.
@@
 For example, CQ12 is not found in the transcripts, but the related counterargument is:  `...you don't have to specifically add a traffic light' (Table~\ref{table:transcripts:traffic-light}, Appendix~\ref{sect:transcripts:excerpts}).
-
-In order to further verify the coherence of our codings, we constructed complete RationalGRL models based on the argument schemes and critical questions found in each transcript.
-An example model is shown in Figure~\ref{fig:transcripts:grl}, and the full models can be found in the repository.
-More examples of RationalGRL models based on our codings will be given in section \ref{sect:overview}.
 
@@
 attack) a goal model element (\textsf{DISABLE}); or it can replace a goal model element (\textsf{REPLACE}).
-Consider, for example, Table~\ref{table:transcripts:traffic-light}.
+Consider, for example, Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts}.
 First, an argument is posed that introduces a number of tasks.
 Counterarguments are then given against two of these tasks (\emph{add traffic light} and \emph{remove intersection}), which are subsequently disabled.
-An example of \textsf{REPLACE} is given in Table~\ref{table:transcript:decomposition}: what used to be an AND-decomposition is changed into an OR-decomposition.
+An example of replacement is given in Table~\ref{table:transcript:decomposition} (Appendix~\ref{sect:transcripts:excerpts}): what used to be an AND-decomposition is changed into an OR-decomposition.
 We will further discuss the possible operations in Section~\ref{sect:overview}.

### 2017-09-26 outtakes.tex
changes up to formal framework, publisher files

--- 2017-09-14 overview.tex
@@
-\section{The RationalGRL Framework}
-\label{sect:overview}
+%I wasn't really sure what to do with the below section, so I commented it out. 
+\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection{RationalGRL Language}
+The RationalGRL language is an extension of GRL, which means that it includes all the GRL elements shown in Figure~\ref{fig:grl_legend}.
+However, there are also new elements (Figure~\ref{fig:rationalgrllegend}):
 
-In this section we present our RationalGRL framework, which allows stakeholders to discuss, analyse and argue about goal models.
+\begin{itemize}
+\item \emph{Argument}: This represents an argument that does not correspond to a GRL element.
+created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
+\item \emph{Rejected GRL element}: If an argument or GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be rejected, meaning that it does not play a role in the analysis of the GRL model.
+The status of arguments and
+\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
+It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
+In these cases, the corresponding GRL element is shown with a striped background.
+\todo{F}{S,M}{Disabled and Refined GRL elements are not in the metamodel}
+\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument or element attacks the target argument or element.
+\end{itemize}
 
-The RationalGRL framework includes two main parts: Argumentation and GRL goal modeling.
-The GRL part of RationalGRL allows for the creation of GRL goal models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
-For the argumentation part, arguments and counterarguments can be put forward about various parts of this goal model.
-These two parts, GRL and argumentation, are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
-For example, answering a critical question \emph{Is the task \texttt{A} possible?} can result in removing or adding a task in the GRL model.
-Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
-
-Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework.
-At the bottom there are two activities: \emph{practical reasoning \& argumentation} and \emph{goal model construction}.
-As already explained, these two activities influence each other: adding elements to a goal model gives rise to new critical questions about these elements, and answering critical questions with arguments may add or delete elements from the GRL model.
-The activities thus provide two models: a RationalGRL argumentation model (left-hand side), which includes arguments for and against goal model elements, and a GRL goal model (right-hand side).
-The two types of models are connected by traceability links.
-In this way, it becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
-
-\begin{figure}[t]
+\begin{figure}[h]
 \centering
-\includegraphics[width=\columnwidth]{img/framework.pdf}
-\caption{The RationalGRL Framework}
-\label{fig:rationalgrl-framework}
+\includegraphics[width=0.35\textwidth]{img/legend}
+\caption{The new elements and link of RationalGRL}
+\label{fig:rationalgrllegend}
 \end{figure}
-
-In the rest of this section, we discuss the individual parts of the GRL framework.
-First we discuss the argument schemes and critical questions that form the basis of our framework (Section~\ref{sect:overview:as}.
-After this, we explain the RationalGRL language and metamodel (Section~\ref{sect:metamodel}).
-\todo{F}{all}{finish once paper done}
-
-\subsection{RationalGRL: Argument Schemes and Critical Questions}
-\label{sect:overview:as}
-
-\begin{table*}[t]
-\centering
-\begin{tabularx}{\textwidth}{|l|l|l|X|l|l|}
-\hline
-\multicolumn{2}{|c|}{\textbf{Argument scheme}} & \multicolumn{2}{c|}{\textbf{Critical Questions}} & \textbf{Effect}\\
-\hline
-AS0 & $a$ is an actor & CQ0 & Is the actor relevant? & DISABLE\\
-\hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & DISABLE\\
-\hline
-AS2 & Actor $a$ can perform task $T$ & CQ2a &Is the task possible? & DISABLE\\
-&& CQ2b & Does the task contribute negatively to some softgoal? & DISABLE\\
-\hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & DISABLE\\
-\hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?& DISABLE\\
-\hline
-\hline
-AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
-& & CQ5b & Does the goal decompose into any other tasks?& REPLACE\\
-\hline
-AS6 & Task $T$ contributes (negatively) to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal?& INTRO \\
-&& CQ6c & Does the task contribute (negatively) to some other softgoal?& INTRO\\
-\hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?& DISABLE\\
-&& CQ7b & Does the goal contribute to some other softgoal?& INTRO\\
-\hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?& DISABLE\\
-\hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?& INTRO\\
-\hline
-AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& REPLACE\\
- &  & CQ10b & Is the decomposition type correct? (AND/OR/XOR)& REPLACE\\
-\hline
-AS11 & Element $IE$ is relevant & CQ11 & Is the element relevant/useful? & DISABLE\\
-\hline
-AS12 & Element $IE$ has name $n$ & CQ12 & Is the name clear/unambiguous? & REPLACE\\
-\hline
-\hline
-- & - & Att & Generic counterargument & ATTACK\\
-\hline
-\end{tabularx}
-\caption{List of argument schemes (AS0-AS13), critical questions (CQ0-CQ12), and the effect of answering them (right column).}
-\label{table:argument-schemes}
-\end{table*}
-
-A core aspect of the RationalGRL framework are the argument schemes that allow us to talk about the goal models.
-Recall from section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
-Using this list as a basis, we further refined our set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
-Note that this list of argument schemes and critical questions is not exhaustive.
-It is an initial list that we have obtained by coding transcripts.
-However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
-
-Schemes AS0-AS4 and AS12-AS13 are arguments for an element of a goal model, and AS5-AS11 are related to links in a goal model.
-The last scheme (Att) is a scheme for a generic counterargument against any type of argument that has been put forward.
-The critical questions presented in Table~\ref{table:argument-schemes} are related to their respective argument schemes.
-As was already discussed at the end of Section~\ref{sect:gmas}, answering a critical question in a certain way can have three different effects on the original argument and the corresponding GRL element:
+\fi%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\subsection{RationalGRL: Examples from the transcripts}
-Having provided the basics of the RationalGRL framework in previous sections -- Section~\ref{sect:background:grl} discusses GRL and Sections \ref{sect:background:pras} and \ref{sect:overview:as} discuss argumentation -- we now turn to some informal examples of the interactions between the argumentative discussions (i.e.
-the left side of the framework in Figure~\ref{fig:rationalgrl-framework}) on the one hand and GRL models (i.e.
-the right side of the framework in Figure~\ref{fig:rationalgrl-framework}).
-The formal grounding for these examples can be found in the RationalGRL Metamodel (Section~\ref{sect:metamodel}) and the logical formalization of RationalGRL (Section~\ref{sect:formalframework}).
-
-\subsubsection{Example 1 - Introducing GRL elements through argument schemes}
-In RationalGRL, each GRL element can be seen as an argument, that is, the instantiation of an argument scheme.
-Take the example in Figure~\ref{fig:example_AS}, which is based on the transcript excerpt shown in Table~\ref{table:transcript:decomposition}.
-On the left side, the arguments found in the transcript are shown, together with the argumentation scheme they are based on (cf.
-Table~\ref{table:argument-schemes}).
-The participants in the discussion argue that \emph{System} has a goal and two tasks, and that the goal AND-decomposes into the two tasks.
-By arguing in this way, new GRL elements are introduced (right side of Figure~\ref{fig:example_AS}).
-The dashed lines indicate the traceability links between the RationalGRL argumentation model on the left and the GRL model on the right.
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=\columnwidth]{img/fig_example_AS.pdf}
-\caption{Arguments (left) and GRL elements (right). Dashed lines indicate traceability links.}
-\label{fig:example_AS}
-\end{figure}
-
-\subsubsection{Example 2: Disable task \texttt{Traffic light}}
-
-The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in the appendix and comes from transcript $t_1$.
-In this example, participants sum up functionality of the traffic simulator, which are tasks that the students can perform in the simulator.
-All these tasks can be formalized and are instantiated from  argument scheme AS2: ``Actor \emph{Student} has tasks $T$'', where $T\in\{$Save map, Open map, Add intersection, Remove intersection, Add road, Add traffic light$\}$''.
-
-Once all these tasks are summed up, participant \texttt{P1} notes that the problem description states that all intersections in the traffic simulator have traffic lights, so the task \texttt{Add traffic light} is not necessary.
-We formalized this using the critical question CQ12: ``Is task \texttt{Add traffic light} useful/relevant?''.
-%TODO I think it is better to say "necessary" and "not necessary" --> change figure 8: useless --> not necessary
-
-We visualize some of the argument schemes, critical questions, and traceability links with GRL model in Figure~\ref{fig:examples:traffic-light}.
-On the left side of the image, we see three of the instantiated argument schemes AS2.
-The bottom one, ``Actor \texttt{Student} has task \texttt{Add traffic light}'', is attacked by another argument generated from applying critical question CQ12: ``\texttt{Add traffic light} is %useless
-not necessary (\emph{All intersections have traffic lights})''. As a result, the corresponding GRL task is disabled. The other two tasks are enabled and have green traceability links.
-
+\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht!]
@@
 \end{figure}
+\fi%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\subsubsection{Example 3: Clarify task \texttt{Road pattern}}
-
-The transcript excerpt of the second example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_3$.
-It consists of a number of clarification steps, resulting in the task \texttt{Choose a road pattern}.
-%TODO where is this task in the model?
-%MvZ: It is in Figure 9, but not in Figure 6.
-Figure 6 is just a single model from our transcripts, but Figure 9 comes from another transcript.
-Do you think it is better to take all of the examples from the same transcript?
-I feel that may look a bit strange since we are completely disregarding the other transcripts then.
-
+\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht!]
@@
 \end{figure}
+\fi%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-The formalized argument schemes and critical questions are shown in Figure~\ref{fig:examples:clarification}.
-The discussion starts with the first instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Create road}''.
-%TODO I don't see it in Figure 6.
-%MvZ: see previous comment, Figure 9 is not part of Figure 6
-This argument is then challenged with critical question CQ12: ``Is the task \texttt{Create road} clear?''.
-Answering this question results in a new instantiation of argument scheme AS2: ``Actor \texttt{Student} has task \texttt{Choose a pattern}''.
-%TODO same %MvZ same
-This process is repeated two more times, resulting in the final argument ``Actor \texttt{Student} has task \texttt{Choose a road pattern}''.
-This final argument is "un-attacked" and has a corresponding intentional element (right image).
-
-What is clearly shown in this example is that a clarifying argument attacks all arguments previously used to describe the element.
-For instance, the final argument on the bottom of Figure~\ref{fig:examples:clarification} attacks all three other arguments for a name of the element.
-If this is not the case, then it may occur that a previous argument is \emph{re-instantiated}, meaning that it becomes accepted again because the argument attacking it, is itself attacked.
-Suppose for instance that the bottom argument ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' did not attack the second argument: ``Action \texttt{Student} has task \texttt{Choose a pattern}''.
-In that case, this argument would be re-instantiated, because its only attacker ``Actor \texttt{Student} has task \texttt{Choose a pattern preference}'' is itself defeated by the bottom argument.
-%TODO check these tasks with the GRL model of Figure 6.
-Also, I recommend to have the changed full  GRL model after we finish the examples to show how the model improves.
-
-\subsubsection{Example 4: Decompose goal \texttt{Simulate}} %TODO In fig 1, the goal is called simulate traffic which makes more sense.
-goal "simulate" is a bit meaningless.
-I suggest to update fig 6 and all the other ones accordingly.
-Also, we are refering to this example in example 1.
-Why can't we merge them?
-
-\todo{Marc}{all}{Check whether the text matches the figure here}
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
-It consists of a discussion about the type of decomposition relationship for the goal \texttt{Simulate}.
-
+\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{figure}[ht]
@@
 \end{figure}
-
-The visualization of this discussion is shown in Figure~\ref{fig:examples:decomposition}.
-Each GRL element on the right has a corresponding argument on the left.
-Moreover, the original argument for an AND-decomposition is attacked by the argument for the OR-decomposition, and the new argument is linked to the decomposition relation in the GRL model.
-%TODO I think we can merge this easily with example 1.
-
-\subsubsection{Example 5: Reinstate actor \texttt{Development team}}
+\fi%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\subsection{RationalGRL Metamodel}
-\label{sect:metamodel}
-
-Figure~\ref{fig:metamodel} depicts the RationalGRL metamodel linking the main elements of our argumentation extension to the main GRL elements.
-We describe the metamodel bottom-up, starting with the GRL package.\todo{F}{all}{Fig.
-6 is now far removed from where it is mentioned in the text, we should fix this when paper is nearly finished}
-
-The GRL package of our metamodel consists of the core GRL concepts, which constitute a part of the URN metamodel from Recommendation Z.151~\cite{URN}.
-These concepts represent the abstract grammar of the language, independently of the notation.
-This metamodel also further defines the GRL concepts and constructs introduced earlier\footnote{Note that for readability, some GRL concepts, such as contribution strength, have been omitted from the figure.}.
-
-The GRL specification consists of \textsf{GRLModelElements}, which can be either \textsf{GRLLinkableElements} or \textsf{ElementLinks}.
-A \textsf{GRLLinkableElement} can again be specialized into an \textsf{Actor} or an \textsf{IntentionalElement} (which is either a \textsf{Softgoal}, \textsf{Goal}, \textsf{Task}, \textsf{Resource}, or a \textsf{Belief}).
-Intentional elements can be part of an actor, and \textsf{GRLLinkableElements} are connected through \textsf{ElementLinks} of different types (i.e., \textsf{Contribution, Decomposition}, or \textsf{Dependency}).
-Note that actors can be connected through links as well, which is done with \textsf{Dependency} links.
-
-The Argumentation package depicts the concepts we introduced in the previous sections.
-An \textsf{ArgumentScheme} represents an (uninstantiated) scheme containing variables that can be replaced with intentional elements.
-\textsf{CriticalQuestions} are possible ways to attack or elaborate an argument scheme.
-As such, each critical question applies to exactly one scheme, but each scheme can be elaborated or attacked through multiple critical questions.
-When an argument scheme is instantiated, we obtain an \textsf{Argument}.
-Therefore, each argument is associated with exactly one scheme, but a scheme can be instantiated in multiple ways.
-When a critical question is answered, we obtain an \textsf{AttackLink}.
-Each \textsf{AttackLink} is associated with at most one critical question, but a critical question can be used to attack multiple arguments.
-Note that an \textsf{AttackLink} can also be associated with no critical questions.
-This allows the user to create attacks between arguments, which do not necessarily correspond to one of the critical questions.
-A \textsf{RationalGRLmodel} is composed out of arguments and attack relations.
-
-\begin{figure*}[t]
-\includegraphics[width=\textwidth]{metamodel/metamodel}
-\caption{The RationalGRL metamodel \todo{F}{M}{Change RationalGRLDiagram to RationalGRLmodel}}
-\label{fig:metamodel}
-\end{figure*}
-
-There is only one link between the \textsf{GRL} package and the \textsf{Argumentation} package, but it is a very important one.
-The link specifies that each \textsf{GRLModelElement} is in fact an argument.
-This means that each model element inherits the \textsf{AcceptStatus} as well, allowing GRL elements to be accepted or rejected.
-This, furthermore, means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme.
-Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements, but can come from an external source, for instance a piece of evidence or an expert opinion.
-We will see various examples of such arguments in the next sections.
-
-%I wasn't really sure what to do with the below section, so I commented it out. 
-\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\subsection{RationalGRL Language}
-The RationalGRL language is an extension of GRL, which means that it includes all the GRL elements shown in Figure~\ref{fig:grl_legend}.
-However, there are also new elements (Figure~\ref{fig:rationalgrllegend}):
-
-\begin{itemize}
-\item \emph{Argument}: This represents an argument that does not correspond to a GRL element.
-created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
-\item \emph{Rejected GRL element}: If an argument or GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be rejected, meaning that it does not play a role in the analysis of the GRL model.
-The status of arguments and
-\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
-It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
-In these cases, the corresponding GRL element is shown with a striped background.
-\todo{F}{S,M}{Disabled and Refined GRL elements are not in the metamodel}
-\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument or element attacks the target argument or element.
-\end{itemize}
-
-\begin{figure}[h]
-\centering
-\includegraphics[width=0.35\textwidth]{img/legend}
-\caption{The new elements and link of RationalGRL}
-\label{fig:rationalgrllegend}
-\end{figure}
-\fi%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{RationalGRL Methodology}
-
-We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework.
-Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
-The rest of the steps are as follows:
-
-\begin{figure*}[ht]
-\centering
-\includegraphics{img/methodology.pdf}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure*}
-
-\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes of the argumentation framework.
-We select an intentional element from the initial GRL model that we want to analyze and we instantiate a relevant argument scheme from the already existing list of argument schemes or by adding a new one.
-For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}".
-When an argument scheme is instantiated, it corresponds to  an argument for or against part of a goal model.
-
-\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions to attack the argument with counter-arguments.
-Each argument scheme includes one or more critical questions.
-For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contribute to the softgoal?} and \emph{Does the goal contribute to some other softgoals?}.
-%It is worth mentioning that, answering a critical question may result in a "conflict" situation. This is out of the scope of our work here.  
-%MvZ: removed this sentence because I don't know what you mean with "conflict". I think We should explain it better or leave it out.
-When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and step (1).
-
-\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the four following cases can occur: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.
-Any of these cases can  impact the arguments and corresponding GRL intentional elements.
-\textsf{INTRO} means that
-a new argument scheme is created.
-That means, the current argument scheme related to the critical question does not get attacked.
-In the case of \textsf{DISABLE}, the intentional element or its related links are disabled or removed from the models.
-\textsf{REPLACE} introduces a new argument and attacks the original argument at the same time.
-This means that the original element of the argument scheme is replaced with a new one.
-\textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when new evidence occurs.
-
-\textbf{(4) Modify GRL Models} -- In this step, we modify the GRL models based on the situation of step (3).
-That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
-This results in a new modified GRL.
-The new GRL model can then impact the argument schemes and instantiate another argument scheme (Step (1)).
-
-We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
-
-\subsubsection{Example 6: Methodology example}
-
-In this section, we present a small example for the RationalGRL methodology.
-This example is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
-We base argument schemes and critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
-%These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
-%TODO again?
-%We provide several more examples in Section~\ref{sect:examples}.
-
-As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders are developing a goal model for a traffic simulator example and they are modeling the goal \texttt{Simulate Traffic} using the RationalGRL methodology.
-This proceeds in the following way:
-\begin{itemize}
-\item
-First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
-with the initial GRL model (Figure~\ref{fig:example-small}) and they invoke the intentional element \texttt{Simulate Traffic}. 
-\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
-They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation"}.
-\item As a result, two argument schemes are created, namely:
-\begin{itemize}
-\item Actor \texttt{Traffic Simulator} has task \texttt{Dynamic Simulation}
-\item Actor \texttt{Traffic Simulator} has task \texttt{Static Simulation}
-\end{itemize}
-\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
-\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
-\begin{itemize}
-\item
-CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
-Thus, the corresponding GRL intentional element is disabled.
-\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
-\end{itemize}
-\end{itemize}
-
-An excerpt of the RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
-%TODO You need to update goal "simulate" in this figure to "Simulate Traffic".
-Also, I don't think we should really call it the rational GRL model.
-it is an excerpts of the model.
+\begin{tabular}{|p{17mm}|p{63mm}|p{70mm}|}
+\hline
+\textit{Speaker} & \textit{Text} & \textit{Coding}\\
+\hline
+0:17:39 (P1) & And in that process there are activities like create a visual map, create a road& \textbf{[14 task (AS2)]} Student has task ``Create road'' \textsf{[INTRO]}\\
+\hline
+0:24:36 (P3)	& And, well interaction.
+Visualization sorry.
+Or interaction, I don't know.
+So create a visual map would have laying out roads and a pattern of their choosing.
+So this would be first, would be choose a pattern.
+& \textbf{[31 critical question CQ12 for 14]} Is Task ``Create road'' clear?\newline
+\textbf{[32a answer to 31]} no, according to the specification the student should choose a pattern. \newline
+\textbf{[32b answer to 31]} ``Create road'' becomes ``Choose a pattern'' \textsf{[REPLACE]}\\
+\hline
+0:24:55 (P1) &	How do you mean, choose a pattern	& \textbf{[33 critical question CQ12 for 32b]} Is ``Choose a pattern'' clear? \\
+\cline{1-2}
+0:24:57 (P3)	& Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing	&\\
+\cline{1-2}
+0:25:07 (P1)	& Yeah I'm not sure if they mean that. I don't know what they mean by pattern in this case. I thought you could just pick roads, varying sizes and like, broads of roads. & 
+\textbf{[34a answer to 33]} No, not sure what they mean by a pattern.\\
+\cline{1-2}
+0:25:26 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \\
+\cline{1-2}
+0:25:36 (P1) & Yeah, selecting a- I don't know &\\
+\cline{1-2}
+0:25:38 (P3)	& Pattern preference maybe?
+As in, maybe we can explain this in the documentation & \textbf{[34b answer to 33]} ``Choose a pattern'' becomes ``Choose a pattern preference'' \textsf{[REPLACE]}\\
+\hline
+0:25:43 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ12 for 34b]} Is ``Choose a pattern preference'' clear?\\
+\cline{1-2}
+0:25:47 (P3) & Maybe, I don't know it's- & \textbf{[36a answer to 35]} No \\
+\cline{1-2}
+0:25:48 (P1)	& [inaudible] a road pattern& \textbf{[36b answer to 35]} ``Choose a pattern preference'' becomes ``Choose a road pattern'' \textsf{[REPLACE]}\\
+\hline
+\end{tabular}
+\caption{Clarifying the name of a task (transcript $t_3$)}
+\label{table:transcript:task-clarification}

### 2017-09-26 overview.tex
changes up to formal framework, publisher files

--- 2017-09-13 overview.tex
@@
 
-In this section we present our RationalGRL framework, which allows stakeholders to discuss, analyse and argue about goal models.
-The RationalGRL framework includes two main parts: Argumentation and GRL goal modeling.
+Having explored the types of arguments and questions in requirements discussions, we now turn to an overview of our RationalGRL framework.
+We will show through a language definition and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
+A more formal, logical rendition of this traceability can be found in Section~\ref{sect:formalframework}.
 
+The RationalGRL framework includes two parts: Argumentation and GRL goal modeling.
 The GRL part of RationalGRL allows for the creation of GRL goal models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
@@
 These two parts, GRL and argumentation, are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
-For example, answering a critical question \emph{Is the task \texttt{A} possible?} can result in removing or adding a task in the GRL model.
+For example, answering a critical question \emph{Is the task \emph{A} possible?} can result in removing or adding a task in the GRL model.
 Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
@@
 As already explained, these two activities influence each other: adding elements to a goal model gives rise to new critical questions about these elements, and answering critical questions with arguments may add or delete elements from the GRL model.
-The activities thus provide two models: a RationalGRL argumentation model (left-hand side), which includes arguments for and against goal model elements, and a GRL goal model (right-hand side).
-The two types of models are connected by traceability links.
-In this way, it becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
+The activities give rise to two different models: a \emph{RationalGRL model} (left-hand side) and a \emph{GRL model} (right-hand side).
+In the RationalGRL model, we have the goal model elements (goals, tasks etc) and the arguments for and against these elements; the GRL model contains only the goal model elements.
+Thus, the class of GRL models is essentially a subset of the class of RationalGRL models.
+Given the framework, it then becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
 
-\begin{figure}[ht]
+\begin{figure}[b]
 \centering
-\includegraphics[scale=0.4]{img/framework}
-\caption{The RationalGRL Framework\todo{F}{M,S}{Can you make this a vector image?}}
+\includegraphics[width=\columnwidth]{img/framework.pdf}
+\caption{The RationalGRL Framework}
 \label{fig:rationalgrl-framework}
@@
 In the rest of this section, we discuss the individual parts of the GRL framework.
-First we discuss the argument schemes and critical questions that form the basis of our framework (Section~\ref{sect:overview:as}.
-After this, we explain the RationalGRL language and metamodel (Section~\ref{sect:metamodel}).
-\todo{F}{all}{finish once paper done}
+In Section~\ref{sect:overview:as}, we continue our discussion of the argument schemes and critical questions for practical reasoning and argumentation, fitting these schemes and questions into our framework.
+In Section~\ref{sect:overview:lang}, we then discuss the language for RationalGRL models.
+In Section~\ref{sect:overview:examples}, we then provide extensive examples from our case study, illustrating the interplay between practical reasoning and argumentation on the one hand and RationalGRL models on the other hand.
 
-\subsection{RationalGRL Argument Schemes and Critical Questions}
+\subsection{The RationalGRL Argument Schemes and Critical Questions for Practical Reasoning}
 \label{sect:overview:as}
@@
 \centering
-\begin{tabularx}{\textwidth}{|l|l|l|X|l|l|}
+\begin{tabularx}{\textwidth}{|l|l|l|X|l|}
 \hline
@@
 \hline
-AS0 & Actor $a$ is relevant & CQ0 &Is the actor relevant? & DISABLE\\
+AS0 & $a$ is an actor & CQ0 & Is the actor relevant? & \textsf{DISABLE} (no)\\
 \hline
-AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & DISABLE\\
+AS1 & Actor $a$ has resource $R$ & CQ1 &Is the resource available? & \textsf{DISABLE} (no)\\
 \hline
-AS2 & Actor $a$ can perform task $T$ & CQ2 &Is the task possible? & DISABLE\\
+AS2 & Actor $a$ can perform task $T$ & CQ2a &Is the task possible? & \textsf{DISABLE} (no)\\
+&& CQ2b & Does the task contribute negatively to some softgoal? & \textsf{DISABLE} (yes)\\
 \hline
-AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & DISABLE\\
+AS3 & Actor $a$ has goal $G$ & CQ3 & Can the desired goal be realized? & \textsf{DISABLE} (no)\\
 \hline
-AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?& DISABLE\\
+AS4 & Actor $a$ has softgoal $S$ & CQ4 & Is the softgoal a legitimate softgoal?& \textsf{DISABLE} (no)\\
 \hline
 \hline
-AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& DISABLE\\
-& & CQ5b & Does the goal decompose into any other tasks?& REPLACE\\
+AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& \textsf{DISABLE} (no)\\
+& & CQ5b & Does the goal decompose into other tasks? & \textsf{REPLACE} (yes)\\
+ &  & CQ5c & Is the decomposition type correct? & \textsf{REPLACE} (no)\\
 \hline
-AS6 & Task $T$ contributes to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& DISABLE\\
-&& CQ6b & Are there alternative ways of contributing to the same softgoal?& INTRO \\
-&& CQ6c & Does the task have a side effect which contribute negatively to some other softgoal?& INTRO\\
-&& CQ6d & Does the task contribute to some other softgoal?& INTRO\\
+AS6 & Task $T$ contributes (negatively) to softgoal $S$& CQ6a & Does the task contribute to the softgoal?& \textsf{DISABLE} (no)\\
+&& CQ6b & Are there alternative ways of contributing to the same softgoal?& \textsf{INTRO} (yes) \\
+&& CQ6c & Does the task contribute (negatively) to some other softgoal?& \textsf{INTRO} (yes)\\
 \hline
-AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?& DISABLE\\
-&& CQ7b & Does the goal contribute to some other softgoal?& INTRO\\
+AS7 & Goal $G$ contributes to softgoal $S$ & CQ7a & Does the goal contribute to the softgoal?& \textsf{DISABLE} (no)\\
+&& CQ7b & Does the goal contribute to some other softgoal?& \textsf{INTRO} (yes)\\
 \hline
-AS8 & Resource $R$ contributes to task $T$ & CQ8 & Is the resource required in order to perform the task?& DISABLE\\
+AS8 & Task $T$ depends on resource $R$ & CQ8 & Is the resource required in order to perform the task?& \textsf{DISABLE} (no)\\
 \hline
-AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?& INTRO\\
+AS9 & Actor $a$ depends on actor $b$ & CQ9 & Does the actor depend on any actors?& \textsf{INTRO} (yes)\\
 \hline
-AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into other tasks?& REPLACE\\
- &  & CQ10b & Is the decomposition type correct? (AND/OR/XOR)& REPLACE\\
+AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into the tasks?& \textsf{DISABLE} (no)\\
+ &  & CQ10b & Does the task decompose into other tasks?& \textsf{REPLACE} (yes)\\
+ &  & CQ10c & Is the decomposition type correct? & \textsf{REPLACE} (no)\\
 \hline
-AS11 & Task $T$ contributes negatively to softgoal $S$& CQ11 & Does the task contribute negatively to the softgoal?& DISABLE\\
+AS11 & Element $IE$ is relevant & CQ11 & Is the element relevant/useful? & \textsf{DISABLE} (no)\\
+\hline
+AS12 & Element $IE$ has name $n$ & CQ12 & Is the name clear/unambiguous? & \textsf{REPLACE} (no)\\
 \hline
 \hline
-AS12 & Element $IE$ is relevant & CQ12 & Is the element relevant/useful? & DISABLE\\
-\hline
-AS13 & Element $IE$ has name $n$ & CQ13 & Is the name clear/unambiguous? & REPLACE\\
-\hline
-\hline
-- & - & Att & Generic counterargument & ATTACK\\
+Att & Generic counterargument & Att & Generic counterargument & \textsf{DISABLE}\\
 \hline
@@
 
-A core aspect of the RationalGRL framework are the argument schemes that allow us to talk about the goal models.
+A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual types of reasoning stakeholders or designers perform in the early requirements engineering phase (requirement 1 of our framework).
 Recall from section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
 Using this list as a basis, we further refined our set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
-
-The argument schemes AS0-AS4 and AS12-AS13 are arguments for an element of a goal model, and AS5-AS11 are related to links in a goal model.
-The last scheme (Att) is a scheme for a generic counterargument against any type of argument that has been put forward.
-
-The critical questions presented in Table~\ref{table:argument-schemes} are related to their respective argument schemes.
-Answering a critical question in a certain way can have four different effects on the original argument and the corresponding GRL element:
-
-\begin{itemize} 
-\item \textsf{INTRO}: Introduce a new goal element or relationship with a corresponding argument.
-This operation does not attack the original argument to which the question pertains, but rather creates a new argument.
-In Figure~\ref{fig:transcripts:grl}, each GRL element can be seen as the instantiation of an argument scheme.
-For instance, the XOR-decomposition from ``Generate Cars'' is an instantiation of AS5 as follows: ``Goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create new cars}''.
-Suppose now that the modelers that created Figure~\ref{fig:transcripts:grl} would continue their analysis by discussing critical question CQ5b: ``Does the goal \texttt{Generate cars} decompose into other tasks?'', and that they would answer this with ``Yes, namely \texttt{Choose randomly}''.
-This then results in the introduction of another task with the name ``Choose randomly'', and the XOR-decomposes would go from ``Generate Cars'' into the three tasks \texttt{Keep same cars}, \texttt{Create new cars}, and \texttt{Choose randomly}.'' %SG: Very good example.
-But should we also show it in the model as a new element?
-\item \textsf{DISABLE:} Disable the element or relationship of an argument scheme to which critical questions pertains.
-This operation does not create a new argument, but it only disables (i.e., attacks) the original one.
-In Figure~\ref{fig:transcripts:grl}, there are several examples of disabled GRL elements.
-The task ``Add traffic light'' (top-left in figure) is attacked by answering critical question CQ2: ``Is task Add traffic light possible?'' negatively, resulting in an argument that disables the GRL element.
-What we also see from Figure~\ref{fig:transcripts:grl} is that actor ``Teacher'' is disabled and thus all the elements that are bound to this actor are disabled as well.
-Furthermore, disabled task ``Dynamic Simulation'' also disables all incoming and outgoing links with this task.
-
-\item \textsf{REPLACE:} Replace the element of the argument scheme with a new element.
-In Figure~\ref{fig:transcripts:grl}, task ``Show map editor'' has been replaced various times, and this is shown in the figure as a \emph{refined} element.
-In this case, participants were discussing the correct naming for this element (CQ13), leading to various replacements of the name.
-While the previous names are not shown in the figure, they show up in the details pane of the corresponding element.
-%SG: why generate car has a star but not this one?
-\item \textsf{ATTACK:} Attack any argument with an argument that cannot be classified as a critical question.
-In Figure~\ref{fig:transcripts:grl}, we see one example of such a counter-attack.
-First, task ``Control car influx per road'' is attack by answering CQ2 (irrelevant task).
-However, after discussing this, participants found that this was not the case, since the problem description stated that it is important that students can control the simulation manually.
-Therefore, the argument that attacked the task is attacked by the counter-argument ``Important for students'', which re-enables the task ``Control car influx per road''.
-We provide a precise semantics for this in Section~\ref{sect:formalframework}.
-\end{itemize}
-
-It is important to note that the list of argument schemes and critical questions we present in this section is not exhaustive.
+Note that this list of argument schemes and critical questions is not exhaustive.
 It is an initial list that we have obtained by coding transcripts.
@@
 
-\subsection{RationalGRL Metamodel}
-\label{sect:metamodel}
+Schemes AS0-AS4 and AS12-AS13 are arguments for an element of a goal model, and AS5-AS11 are related to links in a goal model.
+The last scheme (Att) is a scheme for a generic counterargument against any type of argument that has been put forward.
+Arguments based on these schemes can be used to discuss a GRL goal model.
+Making an argument based on one of the schemes effectively adds the corresponding GRL element to the model.
+See, for example, Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts}: the participants argue for the addition of several tasks to the goal model using argument scheme AS2.
 
-Figure~\ref{fig:metamodel} depicts the RationalGRL metamodel linking the main elements of our argumentation extension to the main GRL elements.
-We describe the metamodel bottom-up, starting with the GRL package.\todo{F}{all}{Fig.
-6 is now far removed from where it is mentioned in the text, we should fix this when paper is nearly finished}
+As we saw in Sections~\ref{sect:background:pras} and \ref{sect:gmas}, an important part of arguing about goal models is asking the right critical questions.
+The critical questions presented in Table~\ref{table:argument-schemes} are therefore related to their respective argument schemes.
+These questions can be answered with ``yes'' or ``no'', and the type of answer has an effect on the original argument (\textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE}).
+This will be further explained in Section~\ref{sect:overview:examples}.
 
-The GRL package of our metamodel consists of the core GRL concepts, which constitute a part of the URN metamodel from Recommendation Z.151~\cite{URN}.
-These concepts represent the abstract grammar of the language, independently of the notation.
-This metamodel also further defines the GRL concepts and constructs introduced earlier\footnote{Note that for readability, some GRL concepts, such as contribution strength, have been omitted from the figure.}.
+\subsection{The RationalGRL Modelling Language}
+\label{sect:overview:lang}\label{sect:metamodel}
+The language used to construct RationalGRL models is an extension of GRL: it includes all the GRL elements shown in Figure~\ref{fig:grl_legend}.
+However, there are also new elements corresponding to argumentation-related concepts.
+Figure~\ref{fig:rationalgrllegend} shows these elements.
+\begin{itemize}
+\item \emph{Argument}: This represents an argument that does not directly correspond to a GRL element.  
+\item \emph{Rejected (Disabled) GRL element}: If an argument or GRL element is attacked by an argument that itself is not attacked, then this GRL element will be rejected. 
+%\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
+It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
+In these cases, the corresponding GRL element is shown with a striped background.
+%%Refined GRL elements are not in the metamodel and the tool also has no different picture for refined elements
+\item \emph{Attack Link}: An attack link can occur between an argument and another argument or GRL element. It means that the source argument attacks the target argument or GRL element.
+\end{itemize}
 
-The GRL specification consists of \textsf{GRLModelElements}, which can be either \textsf{GRLLinkableElements} or \textsf{ElementLinks}.
+\begin{figure}[b]
+\centering
+\includegraphics{img/legend.pdf}
+\caption{The new elements and link of RationalGRL}
+\label{fig:rationalgrllegend}
+\end{figure}
+
+The complete metamodel of the language can be found in Figure~\ref{fig:metamodel}.
+This metamodel represents the abstract grammar of the language, independently of the notation.
+The metamodel also formalizes the GRL concepts and constructs introduced in Section~\ref{sect:background:grl}\footnote{Note that for readability, some GRL concepts, such as contribution strength, have been omitted from the figure.}.
+
+The metamodel consists of two packages, \emph{practical reasoning \& argumentation} and \emph{goal model construction}, which correspond to the relevant activities in the RationalGRL framework (cf.
+Figure~\ref{fig:rationalgrl-framework}).
+The goal model construction package consists of \textsf{GRLModelElements}, which can be either \textsf{GRLLinkableElements} or \textsf{ElementLinks}.
 A \textsf{GRLLinkableElement} can again be specialized into an \textsf{Actor} or an \textsf{IntentionalElement} (which is either a \textsf{Softgoal}, \textsf{Goal}, \textsf{Task}, \textsf{Resource}, or a \textsf{Belief}).
 Intentional elements can be part of an actor, and \textsf{GRLLinkableElements} are connected through \textsf{ElementLinks} of different types (i.e., \textsf{Contribution, Decomposition}, or \textsf{Dependency}).
-Note that actors can be connected through links as well, which is done with \textsf{Dependency} links.
+Finally, a \textsf{GRLmodel} is composed of \textsf{GRLModelElements}.
 
-The Argumentation package depicts the concepts we introduced in the previous sections.
-An \textsf{ArgumentScheme} represents an (uninstantiated) scheme containing variables that can be replaced with intentional elements.
-\textsf{CriticalQuestions} are possible ways to attack or elaborate an argument scheme.
-As such, each critical question applies to exactly one scheme, but each scheme can be elaborated or attacked through multiple critical questions.
+\begin{figure*}[t]
+\includegraphics[width=\textwidth]{metamodel/metamodel}
+\caption{The RationalGRL metamodel}
+\label{fig:metamodel}
+\end{figure*}
+
+The practical reasoning and argumentation package depicts the concepts we introduced in Section~\ref{sect:overview:as}.
+An \textsf{ArgumentScheme} represents a scheme containing variables.
+\textsf{CriticalQuestions} are possible ways to attack or elaborate an argument based on a scheme; each critical question applies to exactly one scheme, but for each scheme there may be more than one applicable critical question.
 When an argument scheme is instantiated, we obtain an \textsf{Argument}.
 Therefore, each argument is associated with exactly one scheme, but a scheme can be instantiated in multiple ways.
-When a critical question is answered, we obtain an \textsf{AttackLink}.
-Each \textsf{AttackLink} is associated with at most one critical question, but a critical question can be used to attack multiple arguments.
-Note that an \textsf{AttackLink} can also be associated with no critical questions.
+When a critical question is answered, we may obtain an \textsf{AttackLink}, an \textsf{Argument} or both, depending on the answer.
+Note that it is also possible to an \textsf{AttackLink} can also be associated with no critical questions.
 This allows the user to create attacks between arguments, which do not necessarily correspond to one of the critical questions.
@@
 
-There is only one link between the \textsf{GRL} package and the \textsf{Argumentation} package, but it is a very important one.
-The link specifies that each \textsf{GRLModelElement} is in fact an argument.
+Notice that there are various \textsf{OperationTypes} in the Argumentation package.
+In RationalGRL, these operations are performed by instantiating an argument scheme or answering a critical question in a certain way.
+An \textsf{INTRO} operation introduces a new RationalGRL element.
+A \textsf{DISABLE} operation creates a new argument that attacks another argument or GRL element, effectively disabling it.
+The \textsf{REPLACE} operation replaces a RationalGRL element with a new element.
+Instantiating an argument scheme from Table~\ref{table:argument-schemes} always leads to an \textsf{INTRO} operation, that is, it always introduces a new RationalGRL element.
+Answering a critical question can have different effects depending on the critical question and the answer.
+Table~\ref{table:argument-schemes} shows these effects for the different critical questions and answers.
+For example, answering CQ0 with ``no'' disables the argument based on AS0.
+
+There are two important links between the \emph{practical reasoning \& argumentation} and \emph{goal model construction} packages.
+First, each \textsf{GRLModelElement} is an \textsf{Argument}.
 This means that each model element inherits the \textsf{AcceptStatus} as well, allowing GRL elements to be accepted or rejected.
 This, furthermore, means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme.
-Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements, but can come from an external source, for instance a piece of evidence or an expert opinion.
-We will see various examples of such arguments in the next sections.
+Finally, the relation between \textsf{GRLModelElement} and \textsf{Argument} means that, as we already briefly indicated when discussing the framework in Figure~\ref{fig:rationalgrl-framework}, the class of \textsf{RatinalGRLmodel} is a superclass of \textsf{GRLmodel}: besides arguments about GRL elements, we can also have arguments that does not relate to any of the GRL elements.
 
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{metamodel/metamodel}
-\caption{The RationalGRL metamodel \todo{F}{M}{Change RationalGRLDiagram to RationalGRLmodel}}
-\label{fig:metamodel}
-\end{figure*}
+\subsection{From Practical Reasoning to RationalGRL Models: examples from the case study}
+\label{sect:overview:examples}
 
-As an example of a RationalGRL model, consider Figure~\ref{fig:transcripts:grl}, which was created on the basis of transcript $t_1$.
-This figure shows a simplified version of the actual model in order to improve the presentation, but the full models can be found back in our repository.
+We now turn to the interactions between the \emph{practical reasoning \& argumentation} (i.e.
+the bottom left element of the framework in Figure~\ref{fig:rationalgrl-framework}) on the one hand, and \emph{RationalGRL models} (i.e.
+the top left element of the framework in Figure~\ref{fig:rationalgrl-framework}) on the other hand.
+We provide informal examples of the links between the practical reasoning found in our case study transcripts and RationalGRL models.
+The formal grounding for the connection between practical reasoning and RationalGRL models can be found in the RationalGRL Metamodel (Section~\ref{sect:metamodel}) and in more detail in the logical formalization of the Argument Schemes and Critical Questions (Section~\ref{sect:algorithms}).
+Furthermore, the connection between the RationalGRL models shown in this section and regular GRL models is further formally defined in section \ref{sect:rationalGRL-GRL}.
 
-\begin{figure*}[h!]
-\includegraphics[width=\textwidth]{img/Fig6}
-\caption{The GRL model constructed from transcript $t_1$.} 
-\label{fig:transcripts:grl}
-\end{figure*}
+\paragraph{Example 1 - Introducing GRL elements with arguments (\textsf{INTRO)}}
 
-%I wasn't really sure what to do with the below section. 
-\iffalse
-\subsection{RationalGRL Language}
-The RationalGRL language is an extension of GRL, which means that it includes all the GRL elements shown in Figure~\ref{fig:grl_legend}.
-However, there are also new elements (Figure~\ref{fig:rationalgrllegend}):
+\begin{figure}[t]
+\centering
+\includegraphics[width=\columnwidth]{img/fig_example_AS.pdf}
+\caption{Introducing new GRL elements with argumentation (\textsf{INTRO)}.}
+\label{fig:example_AS}
+\end{figure}
 
-\begin{itemize}
-\item \emph{Argument}: This represents an argument that does not correspond to a GRL element.
-created by answering one of the critical questions which disables a GRL element, or counter-attacks a previous argument.
-\item \emph{Rejected GRL element}: If an argument or GRL element is attacked by an argument, which itself is not attacked, then this GRL element will be rejected, meaning that it does not play a role in the analysis of the GRL model.
-The status of arguments and
-\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
-It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
-In these cases, the corresponding GRL element is shown with a striped background.
-\todo{F}{S,M}{Disabled and Refined GRL elements are not in the metamodel}
-\item \emph{Attack Link}: An attack link can occur between an argument and a GRL element, or between two elements. It means that the source argument or element attacks the target argument or element.
-\end{itemize}
+We start by showing how instantiating argument schemes leads to the introduction of new RationalGRL elements in a model.
+Take the example in Figure~\ref{fig:example_AS}, which is based on the transcript excerpt shown in Table~\ref{table:transcript:decomposition}.
+On the left side, the arguments found in the transcript are shown, together with the argument scheme they are based on.
+The participants in the discussion argue that \emph{System} has a goal and two tasks, and that the goal AND-decomposes into the two tasks.
+By arguing in this way, new GRL elements are introduced.
+These GRL elements are shown on the right side of Figure~\ref{fig:example_AS}; the dashed arrows indicate the links between the practical reasoning and argumentation on the left and the RationalGRL model on the right.
 
-\begin{figure}[h]
+\paragraph{Example 2: Disabling GRL elements by answering critical questions (\textsf{DISABLE)}}
+
+\begin{figure}[t]
 \centering
-\includegraphics[width=0.35\textwidth]{img/legend}
-\caption{The new elements and link of RationalGRL}
-\label{fig:rationalgrllegend}
+\includegraphics[]{img/fig_example_disable.pdf}
+\caption{Disabling GRL tasks with argumentation (\textsf{DISABLE)}.}
+\label{fig:example_disable}
 \end{figure}
-\fi
 
-\subsection{RationalGRL Methodology}
+The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_1$.
+In this example, participants first sum up functionality of the traffic simulator, which can be captured as instantiations of AS2.
+On left side of Figure~\ref{fig:example_disable} one such instantiation is shown, which leads to the addition of the task \emph{Add traffic light} in the RationalGRL model on the right side of Figure~\ref{fig:example_disable}.
+However, participant P1 notes that the problem description states that all intersections have traffic lights by default, so the task \emph{Add traffic light} is not necessary.
+This is captured using critical question CQ11.
+A negative answer to this question (cf.
+Table~\ref{table:argument-schemes}) should disable the original argument based on AS2 by attacking it.
+On the left side of Figure~\ref{fig:example_disable} a new argument (CQ11) attacks the original argument based on (AS2).
+This new argument is also added to the RationalGRL model on the right of Figure~\ref{fig:example_disable}, where it attacks the original task \emph{Add traffic light}.
+This attack leads to the original argument being \emph{rejected} (cf.
+Section~\ref{sect:background:pras} and Section~\ref{sect:goalmodeling:argumentationsemantics}), indicated by it being greyed out.
+As a result of this, the corresponding GRL task \emph{Add traffic light} is also disabled.
 
-We propose the following methodology (shown in Figure~\ref{fig:rationalgrl-methodology}) to develop an instance of the RationalGRL framework.
-Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
-The rest of the steps are as follows:
+\paragraph{Example 3: Changing a decomposition type by answering critical questions (\textsf{REPLACE})}
 
-\begin{figure*}[ht]
+\begin{figure}[t]
 \centering
-\includegraphics[scale=0.4]{img/methodology}
-\caption{The RationalGRL Methodology\todo{F}{M,S}{Can you make this a vector image?}}
-\label{fig:rationalgrl-methodology}
-\end{figure*}
+\includegraphics[]{img/fig_example_replace.pdf}
+\caption{Replacing a GRL decomposition with argumentation (\textsf{REPLACE)}.}
+\label{fig:examples:decomposition}
+\end{figure}
 
-\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes of the argumentation framework.
-We select an intentional element from the initial GRL model that we want to analyze and we instantiate a relevant argument scheme from the already existing list of argument schemes or by adding a new one.
-For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}".
-When an argument scheme is instantiated, it corresponds to  an argument for or against part of a goal model.
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
+It consists of a discussion about the type of decomposition relationship for the goal \emph{Simulate Traffic}.
+Recall that in Example 1, an AND-decomposition was introduced for this goal with AS5 (Figure~\ref{fig:example_AS}).
+In the discussion CQ10b -- ``Is the decomposition type correct?'' -- is explicitly asked.
+The answer is ``No, it should be OR''.
+The original argument for AND-decomposition is now attacked  by the argument for the OR-decomposition, and the new argument is linked to the OR-decomposition in the RationalGRL model (Figure~\ref{fig:examples:decomposition}).
 
-\textbf{(2) Answer Critical Questions (CQs)} -- After instantiating an argument scheme, we invoke related critical questions to attack the argument with counter-arguments.
-Each argument scheme includes one or more critical questions.
-For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contribute to the softgoal?} and \emph{Does the goal contribute to some other softgoals?}.
-%It is worth mentioning that, answering a critical question may result in a "conflict" situation. This is out of the scope of our work here.  
-%MvZ: removed this sentence because I don't know what you mean with "conflict". I think We should explain it better or leave it out.
-When the analyst answers  a critical question, a new argument scheme may be instantiated.  Thus, it is possible to go back and forth between this step and step (1).
+\paragraph{Example 4: Clarifying a task by answering critical questions (\textsf{REPLACE})}
 
-\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the four following cases can occur: \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE} or \textsf{ATTACK}.
-Any of these cases can  impact the arguments and corresponding GRL intentional elements.
-\textsf{INTRO} means that
-a new argument scheme is created.
-That means, the current argument scheme related to the critical question does not get attacked.
-In the case of \textsf{DISABLE}, the intentional element or its related links are disabled or removed from the models.
-\textsf{REPLACE} introduces a new argument and attacks the original argument at the same time.
-This means that the original element of the argument scheme is replaced with a new one.
-\textsf{ATTACK} is a generic counterargument which attacks any argument with another argument when new evidence occurs.
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_1$.
+The discussion starts with an instantiation of argument scheme AS2: ``Actor \emph{Student} has task \emph{Set car influx}''.
+This argument is then challenged with critical question CQ12: ``Is the task \emph{Set car influx} specific enough?''.
+This is answered negatively, creating a new argument ``Actor \emph{Student} has task \emph{Set car influx per road}'', which attacks the original argument for \emph{Set car influx} and rejects (disables) the corresponding original RationalGRL task (Figure~\ref{fig:examples:clarification}).
 
-\textbf{(4) Modify GRL Models} -- In this step, we modify the GRL models based on the situation of step (3).
-That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
-This results in a new modified GRL.
-The new GRL model can then impact the argument schemes and instantiate another argument scheme (Step (1)).
+\begin{figure}[b]
+\centering
+\includegraphics[]{img/fig_example_rename.pdf}
+\caption{Renaming a GRL task with argumentation (\textsf{REPLACE)}.}
+\label{fig:examples:clarification}
+\end{figure}
 
-We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
+The discussion about the task then continues.
+Again, there is the question (CQ12) whether the new task \emph{Set car influx per road} is clear enough.
+This time the answer is positive, and participant P1 also gives a reason for this positive answer: setting car influx per road was in the original requirements (cf.
+requirement 4 in Appendix~\ref{sect:designprompt}).
+This reason is then introduced as a belief in the RationalGRL diagram, with a correlation link to the task it is about.
+Thus, beliefs can be used as they are intended, namely as rationales for certain choices.
+
+\paragraph{Example 5: Defending the addition of an actor (\textsf{DISABLE)}}
+
+\begin{figure}[t]
+\centering
+\includegraphics[]{img/reinstate1.pdf}
+\caption{Disabling a GRL actor with argumentation (\textsf{DISABLE)}.}
+\label{fig:examples:relevant-actor}
+\end{figure}
+
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
+First participant \emph{P1} puts forth the suggestion to include actor \emph{Development team} in the model.
+This is, then, questioned by participant \emph{P2}, who argues that the professor will develop the software, so there will not be any development team.
+This is shown in Figure~\ref{fig:examples:relevant-actor}: first the actor \emph{Development Team} is introduced with an argument based on AS0.
+This argument is then attacked by answering critical question CQ0.
+
+\begin{figure}[b!]
+\centering
+\includegraphics[]{img/reinstate2.pdf}
+\caption{Defending a GRL actor by attacking its disabling attacker (\textsf{DISABLE)}.}
+\label{fig:examples:relevant-actor2}
+\end{figure}
+
+Further in the discussion, participant \emph{P2} argues that the development team should be considered, since the professor does not develop the software.
+This is captured using a generic counterargument (\emph{Att} in Table~\ref{table:argument-schemes}), which attacks the earlier argument based on CQ0.
+Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counterargument has been put forward: the argument (Att) now attacks the argument (CQ0), which in turn attacks the original argument (AS0).
+As a result, the argument (AS0) is acceptable (cf.
+Section~\ref{sect:background:pras} and Section~\ref{sect:goalmodeling:argumentationsemantics}), which causes the actor in the RationalGRL model to be enabled again.

### 2017-09-26 tool.tex
many improvements (see email)
changes up to formal framework, publisher files

@@
-\section{The RationalGRL Tool}
-\label{sect:implementation}
+\section{The RationalGRL Methodology and Tool Support}
+\label{sect:tool}
 
+In the previous sections, we have shown how the RationalGRL framework can capture stakeholder discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
+In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a methodology\footnote{The methodology presented here was first presented at the 2017 iStar workshop \cite{iSTar}.} (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
+
+\subsection{RationalGRL Methodology}
+\label{sect:methodology}
+
+We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a (Rational)GRL model.
+%Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+The rest of the steps are as follows:
+
+\begin{figure*}[ht]
+\centering
+\includegraphics{img/methodology.pdf}
+\caption{The RationalGRL Methodology}
+\label{fig:rationalgrl-methodology}
+\end{figure*}
+
+\textbf{(1) Instantiate Argument Schemes (AS)} -- First, we start with our list of argument schemes (Table~\ref{table:argument-schemes}).
+Whilst discussing the requirements, we then select schemes from the list and instantiate them to form arguments for GRL model elements.
+In this way we build or modify the GRL model by introducing new GRL elements (\textsf{INTRO}, see example 1 in Section~\ref{sect:overview:examples}).
+Note that it is also possible to start modifying an existing GRL model which was not built using the RationalGRL methodology: each GRL element corresponds to an argument (i.e.
+an instantiated scheme), so it is possible to instantiate argument schemes based on an existing GRL model.
+
+\textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we can start asking the relevant critical questions.
+Because each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.
+
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of three operations are performed on the GRL model: \textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE}.
+Any of these operations impact the arguments and corresponding GRL intentional elements, modifying the initial GRL model into a RationalGRL model (see examples in Section~\ref{sect:overview:examples}).
+After these modifications, we can keep on asking critical questions (e.g.
+about elements that were introduced by previously answering a critical question) until we are satisfied with our model.
+
+\textbf{(4) Modify GRL Models} -- In this step, we modify the regular GRL model based on the RationalGRL model of step (3).
+That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
+This results in a new, modified GRL model, which can be used as the basis for another cycle of the methodology.
+
+We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
+In the next section, we will give an example of how our tool can be used together with the methodology to build a GRL model.
+
+\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Example 6: Methodology example}
+
+In this section, we present a small example for the RationalGRL methodology.
+This example is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
+We base argument schemes and critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
+%These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+%TODO again?
+%We provide several more examples in Section~\ref{sect:examples}.
+
+As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders are developing a goal model for a traffic simulator example and they are modeling the goal \emph{Simulate Traffic} using the RationalGRL methodology.
+This proceeds in the following way:
+\begin{itemize}
+\item
+First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
+with the initial GRL model (Figure~\ref{fig:example-small}) and they invoke the intentional element \emph{Simulate Traffic}. 
+\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
+They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation"}.
+\item As a result, two argument schemes are created, namely:
+\begin{itemize}
+\item Actor \emph{Traffic Simulator} has task \emph{Dynamic Simulation}
+\item Actor \emph{Traffic Simulator} has task \emph{Static Simulation}
+\end{itemize}
+\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
+\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
+\begin{itemize}
+\item
+CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
+Thus, the corresponding GRL intentional element is disabled.
+\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
+\end{itemize}
+\end{itemize}
+
+An excerpt of the RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
+%TODO You need to update goal "simulate" in this figure to "Simulate Traffic".
+Also, I don't think we should really call it the rational GRL model.
+it is an excerpts of the model.
+\fi
+
+\subsection{The RationalGRL Tool}
 \todo{Marc}{Marc}{Currently this section comes from my thesis where I present the tool as future work. Here we should explain it}

### 2017-09-26 transcripts.tex
changes up to formal framework, publisher files

--- 2017-09-13 transcripts.tex
@@
-\section{Transcripts excerpts}
+\section{Transcript excerpts}
 \label{sect:transcripts:excerpts}
@@
 \hline
-0:15:42 (P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{70mm}{\textbf{[21 critical question CQ12 for 20]} Is the task ``Add traffic light" useful/relevant?\newline
-\textbf{[22 answer to 21]} Not useful, because according to the specification all intersections have traffic lights. \textsf{[ATTACK]}}\\
+0:15:42 (P1) & Well, all intersection should have traffic lights so it's & \multirow{4}{70mm}{\textbf{[21 critical question CQ11 for 20]} Is the task ``Add traffic light" useful/relevant?\newline\newline
+\textbf{[22 answer to 21]} Not useful, because according to the specification all intersections have traffic lights. \textsf{[DISABLE]}}\\
 \cline{1-2}
@@
 \cline{1-2}
-(P2)	& They need-&\\
+0:15:51 (P2)	& They need-&\\
 \hline
 \end{tabular}
-\caption{Adding tasks, disabling useless task ``Add traffic light'' (transcript $t_1$)}
+\caption{Adding tasks, disabling unnecessary task ``Add traffic light'' (transcript $t_1$)}
 \label{table:transcripts:traffic-light}
@@
 \hline
-0:17:39 (P1) & And in that process there are activities like create a visual map, create a road& \textbf{[14 task (AS2)]} Student has task ``Create road'' \textsf{[INTRO]}\\
+0:22:52 (P1) & We also have to be able to change the inflow of cars. How many cars come out in here on the side & \textbf{[31 task (AS2)]} Student has task ``Set car influx'' \textsf{[INTRO]}\\
+\cline{1-2}
+0:23:20 (P1)	& So, sets, yeah, car influx. & \\
 \hline
-0:24:36 (P3)	& And, well interaction.
-Visualization sorry.
-Or interaction, I don't know.
-So create a visual map would have laying out roads and a pattern of their choosing.
-So this would be first, would be choose a pattern.
-& \textbf{[31 critical question CQ13 for 14]} Is Task ``Create road'' clear?\newline
-\textbf{[32a answer to 31]} no, according to the specification the student should choose a pattern. \newline
-\textbf{[32b answer to 31]} ``Create road'' becomes ``Choose a pattern'' \textsf{[REPLACE]}\\
+0:23:41 (P2) &	If you can only control the set amount of influx from any side of this sort of random distribution, I think that is going to be less interesting than when you can say something like, this road is frequently travelled.
+	& \textbf{[32 critical question CQ12 for 31]} Is ``Set car influx'' specific enough?
+\newline \textbf{[33 answer to 32]} No, ``Set car influx'' becomes ``Set car influx per road'' \textsf{[REPLACE]}\\
 \hline
-0:24:55 (P1) &	How do you mean, choose a pattern	& \textbf{[33 critical question CQ13 for 32b]} Is ``Choose a pattern'' clear? \\
-\cline{1-2}
-0:24:57 (P3)	& Students must be able to create a visual map of an area, laying out roads in a pattern of their choosing	&\\
-\cline{1-2}
-0:25:07 (P1)	& Yeah I'm not sure if they mean that. I don't know what they mean by pattern in this case. I thought you could just pick roads, varying sizes and like, broads of roads. & 
-\textbf{[34a answer to 33]} No, not sure what they mean by a pattern.\\
-\cline{1-2}
-0:25:26 (P3) & No yeah exactly, but you would have them provide, it's a pattern, it's a different type of road but essentially you would select- how would you call them, selecting a- & \multirow{3}{70mm}{\textbf{[34b answer to 33]} ``Choose a pattern'' becomes ``Choose a pattern preference'' \textsf{[REPLACE]}}\\
-\cline{1-2}
-0:25:36 (P1) & Yeah, selecting a- I don't know &\\
-\cline{1-2}
-0:25:38 (P3)	& Pattern preference maybe? As in, maybe we can explain this in the documentation &\\
+0:24:12 (P2)	& So setting it per road, I think is something we want	& \textbf{[34 critical question CQ12 for 33]} Is ``Set car influx per road'' clear? \\
 \hline
-0:25:43 (P1) & What kind of patterns though. Would you be able to select & \textbf{[35 critical question CQ13 for 34b]} Is ``Choose a pattern preference'' clear?\newline
-\textbf{[36a answer to 35]} no, what kind of pattern?\\
-\cline{1-2}
-0:25:47 (P3) & Maybe, I don't know it's- & \\
-\cline{1-2}
-0:25:48 (P1)	& [inaudible] a road pattern& \multirow{2}{70mm}{\textbf{[36b answer to 35]} ``Choose a pattern preference'' becomes ``Choose a road pattern'' \textsf{[REPLACE]}}\\
+0:24:15 (P1)	& Yeah, that was also one of the constraints I believe &  \textbf{[35 answer to 34]} Yes, ``Set car influx per road'' was one of the original constraints.\\
 \hline
 \end{tabular}
-\caption{Clarifying the name of a task (transcript $t_3$)}
+\caption{Clarifying the name of a task (transcript $t_1$)}
 \label{table:transcript:task-clarification}

### 2017-10-02 formalframework.tex
changes up to formal framework, publisher files
notes formal framework and small changes

--- 2017-09-02 formalframework.tex
@@
 
-In Section~\ref{sect:gmas} we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of a traffic simulator.
-The resulting list is shown in Table~\ref{table:argument-schemes}.
-We also discussed various examples of applications of the different critical questions and all four different effects (right column of Table~\ref{table:argument-schemes}): \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE}, and \textsf{ATTACK}.
+In Section~\ref{sect:overview}, we have shown through a language definition and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (requirement 2), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (requirement 3).
+A more formal, logical rendition of this traceability will be presented in this section.
+Our main approach here is to use formal argumentation, which allows us to use many of the techniques developed in that area directly (see Section~\ref{sect:background:pras}).
 
-The examples and corresponding visualizations of the previous sections provide some insight on how our RationalGRL framework can be used.
-In this section, we present a more precise formalization of our framework.
-In the first two subsections we formalize a static representation of our framework: We develop a formal language to specify a GRL model (Section \ref{sect:goalmodeling:logicallanguage}) and we formalize the notion of an argument (Section \ref{sect:goalmodeling:argumentationsemantics}).
-In the third subsection we formalize the dynamics of our framework: we develop algorithms for instantiating argument schemes and for answering critical questions.
-The formal framework we develop in this section serves as the underlying model for the prototype implementation we present in the next section.
+In Section~\ref{sect:goalmodeling:logicallanguage} we present the basic logical language of our formalism, based on the metamodel from Section~\ref{sect:overview:lang}.
+Section~\ref{sect:goalmodeling:argumentationsemantics} then presents formal definitions of arguments and how the status of arguments can be computed, which was already briefly informally discussed in Section~\ref{sect:background:pras}.
+In Section~\ref{sect:algorithms} we then develop algorithms for instantiating argument schemes and for answering critical questions, formally capturing the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations of which examples were previously given in Section~\ref{sect:overview:examples}.
+Finally, in Section~\ref{sect:rationalGRL-GRL} we define the link between RationalGRL models on the one hand and GRL models -- which form a subset of the set of RationalGRL models -- on the other hand, formally capturing the export function of our tool (cf.
+Section~\ref{sect:tool}).
 
@@
 An \emph{atom} is a formula with no logical connectives ($\vee, \wedge, \rightarrow, \leftrightarrow$) or negation ($\neg$).
-We first define the logical language for RationalGRL.
+A \emph{ground atom} is an atom without variables.
+
+We first define the logical language for RationalGRL. This logical language is largely based on the metamodel (Section~\ref{sect:overview:lang}).
 
@@
 \begin{itemize}
-\item $actor(i)$: Identifier $i$ is an actor.
-\item $softgoal(i)$: Identifier $i$ is a softgoal.
-\item $goal(i)$: Identifier $i$ is a goal.
-\item $task(i)$: Identifier $i$ is a task.
-\item $resource(i)$: Identifier $i$ is a resource.
-\item $decomp(k,i,J,dtype)$: Identifier $k$ is a $dtype$-decomposition (and/or/xor) from the element corresponding to identifier $i$ to the set of elements corresponding to identifiers $J$.
-\item $dep(k,i,j)$: Identifier $k$ is a dependency link from the element corresponding to identifier $i$ to the element corresponding to identifier $j$.
-\item $contr(k,i,j,ctype)$: Identifier $k$ is a $ctype$-contribution (+/-) from the element corresponding to identifier $i$ to the element corresponding to identifier $j$.
+\item $actor(i,n)$: Identifier $i$ is an actor with name $n$.
+\item $IE(i,n,t,dt)$: Element with identifier $i$ is an intentional element with name $n$ of type $t$ and with decomposition-type $dt$, where: 
+\begin{itemize}
+\item $t$ is one of $softgoal$, $goal$, $task$, $resource$ or $belief$
+\item $dt$ is one of $and$, $or$, $xor$ or $none$.
+\end{itemize}
+\item $EL(k,i,j,lt)$: The element link with identifier $k$ is of link-type $lt$ and links the element with identifier $i$ to the element with identifier $j$, where:
+\begin{itemize}
+\item $t$ is one of $decomposition$, $dependency$ or $contribution$
+\end{itemize}
 \item $has(i,j)$: Identifier $i$ (which is an actor) has the element corresponding to identifier $j$.
-\item $disabled(i)$: The element or relationships corresponding to identifier $i$ is disabled.
-\item $name(i,p)$: The name of the element corresponding to identifier $i$ is $p$.
+\item $arg(i,n)$: Element with identifier $i$ is a generic argument with name $n$
+\item $disable(i,r)$: The element or relationships corresponding to identifier $i$ should be disabled because reason $r$.
 \end{itemize}
 
-We use integers $i,j,k\in\mathbb{N}$ as identifiers for actors and GRL elements. For all of the above definitions we assume that all identifiers used in each atom are different, i.e.,
-\begin{itemize}
-\item $k\not=i$,$k\not=i$,$k\not=j$
-\item forall $j\in J: k\not=j, i\not=j$
-\end{itemize}
+\noindent Identifiers $i,j,k$ are natural numbers. Names $n$, types $t, dt, lt$ and reasons $r$ are strings. We assume that links and elements have unique identifiers and names. 
 \end{definition}
 
-Using the language from Definition 1, we can define GRL models. A \emph{ground atom} is an atom without variables.
+\todo{F}{M}{Ik ben niet helemaal gelukkig met IE en EL.
+Ik had eerst $intentional\_element(i,n,t,dt)$ en $element\_link(k,i,j,lt)$ maar dat komt slecht uit in voorbeelden en algoritmes omdat dat lange namen zijn.
+Wat vind jij?}
 
-\begin{definition}[GRL model] \label{def:argument}
-A \emph{GRL model} $\mathcal{M}$ is a set of ground atoms based on $\mathcal{L}$.
+Using the language from Definition 1, we can define GRL models.
+
+\begin{definition}[GRL model] \label{def:rationaGRLmodel}
+A \emph{GRL model} $\mathcal{M}$ is a set of ground atoms based on $\mathcal{L}$, without atoms of the form $arg(i,n)$ and $disable(i,r)$.
 \end{definition}
 
-An example of the specification of the GRL model in Figure~\ref{fig:examples:decomposition} is shown in Table~\ref{table:grl_atom_spec}, written in logic programming style.
-In this article we do not make use of any logic programming techniques but this would be interesting future work (see Section~\ref{sect:goalmodeling:openissues}).
-A more elaborate example of a specification is shown in Appendix~\ref{ch:grlspec}, showing a complete specification of the traffic simulator GRL model in Figure~\ref{fig:trafficsim}.
+An example of the specification of the RationalGRL model in Figure~\ref{fig:example_AS} is shown in Table~\ref{table:grl_atom_spec}, written in logic programming style.
+A more elaborate example of a specification is shown in Appendix~\ref{ch:grlspec}, showing a complete specification of the traffic simulator GRL model in Figure~\ref{fig:trafficsim}\todo{F}{M}{change Appendix specification to match the still-to-be-created example in section~\ref{sect:tool}}.
+
+Notice that Definition~\ref{def:rationaGRLmodel} does not put any constraints on a RationalGRL model.
+If desired, constraints can be captured using simple axioms which enforce, for example, that each element link has one source and one destination element, both of which are intentional elements.
+For simplicity's sake, we will leave these constraints implicit in this article.
 
@@
 \centering
-\begin{tabularx}{0.4\textwidth}{|X|}
+\begin{tabularx}{\columnwidth}{|X|}
 \hline
-\texttt{goal(0).}\\
-\texttt{task(1).}\\
-\texttt{task(2).}\\
-\texttt{name(0,simulate).}\\
-\texttt{name(1,static\_simulation).}\\
-\texttt{name(2,dynamic\_simulation).}\\
-\texttt{decomp(3,0,[1,2],or).}\\
+\texttt{actor(0,system).}\\
+\texttt{IE(1,simulate\_traffic, goal, or).}\\
+\texttt{IE(2,static\_simulation, task, none).}\\
+\texttt{IE(3,dynamic\_simulation, task, none).}\\
+\texttt{has(0,1).}\\ 
+\texttt{has(0,2).}\\ 
+\texttt{has(0,3).}\\
+\texttt{EL(4,2,1,decomposition).}\\
+\texttt{EL(5,3,1,decomposition).}\\
+
 \hline
 \end{tabularx}
-\caption{Specification of the GRL model in Figure~\ref{fig:examples:decomposition}}
+\caption{Specification of the RationalGRL model in Figure~\ref{fig:example_AS}}
 \label{table:grl_atom_spec}
@@
 
-In the previous subsection we introduced an atomic language to specify a GRL model. In this subsection we give a formal definition of an argument, which is simply a set of atoms from our language.
+In the previous subsection we introduced a language to specify a RationalGRL model. In this subsection we give a formal definition of an argument.
 
@@
 
-This simple definition allows us to form arguments about (parts of) a GRL model.
-For instance, $$\{goal(0),name(0,development\_team)\},$$ is an argument stating that identifier 0 is a goal element with name ``development team''.
-Notice that an argument and a GRL model are essentially the same: as was discussed in Section \ref{sect:metamodel}, each element of a GRL model is an argument.
+This simple definition allows us to form arguments about (parts of) a RationalGRL model.
+For instance, $A_1$: $\{actor(0,system)\}$ is an argument stating that there is an actor \emph{system}, $A_2$: $IE(1,goal,simulate\_traffic,none)$ is an argument stating that there is a goal \emph{simulate\_traffic}, $A_3$: $\{has(0,1)\}$ is an argument stating that the actor with id $0$ (in this case \emph{system}) has has goal \emph{simulate\_traffic}, and $A_4 = A_1 \cup A_2 \cup A_3$ is one argument for the fact that actor \emph{system} has goal \emph{simulate\_traffic}.
 
-Definition \ref{def:argument} does not put any restrictions on the content of an argument.
-This means it is possible to form arguments that do not correspond to correct or sensible GRL models.
-Consider for instance the following argument: $$\{goal(0), task(0), name(0,static\_simulation)\}.$$ This argument states that identifier 0 is both a goal and a task with name ``static simulation'', which is clearly undesirable since it does not correspond to a valid GRL model.
+Note that, like a RationalGRL model, an argument is a set of atoms from our language.
+This makes sense: as the metamodel (Figure~\ref{fig:metamodel}) shows, a RationalGRL model is comprised of arguments, that is, it is the union of all arguments.
+Furthermore, similar to Definition~\ref{def:rationaGRLmodel} for RationalGRL models, Definition~\ref{def:argument} does not put any restrictions on the content of an argument, which makes it possible to form arguments that do not correspond to correct or sensible RationalGRL models.
+For example, argument $A_3$ from the above example only makes sense if there is an actor with id $0$ and an intentional element with id $1$.
 While we could have chosen to put logical constraints on the content of an argument, we choose not to do so here.
 Instead, we put these restrictions on the algorithms in section~\ref{sect:algorithms}.
-The algorithms are constructed in such a way that they only allow the construction of arguments that correspond to valid GRL models.
+The algorithms are constructed in such a way that they only allow the construction of arguments that correspond to valid RationalGRL models.
 The advantage of this approach is that it keeps the presentation of the formal part relatively simple.
@@
 In order to determine which arguments are accepted or not, we define \emph{argumentation semantics}.
-We used the standard approach here, which is known as \emph{Dung's semantics}.
-The following notions are preliminary.
 
-\begin{definition}[Attack, conflict-freeness, defense, and admissibility~\cite{Dung1995}] Suppose an argumentation framework $AF=(Arg,Att)$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
-We say that
+\begin{definition}[Argumentation Semantics] Suppose an argumentation framework $AF=(Arg,Att)$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$. We say that
 \begin{itemize}
@@
 \item $S$ is \emph{admissible} if $S$ is conflict-free and defends each argument in it.
+\item $S$ is a \emph{preferred extension} of $AF$ if $S$ is a maximal (w.r.t. set inclusion) admissible set of $AF$.
 \end{itemize}
@@
 Sets containing both $A_2$ and either $A_1$ or $A_3$ are not conflict free, and the sets $\{A_1\}$ and $\{A_2\}$ do not defend themselves against $A_2$ and $A_3$, respectively.
+Given the notion of admissible sets, we can then define our extensions.
+There are a large number of different semantics to determine which arguments are in an extension; in this article, we choose preferred semantics.
+In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_1,A_3\}$.
+Based on this extension, we can define the status of arguments, acceptable or rejected.
 
-Given the notion of admissible sets, we can then define our argumentation semantics.
-There are a large number of different semantics to determine which arguments are acceptable; in this article, we choose preferred semantics.
-
-\begin{definition}[Preferred semantics~\cite{Dung1995}] \label{def:semantics}
-A preferred extension of an argumentation framework $AF$ is a maximal (w.r.t. set inclusion) admissible set of $AF$.
+\begin{definition}[Acceptable and Rejected Arguments] Given an argumentation framework $AF=(Arg,Att)$, an argument $A$ is \emph{acceptable} if $A$ is in the preferred extension of $AF$, and rejected otherwise.
 \end{definition}
 
-In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_1,A_3\}$.
+This means that, given the argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, arguments $A_1$ and $A_3$ are acceptable and $A_2$ is rejected.
 
 \subsection{Algorithms for argument schemes and critical questions}
-\label[sect:algorithms}
+\label{sect:algorithms}
 
-Let us briefly take stock:
-\begin{itemize}
-\item We developed an initial set of argument schemes and critical questions in Section~\ref{sect:gmas}, by annotating transcripts.
-\item We developed a visual notation to illustrate some argument schemes and critical questions in Section~\ref{sect:examples}.
-\item We proposed a logical language consisting of atomic sentence to describe a GRL model in Section~\ref{sect:goalmodeling:logicallanguage}.
-\item We formalized an ``argument'' as a set of atoms from our language, and we introduced \emph{argumentation semantics} to compute sets of accepted arguments in Section~\ref{sect:goalmodeling:argumentationsemantics}.
-\end{itemize}
-
-In this subsection we develop algorithms for applying argument schemes and critical questions.
-The process of applying these algorithms and generating goal models from them is shown in context of the RationalGRL methodology in Figure~\ref{fig:methodologyalgorithm} \todo{Marc}{Marc}{Create an extended version of the methodology, containing where which parts of the formal framework are used.} These algorithms produce arguments (Definition~\ref{def:argument}) and attack relations, forming an argumentation framework (Definition~\ref{def:argumentationframework}).
-One can then use argumentation semantics (Definition~\ref{def:semantics}) to compute sets of accepted arguments.
-The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.
-
-In the following algorithms, we assume the following global variables:
-
-\begin{definition}[Global variables] The following variables are intended to have a \emph{global} scope and have a \emph{static} value, meaning the lifetime of the variable extends across the entire run of the program.
-\begin{itemize}
-\item $id$: the current highest identifier of the elements. This variable is increased for each new element that is added.
-\item $Args$: contains the set of all arguments.
-\item $Att$: contains the set of all attack relations.
-\end{itemize}
-\end{definition}
+In Section~\ref{sect:goalmodeling:logicallanguage} we have developed a simple logical language with which to capture RationalGRL models.
+We then (Secton~\ref{sect:goalmodeling:argumentationsemantics} formalized an argument as a set of atoms from our language, and we introduced \emph{argumentation semantics} to compute sets of accepted arguments.
+In this section we develop algorithms for applying argument schemes and critical questions.
+These algorithms define which arguments should be introduced whenever an argument scheme is instantiated or a critical question is answered.
+In other words, the algorithms define the exact application of the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations for the different argument schemes and critical questions.
+The arguments and attack relations produced by the algorithms then together form an argumentation framework with an associated preferred extension, which can be interpreted as a RationalGRL model.
 
 \subsubsection*{Algorithms for argument schemes}
+
+In the following algorithms, $id$ is the current highest identifier of the elements, which is increased for each new element that is added, $Args$ is the set of all arguments and $Att$is the set of all attack relations.
+
 The algorithms for the argument schemes consist of forming a new argument and adding it to the set of arguments. No attack relations are introduced.
@@
 \begin{algorithm}[h]
-  \caption{Applying AS0: Actor $a$ is relevant}\label{alg:as0}
+  \caption{Applying AS0: Actor $n_{a}$ is relevant}\label{alg:as0}
   \begin{algorithmic}[1]
-    \Procedure{$AS_0$}{$a$}
+    \Procedure{$AS_0$}{$n_{a}$}
     \State $id\gets id+1$\label{alg:as0:1}
-    \State $A \gets \{actor(id), name(id,a)\}$\label{alg:as0:2}
+    \State $A \gets \{actor(id, n_{a})\}$\label{alg:as0:2}
     \State $Args \gets Args \cup A$\label{alg:as0:3}
@@
 
-\emph{Algortihm~\ref{alg:as0} for argument scheme AS0}: The algorithm takes one argument, namely the name of the actor $a$.
+\noindent\textbf{Algorithm~\ref{alg:as0} for argument scheme AS0}: The algorithm takes one argument, namely the name $n_{a}$ of the actor.
 On line~\ref{alg:as0:1} of the algorithm, the global variable $id$ is increased by one.
 This ensures that each new argument has a unique identifier.
-On line~\ref{alg:as0:2}, the argument for the actor is formed, consisting of two statements, stating respectively that $id$ is an actor, and that the name of this actor is $a$.
-Finally, on line~\ref{alg:as0:3} this argument is added to the global set of arguments $Args$.
-In Figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_0(\text{Development team})$, results in one argument: $$Args = \{\{actor(0), name(0, \text{Development team})\}\}.$$
+On line~\ref{alg:as0:2}, the argument for the actor is formed consisting of the statement that there is an actor with $id$ and name $n_{a}$ and on line~\ref{alg:as0:3} this argument is added to the set of arguments $Args$.
+In Figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_0(\text{Development team})$, results in one argument: $\{actor(0, \text{Development team})\}$.
 
 \begin{algorithm}[h]
-  \caption{Applying AS1: Actor $a_{id}$ has resource $n$}\label{alg:as1}
+  \caption{Applying AS1-AS4: Actor $id_{a}$ has intentional element $n_{ie}$ of type $t_{ie}$}\label{alg:as1}
   \begin{algorithmic}[1]
-    \Procedure{$AS_1$}{$a_{id}, n$}
+    \Procedure{$AS_1$}{$id_{a}$, $n_{ie}$, $t_{ie}$}
     \State $id\gets id+1$
-    \State $A\gets \{resource(id), name(id,n), has(a_{id},id)\}$
+    \State $A\gets \{IE(id,n_{ie},t_{ie},none), has(id_{a},id)\}$
     \State $Args \gets Args\cup A$
@@
 \end{algorithm}
-
-\emph{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
+\todo{F}{M}{fix tekst bij aangepast algoritme}
+\noindent\textbf{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
 The algorithm is similar to the previous one, with the difference that the newly added argument contains the statement $has(a_{id},id)$ as well, meaning that the actor with id $a_{id}$ has element $id$ (which is a resource).
@@
 
-\emph{Algorithm~\ref{alg:as5} for argument scheme AS5:} The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
+\noindent\textbf{Algorithm~\ref{alg:as5} for argument scheme AS5:}\todo{F}{M}{goal decomposes in 1 task, dtype is aangegevebn op de goal en niet op de decomp, for-loop kan eruit als we aannemen dat de task al bestaat (we nemen in alg 2 trouwens ook gewoon aan dat de actor al bestaat).}
+The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
 The difficulty of this algorithm is that each of the tasks are stated in natural language, and it is not directly clear whether these tasks are already in the GRL model or not.
@@
 
-\emph{Algorithm~\ref{alg:as6} for argument scheme AS6:} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
+\noindent\textbf{Algorithm~\ref{alg:as6} for argument scheme AS6:} \todo{F}{M}{if-statement kan eruit als we aannemen dat de softgoal al bestaat.}
+The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
 The idea behind this algorithm is very similar to the previous one, with the difference that in the current algorithm we create a single relation, while we created a set of relations in the previous algorithm.
@@
 
-\emph{Algorithms for argument schemes AS7-AS11:} The algorithms for $AS7$ to $AS11$ all have a very similar structure as Algorithm~\ref{alg:as6} and have therefore been omitted.
-Again, we assume the reader can reconstruct them straightforwardly.
+\emph{Algorithms for argument schemes AS7-AS11:} \todo{F}{M}{kunnen algoritme \ref{alg:as6} en de algoritmes voor AS7-AS11 niet in 1 algoritme voor $EL$?
+(op dezelfde manier als alg.
+\ref{alg:as1} voor alle IE)}
+The algorithms for $AS7$ to $AS11$ all have a very similar structure as Algorithm~\ref{alg:as6} and have therefore been omitted. Again, we assume the reader can reconstruct them straightforwardly.
 
@@
 
-We now develop algorithms for our critical questions. Recall that answering a critical question can have four effects, and we discuss each of these effects separately.
+We now develop algorithms for our critical questions.
+\todo{F}{M}{check of alle CQ nummers e.d.
+nog kloppen.} Recall that answering a critical question can have four effects, and we discuss each of these effects separately.
 
@@
 
-\emph{Algorithm~\ref{alg:disable} (\textsf{DISABLE}) for critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12:} The disable operation consists of adding an argument stating the GRL element with identifier $i$ is disabled.
+\noindent\textbf{Algorithm~\ref{alg:disable} (\textsf{DISABLE}) for critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12:} 
+\todo{F}{M}{Wat ik dacht is dat we zeggen dat ``element $n$ is disabled because $r$'' en dan voegen we een nieuwe ARG toe met $disable(id,r)$ welke het argument met naam $n$ aanvalt}
+The disable operation consists of adding an argument stating the GRL element with identifier $i$ is disabled.
 Let us reconsider the example of Figure~\ref{fig:examples:relevant-actor}.
@@
 
-\emph{Algorithm~\ref{alg:cq5b} (\textsf{REPLACE}) for critical questions CQ5b:} This algorithm is executed when critical question CQ5b is answered, which is a critical question for argument scheme AS5.
+\noindent\textbf{Algorithm~\ref{alg:cq5b} (\textsf{REPLACE}) for critical questions CQ5b:} \todo{F}{M}{Kan verandert worden: decmp relaties zijn per task en je introduceert altijd een nieuwe dus de check of hij al betsaat kan eraf.} This algorithm is executed when critical question CQ5b is answered, which is a critical question for argument scheme AS5.
 Therefore, it assumes an argument for a goal decomposition already exists of the following form (see Algorithm~\ref{alg:as5}): $$\{decomp(d, g_{id}, \{i_1,\ldots,i_n\},type).$$  The goal of the algorithm is to generate a new argument of the form $decomp(d, g_{id}, \{i_1,\ldots,i_5\}\cup \{j_1,\ldots,j_k\}, type)$, where $\{j_1,\ldots,j_k\}$ are the identifiers of the additional decomposing tasks $\{t_1,\ldots,t_k\}$.
@@
 
-\emph{Algorithms for critical questions CQ10a and CQ10b (REPLACE)}: These algorithms have a very similar structure as Algorithm~\ref{alg:cq5b} and have therefore been omitted.
+\noindent\textbf{Algorithms for critical questions CQ10a and CQ10b (REPLACE)}: These algorithms have a very similar structure as Algorithm~\ref{alg:cq5b} and have therefore been omitted.
 
@@
 
-\emph{Algorithms for critical question CQ13 (REPLACE):} This algorithm is used to clarify/change the name of an element.
+\noindent\textbf{Algorithms for critical question CQ13 (REPLACE):} \todo{F}{M}{deze moet aangepast worden nu de naam in het element staat} This algorithm is used to clarify/change the name of an element.
 It takes two parameters: the element identifier $i$ and the new name $n$.
@@
 
-\emph{Algorithms for critical questions CQ6b, CQ6c, CQ6d, CQ7b, and CQ9 (INTRO):} The introduction algorithms for the critical questions are all very similar to the INTRO algorithms for argument schemes (Algorithm~\ref{alg:as1}).
+\noindent\textbf{Algorithms for critical questions CQ6b, CQ6c, CQ6d, CQ7b, and CQ9 (INTRO):} The introduction algorithms for the critical questions are all very similar to the INTRO algorithms for argument schemes (Algorithm~\ref{alg:as1}).
 They have therefore been omitted.
@@
 
-\emph{Algorithm for $Att$ (Generic counter argument:} Applying a generic counter argument is very simple, and simply results on an attack on the original argument.
+\noindent\textbf{Algorithm for $Att$ (Generic counter argument:} \todo{F}{M}{zoals je in Def 1 kunt zien heb ik generic arguments een id en een naam gegeven.
+Je moet ze namelijk kunnen aanvallen!
+(en de lege set kun je niet aanvallen want die is niet uniek te identificeren).
+algoritme moet dus aangepast worden}Applying a generic counter argument is very simple, and simply results on an attack on the original argument.
 We illustrate this by continuing our example from Figure~\ref{fig:examples:relevant-actor2:formal} (Algorithm~\ref{alg:as0}).
@@
 \subsection{Constructing GRL models}
+\label{sect:rationalGRL-GRL}
+\todo{F}{F,M}{Dit is snel getypt, moet nog checken of het klopt.}
 
-\todo{F}{F,M}{It would be good to have a formal definition of how to construct models given an argumentation framework.
-If we define a model M at the start of this section, this should be fairly easy.}
+\begin{definition}[RationalGRL model] A RationalGRL model \mathcal{R} is an argumentation framework $AF=(Args,Att)$ where all arguments in $Args$ and attacks in $Att$ are based on algorithm X-XX.  
+\end{definition}
+
+\begin{definition}[GRL model] A GRL model $\mathcal{M}$ based on a RationalGRL model $\mathcal{R}$ is the preferred extension of $\mathcal{R}$.  
+\end{definition}
 

### 2017-10-02 overview.tex
notes formal framework and small changes

@@
 \hline
-AS5 & Goal $G$ decomposes into tasks $T_1,\ldots,T_n$ & CQ5a & Does the goal decompose into the tasks?& \textsf{DISABLE} (no)\\
-& & CQ5b & Does the goal decompose into other tasks? & \textsf{REPLACE} (yes)\\
+AS5 & Goal $G$ decomposes into task $T$ & CQ5a & Does the goal decompose into the task?& \textsf{DISABLE} (no)\\
+& & CQ5b & Does the goal decompose into other tasks? & \textsf{INTRO} (yes)\\
  &  & CQ5c & Is the decomposition type correct? & \textsf{REPLACE} (no)\\
@@
 \hline
-AS10 & Task $T_1$ decomposes into tasks $T_2,\ldots,T_n$ & CQ10a & Does the task decompose into the tasks?& \textsf{DISABLE} (no)\\
- &  & CQ10b & Does the task decompose into other tasks?& \textsf{REPLACE} (yes)\\
+AS10 & Task $T_i$ decomposes into task $T_j$ & CQ10a & Does the task decompose into the task? & \textsf{DISABLE} (no)\\
+ &  & CQ10b & Does the task decompose into other tasks?& \textsf{INTRO} (yes)\\
  &  & CQ10c & Is the decomposition type correct? & \textsf{REPLACE} (no)\\
@@
 This, furthermore, means that argument schemes can be applied to all GRL elements, capturing the intuition that each GRL element can be regarded as an instantiated argument scheme.
+Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements.
 Finally, the relation between \textsf{GRLModelElement} and \textsf{Argument} means that, as we already briefly indicated when discussing the framework in Figure~\ref{fig:rationalgrl-framework}, the class of \textsf{RatinalGRLmodel} is a superclass of \textsf{GRLmodel}: besides arguments about GRL elements, we can also have arguments that does not relate to any of the GRL elements.
@@
 The formal grounding for the connection between practical reasoning and RationalGRL models can be found in the RationalGRL Metamodel (Section~\ref{sect:metamodel}) and in more detail in the logical formalization of the Argument Schemes and Critical Questions (Section~\ref{sect:algorithms}).
-Furthermore, the connection between the RationalGRL models shown in this section and regular GRL models is further formally defined in section \ref{sect:rationalGRL-GRL}.
+The connection between the RationalGRL models shown in this section and regular GRL models is further formally defined in section \ref{sect:rationalGRL-GRL}.
 
@@
 This argument is then challenged with critical question CQ12: ``Is the task \emph{Set car influx} specific enough?''.
-This is answered negatively, creating a new argument ``Actor \emph{Student} has task \emph{Set car influx per road}'', which attacks the original argument for \emph{Set car influx} and rejects (disables) the corresponding original RationalGRL task (Figure~\ref{fig:examples:clarification}).
+This is answered negatively, creating a new argument ``Actor \emph{Student} has task \emph{Set car influx per road}'', which attacks the original argument for \emph{Set car influx}.
+Note how the new task \emph{Set car influx per road} also attacks (and disables) the original RationalGRL task \emph{Set car influx} (Figure~\ref{fig:examples:clarification}).
 

### 2017-10-08 formalframework.tex
added small note formal framework
improved formal framework

--- 2017-09-02 formalframework.tex
@@
 The resulting list is shown in Table~\ref{table:argument-schemes}.
-We also discussed various examples of applications of the different critical questions and all four different effects (right column of Table~\ref{table:argument-schemes}): \textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE}, and \textsf{ATTACK}.
+We then presented the RationalGRL framework in secton~\ref{sect:overview}, consisting of a modeling language, a metamodel, and various examples of using the argument schemes and critical questions.
+In section~\ref{sect:methodology} we explained how this framework can be used by practitioners by explaining the methodology and tool support.
 
-The examples and corresponding visualizations of the previous sections provide some insight on how our RationalGRL framework can be used.
-In this section, we present a more precise formalization of our framework.
-In the first two subsections we formalize a static representation of our framework: We develop a formal language to specify a GRL model (Section \ref{sect:goalmodeling:logicallanguage}) and we formalize the notion of an argument (Section \ref{sect:goalmodeling:argumentationsemantics}).
-In the third subsection we formalize the dynamics of our framework: we develop algorithms for instantiating argument schemes and for answering critical questions.
-The formal framework we develop in this section serves as the underlying model for the prototype implementation we present in the next section.
+In this section we present a formalization of RationalGRL based on formal logic.
+This is done for multiple reasons: (i) Most approaches in formal argumentation use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated in to valid GRL models in a very precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support.
 
-\subsection{Logical Language for RationalGRL}
+In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:goalmodeling:logicallanguage}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
+In the third subsection we turn to the dynamics of our framework.
+We develop algorithms for instantiating argument schemes and for answering critical questions.
+We show that each algorithm is correct, meaning that if the RationalGRL model was valid before applying the algorithm, it is also valid afterwards.
+
+\subsection{Formal Specification of GRL}
 \label{sect:goalmodeling:logicallanguage}
 
-We assume familiarity with basic notions from propositional logic and set theory.
-An \emph{atom} is a formula with no logical connectives ($\vee, \wedge, \rightarrow, \leftrightarrow$) or negation ($\neg$).
-We first define the logical language for RationalGRL.
+In this subsection we formalize a GRL model based on the GRL metamodel and the jUCMNav implementation.
+We first formalize elements of GRL (intentional elements and actors), and then formalize the links.
 
-\begin{definition}[RationalGRL Language]
-The \emph{RationalGRL language} $\mathcal{L}$ contains the following atoms:
+\subsubsection{Intentional Element and Actors}
+
+We start with the definition of various sets of elements that we use in subsequent definitions.
+
+\begin{definition}[Set definitions]
+\label{def:set-definitions}
+Throughout this section, we adopt the convention that variables start with a lowercase letter, and sets and constants start with an uppercase.
+\label{def:ie}
+We define the following sets:
 \begin{itemize}
-\item $actor(i)$: Identifier $i$ is an actor.
-\item $softgoal(i)$: Identifier $i$ is a softgoal.
-\item $goal(i)$: Identifier $i$ is a goal.
-\item $task(i)$: Identifier $i$ is a task.
-\item $resource(i)$: Identifier $i$ is a resource.
-\item $decomp(k,i,J,dtype)$: Identifier $k$ is a $dtype$-decomposition (and/or/xor) from the element corresponding to identifier $i$ to the set of elements corresponding to identifiers $J$.
-\item $dep(k,i,j)$: Identifier $k$ is a dependency link from the element corresponding to identifier $i$ to the element corresponding to identifier $j$.
-\item $contr(k,i,j,ctype)$: Identifier $k$ is a $ctype$-contribution (+/-) from the element corresponding to identifier $i$ to the element corresponding to identifier $j$.
-\item $has(i,j)$: Identifier $i$ (which is an actor) has the element corresponding to identifier $j$.
-\item $disabled(i)$: The element or relationships corresponding to identifier $i$ is disabled.
-\item $name(i,p)$: The name of the element corresponding to identifier $i$ is $p$.
-\end{itemize}
-
-We use integers $i,j,k\in\mathbb{N}$ as identifiers for actors and GRL elements. For all of the above definitions we assume that all identifiers used in each atom are different, i.e.,
-\begin{itemize}
-\item $k\not=i$,$k\not=i$,$k\not=j$
-\item forall $j\in J: k\not=j, i\not=j$
+\item $Type = \{Softgoal, Goal, Task, Resource\}$,
+\item $Names$ is a finite set of string.
+\item $DecompType = \{AND,OR,XOR\}$.
+\item $ContribValue = \{Break, Hurt, Some\ nega-$ $tive, Unknown, Some\  positive, Help, Make\},$
 \end{itemize}
@@
 
-Using the language from Definition 1, we can define GRL models. A \emph{ground atom} is an atom without variables.
+Next we define an intentional element.
 
-\begin{definition}[GRL model] \label{def:argument}
-A \emph{GRL model} $\mathcal{M}$ is a set of ground atoms based on $\mathcal{L}$.
+\begin{definition}[Intentional Element]
+\label{def:ie}
+An intentional element $ie\in \mathbb{N}\times Type\times Names\times DecompType$ is a relation, where $ie = (id, type, name, decomptype)$ means:
+\begin{itemize}
+\item $id\in \mathbb{N}$ is a unique identifier for the element,
+\item $type\in Type$ specifies the type of the element,
+\item $name \in Names$ is a string description of the element,
+\item $decomptype\in DecompType$ refers to the type of decomposition.\footnote{Note that the decomposition type is relevant only if the element is in fact decomposed into other elements, but since the jUCMNav implementation defines the decomposition type on the element, we do the same here.}
+\end{itemize}
+A set of intentional elements is denoted by $IE$.
 \end{definition}
 
-An example of the specification of the GRL model in Figure~\ref{fig:examples:decomposition} is shown in Table~\ref{table:grl_atom_spec}, written in logic programming style.
-In this article we do not make use of any logic programming techniques but this would be interesting future work (see Section~\ref{sect:goalmodeling:openissues}).
-A more elaborate example of a specification is shown in Appendix~\ref{ch:grlspec}, showing a complete specification of the traffic simulator GRL model in Figure~\ref{fig:trafficsim}.
+The definition above is sufficient to capture all intentional elements used in GRL.
+However, we present some syntactic sugar in the next definition by abbreviating the definition above in various ways.
+This does not add anything new to the previous definition, but it simplifies some of the notation.
 
-\begin{table}[h!]
-\centering
-\begin{tabularx}{0.4\textwidth}{|X|}
-\hline
-\texttt{goal(0).}\\
-\texttt{task(1).}\\
-\texttt{task(2).}\\
-\texttt{name(0,simulate).}\\
-\texttt{name(1,static\_simulation).}\\
-\texttt{name(2,dynamic\_simulation).}\\
-\texttt{decomp(3,0,[1,2],or).}\\
-\hline
-\end{tabularx}
-\caption{Specification of the GRL model in Figure~\ref{fig:examples:decomposition}}
-\label{table:grl_atom_spec}
-\end{table}
-
-\subsection{Formal argumentation}
-\label{sect:goalmodeling:argumentationsemantics}
-
-In the previous subsection we introduced an atomic language to specify a GRL model. In this subsection we give a formal definition of an argument, which is simply a set of atoms from our language.
-
-\begin{definition}[Argument] \label{def:argument}
-An \emph{argument} $A\subseteq Args$ is a set of ground atoms based on $\mathcal{L}$.
+\begin{definition}[Notation]
+\label{def:notation}
+We adopt three conventions simplifying our notation:
+\begin{itemize}
+\item
+We refer to the element of an intention element $ie$ using the dot (".") notation.
+That is, we may refer to the id, type, name and decomposition type with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
+\item 
+We refer to an intentional element with $ie.id=i$ simply by $ie_i$.\footnote{Note that the reference $ie_i$ is never ambiguous in GRL since identifiers for intentional elements are always unique (see Definition~\ref{def:grlmodel}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
+\item
+We can also refer to intention elements of a specific type simply by $type_{id}$.
+For instance, we can abbreviate the element in the previous item with $goal_0$ and write $goal_0.name = \text{Make profit}$, and $goal_0.decomptype = AND$.
+\end{itemize}
 \end{definition}
 
-This simple definition allows us to form arguments about (parts of) a GRL model.
-For instance, $$\{goal(0),name(0,development\_team)\},$$ is an argument stating that identifier 0 is a goal element with name ``development team''.
-Notice that an argument and a GRL model are essentially the same: as was discussed in Section \ref{sect:metamodel}, each element of a GRL model is an argument.
+\begin{definition}[Actor]
+\label{def:actor}
+An actor $act \in \mathbb{N}\times Names$ is a relation where $act=(id,name)$ means:
+\begin{itemize}
+\item $id\in\mathbb{N}$ refers to the unique identifier of the actor, 
+\item $name$ refers to a string description of its name.
+\end{itemize}
 
-Definition \ref{def:argument} does not put any restrictions on the content of an argument.
-This means it is possible to form arguments that do not correspond to correct or sensible GRL models.
-Consider for instance the following argument: $$\{goal(0), task(0), name(0,static\_simulation)\}.$$ This argument states that identifier 0 is both a goal and a task with name ``static simulation'', which is clearly undesirable since it does not correspond to a valid GRL model.
-While we could have chosen to put logical constraints on the content of an argument, we choose not to do so here.
-Instead, we put these restrictions on the algorithms in section~\ref{sect:algorithms}.
-The algorithms are constructed in such a way that they only allow the construction of arguments that correspond to valid GRL models.
-The advantage of this approach is that it keeps the presentation of the formal part relatively simple.
+Similar to intentional elements, we may refer to $act = (id,name)$ with $act_{id}$ and write $act_{id}.name$ to refer to its name.
 
-In line with the main approaches in formal argumentation, we next introduce an argumentation framework as a set of arguments and an attack relation between the arguments.
+A set of actors is denoted by $Act$.
+\end{definition}
 
-\begin{definition}[Argumentation framework~\cite{Dung1995}] \label{def:argumentationframework}
-An argumentation framework $AF=(Args,Att)$ consists of a set of arguments $Args$ and an attack relationship $Att:Args\times Args$, where $(A_1,A_2)\in Att$ means that argument $A_1\in Args$ attacks arguments $A_2\in Args$.
+In the next definition we specify relations between actors and intentional elements.
+Each actor can own one or multiple intentional elements.
+We chose to define this simply as a actor-element relation, which is in line with the jUCMNav implementation.
+
+\begin{definition}[Actor-IE Relations]
+\label{def:act-ie-relation}
+An \emph{Actor-IE relation} $r_{ActIE}\in\mathbb{N}\times\mathbb{N}$ is a relation $(i,j)$ meaning that an actor with id $i$ has intentional element with id $j$.
+
+A set of Actor-IE relations is denoted by $R_{ActIE}$.
 \end{definition}
+
+\subsubsection{Links}
+
+At this point we have defined all intentional elements in GRL and a containment relation between actors and intentional elements. We now turn to the GRL links.
+
+\begin{definition}[Contribution Link]
+\label{def:link:contrib}
+A \emph{contribution} link: $contrib\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}\times Contribvalue$ is a relation such that $contrib = (i,j,k,value)$ means
+\begin{itemize}
+\item $i$ is the unique identifier of the link,
+\item $j$ is the unique identifier of the IE from which the contribution originates,
+\item $k$ is the unique identifier of the IE to which is contributed.
+\end{itemize}
+Intuitively, $contrib = (i,j,k,value)$ means that $ie_i$ contributes to $ie_j$ with $value$.
+
+A set of contribution links is denoted by $Contrib$.
+\end{definition}
+
+\begin{definition}[Decomposition Link]
+\label{def:link:decomp}
+A \emph{decomposition} link: $decomp\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $decomp = (i,j,k)$ means
+\begin{itemize}
+\item $i$ is the unique identifier of the link,
+\item $j$ is the unique identifier of the decomposing IE,
+\item $k$ is the unique identifier of the IE that is being decomposed.
+\end{itemize}
+Intuitively, $decomp = (i,j,k)$ means that $ie_i$ decomposes $ie_j$.\footnote{Note that the decomposition type is defined on $IE_{k}$, see Definition~\ref{def:ie}.}
+
+A set of decomposition links is denoted by $Decomp$.
+\end{definition}
+
+\begin{definition}[Dependency Link]
+\label{def:link:dependency}
+A \emph{dependency} link: $dep\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $dep = (i,j,k)$ means
+\begin{itemize}
+\item $i$ is the unique identifier of the link,
+\item $j$ is the unique identifier of the dependee IE,
+\item $k$ is the unique identifier of the dependent IE.
+\end{itemize}
+Intuitively, $dep = (i,j,k)$ means that $ie_j$ depends on $ie_j$. \todo{Marc}{Sepideh}{Is this correct? For a dependency relation $A\rightarrow B$, who is dependent on who?}
+
+A set of dependency links is denoted by $Dep$.
+\end{definition}
+
+Similar to intentional elements and actors, we may refer to links with an id using the subscript notation, e.g., $contrib_0$ is a contribution link with $contrib.id = 0$.
+
+The definitions above come with various implicit assumptions in order to form a valid GRL model.
+For instance, if an Actor-IE relation $(i,j)$ occurs in a GRL model, then there should exist an actor with id $i$, and there also should exist an IE with id $j$.
+We make these assumptions explicit in the following definition.
+
+\begin{definition}[GRL Model]
+\label{def:grl-model}
+A \emph{GRL model} $GRL=(IE, Act, R_{ActIE}, Contr, Decomp, Dep)$ consists of:
+\begin{itemize}
+\item A set $IE$ of intentional elements (Def.~\ref{def:ie}),
+\item A set $Act$ of actors (Def.~\ref{def:actor}),
+\item A set $R_{ActIE}$ of Actor-IE relations (Def.~\ref{def:act-ie-relation}),
+\item A set $Contr$ of contribution links (Def.~\ref{def:link:contrib}),
+\item A set $Decomp$ of decomposition links (Def.~\ref{def:link:decomp}),
+\item A set $Dep$ of dependency links (Def.~\ref{def:link:dependency}).
+\end{itemize}
+
+Furthermore, the following conditions are satisfied:
+\begin{enumerate}
+\item ids are globally unique across IEs, Links, and Actors, i.e., let $X,Y\in \{IE,Act, Contr, Decomp, Dep\}$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
+\item All intentional elements of actors exist: $\forall (i,j)\in R_{ActIE}: act_i \in Act \wedge ie_j \in IE$.
+\item An intentional element belongs at most to one actor: $\forall ie_i\in IE: |\{(k,i)\in R_{ActIE}\}| \le 1$.
+\item Contribution links connect intentional elements: $\forall (i,j,k,value)\in Contrib: \{ie_j,ie_k\}\subseteq IE$.
+\item Decomposition links connect intentional elements: $\forall (i,j,k)\in Decomp: \{ie_j,ie_k\}\subseteq IE$.
+\item Dependency links connect intentional elements: $\forall (i,j,)\in Dep: \{ie_j,ie_k\}\subseteq IE$.
+\end{enumerate}
+\end{definition}
+
+\todo{Marc}{Marc}{Provide explanation and example here.}
+
+\subsection{Formal specification of RationalGRL}
+
+In order to develop a logical framework for RationalGRL, we extend the GRL logical framework of the previous section by adding three elements (see Figure~\ref{fig:rationalgrllegend}):
+\begin{itemize}
+\item A new element called \emph{generic argument},
+\item A new link called \emph{attack link},
+\item An labeling of the elements, assigning to each element (an IE, link, or generic argument) the label IN (accepted) or OUT (reject).
+\end{itemize}
+
+In the next section we explain how the labeling is computed, but in this section we only provide a formal definition.
+
+We start with the new element which we call the \emph{Generic Argument}.
+
+\begin{definition}[Generic Argument]
+\label{def:generic-argument}
+A generic argument $arg \in \mathbb{N}\times Names$ is a relation such that $arg=(id, name)$ is an argument with unique identifier $id$ and name $name$.
+We may refer to the argument with id $i$ simply with $arg_i$.
+A set of arguments is denoted by $Arg$.
+\end{definition}
+
+Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link:contrib},~\ref{def:link:decomp}, and~\ref{def:link:dependency}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links.
+This is correct, since generic arguments are not part of GRL, so they should also not occur in links.
+
+We next define the attack link, which is the only link that RationalGRL adds to GRL.
+
+\begin{definition}[Attack Link]
+\label{def:link:attack}
+An attack link $att\in \mathbb{N}\times \mathbb{N}$ is a relation such that $att=(i,j)$ means:
+\begin{itemize}
+\item $i$ is the unique identifier of the element that is performing that attack,
+\item $j$ is the unique identifier of the element that is being attacked.
+\end{itemize}
+Intuitively, $att=(i,j)$ means that element $i$ is attacking element $j$.
+\end{definition}
+
+Note that the previous definition did not put any constraint on which elements can be connected with attack links. This is what we make explicit in the definition of a RationalGRL model later.
+
+We next add the last part of the RationalGRL addition, namely the labeling of elements.
+This labeling determines whether an element is accepted (which means it shows up in the GRL model if it is an IE or a link), or rejected (which means it won't be shown).
+
+\begin{definition}[RationalGRL labeling]
+\label{def:labeling}
+A RationalGRL labeling function $Lab:\mathcal{N}\rightarrow {in,out}$ that is total on the ids of the RationalGRL elements.
+This means that $Lab(id)$ is defined for each $ie.id, contr.id, decomp.id, dep.id, arg.id$.
+\end{definition}
+
+\begin{definition}[RationalGRL Model]
+\label{def:rationalgrl-model}
+A \emph{RationalGRL model} $RationalGRL=(IE, Act, R_{ActIE}, Contr, Decomp, Dep, Arg, Att, Label)$ consists of:
+\begin{itemize}
+\item A GRL model $(IE, Act, R_{ActIE}, Contr, Decomp,$ $Dep)$
+\item A set of generic arguments $Arg$ (Def.~\ref{def:generic-argument}),
+\item A set of attack links (Def.~\ref{def:link:attack}).
+\item A labeling function $Lab$ (Def.~\ref{def:labeling}).
+\end{itemize}
+
+Furthermore, the following conditions are satisfied. In the following, let $Elems = \{IE, Act, Contr, Decomp, Dep, Arg\}$:
+\begin{enumerate}
+\item All conditions on a GRL model are satisfied, (Def.~\ref{def:grl-model}),
+\item ids are globally unique across IEs, Links, Actors, and Generic arguments, i.e., let $X,Y\in Elems$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
+\item Attack links connect generic arguments, IEs, actors, and links with each other, i.e., let   $\forall (i,j)\in Att: \exists X\in Elems$ s.t. $X_i\in X$ and $\exists Y\in Elems$ s.t. $Y_i\in Y$.
+\end{enumerate}
+\end{definition}
+
+\todo{Marc}{Marc}{Explain conditions here and give some examples}
+
+\subsubsection{Argumentation semantics}
 
@@
 
-\subsection{Algorithms for argument schemes and critical questions}
-\label[sect:algorithms}
+\todo{Marc}{Marc}{Give example with concrete RationalGRL model}
 
-Let us briefly take stock:
-\begin{itemize}
-\item We developed an initial set of argument schemes and critical questions in Section~\ref{sect:gmas}, by annotating transcripts.
-\item We developed a visual notation to illustrate some argument schemes and critical questions in Section~\ref{sect:examples}.
-\item We proposed a logical language consisting of atomic sentence to describe a GRL model in Section~\ref{sect:goalmodeling:logicallanguage}.
-\item We formalized an ``argument'' as a set of atoms from our language, and we introduced \emph{argumentation semantics} to compute sets of accepted arguments in Section~\ref{sect:goalmodeling:argumentationsemantics}.
-\end{itemize}
+\subsection{Algorithms for argument schemes}
+\label{sect:algorithms}
 
-In this subsection we develop algorithms for applying argument schemes and critical questions.
-The process of applying these algorithms and generating goal models from them is shown in context of the RationalGRL methodology in Figure~\ref{fig:methodologyalgorithm} \todo{Marc}{Marc}{Create an extended version of the methodology, containing where which parts of the formal framework are used.} These algorithms produce arguments (Definition~\ref{def:argument}) and attack relations, forming an argumentation framework (Definition~\ref{def:argumentationframework}).
-One can then use argumentation semantics (Definition~\ref{def:semantics}) to compute sets of accepted arguments.
+In the previous section we formalize a \emph{static} representation of the RationalGRL framework.
+In this section we formalize the \emph{dynamics}.
+We do so by developing algorithms for applying argument schemes and critical questions in the context of a RationalGRL model~\ref{def:rationalgrl-model}.
+These algorithms produce new elements and attack relations.
+We can then use argumentation semantics (Definition~\ref{def:semantics}) to compute sets of accepted arguments.
 The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.
 
-In the following algorithms, we assume the following global variables:
-
-\begin{definition}[Global variables] The following variables are intended to have a \emph{global} scope and have a \emph{static} value, meaning the lifetime of the variable extends across the entire run of the program.
-\begin{itemize}
-\item $id$: the current highest identifier of the elements. This variable is increased for each new element that is added.
-\item $Args$: contains the set of all arguments.
-\item $Att$: contains the set of all attack relations.
-\end{itemize}
-\end{definition}
-
-\subsubsection*{Algorithms for argument schemes}
 The algorithms for the argument schemes consist of forming a new argument and adding it to the set of arguments. No attack relations are introduced.
@@
   \begin{algorithmic}[1]
+    \State $RationalGRL=(IE, Act, R_{ActIE}, Contr, Decomp,$
+    \State \qquad\qquad\qquad\qquad$Dep, Arg, Att, Label)$
     \Procedure{$AS_0$}{$a$}
-    \State $id\gets id+1$\label{alg:as0:1}
-    \State $A \gets \{actor(id), name(id,a)\}$\label{alg:as0:2}
-    \State $Args \gets Args \cup A$\label{alg:as0:3}
+    \State $id\gets mintId()$\label{alg:as0:1}
+    \State $Act \gets Act \cup (id, a)$\label{alg:as0:2}
     \EndProcedure
@@
 
-\emph{Algortihm~\ref{alg:as0} for argument scheme AS0}: The algorithm takes one argument, namely the name of the actor $a$.
-On line~\ref{alg:as0:1} of the algorithm, the global variable $id$ is increased by one.
-This ensures that each new argument has a unique identifier.
-On line~\ref{alg:as0:2}, the argument for the actor is formed, consisting of two statements, stating respectively that $id$ is an actor, and that the name of this actor is $a$.
-Finally, on line~\ref{alg:as0:3} this argument is added to the global set of arguments $Args$.
+\subsubsection{Algortihm~\ref{alg:as0} for argument scheme AS0}
+
+\paragraph{Explanation} For all of the argument schemes and critical questions algorithms, we assume an underlying RationalGRL model as defined in Def.~\ref{def:rationalgrl-model}.
+This model is shown in the first two lines.
+
+The algorithm takes one argument, namely the name of the actor $a$.
+On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted.
+On line~\ref{alg:as0:2}, the the tuple $(id,a)$ is added to the set of actors $Act$, meaning that a new actor with id $id$ and name $a$ is added to the RationalGRL model (Def.~\ref{def:actor}).
 In Figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_0(\text{Development team})$, results in one argument: $$Args = \{\{actor(0), name(0, \text{Development team})\}\}.$$
@@
 
-\emph{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
+\paragraph{Example} \emph{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
 The algorithm is similar to the previous one, with the difference that the newly added argument contains the statement $has(a_{id},id)$ as well, meaning that the actor with id $a_{id}$ has element $id$ (which is a resource).
@@
 First, we assume some id is associated with the actor: $$\{actor(0),name(0,traffic\_tycoon)\}.$$ Then we can formalize the argument for the resource as follows: $$\{resource(1),name(1,external\_library),has(0,1)\}.$$
+
+\paragraph{Proof of Correctness} In order to show that this algorithm is correct, we should show that the new RationalGRL model still satisfies all the conditions of Def.~\ref{def:grl-model}.
+This is clearly the case, since adding a single element does not invalidate any condition.
 
@@
 
-\emph{Algorithm~\ref{alg:as5} for argument scheme AS5:} The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
+\subsubsection{Algorithm~\ref{alg:as5} for argument scheme AS5}
+
+\paragraph{Explanation} The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
 The difficulty of this algorithm is that each of the tasks are stated in natural language, and it is not directly clear whether these tasks are already in the GRL model or not.
@@
 
+\paragraph{Example}
 Let us explain this algorithm with the XOR-decomposition of goal \texttt{Generate cars} of Figure~\ref{fig:transcripts:grl}. Suppose the following arguments are constructed already:
@@
 
-\emph{Algorithm~\ref{alg:as6} for argument scheme AS6:} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
+\paragraph{Proof of Correctness}
+
+\subsubsection{Algorithm~\ref{alg:as6} for argument scheme AS6}
+
+\paragraph{Explanation} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
 The idea behind this algorithm is very similar to the previous one, with the difference that in the current algorithm we create a single relation, while we created a set of relations in the previous algorithm.
@@
 
-Let us again illustrate this with a simple example from Figure~\ref{fig:transcripts:grl}.
+\paragraph{Example} Let us again illustrate this with a simple example from Figure~\ref{fig:transcripts:grl}.
 Suppose the following argument exists already: $\{task(0),name(0,keep\_same\_cars\}$, and suppose someone would like to add an argument that the task \texttt{Keep same cars} contributes positively to softgoal \texttt{Dynamic simulation}, i.e.
@@
 Again, we assume the reader can reconstruct them straightforwardly.
+
+\paragraph{Proof of correctness}
 
@@
 \subsection{Constructing GRL models}
-
+\label{sect:rationalGRL-GRL}
 \todo{F}{F,M}{It would be good to have a formal definition of how to construct models given an argumentation framework.

### 2017-10-08 discussion.tex
1st update related research

@@
 \label{sect:goalmodeling:relatedwork}
+
+The need for justifications of modeling choices plays an important role in different requirements engineering methods using goal models.
+The belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} is meant to capture the rationales behind the inclusion of goals and tasks in the model.
+Furthermore, relations between elements in a goal model also provide justifications: high-level goals are reasons for lower-level goals, tasks and resources.
+Hence, refinement and decomposition techniques used in requirements engineering \cite{van2001goal} can be seen as explicit argumentation steps in the goal modeling process.
+Take, for example, CQ2 of PRAS (Section~\ref{sect:background:pras}), which asks whether there are alternative ways of realizing the same goal.
+Providing an alternative sub-goal or -task in a goal model then an explicit argumentative move in the discussion.
+
+So in a sense, a goal model provides a justification for itself, particularly if we include belief elements for extra design rationalization.
+This idea is also prevalent in our RationalGRL framework: many arguments are in fact GRL elements, and many critical questions can be answered by introducing (\textsf{INTRO}) new GRL elements.
+However, as was already argued in Section~\ref{sect:introduction} (and also by \cite{Jureta:RE2008}), argumentation produces different, richer and complementary information to just the goal model.
+The goal model is the product of a process of argumentation, and does not include, for example, goals and tasks that were at some point considered discarded.
+Furthermore, for goal models it is only possible to determine the satisfiability of goals given the possible tasks and resources; what cannot be determined is the acceptability of goals, that is, whether they are acceptable given potentially contradictory opinions of stakeholders.
+
+Closely related to goal modeling and design justification are frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
+DR looks at issues, options and arguments for and against these options in the design of, for example, a software system.
+The similarity with goal modeling is that much of the traditional DR literature provides modeling languages and tool support for building DR diagrams.
+The differences lie in the concepts used: in DR, the focus is more on the rationales, why certain design options were chosen, than on the functional and non-functional requirements of the system, which are expressed in goal models.
+The RationalGRL framework in a sense incorporates the ideas from DR into goal models by including arguments and issues (critical questions) into the RatinalGRL model.
+
+Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
+However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
+Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
+Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
+
+There are several contributions to the literature that relate argumentation-based techniques with goal modeling.
+The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
+Jureta \emph{et al.} propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
+One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
+In this sense, our RationalGRL framework can be seen as an instantiation and implementation of  part of the GAM.
+The main difference between our approach and GAM is that we integrate arguments and goal models using argument schemes, and that we develop these argument schemes by analyzing transcripts.
+GAM instead uses structured argumentation.
 
@@
 Connecting argumentation and goal modelling by providing a mapping between two types of diagram was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
-One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion thus far has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
+One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
 Furthermore, in our previous work we did not provide the specific critical questions for goal models (see Section \ref{sect:gmas}).
@@
 Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling (requirement 4).
-
-The need for justifications of modeling choices plays an important role in different requirements engineering methods using goal models.
-High-level goals are often understood as reasons for representing lower-level goals (in other words, the need for low-level goals is justified by having high-level goals) and other elements in a goal model such as tasks and resources.
-Various refinements and decomposition techniques,  often used in requirements engineer (See~\cite{van2001goal} for an overview), can be seen as incorporating argumentation and justification, in that sub-goals could be understood as arguments supporting parent goals.
-In that case, a refinement alternative is justified if there are no conflicts between sub-goals (i.e., it is consistent), as few obstacles as possible sub-goals harm sub-goal achievement, there are no superfluous sub-goals (the refinement is minimal), and the achievement of sub-goals can be verified to lead to achieving the parent goal (if refinement is formal~\cite{Darimont:1996:FRP:239098.239131}).
-This interpretation is one of the founding ideas of goal modeling.
-However, while this interpretation may seem satisfactory, argumentation and justification processes differ from and are complementary to refinement in several respects, such as limited possibilities for rationalization and lack of semantics (see Jureta~\cite{Jureta:RE2008} for more details).
-
-There are several contributions that relate argumentation-based techniques with goal modeling.
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-Jureta \emph{et al.} propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our RationalGRL framework can be seen as an instantiation and implementation of  part of the GAM.
-The main difference between our approach and GAM is that we integrate arguments and goal models using argument schemes, and that we develop these argument schemes by analyzing transcripts.
-GAM instead uses structured argumentation.
-
-The RationalGRL framework is also closely related to frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against the various options in the design of, for example, a software system, and provides direct tool support for building and analyzing DR graphs.
-One of the main improvements of RationalGRL over DR approaches is that RationalGRL incorporates the formal semantics for both argument acceptability and goal satisfiability, which allow for a partly automated evaluation of goals and the rationales for these goals.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
 
@@
 \item more details on Murukannaiah [RE15]
-\item related work on argumentation theory
+\item related work on argumentation theory - check Serena Villata on goal models
+\item Prakken and Wierenga
 \item First paper we wrote they also said something about Bashar Nuseibeh
-\item Also mention argumentation in AI and how that relates to our work
-\item Also paper by Daniel Amyot that he once sent to us:
-  "Synergy between Activity Theory and goal/scenario modeling for requirements elicitation, analysis, and evolution"
+\item Also mention argumentation in AI - arg diagramming - and how that relates to our work
+\item Software design discussions Black et al.
 \item Maybe this one:
@@
   Anitha Murugesan, Michael Whalen, Elaheh Ghassabani and Mats Heimdahl
-  (University of Minnesota, United States)
-
-\item  At least we can say our work is different from this if it is unrelated..
-\item We should make the comparison more in depth.
-\item Look at related work of Jureta and see how it is related with us.
-\item Maybe the papers above (e.g., Bashar) have related work that relates to us as well.
   \end{itemize}

### 2017-10-08 formalframework.tex
1st update related research

--- 2017-10-02 formalframework.tex
@@
 After applying $DISABLE(0)$ we obtain the arguments: $$Args=\{\{actor(0),name(0,dev\_team\}, \{disabled(0)\}\}.$$ Note that our implementation of this attack does not lead to an actual attack in the argumentation framework.
-
+\todo{F}{M}{algoritmes voor het introduceren van beliefs als je antwoord geeft op een CQ}
 \begin{algorithm}[h]
@@
 
-\begin{definition}[RationalGRL model] A RationalGRL model \mathcal{R} is an argumentation framework $AF=(Args,Att)$ where all arguments in $Args$ and attacks in $Att$ are based on algorithm X-XX.  
+\begin{definition}[RationalGRL model] A RationalGRL model $\mathcal{R}$ is an argumentation framework $AF=(Args,Att)$ where all arguments in $Args$ and attacks in $Att$ are based on algorithm X-XX.  
 \end{definition}

### 2017-10-08 formalframework.tex

--- 2017-10-08 formalframework.tex
@@
 \label{sect:rationalGRL-GRL}
+\todo{F}{F,M}{Dit is snel getypt, moet nog checken of het klopt.}
+
+\begin{definition}[RationalGRL model] A RationalGRL model $\mathcal{R}$ is an argumentation framework $AF=(Args,Att)$ where all arguments in $Args$ and attacks in $Att$ are based on algorithm X-XX.  
+\end{definition}
+
+\begin{definition}[GRL model] A GRL model $\mathcal{M}$ based on a RationalGRL model $\mathcal{R}$ is the preferred extension of $\mathcal{R}$.  
+\end{definition}
+=======
 \todo{F}{F,M}{It would be good to have a formal definition of how to construct models given an argumentation framework.

### 2017-10-08 formalframework.tex
updated into formal framework

@@
 We then presented the RationalGRL framework in secton~\ref{sect:overview}, consisting of a modeling language, a metamodel, and various examples of using the argument schemes and critical questions.
-In section~\ref{sect:methodology} we explained how this framework can be used by practitioners by explaining the methodology and tool support.
+In section~\ref{sect:meth+tool} we explained how this framework can be used by practitioners by explaining the methodology and tool support.
 
@@
 In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:goalmodeling:logicallanguage}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
-In the third subsection we turn to the dynamics of our framework.
-We develop algorithms for instantiating argument schemes and for answering critical questions.
-We show that each algorithm is correct, meaning that if the RationalGRL model was valid before applying the algorithm, it is also valid afterwards.
+We then provide a formal translation procedure from a RationalGRL model to a GRL model.
+
+In the third subsection we turn to the dynamics of our framework. We develop algorithms for instantiating argument schemes and for answering critical questions.
+
+The two main results of this section are that (i) the translation procedure from a RationalGRL model to a GRL model is correct, meaning that if the RationalGRL model is valid, the resulting GRL model is also valid, (ii) the algorithms for applying argument schemes and critical questions are correct as well, meaning that if the Rational model is valid before, it is also valid afterwards.
+These two results together entail that our framework and the corresponding operations on it are consistent.
 
@@
 
-We start with the definition of various sets of elements that we use in subsequent definitions.
+We start with some general definition that we use in subsequent definitions.
 
-\begin{definition}[Set definitions]
+\begin{definition}[General definitions]
 \label{def:set-definitions}
-Throughout this section, we adopt the convention that variables start with a lowercase letter, and sets and constants start with an uppercase.
-\label{def:ie}
+Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$, and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
+
 We define the following sets:
@@
 \end{definition}
+
+\paragraph{Example} \todo{Marc}{Marc}{Give example here}
 

### 2017-10-09 formalframework.tex
improved formalisation of RationalGRL

@@
 \item An labeling of the elements, assigning to each element (an IE, link, or generic argument) the label IN (accepted) or OUT (reject).
+In the next section we explain how the labeling is computed, but in this section we only provide a formal definition.
 \end{itemize}
 
-In the next section we explain how the labeling is computed, but in this section we only provide a formal definition.
+Furthermore, we also change the representation of IE, links, and actors.
+In GRL, these elements have a single identifier.
+However, in RationalGRL, we allow the user to \emph{replace} such elements when answering a critical question with effect REPLACE.
+Therefore, for each GRL element, we maintain a history, which is a list in which the last element of the list corresponds to the element that ends up in the GRL model.
 
@@
 
-We next add the last part of the RationalGRL addition, namely the labeling of elements.
+We next add the the labeling of elements.
 This labeling determines whether an element is accepted (which means it shows up in the GRL model if it is an IE or a link), or rejected (which means it won't be shown).
@@
 
+Finally we add the last part of the RationalGRL extension. We define a list of intentional elements, contribution links, and actors.
+
+\begin{definition}[IE History, Contribution History, Actor History]
+\label{def:element-list}
+\begin{itemize}
+\item An \emph{IE History for element with id $i$}, denoted $\overline{ie}_i=(ie_0,\ldots,ie_n)$, is a list of IEs $ie_0,\ldots,ie_n$ such that $ie_1.id=\ldots=ie_n.id=i$. 
+\item An \emph{Actor history for actor with id $j$}, denoted $\overline{act}_j=(actor_0,\ldots,actor_n)$, is a list of actors $actor_0,\ldots,actor_n$ such that $actor_0.id=\ldots=actor_n.id=j$. 
+\item A \emph{Contribution history for link with id $k$}, denoted $\overline{contrib}_j=(contrib_0,\ldots,contrib_n)$, is a list of contribution links $contrib_0,\ldots,contrib_n$ such that $contrib_0.id=\ldots=contrib_n.id=k$.
+\end{itemize}
+If the id in the list is not relevant we may omit the subscript notation, e.g., we may write $\overline{ie}$ instead of $\overline{ie}_i$.
+
+Sets of histories for IEs, actors, and contribution links are denoted with respectively $\overline{IE}$, $\overline{Act}$, and $\overline{Contrib}$.
+
+The last element of a list is called the \emph{head element}.
+\end{definition}
+
+Note that we do not require a history for decomposition links or dependency links. This is because these links do not have associated REPLACE operations.
+
 \begin{definition}[RationalGRL Model]
 \label{def:rationalgrl-model}
-A \emph{RationalGRL model} $RationalGRL=(IE, Act, R_{ActIE}, Contr, Decomp, Dep, Arg, Att, Label)$ consists of:
+A \emph{RationalGRL model} $RationalGRL=(\overline{IE}, \overline{Act}, R_{ActIE}, \overline{Contrib}, Decomp, Dep, Arg, Att, Label)$ consists of:
 \begin{itemize}
-\item A GRL model $(IE, Act, R_{ActIE}, Contr, Decomp,$ $Dep)$
+\item A set $\overline{IE}$ of lists of intentional elements (Def.~\ref{def:element-list}),
+\item A set $\overline{Act}$ of lists of actors (Def.~\ref{def:element-list}),
+\item A set $R_{ActIE}$ of Actor-IE relations (Def.~\ref{def:act-ie-relation}),
+\item A set $\overline{Contr}$ of lists of contribution links (Def.~\ref{def:element-list}),
+\item A set $Decomp$ of decomposition links (Def.~\ref{def:link:decomp}),
+\item A set $Dep$ of dependency links (Def.~\ref{def:link:dependency}).
 \item A set of generic arguments $Arg$ (Def.~\ref{def:generic-argument}),
@@
 \begin{enumerate}
-\item All conditions on a GRL model are satisfied, (Def.~\ref{def:grl-model}),
-\item ids are globally unique across IEs, Links, Actors, and Generic arguments, i.e., let $X,Y\in Elems$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
+\item ids are globally unique across IEs, Links, and Actors, i.e., let $X,Y\in \{\overline{IE},\overline{Act}, \overline{Contr}, Decomp, Dep\}$.
+For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
+\item All intentional elements of actors exist: $\forall (i,j)\in R_{ActIE}: \overline{act}_i \in \overline{Act} \wedge ie_j \in IE$.
+\item An intentional element belongs at most to one actor: $\forall ie_i\in IE: |\{(k,i)\in R_{ActIE}\}| \le 1$.
+\item Contribution links connect intentional elements: $\forall (i,j,k,value)\in \overline{Contrib}: \{ie_j,ie_k\}\subseteq IE$.
+\item Decomposition links connect intentional elements: $\forall (i,j,k)\in Decomp: \{ie_j,ie_k\}\subseteq IE$.
+\item Dependency links connect intentional elements: $\forall (i,j,)\in Dep: \{ie_j,ie_k\}\subseteq IE$.
 \item Attack links connect generic arguments, IEs, actors, and links with each other, i.e., let   $\forall (i,j)\in Att: \exists X\in Elems$ s.t. $X_i\in X$ and $\exists Y\in Elems$ s.t. $Y_i\in Y$.
@@
 
+\begin{algorithm}[h]
+  \caption{Translation Algorithm}\label{alg:translation}
+  \begin{algorithmic}[1]
+    \State TODO
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{theorem}
+If a RationalGRL model $m$ satisfies the conditions of a RationalGRL model (Def.~\ref{def:rationalgrl-model}), then $Trans(m)$ satisfies the conditions of a GRL model (Def.~\ref{def:grl-model}).
+\end{theorem}
+
+\begin{proof}
+TODO
+\end{proof}
+
 \subsection{Algorithms for argument schemes}
@@
 
-In the previous section we formalize a \emph{static} representation of the RationalGRL framework.
+In the previous section we formalized a \emph{static} representation of the RationalGRL framework.
 In this section we formalize the \emph{dynamics}.
-We do so by developing algorithms for applying argument schemes and critical questions in the context of a RationalGRL model~\ref{def:rationalgrl-model}.
+We do so by developing algorithms for applying argument schemes and critical questions in the context of a RationalGRL model (Def.~\ref{def:rationalgrl-model}).
 These algorithms produce new elements and attack relations.
-We can then use argumentation semantics (Definition~\ref{def:semantics}) to compute sets of accepted arguments.
+We can then use argumentation semantics (Def.~\ref{def:semantics}) to compute sets of accepted arguments.
 The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.
@@
   \begin{algorithmic}[1]
-    \State $RationalGRL=(IE, Act, R_{ActIE}, Contr, Decomp,$
+    \State $RationalGRL=(\overline{IE}, \overline{Act}, R_{ActIE}, \overline{Contr}, Decomp,$
     \State \qquad\qquad\qquad\qquad$Dep, Arg, Att, Label)$
@@
     \State $id\gets mintId()$\label{alg:as0:1}
-    \State $Act \gets Act \cup (id, a)$\label{alg:as0:2}
+    \State $\overline{Act} \gets \overline{Act} \cup ((id, a))$\label{alg:as0:2}
     \EndProcedure
@@
 On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted.
-On line~\ref{alg:as0:2}, the the tuple $(id,a)$ is added to the set of actors $Act$, meaning that a new actor with id $id$ and name $a$ is added to the RationalGRL model (Def.~\ref{def:actor}).
-In Figure~\ref{fig:examples:relevant-actor2}, the application of the argument scheme $AS_0(\text{Development team})$, results in one argument: $$Args = \{\{actor(0), name(0, \text{Development team})\}\}.$$
+On line~\ref{alg:as0:2}, then a list containing a single tuple $(id,a)$ is added to the set of lists of actors $\overline{Act}$, meaning that a new actor with id $id$ and name $a$ is added to the RationalGRL model (Def.~\ref{def:element-list}).
+In Figure~\ref{fig:examples:relevant-actor2}, ..
+TODO
 
 \begin{algorithm}[h]
-  \caption{Applying AS1: Actor $a_{id}$ has resource $n$}\label{alg:as1}
+  \caption{Applying AS1: Actor $a_{id}$ has resource $r_{id}$}\label{alg:as1}
   \begin{algorithmic}[1]
-    \Procedure{$AS_1$}{$a_{id}, n$}
-    \State $id\gets id+1$
-    \State $A\gets \{resource(id), name(id,n), has(a_{id},id)\}$
-    \State $Args \gets Args\cup A$
+     \State $RationalGRL=(\overline{IE}, \overline{Act}, R_{ActIE}, \overline{Contr}, Decomp,$
+    \State \qquad\qquad\qquad\qquad$Dep, Arg, Att, Label)$
+    \Procedure{$AS_1$}{$a_{id}, r_{id}$}
+    \State $R_{ActIE}\gets (a_{id}, r_{id}$
     \EndProcedure
@@
 
-\paragraph{Example} \emph{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
+\paragraph{Example} \todo{Marc}{Marc}{Clarify that the actor and resource exist already. If not, we should first execute Algorithm 2, and then Algorithm 3. No need to combine them}
+
+\emph{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
 The algorithm is similar to the previous one, with the difference that the newly added argument contains the statement $has(a_{id},id)$ as well, meaning that the actor with id $a_{id}$ has element $id$ (which is a resource).
@@
 
+\todo{Marc}{Marc}{This algorithm does not make much sense to me. We simply add a decomposition link, no need to make it so complicated. I actually think we can omit this algorithm.}
+
 \subsubsection{Algorithm~\ref{alg:as5} for argument scheme AS5}
 
-\paragraph{Explanation} The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
+\paragraph{Explanation} \todo{Marc}{Marc}{I think we should remove all this stuff. In this part we are simply adding GRL elements, which is trivial. The interesting part are the critical questions}
+
+The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
 The difficulty of this algorithm is that each of the tasks are stated in natural language, and it is not directly clear whether these tasks are already in the GRL model or not.
@@
 
+\todo{Marc}{Marc}{This one is also not interesting.. just add a contribution link}
 \paragraph{Explanation} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
@@
 \end{algorithm}
+
+\todo{Marc}{Marc}{Here we would add an argument and an attack from the argument to the intentional element. We don't add ``disabled'' because this is what argumentation semantics should do for us.}
 

### 2017-10-10 discussion.tex
Related work & references cleanup

@@
 
-The need for justifications of modeling choices plays an important role in different requirements engineering methods using goal models.
-The belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} is meant to capture the rationales behind the inclusion of goals and tasks in the model.
+\paragraph{Design Rationale} Argumentation in software design has for some time now been the subject of the work on so-called \emph{design rationale} (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or software architecture.
+DR looks at issues, options and arguments for and against these options in the design of, for example, a software system.
+Similar to the literature on goal modeling, much of the traditional DR literature provides modeling languages and diagramming tool support for building design rationales.
+It is in this diagramming functionality that the link with argument diagrams from philosophy, law and AI~\cite{reed2004araucaria,gordon2007visualizing} has been made, where argument diagrams represent reasoning from premises to conclusions.
+More recent work on DR moves away from the idea that all decision have to be explicitly diagrammed and focuses more on empirically investigating how critical reflection can help when designing \cite{razavian2016two,SchriekEtal2016}, or which parts of the design process are best explicitly documented \cite{falessi2013value}.
+
+Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE.
+The work on the RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments pro and con the various options into the goal model, and by proposing a methodology and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
+
+\paragraph{Requirements Engineering} There are a number of general approaches in the field of requirements engineering that explicitly take into account arguments.
+One early example comes from Haley et al.~\cite{haley2008security}, who use formal logical arguments to show that the system behavior satisfies certain security requirements, and more informal arguments to capture and validate the assumptions underlying the system behavior.
+This system behavior is defined by the tasks it executes and thus arguments are given for and against system tasks, similar to the way beliefs and counterarguments can be provided for tasks in RationalGRL.
+What Haley et al.
+leave implicit in their argumentation are the goals of the stakeholders on which the system tasks depend -- they include the goals in their framework and mention that there will often be conflicting goals between stakeholders, but do not explicitly model them.
+Furthermore, the argumentative part of their framework does not include formal semantics for resolving conflict between arguments or determining the acceptability of arguments.
+Yu et al.~\cite{yu2015automated} further extend the framework by Haley et al., including algorithms for Dung-style \cite{Dung1995} argumentation semantics and a database of specific ways in which to attack (or mitigate) risks, which can be likened to a set of critical questions for risks and security requirements (cf.
+Yu et al.~\cite{yu2015automated} Section 3.1).
+
+Another recent example of the use of arguments in goal-oriented requirements engineering is the work by Murukannaiah et al.~\cite{murukannaiah2015}, who propose Arg-ACH, an approach 
+in which the beliefs underlying conflicting goals can be made explicit using argumentation.
+Murukannaiah et al.
+start with the basic technique of Analysis of Competing Hypotheses (ACH), where for conflicting goals the beliefs that are consistent and inconsistent with these goals are included in a matrix and counted.
+They then extend this technique into Arg-ACH: instead of just indicating whether a belief is consistent or inconsistent with a goal, each belief becomes an argument for or against the goal, which is then diagrammed using the Carneades tool \cite{gordon2007visualizing}.
+Belief scores are assigned to arguments, which can be aggregated to provide one's belief in a goal.
+The arguments for and against goals can be based on argument schemes, and critical questions can be used to find new arguments for or against the goals or the existing arguments.
+One example provided in the paper is the argument scheme from expert opinion, which allows one to draw conclusions based on expert statements and subsequently question, for example, the objectivity and veracity of the expert using critical questions.
+Murukannaiah et al.
+conducted an experiment in which they had two groups, one with ACH and one with Arg-ACH, perform an analysis of several conflicting goals regarding security at transport hubs.
+They found that, while the group that used Arg-ACH took longer, they also covered more possibilities in their belief search and the conclusions were more consistent among the group.
+
+One other example of argumentation in RE concerns the use of argumentation in requirements elicitation.
+Ionita et al.~\cite{ionita2014argumentation} propose an simple argumentation dialogue game in which risk assessors try to attack each other's arguments for certain risks attached to a system design.
+Dung's semantics \cite{Dung1995} are then used to determine the risks that are still valid, and those that have been successfully rejected.
+Elrakaiby et al.~\cite{ElrakaibyFSGN17} use argumentation to explain ambiguity.
+They define when a statement by a client who is being interviewed about the requirements of a system presents an inconsistency (either with the client's previous statements or with the requirement engineer's beliefs) or an insufficiency, that is, when an analyst needs more information from the client to accept a client's statement.
+The inconsistencies are then captured as mutually attacking arguments, and the insufficiencies as arguments against the original statements saying that, for example, the functionality expressed in the statement cannot be realised or is irrelevant.
+Elrakaiby et al.
+coded the data from 34 requirement elicitation interviews, identifying 39 inconsistencies (i.e.
+at least two arguments that mutually attack) and 29 insufficiencies (i.e., at least one argument attacking another).
+
+It is clear from this literature that arguments play a core role in RE.
+Murukannaiah et al.~\cite{murukannaiah2015} show that critical reflection using argument schemes and critical questions -- in the same way that RationalGRL proposes -- improves the quality of the reasoning in the RE process.
+Elrakaiby et al.~\cite{ElrakaibyFSGN17} provide a case study similar to the current one, identifying, as we did, many counterarguments specifically with respect to realisability, relevance and clarity (cf.
+CQ2a, CQ3, CQ11 and CQ12 in Table~\ref{table:argument-schemes}).
+Like in RationalGRL, the use of Dung-style argumentation semantics to compute the acceptable claims in RE is further also advocated by the literature \cite{yu2015automated,ionita2014argumentation,ElrakaibyFSGN17}.
+
+The current work on RationalGRL puts the insights from the above-mentioned literature in a broader framework.
+For example, there is a specific focus on security requirements \cite{haley2008security,yu2015automated,ionita2014argumentation} or the reasoning is about single goals or tasks instead of about the wider context as represented in a goal model \cite{haley2008security,yu2015automated,murukannaiah2015,ElrakaibyFSGN17}.
+RationalGRL provides a generic and extensible framework for arguing about goals and tasks in RE.
+At the moment, there is only a ``generic argument'' in addition to arguments about goals and tasks.
+However, new argument schemes and critical questions about for example, security risks or expert opinions, can be easily added: the metamodel (Figure~\ref{fig:metamodel}) accommodates this and the formal specification in Section~\ref{sect:formalframework} is set up in such a way that  extending the definition of argument and adding new algorithms for specific critical questions is easy.
+
+\paragraph{Goal Modeling} Argumentation has been included -- both explicitly and implicitly -- in existing goal modeling languages.
+For example, the belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} is meant to capture the rationales (reasons, arguments) behind the inclusion of goals and tasks in the model.
 Furthermore, relations between elements in a goal model also provide justifications: high-level goals are reasons for lower-level goals, tasks and resources.
@@
 Providing an alternative sub-goal or -task in a goal model then an explicit argumentative move in the discussion.
-
-So in a sense, a goal model provides a justification for itself, particularly if we include belief elements for extra design rationalization.
-This idea is also prevalent in our RationalGRL framework: many arguments are in fact GRL elements, and many critical questions can be answered by introducing (\textsf{INTRO}) new GRL elements.
+So in this sense, a goal model provides a justification for itself, particularly if we include belief elements for extra design rationalization.
+This idea is also prevalent in our RationalGRL framework: many arguments are in fact GRL elements, and many critical questions can be answered by introducing new GRL elements.
 However, as was already argued in Section~\ref{sect:introduction} (and also by \cite{Jureta:RE2008}), argumentation produces different, richer and complementary information to just the goal model.
-The goal model is the product of a process of argumentation, and does not include, for example, goals and tasks that were at some point considered discarded.
+The goal model is the product of a process of argumentation, and does not include, for example, goals and tasks that were at some point considered but discarded.
 Furthermore, for goal models it is only possible to determine the satisfiability of goals given the possible tasks and resources; what cannot be determined is the acceptability of goals, that is, whether they are acceptable given potentially contradictory opinions of stakeholders.
-
-Closely related to goal modeling and design justification are frameworks that aim to provide a design rationale (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or artefact.
-DR looks at issues, options and arguments for and against these options in the design of, for example, a software system.
-The similarity with goal modeling is that much of the traditional DR literature provides modeling languages and tool support for building DR diagrams.
-The differences lie in the concepts used: in DR, the focus is more on the rationales, why certain design options were chosen, than on the functional and non-functional requirements of the system, which are expressed in goal models.
-The RationalGRL framework in a sense incorporates the ideas from DR into goal models by including arguments and issues (critical questions) into the RatinalGRL model.
-
-Arguments and requirements engineering approaches have been combined by, among others, Haley \emph{et al.}~\cite{haley2005arguing}, who use structured arguments to capture and validate the rationales for security requirements.
-However, they do not use goal models, and thus, there is no explicit trace from arguments to goals and tasks.
-Furthermore, like~\cite{Jureta:RE2008}, the argumentative part of their work does not include formal semantics for determining the acceptability of arguments, and the proposed frameworks are not actually implemented.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015resolving} propose Arg-ACH, an approach to capture inconsistencies between stakeholders' beliefs and goals, and resolve goal conflicts using argumentation techniques.
 
@@
 
-\todo{Marc}{all}{Review the following notes:
+One idea is to capture requirements engineering and software design processes as explicit dialogues between parties~\cite{finkelstein1989multiparty}.
+
+\todo{F}{F}{Review the following:
 
 \begin{itemize}
-\item RE17 paper: merging argumentation theory with law
-  "Using Argumentation to Explain Ambiguity in Requirements Elicitation Interviews" Yehia Elrakaiby, Alessio Ferrari, Paola Spoletini, Stefania Gnesi and Bashar Nuseibeh
-\item more details on Murukannaiah [RE15]
-\item related work on argumentation theory - check Serena Villata on goal models
-\item Prakken and Wierenga
-\item First paper we wrote they also said something about Bashar Nuseibeh
-\item Also mention argumentation in AI - arg diagramming - and how that relates to our work
-\item Software design discussions Black et al.
-\item Maybe this one:
-  "Complete Traceability for Requirements in Satisfaction Arguments"
-  Anitha Murugesan, Michael Whalen, Elaheh Ghassabani and Mats Heimdahl
+\item Mirbel and Villata on Dung \& goal models
+\item Software design discussions Black et al., Prakken and Wierenga, \cite{finkelstein1989multiparty}
   \end{itemize}

### 2017-10-10 introduction.tex
Related work & references cleanup

@@
 These activities fall under the umbrella of goal modeling.
-There are a large number of established RE methods using goal models in the early stage of requirements analysis (e.g.,~\cite{liu2004designing,donzelli2004goal,dardenne1993goal,chung2012non,castro2002towards}, overviews can be found in~\cite{van2001goal,kavakliL05}).
+There are a large number of established RE methods using goal models in the early stage of requirements analysis (overviews can be found in~\cite{kavakliL05,van2001goal}).
 Several goal modeling languages have been developed in the last two decades as well.
-The most popular ones include $i*$~\cite{Yu:1997:TMR:827255.827807}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}.
+The most popular ones include $i*$~\cite{yu1997towards}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}.
 

### 2017-10-10 main.tex
Section 1-4
Small changes 1, 2, some work on case study section (3); section 4 is
not finished yet!
Finished section 4 (for now)
changes up to formal framework, publisher files
notes formal framework and small changes
Related work & references cleanup

--- 2017-03-11 main.tex
@@
  \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2},
+        Floris Bex\Mark{2}
         and Sepideh Ghanavati\Mark{3}\\[1em]
@@
   \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science  \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University\tabularnewline
+  University of Luxembourg & Utrecht University & Texas Tech University \tabularnewline
   \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
@@
 \input{background}
+\input{gmas}
 \input{overview}
-\input{gmas}
+%\input{examples}
+\input{tool}
 \input{formalframework}
-\input{tool}
 \input{discussion}

### 2017-10-16 formalframework.tex
Updated formal framework

@@
 \subsection{Formal Specification of GRL}
-\label{sect:goalmodeling:logicallanguage}
+\label{sect:formalframework:grl}
 
@@
 \label{def:set-definitions}
-Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$, and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
+Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
 
@@
 \item $Type = \{Softgoal, Goal, Task, Resource\}$,
-\item $Names$ is a finite set of string.
+\item $Names$ is a finite set of strings.
 \item $DecompType = \{AND,OR,XOR\}$.
@@
 \item 
-We refer to an intentional element with $ie.id=i$ simply by $ie_i$.\footnote{Note that the reference $ie_i$ is never ambiguous in GRL since identifiers for intentional elements are always unique (see Definition~\ref{def:grlmodel}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
+We refer to the set of intentional elements with $ie.id=i$ by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
 \item
@@
 \begin{itemize}
-\item $id\in\mathbb{N}$ refers to the unique identifier of the actor, 
-\item $name$ refers to a string description of its name.
+\item $id\in\mathbb{N}$ is the identifier of the actor, 
+\item $name\in Names$ is a string description of its name.
 \end{itemize}
@@
 \begin{itemize}
-\item $i$ is the unique identifier of the link,
-\item $j$ is the unique identifier of the IE from which the contribution originates,
-\item $k$ is the unique identifier of the IE to which is contributed.
+\item $i\in \mathbb{N}$ is the unique identifier of the link,
+\item $j\in \mathbb{N}$ is the unique identifier of the IE from which the contribution originates,
+\item $k\in \mathbb{N}$ is the unique identifier of the IE to which is contributed.
 \end{itemize}
@@
 \begin{itemize}
-\item $i$ is the unique identifier of the link,
-\item $j$ is the unique identifier of the decomposing IE,
-\item $k$ is the unique identifier of the IE that is being decomposed.
+\item $i\in \mathbb{N}$ is the unique identifier of the link,
+\item $j\in \mathbb{N}$ is the unique identifier of the decomposing IE,
+\item $k\in \mathbb{N}$ is the unique identifier of the IE that is being decomposed.
 \end{itemize}
@@
 \begin{itemize}
-\item $i$ is the unique identifier of the link,
-\item $j$ is the unique identifier of the dependee IE,
-\item $k$ is the unique identifier of the dependent IE.
+\item $i\in \mathbb{N}$ is the unique identifier of the link,
+\item $j\in \mathbb{N}$ is the unique identifier of the dependee IE,
+\item $k\in \mathbb{N}$ is the unique identifier of the dependent IE.
 \end{itemize}
@@
 Similar to intentional elements and actors, we may refer to links with an id using the subscript notation, e.g., $contrib_0$ is a contribution link with $contrib.id = 0$.
-
-The definitions above come with various implicit assumptions in order to form a valid GRL model.
-For instance, if an Actor-IE relation $(i,j)$ occurs in a GRL model, then there should exist an actor with id $i$, and there also should exist an IE with id $j$.
-We make these assumptions explicit in the following definition.
 
@@
 \end{itemize}
+\end{definition}
 
-Furthermore, the following conditions are satisfied:
+The definitions above come with various implicit assumptions in order to form a valid GRL model.
+For instance, if an Actor-IE relation $(i,j)$ occurs in a GRL model, then there should exist an actor with id $i$, and there also should exist an IE with id $j$.
+We make these assumptions explicit in the following definition.
+
+\begin{definition}[Valid GRL Model]
+\label{def:valid-grl-model}
+A GRL model $GRL=(IE, Act, R_{ActIE}, Contr, Decomp, Dep)$ (Def.~\ref{def:grl-model}) is a \emph{valid GRL model} iff the following conditions are satisfied:
 \begin{enumerate}
@@
 \subsection{Formal specification of RationalGRL}
+\label{sect:formalframework:rationalgrl}
 
@@
 \item A new element called \emph{generic argument},
-\item A new link called \emph{attack link},
-\item An labeling of the elements, assigning to each element (an IE, link, or generic argument) the label IN (accepted) or OUT (reject).
-In the next section we explain how the labeling is computed, but in this section we only provide a formal definition.
+\item A new link called \emph{attack link}
 \end{itemize}
 
-Furthermore, we also change the representation of IE, links, and actors.
-In GRL, these elements have a single identifier.
-However, in RationalGRL, we allow the user to \emph{replace} such elements when answering a critical question with effect REPLACE.
-Therefore, for each GRL element, we maintain a history, which is a list in which the last element of the list corresponds to the element that ends up in the GRL model.
+Following the RationalGRL metamodel (Figure~\ref{fig:metamodel}), we define an \emph{Argument} as either a generic argument or any of the GRL elements, and we define attack relations between arguments.
+Thus, a RationalGRL model is simply a set of arguments and attack relations.
+We use \emph{argumentation semantics} to compute extensions, which are sets of accepted arguments.
 
@@
 \label{def:generic-argument}
-A generic argument $arg \in \mathbb{N}\times Names$ is a relation such that $arg=(id, name)$ is an argument with unique identifier $id$ and name $name$.
-We may refer to the argument with id $i$ simply with $arg_i$.
-A set of arguments is denoted by $Arg$.
+A generic argument $ga \in \mathbb{N}\times Names$ is a relation such that $ga=(id, name)$ means
+\begin{itemize}
+\item $id\in \mathbb{N}$ is the identifier of the generic argument,
+\item $name\in Names$ is a string description of its name.
+\end{itemize}
+We may refer to the argument with id $i$ simply with $arg_i$. A set of arguments is denoted by $GA$.
 \end{definition}
@@
 \begin{itemize}
-\item $i$ is the unique identifier of the element that is performing that attack,
-\item $j$ is the unique identifier of the element that is being attacked.
+\item $i\in \mathbb{N}$ is the unique identifier of the element that is performing that attack,
+\item $j\in \mathbb{N}$ is the unique identifier of the element that is being attacked.
 \end{itemize}
-Intuitively, $att=(i,j)$ means that element $i$ is attacking element $j$.
+Intuitively, $att=(i,j)$ means that element $i$ is attacking element $j$. A set of attack links is denoted by $Att$.
 \end{definition}
 
-Note that the previous definition did not put any constraint on which elements can be connected with attack links. This is what we make explicit in the definition of a RationalGRL model later.
+Note that the previous definition did not put any constraint on which elements can be connected with attack links. This is what we make explicit in the definition of a valid RationalGRL model later.
 
-We next add the the labeling of elements.
-This labeling determines whether an element is accepted (which means it shows up in the GRL model if it is an IE or a link), or rejected (which means it won't be shown).
+Next we define an argument as either a generic argument or any of the GRL elements or links.
+This captures the specification in the RationalGRL metamodel (Fig.~\ref{fig:metamodel}) in which the class \textsf{Argument} is a superclass of \textsf{GenericArgument} and \textsf{GRLModelElement}, which again is a superclass of both \textsf{ElementLink} and \textsf{GRLLinkableElement}.
+A \textsf{GRLLinkableElement} is in turn a superclass of \textsf{Actor} and \textsf{IntentionalElement}.
+In sum, we define an argument simply as any one of the GRL elements or links, or a generic argument.
 
-\begin{definition}[RationalGRL labeling]
-\label{def:labeling}
-A RationalGRL labeling function $Lab:\mathcal{N}\rightarrow {in,out}$ that is total on the ids of the RationalGRL elements.
-This means that $Lab(id)$ is defined for each $ie.id, contr.id, decomp.id, dep.id, arg.id$.
+\begin{definition}[Argument]
+\label{def:argument}
+A set of arguments $Arg = IE\cup Act\cup R_{ActIE}\cup Contr\cup Decomp \cup Dep\cup GA$ consists of:
+\begin{itemize}
+\item a set of intentional elements $IE$ (Def.~\ref{def:ie}), 
+\item a set of actors $Act$ (Def.~\ref{def:actor}),
+\item a set of Actor-IE relations $R_{ActIE}$ (Def.~\ref{def:act-ie-relation}), 
+\item a set of contribution links $Contr$ (Def.~\ref{def:link:contrib}), 
+\item a set of decomposition links $Decomp$ (Def.~\ref{def:link:decomp}), 
+\item a set of dependency links $Dep$ (Def.~\ref{def:link:dependency}), 
+\item a set of a generic arguments $GA$ (Def.~\ref{def:generic-argument}).
+\end{itemize}
 \end{definition}
-
-Finally we add the last part of the RationalGRL extension. We define a list of intentional elements, contribution links, and actors.
-
-\begin{definition}[IE History, Contribution History, Actor History]
-\label{def:element-list}
-\begin{itemize}
-\item An \emph{IE History for element with id $i$}, denoted $\overline{ie}_i=(ie_0,\ldots,ie_n)$, is a list of IEs $ie_0,\ldots,ie_n$ such that $ie_1.id=\ldots=ie_n.id=i$. 
-\item An \emph{Actor history for actor with id $j$}, denoted $\overline{act}_j=(actor_0,\ldots,actor_n)$, is a list of actors $actor_0,\ldots,actor_n$ such that $actor_0.id=\ldots=actor_n.id=j$. 
-\item A \emph{Contribution history for link with id $k$}, denoted $\overline{contrib}_j=(contrib_0,\ldots,contrib_n)$, is a list of contribution links $contrib_0,\ldots,contrib_n$ such that $contrib_0.id=\ldots=contrib_n.id=k$.
-\end{itemize}
-If the id in the list is not relevant we may omit the subscript notation, e.g., we may write $\overline{ie}$ instead of $\overline{ie}_i$.
-
-Sets of histories for IEs, actors, and contribution links are denoted with respectively $\overline{IE}$, $\overline{Act}$, and $\overline{Contrib}$.
-
-The last element of a list is called the \emph{head element}.
-\end{definition}
-
-Note that we do not require a history for decomposition links or dependency links. This is because these links do not have associated REPLACE operations.
 
@@
 \label{def:rationalgrl-model}
-A \emph{RationalGRL model} $RationalGRL=(\overline{IE}, \overline{Act}, R_{ActIE}, \overline{Contrib}, Decomp, Dep, Arg, Att, Label)$ consists of:
-\begin{itemize}
-\item A set $\overline{IE}$ of lists of intentional elements (Def.~\ref{def:element-list}),
-\item A set $\overline{Act}$ of lists of actors (Def.~\ref{def:element-list}),
-\item A set $R_{ActIE}$ of Actor-IE relations (Def.~\ref{def:act-ie-relation}),
-\item A set $\overline{Contr}$ of lists of contribution links (Def.~\ref{def:element-list}),
-\item A set $Decomp$ of decomposition links (Def.~\ref{def:link:decomp}),
-\item A set $Dep$ of dependency links (Def.~\ref{def:link:dependency}).
-\item A set of generic arguments $Arg$ (Def.~\ref{def:generic-argument}),
-\item A set of attack links (Def.~\ref{def:link:attack}).
-\item A labeling function $Lab$ (Def.~\ref{def:labeling}).
-\end{itemize}
-
-Furthermore, the following conditions are satisfied. In the following, let $Elems = \{IE, Act, Contr, Decomp, Dep, Arg\}$:
-\begin{enumerate}
-\item ids are globally unique across IEs, Links, and Actors, i.e., let $X,Y\in \{\overline{IE},\overline{Act}, \overline{Contr}, Decomp, Dep\}$.
-For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
-\item All intentional elements of actors exist: $\forall (i,j)\in R_{ActIE}: \overline{act}_i \in \overline{Act} \wedge ie_j \in IE$.
-\item An intentional element belongs at most to one actor: $\forall ie_i\in IE: |\{(k,i)\in R_{ActIE}\}| \le 1$.
-\item Contribution links connect intentional elements: $\forall (i,j,k,value)\in \overline{Contrib}: \{ie_j,ie_k\}\subseteq IE$.
-\item Decomposition links connect intentional elements: $\forall (i,j,k)\in Decomp: \{ie_j,ie_k\}\subseteq IE$.
-\item Dependency links connect intentional elements: $\forall (i,j,)\in Dep: \{ie_j,ie_k\}\subseteq IE$.
-\item Attack links connect generic arguments, IEs, actors, and links with each other, i.e., let   $\forall (i,j)\in Att: \exists X\in Elems$ s.t. $X_i\in X$ and $\exists Y\in Elems$ s.t. $Y_i\in Y$.
-\end{enumerate}
+A \emph{RationalGRL model} $RationalGRL=(Arg, Att)$ consists of a set of arguments $Args$ (Def.~\ref{def:argument}) and a set of attack links $Att$ (Def.~\ref{def:link:attack}).
 \end{definition}
 
-\todo{Marc}{Marc}{Explain conditions here and give some examples}
-
-\subsubsection{Argumentation semantics}
-
-In order to determine which arguments are accepted or not, we define \emph{argumentation semantics}.
+The definition of a RationalGRL model of above is very general, and does not put any constraints on the model.
+In order to define the notion of a \emph{valid} RationalGRL model (similar to that of a valid GRL model, Def.~\ref{def:valid-grl-model}), we should be able to determine precisely when an argument is accepted and when not.
+This is what we do using argumentation semantics.
 We used the standard approach here, which is known as \emph{Dung's semantics}.
@@
 
-\begin{definition}[Attack, conflict-freeness, defense, and admissibility~\cite{Dung1995}] Suppose an argumentation framework $AF=(Arg,Att)$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
+\begin{definition}[Attack, conflict-freeness, defense, and admissibility~\cite{Dung1995}] Suppose a RationalGRL model $\emph{RationalGRL=(Arg,Att)}$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
 We say that
@@
 \begin{definition}[Preferred semantics~\cite{Dung1995}] \label{def:semantics}
-A preferred extension of an argumentation framework $AF$ is a maximal (w.r.t. set inclusion) admissible set of $AF$.
+A preferred extension of a RationalGRL model $R$ is a maximal (w.r.t. set inclusion) admissible set of $R$.
 \end{definition}
@@
 
-\begin{algorithm}[h]
-  \caption{Translation Algorithm}\label{alg:translation}
-  \begin{algorithmic}[1]
-    \State TODO
-  \end{algorithmic}
-\end{algorithm}
+The previous definition of a preferred extension allows us to compute accepted arguments in a RationalGRL model.
+However, it does not yet put constraints on the resulting accepted arguments, and thus does not allow us to transform it into a valid GRL model.
+This is what we do in the following definition.
 
-\begin{theorem}
-If a RationalGRL model $m$ satisfies the conditions of a RationalGRL model (Def.~\ref{def:rationalgrl-model}), then $Trans(m)$ satisfies the conditions of a GRL model (Def.~\ref{def:grl-model}).
-\end{theorem}
+\begin{definition}[Valid RationalGRL Model]
+A RationalGRL model \emph{R = $(Arg, Att)$} is a \emph {valid RationalGRL model} iff the preferred extension $E = IE\cup Act\cup R_{ActIE}\cup Contr\cup Decomp \cup Dep\cup GA$ forms a valid GRL model $(IE,  Act, R_{ActIE}, Contr, Decomp, Dep)$
+\end{definition}
 
-\begin{proof}
-TODO
-\end{proof}
+\todo{Marc}{Marc}{Explain conditions here and give some examples}
 

### 2017-10-18 formalframework.tex
Finished formalization of GRL

@@
 In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:goalmodeling:logicallanguage}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
-We then provide a formal translation procedure from a RationalGRL model to a GRL model.
 
-In the third subsection we turn to the dynamics of our framework. We develop algorithms for instantiating argument schemes and for answering critical questions.
+In the third subsection we develop algorithms in order to translate a GRL model into a RationalGRL model, and visa versa.
 
-The two main results of this section are that (i) the translation procedure from a RationalGRL model to a GRL model is correct, meaning that if the RationalGRL model is valid, the resulting GRL model is also valid, (ii) the algorithms for applying argument schemes and critical questions are correct as well, meaning that if the Rational model is valid before, it is also valid afterwards.
-These two results together entail that our framework and the corresponding operations on it are consistent.
+In the fourth and final subsection we turn to the dynamics of our framework. We develop algorithms for instantiating argument schemes and for answering critical questions.
 
@@
 
-\subsubsection{Intentional Element and Actors}
+\subsubsection{Intentional Element, beliefs and actors}
 
@@
 \begin{itemize}
-\item $Type = \{Softgoal, Goal, Task, Resource\}$,
+\item $Types = \{Softgoal, Goal, Task, Resource\}$,
+\item $Elements = Type \cup \{Actor, Contr, Dependency, Decomp, Belief\}$
 \item $Names$ is a finite set of strings.
-\item $DecompType = \{AND,OR,XOR\}$.
-\item $ContribValue = \{Break, Hurt, Some\ nega-$ $tive, Unknown, Some\  positive, Help, Make\},$
+\item $DecompTypes = \{AND,OR,XOR\}$.
+\item $ContribValues = \{Break, Hurt, Some\ nega-$ $tive, Unknown, Some\  positive, Help, Make\},$
 \end{itemize}
@@
 \label{def:ie}
-An intentional element $ie\in \mathbb{N}\times Type\times Names\times DecompType$ is a relation, where $ie = (id, type, name, decomptype)$ means:
+An intentional element $ie\in \mathbb{N}\times Types\times Names\times DecompType$ is a relation, where $ie = (id, type, name, decomptype)$ means:
 \begin{itemize}
 \item $id\in \mathbb{N}$ is a unique identifier for the element,
-\item $type\in Type$ specifies the type of the element,
+\item $type\in Types$ specifies the type of the element,
 \item $name \in Names$ is a string description of the element,
@@
 \end{definition}
-
-\paragraph{Example} \todo{Marc}{Marc}{Give example here}
 
@@
 \item
-We refer to the element of an intention element $ie$ using the dot (".") notation.
-That is, we may refer to the id, type, name and decomposition type with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
+We refer to the element of a tuple using the dot (".") notation.
+That is, we may for instance refer to the id, type, name and decomposition type of an $IE$ with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
 \item 
-We refer to the set of intentional elements with $ie.id=i$ by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
+We refer to a set of elements with the same id $i$  using the $i$ subscript on the set.
+For instance a set of IEs with id $i$ is denoted by $E_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
 \item
@@
 \end{definition}
+
+\paragraph{Rationale and example} Throughout this section we use the example in Figure~\ref{fig:example-small}, which we reproduce in Figure~\ref{fig:example-small2}.
+Our formalization of GRL is very much in line with the way in which GRL models are represented in the open-source Eclipse-based tool jUCMNav.\footnote{See \url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}} This tool is actively developed and has a rich number of features for the analysis of both GRL and URN models.
+By keeping our formalization in line with this tool, we simplify the translation step from RationalGRL to GRL.
+
+Some of the IEs in Figure~\ref{fig:example-small2} can be formalized using Def.~\ref{def:ie} as follows:
+\begin{itemize}
+\item $(2, goal, \text{Generate traffic}, XOR)$: This represents the goal ``Generate traffic'', which is XOR-decomposed. Note the identifier is not shown in the GRL model, but it is stored internally.
+\item $(5, resource, \text{Car objects}, AND)$: Although this element does not decompose into any other elements, it still contains an AND-decomposition.
+Our motivation for this choice is that it simplifies the formalization of decomposition links.
+It is in line with jUCMNav implementation, the open source implementation of GRL.
+\end{itemize}
+
+Using the short-hand notation defined in Def.~\ref{def:notation}, we can make the following statements:
+\begin{itemize}
+\item $ie_2.name = \text{Generate traffic}$
+\item $goal_2.decomptype = XOR$
+\item $resource_5.name = \text{Car objects}$
+\end{itemize}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{img/Example1.pdf}
+\caption{Partial GRL Model of the traffic simulator example}
+\label{fig:example-small2}
+\end{figure}
+
+\begin{definition}[Belief]
+\label{def:belief}
+A belief $bel\in \mathbb{N}\times Names$ is a relation where $bel=(id,name)$ means:
+\begin{itemize}
+\item $id\in\mathbb{N}$ is the identifier of the belief,
+\item $name\in Names$ is a string description of the belief.
+\end{itemize}
+\end{definition}
+
+\paragraph{Rational and example} Beliefs are different from IEs in the sense that they are not being used in decompositions, dependencies, or contributions.
+They merely provide additional explanation for an element.
+The only belief in Figure~\ref{fig:example-small2} is formalized as $(1, \text{Not scientifically correct})$.
 
@@
 
-In the next definition we specify relations between actors and intentional elements.
-Each actor can own one or multiple intentional elements.
-We chose to define this simply as a actor-element relation, which is in line with the jUCMNav implementation.
+\paragraph{Rationale and example} An actor is simply a pair consisting of its identifier and its name. The relation between actors and their intentional element is formalized in the next definition.
+
+We can formalize the actor of Figure~\ref{fig:example-small2} as $act=(0,\text{Traffic Simulator})$, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
 
@@
 
+\paragraph{Rationale and example} The previous definition specifies relations between actors and intentional elements.
+Each actor can own one or multiple intentional elements, but each intentional element can only be owned by at most one actor (we will make these assumptions explicit in Def.~\ref{def:valid-grl-model}).
+
+In Figure~\ref{fig:example-small2}, recall we formalized two IEs and the actor as follows:
+\begin{itemize}
+\item $(2, goal, \text{Generate traffic}, XOR)$
+\item $(5, resource, \text{Car objects}, AND)$
+\item $(0, \text{Traffic Simulator})$
+\end{itemize}
+
+We can then formalize the Actor-IE relationships for these elements as: $(0,2)$ and $(0,5)$.
+
 \subsubsection{Links}
@@
 
+\begin{definition}[Belief link]
+\label{def:link:belief}
+A \emph{belief} link: $bellink\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $dep = (i,j,k)$ means
+\begin{itemize}
+\item $i\in \mathbb{N}$ is the unique identifier of the link,
+\item $j\in \mathbb{N}$ is the unique identifier of the belief element,
+\item $k\in \mathbb{N}$ is the unique identifier of the IE.
+\end{itemize}
+Intuitively, $bellink = (i,j,k)$ means that $bel_i$ is a belief for $ie_j$.
+
+A set of belief links is denoted by $Bellink$.
+\end{definition}
+
+\paragraph{Rationale and example} The belief links can only be used to connect a belief with an intentional element.
+However, this is not made explicit in the definition above.
+We make this assumption, together with various assumption, explicit in the definition of a \emph{valid GRL model} (Def.~\ref{def:valid-grl-model}).
+Figure~\ref{fig:example-small2} does not contain any belief links.
+
 \begin{definition}[Contribution Link]
 \label{def:link:contrib}
-A \emph{contribution} link: $contrib\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}\times Contribvalue$ is a relation such that $contrib = (i,j,k,value)$ means
+A \emph{contribution} link: $contrib\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}\times Contribvalues$ is a relation such that $contrib = (i,j,k,value)$ means
 \begin{itemize}
@@
 \item $k\in \mathbb{N}$ is the unique identifier of the IE to which is contributed.
+\item $value\in ContribValues$ is the strength of the contribution.
 \end{itemize}
@@
 \end{definition}
+
+\paragraph{Rationale and example} Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small2}, there are two contribution links. First, let us formalize the related IEs:
+\begin{itemize}
+\item $(3, softgoal, \text{Realisitc simulation}, AND)$
+\item $(4, softgoal, \text{Simple design}, AND)$
+\item $(6, task, \text{Create new cars}, AND)$
+\end{itemize}
+
+We can formalize the contribution links as $(9, 6, 3, \text{Help})$ and $(10, 6, 4, \text{Hurt})$.
 
@@
 
+\paragraph{Rationale and example} Decomposition links are simpler than contribution links, because they only contain information about which elements are being connected.
+Recall that the decomposition type is stored in the element that is being decomposed.
+In order to formalize the decomposition links in Figure~\ref{fig:example-small2}, let us first formalize the related IEs:
+\begin{itemize}
+\item $(5, goal, \text{Generate traffic}, XOR)$
+\item $(6, task, \text{Create new cars}, AND)$
+\item $(7, task, \text{Keep same cars}, AND)$
+\end{itemize}
+The decomposition links can be formalized as $(11, 6, 5)$ and $(12, 7, 5)$.
+
 \begin{definition}[Dependency Link]
@@
 \end{itemize}
-Intuitively, $dep = (i,j,k)$ means that $ie_j$ depends on $ie_j$. \todo{Marc}{Sepideh}{Is this correct? For a dependency relation $A\rightarrow B$, who is dependent on who?}
+Intuitively, $dep = (i,j,k)$ means that $ie_j$ depends on $ie_j$.
 
@@
 
-Similar to intentional elements and actors, we may refer to links with an id using the subscript notation, e.g., $contrib_0$ is a contribution link with $contrib.id = 0$.
+\paragraph{Rationale and example} The representation of dependency links is exactly the same as decomposition links.
+In Figure~\ref{fig:example-small2}, let us formalize the IEs related to the only dependency link:
+\begin{itemize}
+\item $(2, resource, \text{Car objects}, AND)$
+\item $(5, goal, \text{Generate traffic}, XOR)$
+\end{itemize}
+
+The dependency link between these two elements is formalized as $(8, 5, 2)$.
 
@@
 \label{def:grl-model}
-A \emph{GRL model} $GRL=(IE, Act, R_{ActIE}, Contr, Decomp, Dep)$ consists of:
+A \emph{GRL model} $GRL=(IE, Bel, Act, R_{ActIE}, Bellink, Contr, Decomp, Dep)$ consists of:
 \begin{itemize}
 \item A set $IE$ of intentional elements (Def.~\ref{def:ie}),
+\item A set $Bel$ of beliefs (Def.~\ref{def:belief}),
 \item A set $Act$ of actors (Def.~\ref{def:actor}),
 \item A set $R_{ActIE}$ of Actor-IE relations (Def.~\ref{def:act-ie-relation}),
+\item A set $Bellink$ of belief links (Def.~\ref{def:link:belief}),
 \item A set $Contr$ of contribution links (Def.~\ref{def:link:contrib}),
@@
 
-The definitions above come with various implicit assumptions in order to form a valid GRL model.
-For instance, if an Actor-IE relation $(i,j)$ occurs in a GRL model, then there should exist an actor with id $i$, and there also should exist an IE with id $j$.
-We make these assumptions explicit in the following definition.
+\paragraph{Rationale and example} The definition of a GRL model collects all the previously defined tuples into a single definition.
+For completeness, we now provide the full specification of Figure~\ref{fig:example-small2}.
+This model is formalized as
+ $GRL=(IE, Bel, Act, R_{ActIE}, Bellink, Contr, Decomp, Dep)$ where:
+\begin{flalign*}
+&IE=&\{&(2, task, \text{Car objects}, AND),&\\
+&   &  &(3, softgoal, \text{Realistic simulation}, AND),&\\
+&   &  &(4, softgoal, \text{Simple design}, AND),&\\
+&   &  &(5, goal, \text{Generate traffic}, XOR),&\\
+&   &  &(6, task, \text{Create new cars}, AND),&\\
+&   &  &(7, task, \text{Keep same cars}, AND)\}&\\
+&Bel=& &\{(1, \text{Not scientifically correct})\}&\\
+&Act=& &\{(0, \text{Traffic Simulator})\}&\\
+&R_{ActIE}=& &\{(0,i)\mid 2\le i\le 7\}&\\
+&Bellink = & &\emptyset&\\
+&Contr=& &\{(9, 6, 3, \text{Help}), (10, 6, 4, \text{Hurt})\}&\\
+&Decomp=& &\{(11, 6, 5), (12, 7, 5)\}&\\
+&Dep=& &\{(8, 5, 2)\}&
+\end{flalign*}
 
@@
 
-\todo{Marc}{Marc}{Provide explanation and example here.}
+\paragraph{Rationale and example} The definition of a GRL model (Def.~\ref{def:grl-model}) comes with various implicit assumptions in order to form a valid GRL model.
+For instance, if an Actor-IE relation $(i,j)$ occurs in a GRL model, then there should exist an actor with id $i$, and there also should exist an IE with id $j$.
+We make these assumptions explicit in this definition.
+
+Let us briefly verify that our previous formalization of Figure~\ref{fig:example-small2} satisfies all the constraints of Def.~\ref{def:valid-grl-model}:
+\begin{enumerate}
+\item All elements in the formalization have different ids, so this constraint is satisfied.
+\item $R_{ActIE}$ contains one element $(0,i)$ for each IE with id $i$, so this constraint is satisfied as well.
+Note that other elements (beliefs and links) are not related to actors.
+This is in line with the jUCMNav implementation.
+\item Since we have only one actor with id 0, and this is the only actor that appears in $R_{ActIE}$, this constraint is satisfied.
+\item The contribution links connect elements with ids 3, 4, and 6, which are all IEs.
+\item The decomposition links connect elements with ids 5, 6, and 7, which are all IEs.
+\item The dependency link connects id 2 with 5, which are both IEs.
+\end{enumerate}
 
@@
 \end{itemize}
-We may refer to the argument with id $i$ simply with $arg_i$. A set of arguments is denoted by $GA$.
+We may refer to the argument with id $i$ simply with $ga_i$. A set of arguments is denoted by $GA$.
 \end{definition}
@@
 \end{definition}
-
-Note that the previous definition did not put any constraint on which elements can be connected with attack links. This is what we make explicit in the definition of a valid RationalGRL model later.
 
@@
 
+In order to identify arguments with the corresponding GRL element, we also add a relation between the element ids and the corresponding element.
+Recall the definition of $Element$ from Def.~\ref{def:set-definitions}.
+
+\begin{definition}[Argument-Element mapping]
+\label{def:arg-elem-mapping}
+An \emph{argument-element mapping} $argElem\in \mathbb{N}\times Element$ is a mapping from an element identifier to its corresponding GRL element. So if $argElem=(i,elem)$, then
+\begin{itemize}
+\item $i\in\mathbb{N}$ is the identifier of the GRL element,
+\item $elem\in Element$ (Def.~\ref{def:set-definitions}) is the GRL element type identified by $i$.
+\end{itemize}
+We denote a set of argument-element mappings with $ArgElem$.
+\end{definition}
+
 \begin{definition}[RationalGRL Model]
 \label{def:rationalgrl-model}
-A \emph{RationalGRL model} $RationalGRL=(Arg, Att)$ consists of a set of arguments $Args$ (Def.~\ref{def:argument}) and a set of attack links $Att$ (Def.~\ref{def:link:attack}).
+A \emph{RationalGRL model} $RatGRL=(Arg, Att, ArgElem)$ consists of a set of arguments $Args$ (Def.~\ref{def:argument}), a set of attack links $Att$ (Def.~\ref{def:link:attack}), and an of argument-element mappings (Def.~\ref{def:arg-elem-mapping}).
 \end{definition}
 
-The definition of a RationalGRL model of above is very general, and does not put any constraints on the model.
-In order to define the notion of a \emph{valid} RationalGRL model (similar to that of a valid GRL model, Def.~\ref{def:valid-grl-model}), we should be able to determine precisely when an argument is accepted and when not.
-This is what we do using argumentation semantics.
-We used the standard approach here, which is known as \emph{Dung's semantics}.
+In order to compute when an argument is accepted and when not we use argumentation semantics.
+We use the standard approach here, which is known as \emph{Dung's semantics}.
 The following notions are preliminary.
 
-\begin{definition}[Attack, conflict-freeness, defense, and admissibility~\cite{Dung1995}] Suppose a RationalGRL model $\emph{RationalGRL=(Arg,Att)}$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
+\begin{definition}[Argumentation Framework\~cite{}]
+\label{def:argumentation-framework}
+An \emph{argumentation framework} $AF=(Arg,Att)$ consists of a set of arguments $Arg$ and a set of attack relations $Att\subseteq Arg\times Arg.$
+\end{definition}
+
+We see that the definition of an argumentation framework is very close to that of a RationalGRL model. This allows us to use the following results directly.
+
+\begin{definition}[Attack, conflict-freeness, defense, and admissibility~\cite{Dung1995}] Suppose an argumentation framework $AF=(Arg,Att)$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
 We say that
@@
 
+\todo{Marc}{Marc}{Say something about what we do with multiple extensions.}
+
 \begin{definition}[Preferred semantics~\cite{Dung1995}] \label{def:semantics}
@@
 
-The previous definition of a preferred extension allows us to compute accepted arguments in a RationalGRL model.
-However, it does not yet put constraints on the resulting accepted arguments, and thus does not allow us to transform it into a valid GRL model.
-This is what we do in the following definition.
+\begin{algorithm}[h]
+  \caption{RationalGRL to GRL Translation}\label{alg:translation}
+  \begin{algorithmic}[1]
+    \Procedure{$Translate$}{$RatGRL$}
+    \State $E \leftarrow ComputeExtension(Arg,Att)$
+    \State $IE\leftarrow\emptyset$, $Act\leftarrow\emptyset$, $R_{ActIE}\leftarrow\emptyset$, $Contr\leftarrow\emptyset$
+    \State $Decomp\leftarrow\emptyset$, $Dep\leftarrow\emptyset$
+    \For{$(A.id, T)\in ArgElem \mid A\in E$}
+      \Switch{$T$}
+          \Case{$T\in Type$}
+            \State $IE\leftarrow IE\cup \{A\}$
+          \EndCase
+          \Case{$T = Actor$}
+            \State $Act\leftarrow Act \cup\{A\}$
+          \EndCase
+          \Case{$T=Contr$}
+            \State $Contr\leftarrow Contr\cup \{A\}$
+          \EndCase
+          \Case{$T=Dep$}
+            \State $Dep\leftarrow Dep\cup \{A\}$
+          \EndCase
+          \Case{$T=Decomp$}
+            \State $Decomp\leftarrow Decomp\cup\{A\}$
+          \EndCase
+      \EndSwitch
+    \EndFor
+    \State \Return $(IE,Act,R_{ActIE}, Contr, Decomp, Dep)$
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+The previous definition of a preferred extension allows us to compute accepted arguments in a RationalGRL model. In the following definition we make precise the notion of a valid RationalGRL model.
 
 \begin{definition}[Valid RationalGRL Model]
-A RationalGRL model \emph{R = $(Arg, Att)$} is a \emph {valid RationalGRL model} iff the preferred extension $E = IE\cup Act\cup R_{ActIE}\cup Contr\cup Decomp \cup Dep\cup GA$ forms a valid GRL model $(IE,  Act, R_{ActIE}, Contr, Decomp, Dep)$
+Suppose a RationalGRL model $RatGRL = (Arg, Att, ArgElem)$ with preferred extension $E=(A_1,\ldots,A_n)$.
+$RatGRL$ is a \emph{valid RationalGRL model} iff $Translate(RatGRL)$ (Alg.~\ref{alg:translation}) is a valid GRL model (Def.~\ref{def:valid-grl-model}).
 \end{definition}
@@
 
-\subsection{Algorithms for argument schemes}
+\subsection{Algorithms for argument schemes and critical questions}
 \label{sect:algorithms}
@@
 
-The algorithms for the argument schemes consist of forming a new argument and adding it to the set of arguments. No attack relations are introduced.
+\paragraph{Argument Schemes}
+
+In all of the following algorithms, we assume the following
+\begin{itemize}
+\item There algorithms are being applied to some valid RationalGRL model $RGRL$,
+\item The procedure $mindId()$ generates a new unique id.
+\end{itemize}
+
+The algorithms for applying argument schemes are short, and consist simply of adding an argument for the element that is being added.
 
 \begin{algorithm}[h]
-  \caption{Applying AS0: Actor $a$ is relevant}\label{alg:as0}
+  \caption{AS0: Add a new actor with name $n$}\label{alg:as0}
   \begin{algorithmic}[1]
-    \State $RationalGRL=(\overline{IE}, \overline{Act}, R_{ActIE}, \overline{Contr}, Decomp,$
-    \State \qquad\qquad\qquad\qquad$Dep, Arg, Att, Label)$
-    \Procedure{$AS_0$}{$a$}
-    \State $id\gets mintId()$\label{alg:as0:1}
-    \State $\overline{Act} \gets \overline{Act} \cup ((id, a))$\label{alg:as0:2}
+    \Procedure{$AS_0$}{$Arg, Att, n$}
+    \State $actor \leftarrow (mintId(), n)$ 
+    \State $Arg\leftarrow Arg \cup \{actor\}$\label{alg:as0:2}
     \EndProcedure

### 2017-10-18 formalframework.tex
Added argument schemes algortihms

@@
 \begin{itemize}
-\item $Types = \{Softgoal, Goal, Task, Resource\}$,
-\item $Elements = Type \cup \{Actor, Contr, Dependency, Decomp, Belief\}$
+\item $IETypes = \{Softgoal, Goal, Task, Resource\}$,
+\item $Types = IETypes \cup \{Actor, Contr, Dependency, Decomp, Belief\}$
 \item $Names$ is a finite set of strings.
@@
 \label{def:ie}
-An intentional element $ie\in \mathbb{N}\times Types\times Names\times DecompType$ is a relation, where $ie = (id, type, name, decomptype)$ means:
+An intentional element $ie\in \mathbb{N}\times IETypes\times Names\times DecompType$ is a relation, where $ie = (id, type, name, decomptype)$ means:
 \begin{itemize}
 \item $id\in \mathbb{N}$ is a unique identifier for the element,
-\item $type\in Types$ specifies the type of the element,
+\item $type\in IETypes$ specifies the type of the element,
 \item $name \in Names$ is a string description of the element,
@@
 
-In order to develop a logical framework for RationalGRL, we extend the GRL logical framework of the previous section by adding three elements (see Figure~\ref{fig:rationalgrllegend}):
+In order to develop a logical framework for RationalGRL, we extend the GRL logical framework of the previous section by adding two elements (see Figure~\ref{fig:rationalgrllegend}):
 \begin{itemize}
@@
 \end{itemize}
-
-Following the RationalGRL metamodel (Figure~\ref{fig:metamodel}), we define an \emph{Argument} as either a generic argument or any of the GRL elements, and we define attack relations between arguments.
-Thus, a RationalGRL model is simply a set of arguments and attack relations.
-We use \emph{argumentation semantics} to compute extensions, which are sets of accepted arguments.
 
@@
 
+\paragraph{Rationale and example}
+
+\todo{Marc}{Marc}{Add extended version of Fig.~\ref{fig:example-small2} with some arguments and disabled elements and use that to explain this section}
+
 Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link:contrib},~\ref{def:link:decomp}, and~\ref{def:link:dependency}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links.
@@
 
-We next define the attack link, which is the only link that RationalGRL adds to GRL.
+\begin{definition}[Argument]
+\label{def:argument}
+An argument $A=(elem,type)$ is a pair such that:
+\begin{itemize}
+\item $elem$ is either an intentional element $ie$ (Def.~\ref{def:ie}), an actor $act$ (Def.~\ref{def:actor}), an Actor-IE relation $r_{ActIE}$ (Def.~\ref{def:act-ie-relation}),  a contribution link $contr$, a decomposition link $decomp$ (Def.~\ref{def:link:decomp}), a dependency link $dep$ (Def.~\ref{def:link:dependency}), or a generic argument $ga$ (Def.~\ref{def:generic-argument}).
+\item $type\in Type$ (Def.~\ref{def:set-definitions}) is the type of the argument.
+\end{itemize}
+\end{definition}
+
+\paragraph{Rationale and example} We define an argument as either a generic argument or any of the GRL elements or links.
+This captures the specification in the RationalGRL metamodel (Fig.~\ref{fig:metamodel}) in which the class \textsf{Argument} is a superclass of \textsf{GenericArgument} and \textsf{GRLModelElement}, which again is a superclass of both \textsf{ElementLink} and \textsf{GRLLinkableElement}.
+A \textsf{GRLLinkableElement} is in turn a superclass of \textsf{Actor} and \textsf{IntentionalElement}.
+In sum, we define an argument simply as any one of the GRL elements or links, or a generic argument.
 
@@
 
-Next we define an argument as either a generic argument or any of the GRL elements or links.
-This captures the specification in the RationalGRL metamodel (Fig.~\ref{fig:metamodel}) in which the class \textsf{Argument} is a superclass of \textsf{GenericArgument} and \textsf{GRLModelElement}, which again is a superclass of both \textsf{ElementLink} and \textsf{GRLLinkableElement}.
-A \textsf{GRLLinkableElement} is in turn a superclass of \textsf{Actor} and \textsf{IntentionalElement}.
-In sum, we define an argument simply as any one of the GRL elements or links, or a generic argument.
-
-\begin{definition}[Argument]
-\label{def:argument}
-A set of arguments $Arg = IE\cup Act\cup R_{ActIE}\cup Contr\cup Decomp \cup Dep\cup GA$ consists of:
-\begin{itemize}
-\item a set of intentional elements $IE$ (Def.~\ref{def:ie}), 
-\item a set of actors $Act$ (Def.~\ref{def:actor}),
-\item a set of Actor-IE relations $R_{ActIE}$ (Def.~\ref{def:act-ie-relation}), 
-\item a set of contribution links $Contr$ (Def.~\ref{def:link:contrib}), 
-\item a set of decomposition links $Decomp$ (Def.~\ref{def:link:decomp}), 
-\item a set of dependency links $Dep$ (Def.~\ref{def:link:dependency}), 
-\item a set of a generic arguments $GA$ (Def.~\ref{def:generic-argument}).
-\end{itemize}
-\end{definition}
-
-In order to identify arguments with the corresponding GRL element, we also add a relation between the element ids and the corresponding element.
-Recall the definition of $Element$ from Def.~\ref{def:set-definitions}.
-
-\begin{definition}[Argument-Element mapping]
-\label{def:arg-elem-mapping}
-An \emph{argument-element mapping} $argElem\in \mathbb{N}\times Element$ is a mapping from an element identifier to its corresponding GRL element. So if $argElem=(i,elem)$, then
-\begin{itemize}
-\item $i\in\mathbb{N}$ is the identifier of the GRL element,
-\item $elem\in Element$ (Def.~\ref{def:set-definitions}) is the GRL element type identified by $i$.
-\end{itemize}
-We denote a set of argument-element mappings with $ArgElem$.
-\end{definition}
+\paragraph{Rationale and example} The attack link is the only link that RationalGRL adds to GRL.
 
@@
 
-\begin{definition}[Argumentation Framework\~cite{}]
+\begin{definition}[Argumentation Framework~\cite{}]
 \label{def:argumentation-framework}
@@
 \begin{itemize}
-\item There algorithms are being applied to some valid RationalGRL model $RGRL$,
+\item There algorithms are being applied to some valid RationalGRL model $RatGRL$,
 \item The procedure $mindId()$ generates a new unique id.
@@
   \begin{algorithmic}[1]
-    \Procedure{$AS_0$}{$Arg, Att, n$}
-    \State $actor \leftarrow (mintId(), n)$ 
-    \State $Arg\leftarrow Arg \cup \{actor\}$\label{alg:as0:2}
+    \Procedure{$AS_0$}{$n$}
+    \State $actor \leftarrow (mintId(), n)$ \label{alg:as0:1}
+    \State $Arg\leftarrow Arg \cup \{(actor,Actor)\}$\label{alg:as0:2}
     \EndProcedure
@@
 
-\subsubsection{Algortihm~\ref{alg:as0} for argument scheme AS0}
-
-\paragraph{Explanation} For all of the argument schemes and critical questions algorithms, we assume an underlying RationalGRL model as defined in Def.~\ref{def:rationalgrl-model}.
-This model is shown in the first two lines.
-
-The algorithm takes one argument, namely the name of the actor $a$.
-On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted.
-On line~\ref{alg:as0:2}, then a list containing a single tuple $(id,a)$ is added to the set of lists of actors $\overline{Act}$, meaning that a new actor with id $id$ and name $a$ is added to the RationalGRL model (Def.~\ref{def:element-list}).
-In Figure~\ref{fig:examples:relevant-actor2}, ..
-TODO
+\paragraph{Rationale and example} The algorithm takes one argument, namely the name of the actor $n$.
+On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$.
+On line~\ref{alg:as0:2} the pair $(actor,Actor)$ is added as a new argument, indicating that an argument of type $Actor$ is added.
 
 \begin{algorithm}[h]
-  \caption{Applying AS1: Actor $a_{id}$ has resource $r_{id}$}\label{alg:as1}
+  \caption{Applying AS1: Actor with id $i$ has resource $n$}\label{alg:as1}
   \begin{algorithmic}[1]
-     \State $RationalGRL=(\overline{IE}, \overline{Act}, R_{ActIE}, \overline{Contr}, Decomp,$
-    \State \qquad\qquad\qquad\qquad$Dep, Arg, Att, Label)$
-    \Procedure{$AS_1$}{$a_{id}, r_{id}$}
-    \State $R_{ActIE}\gets (a_{id}, r_{id}$
+    \Procedure{$AS_1$}{$i, n$}
+    \State $rId\gets mintId()$
+    \State $r\leftarrow (rId, Resource, n, AND)$
+    \State $Arg\gets Arg\cup \{(r, Resource),(i,rId)\}$
     \EndProcedure
@@
 
-\paragraph{Example} \todo{Marc}{Marc}{Clarify that the actor and resource exist already. If not, we should first execute Algorithm 2, and then Algorithm 3. No need to combine them}
-
-\emph{Algorithm~\ref{alg:as1} for argument scheme AS1:} This argument scheme takes two arguments, the identifier $a_{id}$ of the actor and the resource name $n$.
-The algorithm is similar to the previous one, with the difference that the newly added argument contains the statement $has(a_{id},id)$ as well, meaning that the actor with id $a_{id}$ has element $id$ (which is a resource).
-As an example, let us formalize the argument corresponding to resource \emph{External library} of actor \emph{Traffic tycoon} in Figure~\ref{fig:transcripts:grl}.
-First, we assume some id is associated with the actor: $$\{actor(0),name(0,traffic\_tycoon)\}.$$ Then we can formalize the argument for the resource as follows: $$\{resource(1),name(1,external\_library),has(0,1)\}.$$
-
-\paragraph{Proof of Correctness} In order to show that this algorithm is correct, we should show that the new RationalGRL model still satisfies all the conditions of Def.~\ref{def:grl-model}.
-This is clearly the case, since adding a single element does not invalidate any condition.
-
-Argument scheme AS1 to AS4 are all very similar, in the sense that they all assert that some element belongs to an actor.
-Therefore, we only provide the algorithm for AS1 and we assume the reader can easily construct the remaining algorithms AS2-AS4.
+\paragraph{Rationale and example} Explain here...
 
 \begin{algorithm}[h]
-  \caption{Applying AS5: Goal $g_{id}$ decomposes into tasks $T_1,\ldots,T_n$}\label{alg:as5}
+  \caption{Applying AS5: Goal with id $i$ decomposes into tasks $t$}\label{alg:as5}
   \begin{algorithmic}[1]
-    \Procedure{$AS_5$}{$g_{id}, \{T_1,\ldots,T_n\}, type$}
-    \State $T_{id} = \emptyset$\label{alg:as5:1}
-    \For{$T_i$ in $\{T_1,\ldots,T_n\}$}
-      \If{$\exists_{A\in Args}\label{alg:as5:2} \{task(t_{id}),name(t_{id},T_i)\}\subseteq A$}
-        \State $T_{id} \gets T_{id} \cup \{t_{id}\}$
-      \Else\label{alg:as5:3}
-        \State $id\gets id+1$
-        \State $A \gets \{task(id),name(id,T_i)\}$
-        \State $Args \gets Args\cup A$
-        \State $T_{id} \gets T_{id} \cup \{id\}$
-      \EndIf
-    \EndFor
-    \State $id\gets id+1$\label{alg:as5:4}
-    \State $A\gets \{decomp(id, g_{id}, T_{id}, type)\}$
-    \State $Args \gets Args\cup A$
+    \Procedure{$AS_5$}{$i, t$}
+    \State $tId\gets mintId()$
+    \State $t\leftarrow (tId, Task, t, AND)$
+    \State $Arg\gets Arg\cup \{(t, Task),(i,tId)\}$
     \EndProcedure
@@
 
-\todo{Marc}{Marc}{This algorithm does not make much sense to me. We simply add a decomposition link, no need to make it so complicated. I actually think we can omit this algorithm.}
+\paragraph{Rationale and example} Explain here.
 
-\subsubsection{Algorithm~\ref{alg:as5} for argument scheme AS5}
-
-\paragraph{Explanation} \todo{Marc}{Marc}{I think we should remove all this stuff. In this part we are simply adding GRL elements, which is trivial. The interesting part are the critical questions}
-
-The procedure in Algorithm~\ref{alg:as5} takes three arguments: $g_{id}$ is the identifier of goal $G$, $T=(T_1,\ldots,T_n)$ is a list of decomposing task names, and $type\in\{and,or,xor\}$ is the decomposition type.
-The difficulty of this algorithm is that each of the tasks are stated in natural language, and it is not directly clear whether these tasks are already in the GRL model or not.
-Therefore, we have to check for each tasks whether it already exists, and if not, we have to create a new task.
-On line~\ref{alg:as5:1}, the set $T_{id}$ is initialized, which will contain the ids of the tasks $T_1,\ldots,T_n$ to decompose into.
-In the for-loop, the if-statement on line~\ref{alg:as5:2} checks whether some argument already exists for the current task $T_i$, and if so, it simply adds the identifier of the task ($t_{id}$) to the set of task identifiers $T_{id}$.\footnote{Note that the existing task may be disabled, but this does not matter, since the decomposition relation will be suppressed for disabled elements.} Otherwise (line~\ref{alg:as5:3}), a new task is created and the new identifier $id$ is added to the set of task identifiers.
-After the for loop on line~\ref{alg:as5:4}, an argument for the decomposition link itself is constructed, and it is added to the set of arguments $Args$.
-
-\paragraph{Example}
-Let us explain this algorithm with the XOR-decomposition of goal \texttt{Generate cars} of Figure~\ref{fig:transcripts:grl}. Suppose the following arguments are constructed already:
-\begin{itemize}
-\item $\{goal(0),name(0,generate\_cars)\},$
-\item $\{taks(1),name(1,keep\_same\_cars\}.$
-\end{itemize}
-Suppose furthermore that someone wants to put forward the argument that goal \texttt{Generate cars} XOR-decomposes into tasks \texttt{Keep same cars} and \texttt{Create news cars}.
-Formally: $AS_5(0,\{generate\_cars,keep\_same\_cars\},xor)$.
-The algorithm will first set $T_{id}=\emptyset$, and then iterate over the two task names.
-For the first task $generate\_cars$, there does not exist an argument $\{task(t_{id}),name(t_{id},generate\_cars)\}$ yet, so a  new argument is created.
-Suppose the following argument is created: $\{task(2),name(2,generate\_cars)\}.$ After this, 2 is added to $T_{id}$.
-For the second task an argument exists already, namely $\{task(1),name(1,keep\_same\_cars)\}$, so 1 is simply added to $T_{id}$.
-After the for loop, we have $T_{id}=\{1,2\}$.
-Next, the decomposition argument is created, which is $\{decomp(3,0,\{1,2\},xor)\}$.
-This argument is added to $Args$ and the algorithm terminates.
-
-\begin{algorithm}[h]
-  \caption{Applying AS6: Task $t_{id}$ contributes to softgoal $s$}\label{alg:as6}
-  \begin{algorithmic}[1]
-    \Procedure{$AS_6$}{$t_{id}, s$}
-    \If{$\exists_{A\in Args} \{softgoal(i),name(i,s)\}\subseteq A$} \label{alg:as6:if}
-        \State $s_{id} \gets i$
-    \Else
-      \State $id\gets id+1$
-      \State $A \gets \{softgoal(id),name(id,t)\}$
-      \State $Args \gets Args\cup A$
-      \State $s_{id} \gets id$
-    \EndIf
-    \State $id\gets id+1$
-    \State $A\gets \{contr(id, t_{id}, s_{id}, pos)\}$
-    \State $Args \gets Args\cup A$
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\paragraph{Proof of Correctness}
-
-\subsubsection{Algorithm~\ref{alg:as6} for argument scheme AS6}
-
-\todo{Marc}{Marc}{This one is also not interesting.. just add a contribution link}
-\paragraph{Explanation} The procedure in Algorithm~\ref{alg:as6} takes two arguments: $t_{id}$ is the identifier of task $T$, and $s$ is the softgoal name that is contributed to.
-The idea behind this algorithm is very similar to the previous one, with the difference that in the current algorithm we create a single relation, while we created a set of relations in the previous algorithm.
-First, the if-statement on line~\ref{alg:as6:if} checks whether the softgoal exists already, and if not, an argument is added for it.
-This ensures that all softgoals have corresponding arguments.
-After the if-statement, the argument for the contribution link is created and it is added to the set of arguments $Args$.
-
-\paragraph{Example} Let us again illustrate this with a simple example from Figure~\ref{fig:transcripts:grl}.
-Suppose the following argument exists already: $\{task(0),name(0,keep\_same\_cars\}$, and suppose someone would like to add an argument that the task \texttt{Keep same cars} contributes positively to softgoal \texttt{Dynamic simulation}, i.e.
-$AS_6(0,dynamic\_simulation)$.
-The algorithm first checks whether an argument already exists for the softgoal, and when it finds out it does not exist, creates the argument $\{softgoal(1),name(1,dynamic\_simulation)\}$ and adds it to $Args$.
-Then, the argument for the contribution is added to $Args$ as well, which is $\{contr(2,0,1,pos)\}$.
-
-\emph{Algorithms for argument schemes AS7-AS11:} The algorithms for $AS7$ to $AS11$ all have a very similar structure as Algorithm~\ref{alg:as6} and have therefore been omitted.
-Again, we assume the reader can reconstruct them straightforwardly.
-
-\paragraph{Proof of correctness}
+The remaining argument schemes are all very similar, in the sense that they all simply add a single argument to the set of arguments containing the element to be added.
+We therefore have omitted them here.
 

### 2017-10-21 formalframework.tex
Added examples to RationalGRL formalization

@@
 In this section we present a formalization of RationalGRL based on formal logic.
-This is done for multiple reasons: (i) Most approaches in formal argumentation use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated in to valid GRL models in a very precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support.
+This is done for multiple reasons: (i) Most approaches in formal argumentation use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated in to valid GRL models and visa versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support.
 
-In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:goalmodeling:logicallanguage}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
+In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:metamodel}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
 
@@
 \item $IETypes = \{Softgoal, Goal, Task, Resource\}$,
-\item $Types = IETypes \cup \{Actor, Contr, Dependency, Decomp, Belief\}$
+\item $Types = IETypes \cup \{Actor, Contr, Dep,$\\ \qquad $Decomp, Belief, GenArg\}$
 \item $Names$ is a finite set of strings.
@@
 We refer to a set of elements with the same id $i$  using the $i$ subscript on the set.
-For instance a set of IEs with id $i$ is denoted by $E_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
+For instance a set of IEs with id $i$ is denoted by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
 \item
@@
 
-\paragraph{Rationale and example} Throughout this section we use the example in Figure~\ref{fig:example-small}, which we reproduce in Figure~\ref{fig:example-small2}.
+\paragraph{Rationale and example} Throughout this section we use the example in Figure~\ref{fig:example-small}, which is shown in Figure~\ref{fig:example-small2} as well.
 Our formalization of GRL is very much in line with the way in which GRL models are represented in the open-source Eclipse-based tool jUCMNav.\footnote{See \url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}} This tool is actively developed and has a rich number of features for the analysis of both GRL and URN models.
-By keeping our formalization in line with this tool, we simplify the translation step from RationalGRL to GRL.
+By keeping our formalization in line with this tool, we simplify the translation step from models in the RationalGRL tool to models in the jUCMNav tool.
 
@@
 \begin{itemize}
-\item $(2, goal, \text{Generate traffic}, XOR)$: This represents the goal ``Generate traffic'', which is XOR-decomposed. Note the identifier is not shown in the GRL model, but it is stored internally.
+\item $(2, goal, \text{Generate traffic}, XOR)$: This represents the goal ``Generate traffic'', which is XOR-decomposed.
+Note the identifier ``2'' is not shown in the GRL model, but it is stored internally.
 \item $(5, resource, \text{Car objects}, AND)$: Although this element does not decompose into any other elements, it still contains an AND-decomposition.
 Our motivation for this choice is that it simplifies the formalization of decomposition links.
-It is in line with jUCMNav implementation, the open source implementation of GRL.
+It is in line with with GRL metamodel and the jUCMNav implementation.
 \end{itemize}
@@
 \begin{itemize}
-\item $ie_2.name = \text{Generate traffic}$
+\item $goal_2.name = \text{Generate traffic}$
 \item $goal_2.decomptype = XOR$
@@
 \centering
-\includegraphics[width=\columnwidth]{img/Example1.pdf}
-\caption{Partial GRL Model of the traffic simulator example}
+\includegraphics[width=\columnwidth]{img/Example1-new.pdf}
+\caption{Example GRL model (reprint of Fig.~\ref{fig:example-small})}
 \label{fig:example-small2}
@@
 \label{def:link:belief}
-A \emph{belief} link: $bellink\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $dep = (i,j,k)$ means
+A \emph{belief} link: $bellink\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $bellink = (i,j)$ means
 \begin{itemize}
-\item $i\in \mathbb{N}$ is the unique identifier of the link,
-\item $j\in \mathbb{N}$ is the unique identifier of the belief element,
-\item $k\in \mathbb{N}$ is the unique identifier of the IE.
+\item $i\in \mathbb{N}$ is the unique identifier of the belief element,
+\item $j\in \mathbb{N}$ is the unique identifier of the IE.
 \end{itemize}
-Intuitively, $bellink = (i,j,k)$ means that $bel_i$ is a belief for $ie_j$.
+Intuitively, $bellink = (i,j)$ means that $bel_i$ is a belief for $ie_j$.
 
@@
 
-\paragraph{Rationale and example} The belief links can only be used to connect a belief with an intentional element.
-However, this is not made explicit in the definition above.
+\paragraph{Rationale and example} The belief links can only be used to connect a belief with an intentional element, and they are not used in any of the algorithms.
+Instead, they can be understood as metadata, they provide additional documentation on the rationale for making certain design choices.
+Therefore, belief links do not require unique identifiers.
+
+The fact that belief links only connect beliefs with IEs is not made explicit in the definition above.
 We make this assumption, together with various assumption, explicit in the definition of a \emph{valid GRL model} (Def.~\ref{def:valid-grl-model}).
-Figure~\ref{fig:example-small2} does not contain any belief links.
+Figure~\ref{fig:example-small2} contains one belief links.
+It connects the following two elements:
+\begin{itemize}
+\item $(1, \text{Not scientifically correct})$,
+\item $(4, softgoal, \text{Simple design}, AND)$.
+\end{itemize}
+The belief link is then formalized as $(1, 4)$.
 
@@
 &R_{ActIE}=& &\{(0,i)\mid 2\le i\le 7\}&\\
-&Bellink = & &\emptyset&\\
+&Bellink = & &\{(1, 4)\}&\\
 &Contr=& &\{(9, 6, 3, \text{Help}), (10, 6, 4, \text{Hurt})\}&\\
@@
 \begin{enumerate}
-\item ids are globally unique across IEs, Links, and Actors, i.e., let $X,Y\in \{IE,Act, Contr, Decomp, Dep\}$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
+\item ids are globally unique across IEs, Beliefs, Links, and Actors, i.e., let $X,Y\in \{IE,Act, Belief, Contr, Decomp, Dep\}$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
 \item All intentional elements of actors exist: $\forall (i,j)\in R_{ActIE}: act_i \in Act \wedge ie_j \in IE$.
@@
 \item Decomposition links connect intentional elements: $\forall (i,j,k)\in Decomp: \{ie_j,ie_k\}\subseteq IE$.
-\item Dependency links connect intentional elements: $\forall (i,j,)\in Dep: \{ie_j,ie_k\}\subseteq IE$.
+\item Dependency links connect intentional elements: $\forall (i,j,k)\in Dep: \{ie_j,ie_k\}\subseteq IE$.
+\item Belief links connect beliefs with intentional elements: $\forall (i,j)\in Bellink: bel_i\in Bel\wedge ie_j\in IE$.
 \end{enumerate}
@@
 \paragraph{Rationale and example} The definition of a GRL model (Def.~\ref{def:grl-model}) comes with various implicit assumptions in order to form a valid GRL model.
-For instance, if an Actor-IE relation $(i,j)$ occurs in a GRL model, then there should exist an actor with id $i$, and there also should exist an IE with id $j$.
-We make these assumptions explicit in this definition.
+We make these assumptions explicit in the definition above.
 
@@
 \item The dependency link connects id 2 with 5, which are both IEs.
+\item The belief link connects id 1 with id 4, which are respectively a belief and an IE.
 \end{enumerate}
@@
 \end{itemize}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{img/Example1-new-attack.pdf}
+\caption{Example RationalGRL model (extension of Fig.~\ref{fig:example-small2})}
+\label{fig:example-small3}
+\end{figure}
+
+We illustrate the new elements using Figure~\ref{fig:example-small3}, which is an extension of Figure~\ref{fig:example-small2}.
 
@@
 
-\paragraph{Rationale and example}
-
-\todo{Marc}{Marc}{Add extended version of Fig.~\ref{fig:example-small2} with some arguments and disabled elements and use that to explain this section}
-
+\paragraph{Rationale and example} A generic counterargument is simply an argument that can be used to attack any previous argument, or an IE.
+It is different from a belief element, since a generic counterargument is used in the computation to determine whether a RationalGRL element is accepted or rejected.
 Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link:contrib},~\ref{def:link:decomp}, and~\ref{def:link:dependency}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links.
 This is correct, since generic arguments are not part of GRL, so they should also not occur in links.
+
+In Fig.~\ref{fig:example-small3}, there is a single generic counterargument which can be formalized as $(13, \text{Redundant})$.
 
@@
 
+In Fig.~\ref{fig:example-small3}, the actor, all the intentional elements, all links with identifiers, and the generic counter argument are arguments.
+We will give a full formalization of this after the definition of a RationalGRL model (Def.~\ref{def:rationalgrl-model}.
+
 \begin{definition}[Attack Link]
@@
 \paragraph{Rationale and example} The attack link is the only link that RationalGRL adds to GRL.
+It is used to attack arguments.
+In Fig~\ref{fig:example-small3}, there is a single attack link that is used to connect the following two elements:
+\begin{itemize}
+\item $(13, \text{Redundant}$
+\item $(2, Resource, \text{Car objects}, AND)$
+\end{itemize}
+
+The attack link is formalized as $(13, 2)$.
 
@@
 \label{def:rationalgrl-model}
-A \emph{RationalGRL model} $RatGRL=(Arg, Att, ArgElem)$ consists of a set of arguments $Args$ (Def.~\ref{def:argument}), a set of attack links $Att$ (Def.~\ref{def:link:attack}), and an of argument-element mappings (Def.~\ref{def:arg-elem-mapping}).
+A \emph{RationalGRL model} $RatGRL=(Arg, Att, ArgElem)$ consists of a set of arguments $Args$ (Def.~\ref{def:argument}), a set of attack links $Att$ (Def.~\ref{def:link:attack}).
 \end{definition}
+
+\paragraph{Rational and example} The definition of a GRL model collects all the previously defined tuples into a single definition.
+For completeness, we now provide the full specification of Figure~\ref{fig:example-small3}.
+This model is formalized as
+ $RationalGRL=(Arg, Att)$ where:
+\begin{flalign*}
+Arg=&\{((2, task, \text{Car objects}, AND), Task),&\\
+   &  ((3, softgoal, \text{Real. sim.}, AND), Softgoal),&\\
+   &  ((4, softgoal, \text{Simple des.}, AND), Softgoal),&\\
+   &  ((5, goal, \text{Generate traffic}, XOR), Goal),&\\
+   &  ((6, task, \text{Create new cars}, AND), Task),&\\
+   &  ((7, task, \text{Keep same cars}, AND)\}, Task),&\\
+   &  ((0, \text{Traffic Simulator}), Actor),&\\
+   &  \{((0,i), ActIE)\mid 2\le i\le 7\},&\\
+   &  ((9, 6, 3, \text{Help}), Contr),&\\
+   &  ((10, 6, 4, \text{Hurt}), Contr),&\\
+   &  ((11, 6, 5), Decomp),&\\
+   &  ((12, 7, 5), Decomp),&\\
+   &  ((8, 5, 2), Dep),&\\
+   &  ((13, \text{Redundant}), GenArg),&\\
+Att=&\{(13,2)\}
+\end{flalign*}
 
@@
 \begin{tikzpicture}
-        \node[minimum size=1cm] (b) [argNodeIN] at (-2,0) {$A_1$};
-        \node[minimum size=1cm] (c) [argNodeIN] at (0,0) {$A_2$};
-        \node[minimum size=1cm] (d) [argNodeIN] at (2,0) {$A_3$};
+        \node[minimum size=1cm] (b) [argNodeIN] at (-3,0) {Car objects ($A_1$)};
+        \node[minimum size=1cm] (c) [argNodeIN] at (0,0) {Redundant ($A_2$)};
+        \node[minimum size=1cm] (d) [argNodeIN] at (3,0) {Necessary ($A_3$)};
 
@@
 
-Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, where $A_2$ attacks $A_1$ and $A_3$ attacks $A_2$.
-There are two admissible sets: $\{A_3\}$ and $\{A_1, A_3\}$, in which $A_3$ defends $A_1$ against its attacker $A_2$.
+Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which contains two of the arguments used in Figure~\ref{fig:example-small3}, ``Car objects'' ($A_1$) and ``Redundant'' ($A_2$), and contains an additional argument ``Necessary'' ($A_3$) attacking argument ``Redundant''.
+In this example, there are two admissible sets: $\{A_3\}$ and $\{A_1, A_3\}$, in which $A_3$ defends $A_1$ against its attacker $A_2$.
 Sets containing both $A_2$ and either $A_1$ or $A_3$ are not conflict free, and the sets $\{A_1\}$ and $\{A_2\}$ do not defend themselves against $A_2$ and $A_3$, respectively.
@@
 
-\todo{Marc}{Marc}{Say something about what we do with multiple extensions.}
-
 \begin{definition}[Preferred semantics~\cite{Dung1995}] \label{def:semantics}
-A preferred extension of a RationalGRL model $R$ is a maximal (w.r.t. set inclusion) admissible set of $R$.
+A preferred extension of an argumentation framework $(Arg,Att)$ is a maximal (w.r.t. set inclusion) admissible set of $R$.
 \end{definition}
@@
 In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_1,A_3\}$.
-
-\todo{Marc}{Marc}{Give example with concrete RationalGRL model}
+This means that arguments ``Car objects'' and ``Necessary'' are both \emph{accepted}, while ``Redundant'' is rejected.
+We say that the argument for ``Car objects'' is \emph{reinstated} by argument ``Necessary'', which attacks its only attacker ``Redundant'', thereby making ``Car objects'' acceptable.
 
@@
     \State $Decomp\leftarrow\emptyset$, $Dep\leftarrow\emptyset$
-    \For{$(A.id, T)\in ArgElem \mid A\in E$}
+    \For{$(Elem, T)\in E$}
       \Switch{$T$}
           \Case{$T\in Type$}
-            \State $IE\leftarrow IE\cup \{A\}$
+            \State $IE\leftarrow IE\cup \{Elem\}$
           \EndCase
           \Case{$T = Actor$}
-            \State $Act\leftarrow Act \cup\{A\}$
+            \State $Act\leftarrow Act \cup\{Elem\}$
           \EndCase
           \Case{$T=Contr$}
-            \State $Contr\leftarrow Contr\cup \{A\}$
+            \State $Contr\leftarrow Contr\cup \{Elem\}$
           \EndCase
           \Case{$T=Dep$}
-            \State $Dep\leftarrow Dep\cup \{A\}$
+            \State $Dep\leftarrow Dep\cup \{Elem\}$
           \EndCase
           \Case{$T=Decomp$}
-            \State $Decomp\leftarrow Decomp\cup\{A\}$
+            \State $Decomp\leftarrow Decomp\cup\{Elem\}$
           \EndCase
@@
 \end{definition}
-
-\todo{Marc}{Marc}{Explain conditions here and give some examples}
 

### 2017-10-21 formalframework.tex
Finished translation algorithms

@@
 
-In Fig.~\ref{fig:example-small3}, there is a single generic counterargument which can be formalized as $(13, \text{Redundant})$.
+In Fig.~\ref{fig:example-small3}, there are two generic counterargument which can be formalized as $(19, \text{Redundant})$ and $(20, \text{Necessary})$.
 
@@
 
-In Fig.~\ref{fig:example-small3}, the actor, all the intentional elements, all links with identifiers, and the generic counter argument are arguments.
-We will give a full formalization of this after the definition of a RationalGRL model (Def.~\ref{def:rationalgrl-model}.
+In Fig.~\ref{fig:example-small3}, the actor, all IEs, all links, and all generic counterarguments are arguments.
+We will give a full formalization of this after the definition of a RationalGRL model (Def.~\ref{def:rationalgrl-model}).
 
@@
 \label{def:link:attack}
-An attack link $att\in \mathbb{N}\times \mathbb{N}$ is a relation such that $att=(i,j)$ means:
+Given a set of arguments $Arg$, an attack link $att\in Arg\times Arg$ is a relation such that $att=(A_i,A_j)$ means:
 \begin{itemize}
-\item $i\in \mathbb{N}$ is the unique identifier of the element that is performing that attack,
-\item $j\in \mathbb{N}$ is the unique identifier of the element that is being attacked.
+\item $A_i\in Arg$ is the unique identifier of the element that is performing that attack,
+\item $A_j\in Arg$ is the unique identifier of the element that is being attacked.
 \end{itemize}
-Intuitively, $att=(i,j)$ means that element $i$ is attacking element $j$. A set of attack links is denoted by $Att$.
+Intuitively, $att=(A_i,A_j)$ means that argument $A_i$ is attacking argument $A_j$. A set of attack links is denoted by $Att$.
 \end{definition}
@@
 \paragraph{Rationale and example} The attack link is the only link that RationalGRL adds to GRL.
-It is used to attack arguments.
-In Fig~\ref{fig:example-small3}, there is a single attack link that is used to connect the following two elements:
+In Fig~\ref{fig:example-small3}, there are three attack links.
+The first two attack links originate from generic counterarguments and involve the following three elements:
 \begin{itemize}
-\item $(13, \text{Redundant}$
-\item $(2, Resource, \text{Car objects}, AND)$
+\item $A_1 = ((1, Resource, \text{Car objects}, AND), Resource)$
+\item $A_{19} = ((19, \text{Necessary}), GenArg)$
+\item $A_{20} = ((20, \text{Redundant}), GenArg)$
 \end{itemize}
 
-The attack link is formalized as $(13, 2)$.
+These two attack links are formalized as $(A_{20},A_{19})$ and $(A_{19}, A_1)$.
+
+The third attack link is an attack link created by \emph{replacing} and existing argument (we will discuss this process in more detail in Sect.~\ref{sect:algorithms}).
+It involves two version of the same IE:
+\begin{itemize}
+\item $A_{4} = ((5, goal, \text{Generate traffic}, AND), Goal)$
+\item $A_{18} = ((5, goal, \text{Generate traffic}, XOR), Goal)$
+\end{itemize}
+
+This attack link is formalized at $(A_{18},A_4)$.
+
+The last example shows an important difference between RationalGRL models and valid GRL models: While a valid GRL model disallows multiple elements with the same identifier (Def.~\ref{def:valid-grl-model}, condition 1), RationalGRL models do not enforce this restriction.
+This is because it is possible to create multiple arguments for the same element, where argument contains different content for the same element.
+This is what the example above also shows.
+However, the set of \emph{accepted} elements in a RationalGRL should all have unique identifier (see Def.~\ref{def:valid-rationalgrl-model}).
 
@@
 \label{def:rationalgrl-model}
-A \emph{RationalGRL model} $RatGRL=(Arg, Att, ArgElem)$ consists of a set of arguments $Args$ (Def.~\ref{def:argument}), a set of attack links $Att$ (Def.~\ref{def:link:attack}).
+A \emph{RationalGRL model} $RatGRL=(Arg, Att)$ consists of a set of arguments $Args$ (Def.~\ref{def:argument}) and a set of attack links $Att$ (Def.~\ref{def:link:attack}).
 \end{definition}
 
-\paragraph{Rational and example} The definition of a GRL model collects all the previously defined tuples into a single definition.
+\paragraph{Rationale and example} The definition of a RationalGRL model collects all the previously defined tuples into a single definition.
 For completeness, we now provide the full specification of Figure~\ref{fig:example-small3}.
-This model is formalized as
- $RationalGRL=(Arg, Att)$ where:
+Let us first enumerate all the arguments used in this example:
 \begin{flalign*}
-Arg=&\{((2, task, \text{Car objects}, AND), Task),&\\
-   &  ((3, softgoal, \text{Real. sim.}, AND), Softgoal),&\\
-   &  ((4, softgoal, \text{Simple des.}, AND), Softgoal),&\\
-   &  ((5, goal, \text{Generate traffic}, XOR), Goal),&\\
-   &  ((6, task, \text{Create new cars}, AND), Task),&\\
-   &  ((7, task, \text{Keep same cars}, AND)\}, Task),&\\
-   &  ((0, \text{Traffic Simulator}), Actor),&\\
-   &  \{((0,i), ActIE)\mid 2\le i\le 7\},&\\
-   &  ((9, 6, 3, \text{Help}), Contr),&\\
-   &  ((10, 6, 4, \text{Hurt}), Contr),&\\
-   &  ((11, 6, 5), Decomp),&\\
-   &  ((12, 7, 5), Decomp),&\\
-   &  ((8, 5, 2), Dep),&\\
-   &  ((13, \text{Redundant}), GenArg),&\\
-Att=&\{(13,2)\}
+A_0 = &((0, Actor, \text{Traffic simulator}), Actor)&\\
+A_1 = &((1, task, \text{Car objects}, AND), Task),&\\
+A_2 = &((2, softgoal, \text{Real. sim.}, AND), Softgoal),&\\
+A_3 = &((3, softgoal, \text{Simple des.}, AND), Softgoal),&\\
+A_4 = &((4, goal, \text{Generate traffic}, AND), Goal),&\\
+A_5 = &((5, task, \text{Create new cars}, AND), Task),&\\
+A_6 = &((6, task, \text{Keep same cars}, AND)\}, Task),&\\
+A_7 = &((7, 4, 1), Dep),&\\
+A_8 = &((8, 5, 2, \text{Help}), Contr),&\\
+A_9 = &((9, 5, 3, \text{Hurt}), Contr),&\\
+A_{10} = &((10, 5, 4), Decomp),&\\
+A_{11} = &((11, 6, 4), Decomp),&\\
+A_i = &((0,i-11),ActIE), \text{for } 12\le i\le 17&\\
+A_{18} = &((4, goal, \text{Generate traffic}, XOR), Goal),&\\
+A_{19} = &((18, \text{Redundant}), GenArg),&\\
+A_{20} = &((19, \text{Necessary}), GenArg),&
 \end{flalign*}
+
+This model is then formalized as $RationalGRL=(Arg, Att)$ where:
+\begin{flalign*}
+Arg = &\{A_0,A_1,\ldots,A_{20}\}\\
+Att=&\{(A_{18},A_4), (A_{19},A_1), (A_{20},A_{19})\}
+\end{flalign*}
+
+All the arguments and the attack relations of this RationalGRL model are shown in Figure~\ref{fig:example-small4}.
+Note the arguments for Actor-IE containment (arguments $A_{12}$ to $A_{20}$) have been omitted from this figure for readability.
+It can be read from the figure that two arguments are currently rejected, namely $A_4$ and $A_{19}$.
+However, we did not yet make precise how exactly this is computed.
+We will do so in the following definitions.
+
+\begin{figure}[ht]
+\centering
+\includegraphics[width=\columnwidth]{img/Example1-new-arguments}
+\caption{Argumentation framework of RationalGRL model from Fig.~\ref{fig:example-small3}}
+\label{fig:example-small4}
+\end{figure}
 
@@
 
-\begin{definition}[Argumentation Framework~\cite{}]
+\begin{definition}[Argumentation Framework~\cite{Dung1995}]
 \label{def:argumentation-framework}
@@
 \begin{tikzpicture}
-        \node[minimum size=1cm] (b) [argNodeIN] at (-3,0) {Car objects ($A_1$)};
-        \node[minimum size=1cm] (c) [argNodeIN] at (0,0) {Redundant ($A_2$)};
-        \node[minimum size=1cm] (d) [argNodeIN] at (3,0) {Necessary ($A_3$)};
+        \node[minimum size=1cm] (a20) [argNodeIN] at (-3,0) {$A_{20}$};
+        \node[minimum size=1cm] (a19) [argNodeIN] at (0,0) {$A_{19}$};
+        \node[minimum size=1cm] (a1) [argNodeIN] at (3,0) {$A_1$};
 
          \path
-    (c) edge [attackLink] (b)
-    (d) edge [attackLink] (c);
+    (a20) edge [attackLink] (a19)
+    (a19) edge [attackLink] (a1);
 
 \end{tikzpicture}
-\caption{Example argumentation framework.}
+\caption{Example argumentation framework, subset of Fig.~\ref{fig:example-small4}.}
 \label{fig:goalmodeling:arg2}
@@
 
-Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which contains two of the arguments used in Figure~\ref{fig:example-small3}, ``Car objects'' ($A_1$) and ``Redundant'' ($A_2$), and contains an additional argument ``Necessary'' ($A_3$) attacking argument ``Redundant''.
-In this example, there are two admissible sets: $\{A_3\}$ and $\{A_1, A_3\}$, in which $A_3$ defends $A_1$ against its attacker $A_2$.
-Sets containing both $A_2$ and either $A_1$ or $A_3$ are not conflict free, and the sets $\{A_1\}$ and $\{A_2\}$ do not defend themselves against $A_2$ and $A_3$, respectively.
+Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of Figure~\ref{fig:example-small4} containing only arguments $A_1, A_{19}$, and $A_{20}$.
+In this example, there are two admissible sets: $\{A_{20}\}$ and $\{A_1, A_{20}\}$.
+In the last admissible set, $A_{20}$ defends $A_1$ against its attacker $A_{19}$.
+Sets containing both $A_{19}$ and either $A_1$ or $A_{20}$ are not conflict free, and the sets $\{A_1\}$ and $\{A_{19}\}$ do not defend themselves against $A_{19}$ and $A_{20}$, respectively.
 
 Given the notion of admissible sets, we can then define our argumentation semantics.
-There are a large number of different semantics to determine which arguments are acceptable; in this article, we choose preferred semantics.
+There are a large number of different semantics to determine which arguments are acceptable; in this article, we focused on the preferred semantics.
 
-\begin{definition}[Preferred semantics~\cite{Dung1995}] \label{def:semantics}
+\begin{definition}[Preferred semantics~\cite{Dung1995}] 
+\label{def:preferred-semantics}
 A preferred extension of an argumentation framework $(Arg,Att)$ is a maximal (w.r.t. set inclusion) admissible set of $R$.
@@
 
-In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_1,A_3\}$.
-This means that arguments ``Car objects'' and ``Necessary'' are both \emph{accepted}, while ``Redundant'' is rejected.
-We say that the argument for ``Car objects'' is \emph{reinstated} by argument ``Necessary'', which attacks its only attacker ``Redundant'', thereby making ``Car objects'' acceptable.
+In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_{20},A_1\}$.
+Returning to our running example of Figure~\ref{fig:example-small3}, this means that the intentional elements \text{Car objects} (task) and the generic counerargument \emph{Necessary} are both \emph{accepted}, while the generic counter argument ``Redundant'' is rejected.
+
+\subsection{Translation algorithms}
+
+Now that we have formalized both a GRL model and a RationalGRL model, we develop algorithm in order to translate between these two models.
+Both of these two translation algorithms are straightforward.
+Their main task is to ensure that in RationalGRL models, the GRL elements and links are wrapped in an argument with the corresponding type.
+
+\paragraph{GRL to RationalGRL Translation} We start with the translation algorithm from GRL to RationalGRL, which is shown in Algorithm~\ref{alg:translation:to-rationalgrl}.
+The translation algorithm translates all of the GRL elements and links into arguments.
+It takes a GRL model as input.
+Recall that a GRL model is defined as $GRL=(IE, Bel, Act, R_{ActIE}, Bellink, Contr, Decomp, Dep)$ (Def.~\ref{def:grl-model}).
+Then, on line~\ref{alg:translation:to-rationalgrl:init}, the set of arguments is initialized with the empty set.
+Each of the five for loops in the algorithm translates a different GRL component into an argument.
+For instance, in the for loop starting at line~\ref{alg:translation:to-rationalgrl:ie}, all IEs are added as pairs consisting of the IE itself, and its corresponding type.
+This is in line with the definition of an argument~\ref{def:argument}).
+Since the other for loops are very similar, we do not explain them here.
+At line~\ref{alg:translation:to-rationalgrl:return}, the algorithm returns a RationalGRL model $(Arg,\emptyset)$ containing all arguments in $Arg$, and an empty set of attack relations.
 
 \begin{algorithm}[h]
-  \caption{RationalGRL to GRL Translation}\label{alg:translation}
+  \caption{GRL to RationalGRL Translation}
+  \label{alg:translation:to-rationalgrl}
   \begin{algorithmic}[1]
-    \Procedure{$Translate$}{$RatGRL$}
+    \Procedure{$TranslateToRationalGRL$}{$GRL$}
+    \State $Arg \leftarrow \emptyset$\label{alg:translation:to-rationalgrl:init}
+    \For{$ie\in IE$}\label{alg:translation:to-rationalgrl:ie}
+      \State $Arg \leftarrow Arg \cup \{(ie,ie.type)\}$
+    \EndFor
+    \For{$act\in Act$}
+      \State $Arg \leftarrow Arg \cup \{(act, Actor)\}$
+    \EndFor
+    \For{$actIE\in ActIE$}
+      \State $Arg \leftarrow Arg \cup \{(actIE, ActIE)\}$
+    \EndFor
+    \For{$contr\in Contr$}
+      \State $Arg \leftarrow Arg \cup \{(contr, Contr)\}$
+    \EndFor
+    \For{$dep\in Dep$}
+      \State $Arg \leftarrow Arg \cup \{(dep, Dep)\}$
+    \EndFor
+    \For{$decomp\in Decomp$}
+      \State $Arg \leftarrow Arg \cup \{(decomp, Decomp)\}$
+    \EndFor
+    \State \Return $(Arg, \emptyset)$\label{alg:translation:to-rationalgrl:return}
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{RationalGRL to GRL translation} The translation from a RationalGRL model to a GRL model is given in Algorithm~\ref{alg:translation:to-grl}.
+It is very similar to the previous algorithm, but then works in the other direction.
+In this case, arguments are unpacked and put in the corresponding IE components.
+The most important difference between the previous algorithm and this one is that in the RationalGRL to GRL translation, argumentation semantics is used to determine the \emph{preferred extension} (Def.~\ref{def:preferred-semantics}).
+This is a set containing all accepted arguments.
+The GRL is then generated from all the accepted arguments.
+This is done by iterating over all arguments in the extension in line~\ref{alg:translation:to-grl:extension-iter}.
+The switch statement on line~\ref{alg:translation:to-grl:switch} then does a case distinction on the type of the arguments, and each case ensures the argument is put in the right GRL component.
+Finally, the algorithm returns a GRL model on line~\ref{alg:translation:to-grl:return}.
+
+\begin{algorithm}[h]
+  \caption{RationalGRL to GRL Translation}
+  \label{alg:translation:to-grl}
+  \begin{algorithmic}[1]
+    \Procedure{$TranslateToGRL$}{$RatGRL$}
     \State $E \leftarrow ComputeExtension(Arg,Att)$
@@
     \State $Decomp\leftarrow\emptyset$, $Dep\leftarrow\emptyset$
-    \For{$(Elem, T)\in E$}
-      \Switch{$T$}
+    \For{$(Elem, T)\in E$}\label{alg:translation:to-grl:extension-iter}
+      \Switch{$T$}\label{alg:translation:to-grl:switch}
           \Case{$T\in Type$}
@@
     \EndFor
-    \State \Return $(IE,Act,R_{ActIE}, Contr, Decomp, Dep)$
+    \State \Return $(IE,Act,R_{ActIE}, Contr, Decomp, Dep)$\label{alg:translation:to-grl:return}
     \EndProcedure
@@
 
-The previous definition of a preferred extension allows us to compute accepted arguments in a RationalGRL model. In the following definition we make precise the notion of a valid RationalGRL model.
+\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Def.~\ref{def:valid-grl-model}), we have not done so for a RationalGRL model yet.
+This is because it depends on the translation procedure.
+We define a RationalGRL model as valid if and only if the RationalGRL to GRL translation results in a valid GRL model.
+In this way, we do not have to reiterate all conditions on a GRL model, but use the translation algorithm.
 
 \begin{definition}[Valid RationalGRL Model]
-Suppose a RationalGRL model $RatGRL = (Arg, Att, ArgElem)$ with preferred extension $E=(A_1,\ldots,A_n)$.
-$RatGRL$ is a \emph{valid RationalGRL model} iff $Translate(RatGRL)$ (Alg.~\ref{alg:translation}) is a valid GRL model (Def.~\ref{def:valid-grl-model}).
+\label{def:valid-rationalgrl-model}
+A RationalGRL model $RatGRL = (Arg, Att)$ with preferred extension $E=(A_1,\ldots,A_n)$ is a \emph{valid RationalGRL model} iff $TranslateToGRL(RatGRL)$ (Alg.~\ref{alg:translation:to-grl}) is a valid GRL model (Def.~\ref{def:valid-grl-model}).
 \end{definition}

### 2017-10-23 discussion.tex
related research finished

@@
 There are several contributions to the literature that relate argumentation-based techniques with goal modeling.
+The first line of work is by Bagrheri and Ensan \cite{bagheri2011consolidating} and Mirbel and Villata \cite{MirbelVillata12} use abstract Dung argumentation frameworks (see Section~\ref{sect:formalframework}) to capture individual goals and the relations between them as modeled in a goal model, so that consistent (i.e., acceptable) subsets of goals can be computed using the relevant argumentation semantics.
+For example, if in the goal model there is a conflict between goals $G_1$ and $G_2$, there is an attack between the arguments representing these goals and hence $G_1$ and $G_2$ cannot be in the same extension.
+Or if there is a dependency link between $G_1$ and $G_2$, then $G_1$ should be in every extension $G_2$ is in and vice versa.
+AND and OR decomposition are also modeled thus, that is, if $G_3$ AND-decomposes into $G_1$ and $G_2$ then $G_1$ and $G_2$ should be in every extension $G_3$ is in, and if $G_3$ OR-decomposes into $G_1$ and $G_2$ then either $G_1$ or $G_2$ should be in every extension $G_3$ is in.
+
+Modeling goal models as argumentation frameworks allows one to compute the possible consistent sets of goals and tasks given a goal model.
+In RationalGRL, we have opted not to provide such an argumentation-theoretic semantics to goal models.
+The reason for this is that GRL already has quite fine-grained satisfiability semantics \cite{Amyot:2010:EGM:1841349.1841356}, which take into account conflicts, decompositions and dependencies.
+RationalGRL focuses on what is not captured in the work discussed above \cite{bagheri2011consolidating,MirbelVillata12}, namely the arguments and beliefs underlying a goal model, and the way these arguments and beliefs interact with the elements of the goal model.
+If desired, however, it would be easy to provide Dung semantics similar to \cite{bagheri2011consolidating,MirbelVillata12} for goal models, as the elements of a goal model are already arguments in RationalGRL (Figure~\ref{fig:metamodel}).
+
 The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
 Jureta \emph{et al.} propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
-One of the elements of GAM is the translation of formal argument models to goal models (similar to ours).
-In this sense, our RationalGRL framework can be seen as an instantiation and implementation of  part of the GAM.
-The main difference between our approach and GAM is that we integrate arguments and goal models using argument schemes, and that we develop these argument schemes by analyzing transcripts.
-GAM instead uses structured argumentation.
+GAM is a high-level decision process, in which alternative solutions for an RE problem are evaluated and compared using argumentation.
+Jureta et al.
+give a simple made-up example of an argumentative discussion regarding a meeting scheduler, in which a goal model is being built by the stakeholders proposing tasks, goals, and alternative solutions for goals.
+They include clarification as an important step in their GAM process, and discuss various types of clarity problems (vagueness, ambiguity, overgenerality, synonymy) and basic techniques for dealing with them (e.g.
+thesaurus checks, labeling vague expressions).
+Jureta et al.
+then present the argumentation part of their framework, where reasons (justifications) for conclusions are given as formal structured arguments\footnote{Informally, a structured argument is similar to the PRAS argument in Section~\ref{sect:background:pras}, i.e., $a, b \vec{therefore} c$.}.
+Arguments and alternatives are then captured as structured arguments or argument diagrams with reasons for or against goals and tasks.
+Given these arguments the set of undefeated (i.e., acceptable) propositions can be computed to determine which alternative solution to a problem is acceptable.
+Thus, the arguments and beliefs underlying a goal model and possible alternative modelings are captured as formal arguments.
+Furthermore, a mapping from goal models to argument diagrams is given, so that it is possible to start arguing about an already existing goal model.
 
-In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we have explored one way of combining PRAS and GRL.
-In this work, which takes a similar approach to \cite{Jureta:RE2008}, argument diagrams are translated to GRL models (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
-The argument diagrams are complex practical reasoning arguments, where the premises are the goals and the conclusion is some action we need to perform to realize those goals.
-Thus, we have essentially two complex diagrams, a practical reasoning argument diagram and a GRL goal diagram, and a mapping between them.
+The GAM process is essentially a generic, high-level process for problem solving, not specifically tailored towards goal modeling.
+In this sense, the RationalGRL methodology in Figure~\ref{fig:rationalgrl-methodology} can be seen as a more specific version of GAM explicitly meant for goal modeling.
+The argument schemes and critical questions in RationalGRL provide more clear handles for goal modeling (cf.
+requirement 4 of our framework, Section~\ref{sect:introduction}).
 
-Connecting argumentation and goal modelling by providing a mapping between two types of diagram was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
+In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we effectively extended Jureta et al.'s work and translated argument diagrams to GRL models\footenote{In \cite{Jureta:RE2008} only a mapping from goal models to structured arguments is given, and the step from structured arguments or argument diagrams to goal models is never formally defined.} (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
+Thus, we have essentially two complex diagrams, an argument diagram and a goal diagram, and a mapping between them.
+This was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
 One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
-Furthermore, in our previous work we did not provide the specific critical questions for goal models (see Section \ref{sect:gmas}).
-This meant that any counterargument had to be constructed from scratch, as no guidance was given as to possible ways to criticize a GRL model.
 So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged.
-Furthermore, not having specific argument schemes and critical questions for goal modelling makes it hard to develop a guiding methodology for the use of argumentation in goal modelling (requirement 4).
 
-One idea is to capture requirements engineering and software design processes as explicit dialogues between parties~\cite{finkelstein1989multiparty}.
-
-\todo{F}{F}{Review the following:
-
-\begin{itemize}
-\item Mirbel and Villata on Dung \& goal models
-\item Software design discussions Black et al., Prakken and Wierenga, \cite{finkelstein1989multiparty}
-  \end{itemize}
-}
 \subsection{Open issues}
@@
 
+One idea is to capture requirements engineering and software design processes as explicit dialogues between parties~\cite{finkelstein1989multiparty}.
+Software design discussions Black et al., Prakken and Wierenga.
+
 \subsubsection*{Empirical study}

### 2017-10-23 formalframework.tex

@@
 \subsubsection*{Algorithms for critical questions}
-
-We now develop algorithms for our critical questions. Recall that answering a critical question can have four effects, and we discuss each of these effects separately.
-
-\begin{algorithm}[h]
-  \caption{Applying DISABLE: Element $i$ is disabled}\label{alg:disable}
-  \begin{algorithmic}[1]
-    \Procedure{DISABLE}{$i$}
-    \State $id\gets id+1$
-    \State $A\gets \{disabled(i))\}$
-    \State $Args \gets Args\cup A$
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\todo{Marc}{Marc}{Here we would add an argument and an attack from the argument to the intentional element. We don't add ``disabled'' because this is what argumentation semantics should do for us.}
-
-\emph{Algorithm~\ref{alg:disable} (\textsf{DISABLE}) for critical questions CQ0-CQ5a, CQ6a, CQ7a, CQ8, CQ11, and CQ12:} The disable operation consists of adding an argument stating the GRL element with identifier $i$ is disabled.
-Let us reconsider the example of Figure~\ref{fig:examples:relevant-actor}.
-This example consists of an instantiation of argument scheme AS0, which is attacked by an argument that resulted from answering critical question.
-The instantiation of AS0 leads to the argument $A=\{actor(0),name(0,dev\_team\}$.
-After applying $DISABLE(0)$ we obtain the arguments: $$Args=\{\{actor(0),name(0,dev\_team\}, \{disabled(0)\}\}.$$ Note that our implementation of this attack does not lead to an actual attack in the argumentation framework.
-
-\begin{algorithm}[h]
-  \caption{Answering CQ5b: ``Does goal $G$ decompose into any other tasks?'' With: ``Yes, namey into tasks $t_1,\ldots,t_k$''}\label{alg:cq5b}
-  \begin{algorithmic}[1]
-    \Procedure{CQ5b}{$g_{id},\{i_1,\ldots,i_n\}, type, \{t_1,\ldots,t_k\}$}
-    \State $T_{id} = \{i_1,\ldots,i_n\}$
-    \For{$t_i$ in $\{t_1,\ldots,t_k\}$}\label{alg:cq5b:for1}
-      \If{$\exists_{A\in Args} \{task(t_{id}),name(t_{id},t_i)\}\subseteq A$}
-        \State $T_{id} \gets T_{id} \cup \{t_{id}\}$
-      \Else
-        \State $id\gets id+1$
-        \State $A \gets \{task(id),name(id,t_i)\}$
-        \State $Args \gets Args\cup A$
-        \State $T_{id} \gets T_{id} \cup \{id\}$
-      \EndIf
-    \EndFor
-    \State $id\gets id+1$
-    \State $A_{new} = \{decomp(id, g_{id}, T_{id}, type)\}$\label{alg:cq5b:newarg}
-    \For{$A$ in $\{decomp(\_,g_{id},\_,\_)\}\subseteq A\mid A\in Args\}$}\label{alg:cq5b:for2}
-      \State $Att\gets Att \cup \{(A_{new},A)\}$
-    \EndFor
-    \State $Args \gets Args \cup \{A_{new}\}$
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\emph{Algorithm~\ref{alg:cq5b} (\textsf{REPLACE}) for critical questions CQ5b:} This algorithm is executed when critical question CQ5b is answered, which is a critical question for argument scheme AS5.
-Therefore, it assumes an argument for a goal decomposition already exists of the following form (see Algorithm~\ref{alg:as5}): $$\{decomp(d, g_{id}, \{i_1,\ldots,i_n\},type).$$  The goal of the algorithm is to generate a new argument of the form $decomp(d, g_{id}, \{i_1,\ldots,i_5\}\cup \{j_1,\ldots,j_k\}, type)$, where $\{j_1,\ldots,j_k\}$ are the identifiers of the additional decomposing tasks $\{t_1,\ldots,t_k\}$.
-
-The algorithm takes as input the goal identifier $g_{id}$, the set of existing decomposing task identifiers $i_1,\ldots,i_n$, the decomposition type, and the names of the new tasks $t_1,\ldots,t_k$ that should be added to the decomposition.
-The first part of the algorithm is familiar from Algorithm~\ref{alg:as5}: For each task name we check whether it already exists as an argument (line 4), and if it doesn't (line 6) we add a new argument for it.
-After the for-loop (line 13), a new argument is created for the new decomposition relation (14).
-After this, the for-loop on line 15 ensures that the new argument attacks all previous arguments for this decomposition (note that the variable ``$\_$'' means ``do not care'').
-Only at the very end the new argument is added (line 18), to ensure it does not attack itself after the for loop of line 15-17.
-
-An example of this algorithm is shown in Figure~\ref{fig:examples:cq5b}.\footnote{Note that part of the arguments (the statements about actors) have been omitted from the figure for readability.} Before the critical question is applied, the following arguments have been put forward:
-\begin{itemize}
-\item $\{goal(0),name(0,show\_simulation)\}$
-\item $\{task(1),name(1,generate\_traffic)\}$
-\item $\{task(2),name(2,compute\_lights)\}$
-\item $\{decomp(3,0,\{1,2\},and)\}$.
-\end{itemize}
-
-Next, Algorithm~\ref{alg:cq5b} is called as follows: $CQ5b(0,\{1,2\},and,\{show\_controls\})$.
-That is, the existing decomposition is challenged by stating that goal $show\_simulation$ not only decomposes into $generate\_traffic$ and $compute\_lights$, but it also decomposes into $show\_controls$.
-Since this task does not exist yet, it is created by the algorithm, which also ensures the new argument for the decomposition link attacks the previous argument for the decomposition link.
-
-\begin{figure*}[ht!]
-\centering
-        \begin{tikzpicture}
-        \node (a0) [argNodeIN,minimum width=1.5cm] at (-5,0) {
-        	\smallargtext{AS3}{$\{goal(0),$\\$name(0,show\_sim\}$}
-        };
-        \node (a1) [argNodeIN] at (0,0) {
-        \smallargtext{AS2}{$\{task(1),$\\$name(1,gen\_traffic\}$}
-        };
-        \node (a2) [argNodeIN] at (5,0) {
-        \smallargtext{AS2}{$\{task(3),$\\$name(2,show\_ctrls\}$}
-        };
-        \node (a2a) [argNodeIN] at (-2.5,-2) {
-        \smallargtext{AS2}{$\{task(2),$\\$name(2,comp\_lights\}$}
-        };
-        \node (a3) [argNodeOUT] at (-3,-4) {$\{decomp(3,0,\{1,2\},and\}$
-        };
-        \node (a4) [argNodeIN] at (3,-4) {$\{decomp(3,0,\{1,2,3\},and\}$
-        };
-
-         \path
-    (a4) edge [attackLink] (a3)
-    (a3) edge [CQLink, bend right=50] node [below,draw=none] {CQ5b} (a4)
-    (a4) edge [CQLink] node [left,draw=none] {CQ5b} (a2);
-\end{tikzpicture}
-\caption{Example of applying critical question CQ5b (Algorithm~\ref{alg:cq5b})}
-\label{fig:examples:cq5b}
-\end{figure*}
-
-\emph{Algorithms for critical questions CQ10a and CQ10b (REPLACE)}: These algorithms have a very similar structure as Algorithm~\ref{alg:cq5b} and have therefore been omitted.
-
-\begin{algorithm}[h]
-  \caption{Answering CQ13: ``Is the name of element $i$ clear?'' With: ``No, it should be $n$''}\label{alg:cq13}
-  \begin{algorithmic}[1]
-    \Procedure{CQ13}{$i, n$}
-    \State $ArgsN \gets\{ A\in Args \mid name(i,x)\in A\}$
-    \State $B\gets B'\backslash \{name(i,\_)\}$ with $B'\in ArgsN$
-    \State $A \gets B \cup \{name(i,n)\}$
-    \State $Args \gets Args \cup \{A\}$
-    \For{$C$ in $ArgsN$}
-      \State $Att\gets Att \cup \{(A,C)\}$
-    \EndFor
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\emph{Algorithms for critical question CQ13 (REPLACE):} This algorithm is used to clarify/change the name of an element.
-It takes two parameters: the element identifier $i$ and the new name $n$.
-The idea behind the algorithm is that we construct a new argument for $n$, and to ensure that this argument attacks all previous arguments for giving a name to this element.
-Here we can use the following fact: Suppose $Args$ is a set of arguments, each containing an atom about the name for an element $i$, i.e.
-for all $A\in Args: name(i,_)\in Args$, then all arguments in $Args$ are the same except for the $name$ atoms, i.e.
-for all $A,B\in Args: A\backslash\{name(i,_)\} = B\backslash\{name(i,_)\}$.
-This means that if we would like to attack every argument of $Args$ with a new argument that replaces the $name$ atom, we can simply take any argument in $Args$, remove the previous $name$ atom, add the new one and attack all arguments in $Args$.
-This is exactly what the algorithm does.
-
-On line 2, all arguments that have been put forward for this element and contain $name(i,x)$ are collected into the set $ArgsN$.
-On line 3, some arguments $B'\in ArgsN$ minus the $name$ statement is assigned to $B$ (note that it does not matter which one we pick), and on line 4 $B$ is joined with the new $name$ statement and stored in $A$, which is then added to the set of arguments $Args$.
-The for-loop on lines 6-8 ensures all previous arguments for names of the element are attacked by the new argument.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}[->]
-        \node (a0a) [argNodeIN] at (-2,2) {
-        	\longargtext{AS0}{$\{actor(0),name(0,student)\}$}
-        };
-        \node (a0) [argNodeOUT] at (-2,0) {
-        	\longargtext{AS2}{$\{task(1),name(1,create\_road),has(0,1)\}$}
-        };
-        \node (a1) [argNodeOUT] at (-2,-2) {
-        	\longargtext{AS2}{$\{task(1),name(1,choose\_pattern),has(0,1)\}$}
-        };
-        \node (a2) [argNodeOUT] at (-2,-4) {
-        	\longargtext{AS2}{$\{task(1),name(1,pattern\_pref),has(0,1)\}$}
-        };
-        \node (a3) [argNodeIN] at (-2,-6) {
-        	\longargtext{AS2}{$\{task(1),name(1,road\_pattern),has(0,1)\}$}
-        };
-\begin{pgfonlayer}{background}
-         \path
-    (a1) edge[attackLink] (a0)
-    (a2) edge[attackLink] (a1)
-    (a2) edge[attackLink, bend right=20] (a0)
-    (a3) edge[attackLink] (a2)
-    (a3) edge[attackLink, bend right=20] (a1)
-    (a3) edge[attackLink, bend right=40] (a0);
-\end{pgfonlayer}
-
-	\path
-	(a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ13} (a1)
-    (a1) edge [CQLink, bend right=50] node [left,draw=none] {CQ13} (a2)
-    (a2) edge [CQLink, bend right=50] node [left,draw=none] {CQ13} (a3);
-\end{tikzpicture}
-\caption{Applying critical question CQ13 (Algorithm~\ref{alg:cq13}) to the example in Figure~\ref{fig:examples:clarification}.}
-\label{fig:examples:clarification:formal}
-\end{figure}
-
-An example of of Algorithm~\ref{alg:cq13} is shown in Figure~\ref{fig:examples:clarification:formal}.
-Let us consider the last application of CQ13 (bottom argument).
-Before this application, the following arguments have been put forward:
-\begin{itemize}
-\item $A_1$: $\{actor(0),name(0,student)\}$
-\item $A_2$:$\{task(1),name(1,create\_road),has(0,1)\}$
-\item $A_3$ $\{task(1),name(1,choose\_pattern),has(0,1)\}$
-\item $A_4$:$\{task(1),name(1,pattern\_pref),has(0,1)\}$
-\end{itemize}
-The algorithm is now called as follows: $CQ13(1,road\_pattern)$, i.e., the new name of the element should be $road\_pattern$.
-Let us briefly run through the algorithm.
-After executing line 2 we obtain $ArgsN=\{A_2,A_3,A_4\}$, since only those arguments contain $name(1,\_)$.
-Next, on line 3, $B=\{task(1),has(0,1)\}$, i.e., $B$ is the general argument for the task without the $name$ statement.
-After line 4 we have $$A=\{task(1),has(0,1),name(1,road\_pattern),$$ which is added to $Args$ and attacks arguments $A_2,A_3$, and $A_4$.
-
-\emph{Algorithms for critical questions CQ6b, CQ6c, CQ6d, CQ7b, and CQ9 (INTRO):} The introduction algorithms for the critical questions are all very similar to the INTRO algorithms for argument schemes (Algorithm~\ref{alg:as1}).
-They have therefore been omitted.
-
-\begin{algorithm}[h]
-  \caption{Generic counterargument to argument $A$}\label{alg:cq6b}
-  \begin{algorithmic}[1]
-    \Procedure{Attack}{$A$}
-    \State $A_{new} = \{\}$
-    \State $Args \gets Args \cup \{A_{new}\}$
-    \State $Att \gets Att \cup \{(A_{new},A)\}$
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\emph{Algorithm for $Att$ (Generic counter argument:} Applying a generic counter argument is very simple, and simply results on an attack on the original argument.
-We illustrate this by continuing our example from Figure~\ref{fig:examples:relevant-actor2:formal} (Algorithm~\ref{alg:as0}).
-The example is shown in Figure~\ref{fig:examples:relevant-actor2:formal}, where we see that a generic counter argument simply attacks the argument to disable the actor.
-
-\begin{figure}[ht!]
-\centering
-        \begin{tikzpicture}
-        \node (a0) [argNodeIN] at (-3.7,0) {
-        	\smallargtext{AS0}{$\{actor(0),name(0,dev\_team\}$}
-        } ;
-        \node (a1) [argNodeOUT] at (-3.7,-2){
-        	\argtext{DISABLE}{$\{disabled(0)\}$}
-        };
-        \node (a2) [argNodeIN] at (-3.7,-4){
-        	\argtext{Att}{$\{\}$}
-        };
-         \path
-    (a0) edge [CQLink, bend right=50] node [left,draw=none] {CQ0} (a1)
-    (a1) edge [CQLink, bend right=50] node [left,draw=none] {Att} (a2)
-    (a2) edge [attackLink] (a1);
-\end{tikzpicture}
-\caption{Formalization of the arguments in Figure~\ref{fig:examples:relevant-actor2}.}
-\label{fig:examples:relevant-actor2:formal}
-\end{figure}
-
-\subsection{Constructing GRL models}
-\label{sect:rationalGRL-GRL}
-\todo{F}{F,M}{Dit is snel getypt, moet nog checken of het klopt.}
-
-\begin{definition}[RationalGRL model] A RationalGRL model $\mathcal{R}$ is an argumentation framework $AF=(Args,Att)$ where all arguments in $Args$ and attacks in $Att$ are based on algorithm X-XX.  
-\end{definition}
-
-\begin{definition}[GRL model] A GRL model $\mathcal{M}$ based on a RationalGRL model $\mathcal{R}$ is the preferred extension of $\mathcal{R}$.  
-\end{definition}
-=======
-\todo{F}{F,M}{It would be good to have a formal definition of how to construct models given an argumentation framework.
-If we define a model M at the start of this section, this should be fairly easy.}
-
-Constructing GRL models from the arguments is extremely simple: We simply compute the extensions of the argumentation frameworks, and collect all atomic sentences in the accepted arguments.
-This forms out GRL model.
-Let us briefly do so for the examples of the previous subsection:
-\begin{itemize}
-\item
-Figure~\ref{fig:examples:relevant-actor:formal}: Since there are no attacks between the arguments, all atomic sentences are accepted. This results in the following specification: 
-\begin{quote}
-\begin{verbatim}
-actor(0).
-name(0,dev_team).
-disabled(0).
-\end{verbatim}
-\end{quote}
-This again corresponds to the GRL model on the right-hand side of Figure~\ref{fig:examples:relevant-actor}.
-\item
-Figure~\ref{fig:examples:cq5b}: There is one rejected argument and five accepted ones. The resulting specification is:
-\begin{quote}
-\begin{verbatim}
-goal(0). name(0,show_simulation).
-task(1). name(1,generate_traffic).
-task(2). name(2,compute_lights).
-task(3). name(3,show_controls).
-decomp(3,0,{1,2,3},and).
-\end{verbatim}
-\end{quote}
-\item Figure~\ref{fig:examples:clarification:formal}: There are only two accepted arguments. The resulting specification is:
-\begin{quote}
-\begin{verbatim}
-actor(0). name(0, student).
-task(1).  name(1,road_pattern).
-has(0,1).
-\end{verbatim}
-\end{quote}
-This corresponds to the right-hand GRL model of Figure~\ref{fig:examples:clarification}.
-\item Figure~\ref{fig:examples:relevant-actor2:formal}.
-There are two accepted arguments, but the \emph{generic counterargument} does not contain any formulas.
-Therefore the resulting specification is:
-\begin{quote}
-\begin{verbatim}
-actor(0).
-name(0,dev_team).
-\end{verbatim}
-\end{quote}
-This corresponds to the right-hand GRL model of Figure~\ref{fig:examples:relevant-actor2}.
-\end{itemize}

### 2017-10-23 formalframework.tex
finished all algorithms

@@
 \begin{itemize}
-\item $A_i\in Arg$ is the unique identifier of the element that is performing that attack,
-\item $A_j\in Arg$ is the unique identifier of the element that is being attacked.
+\item $A_i\in Arg$ is the argument performing that attack,
+\item $A_j\in Arg$ is the argument being attacked.
 \end{itemize}
@@
 Their main task is to ensure that in RationalGRL models, the GRL elements and links are wrapped in an argument with the corresponding type.
+For each algorithm, we provide an explanation an an example of a translation.
 
@@
   \begin{algorithmic}[1]
-    \Procedure{$TranslateToRationalGRL$}{$GRL$}
+    \Procedure{$ToRationalGRL$}{$GRL$}
     \State $Arg \leftarrow \emptyset$\label{alg:translation:to-rationalgrl:init}
@@
   \begin{algorithmic}[1]
-    \Procedure{$TranslateToGRL$}{$RatGRL$}
+    \Procedure{$ToGRL$}{$RatGRL$}
     \State $E \leftarrow ComputeExtension(Arg,Att)$
@@
 
+The following theorem follows directly from the translation algorithms.
+
+\begin{theorem}[Translation correctness]
+\label{th:translation-correctness}
+Given some valid GRL model (Def.~\ref{def:valid-grl-model}, translating it to RationalGRL and then back to GRL results in the same valid GRL mode, i.e.
+$$ToGRL(ToRationalGRL(GRL)) = GRL.$$
+\end{theorem}
+
 \subsection{Algorithms for argument schemes and critical questions}
@@
 
-In the previous section we formalized a \emph{static} representation of the RationalGRL framework.
+In the previous subsection we formalized a \emph{static} representation of the RationalGRL framework.
 In this section we formalize the \emph{dynamics}.
 We do so by developing algorithms for applying argument schemes and critical questions in the context of a RationalGRL model (Def.~\ref{def:rationalgrl-model}).
-These algorithms produce new elements and attack relations.
+These algorithms are applied to RationalGRL models and produce new arguments and attack relations.
 We can then use argumentation semantics (Def.~\ref{def:semantics}) to compute sets of accepted arguments.
@@
 
-\paragraph{Argument Schemes}
+As discussion in Section~\ref{sect:overview}, ll the argument schemes and critical questions of Table~\ref{table:argument-schemes} fall into one of the following four categories:
+\begin{itemize}
+\item \emph{INTRO}: These algorithms add one or multiple new arguments to the RationalGRL model, without creating any attack links. The arguments are for new GRL elements or links.
+\item \emph{DISABLE}: These algorithms add a single argument to the RationalGRL model, which attacks one of multiple existing arguments.
+The counterargument that is added does not correspond to any GRL element, but instead disables an existing one by attacking all arguments for it.
+\item \emph{REPLACE}: These algorithms can be seen as a combination of \emph{INTRO} and \emph{DISABLE}.
+They add a new argument corresponding to a GRL element or link, and this new argument attacks a previous version of the element or link.
+\end{itemize}
+
+The three subsections of this section correspond to these three types of arguments.
 
@@
 \begin{itemize}
-\item There algorithms are being applied to some valid RationalGRL model $RatGRL$,
+\item There algorithms are being applied to some valid RationalGRL model $RatGRL$ (Def.~\ref{def:valid-rationalgrl-model}),
 \item The procedure $mindId()$ generates a new unique id.
@@
 
-The algorithms for applying argument schemes are short, and consist simply of adding an argument for the element that is being added.
+\subsubsection{INTRO algorithms}
+
+The following arguments schemes and critical questions of Table~\ref{table:argument-schemes} fall into this category:
+\begin{itemize}
+\item AS0-AS12
+\item CQ5b, CQ6b, CQ6c, CQ7b, CQ9, CQ10b
+\end{itemize}
+
+These type of algorithms are short, and consist simply of adding an argument for the element that is being added.
+Since the structure of these algorithms are all very similar, we do not discuss them all.
 
 \begin{algorithm}[h]
-  \caption{AS0: Add a new actor with name $n$}\label{alg:as0}
+  \caption{AS0: $a$ is an actor}\label{alg:add-actor}
   \begin{algorithmic}[1]
     \Procedure{$AS_0$}{$n$}
-    \State $actor \leftarrow (mintId(), n)$ \label{alg:as0:1}
+    \State $actor \leftarrow (mintId(), a)$ \label{alg:as0:1}
     \State $Arg\leftarrow Arg \cup \{(actor,Actor)\}$\label{alg:as0:2}
@@
 
-\paragraph{Rationale and example} The algorithm takes one argument, namely the name of the actor $n$.
+\paragraph{Rationale and example} Algorithm takes one argument, namely the name of the actor $a$.
 On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$.
@@
 
+Consider for instance...\todo{Marc}{Marc}{Add example}
+
 \begin{algorithm}[h]
-  \caption{Applying AS1: Actor with id $i$ has resource $n$}\label{alg:as1}
+  \caption{AS1: Actor with id $i$ has resource $R$}\label{alg:add-actor}
   \begin{algorithmic}[1]
-    \Procedure{$AS_1$}{$i, n$}
-    \State $rId\gets mintId()$
-    \State $r\leftarrow (rId, Resource, n, AND)$
-    \State $Arg\gets Arg\cup \{(r, Resource),(i,rId)\}$
+    \Procedure{$AS_1$}{$i, R$}
+    \State $res\_id\gets mintId()$\label{alg:add-actor:id}
+    \State $res\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-actor:res}
+    \State $Arg\gets Arg\cup \{(res, Resource),(i,res\_id)\}$\label{alg:add-actor:arg}
     \EndProcedure
@@
 
-\paragraph{Rationale and example} Explain here...
+\paragraph{Rationale and example} The original argument schemes AS1 is phrased as ``Actor $a$ has resource $R$''.
+We have slightly reworded this in Algorithm~\ref{alg:add-actor} to ``Actor with id $i$ has resource $R$''.
+The reason is that in GRL, actors have identifiers, so since we assume actor $a$ exists already, we also assume an identifier for it exists.
+It would be straightforward to introduce a mapping from actor names to identifiers, but we have no done so here for simplicity.
+
+The algorithm itself runs as follows: on line~\ref{alg:add-actor:id}, a unique id is assigned to variabel $res\_id$ (``resource identifier'').
+On line~\ref{alg:add-actor:res}, an argument for a resource with name $R$ and identifier $res\_id$ is assigned to $A$.
+Finally, on line~\ref{alg:add-actor:arg}, two arguments are added to the set of arguments $Arg$: one for the resource, and one for the containment relations between the actor and the resource (Def.~\ref{def:act-ie-relation}.
+
+Since arguments schemes AS2-AS4 are very similar to AS1, we have omitted it here.
 
 \begin{algorithm}[h]
-  \caption{Applying AS5: Goal with id $i$ decomposes into tasks $t$}\label{alg:as5}
+  \caption{AS5: Goal with id $i$ decomposes into task $t$}\label{alg:add-decomp}
   \begin{algorithmic}[1]
     \Procedure{$AS_5$}{$i, t$}
-    \State $tId\gets mintId()$
-    \State $t\leftarrow (tId, Task, t, AND)$
-    \State $Arg\gets Arg\cup \{(t, Task),(i,tId)\}$
+    \State $task\_id\gets mintId()$\label{alg:add-decomp:task-id}
+    \State $task\leftarrow (task\_id, Task, t, AND)$\label{alg:add-decomp:task}
+    \State $Arg\gets Arg\cup \{(task, Task),((mintId(),i,task\_id), Decomp)\}$\label{alg:add-decomp:arg}
     \EndProcedure
@@
 
-\paragraph{Rationale and example} Explain here.
+\paragraph{Rationale and example} Similar to the previous algorithm, we have slightly reworded critical question AS5.
+We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $t$ is a decomposition of $G$.
 
-The remaining argument schemes are all very similar, in the sense that they all simply add a single argument to the set of arguments containing the element to be added.
-We therefore have omitted them here.
+In Algorithm~\ref{alg:add-decomp}, on line~\ref{alg:add-decomp:task-id} a unique identifier is created for the task, which is created on line~\ref{alg:add-decomp:task}.
+Finally, on line~\ref{alg:add-decomp:arg}, two arguments are added to the set of arguments $Arg$.
+The first argument $(task,Task)$ is simply an argument for the task that is added.
+The second argument $((mintId(),i,task\_id),Decomp)$ is an argument for the decomposition link $(mintId(), i, task\_id)$ (Def.~\ref{def:link:decomp}, going from the existing goal with identifier $i$ to the new task with identifier $task\_id$.
 
-\subsubsection*{Algorithms for critical questions}
+The remaining argument schemes AS6-AS-12 are all very similar to the previous algorithms and have been omitted here.
+
+The critical questions of type \emph{INTRO} are very similar as well, with one exception: the require an answer.
+For instance, suppose CQ5b: ``Does goal $G$ decompose into other tasks?'' is answered with: ``Yes, namely into task $T$''.
+In this case, we simply obtain an instantiation of argument scheme AS5: ``Goal $G$ decomposes into task $T$'', which can be executed with Algorithm~\ref{alg:add-decomp}.
+This is the same for all the other critical questions of type \emph{INTRO} as well.
+Therefore, we have omitted them here as well.
+
+\subsubsection{DISABLE algorithms}
+
+As discussed before, algorithms of type \emph{DISABLE} consist of adding a new argument attacking an existing argument, which is an argument for an existing GRL element or link.
+The argument that is added is itself not an argument for a GRL element or link.
+
+In all of these algorithms, we assume the critical question is answered affirmatively.
+For instance, for critical question CQ0 ``Is the actor relevant?'', we assume it is answered with ``No'' (see right-most column of Table~\ref{table:argument-schemes}).
+
+\begin{algorithm}[h]
+  \caption{CQ0: Is actor with id $i$ relevant? No}\label{alg:actor-not-relevant}
+  \begin{algorithmic}[1]
+    \Procedure{$CQ_0$}{$i$}
+    \State $A \leftarrow ((mintId(),CQ0),GenArg)$\label{alg:actor-not-relevant:genarg}
+    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:actor-not-relevant:genarg2}
+    \For{$A_i\in \{(actor_i,Actor)\in Arg\}$}\label{alg:actor-not-relevant:for}
+      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:actor-not-relevant:att}
+    \EndFor
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{Rationale and example} Algorithm~\ref{alg:actor-not-relevant} is executed when critical question CQ0 is answered affirmatively (i.e., with ``No'').
+First, on lines~\ref{alg:actor-not-relevant:genarg} and~\ref{alg:actor-not-relevant:genarg2}, an argument is created for the critical question and added to the set of arguments $Arg$.
+Since this argument is not an argument for a GRL element or link, it is formalized as a \emph{generic counterargument} $((mintId(), CQ0), GenArg)$ (Def.~\ref{def:generic-argument}).
+The for loop starting at line~\ref{alg:actor-not-relevant:for} then iterates over ll arguments $(actor_i,Actor)$, where $i$ is the id of the actor that is no longer relevant.
+Recall from Def.~\ref{def:notation} that we denote with $actor_i$ an actor with identifier $i$.
+Thus, this for loop iterates over all arguments for actors with identifier $i$.
+The reason why there could be multiple of such actors is that the actor can be refined by an algorithm of type \emph{REPLACE}.
+We will explain this in more detail in the example below.
+Then, on line~\ref{alg:actor-not-relevant:att}, an attack link is created from the generic argument $A$ that is created to the argument for the actor $A_i$.
+After executing the algorithm, all existing arguments for the actor with identifier $i$ are attacked by a newly created argument $A$.
+
+Consider for example...\todo{Marc}{Marc}{Add example here}
+
+Critical questions C1-CQ3 are all very similar to CQ0 and have therefore been omitted here.
+
+\begin{algorithm}[h]
+  \caption{CQ5a: Does the goal with id $g\_id$ decompose into task with id $t\_id$? No}\label{alg:no-decomp}
+  \begin{algorithmic}[1]
+    \Procedure{$CQ5a$}{$g\_id,t\_id$}
+    \State $A \leftarrow ((mintId(),CQ5a),GenArg)$\label{alg:no-decomp:genarg}
+    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:no-decomp:genarg2}
+    \For{$A_i\in \{((k,m,n),Decomp)\in Arg\mid $\\
+    \qquad\qquad\qquad $m=g\_id, n=t\_id\}$}\label{alg:no-decomp:for}
+      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:no-decomp:att}
+    \EndFor
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{Rationale and example} Algorithm~\ref{alg:no-decomp} is structurally very similar to Algorithm~\ref{alg:actor-not-relevant}, with the only difference that instead of iterating over actors, we now iterate over decomposition links.
+This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links $(k,m,n)$ (Def.~\ref{def:link:decomp}) such that $m$ equals the goal identifier $g\_id$, and $n$ equals the task identifier $t\_id$.
+This means we iterate over all decomposition links from the goal to the task, and we attack all arguments for these links on line~\ref{alg:no-decomp:att}.
+
+Almost all of the remaining critical questions are similar in structure.
+CQ11 (``Is the element relevant/useful'') is slightly different since the attack element is not of a specific type, but is simply any GRL element.
+However, the resulting algorithm is very similar to the previous two and has therefore been omitted.
+The only algorithm of type \emph{DISABLE} that we still discuss is $Att$.
+
+\begin{algorithm}[h]
+  \caption{Att: Generic counter-argument on arguments $A_1,\ldots,A_n$}\label{alg:generic}
+  \begin{algorithmic}[1]
+    \Procedure{$Att$}{$A_1,\ldots,A_n$}
+    \State $A \leftarrow ((mintId(),Att),GenArg)$\label{alg:generic:arg}
+    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:generic:addarg}
+    \For{$A_i\in \{A_1,\ldots,A_n\}$}\label{alg:generic:for}
+      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:generic:att}
+    \EndFor
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{Rationale and example} Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counter-arguments to arguments.
+In all of the previous \emph{DISABLE} algorithm, the attack was on a specific type of argument, for instance an argument for an actor or an argument for a decomposition.
+In this algorithm, however, \emph{any} set of arguments can be attacked by a new argument.
+
+\todo{Marc}{Marc}{Give example}
+
+\subsubsection{REPLACE algorithms}
+
+Recall that the \emph{REPLACE} algorithms both attack all arguments for an existing element, and at the same time create a new argument that contains a different version of the element being attacked.
+
+We present two algorithms for replacing elements: the first one replaces the decomposition type (CQ5c, CQ10c), and the second one replaces the name.
+
+\begin{algorithm}[h]
+  \caption{CQ5c: Is the decomposition type of element $ie_i$ correct? No, it should be $X$ }\label{alg:replace1}
+  \begin{algorithmic}[1]
+    \Procedure{$CQ5c$}{$ie_i, X$}
+    \State $A \leftarrow ((ie_i.id, ie_i.type, ie_i.name, X),ie_i.type)$\label{alg:replace1:arg}
+    \State $IEArgs\leftarrow \{(ie_i,ie_i.type)\in Arg\}$\label{alg:replace1:ieargs}
+    \For{$A_i\in IEArgs$}\label{alg:replace1:for1}
+      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:replace1:att}
+    \EndFor
+    \For{$\{(A_i,A_j)\in Att\mid A_j\in IEArgs\}$}\label{alg:replace1:for2}
+      \State $Att\leftarrow (A_i,A)$
+    \EndFor
+    \For{$\{(A_i,A_j)\in Att\mid A_i\in IEArgs\}$}\label{alg:replace1:for3}
+      \State $Att\leftarrow (A,A_j)$
+    \EndFor
+    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:replace1:addarg}
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{Rationale and example} While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm~\ref{alg:replace1} is more generally applicable to any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}.
+
+Let us go through this algorithm step by step.
+On line~\ref{alg:replace1:arg}, a new argument $((ie_i.id, ie_i.type, ie_i.name, X),ie_i.type)$ is created.
+Recall from Def.~\ref{def:notation} that we may refer to elements of an IE with the ``.'' notation.
+Thus, this argument is an argument for the IE that was input to the algorithm, except that the decomposition type is set to $X$.
+On line~\ref{alg:replace1:ieargs}, the set $IEArgs$ is assigned with all existing arguments for the input IE.
+Then, in the first for loop on line~\ref{alg:replace1:for1}, we add attack links from the argument that has just been created to all existing arguments for the IE.
+The next two for loops on respectively lines~\ref{alg:replace1:for2} and~\ref{alg:replace1:for3} ensure that all attack links that existing from and to the previous versions of the IE are also carried over to the new argument $A$.
+Finally, one line~\ref{alg:replace1:addarg}, the new argument is added to the set of arguments.
+
+Let us explain this with an example. \todo{Marc}{Marc}{Add example}
+
+The other \emph{REPLACE} algorithms are very similar to Algorithm~\ref{alg:replace1}.
+It can be used directly for CQ10c, but for CQ12 we should make a small modification.
+Instead of replacing the decomposition type of the IE, we should replace its name.
+Since this is a very minor modification we have omitted it here.

### 2017-10-23 methodology.tex
finished all algorithms

@@
 
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
+In the previous sections, we have shown how the RationalGRL framework can capture stakeholder discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
+In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a methodology\footnote{The methodology presented here was first presented at the 2017 iStar workshop \cite{iSTar}.} (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
 
-\begin{figure}[ht]
+\subsection{RationalGRL Methodology}
+\label{sect:methodology}
+
+We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a (Rational)GRL model.
+%Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+The rest of the steps are as follows:
+
+\begin{figure*}[ht]
 \centering
-\includegraphics[scale=0.4]{img/methodology}
+\includegraphics{img/methodology.pdf}
 \caption{The RationalGRL Methodology}
 \label{fig:rationalgrl-methodology}
-\end{figure}
+\end{figure*}
+
+\textbf{(1) Instantiate Argument Schemes (AS)} -- First, we start with our list of argument schemes (Table~\ref{table:argument-schemes}).
+Whilst discussing the requirements, we then select schemes from the list and instantiate them to form arguments for GRL model elements.
+In this way we build or modify the GRL model by introducing new GRL elements (\textsf{INTRO}, see example 1 in Section~\ref{sect:overview:examples}).
+Note that it is also possible to start modifying an existing GRL model which was not built using the RationalGRL methodology: each GRL element corresponds to an argument (i.e.
+an instantiated scheme), so it is possible to instantiate argument schemes based on an existing GRL model.
+
+\textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we can start asking the relevant critical questions.
+Because each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.
+
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of three operations are performed on the GRL model: \textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE}.
+Any of these operations impact the arguments and corresponding GRL intentional elements, modifying the initial GRL model into a RationalGRL model (see examples in Section~\ref{sect:overview:examples}).
+After these modifications, we can keep on asking critical questions (e.g.
+about elements that were introduced by previously answering a critical question) until we are satisfied with our model.
+
+\textbf{(4) Modify GRL Models} -- In this step, we modify the regular GRL model based on the RationalGRL model of step (3).
+That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
+This results in a new, modified GRL model, which can be used as the basis for another cycle of the methodology.
+
+We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
+In the next section, we will give an example of how our tool can be used together with the methodology to build a GRL model.
+
+\iffalse%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsubsection{Example 6: Methodology example}
+
+In this section, we present a small example for the RationalGRL methodology.
+This example is based on the traffic simulator example described in Section~\ref{sect:goals:runningexample}.
+We base argument schemes and critical questions of this example on a transcript containing discussions about the development of this traffic simulator system.
+%These transcripts are created as part of two master theses on improving design reasoning~\cite{masterthesis1,masterthesis2}.
+%TODO again?
+%We provide several more examples in Section~\ref{sect:examples}.
+
+As mentioned in Section~\ref{sect:goals:runningexample}, a group of stakeholders are developing a goal model for a traffic simulator example and they are modeling the goal \emph{Simulate Traffic} using the RationalGRL methodology.
+This proceeds in the following way:
+\begin{itemize}
+\item
+First they start %at step \emph{Modify GRL models} (Figure~\ref{fig:rationalgrl-methodology}), 
+with the initial GRL model (Figure~\ref{fig:example-small}) and they invoke the intentional element \emph{Simulate Traffic}. 
+\item Next, they switch to the argumentation part (step \emph{Instantiate arguments schemes}).
+They answer critical question \emph{Does Simulate Traffic AND-decompose into any tasks?} with \emph{Yes: namely tasks ``Dynamic Simulation'' and ``Static Simulation"}.
+\item As a result, two argument schemes are created, namely:
+\begin{itemize}
+\item Actor \emph{Traffic Simulator} has task \emph{Dynamic Simulation}
+\item Actor \emph{Traffic Simulator} has task \emph{Static Simulation}
+\end{itemize}
+\item In the GRL model, this corresponds to the addition of two tasks, and an AND-decomposition from the goal \emph{Simulate Traffic} to these two tasks.
+\item Next, the stakeholders test the validity of their goal model by answering more critical question. They answer two critical questions:
+\begin{itemize}
+\item
+CQ \emph{Is task ``Dynamic Simulation'' relevant} is answered with ``No, it is not relevant since the problem specification explicitly states dynamic simulations are not required''.
+Thus, the corresponding GRL intentional element is disabled.
+\item The decomposition is changed from AND to OR, since it turned out not both tasks can be implemented together.
+\end{itemize}
+\end{itemize}
+
+An excerpt of the RationalGRL model is shown in Figure~\ref{fig:examples:decomposition}.
+%TODO You need to update goal "simulate" in this figure to "Simulate Traffic".
+Also, I don't think we should really call it the rational GRL model.
+it is an excerpts of the model.
+\fi

### 2017-10-25 discussion.tex
removed beliefs, change formalism

@@
 
-In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we effectively extended Jureta et al.'s work and translated argument diagrams to GRL models\footenote{In \cite{Jureta:RE2008} only a mapping from goal models to structured arguments is given, and the step from structured arguments or argument diagrams to goal models is never formally defined.} (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
+In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we effectively extended Jureta et al.'s work and translated argument diagrams to GRL models\footnote{In \cite{Jureta:RE2008} only a mapping from goal models to structured arguments is given, and the step from structured arguments or argument diagrams to goal models is never formally defined.} (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
 Thus, we have essentially two complex diagrams, an argument diagram and a goal diagram, and a mapping between them.

### 2017-10-25 formalframework.tex
removed beliefs, change formalism

@@
 \item $IETypes = \{Softgoal, Goal, Task, Resource\}$,
-\item $Types = IETypes \cup \{Actor, Contr, Dep,$\\ \qquad $Decomp, Belief, GenArg\}$
-\item $Names$ is a finite set of strings.
+\item $LinkTypes = \{PosContr, NegContr, Dep, Decomp\}$\footnote{Note that for decomposition links, we only distinguish between a positive contribution ($PosContr$) and a negative contribution ($NegContr$).
+The GRL metamodel specifies a total of seven contribution values.
+We instead have chosen only two here to simplify some of the notation, but it is straightforward to extend this to seven values.},
+\item $Types = IETypes \cup LinkTypes\cup\{Actor, GenArg, ActIE\}$,
+\item $Names$ is a finite set of strings,
 \item $DecompTypes = \{AND,OR,XOR\}$.
-\item $ContribValues = \{Break, Hurt, Some\ nega-$ $tive, Unknown, Some\  positive, Help, Make\},$
 \end{itemize}
@@
 \label{def:ie}
-An intentional element $ie\in \mathbb{N}\times IETypes\times Names\times DecompType$ is a relation, where $ie = (id, type, name, decomptype)$ means:
+An intentional element $ie\in \mathbb{N}\times IETypes\times Names\times DecompType$ is a tuple, where $ie = (id, type, name, decomptype)$ means:
 \begin{itemize}
@@
 \begin{itemize}
-\item $(2, goal, \text{Generate traffic}, XOR)$: This represents the goal ``Generate traffic'', which is XOR-decomposed.
+\item $(4, Goal, \text{Generate traffic}, XOR)$: This represents the goal ``Generate traffic'', which is XOR-decomposed.
 Note the identifier ``2'' is not shown in the GRL model, but it is stored internally.
-\item $(5, resource, \text{Car objects}, AND)$: Although this element does not decompose into any other elements, it still contains an AND-decomposition.
+\item $(1, Resource, \text{Car objects}, AND)$: Although this element does not decompose into any other elements, it still contains an AND-decomposition.
 Our motivation for this choice is that it simplifies the formalization of decomposition links.
@@
 \begin{itemize}
-\item $goal_2.name = \text{Generate traffic}$
-\item $goal_2.decomptype = XOR$
-\item $resource_5.name = \text{Car objects}$
+\item $goal_4.name = \text{Generate traffic}$
+\item $goal_4.decomptype = XOR$
+\item $resource_1.name = \text{Car objects}$
 \end{itemize}
@@
 
-\begin{definition}[Belief]
-\label{def:belief}
-A belief $bel\in \mathbb{N}\times Names$ is a relation where $bel=(id,name)$ means:
-\begin{itemize}
-\item $id\in\mathbb{N}$ is the identifier of the belief,
-\item $name\in Names$ is a string description of the belief.
-\end{itemize}
-\end{definition}
-
-\paragraph{Rational and example} Beliefs are different from IEs in the sense that they are not being used in decompositions, dependencies, or contributions.
-They merely provide additional explanation for an element.
-The only belief in Figure~\ref{fig:example-small2} is formalized as $(1, \text{Not scientifically correct})$.
-
 \begin{definition}[Actor]
 \label{def:actor}
-An actor $act \in \mathbb{N}\times Names$ is a relation where $act=(id,name)$ means:
+An actor $act \in \mathbb{N}\times {Actor}\times Names$ is a tuple where $act=(id,Actor, name)$ means:
 \begin{itemize}
 \item $id\in\mathbb{N}$ is the identifier of the actor, 
+\item $Actor$ states that this tuple is an actor,\footnote{While this element may look redundant in this tuple, it will become important when translating between RationalGRL models and GRL models.
+Otherwise it would not be possible to identify this tuple as an actor.}
 \item $name\in Names$ is a string description of its name.
@@
 
-We can formalize the actor of Figure~\ref{fig:example-small2} as $act=(0,\text{Traffic Simulator})$, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
+We can formalize the actor of Figure~\ref{fig:example-small2} as $act=(0,Actor,\text{Traffic Simulator})$, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
 
@@
 \label{def:act-ie-relation}
-An \emph{Actor-IE relation} $r_{ActIE}\in\mathbb{N}\times\mathbb{N}$ is a relation $(i,j)$ meaning that an actor with id $i$ has intentional element with id $j$.
+An \emph{Actor-IE relation} $r_{ActIE}\in\mathbb{N}\times\{ActIE\}\times\mathbb{N}\times\mathbb{N}$ is a tuple where $r_{ActIE}=(id,ActIE, i, j)$ means:
+\begin{itemize}
+\item $id\in\mathbb{N}$ is the identifier of the relation, 
+\item $ActIE$ states that this tuple is an Actor-IE relation,
+\item $i\in\mathbb{N}$ is an id of an actor,
+\item $j\in\mathbb{N}$ is an id of an IE.
+\end{itemize}
 
@@
 \begin{itemize}
-\item $(2, goal, \text{Generate traffic}, XOR)$
-\item $(5, resource, \text{Car objects}, AND)$
+\item $(1, Resource, \text{Car objects}, AND)$
+\item $(4, Goal, \text{Generate traffic}, XOR)$
 \item $(0, \text{Traffic Simulator})$
@@
 
-We can then formalize the Actor-IE relationships for these elements as: $(0,2)$ and $(0,5)$.
+We can then formalize the Actor-IE relationships for these elements as:
+\begin{itemize}
+\item $(12, ActIE, 0, 1)$
+\item $(15, ActIE, 0, 4)$
+\end{itemize}
 
@@
 
-\begin{definition}[Belief link]
-\label{def:link:belief}
-A \emph{belief} link: $bellink\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $bellink = (i,j)$ means
+\begin{definition}[GRL Link]
+\label{def:link}
+A \emph{GRL link}: $link\in \mathbb{N}\times LinkTypes\times \mathbb{N}\times\mathbb{N}$ is a tuple such that $link = (id,type,src,dest)$ means
 \begin{itemize}
-\item $i\in \mathbb{N}$ is the unique identifier of the belief element,
-\item $j\in \mathbb{N}$ is the unique identifier of the IE.
+\item $id\in \mathbb{N}$ is the unique identifier of the link,
+\item $type\in LinkTypes$ specifies the type of the link, 
+\item $src\in \mathbb{N}$ is the identifier of the IE from which the link starts,
+\item $dest\in \mathbb{N}$ is the identifier of the IE at which the link ends.
 \end{itemize}
-Intuitively, $bellink = (i,j)$ means that $bel_i$ is a belief for $ie_j$.
 
-A set of belief links is denoted by $Bellink$.
+A set of links is denoted by $Link$.
 \end{definition}
 
-\paragraph{Rationale and example} The belief links can only be used to connect a belief with an intentional element, and they are not used in any of the algorithms.
-Instead, they can be understood as metadata, they provide additional documentation on the rationale for making certain design choices.
-Therefore, belief links do not require unique identifiers.
+We also introduce some simplified notation for the links.
 
-The fact that belief links only connect beliefs with IEs is not made explicit in the definition above.
-We make this assumption, together with various assumption, explicit in the definition of a \emph{valid GRL model} (Def.~\ref{def:valid-grl-model}).
-Figure~\ref{fig:example-small2} contains one belief links.
-It connects the following two elements:
+\begin{definition}[Notation]
+\label{def:notation2}
 \begin{itemize}
-\item $(1, \text{Not scientifically correct})$,
-\item $(4, softgoal, \text{Simple design}, AND)$.
+\item
+We can refer to links of a specific type simply by $type_{id}$. For instance, we can abbreviate the link $(0,PosContr,2,3)$ with $posContr_0$. 
+\item 
+We can refer to a set of positive contribution links, negative contribution links, dependency links, and decomposition link with respectively $PosContr, NegContr, Dep,$, and $Decomp$. 
+\item
+We can refer to a set of contribution links that can be either positive or negative simply with $Contr$.
 \end{itemize}
-The belief link is then formalized as $(1, 4)$.
-
-\begin{definition}[Contribution Link]
-\label{def:link:contrib}
-A \emph{contribution} link: $contrib\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}\times Contribvalues$ is a relation such that $contrib = (i,j,k,value)$ means
-\begin{itemize}
-\item $i\in \mathbb{N}$ is the unique identifier of the link,
-\item $j\in \mathbb{N}$ is the unique identifier of the IE from which the contribution originates,
-\item $k\in \mathbb{N}$ is the unique identifier of the IE to which is contributed.
-\item $value\in ContribValues$ is the strength of the contribution.
-\end{itemize}
-Intuitively, $contrib = (i,j,k,value)$ means that $ie_i$ contributes to $ie_j$ with $value$.
-
-A set of contribution links is denoted by $Contrib$.
 \end{definition}
 
-\paragraph{Rationale and example} Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small2}, there are two contribution links. First, let us formalize the related IEs:
+\paragraph{Rationale and example} Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small2}, there are five links. First, let us formalize the related IEs:
 \begin{itemize}
-\item $(3, softgoal, \text{Realisitc simulation}, AND)$
-\item $(4, softgoal, \text{Simple design}, AND)$
-\item $(6, task, \text{Create new cars}, AND)$
+\item (1, Resource, \text{Car objects}, AND),
+\item (2, Softgoal, \text{Realistic simulation}, AND),
+\item (3, Softgoal, \text{Simple design}, AND),
+\item (4, Goal, \text{Generate traffic}, XOR),
+\item (5, Task, \text{Create new cars}, AND),
+\item (6, Task, \text{Keep same cars}, AND).
 \end{itemize}
 
-We can formalize the contribution links as $(9, 6, 3, \text{Help})$ and $(10, 6, 4, \text{Hurt})$.
-
-\begin{definition}[Decomposition Link]
-\label{def:link:decomp}
-A \emph{decomposition} link: $decomp\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $decomp = (i,j,k)$ means
+We can formalize the links as follows:
 \begin{itemize}
-\item $i\in \mathbb{N}$ is the unique identifier of the link,
-\item $j\in \mathbb{N}$ is the unique identifier of the decomposing IE,
-\item $k\in \mathbb{N}$ is the unique identifier of the IE that is being decomposed.
-\end{itemize}
-Intuitively, $decomp = (i,j,k)$ means that $ie_i$ decomposes $ie_j$.\footnote{Note that the decomposition type is defined on $IE_{k}$, see Definition~\ref{def:ie}.}
-
-A set of decomposition links is denoted by $Decomp$.
-\end{definition}
-
-\paragraph{Rationale and example} Decomposition links are simpler than contribution links, because they only contain information about which elements are being connected.
-Recall that the decomposition type is stored in the element that is being decomposed.
-In order to formalize the decomposition links in Figure~\ref{fig:example-small2}, let us first formalize the related IEs:
-\begin{itemize}
-\item $(5, goal, \text{Generate traffic}, XOR)$
-\item $(6, task, \text{Create new cars}, AND)$
-\item $(7, task, \text{Keep same cars}, AND)$
-\end{itemize}
-The decomposition links can be formalized as $(11, 6, 5)$ and $(12, 7, 5)$.
-
-\begin{definition}[Dependency Link]
-\label{def:link:dependency}
-A \emph{dependency} link: $dep\in \mathbb{N}\times\mathbb{N}\times\mathbb{N}$ is a relation such that $dep = (i,j,k)$ means
-\begin{itemize}
-\item $i\in \mathbb{N}$ is the unique identifier of the link,
-\item $j\in \mathbb{N}$ is the unique identifier of the dependee IE,
-\item $k\in \mathbb{N}$ is the unique identifier of the dependent IE.
-\end{itemize}
-Intuitively, $dep = (i,j,k)$ means that $ie_j$ depends on $ie_j$.
-
-A set of dependency links is denoted by $Dep$.
-\end{definition}
-
-\paragraph{Rationale and example} The representation of dependency links is exactly the same as decomposition links.
-In Figure~\ref{fig:example-small2}, let us formalize the IEs related to the only dependency link:
-\begin{itemize}
-\item $(2, resource, \text{Car objects}, AND)$
-\item $(5, goal, \text{Generate traffic}, XOR)$
+\item $(7, Dep, 4, 1)$
+\item $(8, PosContr, 5, 2)$
+\item $(9, NegContr, 5, 3)$
+\item $(10, Decomp, 5, 4)$
+\item $(11, Decomp, 6, 4)$
 \end{itemize}
 
-The dependency link between these two elements is formalized as $(8, 5, 2)$.
+Let us next provide some examples of the notation of Definition~\ref{def:notation2}:
+\begin{itemize}
+\item $dep_7 = (7, Dep, 4, 1)$
+\item $Decomp = \{(10, Decomp, 5, 4), (11, Decomp, 6, 4)\}$
+\item $PosContr = \{(8, PosContr, 5, 2)\}$
+\item $Contr = \{(8, PosContr, 5, 2), (9, NegContr, 5, 3)\}$
+\end{itemize}
 
@@
 \label{def:grl-model}
-A \emph{GRL model} $GRL=(IE, Bel, Act, R_{ActIE}, Bellink, Contr, Decomp, Dep)$ consists of:
+A \emph{GRL model} $GRL=(IE, Act, R_{ActIE}, Link)$ consists of:
 \begin{itemize}
 \item A set $IE$ of intentional elements (Def.~\ref{def:ie}),
-\item A set $Bel$ of beliefs (Def.~\ref{def:belief}),
 \item A set $Act$ of actors (Def.~\ref{def:actor}),
 \item A set $R_{ActIE}$ of Actor-IE relations (Def.~\ref{def:act-ie-relation}),
-\item A set $Bellink$ of belief links (Def.~\ref{def:link:belief}),
-\item A set $Contr$ of contribution links (Def.~\ref{def:link:contrib}),
-\item A set $Decomp$ of decomposition links (Def.~\ref{def:link:decomp}),
-\item A set $Dep$ of dependency links (Def.~\ref{def:link:dependency}).
+\item A set $Link$ of GRL links (Def.~\ref{def:link}).
 \end{itemize}
@@
 For completeness, we now provide the full specification of Figure~\ref{fig:example-small2}.
-This model is formalized as
- $GRL=(IE, Bel, Act, R_{ActIE}, Bellink, Contr, Decomp, Dep)$ where:
+This model is formalized as $GRL=(IE, Act, R_{ActIE}, Link)$, where:
 \begin{flalign*}
-&IE=&\{&(2, task, \text{Car objects}, AND),&\\
-&   &  &(3, softgoal, \text{Realistic simulation}, AND),&\\
-&   &  &(4, softgoal, \text{Simple design}, AND),&\\
-&   &  &(5, goal, \text{Generate traffic}, XOR),&\\
-&   &  &(6, task, \text{Create new cars}, AND),&\\
-&   &  &(7, task, \text{Keep same cars}, AND)\}&\\
-&Bel=& &\{(1, \text{Not scientifically correct})\}&\\
-&Act=& &\{(0, \text{Traffic Simulator})\}&\\
-&R_{ActIE}=& &\{(0,i)\mid 2\le i\le 7\}&\\
-&Bellink = & &\{(1, 4)\}&\\
-&Contr=& &\{(9, 6, 3, \text{Help}), (10, 6, 4, \text{Hurt})\}&\\
-&Decomp=& &\{(11, 6, 5), (12, 7, 5)\}&\\
-&Dep=& &\{(8, 5, 2)\}&
+&IE=&\{&(1, Resource, \text{Car objects}, AND),&\\
+&   &  &(2, Softgoal, \text{Realistic simulation}, AND),&\\
+&   &  &(3, Softgoal, \text{Simple design}, AND),&\\
+&   &  &(4, Goal, \text{Generate traffic}, XOR),&\\
+&   &  &(5, Task, \text{Create new cars}, AND),&\\
+&   &  &(6, Task, \text{Keep same cars}, AND)\}&\\
+&Act=& &\{(0, Actor, \text{Traffic Simulator})\}&\\
+&R_{ActIE}=& &\{(i+11, ActIE, 0,i)\mid 1\le i\le 6\}&\\
+&Link=&\{&(7, Dep, 4, 1)&\\
+&     & &(8, PosContr, 5, 2)&\\
+&     & &(9, NegContr, 5, 3)&\\
+&     & &(10, Decomp, 5, 4)&\\
+&     & &(11, Decomp, 6, 4)\}&\\
 \end{flalign*}
@@
 \item ids are globally unique across IEs, Beliefs, Links, and Actors, i.e., let $X,Y\in \{IE,Act, Belief, Contr, Decomp, Dep\}$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
-\item All intentional elements of actors exist: $\forall (i,j)\in R_{ActIE}: act_i \in Act \wedge ie_j \in IE$.
-\item An intentional element belongs at most to one actor: $\forall ie_i\in IE: |\{(k,i)\in R_{ActIE}\}| \le 1$.
-\item Contribution links connect intentional elements: $\forall (i,j,k,value)\in Contrib: \{ie_j,ie_k\}\subseteq IE$.
-\item Decomposition links connect intentional elements: $\forall (i,j,k)\in Decomp: \{ie_j,ie_k\}\subseteq IE$.
-\item Dependency links connect intentional elements: $\forall (i,j,k)\in Dep: \{ie_j,ie_k\}\subseteq IE$.
-\item Belief links connect beliefs with intentional elements: $\forall (i,j)\in Bellink: bel_i\in Bel\wedge ie_j\in IE$.
+\item Intentional elements of actors exist: $\forall (k, ActIE, i,j)\in R_{ActIE}: act_i \in Act \wedge ie_j \in IE$.
+\item An intentional element belongs at most to one actor: $\forall ie_i\in IE: |\{(k,ActIE,j,i)\in R_{ActIE}\}| \le 1$.
+\item Links connect intentional elements: $\forall (i,type, j,k)\in Link: \{ie_j,ie_k\}\subseteq IE$.
 \end{enumerate}
@@
 \item All elements in the formalization have different ids, so this constraint is satisfied.
-\item $R_{ActIE}$ contains one element $(0,i)$ for each IE with id $i$, so this constraint is satisfied as well.
-Note that other elements (beliefs and links) are not related to actors.
-This is in line with the jUCMNav implementation.
+\item $R_{ActIE}$ contains one element for each IE with id $i$, so this constraint is satisfied as well. Note that links are not related to actors. This is in line with the GRL metamodel.
 \item Since we have only one actor with id 0, and this is the only actor that appears in $R_{ActIE}$, this constraint is satisfied.
-\item The contribution links connect elements with ids 3, 4, and 6, which are all IEs.
-\item The decomposition links connect elements with ids 5, 6, and 7, which are all IEs.
-\item The dependency link connects id 2 with 5, which are both IEs.
-\item The belief link connects id 1 with id 4, which are respectively a belief and an IE.
+\item All links connect IEs: the contribution links connect elements with ids 2, 3, and 5, which are all IEs; the decomposition links connect elements with ids 4, 5, and 6, which are all IEs; and the dependency link connects id 1 with 4, which are both IEs.
 \end{enumerate}
@@
 \label{def:generic-argument}
-A generic argument $ga \in \mathbb{N}\times Names$ is a relation such that $ga=(id, name)$ means
+A generic argument $ga \in \mathbb{N}\times \{GenArg\}\times Names$ is a tuple such that $ga=(id, GenArg, name)$ means
 \begin{itemize}
 \item $id\in \mathbb{N}$ is the identifier of the generic argument,
+\item $GenArg$ states that the tuple is a generic argument,
 \item $name\in Names$ is a string description of its name.
@@
 \paragraph{Rationale and example} A generic counterargument is simply an argument that can be used to attack any previous argument, or an IE.
-It is different from a belief element, since a generic counterargument is used in the computation to determine whether a RationalGRL element is accepted or rejected.
-Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link:contrib},~\ref{def:link:decomp}, and~\ref{def:link:dependency}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links.
-This is correct, since generic arguments are not part of GRL, so they should also not occur in links.
+Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links.
+This is desired, since generic arguments are not part of GRL, so they should also not be connected using GRL links.
 
-In Fig.~\ref{fig:example-small3}, there are two generic counterargument which can be formalized as $(19, \text{Redundant})$ and $(20, \text{Necessary})$.
+In Fig.~\ref{fig:example-small3}, there are two generic counterargument which can be formalized as $(18,GenArg, \text{Redundant})$ and $(19, GenArg, \text{Necessary})$.
 
@@
 \label{def:argument}
-An argument $A=(elem,type)$ is a pair such that:
+An argument $A$ is on of the following tuples:
 \begin{itemize}
-\item $elem$ is either an intentional element $ie$ (Def.~\ref{def:ie}), an actor $act$ (Def.~\ref{def:actor}), an Actor-IE relation $r_{ActIE}$ (Def.~\ref{def:act-ie-relation}),  a contribution link $contr$, a decomposition link $decomp$ (Def.~\ref{def:link:decomp}), a dependency link $dep$ (Def.~\ref{def:link:dependency}), or a generic argument $ga$ (Def.~\ref{def:generic-argument}).
-\item $type\in Type$ (Def.~\ref{def:set-definitions}) is the type of the argument.
+\item An intentional element (Def.~\ref{def:ie}), 
+\item An actor $act$ (Def.~\ref{def:actor}),
+\item An Actor-IE relation $r_{ActIE}$ (Def.~\ref{def:act-ie-relation}), 
+\item A link $link$ (Def.~\ref{def:link}),
+\item A generic argument $ga$ (Def.~\ref{def:generic-argument}).
 \end{itemize}
@@
 In Fig~\ref{fig:example-small3}, there are three attack links.
-The first two attack links originate from generic counterarguments and involve the following three elements:
+The first two attack links originate from generic counterarguments and involve the following three arguments:
 \begin{itemize}
-\item $A_1 = ((1, Resource, \text{Car objects}, AND), Resource)$
-\item $A_{19} = ((19, \text{Necessary}), GenArg)$
-\item $A_{20} = ((20, \text{Redundant}), GenArg)$
+\item $A_1 = (1, Resource, \text{Car objects}, AND)$
+\item $A_{18} = (18, GenArg, \text{Redundant})$
+\item $A_{19} = (19, GenArg, \text{Necessary})$
 \end{itemize}
 
-These two attack links are formalized as $(A_{20},A_{19})$ and $(A_{19}, A_1)$.
+These two attack links are formalized as $(A_{18},A_{1})$ and $(A_{19}, A_{18})$.
 
@@
 \begin{itemize}
-\item $A_{4} = ((5, goal, \text{Generate traffic}, AND), Goal)$
-\item $A_{18} = ((5, goal, \text{Generate traffic}, XOR), Goal)$
+\item $A_{4} = ((4, Goal, \text{Generate traffic}, AND), Goal)$
+\item $A_{20} = ((4, Goal, \text{Generate traffic}, XOR), Goal)$
 \end{itemize}
 
-This attack link is formalized at $(A_{18},A_4)$.
+This attack link is formalized at $(A_{20},A_4)$.
 
@@
 \begin{flalign*}
-A_0 = &((0, Actor, \text{Traffic simulator}), Actor)&\\
-A_1 = &((1, task, \text{Car objects}, AND), Task),&\\
-A_2 = &((2, softgoal, \text{Real. sim.}, AND), Softgoal),&\\
-A_3 = &((3, softgoal, \text{Simple des.}, AND), Softgoal),&\\
-A_4 = &((4, goal, \text{Generate traffic}, AND), Goal),&\\
-A_5 = &((5, task, \text{Create new cars}, AND), Task),&\\
-A_6 = &((6, task, \text{Keep same cars}, AND)\}, Task),&\\
-A_7 = &((7, 4, 1), Dep),&\\
-A_8 = &((8, 5, 2, \text{Help}), Contr),&\\
-A_9 = &((9, 5, 3, \text{Hurt}), Contr),&\\
-A_{10} = &((10, 5, 4), Decomp),&\\
-A_{11} = &((11, 6, 4), Decomp),&\\
-A_i = &((0,i-11),ActIE), \text{for } 12\le i\le 17&\\
-A_{18} = &((4, goal, \text{Generate traffic}, XOR), Goal),&\\
-A_{19} = &((18, \text{Redundant}), GenArg),&\\
-A_{20} = &((19, \text{Necessary}), GenArg),&
+A_0 = &(0, Actor, \text{Traffic simulator})&\\
+A_1 = &(1, Task, \text{Car objects}, AND),&\\
+A_2 = &(2, Softgoal, \text{Real. sim.}, AND),&\\
+A_3 = &(3, Softgoal, \text{Simple des.}, AND),&\\
+A_4 = &(4, Goal, \text{Generate traffic}, AND),&\\
+A_5 = &(5, Task, \text{Create new cars}, AND),&\\
+A_6 = &(6, Task, \text{Keep same cars}, AND),&\\
+A_7 = &(7, Dep, 4, 1),&\\
+A_8 = &(8, PosContr, 5, 2),&\\
+A_9 = &(9, NegContr, 5, 3),&\\
+A_{10} = &(10, Decomp, 5, 4),&\\
+A_{11} = &(11, Decomp, 6, 4),&\\
+A_i = &(i, ActIE, 0,i+11),ActIE), \text{for } 1\le i\le 6&\\
+A_{18} = &(18, GenArg, \text{Redundant}),&\\
+A_{19} = &(19, GenArg, \text{Necessary}),&\\
+A_{20} = &(4, Goal, \text{Generate traffic}, XOR),&\\
 \end{flalign*}
@@
 Arg = &\{A_0,A_1,\ldots,A_{20}\}\\
-Att=&\{(A_{18},A_4), (A_{19},A_1), (A_{20},A_{19})\}
+Att=&\{(A_{18},A_1), (A_{19},A_{18}), (A_{20},A_{4})\}
 \end{flalign*}
@@
 All the arguments and the attack relations of this RationalGRL model are shown in Figure~\ref{fig:example-small4}.
-Note the arguments for Actor-IE containment (arguments $A_{12}$ to $A_{20}$) have been omitted from this figure for readability.
+Note the arguments for Actor-IE containment (arguments $A_{12}$ to $A_{17}$) have been omitted from this figure for readability.
 It can be read from the figure that two arguments are currently rejected, namely $A_4$ and $A_{19}$.
@@
 \begin{tikzpicture}
-        \node[minimum size=1cm] (a20) [argNodeIN] at (-3,0) {$A_{20}$};
-        \node[minimum size=1cm] (a19) [argNodeIN] at (0,0) {$A_{19}$};
+        \node[minimum size=1cm] (a19) [argNodeIN] at (-3,0) {$A_{19}$};
+        \node[minimum size=1cm] (a18) [argNodeIN] at (0,0) {$A_{18}$};
         \node[minimum size=1cm] (a1) [argNodeIN] at (3,0) {$A_1$};
@@
          \path
-    (a20) edge [attackLink] (a19)
-    (a19) edge [attackLink] (a1);
+    (a19) edge [attackLink] (a18)
+    (a18) edge [attackLink] (a1);
 
@@
 
-Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of Figure~\ref{fig:example-small4} containing only arguments $A_1, A_{19}$, and $A_{20}$.
-In this example, there are two admissible sets: $\{A_{20}\}$ and $\{A_1, A_{20}\}$.
-In the last admissible set, $A_{20}$ defends $A_1$ against its attacker $A_{19}$.
-Sets containing both $A_{19}$ and either $A_1$ or $A_{20}$ are not conflict free, and the sets $\{A_1\}$ and $\{A_{19}\}$ do not defend themselves against $A_{19}$ and $A_{20}$, respectively.
+Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of Figure~\ref{fig:example-small4} containing only arguments $A_1, A_{18}$, and $A_{19}$.
+In this example, there are two admissible sets: $\{A_{19}\}$ and $\{A_1, A_{19}\}$.
+In the last admissible set, $A_{19}$ defends $A_1$ against its attacker $A_{18}$.
+Sets containing both $A_{18}$ and either $A_1$ or $A_{19}$ are not conflict free, and the sets $\{A_1\}$ and $\{A_{18}\}$ do not defend themselves against $A_{18}$ and $A_{19}$, respectively.
 
@@
 
-In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_{20},A_1\}$.
+In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_{19},A_1\}$.
 Returning to our running example of Figure~\ref{fig:example-small3}, this means that the intentional elements \text{Car objects} (task) and the generic counerargument \emph{Necessary} are both \emph{accepted}, while the generic counter argument ``Redundant'' is rejected.
 
-\subsection{Translation algorithms}
-
-Now that we have formalized both a GRL model and a RationalGRL model, we develop algorithm in order to translate between these two models.
-Both of these two translation algorithms are straightforward.
-Their main task is to ensure that in RationalGRL models, the GRL elements and links are wrapped in an argument with the corresponding type.
-For each algorithm, we provide an explanation an an example of a translation.
-
-\paragraph{GRL to RationalGRL Translation} We start with the translation algorithm from GRL to RationalGRL, which is shown in Algorithm~\ref{alg:translation:to-rationalgrl}.
-The translation algorithm translates all of the GRL elements and links into arguments.
-It takes a GRL model as input.
-Recall that a GRL model is defined as $GRL=(IE, Bel, Act, R_{ActIE}, Bellink, Contr, Decomp, Dep)$ (Def.~\ref{def:grl-model}).
-Then, on line~\ref{alg:translation:to-rationalgrl:init}, the set of arguments is initialized with the empty set.
-Each of the five for loops in the algorithm translates a different GRL component into an argument.
-For instance, in the for loop starting at line~\ref{alg:translation:to-rationalgrl:ie}, all IEs are added as pairs consisting of the IE itself, and its corresponding type.
-This is in line with the definition of an argument~\ref{def:argument}).
-Since the other for loops are very similar, we do not explain them here.
-At line~\ref{alg:translation:to-rationalgrl:return}, the algorithm returns a RationalGRL model $(Arg,\emptyset)$ containing all arguments in $Arg$, and an empty set of attack relations.
-
-\begin{algorithm}[h]
-  \caption{GRL to RationalGRL Translation}
-  \label{alg:translation:to-rationalgrl}
-  \begin{algorithmic}[1]
-    \Procedure{$ToRationalGRL$}{$GRL$}
-    \State $Arg \leftarrow \emptyset$\label{alg:translation:to-rationalgrl:init}
-    \For{$ie\in IE$}\label{alg:translation:to-rationalgrl:ie}
-      \State $Arg \leftarrow Arg \cup \{(ie,ie.type)\}$
-    \EndFor
-    \For{$act\in Act$}
-      \State $Arg \leftarrow Arg \cup \{(act, Actor)\}$
-    \EndFor
-    \For{$actIE\in ActIE$}
-      \State $Arg \leftarrow Arg \cup \{(actIE, ActIE)\}$
-    \EndFor
-    \For{$contr\in Contr$}
-      \State $Arg \leftarrow Arg \cup \{(contr, Contr)\}$
-    \EndFor
-    \For{$dep\in Dep$}
-      \State $Arg \leftarrow Arg \cup \{(dep, Dep)\}$
-    \EndFor
-    \For{$decomp\in Decomp$}
-      \State $Arg \leftarrow Arg \cup \{(decomp, Decomp)\}$
-    \EndFor
-    \State \Return $(Arg, \emptyset)$\label{alg:translation:to-rationalgrl:return}
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\paragraph{RationalGRL to GRL translation} The translation from a RationalGRL model to a GRL model is given in Algorithm~\ref{alg:translation:to-grl}.
-It is very similar to the previous algorithm, but then works in the other direction.
-In this case, arguments are unpacked and put in the corresponding IE components.
-The most important difference between the previous algorithm and this one is that in the RationalGRL to GRL translation, argumentation semantics is used to determine the \emph{preferred extension} (Def.~\ref{def:preferred-semantics}).
-This is a set containing all accepted arguments.
-The GRL is then generated from all the accepted arguments.
-This is done by iterating over all arguments in the extension in line~\ref{alg:translation:to-grl:extension-iter}.
-The switch statement on line~\ref{alg:translation:to-grl:switch} then does a case distinction on the type of the arguments, and each case ensures the argument is put in the right GRL component.
-Finally, the algorithm returns a GRL model on line~\ref{alg:translation:to-grl:return}.
-
-\begin{algorithm}[h]
-  \caption{RationalGRL to GRL Translation}
-  \label{alg:translation:to-grl}
-  \begin{algorithmic}[1]
-    \Procedure{$ToGRL$}{$RatGRL$}
-    \State $E \leftarrow ComputeExtension(Arg,Att)$
-    \State $IE\leftarrow\emptyset$, $Act\leftarrow\emptyset$, $R_{ActIE}\leftarrow\emptyset$, $Contr\leftarrow\emptyset$
-    \State $Decomp\leftarrow\emptyset$, $Dep\leftarrow\emptyset$
-    \For{$(Elem, T)\in E$}\label{alg:translation:to-grl:extension-iter}
-      \Switch{$T$}\label{alg:translation:to-grl:switch}
-          \Case{$T\in Type$}
-            \State $IE\leftarrow IE\cup \{Elem\}$
-          \EndCase
-          \Case{$T = Actor$}
-            \State $Act\leftarrow Act \cup\{Elem\}$
-          \EndCase
-          \Case{$T=Contr$}
-            \State $Contr\leftarrow Contr\cup \{Elem\}$
-          \EndCase
-          \Case{$T=Dep$}
-            \State $Dep\leftarrow Dep\cup \{Elem\}$
-          \EndCase
-          \Case{$T=Decomp$}
-            \State $Decomp\leftarrow Decomp\cup\{Elem\}$
-          \EndCase
-      \EndSwitch
-    \EndFor
-    \State \Return $(IE,Act,R_{ActIE}, Contr, Decomp, Dep)$\label{alg:translation:to-grl:return}
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Def.~\ref{def:valid-grl-model}), we have not done so for a RationalGRL model yet.
-This is because it depends on the translation procedure.
-We define a RationalGRL model as valid if and only if the RationalGRL to GRL translation results in a valid GRL model.
-In this way, we do not have to reiterate all conditions on a GRL model, but use the translation algorithm.
+Now that we have defined the semantics for RationalGRL, we can use this to define the notion of a valid RationalGRL model.
 
@@
 \label{def:valid-rationalgrl-model}
-A RationalGRL model $RatGRL = (Arg, Att)$ with preferred extension $E=(A_1,\ldots,A_n)$ is a \emph{valid RationalGRL model} iff $TranslateToGRL(RatGRL)$ (Alg.~\ref{alg:translation:to-grl}) is a valid GRL model (Def.~\ref{def:valid-grl-model}).
+A RationalGRL model $RatGRL = (Arg, Att)$ with preferred extension $E=(A_1,\ldots,A_n)$ is a \emph{valid RationalGRL model} iff the GRL model constructed from $E$ is a valid GRL model (Def.~\ref{def:valid-grl-model}).
 \end{definition}
 
-The following theorem follows directly from the translation algorithms.
-
-\begin{theorem}[Translation correctness]
-\label{th:translation-correctness}
-Given some valid GRL model (Def.~\ref{def:valid-grl-model}, translating it to RationalGRL and then back to GRL results in the same valid GRL mode, i.e.
-$$ToGRL(ToRationalGRL(GRL)) = GRL.$$
-\end{theorem}
+\todo{Marc}{Marc}{Perhaps say something about how to construct it.
+It is very trivial: Just remove the generic arguments and attack links, only keep the other arguments which directly translate to GRL elements and links.}
 
@@
     \Procedure{$AS_0$}{$n$}
-    \State $actor \leftarrow (mintId(), a)$ \label{alg:as0:1}
-    \State $Arg\leftarrow Arg \cup \{(actor,Actor)\}$\label{alg:as0:2}
+    \State $A \leftarrow (mintId(), Actor, a)$ \label{alg:as0:1}
+    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:as0:2}
     \EndProcedure
@@
 On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$.
-On line~\ref{alg:as0:2} the pair $(actor,Actor)$ is added as a new argument, indicating that an argument of type $Actor$ is added.
+On line~\ref{alg:as0:2} the actor is added as a new argument.
 
@@
     \State $res\_id\gets mintId()$\label{alg:add-actor:id}
-    \State $res\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-actor:res}
-    \State $Arg\gets Arg\cup \{(res, Resource),(i,res\_id)\}$\label{alg:add-actor:arg}
+    \State $A_1\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-actor:arg1}
+    \State $A_2\leftarrow (mintId(),ActIE, i, res\_id)$\label{alg:add-actor:arg2}
+    \State $Arg\gets Arg\cup \{A_1,A_2\}$\label{alg:add-actor:add-arg}
     \EndProcedure
@@
 
-The algorithm itself runs as follows: on line~\ref{alg:add-actor:id}, a unique id is assigned to variabel $res\_id$ (``resource identifier'').
-On line~\ref{alg:add-actor:res}, an argument for a resource with name $R$ and identifier $res\_id$ is assigned to $A$.
-Finally, on line~\ref{alg:add-actor:arg}, two arguments are added to the set of arguments $Arg$: one for the resource, and one for the containment relations between the actor and the resource (Def.~\ref{def:act-ie-relation}.
+The algorithm itself runs as follows: on line~\ref{alg:add-actor:id}, a unique id is assigned to variable $res\_id$ (``resource identifier'').
+On line~\ref{alg:add-actor:arg1}, an argument for a resource with name $R$ and identifier $res\_id$ is assigned to $A_1$.
+On line~\ref{alg:add-actor:arg2}, a second argument $A_2$ is created, which is an argument for an Actor-IE relation (Def.~\ref{def:act-ie-relation}) between the input actor $i$ and the newly created element $res\_id$.
+Finally, on line~\ref{alg:add-actor:add-arg}, the arguments are added to the set of arguments $Arg$.
 
-Since arguments schemes AS2-AS4 are very similar to AS1, we have omitted it here.
+Since arguments schemes AS2-AS4 are very similar to AS1, we have omitted algorithms for them here.
 
 \begin{algorithm}[h]
-  \caption{AS5: Goal with id $i$ decomposes into task $t$}\label{alg:add-decomp}
+  \caption{AS5: Goal with id $i$ decomposes into task $T$}\label{alg:add-decomp}
   \begin{algorithmic}[1]
-    \Procedure{$AS_5$}{$i, t$}
+    \Procedure{$AS_5$}{$i, T$}
     \State $task\_id\gets mintId()$\label{alg:add-decomp:task-id}
-    \State $task\leftarrow (task\_id, Task, t, AND)$\label{alg:add-decomp:task}
-    \State $Arg\gets Arg\cup \{(task, Task),((mintId(),i,task\_id), Decomp)\}$\label{alg:add-decomp:arg}
+    \State $A_1\leftarrow (task\_id, Task, t, AND)$\label{alg:add-decomp:arg1}
+    \State $A_2\leftarrow (mintId(), Decomp, i, task\_id)$\label{alg:add-decomp:arg2}
+    \State $Arg\gets Arg\cup \{A_1,A_2\}$\label{alg:add-decomp:add-arg}
     \EndProcedure
@@
 \paragraph{Rationale and example} Similar to the previous algorithm, we have slightly reworded critical question AS5.
-We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $t$ is a decomposition of $G$.
+We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $T$ is a decomposition of $G$.
 
-In Algorithm~\ref{alg:add-decomp}, on line~\ref{alg:add-decomp:task-id} a unique identifier is created for the task, which is created on line~\ref{alg:add-decomp:task}.
-Finally, on line~\ref{alg:add-decomp:arg}, two arguments are added to the set of arguments $Arg$.
-The first argument $(task,Task)$ is simply an argument for the task that is added.
-The second argument $((mintId(),i,task\_id),Decomp)$ is an argument for the decomposition link $(mintId(), i, task\_id)$ (Def.~\ref{def:link:decomp}, going from the existing goal with identifier $i$ to the new task with identifier $task\_id$.
+In Algorithm~\ref{alg:add-decomp}, on line~\ref{alg:add-decomp:task-id} a unique identifier is created for the task, which is created on line~\ref{alg:add-decomp:arg1}.
+On line~\ref{alg:add-decomp:arg2} an argument is created for the decomposition link $(mintId(), Decomp, i, task\_id)$ (Def.~\ref{def:link}, going from the existing goal with identifier $i$ to the new task with identifier $task\_id$.
+On line~\ref{alg:add-decomp:add-arg} the two arguments are added to the set of arguments $Arg$.
 
-The remaining argument schemes AS6-AS-12 are all very similar to the previous algorithms and have been omitted here.
+The remaining argument schemes AS6-AS12 are all very similar to the previous algorithms and have been omitted here.
 
@@
     \Procedure{$CQ_0$}{$i$}
-    \State $A \leftarrow ((mintId(),CQ0),GenArg)$\label{alg:actor-not-relevant:genarg}
+    \State $A \leftarrow (mintId(),GenArg,CQ0)$\label{alg:actor-not-relevant:genarg}
     \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:actor-not-relevant:genarg2}
-    \For{$A_i\in \{(actor_i,Actor)\in Arg\}$}\label{alg:actor-not-relevant:for}
-      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:actor-not-relevant:att}
+    \For{$A_j\in \{A=(i,Actor,n)\mid A\in Arg\}$}\label{alg:actor-not-relevant:for}
+      \State $Att \leftarrow Att \cup \{(A,A_j)\}$\label{alg:actor-not-relevant:att}
     \EndFor
@@
 First, on lines~\ref{alg:actor-not-relevant:genarg} and~\ref{alg:actor-not-relevant:genarg2}, an argument is created for the critical question and added to the set of arguments $Arg$.
-Since this argument is not an argument for a GRL element or link, it is formalized as a \emph{generic counterargument} $((mintId(), CQ0), GenArg)$ (Def.~\ref{def:generic-argument}).
-The for loop starting at line~\ref{alg:actor-not-relevant:for} then iterates over ll arguments $(actor_i,Actor)$, where $i$ is the id of the actor that is no longer relevant.
-Recall from Def.~\ref{def:notation} that we denote with $actor_i$ an actor with identifier $i$.
-Thus, this for loop iterates over all arguments for actors with identifier $i$.
+Since this argument is not an argument for a GRL element or link, it is formalized as a \emph{generic counterargument} $(mintId(), GenArg, CQ0)$ (Def.~\ref{def:generic-argument}).
+The for loop starting at line~\ref{alg:actor-not-relevant:for} then iterates over all arguments for actors, where $i$ is the id of the actor that is no longer relevant.
 The reason why there could be multiple of such actors is that the actor can be refined by an algorithm of type \emph{REPLACE}.
 We will explain this in more detail in the example below.
-Then, on line~\ref{alg:actor-not-relevant:att}, an attack link is created from the generic argument $A$ that is created to the argument for the actor $A_i$.
+Then, on line~\ref{alg:actor-not-relevant:att}, an attack link is created from the generic argument $A$ that is created to the argument for the actor $A_j$.
 After executing the algorithm, all existing arguments for the actor with identifier $i$ are attacked by a newly created argument $A$.
@@
     \Procedure{$CQ5a$}{$g\_id,t\_id$}
-    \State $A \leftarrow ((mintId(),CQ5a),GenArg)$\label{alg:no-decomp:genarg}
+    \State $A \leftarrow (mintId(),GenArg,CQ5a)$\label{alg:no-decomp:genarg}
     \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:no-decomp:genarg2}
-    \For{$A_i\in \{((k,m,n),Decomp)\in Arg\mid $\\
-    \qquad\qquad\qquad $m=g\_id, n=t\_id\}$}\label{alg:no-decomp:for}
-      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:no-decomp:att}
+    \For{$A_j\in \{(k,Decomp,g\_id,t\_id)\in Arg\}$}\label{alg:no-decomp:for}
+      \State $Att \leftarrow Att \cup \{(A,A_j)\}$\label{alg:no-decomp:att}
     \EndFor
@@
 \paragraph{Rationale and example} Algorithm~\ref{alg:no-decomp} is structurally very similar to Algorithm~\ref{alg:actor-not-relevant}, with the only difference that instead of iterating over actors, we now iterate over decomposition links.
-This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links $(k,m,n)$ (Def.~\ref{def:link:decomp}) such that $m$ equals the goal identifier $g\_id$, and $n$ equals the task identifier $t\_id$.
-This means we iterate over all decomposition links from the goal to the task, and we attack all arguments for these links on line~\ref{alg:no-decomp:att}.
+This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links with source id $g\_id$ and target id $t\_id$ (Def.~\ref{def:link}).
+This means we iterate over all decomposition links from the goal to the task, and we attack all arguments for these links on line~\ref{alg:no-decomp:att} with the newly created argument.
 
-Almost all of the remaining critical questions are similar in structure.
+Almost all of the remaining critical questions of type DISABLE are similar in structure.
 CQ11 (``Is the element relevant/useful'') is slightly different since the attack element is not of a specific type, but is simply any GRL element.
@@
 \begin{algorithm}[h]
-  \caption{Att: Generic counter-argument on arguments $A_1,\ldots,A_n$}\label{alg:generic}
+  \caption{Att: Generic counter-argument on arguments $A_1,\ldots,A_n$ with name $N$}\label{alg:generic}
   \begin{algorithmic}[1]
-    \Procedure{$Att$}{$A_1,\ldots,A_n$}
-    \State $A \leftarrow ((mintId(),Att),GenArg)$\label{alg:generic:arg}
+    \Procedure{$Att$}{$A_1,\ldots,A_n$, $N$}
+    \State $A \leftarrow (mintId(),GenArg, N)$\label{alg:generic:arg}
     \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:generic:addarg}
-    \For{$A_i\in \{A_1,\ldots,A_n\}$}\label{alg:generic:for}
-      \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:generic:att}
+    \For{$A_j\in \{A_1,\ldots,A_n\}$}\label{alg:generic:for}
+      \State $Att \leftarrow Att \cup \{(A,A_j)\}$\label{alg:generic:att}
     \EndFor
@@
 
-Recall that the \emph{REPLACE} algorithms both attack all arguments for an existing element, and at the same time create a new argument that contains a different version of the element being attacked.
-
-We present two algorithms for replacing elements: the first one replaces the decomposition type (CQ5c, CQ10c), and the second one replaces the name.
+Recall that the \emph{REPLACE} algorithms both attacks all arguments for an existing element, and at the same time create a new argument that contains a different version of the element being attacked.
 
@@
     \Procedure{$CQ5c$}{$ie_i, X$}
-    \State $A \leftarrow ((ie_i.id, ie_i.type, ie_i.name, X),ie_i.type)$\label{alg:replace1:arg}
-    \State $IEArgs\leftarrow \{(ie_i,ie_i.type)\in Arg\}$\label{alg:replace1:ieargs}
+    \State $A \leftarrow ie_i$\label{alg:replace1:arg}
+    \State $A.decomptype\leftarrow X$\label{alg:replace1:decompchange}
+    \State $IEArgs\leftarrow IE_i\subseteq  Arg$\label{alg:replace1:ieargs}
     \For{$A_i\in IEArgs$}\label{alg:replace1:for1}
@@
 
-\paragraph{Rationale and example} While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm~\ref{alg:replace1} is more generally applicable to any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}.
+\paragraph{Rationale and example} While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm~\ref{alg:replace1} is more generally applicable to any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}).
 
 Let us go through this algorithm step by step.
-On line~\ref{alg:replace1:arg}, a new argument $((ie_i.id, ie_i.type, ie_i.name, X),ie_i.type)$ is created.
+On line~\ref{alg:replace1:arg}, a new argument $A$ is created which is identical to original IE.
 Recall from Def.~\ref{def:notation} that we may refer to elements of an IE with the ``.'' notation.
-Thus, this argument is an argument for the IE that was input to the algorithm, except that the decomposition type is set to $X$.
-On line~\ref{alg:replace1:ieargs}, the set $IEArgs$ is assigned with all existing arguments for the input IE.
+This is what we do on line~\ref{alg:replace1:decompchange}, where the decomposition type of the argument is changed to $X$.
+On line~\ref{alg:replace1:ieargs}, the set $IEArgs$ is assigned with all existing arguments for the input IE (recall the notation of $IE_i$ as the set of all IEs with id $i$, see Def.~\ref{def:notation}).
 Then, in the first for loop on line~\ref{alg:replace1:for1}, we add attack links from the argument that has just been created to all existing arguments for the IE.

### 2017-10-28 background.tex
notes formal framework and small changes
finished all algorithms
Examples updated (beliefs), SpringerJournal cls used

@@
 A GRL model is a connected graph of intentional elements that optionally are part of the actors.
-All the elements and relationships used in GRL are shown in Figure~\ref{fig:grl_legend}.
+The GRL elements and relationships used in this paper are shown in Figure~\ref{fig:grl_legend}.
+Note that we only consider a subset of the full GRL specification in this paper.
+In particular, the full GRL specification considers different levels of positive and negative contributions and satisfaction levels of goals and tasks, which are not directly relevant for this work.
+Furthermore, the full GRL specification includes beliefs to capture design rationale, whereas in this paper, the reasoning and rationales behind the models are captured using arguments -- for a brief discussion on beliefs, see Section~\ref{sect:goalmodeling:relatedwork}.
 
-\begin{figure}[ht]
+\begin{figure*}[ht]
+\centering
+\includegraphics{img/grl_legend.pdf}
+\caption{Basic elements and relationships of GRL}
+\label{fig:grl_legend}
+\end{figure*}
+
+\begin{figure}[b]
 \centering
@@
 
-\begin{figure*}[ht]
-\centering
-\includegraphics[scale=0.6]{img/grl_legend}
-\caption{Basic elements and relationships of GRL}
-\label{fig:grl_legend}
-\end{figure*}
-
 Figure~\ref{fig:example-small} illustrates a simplified GRL diagram from the traffic simulator design exercise.
-An actor (\includegraphics[scale=1]{img/actor}) represents a stakeholder of a system or the system itself (\emph{Traffic Simulator}, Figure~\ref{fig:example-small}).
+An actor represents a stakeholder of a system or the system itself (\emph{Traffic Simulator}, Figure~\ref{fig:example-small}).
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals (\includegraphics[scale=1]{img/softgoal}) differentiate themselves from goals (\includegraphics[scale=1]{img/goal}) in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
+Softgoals differentiate themselves from goals in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
 Softgoals (e.g.
 \emph{Realistic simulation}) are often related to non-functional requirements, whereas goals (such as \emph{Generate Cars}) are related to functional requirements.
-Tasks (\includegraphics[scale=1]{img/task}) represent solutions to (or operationalizations of) goals and softgoals.
+Tasks represent solutions to (or operationalizations of) goals and softgoals.
 In Figure~\ref{fig:example-small}, we have the two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve goal \emph{Generate cars}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off screen.
-In order to be achieved or completed, softgoals, goals, and tasks may require resources (\includegraphics[scale=1]{img/resource}) to be available (e.g., \emph{Car Objects}).
-Finally, design rationale can be captured using beliefs (\includegraphics[scale=0.3]{img/belief}).
-For example, the idea that the simulation should be \emph{Not scientifically correct} is captured as a belief in the model.
+In order to be achieved or completed, softgoals, goals, and tasks may require resources to be available (e.g., \emph{Car Objects}).
 
 Different links connect the elements in a GRL model.
-AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links (\includegraphics[scale=1]{img/decomposition}) allow an element to be decomposed into sub-elements.
+AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links allow an element to be decomposed into sub-elements.
 In Figure~\ref{fig:example-small}, the goal \emph{Generate cars} is XOR-decomposed to the tasks \emph{Create new cars} and \emph{Keep same cars}, as they are alternative ways of achieving the goal \emph{Generate cars}.
-Contribution links (\includegraphics[scale=1]{img/contribution}) indicate desired impacts of one element on another element.
-A contribution link has a qualitative contribution type or a quantitative contribution.
-Task \emph{Create new cars} has a \emph{help} qualitative contribution to the task \emph{Realistic simulation}, and a \emph{hurt} qualitative contribution to the task \emph{Simple design}.
-Dependency links (\includegraphics[scale=1]{img/dependency}) model relationships between actors or resources.
+Contribution links indicate impacts of one element on another element, which can be positive or negative.
+Task \emph{Create new cars} has a positive contribution to the task \emph{Realistic simulation}, and a negative contribution to the task \emph{Simple design}.
+Dependency links model relationships between actors or resources.
 Here, the goal \emph{Generate cars} depends on the resource \emph{Car objects}.
@@
 Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-GRL is also linked to Use Case Maps via URNLinks (\includegraphics[scale=1]{img/urnlink}), which provide traceability between concepts and instances of the goal model and behavioral design models.
+GRL is also linked to Use Case Maps, which provide traceability between concepts and instances of the goal model and behavioral design models.
 Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
 
-GRL has six evaluation algorithms which are semi-automated and allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements quantitatively, qualitatively or in a hybrid way.
+GRL has six evaluation algorithms which allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements quantitatively, qualitatively or in a hybrid way.
 jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
@@
 \begin{itemize}
-	\item Why is belief \emph{Not scientifically correct} not linked to any of the goals or tasks? 
+	\item Why is softgoal \emph{Easy to use} not linked to any of the goals or tasks? 
 	\item What does \emph{Keep same cars} mean?
@@
 \item[] Performing action \emph{Keep same cars} realizes goal \emph{Generate cars}, 
-\item[] \textit{Therefore} 
+\item[] \underline{Therefore} 
 \item[] Action \emph{Keep same cars} should be performed
@@
 These counterarguments also follow a scheme; for example, a negative answer to CQ1 follows the scheme ``Action $A$ will not realize goal $G$'', which can be instantiated (e.g.
-``\emph{Keep same cars} does not realize \emph{Generate cars''}) to forma counterargument to the original argument.
+``\emph{Keep same cars} does not realize \emph{Generate cars''}) to form a counterargument to the original argument.
 
@@
 For example, while the action \emph{Create new cars} realizes the goal \emph{Generate cars}, it has a negative side effect, namely hurting \emph{Simple design}: having the simulation constantly create new cars is fairly complex design choice.
+
+\begin{figure}[t]
+\centering
+\includegraphics[]{img/fig_AF1.pdf}
+\caption{PRAS arguments and attacks in the traffic simulation example.}
+\label{fig:pras:example}
+\end{figure}
 
@@
 
-\begin{figure}[h]
-\centering
-\includegraphics[]{img/fig_AF1.pdf}
-\caption{PRAS arguments and attacks in the traffic simulation example.}
-\label{fig:pras:example}
-\end{figure}
-
-\iffalse
-\begin{figure}[ht!]
-\centering
-\begin{tikzpicture}
-        \node[minimum size=1cm] (att3) [argNodeIN] at (-0.5,-3.4) {\argtext{A1}{Action \emph{Keep Same Cars} realizes goal \emph{Generate Cars}\\ \emph{\underline{Therefore}} Action \emph{Keep Same Cars} should be performed}};
-        \node[minimum size=1cm] (att1) [argNodeOUT] at (-0.5,0) {\argtext{A2}{Action \emph{Create New Cars} realizes goal \emph{Generate Cars}\\ \emph{\underline{Therefore}} Action  \emph{Create New Cars} should be performed}};
-        \node[minimum size=1cm] (att2) [argNodeIN] at (3.5,0) {\argtext{A3}{\emph{Create New Cars} has a negative effect on \emph{Static Simulation}}};
-         \path
-    (att2) edge [attackLink] (att1)
-    (att1) edge [attackLink,<->] (att3);
-\end{tikzpicture}
-\caption{Arguments and attacks in the traffic simulation example.}
-\label{fig:pras:example}
-\end{figure}
-\fi
-
 Looking at PRAS and its critical questions, one can see how it could be used to argue about goals and actions or, more specifically, about goal models.

### 2017-10-28 discussion.tex
Examples updated (beliefs), SpringerJournal cls used

@@
 Jureta et al.
-then present the argumentation part of their framework, where reasons (justifications) for conclusions are given as formal structured arguments\footnote{Informally, a structured argument is similar to the PRAS argument in Section~\ref{sect:background:pras}, i.e., $a, b \vec{therefore} c$.}.
+then present the argumentation part of their framework, where reasons (justifications) for conclusions are given as formal structured arguments\footnote{Informally, a structured argument is similar to the PRAS argument in Section~\ref{sect:background:pras}, i.e., $a, b \xrightarrow{therefore} c$.}.
 Arguments and alternatives are then captured as structured arguments or argument diagrams with reasons for or against goals and tasks.

### 2017-10-28 formalframework.tex
Examples updated (beliefs), SpringerJournal cls used

--- 2017-10-23 formalframework.tex
@@
 
-\paragraph{Rationale and example} Throughout this section we use the example in Figure~\ref{fig:example-small}, which is shown in Figure~\ref{fig:example-small2} as well.
+\paragraph{Rationale and example} Throughout this section we use a slightly adapted version of the example from Figure~\ref{fig:example-small}, which is shown in Figure~\ref{fig:example-small2}.
 Our formalization of GRL is very much in line with the way in which GRL models are represented in the open-source Eclipse-based tool jUCMNav.\footnote{See \url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}} This tool is actively developed and has a rich number of features for the analysis of both GRL and URN models.
@@
 \includegraphics[width=\columnwidth]{img/Example1-new.pdf}
-\caption{Example GRL model (reprint of Fig.~\ref{fig:example-small})}
+\caption{Example GRL model}
 \label{fig:example-small2}

### 2017-10-28 introduction.tex
Examples updated (beliefs), SpringerJournal cls used

@@
 
-Requirements Engineering (RE) is an approach to assess the role of a future information system within a human or automated environment.
+Requirements Engineering (RE) is an approach to assess the role of a future information system within its environment.
 An important goal in RE is to produce a consistent and comprehensive set of requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
@@
 
-The first difficulty is that the goal modeling phase, particularly in large projects, is dynamic and that goal models continuously change and evolve.
-Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}, and stakeholders may change their opinion about a modeling decision in between two goal modeling sessions, which in turn may require revisions of the goal model.
+The first difficulty is that the goal modeling phase, particularly in large projects, is dynamic: goal models continuously change and evolve.
+Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}, and stakeholders may change their opinion about a modeling decision in between two modeling sessions, which may require revisions of the goal model.
 If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially have led to different goal models are lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
@@
 
-Following on from our previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop a framework called \emph{RationalGRL}, which combines the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory and discourse modeling called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
+Following on from our previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop a framework called \emph{RationalGRL}, which combines the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
 Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.

### 2017-10-28 methodology.tex
Examples updated (beliefs), SpringerJournal cls used

@@
 
-\textbf{(1) Instantiate Argument Schemes (AS)} -- First, we start with our list of argument schemes (Table~\ref{table:argument-schemes}).
-Whilst discussing the requirements, we then select schemes from the list and instantiate them to form arguments for GRL model elements.
-In this way we build or modify the GRL model by introducing new GRL elements (\textsf{INTRO}, see example 1 in Section~\ref{sect:overview:examples}).
+\textbf{(1) Instantiate Argument Schemes (AS)} -- We start with the list of argument schemes (Table~\ref{table:argument-schemes}).
+Whilst discussing the requirements, we select schemes from the list and instantiate them to form arguments for GRL model elements.
+In this way we build or modify the GRL model by introducing new elements (\textsf{INTRO}).
 Note that it is also possible to start modifying an existing GRL model which was not built using the RationalGRL methodology: each GRL element corresponds to an argument (i.e.
@@
 
-\textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we can start asking the relevant critical questions.
+\textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we start asking the critical questions.
 Because each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.
@@
 \fi
+
+\subsection{The RationalGRL Tool}
+\label{sect:tool}
+\todo{Marc}{Marc}{Currently this section comes from my thesis where I present the tool as future work. Here we should explain it}
+
+GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}.
+This tool is an extension to Eclipse.
+Although it is a rich tool with many features, we also believe it is not very easy to set it up.
+This seriously harms the exposure of the language, as well as the ability for practitioners to use it.
+We have started to implement a simple version of GRL as an online tool in Javascript.
+This makes it usable from the browser, without requiring the installation of any tool.
+The tool can be used from the following address:
+
+\begin{quote}
+\url{http://marcvanzee.nl/RationalGRL/editor}
+\end{quote}
+
+A screenshot of the tool is shown in Figure~\ref{fig:goalmodeling:tool}.
+As shown, there are two tabs in the tool, one for ``Argument'' and one for ``GRL''.
+The argument part has not been implemented yet, and the GRL part only partly, but the idea behind the tool should be clear.
+Users are able to work on argumentation and on goal modeling in parallel, where the argumentation consists of forming arguments and counterarguments by instantiating argument schemes and answering critical questions.
+
+\begin{figure*}[h!]
+\centering
+\includegraphics[scale=0.5]{img/tool_screen}
+\caption{Screenshot of the prototype tool}
+\label{fig:goalmodeling:tool}
+\end{figure*}
+
+An important aspect of the tool is that users can switch freely between these two ways of modeling the problem.
+One can model the entire problem in GRL, or one can do everything using argumentation.
+However, we believe the most powerful way to do so is to switch back and forth.
+For instance, one can create a simple goal model in GRL, and then turn to the argumentation part, which the users can look at the various critical questions for the elements, which may trigger discussions.
+These discussions results in new arguments for and against the elements in the goal model.
+Once this process is completed, one may switch to the goal model again, and so on.
+We believe that in this way, there is a close and natural coupling between modeling the goals of an organization as well as rationalizing them with arguments.

### 2017-10-28 overview.tex
finished all algorithms
Examples updated (beliefs), SpringerJournal cls used

@@
 The RationalGRL framework includes two parts: Argumentation and GRL goal modeling.
-The GRL part of RationalGRL allows for the creation of GRL goal models by analyzing the non-functional requirements in the requirements specification document and by refining the high-level goals into operationalized tasks.
+The GRL part of RationalGRL allows for the creation of goal models by analyzing the non-functional requirements and refining the high-level goals into operationalized tasks.
 For the argumentation part, arguments and counterarguments can be put forward about various parts of this goal model.
-These two parts, GRL and argumentation, are developed iteratively and each side can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
+These two parts, GRL and argumentation, can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
 For example, answering a critical question \emph{Is the task \emph{A} possible?} can result in removing or adding a task in the GRL model.
 Similarly,  if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
-
-Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework.
-At the bottom there are two activities: \emph{practical reasoning \& argumentation} and \emph{goal model construction}.
-As already explained, these two activities influence each other: adding elements to a goal model gives rise to new critical questions about these elements, and answering critical questions with arguments may add or delete elements from the GRL model.
-The activities give rise to two different models: a \emph{RationalGRL model} (left-hand side) and a \emph{GRL model} (right-hand side).
-In the RationalGRL model, we have the goal model elements (goals, tasks etc) and the arguments for and against these elements; the GRL model contains only the goal model elements.
-Thus, the class of GRL models is essentially a subset of the class of RationalGRL models.
-Given the framework, it then becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
 
@@
 \end{figure}
-
-In the rest of this section, we discuss the individual parts of the GRL framework.
-In Section~\ref{sect:overview:as}, we continue our discussion of the argument schemes and critical questions for practical reasoning and argumentation, fitting these schemes and questions into our framework.
-In Section~\ref{sect:overview:lang}, we then discuss the language for RationalGRL models.
-In Section~\ref{sect:overview:examples}, we then provide extensive examples from our case study, illustrating the interplay between practical reasoning and argumentation on the one hand and RationalGRL models on the other hand.
-
-\subsection{The RationalGRL Argument Schemes and Critical Questions for Practical Reasoning}
-\label{sect:overview:as}
 
@@
 AS2 & Actor $a$ can perform task $T$ & CQ2a &Is the task possible? & \textsf{DISABLE} (no)\\
-&& CQ2b & Does the task contribute negatively to some softgoal? & \textsf{DISABLE} (yes)\\
+&& CQ2b & Does the task have negative side-effects? & \textsf{DISABLE} (yes)\\
 \hline
@@
 
+Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework.
+At the bottom there are two activities: \emph{practical reasoning \& argumentation} and \emph{goal model construction}.
+As already explained, these two activities influence each other: adding elements to a goal model gives rise to new critical questions about these elements, and answering critical questions with arguments may add or delete elements from the GRL model.
+The activities give rise to two different models: a \emph{RationalGRL model} (left-hand side) and a \emph{GRL model} (right-hand side).
+In the RationalGRL model, we have the goal model elements (goals, tasks etc) and the arguments for and against these elements; the GRL model contains only the goal model elements.
+Thus, the class of GRL models is essentially a subset of the class of RationalGRL models.
+Given the framework, it then becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
+
+In the rest of this section, we discuss the individual parts of the GRL framework.
+In Section~\ref{sect:overview:as}, we continue our discussion of the argument schemes and critical questions for practical reasoning and argumentation, fitting these schemes and questions into our framework.
+In Section~\ref{sect:overview:lang}, we then discuss the language for RationalGRL models.
+In Section~\ref{sect:overview:examples}, we then provide extensive examples from our case study, illustrating the interplay between practical reasoning and argumentation on the one hand and RationalGRL models on the other hand.
+
+\subsection{The RationalGRL Argument Schemes and Critical Questions for Practical Reasoning}
+\label{sect:overview:as}
+
 A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual types of reasoning stakeholders or designers perform in the early requirements engineering phase (requirement 1 of our framework).
@@
 The last scheme (Att) is a scheme for a generic counterargument against any type of argument that has been put forward.
-Arguments based on these schemes can be used to discuss a GRL goal model.
+Arguments based on these schemes can be used to discuss a goal model.
 Making an argument based on one of the schemes effectively adds the corresponding GRL element to the model.
@@
 
-As we saw in Sections~\ref{sect:background:pras} and \ref{sect:gmas}, an important part of arguing about goal models is asking the right critical questions.
+An important part of arguing about goal models is asking the right critical questions.
 The critical questions presented in Table~\ref{table:argument-schemes} are therefore related to their respective argument schemes.
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
 \centering
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
 \centering
@@
 
-The discussion about the task then continues.
-Again, there is the question (CQ12) whether the new task \emph{Set car influx per road} is clear enough.
-This time the answer is positive, and participant P1 also gives a reason for this positive answer: setting car influx per road was in the original requirements (cf.
-requirement 4 in Appendix~\ref{sect:designprompt}).
-This reason is then introduced as a belief in the RationalGRL diagram, with a correlation link to the task it is about.
-Thus, beliefs can be used as they are intended, namely as rationales for certain choices.
-
 \paragraph{Example 5: Defending the addition of an actor (\textsf{DISABLE)}}
@@
 
-\begin{figure}[b!]
+\begin{figure}[t]
 \centering

### 2017-10-28 transcripts.tex
Examples updated (beliefs), SpringerJournal cls used

@@
 	& \textbf{[32 critical question CQ12 for 31]} Is ``Set car influx'' specific enough?
-\newline \textbf{[33 answer to 32]} No, ``Set car influx'' becomes ``Set car influx per road'' \textsf{[REPLACE]}\\
+\\
 \hline
-0:24:12 (P2)	& So setting it per road, I think is something we want	& \textbf{[34 critical question CQ12 for 33]} Is ``Set car influx per road'' clear? \\
-\hline
-0:24:15 (P1)	& Yeah, that was also one of the constraints I believe &  \textbf{[35 answer to 34]} Yes, ``Set car influx per road'' was one of the original constraints.\\
+0:24:12 (P2)	& So setting it per road, I think is something we want	& \textbf{[33 answer to 32]} No, ``Set car influx'' becomes ``Set car influx per road'' \textsf{[REPLACE]} \\
 \hline

### 2017-10-29 formalframework.tex
added tool screenshots

--- 2017-10-25 formalframework.tex
@@
 
-Now that we have defined the semantics for RationalGRL, we can use this to define the notion of a valid RationalGRL model.
+\subsection{Translation algorithms}
+
+Now that we have formalized both a GRL model and a RationalGRL model, we develop algorithm in order to translate between these two models.
+Both of these two translation algorithms are straightforward, which is result of the fact that the two models are formalized in a very similar way.
+
+\paragraph{GRL to RationalGRL Translation} We start with the translation algorithm from GRL to RationalGRL, which is shown in Algorithm~\ref{alg:translation:to-rationalgrl}.
+The translation algorithm takes a GRL model $GR=(IE, Act, R_{ActIE}, Link)$ (Def.~\ref{def:grl-model}) as input.
+On line~\ref{alg:translation:to-rationalgrl:args}, it collects all the elements of the GRL model into a set $Arg$, which represents the set of arguments of the RationalGRL model.
+On line~\ref{alg:translation:to-rationalgrl:att}, the set of attack links is initialized with an empty set, and the new RationalGRL model $(Arg, Att)$ is returned on line~\ref{alg:translation:to-rationalgrl:return}.
+
+\begin{algorithm}[h]
+  \caption{GRL to RationalGRL Translation}
+  \label{alg:translation:to-rationalgrl}
+  \begin{algorithmic}[1]
+    \Procedure{$ToRationalGRL$}{$GRL$}
+    \State $Arg \leftarrow IE\cup Act \cup R_{ActIE}\cup Link$\label{alg:translation:to-rationalgrl:args}
+    \State $Att \leftarrow \emptyset$\label{alg:translation:to-rationalgrl:att}
+    \State \Return $(Arg, Att)$\label{alg:translation:to-rationalgrl:return}
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{RationalGRL to GRL translation} The translation from a RationalGRL model to a GRL model is given in Algorithm~\ref{alg:translation:to-grl}.
+It is very similar to the previous algorithm, but then works in the other direction.
+In this case, arguments are unpacked and put in the corresponding GRL components.
+On line~\ref{alg:translation:to-grl:extension}, argumentation semantics is used to determine the \emph{preferred extension} (Def.~\ref{def:preferred-semantics}).
+This is a set containing all accepted arguments.
+The GRL model is then generated from all the accepted arguments.
+This is done by iterating over all arguments in the extension in line~\ref{alg:translation:to-grl:for}.
+The switch statement on line~\ref{alg:translation:to-grl:switch} then does a case distinction on the type of the arguments, and each case ensures the argument is put in the right GRL component.
+Finally, the algorithm returns a GRL model on line~\ref{alg:translation:to-grl:return}.
+
+\begin{algorithm}[h]
+  \caption{RationalGRL to GRL Translation}
+  \label{alg:translation:to-grl}
+  \begin{algorithmic}[1]
+    \Procedure{$ToGRL$}{$RatGRL$}
+    \State $E \leftarrow ComputeExtension(Arg,Att)$\label{alg:translation:to-grl:extension}
+    \State $IE\leftarrow\emptyset$, $Act\leftarrow\emptyset$, $R_{ActIE}\leftarrow\emptyset$, $Link\leftarrow \emptyset$
+    \For{$A\in E$}\label{alg:translation:to-grl:for}
+      \Switch{$A.type$}\label{alg:translation:to-grl:switch}
+          \Case{$IETypes$}
+            \State $IE\leftarrow IE\cup \{A\}$
+          \EndCase
+          \Case{$Actor$}
+            \State $Actor\leftarrow Actor\cup \{A\}$
+          \EndCase
+          \Case{$ActIE$}
+            \State $ActIE\leftarrow ActIE\cup \{A\}$
+          \EndCase
+          \Case{$LinkTypes$}
+            \State $Link\leftarrow Link \cup\{A\}$
+          \EndCase
+      \EndSwitch
+    \EndFor
+    \State \Return $(IE,Act,R_{ActIE}, Link)$\label{alg:translation:to-grl:return}
+    \EndProcedure
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Def.~\ref{def:valid-grl-model}), we have no  done so for a RationalGRL model yet.
+This is because it depends on the translation procedure.
+We define a RationalGRL model as valid if and only if the RationalGRL to GRL translation results in a valid GRL model.
+In this way, we do not have to reiterate all conditions on a GRL model, but use the translation algorithm.
 
@@
 \label{def:valid-rationalgrl-model}
-A RationalGRL model $RatGRL = (Arg, Att)$ with preferred extension $E=(A_1,\ldots,A_n)$ is a \emph{valid RationalGRL model} iff the GRL model constructed from $E$ is a valid GRL model (Def.~\ref{def:valid-grl-model}).
+A RationalGRL model $RatGRL = (Arg, Att)$ (Def.~\ref{def:rationalgrl-model})
+is a \emph{valid RationalGRL model} iff $ToGRL(RatGRL)$ (Algorithm~\ref{alg:translation:to-grl}) is a valid GRL model (Def.~\ref{def:valid-grl-model}).
 \end{definition}
-
-\todo{Marc}{Marc}{Perhaps say something about how to construct it.
-It is very trivial: Just remove the generic arguments and attack links, only keep the other arguments which directly translate to GRL elements and links.}
 
@@
 These algorithms are applied to RationalGRL models and produce new arguments and attack relations.
-We can then use argumentation semantics (Def.~\ref{def:semantics}) to compute sets of accepted arguments.
+We can then use argumentation semantics (Def.~\ref{def:preferred-semantics}) to compute sets of accepted arguments.
 The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.
 
-As discussion in Section~\ref{sect:overview}, ll the argument schemes and critical questions of Table~\ref{table:argument-schemes} fall into one of the following four categories:
+As discussed in Section~\ref{sect:overview}, the argument schemes and critical questions of Table~\ref{table:argument-schemes} fall into one of the following four categories:
 \begin{itemize}
@@
 
-\paragraph{Rationale and example} Algorithm takes one argument, namely the name of the actor $a$.
+\paragraph{Rationale and example} Algorithm~\ref{alg:add-actor} takes one argument, namely the name of the actor $a$.
 On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$.
@@
 
-Consider for instance...\todo{Marc}{Marc}{Add example}
+As an example, we can formalize the addition of actor \emph{Traffic Simulator} in Figure~\ref{fig:example-small2} as executing the algorithm $AS_0($Traffic simulator), which results in a RationalGRL model $RatGRL=(Arg,Att)$ with argument $(i, Actor, \text{Traffic simulator})\in Arg$.
 
 \begin{algorithm}[h]
-  \caption{AS1: Actor with id $i$ has resource $R$}\label{alg:add-actor}
+  \caption{AS1: Actor with id $i$ has resource $R$}\label{alg:add-resource}
   \begin{algorithmic}[1]
     \Procedure{$AS_1$}{$i, R$}
-    \State $res\_id\gets mintId()$\label{alg:add-actor:id}
-    \State $A_1\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-actor:arg1}
-    \State $A_2\leftarrow (mintId(),ActIE, i, res\_id)$\label{alg:add-actor:arg2}
-    \State $Arg\gets Arg\cup \{A_1,A_2\}$\label{alg:add-actor:add-arg}
+    \State $res\_id\gets mintId()$\label{alg:add-resource:id}
+    \State $A_1\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-resource:arg1}
+    \State $A_2\leftarrow (mintId(),ActIE, i, res\_id)$\label{alg:add-resource:arg2}
+    \State $Arg\gets Arg\cup \{A_1,A_2\}$\label{alg:add-resource:add-arg}
     \EndProcedure
@@
 
-The algorithm itself runs as follows: on line~\ref{alg:add-actor:id}, a unique id is assigned to variable $res\_id$ (``resource identifier'').
-On line~\ref{alg:add-actor:arg1}, an argument for a resource with name $R$ and identifier $res\_id$ is assigned to $A_1$.
-On line~\ref{alg:add-actor:arg2}, a second argument $A_2$ is created, which is an argument for an Actor-IE relation (Def.~\ref{def:act-ie-relation}) between the input actor $i$ and the newly created element $res\_id$.
-Finally, on line~\ref{alg:add-actor:add-arg}, the arguments are added to the set of arguments $Arg$.
+The algorithm itself runs as follows: on line~\ref{alg:add-resource:id}, a unique id is assigned to variable $res\_id$ (``resource identifier'').
+On line~\ref{alg:add-resource:arg1}, an argument for a resource with name $R$ and identifier $res\_id$ is assigned to $A_1$.
+On line~\ref{alg:add-resource:arg2}, a second argument $A_2$ is created, which is an argument for an Actor-IE relation (Def.~\ref{def:act-ie-relation}) between the input actor $i$ and the newly created element $res\_id$.
+Finally, on line~\ref{alg:add-resource:add-arg}, the arguments are added to the set of arguments $Arg$.
+
+As an example, consider the addition of resource ``Car objects'' in Figure~\ref{fig:example-small2}, and suppose the actor is formalized as the argument $(i, Actor, \text{Traffic simulator})$.
+Adding the resource can then be formalized as executing $AS1(i, \text{Car objects})$, resulting in the addition of two arguments $(j, Resource, \text{Car object}, AND)$ and $(k, ActIE, i, j)$ which are added to the set of arguments $Arg$.
 
@@
     \State $task\_id\gets mintId()$\label{alg:add-decomp:task-id}
-    \State $A_1\leftarrow (task\_id, Task, t, AND)$\label{alg:add-decomp:arg1}
+    \State $A_1\leftarrow (task\_id, Task, T, AND)$\label{alg:add-decomp:arg1}
     \State $A_2\leftarrow (mintId(), Decomp, i, task\_id)$\label{alg:add-decomp:arg2}
@@
 On line~\ref{alg:add-decomp:add-arg} the two arguments are added to the set of arguments $Arg$.
+
+As an example, suppose we are adding the fact that goal ``Generate traffic'' is decomposing into task ``Keep same cars'' (Figure~\ref{fig:example-small2}), and suppose the following argument for the goal exists already: $(i, Goal,\text{Generate traffic}, AND)$.
+Adding the decomposition can be formalized as executing $AS_5(i, \text{Keep same cars})$, and results in two new arguments: $(task\_id, Task, \text{Keep same cars}, AND)$ and $(k, Decomp, i, task\_id)$, which are both added to the set of arguments $Arg$.
 
@@
 
-Consider for example...\todo{Marc}{Marc}{Add example here}
+Consider for example the RationalGRL model in Figures~\ref{fig:examples:relevant-actor}, which consists of an actor and a generic counterargument.
+Let us reconstruct this model using the application of argument schemes and critical questions.
+Initially, the RationalGRL model $(Arg,Att)$ is empty, i.e.
+$Arg=Att=\emptyset$.
+Next, the algorithm $AS_0(\text{Development team})$ is executed, which results in $A_1=\{(0,Actor,\text{Development team}\})$.
+After this, critical question CQ0: ``Is the actor Development team relevant?'' is answered with ``No, because the professor develops the software''.
+We can formalize this as executing algorithm $CQ_0(0)$ (Algorithm~\ref{alg:actor-not-relevant}), which results in adding a generic counter argument $A_2=(1, GenArg, \text{CQ0: Professor develops software})$ and an attack link $(A_2,A_1)$.
+Note that the RationalGRL model now consists of two arguments, and that the preferred extension is $\{A_2\}$, which means that the resulting GRL model is empty, because generic arguments are not translated to GRL.
 
@@
 
-\todo{Marc}{Marc}{Give example}
+Let us reconsider Figure~\ref{fig:examples:relevant-actor}, which we previously formalized as $RatGRL=(Arg,Att)$, where $Arg=\{A_1,A_2\}$, $A_1=\{(0,Actor,\text{Development team}\}), A_2=(1, GenArg, \text{CQ0: Professor develops software})$, and $Att=(A_2,A_1)$.
+Suppose we execute algorithm $Att(\{A_2\}, \text{Professor does not develop software})$.
+This results in $A_3=(2, GenArg, \text{CQ0: Professor does not develop software})$ and a new attack link $(A_3,A_2)$ (Figure~\ref{fig:examples:relevant-actor2}).
+Since the argument for the goal is now part of the preferred extension, it does show up in the resulting GRL model after the translation of Algorithm~\ref{alg:translation:to-grl}.
 
@@
   \begin{algorithmic}[1]
-    \Procedure{$CQ5c$}{$ie_i, X$}
+    \Procedure{$CQ_{Decomp}$}{$ie_i, X$}
     \State $A \leftarrow ie_i$\label{alg:replace1:arg}
@@
     \State $IEArgs\leftarrow IE_i\subseteq  Arg$\label{alg:replace1:ieargs}
-    \For{$A_i\in IEArgs$}\label{alg:replace1:for1}
+    \For{$\{(A_i,A_j)\in Att\mid A_j\in IEArgs\}$}\label{alg:replace1:for1}
+      \State $Att\leftarrow (A_i,A)$
+    \EndFor
+    \For{$\{(A_i,A_j)\in Att\mid A_i\in IEArgs\}$}\label{alg:replace1:for2}
+      \State $Att\leftarrow (A,A_j)$
+    \For{$A_i\in IEArgs$}\label{alg:replace1:for3}
       \State $Att \leftarrow Att \cup \{(A,A_i)\}$\label{alg:replace1:att}
     \EndFor
-    \For{$\{(A_i,A_j)\in Att\mid A_j\in IEArgs\}$}\label{alg:replace1:for2}
-      \State $Att\leftarrow (A_i,A)$
-    \EndFor
-    \For{$\{(A_i,A_j)\in Att\mid A_i\in IEArgs\}$}\label{alg:replace1:for3}
-      \State $Att\leftarrow (A,A_j)$
     \EndFor
@@
 
-\paragraph{Rationale and example} While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm~\ref{alg:replace1} is more generally applicable to any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}).
+\paragraph{Rationale and example} While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm $CQ_{Decomp}$~\ref{alg:replace1} is more generally applicable to the decomposition type of any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}).
 
@@
 On line~\ref{alg:replace1:ieargs}, the set $IEArgs$ is assigned with all existing arguments for the input IE (recall the notation of $IE_i$ as the set of all IEs with id $i$, see Def.~\ref{def:notation}).
-Then, in the first for loop on line~\ref{alg:replace1:for1}, we add attack links from the argument that has just been created to all existing arguments for the IE.
-The next two for loops on respectively lines~\ref{alg:replace1:for2} and~\ref{alg:replace1:for3} ensure that all attack links that existing from and to the previous versions of the IE are also carried over to the new argument $A$.
+The first two for loops on respectively lines~\ref{alg:replace1:for1} and~\ref{alg:replace1:for2} ensure that all attack links that existing from and to the previous versions of the IE are also carried over to the new argument $A$.
+Then, in the third for loop on line~\ref{alg:replace1:for3}, we add attack links from the argument that has just been created to all existing arguments for the IE.
+This ensures that all previous version of the IE are not part of the preferred extension, and as a result do not show up in the resulting GRL model.
 Finally, one line~\ref{alg:replace1:addarg}, the new argument is added to the set of arguments.
 
-Let us explain this with an example. \todo{Marc}{Marc}{Add example}
+Let us explain this algorithm with the example in Figure~\ref{fig:examples:decomposition}.
+The initial RationalGRL model (before applying CQ10b), can be formalized as follows: $RatGRL=(Arg,Att)$, with $Arg=\{A_1,A_2,A_3,A_4,A_5\}$ and $Att=\emptyset$, where:
+\begin{itemize}
+\item $A_1=(0,Goal,\text{Simulate traffic},AND)$
+\item $A_2=(1,Task,\text{Dynamic simulation},AND)$
+\item $A_3=(2,Task,\text{Static simulation},AND)$
+\item $A_4=(3,Decomp,0,1)$
+\item $A_5=(4,Decomp,0,2)$
+\end{itemize}
+
+Suppose algorithm $CQ_{Decomp}(0, OR)$ is executing for this RationalGRL model.
+On lines~\ref{alg:replace1:arg} and \ref{alg:replace1:decompchange} a new argument $A_6=(0,Goal,\text{Simulate traffic},OR)$ is created.
+On line~\ref{alg:replace1:ieargs}, $IEArgs$ contains all arguments for element 0, which is $\{A_1\}$ (note this does not yet include argument $A_6$, since it is not yet added to the RationalGRL model).
+The first two for loops on lines~\ref{alg:replace1:for1} and~\ref{alg:replace1:for2} do not do anything in the case of our example, since $A_1$ is itself not attack or does not attack any other argument.
+The for loop on line~\ref{alg:replace1:for3} ensures all previous versions of the element with id 0 are now attacked, which means in our case the $(A_6,A_1)$ is added to $Att$.
+Finally, on line~\ref{alg:replace1:addarg} the new argument $A_6$ is added to $Arg$.
 

### 2017-10-29 tool-screenshots.tex
added tool screenshots

\section{Tool screenshots}
\label{sect:tool-screenshots}

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{img/tool-overview}
\caption{Overview of the RationalGRL tool}
\label{fig:tool:overview}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{img/tool-cqdetails}
\caption{Critical question details pane for ``Is the resource relevant/useful?''}
\label{fig:tool:cqdetails}
\end{figure}

\newpage

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{img/tool-cqeffect}
\caption{Effect of answering CQ11: ``Is the resource reelevant/useful?'' with ``No''}
\label{fig:tool:cqeffect}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[scale=0.5]{img/tool-argument}
\caption{Argument details pane for generic argument ``Necessary''}
\label{fig:tool:argument}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[scale=0.5]{img/tool-linkdetails}
\caption{Link details pane for decomposition link}
\label{fig:tool:cqdetails}
\end{figure}

\newpage

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{img/tool-figfrompaper}
\caption{RationalGRL model of Figure~\ref{fig:example-small2} in the tool}
\label{fig:tool:figfrompaper}
\end{figure}

\begin{figure}[ht]
\centering
\includegraphics[scale=0.6]{img/tool-multipleattack}
\caption{Adding argument attack multiple arguments to Figure~\ref{fig:tool:figfrompaper}}
\label{fig:tool:multipleattack}
\end{figure}

### 2017-10-29 background.tex
Small amendmends thoughout the paper.

@@
 
-We use a traffic simulator design case to explain the concepts and framework in this paper.
-Our examples and case study are based on a recent series of experiments by Schriek et al.
-\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning assignment in software engineering.
-In this assignment (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes: The client of the project is Professor E, who teaches civil engineering courses at an American university.
+Our examples and case study are based on the data produced by a recent series of experiments by Schriek et al.
+\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning assignment in software engineering about a traffic simulator.
+In this assignment (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes.
+The client of the project is Professor E, who teaches civil engineering courses at an American university.
 In order for the professor to teach students the various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
-Schriek and colleagues asked designers (groups of students) to discuss the requirements of this traffic simulator.
-These discussions were recorded and transcribed, and we used these transcripts for an extensive case study (Section \ref{sect:} on the basis of which we developed our RationalGRL framework (Section \ref{sect:gmas}).
-Furthermore, we also use the traffic simulator case for a simple running example in this section (Figures \ref{fig:example-small}, \ref{fig:pras:example}).
+Schriek et al.
+asked designers (groups of students) to discuss the requirements of this traffic simulator.
+These discussions were recorded and transcribed, and we used these transcripts for an extensive case study on the basis of which we developed our RationalGRL framework.
+Furthermore, we also use the traffic simulator case for the running examples throughout this paper.
 
@@
 
-GRL is based on $i*$~\cite{Yu:1997:TMR:827255.827807} and the NFR Framework~\cite{chung2012non}, but it is not as restrictive as $i*$.
-Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors, i.e., stakeholders, and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
+GRL is based on $i*$~\cite{yu1997towards} and the NFR Framework~\cite{chung2012non}, but it is less restrictive.
+Intentional elements and links can be more freely combined, the notion of agents is replaced with the more general notion of actors and a task does not necessarily have to be an activity performed by an actor, but may also describe properties of a solution.
 GRL has a well-defined syntax and semantics.
@@
 These are the type of the questions that we cannot answer by just looking at GRL models.
-The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the resulting model, such as various clarification steps for the naming, or alternatives that have been considered for the relationships.
+The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the model, such as clarification steps for the naming, or alternatives that have been considered for the relationships.
 The idea behind the original GRL specification is that beliefs can be used to capture such design rationales that make later justification and review of a model easier.
@@
 %\item[] Which will contribute positively to the softgoal $S$
-\item[] \textit{Therefore} 
+\item[] \underline{Therefore} 
 \item[] Action $A$ should be performed

### 2017-10-29 formalframework.tex
Small amendmends thoughout the paper.

@@
 
-\subsubsection{Intentional Element, beliefs and actors}
+\subsubsection{Intentional Element, Beliefs and Actors}
 
-We start with some general definition that we use in subsequent definitions.
+Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
+We start by defining a few general elements of our language that we use in subsequent definitions.
 
@@
 \label{def:set-definitions}
-Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
-
 We define the following sets:
@@
 \item $IETypes = \{Softgoal, Goal, Task, Resource\}$,
-\item $LinkTypes = \{PosContr, NegContr, Dep, Decomp\}$\footnote{Note that for decomposition links, we only distinguish between a positive contribution ($PosContr$) and a negative contribution ($NegContr$).
-The GRL metamodel specifies a total of seven contribution values.
-We instead have chosen only two here to simplify some of the notation, but it is straightforward to extend this to seven values.},
+\item $LinkTypes = \{PosContr, NegContr, Dep, Decomp\}$\footnote{Recall that for decomposition links we only distinguish between positive and negative contribution.
+Extending the formalization model to include more values is straightforward.},
 \item $Types = IETypes \cup LinkTypes\cup\{Actor, GenArg, ActIE\}$,
@@
 \end{definition}
-
-Next we define an intentional element.
+Next, we define an intentional element.
 
@@
 \label{def:ie}
-An intentional element $ie\in \mathbb{N}\times IETypes\times Names\times DecompType$ is a tuple, where $ie = (id, type, name, decomptype)$ means:
+An intentional element is a tuple $ie = (id, type, name, decomptype)$, where:
 \begin{itemize}
@@
 \item $name \in Names$ is a string description of the element,
-\item $decomptype\in DecompType$ refers to the type of decomposition.\footnote{Note that the decomposition type is relevant only if the element is in fact decomposed into other elements, but since the jUCMNav implementation defines the decomposition type on the element, we do the same here.}
+\item $decomptype\in DecompType$ refers to the type of decomposition.\footnote{Note that, like in the metamodel, the decomposition type is defined on the element.}
 \end{itemize}
-A set of intentional elements is denoted by $IE$.
+The set of intentional elements is denoted by $IE$.
 \end{definition}
 
-The definition above is sufficient to capture all intentional elements used in GRL.
-However, we present some syntactic sugar in the next definition by abbreviating the definition above in various ways.
-This does not add anything new to the previous definition, but it simplifies some of the notation.
-
-\begin{definition}[Notation]
-\label{def:notation}
-We adopt three conventions simplifying our notation:
+The definition above is sufficient to capture all intentional elements used in GRL. %In the next definition, we present some simplifying notational conventions.
+%\begin{definition}[Notation]
+%\label{def:notation}
+%We adopt three conventions simplifying our notation:
+%\begin{itemize}
+%\item
+%We refer to the element of a tuple using the dot (".") notation.
+That is, we may for instance refer to the id, type, name and decomposition type of an $IE$ with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
+%\item 
+%We refer to a set of elements with the same id $i$  using the $i$ subscript on the set.
+For instance a set of IEs with id $i$ is denoted by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
+%\item
+%We can also refer to intentional elements of a specific type simply by $type_{id}$.
+For instance, we can abbreviate the element in the previous item with $goal_0$ and write $goal_0.name = \text{Make profit}$, and $goal_0.decomptype = AND$.
+%\end{itemize}
+%\end{definition}
+Throughout this section we use the example in Figure~\ref{fig:example-small}. The IEs in this example can be formalized using Def.~\ref{def:ie} as follows:
 \begin{itemize}
-\item
-We refer to the element of a tuple using the dot (".") notation.
-That is, we may for instance refer to the id, type, name and decomposition type of an $IE$ with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
-\item 
-We refer to a set of elements with the same id $i$  using the $i$ subscript on the set.
-For instance a set of IEs with id $i$ is denoted by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
-\item
-We can also refer to intention elements of a specific type simply by $type_{id}$.
-For instance, we can abbreviate the element in the previous item with $goal_0$ and write $goal_0.name = \text{Make profit}$, and $goal_0.decomptype = AND$.
+\item (1, Resource, \text{Car objects}, AND),
+\item (2, Softgoal, \text{Realistic simulation}, AND),
+\item (3, Softgoal, \text{Simple design}, AND),
+\item (4, Goal, \text{Generate traffic}, XOR),
+\item (5, Task, \text{Create new cars}, AND),
+\item (6, Task, \text{Keep same cars}, AND).
 \end{itemize}
-\end{definition}
-
-\paragraph{Rationale and example} Throughout this section we use the example in Figure~\ref{fig:example-small}, which is shown in Figure~\ref{fig:example-small2} as well.
-Our formalization of GRL is very much in line with the way in which GRL models are represented in the open-source Eclipse-based tool jUCMNav.\footnote{See \url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}} This tool is actively developed and has a rich number of features for the analysis of both GRL and URN models.
-By keeping our formalization in line with this tool, we simplify the translation step from models in the RationalGRL tool to models in the jUCMNav tool.
-
-Some of the IEs in Figure~\ref{fig:example-small2} can be formalized using Def.~\ref{def:ie} as follows:
-\begin{itemize}
-\item $(4, Goal, \text{Generate traffic}, XOR)$: This represents the goal ``Generate traffic'', which is XOR-decomposed.
-Note the identifier ``2'' is not shown in the GRL model, but it is stored internally.
-\item $(1, Resource, \text{Car objects}, AND)$: Although this element does not decompose into any other elements, it still contains an AND-decomposition.
-Our motivation for this choice is that it simplifies the formalization of decomposition links.
-It is in line with with GRL metamodel and the jUCMNav implementation.
-\end{itemize}
-
-Using the short-hand notation defined in Def.~\ref{def:notation}, we can make the following statements:
-\begin{itemize}
-\item $goal_4.name = \text{Generate traffic}$
-\item $goal_4.decomptype = XOR$
-\item $resource_1.name = \text{Car objects}$
-\end{itemize}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[width=\columnwidth]{img/Example1-new.pdf}
-\caption{Example GRL model (reprint of Fig.~\ref{fig:example-small})}
-\label{fig:example-small2}
-\end{figure}
+%Using the short-hand notation defined in Def.~\ref{def:notation}, we can make the following statements:
+%\begin{itemize}
+%\item $goal_4.name = \text{Generate traffic}$
+%\item $goal_4.decomptype = XOR$
+%\item $resource_1.name = \text{Car objects}$
+%\end{itemize}
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=\columnwidth]{img/Example1-new.pdf}
+%\caption{Example GRL model (reprint of Fig.~\ref{fig:example-small})}
+%\label{fig:example-small2}
+%\end{figure} 
+We now define actors.
 
@@
 \label{def:actor}
-An actor $act \in \mathbb{N}\times {Actor}\times Names$ is a tuple where $act=(id,Actor, name)$ means:
+An actor is a tuple $act=(id,type, name)$, where:
 \begin{itemize}
 \item $id\in\mathbb{N}$ is the identifier of the actor, 
-\item $Actor$ states that this tuple is an actor,\footnote{While this element may look redundant in this tuple, it will become important when translating between RationalGRL models and GRL models.
-Otherwise it would not be possible to identify this tuple as an actor.}
+\item $type = Actor$ states that this tuple is an actor.
 \item $name\in Names$ is a string description of its name.
 \end{itemize}
-
-Similar to intentional elements, we may refer to $act = (id,name)$ with $act_{id}$ and write $act_{id}.name$ to refer to its name.
-
 A set of actors is denoted by $Act$.
 \end{definition}
+An actor is simply a pair consisting of its identifier and its name.
+Similar to intentional elements, we may refer to $act = (id,name)$ with $act_{id}$ and write $act_{id}.name$ to refer to its name.
+We can formalize the actor of Figure~\ref{fig:example-small} as $act=(0,Actor,\text{Traffic Simulator})$.%, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
 
-\paragraph{Rationale and example} An actor is simply a pair consisting of its identifier and its name. The relation between actors and their intentional element is formalized in the next definition.
-
-We can formalize the actor of Figure~\ref{fig:example-small2} as $act=(0,Actor,\text{Traffic Simulator})$, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
+The relation between actors and their intentional element is formalized as follows.
 
@@
 \label{def:act-ie-relation}
-An \emph{Actor-IE relation} $r_{ActIE}\in\mathbb{N}\times\{ActIE\}\times\mathbb{N}\times\mathbb{N}$ is a tuple where $r_{ActIE}=(id,ActIE, i, j)$ means:
+An \emph{Actor-IE relation} is a tuple $(i, j)$, where:
 \begin{itemize}
-\item $id\in\mathbb{N}$ is the identifier of the relation, 
-\item $ActIE$ states that this tuple is an Actor-IE relation,
+%\item $id\in\mathbb{N}$ is the identifier of the relation, 
+%\item $ActIE$ states that this tuple is an Actor-IE relation,
 \item $i\in\mathbb{N}$ is an id of an actor,
@@
 
-\paragraph{Rationale and example} The previous definition specifies relations between actors and intentional elements.
-Each actor can own one or multiple intentional elements, but each intentional element can only be owned by at most one actor (we will make these assumptions explicit in Def.~\ref{def:valid-grl-model}).
-
-In Figure~\ref{fig:example-small2}, recall we formalized two IEs and the actor as follows:
-\begin{itemize}
-\item $(1, Resource, \text{Car objects}, AND)$
-\item $(4, Goal, \text{Generate traffic}, XOR)$
-\item $(0, \text{Traffic Simulator})$
-\end{itemize}
-
-We can then formalize the Actor-IE relationships for these elements as:
-\begin{itemize}
-\item $(12, ActIE, 0, 1)$
-\item $(15, ActIE, 0, 4)$
-\end{itemize}
+Note that Actor-IE relations do not have an $id$, as they signify relations between elements of a GRL model (i.e., intentional elements and actors) and are themselves not elements of a GRL model.
+The relation between the \emph{Traffic Simulator} actor ($id = 0$) and the \emph{Car objects} resource, for example, can be formalized as $(0, 1)$.
 
@@
 \label{def:link}
-A \emph{GRL link}: $link\in \mathbb{N}\times LinkTypes\times \mathbb{N}\times\mathbb{N}$ is a tuple such that $link = (id,type,src,dest)$ means
+A \emph{GRL link} is a tuple $link = (id,type,src,dest)$, where:
 \begin{itemize}
@@
 
-We also introduce some simplified notation for the links.
+%We also introduce some simplified notation for the links.
 
-\begin{definition}[Notation]
-\label{def:notation2}
-\begin{itemize}
-\item
-We can refer to links of a specific type simply by $type_{id}$. For instance, we can abbreviate the link $(0,PosContr,2,3)$ with $posContr_0$. 
-\item 
-We can refer to a set of positive contribution links, negative contribution links, dependency links, and decomposition link with respectively $PosContr, NegContr, Dep,$, and $Decomp$. 
-\item
-We can refer to a set of contribution links that can be either positive or negative simply with $Contr$.
-\end{itemize}
-\end{definition}
+%\begin{definition}[Notation]
+%\label{def:notation2}
+%\begin{itemize}
+%\item
+%We can refer to links of a specific type simply by $type_{id}$. For instance, we can abbreviate the link $(0,PosContr,2,3)$ with $posContr_0$. 
+%\item 
+%We can refer to a set of positive contribution links, negative contribution links, dependency links, and decomposition link with respectively $PosContr, NegContr, Dep,$, and $Decomp$. 
+%\item
+%We can refer to a set of contribution links that can be either positive or negative simply with $Contr$.
+%\end{itemize}
+%\end{definition}
 
-\paragraph{Rationale and example} Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small2}, there are five links. First, let us formalize the related IEs:
-\begin{itemize}
-\item (1, Resource, \text{Car objects}, AND),
-\item (2, Softgoal, \text{Realistic simulation}, AND),
-\item (3, Softgoal, \text{Simple design}, AND),
-\item (4, Goal, \text{Generate traffic}, XOR),
-\item (5, Task, \text{Create new cars}, AND),
-\item (6, Task, \text{Keep same cars}, AND).
-\end{itemize}
+Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small}, there are five links:
 
-We can formalize the links as follows:
 \begin{itemize}
@@
 
-Let us next provide some examples of the notation of Definition~\ref{def:notation2}:
-\begin{itemize}
-\item $dep_7 = (7, Dep, 4, 1)$
-\item $Decomp = \{(10, Decomp, 5, 4), (11, Decomp, 6, 4)\}$
-\item $PosContr = \{(8, PosContr, 5, 2)\}$
-\item $Contr = \{(8, PosContr, 5, 2), (9, NegContr, 5, 3)\}$
-\end{itemize}
+%Let us next provide some examples of the notation of Definition~\ref{def:notation2}:
+%\begin{itemize}
+%\item $dep_7 = (7, Dep, 4, 1)$
+%\item $Decomp = \{(10, Decomp, 5, 4), (11, Decomp, 6, 4)\}$
+%\item $PosContr = \{(8, PosContr, 5, 2)\}$
+%\item $Contr = \{(8, PosContr, 5, 2), (9, NegContr, 5, 3)\}$
+%\end{itemize}
+
+\subsubsection{GRL models}
 
@@
 
-\paragraph{Rationale and example} The definition of a GRL model collects all the previously defined tuples into a single definition.
-For completeness, we now provide the full specification of Figure~\ref{fig:example-small2}.
+The definition of a GRL model collects all the previously defined tuples into a single definition.
+For completeness, we now provide the full specification of Figure~\ref{fig:example-small}.
 This model is formalized as $GRL=(IE, Act, R_{ActIE}, Link)$, where:
@@
 &Act=& &\{(0, Actor, \text{Traffic Simulator})\}&\\
-&R_{ActIE}=& &\{(i+11, ActIE, 0,i)\mid 1\le i\le 6\}&\\
+&R_{ActIE}=& &\{(0, 1), (0,2), (0,3), (0, 4), (0,5),(0,6)\}&\\
 &Link=&\{&(7, Dep, 4, 1)&\\
@@
 \end{flalign*}
+
+The definition of a GRL model (Def.~\ref{def:grl-model}) only sums up the elements of the model and not the constraints that make a \emph{valid} GRL model, viz.
 
@@
 
-\paragraph{Rationale and example} The definition of a GRL model (Def.~\ref{def:grl-model}) comes with various implicit assumptions in order to form a valid GRL model.
-We make these assumptions explicit in the definition above.
-
-Let us briefly verify that our previous formalization of Figure~\ref{fig:example-small2} satisfies all the constraints of Def.~\ref{def:valid-grl-model}:
+Let us briefly verify that our previous formalization of Figure~\ref{fig:example-small} satisfies all the constraints of Def.~\ref{def:valid-grl-model}:
 \begin{enumerate}
@@
 
-\paragraph{Rationale and example} A generic counterargument is simply an argument that can be used to attack any previous argument, or an IE.
+A generic counterargument is simply an argument that can be used to attack any previous argument, or an IE.
 Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links.
@@
 \subsection{Translation algorithms}
+\label{sect:formalframework:translation}
 

### 2017-10-29 gmas.tex
Examples updated (beliefs), SpringerJournal cls used
Small amendmends thoughout the paper.

@@
 Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
-In order to get a sense of such discussions, we performed an extensive case study.
-The objective of our case study was to see which types of discourse are used during a discussion of system requirements, and how these discourse types can be captured as argument schemes and critical questions regarding goals and tasks.
-In order to study this, we manually coded transcripts of such discussions.
-For our coding, we used a list of argument schemes and critical questions based on GRL and PRAS.
+To get a sense of such discussions, we performed an extensive case study to see which types of discourse are used during a discussion of system requirements, and how these discourse types can be captured as argument schemes and critical questions regarding goals and tasks.
+We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS.
 In this section we present our case study.
@@
 \cite{SchriekEtal2016}.
-In these experiments, Schriek and colleagues gave the traffic simulator assignment (Appendix~\ref{sect:designprompt}) to 12 groups of two or three students in a Software Architecture course at MSc level.
+In these experiments, 12 groups of two or three students in a Software Architecture course at MSc level were given the traffic simulator assignment (Appendix~\ref{sect:designprompt}).
 These groups had a maximum of two hours to design a traffic simulator, which included a discussion of the requirements of this traffic simulator.
@@
 We used three of these transcripts, totaling 153 pages, for our case study.
-
-Before we started coding the transcripts, we drew up an initial list of 10 argument schemes (AS0-AS9 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
-AS0 to AS4 are schemes that concern a single element of the goal model.
-So, for example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'.
-AS5 to AS9 are claims about the links between GRL elements.
-Our initial list also contained 18 critical questions, inspired by the questions associated with the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}.
-CQ2 to CQ6b (Table~\ref{table:transcripts:results:argumentschemes}) are examples of these critical questions, other examples are `Is the softgoal legitimate?' and `Are there alternative ways to contribute to the same softgoal?'.
 
@@
 \end{table}
+
+Before we started coding the transcripts, we drew up an initial list of 10 argument schemes (AS0-AS9 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
+AS0 to AS4 are schemes that concern a single element of the goal model.
+So, for example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'.
+AS5 to AS9 are claims about the links between GRL elements.
+Our initial list also contained 18 critical questions, inspired by the questions associated with the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}.
+CQ2 to CQ6b (Table~\ref{table:transcripts:results:argumentschemes}) are examples of these critical questions, other examples are `Is the softgoal legitimate?' and `Are there alternative ways to contribute to the same softgoal?'.
+
 %%TODO check numbers against original annotation docs
@@
 For example, we found that the discussants often talk about tasks decomposing into sub-tasks, so we added AS10 and CQ10a.
-Furthermore, because there were many discussions on the relevance and the clarity of the names of elements (goals, tasks, etc.), two generic critical questions CQ12 and CQ13 were added.
-The final results of the coding can be found in Table~\ref{table:transcripts:results:argumentschemes}; for results per transcript, please consult the online repository.
-We found a total of 159 instantiation of argument schemes AS0-AS11 in transcripts.
+Furthermore, because there were many discussions on the relevance and the clarity of the names of elements, two generic critical questions CQ12 and CQ13 were added.
+The final results of the coding can be found in Table~\ref{table:transcripts:results:argumentschemes}.
+We found a total of 159 instantiations of argument schemes AS0-AS11.
 The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice.

### 2017-10-29 introduction.tex
Small amendmends thoughout the paper.

@@
 
-Among the initial activities in RE are the ``early-phase'' requirements engineering activities, which include those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what the implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
+Among the ``early-phase'' requirements engineering activities are those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what the implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
 These activities fall under the umbrella of goal modeling.
@@
 
-A further problem is that the rationale behind goal modeling decisions is usually static, that is, current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
+A further problem is that current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
 A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
@@
 To show that the RationalGRL methodology can be used in a real case, we illustrate the steps of our methodology with the traffic simulator case study.
-Finally, we have developed a web-based prototype\footnote{\url{insertURL}} for building goal models and arguing about them, which acts as a supporting tool to the RationalGRL methodology (\textbf{requirement 5}).
-
-\todo{F}{all}{check text below when we finish the paper}
+Finally, we have developed a web-based prototype\footnote{\url{http://www.rationalgrl.com}} for building goal models and arguing about them, which acts as a supporting tool to the RationalGRL methodology (\textbf{requirement 5}).
 
 The rest of this article is organized as follows.
-Section~\ref{sect:background} introduces our running example, the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, the Practical Reasoning Argument Scheme (PRAS)~\cite{atkinson2007} and discusses some of our previous work on combining GRL and PRAS.
-Section~\ref{sect:overview} provides a brief and high-level overview of our framework, together with a metamodel and the methodology.
-Section~\ref{sect:gmas} contains an in depth explanation of how we obtained an initial set of argument schemes and critical questions by annotating transcripts from discussions about an information system, and in Section~\ref{sect:examples} we provide several examples of these schemes and questions.
-In Section~\ref{sect:formalframework} we provide formal semantics for GRL and show how argumentation semantics \cite{Dung1995} can be used to compute which arguments are accepted and which are rejected.
-We also develop various algorithms for the argument schemes in this section.
-In Section~\ref{sect:tool} we provide a brief overview of the prototype tool we developed for RationalGRL.
-Finally, Section~\ref{sect:discussion} contains a discussion, covering related work, future work, and a conclusion.
+Section~\ref{sect:background} introduces our case study and briefly discusses the basics of GRL and argumentation about goals using argument schemes.
+Section~\ref{sect:gmas} contains the case study, an in-depth explanation of how we obtained an initial set of argument schemes and critical questions by coding transcripts from discussions about an information system.
+Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL language metamodel, and examples from the case study that illustrate the framework and language.
+In Section~\ref{sect:formalframework} we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and develop various algorithms that change a RationalGRL model according to the argument scheme or critical question that is applied.
+Section~\ref{sect:methodology} discusses the RationalGRL methodology and a discussion of the prototype tool we developed for RationalGRL.
+Finally, Section~\ref{sect:discussion} covers related work, future work, and a conclusion.

### 2017-10-29 main.tex
Finished formalization of GRL
removed old grl specification
finished all algorithms
Examples updated (beliefs), SpringerJournal cls used
added tool screenshots
Small amendmends thoughout the paper.

@@
-\documentclass[11.5pt,two column]{llncs}
+\documentclass[twocolumn]{svjour3}
 
@@
 \usepackage{multirow}
-\usepackage{sectsty}
+%\usepackage{sectsty}
 \usepackage{subcaption}\captionsetup{compatibility=false}
@@
 
-\addtolength{\oddsidemargin}{-0.85in}
-\addtolength{\evensidemargin}{-0.85in}
-\addtolength{\textwidth}{1.7in}
-\addtolength{\topmargin}{-.7in}
-\addtolength{\textheight}{1.3in}
-\setlength{\columnsep}{0.3in}
+%\addtolength{\oddsidemargin}{-0.85in}
+%\addtolength{\evensidemargin}{-0.85in}
+%\addtolength{\textwidth}{1.7in}
+%\addtolength{\topmargin}{-.7in}
+%\addtolength{\textheight}{1.3in}
+%\setlength{\columnsep}{0.3in}
 \newcommand{\Mark}[1]{\textsuperscript{#1}}
@@
 
-% TO TURN OFF TODOS UNCOMMENT THE FOLLOWING:
-% \renewcommand\todo[4][]{}
+\algnewcommand\algorithmicswitch{\textbf{switch}}
+\algnewcommand\algorithmiccase{\textbf{case}}
+\algnewcommand\algorithmicassert{\texttt{assert}}
+\algnewcommand\Assert[1]{\State \algorithmicassert(#1)}%
+% New "environments"
+\algdef{SE}[SWITCH]{Switch}{EndSwitch}[1]{\algorithmicswitch\ #1\ \algorithmicdo}{\algorithmicend\ \algorithmicswitch}%
+\algdef{SE}[CASE]{Case}{EndCase}[1]{\algorithmiccase\ #1}{\algorithmicend\ \algorithmiccase}%
+\algtext*{EndSwitch}%
+\algtext*{EndCase}%
 
-% use like this:
-% \todo{floris}{marc}{Can you please update figure 2?} - Floris is asking Marc to update figure 2, this will show in red in the PDF
-% \todo[resolved]{floris}{marc}{done} - Marc has updated fig 2 (or gives a reason why he doesn't want to uppdate the figure. This will show only in the tex
+%\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
+%\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
 
-\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
-\renewcommand\Affilfont{\fontsize{9}{10.8}\selectfont}
+%\sectionfont{\fontsize{12}{15}\selectfont}
+%\subsectionfont{\fontsize{10}{15}\selectfont}
 
-\sectionfont{\fontsize{12}{15}\selectfont}
-\subsectionfont{\fontsize{10}{15}\selectfont}
+\begin{document}\sloppy
 
-\begin{document}
-\twocolumn[{%
- \centering
- \LARGE RationalGRL: A Framework for Argumentation and Goal Modeling \\[1.5em]
- \large Marc van Zee\Mark{1},
-        Floris Bex\Mark{2}
-        and Sepideh Ghanavati\Mark{3}\\[1em]
- \normalsize
- \begin{tabular}{*{3}{>{\centering}p{.33\textwidth}}}
-  \Mark{1}Computer Science and Communication (CSC) & \Mark{2}Department of Information and Computing Sciences & \Mark{3}Department of Computer Science  \tabularnewline
-  University of Luxembourg & Utrecht University & Texas Tech University \tabularnewline
-  \url{marcvanzee@gmail.com} & \url{f.j.bex@uu.nl} & \url{sepideh.ghanavati@ttu.edu}
- \end{tabular}\\[3em] % some more space after the title part
-}]
+\title{RationalGRL: A Framework for Argumentation and Goal Modeling}
+\author{Marc van Zee \and Floris J. Bex \and Sepideh Ghanavati}
+\institute{Marc van Zee \at Google Research, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} \and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands \\\email{f.j.bex@uu.nl} \and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu}}
+
+\maketitle
 
 \input{abstract}
-
-\newpage
-
-\paragraph{Keywords} Goal modeling $\cdot$ Argumentation $\cdot$ Practical Reasoning $\cdot$ Goal-oriented requirements engineering
 
@@
 %\input{examples}
-\input{tool}
 \input{formalframework}
+\input{methodology}
+%\input{tool}
 \input{discussion}
-
-\section*{Acknowledgments}
-Marc van Zee is funded by the National Research Fund (FNR), Luxembourg, by the Rational Architecture project.
 
@@
 \input{designprompt}
+\newpage
 \input{transcripts}
-\input{grl_specification}
+\newpage
+\input{tool-screenshots}
 

### 2017-10-29 methodology.tex
Small amendmends thoughout the paper.

--- 2017-01-21 methodology.tex
@@
-\section{The RationalGRL Methodology}
+\section{The RationalGRL Methodology and Tool}
 \label{sect:methodology}
 
-The methodology we propose in this paper is visualized in figure~\ref{fig:rationalgrl-methodology}.
-There are two main activities, depicted in grey, namely "goal modeling" and "argumentation".
-These are two separate activities that are being done in parallel.
+In previous sections, we have shown how the RationalGRL framework can capture stakeholder discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
+In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a methodology\footnote{The methodology presented here was first presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.} (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
 
-\begin{figure}[ht]
+\subsection{RationalGRL Methodology}
+\label{sect:methodology}
+
+We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a (Rational)GRL model.
+%Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+The rest of the steps are as follows:
+
+\begin{figure*}[ht]
 \centering
-\includegraphics[scale=0.4]{img/methodology}
+\includegraphics{img/methodology.pdf}
 \caption{The RationalGRL Methodology}
 \label{fig:rationalgrl-methodology}
-\end{figure}
+\end{figure*}
+
+\textbf{(1) Instantiate Argument Schemes (AS)} -- We start with the list of argument schemes (Table~\ref{table:argument-schemes}).
+Whilst discussing the requirements, we select schemes from the list and instantiate them to form arguments for GRL model elements.
+In this way we build or modify the GRL model by introducing new GRL elements (\textsf{INTRO}).
+Note that it is also possible to start modifying an existing GRL model which was not built using the RationalGRL methodology: each GRL element corresponds to an argument (i.e.
+an instantiated scheme), so it is possible to instantiate argument schemes based on an existing GRL model.
+
+\textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we ask the relevant critical questions.
+Because each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.
+
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of three operations are performed on the GRL model: \textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE}.
+Any of these operations impact the arguments and corresponding GRL intentional elements, modifying the initial GRL model into a RationalGRL model.
+After these modifications, we can keep on asking critical questions (e.g.
+about elements that were introduced by previously answering a critical question) until we are satisfied with our model.
+
+\textbf{(4) Modify GRL Models} -- In this step, we modify the regular GRL model based on the RationalGRL model of step (3).
+That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
+This results in a new, modified GRL model, which can be used as the basis for another cycle of the methodology.
+
+We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
+In the next section, we will give an example of how our tool can be used together with the methodology to build a GRL model.
+
+\subsection{The RationalGRL Tool}
+\label{sect:tool}
+
+An important requirement of our framework is that it has tool support.
+This is for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, there does not yet exist any tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent verification to ensure our formal framework and algorithms actually work, iv) having a light-weight web-based version of GRL is useful for the community in general.
+
+GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}.
+This tool is an extension to Eclipse.
+Although it is a rich tool with many features, it is implemented as an Eclipse plugin.
+We have experienced on various occasions that setting this up properly can take up a significant time of experiments, leaving less time for actual modeling.
+
+Our formalization of GRL is very much in line with the way in which GRL models are represented in the open-source Eclipse-based tool jUCMNav.\footnote{See \url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}} By keeping our formalization in line with this tool, we simplify the translation step from models in the RationalGRL tool to models in the jUCMNav tool.
+
+In this section we present our web-based prototype of RationalGRL. It can be accessed from:
+\begin{quote}
+\url{http://www.rationalgrl.com}
+\end{quote}
+
+The tool is written in Javascript, and runs on all modern browsers. The tool is open-source, and the source can be accessed from:
+\begin{quote}
+\url{http://www.github.com/marcvanzee/RationalGRL}
+\end{quote}
+
+In this section we briefly highlight some features of the language, and we explain some of its current limitations.
+
+\todo{all}{marc}{Explain figures in Appendix~\ref{sect:tool-screenshots}}

### 2017-10-29 overview.tex
Small amendmends thoughout the paper.

@@
 
-Having explored the types of arguments and questions in requirements discussions, we now turn to an overview of our RationalGRL framework.
+We now turn to an overview of our RationalGRL framework.
 We will show through a language definition and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
@@
 
-The RationalGRL framework includes two parts: Argumentation and GRL goal modeling.
+Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework.
+There are two activities (bottom), \emph{practical reasoning \& argumentation} and \emph{goal model construction}, which give rise to two different models (top), a \emph{RationalGRL model} and a \emph{GRL model}.
 The GRL part of RationalGRL allows for the creation of goal models by analyzing the non-functional requirements and refining the high-level goals into operationalized tasks.
@@
 Similarly, if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
+Thus, in the framework it becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
 
-\begin{figure}[b]
+\begin{figure}[t]
 \centering
@@
 
-Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework.
-At the bottom there are two activities: \emph{practical reasoning \& argumentation} and \emph{goal model construction}.
-As already explained, these two activities influence each other: adding elements to a goal model gives rise to new critical questions about these elements, and answering critical questions with arguments may add or delete elements from the GRL model.
-The activities give rise to two different models: a \emph{RationalGRL model} (left-hand side) and a \emph{GRL model} (right-hand side).
-In the RationalGRL model, we have the goal model elements (goals, tasks etc) and the arguments for and against these elements; the GRL model contains only the goal model elements.
-Thus, the class of GRL models is essentially a subset of the class of RationalGRL models.
-Given the framework, it then becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
-
 In the rest of this section, we discuss the individual parts of the GRL framework.
@@
 
-A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual types of reasoning stakeholders or designers perform in the early requirements engineering phase (requirement 1 of our framework).
+A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual types of reasoning stakeholders or designers perform (requirement 1 of our framework).
 Recall from section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
-Using this list as a basis, we further refined our set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
-Note that this list of argument schemes and critical questions is not exhaustive.
-It is an initial list that we have obtained by coding transcripts.
-However, our framework is fully extensible, meaning that new argument schemes and critical questions can be added depending on the problem domain.
+Using this list as a basis, we further refined the set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
+Note that this list is not exhaustive, and that new argument schemes and critical questions can be added depending on the problem domain.
 
@@
 \label{sect:overview:lang}\label{sect:metamodel}
-The language used to construct RationalGRL models is an extension of GRL: it includes all the GRL elements shown in Figure~\ref{fig:grl_legend}.
+RationalGRL is an extension of GRL and includes all the elements shown in Figure~\ref{fig:grl_legend}.
 However, there are also new elements corresponding to argumentation-related concepts.
@@
 This metamodel represents the abstract grammar of the language, independently of the notation.
-The metamodel also formalizes the GRL concepts and constructs introduced in Section~\ref{sect:background:grl}\footnote{Note that for readability, some GRL concepts, such as contribution strength, have been omitted from the figure.}.
+The metamodel also formalizes the GRL concepts and constructs introduced in Section~\ref{sect:background:grl}.
+
+\begin{figure*}[t]
+\includegraphics[width=\textwidth]{metamodel/metamodel}
+\caption{The RationalGRL metamodel}
+\label{fig:metamodel}
+\end{figure*}
 
@@
 Finally, a \textsf{GRLmodel} is composed of \textsf{GRLModelElements}.
-
-\begin{figure*}[t]
-\includegraphics[width=\textwidth]{metamodel/metamodel}
-\caption{The RationalGRL metamodel}
-\label{fig:metamodel}
-\end{figure*}
 
@@
 
-Notice that there are various \textsf{OperationTypes} in the Argumentation package.
+Notice that there are \textsf{OperationTypes} in the Argumentation package.
 In RationalGRL, these operations are performed by instantiating an argument scheme or answering a critical question in a certain way.
@@
 The formal grounding for the connection between practical reasoning and RationalGRL models can be found in the RationalGRL Metamodel (Section~\ref{sect:metamodel}) and in more detail in the logical formalization of the Argument Schemes and Critical Questions (Section~\ref{sect:algorithms}).
-The connection between the RationalGRL models shown in this section and regular GRL models is further formally defined in section \ref{sect:rationalGRL-GRL}.
+The connection between the RationalGRL models shown in this section and regular GRL models is further formally defined in Section \ref{sect:formalframework:translation}.
 
@@
 We start by showing how instantiating argument schemes leads to the introduction of new RationalGRL elements in a model.
-Take the example in Figure~\ref{fig:example_AS}, which is based on the transcript excerpt shown in Table~\ref{table:transcript:decomposition}.
+Take the example in Figure~\ref{fig:example_AS}, which is based on the excerpt from transcript $t_1$ shown in Table~\ref{table:transcript:decomposition} in Appendix~\ref{sect:transcripts:excerpts}.
 On the left side, the arguments found in the transcript are shown, together with the argument scheme they are based on.
@@
 
-The transcript excerpt of this example is shown in Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_1$.
+The excerpt from transcript $t_1$ used in this example is shown in Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts}.
 In this example, participants first sum up functionality of the traffic simulator, which can be captured as instantiations of AS2.
@@
 This attack leads to the original argument being \emph{rejected} (cf.
-Section~\ref{sect:background:pras} and Section~\ref{sect:goalmodeling:argumentationsemantics}), indicated by it being greyed out.
+Section~\ref{sect:background:pras} and Section~\ref{sect:formalframework:rationalgrl}), indicated by it being greyed out.
 As a result of this, the corresponding GRL task \emph{Add traffic light} is also disabled.
@@
 
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:decomposition} in the appendix and comes from transcript $t_3$.
-It consists of a discussion about the type of decomposition relationship for the goal \emph{Simulate Traffic}.
+The excerpt of transcript $t_3$ used for this example is shown in Table~\ref{table:transcript:decomposition} in Appendix~\ref{sect:transcripts:excerpts}.
+It consists of a discussion about the type of decomposition relationship for the goal \emph{Simulate Traffic} (Figure~\ref{fig:examples:decomposition}).
 Recall that in Example 1, an AND-decomposition was introduced for this goal with AS5 (Figure~\ref{fig:example_AS}).
@@
 The answer is ``No, it should be OR''.
-The original argument for AND-decomposition is now attacked  by the argument for the OR-decomposition, and the new argument is linked to the OR-decomposition in the RationalGRL model (Figure~\ref{fig:examples:decomposition}).
-
-\paragraph{Example 4: Clarifying a task by answering critical questions (\textsf{REPLACE})}
-
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_1$.
-The discussion starts with an instantiation of argument scheme AS2: ``Actor \emph{Student} has task \emph{Set car influx}''.
-This argument is then challenged with critical question CQ12: ``Is the task \emph{Set car influx} specific enough?''.
-This is answered negatively, creating a new argument ``Actor \emph{Student} has task \emph{Set car influx per road}'', which attacks the original argument for \emph{Set car influx}.
-Note how the new task \emph{Set car influx per road} also attacks (and disables) the original RationalGRL task \emph{Set car influx} (Figure~\ref{fig:examples:clarification}).
+The original argument for AND-decomposition is now attacked  by the argument for the OR-decomposition, and the new argument is linked to the OR-decomposition in the RationalGRL model.
 
@@
 \end{figure}
+
+\paragraph{Example 4: Clarifying a task by answering critical questions (\textsf{REPLACE})}
+
+The transcript excerpt of this example is shown in Table~\ref{table:transcript:task-clarification} in Appendix~\ref{sect:transcripts:excerpts} and comes from transcript $t_1$.
+The discussion starts with an instantiation of argument scheme AS2: ``Actor \emph{Student} has task \emph{Set car influx}'' (Figure~\ref{fig:examples:clarification}).
+This argument is then challenged with critical question CQ12: ``Is the task \emph{Set car influx} specific enough?''.
+This is answered negatively, creating a new argument ``Actor \emph{Student} has task \emph{Set car influx per road}'', which attacks the original argument for \emph{Set car influx}.
+Note how the new task \emph{Set car influx per road} also attacks (and disables) the original RationalGRL task \emph{Set car influx}.
 
@@
 
-The transcript excerpt of this example is shown in Table~\ref{table:transcript:irrelevant-actor} in the appendix and comes from transcript $t_3$.
-First participant \emph{P1} puts forth the suggestion to include actor \emph{Development team} in the model.
-This is, then, questioned by participant \emph{P2}, who argues that the professor will develop the software, so there will not be any development team.
-This is shown in Figure~\ref{fig:examples:relevant-actor}: first the actor \emph{Development Team} is introduced with an argument based on AS0.
-This argument is then attacked by answering critical question CQ0.
+The excerpt from transcript $t_3$ used in this example is shown in Table~\ref{table:transcript:irrelevant-actor} in Appendix~\ref{sect:transcripts:excerpts}.
+Actor \emph{Development Team} is introduced with an argument based on AS0 (Figure~\ref{fig:examples:relevant-actor}).
+This is then attacked by arguing that the professor will develop the software, so there will not be any development team (CQ0).
 
-\begin{figure}[t]
+\begin{figure}[b]
 \centering
@@
 
-Further in the discussion, participant \emph{P2} argues that the development team should be considered, since the professor does not develop the software.
-This is captured using a generic counterargument (\emph{Att} in Table~\ref{table:argument-schemes}), which attacks the earlier argument based on CQ0.
+Further in the discussion, it is then argued that the development team should be considered, since the professor does not develop the software.
+This is captured using a generic counterargument (\emph{Att}, which attacks the earlier argument based on CQ0.
 Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counterargument has been put forward: the argument (Att) now attacks the argument (CQ0), which in turn attacks the original argument (AS0).
 As a result, the argument (AS0) is acceptable (cf.
-Section~\ref{sect:background:pras} and Section~\ref{sect:goalmodeling:argumentationsemantics}), which causes the actor in the RationalGRL model to be enabled again.
+Section~\ref{sect:background:pras} and Section~\ref{sect:formalframework:rationalgrl}), which causes the actor in the RationalGRL model to be enabled again.

### 2017-10-30 discussion.tex
updated future work

@@
 Furthermore, the argumentative part of their framework does not include formal semantics for resolving conflict between arguments or determining the acceptability of arguments.
-Yu et al.~\cite{yu2015automated} further extend the framework by Haley et al., including algorithms for Dung-style \cite{Dung1995} argumentation semantics and a database of specific ways in which to attack (or mitigate) risks, which can be likened to a set of critical questions for risks and security requirements (cf.
+Yu et al.~\cite{yu2015automated} further extend the framework by Haley et al., including algorithms for Dung-style argumentation semantics \cite{Dung1995} and a database of specific ways in which to attack (or mitigate) risks, which can be likened to a set of critical questions for risks and security requirements (cf.
 Yu et al.~\cite{yu2015automated} Section 3.1).
 
-Another recent example of the use of arguments in goal-oriented requirements engineering is the work by Murukannaiah et al.~\cite{murukannaiah2015}, who propose Arg-ACH, an approach 
-in which the beliefs underlying conflicting goals can be made explicit using argumentation.
+Another recent example of the use of arguments in goal-oriented requirements engineering is the work by Murukannaiah et al.~\cite{murukannaiah2015}, who propose Arg-ACH, where the beliefs underlying conflicting goals can be made explicit using argumentation.
 Murukannaiah et al.
@@
 \paragraph{Goal Modeling} Argumentation has been included -- both explicitly and implicitly -- in existing goal modeling languages.
-For example, the belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} is meant to capture the rationales (reasons, arguments) behind the inclusion of goals and tasks in the model.
+For example, the belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} is meant to capture the rationales behind the inclusion of goals and tasks in the model.
 Furthermore, relations between elements in a goal model also provide justifications: high-level goals are reasons for lower-level goals, tasks and resources.
@@
 There are several contributions to the literature that relate argumentation-based techniques with goal modeling.
-The first line of work is by Bagrheri and Ensan \cite{bagheri2011consolidating} and Mirbel and Villata \cite{MirbelVillata12} use abstract Dung argumentation frameworks (see Section~\ref{sect:formalframework}) to capture individual goals and the relations between them as modeled in a goal model, so that consistent (i.e., acceptable) subsets of goals can be computed using the relevant argumentation semantics.
+The first line of work is by Bagrheri and Ensan \cite{bagheri2011consolidating} and Mirbel and Villata \cite{MirbelVillata12} use abstract argumentation frameworks (see Definition~\ref{def:argumentation-framework}) to capture individual goals and the relations between them as modeled in a goal model, so that consistent (i.e., acceptable) subsets of goals can be computed using the relevant argumentation semantics \cite{Dung1995}.
 For example, if in the goal model there is a conflict between goals $G_1$ and $G_2$, there is an attack between the arguments representing these goals and hence $G_1$ and $G_2$ cannot be in the same extension.
@@
 
-Modeling goal models as argumentation frameworks allows one to compute the possible consistent sets of goals and tasks given a goal model.
+Modeling goal models as argumentation frameworks allows one to compute the consistent sets of goals and tasks given a goal model.
 In RationalGRL, we have opted not to provide such an argumentation-theoretic semantics to goal models.
@@
 
-In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we effectively extended Jureta et al.'s work and translated argument diagrams to GRL models\footnote{In \cite{Jureta:RE2008} only a mapping from goal models to structured arguments is given, and the step from structured arguments or argument diagrams to goal models is never formally defined.} (an automatic translation tool is discussed in~\cite{vanZee-etal:comma2016}).
+In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we extended Jureta et al.'s work and translated argument diagrams to GRL models\footnote{In \cite{Jureta:RE2008} only a mapping from goal models to structured arguments is given, and the step from structured arguments or argument diagrams to goal models is never formally defined.} (an automatic translation tool is discussed in~\cite{vanZee-etal:er2016}).
 Thus, we have essentially two complex diagrams, an argument diagram and a goal diagram, and a mapping between them.
@@
 
-We see a large number of open issues that we hope will be explored in future research. We discuss five promising directions here.
+RationalGRL lays down a basic framework for argumentation in requirements engineering, and all aspects of this framework are intended to be fully extensible.
+We see a large number of open issues to be explored in future research.
 
-\subsubsection*{Architecture principles} 
-One aspect of enterprise architecture that we did not touch upon in this article are \emph{(enterprise) architecture principles}.
-Architecture principles are general rules and guidelines, intended to be enduring and seldom amended, that inform and support the way in which an organization sets about fulfilling its mission~\cite{Lankhorst2005,OptLand2007a,OG2009}.
-They reflect a level of consensus among the various elements of the enterprise, and form the basis for making future IT decisions.
-Two characteristics of architecture principles are:
-\begin{itemize}
-\item There are usually a small number of principles (around 10) for an entire organization.
-These principles are developed by enterprise architecture, through discussions with stakeholders or the executive board.
-Such a small list is intended to be understood \emph{throughout the entire organization}.
-All employees should keep these principles in the back of their hard when making a decision.
-\item Principles are meant to guide decision making, and if someone decides to deviate from them, he or she should have a good reason for this and explain why this is the case.
-As such, they play a normative role in the organization.
-\end{itemize}
+\paragraph{Specific Domains}
+At the moment, the argumentation schemes and critical questions of RationalGRL are focused only on the core goal modeling process, that is, the discussion about the goals and functionality of an information system.
+In the RE process, there are many more domain specific discussions that also take place, such as discussions involving expert opinions \cite{murukannaiah2015}, discussions about security requirements \cite{haley2008security,yu2015automated,ionita2014argumentation}, architectural principles \cite{marosin-etal:caise2016}, legal compliance \cite{Ghanavati2013}, and so forth.
+From a knowledge engineering perspective, including the argument schemes and critical questions associated with these domains in the RationalGRL framework is a time-consuming task.
+However, from a formal perspective adding new schemes and questions is easy: as already discussed by Bex et al.~\cite{bexEtal2003}, critical questions for argument schemes will always lead to either new information being introduced, new information replacing old information, or new information attacking old information.
+This corresponds to the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations in the RationalGRL framework, which have been formalized in Algorithms 3-9 in Section~\ref{sect:algorithms}, which would hence be suitable for other types of schemes and questions as well.
 
-Looking at these two characteristics, we see that argumentation, or justification, plays an important role in both forming the principles and adhering to them:
-\begin{itemize}
-\item Architecture principles are \emph{formed} based on underlying arguments, which can be the goals and values of the organization, preferences of stakeholders, environmental constraints, etc.
-\item If architecture principles are \emph{violated}, this violation has to be explained by underlying arguments, which can be project-specific details or lead to a change in the principle.
-\end{itemize}
+\paragraph{Formal argumentation}
+The amount of theory from computational argumentation used in this article has been relatively small.
+Our intention is to create a bridge between the formal theories in argumentation and the practical tools in requirements engineering.
+Now that the initial framework has been developed, however, it is worth exploring what additional techniques computational argumentation has to offer in more detail.
+For instance, in our framework we have not included the possibility of expressing preferences between mutually inconsistent arguments (e.g.
+like in Figure~\ref{fig:XX}).
+Using the work by Modgil~\cite{modgil2009}, it is possible to provide explicit arguments for preferences, thus allowing us to make a reasoned choice between two options.
 
-In a previous paper, we~\cite{marosin-etal:caise2016} propose an extension to GRL based on enterprise architecture principles.
-We present a set of requirements for improving the clarity of definitions and develop a framework to formalize architecture principles in GRL.
-We introduce an extension of the language with the required constructs and establish modeling rules and constraints.
-This allows one to automatically reason about the soundness, completeness and consistency of a set of architecture principles.
-Moreover, principles can be traced back to high-level goals.
+\paragraph{Tool support}
 
-It would be very interesting future work to combine the architecture principles extension with the argumentation extension.
-This would lead to a framework in which principles cannot only be traced back to goals, but also to underlying arguments by the stakeholders.
-
-\subsubsection*{Extensions for argumentation}
-
-The amount of argumentation theory we used in this article has been rather small.
-Our intention was to create a bridge between the formal theories in argumentation and the rather practical tools in requirements engineering.
-Now that the initial framework has been developed, is it worth exploring what tools and variations formal argumentation has to offer in more detail.
-
-For instance, until now we have assumed that every argument put forward by a critical questions always defeats the argument it questions, but this is a rather strong assumption.
-In some cases, it is more difficult to determine whether or not an argument is defeated.
-Take, for example, the argumentation framework in Figure~\ref{fig:goalmodeling:futureargs} with just A1 and A2.
-These two arguments attack each other, they are alternatives and without any explicit preference, and it is impossible to choose between the two.
-It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}.
-If we say that we prefer the action  \texttt{Create new cars} (A2) over the action  \texttt{Keep same cars} (A1), we remove the attack from A1 to A2.
-This makes A2 the only undefeated argument, whereas A1 is now defeated.
-It is also possible to give explicit arguments for preferences~\cite{modgil2009}.
-These arguments are then essentially attacks on attacks.
-For example, say we prefer A3 over A1 because `it is important to have realistic traffic flows' (A4).
-This can be rendered as a separate argument that attacks the attack from A1 to A3, removing this attack and making $\{$A3, A4$\}$ the undefeated set of arguments.
-
-Allowing undefeated attacks also make the question of which semantics to choose more interested.
-In our current (a-cyclic) setting, all semantics coincide, and we always have the same set of accepted arguments.
-However, once we allow for cycles, we may choose accepted arguments based on semantics which, for instance, try to accept/reject as many arguments as possible (preferred semantics), or just do not make any choice once there are multiple choices (grounded).
-Another interesting element of having cycles is that one can have multiple extensions.
-This corresponds to various \emph{positions} are possible, representing various sets of possibly accepted arguments.
-Such sets can then be shown to the user, who can then argue about which one they deem most appropriate.
-
-\begin{figure}[ht]
-\centering
-\begin{tikzpicture}
-        \node[minimum size=1cm] (a1) [argNodeIN] at (0,0) {$A_1$};
-        \node[minimum size=1cm] (a2) [argNodeIN] at (3,0) {$A_2$};
-
-         \path
-    (a2) edge [attackLink] (a1)
-    (a1) edge [attackLink] (a2);
-\end{tikzpicture}
-\caption{Preferences between arguments}
-\label{fig:goalmodeling:futureargs}
-\end{figure}
-
-Finally, in this article we have only explored one single argument scheme, but there are many other around.
-In his famous book ``Argumentation schemes'', Walton describes a total of 96 schemes.
-Murukannaiah \emph{et al.}~\cite{murukannaiah2015} already explain how some of these schemes may be use for resolving goal conflicts, and it is worth studying what this would look like in our framework as well.
+The first possibility is to consider using argumentation semantics \cite{Dung1995} as a way to combine multiple different, possibly inconsistent goal models (cf.
+\cite{bagheri2011consolidating,MirbelVillata12})
 
@@
 
-\subsubsection*{Empirical study}
-
-Although we develop our argument schemes and critical questions with some empirical data, we did not yet validate the outcome.
-This is an important part, because it will allow us to understand whether adding arguments to goal modeling is actually useful.
-We have developed an experimental setup for our experiment, which we intend to do during courses at various universities.
-However, we cannot carry out this experiment until the tool is finished.
-
-\subsubsection*{Formal framework}
-
-The formal framework we present in this article is very simple, and does not provide a lot of detail.
-We believe it would be interesting to develop a more robust characterization of a GRL model using logical formulas.
-Right now, we have no way to verify whether the goal models we obtain through out algorithms are actually valid GRL models.
-This is because we allow any set of atoms to be a GRL model, which is clearly very permissive and incorrect.
-Once we develop a number of such constraints, we can ensure (and even proof) our algorithms do not generate invalid GRL models.
-
-For instance, suppose we assert that an \emph{intentional element} is a goal, softgoal, task, or resource: \begin{align*}
-(softgoal(i)\vee goal(i)\vee task(i)\\
-\vee resource(i)\rightarrow IE(i).
-\end{align*}
-We can then formalize an intuition such as: ``Only intentional elements can be used in contribution relations'' as follows
-\begin{align*}
-contrib(k,i,j,ctype)\rightarrow \\
-(IE(i)\wedge IE(j)\wedge IE(j).
-\end{align*}
-
-Interestingly, such constraints are very comparable to \emph{logic programming} rules.
-We therefore see it as interesting future research to explore this further, specifically in the following two ways:
-\begin{itemize}
-\item Develop a set of constraints on sets of atoms of our language, which correctly describe a GRL model.
-Show formally that using our algorithms, each extension of the resulting argumentation framework corresponds to a valid GRL model, i.e., a GRL model that does not violate any of the constraints.
-\item Implement the constraints as a logic program, and use a logic programming language to compute the resulting GRL model.
-\end{itemize}
+\paragraph{Empirical validation}
 

### 2017-10-30 introduction.tex
updated future work

@@
 
-One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for, among other things, dialogues about safety critical actions \cite{tolchinsky2012deliberation} and software design discussions~\cite{BlackEtal2013}.
+One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for, among other things, modeling software design discussions~\cite{BlackEtal2013}.
 Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyse and guide stakeholders' discussions about goal models.

### 2017-10-30 outtakes.tex
updated future work

@@
+\subsubsection*{Architecture principles} 
+One aspect of enterprise architecture that we did not touch upon in this article are \emph{(enterprise) architecture principles}.
+Architecture principles are general rules and guidelines, intended to be enduring and seldom amended, that inform and support the way in which an organization sets about fulfilling its mission~\cite{Lankhorst2005,OptLand2007a,OG2009}.
+They reflect a level of consensus among the various elements of the enterprise, and form the basis for making future IT decisions.
+Two characteristics of architecture principles are:
+\begin{itemize}
+\item There are usually a small number of principles (around 10) for an entire organization.
+These principles are developed by enterprise architecture, through discussions with stakeholders or the executive board.
+Such a small list is intended to be understood \emph{throughout the entire organization}.
+All employees should keep these principles in the back of their hard when making a decision.
+\item Principles are meant to guide decision making, and if someone decides to deviate from them, he or she should have a good reason for this and explain why this is the case.
+As such, they play a normative role in the organization.
+\end{itemize}
+
+Looking at these two characteristics, we see that argumentation, or justification, plays an important role in both forming the principles and adhering to them:
+\begin{itemize}
+\item Architecture principles are \emph{formed} based on underlying arguments, which can be the goals and values of the organization, preferences of stakeholders, environmental constraints, etc.
+\item If architecture principles are \emph{violated}, this violation has to be explained by underlying arguments, which can be project-specific details or lead to a change in the principle.
+\end{itemize}
+
+In a previous paper, we~\cite{marosin-etal:caise2016} propose an extension to GRL based on enterprise architecture principles.
+We present a set of requirements for improving the clarity of definitions and develop a framework to formalize architecture principles in GRL.
+We introduce an extension of the language with the required constructs and establish modeling rules and constraints.
+This allows one to automatically reason about the soundness, completeness and consistency of a set of architecture principles.
+Moreover, principles can be traced back to high-level goals.
+
+It would be very interesting future work to combine the architecture principles extension with the argumentation extension.
+This would lead to a framework in which principles cannot only be traced back to goals, but also to underlying arguments by the stakeholders.
+
+\subsubsection*{Extensions for argumentation}
+
+The amount of argumentation theory we used in this article has been rather small.
+Our intention was to create a bridge between the formal theories in argumentation and the rather practical tools in requirements engineering.
+Now that the initial framework has been developed, is it worth exploring what tools and variations formal argumentation has to offer in more detail.
+
+For instance, until now we have assumed that every argument put forward by a critical questions always defeats the argument it questions, but this is a rather strong assumption.
+In some cases, it is more difficult to determine whether or not an argument is defeated.
+Take, for example, the argumentation framework in Figure~\ref{fig:goalmodeling:futureargs} with just A1 and A2.
+These two arguments attack each other, they are alternatives and without any explicit preference, and it is impossible to choose between the two.
+It is, however, possible to include explicit preferences between arguments when determining argument acceptability \cite{amgoud2002reasoning}.
+If we say that we prefer the action  \texttt{Create new cars} (A2) over the action  \texttt{Keep same cars} (A1), we remove the attack from A1 to A2.
+This makes A2 the only undefeated argument, whereas A1 is now defeated.
+It is also possible to give explicit arguments for preferences~\cite{modgil2009}.
+These arguments are then essentially attacks on attacks.
+For example, say we prefer A3 over A1 because `it is important to have realistic traffic flows' (A4).
+This can be rendered as a separate argument that attacks the attack from A1 to A3, removing this attack and making $\{$A3, A4$\}$ the undefeated set of arguments.
+
+Allowing undefeated attacks also make the question of which semantics to choose more interested.
+In our current (a-cyclic) setting, all semantics coincide, and we always have the same set of accepted arguments.
+However, once we allow for cycles, we may choose accepted arguments based on semantics which, for instance, try to accept/reject as many arguments as possible (preferred semantics), or just do not make any choice once there are multiple choices (grounded).
+Another interesting element of having cycles is that one can have multiple extensions.
+This corresponds to various \emph{positions} are possible, representing various sets of possibly accepted arguments.
+Such sets can then be shown to the user, who can then argue about which one they deem most appropriate.
+
+\begin{figure}[ht]
+\centering
+\begin{tikzpicture}
+        \node[minimum size=1cm] (a1) [argNodeIN] at (0,0) {$A_1$};
+        \node[minimum size=1cm] (a2) [argNodeIN] at (3,0) {$A_2$};
+
+         \path
+    (a2) edge [attackLink] (a1)
+    (a1) edge [attackLink] (a2);
+\end{tikzpicture}
+\caption{Preferences between arguments}
+\label{fig:goalmodeling:futureargs}
+\end{figure}
+
+Finally, in this article we have only explored one single argument scheme, but there are many other around.
+In his famous book ``Argumentation schemes'', Walton describes a total of 96 schemes.
+Murukannaiah \emph{et al.}~\cite{murukannaiah2015} already explain how some of these schemes may be use for resolving goal conflicts, and it is worth studying what this would look like in our framework as well.
+
+One idea is to capture requirements engineering and software design processes as explicit dialogues between parties~\cite{finkelstein1989multiparty}.
+Software design discussions Black et al., Prakken and Wierenga.
+
+\subsubsection*{Empirical study}
+
+Although we develop our argument schemes and critical questions with some empirical data, we did not yet validate the outcome.
+This is an important part, because it will allow us to understand whether adding arguments to goal modeling is actually useful.
+We have developed an experimental setup for our experiment, which we intend to do during courses at various universities.
+However, we cannot carry out this experiment until the tool is finished.
+
+\subsubsection*{Formal framework}
+
+The formal framework we present in this article is very simple, and does not provide a lot of detail.
+We believe it would be interesting to develop a more robust characterization of a GRL model using logical formulas.
+Right now, we have no way to verify whether the goal models we obtain through out algorithms are actually valid GRL models.
+This is because we allow any set of atoms to be a GRL model, which is clearly very permissive and incorrect.
+Once we develop a number of such constraints, we can ensure (and even proof) our algorithms do not generate invalid GRL models.
+
+For instance, suppose we assert that an \emph{intentional element} is a goal, softgoal, task, or resource: \begin{align*}
+(softgoal(i)\vee goal(i)\vee task(i)\\
+\vee resource(i)\rightarrow IE(i).
+\end{align*}
+We can then formalize an intuition such as: ``Only intentional elements can be used in contribution relations'' as follows
+\begin{align*}
+contrib(k,i,j,ctype)\rightarrow \\
+(IE(i)\wedge IE(j)\wedge IE(j).
+\end{align*}
+
+Interestingly, such constraints are very comparable to \emph{logic programming} rules.
+We therefore see it as interesting future research to explore this further, specifically in the following two ways:
+\begin{itemize}
+\item Develop a set of constraints on sets of atoms of our language, which correctly describe a GRL model.
+Show formally that using our algorithms, each extension of the resulting argumentation framework corresponds to a valid GRL model, i.e., a GRL model that does not violate any of the constraints.
+\item Implement the constraints as a logic program, and use a logic programming language to compute the resulting GRL model.
+\end{itemize}
+
 %I wasn't really sure what to do with the below section, so I commented it out. 

### 2017-10-31 discussion.tex
future work finished

@@
 \paragraph{Tool support}
-
-The first possibility is to consider using argumentation semantics \cite{Dung1995} as a way to combine multiple different, possibly inconsistent goal models (cf.
-\cite{bagheri2011consolidating,MirbelVillata12})
-
-One idea is to capture requirements engineering and software design processes as explicit dialogues between parties~\cite{finkelstein1989multiparty}.
-Software design discussions Black et al., Prakken and Wierenga.
+The current tool is a prototype that implements the RationalGRL framework in a fairly simple way, that is, by offering the possibility to argue and ask critical questions about a goal model, and then export this goal model for further analysis in jUCMNav.
+Apart from implementing the tool to include the new fatures described above (e.g.
+new argument schemes, reasoning with preferences), it is also interesting to think about further possibilities.
+One idea is to use the tool for collaborative on-line goal modeling, similar to Github: because the renaming (i.e.
+replacement) and deletion (i.e.
+disabling) of elements are all logged, it is easy for a stakeholder to continue working on a model that was made by another stakeholder.
+Furthermore, critical questions can be included for other stakeholders to answer at a later date.
+As a formal underpinning of this asynchronous communication between users, it would make sense to capture requirements engineering and software design processes as dialogs between parties~\cite{finkelstein1989multiparty,BlackEtal2013}, which are a natural fit with the question-answer format employed in the RationalGRL framework.
 
 \paragraph{Empirical validation}
+We have performed a large case study and found more than 200 incidences of arguments and questions.
+However, what is still lacking is further empirical investigation of whether, and how, the tools and techniques provided by the RationalGRL framework really improve early phase requirements engineering.
+The complexity of the domain makes focused experiments difficult, but not impossible.
+For example, Murakannaiah et al.~\cite{murukannaiah2015} provided a realistic but bounded problem for their subjects to solve, and similar experiments are imaginable to test RationalGRL: have two groups solve a goal modelling problem, one using only GRL and one using RationalGRL, and see how they perform.
+Naturally, the type of problem will greatly influence the outcome.
+For a simple problem which two people have to solve in an hour, RationalGRL will probably not add much.
+However, a more complex problem on which multiple persons have to work asynchronously might benefit from the extra tools offered by RationalGRL.
+Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the methodology, the tool, or a list of critical questions to serve as reminders during the goal modelling process (cf.
+\cite{SchriekEtal2016}}).
 

### 2017-11-01 formalframework.tex
formalframework cleanup & new example

@@
 
-In Section~\ref{sect:gmas} we developed a list of critical questions and argument schemes by analyzing transcripts of discussions about the development of a traffic simulator.
-The resulting list is shown in Table~\ref{table:argument-schemes}.
-We then presented the RationalGRL framework in secton~\ref{sect:overview}, consisting of a modeling language, a metamodel, and various examples of using the argument schemes and critical questions.
-In section~\ref{sect:meth+tool} we explained how this framework can be used by practitioners by explaining the methodology and tool support.
+In Section~\ref{sect:overview}, we have shown through a language definition and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
+A formal rendition of this traceability will be presented in this section, in which we present a logical formalization of RationalGRL.
+This is done for multiple reasons: (i) Most approaches in formal argumentation (cf.
+Section~\ref{sect:background:pras}) use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated in to valid GRL models and visa versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support.
 
-In this section we present a formalization of RationalGRL based on formal logic.
-This is done for multiple reasons: (i) Most approaches in formal argumentation use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated in to valid GRL models and visa versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support.
-
-In the first two subsections we formalize a static representation of our framework: We provide a formal specification of a GRL model based on the GRL metamodel (Section \ref{sect:metamodel}) in the first subsection, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
-
-In the third subsection we develop algorithms in order to translate a GRL model into a RationalGRL model, and visa versa.
-
-In the fourth and final subsection we turn to the dynamics of our framework. We develop algorithms for instantiating argument schemes and for answering critical questions.
+In Sections \ref{sect:formalframework:grl} and \ref{sect:formalframework:rationalgrl} we formalize a static representation of our framework based on the metamodel Figure~\ref{fig:metamodel}).
+We first provide a formal specification of a GRL model in Section~\ref{sect:formalframework:grl}, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
+In Section~\ref{sect:formalframework:translation} we then present algorithms in order to translate a GRL model into a RationalGRL model, and vice versa.
+Finally, in Section~\ref{sect:algorithms} we turn to the dynamics of our framework, developing algorithms for instantiating argument schemes and answering critical questions and thus formally capturing the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations of which examples were previously given in Section~\ref{sect:overview:examples}.
 
@@
 
-In this subsection we formalize a GRL model based on the GRL metamodel and the jUCMNav implementation.
-We first formalize elements of GRL (intentional elements and actors), and then formalize the links.
+We formalize a GRL model based on the metamodel (Figure~\ref{fig:metamodel}), starting with intentional elements and actors.
 
-\subsubsection{Intentional Element, Beliefs and Actors}
-
-Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id, i, j, name, ie, goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
-We start by defining a few general elements of our language that we use in subsequent definitions.
+Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id$, $i$, $j$, $name$, $goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
+We start by defining general elements of our language that we use in subsequent definitions.
 
@@
 Extending the formalization model to include more values is straightforward.},
-\item $Types = IETypes \cup LinkTypes\cup\{Actor, GenArg, ActIE\}$,
+\item $Types = IETypes \cup LinkTypes\cup\{Actor, ActIE, GenArg\}$,
 \item $Names$ is a finite set of strings,
@@
 
-The definition above is sufficient to capture all intentional elements used in GRL. %In the next definition, we present some simplifying notational conventions.
+The definition above is sufficient to capture all intentional elements (IEs) used in GRL. %In the next definition, we present some simplifying notational conventions.
 %\begin{definition}[Notation]
@@
 %\end{definition}
-Throughout this section we use the example in Figure~\ref{fig:example-small}. The IEs in this example can be formalized using Def.~\ref{def:ie} as follows:
-\begin{itemize}
-\item (1, Resource, \text{Car objects}, AND),
-\item (2, Softgoal, \text{Realistic simulation}, AND),
-\item (3, Softgoal, \text{Simple design}, AND),
-\item (4, Goal, \text{Generate traffic}, XOR),
-\item (5, Task, \text{Create new cars}, AND),
-\item (6, Task, \text{Keep same cars}, AND).
-\end{itemize}
+Throughout this section we use the example in Figure~\ref{fig:example-small}. This example contains various IEs which can be formalized using Definition~\ref{def:ie}, for example, 
+$(1, Resource, \text{Car objects}, AND)$, $(2, Softgoal,$ Realistic simulation$, AND)$ and $(5, Goal,$ Generate traffic$, XOR)$.
+
 %Using the short-hand notation defined in Def.~\ref{def:notation}, we can make the following statements:
@@
 \end{definition}
-An actor is simply a pair consisting of its identifier and its name.
-Similar to intentional elements, we may refer to $act = (id,name)$ with $act_{id}$ and write $act_{id}.name$ to refer to its name.
-We can formalize the actor of Figure~\ref{fig:example-small} as $act=(0,Actor,\text{Traffic Simulator})$.%, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
+We can now formalize the actor in our example (Figure~\ref{fig:example-small}) as $(0,Actor,\text{Traffic Simulator})$.%, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
 
@@
 \label{def:act-ie-relation}
-An \emph{Actor-IE relation} is a tuple $(i, j)$, where:
+An \emph{Actor-IE relation} is a tuple $(type, i, j)$, where:
 \begin{itemize}
 %\item $id\in\mathbb{N}$ is the identifier of the relation, 
-%\item $ActIE$ states that this tuple is an Actor-IE relation,
+\item $type = ActIE$ states that this tuple is an Actor-IE relation,
 \item $i\in\mathbb{N}$ is an id of an actor,
@@
 
-Note that Actor-IE relations do not have an $id$, as they signify relations between elements of a GRL model (i.e., intentional elements and actors) and are themselves not elements of a GRL model.
-The relation between the \emph{Traffic Simulator} actor ($id = 0$) and the \emph{Car objects} resource, for example, can be formalized as $(0, 1)$.
-
-\subsubsection{Links}
+Note that Actor-IE relations do not have an $id$: they are themselves not elements of a GRL model but rather indicate relations between elements of a GRL model (i.e., intentional elements and actors).
+The relation between the \emph{Traffic Simulator} actor ($id = 0$) and the \emph{Car objects} resource ($id=1$), for example, can be formalized as $(0, 1)$.
 
@@
 \item $type\in LinkTypes$ specifies the type of the link, 
-\item $src\in \mathbb{N}$ is the identifier of the IE from which the link starts,
-\item $dest\in \mathbb{N}$ is the identifier of the IE at which the link ends.
+\item $src\in \mathbb{N}$ is the identifier of the source IE,
+\item $dest\in \mathbb{N}$ is the identifier of the destination IE.
 \end{itemize}
@@
 
-Similar to IEs, links have identifiers as well. In Fig.~\ref{fig:example-small}, there are five links:
-
-\begin{itemize}
-\item $(7, Dep, 4, 1)$
-\item $(8, PosContr, 5, 2)$
-\item $(9, NegContr, 5, 3)$
-\item $(10, Decomp, 5, 4)$
-\item $(11, Decomp, 6, 4)$
-\end{itemize}
+Similar to IEs, links have identifiers as well. Some of the links in our example (Figure~\ref{fig:example-small}) are $(8, Dep, 4, 1)$, $(9, PosContr, 5, 2)$ and $(11, Decomp, 5, 4)$.
 
@@
 
-\subsubsection{GRL models}
+We can now form GRL models as follows.
 
@@
 
-The definition of a GRL model collects all the previously defined tuples into a single definition.
-For completeness, we now provide the full specification of Figure~\ref{fig:example-small}.
-This model is formalized as $GRL=(IE, Act, R_{ActIE}, Link)$, where:
+The full specification of Figure~\ref{fig:example-small} is as follows.
+
 \begin{flalign*}
@@
 &   &  &(3, Softgoal, \text{Simple design}, AND),&\\
-&   &  &(4, Goal, \text{Generate traffic}, XOR),&\\
-&   &  &(5, Task, \text{Create new cars}, AND),&\\
-&   &  &(6, Task, \text{Keep same cars}, AND)\}&\\
+&   &  &(4, Softgoal, \text{Easy to use}, AND),&\\
+&   &  &(5, Goal, \text{Generate traffic}, XOR),&\\
+&   &  &(6, Task, \text{Create new cars}, AND),&\\
+&   &  &(7, Task, \text{Keep same cars}, AND)\}&\\
 &Act=& &\{(0, Actor, \text{Traffic Simulator})\}&\\
-&R_{ActIE}=& &\{(0, 1), (0,2), (0,3), (0, 4), (0,5),(0,6)\}&\\
-&Link=&\{&(7, Dep, 4, 1)&\\
-&     & &(8, PosContr, 5, 2)&\\
-&     & &(9, NegContr, 5, 3)&\\
-&     & &(10, Decomp, 5, 4)&\\
-&     & &(11, Decomp, 6, 4)\}&\\
+&R_{ActIE}=& &\{(ActIE, 0, 1), \ldots, (ActIE,0,7)\}&\\
+&Link=&\{&(8, Dep, 4, 1)&\\
+&     & &(9, PosContr, 6, 2)&\\
+&     & &(10, NegContr, 6, 3)&\\
+&     & &(11, Decomp, 6, 5)&\\
+&     & &(12, Decomp, 7, 5)\}&\\
 \end{flalign*}
 
-The definition of a GRL model (Def.~\ref{def:grl-model}) only sums up the elements of the model and not the constraints that make a \emph{valid} GRL model, viz.
+Definition \ref{def:grl-model} only sums up the elements of the model and not the constraints that make a \emph{valid} GRL model, viz.
 
@@
 
-Let us briefly verify that our previous formalization of Figure~\ref{fig:example-small} satisfies all the constraints of Def.~\ref{def:valid-grl-model}:
+Let us briefly verify that our previous formalization of Figure~\ref{fig:example-small} satisfies all the constraints of Definition~\ref{def:valid-grl-model}:
 \begin{enumerate}
 \item All elements in the formalization have different ids, so this constraint is satisfied.
-\item $R_{ActIE}$ contains one element for each IE with id $i$, so this constraint is satisfied as well. Note that links are not related to actors. This is in line with the GRL metamodel.
-\item Since we have only one actor with id 0, and this is the only actor that appears in $R_{ActIE}$, this constraint is satisfied.
+\item $R_{ActIE}$ contains one element for each IE with id $i$, so this constraint is satisfied as well. Note that, in line with the GRL metamodel, links are not related to actors. 
+\item There is one actor ($id=0$) and this is the only actor that appears in $R_{ActIE}$ so this constraint is satisfied.
 \item All links connect IEs: the contribution links connect elements with ids 2, 3, and 5, which are all IEs; the decomposition links connect elements with ids 4, 5, and 6, which are all IEs; and the dependency link connects id 1 with 4, which are both IEs.
@@
 
-In order to develop a logical framework for RationalGRL, we extend the GRL logical framework of the previous section by adding two elements (see Figure~\ref{fig:rationalgrllegend}):
-\begin{itemize}
-\item A new element called \emph{generic argument},
-\item A new link called \emph{attack link}
-\end{itemize}
+In order to develop a logical framework for RationalGRL, we extend the GRL logical framework of the previous section by adding two elements (see Figure~\ref{fig:rationalgrllegend}), namely \emph{generic arguments} and \emph{attack links}.
+We illustrate the new elements using the RationalGRL model in Figure~\ref{fig:example-small3}, which is an extension of Figure~\ref{fig:example-small}.
 
-\begin{figure}[ht]
+\begin{figure}[b]
 \centering
 \includegraphics[width=\columnwidth]{img/Example1-new-attack.pdf}
-\caption{Example RationalGRL model (extension of Fig.~\ref{fig:example-small2})}
+\caption{Example RationalGRL model (extension of Fig.~\ref{fig:example-small})}
 \label{fig:example-small3}
@@
 
-We illustrate the new elements using Figure~\ref{fig:example-small3}, which is an extension of Figure~\ref{fig:example-small2}.
-
-We start with the new element which we call the \emph{Generic Argument}.
-
 \begin{definition}[Generic Argument]
 \label{def:generic-argument}
-A generic argument $ga \in \mathbb{N}\times \{GenArg\}\times Names$ is a tuple such that $ga=(id, GenArg, name)$ means
+A generic argument is a tuple $ga=(id, type, name)$, where:
 \begin{itemize}
 \item $id\in \mathbb{N}$ is the identifier of the generic argument,
-\item $GenArg$ states that the tuple is a generic argument,
+\item $type = GenArg$ states that the tuple is a generic argument,
 \item $name\in Names$ is a string description of its name.
 \end{itemize}
-We may refer to the argument with id $i$ simply with $ga_i$. A set of arguments is denoted by $GA$.
+The set of generic arguments is denoted by $AE$.
 \end{definition}
 
-A generic counterargument is simply an argument that can be used to attack any previous argument, or an IE.
-Note that one of the constraints of a GRL model (Def.~\ref{def:grl-model}) is that GRL links (Def.~\ref{def:link}) should connect IEs, which means that in GRL generic arguments cannot be connected with GRL links.
-This is desired, since generic arguments are not part of GRL, so they should also not be connected using GRL links.
+So a generic argument is any element in the RationalGRL model that is not an intentional element or an actor.
+In the example, some of the generic arguments are $(13, GenArg, \text{Redundant})$ and $(14,$ $GenArg,$ Necessary$)$.
+Note that a constraint of a GRL model (Definition~\ref{def:grl-model}) is that GRL links should connect IEs, which means that in generic arguments cannot be connected with GRL links.
 
-In Fig.~\ref{fig:example-small3}, there are two generic counterargument which can be formalized as $(18,GenArg, \text{Redundant})$ and $(19, GenArg, \text{Necessary})$.
+We can now define \emph{arguments}.
 
@@
 
-\paragraph{Rationale and example} We define an argument as either a generic argument or any of the GRL elements or links.
-This captures the specification in the RationalGRL metamodel (Fig.~\ref{fig:metamodel}) in which the class \textsf{Argument} is a superclass of \textsf{GenericArgument} and \textsf{GRLModelElement}, which again is a superclass of both \textsf{ElementLink} and \textsf{GRLLinkableElement}.
-A \textsf{GRLLinkableElement} is in turn a superclass of \textsf{Actor} and \textsf{IntentionalElement}.
+This definition captures the specification in the RationalGRL metamodel (Figure~\ref{fig:metamodel}) in which the class \textsf{Argument} is a superclass of \textsf{GenericArgument} and \textsf{GRLModelElement}.
 In sum, we define an argument simply as any one of the GRL elements or links, or a generic argument.
 
-In Fig.~\ref{fig:example-small3}, the actor, all IEs, all links, and all generic counterarguments are arguments.
-We will give a full formalization of this after the definition of a RationalGRL model (Def.~\ref{def:rationalgrl-model}).
+Examples of arguments in Figure~\ref{fig:example-small3} are $A_5 = (5,$ $Goal,$ Generate traffic$, AND)$, $A'_5 = (5, Goal,$ Generate traffic$, XOR)$, $A_4 = (4, Softgoal,$ Easy to use$, AND)$ and $A_{13}=(13,$ $ArgElem,$ Redundant$)$.
+Note that the two arguments $A_5$ and $A'_5$ show an important difference between RationalGRL models and valid GRL models: While a valid GRL model disallows multiple elements with the same identifier (Definition~\ref{def:valid-grl-model}, condition 1), RationalGRL models do not enforce this restriction.
+This is because it is possible to create multiple arguments for the same element, where each argument contains different content for the same element.
+However, the set of \emph{accepted} elements in a RationalGRL should all have unique identifier (see Definition~\ref{def:valid-rationalgrl-model}).
 
@@
 \label{def:link:attack}
-Given a set of arguments $Arg$, an attack link $att\in Arg\times Arg$ is a relation such that $att=(A_i,A_j)$ means:
+Given a set of arguments $Arg$, an attack link is a relation such that $att=(A_i,A_j)$ means:
 \begin{itemize}
@@
 \end{itemize}
-Intuitively, $att=(A_i,A_j)$ means that argument $A_i$ is attacking argument $A_j$. A set of attack links is denoted by $Att$.
+The set of attack links is denoted by $Att$.
 \end{definition}
 
-\paragraph{Rationale and example} The attack link is the only link that RationalGRL adds to GRL.
-In Fig~\ref{fig:example-small3}, there are three attack links.
-The first two attack links originate from generic counterarguments and involve the following three arguments:
-\begin{itemize}
-\item $A_1 = (1, Resource, \text{Car objects}, AND)$
-\item $A_{18} = (18, GenArg, \text{Redundant})$
-\item $A_{19} = (19, GenArg, \text{Necessary})$
-\end{itemize}
+As an example, take the arguments $A_5$ for the goal (Generate traffic, $AND$), $A'_5$ for the goal (Generate traffic, $XOR$), $A_4$ for the softgoal (Easy to use) and the generic argument $A_{13}$ (Redundant) against the softgoal (Easy to use).
+Given these arguments there are two attack links (Figure~\ref{fig:example-small3}), namely $(A'_{5},A_{5})$ and $(A_{13}, A_{4})$.
 
-These two attack links are formalized as $(A_{18},A_{1})$ and $(A_{19}, A_{18})$.
-
-The third attack link is an attack link created by \emph{replacing} and existing argument (we will discuss this process in more detail in Sect.~\ref{sect:algorithms}).
-It involves two version of the same IE:
-\begin{itemize}
-\item $A_{4} = ((4, Goal, \text{Generate traffic}, AND), Goal)$
-\item $A_{20} = ((4, Goal, \text{Generate traffic}, XOR), Goal)$
-\end{itemize}
-
-This attack link is formalized at $(A_{20},A_4)$.
-
-The last example shows an important difference between RationalGRL models and valid GRL models: While a valid GRL model disallows multiple elements with the same identifier (Def.~\ref{def:valid-grl-model}, condition 1), RationalGRL models do not enforce this restriction.
-This is because it is possible to create multiple arguments for the same element, where argument contains different content for the same element.
-This is what the example above also shows.
-However, the set of \emph{accepted} elements in a RationalGRL should all have unique identifier (see Def.~\ref{def:valid-rationalgrl-model}).
+We now define RationalGRL models.
 
@@
 
-\paragraph{Rationale and example} The definition of a RationalGRL model collects all the previously defined tuples into a single definition.
+The definition of a RationalGRL model collects all the previously defined GRL and RationalGRL elements in a single definition.
 For completeness, we now provide the full specification of Figure~\ref{fig:example-small3}.
 Let us first enumerate all the arguments used in this example:
+
 \begin{flalign*}
@@
 A_1 = &(1, Task, \text{Car objects}, AND),&\\
-A_2 = &(2, Softgoal, \text{Real. sim.}, AND),&\\
-A_3 = &(3, Softgoal, \text{Simple des.}, AND),&\\
-A_4 = &(4, Goal, \text{Generate traffic}, AND),&\\
-A_5 = &(5, Task, \text{Create new cars}, AND),&\\
-A_6 = &(6, Task, \text{Keep same cars}, AND),&\\
-A_7 = &(7, Dep, 4, 1),&\\
-A_8 = &(8, PosContr, 5, 2),&\\
-A_9 = &(9, NegContr, 5, 3),&\\
-A_{10} = &(10, Decomp, 5, 4),&\\
-A_{11} = &(11, Decomp, 6, 4),&\\
-A_i = &(i, ActIE, 0,i+11),ActIE), \text{for } 1\le i\le 6&\\
-A_{18} = &(18, GenArg, \text{Redundant}),&\\
-A_{19} = &(19, GenArg, \text{Necessary}),&\\
-A_{20} = &(4, Goal, \text{Generate traffic}, XOR),&\\
+A_2 = &(2, Softgoal, \text{Realistic simulation}, AND),&\\
+A_3 = &(3, Softgoal, \text{Simple design}, AND),&\\
+A_4 = &(4, Softgoal, \text{Easy to use}, AND),&\\
+A_5 = &(5, Goal, \text{Generate traffic}, AND),&\\
+A'_5 = &(5, Goal, \text{Generate traffic}, XOR),&\\
+A_6 = &(6, Task, \text{Wrap cars}, AND),&\\
+A'_6 = &(6, Task, \text{Create new cars}, AND),&\\
+A_7 = &(7, Task, \text{Keep same cars}, AND),&\\
+A_8 = &(8, Dep, 4, 1),&\\
+A_9 = &(9, PosContr, 6, 2),&\\
+A_{10} = &(10, NegContr, 6, 3),&\\
+A_{11} = &(11, Decomp, 6, 5),&\\
+A_{12} = &(12, Decomp, 7, 5),&\\
+A_{13} = &(13, GenArg, \text{Redundant}),&\\
+A_{14} = &(14, GenArg, \text{Necessary}),&\\
+A_{15} = &(15, GenArg, \text{No effect}),&\\
+A_{16} = &(16, GenArg, \text{Changes traffic flow}),&\\
+A_{17} = &(ActIE,0,1),\ldots,A_{23}=(ActIE,0,7)&\\
+\end{flalign*}
+This model is then formalized as $RatGRL=(Arg, Att)$ where:
+\begin{flalign*}
+Arg = &\{A_0,\ldots,A_5, A'_5,A_6,A'_6,A_7,\ldots,A_{23}\}\\
+Att= &\{(A'_{5},A_5), (A'_{6},A_{6}), (A_{13},A_{4}), (A_{14},A_{13}),\\
+     & (A_{15},A_{9}),(A_{16},A_{15})\}\\
 \end{flalign*}
 
-This model is then formalized as $RationalGRL=(Arg, Att)$ where:
-\begin{flalign*}
-Arg = &\{A_0,A_1,\ldots,A_{20}\}\\
-Att=&\{(A_{18},A_1), (A_{19},A_{18}), (A_{20},A_{4})\}
-\end{flalign*}
-
-All the arguments and the attack relations of this RationalGRL model are shown in Figure~\ref{fig:example-small4}.
+%All the arguments and the attack relations of this RationalGRL model are shown in Figure~\ref{fig:example-small4}.
 Note the arguments for Actor-IE containment (arguments $A_{12}$ to $A_{17}$) have been omitted from this figure for readability.
@@
 
-\begin{figure}[ht]
+%\begin{figure}[ht]
+%\centering
+%\includegraphics[width=\columnwidth]{img/Example1-new-arguments}
+%\caption{Argumentation framework of RationalGRL model from Fig.~\ref{fig:example-small3}}
+%\label{fig:example-small4}
+%\end{figure}
+
+From Figure~\ref{fig:example-small3}, it can be read from that arguments $A_5$, $A_6$, $A_{13}$ and $A_{15}$ are currently disabled (rejected).
+However, we did not yet make precise how exactly this is computed.
+We will do so in the following definitions.
+\begin{figure}[b]
 \centering
-\includegraphics[width=\columnwidth]{img/Example1-new-arguments}
-\caption{Argumentation framework of RationalGRL model from Fig.~\ref{fig:example-small3}}
-\label{fig:example-small4}
+\includegraphics[]{img/Example1-new-arguments.pdf}
+\caption{Example argumentation framework, subset of the RationalGRL model from Figure~\ref{fig:example-small3}.}
+\label{fig:goalmodeling:arg2}
 \end{figure}
+In order to compute when an argument is accepted and when not we use argumentation semantics.  We use the standard approach by Dung~\cite{Dung1995}.
 
-In order to compute when an argument is accepted and when not we use argumentation semantics.
-We use the standard approach here, which is known as \emph{Dung's semantics}.
-The following notions are preliminary.
-
-\begin{definition}[Argumentation Framework~\cite{Dung1995}]
+\begin{definition}[Argumentation Framework]
 \label{def:argumentation-framework}
@@
 
-We see that the definition of an argumentation framework is very close to that of a RationalGRL model. This allows us to use the following results directly.
+Note that the definition of a RationalGRL model (Definition~\ref{def:rationalgrl-model}) is exactly the same as the definition for an argumentation framework.
+This allows us to use the following definitions directly.
 
-\begin{definition}[Attack, conflict-freeness, defense, and admissibility~\cite{Dung1995}] Suppose an argumentation framework $AF=(Arg,Att)$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
+\begin{definition}[Attack, conflict-freeness, defense, admissibility, preferred extension] \label{def:semantics}Suppose an argumentation framework $AF=(Arg,Att)$, two sets of arguments $S\cup S'\subseteq Arg$, and some argument $A\in Arg$.
 We say that
 \begin{itemize}
-\item $S$ \emph{attacks} $A$ if some argument in $S$ attacks $A$,
-\item $S$ \emph{attacks} $S'$ if some argument in $S$ attacks some argument in $S'$,
-\item $S$ is \emph{conflict-free} if it does not attack itself,
-\item $S$ \emph{defends} $A$ if for each $B$ such that $B$ attacks $A$, $S$ attacks $B$,
-\item $S$ is \emph{admissible} if $S$ is conflict-free and defends each argument in it.
+\item $S$ \emph{attacks} $A$ iff some argument in $S$ attacks $A$,
+\item $S$ \emph{attacks} $S'$ iff some argument in $S$ attacks some argument in $S'$,
+\item $S$ is \emph{conflict-free} iff it does not attack itself,
+\item $S$ \emph{defends} $A$ iff for each $B$ such that $B$ attacks $A$, $S$ attacks $B$,
+\item $S$ is \emph{admissible} iff $S$ is conflict-free and defends each argument in it.
+\item $S$ is a \emph{preferred extension} iff it is a maximal (w.r.t. set inclusion) admissible set.
 \end{itemize}
@@
 
-\begin{figure}[ht!]
+Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of our RationalGRL model from Figure~\ref{fig:example-small3}, containing only arguments $A_4, A_{5},A'_{5},A_{13}$, and $A_{14}$.
+In this example, there are five admissible sets: $\{A'_{5}\}$, $\{A_{14}\}$, $\{A'_{5},A_{14}\}$, $\{A_4, A_{14}\}$ and $\{A_4, A'_5, A_{14}\}$.
+In the admissible sets that contain both $A_{4}$ and $A_{14}$, we say that $A_{14}$ defends $A_4$ against its attacker $A_{13}$.
+Sets containing both $A_{5}$ and $A'_{5}$, or both $A_{13}$ and either $A_4$ or $A_{14}$ are not conflict free.
+Sets containing $A_5$ are not admissible, as they do not defend $A_5$.
+Similarly, sets containing $A_4$ but not $A_{14}$ are not admissible as they do not defend themselves against $A_{13}$.
+
+The notion of admissible sets gives rise to various possible extensions of an argumentation framework; in this article, we use the preferred extension.
+In Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_4, A'_5, A_{14}\}$.
+It is possible to have multiple preferred extensions in cases where the argumentation framework contains cycles.
+A simple example of such an argumentation framework is shown in Figure~\ref{fig:goalmodeling:arg3}, where arguments $A_x$ and $A_y$ attack each other.
+There are two preferred extensions, namely $\{A_x\}$ and $\{A_y\}$.
+Although the algorithms in Section~\ref{sect:algorithms} do not generate mutually attacking arguments such as in Figure~\ref{fig:goalmodeling:arg3}, our formal framework does not explicitly forbid attack cycles in RationalGRL models.
+
+\begin{figure}[b]
 \centering
-\begin{tikzpicture}
-        \node[minimum size=1cm] (a19) [argNodeIN] at (-3,0) {$A_{19}$};
-        \node[minimum size=1cm] (a18) [argNodeIN] at (0,0) {$A_{18}$};
-        \node[minimum size=1cm] (a1) [argNodeIN] at (3,0) {$A_1$};
-
-         \path
-    (a19) edge [attackLink] (a18)
-    (a18) edge [attackLink] (a1);
-
-\end{tikzpicture}
-\caption{Example argumentation framework, subset of Fig.~\ref{fig:example-small4}.}
-\label{fig:goalmodeling:arg2}
+\includegraphics[]{img/example-mutual-attack.pdf}
+\caption{Example of a cyclic argumentation framework.}
+\label{fig:goalmodeling:arg3}
 \end{figure}
 
-Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of Figure~\ref{fig:example-small4} containing only arguments $A_1, A_{18}$, and $A_{19}$.
-In this example, there are two admissible sets: $\{A_{19}\}$ and $\{A_1, A_{19}\}$.
-In the last admissible set, $A_{19}$ defends $A_1$ against its attacker $A_{18}$.
-Sets containing both $A_{18}$ and either $A_1$ or $A_{19}$ are not conflict free, and the sets $\{A_1\}$ and $\{A_{18}\}$ do not defend themselves against $A_{18}$ and $A_{19}$, respectively.
+The status of individual arguments can now be determined on the basis of the preferred extensions.
+Recall from the metamodel (Figure~\ref{fig:metamodel}) that an argument can be \emph{acceptable}, \emph{undecided} or \emph{rejected}.
 
-Given the notion of admissible sets, we can then define our argumentation semantics.
-There are a large number of different semantics to determine which arguments are acceptable; in this article, we focused on the preferred semantics.
-
-\begin{definition}[Preferred semantics~\cite{Dung1995}] 
-\label{def:preferred-semantics}
-A preferred extension of an argumentation framework $(Arg,Att)$ is a maximal (w.r.t. set inclusion) admissible set of $R$.
+\begin{definition}[Argument Acceptability] 
+\label{def:acceptability}
+\begin{itemize}
+\item An argument is \emph{acceptable} w.r.t. an argumentation framework $AF$ if it is in every preferred extension of $AF$. 
+\item An argument is \emph{undecided} w.r.t. an argumentation framework $AF$ if it is in at least one but not all preferred extensions of $AF$. 
+\item An argument is \emph{rejected} w.r.t. an argumentation framework $AF$ if it is in no preferred extension of $AF$. 
+\end{itemize}
 \end{definition}
 
-In our example from Figure~\ref{fig:goalmodeling:arg2}, there is one preferred extension, namely $\{A_{19},A_1\}$.
-Returning to our running example of Figure~\ref{fig:example-small3}, this means that the intentional elements \text{Car objects} (task) and the generic counerargument \emph{Necessary} are both \emph{accepted}, while the generic counter argument ``Redundant'' is rejected.
+Take our two simple examples.
+In Figure~\ref{fig:goalmodeling:arg2}, we have that arguments $A_4$, $A'_5$ and $A_{14}$ are justified and arguments $A_5$ and $A_{13}$ are rejected.
+In Figure~\ref{fig:goalmodeling:arg3}, both arguments $A_x$ and $A_y$ are undecided.
+Returning to our larger example from Figure~\ref{fig:example-small3}, we can see that arguments $A_5$, $A_6$, $A_{13}$ and $A_{15}$ are rejected and all the other arguments are acceptable.
 
-\subsection{Translation algorithms}
+\subsection{Translating between RationalGRL and GRL}
 \label{sect:formalframework:translation}
 
-Now that we have formalized both a GRL model and a RationalGRL model, we develop algorithm in order to translate between these two models.
+Now that we have formalized both a GRL model and a RationalGRL model, we present algorithms to translate between these two models.
 Both of these two translation algorithms are straightforward, which is result of the fact that the two models are formalized in a very similar way.
@@
 \paragraph{GRL to RationalGRL Translation} We start with the translation algorithm from GRL to RationalGRL, which is shown in Algorithm~\ref{alg:translation:to-rationalgrl}.
-The translation algorithm takes a GRL model $GR=(IE, Act, R_{ActIE}, Link)$ (Def.~\ref{def:grl-model}) as input.
+The translation algorithm takes a GRL model $GR=(IE, Act, R_{ActIE}, Link)$ (Definition~\ref{def:grl-model}) as input.
 On line~\ref{alg:translation:to-rationalgrl:args}, it collects all the elements of the GRL model into a set $Arg$, which represents the set of arguments of the RationalGRL model.
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[t]
   \caption{GRL to RationalGRL Translation}
@@
 \paragraph{RationalGRL to GRL translation} The translation from a RationalGRL model to a GRL model is given in Algorithm~\ref{alg:translation:to-grl}.
-It is very similar to the previous algorithm, but then works in the other direction.
-In this case, arguments are unpacked and put in the corresponding GRL components.
-On line~\ref{alg:translation:to-grl:extension}, argumentation semantics is used to determine the \emph{preferred extension} (Def.~\ref{def:preferred-semantics}).
-This is a set containing all accepted arguments.
-The GRL model is then generated from all the accepted arguments.
+It is very similar to the previous algorithm, but works in the other direction.
+In this case, the arguments are each put in the corresponding GRL component sets.
+On line~\ref{alg:translation:to-grl:extension}, the procedure $ComputeExtensions(Arg,Att)$ computes the extensions of the RationalGRL model -- in our case, the preferred extensions (Definition~\ref{def:semantics}).
+A GRL model is then generated from each of the preferred extensions.
 This is done by iterating over all arguments in the extension in line~\ref{alg:translation:to-grl:for}.
-The switch statement on line~\ref{alg:translation:to-grl:switch} then does a case distinction on the type of the arguments, and each case ensures the argument is put in the right GRL component.
+The switch statement on line~\ref{alg:translation:to-grl:switch} then does a case distinction on the type of the arguments -- here, $Arg.type$ refers to the $type$ element in an argument tuple $Arg$ -- and each case ensures the argument is put in the right GRL component.
 Finally, the algorithm returns a GRL model on line~\ref{alg:translation:to-grl:return}.
 
-\begin{algorithm}[h]
+\begin{algorithm}[b]
   \caption{RationalGRL to GRL Translation}
@@
     \Procedure{$ToGRL$}{$RatGRL$}
-    \State $E \leftarrow ComputeExtension(Arg,Att)$\label{alg:translation:to-grl:extension}
+    \State $Ext \leftarrow ComputeExtensions(Arg,Att)$\label{alg:translation:to-grl:extension}
+    \For{$E \in Ext$} 
     \State $IE\leftarrow\emptyset$, $Act\leftarrow\emptyset$, $R_{ActIE}\leftarrow\emptyset$, $Link\leftarrow \emptyset$
@@
     \State \Return $(IE,Act,R_{ActIE}, Link)$\label{alg:translation:to-grl:return}
+    \EndFor
     \EndProcedure
@@
 
-\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Def.~\ref{def:valid-grl-model}), we have no  done so for a RationalGRL model yet.
+\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Definition~\ref{def:valid-grl-model}), we have not done so for a RationalGRL model yet.
 This is because it depends on the translation procedure.
@@
 
-In the previous subsection we formalized a \emph{static} representation of the RationalGRL framework.
-In this section we formalize the \emph{dynamics}.
-We do so by developing algorithms for applying argument schemes and critical questions in the context of a RationalGRL model (Def.~\ref{def:rationalgrl-model}).
-These algorithms are applied to RationalGRL models and produce new arguments and attack relations.
-We can then use argumentation semantics (Def.~\ref{def:preferred-semantics}) to compute sets of accepted arguments.
-The content of these arguments is then used to compute the resulting RationalGRL model, together with enabled and disabled GRL elements and their underlying arguments.
+We have now formalized a \emph{static} representation of the RationalGRL framework.
+In this section we formalize the \emph{dynamics} by developing algorithms for applying argument schemes and critical questions.
+These algorithms are applied to RationalGRL models (Definition~\ref{def:rationalgrl-model}) and produce new arguments and attack relations.
 
-As discussed in Section~\ref{sect:overview}, the argument schemes and critical questions of Table~\ref{table:argument-schemes} fall into one of the following four categories:
+As discussed in Section~\ref{sect:overview}, the argument schemes and critical questions of Table~\ref{table:argument-schemes} all lead to one of three operations: \textsf{INTRO}
+introduces a new RationalGRL element and \textsf{DISABLE} creates a new argument that attacks another argument.
+The \textsf{REPLACE} operation is can be seen as a combination of \emph{INTRO} and \emph{DISABLE}: a new argument corresponding to a GRL element is added and this new argument attacks a previous version of the element or link.
+We will discuss all three operations in separate sections below.
+
+In all of the following algorithms, we assume the following.
 \begin{itemize}
-\item \emph{INTRO}: These algorithms add one or multiple new arguments to the RationalGRL model, without creating any attack links. The arguments are for new GRL elements or links.
-\item \emph{DISABLE}: These algorithms add a single argument to the RationalGRL model, which attacks one of multiple existing arguments.
-The counterargument that is added does not correspond to any GRL element, but instead disables an existing one by attacking all arguments for it.
-\item \emph{REPLACE}: These algorithms can be seen as a combination of \emph{INTRO} and \emph{DISABLE}.
-They add a new argument corresponding to a GRL element or link, and this new argument attacks a previous version of the element or link.
-\end{itemize}
-
-The three subsections of this section correspond to these three types of arguments.
-
-In all of the following algorithms, we assume the following
-\begin{itemize}
-\item There algorithms are being applied to some valid RationalGRL model $RatGRL$ (Def.~\ref{def:valid-rationalgrl-model}),
+\item The algorithms are applied to some valid RationalGRL model $RatGRL$ (Def.~\ref{def:valid-rationalgrl-model}),
 \item The procedure $mindId()$ generates a new unique id.
@@
 These type of algorithms are short, and consist simply of adding an argument for the element that is being added.
-Since the structure of these algorithms are all very similar, we do not discuss them all.
 
@@
 
-\paragraph{Rationale and example} Algorithm~\ref{alg:add-actor} takes one argument, namely the name of the actor $a$.
+Algorithm~\ref{alg:add-actor} takes one argument, namely the name of the actor $a$.
 On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$.
 On line~\ref{alg:as0:2} the actor is added as a new argument.
-
-As an example, we can formalize the addition of actor \emph{Traffic Simulator} in Figure~\ref{fig:example-small2} as executing the algorithm $AS_0($Traffic simulator), which results in a RationalGRL model $RatGRL=(Arg,Att)$ with argument $(i, Actor, \text{Traffic simulator})\in Arg$.
+As an example, we can formalize the addition of actor \emph{Traffic Simulator} to a RationalGRL model as executing the algorithm $AS_0$(Traffic simulator), which results in a RationalGRL model $RatGRL=(Arg,Att)$ with argument $A_0 = (0, Actor, \text{Traffic simulator})\in Arg$.
 
@@
     \State $A_1\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-resource:arg1}
-    \State $A_2\leftarrow (mintId(),ActIE, i, res\_id)$\label{alg:add-resource:arg2}
+    \State $A_2\leftarrow (ActIE, i, res\_id)$\label{alg:add-resource:arg2}
     \State $Arg\gets Arg\cup \{A_1,A_2\}$\label{alg:add-resource:add-arg}
@@
 
-\paragraph{Rationale and example} The original argument schemes AS1 is phrased as ``Actor $a$ has resource $R$''.
-We have slightly reworded this in Algorithm~\ref{alg:add-actor} to ``Actor with id $i$ has resource $R$''.
-The reason is that in GRL, actors have identifiers, so since we assume actor $a$ exists already, we also assume an identifier for it exists.
-It would be straightforward to introduce a mapping from actor names to identifiers, but we have no done so here for simplicity.
+In Algorithm~\ref{alg:add-actor}, we have slightly reworded the original AS1 from Table~\ref{table:argument-schemes} to ``Actor with id $i$ has resource $R$''.
+Furthermore, we assume actor $a$ already exists when Algorithm~\ref{alg:add-actor} is run.
 
@@
 
-As an example, consider the addition of resource ``Car objects'' in Figure~\ref{fig:example-small2}, and suppose the actor is formalized as the argument $(i, Actor, \text{Traffic simulator})$.
-Adding the resource can then be formalized as executing $AS1(i, \text{Car objects})$, resulting in the addition of two arguments $(j, Resource, \text{Car object}, AND)$ and $(k, ActIE, i, j)$ which are added to the set of arguments $Arg$.
+As an example, consider the addition of resource ``Car objects'' to a RationalGRL model, and suppose the actor is formalized as the argument $A_0 = (0, Actor, \text{Traffic simulator})$.
+Adding the resource can then be formalized as executing $AS1(0, \text{Car objects})$, resulting in the addition of two arguments $(1, Resource, \text{Car object}, AND)$ and $(ActIE, 0, 1)$ which are added to the set of arguments $Arg$.
 
-Since arguments schemes AS2-AS4 are very similar to AS1, we have omitted algorithms for them here.
+Since arguments schemes AS2-AS4 are very similar to AS1, we omit these algorithms here.
 
@@
 
-\paragraph{Rationale and example} Similar to the previous algorithm, we have slightly reworded critical question AS5.
+Similar to the previous algorithm, we have slightly reworded critical question AS5.
 We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $T$ is a decomposition of $G$.
@@
 
-As an example, suppose we are adding the fact that goal ``Generate traffic'' is decomposing into task ``Keep same cars'' (Figure~\ref{fig:example-small2}), and suppose the following argument for the goal exists already: $(i, Goal,\text{Generate traffic}, AND)$.
-Adding the decomposition can be formalized as executing $AS_5(i, \text{Keep same cars})$, and results in two new arguments: $(task\_id, Task, \text{Keep same cars}, AND)$ and $(k, Decomp, i, task\_id)$, which are both added to the set of arguments $Arg$.
+As an example, suppose we are adding the fact that goal ``Generate traffic'' is decomposing into task ``Keep same cars'', and suppose we have the following argument for the goal: $A_5 = (5,$ $Goal,$ Generate traffic$, AND)$.
+Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars}, AND)$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.
 
@@
 
-The critical questions of type \emph{INTRO} are very similar as well, with one exception: the require an answer.
+The critical questions of type \emph{INTRO} are very similar as well, with one exception: they require an answer.
 For instance, suppose CQ5b: ``Does goal $G$ decompose into other tasks?'' is answered with: ``Yes, namely into task $T$''.
 In this case, we simply obtain an instantiation of argument scheme AS5: ``Goal $G$ decomposes into task $T$'', which can be executed with Algorithm~\ref{alg:add-decomp}.
-This is the same for all the other critical questions of type \emph{INTRO} as well.
+This is the same for all the other critical questions of type \emph{INTRO}.
 Therefore, we have omitted them here as well.
@@
 
-As discussed before, algorithms of type \emph{DISABLE} consist of adding a new argument attacking an existing argument, which is an argument for an existing GRL element or link.
-The argument that is added is itself not an argument for a GRL element or link.
+As discussed before, algorithms of type \emph{DISABLE} consist of adding a new argument attacking an existing argument.
+The argument that is added is itself not an argument for a GRL element or link, but rather a generic argument (Definition~\ref{def:generic-argument}).
 
-In all of these algorithms, we assume the critical question is answered affirmatively.
-For instance, for critical question CQ0 ``Is the actor relevant?'', we assume it is answered with ``No'' (see right-most column of Table~\ref{table:argument-schemes}).
+In all of these algorithms, we assume the critical question is answered as indicated in the right-most column of Table~\ref{table:argument-schemes}.
+For instance, for critical question CQ0 ``Is the actor relevant?'', we assume it is answered with ``No''.
 
@@
 
-\paragraph{Rationale and example} Algorithm~\ref{alg:actor-not-relevant} is executed when critical question CQ0 is answered affirmatively (i.e., with ``No'').
+Algorithm~\ref{alg:actor-not-relevant} is executed when critical question CQ0 is answered affirmatively (i.e., with ``No'').
 First, on lines~\ref{alg:actor-not-relevant:genarg} and~\ref{alg:actor-not-relevant:genarg2}, an argument is created for the critical question and added to the set of arguments $Arg$.
@@
 
-\paragraph{Rationale and example} Algorithm~\ref{alg:no-decomp} is structurally very similar to Algorithm~\ref{alg:actor-not-relevant}, with the only difference that instead of iterating over actors, we now iterate over decomposition links.
+Algorithm~\ref{alg:no-decomp} is structurally very similar to Algorithm~\ref{alg:actor-not-relevant}, with the only difference that instead of iterating over actors, we now iterate over decomposition links.
 This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links with source id $g\_id$ and target id $t\_id$ (Def.~\ref{def:link}).
@@
 
-\paragraph{Rationale and example} Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counter-arguments to arguments.
+Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counter-arguments to arguments.
 In all of the previous \emph{DISABLE} algorithm, the attack was on a specific type of argument, for instance an argument for an actor or an argument for a decomposition.
@@
 
-Let us reconsider Figure~\ref{fig:examples:relevant-actor}, which we previously formalized as $RatGRL=(Arg,Att)$, where $Arg=\{A_1,A_2\}$, $A_1=\{(0,Actor,\text{Development team}\}), A_2=(1, GenArg, \text{CQ0: Professor develops software})$, and $Att=(A_2,A_1)$.
+Let us reconsider Figure~\ref{fig:examples:relevant-actor}, which we previously formalized as $RatGRL=(Arg,Att)$, where $Arg=\{A_1,A_2\}$, $A_1 = \{(0,Actor,\text{Development team}\})$, $A_2= (1,$ $GenArg,$ CQ0: Professor develops software$)$ and $Att=(A_2,A_1)$.
 Suppose we execute algorithm $Att(\{A_2\}, \text{Professor does not develop software})$.
-This results in $A_3=(2, GenArg, \text{CQ0: Professor does not develop software})$ and a new attack link $(A_3,A_2)$ (Figure~\ref{fig:examples:relevant-actor2}).
+This results in $A_3=(2, GenArg,$ CQ0: Professor does not develop software$)$ and a new attack link $(A_3,A_2)$ (Figure~\ref{fig:examples:relevant-actor2}).
 Since the argument for the goal is now part of the preferred extension, it does show up in the resulting GRL model after the translation of Algorithm~\ref{alg:translation:to-grl}.
@@
 
-\paragraph{Rationale and example} While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm $CQ_{Decomp}$~\ref{alg:replace1} is more generally applicable to the decomposition type of any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}).
+While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm $CQ_{Decomp}$~\ref{alg:replace1} is more generally applicable to the decomposition type of any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}).
 

### 2017-11-01 overview.tex
formalframework cleanup & new example

@@
 
-A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual types of reasoning stakeholders or designers perform (requirement 1 of our framework).
+A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual types of reasoning stakeholders perform (requirement 1).
 Recall from section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
 Using this list as a basis, we further refined the set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
-Note that this list is not exhaustive, and that new argument schemes and critical questions can be added depending on the problem domain.
+Note that this list is not exhaustive, and that new argument schemes and questions can be added depending on the problem domain.
 

### 2017-11-03 formalframework.tex
tool section and small changes

@@
 \paragraph{RationalGRL to GRL translation} The translation from a RationalGRL model to a GRL model is given in Algorithm~\ref{alg:translation:to-grl}.
-It is very similar to the previous algorithm, but works in the other direction.
-In this case, the arguments are each put in the corresponding GRL component sets.
-On line~\ref{alg:translation:to-grl:extension}, the procedure $ComputeExtensions(Arg,Att)$ computes the extensions of the RationalGRL model -- in our case, the preferred extensions (Definition~\ref{def:semantics}).
+First.
+the arguments are each put in the corresponding GRL component sets.
+The procedure $ComputeExtensions(Arg,Att)$ (line~\ref{alg:translation:to-grl:extension}) computes the (preferred) extensions of the RationalGRL model (Definition~\ref{def:semantics}).
 A GRL model is then generated from each of the preferred extensions.
@@
 Finally, the algorithm returns a GRL model on line~\ref{alg:translation:to-grl:return}.
+As an example, compare the RationalGRL model in Figure~\ref{fig:example-small3} to the GRL model in Figure~\ref{fig:example-small}, where the latter is a translation of the former.
+
+\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Definition~\ref{def:valid-grl-model}), we have not done so for a RationalGRL model yet.
+We define a RationalGRL model as valid if and only if the RationalGRL to GRL translation results in a valid GRL model.
+Thus, we do not have to reiterate all conditions on a GRL model, but use the translation algorithm.
 
@@
 
-\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Definition~\ref{def:valid-grl-model}), we have not done so for a RationalGRL model yet.
-This is because it depends on the translation procedure.
-We define a RationalGRL model as valid if and only if the RationalGRL to GRL translation results in a valid GRL model.
-In this way, we do not have to reiterate all conditions on a GRL model, but use the translation algorithm.
-
 \begin{definition}[Valid RationalGRL Model]
@@
 
-In all of the following algorithms, we assume the following.
+In all of the following algorithms, we assume that:
 \begin{itemize}
-\item The algorithms are applied to some valid RationalGRL model $RatGRL$ (Def.~\ref{def:valid-rationalgrl-model}),
-\item The procedure $mindId()$ generates a new unique id.
+\item The algorithms are applied to some valid RationalGRL model $RatGRL$ (Definition~\ref{def:valid-rationalgrl-model}),
+\item The procedure $mintId()$ generates a new unique id.
 \end{itemize}
@@
 \subsubsection{INTRO algorithms}
+\label{sect:formalframework:intro}
 
@@
 On line~\ref{alg:as0:2} the actor is added as a new argument.
-As an example, we can formalize the addition of actor \emph{Traffic Simulator} to a RationalGRL model as executing the algorithm $AS_0$(Traffic simulator), which results in a RationalGRL model $RatGRL=(Arg,Att)$ with argument $A_0 = (0, Actor, \text{Traffic simulator})\in Arg$.
+As an example, we can formalize the addition of actor ``Traffic Simulator'' to an empty RationalGRL model (i.e.
+$Arg=\emptyset$ and $Att=\emptyset$)as executing the algorithm $AS_0$(Traffic simulator), which results in a RationalGRL model with argument $A_0 = (0, Actor, \text{Traffic simulator})\in Arg$.
 
@@
 
-In Algorithm~\ref{alg:add-actor}, we have slightly reworded the original AS1 from Table~\ref{table:argument-schemes} to ``Actor with id $i$ has resource $R$''.
+In Algorithm~\ref{alg:add-resource}, we have slightly reworded the original AS1 from Table~\ref{table:argument-schemes} to ``Actor with id $i$ has resource $R$''.
 Furthermore, we assume actor $a$ already exists when Algorithm~\ref{alg:add-actor} is run.
-
 The algorithm itself runs as follows: on line~\ref{alg:add-resource:id}, a unique id is assigned to variable $res\_id$ (``resource identifier'').
@@
 
-As an example, consider the addition of resource ``Car objects'' to a RationalGRL model, and suppose the actor is formalized as the argument $A_0 = (0, Actor, \text{Traffic simulator})$.
+As an example, consider the addition of resource ``Car objects'' to the RationalGRL model we just discussed (where $A_0 = (0, Actor, \text{Traffic simulator})$).
 Adding the resource can then be formalized as executing $AS1(0, \text{Car objects})$, resulting in the addition of two arguments $(1, Resource, \text{Car object}, AND)$ and $(ActIE, 0, 1)$ which are added to the set of arguments $Arg$.
@@
 
-Similar to the previous algorithm, we have slightly reworded critical question AS5.
+Similar to the previous algorithm, we have slightly reworded critical question AS5 in Algoritm~\ref{alg:add-decomp}.
 We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $T$ is a decomposition of $G$.
-
-In Algorithm~\ref{alg:add-decomp}, on line~\ref{alg:add-decomp:task-id} a unique identifier is created for the task, which is created on line~\ref{alg:add-decomp:arg1}.
+In the algorithm, on line~\ref{alg:add-decomp:task-id} a unique identifier is created for the task, which is created on line~\ref{alg:add-decomp:arg1}.
 On line~\ref{alg:add-decomp:arg2} an argument is created for the decomposition link $(mintId(), Decomp, i, task\_id)$ (Def.~\ref{def:link}, going from the existing goal with identifier $i$ to the new task with identifier $task\_id$.
@@
 
-As an example, suppose we are adding the fact that goal ``Generate traffic'' is decomposing into task ``Keep same cars'', and suppose we have the following argument for the goal: $A_5 = (5,$ $Goal,$ Generate traffic$, AND)$.
+As an example, suppose we are adding the decomposition of goal ``Generate traffic'', expressed as argument $A_5 = (5,$ $Goal,$ Generate traffic$, AND)$, into task ``Keep same cars'', and suppose we have the following argument for the goal: .
 Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars}, AND)$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.
 
-The remaining argument schemes AS6-AS12 are all very similar to the previous algorithms and have been omitted here.
+The remaining argument schemes AS6-AS12 are all similar to algorithms~\ref{alg:add-actor}-\ref{alg:add-decomp} and have been omitted here.
 
@@
 \subsubsection{DISABLE algorithms}
-
+\label{sect:formalframework:disable}
 As discussed before, algorithms of type \emph{DISABLE} consist of adding a new argument attacking an existing argument.
@@
 Consider for example the RationalGRL model in Figures~\ref{fig:examples:relevant-actor}, which consists of an actor and a generic counterargument.
-Let us reconstruct this model using the application of argument schemes and critical questions.
-Initially, the RationalGRL model $(Arg,Att)$ is empty, i.e.
-$Arg=Att=\emptyset$.
-Next, the algorithm $AS_0(\text{Development team})$ is executed, which results in $A_1=\{(0,Actor,\text{Development team}\})$.
+Let us reconstruct this model using the application of argument schemes and critical questions to an initially empty RationalGRL model ($Arg=\emptyset$ and $Att=\emptyset$).
+The algorithm $AS_0(\text{Development team})$ is executed, which results in $A_1=\{(0,Actor,\text{Development team}\})$.
 After this, critical question CQ0: ``Is the actor Development team relevant?'' is answered with ``No, because the professor develops the software''.
 We can formalize this as executing algorithm $CQ_0(0)$ (Algorithm~\ref{alg:actor-not-relevant}), which results in adding a generic counter argument $A_2=(1, GenArg, \text{CQ0: Professor develops software})$ and an attack link $(A_2,A_1)$.
-Note that the RationalGRL model now consists of two arguments, and that the preferred extension is $\{A_2\}$, which means that the resulting GRL model is empty, because generic arguments are not translated to GRL.
+The RationalGRL model now consists of two arguments, and that the preferred extension is $\{A_2\}$, which means that the resulting GRL model is empty, because generic arguments are not translated to GRL.
 
@@
 Algorithm~\ref{alg:no-decomp} is structurally very similar to Algorithm~\ref{alg:actor-not-relevant}, with the only difference that instead of iterating over actors, we now iterate over decomposition links.
-This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links with source id $g\_id$ and target id $t\_id$ (Def.~\ref{def:link}).
+This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links with source id $g\_id$ and target id $t\_id$ (Definition~\ref{def:link}).
 This means we iterate over all decomposition links from the goal to the task, and we attack all arguments for these links on line~\ref{alg:no-decomp:att} with the newly created argument.
@@
 However, the resulting algorithm is very similar to the previous two and has therefore been omitted.
-The only algorithm of type \emph{DISABLE} that we still discuss is $Att$.
+The only other algorithm of type \emph{DISABLE} that we discuss is $Att$.
 
-\begin{algorithm}[h]
+\begin{algorithm}[b]
   \caption{Att: Generic counter-argument on arguments $A_1,\ldots,A_n$ with name $N$}\label{alg:generic}
@@
 
-Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counter-arguments to arguments.
+Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counterarguments to arguments.
 In all of the previous \emph{DISABLE} algorithm, the attack was on a specific type of argument, for instance an argument for an actor or an argument for a decomposition.
@@
 
-Let us reconsider Figure~\ref{fig:examples:relevant-actor}, which we previously formalized as $RatGRL=(Arg,Att)$, where $Arg=\{A_1,A_2\}$, $A_1 = \{(0,Actor,\text{Development team}\})$, $A_2= (1,$ $GenArg,$ CQ0: Professor develops software$)$ and $Att=(A_2,A_1)$.
-Suppose we execute algorithm $Att(\{A_2\}, \text{Professor does not develop software})$.
+Let us reconsider Figure~\ref{fig:examples:relevant-actor}, which we previously formalized as a RationalGRL model where $Arg=\{A_1,A_2\}$, $A_1 = \{(0,Actor,\text{Development team}\})$, $A_2= (1,$ $GenArg,$ CQ0: Professor develops software$)$ and $Att=(A_2,A_1)$.
+Suppose we execute algorithm $Att(\{A_2\},$ Professor does not develop software$)$.
 This results in $A_3=(2, GenArg,$ CQ0: Professor does not develop software$)$ and a new attack link $(A_3,A_2)$ (Figure~\ref{fig:examples:relevant-actor2}).
@@
 
+\begin{figure*}[b]
+\centering
+\includegraphics{img/methodology.pdf}
+\caption{The RationalGRL Methodology}
+\label{fig:rationalgrl-methodology}
+\end{figure*}
+
 \subsubsection{REPLACE algorithms}
+\label{sect:formalframework:replace}
 
-Recall that the \emph{REPLACE} algorithms both attacks all arguments for an existing element, and at the same time create a new argument that contains a different version of the element being attacked.
+Recall that \emph{REPLACE} algorithms create a new argument that attacks all arguments for an existing element.
 
@@
 
-While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm $CQ_{Decomp}$~\ref{alg:replace1} is more generally applicable to the decomposition type of any IE, since all IEs have a decomposition type in their definition (Def.~\ref{def:ie}).
+While the original critical question CQ5c is specific to the decomposition between a goal and a task, Algorithm~\ref{alg:replace1} is more generally applicable to the decomposition type of any IE, since all IEs have a decomposition type (Definition~\ref{def:ie}).
 
@@
 On line~\ref{alg:replace1:arg}, a new argument $A$ is created which is identical to original IE.
-Recall from Def.~\ref{def:notation} that we may refer to elements of an IE with the ``.'' notation.
-This is what we do on line~\ref{alg:replace1:decompchange}, where the decomposition type of the argument is changed to $X$.
-On line~\ref{alg:replace1:ieargs}, the set $IEArgs$ is assigned with all existing arguments for the input IE (recall the notation of $IE_i$ as the set of all IEs with id $i$, see Def.~\ref{def:notation}).
+On line~\ref{alg:replace1:decompchange} the decomposition type of the argument is changed to $X$ -- here, $Arg.decomptype$ refers to the $decomptype$ element in an argument tuple $Arg$.
+On line~\ref{alg:replace1:ieargs}, the set $IEArgs$ is assigned with all existing arguments for the input IE -- here $IE_i$ is the set of all IEs with id $i$.
 The first two for loops on respectively lines~\ref{alg:replace1:for1} and~\ref{alg:replace1:for2} ensure that all attack links that existing from and to the previous versions of the IE are also carried over to the new argument $A$.
@@
 
-Let us explain this algorithm with the example in Figure~\ref{fig:examples:decomposition}.
+As an example, take Figure~\ref{fig:examples:decomposition}.
 The initial RationalGRL model (before applying CQ10b), can be formalized as follows: $RatGRL=(Arg,Att)$, with $Arg=\{A_1,A_2,A_3,A_4,A_5\}$ and $Att=\emptyset$, where:
@@
 \item $A_3=(2,Task,\text{Static simulation},AND)$
-\item $A_4=(3,Decomp,0,1)$
-\item $A_5=(4,Decomp,0,2)$
+\item $A_4=(3,Decomp,1,0)$
+\item $A_5=(4,Decomp,2,0)$
 \end{itemize}
 
-Suppose algorithm $CQ_{Decomp}(0, OR)$ is executing for this RationalGRL model.
+Suppose algorithm $CQ_{Decomp}(0, XOR)$ is executed for this RationalGRL model.
 On lines~\ref{alg:replace1:arg} and \ref{alg:replace1:decompchange} a new argument $A_6=(0,Goal,\text{Simulate traffic},OR)$ is created.
@@
 
-The other \emph{REPLACE} algorithms are very similar to Algorithm~\ref{alg:replace1}.
-It can be used directly for CQ10c, but for CQ12 we should make a small modification.
-Instead of replacing the decomposition type of the IE, we should replace its name.
+The other \emph{REPLACE} algorithms are similar to Algorithm~\ref{alg:replace1}, which can be used directly for CQ10c.
+For CQ12 we should make a small modification: instead of replacing the decomposition type of the IE, we replace its name.
 Since this is a very minor modification we have omitted it here.

### 2017-11-03 methodology.tex
tool section and small changes

@@
 In previous sections, we have shown how the RationalGRL framework can capture stakeholder discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
-In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a methodology\footnote{The methodology presented here was first presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.} (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
+In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a methodology (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
 
@@
 
-We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a (Rational)GRL model.
+We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a (Rational)GRL model, an earlier version of which was presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.
 %Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
 The rest of the steps are as follows:
-
-\begin{figure*}[ht]
-\centering
-\includegraphics{img/methodology.pdf}
-\caption{The RationalGRL Methodology}
-\label{fig:rationalgrl-methodology}
-\end{figure*}
 
@@
 
-\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of three operations are performed on the GRL model: \textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE}.
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the \textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE} are performed on the GRL model.
 Any of these operations impact the arguments and corresponding GRL intentional elements, modifying the initial GRL model into a RationalGRL model.
@@
 
-\textbf{(4) Modify GRL Models} -- In this step, we modify the regular GRL model based on the RationalGRL model of step (3).
-That is, one of the following situation can happen with respect to the initial GRL model: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
+\textbf{(4) Modify GRL Models} -- Based on the RationalGRL model of step (3), the GRL model can be modified: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
 This results in a new, modified GRL model, which can be used as the basis for another cycle of the methodology.
@@
 
+\begin{figure*}[t]
+\centering
+\includegraphics[scale=0.8]{img/tool/goal_details}
+\caption{Overview of the RationalGRL tool}
+\label{fig:tool:overview}
+\end{figure*}
+
 \subsection{The RationalGRL Tool}
@@
 
-An important requirement of our framework is that it has tool support.
+An important requirement of our framework is that it has tool support (\textbf{requirement 5}.
 This is for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, there does not yet exist any tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent verification to ensure our formal framework and algorithms actually work, iv) having a light-weight web-based version of GRL is useful for the community in general.
+In this section we briefly highlight some features of the tool, and we explain some of its current limitations.
 
-GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}.
-This tool is an extension to Eclipse.
-Although it is a rich tool with many features, it is implemented as an Eclipse plugin.
-We have experienced on various occasions that setting this up properly can take up a significant time of experiments, leaving less time for actual modeling.
+GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}, which is implemented as an Eclipse plugin.
+Although it is a rich tool with many features, it can take significant time to set it up properly and is quite feature-heavy.
+Therefore, we have decided to implement our own lightweight modelling tool as our proof-of-concept.
+The tool can be accessed from:
 
-Our formalization of GRL is very much in line with the way in which GRL models are represented in the open-source Eclipse-based tool jUCMNav.\footnote{See \url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}} By keeping our formalization in line with this tool, we simplify the translation step from models in the RationalGRL tool to models in the jUCMNav tool.
-
-In this section we present our web-based prototype of RationalGRL. It can be accessed from:
 \begin{quote}
@@
 
-The tool is written in Javascript, and runs on all modern browsers. The tool is open-source, and the source can be accessed from:
-\begin{quote}
-\url{http://www.github.com/marcvanzee/RationalGRL}
-\end{quote}
+The RationalGRL tool is an open-source\footnote{\url{http://www.github.com/marcvanzee/RationalGRL}}, web-based Javascript application, which runs on all modern browsers.
+The tool is based on our formalization in Section~\ref{sect:formalframework}.
+Because this formalization is very close to the GRL metamodel, the models can be easily exported to jUCMNav.
 
-In this section we briefly highlight some features of the language, and we explain some of its current limitations.
+When the tool starts, the user is presented with a screen as in Figure~\ref{fig:tool:overview}.
+This screen shows the palette of elements and links to the left, a canvas on which RationalGRL models can be built, and a pane with details of the currently selected elements on the canvas.
+Elements and links can be added to the canvas by selecting them on the left and clicking on the canvas, thus capturing the \textsf{INTRO} operation (Section~\ref{sect:formalframework:intro}).
+This makes it possible to build and modify regular GRL models using the tool.
 
-\todo{all}{marc}{Explain figures in Appendix~\ref{sect:tool-screenshots}}
+The details panel can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal element ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
+Clicking the ``Answer'' button next to a critical question allows the user to answer the question.
+As an example, consider Figure~\ref{fig:tool:cqdetails}, where the softgoal ``Easy to use'' is questioned with the relevancy question (CQ11).
+It is possible to select the answer and provide an explanation for the answer.
+
+Assume that in the example of Figure~\ref{fig:tool:cqdetails} the user selects ``No'' and clicks the ``Answer question'' button.
+A new argument is then automatically created that attacks the softgoal and the details pane shows the critical question as answered (Figure~\ref{fig:tool:cqeffect}).
+This is an implementation of a \textsf{DISABLE} algorithm (Section~\ref{sect:formalframework:disable}) similar to Algorithm~\ref{alg:actor-not-relevant}: a new argument is added that attacks the existing argument.
+It is also possible to attack arguments by adding an ``Argument'' element and an ``Attack'' link manually.
+Consider, for example, Figure~\ref{fig:tool:argument}.
+Here, a new argument ``Necessary'', which attacks the previously generated argument based on CQ11, has been added by the user.
+As the details pane shows, this new argument is not based on a CQ.
+It is further worth noting that it is possible to provide further a explanation in argument elements, allowing for more fine-grained rationalizations.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{img/tool/details_softgoal}
+\caption{Critical question details pane for ``Is the softgoal relevant/useful?''}
+\label{fig:tool:cqdetails}
+\end{figure}
+
+\begin{figure}[b]
+\centering
+\includegraphics[width=\columnwidth]{img/tool/attack_softgoal}
+\caption{Effect of answering CQ11: ``Is the softgoal relevant/useful?'' with ``No''}
+\label{fig:tool:cqeffect}
+\end{figure}
+
+The RationalGRL tool computes the acceptability of arguments on the fly.
+In the example of Figure~\ref{fig:tool:cqeffect}, the original (argument for) softgoal ``Easy to use'' is rejected because it's only attacker ``CQ11 - Redundant'' is accepted.
+However, if we then attack ``CQ11 - Redundant'' with a new argument ``Necessary'', the original argument for ``Easy to use'' is again accepted because its only attacker is rejected  (Figure~\ref{fig:tool:argument}).
+Note that when computing the acceptability of arguments, the RationalGRL tool makes the assumption that there are no attack cycles in the model and that hence there is one unique preferred extension (cf.
+Section~\ref{sect:formalframework:rationalgrl}) -- if the user creates an attack cycle, an error message is shown.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{img/tool/reinstate_softgoal}
+\caption{Attacking an argument with a generic argument ``Necessary''}
+\label{fig:tool:argument}
+\end{figure}
+
+The final operation that has been captured in the RationalGRL tool is \textsf{REPLACE}.
+For usability purposes, this operation has been implemented differently than in our formal framework (Section~\ref{sect:formalframework:disable}).
+In the formal framework, a \textsf{REPLACE} introduces a new argument that replaces (and therefore attacks) all previous arguments with the same identifier.
+Including \textsf{REPLACE} like this in the tool would mean that, when the name of an element is changed, we would need to render all the previous versions of the element on the canvas of the tool.
+Because this would quickly become very cumbersome, we decided to implement a naming history for each element.
+For example, in Figure~\ref{fig:tool:overview}, the goal ``Generate cars'' was previously named ``Add cars''.
+In this way, a history is kept of how an element name was clarified without cluttering the model canvas.
+Similarly, the decomposition type can be changed in the details pane (Figure~\ref{fig:tool:overview}) without introducing new elements that attack the original ones (cf.
+Figure~\ref{fig:example-small3}, where the old decomposition type $AND$ of ``Generate traffic'' is shown as an explicit argument).
+
+The tool further implements the RationalGRL to GRL translation (Algorithm~\ref{alg:translation:to-grl}) through an export function.
+RationalGRL models built in the RationalGRL tool can be exported to the \texttt{.grl} file format, which can be imported in jUCMNav.
+Two examples of this export are provided in Appendix~\ref{sect:tool-screenshots}.
+Figure~\ref{fig:tool:figfrompaper} shows the model that was discussed earlier in this paper (Figure~\ref{fig:example-small3}) in the RationalGRL tool.
+Recall that translating this model to GRL provided us with the model in Figure~\ref{fig:example-small}.
+This is also what follows from our export: if we export the RationalGRL model and then import the resulting GRL model in jUCMNav, we get the model in Figure~\ref{fig:tool:figfrompaper1}, which is the same as Figure~\ref{fig:example-small}.
+
+Our translation and export function uses the argument acceptability as a way of determining the GRL model.
+Take for example, the RationalGRL model in Figure~\ref{fig:tool:multipleattack}.
+A positive contribution is added from ``Keep same cars'' to ``Simple design''.
+More inportantly, an argument ``not enough CPU'' has been added to the model.
+This argument attacks the ``Realistic simulation'' softgoal and the ``Create new cars'' task, arguing that there is not enough processing power for either of these to be feasible in the traffic simulator.
+Now, if we export this to GRL, the GRL elements that are rejected or disabled (greyed out in Figure~\ref{fig:tool:multipleattack}) are not included.
+This can be seen in Figure~\ref{fig:tool:multipleattack1}, where the jUCMNav GRL model that was exported from the RationalGRL tool is shown.
+The pair of figures \ref{fig:tool:multipleattack} and \ref{fig:tool:multipleattack1} also nicely shows the added value of RationalGRL: Figure~\ref{fig:tool:multipleattack} shows that there can be a larger discussion and rationalization underlying even a fairly simple goal model such as the one in Figure~\ref{fig:tool:multipleattack1}.

### 2017-11-03 tool-screenshots.tex
tool section and small changes

@@
-\section{Tool screenshots}
+\section{From RationalGRL to GRL - examples from the tool}
 \label{sect:tool-screenshots}
@@
 \centering
-\includegraphics[scale=0.6]{img/tool-overview}
-\caption{Overview of the RationalGRL tool}
-\label{fig:tool:overview}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.6]{img/tool-cqdetails}
-\caption{Critical question details pane for ``Is the resource relevant/useful?''}
-\label{fig:tool:cqdetails}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.6]{img/tool-cqeffect}
-\caption{Effect of answering CQ11: ``Is the resource reelevant/useful?'' with ``No''}
-\label{fig:tool:cqeffect}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.5]{img/tool-argument}
-\caption{Argument details pane for generic argument ``Necessary''}
-\label{fig:tool:argument}
-\end{figure}
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.5]{img/tool-linkdetails}
-\caption{Link details pane for decomposition link}
-\label{fig:tool:cqdetails}
-\end{figure}
-
-\newpage
-
-\begin{figure}[ht]
-\centering
-\includegraphics[scale=0.6]{img/tool-figfrompaper}
-\caption{RationalGRL model of Figure~\ref{fig:example-small2} in the tool}
+\includegraphics[scale=0.8]{img/tool/model1a}
+\caption{RationalGRL model of Figure~\ref{fig:example-small3} in the tool}
 \label{fig:tool:figfrompaper}
@@
 \centering
-\includegraphics[scale=0.6]{img/tool-multipleattack}
-\caption{Adding argument attack multiple arguments to Figure~\ref{fig:tool:figfrompaper}}
+\includegraphics[scale=0.6]{img/tool/GRLmodel1a}
+\caption{Exporting the model Figure~\ref{fig:tool:figfrompaper} to GRL gives the the above model in jUCMNAv}
+\label{fig:tool:figfrompaper1}
+\end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.8]{img/tool/model2}
+\caption{Adding an argument that attacks multiple GRL elements to Figure~\ref{fig:tool:figfrompaper}}
 \label{fig:tool:multipleattack}
 \end{figure}
+
+\begin{figure}[ht]
+\centering
+\includegraphics[scale=0.6]{img/tool/GRLmodel2}
+\caption{Exporting the model Figure~\ref{fig:tool:multipleattack} to GRL gives the the above model in jUCMNAv}
+\label{fig:tool:multipleattack1}
+\end{figure}

### 2017-11-03 introduction.tex
Introduction
Rewrote part of the introduction to make it sense for RE community and
remove all the grammatical mistakes.

@@
 The most popular ones include $i*$~\cite{yu1997towards}, Keep All Objects Satisfied (KAOS)~\cite{van2008requirements}, the NFR framework~\cite{chung2012non}, \textsc{Tropos}~\cite{giorgini2005goal}, the Business Intelligence Model (BIM)~\cite{horkoff2014strategic}, and the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}.
-
-A goal model is often the result of a discussion process between a group of stakeholders.
-For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
-Therefore, it is often not necessary to record all of the details of the discussion process that led to the final goal model.
-However, goal models for many complex, real-world information systems -- e.g., air-traffic management systems, systems that support industrial production processes, or government and healthcare services -- are not constructed in a short amount of time, but rather over the course of several workshops with stakeholders and requirements engineers.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development.
 
@@
 
-The first difficulty is that the goal modeling phase, particularly in large projects, is dynamic: goal models continuously change and evolve.
-Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}, and stakeholders may change their opinion about a modeling decision in between two modeling sessions, which may require revisions of the goal model.
-If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially have led to different goal models are lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
-Furthermore, other stakeholders, such as developers who were not the original authors of the goal model, may have to make sense of a goal model in order to, for example, use it as input in a later RE stage or in the development phase.
-If preferences, opinions and rationales behind the goal models are not stored explicitly, it may not only be more difficult to understand the model, but the stakeholders may also end up having the same unnecessary discussions throughout the goal modeling phase.
+A goal model is often the result of a discussion process between a group of stakeholders.
+For small-sized systems, goal models are usually constructed in a short amount of time, involving stakeholders with a similar background.
+Therefore, it is often not necessary to record all of the details of the discussion process that led to the final goal model.
+However, goal models for many complex, real-world information systems -- e.g., air-traffic management systems, systems that support industrial production processes, or government and healthcare services -- are not constructed in a short amount of time, but rather over the course of several workshops with stakeholders and requirements engineers.
+%SG: I reword and add the following sentences as the paragraphs below are very disjunct.
+Thus, developing goal models for a complex and large system is not a trivial task and can be very cumbersome.
+In such situations, failure in recording the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development.
 
-A further problem is that current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on the goal model.
-Similarly, with existing goal modelling languages one can change a part of the goal model, but it is not possible to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
-This becomes even more problematic if the stakeholders constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs of another group of stakeholders.
-The disconnect between the goal models and their underlying beliefs and opinions may further lead to a poor understanding of the problem and solution domain, which is an important reason of RE project failure~\cite{curtis1988field}.
+The first challenge with the goal modeling phase, particularly in large projects, is related to their dynamic nature: goal models continuously change and evolve.
+Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
+Stakeholders may change their opinion about a modeling decision in between two modeling sessions, which may require revisions of a goal model.
+If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially lead to different goal models might be lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
+Furthermore, other stakeholders, such as developers who were not the original authors of the goal model, may have to make sense of a goal model in order to, for example, use it as an input in a later RE stage or in the development phase.
+If preferences, opinions and rationales behind the goal models are not stored explicitly, it may not only be more difficult to understand the model, but the other stakeholders may end up having similar discussions throughout the design and development phase as well.
 
-To summarize, what is needed is a way of recording the rationales (beliefs, opinions, discussions, ideas) underlying a goal model.
-It should be possible to see how these rationales changed during the goal modeling process, and the rationales should be clearly linked to the various elements of the resulting goal model.
-In order to be able to do this, we propose a framework with tool-support that combines traditional goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{BenchCaponDunne2007}.
+Another challenge is that current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
+A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on a goal model.
+Similarly, with existing goal modeling languages one can change a part of a goal model without being able to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
+This becomes even more challenging if the stakeholders constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs of another group of stakeholders.
+The disconnect between goal models and their underlying beliefs and opinions may further lead to a poor understanding of the problem and solution domain, which is an important reason of RE project failure~\cite{curtis1988field}.
+
+To summarize, what is needed is a way to record the rationales (beliefs, opinions, discussions, ideas) underlying a goal model.
+It should be possible to see how these rationales change during the goal modeling process, and they  should be clearly linked to the intentional elements of the resulting goal model.
+In order to do this, we propose a framework with its tool-support which combines traditional goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{BenchCaponDunne2007}.
 We have identified \textbf{five important requirements} for our framework:
@@
 \item 
-The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
+The argumentation techniques must capture the actual discussions of the stakeholders or designers in the early requirements engineering phase.
 \item 
-The framework must have formal traceability links between elements of the goal model and underlying arguments.
+The framework must have formal traceability links between elements of the goal model and their underlying arguments.
 \item 
@@
 \item 
-There should be a methodology for the framework to guide the practitioners in its application in real cases.
+There must be a methodology for the framework to guide the practitioners in its application in real cases.
 \item 
-The framework must have software tool support.
+The framework must have a tool-support.
 \end{enumerate}
@@
 
-Following on from our previous work \cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop a framework called \emph{RationalGRL}, which combines the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
-Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.
-Associated with argumentation schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way.
-Argument schemes they are very well suited for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}.
+Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop our \emph{RationalGRL} framework, which combines the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
+The \emph{RationalGRL} framework includes argument schemes and their critical questions, a meta-model, a tool-support and a methodology on how to develop such framework.
+
+Argument schemes are reusable patterns of reasoning that capture the common approaches in which humans argue and reason.
+\emph{Critical questions} which are associated with argument schemes, point to typical sources of doubt or implicit assumptions people make when raising an arguement.
+Argument schemes are a good choice for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}.
 
 One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for, among other things, modeling software design discussions~\cite{BlackEtal2013}.
-Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyse and guide stakeholders' discussions about goal models.
-Our approach thus provides a rationalization to the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted and others have been rejected.
+Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyze and guide stakeholders' discussions about goal models.
+Our approach, thus, provides a rationalization to the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected.
 Our list of argument schemes was constructed by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
-This ensures that the argumentation schemes we propose are close to actual real-world discussions stakeholders have (\textbf{requirement 1}).
+This ensures that the argument schemes we propose are close to actual real-world discussions the stakeholders have (\textbf{requirement 1}).
 
+%SG: Need to come back to this later.
 The meta-model of the RationalGRL framework clearly specifies the traceability links between the arguments based on the schemes and the GRL models (\textbf{requirement 2}).
-In addition to this meta-model, we provide formal semantics for RationalGRL by formalising the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
-We then formally capture the link between argumentation and goal modelling as a set of algorithms for applying argument schemes and critical questions about goal models.
+In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
+We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.
 These formal traceability links allow us to compute the effect of the arguments and counterarguments proposed in a discussion on a GRL model (\textbf{requirement 3}).
@@
 
-Because we want RE practitioners in the field to be able to use our RationalGRL framework (\textbf{requirement 4}), we also propose a methodology for using RationalGRL, which consists of developing goal models and posing arguments based on schemes in an integrated way.
-To show that the RationalGRL methodology can be used in a real case, we illustrate the steps of our methodology with the traffic simulator case study.
-Finally, we have developed a web-based prototype\footnote{\url{http://www.rationalgrl.com}} for building goal models and arguing about them, which acts as a supporting tool to the RationalGRL methodology (\textbf{requirement 5}).
+One of our major goals is to provide means for RE practitioners to capture the underlying arguments of goal models by using our RationalGRL framework (\textbf{requirement 4}).
+Thus, we propose a methodology, which consists of developing goal models and posing arguments based on schemes in an integrated way.
+To show the applicability of the RationalGRL methodology, we illustrate its steps with our traffic simulator case study.
+Finally, we have implemented the Rational-GRL tool-support, a web-based prototype\footnote{\url{http://www.rationalgrl.com}}, for modeling argument schemes and critical questions and for reasoning about goal models with respect to them (\textbf{requirement 5}).
 
+%SG: Need to reword this later.
 The rest of this article is organized as follows.
-Section~\ref{sect:background} introduces our case study and briefly discusses the basics of GRL and argumentation about goals using argument schemes.
+Section~\ref{sect:background} introduces our running example and briefly discusses the basics of GRL and argumentation about goals using argument schemes.
 Section~\ref{sect:gmas} contains the case study, an in-depth explanation of how we obtained an initial set of argument schemes and critical questions by coding transcripts from discussions about an information system.
-Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL language metamodel, and examples from the case study that illustrate the framework and language.
-In Section~\ref{sect:formalframework} we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and develop various algorithms that change a RationalGRL model according to the argument scheme or critical question that is applied.
-Section~\ref{sect:methodology} discusses the RationalGRL methodology and a discussion of the prototype tool we developed for RationalGRL.
+Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL language metamodel, and examples from the case study that illustrate the framework and its language.
+In Section~\ref{sect:formalframework}, we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and develop various algorithms that change a RationalGRL model according to an argument scheme or a critical question that is applied.
+Section~\ref{sect:methodology} discusses the RationalGRL methodology and a discussion of the prototype tool we have developed for RationalGRL.
 Finally, Section~\ref{sect:discussion} covers related work, future work, and a conclusion.

### 2017-11-03 main.tex
tool section and small changes
Introduction
Rewrote part of the introduction to make it sense for RE community and
remove all the grammatical mistakes.

@@
 
+\usepackage{dblfloatfix}
 \setcounter{tocdepth}{3}
@@
 \title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee \and Floris J. Bex \and Sepideh Ghanavati}
-\institute{Marc van Zee \at Google Research, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} \and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands \\\email{f.j.bex@uu.nl} \and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu}}
+\author{Marc van Zee  \and Sepideh Ghanavati \and Floris J. Bex}
+\institute{Marc van Zee \at Google Research, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com}  \and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu} \and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands \\\email{f.j.bex@uu.nl}}
 

### 2017-11-06 abstract.tex
Small amendmends thoughout the paper.
MASTERFIX

@@
 The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found in the resulting model.
+While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
 For instance, the discussion on whether to accept or reject a certain goal and the ultimate rationale for acceptance or rejection cannot be captured in current goal modeling languages.
@@
 RationalGRL provides a methodology, formal semantics and tool support to capture the discussions and outcomes of the argumentation process that leads to a goal model.
+
+\keywords{Goal Modeling \and Argumentation \and Goal-oriented requirements engineering \and Practical Reasoning} 
 \end{abstract}

### 2017-11-06 background.tex
MASTERFIX

@@
 In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.
+In this article, we consider a slightly simplified version of GRL.
+We motivate this and explain which subset of GRL we consider.
 Next, we introduce argumentation: we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
@@
 Our examples and case study are based on the data produced by a recent series of experiments by Schriek et al.
-\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}, which presents a well-known design reasoning assignment in software engineering about a traffic simulator.
+\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}.
+This experiment contains a well-known design reasoning assignment in software engineering about a traffic simulator.
 In this assignment (see Appendix~\ref{sect:designprompt}), designers are provided with a problem description, requirements, and a description of the desired outcomes.
@@
 asked designers (groups of students) to discuss the requirements of this traffic simulator.
-These discussions were recorded and transcribed, and we used these transcripts for an extensive case study on the basis of which we developed our RationalGRL framework.
+These discussions were recorded and transcribed, and we used these transcripts for an extensive case study on the basis of which we develop our RationalGRL framework.
 Furthermore, we also use the traffic simulator case for the running examples throughout this paper.
@@
 The GRL elements and relationships used in this paper are shown in Figure~\ref{fig:grl_legend}.
-Note that we only consider a subset of the full GRL specification in this paper.
-In particular, the full GRL specification considers different levels of positive and negative contributions and satisfaction levels of goals and tasks, which are not directly relevant for this work.
-Furthermore, the full GRL specification includes beliefs to capture design rationale, whereas in this paper, the reasoning and rationales behind the models are captured using arguments -- for a brief discussion on beliefs, see Section~\ref{sect:goalmodeling:relatedwork}.
 
@@
 Tasks represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, we have the two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve goal \emph{Generate cars}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off screen.
+In Figure~\ref{fig:example-small}, there are two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve goal \emph{Generate traffic}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off screen.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources to be available (e.g., \emph{Car Objects}).
@@
 AND, IOR (Inclusive OR), and XOR (eXclusive OR) decomposition links allow an element to be decomposed into sub-elements.
-In Figure~\ref{fig:example-small}, the goal \emph{Generate cars} is XOR-decomposed to the tasks \emph{Create new cars} and \emph{Keep same cars}, as they are alternative ways of achieving the goal \emph{Generate cars}.
+In Figure~\ref{fig:example-small}, the goal \emph{Generate traffic} is XOR-decomposed to the tasks \emph{Create new cars} and \emph{Keep same cars}, as they are alternative ways of achieving the goal \emph{Generate traffic}.
 Contribution links indicate impacts of one element on another element, which can be positive or negative.
-Task \emph{Create new cars} has a positive contribution to the task \emph{Realistic simulation}, and a negative contribution to the task \emph{Simple design}.
+Task \emph{Create new cars} has a positive contribution to the softgoal \emph{Realistic simulation}, and a negative contribution to the softgoal \emph{Simple design}.
 Dependency links model relationships between actors or resources.
-Here, the goal \emph{Generate cars} depends on the resource \emph{Car objects}.
+Here, the goal \emph{Generate traffic} depends on the resource \emph{Car objects}.
 
@@
 Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
+GRL has six evaluation algorithms which allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements quantitatively, qualitatively or in a hybrid way.
 
-GRL has six evaluation algorithms which allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements quantitatively, qualitatively or in a hybrid way.
-jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
+GRL is implemented in the open-source tool called jUCMNav~\cite{jUCMNav}, which is actively maintained\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}.
+jUCMNav currently has over 2.000 commits made by over 40 contributors, representing over 200.000 lines of code.
+jUCMNav is used actively in research, it is mentioned in over 400 papers according to Google Scholar.
+As jUCMNav is open, other research teams have also extended the tool in many directions in the past few years, e.g., for: modeling and analyzing security misuse cases, supporting activity theory, Web-based goal modeling, and enterprise architecture modeling.
+jUCMNav also allows for adding new GRL evaluation algorithms.
 GRL also has the capability to be extended through metadata, links, and external OCL constraints.
-This allows GRL to be used in many domains without the need to change the whole modeling language.
+This allows GRL to be used in many domains without the need to change the modeling language itself.
 
@@
 	\item Why does task \emph{Create new cars} contribute negatively to \emph{Simple design} and positively to \emph{Realistic simulation}?
-	\item Why does \emph{Generate cars} XOR-decompose into two tasks?
+	\item Why does \emph{Generate traffic} XOR-decompose into two tasks?
 \end{itemize}
@@
 The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the model, such as clarification steps for the naming, or alternatives that have been considered for the relationships.
+
+In this article, we consider a subset of the full GRL specification.
+In particular, the full GRL specification considers different levels of positive and negative contributions and satisfaction levels of goals and tasks, which are not directly relevant for this work.
+Therefore, we consider only two kinds of contribution values in our framework: \emph{positive} and \emph{negative}.
+It is straightforward to extend this to multiple values, but we have not done so to keep our presentation concise.
+
+Furthermore, the full GRL specification includes beliefs to capture design rationale, whereas in this paper, the reasoning and rationales behind the models are captured using arguments.
 The idea behind the original GRL specification is that beliefs can be used to capture such design rationales that make later justification and review of a model easier.
-However, beliefs cannot be connected to links - this makes answering the third and fourth question above impossible.
-Furthermore, beliefs are after-the-fact design rationales and do not capture the types of critical questions given above.
+While beliefs can indeed be used for rationalization by connecting them to IEs, jUCMNav does not use them in the evaluation algorithms.
+Moreover, a belief can only be used to \emph{support} IEs, not to reject them.
+For a more detailed comparison of our framework with GRL beliefs, see Section~\ref{sect:goalmodeling:relatedwork}.
 
@@
 \begin{itemize}
-\item[] \emph{Generate Cars} is a goal,
-\item[] Performing action \emph{Keep same cars} realizes goal \emph{Generate cars}, 
+\item[] \emph{Generate traffic} is a goal,
+\item[] Performing action \emph{Keep same cars} realizes goal \emph{Generate traffic}, 
 \item[] \underline{Therefore} 
@@
 
-Note that PRAS is an argument scheme that captures a full inference step: ``$G$, $A$ realizes $G$ \emph{Therefore} $A$''.
+Note that PRAS is an argument scheme that captures a full inference step: ``$G$, $A$ realizes $G$, \emph{Therefore} $A$''.
 There are, however, also schemes that capture simpler reasoning patterns, such as claims of the form ``$A$ does not realize $G$''.
@@
 In argumentation, conclusions which are at one point acceptable can later be rejected because of new information.
-For example, we may argue that, in fact, performing action \emph{Keep same cars} does not realize goal \emph{Generate cars}, thus giving a counterargument to the above instantiation of PRAS.
-Atkinson et al.~\cite{atkinson2007} define a set of so-called critical questions that point to typical ways in which an argument based on PRAS can be criticized by.
+For example, we may argue that, in fact, performing action \emph{Keep same cars} does not realize goal \emph{Generate traffic}, thus giving a counterargument to the above instantiation of PRAS.
+Atkinson et al.~\cite{atkinson2007} define a set of so-called critical questions that point to typical ways in which an argument based on PRAS can be criticized.
 Some examples of critical questions are as follows.
@@
 These counterarguments also follow a scheme; for example, a negative answer to CQ1 follows the scheme ``Action $A$ will not realize goal $G$'', which can be instantiated (e.g.
-``\emph{Keep same cars} does not realize \emph{Generate cars''}) to form a counterargument to the original argument.
+``\emph{Keep same cars} does not realize \emph{Generate traffic''}) to form a counterargument to the original argument.
 
 Another way to criticize an argument for an action is to suggest an alternative action that realizes the same goal (CQ2).
-For example, we can argue that performing \emph{Create new cars} also realizes the goal \emph{Generate cars}.
+For example, we can argue that performing \emph{Create new cars} also realizes the goal \emph{Generate traffic}.
 Also, it is possible that performing an action has a negative side effect (CQ3).
-For example, while the action \emph{Create new cars} realizes the goal \emph{Generate cars}, it has a negative side effect, namely hurting \emph{Simple design}: having the simulation constantly create new cars is fairly complex design choice.
+For example, while the action \emph{Create new cars} realizes the goal \emph{Generate traffic}, it has a negative side effect, namely hurting \emph{Simple design}: having the simulation constantly create new cars is fairly complex design choice.
 
@@
 In argumentation, counterarguments are said to \emph{attack} the original arguments.
-Given a set of arguments and attacks between these arguments, we can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Formal definitions of argumentation frameworks and semantics will be given in section \ref{sect:gmas}.
-In this section, we will briefly discuss the intuitions behind these concepts.}.
+Given a set of arguments and attacks between these arguments, we can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Formal definitions of argumentation frameworks and semantics will be given in Section~\ref{sect:gmas}.
+In this section, we briefly discuss the intuitions behind these concepts.}.
 Figure \ref{fig:pras:example} shows three arguments from the traffic simulation example, where arguments are rendered as boxes and attack relations as arrows.
 There are two arguments based on PRAS: argument A1 for \emph{Keep Same Cars} and argument A2 for \emph{Create new cars}.
-Argument A2 proposes an alternative way of realizing the same goal \emph{Generate cars} with respect to argument A1 and vice versa (cf.
+Argument A2 proposes an alternative way of realizing the same goal \emph{Generate traffic} with respect to argument A1 and vice versa (cf.
 CQ2), so A1 and A2 mutually attack each other, denoted by the arrows between A1 and A2.
@@
 So if we would use the version of PRAS presented in this section for our argumentation, we would violate requirement 1: The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
-Our solution was to develop our own set of argument schemes and critical questions by analyzing transcripts of discussions about the traffic simulator.
+Our solution is to develop our own set of argument schemes and critical questions by analyzing transcripts of discussions about the traffic simulator.
 This set of schemes and questions and our case study are described in the next section.

### 2017-11-06 casestudy.tex
MASTERFIX

--- 2017-10-29 gmas.tex
@@
 Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
-To get a sense of such discussions, we performed an extensive case study to see which types of discourse are used during a discussion of system requirements, and how these discourse types can be captured as argument schemes and critical questions regarding goals and tasks.
+To get a sense of such discussions, we performed an extensive case study to see which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
 We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS.
 In this section we present our case study.
-All original transcripts, codings, and models are available in our online repository\footnote{\label{foot:url}\todo{F}{M}{\url{insertURL}}}.
+All original transcripts, codings, and models are available in our online repository, which can be found at:
+
+\begin{quote}
+\rationalgrlurl{}
+\end{quote}
 
@@
 The students did not use any goal modeling technique in the course or during the discussions.
-The students were asked to record their design sessions, and the recordings were subsequently transcribed.
+They were asked to record their design sessions, and the recordings were subsequently transcribed.
 We used three of these transcripts, totaling 153 pages, for our case study.
@@
 \hline
-- & Generic counterargument & 0& 2 & 2 & 4\\
+Gen & Generic counterargument & 0& 2 & 2 & 4\\
 \hline
@@
 
-Before we started coding the transcripts, we drew up an initial list of 10 argument schemes (AS0-AS9 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
+Before we started coding the transcripts, we drew up an initial list of 11 argument schemes (AS0-AS10 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
 AS0 to AS4 are schemes that concern a single element of the goal model.
 So, for example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'.
-AS5 to AS9 are claims about the links between GRL elements.
+AS5 to AS10 are claims about the links between GRL elements.
 Our initial list also contained 18 critical questions, inspired by the questions associated with the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}.
-CQ2 to CQ6b (Table~\ref{table:transcripts:results:argumentschemes}) are examples of these critical questions, other examples are `Is the softgoal legitimate?' and `Are there alternative ways to contribute to the same softgoal?'.
+CQ2 to CQ12 (Table~\ref{table:transcripts:results:argumentschemes}) are examples of these critical questions, other examples are `Is the softgoal legitimate?' and `Are there alternative ways to contribute to the same softgoal?'.
 
-%%TODO check numbers against original annotation docs
 Using the initial list of arguments and critical questions, we coded three transcripts of requirements discussions.
@@
 Rather, they say things such as `...we have a set of actions.
-Save map, open map, ...' (Table~\ref{table:transcripts:traffic-light}, Appendix~\ref{sect:transcripts:excerpts}) and `...
-in that process there are activities like create a visual map, create a road' (Table~\ref{table:transcript:task-clarification}, Appendix~\ref{sect:transcripts:excerpts}).
+Save map, open map, ...' (Table~\ref{table:transcripts:traffic-light}, Appendix~\ref{sect:transcripts:excerpts}) and `We also have to be able to change the inflow of cars.
+How many car come out in here on the side' (Table~\ref{table:transcript:task-clarification}, Appendix~\ref{sect:transcripts:excerpts}).
 Furthermore, in some cases the critical questions are explicit.
@@
 In other cases, however, the question remains implicit but we added it in the coding.
-For example, CQ12 is not found in the transcripts, but the related counterargument is:  `...you don't have to specifically add a traffic light' (Table~\ref{table:transcripts:traffic-light}, Appendix~\ref{sect:transcripts:excerpts}).
+For example, CQ11 (`Is the element relevant/useful?') is not found directly in the transcripts, but it can be inferred from statements such as `...you don't have to specifically add a traffic light' (Table~\ref{table:transcripts:traffic-light}, Appendix~\ref{sect:transcripts:excerpts}).
 
@@
 The final results of the coding can be found in Table~\ref{table:transcripts:results:argumentschemes}.
-We found a total of 159 instantiations of argument schemes AS0-AS11.
+We found a total of 159 instantiations of argument schemes AS0-AS10.
 The most used argument scheme was AS2: ``Actor $A$ has task $T$'', however, each argument scheme is found in transcripts at least twice.
@@
 We coded 41 applications of critical questions.
-Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, CQ13).
+Many critical questions (about 55\%) involved clarifying the name of an element, or discussing its relevance (CQ12, Gen).
 
@@
 First, an argument is posed that introduces a number of tasks.
-Counterarguments are then given against two of these tasks (\emph{add traffic light} and \emph{remove intersection}), which are subsequently disabled.
+A counterargument is then given against one of these tasks (\emph{add traffic light}), which is subsequently disabled.
 An example of replacement is given in Table~\ref{table:transcript:decomposition} (Appendix~\ref{sect:transcripts:excerpts}): what used to be an AND-decomposition is changed into an OR-decomposition.
-We will further discuss the possible operations in Section~\ref{sect:overview}.
-
-%MvZ: I completely removed the subsection below. I think it doesn't add much and it is confusing. What do you think?
-
-%SG: I agree... It was more confusing and downgrade our approach
-
-%\subsection{Analysis}
-%\label{sect:gmas:transcripts:analysis}
-
-%\paragraph{Analysis of Argument Schemes}
-%Recall that our initial list of argument schemes consists of AS1-AS4, AS6-AS9 (Table~\ref{table:argument-schemes}).
-Therefore, the difference between the initial list of argument schemes and those found in transcripts is quite small.
-%TODO The next few sentences are weird... they make our work very trivial.. rewrite.
-%We found it surprising that we were able to find back all the schemes in the transcript at least twice, even more since the topic of discussion was not goal models, but more generally the architecture of an information system.
-This gives us an indication that it is possible to capture (parts of the) arguments used in those type of discussions using argument schemes.
-
-%We observed that our initial list is rather limited, which is a consequence of the fact that it is derived from PRAS.
-Since PRAS only considers very specific types of relationships, we are not able to capture many other relationships existing in GRL.
-GRL has four types of intentional elements (softgoal, goal, task, resource) and four types of relationships (positive contribution, negative contribution\footnote{In fact, a contribution can be any integer in the domain [-100,100], but for the sake of simplicity we only consider two kinds of contributions here.}, dependency, decomposition), allowing theoretically $4^3=64$ different types of argument schemes, of which we currently only consider 11.
-%TODO which analysis? why aren't they used? How did we empirically validate it? How are we confident in the result?
-%Our analysis, however, shows that many of these schemes are not often used, and thus, gives us some confidence in the resulting list.
-However, additional argument schemes and critical questions can be easily added to our framework.
-Beside, our list is not meant to be exhaustive.
-
-%\paragraph{Analysis of the Critical Questions} The difference between the initial list of critical questions and those we found in transcripts is much larger than for the critical questions.
-%TODO  I don't get this sentence.
-Also, the ones below.
-%We found few of the critical questions we initially proposed. %TODO ???
-%However, this does not mean that they were not implicitly used in the minds of the participants.
-If a participant, for instance, forms an argument for a contribution from a task to a softgoal, it may very well be that she/he was asking her/him-self the question ``Does the task contributes to some other softgoals?''.
-However, many of these critical questions are not mentioned explicitly.
-If we assume this explanation is at least partially correct, then this would mean that critical questions may still play a role when formalizing the discussions leading up to a goal model, and it would be limiting to leave them out of our framework.
-In the context of a tool support, we believe that having these critical questions available may stimulate discussions.
-%TODO what does it mean in the context of a tool support?
+We will further discuss these operations in Section~\ref{sect:overview} with various examples, and w formalize the operations in Section~\ref{sect:formalframework} with algorithms.

### 2017-11-06 discussion.tex
formalframework cleanup & new example
MASTERFIX

@@
 \subsection{Related work}
-\label{sect:goalmodeling:relatedwork}
+\label{sect:discussions:relatedwork}
 
@@
 Jureta et al.
-give a simple made-up example of an argumentative discussion regarding a meeting scheduler, in which a goal model is being built by the stakeholders proposing tasks, goals, and alternative solutions for goals.
+use a well-known fictitious example of an argumentative discussion regarding a meeting scheduler, in which a goal model is being built by the stakeholders proposing tasks, goals, and alternative solutions for goals.
 They include clarification as an important step in their GAM process, and discuss various types of clarity problems (vagueness, ambiguity, overgenerality, synonymy) and basic techniques for dealing with them (e.g.
@@
 One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
-So the previous iterations of the RationalGRL framework violated requirement 1: argument diagrams do not closely mirror the actual discussions of stakeholders in which ideas are proposed and challenged.
+RationalGRL overcomes this problem by integrating argumentation and goal modeling in a single modeling language.
 
-\subsection{Open issues}
-\label{sect:goalmodeling:openissues}
+\subsection{Future work}
+\label{sect:discussion:futurework}
 
@@
 \paragraph{Specific Domains}
-At the moment, the argumentation schemes and critical questions of RationalGRL are focused only on the core goal modeling process, that is, the discussion about the goals and functionality of an information system.
+The argumentation schemes and critical questions presented in this paper are focused only on the core goal modeling process, that is, the discussion about the goals and functionality of an information system.
 In the RE process, there are many more domain specific discussions that also take place, such as discussions involving expert opinions \cite{murukannaiah2015}, discussions about security requirements \cite{haley2008security,yu2015automated,ionita2014argumentation}, architectural principles \cite{marosin-etal:caise2016}, legal compliance \cite{Ghanavati2013}, and so forth.
@@
 For instance, in our framework we have not included the possibility of expressing preferences between mutually inconsistent arguments (e.g.
-like in Figure~\ref{fig:XX}).
+such as in Figure~\ref{fig:goalmodeling:arg3}).
 Using the work by Modgil~\cite{modgil2009}, it is possible to provide explicit arguments for preferences, thus allowing us to make a reasoned choice between two options.
@@
 \paragraph{Empirical validation}
-We have performed a large case study and found more than 200 incidences of arguments and questions.
+We have performed a case study and found more than 200 incidences of arguments and questions.
 However, what is still lacking is further empirical investigation of whether, and how, the tools and techniques provided by the RationalGRL framework really improve early phase requirements engineering.
@@
 Naturally, the type of problem will greatly influence the outcome.
-For a simple problem which two people have to solve in an hour, RationalGRL will probably not add much.
+For a simple problem which two people have to solve in an hour, RationalGRL will most likely be of little benefit.
 However, a more complex problem on which multiple persons have to work asynchronously might benefit from the extra tools offered by RationalGRL.
 Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the methodology, the tool, or a list of critical questions to serve as reminders during the goal modelling process (cf.
-\cite{SchriekEtal2016}}).
+\cite{SchriekEtal2016}).
 
 \subsection{Conclusion}
+\label{sect:discussion:conclusion}
 
-\todo{Marc}{all}{Finish this}
+This paper presents RationalGRL: a framework for integrated goal modeling and argumentation.
+We extend the well-known goal modeling language GRL with argument schemes and critical questions.
+\emph{Argument schemes} are reusable patterns of reasoning that capture the common approaches in which humans argue and reason.
+\emph{Critical questions}, which are associated with argument schemes, point to typical sources of doubt or implicit assumptions people make when raising an argument.
 
-The introduction of this article contains five requirements we identified for our framework. We use the conclusion to discuss how RationalGRL meets our initial requirements.
+We argue that the argument scheme for practical reasoning~\cite{walton1990,atkinson2007} is particularly useful when reasoning about goals.
+Inspired by the work on practical reasoning from Artificial Intelligence, most notably by Atkinson and Bench-Capon~\cite{atkinson2007}, we develop a list of argument schemes that can be used to analyze and guide stakeholders' discussions about goal models.
+Our approach, thus, provides a rationalization for the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected.
+Our list of argument schemes was constructed by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
+This ensures that the argument schemes we propose are close to actual discussions the stakeholders have (\textbf{requirement 1}).
 
-\paragraph{1. The argumentation techniques should be close to actual discussions stakeholders or designers have.}
+The meta-model of the RationalGRL framework clearly specifies the traceability links between the arguments based on the schemes and the GRL models (\textbf{requirement 2}).
+In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
+We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.
+These formal traceability links allow us to compute the effect of the arguments and counterarguments proposed in a discussion on a GRL model (\textbf{requirement 3}).
+In other words, we can determine whether the elements of a GRL model are acceptable given potentially contradictory opinions of stakeholders.
+Thus, we add a new formal evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (in addition to their \emph{satisfiability}~\cite{Amyot:2010:EGM:1841349.1841356}).
 
-We analyze a set of transcripts containing discussions about the architecture of an information system.
-
-\paragraph{2. The framework must have the means to formally model parts of the discussion process.}
-In order to generate goal models based on formalized discussions (requirement 2), we, first, formalize the list of arguments from requirement 1 in an argumentation framework.
-We formalize the critical questions as algorithms modifying the argumentation framework.
-We use argumentation techniques from AI in order to determine which arguments are accepted and which are rejected.
-We propose an algorithm to generate a GRL model based on the accepted arguments.
-This helps providing traceability links from GRL elements to the underlying arguments (requirement 3).
-
-We implement our framework in an online tool called RationalGRL (requirement 4).
-The tool is implemented using Javascript.
-It contains  two parts, goal modeling and argumentation.
-The goal modeling part is a simplified version of GRL, leaving out features such as evaluation algorithms and key performance indicators.
-The argumentation part is new, and we develop a modeling language for the arguments and critical questions.
-The created GRL models in RationalGRL can be exported to jUCMNav~\cite{} %TODO SG: add citation
- the Eclipse-based tool for GRL modeling, for further evaluation and analysis.
-
-Our final contribution is a methodology on how to develop goal models that are linked to underlying discussions.
-The methodology consists of two parts, namely argumentation and goal modeling.
-In the argumentation part, one puts forward arguments and counter-arguments by applying critical questions.
-When switching to the goal modeling part, the accepted arguments are used to create a goal model.
-In the goal modeling part, one simply modifies goal models, which may have an effect on the underlying arguments.
-This might mean that the underlying arguments are no longer consistent with the goal models.
+One of our main goals is to provide means for RE practitioners to capture the underlying arguments of goal models by using the RationalGRL framework (\textbf{requirement 4}).
+To this end, we propose a methodology, which consists of developing goal models and posing arguments based on schemes in an integrated way.
+Finally, we have implemented the RationalGRL tool, a web-based prototype\footnote{\url{http://www.rationalgrl.com}}, for modeling argument schemes and critical questions and for reasoning about goal models with respect to them (\textbf{requirement 5}).

### 2017-11-06 formalframework.tex
MASTERFIX

@@
 This is done for multiple reasons: (i) Most approaches in formal argumentation (cf.
-Section~\ref{sect:background:pras}) use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated in to valid GRL models and visa versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support.
+Section~\ref{sect:background:pras}) use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated to valid GRL models and visa versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support, which we present in the next section.
 
-In Sections \ref{sect:formalframework:grl} and \ref{sect:formalframework:rationalgrl} we formalize a static representation of our framework based on the metamodel Figure~\ref{fig:metamodel}).
+In Sections \ref{sect:formalframework:grl} and \ref{sect:formalframework:rationalgrl} we formalize a static representation of our framework based on the metamodel (Figure~\ref{fig:metamodel}).
 We first provide a formal specification of a GRL model in Section~\ref{sect:formalframework:grl}, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
@@
 \item $IETypes = \{Softgoal, Goal, Task, Resource\}$,
-\item $LinkTypes = \{PosContr, NegContr, Dep, Decomp\}$\footnote{Recall that for decomposition links we only distinguish between positive and negative contribution.
-Extending the formalization model to include more values is straightforward.},
+\item $LinkTypes = \{PosContr, NegContr, Dep, Decomp\}$\footnote{Recall from Section~\ref{sect:background} that for contribution links we only distinguish between positive and negative contributions.
+Extending the formalization model to include all the GRL values is straightforward but has been omitted here for conciseness.},
 \item $Types = IETypes \cup LinkTypes\cup\{Actor, ActIE, GenArg\}$,
@@
 \item $name \in Names$ is a string description of the element,
-\item $decomptype\in DecompType$ refers to the type of decomposition.\footnote{Note that, like in the metamodel, the decomposition type is defined on the element.}
+\item $decomptype\in DecompTypes$ refers to the type of decomposition.\footnote{Note that, like in the metamodel, the decomposition type is defined on the element.}
 \end{itemize}
-The set of intentional elements is denoted by $IE$.
+A set of intentional elements is denoted by $IE$.
 \end{definition}
@@
 \end{definition}
-We can now formalize the actor in our example (Figure~\ref{fig:example-small}) as $(0,Actor,\text{Traffic Simulator})$.%, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
+We can formalize the actor in our example (Figure~\ref{fig:example-small}) as $(0,Actor,\text{Traffic Simulator})$.%, and we can for instance state $act_0.name = \text{Traffic Simulator}$.
 
@@
 Note that Actor-IE relations do not have an $id$: they are themselves not elements of a GRL model but rather indicate relations between elements of a GRL model (i.e., intentional elements and actors).
-The relation between the \emph{Traffic Simulator} actor ($id = 0$) and the \emph{Car objects} resource ($id=1$), for example, can be formalized as $(0, 1)$.
+The relation between the \emph{Traffic Simulator} actor ($id = 0$) and the \emph{Car objects} resource ($id=1$), for example, can be formalized as $(ActIE, 0, 1)$.
 
@@
 \end{definition}
-
-%We also introduce some simplified notation for the links.
 
@@
 
-Definition \ref{def:grl-model} only sums up the elements of the model and not the constraints that make a \emph{valid} GRL model, viz.
+Definition \ref{def:grl-model} only sums up the elements of the model and not the constraints that make a \emph{valid} GRL model.
+We will do so in the next definition.
+Note that in this definition, we use a subscript notation to refer to an element with a specific id.
+That is, $IE_i, Link_j$, and $Act_k$ refer respectively to the IE with id $i$, the link with id $j$, and the actor with id $k$.
 
@@
 \label{def:valid-grl-model}
-A GRL model $GRL=(IE, Act, R_{ActIE}, Contr, Decomp, Dep)$ (Def.~\ref{def:grl-model}) is a \emph{valid GRL model} iff the following conditions are satisfied:
+A GRL model $GRL=(IE, Act, R_{ActIE}, Link)$ (Def.~\ref{def:grl-model}) is a \emph{valid GRL model} iff the following conditions are satisfied:
 \begin{enumerate}
-\item ids are globally unique across IEs, Beliefs, Links, and Actors, i.e., let $X,Y\in \{IE,Act, Belief, Contr, Decomp, Dep\}$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
-\item Intentional elements of actors exist: $\forall (k, ActIE, i,j)\in R_{ActIE}: act_i \in Act \wedge ie_j \in IE$.
-\item An intentional element belongs at most to one actor: $\forall ie_i\in IE: |\{(k,ActIE,j,i)\in R_{ActIE}\}| \le 1$.
+\item ids are globally unique across IEs, Links, and Actors, i.e., let $X,Y\in \{IE,Act, Link\}$. For all $X_i$ and $Y_j$: if $i=j$ then $X=Y$ and $X_i=Y_j$.
+\item Intentional elements of actors exist: $\forall (ActIE, i,j)\in R_{ActIE}: \exists act_i \in Act \wedge \exists ie_j \in IE$.
+\item An intentional element belongs at most to one actor: $\forall ie_i\in IE: |\{(ActIE,j,i)\in R_{ActIE}\}| \le 1$.
 \item Links connect intentional elements: $\forall (i,type, j,k)\in Link: \{ie_j,ie_k\}\subseteq IE$.
@@
 \item All elements in the formalization have different ids, so this constraint is satisfied.
-\item $R_{ActIE}$ contains one element for each IE with id $i$, so this constraint is satisfied as well. Note that, in line with the GRL metamodel, links are not related to actors. 
+\item $R_{ActIE}$ contains one element for each IE with id $i$, so this constraint is satisfied as well. Note that, in line with the GRL metamodel, $R_{ActIE}$ does not relate links with actors. 
 \item There is one actor ($id=0$) and this is the only actor that appears in $R_{ActIE}$ so this constraint is satisfied.
-\item All links connect IEs: the contribution links connect elements with ids 2, 3, and 5, which are all IEs; the decomposition links connect elements with ids 4, 5, and 6, which are all IEs; and the dependency link connects id 1 with 4, which are both IEs.
+\item All links connect IEs: the contribution links connect elements with ids 2, 3, and 6, which are all IEs; the decomposition links connect elements with ids 5, 6, and 7, which are all IEs; and the dependency link connects id 1 with 4, which are both IEs.
 \end{enumerate}
@@
 \begin{itemize}
-\item An intentional element (Def.~\ref{def:ie}), 
+\item An intentional element $ie$ (Def.~\ref{def:ie}), 
 \item An actor $act$ (Def.~\ref{def:actor}),
@@
 \label{def:link:attack}
-Given a set of arguments $Arg$, an attack link is a relation such that $att=(A_i,A_j)$ means:
+Given a set of arguments $Arg$, an attack link $att=(A_i,A_j)$ means:
 \begin{itemize}
@@
 \end{itemize}
-The set of attack links is denoted by $Att$.
+A set of attack links is denoted by $Att$.
 \end{definition}
 
-As an example, take the arguments $A_5$ for the goal (Generate traffic, $AND$), $A'_5$ for the goal (Generate traffic, $XOR$), $A_4$ for the softgoal (Easy to use) and the generic argument $A_{13}$ (Redundant) against the softgoal (Easy to use).
+As an example, take the arguments $A_5$ for the goal `Generate traffic' ($AND$ decomposition), $A'_5$ for the goal `Generate traffic' ($XOR$ decomposition), $A_4$ for the softgoal `Easy to use' and the generic argument $A_{13}$ (`Redundant') against the softgoal (`Easy to use').
 Given these arguments there are two attack links (Figure~\ref{fig:example-small3}), namely $(A'_{5},A_{5})$ and $(A_{13}, A_{4})$.
 
-We now define RationalGRL models.
+We now define a RationalGRL model.
 
@@
 A'_5 = &(5, Goal, \text{Generate traffic}, XOR),&\\
-A_6 = &(6, Task, \text{Wrap cars}, AND),&\\
-A'_6 = &(6, Task, \text{Create new cars}, AND),&\\
-A_7 = &(7, Task, \text{Keep same cars}, AND),&\\
+A_6 = &(6, Task, \text{Create new cars}, AND),&\\
+A_7 = &(7, Task, \text{Wrap cars}, AND),&\\
+A'_7 = &(7, Task, \text{Keep same cars}, AND),&\\
 A_8 = &(8, Dep, 4, 1),&\\
@@
 \begin{flalign*}
-Arg = &\{A_0,\ldots,A_5, A'_5,A_6,A'_6,A_7,\ldots,A_{23}\}\\
-Att= &\{(A'_{5},A_5), (A'_{6},A_{6}), (A_{13},A_{4}), (A_{14},A_{13}),\\
+Arg = &\{A_0,\ldots,A_5, A'_5,A_6,A_7,A'_7\ldots,A_{23}\}\\
+Att= &\{(A'_{5},A_5), (A'_{7},A_{7}), (A_{13},A_{4}), (A_{14},A_{13}),\\
      & (A_{15},A_{9}),(A_{16},A_{15})\}\\
@@
 
-From Figure~\ref{fig:example-small3}, it can be read from that arguments $A_5$, $A_6$, $A_{13}$ and $A_{15}$ are currently disabled (rejected).
+From Figure~\ref{fig:example-small3}, it can be read from that arguments $A_5$, $A_7$, $A_{13}$ and $A_{15}$ are currently disabled (rejected).
 However, we did not yet make precise how exactly this is computed.
@@
 Although the algorithms in Section~\ref{sect:algorithms} do not generate mutually attacking arguments such as in Figure~\ref{fig:goalmodeling:arg3}, our formal framework does not explicitly forbid attack cycles in RationalGRL models.
+We discuss this in more detail in Section~\ref{sect:discussion:futurework}.
 
-\begin{figure}[b]
+\begin{figure}[ht]
 \centering
@@
 Take our two simple examples.
-In Figure~\ref{fig:goalmodeling:arg2}, we have that arguments $A_4$, $A'_5$ and $A_{14}$ are justified and arguments $A_5$ and $A_{13}$ are rejected.
+In Figure~\ref{fig:goalmodeling:arg2}, we have that arguments $A_4$, $A'_5$ and $A_{14}$ are acceptable and arguments $A_5$ and $A_{13}$ are rejected.
 In Figure~\ref{fig:goalmodeling:arg3}, both arguments $A_x$ and $A_y$ are undecided.
-Returning to our larger example from Figure~\ref{fig:example-small3}, we can see that arguments $A_5$, $A_6$, $A_{13}$ and $A_{15}$ are rejected and all the other arguments are acceptable.
+Returning to our larger example from Figure~\ref{fig:example-small3}, we can see that arguments $A_5$, $A_7$, $A_{13}$ and $A_{15}$ are rejected and all the other arguments are acceptable.
 
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}[ht]
   \caption{GRL to RationalGRL Translation}
@@
 introduces a new RationalGRL element and \textsf{DISABLE} creates a new argument that attacks another argument.
-The \textsf{REPLACE} operation is can be seen as a combination of \emph{INTRO} and \emph{DISABLE}: a new argument corresponding to a GRL element is added and this new argument attacks a previous version of the element or link.
+The \textsf{REPLACE} operation is can be seen as a combination of \textsf{INTRO} and \textsf{DISABLE}: a new argument corresponding to a GRL element is added and this new argument attacks a previous version of the element or link.
 We will discuss all three operations in separate sections below.
@@
 As an example, we can formalize the addition of actor ``Traffic Simulator'' to an empty RationalGRL model (i.e.
-$Arg=\emptyset$ and $Att=\emptyset$)as executing the algorithm $AS_0$(Traffic simulator), which results in a RationalGRL model with argument $A_0 = (0, Actor, \text{Traffic simulator})\in Arg$.
+$Arg=\emptyset$ and $Att=\emptyset$) as executing the algorithm $AS_0$(Traffic simulator), which results in a RationalGRL model with argument $A_0 = (0, Actor, \text{Traffic simulator})\in Arg$.
 

### 2017-11-06 introduction.tex
MASTERFIX

@@
 %SG: I reword and add the following sentences as the paragraphs below are very disjunct.
-Thus, developing goal models for a complex and large system is not a trivial task and can be very cumbersome.
-In such situations, failure in recording the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development.
+Developing goal models for such complex and large systems is not a trivial task and can be very cumbersome.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development.
 
@@
 Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
-Stakeholders may change their opinion about a modeling decision in between two modeling sessions, which may require revisions of a goal model.
+Stakeholders may change their opinions about a modeling decision in between two modeling sessions, which may require revisions of a goal model.
 If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially lead to different goal models might be lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
@@
 It should be possible to see how these rationales change during the goal modeling process, and they  should be clearly linked to the intentional elements of the resulting goal model.
-In order to do this, we propose a framework with its tool-support which combines traditional goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{BenchCaponDunne2007}.
+In order to do this, we propose a framework with tool support which combines traditional goal modeling approaches with argumentation techniques from Artificial Intelligence (AI) research~\cite{BenchCaponDunne2007}.
 We have identified \textbf{five important requirements} for our framework:
@@
 \item 
-There must be a methodology for the framework to guide the practitioners in its application in real cases.
+A methodology must be identified that can guide practitioners in using the framework.
 \item 
-The framework must have a tool-support.
+The framework must have a tool support.
 \end{enumerate}
@@
 
-Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop our \emph{RationalGRL} framework, which combines the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356} with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
-The \emph{RationalGRL} framework includes argument schemes and their critical questions, a meta-model, a tool-support and a methodology on how to develop such framework.
+%MvZ: I shortened the paragraph below and put the full version in the conclusion.
+I think it makes more sense to reflect on the requirements in the conclusion, and to only give a brief overview of our framework in the introduction.
+Too much detail about argument schemes etc may confuse the reader.
 
+Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing framework for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
 Argument schemes are reusable patterns of reasoning that capture the common approaches in which humans argue and reason.
-\emph{Critical questions} which are associated with argument schemes, point to typical sources of doubt or implicit assumptions people make when raising an arguement.
-Argument schemes are a good choice for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{murukannaiah2015}.
-
-One argument scheme that is important when reasoning about goals is the argument scheme for practical reasoning~\cite{walton1990,atkinson2007}, which has been used for, among other things, modeling software design discussions~\cite{BlackEtal2013}.
-Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyze and guide stakeholders' discussions about goal models.
-Our approach, thus, provides a rationalization to the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected.
-Our list of argument schemes was constructed by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
-This ensures that the argument schemes we propose are close to actual real-world discussions the stakeholders have (\textbf{requirement 1}).
-
-%SG: Need to come back to this later.
-The meta-model of the RationalGRL framework clearly specifies the traceability links between the arguments based on the schemes and the GRL models (\textbf{requirement 2}).
+We develop a list of argument schemes by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
+In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models.
 In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
 We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.
-These formal traceability links allow us to compute the effect of the arguments and counterarguments proposed in a discussion on a GRL model (\textbf{requirement 3}).
-In other words, we can determine whether the elements of a GRL model are acceptable given potentially contradictory opinions of stakeholders.
-Thus, we add a new formal evaluation technique for goal models that allows us to assess the \emph{acceptability} of elements of a goal model (in addition to their \emph{satisfiability}~\cite{Amyot:2010:EGM:1841349.1841356}).
 
-One of our major goals is to provide means for RE practitioners to capture the underlying arguments of goal models by using our RationalGRL framework (\textbf{requirement 4}).
-Thus, we propose a methodology, which consists of developing goal models and posing arguments based on schemes in an integrated way.
-To show the applicability of the RationalGRL methodology, we illustrate its steps with our traffic simulator case study.
-Finally, we have implemented the Rational-GRL tool-support, a web-based prototype\footnote{\url{http://www.rationalgrl.com}}, for modeling argument schemes and critical questions and for reasoning about goal models with respect to them (\textbf{requirement 5}).
+In order to support practitioners in using the RationalGRL framework, we propose a methodology, which consists of developing goal models and posing arguments based on schemes in an integrated way.
+We  implemented a web-based tool for the RationalGRL framework as well, which works on any modern browser and is developed in Javascript.\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
 
-%SG: Need to reword this later.
 The rest of this article is organized as follows.
 Section~\ref{sect:background} introduces our running example and briefly discusses the basics of GRL and argumentation about goals using argument schemes.
-Section~\ref{sect:gmas} contains the case study, an in-depth explanation of how we obtained an initial set of argument schemes and critical questions by coding transcripts from discussions about an information system.
-Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL language metamodel, and examples from the case study that illustrate the framework and its language.
-In Section~\ref{sect:formalframework}, we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and develop various algorithms that change a RationalGRL model according to an argument scheme or a critical question that is applied.
-Section~\ref{sect:methodology} discusses the RationalGRL methodology and a discussion of the prototype tool we have developed for RationalGRL.
+Section~\ref{sect:gmas} contains the case study and an explanation of how we obtained an initial set of argument schemes and critical questions by coding transcripts from discussions about an information system.
+Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL metamodel, and examples from the case study that illustrate the framework and its language.
+In Section~\ref{sect:formalframework}, we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and visa versa, and we develop various algorithms that change a RationalGRL model according to an argument scheme or a critical question that is applied.
+Section~\ref{sect:methodology+tool} discusses the RationalGRL methodology and explain various features of the tool we developed.
 Finally, Section~\ref{sect:discussion} covers related work, future work, and a conclusion.

### 2017-11-06 main.tex
MASTERFIX

--- 2017-10-29 main.tex
@@
 
+\newcommand{\rationalgrlurl}{\url{http://www.rationalgrl.com}}
+
 %\renewcommand\Authfont{\fontsize{10}{14.4}\selectfont}
@@
 \title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee \and Floris J. Bex \and Sepideh Ghanavati}
-\institute{Marc van Zee \at Google Research, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} \and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands \\\email{f.j.bex@uu.nl} \and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu}}
+\author{Marc van Zee  \and Floris J. Bex\and Sepideh Ghanavati}
+\institute{Marc van Zee \at Google Research, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} 
+\and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands \\\email{f.j.bex@uu.nl}
+\and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu} }
 
@@
 \input{background}
-\input{gmas}
+\input{casestudy}
 \input{overview}

### 2017-11-06 methodology.tex
MASTERFIX

@@
 \section{The RationalGRL Methodology and Tool}
-\label{sect:methodology}
+\label{sect:methodology+tool}
 
 In previous sections, we have shown how the RationalGRL framework can capture stakeholder discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
+The previous section contained a formalization of RationalGRL, which forms the starting point of this section.
 In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a methodology (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
@@
 
-We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a (Rational)GRL model, an earlier version of which was presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.
+We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a RationalGRL model, an earlier version of which was presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.
 %Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
@@
 
-An important requirement of our framework is that it has tool support (\textbf{requirement 5}.
-This is for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, there does not yet exist any tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent verification to ensure our formal framework and algorithms actually work, iv) having a light-weight web-based version of GRL is useful for the community in general.
+The final requirement of our framework is that it has tool support (\textbf{requirement 5}).
+This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of integrated tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent verification to ensure our formal framework and algorithms actually work.
 In this section we briefly highlight some features of the tool, and we explain some of its current limitations.
 
-GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}, which is implemented as an Eclipse plugin.
-Although it is a rich tool with many features, it can take significant time to set it up properly and is quite feature-heavy.
-Therefore, we have decided to implement our own lightweight modelling tool as our proof-of-concept.
+\paragraph{Relation to jUCMNav}
+
+As we discussed in Section~\ref{sect:background:grl}, GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}, which is implemented as an Eclipse plugin.
+Since this tool is open-source, many researchers have chosen to extend it in various way.
+However, our tool is not yet integrated into jUCMNav, but consists of a stand-alone web-based version.
+While we aim to integrate our framework into jUCMNav in the future, a web-based version of RationalGRL (and GRL) is useful by itself as well.
+The authors of this article have been involved in a large number of empirical studies involving jUCMNav, and our experience is that it can take significant time for participants to setup the Eclipse environment and install the jUCMNav plugin properly.
+Since our extension does not make use of the many features of jUCMNav, we believe a light-weight web-based version is useful for future empirical research, and it may benefit the community in general.
+Since RationalGRL is an extension of GRL, our web-based version allows for (limited) GRL modeling.
 The tool can be accessed from:
@@
 
-The RationalGRL tool is an open-source\footnote{\url{http://www.github.com/marcvanzee/RationalGRL}}, web-based Javascript application, which runs on all modern browsers.
-The tool is based on our formalization in Section~\ref{sect:formalframework}.
-Because this formalization is very close to the GRL metamodel, the models can be easily exported to jUCMNav.
+\paragraph{Tool overview} The RationalGRL tool is an open-source web-based Javascript application, which runs on all modern browsers.
+It is based on our formalization in Section~\ref{sect:formalframework} and provides export functionality to jUCMNav, using the translation algorithm (Algorithm~\ref{alg:translation:to-grl}).
+It contains all of the GRL elements and links, except for beliefs and actors.
+We have argued in Section~\ref{sect:background:grl} that arguments can be seen as an extension to beliefs, which is the reason why we did not implement them.
+Actors are also missing, but they will be added in the future.
 
 When the tool starts, the user is presented with a screen as in Figure~\ref{fig:tool:overview}.
-This screen shows the palette of elements and links to the left, a canvas on which RationalGRL models can be built, and a pane with details of the currently selected elements on the canvas.
+This screen shows the palette of elements and links (top left pane), a canvas on which RationalGRL models can be built (top right pane), an `Export to GRL' button (bottom left pane), and a details pane of the currently selected elements (bottom right pane).
 Elements and links can be added to the canvas by selecting them on the left and clicking on the canvas, thus capturing the \textsf{INTRO} operation (Section~\ref{sect:formalframework:intro}).
-This makes it possible to build and modify regular GRL models using the tool.
+This makes it possible to build and modify GRL models using the tool.
 
-The details panel can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal element ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
-Clicking the ``Answer'' button next to a critical question allows the user to answer the question.
+In total there are four different types of details panes, which we now explain in turn.
+
+\paragraph{IE and links details pane} Figure~\ref{fig:tool:overview} shows the details pane for the goal `Generate cars'.
+The user can change the name of the element by changing the name and clicking `Rename', which will update the naming history on the bottom of the pane.
+The naming history is (simplified) implemenetation of the \textsf{REPLACE} operation on IEs.
+If the IE has decomposition links, the user can change the decomposition type as well.
+Each IE and link has a set of associated critical questions, which can be answered in the details pane by click on the button next to the question.
+
+\paragraph{Critical question details pane} The details panel can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal element ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
 As an example, consider Figure~\ref{fig:tool:cqdetails}, where the softgoal ``Easy to use'' is questioned with the relevancy question (CQ11).
@@
 This is an implementation of a \textsf{DISABLE} algorithm (Section~\ref{sect:formalframework:disable}) similar to Algorithm~\ref{alg:actor-not-relevant}: a new argument is added that attacks the existing argument.
-It is also possible to attack arguments by adding an ``Argument'' element and an ``Attack'' link manually.
+
+\paragraph{Argument details pane} It is also possible to attack arguments by adding an ``Argument'' element and an ``Attack'' link manually.
 Consider, for example, Figure~\ref{fig:tool:argument}.
@@
 
-The RationalGRL tool computes the acceptability of arguments on the fly.
+\paragraph{Argumentation semantics} The RationalGRL tool computes the acceptability of arguments on the fly.
 In the example of Figure~\ref{fig:tool:cqeffect}, the original (argument for) softgoal ``Easy to use'' is rejected because it's only attacker ``CQ11 - Redundant'' is accepted.
@@
 
-The final operation that has been captured in the RationalGRL tool is \textsf{REPLACE}.
-For usability purposes, this operation has been implemented differently than in our formal framework (Section~\ref{sect:formalframework:disable}).
-In the formal framework, a \textsf{REPLACE} introduces a new argument that replaces (and therefore attacks) all previous arguments with the same identifier.
-Including \textsf{REPLACE} like this in the tool would mean that, when the name of an element is changed, we would need to render all the previous versions of the element on the canvas of the tool.
-Because this would quickly become very cumbersome, we decided to implement a naming history for each element.
-For example, in Figure~\ref{fig:tool:overview}, the goal ``Generate cars'' was previously named ``Add cars''.
-In this way, a history is kept of how an element name was clarified without cluttering the model canvas.
-Similarly, the decomposition type can be changed in the details pane (Figure~\ref{fig:tool:overview}) without introducing new elements that attack the original ones (cf.
-Figure~\ref{fig:example-small3}, where the old decomposition type $AND$ of ``Generate traffic'' is shown as an explicit argument).
+\paragraph{Export to jUCMNav} The tool implements the RationalGRL to GRL translation (Algorithm~\ref{alg:translation:to-grl}) through an export function.
+RationalGRL models built in the RationalGRL tool can be exported to the \texttt{.grl} file format, which can be imported in jUCMNav.
+Note that in order for the GRL models to be rendered correctly, we recommend using the Graphviz dot\footnote{\url{http://www.graphviz.org/}} as the auto-layout tool (this can be selected under ``Auto layout preferences'' when importing the file in jUCMNav).
 
-The tool further implements the RationalGRL to GRL translation (Algorithm~\ref{alg:translation:to-grl}) through an export function.
-RationalGRL models built in the RationalGRL tool can be exported to the \texttt{.grl} file format, which can be imported in jUCMNav.
 Two examples of this export are provided in Appendix~\ref{sect:tool-screenshots}.
@@
 The pair of figures \ref{fig:tool:multipleattack} and \ref{fig:tool:multipleattack1} also nicely shows the added value of RationalGRL: Figure~\ref{fig:tool:multipleattack} shows that there can be a larger discussion and rationalization underlying even a fairly simple goal model such as the one in Figure~\ref{fig:tool:multipleattack1}.
+
+\paragraph{Limitations}
+For usability purposes, the \textsf{REPLACE} operation has been implemented differently than in our formal framework (Section~\ref{sect:formalframework:disable}).
+In the formal framework, a \textsf{REPLACE} introduces a new argument that replaces (and therefore attacks) all previous arguments with the same identifier.
+Including \textsf{REPLACE} like this in the tool would mean that, when the name of an element is changed, we would need to render all the previous versions of the element on the canvas of the tool.
+Because this would quickly become very cumbersome, we decided to implement a naming history for each element.
+For example, in Figure~\ref{fig:tool:overview}, the goal ``Generate cars'' was previously named ``Add cars''.
+In this way, a history is kept of how an element name was clarified without cluttering the model canvas.
+Similarly, the decomposition type can be changed in the details pane (Figure~\ref{fig:tool:overview}) without introducing new elements that attack the original ones (cf.
+Figure~\ref{fig:example-small3}, where the old decomposition type $AND$ of ``Generate traffic'' is shown as an explicit argument).

### 2017-11-06 overview.tex
MASTERFIX

@@
 We now turn to an overview of our RationalGRL framework.
-We will show through a language definition and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
-A more formal, logical rendition of this traceability can be found in Section~\ref{sect:formalframework}.
+We will show through a metamodel and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
+A formalization of our framework using formal logic can be found in Section~\ref{sect:formalframework}.
 
-Figure~\ref{fig:rationalgrl-framework} presents an overview of RationalGRL framework.
-There are two activities (bottom), \emph{practical reasoning \& argumentation} and \emph{goal model construction}, which give rise to two different models (top), a \emph{RationalGRL model} and a \emph{GRL model}.
+Figure~\ref{fig:rationalgrl-framework} presents an overview of the RationalGRL framework.
+There are two activities (bottom), \emph{Practical reasoning \& argumentation} and \emph{Goal model construction}, which give rise to two different models (top), a \emph{RationalGRL model} and a \emph{GRL model}.
 The GRL part of RationalGRL allows for the creation of goal models by analyzing the non-functional requirements and refining the high-level goals into operationalized tasks.
@@
 Similarly, if, for example, we add a new intentional element to the GRL model, it can lead to a new critical question relevant to this intentional element and its relationships.
-Thus, in the framework it becomes possible to trace a goal model back to the original argumentative discussion about goals, tasks and requirements.
+Thus, in the framework it is possible to trace a goal model back to the original discussions about goals, tasks and requirements.
 
@@
 
-In the rest of this section, we discuss the individual parts of the GRL framework.
+In the rest of this section we discuss the individual parts of the GRL framework.
 In Section~\ref{sect:overview:as}, we continue our discussion of the argument schemes and critical questions for practical reasoning and argumentation, fitting these schemes and questions into our framework.
-In Section~\ref{sect:overview:lang}, we then discuss the language for RationalGRL models.
-In Section~\ref{sect:overview:examples}, we then provide extensive examples from our case study, illustrating the interplay between practical reasoning and argumentation on the one hand and RationalGRL models on the other hand.
+In Section~\ref{sect:overview:lang}, we then discuss the language for RationalGRL models an we provide a metamodel, linking the new concepts of our language to GRL.
+In Section~\ref{sect:overview:examples}, we provide extensive examples from our case study.
 
@@
 
-A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual types of reasoning stakeholders perform (requirement 1).
-Recall from section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
+A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual discussions of stakeholders (\textbf{requirement 1}).
+Recall from Section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
 Using this list as a basis, we further refined the set of argument schemes and critical questions for RationalGRL into the list shown in Table~\ref{table:argument-schemes}.
-Note that this list is not exhaustive, and that new argument schemes and questions can be added depending on the problem domain.
 
-Schemes AS0-AS4 and AS12-AS13 are arguments for an element of a goal model, and AS5-AS11 are related to links in a goal model.
+It is important to note that the list we provide here is not exhaustive.
+It is merely the result of our empirical study, but new new argument schemes and questions can be added depending on the problem domain.
+In fact, our tool (Section~\ref{sect:tool}) contains many more critical questions, and it is fully extensible, meaning that new argument schemes and critical questions can be added easily.
+See Section~\ref{sect:discussion:futurework} for more details.
+
+Schemes AS0-AS4 and AS11-AS12 are arguments for an element of a goal model, and AS5-AS10 are related to links in a goal model.
 The last scheme (Att) is a scheme for a generic counterargument against any type of argument that has been put forward.
-Arguments based on these schemes can be used to discuss a goal model.
+Arguments based on the schemes in Table~\ref{table:argument-schemes} can be used to form arguments about elements in a GRL model.
 Making an argument based on one of the schemes effectively adds the corresponding GRL element to the model.
@@
 These questions can be answered with ``yes'' or ``no'', and the type of answer has an effect on the original argument (\textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE}).
-This will be further explained in Section~\ref{sect:overview:examples}.
+This will be further explained in Section~\ref{sect:overview:examples} an in Section~\ref{sect:formalframework}.
 
-\subsection{The RationalGRL Modelling Language}
-\label{sect:overview:lang}\label{sect:metamodel}
+\subsection{The RationalGRL Modeling Language}
+\label{sect:overview:lang}
+\label{sect:metamodel}
 RationalGRL is an extension of GRL and includes all the elements shown in Figure~\ref{fig:grl_legend}.
@@
 \item \emph{Rejected (Disabled) GRL element}: If an argument or GRL element is attacked by an argument that itself is not attacked, then this GRL element will be rejected.
-%\item \emph{Refined GRL Element}: Not all critical questions attack a GRL element.
-It is also possible that a critical question \emph{replaces} an existing element (for instance, by clarifying the name of the element), or that it leads to the \emph{introduction} of a new element.
-In these cases, the corresponding GRL element is shown with a striped background.
-%%Refined GRL elements are not in the metamodel and the tool also has no different picture for refined elements
+Note Figure~\ref{fig:rationalgrllegend} only shows one type of disabled element, a Task, but all the elements (IEs and links) can be disabled in RationalGRL.
 \item \emph{Attack Link}: An attack link can occur between an argument and another argument or GRL element. It means that the source argument attacks the target argument or GRL element.
@@
 \includegraphics{img/legend.pdf}
-\caption{The new elements and link of RationalGRL}
+\caption{The new elements and link of RationalGRL.
+Generic argument (left), attack link (middle), and disabled element (right).
+The disabled element in this figure is a task, but all IEs and links can be disabled in RationalGRL}
 \label{fig:rationalgrllegend}
@@
 \includegraphics[width=\textwidth]{metamodel/metamodel}
-\caption{The RationalGRL metamodel}
+\caption{The RationalGRL metamodel. The \emph{Goal model construction} package (bottom) is the GRL metamodel. The \emph{Practical reasoning \& argumentation} package is the RationalGRL extension}
 \label{fig:metamodel}
@@
 
-The metamodel consists of two packages, \emph{practical reasoning \& argumentation} and \emph{goal model construction}, which correspond to the relevant activities in the RationalGRL framework (cf.
+The metamodel consists of two packages, \emph{Practical reasoning \& argumentation} and \emph{Goal model construction}, which correspond to the relevant activities in the RationalGRL framework (cf.
 Figure~\ref{fig:rationalgrl-framework}).
-The goal model construction package consists of \textsf{GRLModelElements}, which can be either \textsf{GRLLinkableElements} or \textsf{ElementLinks}.
+The goal model construction package is simply the GRL metamodel.
+It consists of \textsf{GRLModelElements}, which can be either \textsf{GRLLinkableElements} or \textsf{ElementLinks}.
 A \textsf{GRLLinkableElement} can again be specialized into an \textsf{Actor} or an \textsf{IntentionalElement} (which is either a \textsf{Softgoal}, \textsf{Goal}, \textsf{Task}, \textsf{Resource}, or a \textsf{Belief}).
@@
 When a critical question is answered, we may obtain an \textsf{AttackLink}, an \textsf{Argument} or both, depending on the answer.
-Note that it is also possible to an \textsf{AttackLink} can also be associated with no critical questions.
+Note that it is also possible for an \textsf{AttackLink} to be associated with no critical questions.
 This allows the user to create attacks between arguments, which do not necessarily correspond to one of the critical questions.
@@
 
-Notice that there are \textsf{OperationTypes} in the Argumentation package.
-In RationalGRL, these operations are performed by instantiating an argument scheme or answering a critical question in a certain way.
+The \textsf{OperationTypes} in the Argumentation package correspond to the operations we informally introduced in Section~\ref{sect:overview}.
+These operations are performed by instantiating an argument scheme or answering a critical question in a certain way.
 An \textsf{INTRO} operation introduces a new RationalGRL element.
@@
 Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements.
-Finally, the relation between \textsf{GRLModelElement} and \textsf{Argument} means that, as we already briefly indicated when discussing the framework in Figure~\ref{fig:rationalgrl-framework}, the class of \textsf{RatinalGRLmodel} is a superclass of \textsf{GRLmodel}: besides arguments about GRL elements, we can also have arguments that does not relate to any of the GRL elements.
+Finally, the relation between \textsf{GRLModelElement} and \textsf{Argument} means that, as we already briefly indicated when discussing the framework in Figure~\ref{fig:rationalgrl-framework}, the class of \textsf{RationalGRLmodel} is a superclass of \textsf{GRLmodel}: besides arguments about GRL elements, we can also have arguments that does not relate to any of the GRL elements.
 
@@
 We provide informal examples of the links between the practical reasoning found in our case study transcripts and RationalGRL models.
-The formal grounding for the connection between practical reasoning and RationalGRL models can be found in the RationalGRL Metamodel (Section~\ref{sect:metamodel}) and in more detail in the logical formalization of the Argument Schemes and Critical Questions (Section~\ref{sect:algorithms}).
+The formal grounding for the connection between practical reasoning and RationalGRL models can be found in the RationalGRL metamodel (Section~\ref{sect:metamodel}) and in more detail in the logical formalization in the next section.
 The connection between the RationalGRL models shown in this section and regular GRL models is further formally defined in Section \ref{sect:formalframework:translation}.
@@
 \includegraphics[width=\columnwidth]{img/fig_example_AS.pdf}
-\caption{Introducing new GRL elements with argumentation (\textsf{INTRO)}.}
+\caption{RationalGRL model (right) with instantiated argument schemes (left): Introducing new elements (operation \textsf{INTRO)}}. 
 \label{fig:example_AS}
@@
 We start by showing how instantiating argument schemes leads to the introduction of new RationalGRL elements in a model.
-Take the example in Figure~\ref{fig:example_AS}, which is based on the excerpt from transcript $t_1$ shown in Table~\ref{table:transcript:decomposition} in Appendix~\ref{sect:transcripts:excerpts}.
-On the left side, the arguments found in the transcript are shown, together with the argument scheme they are based on.
-The participants in the discussion argue that \emph{System} has a goal and two tasks, and that the goal AND-decomposes into the two tasks.
-By arguing in this way, new GRL elements are introduced.
+Take the example in Figure~\ref{fig:example_AS}, which is based on the excerpt from transcript $t_3$ shown in Table~\ref{table:transcript:decomposition} in Appendix~\ref{sect:transcripts:excerpts}.
+On the left side of the image, the arguments found in the transcript are shown, together with the argument scheme they are based on.
+The participants in the discussion argue that actor \emph{System} has a goal and two tasks, and that the goal AND-decomposes into the two tasks.
+By putting forward these arguments, new GRL elements are introduced.
 These GRL elements are shown on the right side of Figure~\ref{fig:example_AS}; the dashed arrows indicate the links between the practical reasoning and argumentation on the left and the RationalGRL model on the right.
@@
 \includegraphics[]{img/fig_example_disable.pdf}
-\caption{Disabling GRL tasks with argumentation (\textsf{DISABLE)}.}
+\caption{Disabling elements (operation \textsf{DISABLE)}.}
 \label{fig:example_disable}
@@
 Table~\ref{table:argument-schemes}) should disable the original argument based on AS2 by attacking it.
-On the left side of Figure~\ref{fig:example_disable} a new argument (CQ11) attacks the original argument based on (AS2).
+On the left side of Figure~\ref{fig:example_disable} a new argument (CQ11) attacks the original argument based on argument scheme AS2.
 This new argument is also added to the RationalGRL model on the right of Figure~\ref{fig:example_disable}, where it attacks the original task \emph{Add traffic light}.
@@
 \includegraphics[]{img/fig_example_replace.pdf}
-\caption{Replacing a GRL decomposition with argumentation (\textsf{REPLACE)}.}
+\caption{Replacing an element (operation \textsf{REPLACE)}.}
 \label{fig:examples:decomposition}
@@
 Recall that in Example 1, an AND-decomposition was introduced for this goal with AS5 (Figure~\ref{fig:example_AS}).
-In the discussion CQ10b -- ``Is the decomposition type correct?'' -- is explicitly asked.
+In the discussion CQ5c -- ``Is the decomposition type correct?'' -- is explicitly asked.
 The answer is ``No, it should be OR''.
@@
 \includegraphics[]{img/fig_example_rename.pdf}
-\caption{Renaming a GRL task with argumentation (\textsf{REPLACE)}.}
+\caption{Renaming an element name (operation \textsf{REPLACE)}.}
 \label{fig:examples:clarification}
@@
 \includegraphics[]{img/reinstate1.pdf}
-\caption{Disabling a GRL actor with argumentation (\textsf{DISABLE)}.}
+\caption{Disabling an element (operation \textsf{DISABLE)}.}
 \label{fig:examples:relevant-actor}
@@
 \includegraphics[]{img/reinstate2.pdf}
-\caption{Defending a GRL actor by attacking its disabling attacker (\textsf{DISABLE)}.}
+\caption{Defending an element by attacking its disabling attacker (operation \textsf{DISABLE)}.}
 \label{fig:examples:relevant-actor2}
@@
 Further in the discussion, it is then argued that the development team should be considered, since the professor does not develop the software.
-This is captured using a generic counterargument (\emph{Att}, which attacks the earlier argument based on CQ0.
+This is captured using a generic counterargument \emph{Att}, which attacks the earlier argument based on CQ0.
 Figure~\ref{fig:examples:relevant-actor2} shows the situation after the counterargument has been put forward: the argument (Att) now attacks the argument (CQ0), which in turn attacks the original argument (AS0).

### 2017-11-06 casestudy.tex
MASTERFIX

@@
 Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
-To get a sense of such discussions, we performed an extensive case study to see which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
+To get a sense of such discussions, we performed a case study to see which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
 We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS.

### 2017-11-06 methodology.tex
MASTERFIX

@@
 
+\paragraph{Argument schemes and critical questions} The tool contains more argument schemes and critical questions than those that we initial collected in Table~\ref{table:argument-schemes}.
+As we mentioned before, our table is not meant to be exhaustive, and it is straightforward to add more argument schemes and critical questions if necessary.
+For instance, the table does not contain critical questions for links between all IEs (e.g., a contribution link from a task to a goal does not occur in the table).
+In the tool, we have added critical questions to all links and IEs.
+We implemented an \emph{argument schemes database}, which is designed to be extended easily.
+
 \paragraph{Export to jUCMNav} The tool implements the RationalGRL to GRL translation (Algorithm~\ref{alg:translation:to-grl}) through an export function.

### 2017-11-07 background.tex
small fixes

--- 2017-10-29 background.tex
@@
 
-In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.
+In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework\footnote{Note that in this article, we consider a slightly simplified version of GRL.}.
 Next, we introduce argumentation: we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
-Finally, we briefly discuss the possibilities integrating PRAS and GRL.
 %This will be our starting point in the next section.
@@
 \label{sect:background:grl}
-GRL is a visual modeling language for specifying intentions, business goals, and non-functional requirements of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
+GRL is a modeling language for specifying intentions, business goals, and non-functional requirements of multiple stakeholders \cite{Amyot:2010:EGM:1841349.1841356}.
 GRL is part of the User Requirements Notation, an ITU-T standard, that combines goals and non-functional requirements with functional and operational requirements (i.e.
@@
 Tasks represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, we have the two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve goal \emph{Generate cars}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off screen.
+In Figure~\ref{fig:example-small}, there are two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve goal \emph{Generate cars}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off screen.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources to be available (e.g., \emph{Car Objects}).
+Finally, the full version of GRL allows design rationale to be captured using beliefs.
+Because we capture the reasoning and rationales behind goal models using arguments, we do not include beliefs in our simplified version of GRL (but see the discussion at the end of this section).
 
@@
 Contribution links indicate impacts of one element on another element, which can be positive or negative.
-Task \emph{Create new cars} has a positive contribution to the task \emph{Realistic simulation}, and a negative contribution to the task \emph{Simple design}.
+Task \emph{Create new cars} has a positive contribution to the softgoal \emph{Realistic simulation}, and a negative contribution to the softgoal \emph{Simple design}.
+Note that the full GRL specification considers different levels of positive and negative contributions, which are not directly relevant for current purposes.
 Dependency links model relationships between actors or resources.
@@
 Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
-GRL is also linked to Use Case Maps, which provide traceability between concepts and instances of the goal model and behavioral design models.
+GRL also has the capability to be extended through metadata, links, and external OCL constraints.
+This allows GRL to be used in many domains without the need to change the whole modeling language.
+For example, GRL is linked to Use Case Maps, which provide traceability between concepts and instances of the goal model and behavioral design models.
 Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
@@
 GRL has six evaluation algorithms which allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements quantitatively, qualitatively or in a hybrid way.
-jUCMNav, GRL tool-support, also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
-GRL also has the capability to be extended through metadata, links, and external OCL constraints.
-This allows GRL to be used in many domains without the need to change the whole modeling language.
+jUCMNav, GRL tool-support (see Section~\ref{sect:tool}), also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
+For simplicity's sake, the version of GRL used in this paper does not take into account satisfaction values.
 
@@
 Furthermore, beliefs are after-the-fact design rationales and do not capture the types of critical questions given above.
+Finally, the jUCMNav tool does not beliefs in the evaluation algorithms, and thus there is no formal connection between goal models and their underlying beliefs and opinions.
+For a more detailed comparison of our framework with GRL beliefs, see Section~\ref{sect:discussions:relatedwork}.
 
@@
 \begin{itemize}
-\item[] \emph{Generate Cars} is a goal,
+\item[] \emph{Generate cars} is a goal,
 \item[] Performing action \emph{Keep same cars} realizes goal \emph{Generate cars}, 
@@
 For example, we may argue that, in fact, performing action \emph{Keep same cars} does not realize goal \emph{Generate cars}, thus giving a counterargument to the above instantiation of PRAS.
-Atkinson et al.~\cite{atkinson2007} define a set of so-called critical questions that point to typical ways in which an argument based on PRAS can be criticized by.
+Atkinson et al.~\cite{atkinson2007} define a set of so-called critical questions that point to typical ways in which an argument based on PRAS can be criticized.
 Some examples of critical questions are as follows.
@@
 In argumentation, counterarguments are said to \emph{attack} the original arguments.
-Given a set of arguments and attacks between these arguments, we can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Formal definitions of argumentation frameworks and semantics will be given in section \ref{sect:gmas}.
-In this section, we will briefly discuss the intuitions behind these concepts.}.
+Given a set of arguments and attacks between these arguments, we can compute which arguments are accepted and which are rejected using different argumentation semantics~\cite{Dung1995}\footnote{Formal definitions of argumentation frameworks and semantics will be given in Section~\ref{sect:gmas}.
+In this section, we briefly discuss the intuitions behind these concepts.}.
 Figure \ref{fig:pras:example} shows three arguments from the traffic simulation example, where arguments are rendered as boxes and attack relations as arrows.
@@
 Looking at PRAS and its critical questions, one can see how it could be used to argue about goals and actions or, more specifically, about goal models.
-However, we cannot literally use PRAS and its critical questions, as there are elements of the GRL language, such as actors and resources, which cannot be found in PRAS.
+In fact, in some of our previous work on RationalGRL \cite{vanzee-etal:renext2015,vanZee-etal:er2016} we used PRAS-arguments such as the ones above to capture reasoning about goals, and provided a translation to goal models.
+However, one problem of this is that we cannot literally use PRAS and its critical questions, as there are elements of the GRL language, such as actors and resources, which cannot be found in PRAS.
 Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson and Bench-Capon~\cite{atkinson2007} actually apply to GRL models.

### 2017-11-07 discussion.tex
small fixes

@@
 
-The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
-Jureta \emph{et al.} propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
+The contribution most closely related to ours is the work by Jureta et al.~\cite{Jureta:RE2008}.
+Jureta et al.
+propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
 GAM is a high-level decision process, in which alternative solutions for an RE problem are evaluated and compared using argumentation.
@@
 thesaurus checks, labeling vague expressions).
-Jureta et al.
-then present the argumentation part of their framework, where reasons (justifications) for conclusions are given as formal structured arguments\footnote{Informally, a structured argument is similar to the PRAS argument in Section~\ref{sect:background:pras}, i.e., $a, b \xrightarrow{therefore} c$.}.
+
+The GAM process is essentially a generic, high-level process for problem solving, not specifically tailored towards goal modeling.
+In this sense, the RationalGRL methodology in Figure~\ref{fig:rationalgrl-methodology} can be seen as a more specific version of GAM explicitly meant for goal modeling.
+The argument schemes and critical questions in RationalGRL provide more clear handles for goal modeling (cf.
+requirement 4 of our framework, Section~\ref{sect:introduction}).
+Furthermore, the RationalGRL framework adds full tool support for the goal modeling and reasoning process, something that GAM is lacking.
+GAM does contain more specific ways of dealing with different types of clarity problems; in RationalGRL clarification is captured as a single critical question (CQ12, Table~\ref{table:argument-schemes}).
+However, as argued above further critical questions can easily be added to RationalGRL if needed.
+
+Jureta et al.'s framework also includes an argumentation part, where reasons (justifications) for conclusions are given as formal structured arguments\footnote{Informally, a structured argument is similar to the PRAS argument in Section~\ref{sect:background:pras}, i.e., $a, b \xrightarrow{therefore} c$.}.
 Arguments and alternatives are then captured as structured arguments or argument diagrams with reasons for or against goals and tasks.
@@
 
-The GAM process is essentially a generic, high-level process for problem solving, not specifically tailored towards goal modeling.
-In this sense, the RationalGRL methodology in Figure~\ref{fig:rationalgrl-methodology} can be seen as a more specific version of GAM explicitly meant for goal modeling.
-The argument schemes and critical questions in RationalGRL provide more clear handles for goal modeling (cf.
-requirement 4 of our framework, Section~\ref{sect:introduction}).
-
-In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we extended Jureta et al.'s work and translated argument diagrams to GRL models\footnote{In \cite{Jureta:RE2008} only a mapping from goal models to structured arguments is given, and the step from structured arguments or argument diagrams to goal models is never formally defined.} (an automatic translation tool is discussed in~\cite{vanZee-etal:er2016}).
-Thus, we have essentially two complex diagrams, an argument diagram and a goal diagram, and a mapping between them.
+RationalGRL allows for two-way translation between arguments about goals and standard goal models (cf.
+Section~\ref{sect:formalframework:translation}).
+In contrast, Jureta et al.
+only provide a mapping from goal models to structured arguments, and the step from structured arguments or argument diagrams to goal models is never formally defined.
+In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we extended Jureta et al.'s work and translated argument diagrams to GRL models (an automatic translation tool is discussed in~\cite{vanZee-etal:er2016}).
+This gives us two complex diagrams, an argument diagram and a goal diagram, and a mapping between them.
 This was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
@@
 The current tool is a prototype that implements the RationalGRL framework in a fairly simple way, that is, by offering the possibility to argue and ask critical questions about a goal model, and then export this goal model for further analysis in jUCMNav.
-Apart from implementing the tool to include the new fatures described above (e.g.
-new argument schemes, reasoning with preferences), it is also interesting to think about further possibilities.
+Integration of RationalGRL elements into jUCMNav, or at least creating an import function to import jUCMNav goal models into the tool would allow us to tap into the large amount of research and development that is being carried out with jUCMNav.
+Furthermore, the tool could be expanded to include the new features described above, such as new argument schemes and reasoning with preferences.
+In addition to such extensions, it is interesting to think about new possibilities.
 One idea is to use the tool for collaborative on-line goal modeling, similar to Github: because the renaming (i.e.
@@
 
-This paper presents RationalGRL: a framework for integrated goal modeling and argumentation.
-We extend the well-known goal modeling language GRL with argument schemes and critical questions.
-\emph{Argument schemes} are reusable patterns of reasoning that capture the common approaches in which humans argue and reason.
-\emph{Critical questions}, which are associated with argument schemes, point to typical sources of doubt or implicit assumptions people make when raising an argument.
+The process of constructing a goal model involves discussions between a requirements engineer and a group of stakeholders.
+While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
+This makes it not only more difficult to understand the model, but other stakeholders may end up having similar discussions throughout the design and development phase as well.
+Furthermore, the disconnect between goal models and their underlying beliefs and opinions may lead to a poor understanding of the problem and solution domain.
 
-We argue that the argument scheme for practical reasoning~\cite{walton1990,atkinson2007} is particularly useful when reasoning about goals.
-Inspired by the work on practical reasoning from Artificial Intelligence, most notably by Atkinson and Bench-Capon~\cite{atkinson2007}, we develop a list of argument schemes that can be used to analyze and guide stakeholders' discussions about goal models.
+In order to solve these problems, we present RationalGRL: a framework for integrated goal modeling and argumentation.
+We extend the well-known goal modeling language GRL with argument schemes and critical questions that can be used to analyze and guide stakeholders' discussions about goal models.
 Our approach, thus, provides a rationalization for the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected.

### 2017-11-07 formalframework.tex
small fixes

@@
 Throughout this section we use the example in Figure~\ref{fig:example-small}. This example contains various IEs which can be formalized using Definition~\ref{def:ie}, for example, 
-$(1, Resource, \text{Car objects}, AND)$, $(2, Softgoal,$ Realistic simulation$, AND)$ and $(5, Goal,$ Generate traffic$, XOR)$.
+$(1, Resource, \text{Car objects}, AND)$, $(2, Softgoal,$ Realistic simulation$, AND)$ and $(5, Goal,$ Generate cars$, XOR)$.
 
@@
 %\begin{itemize}
-%\item $goal_4.name = \text{Generate traffic}$
+%\item $goal_4.name = \text{Generate cars}$
 %\item $goal_4.decomptype = XOR$
@@
 &   &  &(4, Softgoal, \text{Easy to use}, AND),&\\
-&   &  &(5, Goal, \text{Generate traffic}, XOR),&\\
+&   &  &(5, Goal, \text{Generate cars}, XOR),&\\
 &   &  &(6, Task, \text{Create new cars}, AND),&\\
@@
 
-Examples of arguments in Figure~\ref{fig:example-small3} are $A_5 = (5,$ $Goal,$ Generate traffic$, AND)$, $A'_5 = (5, Goal,$ Generate traffic$, XOR)$, $A_4 = (4, Softgoal,$ Easy to use$, AND)$ and $A_{13}=(13,$ $ArgElem,$ Redundant$)$.
+Examples of arguments in Figure~\ref{fig:example-small3} are $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, $A'_5 = (5, Goal,$ Generate cars$, XOR)$, $A_4 = (4, Softgoal,$ Easy to use$, AND)$ and $A_{13}=(13,$ $ArgElem,$ Redundant$)$.
 Note that the two arguments $A_5$ and $A'_5$ show an important difference between RationalGRL models and valid GRL models: While a valid GRL model disallows multiple elements with the same identifier (Definition~\ref{def:valid-grl-model}, condition 1), RationalGRL models do not enforce this restriction.
@@
 
-As an example, take the arguments $A_5$ for the goal `Generate traffic' ($AND$ decomposition), $A'_5$ for the goal `Generate traffic' ($XOR$ decomposition), $A_4$ for the softgoal `Easy to use' and the generic argument $A_{13}$ (`Redundant') against the softgoal (`Easy to use').
+As an example, take the arguments $A_5$ for the goal `Generate cars' ($AND$ decomposition), $A'_5$ for the goal `Generate cars' ($XOR$ decomposition), $A_4$ for the softgoal `Easy to use' and the generic argument $A_{13}$ (`Redundant') against the softgoal (`Easy to use').
 Given these arguments there are two attack links (Figure~\ref{fig:example-small3}), namely $(A'_{5},A_{5})$ and $(A_{13}, A_{4})$.
@@
 A_4 = &(4, Softgoal, \text{Easy to use}, AND),&\\
-A_5 = &(5, Goal, \text{Generate traffic}, AND),&\\
-A'_5 = &(5, Goal, \text{Generate traffic}, XOR),&\\
+A_5 = &(5, Goal, \text{Generate cars}, AND),&\\
+A'_5 = &(5, Goal, \text{Generate cars}, XOR),&\\
 A_6 = &(6, Task, \text{Create new cars}, AND),&\\
@@
 
-As an example, suppose we are adding the decomposition of goal ``Generate traffic'', expressed as argument $A_5 = (5,$ $Goal,$ Generate traffic$, AND)$, into task ``Keep same cars'', and suppose we have the following argument for the goal: .
+As an example, suppose we are adding the decomposition of goal ``Generate cars'', expressed as argument $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, into task ``Keep same cars'', and suppose we have the following argument for the goal: .
 Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars}, AND)$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.

### 2017-11-07 introduction.tex
small fixes

@@
 
-The first challenge with the goal modeling phase, particularly in large projects, is related to their dynamic nature: goal models continuously change and evolve.
+The first challenge for the goal modeling phase, particularly in large projects, is related to its dynamic nature: goal models continuously change and evolve.
 Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
@@
 If the rationales behind these revisions are not properly documented, alternative ideas and opposing views that could potentially lead to different goal models might be lost, as the resulting goal model only shows the end product of a long process and not the discussions during the modeling process.
-Furthermore, other stakeholders, such as developers who were not the original authors of the goal model, may have to make sense of a goal model in order to, for example, use it as an input in a later RE stage or in the development phase.
-If preferences, opinions and rationales behind the goal models are not stored explicitly, it may not only be more difficult to understand the model, but the other stakeholders may end up having similar discussions throughout the design and development phase as well.
+Furthermore, other stakeholders, such as developers who were not the original authors of the goal model, may have to make sense of a goal model in order to, for example, use it as an input in a later RE or development phase.
+If the preferences, opinions and rationales behind the goal models are not stored explicitly, it may not only be more difficult to understand the model, but the other stakeholders may end up having similar discussions throughout the design and development phase as well.
 
 Another challenge is that current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
-A stakeholder may change his or her opinion, but it is not always directly clear what its effect is on a goal model.
+A stakeholder may change his or her opinion, but it is not always directly clear what the effect of this is on a goal model.
 Similarly, with existing goal modeling languages one can change a part of a goal model without being able to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
@@
 \item 
-The framework must have a tool support.
+The framework must have tool support. %"a tool support" is incorrect
 \end{enumerate}
@@
 Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing framework for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
-Argument schemes are reusable patterns of reasoning that capture the common approaches in which humans argue and reason.
-We develop a list of argument schemes by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
+Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.
+Associated with argumentation schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way.
+We believe that argument schemes are are a good choice for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{bexEtal2003,murukannaiah2015}.
+
+Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyze and guide stakeholders' discussions about goal models.
+This list of argument schemes is based on an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
+
 In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models.
@@
 In order to support practitioners in using the RationalGRL framework, we propose a methodology, which consists of developing goal models and posing arguments based on schemes in an integrated way.
-We  implemented a web-based tool for the RationalGRL framework as well, which works on any modern browser and is developed in Javascript.\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
+We implement a web-based tool for the RationalGRL framework as well, which works on any modern browser and is developed in Javascript.\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
 

### 2017-11-07 methodology.tex
small fixes

@@
 In this section we briefly highlight some features of the tool, and we explain some of its current limitations.
-
-\paragraph{Relation to jUCMNav}
-
-As we discussed in Section~\ref{sect:background:grl}, GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}, which is implemented as an Eclipse plugin.
-Since this tool is open-source, many researchers have chosen to extend it in various way.
-However, our tool is not yet integrated into jUCMNav, but consists of a stand-alone web-based version.
-While we aim to integrate our framework into jUCMNav in the future, a web-based version of RationalGRL (and GRL) is useful by itself as well.
-The authors of this article have been involved in a large number of empirical studies involving jUCMNav, and our experience is that it can take significant time for participants to setup the Eclipse environment and install the jUCMNav plugin properly.
-Since our extension does not make use of the many features of jUCMNav, we believe a light-weight web-based version is useful for future empirical research, and it may benefit the community in general.
-Since RationalGRL is an extension of GRL, our web-based version allows for (limited) GRL modeling.
 The tool can be accessed from:
@@
 \end{quote}
+
+\paragraph{Relation to jUCMNav}
+
+GRL is implemented in the open-source tool jUCMNav~\cite{jUCMNav}, which is actively maintained\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}.
+jUCMNav currently has over 2.000 commits made by over 40 contributors, representing over 200.000 lines of code.
+jUCMNav is used actively in research, and the tool has been extended in many directions over the past few years, such as for modeling and analyzing security misuse cases \cite{}, supporting activity theory \cite{}, Web-based goal modeling \cite{}, and enterprise architecture modeling \cite{}.
+\todo{Floris}{Marc}{references need to be added here!} Furthermore, jUCMNav includes Use Case modelling and contains various GRL evaluation algorithms with which the satisfiability of goals can be determined.
+
+The jUCMNav tool is implemented as an Eclipse plugin.
+While we aim to integrate our framework into jUCMNav in the future, a web-based version of RationalGRL (and GRL) is more suited for current purposes.
+For example, it is our experience is that it can take significant time to set up the Eclipse environment and install the jUCMNav plugin, which hampers quick uptake by practitioners and also makes setting up empirical evaluations more difficult.
+Since RationalGRL does not make use of the many features of jUCMNav, we believe a light-weight version is sufficient for our current aims, and may benefit the community in general.
 

### 2017-11-07 discussion.tex
small fixes conclusion

@@
 Our approach, thus, provides a rationalization for the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected.
-Our list of argument schemes was constructed by performing an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
-This ensures that the argument schemes we propose are close to actual discussions the stakeholders have (\textbf{requirement 1}).
+In the introduction, we identified five important requirements for our framework.
+Below, we discuss how the RationalGRL framework meets these requirements.
 
-The meta-model of the RationalGRL framework clearly specifies the traceability links between the arguments based on the schemes and the GRL models (\textbf{requirement 2}).
-In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
+Our list of argument schemes and critical questions in Table~\ref{table:argument-schemes} was constructed by performing an extensive case study in which we analyzed 153 pages of transcripts of discussions among designers of a traffic simulator information system.
+In this way, we ensure that he argumentation techniques capture the actual discussions of the stakeholders or designers in the early requirements engineering phase (\textbf{requirement 1}).
+
+The metamodel of the RationalGRL framework in Figure~\ref{fig:metamodel} clearly specifies the formal traceability links elements of the GRL goal model and the underlying arguments.
+(\textbf{requirement 2}).
+In addition to this metamodel, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
 We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.

### 2017-11-09 introduction.tex
Update introduction.tex
very small change (mention semantics rights after reqs)

@@
 
-Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing framework for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with a technique from argumentation theory called \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}).
+Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing framework for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}.
 Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.
-Associated with argumentation schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way.
-We believe that argument schemes are are a good choice for modeling discussions about a goal model, as they can guide users in systematically deriving conclusions and making assumptions explicit~\cite{bexEtal2003,murukannaiah2015}.
+Associated with argument schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way.
+These critical questions can point to counterarguments, and the acceptability of sets of arguments can be computed using argumentation semantics.
+Thus computational argumentation based on schemes can guide users in systematically deriving conclusions and making assumptions explicit~\cite{bexEtal2003,murukannaiah2015}.
 
@@
 In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models.
-In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework~\cite{Dung1995}.
+In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework to which we can apply argumentation semantics~\cite{Dung1995}.
 We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.

### 2017-11-12 casestudy.tex
Minor changes, typos, improvements

@@
 Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
-To get a sense of such discussions, we performed a case study to see which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
+To get a sense of such discussions, we performed a case study to examine which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
 We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS.
@@
 
-Before we started coding the transcripts, we drew up an initial list of 11 argument schemes (AS0-AS10 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
+Before we started coding the transcripts, came up with an initial list of 11 argument schemes (AS0-AS10 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
+We used no particular method to develop this initial list besides our own intuition about which part of a goal model would be likely to be discussed.
+
 AS0 to AS4 are schemes that concern a single element of the goal model.
-So, for example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'.
+For example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'.
 AS5 to AS10 are claims about the links between GRL elements.
@@
 For example, we found that the discussants often talk about tasks decomposing into sub-tasks, so we added AS10 and CQ10a.
-Furthermore, because there were many discussions on the relevance and the clarity of the names of elements, two generic critical questions CQ12 and CQ13 were added.
+Furthermore, since there were many discussions on the relevance and the clarity of the names of elements, two generic critical questions CQ12 and CQ13 were added.
 The final results of the coding can be found in Table~\ref{table:transcripts:results:argumentschemes}.
@@
 
-Our coding further led us to identify three different operations, that is, different effects an argument or critical question can have on a goal model: an argument can introduce a new goal model element (\textsf{INTRO}); it can disable (i.e.
-attack) a goal model element (\textsf{DISABLE}); or it can replace a goal model element (\textsf{REPLACE}).
+Our coding further led us to identify three different operations, that represent different effects an argument or critical question can have on a goal model: an argument can introduce a new element in the goal model (\textsf{INTRO}); it can disable (i.e.
+attack) a goal model element (\textsf{DISABLE}); or it can replace an element in the goal model with another one (\textsf{REPLACE}).
 Consider, for example, Table~\ref{table:transcripts:traffic-light} in Appendix~\ref{sect:transcripts:excerpts}.

### 2017-11-12 discussion.tex
Update discussion.tex
1 sentence about arg semantics
Minor changes, typos, improvements

@@
 
-\paragraph{Design Rationale} Argumentation in software design has for some time now been the subject of the work on so-called \emph{design rationale} (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or software architecture.
+\paragraph{Design Rationale} Argumentation in software design has been the subject of the research on so-called \emph{design rationale} (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or software architecture.
 DR looks at issues, options and arguments for and against these options in the design of, for example, a software system.
@@
 It is in this diagramming functionality that the link with argument diagrams from philosophy, law and AI~\cite{reed2004araucaria,gordon2007visualizing} has been made, where argument diagrams represent reasoning from premises to conclusions.
-More recent work on DR moves away from the idea that all decision have to be explicitly diagrammed and focuses more on empirically investigating how critical reflection can help when designing \cite{razavian2016two,SchriekEtal2016}, or which parts of the design process are best explicitly documented \cite{falessi2013value}.
+More recent research on DR moves away from the idea that all decision have to be explicitly diagrammed and focuses more on empirically investigating how critical reflection can help when designing \cite{razavian2016two,SchriekEtal2016}, or which parts of the design process are best explicitly documented \cite{falessi2013value}.
 
 Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE.
-The work on the RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments pro and con the various options into the goal model, and by proposing a methodology and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
+The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a methodology and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
 
 \paragraph{Requirements Engineering} There are a number of general approaches in the field of requirements engineering that explicitly take into account arguments.
-One early example comes from Haley et al.~\cite{haley2008security}, who use formal logical arguments to show that the system behavior satisfies certain security requirements, and more informal arguments to capture and validate the assumptions underlying the system behavior.
+One early example comes from Haley \emph{et al.}~\cite{haley2008security}, who use formal logical arguments to show that the system behavior satisfies certain security requirements, and more informal arguments to capture and validate the assumptions underlying the system behavior.
 This system behavior is defined by the tasks it executes and thus arguments are given for and against system tasks, similar to the way beliefs and counterarguments can be provided for tasks in RationalGRL.
-What Haley et al.
-leave implicit in their argumentation are the goals of the stakeholders on which the system tasks depend -- they include the goals in their framework and mention that there will often be conflicting goals between stakeholders, but do not explicitly model them.
+What Haley \emph{et al.} leave implicit in their argumentation are the goals of the stakeholders on which the system tasks depend -- they include the goals in their framework and mention that there will often be conflicting goals between stakeholders, but they do not explicitly model them.
 Furthermore, the argumentative part of their framework does not include formal semantics for resolving conflict between arguments or determining the acceptability of arguments.
-Yu et al.~\cite{yu2015automated} further extend the framework by Haley et al., including algorithms for Dung-style argumentation semantics \cite{Dung1995} and a database of specific ways in which to attack (or mitigate) risks, which can be likened to a set of critical questions for risks and security requirements (cf.
-Yu et al.~\cite{yu2015automated} Section 3.1).
+Yu \emph{et al.}~\cite{yu2015automated} further extend the framework by Haley \emph{et al.}, including algorithms for Dung-style argumentation semantics \cite{Dung1995} and a database of specific ways in which to attack (or mitigate) risks.
+This extension can be likened to a set of critical questions for risks and security requirements (cf.
+Yu \emph{et al.}~\cite{yu2015automated} Section 3.1).
 
-Another recent example of the use of arguments in goal-oriented requirements engineering is the work by Murukannaiah et al.~\cite{murukannaiah2015}, who propose Arg-ACH, where the beliefs underlying conflicting goals can be made explicit using argumentation.
-Murukannaiah et al.
-start with the basic technique of Analysis of Competing Hypotheses (ACH), where for conflicting goals the beliefs that are consistent and inconsistent with these goals are included in a matrix and counted.
+Another recent example of the use of arguments in goal-oriented requirements engineering is the work by Murukannaiah \emph{et al.}~\cite{murukannaiah2015}, who propose Arg-ACH, where the beliefs underlying conflicting goals can be made explicit using argumentation.
+Murukannaiah \emph{et al.} start with the basic technique of Analysis of Competing Hypotheses (ACH), where for conflicting goals the beliefs that are consistent and inconsistent with these goals are included in a matrix and counted.
 They then extend this technique into Arg-ACH: instead of just indicating whether a belief is consistent or inconsistent with a goal, each belief becomes an argument for or against the goal, which is then diagrammed using the Carneades tool \cite{gordon2007visualizing}.
@@
 One example provided in the paper is the argument scheme from expert opinion, which allows one to draw conclusions based on expert statements and subsequently question, for example, the objectivity and veracity of the expert using critical questions.
-Murukannaiah et al.
-conducted an experiment in which they had two groups, one with ACH and one with Arg-ACH, perform an analysis of several conflicting goals regarding security at transport hubs.
+Murukannaiah \emph{et al.} conducted an experiment in which they had two groups, one with ACH and one with Arg-ACH, perform an analysis of several conflicting goals regarding security at transport hubs.
 They found that, while the group that used Arg-ACH took longer, they also covered more possibilities in their belief search and the conclusions were more consistent among the group.
@@
 One other example of argumentation in RE concerns the use of argumentation in requirements elicitation.
-Ionita et al.~\cite{ionita2014argumentation} propose an simple argumentation dialogue game in which risk assessors try to attack each other's arguments for certain risks attached to a system design.
+Ionita \emph{et al.}~\cite{ionita2014argumentation} propose a simple argumentation dialogue game in which risk assessors try to attack each other's arguments for certain risks attached to a system design.
 Dung's semantics \cite{Dung1995} are then used to determine the risks that are still valid, and those that have been successfully rejected.
-Elrakaiby et al.~\cite{ElrakaibyFSGN17} use argumentation to explain ambiguity.
-They define when a statement by a client who is being interviewed about the requirements of a system presents an inconsistency (either with the client's previous statements or with the requirement engineer's beliefs) or an insufficiency, that is, when an analyst needs more information from the client to accept a client's statement.
-The inconsistencies are then captured as mutually attacking arguments, and the insufficiencies as arguments against the original statements saying that, for example, the functionality expressed in the statement cannot be realised or is irrelevant.
-Elrakaiby et al.
-coded the data from 34 requirement elicitation interviews, identifying 39 inconsistencies (i.e.
+Elrakaiby \emph{et al.}~\cite{ElrakaibyFSGN17} use argumentation to explain ambiguity.
+They define when a statement by a client who is being interviewed about the requirements of a system presents an inconsistency or an insufficiency.
+An inconsistency can be either with the client's previous statements or with the requirement engineer's beliefs.
+An insufficiency occurs when an analyst needs more information from the client to accept a client's statement.
+The inconsistencies are then captured as mutually attacking arguments, and the insufficiencies as arguments against the original statements.
+These counterarguments say that, for example, the functionality expressed in the statement cannot be realized or is irrelevant.
+Elrakaiby \emph{et al.} coded the data from 34 requirement elicitation interviews.
+They identified 39 inconsistencies (i.e.
 at least two arguments that mutually attack) and 29 insufficiencies (i.e., at least one argument attacking another).
@@
 It is clear from this literature that arguments play a core role in RE.
-Murukannaiah et al.~\cite{murukannaiah2015} show that critical reflection using argument schemes and critical questions -- in the same way that RationalGRL proposes -- improves the quality of the reasoning in the RE process.
-Elrakaiby et al.~\cite{ElrakaibyFSGN17} provide a case study similar to the current one, identifying, as we did, many counterarguments specifically with respect to realisability, relevance and clarity (cf.
+Murukannaiah \emph{et al.}~\cite{murukannaiah2015} show that critical reflection using argument schemes and critical questions -- in the same way that RationalGRL proposes -- improves the quality of the reasoning in the RE process.
+Elrakaiby \emph{et al.}~\cite{ElrakaibyFSGN17} provide a case study similar to our research, identifying, many counterarguments specifically with respect to realisability, relevance and clarity (cf.
 CQ2a, CQ3, CQ11 and CQ12 in Table~\ref{table:argument-schemes}).
-Like in RationalGRL, the use of Dung-style argumentation semantics to compute the acceptable claims in RE is further also advocated by the literature \cite{yu2015automated,ionita2014argumentation,ElrakaibyFSGN17}.
+Similar to RationalGRL, the use of Dung-style argumentation semantics to compute the acceptable claims in RE is also advocated by the literature \cite{yu2015automated,ionita2014argumentation,ElrakaibyFSGN17}.
 
@@
 At the moment, there is only a ``generic argument'' in addition to arguments about goals and tasks.
-However, new argument schemes and critical questions about for example, security risks or expert opinions, can be easily added: the metamodel (Figure~\ref{fig:metamodel}) accommodates this and the formal specification in Section~\ref{sect:formalframework} is set up in such a way that extending the definition of argument and adding new algorithms for specific critical questions is easy.
+However, new argument schemes and critical questions about, for example, security risks or expert opinions, can be easily added: the metamodel (Figure~\ref{fig:metamodel}) accommodates this and the formal specification in Section~\ref{sect:formalframework} is set up in such a way that extending the definition of argument and adding new algorithms for specific critical questions is easy.
 
@@
 Take, for example, CQ2 of PRAS (Section~\ref{sect:background:pras}), which asks whether there are alternative ways of realizing the same goal.
-Providing an alternative sub-goal or -task in a goal model then an explicit argumentative move in the discussion.
-So in this sense, a goal model provides a justification for itself, particularly if we include belief elements for extra design rationalization.
+Providing an alternative subgoal or task in a goal model is then an explicit argumentative move in the discussion.
+In this sense, a goal model provides a justification for itself, particularly if we include belief elements for extra design rationalization.
 This idea is also prevalent in our RationalGRL framework: many arguments are in fact GRL elements, and many critical questions can be answered by introducing new GRL elements.
-However, as was already argued in Section~\ref{sect:introduction} (and also by \cite{Jureta:RE2008}), argumentation produces different, richer and complementary information to just the goal model.
+However, as was already argued in Section~\ref{sect:introduction} (and also by \cite{Jureta:RE2008}), argumentation produces different, richer and complementary information to the goal model.
 The goal model is the product of a process of argumentation, and does not include, for example, goals and tasks that were at some point considered but discarded.
-Furthermore, for goal models it is only possible to determine the satisfiability of goals given the possible tasks and resources; what cannot be determined is the acceptability of goals, that is, whether they are acceptable given potentially contradictory opinions of stakeholders.
+Furthermore, for goal models, it is only possible to determine the satisfiability of goals given the possible tasks and resources; what cannot be determined is the acceptability of goals, that is, whether they are acceptable given potentially contradictory opinions of stakeholders.
 
 There are several contributions to the literature that relate argumentation-based techniques with goal modeling.
-The first line of work is by Bagrheri and Ensan \cite{bagheri2011consolidating} and Mirbel and Villata \cite{MirbelVillata12} use abstract argumentation frameworks (see Definition~\ref{def:argumentation-framework}) to capture individual goals and the relations between them as modeled in a goal model, so that consistent (i.e., acceptable) subsets of goals can be computed using the relevant argumentation semantics \cite{Dung1995}.
+The first line of work is by Bagheri and Ensan \cite{bagheri2011consolidating} and Mirbel and Villata \cite{MirbelVillata12} use abstract argumentation frameworks (see Definition~\ref{def:argumentation-framework}) to capture individual goals and the relations between them as modeled in a goal model, so that consistent (i.e., acceptable) subsets of goals can be computed using the relevant argumentation semantics \cite{Dung1995}.
 For example, if in the goal model there is a conflict between goals $G_1$ and $G_2$, there is an attack between the arguments representing these goals and hence $G_1$ and $G_2$ cannot be in the same extension.
 Or if there is a dependency link between $G_1$ and $G_2$, then $G_1$ should be in every extension $G_2$ is in and vice versa.
-AND and OR decomposition are also modeled thus, that is, if $G_3$ AND-decomposes into $G_1$ and $G_2$ then $G_1$ and $G_2$ should be in every extension $G_3$ is in, and if $G_3$ OR-decomposes into $G_1$ and $G_2$ then either $G_1$ or $G_2$ should be in every extension $G_3$ is in.
+AND and OR decomposition are also modeled as such, that is, if $G_3$ AND-decomposes into $G_1$ and $G_2$ then $G_1$ and $G_2$ should be in every extension $G_3$ is in, and if $G_3$ OR-decomposes into $G_1$ and $G_2$ then either $G_1$ or $G_2$ should be in every extension $G_3$ is in.
 
@@
 
-The contribution most closely related to ours is the work by Jureta et al.~\cite{Jureta:RE2008}.
-Jureta et al.
-propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
+The contribution most closely related to ours is the work by Jureta \emph{et al.}~\cite{Jureta:RE2008}.
+Jureta \emph{et al.} propose ``Goal Argumentation Method (GAM)'' to guide argumentation and justification of modeling choices during the construction of goal models.
 GAM is a high-level decision process, in which alternative solutions for an RE problem are evaluated and compared using argumentation.
-Jureta et al.
-use a well-known fictitious example of an argumentative discussion regarding a meeting scheduler, in which a goal model is being built by the stakeholders proposing tasks, goals, and alternative solutions for goals.
+Jureta \emph{et al.} use a well-known fictitious example of an argumentative discussion regarding a meeting scheduler, in which a goal model is being built by the stakeholders proposing tasks, goals, and alternative solutions for goals.
 They include clarification as an important step in their GAM process, and discuss various types of clarity problems (vagueness, ambiguity, overgenerality, synonymy) and basic techniques for dealing with them (e.g.
@@
 In this sense, the RationalGRL methodology in Figure~\ref{fig:rationalgrl-methodology} can be seen as a more specific version of GAM explicitly meant for goal modeling.
-The argument schemes and critical questions in RationalGRL provide more clear handles for goal modeling (cf.
+The argument schemes and critical questions in RationalGRL provide clearer handles for goal modeling (cf.
 requirement 4 of our framework, Section~\ref{sect:introduction}).
-Furthermore, the RationalGRL framework adds full tool support for the goal modeling and reasoning process, something that GAM is lacking.
+Furthermore, the RationalGRL framework adds full tool support for the goal modeling and reasoning process, which GAM lacks.
 GAM does contain more specific ways of dealing with different types of clarity problems; in RationalGRL clarification is captured as a single critical question (CQ12, Table~\ref{table:argument-schemes}).
@@
 
-Jureta et al.'s framework also includes an argumentation part, where reasons (justifications) for conclusions are given as formal structured arguments\footnote{Informally, a structured argument is similar to the PRAS argument in Section~\ref{sect:background:pras}, i.e., $a, b \xrightarrow{therefore} c$.}.
+Jureta \emph{et al.}'s framework also includes an argumentation part, where reasons (justifications) for conclusions are given as formal structured arguments\footnote{Informally, a structured argument is similar to the PRAS argument in Section~\ref{sect:background:pras}, i.e., $a, b \xrightarrow{therefore} c$.}.
 Arguments and alternatives are then captured as structured arguments or argument diagrams with reasons for or against goals and tasks.
@@
 Section~\ref{sect:formalframework:translation}).
-In contrast, Jureta et al.
-only provide a mapping from goal models to structured arguments, and the step from structured arguments or argument diagrams to goal models is never formally defined.
-In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we extended Jureta et al.'s work and translated argument diagrams to GRL models (an automatic translation tool is discussed in~\cite{vanZee-etal:er2016}).
+In contrast, Jureta \emph{et al.} only provide a mapping from goal models to structured arguments, and the step from structured arguments or argument diagrams to goal models is never formally defined.
+In previous work on the RationalGRL framework~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we extended Jureta \emph{et al.}'s work and translated argument diagrams to GRL models (an automatic translation tool is discussed in~\cite{vanZee-etal:er2016}).
 This gives us two complex diagrams, an argument diagram and a goal diagram, and a mapping between them.
 This was, in our opinion, ultimately an unsatisfying solution given the problems and requirements described in Section \ref{sect:introduction}.
-One problem is that the argument diagram is at least as complex as the GRL diagram, so any stakeholder trying to understand the discussion has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
+One problem is that the argument diagram is at least as complex as the GRL diagram.
+This means that any stakeholder trying to understand the discussion has to parse two complex diagrams containing goals, alternative solutions, tasks, and so forth.
 RationalGRL overcomes this problem by integrating argumentation and goal modeling in a single modeling language.
@@
 RationalGRL lays down a basic framework for argumentation in requirements engineering, and all aspects of this framework are intended to be fully extensible.
-We see a large number of open issues to be explored in future research.
+We envision a large number of open issues to be explored in future research.
 
@@
 From a knowledge engineering perspective, including the argument schemes and critical questions associated with these domains in the RationalGRL framework is a time-consuming task.
-However, from a formal perspective adding new schemes and questions is easy: as already discussed by Bex et al.~\cite{bexEtal2003}, critical questions for argument schemes will always lead to either new information being introduced, new information replacing old information, or new information attacking old information.
-This corresponds to the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations in the RationalGRL framework, which have been formalized in Algorithms 3-9 in Section~\ref{sect:algorithms}, which would hence be suitable for other types of schemes and questions as well.
+However, from a formal perspective adding new schemes and questions is easy: as already discussed by Bex \emph{et al.}~\cite{bexEtal2003}, critical questions for argument schemes will always lead to either new information being introduced, new information replacing old information, or new information attacking old information.
+This corresponds to the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations in the RationalGRL framework, which have been formalized in Algorithms 3-9 in Section~\ref{sect:algorithms}, hence it would be suitable for other types of schemes and questions as well.
 
@@
 Now that the initial framework has been developed, however, it is worth exploring what additional techniques computational argumentation has to offer in more detail.
-For instance, in our framework we have not included the possibility of expressing preferences between mutually inconsistent arguments (e.g.
+For instance, in our framework we did not include the possibility of expressing preferences between mutually inconsistent arguments (e.g.
 such as in Figure~\ref{fig:goalmodeling:arg3}).
@@
 \paragraph{Tool support}
-The current tool is a prototype that implements the RationalGRL framework in a fairly simple way, that is, by offering the possibility to argue and ask critical questions about a goal model, and then export this goal model for further analysis in jUCMNav.
+The current tool is a prototype that implements the RationalGRL framework in a fairly simple way.
+That is, by offering the possibility to argue and ask critical questions about a goal model, and then export this goal model for further analysis in jUCMNav.
 Integration of RationalGRL elements into jUCMNav, or at least creating an import function to import jUCMNav goal models into the tool would allow us to tap into the large amount of research and development that is being carried out with jUCMNav.
@@
 In addition to such extensions, it is interesting to think about new possibilities.
-One idea is to use the tool for collaborative on-line goal modeling, similar to Github: because the renaming (i.e.
+One idea is to use the tool for collaborative on-line goal modeling, similar to Github: since the renaming (i.e.
 replacement) and deletion (i.e.
@@
 The complexity of the domain makes focused experiments difficult, but not impossible.
-For example, Murakannaiah et al.~\cite{murukannaiah2015} provided a realistic but bounded problem for their subjects to solve, and similar experiments are imaginable to test RationalGRL: have two groups solve a goal modelling problem, one using only GRL and one using RationalGRL, and see how they perform.
+For example, Murakannaiah \emph{et al.}~\cite{murukannaiah2015} provided a realistic but bounded problem for their subjects to solve, and similar experiments are imaginable to test RationalGRL: have two groups solve a goal modeling problem, one using only GRL and one using RationalGRL, and examine how they perform.
 Naturally, the type of problem will greatly influence the outcome.
@@
 However, a more complex problem on which multiple persons have to work asynchronously might benefit from the extra tools offered by RationalGRL.
-Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the methodology, the tool, or a list of critical questions to serve as reminders during the goal modelling process (cf.
+Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the methodology, the tool, or a list of critical questions to serve as reminders during the goal modeling process (cf.
 \cite{SchriekEtal2016}).
@@
 We extend the well-known goal modeling language GRL with argument schemes and critical questions that can be used to analyze and guide stakeholders' discussions about goal models.
+Furthermore, we provide formal argumentation semantics for the new RationalGRL language.
 Our approach, thus, provides a rationalization for the elements of the goal model in terms of underlying arguments, and helps in understanding why parts of the model have been accepted while others have been rejected.

### 2018-02-16 Response to reviewers.txt
Response to reviewers & small changes

%*Editor: Please ensure also that your manuscript shows clearly how your work relates to relevant works as they may have been published in the Requirements Engineering or presented at the Requirements Engineering series of conferences.

%FB: perhaps the editor would like us to add more references to RE and REJ? It is true that the majority of references is now outside of RE. **Marc, Sepideh**

-----------------------
%R1: The integrated meta model looks reasonable but I have some doubts about the inclusion of "critical question" in the schema. Critical questions are something that are used to identify arguments and attack links, but  I don't see them as a concept to be included in the meta model. They are part of the methodology used to build the goal model rather concepts of the language.

%FB: **Marc**

-----------------------
%R1: The proposed algorithms are very simple and they do not add much to the overall presentation, I think they can be summarized in the section describing the tool. 
%R2: I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?

Reviewer 1 argues that the algorithms are fairly simple. We agree that many algorithms are very much the same, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two. Reviewer 2 asks whether the automated reasoning can be handled by propositional logic instead of the algorithms. The answer is that propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE. Furthermore, the algorithms also introduce arguments and attacks, formal notions of which are not in basic propositional logic. The idea is that the formalization is needed for the algorithms, but that just providing a formal semantics of GRL does not directly tell you how to compute/reason with GRL models.

%FB: **Marc, can you work on shortening sections 5.4.1-5.4.3?**

-----------------------
%R1: The methodology and the tools looks quite preliminary and without a real validation of the approach. I think this is the weakest part of the paper. Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word. The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.
%R2: I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable. However, this remains to be shown yet in an industrial environment. Connecting them with an existing approach for goal modeling is also a pro in this respect.

%FB: Depending on the editor's answer we can leave this as is in the paper. However, we do need to come up withy a satisfactory example in the response to reviewers. Also, I think we need to include some references on empirical work that has been done on goal modelling ***Sepideh?***

-----------------------
%R2: the statement on actually using a reduced GRL should be made in the Abstract and Introduction already, not just in a footnote later.

Reviewer 2 comments that the connection to an existing approach, GRL is commendable, but that we should mention in the abstract and the introduction that we use a simplified version of GRL. We have included in the introduction that we use a simplified version. Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution. This simplification does not in any way influence the case study or any of the other results. Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would be trivial.

%R2:I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL.
In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision? Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?

%FB: **Floris to write on GRL beliefs in 7.1**

-----------------------
%R2: How would you compare this specific approach with the old gIBIS?
%FB: **Floris to write on gIBIS in 7.1**

-----------------------
%R2: Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?
%FB: **Floris to write on this in 7.1**

-----------------------
%R2: I am also missing a discussion about the consequences of this binary "nature" of argumentation with respect to argumentation in reality.
%FB: **Floris to write on this in 5.2**

-----------------------
%R2: Regarding terminology, I wonder about "valid" GRL models here, which seem to mean that certain criteria are fulfilled. Validation would mean that a goal model makes sense in the real world as having the "right" goals and relationships between them.
"Verification" in 6.2 appears to stand for evaluation. This cannot be "verified".

%FB: **Marc**

-----------------------
%R2: I also do not think that Figure 16 shows a "methodology". It does not even look like a "method".

%FB: **Sepideh, Marc**

-----------------------
%R2: please reconsider the very first sentences in the Abstract and Introduction each, especially the latter statement on RE.

%FB: who originally wrote these sentences? **Marc, Sepideh**

-----------------------
%R2: Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.

%FB: Reference to figure is line 56 p. 19 **Marc**
%FB: A quick scan in my TeX environment with UK spelling shows me no spelling mistakes, and I believe gramatically there's not too much wrong with the paper either, but let's check **Marc, Floris, Sepideh**

### 2018-02-16 background.tex
Response to reviewers & small changes

@@
-\section{Background: Goal-oriented Requirements Language and Argumentation}
+\section{Background: Goal-oriented Requirements Language and Practical Reasoning}
 \label{sect:background}
 
-In this section, we first introduce our running example, after which we introduce the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework\footnote{Note that in this article, we consider a slightly simplified version of GRL.}.
-Next, we introduce argumentation: we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
+In this section, we first introduce our running example, after which we give a brief overview of the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework\footnote{For simplicity's sake, we consider a slightly cut-down version of GRL.
+All the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution.
+This simplification does not in any way influence the case study.
+Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would be trivial.}.
+Next, we introduce argumentation; we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
 %This will be our starting point in the next section.
@@
 asked designers (groups of students) to discuss the requirements of this traffic simulator.
-These discussions were recorded and transcribed, and we used these transcripts for an extensive case study on the basis of which we develop our RationalGRL framework.
-Furthermore, we also use the traffic simulator case for the running examples throughout this paper.
+These discussions were recorded and transcribed.
+We used these transcripts for an extensive case study on the basis of which we develop our RationalGRL framework.
+Furthermore, we also use the traffic simulator case as running examples throughout this paper.
 
@@
 Actors are holders of intentions; they are the active entities in the system or its environment who want goals to be achieved, tasks to be performed, resources to be available, and softgoals to be satisfied.
-Softgoals differentiate themselves from goals in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable, often in a binary way.
+Softgoals differentiate themselves from goals in that there is no clear, objective measure of satisfaction for a softgoal whereas a goal is quantifiable.
 Softgoals (e.g.
@@
 Tasks represent solutions to (or operationalizations of) goals and softgoals.
-In Figure~\ref{fig:example-small}, there are two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve goal \emph{Generate cars}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off screen.
+In Figure~\ref{fig:example-small}, there are two tasks \emph{Create new cars} and \emph{Keep same cars}: in order to achieve the goal \emph{Generate cars}, the simulation can either constantly generate new ones or keep the same cars and have them reappear after they disappear off-screen.
 In order to be achieved or completed, softgoals, goals, and tasks may require resources to be available (e.g., \emph{Car Objects}).
 Finally, the full version of GRL allows design rationale to be captured using beliefs.
-Because we capture the reasoning and rationales behind goal models using arguments, we do not include beliefs in our simplified version of GRL (but see the discussion at the end of this section).
+Since we capture the reasoning and rationales behind goal models using arguments, we do not include beliefs in our simplified version of GRL (but see the discussion at the end of this section).
 
@@
 Task \emph{Create new cars} has a positive contribution to the softgoal \emph{Realistic simulation}, and a negative contribution to the softgoal \emph{Simple design}.
-Note that the full GRL specification considers different levels of positive and negative contributions, which are not directly relevant for current purposes.
-Dependency links model relationships between actors or resources.
+Note that the full GRL specification considers different levels of positive and negative contribution values (both quantitative and qualitative), which are not directly relevant for current purposes.
+Dependency links are relationships between IEs, which can model dependencies between actors.
 Here, the goal \emph{Generate cars} depends on the resource \emph{Car objects}.
@@
 GRL has a well-defined syntax and semantics.
-Furthermore, GRL provides support for providing a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
+Furthermore, GRL provides support for a scalable and consistent representation of multiple views/diagrams of the same goal model (see~\cite[Ch.2]{Ghanavati2013} for more details).
 GRL also has the capability to be extended through metadata, links, and external OCL constraints.
 This allows GRL to be used in many domains without the need to change the whole modeling language.
-For example, GRL is linked to Use Case Maps, which provide traceability between concepts and instances of the goal model and behavioral design models.
+For example, GRL is linked to Use Case Maps, which provides traceability between concepts and instances of the goal model and behavioral design models.
 Multiple views and traceability links are a good fit with our current research: we aim to add traceability links between intentional elements and their underlying arguments.
-
-GRL has six evaluation algorithms which allow the analysis of alternatives and design decisions by calculating the satisfaction value of the intentional elements quantitatively, qualitatively or in a hybrid way.
-jUCMNav, GRL tool-support (see Section~\ref{sect:tool}), also allows for adding new GRL evaluation algorithms~\cite{jUCMNav}.
-For simplicity's sake, the version of GRL used in this paper does not take into account satisfaction values.
 
@@
 	\item What does \emph{Keep same cars} mean?
-	\item Why does task \emph{Create new cars} contribute negatively to \emph{Simple design} and positively to \emph{Realistic simulation}?
+	\item Why does the task \emph{Create new cars} contribute negatively to \emph{Simple design} and positively to \emph{Realistic simulation}?
 	\item Why does \emph{Generate cars} XOR-decompose into two tasks?
@@
 
-These are the type of the questions that we cannot answer by just looking at GRL models.
+These are the types of the questions that we cannot answer by just looking at GRL models.
 The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the model, such as clarification steps for the naming, or alternatives that have been considered for the relationships.
@@
 However, beliefs cannot be connected to links - this makes answering the third and fourth question above impossible.
-Furthermore, beliefs are after-the-fact design rationales and do not capture the types of critical questions given above.
-Finally, the jUCMNav tool does not beliefs in the evaluation algorithms, and thus there is no formal connection between goal models and their underlying beliefs and opinions.
+Furthermore, beliefs are after-the-fact design rationales and do not capture the types of questions given above.
+Finally, the jUCMNav tool does not consider beliefs in the evaluation algorithms, and thus there is no formal connection between goal models and their underlying beliefs and opinions.
 For a more detailed comparison of our framework with GRL beliefs, see Section~\ref{sect:discussions:relatedwork}.
@@
 In fact, in some of our previous work on RationalGRL \cite{vanzee-etal:renext2015,vanZee-etal:er2016} we used PRAS-arguments such as the ones above to capture reasoning about goals, and provided a translation to goal models.
-However, one problem of this is that we cannot literally use PRAS and its critical questions, as there are elements of the GRL language, such as actors and resources, which cannot be found in PRAS.
+However, one problem of this approach is that we cannot literally use PRAS and its critical questions, as there are elements in the GRL language, such as actors and resources, which cannot be found in PRAS.
 Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson and Bench-Capon~\cite{atkinson2007} actually apply to GRL models.
@@
 A typical discussion is much more unstructured, as is clear from the transcript excerpts in Appendix~\ref{sect:transcripts:excerpts}.
-So if we would use the version of PRAS presented in this section for our argumentation, we would violate requirement 1: The argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
+Thus, if we would use the version of PRAS presented in this section for our argumentation, we would violate requirement 1: The argumentation techniques must capture the actual discussions of the stakeholders or designers in the early requirements engineering phase.
 Our solution is to develop our own set of argument schemes and critical questions by analyzing transcripts of discussions about the traffic simulator.

### 2018-02-16 introduction.tex
Response to reviewers & small changes

@@
 Developing goal models for such complex and large systems is not a trivial task and can be very cumbersome.
-In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of system development.
+In such situations, failing to record the discussions underlying a goal model in a structured manner may harm the success of the RE phase of the system development process.
 
@@
 
-Another challenge is that current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effect on the goal model.
+Another challenge is that current goal modeling languages have limited support for reasoning about changing beliefs and opinions, and their effects on the goal model.
 A stakeholder may change his or her opinion, but it is not always directly clear what the effect of this is on a goal model.
-Similarly, with existing goal modeling languages one can change a part of a goal model without being able to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
+Similarly, with existing goal modeling languages one might change a part of a goal model without being able to reason about whether or not this new goal model is consistent with the underlying beliefs and arguments.
 This becomes even more challenging if the stakeholders constructing the goal model change, since modeling decisions made by one group of stakeholders may conflict with the underlying beliefs of another group of stakeholders.
-The disconnect between goal models and their underlying beliefs and opinions may further lead to a poor understanding of the problem and solution domain, which is an important reason of RE project failure~\cite{curtis1988field}.
+The disconnect between goal models and their underlying beliefs and opinions may further lead to a poor understanding of the problem and its solution, which is an important reason of RE project failure~\cite{curtis1988field}.
 
-To summarize, what is needed is a way to record the rationales (beliefs, opinions, discussions, ideas) underlying a goal model.
+To summarize, what is needed is a systematic approach to record the rationales (beliefs, opinions, discussions, ideas) underlying a goal model.
 It should be possible to see how these rationales change during the goal modeling process, and they  should be clearly linked to the intentional elements of the resulting goal model.
@@
 \item 
-Using these traceability links, it must be possible to compute the effect of changes in the underlying argumentation on the goal model, and vice versa.
+Using these traceability links, it must be possible to compute the effect of changes in the underlying arguments about the goal model, and vice versa.
 \item 
-A methodology must be identified that can guide practitioners in using the framework.
+There must be a methodology to guide practitioners in using the framework.
 \item 
-The framework must have tool support. %"a tool support" is incorrect
+The framework must have tool support.
 \end{enumerate}
 
-%In this context, the main research question is:
-
-%\begin{quote}
-%\textbf{RQ.} What are the constructs, mechanisms and rules for developing a framework that formally captures the discussions between stakeholders such that it can generate goal models?
-%\end{quote}
-
-%FB: I deleted the research question.
-I think that the requirements already nicely capture exactly what our goal is in the paper.
-We can rephrase this goal as an RQ, but what does this buy us?
-Also, the RQ introduces new terminology (construct, mechanisms, rules), to which we have to explicitly refer later in the paper (e.g.
-we found mechanisms A, B, rules R1, R2, etc.).
-
-%MvZ: I shortened the paragraph below and put the full version in the conclusion.
-I think it makes more sense to reflect on the requirements in the conclusion, and to only give a brief overview of our framework in the introduction.
-Too much detail about argument schemes etc may confuse the reader.
-
-Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing framework for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}.
+Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework.
+RationalGRL extends a slightly simplified existing approach for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}.
 Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.
-Associated with argument schemes are so-called \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a particular way.
+Argument schemes are associated with \emph{critical questions}, which can point to typical sources of doubt or implicit assumptions people make when arguing in a certain way.
 These critical questions can point to counterarguments, and the acceptability of sets of arguments can be computed using argumentation semantics.
@@
 In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework to which we can apply argumentation semantics~\cite{Dung1995}.
-We, then, formally capture the link between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.
+We, then, formally capture the links between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.
 
@@
 Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL metamodel, and examples from the case study that illustrate the framework and its language.
-In Section~\ref{sect:formalframework}, we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and visa versa, and we develop various algorithms that change a RationalGRL model according to an argument scheme or a critical question that is applied.
-Section~\ref{sect:methodology+tool} discusses the RationalGRL methodology and explain various features of the tool we developed.
-Finally, Section~\ref{sect:discussion} covers related work, future work, and a conclusion.
+In Section~\ref{sect:formalframework}, we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and vice versa, and we develop various algorithms that change a RationalGRL model according to an argument scheme or a critical question.
+Section~\ref{sect:methodology+tool} discusses the RationalGRL methodology and explains various features of the tool we developed.
+Finally, Section~\ref{sect:discussion} covers the related work, the future work, and the conclusion.

### 2018-05-01 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/abstract.tex
submitted files

--- 2017-11-06 abstract.tex
@@
 Based on a case study in which stakeholders discuss requirements for a Traffic Simulator, we apply argumentation techniques from artificial intelligence to a goal modeling approach.
-Thus, we combine a traditional goal modelling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).
+Thus, we combine a traditional goal modeling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).
 RationalGRL provides a methodology, formal semantics and tool support to capture the discussions and outcomes of the argumentation process that leads to a goal model.

### 2018-05-01 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/background.tex
submitted files

--- 2018-02-16 background.tex
@@
 
-In this section, we first introduce our running example, after which we give a brief overview of the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework\footnote{For simplicity's sake, we consider a slightly cut-down version of GRL.
-All the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution.
-This simplification does not in any way influence the case study.
-Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would be trivial.}.
+In this section, we first introduce our running example, after which we give a brief overview of the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework\footnote{Note that in this article, we consider a slightly simplified version of GRL.}.
 Next, we introduce argumentation; we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.

### 2018-05-01 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/introduction.tex
submitted files

--- 2018-02-16 introduction.tex
@@
 
-Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework.
-RationalGRL extends a slightly simplified existing approach for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}.
+Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework, which extends an existing approach for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}.
 Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.

### 2018-05-01 Response to reviewers.txt
Revision for final version

@@
-%*Editor: Please ensure also that your manuscript shows clearly how your work relates to relevant works as they may have been published in the Requirements Engineering or presented at the Requirements Engineering series of conferences.
+######## EDITOR ########
 
-%FB: perhaps the editor would like us to add more references to RE and REJ? It is true that the majority of references is now outside of RE. **Marc, Sepideh**
+===== COMMENT #1
+Please ensure also that your manuscript shows clearly how your work relates to relevant works as they may have been published in the Requirements Engineering or presented at the Requirements Engineering series of conferences.
 
------------------------
-%R1: The integrated meta model looks reasonable but I have some doubts about the inclusion of "critical question" in the schema.
+===== REPLY #1
+Currently 6 of the 44 references in the article come from either REJ or the RE conference:
+- Y.
+Elrakaiby, A.
+Ferrari, P.
+Spoletini, S.
+Gnesi, and B.
+Nuseibeh.
+Using argumentation to explain ambiguity in requirements elici- tation interviews.
+In Proceedings of the 25th IEEE International Requirements Engineering Conference (RE 2017), pages 51–60.
+IEEE Press, 2017.
+- I. Jureta, S. Faulkner, and P. Schobbens. Clear justification of mod- eling decisions for goal-oriented requirements engineering. Re- quirements Engineering, 13(2):87–115, May 2008.
+- Y.
+Liu, Y.
+Su, X.
+Yin, and G.
+Mussbacher.
+Combined goal and feature model reasoning with the user requirements notation and jucmnav.
+In Requirements Engineering Conference (RE), 2014 IEEE 22nd International, pages 321–322.
+IEEE, 2014.
+- P.
+K.
+Murukannaiah, A.
+K.
+Kalia, P.
+R.
+Telangy, and M.
+P.
+Singh.
+Resolving goal conflicts via argumentation-based analysis of com- peting hypotheses.
+In Proceedings of the 23rd International Re- quirements Engineering Conference (RE 2015), pages 156–165.
+IEEE, 2015.
+- M.
+van Zee, F.
+Bex, and S.
+Ghanavati.
+Rationalization of Goal Models in GRL using Formal Argumentation.
+In Proceedings of the 23rd International Requirements Engineering Conference (RE’15), pages 220–225.
+IEEE Press, August 2015
+- E. S. K. Yu. Towards Modeling and Reasoning Support for Early-Phase Requirements Engineering. In Proceedings of the 3rd IEEE International Symposium on Requirements Engineering, pages 226–235, 1997.
+
+We have searched through the journal and conference database for more references, and we found 4 more references that are closely related to our work:
+- S. R. Smith. Constructing and revising formal arguments in re- quirements engineering. Requir. Eng., 2(4):199–216, Dec. 1997.
+- J.Lockerbie,N.A.M.Maiden,J.Engmann,D.Randall,S.Jones, and D.
+Bush.
+Exploring the impact of software requirements on system-wide goals: a method using satisfaction arguments and i* goal modelling.
+Requirements Engineering, 17(3):227–254, 2012.
+- J. Horkoff and E. Yu. Interactive goal model analysis for early requirements engineering. Requirements Engineering, 21(1):29– 61, 2016.
+- J. Hassine and D. Amyot. A questionnaire-based survey method- ology for systematically validating goal-oriented models. Require- ments Engineering, 21(2):285–308, 2016.
+
+This means we currently have 10 references from the Requirements Engineering Journal or presented at the Requirements Engineering series of conferences.
+
+######## REVIEWER #1 ########
+
+We thank the reviewer for the insightful comments.
+Especially the comment about the missing validation was useful since it motivated us to perform an small-scale in-house user study, which allowed us to strengthen our claims for the usefulness and usability of our framework.
+We’ve replied to each separate comment below.
+
+===== COMMENT #1
+The integrated meta model looks reasonable but I have some doubts about the inclusion of "critical question" in the schema.
 Critical questions are something that are used to identify arguments and attack links, but  I don't see them as a concept to be included in the meta model.
@@
 
-%FB: **Marc**
+===== REPLY #1
+We thank the reviewer for pointing this out.
+We indeed believe critical questions should not be part of the metamodel, and we have removed them accordingly.
+We initially had put them into the metamodel because they play an important part in the implementation, but they are by themselves not part of the RationalGRL language.
+We also have updated the text accordingly.
 
------------------------
-%R1: The proposed algorithms are very simple and they do not add much to the overall presentation, I think they can be summarized in the section describing the tool. 
-%R2: I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?
+===== COMMENT #2
+The proposed algorithms are very simple and they do not add much to the overall presentation, I think they can be summarized in the section describing the tool.
 
-Reviewer 1 argues that the algorithms are fairly simple.
-We agree that many algorithms are very much the same, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two.
-Reviewer 2 asks whether the automated reasoning can be handled by propositional logic instead of the algorithms.
-The answer is that propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE.
+===== REPLY #2
+We agree that many algorithms are very much the same, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two or more.
+
+===== COMMENT #3
+The methodology and the tools looks quite preliminary and without a real validation of the approach.
+I think this is the weakest part of the paper.
+Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word.
+The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.
+
+===== REPLY #3
+TODO: WRITE WHEN ANALYSIS IS DONE
+
+######## REVIEWER #2 ########
+
+We thank the reviewer for reading our article in detail, and providing many insightful and useful comments. We’ve replied to each separate comment below.
+
+===== COMMENT #1
+I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable.
+However, this remains to be shown yet in an industrial environment.
+
+===== REPLY #1
+TODO: WRITE WHEN ANALYSIS IS DONE
+
+===== COMMENT #2
+The statement on actually using a reduced GRL should be made in the Abstract and Introduction already, not just in a footnote later.
+
+===== REPLY #2
+We have included in the abstract and introduction that we use a simplified version.
+Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution.
+This simplification does not in any way influence the case study or any of the other results.
+Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would not change the essence of our formalism.
+
+===== COMMENT #3
+I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL.
+In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision?
+Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?
+
+===== REPLY #3
+%FB: **Floris to write on GRL beliefs in 7.1**
+
+===== COMMENT #4
+How would you compare this specific approach with the old gIBIS?
+
+===== REPLY #4
+%FB: **Floris to write on gIBIS in 7.1**
+
+===== COMMENT #5
+Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?
+
+===== REPLY #5
+%FB: **Floris to write on this in 7.1**
+
+===== COMMENT #6
+I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?
+
+===== REPLY #6
+Propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE.
 Furthermore, the algorithms also introduce arguments and attacks, formal notions of which are not in basic propositional logic.
@@
 
-%FB: **Marc, can you work on shortening sections 5.4.1-5.4.3?**
+===== COMMENT #7
+I am also missing a discussion about the consequences of this binary "nature" of argumentation with respect to argumentation in reality.
 
------------------------
-%R1: The methodology and the tools looks quite preliminary and without a real validation of the approach.
-I think this is the weakest part of the paper.
-Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word.
-The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.
-%R2: I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable.
-However, this remains to be shown yet in an industrial environment.
-Connecting them with an existing approach for goal modeling is also a pro in this respect.
-
-%FB: Depending on the editor's answer we can leave this as is in the paper.
-However, we do need to come up withy a satisfactory example in the response to reviewers.
-Also, I think we need to include some references on empirical work that has been done on goal modelling ***Sepideh?***
-
------------------------
-%R2: the statement on actually using a reduced GRL should be made in the Abstract and Introduction already, not just in a footnote later.
-
-Reviewer 2 comments that the connection to an existing approach, GRL is commendable, but that we should mention in the abstract and the introduction that we use a simplified version of GRL.
-We have included in the introduction that we use a simplified version.
-Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution.
-This simplification does not in any way influence the case study or any of the other results.
-Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would be trivial.
-
-%R2:I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL.
-In general, how does this argumentation approach really differ from one based on beliefs and, in particular, belief revision?
-Assume that an argument is a belief of some stakeholder revised by another argument, which is a belief of some other stakeholder?
-
-%FB: **Floris to write on GRL beliefs in 7.1**
-
------------------------
-%R2: How would you compare this specific approach with the old gIBIS?
-%FB: **Floris to write on gIBIS in 7.1**
-
------------------------
-%R2: Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?
-%FB: **Floris to write on this in 7.1**
-
------------------------
-%R2: I am also missing a discussion about the consequences of this binary "nature" of argumentation with respect to argumentation in reality.
+===== REPLY #7
 %FB: **Floris to write on this in 5.2**
 
------------------------
-%R2: Regarding terminology, I wonder about "valid" GRL models here, which seem to mean that certain criteria are fulfilled.
+===== COMMENT #8
+Regarding terminology, I wonder about "valid" GRL models here, which seem to mean that certain criteria are fulfilled.
 Validation would mean that a goal model makes sense in the real world as having the "right" goals and relationships between them.
+
+===== REPLY #8
+We agree that the term “valid” is confusing in this respect, since we do not want to express that the goal model makes sense in the real world.
+Instead, we’d like to express that certain criteria are fulfilled.
+Therefore, we’ve renamed “valid” to “consistent”.
+TODO(marcvanzee): Implement this renaming in the article.
+
+===== COMMENT #9
 "Verification" in 6.2 appears to stand for evaluation. This cannot be "verified".
 
-%FB: **Marc**
+===== REPLY #9
+We’ve renamed “verification” to “demonstration” and slightly reworded the sentence.
 
------------------------
-%R2: I also do not think that Figure 16 shows a "methodology". It does not even look like a "method".
+===== COMMENT #10
+I also do not think that Figure 16 shows a "methodology". It does not even look like a "method".
 
-%FB: **Sepideh, Marc**
+===== REPLY #10
+**Sepideh**
 
------------------------
-%R2: please reconsider the very first sentences in the Abstract and Introduction each, especially the latter statement on RE.
+===== COMMENT #11
+please reconsider the very first sentences in the Abstract and Introduction each, especially the latter statement on RE
 
-%FB: who originally wrote these sentences? **Marc, Sepideh**
+===== REPLY #11
+**Sepideh**
 
------------------------
-%R2: Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.
+===== COMMENT #12
+Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.
 
-%FB: Reference to figure is line 56 p. 19 **Marc**
-%FB: A quick scan in my TeX environment with UK spelling shows me no spelling mistakes, and I believe gramatically there's not too much wrong with the paper either, but let's check **Marc, Floris, Sepideh**
+===== REPLY #11
+We’ve ran a spell-checker and proof-read the paper carefully. There was one unresolved reference to a figure, which a corrected. Thank you for pointing this out.

### 2018-05-01 discussion.tex
Revision for final version

@@
 An insufficiency occurs when an analyst needs more information from the client to accept a client's statement.
-The inconsistencies are then captured as mutually attacking arguments, and the insufficiencies as arguments against the original statements.
-These counterarguments say that, for example, the functionality expressed in the statement cannot be realized or is irrelevant.
-Elrakaiby \emph{et al.} coded the data from 34 requirement elicitation interviews.
-They identified 39 inconsistencies (i.e.
-at least two arguments that mutually attack) and 29 insufficiencies (i.e., at least one argument attacking another).
+
+Hassin and Amyot~\cite{hassine2016questionnaire} analyze the correctness of GRL models by performing statistical analysis on generated questionnaires.
+This allows early detection of potential conflicts between stakeholders of the system.
+Instead of questionnaires, we use arguments and make them explicitly traceable to elements of the goal model.
+Horkoff and Yu~\cite{horkoff2016interactive} introduce analysis procedures for the $i*$ goal-oriented framework, allowing users to ask questions such as ``what if?'', or ``are certain goals achievable?
+how?
+or why not?''.
+They formally defined their procedures and describe a framework implementation.
+Also in $i*$, Lockerbie \emph{et al.}~\cite{lockerbie2012exploring} describe procedures that analysts follow to use satisfaction arguments.
+They demonstrate their results in two case studies, providing evidence towards the claim that the integration of $i*$ and satisfaction arguments is useful and usable.
+About 30 years ago, Smith~\cite{Smith:1997:CRF:2737426.2737493} already proposed a framework to develop arguments for the confidences that a software-based system will be successful in a given environment.
+He used a simple temporal logic with compositional reasoning, allowing arguments to be constructed, and revised.
 

### 2018-05-01 methodology.tex
Revision for final version

@@
 
-In previous sections, we have shown how the RationalGRL framework can capture stakeholder discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
+In previous sections, we have shown how the RationalGRL framework can capture stakeholders discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
 The previous section contained a formalization of RationalGRL, which forms the starting point of this section.
@@
 
-We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a RationalGRL model, an earlier version of which was presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.
-%Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+We propose the methodology shown in Figure~\ref{fig:rationalgrl-methodology} to develop a RationalGRL model, a version of which was presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.
+Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
 The rest of the steps are as follows:
 
-\textbf{(1) Instantiate Argument Schemes (AS)} -- We start with the list of argument schemes (Table~\ref{table:argument-schemes}).
-Whilst discussing the requirements, we select schemes from the list and instantiate them to form arguments for GRL model elements.
-In this way we build or modify the GRL model by introducing new GRL elements (\textsf{INTRO}).
-Note that it is also possible to start modifying an existing GRL model which was not built using the RationalGRL methodology: each GRL element corresponds to an argument (i.e.
-an instantiated scheme), so it is possible to instantiate argument schemes based on an existing GRL model.
+\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes (Table~\ref{table:argument-schemes}).
+Whilst discussing the requirements, we select schemes from the list and instantiate them to form arguments for GRL IEs or links.
+In this way we build the GRL model by introducing new GRL elements (\textsf{INTRO}).
+For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}".
+When an argument scheme is instantiated, it corresponds to an argument for or against part of a goal model.
 
 \textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we ask the relevant critical questions.
-Because each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.
+Since each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.
+For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}.
+When the analyst answers  a critical question, a new argument scheme may be instantiated.
+This is done in step (3) when the operation INTRO is executed.
 
-\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the \textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE} are performed on the GRL model.
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the three operations (\textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE}) is applied to the GRL model.
 Any of these operations impact the arguments and corresponding GRL intentional elements, modifying the initial GRL model into a RationalGRL model.
-After these modifications, we can keep on asking critical questions (e.g.
-about elements that were introduced by previously answering a critical question) until we are satisfied with our model.
+Recall that \textsf{INTRO} means that
+a new argument scheme is created.
+That means, the current argument related to the critical question does not get attacked.
+In the case of \textsf{DISABLE}, the intentional element or its related links are disabled in the model.
+\textsf{REPLACE} introduces a new argument and attacks the original argument at the same time.
+This means that the original element of the argument scheme is replaced with a new one.
 
@@
 The final requirement of our framework is that it has tool support (\textbf{requirement 5}).
-This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of integrated tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent verification to ensure our formal framework and algorithms actually work.
+This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.
 In this section we briefly highlight some features of the tool, and we explain some of its current limitations.
@@
 GRL is implemented in the open-source tool jUCMNav~\cite{jUCMNav}, which is actively maintained\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}.
-jUCMNav currently has over 2.000 commits made by over 40 contributors, representing over 200.000 lines of code.
-jUCMNav is used actively in research, and the tool has been extended in many directions over the past few years, such as for modeling and analyzing security misuse cases \cite{}, supporting activity theory \cite{}, Web-based goal modeling \cite{}, and enterprise architecture modeling \cite{}.
-\todo{Floris}{Marc}{references need to be added here!} Furthermore, jUCMNav includes Use Case modelling and contains various GRL evaluation algorithms with which the satisfiability of goals can be determined.
+jUCMNav currently has over 2,000 commits made by over 40 contributors, representing over 200,000 lines of code.
+jUCMNav is used actively in research, and the tool has been extended in many directions over the past few years, such as for modeling and analyzing security misuse cases~\cite{daramola2012ontology}, supporting activity theory \cite{georg2015synergy}, combined goal and feature model reasoning~\cite{liu2014combined}, and enterprise architecture modeling \cite{marosin-etal:caise2016}.
+Furthermore, jUCMNav includes Use Case modeling and contains various GRL evaluation algorithms with which the satisfiability of goals can be determined.
 
@@
 While we aim to integrate our framework into jUCMNav in the future, a web-based version of RationalGRL (and GRL) is more suited for current purposes.
-For example, it is our experience is that it can take significant time to set up the Eclipse environment and install the jUCMNav plugin, which hampers quick uptake by practitioners and also makes setting up empirical evaluations more difficult.
 Since RationalGRL does not make use of the many features of jUCMNav, we believe a light-weight version is sufficient for our current aims, and may benefit the community in general.
+Our tool is thus not meant to replace jUCMNav.
+For simplicity we have a web-based version, and it has the functionality to export to jUCMNav.
 
@@
 \paragraph{IE and links details pane} Figure~\ref{fig:tool:overview} shows the details pane for the goal `Generate cars'.
-The user can change the name of the element by changing the name and clicking `Rename', which will update the naming history on the bottom of the pane.
-The naming history is (simplified) implemenetation of the \textsf{REPLACE} operation on IEs.
+The user can change the name of the IE by changing the name and clicking `Rename', which will update the naming history on the bottom of the pane.
+The naming history is (simplified) implementation of the \textsf{REPLACE} operation on IEs, since it is a very simple form of \emph{clarification}.
+The main difference with jUCMNav is that we have a renaming history, so the user can see which names the IE had prior to the current one.
 If the IE has decomposition links, the user can change the decomposition type as well.
@@
 
-\paragraph{Critical question details pane} The details panel can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal element ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
+\paragraph{Critical question details pane} The details panel can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal IE ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
 As an example, consider Figure~\ref{fig:tool:cqdetails}, where the softgoal ``Easy to use'' is questioned with the relevancy question (CQ11).
@@
 \paragraph{Argumentation semantics} The RationalGRL tool computes the acceptability of arguments on the fly.
-In the example of Figure~\ref{fig:tool:cqeffect}, the original (argument for) softgoal ``Easy to use'' is rejected because it's only attacker ``CQ11 - Redundant'' is accepted.
+In the example of Figure~\ref{fig:tool:cqeffect}, the original (argument for) softgoal ``Easy to use'' is rejected because its only attacker ``CQ11 - Redundant'' is accepted.
 However, if we then attack ``CQ11 - Redundant'' with a new argument ``Necessary'', the original argument for ``Easy to use'' is again accepted because its only attacker is rejected  (Figure~\ref{fig:tool:argument}).
@@
 Including \textsf{REPLACE} like this in the tool would mean that, when the name of an element is changed, we would need to render all the previous versions of the element on the canvas of the tool.
-Because this would quickly become very cumbersome, we decided to implement a naming history for each element.
+Since this would quickly become very cumbersome, we decided to implement a naming history for each element.
 For example, in Figure~\ref{fig:tool:overview}, the goal ``Generate cars'' was previously named ``Add cars''.
 In this way, a history is kept of how an element name was clarified without cluttering the model canvas.
+This feature is currently very limited and can be extended in various was, for instance by changing the IE type.
 Similarly, the decomposition type can be changed in the details pane (Figure~\ref{fig:tool:overview}) without introducing new elements that attack the original ones (cf.

### 2018-05-02 Response to reviewers.txt
some references added

--- 2018-02-16 Response to reviewers.txt
@@
 However, we do need to come up withy a satisfactory example in the response to reviewers.
-Also, I think we need to include some references on empirical work that has been done on goal modelling ***Sepideh?***
+Also, I think we need to include some references on empirical work that has been done on goal modelling
+
+I asked Fabiano and he gave me the following two references, we can discuss them.
+
+Estrada, H., Rebollar, A.
+M., Pastor, O., & Mylopoulos, J.
+(2006, June).
+An empirical evaluation of the i* framework in a model-based software generation environment.
+In International Conference on Advanced Information Systems Engineering (pp.
+513-527).
+Springer, Berlin, Heidelberg.
+
+Morandini, M., Marchetto, A., & Perini, A.
+(2011, August).
+Requirements comprehension: A controlled experiment on conceptual modeling methods.
+In Empirical Requirements Engineering (EmpiRE), 2011 First International Workshop on (pp.
+53-60).
+IEEE
+
+***Sepideh***
 

### 2018-05-02 Response to reviewers.txt

--- 2018-05-01 Response to reviewers.txt
@@
 
+%FB: Marc to write on Enterprise Architects
+
+%FB: Marc/Floris to write on the below
+I think we need to include some references on empirical work that has been done on goal modelling. I asked Fabiano and he gave me the following two references, we can discuss them.
+
+Estrada, H., Rebollar, A.
+M., Pastor, O., & Mylopoulos, J.
+(2006, June).
+An empirical evaluation of the i* framework in a model-based software generation environment.
+In International Conference on Advanced Information Systems Engineering (pp.
+513-527).
+Springer, Berlin, Heidelberg.
+
+Morandini, M., Marchetto, A., & Perini, A.
+(2011, August).
+Requirements comprehension: A controlled experiment on conceptual modeling methods.
+In Empirical Requirements Engineering (EmpiRE), 2011 First International Workshop on (pp.
+53-60).
+IEEE
+
 ######## REVIEWER #2 ########

### 2018-05-02 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/designprompt.tex

--- 2017-06-27 designprompt.tex
@@
 Broadly, the tool should be easy to use, and should encourage students to explore multiple alternative approaches.
-Stu- dents should be able to observe any problems with their map's timing scheme, alter it, and see the results of their changes on the traffic patterns.
+Students should be able to observe any problems with their map's timing scheme, alter it, and see the results of their changes on the traffic patterns.
 This program is not meant to be an exact, scientific simulation, but aims to simply illustrate the basic effect that traffic signal timing has on traffic.
@@
 
-Your design will primarily be evaluated based on its elegance and clarity – both in its overall solution and envisioned implementation structure.
-
-\subsubsection*{Desired Outcomes}
-
-Your work on this design should focus on two main issues:
-
-\begin{enumerate}
-\item
-You must design the interaction that the students will have with the system.
-You should design the basic appearance of the program, as well as the means by which the user creates a map, sets traffic timing schemes, and views traffic simulations.
-\item
-You must design the basic structure of the code that will be used to implement this system.
-You should focus on the important design decisions that form the foundation of the implementation, and work those out to the depth you believe is needed.
-\end{enumerate}
-
-The result of this session should be: \emph{the ability to present your design to a team of software developers who will be tasked with actually implementing it.} The level of competency you can expect is that of students who just completed a basic computer science or software engineering undergraduate degree.
-You do not need to create a complete, final diagram to be handed off to an implementation team.
-But you should have an understanding that is sufficient to explain how to implement the system to competent developers, without requiring them to make many high-level design decisions on their own.
-
-To simulate this hand-off, you will be asked to briefly explain the above two aspects of your design after the design session is over.
-
-\subsubsection*{Timeline}
-\begin{itemize}
-\item 1 hour and 50 minutes: Design session
-\item 10 minutes: Break / collect thoughts
-\item 10 minutes: Explanation of your design 
-\item 10 minutes: Exit questionnaire
-\end{itemize}
+%I deleted the Desired Outcomes and Timeline, as the requirements are most important for our story.

### 2018-05-02 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/examples.tex

--- 2017-08-15 examples.tex
@@
-\section{Examples}
+\section{Examples from the Transcripts}
 \label{sect:examples}
 
-We now discuss various instantiations of argument schemes and the result of answering critical questions in more detail.
+In the previous section we discussed the results of our empirical analysis, and we used a simplified version of one of the transcripts as a motivating example.
+In this section, we provide examples from all of the three transcripts in more detail.
+Note that the examples in this section do not always come from Figure~\ref{fig:transcripts:grl}, but they may come from any of the three transcripts.
+
 For each example, we provide transcript excerpts, a visualization of arguments, and the corresponding goal model elements.
-Figure~\ref{fig:transcripts:grl} in the previous section shows an example of how a RationalGRL may look like when it is developed in our prototype tool.
+As we explained in the previous section, Figure~\ref{fig:transcripts:grl} shows an example of how a RationalGRL model may look like when it is developed in our prototype tool.
 However, the argument schemes that are instantiated are left implicit in this figure (in the tool, they can be seen in the details pane, see Section~\ref{sect:tool}).

### 2018-05-02 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/formalframework.tex

--- 2017-11-07 formalframework.tex
@@
 This is done for multiple reasons: (i) Most approaches in formal argumentation (cf.
-Section~\ref{sect:background:pras}) use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated to valid GRL models and visa versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support, which we present in the next section.
+Section~\ref{sect:background:pras}) use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated to valid GRL models and vice versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support, which we present in the next section.
 
 In Sections \ref{sect:formalframework:grl} and \ref{sect:formalframework:rationalgrl} we formalize a static representation of our framework based on the metamodel (Figure~\ref{fig:metamodel}).
-We first provide a formal specification of a GRL model in Section~\ref{sect:formalframework:grl}, and we extend this with arguments and attack links in the second subsection, hereby obtaining a formal specification of a RationalGRL model.
+We first provide a formal specification of a GRL model in Section~\ref{sect:formalframework:grl}, and we extend this with arguments and attack links in the Section~\ref{sect:formalframework:rationalgrl}, hereby obtaining a formal specification of a RationalGRL model.
 In Section~\ref{sect:formalframework:translation} we then present algorithms in order to translate a GRL model into a RationalGRL model, and vice versa.
-Finally, in Section~\ref{sect:algorithms} we turn to the dynamics of our framework, developing algorithms for instantiating argument schemes and answering critical questions and thus formally capturing the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations of which examples were previously given in Section~\ref{sect:overview:examples}.
+Finally, in Section~\ref{sect:algorithms} we develop algorithms for instantiating argument schemes and answering critical questions and thus formally capture the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations of which examples were previously given in Section~\ref{sect:overview:examples}.
 
@@
 Throughout this section, we adopt the convention that variables start with a lowercase letter (e.g, $id$, $i$, $j$, $name$, $goal$), and sets and constants start with an uppercase letter (e.g., $Type, AND, Goal$).
-We start by defining general elements of our language that we use in subsequent definitions.
+We begin by defining general elements of our language that we use in subsequent definitions.
 
@@
 \item $name \in Names$ is a string description of the element,
-\item $decomptype\in DecompTypes$ refers to the type of decomposition.\footnote{Note that, like in the metamodel, the decomposition type is defined on the element.}
+\item $decomptype\in DecompTypes$ refers to the type of decomposition.
 \end{itemize}
@@
 
-The definition above is sufficient to capture all intentional elements (IEs) used in GRL. %In the next definition, we present some simplifying notational conventions.
-%\begin{definition}[Notation]
-%\label{def:notation}
-%We adopt three conventions simplifying our notation:
-%\begin{itemize}
-%\item
-%We refer to the element of a tuple using the dot (".") notation.
-That is, we may for instance refer to the id, type, name and decomposition type of an $IE$ with respectively $ie.id$, $ie.type$, $ie.name$, and $ie.decomptype$.
-%\item 
-%We refer to a set of elements with the same id $i$  using the $i$ subscript on the set.
-For instance a set of IEs with id $i$ is denoted by $IE_i$, and if this is a single element, we denote it by $ie_i$\footnote{In GRL, there always exists at most one element for every id (see Def.~\ref{def:grl-model}), in RationalGRL, however, this condition does not hold (see Def.~\ref{def:rationalgrl-model}).} For instance, we may refer to the intentional element $ie = (0, Goal, \text{Make profit}, AND)$ with $ie_0$ and write $ie_0.type = Goal$, $ie_0.name = \text{Make profit}$, and $ie_0.decompositiontype = AND$.
-%\item
-%We can also refer to intentional elements of a specific type simply by $type_{id}$.
-For instance, we can abbreviate the element in the previous item with $goal_0$ and write $goal_0.name = \text{Make profit}$, and $goal_0.decomptype = AND$.
-%\end{itemize}
-%\end{definition}
+The definition above is sufficient to capture all intentional elements (IEs) used in GRL.
+Note that according to the definition above, all IEs have a decomposition type, even when the IE is not decomposed into other IEs.
+This is in line with the GRL metamodel.
+However, in the following we sometimes leave out the $decomptype$ if the IE is not decomposed to simplify notation and improve readability.
+
 Throughout this section we use the example in Figure~\ref{fig:example-small}. This example contains various IEs which can be formalized using Definition~\ref{def:ie}, for example, 
-$(1, Resource, \text{Car objects}, AND)$, $(2, Softgoal,$ Realistic simulation$, AND)$ and $(5, Goal,$ Generate cars$, XOR)$.
+$(1, Resource, \text{Car objects})$, $(2, Softgoal,$ Realistic simulation$)$ and $(5, Goal,$ Generate cars$, XOR)$.
 
-%Using the short-hand notation defined in Def.~\ref{def:notation}, we can make the following statements:
-%\begin{itemize}
-%\item $goal_4.name = \text{Generate cars}$
-%\item $goal_4.decomptype = XOR$
-%\item $resource_1.name = \text{Car objects}$
-%\end{itemize}
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=\columnwidth]{img/Example1-new.pdf}
-%\caption{Example GRL model (reprint of Fig.~\ref{fig:example-small})}
-%\label{fig:example-small2}
-%\end{figure} 
 We now define actors.
@@
 
-At this point we have defined all intentional elements in GRL and a containment relation between actors and intentional elements. We now turn to the GRL links.
+At this point we have defined all intentional elements in GRL and a containment relation between actors and intentional elements. We now discuss the GRL links.
 
@@
 
-%\begin{definition}[Notation]
-%\label{def:notation2}
-%\begin{itemize}
-%\item
-%We can refer to links of a specific type simply by $type_{id}$. For instance, we can abbreviate the link $(0,PosContr,2,3)$ with $posContr_0$. 
-%\item 
-%We can refer to a set of positive contribution links, negative contribution links, dependency links, and decomposition link with respectively $PosContr, NegContr, Dep,$, and $Decomp$. 
-%\item
-%We can refer to a set of contribution links that can be either positive or negative simply with $Contr$.
-%\end{itemize}
-%\end{definition}
-
 Similar to IEs, links have identifiers as well. Some of the links in our example (Figure~\ref{fig:example-small}) are $(8, Dep, 4, 1)$, $(9, PosContr, 5, 2)$ and $(11, Decomp, 5, 4)$.
-
-%Let us next provide some examples of the notation of Definition~\ref{def:notation2}:
-%\begin{itemize}
-%\item $dep_7 = (7, Dep, 4, 1)$
-%\item $Decomp = \{(10, Decomp, 5, 4), (11, Decomp, 6, 4)\}$
-%\item $PosContr = \{(8, PosContr, 5, 2)\}$
-%\item $Contr = \{(8, PosContr, 5, 2), (9, NegContr, 5, 3)\}$
-%\end{itemize}
 
@@
 \begin{flalign*}
-&IE=&\{&(1, Resource, \text{Car objects}, AND),&\\
-&   &  &(2, Softgoal, \text{Realistic simulation}, AND),&\\
-&   &  &(3, Softgoal, \text{Simple design}, AND),&\\
-&   &  &(4, Softgoal, \text{Easy to use}, AND),&\\
+&IE=&\{&(1, Resource, \text{Car objects}),&\\
+&   &  &(2, Softgoal, \text{Realistic simulation}),&\\
+&   &  &(3, Softgoal, \text{Simple design}),&\\
+&   &  &(4, Softgoal, \text{Easy to use}),&\\
 &   &  &(5, Goal, \text{Generate cars}, XOR),&\\
-&   &  &(6, Task, \text{Create new cars}, AND),&\\
-&   &  &(7, Task, \text{Keep same cars}, AND)\}&\\
+&   &  &(6, Task, \text{Create new cars}),&\\
+&   &  &(7, Task, \text{Keep same cars})\}&\\
 &Act=& &\{(0, Actor, \text{Traffic Simulator})\}&\\
@@
 \end{itemize}
-The set of generic arguments is denoted by $AE$.
+A set of generic arguments is denoted by $GA$.
 \end{definition}
 
-So a generic argument is any element in the RationalGRL model that is not an intentional element or an actor.
+Thus, a generic argument is any element in the RationalGRL model that is not an intentional element or an actor.
 In the example, some of the generic arguments are $(13, GenArg, \text{Redundant})$ and $(14,$ $GenArg,$ Necessary$)$.
@@
 This definition captures the specification in the RationalGRL metamodel (Figure~\ref{fig:metamodel}) in which the class \textsf{Argument} is a superclass of \textsf{GenericArgument} and \textsf{GRLModelElement}.
-In sum, we define an argument simply as any one of the GRL elements or links, or a generic argument.
+In sum, we define an argument simply as any one of the GRL IEs or links, or a generic argument.
 
-Examples of arguments in Figure~\ref{fig:example-small3} are $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, $A'_5 = (5, Goal,$ Generate cars$, XOR)$, $A_4 = (4, Softgoal,$ Easy to use$, AND)$ and $A_{13}=(13,$ $ArgElem,$ Redundant$)$.
+Examples of arguments in Figure~\ref{fig:example-small3} are $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, $A'_5 = (5, Goal,$ Generate cars$, XOR)$, $A_4 = (4, Softgoal,$ Easy to use) and $A_{13}=(13,$ $ArgElem,$ Redundant$)$.
 Note that the two arguments $A_5$ and $A'_5$ show an important difference between RationalGRL models and valid GRL models: While a valid GRL model disallows multiple elements with the same identifier (Definition~\ref{def:valid-grl-model}, condition 1), RationalGRL models do not enforce this restriction.
@@
 A_0 = &(0, Actor, \text{Traffic simulator})&\\
-A_1 = &(1, Task, \text{Car objects}, AND),&\\
-A_2 = &(2, Softgoal, \text{Realistic simulation}, AND),&\\
-A_3 = &(3, Softgoal, \text{Simple design}, AND),&\\
-A_4 = &(4, Softgoal, \text{Easy to use}, AND),&\\
+A_1 = &(1, Task, \text{Car objects}),&\\
+A_2 = &(2, Softgoal, \text{Realistic simulation}),&\\
+A_3 = &(3, Softgoal, \text{Simple design}),&\\
+A_4 = &(4, Softgoal, \text{Easy to use}),&\\
 A_5 = &(5, Goal, \text{Generate cars}, AND),&\\
 A'_5 = &(5, Goal, \text{Generate cars}, XOR),&\\
-A_6 = &(6, Task, \text{Create new cars}, AND),&\\
-A_7 = &(7, Task, \text{Wrap cars}, AND),&\\
-A'_7 = &(7, Task, \text{Keep same cars}, AND),&\\
+A_6 = &(6, Task, \text{Create new cars}),&\\
+A_7 = &(7, Task, \text{Wrap cars}),&\\
+A'_7 = &(7, Task, \text{Keep same cars}),&\\
 A_8 = &(8, Dep, 4, 1),&\\
@@
 
-%All the arguments and the attack relations of this RationalGRL model are shown in Figure~\ref{fig:example-small4}.
-Note the arguments for Actor-IE containment (arguments $A_{12}$ to $A_{17}$) have been omitted from this figure for readability.
-It can be read from the figure that two arguments are currently rejected, namely $A_4$ and $A_{19}$.
-However, we did not yet make precise how exactly this is computed.
-We will do so in the following definitions.
-
-%\begin{figure}[ht]
-%\centering
-%\includegraphics[width=\columnwidth]{img/Example1-new-arguments}
-%\caption{Argumentation framework of RationalGRL model from Fig.~\ref{fig:example-small3}}
-%\label{fig:example-small4}
-%\end{figure}
-
-From Figure~\ref{fig:example-small3}, it can be read from that arguments $A_5$, $A_7$, $A_{13}$ and $A_{15}$ are currently disabled (rejected).
-However, we did not yet make precise how exactly this is computed.
+In Figure~\ref{fig:example-small3}, it can be read from that arguments $A_5$, $A_7$, $A_{13}$ and $A_{15}$ are currently disabled (rejected).
+However, we have not yet make it clear how exactly this is computed.
 We will do so in the following definitions.
@@
 \end{figure}
-In order to compute when an argument is accepted and when not we use argumentation semantics.  We use the standard approach by Dung~\cite{Dung1995}.
+In order to compute when an argument is accepted and when it is not we use argumentation semantics.  We use the standard approach by Dung~\cite{Dung1995}.
 
@@
 We have now formalized a \emph{static} representation of the RationalGRL framework.
-In this section we formalize the \emph{dynamics} by developing algorithms for applying argument schemes and critical questions.
+In this section we formalize algorithms for applying argument schemes and critical questions.
 These algorithms are applied to RationalGRL models (Definition~\ref{def:rationalgrl-model}) and produce new arguments and attack relations.
@@
 introduces a new RationalGRL element and \textsf{DISABLE} creates a new argument that attacks another argument.
-The \textsf{REPLACE} operation is can be seen as a combination of \textsf{INTRO} and \textsf{DISABLE}: a new argument corresponding to a GRL element is added and this new argument attacks a previous version of the element or link.
+The \textsf{REPLACE} operation can be seen as a combination of \textsf{INTRO} and \textsf{DISABLE}: a new argument corresponding to a GRL element is added and this new argument attacks a previous version of the element or link.
 We will discuss all three operations in separate sections below.
@@
 On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$.
-On line~\ref{alg:as0:2} the actor is added as a new argument.
+On line~\ref{alg:as0:2}, the actor is added as a new argument.
 As an example, we can formalize the addition of actor ``Traffic Simulator'' to an empty RationalGRL model (i.e.
@@
     \State $res\_id\gets mintId()$\label{alg:add-resource:id}
-    \State $A_1\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-resource:arg1}
+    \State $A_1\leftarrow (res\_id, Resource, R)$\label{alg:add-resource:arg1}
     \State $A_2\leftarrow (ActIE, i, res\_id)$\label{alg:add-resource:arg2}
@@
 As an example, consider the addition of resource ``Car objects'' to the RationalGRL model we just discussed (where $A_0 = (0, Actor, \text{Traffic simulator})$).
-Adding the resource can then be formalized as executing $AS1(0, \text{Car objects})$, resulting in the addition of two arguments $(1, Resource, \text{Car object}, AND)$ and $(ActIE, 0, 1)$ which are added to the set of arguments $Arg$.
+Adding the resource can then be formalized as executing $AS1(0, \text{Car objects})$, resulting in the addition of two arguments $(1, Resource, \text{Car object})$ and $(ActIE, 0, 1)$ which are added to the set of arguments $Arg$.
 
-Since arguments schemes AS2-AS4 are very similar to AS1, we omit these algorithms here.
+Since arguments schemes AS2-AS4 are very similar to AS1, we don't show these algorithms here.
 
@@
     \State $task\_id\gets mintId()$\label{alg:add-decomp:task-id}
-    \State $A_1\leftarrow (task\_id, Task, T, AND)$\label{alg:add-decomp:arg1}
+    \State $A_1\leftarrow (task\_id, Task, T)$\label{alg:add-decomp:arg1}
     \State $A_2\leftarrow (mintId(), Decomp, i, task\_id)$\label{alg:add-decomp:arg2}
@@
 As an example, suppose we are adding the decomposition of goal ``Generate cars'', expressed as argument $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, into task ``Keep same cars'', and suppose we have the following argument for the goal: .
-Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars}, AND)$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.
+Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars})$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.
 
-The remaining argument schemes AS6-AS12 are all similar to algorithms~\ref{alg:add-actor}-\ref{alg:add-decomp} and have been omitted here.
+The remaining argument schemes AS6-AS12 are all similar to algorithms~\ref{alg:add-actor}-\ref{alg:add-decomp} and are therefore not shown here.
 
@@
 This is the same for all the other critical questions of type \emph{INTRO}.
-Therefore, we have omitted them here as well.
+Therefore, we are not shown here as well.
 
@@
 As discussed before, algorithms of type \emph{DISABLE} consist of adding a new argument attacking an existing argument.
-The argument that is added is itself not an argument for a GRL element or link, but rather a generic argument (Definition~\ref{def:generic-argument}).
+The argument that is added is not an argument for a GRL element or link, but it is rather a generic argument (Definition~\ref{def:generic-argument}).
 
-In all of these algorithms, we assume the critical question is answered as indicated in the right-most column of Table~\ref{table:argument-schemes}.
+In all of these algorithms, we assume the critical question is answered affirmatively, as indicated in the right-most column of Table~\ref{table:argument-schemes}.
 For instance, for critical question CQ0 ``Is the actor relevant?'', we assume it is answered with ``No''.
+In this case, an action is required.
+In contrary, if the answer to the question is ``Yes'', no action is required.
 
@@
 
-Algorithm~\ref{alg:actor-not-relevant} is executed when critical question CQ0 is answered affirmatively (i.e., with ``No'').
+Algorithm~\ref{alg:actor-not-relevant} is executed when critical question CQ0 is answered with ``No''.
 First, on lines~\ref{alg:actor-not-relevant:genarg} and~\ref{alg:actor-not-relevant:genarg2}, an argument is created for the critical question and added to the set of arguments $Arg$.
@@
 We will explain this in more detail in the example below.
-Then, on line~\ref{alg:actor-not-relevant:att}, an attack link is created from the generic argument $A$ that is created to the argument for the actor $A_j$.
+Then, on line~\ref{alg:actor-not-relevant:att}, an attack link is created from the generic argument $A$ to the argument for the actor $A_j$.
 After executing the algorithm, all existing arguments for the actor with identifier $i$ are attacked by a newly created argument $A$.
@@
 
-Critical questions C1-CQ3 are all very similar to CQ0 and have therefore been omitted here.
+Critical questions C1-CQ3 are all very similar to CQ0 and are therefore not shown here.
 
@@
 CQ11 (``Is the element relevant/useful'') is slightly different since the attack element is not of a specific type, but is simply any GRL element.
-However, the resulting algorithm is very similar to the previous two and has therefore been omitted.
+However, the resulting algorithm is very similar to the previous two and is therefore now shown here.
 The only other algorithm of type \emph{DISABLE} that we discuss is $Att$.
@@
 Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counterarguments to arguments.
-In all of the previous \emph{DISABLE} algorithm, the attack was on a specific type of argument, for instance an argument for an actor or an argument for a decomposition.
+In all of the previous \emph{DISABLE} algorithms, the attack was on a specific type of argument, for instance an argument for an actor or an argument for a decomposition.
 In this algorithm, however, \emph{any} set of arguments can be attacked by a new argument.
@@
 For CQ12 we should make a small modification: instead of replacing the decomposition type of the IE, we replace its name.
-Since this is a very minor modification we have omitted it here.
+Since this is a very minor modification we don't show it here.

### 2018-05-02 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/methodology.tex

--- 2018-05-01 methodology.tex
@@
 The final requirement of our framework is that it has tool support (\textbf{requirement 5}).
-This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.
+This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent verification to ensure our formal framework and algorithms actually work.
 In this section we briefly highlight some features of the tool, and we explain some of its current limitations.

### 2018-05-02 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/overview.tex

--- 2017-11-06 overview.tex
@@
 
-We now turn to an overview of our RationalGRL framework.
-We will show through a metamodel and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
+We now give an overview of our RationalGRL framework.
+Through a metamodel and informal examples from our case study we will show that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
 A formalization of our framework using formal logic can be found in Section~\ref{sect:formalframework}.
@@
 There are two activities (bottom), \emph{Practical reasoning \& argumentation} and \emph{Goal model construction}, which give rise to two different models (top), a \emph{RationalGRL model} and a \emph{GRL model}.
-The GRL part of RationalGRL allows for the creation of goal models by analyzing the non-functional requirements and refining the high-level goals into operationalized tasks.
-For the argumentation part, arguments and counterarguments can be put forward about various parts of this goal model.
+A RationalGRL model is a model in the language that we will explain in the next subsection, while a GRL model is a model in the language we described in Section~\ref{sect:background:grl}.
+The \emph{Goal model construction} part of the RationalGRL framework allows for the creation of goal models by analyzing the non-functional requirements and refining the high-level goals into operationalized tasks.
+In the \emph{Practical reasoning \& argumentation} part, arguments and counterarguments can be put forward about various parts of the goal model.
 These two parts, GRL and argumentation, can impact the other side so that the models can be refined or new critical questions and argument schemes can be instantiated.
@@
 Thus, in the framework it is possible to trace a goal model back to the original discussions about goals, tasks and requirements.
+The GRL model is shown  on the right-hand side of the framework while the argumentation model is on the left-hand side.
+The links between the two sides illustrate the impacts and relationships between two sides.
+Note that answer to critical questions and argument schemes that are instantiated during the analysis phase of the GRL model are documented with the GRL model and can be referred to in the future.
 
@@
 \end{tabularx}
-\caption{List of argument schemes (AS0-AS13), critical questions (CQ0-CQ12), and the effect of answering them (right column).}
+\caption{List of argument schemes (AS0-AS13), critical questions (CQ0-CQ12), and the effect of answering the CQs (right column).
+The effects are for the CQs only, and do not apply to the argument schemes.}
 \label{table:argument-schemes}
@@
 In Section~\ref{sect:overview:as}, we continue our discussion of the argument schemes and critical questions for practical reasoning and argumentation, fitting these schemes and questions into our framework.
-In Section~\ref{sect:overview:lang}, we then discuss the language for RationalGRL models an we provide a metamodel, linking the new concepts of our language to GRL.
+In Section~\ref{sect:overview:lang}, we discuss the language for RationalGRL models and we provide a metamodel, linking the new concepts of our language to GRL.
 In Section~\ref{sect:overview:examples}, we provide extensive examples from our case study.
@@
 It is important to note that the list we provide here is not exhaustive.
-It is merely the result of our empirical study, but new new argument schemes and questions can be added depending on the problem domain.
+It is merely the result of our empirical study, but new argument schemes and questions can be added depending on the problem domain.
 In fact, our tool (Section~\ref{sect:tool}) contains many more critical questions, and it is fully extensible, meaning that new argument schemes and critical questions can be added easily.
@@
 These questions can be answered with ``yes'' or ``no'', and the type of answer has an effect on the original argument (\textsf{INTRO}, \textsf{DISABLE}, \textsf{REPLACE}).
-This will be further explained in Section~\ref{sect:overview:examples} an in Section~\ref{sect:formalframework}.
+This will be further explained in Section~\ref{sect:overview:examples} and in Section~\ref{sect:formalframework}.
 
@@
 \item \emph{Argument}: This represents an argument that does not directly correspond to a GRL element.  
-\item \emph{Rejected (Disabled) GRL element}: If an argument or GRL element is attacked by an argument that itself is not attacked, then this GRL element will be rejected.
+\item \emph{Rejected (Disabled) GRL element}: If an argument or a GRL element is attacked by an argument that itself is not attacked, then this GRL element will be rejected.
 Note Figure~\ref{fig:rationalgrllegend} only shows one type of disabled element, a Task, but all the elements (IEs and links) can be disabled in RationalGRL.
-\item \emph{Attack Link}: An attack link can occur between an argument and another argument or GRL element. It means that the source argument attacks the target argument or GRL element.
+\item \emph{Attack Link}: An attack link can occur between an argument and another argument or GRL element. It means that the source argument attacks the target argument or the GRL element.
 \end{itemize}
@@
 When an argument scheme is instantiated, we obtain an \textsf{Argument}.
-Therefore, each argument is associated with exactly one scheme, but a scheme can be instantiated in multiple ways.
+Therefore, each argument is associated with at most one scheme, but a scheme can be instantiated in multiple ways.
 When a critical question is answered, we may obtain an \textsf{AttackLink}, an \textsf{Argument} or both, depending on the answer.
@@
 
-\subsection{From Practical Reasoning to RationalGRL Models: examples from the case study}
+\subsection{From Practical Reasoning to RationalGRL Models: Examples from the Case Study}
 \label{sect:overview:examples}
 
-We now turn to the interactions between the \emph{practical reasoning \& argumentation} (i.e.
+We now discuss the interactions between the \emph{practical reasoning \& argumentation} (i.e.
 the bottom left element of the framework in Figure~\ref{fig:rationalgrl-framework}) on the one hand, and \emph{RationalGRL models} (i.e.
@@
 This attack leads to the original argument being \emph{rejected} (cf.
-Section~\ref{sect:background:pras} and Section~\ref{sect:formalframework:rationalgrl}), indicated by it being greyed out.
+Section~\ref{sect:background:pras} and Section~\ref{sect:formalframework:rationalgrl}), indicated by being greyed out.
 As a result of this, the corresponding GRL task \emph{Add traffic light} is also disabled.
@@
 \caption{Replacing an element (operation \textsf{REPLACE)}.}
-\label{fig:examples:decomposition}
+\label{fig:examples:replace}
 \end{figure}
@@
 The excerpt of transcript $t_3$ used for this example is shown in Table~\ref{table:transcript:decomposition} in Appendix~\ref{sect:transcripts:excerpts}.
-It consists of a discussion about the type of decomposition relationship for the goal \emph{Simulate Traffic} (Figure~\ref{fig:examples:decomposition}).
-Recall that in Example 1, an AND-decomposition was introduced for this goal with AS5 (Figure~\ref{fig:example_AS}).
-In the discussion CQ5c -- ``Is the decomposition type correct?'' -- is explicitly asked.
+It consists of a discussion about the type of decomposition relationship for the goal \emph{Simulate Traffic} (Figure~\ref{fig:examples:replace}).
+Recall that in Example 1, an AND-decomposition was introduced for this goal with AS5.
+In the discussion of the current example, CQ5c -- ``Is the decomposition type correct?'' -- is explicitly asked.
 The answer is ``No, it should be OR''.

### 2018-05-02 SUBMITTED_FILES/REJ_Submission_MarcvanZee_RationalGRL/tool.tex

\section{The RationalGRL Tool}
\label{sect:tool}

An important requirement of our framework is that it has tool support. This is for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, there does not yet exist any tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent verification to ensure our formal framework and algorithms actually work, iv) having a light-weight web-based version of GRL is useful for the community in general.

GRL has a well-documented and well-maintained tool called jUCMNav~\cite{jUCMNav}. This tool is an extension to Eclipse. Although it is a rich tool with many features, it is implemented as an Eclipse plugin. We have experienced on various occasions that setting this up properly can take up a significant time of experiments, leaving less time for actual modeling.

In this section we present our web-based prototype of RationalGRL. It can be accessed from:
\begin{quote}
\url{http://www.rationalgrl.com}
\end{quote}

The tool is written in Javascript, and runs on all modern browsers. The tool is open-source, and the source can be accessed from:
\begin{quote}
\url{http://www.github.com/marcvanzee/RationalGRL}
\end{quote}

In this section we briefly highlight some features of the language, and we explain some of its current limitations.

\todo{all}{marc}{Explain figures in Appendix~\ref{sect:tool-screenshots}}

### 2018-05-02 abstract.tex

@@
 Based on a case study in which stakeholders discuss requirements for a Traffic Simulator, we apply argumentation techniques from artificial intelligence to a goal modeling approach.
-Thus, we combine a traditional goal modelling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).
+We combine a simplified version of a traditional goal modeling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).
 RationalGRL provides a methodology, formal semantics and tool support to capture the discussions and outcomes of the argumentation process that leads to a goal model.

### 2018-05-02 formalframework.tex

@@
 
-These type of algorithms are short, and consist simply of adding an argument for the element that is being added.
-
-\begin{algorithm}[h]
-  \caption{AS0: $a$ is an actor}\label{alg:add-actor}
-  \begin{algorithmic}[1]
-    \Procedure{$AS_0$}{$n$}
-    \State $A \leftarrow (mintId(), Actor, a)$ \label{alg:as0:1}
-    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:as0:2}
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-Algorithm~\ref{alg:add-actor} takes one argument, namely the name of the actor $a$.
-On line~\ref{alg:as0:1} of the algorithm, a new (unique) id is minted as the identifier of the new actor, which is assigned with its corresponding name to the variable $actor$.
-On line~\ref{alg:as0:2} the actor is added as a new argument.
-As an example, we can formalize the addition of actor ``Traffic Simulator'' to an empty RationalGRL model (i.e.
-$Arg=\emptyset$ and $Att=\emptyset$) as executing the algorithm $AS_0$(Traffic simulator), which results in a RationalGRL model with argument $A_0 = (0, Actor, \text{Traffic simulator})\in Arg$.
-
-\begin{algorithm}[h]
-  \caption{AS1: Actor with id $i$ has resource $R$}\label{alg:add-resource}
-  \begin{algorithmic}[1]
-    \Procedure{$AS_1$}{$i, R$}
-    \State $res\_id\gets mintId()$\label{alg:add-resource:id}
-    \State $A_1\leftarrow (res\_id, Resource, R, AND)$\label{alg:add-resource:arg1}
-    \State $A_2\leftarrow (ActIE, i, res\_id)$\label{alg:add-resource:arg2}
-    \State $Arg\gets Arg\cup \{A_1,A_2\}$\label{alg:add-resource:add-arg}
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-In Algorithm~\ref{alg:add-resource}, we have slightly reworded the original AS1 from Table~\ref{table:argument-schemes} to ``Actor with id $i$ has resource $R$''.
-Furthermore, we assume actor $a$ already exists when Algorithm~\ref{alg:add-actor} is run.
-The algorithm itself runs as follows: on line~\ref{alg:add-resource:id}, a unique id is assigned to variable $res\_id$ (``resource identifier'').
-On line~\ref{alg:add-resource:arg1}, an argument for a resource with name $R$ and identifier $res\_id$ is assigned to $A_1$.
-On line~\ref{alg:add-resource:arg2}, a second argument $A_2$ is created, which is an argument for an Actor-IE relation (Def.~\ref{def:act-ie-relation}) between the input actor $i$ and the newly created element $res\_id$.
-Finally, on line~\ref{alg:add-resource:add-arg}, the arguments are added to the set of arguments $Arg$.
-
-As an example, consider the addition of resource ``Car objects'' to the RationalGRL model we just discussed (where $A_0 = (0, Actor, \text{Traffic simulator})$).
-Adding the resource can then be formalized as executing $AS1(0, \text{Car objects})$, resulting in the addition of two arguments $(1, Resource, \text{Car object}, AND)$ and $(ActIE, 0, 1)$ which are added to the set of arguments $Arg$.
-
-Since arguments schemes AS2-AS4 are very similar to AS1, we omit these algorithms here.
+These type of algorithms are short, and consist simply of adding an argument for the element that is being added. We illustrate this with a single example algorithm for AS5.
 
@@
 
-Similar to the previous algorithm, we have slightly reworded critical question AS5 in Algoritm~\ref{alg:add-decomp}.
+We have slightly reworded critical question AS5 in Algoritm~\ref{alg:add-decomp}.
 We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $T$ is a decomposition of $G$.
@@
 As an example, suppose we are adding the decomposition of goal ``Generate cars'', expressed as argument $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, into task ``Keep same cars'', and suppose we have the following argument for the goal: .
-Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars}, AND)$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.
+Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars})$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.
 
-The remaining argument schemes AS6-AS12 are all similar to algorithms~\ref{alg:add-actor}-\ref{alg:add-decomp} and have been omitted here.
-
-The critical questions of type \emph{INTRO} are very similar as well, with one exception: they require an answer.
+The critical questions of type \emph{INTRO} are all very similar, with one exception: they require an answer.
 For instance, suppose CQ5b: ``Does goal $G$ decompose into other tasks?'' is answered with: ``Yes, namely into task $T$''.
@@
 This is the same for all the other critical questions of type \emph{INTRO}.
-Therefore, we have omitted them here as well.
+Therefore, they are not shown here as well.
 
@@
 As discussed before, algorithms of type \emph{DISABLE} consist of adding a new argument attacking an existing argument.
-The argument that is added is itself not an argument for a GRL element or link, but rather a generic argument (Definition~\ref{def:generic-argument}).
+The argument that is added is not an argument for a GRL element or link, but it is rather a generic argument (Definition~\ref{def:generic-argument}).
 
-In all of these algorithms, we assume the critical question is answered as indicated in the right-most column of Table~\ref{table:argument-schemes}.
+In the following algorithm, we assume the critical question is answered affirmatively, as indicated in the right-most column of Table~\ref{table:argument-schemes}.
 For instance, for critical question CQ0 ``Is the actor relevant?'', we assume it is answered with ``No''.
+In this case, an action is required.
+In contrary, if the answer to the question is ``Yes'', no action is required.
 
@@
 
-Critical questions C1-CQ3 are all very similar to CQ0 and have therefore been omitted here.
-
-\begin{algorithm}[h]
-  \caption{CQ5a: Does the goal with id $g\_id$ decompose into task with id $t\_id$? No}\label{alg:no-decomp}
-  \begin{algorithmic}[1]
-    \Procedure{$CQ5a$}{$g\_id,t\_id$}
-    \State $A \leftarrow (mintId(),GenArg,CQ5a)$\label{alg:no-decomp:genarg}
-    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:no-decomp:genarg2}
-    \For{$A_j\in \{(k,Decomp,g\_id,t\_id)\in Arg\}$}\label{alg:no-decomp:for}
-      \State $Att \leftarrow Att \cup \{(A,A_j)\}$\label{alg:no-decomp:att}
-    \EndFor
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-Algorithm~\ref{alg:no-decomp} is structurally very similar to Algorithm~\ref{alg:actor-not-relevant}, with the only difference that instead of iterating over actors, we now iterate over decomposition links.
-This is done in line~\ref{alg:no-decomp:for}, where we iterate over all decomposition links with source id $g\_id$ and target id $t\_id$ (Definition~\ref{def:link}).
-This means we iterate over all decomposition links from the goal to the task, and we attack all arguments for these links on line~\ref{alg:no-decomp:att} with the newly created argument.
-
-Almost all of the remaining critical questions of type DISABLE are similar in structure.
-CQ11 (``Is the element relevant/useful'') is slightly different since the attack element is not of a specific type, but is simply any GRL element.
-However, the resulting algorithm is very similar to the previous two and has therefore been omitted.
-The only other algorithm of type \emph{DISABLE} that we discuss is $Att$.
-
-\begin{algorithm}[b]
-  \caption{Att: Generic counter-argument on arguments $A_1,\ldots,A_n$ with name $N$}\label{alg:generic}
-  \begin{algorithmic}[1]
-    \Procedure{$Att$}{$A_1,\ldots,A_n$, $N$}
-    \State $A \leftarrow (mintId(),GenArg, N)$\label{alg:generic:arg}
-    \State $Arg\leftarrow Arg \cup \{A\}$\label{alg:generic:addarg}
-    \For{$A_j\in \{A_1,\ldots,A_n\}$}\label{alg:generic:for}
-      \State $Att \leftarrow Att \cup \{(A,A_j)\}$\label{alg:generic:att}
-    \EndFor
-    \EndProcedure
-  \end{algorithmic}
-\end{algorithm}
-
-Algorithm~\ref{alg:generic} can be regarded as the most general way of providing counterarguments to arguments.
-In all of the previous \emph{DISABLE} algorithm, the attack was on a specific type of argument, for instance an argument for an actor or an argument for a decomposition.
-In this algorithm, however, \emph{any} set of arguments can be attacked by a new argument.
-
-Let us reconsider Figure~\ref{fig:examples:relevant-actor}, which we previously formalized as a RationalGRL model where $Arg=\{A_1,A_2\}$, $A_1 = \{(0,Actor,\text{Development team}\})$, $A_2= (1,$ $GenArg,$ CQ0: Professor develops software$)$ and $Att=(A_2,A_1)$.
-Suppose we execute algorithm $Att(\{A_2\},$ Professor does not develop software$)$.
-This results in $A_3=(2, GenArg,$ CQ0: Professor does not develop software$)$ and a new attack link $(A_3,A_2)$ (Figure~\ref{fig:examples:relevant-actor2}).
-Since the argument for the goal is now part of the preferred extension, it does show up in the resulting GRL model after the translation of Algorithm~\ref{alg:translation:to-grl}.
+Algorithms for the remaining critical questions are omitted here since they are not very complex.
 
@@
 
-As an example, take Figure~\ref{fig:examples:decomposition}.
+As an example, take Figure~\ref{fig:example_AS}.
 The initial RationalGRL model (before applying CQ10b), can be formalized as follows: $RatGRL=(Arg,Att)$, with $Arg=\{A_1,A_2,A_3,A_4,A_5\}$ and $Att=\emptyset$, where:

### 2018-05-02 overview.tex

@@
 \begin{figure*}[t]
-\includegraphics[width=\textwidth]{metamodel/metamodel}
+\includegraphics[width=\textwidth]{metamodel/metamodel2}
 \caption{The RationalGRL metamodel. The \emph{Goal model construction} package (bottom) is the GRL metamodel. The \emph{Practical reasoning \& argumentation} package is the RationalGRL extension}
@@
 An \textsf{ArgumentScheme} represents a scheme containing variables.
-\textsf{CriticalQuestions} are possible ways to attack or elaborate an argument based on a scheme; each critical question applies to exactly one scheme, but for each scheme there may be more than one applicable critical question.
 When an argument scheme is instantiated, we obtain an \textsf{Argument}.
-Therefore, each argument is associated with exactly one scheme, but a scheme can be instantiated in multiple ways.
-When a critical question is answered, we may obtain an \textsf{AttackLink}, an \textsf{Argument} or both, depending on the answer.
-Note that it is also possible for an \textsf{AttackLink} to be associated with no critical questions.
-This allows the user to create attacks between arguments, which do not necessarily correspond to one of the critical questions.
+Therefore, each argument is associated with at most one scheme, but a scheme can be instantiated in multiple ways.
 A \textsf{RationalGRLmodel} is composed out of arguments and attack relations.

### 2018-05-02 Response to reviewers.txt
Addressed some more comments by reviewers

@@
 ===== REPLY #1
-Currently 6 of the 44 references in the article come from either REJ or the RE conference:
-- Y.
-Elrakaiby, A.
-Ferrari, P.
-Spoletini, S.
-Gnesi, and B.
-Nuseibeh.
-Using argumentation to explain ambiguity in requirements elici- tation interviews.
-In Proceedings of the 25th IEEE International Requirements Engineering Conference (RE 2017), pages 51–60.
-IEEE Press, 2017.
-- I. Jureta, S. Faulkner, and P. Schobbens. Clear justification of mod- eling decisions for goal-oriented requirements engineering. Re- quirements Engineering, 13(2):87–115, May 2008.
-- Y.
-Liu, Y.
-Su, X.
-Yin, and G.
-Mussbacher.
-Combined goal and feature model reasoning with the user requirements notation and jucmnav.
-In Requirements Engineering Conference (RE), 2014 IEEE 22nd International, pages 321–322.
-IEEE, 2014.
-- P.
-K.
-Murukannaiah, A.
-K.
-Kalia, P.
-R.
-Telangy, and M.
-P.
-Singh.
-Resolving goal conflicts via argumentation-based analysis of com- peting hypotheses.
-In Proceedings of the 23rd International Re- quirements Engineering Conference (RE 2015), pages 156–165.
-IEEE, 2015.
-- M.
-van Zee, F.
-Bex, and S.
-Ghanavati.
-Rationalization of Goal Models in GRL using Formal Argumentation.
-In Proceedings of the 23rd International Requirements Engineering Conference (RE’15), pages 220–225.
-IEEE Press, August 2015
-- E. S. K. Yu. Towards Modeling and Reasoning Support for Early-Phase Requirements Engineering. In Proceedings of the 3rd IEEE International Symposium on Requirements Engineering, pages 226–235, 1997.
-
-We have searched through the journal and conference database for more references, and we found 4 more references that are closely related to our work:
-- S. R. Smith. Constructing and revising formal arguments in re- quirements engineering. Requir. Eng., 2(4):199–216, Dec. 1997.
-- J.Lockerbie,N.A.M.Maiden,J.Engmann,D.Randall,S.Jones, and D.
-Bush.
-Exploring the impact of software requirements on system-wide goals: a method using satisfaction arguments and i* goal modelling.
-Requirements Engineering, 17(3):227–254, 2012.
-- J. Horkoff and E. Yu. Interactive goal model analysis for early requirements engineering. Requirements Engineering, 21(1):29– 61, 2016.
-- J. Hassine and D. Amyot. A questionnaire-based survey method- ology for systematically validating goal-oriented models. Require- ments Engineering, 21(2):285–308, 2016.
-
+Currently 6 of the 44 references in the article come from either REJ or the RE conference [1,2,3,4,5,6].
+We have searched through the journal and conference database for more references, and we found 4 more references that are closely related to our work [7,8,9,10].
+We have added a discussion of these articles to the related work section, at the end of the paragraph "Requirements Engineering".
 This means we currently have 10 references from the Requirements Engineering Journal or presented at the Requirements Engineering series of conferences.
@@
 ===== REPLY #1
-We thank the reviewer for pointing this out.
-We indeed believe critical questions should not be part of the metamodel, and we have removed them accordingly.
-We initially had put them into the metamodel because they play an important part in the implementation, but they are by themselves not part of the RationalGRL language.
-We also have updated the text accordingly.
+We thank the reviewer for pointing this out. We agree critical questions should not be part of the metamodel, and we have removed them accordingly. We also have updated the text accordingly.
 
@@
 ===== REPLY #2
-We agree that many algorithms are very much the same, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two or more.
+We agree that many algorithms are similar, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two or more.
 
@@
 ===== REPLY #3
-TODO: WRITE WHEN ANALYSIS IS DONE
+We agree with the reviewer that a real validation of the approach is a necessary next step.
+However, we also would like to point out that an industrial strength case study is a significant amount of work, and would most likely have enough new insights for a separate paper.
+To give an example, it took Estrada et al.
+[11] about 9 months to conduct the case study.
+
+Since the main focus of our article is to explain the noval RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content.
+Still, we agree that having no validation at all is limiting, so we have performed two small-scale user studies in order to provide an initial validation of our framework.
+These two studies are as follows:
+1) We evaluated the RationalGRL tool with two Enterprise Architects from a large bank in the Netherlands.
+The architects found the RationalGRL models very useful in that they could visualize the arguments and were easier for them to understand the links between the rationale, and the resulting goal model elements.
+2) We performed an empirical study amongst ==> TODO(marcvanzee): ADD NUMBER HERE  <== participants, each with a computer science background.
+
+We have added a section "Validation" (Section 7) where we describe the results of the validation.
 
@@
 ===== REPLY #1
-TODO: WRITE WHEN ANALYSIS IS DONE
+We have added a section "Validation" (Section 7) where we describe the results of a validation. Please see REPLY #3 to REVIEWER #1 for more details.
 
@@
 ===== REPLY #3
-%FB: **Floris to write on GRL beliefs in 7.1**
+We have added a comparison between arguments in RationalGRL and beliefs in GRL at the end of the related work section, in the paragraph "Beliefs in GRL".
+
+The comparison between argumentation and belief revision is interesting, especially from a theoretical point of view.
+In the area of knowledge representation, obtaining correspondences between beliefs revision and the dynamics of argumentation theory is a popular topic of research [12,13,14,15].
+However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a theoretical point of view, we have left out the theoretical comparison.
+Instead, we chose to compare it with beliefs in GRL.
 
@@
 ===== REPLY #9
-We’ve renamed “verification” to “demonstration” and slightly reworded the sentence.
+We have renamed “verification” to “demonstration” and slightly reworded the sentence.
 
@@
 ===== REPLY #10
-**Sepideh**
+We agree that the terminology is wrong.
+We have renamed the methodology to "The RationalGRL development process", since it defines the steps on how to create a RationalGRL model.
+We have renamed this throughout the entire article.
 
@@
 ===== REPLY #11
-**Sepideh**
+We agree with the reviewer that the two mentioned sentences are confusing. We have improved this as follows:
+- We have rewritten the first sentence of the abstract as follows: "Goal-oriented requirements modeling approaches aim at capturing intentions of various stakeholders involved in the development of an information system and refining them into low level goals and tasks."
+- We have deleted the first sentence of the introduction, since we feel that it is not necessary to explain the definition of Requirements Engineering in the Requirements Engineering Journal.
 
@@
 We’ve ran a spell-checker and proof-read the paper carefully. There was one unresolved reference to a figure, which a corrected. Thank you for pointing this out.
+
+######## REFERENCES ########
+[1] Y.
+Elrakaiby, A.
+Ferrari, P.
+Spoletini, S.
+Gnesi, and B.
+Nuseibeh.
+Using argumentation to explain ambiguity in requirements elici- tation interviews.
+In Proceedings of the 25th IEEE International Requirements Engineering Conference (RE 2017), pages 51–60.
+IEEE Press, 2017.
+[2] I. Jureta, S. Faulkner, and P. Schobbens. Clear justification of mod- eling decisions for goal-oriented requirements engineering. Re- quirements Engineering, 13(2):87–115, May 2008.
+[3] Y.
+Liu, Y.
+Su, X.
+Yin, and G.
+Mussbacher.
+Combined goal and feature model reasoning with the user requirements notation and jucmnav.
+In Requirements Engineering Conference (RE), 2014 IEEE 22nd International, pages 321–322.
+IEEE, 2014.
+[4] P.
+K.
+Murukannaiah, A.
+K.
+Kalia, P.
+R.
+Telangy, and M.
+P.
+Singh.
+Resolving goal conflicts via argumentation-based analysis of com- peting hypotheses.
+In Proceedings of the 23rd International Re- quirements Engineering Conference (RE 2015), pages 156–165.
+IEEE, 2015.
+[5] M.
+van Zee, F.
+Bex, and S.
+Ghanavati.
+Rationalization of Goal Models in GRL using Formal Argumentation.
+In Proceedings of the 23rd International Requirements Engineering Conference (RE’15), pages 220–225.
+IEEE Press, August 2015
+[6] E.
+S.
+K.
+Yu.
+Towards Modeling and Reasoning Support for Early-Phase Requirements Engineering.
+In Proceedings of the 3rd IEEE International Symposium on Requirements Engineering, pages 226–235, 1997.
+
+[7] S. R. Smith. Constructing and revising formal arguments in re- quirements engineering. Requir. Eng., 2(4):199–216, Dec. 1997.
+[8] J.Lockerbie,N.A.M.Maiden,J.Engmann,D.Randall,S.Jones, and D.
+Bush.
+Exploring the impact of software requirements on system-wide goals: a method using satisfaction arguments and i* goal modelling.
+Requirements Engineering, 17(3):227–254, 2012.
+[9] J. Horkoff and E. Yu. Interactive goal model analysis for early requirements engineering. Requirements Engineering, 21(1):29– 61, 2016.
+[10] J. Hassine and D. Amyot. A questionnaire-based survey method- ology for systematically validating goal-oriented models. Require- ments Engineering, 21(2):285–308, 2016.
+
+[11] Estrada, H., Rebollar, A.
+M., Pastor, O., & Mylopoulos, J.
+(2006, June).
+An empirical evaluation of the i* framework in a model-based software generation environment.
+In International Conference on Advanced Information Systems Engineering (pp.
+513-527).
+Springer, Berlin, Heidelberg.
+
+[12] Falappa, Marcelo Alejandro, Gabriele Kern-Isberner, and Guillermo Ricardo Simari.
+"Belief revision and argumentation theory." Argumentation in artificial intelligence.
+Springer, Boston, MA, 2009.
+341-360.
+[13] Liao, Beishui, Li Jin, and Robert C. Koons. "Dynamics of argumentation systems: A division-based method." Artificial Intelligence 175.11 (2011): 1790-1814.
+[14] Dix, Jürgen, et al. "Belief change and argumentation in multi-agent scenarios (Dagstuhl seminar 13231)." Dagstuhl Reports. Vol. 3. No. 6. Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik, 2013.
+[15] Paglieri, Fabio, and Cristiano Castelfranchi.
+"Revising beliefs through arguments: Bridging the gap between argumentation and belief revision in MAS." International Workshop on Argumentation in Multi-Agent Systems.
+Springer, Berlin, Heidelberg, 2004.

### 2018-05-02 abstract.tex
Addressed some more comments by reviewers

@@
 \begin{abstract}
-Goal modeling languages capture the relations between an information system and its environment using high-level goals and their relationships with lower level goals and tasks.
+Goal-oriented requirements modeling approaches aim at capturing intentions of various stakeholders involved in the development of an information system and refining them into low level goals and tasks.
 The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
@@
 We combine a simplified version of a traditional goal modeling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).
-RationalGRL provides a methodology, formal semantics and tool support to capture the discussions and outcomes of the argumentation process that leads to a goal model.
+RationalGRL provides a formal semantics and tool support to capture the discussions and outcomes of the argumentation process that leads to a goal model.
+We also define the RationalGRL development process to create a RationalGRL model.
 

### 2018-05-02 discussion.tex
Addressed some more comments by reviewers

@@
 Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE.
-The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a methodology and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
+The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
 
@@
 The GAM process is essentially a generic, high-level process for problem solving, not specifically tailored towards goal modeling.
-In this sense, the RationalGRL methodology in Figure~\ref{fig:rationalgrl-methodology} can be seen as a more specific version of GAM explicitly meant for goal modeling.
+In this sense, the RationalGRL Development Process in Figure~\ref{fig:rationalgrl-methodology} can be seen as a more specific version of GAM explicitly meant for goal modeling.
 The argument schemes and critical questions in RationalGRL provide clearer handles for goal modeling (cf.
@@
 RationalGRL overcomes this problem by integrating argumentation and goal modeling in a single modeling language.
+
+\paragraph{Beliefs in GRL and Belief Revision} As mentioned already in the introduction, it is possible in GRL to rationalize models
+through so-called \emph{belief} elements.
+These beliefs are elements that can be attached to an IE with a \emph{belief link}.
+The content of a belief is a textual description, which can explain, for instance, the rationale for including an IE in the resulting goal model.
+There are several important differences between GRL beliefs and RationalGRL arguments.
+
+First of all, beliefs in GRL are not used in any of the algorithms to compute the satisfiability of the IEs.
+This means that a GRL belief can be understood as metadata that can be attached to an element, but it does not have semantics.
+Argument in RationalGRL, in contrast, play an important role in forming the resulting goal model.
+If an IE is defeated by an argument, it is not part of the resulting goal model.
+
+Secondly, arguments in RationalGRL can not only provide rationale for IEs in the resulting goal model, but they can also provide rational for those that are rejected, and they can be attached to links as well.
+This means that Rational arguments are strictly more expressive than GRL beliefs.
+
+Finally, the formation of arguments and counterarguments in RationalGRL is guided by critical questions, which stimulate discussions between stakeholders. This is not possible for GRL beliefs.
 
@@
 However, a more complex problem on which multiple persons have to work asynchronously might benefit from the extra tools offered by RationalGRL.
-Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the methodology, the tool, or a list of critical questions to serve as reminders during the goal modeling process (cf.
+Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the development process, the tool, or a list of critical questions to serve as reminders during the goal modeling process (cf.
 \cite{SchriekEtal2016}).
@@
 One of our main goals is to provide means for RE practitioners to capture the underlying arguments of goal models by using the RationalGRL framework (\textbf{requirement 4}).
-To this end, we propose a methodology, which consists of developing goal models and posing arguments based on schemes in an integrated way.
+To this end, we propose a development process, which consists of developing goal models and posing arguments based on schemes in an integrated way.
 Finally, we have implemented the RationalGRL tool, a web-based prototype\footnote{\url{http://www.rationalgrl.com}}, for modeling argument schemes and critical questions and for reasoning about goal models with respect to them (\textbf{requirement 5}).

### 2018-05-02 introduction.tex
Addressed some more comments by reviewers

@@
 
-Requirements Engineering (RE) is an approach to assess the role of a future information system within its environment.
-An important goal in RE is to produce a consistent and comprehensive set of requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
-
+An important goal in Requirements Engineering (RE) is to produce a consistent and comprehensive set of requirements covering different aspects of the system, such as general functional requirements, operational environment constraints, and so-called non-functional requirements such as security and performance.
 Among the ``early-phase'' requirements engineering activities are those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what the implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
@@
 \item 
-There must be a methodology to guide practitioners in using the framework.
+A process explaining how practitioners can develop models in the framework must be identified.
 \item 
@@
 
-In order to support practitioners in using the RationalGRL framework, we propose a methodology, which consists of developing goal models and posing arguments based on schemes in an integrated way.
+In order to support practitioners in using the RationalGRL framework, we propose the RationalGRL development process, which explains steps for  developing goal models and posing arguments based on schemes in an integrated way.
 We implement a web-based tool for the RationalGRL framework as well, which works on any modern browser and is developed in Javascript.\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
@@
 In Section~\ref{sect:formalframework}, we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and vice versa, and we develop various algorithms that change a RationalGRL model according to an argument scheme or a critical question.
-Section~\ref{sect:methodology+tool} discusses the RationalGRL methodology and explains various features of the tool we developed.
+Section~\ref{sect:methodology+tool} discusses the RationalGRL development process and explains various features of the tool we developed.
 Finally, Section~\ref{sect:discussion} covers the related work, the future work, and the conclusion.

### 2018-05-02 tool.tex
notes formal framework and small changes
finished all algorithms
Addressed some more comments by reviewers

--- 2017-01-21 tool.tex
@@
-\section{The RationalGRL Tool}
-\label{sect:implementation}
+\section{The RationalGRL Development Process and Tool}
+\label{sect:methodology+tool}
+
+In previous sections, we have shown how the RationalGRL framework can capture stakeholders discussions, and how interactions between two types of reasoning, practical reasoning and goal modeling, leads to two interlinked models, RationalGRL and GRL models.
+The previous section contained a formalization of RationalGRL, which forms the starting point of this section.
+In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a development process (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
+
+\subsection{RationalGRL Development Process}
+\label{sect:methodology}
+
+We propose the development process shown in Figure~\ref{fig:rationalgrl-methodology} to develop a RationalGRL model, a version of which was presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.
+Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+The rest of the steps are as follows:
+
+\textbf{(1) Instantiate Argument Schemes (AS)} -- In this step, we start from the list of arguments schemes (Table~\ref{table:argument-schemes}).
+Whilst discussing the requirements, we select schemes from the list and instantiate them to form arguments for GRL IEs or links.
+In this way we build the GRL model by introducing new GRL elements (\textsf{INTRO}).
+For example, an argument scheme can be "Goal \emph{G} contributes to softgoal \emph{S}".
+When an argument scheme is instantiated, it corresponds to an argument for or against part of a goal model.
+
+\textbf{(2) Answer Critical Questions (CQs)} -- After building or modifying the initial GRL model, we ask the relevant critical questions.
+Since each element in the GRL model corresponds to an instantiated scheme, we can look at Table~\ref{table:argument-schemes}) to see which questions are relevant given our GRL model.
+For example, for the argument scheme, "Goal \emph{G} contributes to softgoal \emph{S}", there are two critical questions as follows:  \emph{Does the goal contributes to the softgoal?} and \emph{Does the goal contributes to some other softgoals?}.
+When the analyst answers  a critical question, a new argument scheme may be instantiated.
+This is done in step (3) when the operation INTRO is executed.
+
+\textbf{(3) Decide on Intentional Elements and their Relationships} -- By answering a critical question, one of the three operations (\textsf{INTRO}, \textsf{DISABLE} or \textsf{REPLACE}) is applied to the GRL model.
+Any of these operations impact the arguments and corresponding GRL intentional elements, modifying the initial GRL model into a RationalGRL model.
+Recall that \textsf{INTRO} means that
+a new argument scheme is created.
+That means, the current argument related to the critical question does not get attacked.
+In the case of \textsf{DISABLE}, the intentional element or its related links are disabled in the model.
+\textsf{REPLACE} introduces a new argument and attacks the original argument at the same time.
+This means that the original element of the argument scheme is replaced with a new one.
+
+\textbf{(4) Modify GRL Models} -- Based on the RationalGRL model of step (3), the GRL model can be modified: 1) a new intentional element or a new link is introduced; 2) an existing intentional element or an existing link gets disabled (removed) from the model; or 3) an existing intentional element or link is replaced by a new one.
+This results in a new, modified GRL model, which can be used as the basis for another cycle of the RationalGRL Development Process.
+
+We can continue these four steps until there is no more intentional element or link to analyze or we reach a satisfactory model.
+In the next section, we will give an example of how our tool can be used together with the methodology to build a GRL model.
+
+\begin{figure*}[t]
+\centering
+\includegraphics[scale=0.8]{img/tool/goal_details}
+\caption{Overview of the RationalGRL tool}
+\label{fig:tool:overview}
+\end{figure*}
+
+\subsection{The RationalGRL Tool}
+\label{sect:tool}
+
+The final requirement of our framework is that it has tool support (\textbf{requirement 5}).
+This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.
+In this section we briefly highlight some features of the tool, and we explain some of its current limitations.
+The tool can be accessed from:
+
+\begin{quote}
+\url{http://www.rationalgrl.com}
+\end{quote}
+
+\paragraph{Relation to jUCMNav}
+
+GRL is implemented in the open-source tool jUCMNav~\cite{jUCMNav}, which is actively maintained\footnote{\url{http://jucmnav.softwareengineering.ca/foswiki/ProjetSEG}}.
+jUCMNav currently has over 2,000 commits made by over 40 contributors, representing over 200,000 lines of code.
+jUCMNav is used actively in research, and the tool has been extended in many directions over the past few years, such as for modeling and analyzing security misuse cases~\cite{daramola2012ontology}, supporting activity theory \cite{georg2015synergy}, combined goal and feature model reasoning~\cite{liu2014combined}, and enterprise architecture modeling \cite{marosin-etal:caise2016}.
+Furthermore, jUCMNav includes Use Case modeling and contains various GRL evaluation algorithms with which the satisfiability of goals can be determined.
+
+The jUCMNav tool is implemented as an Eclipse plugin.
+While we aim to integrate our framework into jUCMNav in the future, a web-based version of RationalGRL (and GRL) is more suited for current purposes.
+Since RationalGRL does not make use of the many features of jUCMNav, we believe a light-weight version is sufficient for our current aims, and may benefit the community in general.
+Our tool is thus not meant to replace jUCMNav.
+For simplicity we have a web-based version, and it has the functionality to export to jUCMNav.
+
+\paragraph{Tool overview} The RationalGRL tool is an open-source web-based Javascript application, which runs on all modern browsers.
+It is based on our formalization in Section~\ref{sect:formalframework} and provides export functionality to jUCMNav, using the translation algorithm (Algorithm~\ref{alg:translation:to-grl}).
+It contains all of the GRL elements and links, except for beliefs and actors.
+We have argued in Section~\ref{sect:background:grl} that arguments can be seen as an extension to beliefs, which is the reason why we did not implement them.
+Actors are also missing, but they will be added in the future.
+
+When the tool starts, the user is presented with a screen as in Figure~\ref{fig:tool:overview}.
+This screen shows the palette of elements and links (top left pane), a canvas on which RationalGRL models can be built (top right pane), an `Export to GRL' button (bottom left pane), and a details pane of the currently selected elements (bottom right pane).
+Elements and links can be added to the canvas by selecting them on the left and clicking on the canvas, thus capturing the \textsf{INTRO} operation (Section~\ref{sect:formalframework:intro}).
+This makes it possible to build and modify GRL models using the tool.
+
+In total there are four different types of details panes, which we now explain in turn.
+
+\paragraph{IE and links details pane} Figure~\ref{fig:tool:overview} shows the details pane for the goal `Generate cars'.
+The user can change the name of the IE by changing the name and clicking `Rename', which will update the naming history on the bottom of the pane.
+The naming history is (simplified) implementation of the \textsf{REPLACE} operation on IEs, since it is a very simple form of \emph{clarification}.
+The main difference with jUCMNav is that we have a renaming history, so the user can see which names the IE had prior to the current one.
+If the IE has decomposition links, the user can change the decomposition type as well.
+Each IE and link has a set of associated critical questions, which can be answered in the details pane by click on the button next to the question.
+
+\paragraph{Critical question details pane} The details panel can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal IE ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
+As an example, consider Figure~\ref{fig:tool:cqdetails}, where the softgoal ``Easy to use'' is questioned with the relevancy question (CQ11).
+It is possible to select the answer and provide an explanation for the answer.
+
+Assume that in the example of Figure~\ref{fig:tool:cqdetails} the user selects ``No'' and clicks the ``Answer question'' button.
+A new argument is then automatically created that attacks the softgoal and the details pane shows the critical question as answered (Figure~\ref{fig:tool:cqeffect}).
+This is an implementation of a \textsf{DISABLE} algorithm (Section~\ref{sect:formalframework:disable}) similar to Algorithm~\ref{alg:actor-not-relevant}: a new argument is added that attacks the existing argument.
+
+\paragraph{Argument details pane} It is also possible to attack arguments by adding an ``Argument'' element and an ``Attack'' link manually.
+Consider, for example, Figure~\ref{fig:tool:argument}.
+Here, a new argument ``Necessary'', which attacks the previously generated argument based on CQ11, has been added by the user.
+As the details pane shows, this new argument is not based on a CQ.
+It is further worth noting that it is possible to provide further a explanation in argument elements, allowing for more fine-grained rationalizations.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{img/tool/details_softgoal}
+\caption{Critical question details pane for ``Is the softgoal relevant/useful?''}
+\label{fig:tool:cqdetails}
+\end{figure}
+
+\begin{figure}[b]
+\centering
+\includegraphics[width=\columnwidth]{img/tool/attack_softgoal}
+\caption{Effect of answering CQ11: ``Is the softgoal relevant/useful?'' with ``No''}
+\label{fig:tool:cqeffect}
+\end{figure}
+
+\paragraph{Argumentation semantics} The RationalGRL tool computes the acceptability of arguments on the fly.
+In the example of Figure~\ref{fig:tool:cqeffect}, the original (argument for) softgoal ``Easy to use'' is rejected because its only attacker ``CQ11 - Redundant'' is accepted.
+However, if we then attack ``CQ11 - Redundant'' with a new argument ``Necessary'', the original argument for ``Easy to use'' is again accepted because its only attacker is rejected  (Figure~\ref{fig:tool:argument}).
+Note that when computing the acceptability of arguments, the RationalGRL tool makes the assumption that there are no attack cycles in the model and that hence there is one unique preferred extension (cf.
+Section~\ref{sect:formalframework:rationalgrl}) -- if the user creates an attack cycle, an error message is shown.
+
+\begin{figure}[t]
+\centering
+\includegraphics[width=0.8\columnwidth]{img/tool/reinstate_softgoal}
+\caption{Attacking an argument with a generic argument ``Necessary''}
+\label{fig:tool:argument}
+\end{figure}
+
+\paragraph{Argument schemes and critical questions} The tool contains more argument schemes and critical questions than those that we initial collected in Table~\ref{table:argument-schemes}.
+As we mentioned before, our table is not meant to be exhaustive, and it is straightforward to add more argument schemes and critical questions if necessary.
+For instance, the table does not contain critical questions for links between all IEs (e.g., a contribution link from a task to a goal does not occur in the table).
+In the tool, we have added critical questions to all links and IEs.
+We implemented an \emph{argument schemes database}, which is designed to be extended easily.
+
+\paragraph{Export to jUCMNav} The tool implements the RationalGRL to GRL translation (Algorithm~\ref{alg:translation:to-grl}) through an export function.
+RationalGRL models built in the RationalGRL tool can be exported to the \texttt{.grl} file format, which can be imported in jUCMNav.
+Note that in order for the GRL models to be rendered correctly, we recommend using the Graphviz dot\footnote{\url{http://www.graphviz.org/}} as the auto-layout tool (this can be selected under ``Auto layout preferences'' when importing the file in jUCMNav).
+
+Two examples of this export are provided in Appendix~\ref{sect:tool-screenshots}.
+Figure~\ref{fig:tool:figfrompaper} shows the model that was discussed earlier in this paper (Figure~\ref{fig:example-small3}) in the RationalGRL tool.
+Recall that translating this model to GRL provided us with the model in Figure~\ref{fig:example-small}.
+This is also what follows from our export: if we export the RationalGRL model and then import the resulting GRL model in jUCMNav, we get the model in Figure~\ref{fig:tool:figfrompaper1}, which is the same as Figure~\ref{fig:example-small}.
+
+Our translation and export function uses the argument acceptability as a way of determining the GRL model.
+Take for example, the RationalGRL model in Figure~\ref{fig:tool:multipleattack}.
+A positive contribution is added from ``Keep same cars'' to ``Simple design''.
+More inportantly, an argument ``not enough CPU'' has been added to the model.
+This argument attacks the ``Realistic simulation'' softgoal and the ``Create new cars'' task, arguing that there is not enough processing power for either of these to be feasible in the traffic simulator.
+Now, if we export this to GRL, the GRL elements that are rejected or disabled (greyed out in Figure~\ref{fig:tool:multipleattack}) are not included.
+This can be seen in Figure~\ref{fig:tool:multipleattack1}, where the jUCMNav GRL model that was exported from the RationalGRL tool is shown.
+The pair of figures \ref{fig:tool:multipleattack} and \ref{fig:tool:multipleattack1} also nicely shows the added value of RationalGRL: Figure~\ref{fig:tool:multipleattack} shows that there can be a larger discussion and rationalization underlying even a fairly simple goal model such as the one in Figure~\ref{fig:tool:multipleattack1}.
+
+\paragraph{Limitations}
+For usability purposes, the \textsf{REPLACE} operation has been implemented differently than in our formal framework (Section~\ref{sect:formalframework:disable}).
+In the formal framework, a \textsf{REPLACE} introduces a new argument that replaces (and therefore attacks) all previous arguments with the same identifier.
+Including \textsf{REPLACE} like this in the tool would mean that, when the name of an element is changed, we would need to render all the previous versions of the element on the canvas of the tool.
+Since this would quickly become very cumbersome, we decided to implement a naming history for each element.
+For example, in Figure~\ref{fig:tool:overview}, the goal ``Generate cars'' was previously named ``Add cars''.
+In this way, a history is kept of how an element name was clarified without cluttering the model canvas.
+This feature is currently very limited and can be extended in various was, for instance by changing the IE type.
+Similarly, the decomposition type can be changed in the details pane (Figure~\ref{fig:tool:overview}) without introducing new elements that attack the original ones (cf.
+Figure~\ref{fig:example-small3}, where the old decomposition type $AND$ of ``Generate traffic'' is shown as an explicit argument).

### 2018-05-02 validation.tex
Addressed some more comments by reviewers

\section{Validation}
\label{sect:validation}

\subsection{Evaluation with Enterprise Architects}

\todo{marcvanzee}{marcvanzee,floris}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects. I'm not sure if we can re-use it. The main problem is that at that time we developed GRL models specifically for the organization in which the Enterprise Architects were working, and we don't have such a model now. What do you think?}

To ensure GRL models captured the essence of the EA principles correctly, we evaluated the models with two Enterprise Architects from eGovernment. The
architects found the GRL models very useful in that they could visualize the principles and were easier for them to understand the links between the rationale, added value, actions and the goals of the principles. Furthermore, since the EA
principles were linked to the high-level goals of the organization, it made them easier to justify and analyze. Also, we captured the traceability links between the related EA principles which helped the architects to document the links and
verify the connections.

The architects also informed us that assigning values for contribution links can be done at design time together with the architects.
We aim to extend our models to more principles and provide analysis of the principles with high-level goals as well as low-level operationalized activities of the organization.

\subsection{Empirical evaluation}

\todo{marcvanzee}{marcvanzee,florisbex}{Write this}

### 2018-05-03 discussion.tex
small changes discussion

--- 2017-11-12 discussion.tex
@@
 Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE.
-The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a methodology and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
+The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
 
-\paragraph{Requirements Engineering} There are a number of general approaches in the field of requirements engineering that explicitly take into account arguments.
+\paragraph{Requirements Engineering} 
+There are numerous examples in the RE literature that make mention of arguments, questions or conflicts in some way.
+Already in 1997, Smith~\cite{Smith:1997:CRF:2737426.2737493} proposed a framework to develop arguments, which relates a model of a proposed software system to a model of its environment, and shows that these together achieve properties described by a model of requirements.
+In a similar vein, Lockerbie \emph{et al.}~\cite{lockerbie2012exploring} describe procedures that analysts follow to use satisfaction arguments in combination with $i*$.
+Also, Horkoff and Yu~\cite{horkoff2016interactive} introduce analysis procedures for the $i*$ goal-oriented framework, allowing users to ask questions such as ``what if?'', or ``are certain goals achievable?
+how?
+or why not?''.
+Finally, Hassin and Amyot~\cite{hassine2016questionnaire} analyze the correctness of GRL models by performing statistical analysis on generated questionnaires, which allows for early detection of potential conflicts between stakeholders of the system.
+The mentioned work, however, does not consider arguments the way RationalGRL does, that is, as explicit opinions of stakeholders.
+
+A number of approaches in the field of requirements engineering explicitly take into account arguments.
 One early example comes from Haley \emph{et al.}~\cite{haley2008security}, who use formal logical arguments to show that the system behavior satisfies certain security requirements, and more informal arguments to capture and validate the assumptions underlying the system behavior.
@@
 An insufficiency occurs when an analyst needs more information from the client to accept a client's statement.
-The inconsistencies are then captured as mutually attacking arguments, and the insufficiencies as arguments against the original statements.
-These counterarguments say that, for example, the functionality expressed in the statement cannot be realized or is irrelevant.
-Elrakaiby \emph{et al.} coded the data from 34 requirement elicitation interviews.
-They identified 39 inconsistencies (i.e.
-at least two arguments that mutually attack) and 29 insufficiencies (i.e., at least one argument attacking another).
 
@@
 \paragraph{Goal Modeling} Argumentation has been included -- both explicitly and implicitly -- in existing goal modeling languages.
-For example, the belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} is meant to capture the rationales behind the inclusion of goals and tasks in the model.
-Furthermore, relations between elements in a goal model also provide justifications: high-level goals are reasons for lower-level goals, tasks and resources.
+For example, the belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356}\footnote{See also the brief discussion at the end of Section \ref{sect:background:grl}.} is meant to capture the rationales behind the inclusion of goals and tasks in the model.
+Such belief elements can be attached to a goal or task with a belief link.
+The content of a belief is a textual description, which can explain, for instance, the rationale for including a goal in the model.
+A GRL belief is thus essentially metadata that can be attached to an element, but it does not have a formal semantics.
+Arguments in RationalGRL, in contrast, are more than just metadata, as they play a role in determining the final goal model: if a goal or task is defeated by an argument, it is not part of the final goal model.
+In this way, RationalGRL arguments are used to actually formally revise the beliefs underlying the goal model.
+Furthermore, arguments in RationalGRL can not only provide rationales that tell us why certain elements are part of the final goal model, but they can also provide reasons for why certain elements were not included, and they can be attached to links as well.
+Hence, RationalGRL arguments are strictly more expressive than GRL beliefs.
+
+Relations between elements in a goal model provide justifications in and of themselves: high-level goals are reasons for lower-level goals, tasks and resources.
 Hence, refinement and decomposition techniques used in requirements engineering \cite{van2001goal} can be seen as explicit argumentation steps in the goal modeling process.

### 2018-05-03 discussion.tex
validation

--- 2018-05-02 discussion.tex
@@
 
-\paragraph{Beliefs in GRL and Belief Revision} As mentioned already in the introduction, it is possible in GRL to rationalize models
+\paragraph{Beliefs in GRL} As mentioned already in the introduction, it is possible in GRL to rationalize models
 through so-called \emph{belief} elements.

### 2018-05-03 validation.tex
validation

@@
 
+\subsection{RationalGRL user evaluation}
+
+Since RationalGRL is targeted at practitioners and not just for academic end users, we performed a brief user evaluation with 20 users.
+In this evaluation, we asked participants to rate and comment the usability of the RationalGRL tool.
+The study consisted of three parts:
+\begin{enumerate}
+\item Explanation: We explained the participants the basics of the RationalGRL methodology and the RationalGRL tool in a document we wrote.
+\item Modelling: We asked the participants to model a summarized discussion in the RationalGRL tool.
+\item Questionnaire: We asked some questions about the usability of the RationalGRL tool, and how it relates to GRL.
+\end{enumerate}
+
+The instructions for these three parts were contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} At the end of the document we provided a link to the survey\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2}.
+
+In the second task, the users were asked to visit rationalgrl.com, where they could start experimenting with the RationalGRL tool directly.
+We presented the users with summarized transcripts of the discussion about the development of the Traffic Controller, and then asked them to model this to the best of their abilities.
+ADD RESULTS HERE
+
+For the third task, we asked the users to fill out a survey. The survey starts by asking about the users’ backgrounds. The participants turned out to have a relatively high level of experience with
+IT, with an average rating of.... Experience
+with goal modeling was less....
+
+Explain rest of results here....
+
 \subsection{Evaluation with Enterprise Architects}
 
-\todo{marcvanzee}{marcvanzee,floris}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects.
+\todo{marcvanzee}{marcvanzee,floris,sepideh}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects.
 I'm not sure if we can re-use it.
@@
 We aim to extend our models to more principles and provide analysis of the principles with high-level goals as well as low-level operationalized activities of the organization.
-
-\subsection{Empirical evaluation}
-
-\todo{marcvanzee}{marcvanzee,florisbex}{Write this}

### 2018-05-03 Response to reviewers.txt

@@
 
-Since the main focus of our article is to explain the noval RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content.
-Still, we agree that having no validation at all is limiting, so we have performed two small-scale user studies in order to provide an initial validation of our framework.
-These two studies are as follows:
-1) We evaluated the RationalGRL tool with two Enterprise Architects from a large bank in the Netherlands.
-The architects found the RationalGRL models very useful in that they could visualize the arguments and were easier for them to understand the links between the rationale, and the resulting goal model elements.
-2) We performed an empirical study amongst ==> TODO(marcvanzee): ADD NUMBER HERE  <== participants, each with a computer science background.
+Since the main focus of our article is to explain the novel RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content.
+Still, we agree that having no validation at all is limiting, so we have performed a small-scale user study in order to provide an initial validation of our framework.
+This empirical study was performed amongst ==> TODO(marcvanzee): ADD NUMBER HERE  <== participants, each with a computer science background.
 
 We have added a section "Validation" (Section 7) where we describe the results of the validation.
-
-%FB: Marc to write on Enterprise Architects
 
@@
 ===== REPLY #3
-We have added a comparison between arguments in RationalGRL and beliefs in GRL at the end of the related work section, in the paragraph "Beliefs in GRL".
+We have added a comparison between arguments in RationalGRL and beliefs in GRL in the related work section, at the beginning of the paragraph "Goal Modelling".
 
@@
 In the area of knowledge representation, obtaining correspondences between beliefs revision and the dynamics of argumentation theory is a popular topic of research [12,13,14,15].
-However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a theoretical point of view, we have left out the theoretical comparison.
-Instead, we chose to compare it with beliefs in GRL.
+However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on beliefs in GRL.
 
@@
 Nuseibeh.
-Using argumentation to explain ambiguity in requirements elici- tation interviews.
+Using argumentation to explain ambiguity in requirements elicitation interviews.
 In Proceedings of the 25th IEEE International Requirements Engineering Conference (RE 2017), pages 51–60.

### 2018-05-03 discussion.tex

--- 2018-05-03 discussion.tex
@@
 
+\paragraph{Beliefs in GRL} As mentioned already in the introduction, it is possible in GRL to rationalize models
+through so-called \emph{belief} elements.
+These beliefs are elements that can be attached to an IE with a \emph{belief link}.
+The content of a belief is a textual description, which can explain, for instance, the rationale for including an IE in the resulting goal model.
+There are several important differences between GRL beliefs and RationalGRL arguments.
+
+First of all, beliefs in GRL are not used in any of the algorithms to compute the satisfiability of the IEs.
+This means that a GRL belief can be understood as metadata that can be attached to an element, but it does not have semantics.
+Argument in RationalGRL, in contrast, play an important role in forming the resulting goal model.
+If an IE is defeated by an argument, it is not part of the resulting goal model.
+
+Secondly, arguments in RationalGRL can not only provide rationale for IEs in the resulting goal model, but they can also provide rational for those that are rejected, and they can be attached to links as well.
+This means that Rational arguments are strictly more expressive than GRL beliefs.
+
+Finally, the formation of arguments and counterarguments in RationalGRL is guided by critical questions, which stimulate discussions between stakeholders. This is not possible for GRL beliefs.
+
 \subsection{Future work}

### 2018-05-04 Response to reviewers.txt
final comments FB fixed

@@
 ===== REPLY #1
-Currently 6 of the 44 references in the article come from either REJ or the RE conference [1,2,3,4,5,6].
-We have searched through the journal and conference database for more references, and we found 4 more references that are closely related to our work [7,8,9,10].
-We have added a discussion of these articles to the related work section, at the end of the paragraph "Requirements Engineering".
+Currently 6 of the 44 references in the article come from either REJ or the RE conference [11,24,26,32,43,47].
+We have searched through the journal and conference database for more references, and we found 4 more articles that are related to our work [20,22,27,39].
+We have added a discussion of these articles to the related work section, at the beginning of the paragraph "Requirements Engineering".
 This means we currently have 10 references from the Requirements Engineering Journal or presented at the Requirements Engineering series of conferences.
@@
 We agree with the reviewer that a real validation of the approach is a necessary next step.
-However, we also would like to point out that an industrial strength case study is a significant amount of work, and would most likely have enough new insights for a separate paper.
-To give an example, it took Estrada et al.
-[11] about 9 months to conduct the case study.
+However, we would like to point the reviewer to the case study in section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com).
+We do agree that further validation of the tool and method is necessary, preferably in an industrial setting.
+However, such an industrial strength case study is a significant amount of work (see for example Estrada et al.
+(2006 "An empirical evaluation of the i* framework in a model-based software generation environment".
+They took 9 months to conduct the case study.), and would most likely yield enough new insights for a separate paper.
 
@@
 We have added a section "Validation" (Section 7) where we describe the results of the validation.
-
-%FB: Marc/Floris to write on the below
-I think we need to include some references on empirical work that has been done on goal modelling. I asked Fabiano and he gave me the following two references, we can discuss them.
-
-Estrada, H., Rebollar, A.
-M., Pastor, O., & Mylopoulos, J.
-(2006, June).
-An empirical evaluation of the i* framework in a model-based software generation environment.
-In International Conference on Advanced Information Systems Engineering (pp.
-513-527).
-Springer, Berlin, Heidelberg.
-
-Morandini, M., Marchetto, A., & Perini, A.
-(2011, August).
-Requirements comprehension: A controlled experiment on conceptual modeling methods.
-In Empirical Requirements Engineering (EmpiRE), 2011 First International Workshop on (pp.
-53-60).
-IEEE
 
@@
 
-The comparison between argumentation and belief revision is interesting, especially from a theoretical point of view.
-In the area of knowledge representation, obtaining correspondences between beliefs revision and the dynamics of argumentation theory is a popular topic of research [12,13,14,15].
+The connection between argumentation and belief revision is interesting, especially from a theoretical point of view, and in knowledge representation and reasoning this is a popular topic of research (see references 5-7 below).
 However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on beliefs in GRL.
 
-===== COMMENT #4
+===== COMMENT #4 & #5
 How would you compare this specific approach with the old gIBIS?
-
-===== REPLY #4
-%FB: **Floris to write on gIBIS in 7.1**
-
-===== COMMENT #5
 Please discuss why you are trying to have detailed, while simple, rationale, although the recent trend in DR is otherwise?
 
-===== REPLY #5
-%FB: **Floris to write on this in 7.1**
+===== REPLY #4 & #5
+We have added some discussion in the 2nd paragraph of section 8.1-Design Rationale.
+A brief comparison with, for example gIBIS, is already in section 8.1-Design Rationale (reference [38] includes gIBIS).
+We felt that an extensive comparison between the ideas behind goal modelling, which we extend with RationalGRL, and design rationale diagramming methods like gIBIS is not within the extent of this paper, where the focus is more on the connection between "normal" goal modelling (GRL) and extended "argumentative" goal modelling (RationalGRL).
 
@@
 ===== REPLY #7
-%FB: **Floris to write on this in 5.2**
+While there is extensive work on less "binary" frameworks for argumentation (see e.g below references 2 and 3), we feel that a discussion on this would not be suitable for the current paper (as it is more of an AI subject than relevant for RE).
+For an extensive discussion on the relation between the formal argumentation proposed in our article and argumentation in reality, we refer the reviewer to reference [4] below.
+Note that we have opted not to include these references in the paper, as we are already at 50 references and want to focus more on an RE/SE audience.
+If the reviewer desires, however, we can add the references to the article.
 
@@
 Therefore, we’ve renamed “valid” to “consistent”.
-TODO(marcvanzee): Implement this renaming in the article.
+
+**TODO(marcvanzee): Implement this renaming in the article.
 
@@
 ######## REFERENCES ########
-[1] Y.
-Elrakaiby, A.
-Ferrari, P.
-Spoletini, S.
-Gnesi, and B.
-Nuseibeh.
-Using argumentation to explain ambiguity in requirements elicitation interviews.
-In Proceedings of the 25th IEEE International Requirements Engineering Conference (RE 2017), pages 51–60.
-IEEE Press, 2017.
-[2] I. Jureta, S. Faulkner, and P. Schobbens. Clear justification of mod- eling decisions for goal-oriented requirements engineering. Re- quirements Engineering, 13(2):87–115, May 2008.
-[3] Y.
-Liu, Y.
-Su, X.
-Yin, and G.
-Mussbacher.
-Combined goal and feature model reasoning with the user requirements notation and jucmnav.
-In Requirements Engineering Conference (RE), 2014 IEEE 22nd International, pages 321–322.
-IEEE, 2014.
-[4] P.
-K.
-Murukannaiah, A.
-K.
-Kalia, P.
-R.
-Telangy, and M.
-P.
-Singh.
-Resolving goal conflicts via argumentation-based analysis of com- peting hypotheses.
-In Proceedings of the 23rd International Re- quirements Engineering Conference (RE 2015), pages 156–165.
-IEEE, 2015.
-[5] M.
-van Zee, F.
-Bex, and S.
-Ghanavati.
-Rationalization of Goal Models in GRL using Formal Argumentation.
-In Proceedings of the 23rd International Requirements Engineering Conference (RE’15), pages 220–225.
-IEEE Press, August 2015
-[6] E.
-S.
-K.
-Yu.
-Towards Modeling and Reasoning Support for Early-Phase Requirements Engineering.
-In Proceedings of the 3rd IEEE International Symposium on Requirements Engineering, pages 226–235, 1997.
 
-[7] S. R. Smith. Constructing and revising formal arguments in re- quirements engineering. Requir. Eng., 2(4):199–216, Dec. 1997.
-[8] J.Lockerbie,N.A.M.Maiden,J.Engmann,D.Randall,S.Jones, and D.
-Bush.
-Exploring the impact of software requirements on system-wide goals: a method using satisfaction arguments and i* goal modelling.
-Requirements Engineering, 17(3):227–254, 2012.
-[9] J. Horkoff and E. Yu. Interactive goal model analysis for early requirements engineering. Requirements Engineering, 21(1):29– 61, 2016.
-[10] J. Hassine and D. Amyot. A questionnaire-based survey method- ology for systematically validating goal-oriented models. Require- ments Engineering, 21(2):285–308, 2016.
-
-[11] Estrada, H., Rebollar, A.
+[1] Estrada, H., Rebollar, A.
 M., Pastor, O., & Mylopoulos, J.
@@
 Springer, Berlin, Heidelberg.
+[2] A. Hunter and M. Thimm (2017) Probabilistic Reasoning with Abstract Argumentation Frameworks, Journal of Artificial Intelligence Research, 59:565--611. 
+[3] Grossi, D., & Modgil, S. (2015, July). On the Graded Acceptability of Arguments. In IJCAI (pp. 868-874).
+[4] van Eemeren, F.H., & Verheij, B. (2017). Argumentation Theory in Formal and Computational Perspective. IFCoLog Journal of Logics and Their Applications 4 (8), 2099-2181.
 
-[12] Falappa, Marcelo Alejandro, Gabriele Kern-Isberner, and Guillermo Ricardo Simari.
+[5] Falappa, Marcelo Alejandro, Gabriele Kern-Isberner, and Guillermo Ricardo Simari.
 "Belief revision and argumentation theory." Argumentation in artificial intelligence.
@@
 341-360.
-[13] Liao, Beishui, Li Jin, and Robert C. Koons. "Dynamics of argumentation systems: A division-based method." Artificial Intelligence 175.11 (2011): 1790-1814.
-[14] Dix, Jürgen, et al. "Belief change and argumentation in multi-agent scenarios (Dagstuhl seminar 13231)." Dagstuhl Reports. Vol. 3. No. 6. Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik, 2013.
-[15] Paglieri, Fabio, and Cristiano Castelfranchi.
+[6] Dix, Jürgen, et al. "Belief change and argumentation in multi-agent scenarios (Dagstuhl seminar 13231)." Dagstuhl Reports. Vol. 3. No. 6. Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik, 2013.
+[7] Paglieri, Fabio, and Cristiano Castelfranchi.
 "Revising beliefs through arguments: Bridging the gap between argumentation and belief revision in MAS." International Workshop on Argumentation in Multi-Agent Systems.

### 2018-05-04 discussion.tex
final comments FB fixed

@@
 Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE.
-The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
+The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model.
+Note that we explicitly connect to the extensive literature on goal modeling, and in this respect may seem to go against the trend in DR research to diagram decisions and processes.
+However, we note that RationalGRL also includes a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system, which connects with the more research ideas on DR \cite{razavian2016two,SchriekEtal2016}.
 

### 2018-05-04 validation.tex
final comments FB fixed

@@
 
-\subsection{Evaluation with Enterprise Architects}
+%FB: I'd leave this out
 
-\todo{marcvanzee}{marcvanzee,floris,sepideh}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects.
+%\subsection{Evaluation with Enterprise Architects}
+%
+%\todo{marcvanzee}{marcvanzee,floris,sepideh}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects.
 I'm not sure if we can re-use it.
@@
 What do you think?}
+%
+%To ensure GRL models captured the essence of the EA principles correctly, we evaluated the models with two Enterprise Architects from eGovernment.
+The architects found the GRL models very useful in that they could visualize the principles and were easier for them to understand the links between the rationale, added value, actions and the goals of the principles.
+Furthermore, since the EA principles were linked to the high-level goals of the organization, it made them easier to justify and analyze.
+Also, we captured the traceability links between the related EA principles which helped the architects to document the links and verify the connections.
 
-To ensure GRL models captured the essence of the EA principles correctly, we evaluated the models with two Enterprise Architects from eGovernment. The
-architects found the GRL models very useful in that they could visualize the principles and were easier for them to understand the links between the rationale, added value, actions and the goals of the principles.
-Furthermore, since the EA
-principles were linked to the high-level goals of the organization, it made them easier to justify and analyze.
-Also, we captured the traceability links between the related EA principles which helped the architects to document the links and
-verify the connections.
-
-The architects also informed us that assigning values for contribution links can be done at design time together with the architects.
+%The architects also informed us that assigning values for contribution links can be done at design time together with the architects.
 We aim to extend our models to more principles and provide analysis of the principles with high-level goals as well as low-level operationalized activities of the organization.

### 2018-05-06 discussion.tex
survey

--- 2018-05-03 discussion.tex
@@
 Software design and requirements engineering are very closely related \cite{nuseibeh2001weaving} and hence the insights from the DR literature are directly applicable to RE.
-The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model, and by proposing a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system.
+The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model.
+Note that we explicitly connect to the extensive literature on goal modeling, and in this respect may seem to go against the trend in DR research to diagram decisions and processes.
+However, we note that RationalGRL also includes a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system, which connects with the more research ideas on DR \cite{razavian2016two,SchriekEtal2016}.
 
-\paragraph{Requirements Engineering} There are a number of general approaches in the field of requirements engineering that explicitly take into account arguments.
+\paragraph{Requirements Engineering} 
+There are numerous examples in the RE literature that make mention of arguments, questions or conflicts in some way.
+Already in 1997, Smith~\cite{Smith:1997:CRF:2737426.2737493} proposed a framework to develop arguments, which relates a model of a proposed software system to a model of its environment, and shows that these together achieve properties described by a model of requirements.
+In a similar vein, Lockerbie \emph{et al.}~\cite{lockerbie2012exploring} describe procedures that analysts follow to use satisfaction arguments in combination with $i*$.
+Also, Horkoff and Yu~\cite{horkoff2016interactive} introduce analysis procedures for the $i*$ goal-oriented framework, allowing users to ask questions such as ``what if?'', or ``are certain goals achievable?
+how?
+or why not?''.
+Finally, Hassin and Amyot~\cite{hassine2016questionnaire} analyze the correctness of GRL models by performing statistical analysis on generated questionnaires, which allows for early detection of potential conflicts between stakeholders of the system.
+The mentioned work, however, does not consider arguments the way RationalGRL does, that is, as explicit opinions of stakeholders.
+
+A number of approaches in the field of requirements engineering explicitly take into account arguments.
 One early example comes from Haley \emph{et al.}~\cite{haley2008security}, who use formal logical arguments to show that the system behavior satisfies certain security requirements, and more informal arguments to capture and validate the assumptions underlying the system behavior.

### 2018-05-06 survey.tex
survey

$\ $\\
\newpage
\section{Modeling Task - Traffic Controller}
\label{sect:survey}

A group of three engineers (E1, E2, and E3) is tasked with designing a traffic flow simulation program. They receive a task specification, focusing on how the simulator will handle the generation of cars in the simulation. Below is an excerpt discussing the early requirements of the system. You are asked to model this discussion in as much detail as possible in the RationalGRL tool (http://www.rationalgrl.com), using all the tools RationalGRL puts at your disposal (GRL elements, critical questions, arguments).

As the RationalGRL prototype does not yet have a save functionality, please save a screenshot of your final model.\\

\begin{tabular}{|p{10mm}|p{140mm}|}
\hline
\textbf{Person} & \textbf{Transcript}\\
\hline
E1 & Let's start by adding the actor "Traffic Simulator". Do you think we need to add more actors?\\
\hline
E2 & I don't think so. Let's start with the softgoals and derive the main goals and tasks from that.\\
\hline
E3 & The task description states that the simulator should be easy to use.\\
\hline
E2 & I'm not sure that we should include that. It seems to me this is redundant since any UI should be easy to use.\\
\hline
E1 & No I disagree, for some UI's it is more important than others, and the fact that they mention it explicitly in the task description means we should probably add it anyway.\\
\hline
E3 &I agree, let's keep it in, it seems necessary.\\
\hline
E1 & Ok so we have one softgoal 'easy to use'. Anything else?\\
\hline
E2 & Here it say the simulation should be realistic, and the design should be simple. I guess those are two softgoals as well.\\
\hline
E3 & Ok sounds good. So one goal of the simulation is to Generate cars. Shall we just add that as a goal?\\
\hline
E1 & Yes let's do that.\\
\hline
E2 & I think we then have two main tasks for that goal: The simulator should be able to create new cars in the simulation, and it should be able to wrap the cars when they go out of the screen.\\
\hline
E3 & So Generate cars AND-decomposes into those two tasks?\\
\hline
E1 & I don't think so actually. I think it is a XOR-decomposition. The simulation either creates new cars or wrap them when they are being reused.\\
\hline
E2 & Hmm makes sense. But then I think we should rename Wrap cars to Keep same cars.\\
\hline
E3 & Yes that is better, then Create new cars and Keep same cars are the two alternatives for the goal Generate cars.\\
\hline
E1 & Hmm what do you think about the contribution links? Should we add any?\\
\hline
E2 & Well I think if we create new cars, the simulation will be more realistic because it becomes less static. At the same time it will probably make the design less simple.\\
\hline
E3 & Yes so create new cars contributes positively to the realistic design, and negatively to simple design?\\
\hline
E2 & Hmm now I think about it, actually I think think creating new cars doesn't have much effect on the realism of the simulation.I don't think the user will notice the difference.\\
\hline
E1 & Why not? If you create new cars it changes the flow of the traffic, so I think the user will definitely notice this.\\
\hline
E2 & Okay well, ok maybe. Yes I think you are right. Let's keep the contribution link in then.\\
\hline
\end{tabular}

### 2018-05-06 validation.tex
survey

--- 2018-05-02 validation.tex
@@
-\section{Validation}
+\section{RationalGRL user evaluation}
 \label{sect:validation}
 
-\subsection{Evaluation with Enterprise Architects}
+Since RationalGRL is targeted at practitioners and not just for academic end users, we performed a brief user evaluation with 16 users.
+In this evaluation, we asked participants to rate and comment the usability of the RationalGRL tool.
+In this section we first present the results of the evaluation in three subsections, and then provide a more high-level analysis in the last subsection.
 
-\todo{marcvanzee}{marcvanzee,floris}{This is copied from the CAiSE paper where we did the original evaluation with two Enterprise Architects.
-I'm not sure if we can re-use it.
-The main problem is that at that time we developed GRL models specifically for the organization in which the Enterprise Architects were working, and we don't have such a model now.
-What do you think?}
+\subsection{Experimental setup}
+The study consists of three parts:
+\begin{enumerate}
+\item Explanation: We explain the participants the basics of the RationalGRL development process and the RationalGRL tool.
+\item Modelling: We ask the participants to model a summarized discussion in the RationalGRL tool.
+\item Survey: We ask some questions about the usability of the RationalGRL tool, and how it relates to other goal modeling languages.
+\end{enumerate}
 
-To ensure GRL models captured the essence of the EA principles correctly, we evaluated the models with two Enterprise Architects from eGovernment. The
-architects found the GRL models very useful in that they could visualize the principles and were easier for them to understand the links between the rationale, added value, actions and the goals of the principles.
-Furthermore, since the EA
-principles were linked to the high-level goals of the organization, it made them easier to justify and analyze.
-Also, we captured the traceability links between the related EA principles which helped the architects to document the links and
-verify the connections.
+The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} At the end of the document we provide a link to the survey.\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} The instructions we provide to the users for task 2 can be found in Appendix~\ref{sect:survey}, and some example RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
 
-The architects also informed us that assigning values for contribution links can be done at design time together with the architects.
-We aim to extend our models to more principles and provide analysis of the principles with high-level goals as well as low-level operationalized activities of the organization.
+Two thirds of the participants had either a PhD or Master degree, 20\% a Bachelor degree, and one respondent responded with ``Other''.
+The participants turned out to have a relatively high level of experience with IT, with an average rating of 6 years, ranging from 0 to 25 years.
+They also judged themselves to be quite experienced with early-phase requirements engineering (average of rating 3.2 out of 5), but experience with goal modeling was less (rating 1.9 out of 5).
+Goal modeling languages the participants had experience with were GRL (2), i* (5), Tropos (1), and KAOS (1).
 
-\subsection{Empirical evaluation}
+\subsection{Usability of the RationalGRL tool}
 
-\todo{marcvanzee}{marcvanzee,florisbex}{Write this}
+There are four questions in the survey which ask users to rate various aspects of the RationalGRL tool on a scale of 1 to 5:
+\begin{itemize}
+\item[Q1] Was it easy or difficult to get started with the RationalGRL tool?
+\item[Q2] Was the details pane (containing details of an element, critical questions, etc) easy or difficult to use?
+\item[Q3] Was it easy or difficult to understand the way in which the status of arguments and other elements is determined?
+\item[Q4] Did you find it easy or difficult to model the example discussion using RationalGRL?
+\end{itemize}
+
+\begin{table*}[t]
+\centering
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
+& very difficult & somewhat difficult & neither difficult nor easy & somewhat easy & very easy & avg. score (1-5) & std. dev.\\
+\hline
+Q1 & 2 (12.5\%) & 1 (6.3\%) & 6 (37.5\%) & 6 (37.5\% & 1 (6.3\%) & 3.2 & 1.1\\
+Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25\%) & 8 (50\%) & 0 (0\%) & 3.1 & 1.1\\
+Q3 & 0 (0\%) & 3 (20\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20\%) & 3.3 & 1.4\\
+Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0\%) & 2.6 & 1.0
+\end{tabularx}
+\caption{User scores for the usability of the RationalGRL tool}
+\label{table:survey:table1}
+\end{table*}
+
+\begin{table*}[t]
+\centering
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
+& very useful & somewhat useful & neither useful nor useless & somewhat useless & very useless & avg. score (1-5) & std. dev.\\
+\hline
+Q5 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.68\\
+Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
+Q7 & 0 (0\%) & 0 (0\%) & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7
+\end{tabularx}
+\caption{User scores for the comparison between RationalGRL and other goal modeling languages}
+\label{table:survey:table2}
+\end{table*}
+
+Table~\ref{table:survey:table1} shows the respondents' answers, the average score, and the standard deviation for each answer.
+On average, the respondent found the tool relatively easy to use, but there was also a number of users that had some trouble getting started (in 3 out of the 4 questions, 12.\%5 rated ``very difficult'').
+We think this is not unexpected, as the cognitive burden for our framework is higher than for other goal modeling languages such as GRL and i*, simply because there are more elements in the language.
+Still, the majority of the users did not found the tool difficult to use.
+
+We also asked the users three open questions related to usability.
+When asked about the strengths and weaknesses of RationalGRL, users where generally positive about RationalGRL's clear UI and the fact that arguments can be combined with goal models.
+For instance, one user stated that RationalGRL successfully ``tries to capture the rational behind the modeling process' and ``has a simple) way to compute the status of the arguments''.
+Users did found it challenging to decide when to use arguments, and to determine when an arguments attacks another argument.
+For instance, user mentioned ``it wasn't very easy because i didn't always know how to attack the elements'', and ``i dont think its necessary to include all the arguments because some of them are a bit trivial''.
+Other users were worried about the cognitive overload of the tool.
+One user mentioned ``I still believe in the value of arguments, but there should be less confusing ways to capture them''.
+
+When asked about improvements in the RationalGRL tool, the users mentioned additional UI functionality such as the possibility to save models, having an ``undo'' function, and flipping the arrow after adding them.
+Users also suggested various language-related improvements.
+Some users mentioned they missed the possibility to attack links, while other mentioned that not all GRL elements are supported (for instance, it is currently not possible to add actors).
+
+When asked whether the users believe a more mature version of RationalGRL can be used to capture early-phase requirements in an actual software development project, most users responded positively.
+This was somewhat surprising, since our tool is still an early prototype.
+Users' responses include ``Yes, having the possibility to add arguments seems quite useful'', ``I think it can, but maybe try to find a way to combine it with more regular/mainstream requirements gathering such as user stories and customer journeys'', ``yes, because it creates a clear scope for the project, what the goals are'', and ``Generally I think its useful to explicitly document arguments of a discussion''.
+The worry that there may be a cognitive overhead did come back here as well: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', and ``the manual input is too complex and takes too much time.
+An automated process of parsing the conversation log would be much more helpful''.
+
+\subsection{Comparison with other goal modeling languages}
+
+We asked the users three questions about the extension of RationalGRL on a scale of 1 to 5:
+\begin{itemize}
+\item[Q5] Do you think the arguments and counterarguments of RationalGRL are a useful extension to standard goal modelling?
+\item[Q6] Do you think the critical questions and answers in the details pane are a useful extension to standard goal modelling?
+\item[Q7] Do you think the automatic determination of the status of arguments and elements is a useful extension to standard goal modelling?
+\end{itemize}
+
+Table~\ref{table:survey:table2} shows the respondents' answers, the average score, and the standard deviation for each answer.
+On average, the users were positive about the usefulness of the RationalGRL extension.
+The rating for the usefulness of determining the status of arguments was particularly positive (rating 4.2 out of 5), which provides some indication to the fact that using argumentation semantics to be able to determine precisely which elements of a goal model are accepted could be useful in practice.
+
+We concluded with two open questions about the comparison between RationalGRL and other goal modeling languages.
+When asked about the advantages of RationalGRL over standard goal modeling languages, many users agreed that making arguments explicit may force end users to have a more structured discussion: ``	Clear communication about argumentation and forcing people to think in those clear terms.'', ``...you can add arguments and that you can answer questions that help you to develop arguments'', ``It's useful that discussion and explanation are close to the diagrams'', ``a way to see how decisions are being shaped.''.
+When asked about the weaknesses, users mentioned the complexity as the most important weakness: ``The apparent increase in complexity might lead to negative perceptions'', ``adding yet another layer of complexity scares me''.
+One user mentioned that this may be a problem with goal modeling in general: ``Goal models are already complex (...) I have worked for years on the effect of context on goal models, and my conclusion is that this was very interesting academic work but with close-to-zero practical implications, unfortunately.''.
+
+\subsection{Analysis}
+Although our empirical evaluation is relatively small-scale, and we cannot draw hard conclusions from it, we do believe that we can extract some interesting observations.
+We list the three most important ones:
+\begin{enumerate}
+\item \emph{High cognitive overload.} A worry that came back often in our empirical evaluation is that, in its current form, RationalGRL has a relatively high cognitive overload.
+Goal modeling is by itself already a relatively high-effort activity, and the fact that we add more elements to the language does not improve this.
+This may mean that in future work we should focus more on the RationalGRL development process.
+It seems users do like to be guided during modeling, and making arguments explicit in this process seems useful, but it could be integrated into the process of goal modeling in a more natural way.
+Critical questions seem to be a natural fit for this.
+They play a key role in the RationalGRL development process, but in its current form they are relatively difficult to access (in the details pane).
+\item \emph{Useful argumentation semantics.} Participants were enthusiastic about adding arguments to a goal modeling language with a clear formal underpinning, and they believed the argumentation semantics we use in the tool is very intuitive.
+This is a positive signal for our formal approach.
+\item \emph{Lightweight tool.} Respondents were positive about the fact that it was very easy to get started with the tool.
+Thus it seems that having a simple web-interface for goal modeling languages is something that could be done more in the future.
+Most comments on the tool were the type of comments that are expected from a prototype implementation.
+\end{enumerate}

### 2018-05-06 Response to reviewers.txt
rebuttal and small changes

@@
 ===== REPLY #1
-Currently 6 of the 44 references in the article come from either REJ or the RE conference [11,24,26,32,43,47].
+In the first submission, 6 of the 44 references in the article came from either REJ or the RE conference [11,24,26,32,43,47].
 We have searched through the journal and conference database for more references, and we found 4 more articles that are related to our work [20,22,27,39].
@@
 We thank the reviewer for the insightful comments.
-Especially the comment about the missing validation was useful since it motivated us to perform an small-scale in-house user study, which allowed us to strengthen our claims for the usefulness and usability of our framework.
+Especially the comment about the missing validation was useful since it motivated us to perform an small-scale in-house user study, which allowed us to make some useful observation regarding the usefulness and usability of our framework.
 We’ve replied to each separate comment below.
@@
 We agree with the reviewer that a real validation of the approach is a necessary next step.
-However, we would like to point the reviewer to the case study in section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com).
+However, we would like to point the reviewer to the case study in Section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com).
 We do agree that further validation of the tool and method is necessary, preferably in an industrial setting.
@@
 Still, we agree that having no validation at all is limiting, so we have performed a small-scale user study in order to provide an initial validation of our framework.
-This empirical study was performed amongst ==> TODO(marcvanzee): ADD NUMBER HERE  <== participants, each with a computer science background.
+This empirical study was performed amongst 16 participants, each with a computer science background.
 
-We have added a section "Validation" (Section 7) where we describe the results of the validation.
+We have added a section "RationalGRL user validation" (Section 7) where we describe our empirical study.
 
@@
 ===== REPLY #1
-We have added a section "Validation" (Section 7) where we describe the results of a validation. Please see REPLY #3 to REVIEWER #1 for more details.
+We have added a section "RationalGRL user validation" (Section 7) where we describe the results of a validation. Please see REPLY #3 to REVIEWER #1 for more details.
 
@@
 The connection between argumentation and belief revision is interesting, especially from a theoretical point of view, and in knowledge representation and reasoning this is a popular topic of research (see references 5-7 below).
-However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on beliefs in GRL.
+However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on a comparison with beliefs in GRL.
 
@@
 ===== REPLY #8
-We agree that the term “valid” is confusing in this respect, since we do not want to express that the goal model makes sense in the real world.
-Instead, we’d like to express that certain criteria are fulfilled.
-Therefore, we’ve renamed “valid” to “consistent”.
-
-**TODO(marcvanzee): Implement this renaming in the article.
+We agree that the term “valid” is an overloaded term that has different meanings in different areas of research.
+In our case, we indeed want to express that certain criteria are fulfilled.
+In order to avoid confusion, we have added a footnote to Definition 7 (Valid GRL Model), clarifying what exactly we mean with a "valid model".
 
@@
 ===== REPLY #10
-We agree that the terminology is wrong.
+We agree that the terminology is inappropriate.
 We have renamed the methodology to "The RationalGRL development process", since it defines the steps on how to create a RationalGRL model.
-We have renamed this throughout the entire article.
+We have renamed this throughout the article.
 

### 2018-05-06 formalframework.tex
Addressed some more comments by reviewers
rebuttal and small changes

@@
 \label{def:valid-grl-model}
-A GRL model $GRL=(IE, Act, R_{ActIE}, Link)$ (Def.~\ref{def:grl-model}) is a \emph{valid GRL model} iff the following conditions are satisfied:
+A GRL model $GRL=(IE, Act, R_{ActIE}, Link)$ (Def.~\ref{def:grl-model}) is a \emph{valid GRL model}\footnote{Note that \emph{validity} is a bit of an overloaded term and has different meanings in different areas of research.
+In our setting, a valid model simply means that certain criteria are satisfied.
+It does not imply any real world ``validation'', in the sense that the model makes sense in the ``real world''.} iff the following conditions are satisfied:
 \begin{enumerate}
@@
 \includegraphics{img/methodology.pdf}
-\caption{The RationalGRL Methodology}
+\caption{The RationalGRL Development Process}
 \label{fig:rationalgrl-methodology}

### 2018-05-07 discussion.tex
small change
rebuttal and small changes
updates

@@
 
-\paragraph{Specific Domains}
+\paragraph{Specific domains}
 The argumentation schemes and critical questions presented in this paper are focused only on the core goal modeling process, that is, the discussion about the goals and functionality of an information system.
@@
 
-\paragraph{Tool support}
+\paragraph{Process and tool support}
 The current tool is a prototype that implements the RationalGRL framework in a fairly simple way.
@@
 Furthermore, the tool could be expanded to include the new features described above, such as new argument schemes and reasoning with preferences.
-In addition to such extensions, it is interesting to think about new possibilities.
+
+Additionally, from the user evaluation of the tool it followed that in order to implement the tool in realistic settings, other ways may have to be found to incorporate arguments in the goal modeling process.
 One idea is to use the tool for collaborative on-line goal modeling, similar to Github: since the renaming (i.e.
@@
 As a formal underpinning of this asynchronous communication between users, it would make sense to capture requirements engineering and software design processes as dialogs between parties~\cite{finkelstein1989multiparty,BlackEtal2013}, which are a natural fit with the question-answer format employed in the RationalGRL framework.
+Finally, it would make sense to further develop the RationalGRL Development Process.
+One idea here is to have stakeholders play a card game (similar to the well-known planning poker) with cards that contain critical, reflective questions (cf.
+\cite{SchriekEtal2016})
 
@@
 We have performed a case study and found more than 200 incidences of arguments and questions.
+\rood{Zeg ook dat we user study gedaan hebben, en noem het idee om in de toekomst een "industrial-strength case-study" te doen}.
 However, what is still lacking is further empirical investigation of whether, and how, the tools and techniques provided by the RationalGRL framework really improve early phase requirements engineering.
@@
 To this end, we propose a development process, which consists of developing goal models and posing arguments based on schemes in an integrated way.
+
 Finally, we have implemented the RationalGRL tool, a web-based prototype\footnote{\url{http://www.rationalgrl.com}}, for modeling argument schemes and critical questions and for reasoning about goal models with respect to them (\textbf{requirement 5}).
+We performed a small empirical evaluation with 16 users.
+The results indicate that while the cognitive overhead of RationalGRL (or goal modeling in general) is high, combining goal modeling with argumentation is perceived as useful.
+Furthermore, using a formal semantics to compute which arguments and ultimately goal models are acceptable given the stakeholders' opinions valid argument was perceived as very useful.

### 2018-05-07 introduction.tex
rebuttal and small changes
updates

@@
 Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyze and guide stakeholders' discussions about goal models.
-This list of argument schemes is based on an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system.
+This list of argument schemes is based on an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system (requirement 1).
 
-In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models.
+In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models (requirement 2).
 In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework to which we can apply argumentation semantics~\cite{Dung1995}.
-We, then, formally capture the links between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models.
+We, then, formally capture the links between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models (requirement 3).
 
-In order to support practitioners in using the RationalGRL framework, we propose the RationalGRL development process, which explains steps for  developing goal models and posing arguments based on schemes in an integrated way.
-We implement a web-based tool for the RationalGRL framework as well, which works on any modern browser and is developed in Javascript.\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
+In order to support practitioners in using the RationalGRL framework, we propose the RationalGRL development process, which explains steps for developing goal models and posing arguments based on schemes in an integrated way (requirement 4).
+We implement and briefly evaluate a web-based tool for the RationalGRL framework, which works on any modern browser and is developed in Javascript (requirement 5).\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
 
@@
 Section~\ref{sect:methodology+tool} discusses the RationalGRL development process and explains various features of the tool we developed.
-Finally, Section~\ref{sect:discussion} covers the related work, the future work, and the conclusion.
+Section~\ref{sect:validation} contains a brief validation of RationalGRL.
+We conducted a small user study (16 participants), where we asked the participants to model a small example in the RationalGRL, and to answer some questions on the usability and usefulness RationalGRL.
+Finally, Section~\ref{sect:discussion} covers related work, future work, and the conclusion.

### 2018-05-07 validation.tex
updates

@@
 
-Since RationalGRL is targeted at practitioners and not just for academic end users, we performed a brief user evaluation with 16 users.
-In this evaluation, we asked participants to rate and comment the usability of the RationalGRL tool.
-In this section we first present the results of the evaluation in three subsections, and then provide a more high-level analysis in the last subsection.
+In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a small user evaluation with 16 users.
+This evaluation concerned  mainly the use of the RationalGRL tool, but more general questions regarding the usefulness of arguments in goal modelling were also asked.
+We first present the details and results of the evaluation, and then provide a more high-level analysis in Section~\ref{sec:eval-an}.
 
@@
 
-The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} At the end of the document we provide a link to the survey.\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} The instructions we provide to the users for task 2 can be found in Appendix~\ref{sect:survey}, and some example RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
+The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} At the end of the document we provide a link to the survey.\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} The instructions we provide to the users for task 2 can be found in Appendix~\ref{sect:survey}.
 
@@
 The participants turned out to have a relatively high level of experience with IT, with an average rating of 6 years, ranging from 0 to 25 years.
-They also judged themselves to be quite experienced with early-phase requirements engineering (average of rating 3.2 out of 5), but experience with goal modeling was less (rating 1.9 out of 5).
+They also judged themselves to be quite experienced with early-phase requirements engineering (average of rating 3.2 out of 5), but experience with goal modeling was less (rating 1.9 out of 5, although two respondents use goal modeling languages almost weekly).
 Goal modeling languages the participants had experience with were GRL (2), i* (5), Tropos (1), and KAOS (1).
+
+In task 2, users clearly made use of the extra features offered by the RationalGRL tool: they used the details pane more often than not (rating 3.2 out of 5) and on average produced 3 arguments from the short transcript.
+Some example RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
 
@@
 When asked about the strengths and weaknesses of RationalGRL, users where generally positive about RationalGRL's clear UI and the fact that arguments can be combined with goal models.
-For instance, one user stated that RationalGRL successfully ``tries to capture the rational behind the modeling process' and ``has a simple) way to compute the status of the arguments''.
+For instance, one user stated that RationalGRL successfully ``tries to capture the rationale behind the modeling process' and ``has a simple way to compute the status of the arguments''.
 Users did found it challenging to decide when to use arguments, and to determine when an arguments attacks another argument.
-For instance, user mentioned ``it wasn't very easy because i didn't always know how to attack the elements'', and ``i dont think its necessary to include all the arguments because some of them are a bit trivial''.
+For instance, user mentioned ``it wasn't very easy because I didn't always know how to attack the elements'', and ``I don't think its necessary to include all the arguments because some of them are a bit trivial''.
 Other users were worried about the cognitive overload of the tool.
@@
 Users' responses include ``Yes, having the possibility to add arguments seems quite useful'', ``I think it can, but maybe try to find a way to combine it with more regular/mainstream requirements gathering such as user stories and customer journeys'', ``yes, because it creates a clear scope for the project, what the goals are'', and ``Generally I think its useful to explicitly document arguments of a discussion''.
-The worry that there may be a cognitive overhead did come back here as well: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', and ``the manual input is too complex and takes too much time.
+The worry that the modeling process may involve too much cognitive overhead was mentioned again though: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', and ``the manual input is too complex and takes too much time.
 An automated process of parsing the conversation log would be much more helpful''.
@@
 
-\subsection{Analysis}
+\subsection{Analysis}\label{sec:eval-an}
 Although our empirical evaluation is relatively small-scale, and we cannot draw hard conclusions from it, we do believe that we can extract some interesting observations.
@@
 \begin{enumerate}
-\item \emph{High cognitive overload.} A worry that came back often in our empirical evaluation is that, in its current form, RationalGRL has a relatively high cognitive overload.
-Goal modeling is by itself already a relatively high-effort activity, and the fact that we add more elements to the language does not improve this.
-This may mean that in future work we should focus more on the RationalGRL development process.
-It seems users do like to be guided during modeling, and making arguments explicit in this process seems useful, but it could be integrated into the process of goal modeling in a more natural way.
-Critical questions seem to be a natural fit for this.
-They play a key role in the RationalGRL development process, but in its current form they are relatively difficult to access (in the details pane).
+\item \emph{High cognitive overhead.} A worry that came back often in our empirical evaluation is that, in its current form, RationalGRL has a relatively high cognitive overhead.
+Goal modeling is by itself already a cognitively high-effort activity, and the fact that we add more elements to the language does not improve this.
+This may mean that in future work we should focus more on the RationalGRL development process\footnote{Note that in order to keep the study simple for the users, we did not explicitly ask the respondents to follow the development process from Section~\ref{sect:methodology}}.
+Users like to be guided during modeling, and making arguments explicit in the modelling process was indicated as useful, but argumentation should be integrated into the \emph{process} of goal modeling more than in the goal models themselves.
+Critical questions seem to be a natural fit for this, as they play a key role in the RationalGRL development process, but in its current form they are only mentioned in the details pane of the tool.
 \item \emph{Useful argumentation semantics.} Participants were enthusiastic about adding arguments to a goal modeling language with a clear formal underpinning, and they believed the argumentation semantics we use in the tool is very intuitive.

### 2018-05-07 abstract.tex
Modified abstract

@@
 The process of constructing a goal model usually involves discussions between a requirements engineer and a group of stakeholders.
-While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
+While it is possible to capture part of this discussion process in a goal model, by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
 For instance, the discussion on whether to accept or reject a certain goal and the ultimate rationale for acceptance or rejection cannot be captured in current goal modeling languages.
-Based on a case study in which stakeholders discuss requirements for a Traffic Simulator, we apply argumentation techniques from artificial intelligence to a goal modeling approach.
+In this paper, we apply argumentation techniques from artificial intelligence to a goal modeling approach by using a case study in which stakeholders discuss requirements for a Traffic Simulator.
 We combine a simplified version of a traditional goal modeling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).

### 2018-05-07 validation.tex
Modified a bit the wordings

@@
 In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a small user evaluation with 16 users.
-This evaluation concerned  mainly the use of the RationalGRL tool, but more general questions regarding the usefulness of arguments in goal modelling were also asked.
+This evaluation concerned mainly with the use of the RationalGRL tool, but more general questions regarding the usefulness of arguments in goal modeling were also asked.
 We first present the details and results of the evaluation, and then provide a more high-level analysis in Section~\ref{sec:eval-an}.
@@
 \begin{enumerate}
-\item Explanation: We explain the participants the basics of the RationalGRL development process and the RationalGRL tool.
-\item Modelling: We ask the participants to model a summarized discussion in the RationalGRL tool.
+\item Explanation: We explain to the participants the basics of the RationalGRL development process and the RationalGRL tool.
+\item Modeling: We ask the participants to model a summarized discussion in the RationalGRL tool.
 \item Survey: We ask some questions about the usability of the RationalGRL tool, and how it relates to other goal modeling languages.
@@
 The participants turned out to have a relatively high level of experience with IT, with an average rating of 6 years, ranging from 0 to 25 years.
-They also judged themselves to be quite experienced with early-phase requirements engineering (average of rating 3.2 out of 5), but experience with goal modeling was less (rating 1.9 out of 5, although two respondents use goal modeling languages almost weekly).
-Goal modeling languages the participants had experience with were GRL (2), i* (5), Tropos (1), and KAOS (1).
+They also judged themselves to be quite experienced with early-phase requirements engineering (average of rating 3.2 out of 5), but experience with goal modeling languages was less (rating 1.9 out of 5, although two respondents use goal modeling languages almost weekly).
+The participants of the study had experience with the following goal modeling languages: No experience (7 users), GRL (2 users), i* (5 users), Tropos (1 user), and KAOS (1 user).
 
 In task 2, users clearly made use of the extra features offered by the RationalGRL tool: they used the details pane more often than not (rating 3.2 out of 5) and on average produced 3 arguments from the short transcript.
-Some example RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
+Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
 
@@
 Table~\ref{table:survey:table1} shows the respondents' answers, the average score, and the standard deviation for each answer.
-On average, the respondent found the tool relatively easy to use, but there was also a number of users that had some trouble getting started (in 3 out of the 4 questions, 12.\%5 rated ``very difficult'').
+On average, the respondent found the tool relatively easy to use, but there were also a number of users that had some trouble getting started (in 3 out of the 4 questions, 12.\%5 rated ``very difficult'').
 We think this is not unexpected, as the cognitive burden for our framework is higher than for other goal modeling languages such as GRL and i*, simply because there are more elements in the language.
@@
 For instance, one user stated that RationalGRL successfully ``tries to capture the rationale behind the modeling process' and ``has a simple way to compute the status of the arguments''.
-Users did found it challenging to decide when to use arguments, and to determine when an arguments attacks another argument.
-For instance, user mentioned ``it wasn't very easy because I didn't always know how to attack the elements'', and ``I don't think its necessary to include all the arguments because some of them are a bit trivial''.
+On the other hand, some of the users found it challenging to decide when to use arguments, and to determine when an arguments attacks another argument.
+For instance, a user mentioned ``it wasn't very easy because I didn't always know how to attack the elements'', and ``I don't think its necessary to include all the arguments because some of them are a bit trivial''.
 Other users were worried about the cognitive overload of the tool.
@@
 
-When asked about improvements in the RationalGRL tool, the users mentioned additional UI functionality such as the possibility to save models, having an ``undo'' function, and flipping the arrow after adding them.
+When asked about improvements in the RationalGRL tool, the users mentioned additional UI functionalities such as the possibility to save models, having an ``undo'' function, and flipping the arrow after adding them.
 Users also suggested various language-related improvements.
-Some users mentioned they missed the possibility to attack links, while other mentioned that not all GRL elements are supported (for instance, it is currently not possible to add actors).
+Some users mentioned they missed the possibility to attack links, while others mentioned that not all GRL elements are supported (for instance, it is currently not possible to add actors).
 
@@
 Users' responses include ``Yes, having the possibility to add arguments seems quite useful'', ``I think it can, but maybe try to find a way to combine it with more regular/mainstream requirements gathering such as user stories and customer journeys'', ``yes, because it creates a clear scope for the project, what the goals are'', and ``Generally I think its useful to explicitly document arguments of a discussion''.
-The worry that the modeling process may involve too much cognitive overhead was mentioned again though: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', and ``the manual input is too complex and takes too much time.
+The concern that the modeling process may involve too much cognitive overhead was mentioned again though: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', and ``the manual input is too complex and takes too much time.
 An automated process of parsing the conversation log would be much more helpful''.
@@
 \begin{itemize}
-\item[Q5] Do you think the arguments and counterarguments of RationalGRL are a useful extension to standard goal modelling?
-\item[Q6] Do you think the critical questions and answers in the details pane are a useful extension to standard goal modelling?
-\item[Q7] Do you think the automatic determination of the status of arguments and elements is a useful extension to standard goal modelling?
+\item[Q5] Do you think the arguments and counterarguments of RationalGRL are a useful extension to standard goal modeling?
+\item[Q6] Do you think the critical questions and answers in the details pane are a useful extension to standard goal modeling?
+\item[Q7] Do you think the automatic determination of the status of arguments and elements is a useful extension to standard goal modeling?
 \end{itemize}
@@
 \begin{enumerate}
-\item \emph{High cognitive overhead.} A worry that came back often in our empirical evaluation is that, in its current form, RationalGRL has a relatively high cognitive overhead.
+\item \emph{High cognitive overhead.} A concern that was raised often in our empirical evaluation is that, in its current form, RationalGRL has a relatively high cognitive overhead.
 Goal modeling is by itself already a cognitively high-effort activity, and the fact that we add more elements to the language does not improve this.
 This may mean that in future work we should focus more on the RationalGRL development process\footnote{Note that in order to keep the study simple for the users, we did not explicitly ask the respondents to follow the development process from Section~\ref{sect:methodology}}.
-Users like to be guided during modeling, and making arguments explicit in the modelling process was indicated as useful, but argumentation should be integrated into the \emph{process} of goal modeling more than in the goal models themselves.
+Users like to be guided during modeling phase, and making arguments explicit in the modeling process was indicated as useful, but argumentation should be integrated into the \emph{process} of goal modeling more than in the goal models themselves.
 Critical questions seem to be a natural fit for this, as they play a key role in the RationalGRL development process, but in its current form they are only mentioned in the details pane of the tool.
@@
 \item \emph{Lightweight tool.} Respondents were positive about the fact that it was very easy to get started with the tool.
-Thus it seems that having a simple web-interface for goal modeling languages is something that could be done more in the future.
+Thus, it seems that having a simple web-interface for goal modeling languages is something that could be done more in the future.
 Most comments on the tool were the type of comments that are expected from a prototype implementation.

### 2018-05-07 discussion.tex
Fixed a bit the conclusion

@@
 
-The process of constructing a goal model involves discussions between a requirements engineer and a group of stakeholders.
+The process of constructing a goal model involves discussions between requirements engineers and a group of stakeholders.
 While it is possible to capture part of this discussion process in a goal model, for instance by specifying alternative solutions for a goal, not all of the arguments can be found back in the resulting model.
@@
 Our list of argument schemes and critical questions in Table~\ref{table:argument-schemes} was constructed by performing an extensive case study in which we analyzed 153 pages of transcripts of discussions among designers of a traffic simulator information system.
-In this way, we ensure that he argumentation techniques capture the actual discussions of the stakeholders or designers in the early requirements engineering phase (\textbf{requirement 1}).
+In this way, we ensure that the argumentation techniques capture the actual discussions of the stakeholders or designers in the early requirements engineering phase (\textbf{requirement 1}).
 
-The metamodel of the RationalGRL framework in Figure~\ref{fig:metamodel} clearly specifies the formal traceability links elements of the GRL goal model and the underlying arguments.
+The metamodel of the RationalGRL framework in Figure~\ref{fig:metamodel} clearly specifies the formal traceability links elements of GRL goal model and the underlying arguments.
 (\textbf{requirement 2}).

### 2018-08-13 discussion.tex
laatste commit

@@
 An insufficiency occurs when an analyst needs more information from the client to accept a client's statement.
-
-Hassin and Amyot~\cite{hassine2016questionnaire} analyze the correctness of GRL models by performing statistical analysis on generated questionnaires.
-This allows early detection of potential conflicts between stakeholders of the system.
-Instead of questionnaires, we use arguments and make them explicitly traceable to elements of the goal model.
-Horkoff and Yu~\cite{horkoff2016interactive} introduce analysis procedures for the $i*$ goal-oriented framework, allowing users to ask questions such as ``what if?'', or ``are certain goals achievable?
-how?
-or why not?''.
-They formally defined their procedures and describe a framework implementation.
-Also in $i*$, Lockerbie \emph{et al.}~\cite{lockerbie2012exploring} describe procedures that analysts follow to use satisfaction arguments.
-They demonstrate their results in two case studies, providing evidence towards the claim that the integration of $i*$ and satisfaction arguments is useful and usable.
-About 30 years ago, Smith~\cite{Smith:1997:CRF:2737426.2737493} already proposed a framework to develop arguments for the confidences that a software-based system will be successful in a given environment.
-He used a simple temporal logic with compositional reasoning, allowing arguments to be constructed, and revised.
 
@@
 We have performed a case study and found more than 200 incidences of arguments and questions.
-\rood{Zeg ook dat we user study gedaan hebben, en noem het idee om in de toekomst een "industrial-strength case-study" te doen}.
-However, what is still lacking is further empirical investigation of whether, and how, the tools and techniques provided by the RationalGRL framework really improve early phase requirements engineering.
+We also did a user with 16 participants, testing the usability and usefulness of RationalGRL compared to other goal modeling languages.
+However, what is still lacking is an industrial-strength case study testing extensively whether, and how, the tools and techniques provided by the RationalGRL framework really improve early phase requirements engineering.
 The complexity of the domain makes focused experiments difficult, but not impossible.

### 2018-08-13 tool.tex
laatste commit

@@
 The final requirement of our framework is that it has tool support (\textbf{requirement 5}).
-This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.
+This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in the next section, exploring the usability and usefulness of RationalGRL compared to other goal modeling languages, and iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.
 In this section we briefly highlight some features of the tool, and we explain some of its current limitations.

### 2018-08-13 background.tex
fix conflicts

@@
 The model in Figure~\ref{fig:example-small} does not contain information about discussions that led to the model, such as clarification steps for the naming, or alternatives that have been considered for the relationships.
-The idea behind the original GRL specification is that beliefs can be used to capture such design rationales that make later justification and review of a model easier.
-However, beliefs cannot be connected to links - this makes answering the third and fourth question above impossible.
-Furthermore, beliefs are after-the-fact design rationales and do not capture the types of questions given above.
-Finally, the jUCMNav tool does not consider beliefs in the evaluation algorithms, and thus there is no formal connection between goal models and their underlying beliefs and opinions.
-For a more detailed comparison of our framework with GRL beliefs, see Section~\ref{sect:discussions:relatedwork}.
+The idea behind the original GRL specification is that so-called \emph{belief} elements can be used to capture such design rationales that make later justification and review of a model easier.
+In the original GRL specification, however, the exact semantics of these belief elements and the ways in which they should be used remain unclear.
+Furthermore, belief elements cannot be attached to links between elements, which makes answering the fourth and fifth question above impossible.
+We further discuss belief elements in light of our proposed argumentation-based framework in Section~\ref{sect:discussions:relatedwork}.
 

### 2018-08-13 discussion.tex
fix conflicts

@@
 
-\paragraph{Goal Modeling} Argumentation has been included -- both explicitly and implicitly -- in existing goal modeling languages.
-For example, the belief element in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} is meant to capture the rationales behind the inclusion of goals and tasks in the model.
-Furthermore, relations between elements in a goal model also provide justifications: high-level goals are reasons for lower-level goals, tasks and resources.
+\paragraph{Goal Modeling} Argumentation has already been included implicitly in existing goal modeling languages.
+At its core, goal modeling is a process of argumentation: high-level goals are reasons for lower-level goals, tasks and resources.
 Hence, refinement and decomposition techniques used in requirements engineering \cite{van2001goal} can be seen as explicit argumentation steps in the goal modeling process.
@@
 Providing an alternative subgoal or task in a goal model is then an explicit argumentative move in the discussion.
-In this sense, a goal model provides a justification for itself, particularly if we include belief elements for extra design rationalization.
+In this sense, a goal model provides a justification for itself.
 This idea is also prevalent in our RationalGRL framework: many arguments are in fact GRL elements, and many critical questions can be answered by introducing new GRL elements.
@@
 The goal model is the product of a process of argumentation, and does not include, for example, goals and tasks that were at some point considered but discarded.
-Furthermore, for goal models, it is only possible to determine the satisfiability of goals given the possible tasks and resources; what cannot be determined is the acceptability of goals, that is, whether they are acceptable given potentially contradictory opinions of stakeholders.
+Furthermore, for goal models, it is only possible to determine the satisfiability of goals given the possible tasks and resources; what cannot be determined is the acceptability of goals, that is, whether they are \emph{acceptable} (cf.
+Definition 14} given potentially contradictory opinions of stakeholders.
 
-There are several contributions to the literature that relate argumentation-based techniques with goal modeling.
+More explicit rationales for goal models are included in the original GRL specification \cite{} in the form of \emph{belief} elements, which can be attached to other intentional elements (goals, tasks) using the different links (positive contribution, negative contribution) of GRL.
+The content of a belief is a textual description, which can explain, for instance, the rationale for including a goal in a model.
+In this regard, they seem like arguments or reasons for the different elements in a goal model.
+However, in the literature the precise usage and semantics of belief elements in GRL remain unclear: according to the metamodel \cite{•} they can be connected to other intentional elements (including beliefs) through any type of link, but in the jUCMNav tool there is a separate \emph{belief link} for connecting beliefs to goals and tasks only, and the beliefs are not used in any of the algorithms to compute the satisfiability of goals and tasks.
+We are also not aware of many examples of the usage of belief elements in the literature.
+So what it seems like is that GRL belief elements can be understood as metadata that can be attached to an element
+
+Ambiguities notwithstanding, there are further advantages of using RationalGRL over GRL with belief elements.
+For example, in none of the GRL specifications is it possible to attach beliefs to links between other elements so it is not possible, for example, to provide a rationale for why some tasks positively contributes to a goal.
+Furthermore, the RationalGRL arguments follow clear schemes with associated critical questions explicitly tailored towards reasoning about goals and actions or tasks, while beliefs in GRL offer no further guidance as to what kinds of rationales can be given for elements in a goal model.
+Furthermore, beliefs in GRL seem not to influence the goal model in any way.
+Arguments in RationalGRL, in contrast, play an important role in forming the resulting goal model: if an element is defeated by an argument, it is not part of the resulting goal model.
+
+Considering beliefs, an interesting avenue for further research would be to look at to connections between argumentation and belief revision \cite{alchourron1985logic}, and how these can play a role in RationalGRL and requirements engineering in general.
+While there is formal work in Artificial Intelligence on the connections between argumentation and belief revision (see e.g.
+\cite{falappa2009belief}, there is very little work on applying the combination to requirements engineering.
+The only work we are aware of that deals with his subject is by Gervasi and Zowghi \cite{gervasi2005reasoning}, who use a logical framework for belief revision to detect inconsistencies in software specifications and then revise those specifications so that the inconsistencies are removed.
+They do not explicitly connect their framework to goal models of goal modelling, however, and do not connect their framework to specific argumentation using schemes in the way RationalGRL does.
+
+\paragraph{Goal Modeling and Argumentation}
+
+There are several contributions to the literature that explicitly relate argumentation-based techniques with goal modeling.
 The first line of work is by Bagheri and Ensan \cite{bagheri2011consolidating} and Mirbel and Villata \cite{MirbelVillata12} use abstract argumentation frameworks (see Definition~\ref{def:argumentation-framework}) to capture individual goals and the relations between them as modeled in a goal model, so that consistent (i.e., acceptable) subsets of goals can be computed using the relevant argumentation semantics \cite{Dung1995}.
@@
 RationalGRL overcomes this problem by integrating argumentation and goal modeling in a single modeling language.
-
-\paragraph{Beliefs in GRL} As mentioned already in the introduction, it is possible in GRL to rationalize models
-through so-called \emph{belief} elements.
-These beliefs are elements that can be attached to an IE with a \emph{belief link}.
-The content of a belief is a textual description, which can explain, for instance, the rationale for including an IE in the resulting goal model.
-There are several important differences between GRL beliefs and RationalGRL arguments.
-
-First of all, beliefs in GRL are not used in any of the algorithms to compute the satisfiability of the IEs.
-This means that a GRL belief can be understood as metadata that can be attached to an element, but it does not have semantics.
-Argument in RationalGRL, in contrast, play an important role in forming the resulting goal model.
-If an IE is defeated by an argument, it is not part of the resulting goal model.
-
-Secondly, arguments in RationalGRL can not only provide rationale for IEs in the resulting goal model, but they can also provide rational for those that are rejected, and they can be attached to links as well.
-This means that Rational arguments are strictly more expressive than GRL beliefs.
-
-Finally, the formation of arguments and counterarguments in RationalGRL is guided by critical questions, which stimulate discussions between stakeholders. This is not possible for GRL beliefs.
 

### 2018-08-13 Response to reviewers2.txt
excel & response to rev

--- 2018-05-06 Response to reviewers.txt
@@
-######## EDITOR ########
-
-===== COMMENT #1
-Please ensure also that your manuscript shows clearly how your work relates to relevant works as they may have been published in the Requirements Engineering or presented at the Requirements Engineering series of conferences.
-
-===== REPLY #1
-In the first submission, 6 of the 44 references in the article came from either REJ or the RE conference [11,24,26,32,43,47].
-We have searched through the journal and conference database for more references, and we found 4 more articles that are related to our work [20,22,27,39].
-We have added a discussion of these articles to the related work section, at the beginning of the paragraph "Requirements Engineering".
-This means we currently have 10 references from the Requirements Engineering Journal or presented at the Requirements Engineering series of conferences.
-
-######## REVIEWER #1 ########
-
-We thank the reviewer for the insightful comments.
-Especially the comment about the missing validation was useful since it motivated us to perform an small-scale in-house user study, which allowed us to make some useful observation regarding the usefulness and usability of our framework.
-We’ve replied to each separate comment below.
-
-===== COMMENT #1
-The integrated meta model looks reasonable but I have some doubts about the inclusion of "critical question" in the schema.
-Critical questions are something that are used to identify arguments and attack links, but  I don't see them as a concept to be included in the meta model.
-They are part of the methodology used to build the goal model rather concepts of the language.
-
-===== REPLY #1
-We thank the reviewer for pointing this out. We agree critical questions should not be part of the metamodel, and we have removed them accordingly. We also have updated the text accordingly.
-
-===== COMMENT #2
-The proposed algorithms are very simple and they do not add much to the overall presentation, I think they can be summarized in the section describing the tool.
-
-===== REPLY #2
-We agree that many algorithms are similar, and we have therefore shortened the sections containing these algorithms: sections 5.4.1 to 5.4.3 now all contain one example of the specific type of algorithm instead of two or more.
-
-===== COMMENT #3
-The methodology and the tools looks quite preliminary and without a real validation of the approach.
-I think this is the weakest part of the paper.
-Authors should demonstrate the applicability, scalability and usefulness of their approach in the real word.
-The tool and the modeling language should be used and evaluated by practitioners while they cope with real requirements engineering problems and complexity issues.
-
-===== REPLY #3
-We agree with the reviewer that a real validation of the approach is a necessary next step.
-However, we would like to point the reviewer to the case study in Section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com).
-We do agree that further validation of the tool and method is necessary, preferably in an industrial setting.
-However, such an industrial strength case study is a significant amount of work (see for example Estrada et al.
-(2006 "An empirical evaluation of the i* framework in a model-based software generation environment".
-They took 9 months to conduct the case study.), and would most likely yield enough new insights for a separate paper.
-
-Since the main focus of our article is to explain the novel RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content.
-Still, we agree that having no validation at all is limiting, so we have performed a small-scale user study in order to provide an initial validation of our framework.
-This empirical study was performed amongst 16 participants, each with a computer science background.
-
-We have added a section "RationalGRL user validation" (Section 7) where we describe our empirical study.
-
-######## REVIEWER #2 ########
-
-We thank the reviewer for reading our article in detail, and providing many insightful and useful comments. We’ve replied to each separate comment below.
-
-===== COMMENT #1
-I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable.
-However, this remains to be shown yet in an industrial environment.
-
-===== REPLY #1
-We have added a section "RationalGRL user validation" (Section 7) where we describe the results of a validation. Please see REPLY #3 to REVIEWER #1 for more details.
-
-===== COMMENT #2
-The statement on actually using a reduced GRL should be made in the Abstract and Introduction already, not just in a footnote later.
-
-===== REPLY #2
-We have included in the abstract and introduction that we use a simplified version.
-Note that all the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution.
-This simplification does not in any way influence the case study or any of the other results.
-Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would not change the essence of our formalism.
-
-===== COMMENT #3
+===== ORIGINAL COMMENT #3
 I think that the paper should compare the proposed argumentation structure at least with the beliefs in GRL.
@@
 
-===== REPLY #3
-We have added a comparison between arguments in RationalGRL and beliefs in GRL in the related work section, at the beginning of the paragraph "Goal Modelling".
+===== ORIGINAL REPLY #3
 
@@
 However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on a comparison with beliefs in GRL.
+
+====== REPLY
+In our first revision, we already added a brief comparison between arguments in RationalGRL and beliefs in GRL in the related work section (under "Beliefs in GRL").
+We have moved these paragraphs to "Goal Modeling" in section 8.1, and further extended and sharpened our discussion and comparison with GRL beliefs.
+Our main argument, which is now fully in the article, is that in GRL, beliefs are more of an afterthought: they have no formal semantics and it is not possible to further reason with them (i.e.
+given indicate conflicts between beliefs compute ways in which beliefs can be revised).
+Furthermore, we briefly discuss belief revision and the possibilities this may offer for future research.
+We still believe that discussing the relations between formal argumentation from AI and belief revision postulates from AI is not directly suited for the current (RE) audience.
+A more detailed comparison between RationalGRL and a framework based on belief revision would require us to develop a new belief revision framework connected to GRL and then model, for example, the case study or an extended example using this new belief revision framework for GRL.
+This would provide enough matrial for a new paper.
 

### 2018-08-15 validation.tex
1st update section 7

@@
 
-In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a small user evaluation with 16 users.
-This evaluation concerned mainly with the use of the RationalGRL tool, but more general questions regarding the usefulness of arguments in goal modeling were also asked.
-We first present the details and results of the evaluation, and then provide a more high-level analysis in Section~\ref{sec:eval-an}.
+In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a user evaluation with 16 users.
+The objective of this evaluation was to determine whether the users found the additions of RationalGRL (arguments, argumentation semantics, critical questions) useful compared to standard GRL, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to standard GRL.
+Regarding the comparison of RationalGRL to GRL, the following two broad hypotheses follow from our main points in this paper:
 
-\subsection{Experimental setup}
-The study consists of three parts:
+\begin{itemize}
+\item[H1] The additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
+\item[H2] The additions of RationalGRL make it easier to express, determine the effect of and communicate to other stakeholders one's opinions and beliefs about a goal model.
+\end{itemize}
+
+In addition, we wanted to know whether the RationalGRL tool was difficult or easy to use.
+Note that we did not aim to conduct a usability study for the RationalGRL tool.
+Rather, we were interested to know if the tool was clearly very easy or difficult to use, as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive twoards a particular modelling language, whereas irritations about a bad or buggy tool will lead to a more negative disposition.
+This leads to the following general hypothesis:
+
+\begin{itemize}
+\item[H3] The usability of the RationalGRL tool is neither (very) easy or (very) difficult. 
+\end{itemize}
+
+In the rest of this section, we will describe our experiment in more detail and further specify these hypotheses.
+
+\subsection{Experiment design}
+
+The idea was to have our participants perform a small modelling task, and then ask them what they thought of RationalGRL and the tool. Thus, our experiment consisted of three parts:
+
 \begin{enumerate}
@@
 \item Modeling: We ask the participants to model a summarized discussion in the RationalGRL tool.
-\item Survey: We ask some questions about the usability of the RationalGRL tool, and how it relates to other goal modeling languages.
+\item Survey: We present the participants with a survey containing questions about the comparison between RationalGRL and GRL, and questions about the usability of the RationalGRL tool.
 \end{enumerate}
 
-The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} At the end of the document we provide a link to the survey.\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} The instructions we provide to the users for task 2 can be found in Appendix~\ref{sect:survey}.
+The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (part 1) starts with a general explanation of GRL and standard goal modelling similar to the explanation provided in section \ref{{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
 
-Two thirds of the participants had either a PhD or Master degree, 20\% a Bachelor degree, and one respondent responded with ``Other''.
-The participants turned out to have a relatively high level of experience with IT, with an average rating of 6 years, ranging from 0 to 25 years.
-They also judged themselves to be quite experienced with early-phase requirements engineering (average of rating 3.2 out of 5), but experience with goal modeling languages was less (rating 1.9 out of 5, although two respondents use goal modeling languages almost weekly).
-The participants of the study had experience with the following goal modeling languages: No experience (7 users), GRL (2 users), i* (5 users), Tropos (1 user), and KAOS (1 user).
+For part 2, we provided a small part of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the tool and take a screenshot of their final model.
+The idea is that the participants are ``present'' at a discussion about the early-stage requirements and are asked to model the goal model corresponding to these requirements.
 
-In task 2, users clearly made use of the extra features offered by the RationalGRL tool: they used the details pane more often than not (rating 3.2 out of 5) and on average produced 3 arguments from the short transcript.
-Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
-
-\subsection{Usability of the RationalGRL tool}
-
-There are four questions in the survey which ask users to rate various aspects of the RationalGRL tool on a scale of 1 to 5:
+For part 3, we provided the participants with a survey\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} .
+The survey starts with general questions about the participant (experience etc.).
+It then asks the following questions about the RationalGRL tool, asking the participants to rate various aspects of the tool on a Likert scale of 1 (very difficult) to 5 (very easy):
 \begin{itemize}
@@
 \end{itemize}
+We also ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could in the future be used in practice.
+
+The second part of the survey concerns the comparison between GRL and RationalGRL and starts with Figure~\ref{fig:example-small}, which models the discussion for part 2 (Appendix~\ref{sect:survey}) in standard GRL, so that the participants can compare their own RationalGRL models and experiences with the standard GRL model.
+The following questions ask the participants to rate RationalGRL on a Likert scale from 1 (very useless) to 5 (very useful):
+\begin{itemize}
+\item[Q5] Do you think the arguments and counterarguments of RationalGRL are a useful extension to standard goal modeling?
+\item[Q6] Do you think the critical questions and answers in the details pane are a useful extension to standard goal modeling?
+\item[Q7] Do you think the automatic determination of the status of arguments and elements is a useful extension to standard goal modeling?
+\end{itemize}
+The following questions ask the participants to rate RationalGRL vs. standard goal modelling on a Likert scale from 1 (much more difficult) to 5 (much easier):
+\item[Q8] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone to express beliefs and opinions in a goal model?
+\item[Q9] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone to determine the effect of beliefs and opinions on the resulting goal model?
+\item [Q10] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone who is not the original author to understand the goal model? 
+\end{itemize}
+We conclude with two open questions regarding the strengths and weaknesses of RationalGRL compared to GRL.
+
+\paragraph{Participants}
+We asked 16 participants in our network to participate in our experiment.
+Most of the participants were therefore either employed at the first author's company or staff and (ex-)students from the second author's university.
+Hence, two thirds of the participants had either a PhD or Master degree, 20\% a Bachelor degree, and one respondent responded with ``Other''.
+All but one participant had a year or more experience with software development.
+The average experience was 6.2 years (standard deviation 6.7), with 10 participants having less than 5 years of experience and 6 participants having 8 or more years of experience, with one participant having 25 years of experience.
+The participants also judged themselves to be quite competent in  early-phase requirements engineering: on average they gave themselves a rating of 3.2 out of 5 (standard deviation 1.3) with half of the participants claiming they were at least ``competent'' (rating 4) or ``very competent'' (rating 5).
+However, experience with goal modeling languages was markedly less: the average rating was 1.9 (standard deviation 1.3), with 9 participants claiming never to have used a goal modeling language (rating 1), and only two participants displaying regular use (monthly, rating 4, or weekly, rating 5).
+The participants that had experience with goal modelling languages had mostly used i* (5 users), with 2 users being familiar with GRL and 2 users having used another goal modelling language.
+
+\subsection{Results}
+
+In task 2, users clearly made use of the extra features offered by the RationalGRL tool: they used the details pane more often than not (rating 3.2 out of 5) and on average produced 3 arguments from the short transcript.
+Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
 
@@
 \begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
-& very difficult & somewhat difficult & neither difficult nor easy & somewhat easy & very easy & avg. score (1-5) & std. dev.\\
+& very useless & useless & neither useful nor useless & useful & very useful & avg. score (1-5) & std. dev.\\
+\hline
+Q5 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.68\\
+Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
+Q7 & 0 (0\%) & 0 (0\%) & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7
+\end{tabularx}
+\caption{Participant ratings of the usefulness of the additions of RationalGRL}
+\label{table:survey:table2}
+\end{table*}
+
+\begin{table*}[t]
+\centering
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
+& much more difficult & more difficult & neither more difficult nor easier & easier & much easier & avg. score (1-5) & std. dev.\\
+\hline
+Q8 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.68\\
+Q9 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
+Q10 & 0 (0\%) & 0 (0\%) & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7
+\end{tabularx}
+\caption{Participant ratings of whether the additions of RationalGRL make reasoning about a goal model easier}
+\label{table:survey:table2}
+\end{table*}
+
+\begin{table*}[t]
+\centering
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
+& very difficult & difficult & neither difficult nor easy & easy & very easy & avg. score (1-5) & std. dev.\\
 \hline
@@
 \label{table:survey:table1}
-\end{table*}
-
-\begin{table*}[t]
-\centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
-& very useful & somewhat useful & neither useful nor useless & somewhat useless & very useless & avg. score (1-5) & std. dev.\\
-\hline
-Q5 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.68\\
-Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
-Q7 & 0 (0\%) & 0 (0\%) & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7
-\end{tabularx}
-\caption{User scores for the comparison between RationalGRL and other goal modeling languages}
-\label{table:survey:table2}
 \end{table*}
@@
 
-\subsection{Comparison with other goal modeling languages}
-
-We asked the users three questions about the extension of RationalGRL on a scale of 1 to 5:
-\begin{itemize}
-\item[Q5] Do you think the arguments and counterarguments of RationalGRL are a useful extension to standard goal modeling?
-\item[Q6] Do you think the critical questions and answers in the details pane are a useful extension to standard goal modeling?
-\item[Q7] Do you think the automatic determination of the status of arguments and elements is a useful extension to standard goal modeling?
-\end{itemize}
-
 Table~\ref{table:survey:table2} shows the respondents' answers, the average score, and the standard deviation for each answer.
@@
 One user mentioned that this may be a problem with goal modeling in general: ``Goal models are already complex (...) I have worked for years on the effect of context on goal models, and my conclusion is that this was very interesting academic work but with close-to-zero practical implications, unfortunately.''.
+
+\subsection{Analysis and discussion}
+In order to test our main hypotheses H1, H2 and H3, we have to formulate null hypotheses and alternative hypotheses that can be tested.
+For H1 and H2, we hypothesize that the participants will, on average, rate the additions of RationalGRL as useful (rating 4) or very useful (rating 5) for Q5, Q6 and Q7, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q8, Q9 and Q10.
+In other words, we say that our null hypothesis and alternative hypothesis for H1 and H2 are as follows:
+\begin{itemize}
+[] H1$_{0}$ and H2$_{0}$: average rating for questions Q5-Q10 is 3 or lower.
+[] H1$_{a}$ and H1$_{a}$: average rating for questions Q5-Q10 is higher than 3.
+\end{itemize}
+
+For hypothesis H3, we do not claim that the usability of the tool is especially easy or difficult: we expect the average rating for questions Q1-Q4 to not be significantly higher or lower than 3 (neither easy or difficult).
+Thus, our null hypothesis and alternative hypothesis for H3 are as follows:
+\begin{itemize}
+[] H3$_{0}$: average rating for questions Q1-Q4 is 3.
+[] H3$_{a}$: average rating for questions Q1-Q4 is not 3.
+\end{itemize}
+
+Below, we will discuss each of the hypotheses in turn.
+
+\paragraph{Hypothesis H1}
+The first analysis concerns hypothesis H1, whether the additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
+Our hypothesis was that on average the participants would rate the additions as either useful (rating 4) or very useful (rating 5) in question Q4, Q5 and Q6.
+For each separate question Q4, Q5, Q6, we can thus formulate the following null hypothesis H1$_{0}$ and alternative hypothesis H1$_{a}$.
+
+\paragraph{Hypothesis H2}
+
+\paragraph{Hypothesis H3}
 

### 2018-08-29 Response to reviewers2.txt
typo's

@@
 ===== ORIGINAL REPLY #3
-
 The connection between argumentation and belief revision is interesting, especially from a theoretical point of view, and in knowledge representation and reasoning this is a popular topic of research (see references 5-7 below).
@@
 Our main argument, which is now fully in the article, is that in GRL, beliefs are more of an afterthought: they have no formal semantics and it is not possible to further reason with them (i.e.
-given indicate conflicts between beliefs compute ways in which beliefs can be revised).
+given indicated conflicts between beliefs compute ways in which beliefs can be revised).
 Furthermore, we briefly discuss belief revision and the possibilities this may offer for future research.
@@
 A more detailed comparison between RationalGRL and a framework based on belief revision would require us to develop a new belief revision framework connected to GRL and then model, for example, the case study or an extended example using this new belief revision framework for GRL.
-This would provide enough matrial for a new paper.
+This would provide enough material for a new paper.
 

### 2018-08-29 validation.tex
typo's

@@
 Note that we did not aim to conduct a usability study for the RationalGRL tool.
-Rather, we were interested to know if the tool was clearly very easy or difficult to use, as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive twoards a particular modelling language, whereas irritations about a bad or buggy tool will lead to a more negative disposition.
+Rather, we were interested to know if the tool was clearly very easy or difficult to use, as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive towards a particular modelling language, whereas irritations about a bad tool will lead to a more negative disposition.
 This leads to the following general hypothesis:
@@
 For part 2, we provided a small part of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the tool and take a screenshot of their final model.
-The idea is that the participants are ``present'' at a discussion about the early-stage requirements and are asked to model the goal model corresponding to these requirements.
+The idea is that the participants act as if they are ``present'' at a discussion about the early-stage requirements and are asked to model the goal model corresponding to these requirements.
 
-For part 3, we provided the participants with a survey\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} .
+For part 3, we provided the participants with a survey\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2}.
 The survey starts with general questions about the participant (experience etc.).

### 2018-09-14 validation.tex
updates experimental results

@@
-\section{RationalGRL user evaluation}
+\section{RationalGRL evaluation experiment}
 \label{sect:validation}
@@
 
+In order to test our main hypotheses H1, H2 and H3, we have to formulate the appropriate null hypotheses and alternative hypotheses.
+For H1 and H2, we hypothesize that the participants will, on average, rate the additions of RationalGRL as useful (rating 4) or very useful (rating 5) for Q5, Q6 and Q7, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q8, Q9 and Q10.
+In other words, we say that our null hypothesis and alternative hypothesis for H1 and H2 are as follows:
+\begin{itemize}
+[] H1$_{0}$ and H2$_{0}$: average rating for questions Q5-Q10 is 3 or lower.
+[] H1$_{a}$ and H2$_{a}$: average rating for questions Q5-Q10 is higher than 3.
+\end{itemize}
+Note that for each question (Q5-Q10) we can test the null and alternative hypothesis separately, giving insight into exactly which additions of RationalGRL were deemed useful.
+
+For hypothesis H3, we do not claim that the usability of the tool is especially easy or difficult: we expect the average rating for questions Q1-Q4 to not be significantly higher or lower than 3 (neither easy or difficult).
+Thus, our null hypothesis and alternative hypothesis for H3 are as follows:
+\begin{itemize}
+[] H3$_{0}$: average rating for questions Q1-Q4 is 3.
+[] H3$_{a}$: average rating for questions Q1-Q4 is not 3.
+\end{itemize}
+Again, we can test these hypothesis separately for each of the questions.
+
 \paragraph{Participants}
@@
 
-\subsection{Results}
+\subsection{Results}\label{sec:survey:results}
 
-In task 2, users clearly made use of the extra features offered by the RationalGRL tool: they used the details pane more often than not (rating 3.2 out of 5) and on average produced 3 arguments from the short transcript.
-Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
+\begin{table*}[t]
+\centering
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
+& very difficult & difficult & neither difficult nor easy & easy & very easy & avg. score (1-5) & std. dev.\\
+\hline
+Q1 & 2 (12.5\%) & 1 (6.3\%) & 6 (37.5\%) & 6 (37.5\% & 1 (6.3\%) & 3.2 & 1.1\\
+Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25\%) & 8 (50\%) & 0 (0\%) & 3.1 & 1.1\\
+Q3 & 0 (0\%) & 3 (20\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20\%) & 3.5 & 1.1\\
+Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0\%) & 2.6 & 1.0
+\end{tabularx}
+\caption{User scores for the usability of the RationalGRL tool}
+\label{table:survey:table1}
+\end{table*}
 
@@
 \hline
-Q5 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.68\\
+Q5 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.7\\
 Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
@@
 \hline
-Q8 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.68\\
+Q8 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.7\\
 Q9 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
@@
 \end{tabularx}
-\caption{Participant ratings of whether the additions of RationalGRL make reasoning about a goal model easier}
-\label{table:survey:table2}
+\caption{Participant ratings of whether the additions of RationalGRL make reasoning about a goal model easier **MARC PLEASE FILL IN THE CORRECT NUMBERS FROM THE SURVEY HERE, TNX}
+\label{table:survey:table3}
 \end{table*}
 
-\begin{table*}[t]
-\centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
-& very difficult & difficult & neither difficult nor easy & easy & very easy & avg. score (1-5) & std. dev.\\
-\hline
-Q1 & 2 (12.5\%) & 1 (6.3\%) & 6 (37.5\%) & 6 (37.5\% & 1 (6.3\%) & 3.2 & 1.1\\
-Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25\%) & 8 (50\%) & 0 (0\%) & 3.1 & 1.1\\
-Q3 & 0 (0\%) & 3 (20\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20\%) & 3.3 & 1.4\\
-Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0\%) & 2.6 & 1.0
-\end{tabularx}
-\caption{User scores for the usability of the RationalGRL tool}
-\label{table:survey:table1}
-\end{table*}
+First, the participants had to perform the modeling task.
+They produced RationalGRL models containing on average 3 arguments (in addition  to goals, tasks, etc.) from the short transcript.
+Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
 
-Table~\ref{table:survey:table1} shows the respondents' answers, the average score, and the standard deviation for each answer.
-On average, the respondent found the tool relatively easy to use, but there were also a number of users that had some trouble getting started (in 3 out of the 4 questions, 12.\%5 rated ``very difficult'').
-We think this is not unexpected, as the cognitive burden for our framework is higher than for other goal modeling languages such as GRL and i*, simply because there are more elements in the language.
-Still, the majority of the users did not found the tool difficult to use.
-
+After the modelling task, the survey questions were asked. Table~\ref{table:survey:table1} shows the respondents' answers, the average score, and the standard deviation for questions Q1-Q4. 
 We also asked the users three open questions related to usability.
-When asked about the strengths and weaknesses of RationalGRL, users where generally positive about RationalGRL's clear UI and the fact that arguments can be combined with goal models.
-For instance, one user stated that RationalGRL successfully ``tries to capture the rationale behind the modeling process' and ``has a simple way to compute the status of the arguments''.
-On the other hand, some of the users found it challenging to decide when to use arguments, and to determine when an arguments attacks another argument.
-For instance, a user mentioned ``it wasn't very easy because I didn't always know how to attack the elements'', and ``I don't think its necessary to include all the arguments because some of them are a bit trivial''.
-Other users were worried about the cognitive overload of the tool.
-One user mentioned ``I still believe in the value of arguments, but there should be less confusing ways to capture them''.
-
-When asked about improvements in the RationalGRL tool, the users mentioned additional UI functionalities such as the possibility to save models, having an ``undo'' function, and flipping the arrow after adding them.
+When asked about the strengths and weaknesses of the RationalGRL tool, users where generally positive about RationalGRL's clear UI, the fact that the tools gives a nice overview of the goals and opinions in a case and that it was straightforward to use and understand.
+With respect to weaknesses and improvements to the RationalGRL tool, the users mentioned additional UI functionalities such as the possibility to save models, having an ``undo'' function, and flipping the arrow after adding them.
 Users also suggested various language-related improvements.
@@
 
-When asked whether the users believe a more mature version of RationalGRL can be used to capture early-phase requirements in an actual software development project, most users responded positively.
-This was somewhat surprising, since our tool is still an early prototype.
+When asked in an open question whether the users believe a more mature version of the RationalGRL tool could be used to capture early-phase requirements in an actual software development project, most users responded positively.
 Users' responses include ``Yes, having the possibility to add arguments seems quite useful'', ``I think it can, but maybe try to find a way to combine it with more regular/mainstream requirements gathering such as user stories and customer journeys'', ``yes, because it creates a clear scope for the project, what the goals are'', and ``Generally I think its useful to explicitly document arguments of a discussion''.
-The concern that the modeling process may involve too much cognitive overhead was mentioned again though: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', and ``the manual input is too complex and takes too much time.
-An automated process of parsing the conversation log would be much more helpful''.
+The concern that the modeling process may involve too much cognitive overhead was mentioned a few times though: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', ``the manual input is too complex and takes too much time.
+An automated process of parsing the conversation log would be much more helpful'', and ``I still believe in the value of arguments, but there should be less confusing ways to capture them''.
 
-Table~\ref{table:survey:table2} shows the respondents' answers, the average score, and the standard deviation for each answer.
-On average, the users were positive about the usefulness of the RationalGRL extension.
-The rating for the usefulness of determining the status of arguments was particularly positive (rating 4.2 out of 5), which provides some indication to the fact that using argumentation semantics to be able to determine precisely which elements of a goal model are accepted could be useful in practice.
-
+Tables~\ref{table:survey:table2} and \ref{table:survey:table3} show the respondents' answers, the average score, and the standard deviation for questions Q5-Q7.
 We concluded with two open questions about the comparison between RationalGRL and other goal modeling languages.
 When asked about the advantages of RationalGRL over standard goal modeling languages, many users agreed that making arguments explicit may force end users to have a more structured discussion: ``	Clear communication about argumentation and forcing people to think in those clear terms.'', ``...you can add arguments and that you can answer questions that help you to develop arguments'', ``It's useful that discussion and explanation are close to the diagrams'', ``a way to see how decisions are being shaped.''.
+Furthermore, one user stated that RationalGRL successfully ``tries to capture the rationale behind the modeling process' and ``has a simple way to compute the status of the arguments''.
 When asked about the weaknesses, users mentioned the complexity as the most important weakness: ``The apparent increase in complexity might lead to negative perceptions'', ``adding yet another layer of complexity scares me''.
@@
 \subsection{Analysis and discussion}
-In order to test our main hypotheses H1, H2 and H3, we have to formulate null hypotheses and alternative hypotheses that can be tested.
-For H1 and H2, we hypothesize that the participants will, on average, rate the additions of RationalGRL as useful (rating 4) or very useful (rating 5) for Q5, Q6 and Q7, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q8, Q9 and Q10.
-In other words, we say that our null hypothesis and alternative hypothesis for H1 and H2 are as follows:
-\begin{itemize}
-[] H1$_{0}$ and H2$_{0}$: average rating for questions Q5-Q10 is 3 or lower.
-[] H1$_{a}$ and H1$_{a}$: average rating for questions Q5-Q10 is higher than 3.
-\end{itemize}
-
-For hypothesis H3, we do not claim that the usability of the tool is especially easy or difficult: we expect the average rating for questions Q1-Q4 to not be significantly higher or lower than 3 (neither easy or difficult).
-Thus, our null hypothesis and alternative hypothesis for H3 are as follows:
-\begin{itemize}
-[] H3$_{0}$: average rating for questions Q1-Q4 is 3.
-[] H3$_{a}$: average rating for questions Q1-Q4 is not 3.
-\end{itemize}
-
-Below, we will discuss each of the hypotheses in turn.
 
@@
 The first analysis concerns hypothesis H1, whether the additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
-Our hypothesis was that on average the participants would rate the additions as either useful (rating 4) or very useful (rating 5) in question Q4, Q5 and Q6.
-For each separate question Q4, Q5, Q6, we can thus formulate the following null hypothesis H1$_{0}$ and alternative hypothesis H1$_{a}$.
+Table~\ref{fig:statsanalysis1} summarizes the average scores, standard deviations, p-values for a right-tailed t-test, effect size and statistical power for the relevant questions Q5-Q7.
+As can be seen, all the averages are significantly higher than the median of 3 with a significance level ($\alpha$) of 0.05, which means that we can accept H1$_{a}$ and reject H1$_{0}$ for all the separate questions Q5-Q7.
+In other words, the fact that the participants rate the additions of RationalGRL as, on average, (very) useful is very unlikely to be due to chance.
+
+For questions Q5 and Q7, the effect sizes are very large\footnote{Where am low p-value indicates whether the fact that average rating of a question is statistically significantly higher than the median 3, the effect size indicates the magnitude of the difference between the measured average and the median.
+Generally, an effect size below 0.2 is seen as low, and effect size of 0.5 is seen as medium, an effect size of 0.8 is seen as large and an effect size of 1.2 is seen as very large.}.
+This means that the addition of arguments and the use of arguments to determine the acceptability of elements of the goal model are deemed to be useful to very useful by the participants.
+This can also be seen in Table~\ref{table:survey:table2}: no participant found the addition of arguments and argument status useless or very useless, and 78.6\% (Q5) respectively 85.7\% (Q7) found the additions useful or very useful.
+Furthermore, these findings fit with the answers to the open questions (see Section \ref{sec:survey:results}).
+Finally, the statistical power for these questions is very high (>0.99), which means that the chances of wrongly rejecting H1$_{0}$ are close to 0.
+
+Interestingly, while for Q6 the average is significantly higher than the median 3, the effect size is only medium and the power is not very high, with the chances of wrongly rejecting H1$_{0}$ being almost 50\%.
+Hence, we can say that the addition of critical questions is perhaps not as useful as we had hypothesized.
+This again fits the descriptive statistics in Table\ref{table:survey:table2}: 21.4\% of respondents found the critical questions ``useless''.
+Furthermore, in the open questions a number of participants indicated that the exact difference between some of the critical questions was unclear, and that some of the critical questions seemed unnecessary, at least for the small modeling exercise given in the experiment.
 
 \paragraph{Hypothesis H2}
+The second hypothesis states that additions of RationalGRL make it easier to express one's opinions and beliefs about a goal model, determine the effect of one's opinions and beliefs about a goal model and communicate one's opinions and beliefs about a goal model to other stakeholders.
+Table~\ref{fig:statsanalysis1} again provides a summary of the statistical analysis of the relevant questions Q8-Q10.
+As can be seen, all the averages are significantly higher than the median of 3 with a significance level ($\alpha$) of 0.05, which means that we can accept H1$_{a}$ and reject H1$_{0}$ for all the separate questions Q8-Q10.
+In other words, the fact that the participants rate the additions of RationalGRL as making it, on average, easier to reason with their beliefs and opinions about a goal model is very unlikely to be due to chance.
+
+With respect to Q8, there is again a very large effect size and high statistical power: participants say that arguments make it much easier to express their opinions.
+None of the participants found expressing their beliefs (much) more difficult, and 69.3\% found it easier or much easier.
+This corresponds to the answers to Q5, where participants said the arguments were a useful addition.
+Like for Q5, for Q8 the chances of wrongly rejecting H1$_{0}$ are close to 0.
+
+For Q9, the effect size is somewhat smaller, though still large: RationalGRL makes it easier to determine the effect of one's beliefs on the goal model.
+Power shows that there is a small (around 7.5\% chance) or wrongly rejecting the null hypothesis.
+However, combining this with the results for question Q7, which asked about the usefulness of determining the status, we can say that the participants overall found this a useful feature which makes working with conflicting beliefs easier.
+One participant remarked that ``The reasoning [in the example] seems simplistic [...] the added value of [the formal argumentation] would be justified if the reasoning is more complex''.
+So the simplistic case in the modeling exercise possibly detracts from the results for Q9 and Q7.
+
+The analysis of the results for Q10 show a medium to large effect size: overall, participants were less sure that arguments would help in communicating the goal model to others.
+Statistical power is also not that high at 77\%.
+This may be due to the fact that in the experiment, communication of opinions to others was not really tested, leaving the participants to guess what this effect would be.
 
 \paragraph{Hypothesis H3}
+\begin{figure}[t]
+\centering
+\includegraphics[width=\textwidth]{img/Table_analysisQ1-Q4.png}
+\caption{Statistical analysis of the results of Q1-Q4 **MARC CAN YOU CONVERT THIS INTO A PROPER LATEX TABLE? THANKS!}
+\label{fig:statsanalysis2}
+\end{figure}
+
+On average, the participants did not find the tool easy or difficult to use, with most average scores being between 2.5 and 3.5.
+There were a number of users that had some trouble getting started (in 3 out of the 4 questions, 12.\%5 rated ``very difficult'').
+We think this is not unexpected, as the cognitive burden for our framework is higher than for other goal modeling languages such as GRL and i*, simply because there are more elements in the language.
 

### 2018-09-14 validation.tex
fixed some compiling errors

@@
 
-The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (part 1) starts with a general explanation of GRL and standard goal modelling similar to the explanation provided in section \ref{{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
+The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (part 1) starts with a general explanation of GRL and standard goal modelling similar to the explanation provided in section \ref{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
 
@@
 The following questions ask the participants to rate RationalGRL vs. standard goal modelling on a Likert scale from 1 (much more difficult) to 5 (much easier):
+\begin{itemize}
 \item[Q8] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone to express beliefs and opinions in a goal model?
@@
 \begin{itemize}
-[] H1$_{0}$ and H2$_{0}$: average rating for questions Q5-Q10 is 3 or lower.
-[] H1$_{a}$ and H2$_{a}$: average rating for questions Q5-Q10 is higher than 3.
+\item H1$_{0}$ and H2$_{0}$: average rating for questions Q5-Q10 is 3 or lower.
+\item H1$_{a}$ and H2$_{a}$: average rating for questions Q5-Q10 is higher than 3.
 \end{itemize}
@@
 \begin{itemize}
-[] H3$_{0}$: average rating for questions Q1-Q4 is 3.
-[] H3$_{a}$: average rating for questions Q1-Q4 is not 3.
+\item H3$_{0}$: average rating for questions Q1-Q4 is 3.
+\item H3$_{a}$: average rating for questions Q1-Q4 is not 3.
 \end{itemize}

### 2018-09-14 Response to reviewers2.txt
revisions for reviewer 2

@@
+<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
+<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<< REPLIES FOR REVIEW #2 >>>>>><<<<>>>>>>>>>>>>>>>>>>>>>>
+<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
+
+===== COMMENT #1
+In the course of the revision, several of my suggestions for improvement have simply been refused. Hence, I do not consider the improvements sufficient.
+
+===== REPLY #1
+We have reconsidered all of your suggestions for improvements in the first revision, and did another cycle of improvements of our paper based on your suggestions.
+We have replied to those suggestions below.
+
+===== COMMENT #2
+To the contrary, the newly added Section 7 is unacceptable from a scientific point of view.
+First, it is called a "small user evaluation".
+Then it has a subsection entitled "Experimental setup", but there are no hypotheses.
+Instead, there is another subsection entitled "Analysis", which starts with "Although our empirical evaluation ...".
+And why is there another subsection about usability of the tool?
+If that was the concern, a usability test or a heuristic evaluation would have had to be done.
+
+===== REPLY #2
+— Please add here Floris —
+
+<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
+<<<<<<<<<<<<<<<<<<<<<<< REVISION OF REPLIES FOR REBUTTAL #1 >>>>>>>>>>>>>>>>>>>>>>>>>>>>
+<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
+
 ===== ORIGINAL COMMENT #3
@@
 In our first revision, we already added a brief comparison between arguments in RationalGRL and beliefs in GRL in the related work section (under "Beliefs in GRL").
-We have moved these paragraphs to "Goal Modeling" in section 8.1, and further extended and sharpened our discussion and comparison with GRL beliefs.
+We have moved these paragraphs to “Belief Revision“ in section 8.2 (Future work), and further extended and sharpened our discussion and comparison with GRL beliefs.
 Our main argument, which is now fully in the article, is that in GRL, beliefs are more of an afterthought: they have no formal semantics and it is not possible to further reason with them (i.e.
@@
 Furthermore, we briefly discuss belief revision and the possibilities this may offer for future research.
-We still believe that discussing the relations between formal argumentation from AI and belief revision postulates from AI is not directly suited for the current (RE) audience.
 A more detailed comparison between RationalGRL and a framework based on belief revision would require us to develop a new belief revision framework connected to GRL and then model, for example, the case study or an extended example using this new belief revision framework for GRL.
-This would provide enough material for a new paper.
+This would be interesting material for a new paper
 
@@
 
-===== REPLY #4 & #5
+===== ORIGINAL REPLY #4 & #5
 We have added some discussion in the 2nd paragraph of section 8.1-Design Rationale.
@@
 
-===== REPLY #6
+===== ORIGINAL REPLY #6
 Propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE.
@@
 
+===== REPLY #6
+We have added a paragraph to the introduction of section 5, where we explain that we are in fact doing automatic reasoning by implementing the argumentation semantics, but that, from a practical point of view, we have chosen to provide algorithms instead of a deduction system.
+
 ===== COMMENT #7
@@
 
-===== REPLY #7
+===== ORIGINAL REPLY #7
 While there is extensive work on less "binary" frameworks for argumentation (see e.g below references 2 and 3), we feel that a discussion on this would not be suitable for the current paper (as it is more of an AI subject than relevant for RE).
@@
 If the reviewer desires, however, we can add the references to the article.
+
+===== REPLY #7
 

### 2018-09-14 discussion.tex
fixed some compiling errors
revisions for reviewer 2

@@
 Furthermore, for goal models, it is only possible to determine the satisfiability of goals given the possible tasks and resources; what cannot be determined is the acceptability of goals, that is, whether they are \emph{acceptable} (cf.
-Definition 14} given potentially contradictory opinions of stakeholders.
+Definition 14 given potentially contradictory opinions of stakeholders).
 
@@
 Arguments in RationalGRL, in contrast, play an important role in forming the resulting goal model: if an element is defeated by an argument, it is not part of the resulting goal model.
-
-Considering beliefs, an interesting avenue for further research would be to look at to connections between argumentation and belief revision \cite{alchourron1985logic}, and how these can play a role in RationalGRL and requirements engineering in general.
-While there is formal work in Artificial Intelligence on the connections between argumentation and belief revision (see e.g.
-\cite{falappa2009belief}, there is very little work on applying the combination to requirements engineering.
-The only work we are aware of that deals with his subject is by Gervasi and Zowghi \cite{gervasi2005reasoning}, who use a logical framework for belief revision to detect inconsistencies in software specifications and then revise those specifications so that the inconsistencies are removed.
-They do not explicitly connect their framework to goal models of goal modelling, however, and do not connect their framework to specific argumentation using schemes in the way RationalGRL does.
 
@@
 
+\paragraph{Belief revision} An interesting avenue for further investigation is the theoretical foundation of RationalGRL.
+Since it is one of our main goals to develop a framework that can be used by practitioners, we developed a formalism that is geared towards implementation using algorithms and a compact propositional logic.
+However, there is a long strand of research in Artificial Intelligence focusing on developing postulates that characterize how an agent can revise its knowledge base in a rational way~\cite{falappa2009belief}.
+The most famous postulates are the so-called AGM postulates~\cite{alchourron1985logic}.
+It would be very interesting to study whether (a variation of) these rationality postulates can be used to characterize the algorithms we have developed in this article, and more generally, what rationality means in our setting.
+While there is formal work in Artificial Intelligence on the connections between argumentation and belief revision (see e.g.
+\cite{falappa2009belief}, there is very little work on applying the combination to requirements engineering.
+The only work we are aware of that deals with his subject is by Gervasi and Zowghi \cite{gervasi2005reasoning}, who use a logical framework for belief revision to detect inconsistencies in software specifications and then revise those specifications so that the inconsistencies are removed.
+They do not explicitly connect their framework to goal models of goal modeling, however, and do not connect their framework to specific argumentation using schemes in the way RationalGRL does.
+
 \paragraph{Process and tool support}

### 2018-09-14 formalframework.tex
revisions for reviewer 2

@@
 Section~\ref{sect:background:pras}) use formal logic, allowing us to employ existing technique directly in order to compute which arguments are accepted and which are rejected, (ii) we can be more precise about how critical questions are answered, (iii) we can show that RationalGRL models can be translated to valid GRL models and vice versa in a precise way, and (iv) the formal approach is a basis for automating the framework in terms of tool support, which we present in the next section.
+
+We use formal logic (propositional logic) to specify our framework, and in order to develop our tool we have implemented an algorithm to compute argumentation extensions based on this formalization.
+However, we do not use a dedicated theorem prover or a logical inference engine.
+The reason is purely practical, it is our aim to develop a tool that practitioners can use in their browser, and it does not seem necessary nor useful to use an external theorem prover for this.
+Instead, in this section we propose algorithms based on our logical specification, which form the basis of our prototype implementation.
 

### 2018-09-14 validation.tex
revisions for reviewer 2

@@
 \centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
-& very difficult & difficult & neither difficult nor easy & easy & very easy & avg. score (1-5) & std. dev.\\
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
+& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p value & power & effect size\\
 \hline
-Q1 & 2 (12.5\%) & 1 (6.3\%) & 6 (37.5\%) & 6 (37.5\% & 1 (6.3\%) & 3.2 & 1.1\\
-Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25\%) & 8 (50\%) & 0 (0\%) & 3.1 & 1.1\\
-Q3 & 0 (0\%) & 3 (20\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20\%) & 3.5 & 1.1\\
-Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0\%) & 2.6 & 1.0
+Q1 & 2 (12.5\%) & 1 (6.3\%)  & 6 (37.5\%) & 6 (37.5\%) & 1 (6.3\%)  & 3.2 & 1.1 & 0.509 & ?? & ??\\
+Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25.0\%) & 8 (50.0\%) & 0 (0.0\%)  & 3.1 & 1.1 & 0.652 & ?? & ??\\
+Q3 & 0 (0.0\%)  & 3 (20.0\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20.0\%) & 3.5 & 1.1 & 0.072 & ?? & ??\\
+Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0.0\%)  & 2.6 & 1.0 & 0.138 & ?? & ??
 \end{tabularx}
@@
 \centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
-& very useless & useless & neither useful nor useless & useful & very useful & avg. score (1-5) & std. dev.\\
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
+& very useless & useless & neutral & useful & very useful & avg (1-5) & std. dev. & p value & power & effect size\\
 \hline
-Q5 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.7\\
-Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
-Q7 & 0 (0\%) & 0 (0\%) & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7
+Q5 & 0 (0.0\%) & 0 (0.0\%)  & 3 (21.4\%) & 8 (57.1\%) & 3 (21.4\%) & 4.0 & 0.7 & 0.000 & 0.999 & 1.5 \\
+Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0 & 0.045 & 0.525 & 0.5\\
+Q7 & 0 (0.0\%) & 0 (0.0\%)  & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7 & 0.000 & 1.000 & 1.7
 \end{tabularx}
@@
 \centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l|l|l}
-& much more difficult & more difficult & neither more difficult nor easier & easier & much easier & avg. score (1-5) & std. dev.\\
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
+& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p value & power & effect size\\
 \hline
-Q8 & 0 (0\%) & 0 (0.0\%) & 3 (21.4\%) & 8 (57.1\% & 3 (21.4\%) & 4.0 & 0.7\\
-Q9 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0\\
-Q10 & 0 (0\%) & 0 (0\%) & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7
+Q8  & 0 (0.0\%) & 0 (0.0\%) & 4 (30.8\%) & 4 (30.8\%) & 5 (38.5\%) & 4.1 & 0.9 & 0.000 & 0.991 & 1.2\\
+Q9  & 0 (0.0\%) & 0 (0.0\%) & 6 (46.2\%) & 5 (38.5\%) & 2 (15.4\%) & 3.7 & 0.8 & 0.003 & 0.925 & 0.9\\
+Q10 & 0 (0.0\%) & 1 (8.3\%) & 4 (33.3\%) & 7 (58.3\%) & 0 (0.0\%)  & 3.5 & 0.7 & 0.013 & 0.772 & 0.7
 \end{tabularx}
-\caption{Participant ratings of whether the additions of RationalGRL make reasoning about a goal model easier **MARC PLEASE FILL IN THE CORRECT NUMBERS FROM THE SURVEY HERE, TNX}
+\caption{Participant ratings of whether the additions of RationalGRL make reasoning about a goal model easier}
 \label{table:survey:table3}
@@
 Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
+The full results of the survey can be downloaded from \url{www.rationalgrl.com}, on the page ``Empirical study''.
 
 After the modelling task, the survey questions were asked. Table~\ref{table:survey:table1} shows the respondents' answers, the average score, and the standard deviation for questions Q1-Q4. 
+\todo{Marc}{Floris}{Wellicht kun je hier het best even uitleggen wat p value, power, en effect size betekenen?}
+
 We also asked the users three open questions related to usability.
@@
 
-Tables~\ref{table:survey:table2} and \ref{table:survey:table3} show the respondents' answers, the average score, and the standard deviation for questions Q5-Q7.
+Table~\ref{table:survey:table2} shows the respondents' answers, the average score, the standard deviation, p-value, power, and effect size for questions Q5-Q7, and table~\ref{table:survey:table3} shows the results for questions Q8-Q10.
 We concluded with two open questions about the comparison between RationalGRL and other goal modeling languages.
@@
 Furthermore, these findings fit with the answers to the open questions (see Section \ref{sec:survey:results}).
-Finally, the statistical power for these questions is very high (>0.99), which means that the chances of wrongly rejecting H1$_{0}$ are close to 0.
+Finally, the statistical power for these questions is very high (\textgreater 0.99), which means that the chances of wrongly rejecting H1$_{0}$ are close to 0.
 
@@
 
-\paragraph{Hypothesis H3}
-\begin{figure}[t]
-\centering
-\includegraphics[width=\textwidth]{img/Table_analysisQ1-Q4.png}
-\caption{Statistical analysis of the results of Q1-Q4 **MARC CAN YOU CONVERT THIS INTO A PROPER LATEX TABLE? THANKS!}
-\label{fig:statsanalysis2}
-\end{figure}
-
 On average, the participants did not find the tool easy or difficult to use, with most average scores being between 2.5 and 3.5.

### 2018-09-14 validation.tex
forgot to push some changes

@@
 
-Table~\ref{table:survey:table2} shows the respondents' answers, the average score, the standard deviation, p-value, power, and effect size for questions Q5-Q7, and table~\ref{table:survey:table3} shows the results for questions Q8-Q10.
+Table~\ref{table:survey:table2} shows the respondents' answers, the average score, the standard deviation, p-value, power, and effect size for questions Q5-Q7, and Table~\ref{table:survey:table3} shows the results for questions Q8-Q10.
 We concluded with two open questions about the comparison between RationalGRL and other goal modeling languages.

### 2018-09-15 background.tex
Small changes references and finished validation.tex

@@
 
-Our examples and case study are based on the data produced by a recent series of experiments by Schriek et al.
-\cite{SchriekEtal2016}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}.
+Our examples and case study are based on the data produced by a recent series of experiments by Tang et al.
+\cite{TangEtal2018}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}.
 This experiment contains a well-known design reasoning assignment in software engineering about a traffic simulator.
@@
 In order for the professor to teach students the various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
-Schriek et al.
+Tang et al.
 asked designers (groups of students) to discuss the requirements of this traffic simulator.

### 2018-09-15 discussion.tex
Small changes references and finished validation.tex

@@
 We have performed a case study and found more than 200 incidences of arguments and questions.
-We also did a user with 16 participants, testing the usability and usefulness of RationalGRL compared to other goal modeling languages.
+We also performed an experiment with 16 participants, testing the usability and usefulness of RationalGRL compared to other goal modeling languages.
 However, what is still lacking is an industrial-strength case study testing extensively whether, and how, the tools and techniques provided by the RationalGRL framework really improve early phase requirements engineering.
 The complexity of the domain makes focused experiments difficult, but not impossible.
-For example, Murakannaiah \emph{et al.}~\cite{murukannaiah2015} provided a realistic but bounded problem for their subjects to solve, and similar experiments are imaginable to test RationalGRL: have two groups solve a goal modeling problem, one using only GRL and one using RationalGRL, and examine how they perform.
+For example, both Murakannaiah \emph{et al.}~\cite{murukannaiah2015} and Tang \emph{et al.}~\cite{TangEtal2018} provided a realistic but bounded problem for their subjects to solve, and similar experiments are imaginable to test RationalGRL: have two groups solve a goal modeling problem, one using only GRL and one using RationalGRL, and examine how they perform.
 Naturally, the type of problem will greatly influence the outcome.
@@
 However, a more complex problem on which multiple persons have to work asynchronously might benefit from the extra tools offered by RationalGRL.
-Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the development process, the tool, or a list of critical questions to serve as reminders during the goal modeling process (cf.
-\cite{SchriekEtal2016}).
+Furthermore, the outcome also depends on which part of the set of tools and techniques offered by RationalGRL is tested: the development process, the tool, or a list of critical questions to serve as reminders during the goal modeling process.
 

### 2018-09-15 main.tex
Addressed some more comments by reviewers
survey
Small changes references and finished validation.tex

@@
 \author{Marc van Zee  \and Floris J. Bex\and Sepideh Ghanavati}
-\institute{Marc van Zee \at Google Research, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} 
-\and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands \\\email{f.j.bex@uu.nl}
+\institute{Marc van Zee \at Google BRAIN, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} 
+\and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands 
+				\at Tilburg Institute for Law, Technology, and Society, Tilburg University, Tilburg, The Netherlands \\\email{f.j.bex@uu.nl}
 \and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu} }
@@
 \input{formalframework}
-\input{methodology}
-%\input{tool}
+\input{tool}
+\input{validation}
 \input{discussion}
@@
 \input{tool-screenshots}
+\newpage
+\input{survey}
+\newpage
+\input{survey-screenshots}
 

### 2018-09-15 validation.tex
Small changes references and finished validation.tex

@@
 \centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
-& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p value & power & effect size\\
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l}
+& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p-value \\
 \hline
-Q1 & 2 (12.5\%) & 1 (6.3\%)  & 6 (37.5\%) & 6 (37.5\%) & 1 (6.3\%)  & 3.2 & 1.1 & 0.509 & ?? & ??\\
-Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25.0\%) & 8 (50.0\%) & 0 (0.0\%)  & 3.1 & 1.1 & 0.652 & ?? & ??\\
-Q3 & 0 (0.0\%)  & 3 (20.0\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20.0\%) & 3.5 & 1.1 & 0.072 & ?? & ??\\
-Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0.0\%)  & 2.6 & 1.0 & 0.138 & ?? & ??
+Q1 & 2 (12.5\%) & 1 (6.3\%)  & 6 (37.5\%) & 6 (37.5\%) & 1 (6.3\%)  & 3.2 & 1.1 & 0.509 \\
+Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25.0\%) & 8 (50.0\%) & 0 (0.0\%)  & 3.1 & 1.1 & 0.652 \\
+Q3 & 0 (0.0\%)  & 3 (20.0\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20.0\%) & 3.5 & 1.1 & 0.072 \\
+Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0.0\%)  & 2.6 & 1.0 & 0.138 
 \end{tabularx}
-\caption{User scores for the usability of the RationalGRL tool}
+\caption{Participant ratings and statistical results for the usability of the RationalGRL tool\todo{Floris}{Marc}{Fix table width}}
 \label{table:survey:table1}
@@
 \begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
-& very useless & useless & neutral & useful & very useful & avg (1-5) & std. dev. & p value & power & effect size\\
+& very useless & useless & neutral & useful & very useful & avg (1-5) & std. dev. & p-value & effect size $d$ & power\\
 \hline
-Q5 & 0 (0.0\%) & 0 (0.0\%)  & 3 (21.4\%) & 8 (57.1\%) & 3 (21.4\%) & 4.0 & 0.7 & 0.000 & 0.999 & 1.5 \\
-Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0 & 0.045 & 0.525 & 0.5\\
-Q7 & 0 (0.0\%) & 0 (0.0\%)  & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7 & 0.000 & 1.000 & 1.7
+Q5 & 0 (0.0\%) & 0 (0.0\%)  & 3 (21.4\%) & 8 (57.1\%) & 3 (21.4\%) & 4.0 & 0.7 & 0.000 & 1.5 & 0.999 \\
+Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0 & 0.045 & 0.5 & 0.525 \\
+Q7 & 0 (0.0\%) & 0 (0.0\%)  & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7 & 0.000 & 1.7 & 1.000
 \end{tabularx}
-\caption{Participant ratings of the usefulness of the additions of RationalGRL}
+\caption{Participant ratings and statistical results of the usefulness of the additions of RationalGRL}
 \label{table:survey:table2}
@@
 \begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
-& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p value & power & effect size\\
+& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p-value & effect size $d$ & power \\
 \hline
-Q8  & 0 (0.0\%) & 0 (0.0\%) & 4 (30.8\%) & 4 (30.8\%) & 5 (38.5\%) & 4.1 & 0.9 & 0.000 & 0.991 & 1.2\\
-Q9  & 0 (0.0\%) & 0 (0.0\%) & 6 (46.2\%) & 5 (38.5\%) & 2 (15.4\%) & 3.7 & 0.8 & 0.003 & 0.925 & 0.9\\
-Q10 & 0 (0.0\%) & 1 (8.3\%) & 4 (33.3\%) & 7 (58.3\%) & 0 (0.0\%)  & 3.5 & 0.7 & 0.013 & 0.772 & 0.7
+Q8  & 0 (0.0\%) & 0 (0.0\%) & 4 (30.8\%) & 4 (30.8\%) & 5 (38.5\%) & 4.1 & 0.9 & 0.000 & 1.2 & 0.991\\
+Q9  & 0 (0.0\%) & 0 (0.0\%) & 6 (46.2\%) & 5 (38.5\%) & 2 (15.4\%) & 3.7 & 0.8 & 0.003 & 0.9 & 0.925\\
+Q10 & 0 (0.0\%) & 1 (8.3\%) & 4 (33.3\%) & 7 (58.3\%) & 0 (0.0\%)  & 3.5 & 0.7 & 0.013 & 0.7 & 0.772
 \end{tabularx}
-\caption{Participant ratings of whether the additions of RationalGRL make reasoning about a goal model easier}
+\caption{Participant ratings and statistical results of whether the additions of RationalGRL make reasoning about a goal model easier}
 \label{table:survey:table3}
@@
 The full results of the survey can be downloaded from \url{www.rationalgrl.com}, on the page ``Empirical study''.
+\todo{Floris}{Marc}{Staan de results op de webpagina?}
 
-After the modelling task, the survey questions were asked. Table~\ref{table:survey:table1} shows the respondents' answers, the average score, and the standard deviation for questions Q1-Q4. 
-\todo{Marc}{Floris}{Wellicht kun je hier het best even uitleggen wat p value, power, en effect size betekenen?}
+After the modelling task, the survey questions were asked.
+Table~\ref{table:survey:table1} shows the respondents' answers to question Q1-Q4 together with some statistical results which will be further discussed in Section \ref{sect:validation:analysis}.
 
@@
 
-Table~\ref{table:survey:table2} shows the respondents' answers, the average score, the standard deviation, p-value, power, and effect size for questions Q5-Q7, and Table~\ref{table:survey:table3} shows the results for questions Q8-Q10.
+Table~\ref{table:survey:table2} shows the respondents' answers to questions Q5-Q7, and Table~\ref{table:survey:table3} shows the answers to questions Q8-Q10.
 We concluded with two open questions about the comparison between RationalGRL and other goal modeling languages.
@@
 
-\subsection{Analysis and discussion}
+\subsection{Analysis or results}\label{sect:validation:analysis}
 
@@
 The first analysis concerns hypothesis H1, whether the additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
-Table~\ref{fig:statsanalysis1} summarizes the average scores, standard deviations, p-values for a right-tailed t-test, effect size and statistical power for the relevant questions Q5-Q7.
-As can be seen, all the averages are significantly higher than the median of 3 with a significance level ($\alpha$) of 0.05, which means that we can accept H1$_{a}$ and reject H1$_{0}$ for all the separate questions Q5-Q7.
+Table~\ref{table:survey:table2} summarizes the average scores, standard deviations, p-values for a right-tailed t-test, effect size and statistical power for the relevant questions Q5-Q7.
+As indicated by the p-values, which are all lower than 0.05, all the averages are significantly higher than the median of 3 with a significance level ($\alpha$) of 0.05, which means that we can accept H1$_{a}$ and reject H1$_{0}$ for all the separate questions Q5-Q7.
 In other words, the fact that the participants rate the additions of RationalGRL as, on average, (very) useful is very unlikely to be due to chance.
 
-For questions Q5 and Q7, the effect sizes are very large\footnote{Where am low p-value indicates whether the fact that average rating of a question is statistically significantly higher than the median 3, the effect size indicates the magnitude of the difference between the measured average and the median.
+For questions Q5 and Q7, the effect sizes are very large\footnote{Where a low p-value for a hypothesis test such as a t-test indicates whether the fact that average rating of a question is statistically significantly higher than the median 3, the effect size indicates the magnitude of the difference between the measured average and the median.
 Generally, an effect size below 0.2 is seen as low, and effect size of 0.5 is seen as medium, an effect size of 0.8 is seen as large and an effect size of 1.2 is seen as very large.}.
 This means that the addition of arguments and the use of arguments to determine the acceptability of elements of the goal model are deemed to be useful to very useful by the participants.
-This can also be seen in Table~\ref{table:survey:table2}: no participant found the addition of arguments and argument status useless or very useless, and 78.6\% (Q5) respectively 85.7\% (Q7) found the additions useful or very useful.
+As can be seen in Table~\ref{table:survey:table2}: no participant found the addition of arguments and argument status useless or very useless, and 78.6\% (Q5) respectively 85.7\% (Q7) found the additions useful or very useful.
 Furthermore, these findings fit with the answers to the open questions (see Section \ref{sec:survey:results}).
-Finally, the statistical power for these questions is very high (\textgreater 0.99), which means that the chances of wrongly rejecting H1$_{0}$ are close to 0.
+Finally, despite the low $n$ the statistical power for these questions is very high (\textgreater 0.99), which means that the chances of wrongly rejecting H1$_{0}$ are close to 0\footnote{Statistical power indicates the probability of correctly rejecting the null hypothesis}.
 
@@
 The second hypothesis states that additions of RationalGRL make it easier to express one's opinions and beliefs about a goal model, determine the effect of one's opinions and beliefs about a goal model and communicate one's opinions and beliefs about a goal model to other stakeholders.
-Table~\ref{fig:statsanalysis1} again provides a summary of the statistical analysis of the relevant questions Q8-Q10.
+Table~\ref{table:survey:table3} provides a summary of the statistical analysis of the relevant questions Q8-Q10.
 As can be seen, all the averages are significantly higher than the median of 3 with a significance level ($\alpha$) of 0.05, which means that we can accept H1$_{a}$ and reject H1$_{0}$ for all the separate questions Q8-Q10.
@@
 
-On average, the participants did not find the tool easy or difficult to use, with most average scores being between 2.5 and 3.5.
-There were a number of users that had some trouble getting started (in 3 out of the 4 questions, 12.\%5 rated ``very difficult'').
-We think this is not unexpected, as the cognitive burden for our framework is higher than for other goal modeling languages such as GRL and i*, simply because there are more elements in the language.
+\paragraph{Hypothesis H3}
+The third hypothesis H3 is about the usability of the RationalGRL tool.
+It is different from the other two hypotheses in that we do not expect or claim that the usability of the tool is very good (or bad).
+Rather, because in the RationalGRL framework the tool is more a means than an end in itself, we hypothesize that the too, is neither easy or difficult to use.
+Table~\ref{table:survey:table1} summarizes the average scores, standard deviations and p-values for a two-tailed t-test for questions Q1-Q4.
+As indicated by the p-values which are all higher than $\alpha=0.05$, none of the results is statistically significant, that is, the fact that there is a difference between the median of 3 and the average scores for the questions could very well be due to chance.
+This means we cannot reject the null hypothesis H3$_{0}$ and that, hence, the usability of the RationalGRL tool is neither (very) easy or (very) difficult.
+This corresponds to the answers to the open questions the participants gave: respondents were positive about the fact that it was very easy to get started with the tool and most comments on the tool were the type of comments that are expected from a prototype implementation.
 
-\subsection{Analysis}\label{sec:eval-an}
-Although our empirical evaluation is relatively small-scale, and we cannot draw hard conclusions from it, we do believe that we can extract some interesting observations.
-We list the three most important ones:
+\subsection{Threats to validity}
+Internal validity is about the validity of the experiment results given our experiment setup and interpretation of findings.
+The main threat in this regard is that the modeling task the participants were asked to do is not realistic.
+The task is quite small, and based on an existing transcript of a requirements discussion.
+In a realistic setting, users of RationalGRL would model much more complex requirements together with the stakeholders.
+That said, the discussions analysed in our case study (Section \ref{sect:gmas}) were of a realistic size and included many arguments and critical questions, so these concepts seem to take a central role in goal modeling and requirements engineering.
+
+Another threat to internal validity is that many of the participants had no experience with goal modelling languages at all.
+Hence, the perceived usefulness of the arguments could be influenced by the fact that the participants, for the first time, were given a goal modeling language to perform structured requirements analysis.
+In order to test the usefulness of arguments and critical questions alone, one would have to compare a test and control group, where the test group uses RationalGRL and the control group uses normal GRL.
+
+External validity concerns the generalizability of the results.
+As the number of participants is fairly small (16), this threatens the generalizability of our findings.
+However, the type of participant is realistic: software engineers having a lot of experience with requirements engineering, as opposed to, for example, engineering students with little experience.
+Furthermore, the results are in line with existing research on design rationale.
+On the one hand, these existing studies, like our study, show that structured modeling languages suffer from the common problem of high cognitive overhead \cite{shum2006hypermedia}.
+One the other hand, studies show that the inclusion of argumentation and critical questions in design and requirements discussions improve the reasoning \cite{razavian2016two,TangEtal2018}.
+
+\subsection{Discussion}
+Although our empirical evaluation is relatively small-scale, the results point in some interesting directions for further discussion and research.
+
 \begin{enumerate}
+\item \emph{Arguments and automatically determining acceptability are useful.} Participants were enthusiastic about adding arguments to a goal modeling language with a clear formal underpinning, and they believed the argumentation semantics we use in the tool are intuitive.
+This is a positive signal for the addition of arguments to goal modeling languages, and also for our formal approach, as just adding arguments as, for example, labels on goals and tasks does not allow one to compute the status of elements in a goal model in the way our approach does.
+Furthermore, automatically determining the status of elements will perhaps have a bigger impact when working with larger, more complex goal models, for which manually determining the impact of arguments is not feasible.
+\item \emph{Critical questions need to be clear and have a prominent place in the process of building a goal model.} While participants overall found the critical questions (CQ's) useful, their impact was markedly less than the addition of arguments.
+This can have various causes.
+Participants mentioned they did not use the details panel of the tool, which can be used to ask and answer critical questions, that much.
+Furthermore, it was also mentioned that the difference between different questions was not always clear.
+Finally, the experiment task, modeling an existing discussion, may also have influenced the CQ's perceived usefulness.
+The CQ's are meant to be asked during the actual RationalGRL Development Process in which goal models are constructed and critically analysed from scratch.
 \item \emph{High cognitive overhead.} A concern that was raised often in our empirical evaluation is that, in its current form, RationalGRL has a relatively high cognitive overhead.
 Goal modeling is by itself already a cognitively high-effort activity, and the fact that we add more elements to the language does not improve this.
-This may mean that in future work we should focus more on the RationalGRL development process\footnote{Note that in order to keep the study simple for the users, we did not explicitly ask the respondents to follow the development process from Section~\ref{sect:methodology}}.
-Users like to be guided during modeling phase, and making arguments explicit in the modeling process was indicated as useful, but argumentation should be integrated into the \emph{process} of goal modeling more than in the goal models themselves.
+\end{enumerate}
+
+Taking the above into account, it seems natural in future work to focus more on the RationalGRL development process\footnote{Note that in order to keep the study simple for the users, we did not explicitly ask the respondents to follow the development process from Section~\ref{sect:methodology}}.
+Users like to be guided during modeling phase, and making arguments explicit in the modeling process was indicated as useful, but argumentation should be integrated into the \emph{process} of goal modeling more than in the goal models themselves, keeping the goal models simple and thus lessening the cognitive overhead.
 Critical questions seem to be a natural fit for this, as they play a key role in the RationalGRL development process, but in its current form they are only mentioned in the details pane of the tool.
-\item \emph{Useful argumentation semantics.} Participants were enthusiastic about adding arguments to a goal modeling language with a clear formal underpinning, and they believed the argumentation semantics we use in the tool is very intuitive.
-This is a positive signal for our formal approach.
-\item \emph{Lightweight tool.} Respondents were positive about the fact that it was very easy to get started with the tool.
-Thus, it seems that having a simple web-interface for goal modeling languages is something that could be done more in the future.
-Most comments on the tool were the type of comments that are expected from a prototype implementation.
-\end{enumerate}

### 2018-09-15 Response to reviewers2.txt
Reviewer comments

@@
 ===== COMMENT #2
-To the contrary, the newly added Section 7 is unacceptable from a scientific point of view.
+The newly added Section 7 is unacceptable from a scientific point of view.
 First, it is called a "small user evaluation".
@@
 ===== REPLY #2
-— Please add here Floris —
+We have significantly reworked section 7, going through the usual steps for an empirical evaluation and including extra statistical results.
+The section now includes the hypotheses that have been tested, results, a brief statistical analysis of the results, possible threats to validity and a discussion.
 
@@
 
-===== ORIGINAL REPLY #3
-The connection between argumentation and belief revision is interesting, especially from a theoretical point of view, and in knowledge representation and reasoning this is a popular topic of research (see references 5-7 below).
-However, since studying relations between argumentation and belief revision does not seem directly related to our current framework and seems more interesting from a formal logical point of view, we have left this out of the discussion in our paper, instead choosing to focus on a comparison with beliefs in GRL.
-
-====== REPLY
+====== REPLY #3
 In our first revision, we already added a brief comparison between arguments in RationalGRL and beliefs in GRL in the related work section (under "Beliefs in GRL").
-We have moved these paragraphs to “Belief Revision“ in section 8.2 (Future work), and further extended and sharpened our discussion and comparison with GRL beliefs.
+In this revision, we have done two things.
+First, we have moved the discussion on beliefs in GRL to "related work" under "goal modelling" and further extended and sharpened our discussion and comparison with GRL beliefs.
 Our main argument, which is now fully in the article, is that in GRL, beliefs are more of an afterthought: they have no formal semantics and it is not possible to further reason with them (i.e.
 given indicated conflicts between beliefs compute ways in which beliefs can be revised).
-Furthermore, we briefly discuss belief revision and the possibilities this may offer for future research.
+
+Second, we have added a new discussion under “Belief Revision“ in section 8.2 (Future work),  where we discuss belief revision and the possibilities this may offer for future research.
 A more detailed comparison between RationalGRL and a framework based on belief revision would require us to develop a new belief revision framework connected to GRL and then model, for example, the case study or an extended example using this new belief revision framework for GRL.
@@
 
-===== ORIGINAL REPLY #4 & #5
-We have added some discussion in the 2nd paragraph of section 8.1-Design Rationale.
-A brief comparison with, for example gIBIS, is already in section 8.1-Design Rationale (reference [38] includes gIBIS).
-We felt that an extensive comparison between the ideas behind goal modelling, which we extend with RationalGRL, and design rationale diagramming methods like gIBIS is not within the extent of this paper, where the focus is more on the connection between "normal" goal modelling (GRL) and extended "argumentative" goal modelling (RationalGRL).
+===== REPLY #4 & #5
+In the first revision, we added some discussion in the 2nd paragraph of section 8.1-Design Rationale, noting that a comparison with gIBIS is contained in this discussion (reference [38] includes gIBIS).
+In this revision, we have made this connection explicit.
+We have clearly indicated that we combine ideas from DR and goal modeling.
 
-===== COMMENT #6
+Furthermore, in this revision we have given new reasons (under 8.1, second par, "There are two reasons for this....") for why we connect to goal diagramming methods, even though the trend in DR is to not specifically diagram reasons.
+
+===== ORIGINAL COMMENT #6
 I wonder about the formalization in propositional logic. Why is it not used for automated reasoning (at least partly instead of the specific algorithms)?
 
-===== ORIGINAL REPLY #6
-Propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE.
+===== REPLY #6
+In the original reply, we indicated that propositional logic in itself does not contain specific algorithms for, for example, INTRO, REPLACE and DISABLE.
 Furthermore, the algorithms also introduce arguments and attacks, formal notions of which are not in basic propositional logic.
@@
 
-===== REPLY #6
-We have added a paragraph to the introduction of section 5, where we explain that we are in fact doing automatic reasoning by implementing the argumentation semantics, but that, from a practical point of view, we have chosen to provide algorithms instead of a deduction system.
+In this revision, we have added a paragraph to the introduction of section 5, where we explain that we are in fact doing automatic reasoning by implementing the argumentation semantics, but that, from a practical point of view, we have chosen to provide algorithms instead of a propositional logic deduction system.
 
@@
 
-===== ORIGINAL REPLY #7
-While there is extensive work on less "binary" frameworks for argumentation (see e.g below references 2 and 3), we feel that a discussion on this would not be suitable for the current paper (as it is more of an AI subject than relevant for RE).
-For an extensive discussion on the relation between the formal argumentation proposed in our article and argumentation in reality, we refer the reviewer to reference [4] below.
-Note that we have opted not to include these references in the paper, as we are already at 50 references and want to focus more on an RE/SE audience.
-If the reviewer desires, however, we can add the references to the article.
+===== REPLY #7
+In our original reply, we gave references to less "binary" frameworks for argumentation but did not include them in the revised paper.
+In this revision, we have included a brief discussion on these frameworks (last paragraph of 5.2), and we also refer to work that discusses the relation between "binary" argumentation and argumentation in the real world in more detail (Rahwan et al.
+2010, also see the added footnote X).
 
-===== REPLY #7
-
-===== COMMENT #8
-Regarding terminology, I wonder about "valid" GRL models here, which seem to mean that certain criteria are fulfilled.
-Validation would mean that a goal model makes sense in the real world as having the "right" goals and relationships between them.
-
-===== REPLY #8
-We agree that the term “valid” is an overloaded term that has different meanings in different areas of research.
-In our case, we indeed want to express that certain criteria are fulfilled.
-In order to avoid confusion, we have added a footnote to Definition 7 (Valid GRL Model), clarifying what exactly we mean with a "valid model".
-
-===== COMMENT #9
-"Verification" in 6.2 appears to stand for evaluation. This cannot be "verified".
-
-===== REPLY #9
-We have renamed “verification” to “demonstration” and slightly reworded the sentence.
-
-===== COMMENT #10
-I also do not think that Figure 16 shows a "methodology". It does not even look like a "method".
-
-===== REPLY #10
-We agree that the terminology is inappropriate.
-We have renamed the methodology to "The RationalGRL development process", since it defines the steps on how to create a RationalGRL model.
-We have renamed this throughout the article.
-
-===== COMMENT #11
-please reconsider the very first sentences in the Abstract and Introduction each, especially the latter statement on RE
-
-===== REPLY #11
-We agree with the reviewer that the two mentioned sentences are confusing. We have improved this as follows:
-- We have rewritten the first sentence of the abstract as follows: "Goal-oriented requirements modeling approaches aim at capturing intentions of various stakeholders involved in the development of an information system and refining them into low level goals and tasks."
-- We have deleted the first sentence of the introduction, since we feel that it is not necessary to explain the definition of Requirements Engineering in the Requirements Engineering Journal.
-
-===== COMMENT #12
-Finally, you should run a spell checker, have a careful proof reading of grammar, and make sure that there is no unresolved reference to a figure in a submitted manuscript.
-
-===== REPLY #11
-We’ve ran a spell-checker and proof-read the paper carefully. There was one unresolved reference to a figure, which a corrected. Thank you for pointing this out.
-
-######## REFERENCES ########
-
-[1] Estrada, H., Rebollar, A.
-M., Pastor, O., & Mylopoulos, J.
-(2006, June).
-An empirical evaluation of the i* framework in a model-based software generation environment.
-In International Conference on Advanced Information Systems Engineering (pp.
-513-527).
-Springer, Berlin, Heidelberg.
-[2] A. Hunter and M. Thimm (2017) Probabilistic Reasoning with Abstract Argumentation Frameworks, Journal of Artificial Intelligence Research, 59:565--611. 
-[3] Grossi, D., & Modgil, S. (2015, July). On the Graded Acceptability of Arguments. In IJCAI (pp. 868-874).
-[4] van Eemeren, F.H., & Verheij, B. (2017). Argumentation Theory in Formal and Computational Perspective. IFCoLog Journal of Logics and Their Applications 4 (8), 2099-2181.
-
-[5] Falappa, Marcelo Alejandro, Gabriele Kern-Isberner, and Guillermo Ricardo Simari.
-"Belief revision and argumentation theory." Argumentation in artificial intelligence.
-Springer, Boston, MA, 2009.
-341-360.
-[6] Dix, Jürgen, et al. "Belief change and argumentation in multi-agent scenarios (Dagstuhl seminar 13231)." Dagstuhl Reports. Vol. 3. No. 6. Schloss Dagstuhl-Leibniz-Zentrum fuer Informatik, 2013.
-[7] Paglieri, Fabio, and Cristiano Castelfranchi.
-"Revising beliefs through arguments: Bridging the gap between argumentation and belief revision in MAS." International Workshop on Argumentation in Multi-Agent Systems.
-Springer, Berlin, Heidelberg, 2004.
+===== ALL OTHER ORIGINAL COMMENTS WERE SMALLER COMMENTS (STYLE, TYPO, NAMING) WHICH WE ADDRESSED IN THE ORIGINAL REVISION

### 2018-09-15 discussion.tex
Reviewer comments

@@
 
-\paragraph{Design Rationale} Argumentation in software design has been the subject of the research on so-called \emph{design rationale} (DR)~\cite{shum2006hypermedia}, an explicit documentation of the reasons behind decisions made when designing a system or software architecture.
+\paragraph{Design Rationale} Argumentation in software design has been the subject of the research on so-called \emph{design rationale} (DR), an explicit documentation of the reasons behind decisions made when designing a system or software architecture.
+Starting already in the 90's with influential issue-driven design methods such as gIBIS and QOC~\cite{shum2006hypermedia},
  DR looks at issues, options and arguments for and against these options in the design of, for example, a software system.
@@
 It is in this diagramming functionality that the link with argument diagrams from philosophy, law and AI~\cite{reed2004araucaria,gordon2007visualizing} has been made, where argument diagrams represent reasoning from premises to conclusions.
-More recent research on DR moves away from the idea that all decision have to be explicitly diagrammed and focuses more on empirically investigating how critical reflection can help when designing \cite{razavian2016two,SchriekEtal2016}, or which parts of the design process are best explicitly documented \cite{falessi2013value}.
+One of the problems of such methods and tools is the cognitive overload that results from having to learn and use such tools at the same time as having to discuss and think about complex software designs [5].
+Furthermore, not much of the research on design reasoning has empirically investigated which techniques and how much information designers use in design reasoning, which information is missing, and how the use of design reasoning techniques affects design discussions and eventually the quality of design discourse.
+More recent research on DR therefore moves away from the idea that all decision have to be explicitly diagrammed and focuses more on empirically investigating how critical reflection can help when designing \cite{razavian2016two,TangEtal2018}, or which parts of the design process are best explicitly documented \cite{falessi2013value}.
 
@@
 The RationalGRL framework essentially incorporates the core ideas from DR into goal-oriented requirements engineering by explicitly including arguments for and against the various options into the goal model.
-Note that we explicitly connect to the extensive literature on goal modeling, and in this respect may seem to go against the trend in DR research to diagram decisions and processes.
-However, we note that RationalGRL also includes a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system, which connects with the more research ideas on DR \cite{razavian2016two,SchriekEtal2016}.
+In this respect, we may seem to go against the trend in DR research to focus less on diagramming and more on guidelines for design processes.
+There are two reasons for this.
+First, as opposed to DR diagramming, goal modeling is still very much an active area of research \cite{GonccalvesEtal2018} with clear practical applications \cite{NeaceEtal2018}.
+Second, we note that RationalGRL also includes a development process and critical questions that encourage reflection when thinking about the possible goals and functionality of a system, which connects with the more research ideas on DR \cite{razavian2016two,TangEtal2018}.
 
@@
 
-More explicit rationales for goal models are included in the original GRL specification \cite{} in the form of \emph{belief} elements, which can be attached to other intentional elements (goals, tasks) using the different links (positive contribution, negative contribution) of GRL.
+More explicit rationales for goal models are included in the original GRL specification \cite{Amyot:2010:EGM:1841349.1841356} in the form of \emph{belief} elements, which can be attached to other intentional elements (goals, tasks) using the different links (positive contribution, negative contribution) of GRL.
 The content of a belief is a textual description, which can explain, for instance, the rationale for including a goal in a model.
 In this regard, they seem like arguments or reasons for the different elements in a goal model.
-However, in the literature the precise usage and semantics of belief elements in GRL remain unclear: according to the metamodel \cite{•} they can be connected to other intentional elements (including beliefs) through any type of link, but in the jUCMNav tool there is a separate \emph{belief link} for connecting beliefs to goals and tasks only, and the beliefs are not used in any of the algorithms to compute the satisfiability of goals and tasks.
+However, in the literature the precise usage and semantics of belief elements in GRL remain unclear: according to the metamodel \cite{Amyot:2010:EGM:1841349.1841356} they can be connected to other intentional elements (including beliefs) through any type of link, but in the jUCMNav tool there is a separate \emph{belief link} for connecting beliefs to goals and tasks only, and the beliefs are not used in any of the algorithms to compute the satisfiability of goals and tasks.
 We are also not aware of many examples of the usage of belief elements in the literature.
@@
 One idea here is to have stakeholders play a card game (similar to the well-known planning poker) with cards that contain critical, reflective questions (cf.
-\cite{SchriekEtal2016})
+\cite{TangEtal2018}).
 

### 2018-09-15 formalframework.tex
Reviewer comments

@@
 In this example, there are five admissible sets: $\{A'_{5}\}$, $\{A_{14}\}$, $\{A'_{5},A_{14}\}$, $\{A_4, A_{14}\}$ and $\{A_4, A'_5, A_{14}\}$.
-In the admissible sets that contain both $A_{4}$ and $A_{14}$, we say that $A_{14}$ defends $A_4$ against its attacker $A_{13}$.
+In the admissible sets that contain both $A_{4}$ and $A_{14}$, we say that $A_{14}$ defends $A_4$ against its attacker $A_{13}$\footnote{In the argumentation literature, argument $A_{14}$ is often said to \emph{reinstate} argument $A_{4}$}.
 Sets containing both $A_{5}$ and $A'_{5}$, or both $A_{13}$ and either $A_4$ or $A_{14}$ are not conflict free.
@@
 Returning to our larger example from Figure~\ref{fig:example-small3}, we can see that arguments $A_5$, $A_7$, $A_{13}$ and $A_{15}$ are rejected and all the other arguments are acceptable.
+
+Note that argument acceptability is essentially a three-valued (or binary if one does not count \emph{undecided}) concept.
+There are many more fine-grained ways of defining the acceptability of arguments using, for example, probabilities \cite{li2011probabilistic} or counting the number of attacks \cite{grossi2015graded}.
+However, the ``basic'' argumentation frameworks, and in particular the  have been shown to be close to how humans reason \cite{rahwan2010behavioral}.
 

### 2018-09-16 formalframework.tex
small changes + reference inference engine

@@
 
-We use formal logic (propositional logic) to specify our framework, and in order to develop our tool we have implemented an algorithm to compute argumentation extensions based on this formalization.
-However, we do not use a dedicated theorem prover or a logical inference engine.
-The reason is purely practical, it is our aim to develop a tool that practitioners can use in their browser, and it does not seem necessary nor useful to use an external theorem prover for this.
-Instead, in this section we propose algorithms based on our logical specification, which form the basis of our prototype implementation.
+We use formal logic (propositional logic) to specify our framework, and we have implemented an algorithm to compute argumentation extensions based on this formalization.
+However, we do not use a dedicated logical inference engine for argumentation (see e.g.
+\cite{hecham2018first}).
+The reason is that, in the current paper, our is to aim to develop a tool that users can use in their browser, and that hence the focus is not on developing a complete logical inference engine for, for example, automatically deriving goals from tasks.
+Instead, we propose algorithms based on our logical specification, which form the basis of our prototype implementation.
 

### 2018-09-16 validation.tex
last todo's

@@
 \centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l}
+\begin{tabularx}{0.8\textwidth}{l|l|l|l|l|l||l|l|l}
 & very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p-value \\
@@
 \end{tabularx}
-\caption{Participant ratings and statistical results for the usability of the RationalGRL tool\todo{Floris}{Marc}{Fix table width}}
+\caption{Participant ratings and statistical results for the usability of the RationalGRL tool}
 \label{table:survey:table1}
@@
 They produced RationalGRL models containing on average 3 arguments (in addition  to goals, tasks, etc.) from the short transcript.
-Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.
-The full results of the survey can be downloaded from \url{www.rationalgrl.com}, on the page ``Empirical study''.
-\todo{Floris}{Marc}{Staan de results op de webpagina?}
+Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.\footnote{The full results of the survey can be downloaded from \url{www.rationalgrl.com}, on the page ``Empirical study''.}
 

### 2018-09-17 Response to reviewers2.txt
response to reviewers

@@
+Thank you for your considerations.
+
+Reviewer 1 proposes to accept the paper.
+Reviewer 2 argues that the new section 7 should be revised - we agree with the second reviewer that the style and wording of the section was too informal and extended and sharpened the discussion (see also reply 2 below).
+Furthermore, reviewer 2 argues that they would like to see more concrete revisions based on some of the points raised in their first review.
+We have extended the paper, and below we discuss these points from the original review in more detail.
+
 <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
@@
 <<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<<>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>>
+
+===== ORIGINAL COMMENT #1
+I think that the relatively simple argument structures are appealing in the sense that the overhead for practical application may be tolerable.
+However, this remains to be shown yet in an industrial environment.
+
+===== REPLY #1
+Here, we reiterate our argument from the original response to reviewers.
+
+"We agree with the reviewer that a real validation of the approach is a necessary next step.
+However, we would like to point the reviewer to the case study in Section 3, in which 153 pages of transcripts of requirements discussions were used (see also the original materials at www.rationalgrl.com).
+We do agree that further validation of the tool and method is necessary, preferably in an industrial setting.
+However, such an industrial strength case study is a significant amount of work (see for example Estrada et al.
+(2006 "An empirical evaluation of the i* framework in a model-based software generation environment".
+They took 9 months to conduct the case study.), and would most likely yield enough new insights for a separate paper.
+
+Since the main focus of our article is to explain the novel RationalGRL framework in detail, we feel that adding such an extensive study to the current article may be too much content.
+Still, we agree that having no validation at all is limiting, so we have performed a small-scale empirical study in order to provide an initial validation of our framework.
+This empirical study was performed amongst 16 participants, each with a computer science background."
 
@@
 
-===== ALL OTHER ORIGINAL COMMENTS WERE SMALLER COMMENTS (STYLE, TYPO, NAMING) WHICH WE ADDRESSED IN THE ORIGINAL REVISION
+===== ALL OTHER ORIGINAL COMMENTS WERE SMALLER COMMENTS (STYLE, TYPO, WORDING) WHICH WE ADDRESSED IN THE ORIGINAL REVISION

### 2018-09-17 background.tex
Section 1-5

@@
 In this section, we first introduce our running example, after which we give a brief overview of the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework\footnote{For simplicity's sake, we consider a slightly cut-down version of GRL.
-All the core elements of GRL are included but, for example, we have included only one level of contribution strength, where GRL contains different levels of positive and negative contribution.
+All the core elements of GRL are included but, for example, we have included only one level of contribution strength, whereas GRL contains different levels of positive and negative contribution.
 This simplification does not in any way influence the case study.
-Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements would be trivial.}.
-Next, we introduce argumentation; we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
+Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements is trivial.}.
+Next, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
 %This will be our starting point in the next section.
@@
 The client of the project is Professor E, who teaches civil engineering courses at an American university.
-In order for the professor to teach students the various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
+In order for the professor to teach students various theories concerning traffic (such as queuing theory), traffic simulator software needs to be developed in which students can create visual maps of an area, regulate traffic, and so forth.
 Tang et al.
@@
 We used these transcripts for an extensive case study on the basis of which we develop our RationalGRL framework.
-Furthermore, we also use the traffic simulator case as running examples throughout this paper.
+Furthermore, we also use the traffic simulator case as a running example throughout this paper.
 
@@
 A GRL model is a connected graph of intentional elements that optionally are part of the actors.
-The GRL elements and relationships used in this paper are shown in Figure~\ref{fig:grl_legend}.
+The GRL intentional elements (IEs) and relationships used in this paper are shown in Figure~\ref{fig:grl_legend}.
 
@@
 \includegraphics{img/grl_legend.pdf}
-\caption{Basic elements and relationships of GRL}
+\caption{Basic Elements and Relationships of GRL}
 \label{fig:grl_legend}
@@
 \includegraphics[width=\columnwidth]{img/Example1.pdf}
-\caption{Partial GRL Model of the traffic simulator example}
+\caption{Partial GRL Model of the Traffic Simulator Example}
 \label{fig:example-small}
@@
 Task \emph{Create new cars} has a positive contribution to the softgoal \emph{Realistic simulation}, and a negative contribution to the softgoal \emph{Simple design}.
-Note that the full GRL specification considers different levels of positive and negative contribution values (both quantitative and qualitative), which are not directly relevant for current purposes.
-Dependency links are relationships between IEs, which can model dependencies between actors.
+Note that the full GRL specification considers different levels of positive and negative contribution values (both quantitative and qualitative), which are not directly relevant for our current purpose.
+Dependency links are relationships between intentional elements, which can model dependencies between actors.
 Here, the goal \emph{Generate cars} depends on the resource \emph{Car objects}.
@@
 However, the rationales and arguments behind certain intentional elements are not shown in the GRL model.
-Some of the questions that might be interesting to know about are the following:
+Some of the questions that might be interesting to reason about are the following:
 
@@
 
-Reasoning about which goals to pursue and actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy and artificial intelligence.
+Reasoning about which goals to pursue and which actions to take is often referred to as \emph{practical reasoning}, and has been studied extensively in philosophy and artificial intelligence.
 One approach is to capture practical reasoning with argument schemes~\cite{walton1990}.
 Applying an argument scheme results in an argument in favor of, for example, taking an action.
-This argument can then be tested with critical questions about, for instance, whether the action is possible given the situation, and a negative answer to such a question leads to a counterargument to the original argument for the action.
+This argument can then be tested with critical questions about, for instance, whether the action is possible given the situation.
+A negative answer to such a question leads to a counterargument to the original argument for the action.
 
@@
 \includegraphics[]{img/fig_AF1.pdf}
-\caption{PRAS arguments and attacks in the traffic simulation example.}
+\caption{PRAS aArguments and Attacks in the Traffic Simulation Example}
 \label{fig:pras:example}
@@
 Argument A2 proposes an alternative way of realizing the same goal \emph{Generate cars} with respect to argument A1 and vice versa (cf.
-CQ2), so A1 and A2 mutually attack each other, denoted by the arrows between A1 and A2.
-Argument A3 says that \emph{Create new cars} has a negative effect on \emph{Static Simulation}, so A3 attacks A2, as it points to a negative side-effect of \emph{Create new cars} (CQ3).
+CQ2).
+Thus, A1 and A2 mutually attack each other, denoted by the arrows between A1 and A2.
+Argument A3 says that \emph{Create new cars} has a negative effect on \emph{Static Simulation}, therefor, A3 attacks A2, as it points to a negative side-effect of \emph{Create new cars} (CQ3).
 The intuition here is that an argument is acceptable if any argument that attacks it is itself rejected.
@@
 In fact, our case study (Section \ref{sect:gmas}) shows that when discussing requirements, people very often do not structure their reasoning nicely in the way that PRAS presents it.
-That is, you do not see the discussants setting up an argument ``We have goal $G$, $A$ realizes $G$ \emph{Therefore} we should perform $A$''.
+That is, it is rarely the case that the discussants setting up an argument ``We have goal $G$, $A$ realizes $G$ \emph{Therefore} we should perform $A$''.
 A typical discussion is much more unstructured, as is clear from the transcript excerpts in Appendix~\ref{sect:transcripts:excerpts}.

### 2018-09-17 casestudy.tex
Small changes references and finished validation.tex
Section 1-5

@@
 We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS.
-In this section we present our case study.
-All original transcripts, codings, and models are available in our online repository, which can be found at:
+In this section, we present our case study.
+All original transcripts, coding, and models are available in our online repository, which can be found at:
 
@@
 
-In order to obtain actual requirements discussions, we turned to a recent series of experiments by Schriek et al.
-\cite{SchriekEtal2016}.
+In order to obtain actual requirements discussions, we turned to a recent series of experiments by Tang et al.
+\cite{TangEtal2018}.
 In these experiments, 12 groups of two or three students in a Software Architecture course at MSc level were given the traffic simulator assignment (Appendix~\ref{sect:designprompt}).
@@
 \end{tabularx}
-\caption{Occurrences of argument schemes and critical questions in the transcripts.}
+\caption{Occurrences of Argument Schemes and Critical Questions in the Transcripts.}
 \label{table:transcripts:results:argumentschemes}
@@
 
-Before we started coding the transcripts, came up with an initial list of 11 argument schemes (AS0-AS10 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about the goal model containing the requirements of the system.
+Before we started coding the transcripts, we came up with an initial list of 11 argument schemes (AS0-AS10 in Table~\ref{table:transcripts:results:argumentschemes}), representing \emph{claims} about a goal model containing the requirements of the system.
 We used no particular method to develop this initial list besides our own intuition about which part of a goal model would be likely to be discussed.
 
-AS0 to AS4 are schemes that concern a single element of the goal model.
+AS0 to AS4 are schemes that concern a single element of a goal model.
 For example, AS0 represents the claim `$a$ is a relevant actor for the system', and AS3 represents the claim `$G$ is a goal for the system'.
-AS5 to AS10 are claims about the links between GRL elements.
+AS5 to AS10 are claims about the links between GRL intentional elements.
 Our initial list also contained 18 critical questions, inspired by the questions associated with the original Practical Reasoning Argumentation Scheme \cite{atkinson2007}.
@@
 Using the initial list of arguments and critical questions, we coded three transcripts of requirements discussions.
-The codings were performed by one author and subsequently checked by another author.
-As the transcripts contain spoken language, the codings involved some interpretation.
+The coding was performed by one author and subsequently checked by other authors.
+As the transcripts contain spoken language, the coding involved some interpretation.
 For example, the students almost never literally say `actor $a$ has task $T$'.
@@
 During the coding, new argument schemes and critical questions were added to the list.
-For example, we found that the discussants often talk about tasks decomposing into sub-tasks, so we added AS10 and CQ10a.
+For example, we found that the discussants often talk about tasks decomposing into sub-tasks, therefore, we added AS10 and CQ10a.
 Furthermore, since there were many discussions on the relevance and the clarity of the names of elements, two generic critical questions CQ12 and CQ13 were added.
@@
 An example of replacement is given in Table~\ref{table:transcript:decomposition} (Appendix~\ref{sect:transcripts:excerpts}): what used to be an AND-decomposition is changed into an OR-decomposition.
-We will further discuss these operations in Section~\ref{sect:overview} with various examples, and w formalize the operations in Section~\ref{sect:formalframework} with algorithms.
+We will further discuss these operations in Section~\ref{sect:overview} with various examples, and we formalize the operations in Section~\ref{sect:formalframework} with algorithms.

### 2018-09-17 formalframework.tex
Section 1-5

@@
 
-In Section~\ref{sect:overview}, we have shown through a language definition and informal examples from our case study that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
-A formal rendition of this traceability will be presented in this section, in which we present a logical formalization of RationalGRL.
+In Section~\ref{sect:overview}, we have shown through a language definition and informal examples from our case study that it is possible to trace elements of a goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on a goal model, and vice versa (\textbf{requirement 3}).
+A formal rendition of this traceability will be presented in this section, in which we present a logical formalization of the RationalGRL framework.
 This is done for multiple reasons: (i) Most approaches in formal argumentation (cf.
@@
 \cite{hecham2018first}).
-The reason is that, in the current paper, our is to aim to develop a tool that users can use in their browser, and that hence the focus is not on developing a complete logical inference engine for, for example, automatically deriving goals from tasks.
+The reason is that, in the current paper, our is to aim to develop a tool that users can use in their browser, and  hence the focus is not on developing a complete logical inference engine for, for example, automatically deriving goals from tasks.
 Instead, we propose algorithms based on our logical specification, which form the basis of our prototype implementation.
@@
 In Sections \ref{sect:formalframework:grl} and \ref{sect:formalframework:rationalgrl} we formalize a static representation of our framework based on the metamodel (Figure~\ref{fig:metamodel}).
-We first provide a formal specification of a GRL model in Section~\ref{sect:formalframework:grl}, and we extend this with arguments and attack links in the Section~\ref{sect:formalframework:rationalgrl}, hereby obtaining a formal specification of a RationalGRL model.
+We first provide a formal specification of a GRL model in Section~\ref{sect:formalframework:grl}, and we extend it with arguments and attack links in the Section~\ref{sect:formalframework:rationalgrl}, hereby obtaining a formal specification of a RationalGRL model.
 In Section~\ref{sect:formalframework:translation} we then present algorithms in order to translate a GRL model into a RationalGRL model, and vice versa.
-Finally, in Section~\ref{sect:algorithms} we develop algorithms for instantiating argument schemes and answering critical questions and thus formally capture the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations of which examples were previously given in Section~\ref{sect:overview:examples}.
+Finally, in Section~\ref{sect:algorithms}, we develop algorithms for instantiating argument schemes and answering critical questions and thus formally capture the \textsf{INTRO}, \textsf{REPLACE} and \textsf{DISABLE} operations of which examples were previously given in Section~\ref{sect:overview:examples}.
 
@@
 
-\begin{definition}[General definitions]
+\begin{definition}[General Definitions]
 \label{def:set-definitions}
@@
 This is in line with the GRL metamodel.
-However, in the following we sometimes leave out the $decomptype$ if the IE is not decomposed to simplify notation and improve readability.
+However, in the following, we sometimes leave out the $decomptype$ if the IE is not decomposed to simplify the notation and improve readability.
 
-Throughout this section we use the example in Figure~\ref{fig:example-small}. This example contains various IEs which can be formalized using Definition~\ref{def:ie}, for example, 
+Throughout this section, we use the example in Figure~\ref{fig:example-small}. This example contains various IEs which can be formalized using Definition~\ref{def:ie}, for example, 
 $(1, Resource, \text{Car objects})$, $(2, Softgoal,$ Realistic simulation$)$ and $(5, Goal,$ Generate cars$, XOR)$.
@@
 
-\subsection{Formal specification of RationalGRL}
+\subsection{Formal Specification of RationalGRL}
 \label{sect:formalframework:rationalgrl}
@@
 \includegraphics[width=\columnwidth]{img/Example1-new-attack.pdf}
-\caption{Example RationalGRL model (extension of Fig.~\ref{fig:example-small})}
+\caption{Example RationalGRL Model (Extension of Fig.~\ref{fig:example-small})}
 \label{fig:example-small3}
@@
 As an example, take the arguments $A_5$ for the goal `Generate cars' ($AND$ decomposition), $A'_5$ for the goal `Generate cars' ($XOR$ decomposition), $A_4$ for the softgoal `Easy to use' and the generic argument $A_{13}$ (`Redundant') against the softgoal (`Easy to use').
-Given these arguments there are two attack links (Figure~\ref{fig:example-small3}), namely $(A'_{5},A_{5})$ and $(A_{13}, A_{4})$.
+Given these arguments, there are two attack links (Figure~\ref{fig:example-small3}), namely $(A'_{5},A_{5})$ and $(A_{13}, A_{4})$.
 
@@
 
-Let us explain these definitions using the example argumentation framework in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of our RationalGRL model from Figure~\ref{fig:example-small3}, containing only arguments $A_4, A_{5},A'_{5},A_{13}$, and $A_{14}$.
+Let us explain these definitions using the example, argumentation framework, in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of our RationalGRL model from Figure~\ref{fig:example-small3}, containing only arguments $A_4, A_{5},A'_{5},A_{13}$, and $A_{14}$.
 In this example, there are five admissible sets: $\{A'_{5}\}$, $\{A_{14}\}$, $\{A'_{5},A_{14}\}$, $\{A_4, A_{14}\}$ and $\{A_4, A'_5, A_{14}\}$.
@@
 \includegraphics[]{img/example-mutual-attack.pdf}
-\caption{Example of a cyclic argumentation framework.}
+\caption{Example of a Cyclic Argumentation Framework.}
 \label{fig:goalmodeling:arg3}
@@
 
-Now that we have formalized both a GRL model and a RationalGRL model, we present algorithms to translate between these two models.
+Now that we have formalized both GRL and RationalGRL models, we present algorithms to translate between these two models.
 Both of these two translation algorithms are straightforward, which is result of the fact that the two models are formalized in a very similar way.
@@
 \paragraph{GRL to RationalGRL Translation} We start with the translation algorithm from GRL to RationalGRL, which is shown in Algorithm~\ref{alg:translation:to-rationalgrl}.
-The translation algorithm takes a GRL model $GR=(IE, Act, R_{ActIE}, Link)$ (Definition~\ref{def:grl-model}) as input.
+The translation algorithm takes a GRL model $GR=(IE, Act, R_{ActIE}, Link)$ (Definition~\ref{def:grl-model}) as an input.
 On line~\ref{alg:translation:to-rationalgrl:args}, it collects all the elements of the GRL model into a set $Arg$, which represents the set of arguments of the RationalGRL model.
@@
 
-\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Definition~\ref{def:valid-grl-model}), we have not done so for a RationalGRL model yet.
+\paragraph{Valid RationalGRL model} While we have defined a notion of a \emph{valid GRL model} (Definition~\ref{def:valid-grl-model}), we have not yet done the same for a RationalGRL model.
 We define a RationalGRL model as valid if and only if the RationalGRL to GRL translation results in a valid GRL model.
@@
 
-\subsection{Algorithms for argument schemes and critical questions}
+\subsection{Algorithms for Argument Schemes and Critical Questions}
 \label{sect:algorithms}
@@
 We have now formalized a \emph{static} representation of the RationalGRL framework.
-In this section we formalize algorithms for applying argument schemes and critical questions.
+In this section, we formalize algorithms for applying argument schemes and critical questions.
 These algorithms are applied to RationalGRL models (Definition~\ref{def:rationalgrl-model}) and produce new arguments and attack relations.
@@
 The \textsf{REPLACE} operation can be seen as a combination of \textsf{INTRO} and \textsf{DISABLE}: a new argument corresponding to a GRL element is added and this new argument attacks a previous version of the element or link.
-We will discuss all three operations in separate sections below.
+We will discuss all three operations in separate subsections below.
 
@@
 
-\subsubsection{INTRO algorithms}
+\subsubsection{INTRO Algorithms}
 \label{sect:formalframework:intro}
@@
 
-We have slightly reworded critical question AS5 in Algoritm~\ref{alg:add-decomp}.
+We have slightly reworded critical question AS5 in Algorithm~\ref{alg:add-decomp}.
 We assume that a goal $G$ exists already with identifier $i$, and that some new task with name $T$ is a decomposition of $G$.
@@
 
-As an example, suppose we are adding the decomposition of goal ``Generate cars'', expressed as argument $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, into task ``Keep same cars'', and suppose we have the following argument for the goal: .
+As an example, suppose we add the decomposition of goal ``Generate cars'', expressed as argument $A_5 = (5,$ $Goal,$ Generate cars$, AND)$, into task ``Keep same cars'', and suppose we have the following argument for the goal: .
 Adding the decomposition can be formalized as executing $AS_5(5, \text{Keep same cars})$, and results in two new arguments: $A_7 = (7, Task, \text{Keep same cars})$ and $A_{12} = (12, Decomp, 7, 5)$, which are both added to the set of arguments $Arg$.
@@
 
-\subsubsection{DISABLE algorithms}
+\subsubsection{DISABLE Algorithms}
 \label{sect:formalframework:disable}
@@
 
-\subsubsection{REPLACE algorithms}
+\subsubsection{REPLACE Algorithms}
 \label{sect:formalframework:replace}
@@
 The other \emph{REPLACE} algorithms are similar to Algorithm~\ref{alg:replace1}, which can be used directly for CQ10c.
-For CQ12 we should make a small modification: instead of replacing the decomposition type of the IE, we replace its name.
-Since this is a very minor modification we don't show it here.
+For CQ12, we should make a small modification: instead of replacing the decomposition type of the IE, we replace its name.
+Since this is a very minor modification we do not show it here.

### 2018-09-17 introduction.tex
Section 1-5

@@
 Among the ``early-phase'' requirements engineering activities are those that consider how the intended system should meet organizational goals, why it is needed, what alternatives may exist, what the implications of the alternatives are for different stakeholders, and how the interests and concerns of stakeholders might be addressed~\cite{yu1997towards}.
-These activities fall under the umbrella of goal modeling.
+These activities fall under the umbrella of goal modeling approaches.
 There are a large number of established RE methods using goal models in the early stage of requirements analysis (overviews can be found in~\cite{kavakliL05,van2001goal}).
@@
 
-The first challenge for the goal modeling phase, particularly in large projects, is related to its dynamic nature: goal models continuously change and evolve.
+The first challenge for goal modeling approaches, particularly in large projects, is related to its dynamic nature: goal models continuously change and evolve.
 Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
@@
 \item 
-Using these traceability links, it must be possible to compute the effect of changes in the underlying arguments about the goal model, and vice versa.
+Using these traceability links, it must be possible to compute the effect of changes in the underlying arguments on the goal model, and vice versa.
 \item 
@@
 \item 
-The framework must have tool support.
+The framework must have a tool support.
 \end{enumerate}
@@
 Following from our previous work~\cite{vanzee-etal:renext2015,vanZee-etal:er2016}, we develop the \emph{RationalGRL} framework.
-RationalGRL extends a slightly simplified existing approach for goal modeling, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}.
+RationalGRL extends a slightly simplified version of an existing goal modeling approach, the Goal-oriented Requirements Language (GRL) \cite{Amyot:2010:EGM:1841349.1841356}, with techniques from argumentation theory, namely \emph{argument schemes} (or argumentation schemes~\cite{walton-etal2008}) and argumentation semantics \cite{Dung1995}.
 Argument schemes are reusable patterns of reasoning that capture the typical ways in which humans argue and reason.
@@
 These critical questions can point to counterarguments, and the acceptability of sets of arguments can be computed using argumentation semantics.
-Thus computational argumentation based on schemes can guide users in systematically deriving conclusions and making assumptions explicit~\cite{bexEtal2003,murukannaiah2015}.
+Thus, computational argumentation based on schemes can guide users in systematically deriving conclusions and making assumptions explicit~\cite{bexEtal2003,murukannaiah2015}.
 
@@
 In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework to which we can apply argumentation semantics~\cite{Dung1995}.
-We, then, formally capture the links between argumentation and goal modeling as a set of algorithms for applying argument schemes and critical questions about goal models (requirement 3).
+We, then, formally capture the links between argumentation and goal models as a set of algorithms for applying argument schemes and critical questions about goal models (requirement 3).
 
@@
 Section~\ref{sect:methodology+tool} discusses the RationalGRL development process and explains various features of the tool we developed.
-Section~\ref{sect:validation} contains a brief validation of RationalGRL.
-We conducted a small user study (16 participants), where we asked the participants to model a small example in the RationalGRL, and to answer some questions on the usability and usefulness RationalGRL.
+Section~\ref{sect:validation} contains a brief validation of the RationalGRL framework.
+We conducted a user study with 16 participants, where we asked the participants to model a small example in RationalGRL, and to answer some questions on the usability and usefulness of RationalGRL.
 Finally, Section~\ref{sect:discussion} covers related work, future work, and the conclusion.

### 2018-09-17 discussion.tex
Section 6-7-8

@@
 
-\subsection{Related work}
+\subsection{Related Work}
 \label{sect:discussions:relatedwork}
@@
 Another recent example of the use of arguments in goal-oriented requirements engineering is the work by Murukannaiah \emph{et al.}~\cite{murukannaiah2015}, who propose Arg-ACH, where the beliefs underlying conflicting goals can be made explicit using argumentation.
-Murukannaiah \emph{et al.} start with the basic technique of Analysis of Competing Hypotheses (ACH), where for conflicting goals the beliefs that are consistent and inconsistent with these goals are included in a matrix and counted.
+Murukannaiah \emph{et al.} start with the basic technique of Analysis of Competing Hypotheses (ACH), where for conflicting goals, the beliefs that are consistent and inconsistent with these goals are included in a matrix and counted.
 They then extend this technique into Arg-ACH: instead of just indicating whether a belief is consistent or inconsistent with a goal, each belief becomes an argument for or against the goal, which is then diagrammed using the Carneades tool \cite{gordon2007visualizing}.
@@
 We are also not aware of many examples of the usage of belief elements in the literature.
-So what it seems like is that GRL belief elements can be understood as metadata that can be attached to an element
+Thus, what it seems like is that GRL belief elements can be understood as metadata that can be attached to an element.
 
 Ambiguities notwithstanding, there are further advantages of using RationalGRL over GRL with belief elements.
-For example, in none of the GRL specifications is it possible to attach beliefs to links between other elements so it is not possible, for example, to provide a rationale for why some tasks positively contributes to a goal.
+For example, in none of the GRL specifications, is it possible to attach beliefs to links between other elements thus it is not possible, for example, to provide a rationale for why some tasks positively contributes to a goal.
 Furthermore, the RationalGRL arguments follow clear schemes with associated critical questions explicitly tailored towards reasoning about goals and actions or tasks, while beliefs in GRL offer no further guidance as to what kinds of rationales can be given for elements in a goal model.
@@
 Or if there is a dependency link between $G_1$ and $G_2$, then $G_1$ should be in every extension $G_2$ is in and vice versa.
-AND and OR decomposition are also modeled as such, that is, if $G_3$ AND-decomposes into $G_1$ and $G_2$ then $G_1$ and $G_2$ should be in every extension $G_3$ is in, and if $G_3$ OR-decomposes into $G_1$ and $G_2$ then either $G_1$ or $G_2$ should be in every extension $G_3$ is in.
+AND- and OR-decomposition are also modeled as such, that is, if $G_3$ AND-decomposes into $G_1$ and $G_2$ then $G_1$ and $G_2$ should be in every extension $G_3$ is in, and if $G_3$ OR-decomposes into $G_1$ and $G_2$ then either $G_1$ or $G_2$ should be in every extension $G_3$ is in.
 
@@
 
-Additionally, from the user evaluation of the tool it followed that in order to implement the tool in realistic settings, other ways may have to be found to incorporate arguments in the goal modeling process.
+Additionally, from the user evaluation of the tool, it followed that in order to implement the tool in realistic settings, other ways may have to be found to incorporate arguments in the goal modeling process.
 One idea is to use the tool for collaborative on-line goal modeling, similar to Github: since the renaming (i.e.

### 2018-09-17 validation.tex
Section 6-7-8

@@
-\section{RationalGRL evaluation experiment}
+\section{RationalGRL Evaluation Experiment}
 \label{sect:validation}
@@
 In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a user evaluation with 16 users.
-The objective of this evaluation was to determine whether the users found the additions of RationalGRL (arguments, argumentation semantics, critical questions) useful compared to standard GRL, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to standard GRL.
+The objective of this evaluation was to determine whether the users found the additions of RationalGRL (arguments, argumentation semantics, critical questions) useful compared to the standard GRL, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to the standard GRL.
 Regarding the comparison of RationalGRL to GRL, the following two broad hypotheses follow from our main points in this paper:
@@
 \begin{itemize}
-\item[H1] The additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
+\item[H1] The additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modeling in GRL.
 \item[H2] The additions of RationalGRL make it easier to express, determine the effect of and communicate to other stakeholders one's opinions and beliefs about a goal model.
@@
 Note that we did not aim to conduct a usability study for the RationalGRL tool.
-Rather, we were interested to know if the tool was clearly very easy or difficult to use, as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive towards a particular modelling language, whereas irritations about a bad tool will lead to a more negative disposition.
+Rather, we were interested to know if the tool was clearly very easy or difficult to use, as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive towards a particular modeling language, whereas irritations about a bad tool will lead to a more negative disposition.
 This leads to the following general hypothesis:
@@
 
-The idea was to have our participants perform a small modelling task, and then ask them what they thought of RationalGRL and the tool. Thus, our experiment consisted of three parts:
+The idea was to have our participants perform a small modeling task, and then ask them what they thought of RationalGRL and the tool. Thus, our experiment consisted of three parts:
 
@@
 
-The instructions for these three parts are contained in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (part 1) starts with a general explanation of GRL and standard goal modelling similar to the explanation provided in section \ref{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
+The instructions for these three parts are provided in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (part 1) starts with a general explanation of GRL and standard goal modeling similar to the explanation provided in Section \ref{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
 
@@
 \end{itemize}
-The following questions ask the participants to rate RationalGRL vs. standard goal modelling on a Likert scale from 1 (much more difficult) to 5 (much easier):
+The following questions ask the participants to rate RationalGRL vs. standard goal modeling language  on a Likert scale from 1 (much more difficult) to 5 (much easier):
 \begin{itemize}
@@
 \end{itemize}
-Note that for each question (Q5-Q10) we can test the null and alternative hypothesis separately, giving insight into exactly which additions of RationalGRL were deemed useful.
+Note that for each question (Q5-Q10,) we can test the null and alternative hypothesis separately, giving insight into exactly which additions of RationalGRL were deemed useful.
 
@@
 However, experience with goal modeling languages was markedly less: the average rating was 1.9 (standard deviation 1.3), with 9 participants claiming never to have used a goal modeling language (rating 1), and only two participants displaying regular use (monthly, rating 4, or weekly, rating 5).
-The participants that had experience with goal modelling languages had mostly used i* (5 users), with 2 users being familiar with GRL and 2 users having used another goal modelling language.
+The participants that had experience with goal modeling languages had mostly used i* (5 users), with 2 users being familiar with GRL and 2 users having used another goal modeling language.
+
+Note that the third author of the paper did not participate in recruiting of the participants, data collection and data processing of the user study due to the limitation of ``Institutional Review Board for the Protection of Human Subjects (IRB)" at her home university.
+Thus, she has not seen the participants information, the raw and processed data.
 
@@
 
-After the modelling task, the survey questions were asked.
+After the modeling task, the survey questions were asked.
 Table~\ref{table:survey:table1} shows the respondents' answers to question Q1-Q4 together with some statistical results which will be further discussed in Section \ref{sect:validation:analysis}.
@@
 
-\subsection{Analysis or results}\label{sect:validation:analysis}
+\subsection{Analysis of Results}\label{sect:validation:analysis}
 
 \paragraph{Hypothesis H1}
-The first analysis concerns hypothesis H1, whether the additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
+The first analysis concerns hypothesis H1, whether the additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modeling in GRL.
 Table~\ref{table:survey:table2} summarizes the average scores, standard deviations, p-values for a right-tailed t-test, effect size and statistical power for the relevant questions Q5-Q7.
@@
 It is different from the other two hypotheses in that we do not expect or claim that the usability of the tool is very good (or bad).
-Rather, because in the RationalGRL framework the tool is more a means than an end in itself, we hypothesize that the too, is neither easy or difficult to use.
+Rather, because in the RationalGRL framework, the tool is more a means than an end in itself, we hypothesize that the too, is neither easy or difficult to use.
 Table~\ref{table:survey:table1} summarizes the average scores, standard deviations and p-values for a two-tailed t-test for questions Q1-Q4.
@@
 In a realistic setting, users of RationalGRL would model much more complex requirements together with the stakeholders.
-That said, the discussions analysed in our case study (Section \ref{sect:gmas}) were of a realistic size and included many arguments and critical questions, so these concepts seem to take a central role in goal modeling and requirements engineering.
+That said, the discussions analyzed in our case study (Section \ref{sect:gmas}) were of a realistic size and included many arguments and critical questions, so these concepts seem to take a central role in goal modeling and requirements engineering.
 
-Another threat to internal validity is that many of the participants had no experience with goal modelling languages at all.
+Another threat to internal validity is that many of the participants had no experience with goal modeling languages at all.
 Hence, the perceived usefulness of the arguments could be influenced by the fact that the participants, for the first time, were given a goal modeling language to perform structured requirements analysis.
@@
 This can have various causes.
-Participants mentioned they did not use the details panel of the tool, which can be used to ask and answer critical questions, that much.
+Participants mentioned they did not use the details pane of the tool, which can be used to ask and answer critical questions, that much.
 Furthermore, it was also mentioned that the difference between different questions was not always clear.
 Finally, the experiment task, modeling an existing discussion, may also have influenced the CQ's perceived usefulness.
-The CQ's are meant to be asked during the actual RationalGRL Development Process in which goal models are constructed and critically analysed from scratch.
+The CQ's are meant to be asked during the actual RationalGRL development process in which goal models are constructed and critically analyzed from scratch.
 \item \emph{High cognitive overhead.} A concern that was raised often in our empirical evaluation is that, in its current form, RationalGRL has a relatively high cognitive overhead.

### 2018-09-17 Response to reviewers2.txt
fixed a typo

@@
 In this revision, we have done two things.
-First, we have moved the discussion on beliefs in GRL to "related work" under "goal modelling" and further extended and sharpened our discussion and comparison with GRL beliefs.
+First, we have moved the discussion on beliefs in GRL to "related work" under "goal modeling" and further extended and sharpened our discussion and comparison with GRL beliefs.
 Our main argument, which is now fully in the article, is that in GRL, beliefs are more of an afterthought: they have no formal semantics and it is not possible to further reason with them (i.e.

### 2018-09-18 main.tex
Section 1-5
last changes

@@
 \author{Marc van Zee  \and Floris J. Bex\and Sepideh Ghanavati}
-\institute{Marc van Zee \at Google BRAIN, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} 
+\institute{Marc van Zee \at University of Luxembourg, Luxembourg (currently at Google Research, Z\"urich, Switzerland) \\\email{marcvanzee@gmail.com} 
 \and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands 
 				\at Tilburg Institute for Law, Technology, and Society, Tilburg University, Tilburg, The Netherlands \\\email{f.j.bex@uu.nl}
-\and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu} }
+\and Sepideh Ghanavati \at School of Computing and Information Science, University of Maine, Bangor, USA \\\email{sepideh.ghanavati@maine.edu} }
 

### 2019-02-02 validation.tex
updated stats & stuff

--- 2018-09-16 validation.tex
@@
 
-In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a user evaluation with 16 users.
-The objective of this evaluation was to determine whether the users found the additions of RationalGRL (arguments, argumentation semantics, critical questions) useful compared to standard GRL, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to standard GRL.
-Regarding the comparison of RationalGRL to GRL, the following two broad hypotheses follow from our main points in this paper:
+In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a user evaluation with 16 expert users.
+The objective of this evaluation was to determine whether the users found the additions of RationalGRL (arguments, argumentation semantics, critical questions) useful, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to standard GRL.
+Regarding the usefulness of the RationalGRl features and the to GRL, the following two broad hypotheses follow from our main points in this paper:
 
@@
 
-In addition, we wanted to know whether the RationalGRL tool was difficult or easy to use.
-Note that we did not aim to conduct a usability study for the RationalGRL tool.
-Rather, we were interested to know if the tool was clearly very easy or difficult to use, as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive towards a particular modelling language, whereas irritations about a bad tool will lead to a more negative disposition.
-This leads to the following general hypothesis:
+In the rest of this section, we will describe our experiment in more detail and further specify these hypotheses.
 
-\begin{itemize}
-\item[H3] The usability of the RationalGRL tool is neither (very) easy or (very) difficult. 
-\end{itemize}
-
-In the rest of this section, we will describe our experiment in more detail and further specify these hypotheses.
+In addition, we wanted to know some basics about user experience.
+The aim here was not to conduct a full user experience study for the RationalGRL tool.
+Rather, we were interested to know if the user experience of the tool was either very good (easy to use) or very bad (difficult to use), as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive towards a particular modelling language, whereas irritations about a bad tool will lead to a more negative disposition.
 
@@
 
-The idea was to have our participants perform a small modelling task, and then ask them what they thought of RationalGRL and the tool. Thus, our experiment consisted of three parts:
+The idea was to have our participants perform a small modelling task, and then ask them what they thought of RationalGRL. Thus, our experiment consisted of three parts:
 
@@
 \item Modeling: We ask the participants to model a summarized discussion in the RationalGRL tool.
-\item Survey: We present the participants with a survey containing questions about the comparison between RationalGRL and GRL, and questions about the usability of the RationalGRL tool.
+\item Survey: We present the participants with a survey containing questions about the features of RationalGRL, and some questions about the usability of the RationalGRL tool.
 \end{enumerate}
@@
 
-For part 2, we provided a small part of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the tool and take a screenshot of their final model.
+For part 2, we provided a small part of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the RationalGRL tool and take a screenshot of their final model.
 The idea is that the participants act as if they are ``present'' at a discussion about the early-stage requirements and are asked to model the goal model corresponding to these requirements.
 
+In order to allow the participants to compare their models to regular GRL models, after the modelling exercise we provide the participants with Figure~\ref{fig:example-small}, which models the discussion for part 2 (Appendix~\ref{sect:survey}) in standard GRL.
+
 For part 3, we provided the participants with a survey\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2}.
-The survey starts with general questions about the participant (experience etc.).
-It then asks the following questions about the RationalGRL tool, asking the participants to rate various aspects of the tool on a Likert scale of 1 (very difficult) to 5 (very easy):
+The survey starts with general questions about the participant (years of experience etc.).
+The second part of the survey concerns the specific features of RationalGRL, asking participants to rate RationalGRL on a Likert scale from 1 (very useless) to 5 (very useful):
 \begin{itemize}
-\item[Q1] Was it easy or difficult to get started with the RationalGRL tool?
-\item[Q2] Was the details pane (containing details of an element, critical questions, etc) easy or difficult to use?
-\item[Q3] Was it easy or difficult to understand the way in which the status of arguments and other elements is determined?
-\item[Q4] Did you find it easy or difficult to model the example discussion using RationalGRL?
-\end{itemize}
-We also ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could in the future be used in practice.
-
-The second part of the survey concerns the comparison between GRL and RationalGRL and starts with Figure~\ref{fig:example-small}, which models the discussion for part 2 (Appendix~\ref{sect:survey}) in standard GRL, so that the participants can compare their own RationalGRL models and experiences with the standard GRL model.
-The following questions ask the participants to rate RationalGRL on a Likert scale from 1 (very useless) to 5 (very useful):
-\begin{itemize}
-\item[Q5] Do you think the arguments and counterarguments of RationalGRL are a useful extension to standard goal modeling?
-\item[Q6] Do you think the critical questions and answers in the details pane are a useful extension to standard goal modeling?
-\item[Q7] Do you think the automatic determination of the status of arguments and elements is a useful extension to standard goal modeling?
+\item[Q1] Do you think the arguments and counterarguments of RationalGRL are a useful extension to standard goal modeling?
+\item[Q2] Do you think the critical questions and answers in the details pane are a useful extension to standard goal modeling?
+\item[Q3] Do you think the automatic determination of the status of arguments and elements is a useful extension to standard goal modeling?
 \end{itemize}
@@
 \begin{itemize}
-\item[Q8] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone to express beliefs and opinions in a goal model?
-\item[Q9] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone to determine the effect of beliefs and opinions on the resulting goal model?
-\item [Q10] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone who is not the original author to understand the goal model? 
+\item[Q4] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone to express beliefs and opinions in a goal model?
+\item[Q5] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone to determine the effect of beliefs and opinions on the resulting goal model?
+\item [Q6] Do you think using RationalGRL instead of a standard goal modeling language makes it easier or more difficult to for someone who is not the original author to understand the goal model? 
 \end{itemize}
-We conclude with two open questions regarding the strengths and weaknesses of RationalGRL compared to GRL.
+We also ask two open questions regarding the strengths and weaknesses of RationalGRL compared to GRL.
 
-In order to test our main hypotheses H1, H2 and H3, we have to formulate the appropriate null hypotheses and alternative hypotheses.
-For H1 and H2, we hypothesize that the participants will, on average, rate the additions of RationalGRL as useful (rating 4) or very useful (rating 5) for Q5, Q6 and Q7, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q8, Q9 and Q10.
+The survey also contains questions about the RationalGRL tool user experience, such as whether it was easy or difficult to get started with the tool and to model the example case.
+We also ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could in the future be used in practice.
+
+In order to test our main hypotheses H1 and H2, we have to formulate the appropriate null hypotheses and alternative hypotheses.
+For H1 and H2, we hypothesize that the participants will, on average, rate the additions of RationalGRL as useful (rating 4) or very useful (rating 5) for Q1, Q2 and Q3, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q4, Q5 and Q6.
 In other words, we say that our null hypothesis and alternative hypothesis for H1 and H2 are as follows:
 \begin{itemize}
-\item H1$_{0}$ and H2$_{0}$: average rating for questions Q5-Q10 is 3 or lower.
-\item H1$_{a}$ and H2$_{a}$: average rating for questions Q5-Q10 is higher than 3.
+\item H1$_{0}$ and H2$_{0}$: median rating for questions Q1-Q6 is 3 or lower.
+\item H1$_{a}$ and H2$_{a}$: median rating for questions Q1-Q6 is higher than 3.
 \end{itemize}
-Note that for each question (Q5-Q10) we can test the null and alternative hypothesis separately, giving insight into exactly which additions of RationalGRL were deemed useful.
-
-For hypothesis H3, we do not claim that the usability of the tool is especially easy or difficult: we expect the average rating for questions Q1-Q4 to not be significantly higher or lower than 3 (neither easy or difficult).
-Thus, our null hypothesis and alternative hypothesis for H3 are as follows:
-\begin{itemize}
-\item H3$_{0}$: average rating for questions Q1-Q4 is 3.
-\item H3$_{a}$: average rating for questions Q1-Q4 is not 3.
-\end{itemize}
-Again, we can test these hypothesis separately for each of the questions.
+Note that for each question (Q1-Q6) we can test the null and alternative hypothesis separately, giving insight into exactly which additions of RationalGRL were deemed useful.
+We set the significance level for each hypothesis at $\aplha = 0.05$
 
@@
 \centering
-\begin{tabularx}{0.8\textwidth}{l|l|l|l|l|l||l|l|l}
-& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p-value \\
+\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
+& very useless & useless & neutral & useful & very useful & median & mode & p-value \\
 \hline
-Q1 & 2 (12.5\%) & 1 (6.3\%)  & 6 (37.5\%) & 6 (37.5\%) & 1 (6.3\%)  & 3.2 & 1.1 & 0.509 \\
-Q2 & 2 (12.5\%) & 2 (12.5\%) & 4 (25.0\%) & 8 (50.0\%) & 0 (0.0\%)  & 3.1 & 1.1 & 0.652 \\
-Q3 & 0 (0.0\%)  & 3 (20.0\%) & 4 (26.7\%) & 5 (33.3\%) & 3 (20.0\%) & 3.5 & 1.1 & 0.072 \\
-Q4 & 2 (12.5\%) & 5 (31.3\%) & 6 (37.5\%) & 3 (18.8\%) & 0 (0.0\%)  & 2.6 & 1.0 & 0.138 
-\end{tabularx}
-\caption{Participant ratings and statistical results for the usability of the RationalGRL tool}
-\label{table:survey:table1}
-\end{table*}
-
-\begin{table*}[t]
-\centering
-\begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
-& very useless & useless & neutral & useful & very useful & avg (1-5) & std. dev. & p-value & effect size $d$ & power\\
-\hline
-Q5 & 0 (0.0\%) & 0 (0.0\%)  & 3 (21.4\%) & 8 (57.1\%) & 3 (21.4\%) & 4.0 & 0.7 & 0.000 & 1.5 & 0.999 \\
-Q6 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 3.5 & 1.0 & 0.045 & 0.5 & 0.525 \\
-Q7 & 0 (0.0\%) & 0 (0.0\%)  & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4.2 & 0.7 & 0.000 & 1.7 & 1.000
+Q1 & 0 (0.0\%) & 0 (0.0\%)  & 3 (21.4\%) & 8 (57.1\%) & 3 (21.4\%) & 4 & 4 & 0.0005 \\
+Q2 & 0 (0.0\%) & 3 (21.4\%) & 3 (21.4\%) & 6 (42.9\%) & 2 (14.3\%) & 4 & 4 & 0.1133 \\
+Q3 & 0 (0.0\%) & 0 (0.0\%)  & 2 (14.3\%) & 7 (50.0\%) & 5 (35.7\%) & 4 & 4 & 0.0002 
 \end{tabularx}
@@
 \begin{tabularx}{0.95\textwidth}{l|l|l|l|l|l||l|l|l|l|l}
-& very difficult & difficult & neutral & easy & very easy & avg (1-5) & std. dev. & p-value & effect size $d$ & power \\
+& very difficult & difficult & neutral & easy & very easy & median & mode & p-value \\
 \hline
-Q8  & 0 (0.0\%) & 0 (0.0\%) & 4 (30.8\%) & 4 (30.8\%) & 5 (38.5\%) & 4.1 & 0.9 & 0.000 & 1.2 & 0.991\\
-Q9  & 0 (0.0\%) & 0 (0.0\%) & 6 (46.2\%) & 5 (38.5\%) & 2 (15.4\%) & 3.7 & 0.8 & 0.003 & 0.9 & 0.925\\
-Q10 & 0 (0.0\%) & 1 (8.3\%) & 4 (33.3\%) & 7 (58.3\%) & 0 (0.0\%)  & 3.5 & 0.7 & 0.013 & 0.7 & 0.772
+Q4  & 0 (0.0\%) & 0 (0.0\%) & 4 (30.8\%) & 4 (30.8\%) & 5 (38.5\%) & 4 & 5 & 0.0020 \\
+Q5  & 0 (0.0\%) & 0 (0.0\%) & 6 (46.2\%) & 5 (38.5\%) & 2 (15.4\%) & 4 & 3 & 0.0078\\
+Q6 & 0 (0.0\%) & 1 (8.3\%) & 4 (33.3\%) & 7 (58.3\%) & 0 (0.0\%)  & 4 & 4 & 0.0352
 \end{tabularx}
@@
 After the modelling task, the survey questions were asked.
-Table~\ref{table:survey:table1} shows the respondents' answers to question Q1-Q4 together with some statistical results which will be further discussed in Section \ref{sect:validation:analysis}.
-
-We also asked the users three open questions related to usability.
-When asked about the strengths and weaknesses of the RationalGRL tool, users where generally positive about RationalGRL's clear UI, the fact that the tools gives a nice overview of the goals and opinions in a case and that it was straightforward to use and understand.
+We first asked the participants some questions about user experience.
+The users where generally positive about RationalGRL's clear UI, the fact that the tools gives a nice overview of the goals and opinions in a case and that it was straightforward to use and understand.
 With respect to weaknesses and improvements to the RationalGRL tool, the users mentioned additional UI functionalities such as the possibility to save models, having an ``undo'' function, and flipping the arrow after adding them.
@@
 
-When asked in an open question whether the users believe a more mature version of the RationalGRL tool could be used to capture early-phase requirements in an actual software development project, most users responded positively.
-Users' responses include ``Yes, having the possibility to add arguments seems quite useful'', ``I think it can, but maybe try to find a way to combine it with more regular/mainstream requirements gathering such as user stories and customer journeys'', ``yes, because it creates a clear scope for the project, what the goals are'', and ``Generally I think its useful to explicitly document arguments of a discussion''.
+The next set of questions concerned the specific features of the RationalGRL modeling language and the difference with standard goal modeling languages.
+Table~\ref{table:survey:table2} shows the respondents' answers to questions Q1-Q3, and Table~\ref{table:survey:table3} shows the answers to questions Q4-Q6.
+We concluded with two open questions about the comparison between RationalGRL and other goal modeling languages.
+When asked about the advantages of RationalGRL over standard goal modeling languages, many users agreed that making arguments explicit may force end users to have a more structured discussion:  ``Yes, having the possibility to add arguments seems quite useful'', `Generally I think its useful to explicitly document arguments of a discussion'', ``Clear communication about argumentation and forcing people to think in those clear terms.'', ``...you can add arguments and that you can answer questions that help you to develop arguments'', ``It's useful that discussion and explanation are close to the diagrams'', ``a way to see how decisions are being shaped.''.
+Furthermore, one user stated that RationalGRL successfully ``tries to capture the rationale behind the modeling process' and ``has a simple way to compute the status of the arguments''.
+
+When asked about the weaknesses, users mentioned the complexity as the most important weakness: ``The apparent increase in complexity might lead to negative perceptions'', ``adding yet another layer of complexity scares me''.
 The concern that the modeling process may involve too much cognitive overhead was mentioned a few times though: ``I think the overhead of inputting a (detailed) discussion in a structured manner into any system makes adoption difficult'', ``the manual input is too complex and takes too much time.
 An automated process of parsing the conversation log would be much more helpful'', and ``I still believe in the value of arguments, but there should be less confusing ways to capture them''.
-
-Table~\ref{table:survey:table2} shows the respondents' answers to questions Q5-Q7, and Table~\ref{table:survey:table3} shows the answers to questions Q8-Q10.
-We concluded with two open questions about the comparison between RationalGRL and other goal modeling languages.
-When asked about the advantages of RationalGRL over standard goal modeling languages, many users agreed that making arguments explicit may force end users to have a more structured discussion: ``	Clear communication about argumentation and forcing people to think in those clear terms.'', ``...you can add arguments and that you can answer questions that help you to develop arguments'', ``It's useful that discussion and explanation are close to the diagrams'', ``a way to see how decisions are being shaped.''.
-Furthermore, one user stated that RationalGRL successfully ``tries to capture the rationale behind the modeling process' and ``has a simple way to compute the status of the arguments''.
-When asked about the weaknesses, users mentioned the complexity as the most important weakness: ``The apparent increase in complexity might lead to negative perceptions'', ``adding yet another layer of complexity scares me''.
-One user mentioned that this may be a problem with goal modeling in general: ``Goal models are already complex (...) I have worked for years on the effect of context on goal models, and my conclusion is that this was very interesting academic work but with close-to-zero practical implications, unfortunately.''.
+One user mentioned that complexity and cognitive overload is a problem with goal modeling in general: ``Goal models are already complex (...) I have worked for years on the effect of context on goal models, and my conclusion is that this was very interesting academic work but with close-to-zero practical implications, unfortunately.''.
 
@@
 The first analysis concerns hypothesis H1, whether the additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
-Table~\ref{table:survey:table2} summarizes the average scores, standard deviations, p-values for a right-tailed t-test, effect size and statistical power for the relevant questions Q5-Q7.
-As indicated by the p-values, which are all lower than 0.05, all the averages are significantly higher than the median of 3 with a significance level ($\alpha$) of 0.05, which means that we can accept H1$_{a}$ and reject H1$_{0}$ for all the separate questions Q5-Q7.
-In other words, the fact that the participants rate the additions of RationalGRL as, on average, (very) useful is very unlikely to be due to chance.
-
-For questions Q5 and Q7, the effect sizes are very large\footnote{Where a low p-value for a hypothesis test such as a t-test indicates whether the fact that average rating of a question is statistically significantly higher than the median 3, the effect size indicates the magnitude of the difference between the measured average and the median.
-Generally, an effect size below 0.2 is seen as low, and effect size of 0.5 is seen as medium, an effect size of 0.8 is seen as large and an effect size of 1.2 is seen as very large.}.
+Table~\ref{table:survey:table2} summarizes the medians, modes, and p-values for a one-tailed single-sample sign test for the relevant questions Q1-Q3.
+We chose the simplest non-parametric test available so as to have to take as few assumptions as possible w.r.t.
+the distribution of answers for each question.
+As is evident from the results, the participants were generally positive about the additions of RationalGRL.
+For Q1 and Q3 the p-values are both significantly lower than $\alpha = 0.05$, which means that for these questions we can accept H1$_{a}$ and reject H1$_{0}$.
 This means that the addition of arguments and the use of arguments to determine the acceptability of elements of the goal model are deemed to be useful to very useful by the participants.
-As can be seen in Table~\ref{table:survey:table2}: no participant found the addition of arguments and argument status useless or very useless, and 78.6\% (Q5) respectively 85.7\% (Q7) found the additions useful or very useful.
+As can be seen in Table~\ref{table:survey:table2}: no participant found the addition of arguments and argument status useless or very useless, and 78.6\% (Q1) respectively 85.7\% (Q3) found the additions useful or very useful.
 Furthermore, these findings fit with the answers to the open questions (see Section \ref{sec:survey:results}).
-Finally, despite the low $n$ the statistical power for these questions is very high (\textgreater 0.99), which means that the chances of wrongly rejecting H1$_{0}$ are close to 0\footnote{Statistical power indicates the probability of correctly rejecting the null hypothesis}.
-
-Interestingly, while for Q6 the average is significantly higher than the median 3, the effect size is only medium and the power is not very high, with the chances of wrongly rejecting H1$_{0}$ being almost 50\%.
-Hence, we can say that the addition of critical questions is perhaps not as useful as we had hypothesized.
+For Q2, the p-value indicates that the result is not significant with $\alpha = 0.05$, which means we cannot reject the null hypothesis H1$_{0}$.
+This means that the participants did not think the addition of critical questions like to the RationalGRL tool was significantly useful.
+So we can say that the addition of critical questions is perhaps not as useful as we had hypothesized.
 This again fits the descriptive statistics in Table\ref{table:survey:table2}: 21.4\% of respondents found the critical questions ``useless''.
@@
 The second hypothesis states that additions of RationalGRL make it easier to express one's opinions and beliefs about a goal model, determine the effect of one's opinions and beliefs about a goal model and communicate one's opinions and beliefs about a goal model to other stakeholders.
-Table~\ref{table:survey:table3} provides a summary of the statistical analysis of the relevant questions Q8-Q10.
-As can be seen, all the averages are significantly higher than the median of 3 with a significance level ($\alpha$) of 0.05, which means that we can accept H1$_{a}$ and reject H1$_{0}$ for all the separate questions Q8-Q10.
+Table~\ref{table:survey:table3} provides a summary of the statistical analysis of the relevant questions Q4-Q6.
+As can be seen, all the scores are significantly higher than the median of 3  for $\alpha = 0.05$, which means that we can accept H2$_{a}$ and reject H2$_{0}$ for all the separate questions Q4-Q6.
 In other words, the fact that the participants rate the additions of RationalGRL as making it, on average, easier to reason with their beliefs and opinions about a goal model is very unlikely to be due to chance.
 
-With respect to Q8, there is again a very large effect size and high statistical power: participants say that arguments make it much easier to express their opinions.
+With respect to Q4, participants say that arguments make it much easier to express their opinions.
 None of the participants found expressing their beliefs (much) more difficult, and 69.3\% found it easier or much easier.
-This corresponds to the answers to Q5, where participants said the arguments were a useful addition.
-Like for Q5, for Q8 the chances of wrongly rejecting H1$_{0}$ are close to 0.
-
-For Q9, the effect size is somewhat smaller, though still large: RationalGRL makes it easier to determine the effect of one's beliefs on the goal model.
-Power shows that there is a small (around 7.5\% chance) or wrongly rejecting the null hypothesis.
-However, combining this with the results for question Q7, which asked about the usefulness of determining the status, we can say that the participants overall found this a useful feature which makes working with conflicting beliefs easier.
+This corresponds to the answers to Q1, where participants said the arguments were a useful addition.
+For Q5, participants agree that RationalGRL makes it easier to determine the effect of one's beliefs on the goal model.
+Combining this with the results for question Q3, which asked about the usefulness of determining the status, we can say that the participants overall found this a useful feature which makes working with conflicting beliefs easier.
 One participant remarked that ``The reasoning [in the example] seems simplistic [...] the added value of [the formal argumentation] would be justified if the reasoning is more complex''.
-So the simplistic case in the modeling exercise possibly detracts from the results for Q9 and Q7.
-
-The analysis of the results for Q10 show a medium to large effect size: overall, participants were less sure that arguments would help in communicating the goal model to others.
-Statistical power is also not that high at 77\%.
+So the simplistic case in the modeling exercise possibly detracts from the results for Q5 and Q3.
+Finally, participants were less sure that arguments would help in communicating the goal model to others.
 This may be due to the fact that in the experiment, communication of opinions to others was not really tested, leaving the participants to guess what this effect would be.
-
-\paragraph{Hypothesis H3}
-The third hypothesis H3 is about the usability of the RationalGRL tool.
-It is different from the other two hypotheses in that we do not expect or claim that the usability of the tool is very good (or bad).
-Rather, because in the RationalGRL framework the tool is more a means than an end in itself, we hypothesize that the too, is neither easy or difficult to use.
-Table~\ref{table:survey:table1} summarizes the average scores, standard deviations and p-values for a two-tailed t-test for questions Q1-Q4.
-As indicated by the p-values which are all higher than $\alpha=0.05$, none of the results is statistically significant, that is, the fact that there is a difference between the median of 3 and the average scores for the questions could very well be due to chance.
-This means we cannot reject the null hypothesis H3$_{0}$ and that, hence, the usability of the RationalGRL tool is neither (very) easy or (very) difficult.
-This corresponds to the answers to the open questions the participants gave: respondents were positive about the fact that it was very easy to get started with the tool and most comments on the tool were the type of comments that are expected from a prototype implementation.
 
@@
 In a realistic setting, users of RationalGRL would model much more complex requirements together with the stakeholders.
-That said, the discussions analysed in our case study (Section \ref{sect:gmas}) were of a realistic size and included many arguments and critical questions, so these concepts seem to take a central role in goal modeling and requirements engineering.
+The problems with complexity and cognitive overload only start to play a role for larger models, so while the participants were generally positive in the ranking questions Q1-Q6, some reservations were expressed in the open questions as to whether this approach would work in a realistic setting.
 
@@
 Hence, the perceived usefulness of the arguments could be influenced by the fact that the participants, for the first time, were given a goal modeling language to perform structured requirements analysis.
-In order to test the usefulness of arguments and critical questions alone, one would have to compare a test and control group, where the test group uses RationalGRL and the control group uses normal GRL.
+In other words, perhaps they found a structured modelling language generally quite useful.
+One way to better check the difference between standard goal modeling and RationalGRL would be to first provide users with a standard goal modeling tool and then have them use RationalGRL.
+An even more thorough way to test the usefulness of RationalGRL over, for example, normal GRL would be to compare a test and control group, where the test group uses RationalGRL and the control group uses normal GRL.
+We leave these types of further empirical studies for future research.
 

### 2019-02-02 formalframework.tex

@@
 
-Let us explain these definitions using the example, argumentation framework, in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of our RationalGRL model from Figure~\ref{fig:example-small3}, containing only arguments $A_4, A_{5},A'_{5},A_{13}$, and $A_{14}$.
+Let us explain these definitions using the example argumentation framework, in Figure~\ref{fig:goalmodeling:arg2}, which is a subset of our RationalGRL model from Figure~\ref{fig:example-small3}, containing only arguments $A_4, A_{5},A'_{5},A_{13}$, and $A_{14}$.
 In this example, there are five admissible sets: $\{A'_{5}\}$, $\{A_{14}\}$, $\{A'_{5},A_{14}\}$, $\{A_4, A_{14}\}$ and $\{A_4, A'_5, A_{14}\}$.

### 2019-02-02 overview.tex

@@
 
-In the rest of this section we discuss the individual parts of the GRL framework.
+In the rest of this section we discuss the individual parts of the RationalGRL framework.
 In Section~\ref{sect:overview:as}, we continue our discussion of the argument schemes and critical questions for practical reasoning and argumentation, fitting these schemes and questions into our framework.
@@
 \includegraphics[width=\textwidth]{metamodel/metamodel2}
-\caption{The RationalGRL metamodel. The \emph{Goal model construction} package (bottom) is the GRL metamodel. The \emph{Practical reasoning \& argumentation} package is the RationalGRL extension}
+\caption{The RationalGRL Metamodel. The \emph{Goal model construction} package (bottom) is the GRL metamodel. The \emph{Practical reasoning \& argumentation} package is the RationalGRL extension}
 \label{fig:metamodel}
@@
 Therefore, each argument is associated with at most one scheme, but a scheme can be instantiated in multiple ways.
-A \textsf{RationalGRLmodel} is composed out of arguments and attack relations.
+A \textsf{RationalGRLmodel} is composed of arguments and attack relations.
 
-The \textsf{OperationTypes} in the Argumentation package correspond to the operations we informally introduced in Section~\ref{sect:overview}.
+The \textsf{OperationTypes} in the argumentation package correspond to the operations we informally introduced in Section~\ref{sect:overview}.
 These operations are performed by instantiating an argument scheme or answering a critical question in a certain way.
@@
 Note that besides arguments about elements of the GRL model, we also have a \textsf{GenericArgument} which is simply a counter-argument to an existing argument that does not relate to any of the GRL elements.
-Finally, the relation between \textsf{GRLModelElement} and \textsf{Argument} means that, as we already briefly indicated when discussing the framework in Figure~\ref{fig:rationalgrl-framework}, the class of \textsf{RationalGRLmodel} is a superclass of \textsf{GRLmodel}: besides arguments about GRL elements, we can also have arguments that does not relate to any of the GRL elements.
+Finally, the relation between \textsf{GRLModelElement} and \textsf{Argument} means that, as we already briefly indicated when discussing the framework in Figure~\ref{fig:rationalgrl-framework}, the class of \textsf{RationalGRLmodel} is a superclass of \textsf{GRLmodel}: besides arguments about GRL elements, we can also have arguments that do not relate to any of the GRL elements.
 
@@
 \includegraphics[width=\columnwidth]{img/fig_example_AS.pdf}
-\caption{RationalGRL model (right) with instantiated argument schemes (left): Introducing new elements (operation \textsf{INTRO)}}. 
+\caption{RationalGRL Model (right) with Instantiated Argument Schemes (left): Introducing New Elements (Operation \textsf{INTRO)}}. 
 \label{fig:example_AS}
@@
 \includegraphics[]{img/fig_example_disable.pdf}
-\caption{Disabling elements (operation \textsf{DISABLE)}.}
+\caption{Disabling Elements (Operation \textsf{DISABLE)}.}
 \label{fig:example_disable}
@@
 On the left side of Figure~\ref{fig:example_disable} a new argument (CQ11) attacks the original argument based on argument scheme AS2.
-This new argument is also added to the RationalGRL model on the right of Figure~\ref{fig:example_disable}, where it attacks the original task \emph{Add traffic light}.
+This new argument is also added to the RationalGRL model on the right-hand side of Figure~\ref{fig:example_disable}, where it attacks the original task \emph{Add traffic light}.
 This attack leads to the original argument being \emph{rejected} (cf.
@@
 \includegraphics[]{img/fig_example_replace.pdf}
-\caption{Replacing an element (operation \textsf{REPLACE)}.}
+\caption{Replacing an Element (Operation \textsf{REPLACE)}.}
 \label{fig:examples:replace}
@@
 \includegraphics[]{img/fig_example_rename.pdf}
-\caption{Renaming an element name (operation \textsf{REPLACE)}.}
+\caption{Renaming an Element Name (Operation \textsf{REPLACE)}.}
 \label{fig:examples:clarification}
@@
 \includegraphics[]{img/reinstate1.pdf}
-\caption{Disabling an element (operation \textsf{DISABLE)}.}
+\caption{Disabling an Element (Operation \textsf{DISABLE)}.}
 \label{fig:examples:relevant-actor}
@@
 Actor \emph{Development Team} is introduced with an argument based on AS0 (Figure~\ref{fig:examples:relevant-actor}).
-This is then attacked by arguing that the professor will develop the software, so there will not be any development team (CQ0).
+This is then attacked by arguing that the professor will develop the software, thus, there will not be any development team (CQ0).
 
@@
 \includegraphics[]{img/reinstate2.pdf}
-\caption{Defending an element by attacking its disabling attacker (operation \textsf{DISABLE)}.}
+\caption{Defending an Element by Attacking its Disabling Attacker (Operation \textsf{DISABLE)}.}
 \label{fig:examples:relevant-actor2}

### 2019-02-02 tool.tex

--- 2018-05-02 tool.tex
@@
 We propose the development process shown in Figure~\ref{fig:rationalgrl-methodology} to develop a RationalGRL model, a version of which was presented at the 2017 iStar workshop \cite{ghanavatiMethodology}.
-Here we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
+Here, we assume that the initial GRL models have been created based on the requirements specification documents and the discussions of the stakeholders.
 The rest of the steps are as follows:
@@
 The final requirement of our framework is that it has tool support (\textbf{requirement 5}).
-This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to do user tests in future research, exploring the difference between GRL and RationalGRL, iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.
-In this section we briefly highlight some features of the tool, and we explain some of its current limitations.
+This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of an existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to conduct user studies in the next section, exploring the usability and usefulness of the RationalGRL framework compared to other goal modeling languages, and iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.
+In this section, we briefly highlight some features of the tool, and we explain some of its current limitations.
 The tool can be accessed from:
@@
 jUCMNav is used actively in research, and the tool has been extended in many directions over the past few years, such as for modeling and analyzing security misuse cases~\cite{daramola2012ontology}, supporting activity theory \cite{georg2015synergy}, combined goal and feature model reasoning~\cite{liu2014combined}, and enterprise architecture modeling \cite{marosin-etal:caise2016}.
-Furthermore, jUCMNav includes Use Case modeling and contains various GRL evaluation algorithms with which the satisfiability of goals can be determined.
+Furthermore, jUCMNav includes Use Case Maps for modeling business processes and scenarios and contains various GRL evaluation algorithms with which the satisfiability of goals can be determined.
 
@@
 Our tool is thus not meant to replace jUCMNav.
-For simplicity we have a web-based version, and it has the functionality to export to jUCMNav.
+For simplicity, we have a web-based version, and it has the functionality to export to jUCMNav.
 
@@
 
-In total there are four different types of details panes, which we now explain in turn.
+In total, there are four different types of details panes, which we now explain in turn.
 
@@
 
-\paragraph{Critical question details pane} The details panel can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal IE ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
+\paragraph{Critical question details pane} The details pane can be used to answer the critical questions of the RationalGRL framework (Table~\ref{table:argument-schemes}): Figure~\ref{fig:tool:overview}, for example, shows two critical questions for the goal IE ``Generate cars'', namely ``Can the goal be realized?'' (CQ3) and ``Is the goal relevant/useful?'' (CQ11).
 As an example, consider Figure~\ref{fig:tool:cqdetails}, where the softgoal ``Easy to use'' is questioned with the relevancy question (CQ11).
@@
 
-Assume that in the example of Figure~\ref{fig:tool:cqdetails} the user selects ``No'' and clicks the ``Answer question'' button.
+Assume that in the example of Figure~\ref{fig:tool:cqdetails}, the user selects ``No'' and clicks the ``Answer question'' button.
 A new argument is then automatically created that attacks the softgoal and the details pane shows the critical question as answered (Figure~\ref{fig:tool:cqeffect}).
@@
 \includegraphics[width=0.8\columnwidth]{img/tool/details_softgoal}
-\caption{Critical question details pane for ``Is the softgoal relevant/useful?''}
+\caption{Critical Question Details Pane for ``Is the softgoal relevant/useful?''}
 \label{fig:tool:cqdetails}
@@
 \includegraphics[width=\columnwidth]{img/tool/attack_softgoal}
-\caption{Effect of answering CQ11: ``Is the softgoal relevant/useful?'' with ``No''}
+\caption{Effect of Answering CQ11: ``Is the softgoal relevant/useful?'' with ``No''}
 \label{fig:tool:cqeffect}
@@
 \includegraphics[width=0.8\columnwidth]{img/tool/reinstate_softgoal}
-\caption{Attacking an argument with a generic argument ``Necessary''}
+\caption{Attacking an Argument with a Generic Argument ``Necessary''}
 \label{fig:tool:argument}
@@
 A positive contribution is added from ``Keep same cars'' to ``Simple design''.
-More inportantly, an argument ``not enough CPU'' has been added to the model.
+More importantly, an argument ``not enough CPU'' has been added to the model.
 This argument attacks the ``Realistic simulation'' softgoal and the ``Create new cars'' task, arguing that there is not enough processing power for either of these to be feasible in the traffic simulator.

### 2019-02-02 validation.tex

@@
 
-The idea was to have our participants perform a small modelling task, and then ask them what they thought of RationalGRL. Thus, our experiment consisted of three parts:
+The idea was to have our participants perform a small modeling task, and then ask them what they thought of RationalGRL. Thus, our experiment consisted of three parts:
 
@@
 We also ask two open questions regarding the strengths and weaknesses of RationalGRL compared to GRL.
-
-The survey also contains questions about the RationalGRL tool user experience, such as whether it was easy or difficult to get started with the tool and to model the example case.
+The survey further contains questions about the RationalGRL tool user experience, such as whether it was easy or difficult to get started with the tool and to model the example case.
 We also ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could in the future be used in practice.
@@
 
-\subsection{Analysis or results}\label{sect:validation:analysis}
+\subsection{Analysis of Results}\label{sect:validation:analysis}
 

### 2019-02-03 Response to reviewers3.txt
response to reviewers

We thank the reviewers for their work.

Reviewer 1 again proposes to accept the paper. Reviewer 2 recommends rejection. The editor proposes that we perform major revisions

We have worked hard to improve the paper since the first submission. We propose a new framework, inserting argument schemes and formal argument semantics into goal modelling in a way that hasn't been done before. We formally ground this framework in a logical framework for argumentation that has been accepted as a de-facto standard for non-monotonic argumentation inside and outside the field of AI. We give an simply development process, and a fully-working tool. We have further performed an extensive qualitative emprirical coding study (which we call case study in our article), plus a modest expert survey.

One remark that has been made since the beginning is that, because we propose a new process/language/tool, we should test these in an industrial-strength case-study. However, we remain convinced that a large case study or experiment (where e.g. one compares test and control groups) would take an unreasonable amount of time and space given the material that is alreadty in the paper. See e.g. Estrada et al. (2006) "An empirical evaluation of the i* framework in a model-based software generation environment", who took 9 months to conduct the case study. Furthermore, there are many papers on goal-modeling and related subjects in REJ which perform less empirical evaluation than we do, and the few that do have industrial-strength case studies evaluate a modeling language or tool that was introduced earlier. We are not directly aware of papers on goal modeling languages doing extensive experimentation with test and control groups. Hence, we believe that asking us to do an industrial strength case-study or test-control group experiment is unreasonable for this paper, more so because this would likely yield enough new insights for a completely new paper.

The (smaller) remarks made by reviewer 2 since Revision 1 have been helpful, and we try to accomodate as many of them as possible. We would like to note that we have never simply refused suggestions or ignored comments. In some cases, we have replied at length to the reviewer in the response letter and have not added anything to the paper, but we do not think this counts as a refusal to take comments into account.

Below we respond to reviewer 2's comments on the latest version of the article in detail.

======= COMMENT #1
First, there are "broad" and "general" hypotheses given, and based on them Null and Alternative Hypotheses. Still, the latter do not contain the level of significance for accepting or rejecting them based on statistical data and their analyses. (It has to be defined before performing an experiment.)
======= REPLY #1
We have added the significance levels below the hypothesis

======= COMMENT #2
H3 (and its specific counterparts) are strange in their formulation, since usability is not difficult, but use is difficult and usability potentially bad. In English, after "neither" there should be a "nor".
The paragraph apparently trying to motivate H3 sounds actually more like motivating user experience (UX) rather than usability, since testing the latter is first of all more about finding usability problems, something never mentioned in this paper. Also the related Null and Alternative Hypotheses are strange with regard to the given motivation. Much as in my previous review, I really wonder about including usability of such a prototype tool here. (As given later in the paper, this tool does not even have an Undo, and the results were completely inconclusive, anyway.)
======= REPLY #2
Our reasons for including the questions and hypothesis on usability (or UX) were not so that we could argue that said UX is particularly good (or bad). We wanted to show that the usability or UX of the tool did, in our opinion, not interfere greatly with what the participants filled in for the other questions. If a tool is very bad (keeps crashing, impossible to use, unclear interface), participants might be more inclined to be very negative about the rest (i.e. the RationalGRL features) as well. In any case, clearly we were not able to explain our ideas clearly enough, so we deleted H3 and only mention the user experience questions in passing.

======= COMMENT #3
For a comparison, actually a user study would have had to be designed, with independent and dependent variables, and participants trying both approaches in two groups (within-subject study design). In such a setting, something should be really measured in addition to subjective questionnaires. Much as indicated by the participants of the reported study (and as known from the literature), a key point of such argumentation systems is the overhead of providing them. There needs to be a return on invest, ideally in a balanced trade-off. I think that this should have been studied, since I am very much in favor of documenting rationale, but in practice developers will only do that if there is a clear benefit in return. The communication of opinions to others was obviously not really included.
======= REPLY #3
As we argue above, we think this is an unreasonable amount of work to ask given what is already in the paper. Also, it now seems as if our (modest) expert survey is worthless, which we think it is not.

======= COMMENT #4
A theoretical problem of all the results reported here is the use of statistics like mean (here obviously called "average") and standard deviation based on data  on a Likert scale, since this is not permitted according to Measurement Theory. (I am aware, that this is often done, anyway, especially in practice, but the theoretical problem remains, especially for results in a prestigious scientific journal.) Such statistics could and should be done based on measured data, which are usually on an absolute scale.
In addition, has the probability distribution been tested before applying the t-test?
The specific results for Q6 are really strange.
Finally, as indicated above already, the results on usability of the tool are virtually useless.
======= REPLY #4
There is actually quite some discussion about whether Likert scales can be analysed as interval data - in academia more and more researchers (who publish in prestigious statistics journals) are starting to agree that one is allowed to do this. However, we agree that for our data it is perhaps better to err on the side of caution and use a non-parametric test which makes as little assumptions as possible (i.e. the sign test). Interestingly, the results are not much different: for Q2 (prev. Q6) we now cannot reject H0, where previously we could, but in the previous version (with the t-tests) we already indicated that Q2/Q6 has results which are clearly different from the other related questions.

======= COMMENT #5
The first argument on internal validity is strange and really not convincing. If the modeling task given in the experiment was not realistic (and this is most likely the case), then the case study does not help, see below.
I agree that a comparison with a control group is necessary!
-       Discussion
And yes,  a realistic study on this approach needs to involve asking CQs in the course of a development process.
Why did you not let users being guided by your process in the course of a study, when users like that? This would have provided evidence or at least insights on the usefulness of this process, while it is just a claim as it stands.
======= REPLY #5
Again: yes there are many ways in which we could have done our experiments and surveys differently, better, bigger etc. However, the reviewer makes it sound as if the current survey is next to useless, and we strongly disagree with this. It is a modest survey with a fairly high response rate, particularly as all 16 participants have significant expertise in software and requirements engineering - we could do an experiment with 80 inexperienced undergraduates, but what would that really tell us? The answers to the questions mainly point in interesting future directions, but overall the experiment gives modest support to the idea that adding arguments to goal models is the way to go.

======= COMMENT #6
In this context, let me also say a few words on the "Case Study":
-       It cannot be taken as evidence for the approach, since studying recorded design sessions (without goal modeling) is reported in this paper as the basis for creating the argumentation structures in the first place. And this was done by the authors themselves based on their intuition.
-       Actually, this is not really a case study in the sense used in scientific literature. As a matter of fact, it also does not contain lessons learned.
======= REPLY #6
We agree that "case study" is perhaps not the most opportune name. It is, if you want to be precise, a qualitative empirical coding study with a priori coding. We have changed the text in the paper to reflect this.

======= COMMENT #7
In addition, I am still not convinced by the argument for dismissing formal reasoning in propositional logic. After all, it is not theoretically difficult and tools exist. So, what is the point of presenting special-purpose "algorithms" here?
======= REPLY #7
The argumentation formalism we propose is a non-monotonic logic, that is, earlier results might not be true when new information is added. Formal reasoning in propositional logic, taking the semantics of classical logic, is inherently monotonic: you can add propositions (arguments, goals), but these will never change the status (true/false) of existing propositions. So what is needed is a new algorithm that shows what happens if new information is added, we cannot simply take an existing tool (theorem prover) for classical logic.

======= COMMENT #8
Finally, "defending" a binary framework simply by a reference is insufficient. Since the authors propose such a framework, they need to clearly point out its limitations in this regard.
======= REPLY #8
We already included some discussion in the previous revision. We have included an extra sentence on why we think numerical, more fine-grained frameworks are not very suitable.

======= COMMENT #9
From an editorial perspective, a careful proof reading of grammar would be necessary.
======= REPLY #9
This remark makes it sound as if there are many grammatical mistakes in the paper.

### 2019-02-03 formalframework.tex
response to reviewers

@@
 Note that argument acceptability is essentially a three-valued (or binary if one does not count \emph{undecided}) concept.
-There are many more fine-grained ways of defining the acceptability of arguments using, for example, probabilities \cite{li2011probabilistic} or counting the number of attacks \cite{grossi2015graded}.
-However, the ``basic'' argumentation frameworks, and in particular the  have been shown to be close to how humans reason \cite{rahwan2010behavioral}.
+Extending the argumentation with preferences between arguments is fairly simple \cite{atkinson2007}, and would give users more control over which arguments are more or less convincing.
+Furthermore, there are many more fine-grained ways of defining the acceptability of arguments using, for example, probabilities \cite{li2011probabilistic} or counting the number of attacks \cite{grossi2015graded}.
+However, the ``basic'' argumentation framework, and in particular the idea that one argument can ``defend'' other arguments (as illustrated by the interaction between $A_4$ and $A_{14}$ in Figure \ref{fig:goalmodeling:arg2}), has been shown to be close to how humans reason \cite{rahwan2010behavioral}.
+Furthermore, more fine-grained numerical frameworks have the disadvantage that these exact numbers have to be estimated, which is a known bottleneck for such approaches \cite{druzdzel2000building}.
 

### 2019-02-03 main.tex
main

--- 2018-09-15 main.tex
@@
 \title{RationalGRL: A Framework for Argumentation and Goal Modeling}
-\author{Marc van Zee  \and Floris J. Bex\and Sepideh Ghanavati}
-\institute{Marc van Zee \at Google BRAIN, Z\"urich, Switzerland \\\email{marcvanzee@gmail.com} 
-\and Floris Bex \at Department of Information and Computing Sciences, Utrecht University, Utrecht, The Netherlands 
-				\at Tilburg Institute for Law, Technology, and Society, Tilburg University, Tilburg, The Netherlands \\\email{f.j.bex@uu.nl}
-\and Sepideh Ghanavati \at Department of Computer Science, Texas Tech University, Lubbock TX, USA \\\email{sepideh.ghanavati@ttu.edu} }
+\author{Authors names anonymized for review}
 

### 2019-02-04 Response to reviewers3.txt
small changes response to reviewers

@@
 
-Reviewer 1 again proposes to accept the paper. Reviewer 2 recommends rejection. The editor proposes that we perform major revisions
+Reviewer 1 again proposes to accept the paper. Reviewer 2 recommends rejection. The editor proposes that we perform major revisions.
 
@@
 We formally ground this framework in a logical framework for argumentation that has been accepted as a de-facto standard for non-monotonic argumentation inside and outside the field of AI.
-We give an simply development process, and a fully-working tool.
+We give a simple development process, and a fully-working tool.
 We have further performed an extensive qualitative emprirical coding study (which we call case study in our article), plus a modest expert survey.
@@
 (2006) "An empirical evaluation of the i* framework in a model-based software generation environment", who took 9 months to conduct the case study.
-Furthermore, there are many papers on goal-modeling and related subjects in REJ which perform less empirical evaluation than we do, and the few that do have industrial-strength case studies evaluate a modeling language or tool that was introduced earlier.
+Furthermore, there are many papers on goal-modeling and related subjects in REJ which perform less empirical evaluation than we do, and the few that do have industrial-strength case studies evaluate a modeling language or tool that was introduced in an earlier paper.
 We are not directly aware of papers on goal modeling languages doing extensive experimentation with test and control groups.
@@
 It is a modest survey with a fairly high response rate, particularly as all 16 participants have significant expertise in software and requirements engineering - we could do an experiment with 80 inexperienced undergraduates, but what would that really tell us?
-The answers to the questions mainly point in interesting future directions, but overall the experiment gives modest support to the idea that adding arguments to goal models is the way to go.
+The answers to the survey questions point in interesting future directions, and overall the experiment gives support to the idea that adding arguments to goal models is the way to go.
 
@@
 ======= REPLY #8
-We already included some discussion in the previous revision. We have included an extra sentence on why we think numerical, more fine-grained frameworks are not very suitable.
+We already included some discussion in the previous revision.
+We have included an extra sentence on why we think numerical, more fine-grained frameworks are not very suitable.
+Furthermore, note that the framework is not "binary" in the sense that arguments can only be true or false (as in classical logic): arguments which are at one point considered defeated or overruled can later become justified again if new arguments are added.
 
@@
 ======= REPLY #9
-This remark makes it sound as if there are many grammatical mistakes in the paper.
+This remark makes it sound as if there are *many* grammatical mistakes in the paper.

### 2019-02-04 Response to reviewers3.txt
final changes

--- 2019-02-03 Response to reviewers3.txt
@@
 We give an simply development process, and a fully-working tool.
-We have further performed an extensive qualitative emprirical coding study (which we call case study in our article), plus a modest expert survey.
+We have further performed an extensive coding analysis, plus a modest expert survey.
 
@@
 However, we remain convinced that a large case study or experiment (where e.g.
-one compares test and control groups) would take an unreasonable amount of time and space given the material that is alreadty in the paper.
+one compares test and control groups) would take an unreasonable amount of time and space given the material that is already in the paper.
 See e.g.
@@
 (2006) "An empirical evaluation of the i* framework in a model-based software generation environment", who took 9 months to conduct the case study.
-Furthermore, there are many papers on goal-modeling and related subjects in REJ which perform less empirical evaluation than we do, and the few that do have industrial-strength case studies evaluate a modeling language or tool that was introduced earlier.
+Furthermore, there are many papers on goal-modeling and related subjects in REJ that perform less empirical evaluation than we do, and the few that do have industrial-strength case studies evaluate a modeling language or tool that was introduced earlier.
 We are not directly aware of papers on goal modeling languages doing extensive experimentation with test and control groups.
@@
 
-The (smaller) remarks made by reviewer 2 since Revision 1 have been helpful, and we try to accomodate as many of them as possible.
+The (smaller) remarks made by reviewer 2 since Revision 1 have been helpful, and we try to accommodate as many of them as possible.
 We would like to note that we have never simply refused suggestions or ignored comments.
@@
 the RationalGRL features) as well.
-In any case, clearly we were not able to explain our ideas clearly enough, so we deleted H3 and only mention the user experience questions in passing.
+In any case, to avoid confusion we deleted H3 and only mention the user experience questions in passing.
 
@@
 As we argue above, we think this is an unreasonable amount of work to ask given what is already in the paper.
-Also, it now seems as if our (modest) expert survey is worthless, which we think it is not.
 
@@
 ======= REPLY #5
-Again: yes there are many ways in which we could have done our experiments and surveys differently, better, bigger etc.
-However, the reviewer makes it sound as if the current survey is next to useless, and we strongly disagree with this.
-It is a modest survey with a fairly high response rate, particularly as all 16 participants have significant expertise in software and requirements engineering - we could do an experiment with 80 inexperienced undergraduates, but what would that really tell us?
-The answers to the questions mainly point in interesting future directions, but overall the experiment gives modest support to the idea that adding arguments to goal models is the way to go.
+We agree that are many ways in which we could have done our experiments and surveys differently and more extensively.
+However, we disagree with the insinuation that the current survey is next to useless.
+We provide a modest survey with a fairly high response rate, particularly as all 16 participants have significant expertise in software and requirements engineering.
+Alternatively, it would be possible to do an experiment with a large number of inexperienced undergraduate students, but we believe our current study provides us with more insights.
+Overall, the experiment gives modest support to the idea that adding arguments to goal models is useful and insightful for modelers.
 
@@
 Formal reasoning in propositional logic, taking the semantics of classical logic, is inherently monotonic: you can add propositions (arguments, goals), but these will never change the status (true/false) of existing propositions.
-So what is needed is a new algorithm that shows what happens if new information is added, we cannot simply take an existing tool (theorem prover) for classical logic.
+Formal reasoning in propositional logic would be possible using, for instance, SAT solvers, but this would require a monotonic logic.
+Theorem provers for non-monotonic logics are significantly more complex, and not the focus of our research.
 
@@
 ======= REPLY #9
-This remark makes it sound as if there are many grammatical mistakes in the paper.
+We have revised our entire article in detail and correct a large number of grammatical mistakes.

### 2019-02-04 abstract.tex
final changes

@@
 For instance, the discussion on whether to accept or reject a certain goal and the ultimate rationale for acceptance or rejection cannot be captured in current goal modeling languages.
-In this paper, we apply argumentation techniques from artificial intelligence to a goal modeling approach by using a case study in which stakeholders discuss requirements for a Traffic Simulator.
+In this paper, we apply argumentation techniques from artificial intelligence to a goal modeling approach by using a coding analysis in which stakeholders discuss requirements for a Traffic Simulator.
 We combine a simplified version of a traditional goal modeling approach, the Goal-oriented Requirements Language (GRL), with a formal Practical Reasoning Argument Scheme (PRAS) for reasoning about goals into a new framework (RationalGRL).

### 2019-02-04 background.tex
final changes

@@
 
-In this section, we first introduce our running example, after which we give a brief overview of the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework\footnote{For simplicity's sake, we consider a slightly cut-down version of GRL.
+In this section, we first introduce our running example, after which we give a brief overview of the Goal-oriented Requirements Language (GRL)~\cite{Amyot:2010:EGM:1841349.1841356}, which is the goal modeling language we use to integrate with the argumentation framework.\footnote{For simplicity's sake, we consider a slightly cut-down version of GRL.
 All the core elements of GRL are included but, for example, we have included only one level of contribution strength, whereas GRL contains different levels of positive and negative contribution.
-This simplification does not in any way influence the case study.
-Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements is trivial.}.
-Next, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
+This simplification does not in any way influence the coding analysis.
+Furthermore, extending the GRL part of RationalGRL to include the full set of GRL elements is straightforward.} Next, we discuss the \emph{practical reasoning argument scheme (PRAS)}~\cite{atkinson2007}, an argument scheme that is used to form arguments and counter-arguments about situations involving goals, and we give informal examples of how argument and counterargument can influence the status of beliefs about goals.
 %This will be our starting point in the next section.
@@
 
-Our examples and case study are based on the data produced by a recent series of experiments by Tang et al.
+Our examples and coding analysis are based on the data produced by a recent series of experiments by Tang et al.
 \cite{TangEtal2018}, who in turn base their work on the so-called Irvine experiment~\cite{UCIworkshop}.
@@
 These discussions were recorded and transcribed.
-We used these transcripts for an extensive case study on the basis of which we develop our RationalGRL framework.
+We performed a coding analysis on these transcripts, which we used to develop our RationalGRL framework.
 Furthermore, we also use the traffic simulator case as a running example throughout this paper.
@@
 \includegraphics[]{img/fig_AF1.pdf}
-\caption{PRAS aArguments and Attacks in the Traffic Simulation Example}
+\caption{PRAS Arguments and Attacks in the Traffic Simulation Example}
 \label{fig:pras:example}
@@
 Furthermore, it is not directly clear whether the critical questions as proposed by Atkinson and Bench-Capon~\cite{atkinson2007} actually apply to GRL models.
-In fact, our case study (Section \ref{sect:gmas}) shows that when discussing requirements, people very often do not structure their reasoning nicely in the way that PRAS presents it.
+In fact, our coding analysis (Section \ref{sect:gmas}) shows that when discussing requirements, people very often do not structure their reasoning nicely in the way that PRAS presents it.
 That is, it is rarely the case that the discussants setting up an argument ``We have goal $G$, $A$ realizes $G$ \emph{Therefore} we should perform $A$''.
 A typical discussion is much more unstructured, as is clear from the transcript excerpts in Appendix~\ref{sect:transcripts:excerpts}.
-Thus, if we would use the version of PRAS presented in this section for our argumentation, we would violate requirement 1: The argumentation techniques must capture the actual discussions of the stakeholders or designers in the early requirements engineering phase.
+Thus, if we would use the version of PRAS presented in this section for our argumentation, we would violate Requirement 1: The argumentation techniques must capture the actual discussions of the stakeholders or designers in the early requirements engineering phase.
 Our solution is to develop our own set of argument schemes and critical questions by analyzing transcripts of discussions about the traffic simulator.
-This set of schemes and questions and our case study are described in the next section.
+This set of schemes and questions and our coding analysis are described in the next section.

### 2019-02-04 casestudy.tex
final changes

@@
-\section{Argument Schemes for Goal Modeling: a Case Study}
+\section{Argument Schemes for Goal Modeling: a Coding Analysis}
 \label{sect:gmas}
 
-Recall that \textbf{requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
-To get a sense of such discussions, we performed a case study to examine which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
+Recall that \textbf{Requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
+To get a sense of such discussions, we performed a coding analysis to examine which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
 We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS.
-In this section, we present our case study.
-All original transcripts, coding, and models are available in our online repository, which can be found at:
+In this section, we present our coding analysis.
+All original transcripts, codings, and models are available in our online repository, which can be found at:
 
@@
 They were asked to record their design sessions, and the recordings were subsequently transcribed.
-We used three of these transcripts, totaling 153 pages, for our case study.
+We used three of these transcripts, totaling 153 pages, for our coding analysis.
 

### 2019-02-04 discussion.tex
final changes

@@
 \paragraph{Empirical validation}
-We have performed a case study and found more than 200 incidences of arguments and questions.
+We have performed a coding analysis and found more than 200 occurrences of arguments and questions.
 We also performed an experiment with 16 participants, testing the usability and usefulness of RationalGRL compared to other goal modeling languages.
@@
 
-Our list of argument schemes and critical questions in Table~\ref{table:argument-schemes} was constructed by performing an extensive case study in which we analyzed 153 pages of transcripts of discussions among designers of a traffic simulator information system.
+Our list of argument schemes and critical questions in Table~\ref{table:argument-schemes} was constructed by performing a coding analysis in which we analyzed 153 pages of transcripts of discussions among designers of a traffic simulator information system.
 In this way, we ensure that the argumentation techniques capture the actual discussions of the stakeholders or designers in the early requirements engineering phase (\textbf{requirement 1}).

### 2019-02-04 formalframework.tex
final changes

@@
 
-In Section~\ref{sect:overview}, we have shown through a language definition and informal examples from our case study that it is possible to trace elements of a goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on a goal model, and vice versa (\textbf{requirement 3}).
+In Section~\ref{sect:overview}, we have shown through a language definition and informal examples from our coding analysis that it is possible to trace elements of a goal model back to their underlying arguments (\textbf{Requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on a goal model, and vice versa (\textbf{Requirement 3}).
 A formal rendition of this traceability will be presented in this section, in which we present a logical formalization of the RationalGRL framework.

### 2019-02-04 introduction.tex
final changes

@@
 However, goal models for many complex, real-world information systems -- e.g., air-traffic management systems, systems that support industrial production processes, or government and healthcare services -- are not constructed in a short amount of time, but rather over the course of several workshops with stakeholders and requirements engineers.
-%SG: I reword and add the following sentences as the paragraphs below are very disjunct.
 Developing goal models for such complex and large systems is not a trivial task and can be very cumbersome.
@@
 
-The first challenge for goal modeling approaches, particularly in large projects, is related to its dynamic nature: goal models continuously change and evolve.
+An important challenge for goal modeling approaches, particularly in large projects, is related to its dynamic nature: goal models continuously change and evolve.
 Stakeholders' preferences are rarely absolute, relevant, stable, or consistent~\cite{march1978bounded}.
@@
 Inspired by the work on practical reasoning from Artificial Intelligence, most notably Atkinson and Bench-Capon~\cite{atkinson2007}, we have developed a list of argument schemes that can be used to analyze and guide stakeholders' discussions about goal models.
-This list of argument schemes is based on an extensive case study in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system (requirement 1).
+This list of argument schemes is based on a coding analysis in which we analyzed a set of transcripts containing more than 4 hours of discussions among designers of a traffic simulator information system (Requirement 1).
 
-In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models (requirement 2).
+In order to specify clearly in what way RationalGRL extends GRL, we develop a metamodel, which specifies the traceability links between the arguments based on the schemes and the GRL models (Requirement 2).
 In addition to this meta-model, we provide formal semantics for RationalGRL by formalizing the GRL language in propositional logic and rendering arguments about a GRL model as a formal argumentation framework to which we can apply argumentation semantics~\cite{Dung1995}.
-We, then, formally capture the links between argumentation and goal models as a set of algorithms for applying argument schemes and critical questions about goal models (requirement 3).
+We, then, formally capture the links between argumentation and goal models as a set of algorithms for applying argument schemes and critical questions about goal models (Requirement 3).
 
-In order to support practitioners in using the RationalGRL framework, we propose the RationalGRL development process, which explains steps for developing goal models and posing arguments based on schemes in an integrated way (requirement 4).
-We implement and briefly evaluate a web-based tool for the RationalGRL framework, which works on any modern browser and is developed in Javascript (requirement 5).\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
+In order to support practitioners in using the RationalGRL framework, we propose the RationalGRL development process, which explains steps for developing goal models and posing arguments based on schemes in an integrated way (Requirement 4).
+We implement and briefly evaluate a web-based tool for the RationalGRL framework, which works on any modern browser and is developed in Javascript (Requirement 5).\footnote{The tool and source code can be found at \rationalgrlurl.} The tool is open-source, and designed such that new argument schemes and critical questions can be added easily.
 
@@
 Section~\ref{sect:background} introduces our running example and briefly discusses the basics of GRL and argumentation about goals using argument schemes.
-Section~\ref{sect:gmas} contains the case study and an explanation of how we obtained an initial set of argument schemes and critical questions by coding transcripts from discussions about an information system.
-Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL metamodel, and examples from the case study that illustrate the framework and its language.
+Section~\ref{sect:gmas} contains the coding analysis and an explanation of how we obtained an initial set of argument schemes and critical questions by coding transcripts from discussions about an information system.
+Section~\ref{sect:overview} provides an overview of the RationalGRL framework and the RationalGRL metamodel, and examples from the coding analysis that illustrate the framework and its language.
 In Section~\ref{sect:formalframework}, we provide formal semantics for GRL and RationalGRL, show how RationalGRL models can be translated to GRL models and vice versa, and we develop various algorithms that change a RationalGRL model according to an argument scheme or a critical question.

### 2019-02-04 overview.tex
final changes

@@
 We now give an overview of our RationalGRL framework.
-Through a metamodel and informal examples from our case study we will show that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{requirement 3}).
+Through a metamodel and informal examples from our coding analysis we will show that it is possible to trace elements of the goal model back to their underlying arguments (\textbf{Requirement 2}), and that it is possible to determine the effect of changes in the underlying argumentation on the goal model, and vice versa (\textbf{Requirement 3}).
 A formalization of our framework using formal logic can be found in Section~\ref{sect:formalframework}.
@@
 In Section~\ref{sect:overview:lang}, we discuss the language for RationalGRL models and we provide a metamodel, linking the new concepts of our language to GRL.
-In Section~\ref{sect:overview:examples}, we provide extensive examples from our case study.
+In Section~\ref{sect:overview:examples}, we provide extensive examples from our coding analysis.
 
@@
 
-A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual discussions of stakeholders (\textbf{requirement 1}).
+A core aspect of the RationalGRL framework are the argument schemes, which should be close to the actual discussions of stakeholders (\textbf{Requirement 1}).
 Recall from Section \ref{sect:gmas} that we ended up with a list of argument schemes and critical questions that were found in the transcripts (Table~\ref{table:transcripts:results:argumentschemes}).
@@
 
-\subsection{From Practical Reasoning to RationalGRL Models: Examples from the Case Study}
+\subsection{From Practical Reasoning to RationalGRL Models: Examples from the Coding Analysis}
 \label{sect:overview:examples}
@@
 the top left element of the framework in Figure~\ref{fig:rationalgrl-framework}) on the other hand.
-We provide informal examples of the links between the practical reasoning found in our case study transcripts and RationalGRL models.
+We provide informal examples of the links between the practical reasoning found in our coding analysis transcripts and RationalGRL models.
 The formal grounding for the connection between practical reasoning and RationalGRL models can be found in the RationalGRL metamodel (Section~\ref{sect:metamodel}) and in more detail in the logical formalization in the next section.

### 2019-02-04 tool.tex
final changes

@@
 The previous section contained a formalization of RationalGRL, which forms the starting point of this section.
-In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a development process (\textbf{requirement 4}) and discussing a prototype RationalGRL tool (\textbf{requirement 5}).
+In this section we clarify how practitioners can actually use the RationalGRL framework by proposing a development process (\textbf{Requirement 4}) and discussing a prototype RationalGRL tool (\textbf{Requirement 5}).
 
@@
 
-The final requirement of our framework is that it has tool support (\textbf{requirement 5}).
+The final requirement of our framework is that it has tool support (\textbf{Requirement 5}).
 This is important for various reasons: i) although there are various approaches attempting to combine goal modeling with argumentation, we are not aware of an existing tool support (see Section~\ref{sect:discussion}), ii) it allows us to conduct user studies in the next section, exploring the usability and usefulness of the RationalGRL framework compared to other goal modeling languages, and iii) tool support is an excellent demonstration to show that our formal framework and algorithms actually work.

### 2019-02-04 validation.tex
main.pdf
final changes

@@
 
-In addition to the empirical case study (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a user evaluation with 16 expert users.
+In addition to the coding analysis (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a user evaluation with 16 expert users.
 The objective of this evaluation was to determine whether the users found the additions of RationalGRL (arguments, argumentation semantics, critical questions) useful, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to standard GRL.
@@
 
-The instructions for these three parts are provided in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (part 1) starts with a general explanation of GRL and standard goal modeling similar to the explanation provided in Section \ref{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
+The instructions for these three parts are provided in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (Part 1) starts with a general explanation of GRL and standard goal modeling similar to the explanation provided in Section \ref{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
 
-For part 2, we provided a small part of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the RationalGRL tool and take a screenshot of their final model.
+For Part 2, we provided a small part of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the RationalGRL tool and take a screenshot of their final model.
 The idea is that the participants act as if they are ``present'' at a discussion about the early-stage requirements and are asked to model the goal model corresponding to these requirements.
 
-In order to allow the participants to compare their models to regular GRL models, after the modelling exercise we provide the participants with Figure~\ref{fig:example-small}, which models the discussion for part 2 (Appendix~\ref{sect:survey}) in standard GRL.
+In order to allow the participants to compare their models to regular GRL models, after the modelling exercise we provide the participants with Figure~\ref{fig:example-small}, which models the discussion for Part 2 (Appendix~\ref{sect:survey}) in standard GRL.
 
-For part 3, we provided the participants with a survey\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2}.
-The survey starts with general questions about the participant (years of experience etc.).
+For Part 3, we provided the participants with a survey.\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} The survey starts with general questions about the participant (years of experience etc.).
 The second part of the survey concerns the specific features of RationalGRL, asking participants to rate RationalGRL on a Likert scale from 1 (very useless) to 5 (very useful):
@@
 The survey further contains questions about the RationalGRL tool user experience, such as whether it was easy or difficult to get started with the tool and to model the example case.
-We also ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could in the future be used in practice.
+We furthermore ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could in the future be used in practice.
 
@@
 For H1 and H2, we hypothesize that the participants will, on average, rate the additions of RationalGRL as useful (rating 4) or very useful (rating 5) for Q1, Q2 and Q3, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q4, Q5 and Q6.
-In other words, we say that our null hypothesis and alternative hypothesis for H1 and H2 are as follows:
+In other words, the null hypothesis and alternative hypothesis for H1 and H2 are as follows:
 \begin{itemize}
@@
 Note that for each question (Q1-Q6) we can test the null and alternative hypothesis separately, giving insight into exactly which additions of RationalGRL were deemed useful.
-We set the significance level for each hypothesis at $\aplha = 0.05$
+We set the significance level for each hypothesis at $\alpha = 0.05$.
 
 \paragraph{Participants}
-We asked 16 participants in our network to participate in our experiment.
-Most of the participants were therefore either employed at the first author's company or staff and (ex-)students from the second author's university.
+We asked 16 participants in our network to participate in the experiment.
+Most of the participants were therefore either employed in industry or staff and (ex-)students from university.
+For those working in industry, all participants took the experiment outside of working hours to avoid conflicts of interest.
 Hence, two thirds of the participants had either a PhD or Master degree, 20\% a Bachelor degree, and one respondent responded with ``Other''.
@@
 First, the participants had to perform the modeling task.
-They produced RationalGRL models containing on average 3 arguments (in addition  to goals, tasks, etc.) from the short transcript.
+They produced RationalGRL models containing on average three arguments (in addition  to goals, tasks, etc.) from the short transcript.
 Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.\footnote{The full results of the survey can be downloaded from \url{www.rationalgrl.com}, on the page ``Empirical study''.}

### 2019-02-04 Response to reviewers3.txt
small sentence & ref coding

@@
 We formally ground this framework in a logical framework for argumentation that has been accepted as a de-facto standard for non-monotonic argumentation inside and outside the field of AI.
-We give an simply development process, and a fully-working tool.
+We give a simple development process, and a fully-working tool.
 We have further performed an extensive coding analysis, plus a modest expert survey.
@@
 ======= REPLY #9
-We have revised our entire article in detail and correct a large number of grammatical mistakes.
+We have revised our entire article in detail and corrected a large number of grammatical mistakes.

### 2019-02-04 casestudy.tex
small sentence & ref coding

@@
 Recall that \textbf{Requirement 1} of our RationalGRL framework is that the argumentation techniques should be close to the actual discussions of stakeholders or designers in the early requirements engineering phase.
-To get a sense of such discussions, we performed a coding analysis to examine which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions.
-We manually coded transcripts of such discussions using a list of argument schemes and critical questions based on GRL and PRAS.
+To get a sense of such discussions, we performed a coding analysis to examine which types of discourse are used during discussions of system requirements, and how these discourse types can be captured as argument schemes and critical questions \cite{saldana2015coding}.
+The type of coding we performed is \emph{a priori coding}: we manually coded transcripts of such discussions using a pre-existing list of argument schemes and critical questions based on GRL and PRAS.
 In this section, we present our coding analysis.

### 2019-02-05 Response to reviewers3.txt
Fixed small things

@@
 
-The (smaller) remarks made by reviewer 2 since Revision 1 have been helpful, and we try to accommodate as many of them as possible.
+The (smaller) remarks made by reviewer 2 since Revision 1 have been helpful, and we tried to accommodate as many of them as possible.
 We would like to note that we have never simply refused suggestions or ignored comments.
@@
 As we argue above, we think this is an unreasonable amount of work to ask given what is already in the paper.
+We reserve this idea for another paper which is tailored to relevant journals such as "Journal of Empirical Software Engineering (EMSE)".
 
@@
 ======= REPLY #9
-We have revised our entire article in detail and corrected a large number of grammatical mistakes.
+We have revised our entire article in detail and correct a large number of grammatical mistakes.

### 2019-02-05 validation.tex
Fixed few things

@@
 In addition to the coding analysis (Section~\ref{sect:gmas}), which forms the basis of the RationalGRL framework, we also performed a user evaluation with 16 expert users.
-The objective of this evaluation was to determine whether the users found the additions of RationalGRL (arguments, argumentation semantics, critical questions) useful, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to standard GRL.
+The objective of this evaluation was to determine whether the users found the components of RationalGRL (arguments, argumentation semantics, critical questions) useful, and whether they found it easier to keep track of and express opinions and beliefs using RationalGRL compared to standard GRL.
 Regarding the usefulness of the RationalGRl features and the to GRL, the following two broad hypotheses follow from our main points in this paper:
@@
 \begin{itemize}
-\item[H1] The additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modeling in GRL.
-\item[H2] The additions of RationalGRL make it easier to express, determine the effect of and communicate to other stakeholders one's opinions and beliefs about a goal model.
+\item[H1] The components of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful extension to standard goal modeling in GRL.
+\item[H2] The components of RationalGRL make it easier to express, determine the effect of and communicate to other stakeholders one's opinions and beliefs about a goal model.
 \end{itemize}
@@
 
-In addition, we wanted to know some basics about user experience.
-The aim here was not to conduct a full user experience study for the RationalGRL tool.
-Rather, we were interested to know if the user experience of the tool was either very good (easy to use) or very bad (difficult to use), as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive towards a particular modelling language, whereas irritations about a bad tool will lead to a more negative disposition.
+In addition, we wanted to know some basics about the user experience.
+The aim here is not to conduct a full user experience study for the RationalGRL tool.
+Rather, we are interested to know if the user experience of the tool is either very good (easy to use) or very bad (difficult to use), as this could influence the participants ideas about the RationalGRL framework and language -- a very nice and easy to use tool tends to make people more positive towards a particular modeling language, whereas irritations about a bad tool will lead to a more negative disposition.
 
@@
 
-The idea was to have our participants perform a small modeling task, and then ask them what they thought of RationalGRL. Thus, our experiment consisted of three parts:
+The idea was to have our participants perform a small modeling task, and then ask them what they thought of RationalGRL. Thus, our experiment consists of three parts:
 
@@
 
-The instructions for these three parts are provided in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (Part 1) starts with a general explanation of GRL and standard goal modeling similar to the explanation provided in Section \ref{sect:background:grl}, then briefly discusses the additions of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
+The instructions for these three parts are provided in a single document that we sent out to the participants.\footnote{See \url{www.rationalgrl.com}, page ``Empirical Study''} The explanation (Part 1) starts with a general explanation of GRL and standard goal modeling similar to the explanation provided in Section \ref{sect:background:grl}, then briefly discusses the components of RationalGRL (argument, attack, critical questions) and finally the tool is briefly described (similar to Section \ref{sect:tool}).
 
-For Part 2, we provided a small part of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the RationalGRL tool and take a screenshot of their final model.
+For Part 2, we provided a small section of a transcript and some context information (see Appendix~\ref{sect:survey}) and asked the participants to model this in the RationalGRL tool and take a screenshot of their final model.
 The idea is that the participants act as if they are ``present'' at a discussion about the early-stage requirements and are asked to model the goal model corresponding to these requirements.
 
-In order to allow the participants to compare their models to regular GRL models, after the modelling exercise we provide the participants with Figure~\ref{fig:example-small}, which models the discussion for Part 2 (Appendix~\ref{sect:survey}) in standard GRL.
+In order to allow the participants to compare their models to standard GRL models, after the modeling exercise we provide the participants with Figure~\ref{fig:example-small}, which models the discussion for Part 2 (Appendix~\ref{sect:survey}) in standard GRL.
 
-For Part 3, we provided the participants with a survey.\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} The survey starts with general questions about the participant (years of experience etc.).
+For Part 3, we provided the participants with a survey.\footnote{https://goo.gl/forms/fDSMUnAV20wy7kbY2} The survey starts with general questions about the participants (years of experience etc.).
 The second part of the survey concerns the specific features of RationalGRL, asking participants to rate RationalGRL on a Likert scale from 1 (very useless) to 5 (very useful):
@@
 \end{itemize}
-The following questions ask the participants to rate RationalGRL vs. standard goal modeling language  on a Likert scale from 1 (much more difficult) to 5 (much easier):
+The following questions ask the participants to rate RationalGRL vs. a standard goal modeling language on a Likert scale from 1 (much more difficult) to 5 (much easier):
 \begin{itemize}
@@
 The survey further contains questions about the RationalGRL tool user experience, such as whether it was easy or difficult to get started with the tool and to model the example case.
-We furthermore ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could in the future be used in practice.
+We, furthermore, ask open questions about the strengths and weaknesses of the tool, possible improvements and whether the participant thinks the tool could be used in practice in future.
 
 In order to test our main hypotheses H1 and H2, we have to formulate the appropriate null hypotheses and alternative hypotheses.
-For H1 and H2, we hypothesize that the participants will, on average, rate the additions of RationalGRL as useful (rating 4) or very useful (rating 5) for Q1, Q2 and Q3, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q4, Q5 and Q6.
+For H1 and H2, we hypothesize that the participants will, on average, rate the components of RationalGRL as useful (rating 4) or very useful (rating 5) for Q1, Q2 and Q3, and as making it easier (rating 4) or significantly easier (rating 5) for questions Q4, Q5 and Q6.
 In other words, the null hypothesis and alternative hypothesis for H1 and H2 are as follows:
@@
 \end{itemize}
-Note that for each question (Q1-Q6) we can test the null and alternative hypothesis separately, giving insight into exactly which additions of RationalGRL were deemed useful.
+Note that for each question (Q1-Q6), we can test the null and alternative hypothesis separately, giving insight into exactly which components of RationalGRL were deemed useful.
 We set the significance level for each hypothesis at $\alpha = 0.05$.
@@
 For those working in industry, all participants took the experiment outside of working hours to avoid conflicts of interest.
-Hence, two thirds of the participants had either a PhD or Master degree, 20\% a Bachelor degree, and one respondent responded with ``Other''.
+Two thirds of the participants had either a PhD or Master degree, 20\% a Bachelor degree, and one respondent responded with ``Other''.
 All but one participant had a year or more experience with software development.
-The average experience was 6.2 years (standard deviation 6.7), with 10 participants having less than 5 years of experience and 6 participants having 8 or more years of experience, with one participant having 25 years of experience.
+The average experience was 6.2 years (standard deviation 6.7), with 10 participants having less than 5 years of experience, 6 participants having 8 or more years of experience and one participant having 25 years of experience.
 The participants also judged themselves to be quite competent in  early-phase requirements engineering: on average they gave themselves a rating of 3.2 out of 5 (standard deviation 1.3) with half of the participants claiming they were at least ``competent'' (rating 4) or ``very competent'' (rating 5).
-However, experience with goal modeling languages was markedly less: the average rating was 1.9 (standard deviation 1.3), with 9 participants claiming never to have used a goal modeling language (rating 1), and only two participants displaying regular use (monthly, rating 4, or weekly, rating 5).
+However, experience with goal modeling languages was markedly less: the average rating was 1.9 (standard deviation 1.3), with 9 participants claiming never used a goal modeling language (rating 1), and only two participants displaying regular use (monthly, rating 4, or weekly, rating 5).
 The participants that had experience with goal modeling languages had mostly used i* (5 users), with 2 users being familiar with GRL and 2 users having used another goal modeling language.
@@
 \end{tabularx}
-\caption{Participant ratings and statistical results of the usefulness of the additions of RationalGRL}
+\caption{Participant ratings and statistical results of the usefulness of the components of RationalGRL}
 \label{table:survey:table2}
@@
 \end{tabularx}
-\caption{Participant ratings and statistical results of whether the additions of RationalGRL make reasoning about a goal model easier}
+\caption{Participant ratings and statistical results of whether the components of RationalGRL make reasoning about a goal model easier}
 \label{table:survey:table3}
@@
 They produced RationalGRL models containing on average three arguments (in addition to goals, tasks, etc.) from the short transcript.
-Some of the examples of RationalGRL models the users created can be found in Appendix~\ref{sect:survey-screenshots}.\footnote{The full results of the survey can be downloaded from \url{www.rationalgrl.com}, on the page ``Empirical study''.}
+Some of the examples of RationalGRL models, created by the users, can be found in Appendix~\ref{sect:survey-screenshots}.\footnote{The full results of the survey can be downloaded from \url{www.rationalgrl.com}, on the page ``Empirical study''.}
 
-After the modelling task, the survey questions were asked.
+After the modeling task, the survey questions were asked.
 We first asked the participants some questions about user experience.
-The users where generally positive about RationalGRL's clear UI, the fact that the tools gives a nice overview of the goals and opinions in a case and that it was straightforward to use and understand.
+The users where generally positive about RationalGRL's clear User Interface (UI), the fact that the tool gives a nice overview of the goals and opinions in a case and that it was straightforward to use and understand.
 With respect to weaknesses and improvements to the RationalGRL tool, the users mentioned additional UI functionalities such as the possibility to save models, having an ``undo'' function, and flipping the arrow after adding them.
@@
 \paragraph{Hypothesis H1}
-The first analysis concerns hypothesis H1, whether the additions of RationalGRL -- arguments, critical questions and determination of the status of arguments -- are a useful addition to standard goal modelling in GRL.
+The first analysis concerns hypothesis H1, whether the components of RationalGRL -- i.e.
+arguments, critical questions and determination of the status of arguments -- are a useful extension to standard goal modelling in GRL.
 Table~\ref{table:survey:table2} summarizes the medians, modes, and p-values for a one-tailed single-sample sign test for the relevant questions Q1-Q3.
@@
 the distribution of answers for each question.
-As is evident from the results, the participants were generally positive about the additions of RationalGRL.
+As it is evident from the results, the participants were generally positive about the components of RationalGRL.
 For Q1 and Q3 the p-values are both significantly lower than $\alpha = 0.05$, which means that for these questions we can accept H1$_{a}$ and reject H1$_{0}$.
 This means that the addition of arguments and the use of arguments to determine the acceptability of elements of the goal model are deemed to be useful to very useful by the participants.
-As can be seen in Table~\ref{table:survey:table2}: no participant found the addition of arguments and argument status useless or very useless, and 78.6\% (Q1) respectively 85.7\% (Q3) found the additions useful or very useful.
+As can be seen in Table~\ref{table:survey:table2}: no participant found the addition of arguments and argument status useless or very useless, and 78.6\% (Q1) respectively 85.7\% (Q3) found the new components useful or very useful.
 Furthermore, these findings fit with the answers to the open questions (see Section \ref{sec:survey:results}).
@@
 This means that the participants did not think the addition of critical questions like to the RationalGRL tool was significantly useful.
-So we can say that the addition of critical questions is perhaps not as useful as we had hypothesized.
-This again fits the descriptive statistics in Table\ref{table:survey:table2}: 21.4\% of respondents found the critical questions ``useless''.
+Thus, we can say that the addition of critical questions is perhaps not as useful as we had hypothesized.
+This again fits the descriptive statistics in Table\ref{table:survey:table2}: 21.4\% of respondents found critical questions ``useless''.
 Furthermore, in the open questions a number of participants indicated that the exact difference between some of the critical questions was unclear, and that some of the critical questions seemed unnecessary, at least for the small modeling exercise given in the experiment.
@@
 \paragraph{Hypothesis H2}
-The second hypothesis states that additions of RationalGRL make it easier to express one's opinions and beliefs about a goal model, determine the effect of one's opinions and beliefs about a goal model and communicate one's opinions and beliefs about a goal model to other stakeholders.
+The second hypothesis states that components of RationalGRL make it easier to express one's opinions and beliefs about a goal model, determine the effect of one's opinions and beliefs about a goal model and communicate one's opinions and beliefs about a goal model to other stakeholders.
 Table~\ref{table:survey:table3} provides a summary of the statistical analysis of the relevant questions Q4-Q6.
 As can be seen, all the scores are significantly higher than the median of 3 for $\alpha = 0.05$, which means that we can accept H2$_{a}$ and reject H2$_{0}$ for all the separate questions Q4-Q6.
-In other words, the fact that the participants rate the additions of RationalGRL as making it, on average, easier to reason with their beliefs and opinions about a goal model is very unlikely to be due to chance.
+In other words, the fact that the participants rate the components of RationalGRL as making it, on average, easier to reason with their beliefs and opinions about a goal model is very unlikely to be due to chance.
 
-With respect to Q4, participants say that arguments make it much easier to express their opinions.
+With respect to Q4, participants stated that arguments make it much easier to express their opinions.
 None of the participants found expressing their beliefs (much) more difficult, and 69.3\% found it easier or much easier.
@@
 One participant remarked that ``The reasoning [in the example] seems simplistic [...] the added value of [the formal argumentation] would be justified if the reasoning is more complex''.
-So the simplistic case in the modeling exercise possibly detracts from the results for Q5 and Q3.
+Thus, the simplistic case in the modeling exercise possibly detracts from the results for Q5 and Q3.
 Finally, participants were less sure that arguments would help in communicating the goal model to others.
@@
 
-Another threat to internal validity is that many of the participants had no experience with goal modelling languages at all.
+Another threat to internal validity is that many of the participants had no experience with goal modeling languages at all.
 Hence, the perceived usefulness of the arguments could be influenced by the fact that the participants, for the first time, were given a goal modeling language to perform structured requirements analysis.
-In other words, perhaps they found a structured modelling language generally quite useful.
+In other words, perhaps they found a structured modeling language generally quite useful.
 One way to better check the difference between standard goal modeling and RationalGRL would be to first provide users with a standard goal modeling tool and then have them use RationalGRL.
-An even more thorough way to test the usefulness of RationalGRL over, for example, normal GRL would be to compare a test and control group, where the test group uses RationalGRL and the control group uses normal GRL.
+An even more thorough way to test the usefulness of RationalGRL over, for example, standard GRL would be to compare a test and control group, where the test group uses RationalGRL and the control group uses standard GRL.
 We leave these types of further empirical studies for future research.
@@
 As the number of participants is fairly small (16), this threatens the generalizability of our findings.
-However, the type of participant is realistic: software engineers having a lot of experience with requirements engineering, as opposed to, for example, engineering students with little experience.
+However, the type of participants is realistic: software engineers having a lot of experience with requirements engineering, as opposed to, for example, engineering students with little experience.
 Furthermore, the results are in line with existing research on design rationale.
 On the one hand, these existing studies, like our study, show that structured modeling languages suffer from the common problem of high cognitive overhead \cite{shum2006hypermedia}.
-One the other hand, studies show that the inclusion of argumentation and critical questions in design and requirements discussions improve the reasoning \cite{razavian2016two,TangEtal2018}.
+One the other hand, these studies show that the inclusion of argumentation and critical questions in design and requirements discussions improve the reasoning \cite{razavian2016two,TangEtal2018}.
 
@@
 This is a positive signal for the addition of arguments to goal modeling languages, and also for our formal approach, as just adding arguments as, for example, labels on goals and tasks does not allow one to compute the status of elements in a goal model in the way our approach does.
-Furthermore, automatically determining the status of elements will perhaps have a bigger impact when working with larger, more complex goal models, for which manually determining the impact of arguments is not feasible.
+Furthermore, automatically determining the status of elements will have a bigger impact when working with larger, more complex goal models, for which manually determining the impact of arguments is not feasible.
 \item \emph{Critical questions need to be clear and have a prominent place in the process of building a goal model.} While participants overall found the critical questions (CQ's) useful, their impact was markedly less than the addition of arguments.
